### 2021-07-07 main.Rmd
adding all files from paper-graclust

---
title: Clustering of residential Australian smart meter customers

# to produce blinded version set to 1
blinded: 0

authors:
- name: Sayani Gupta
  affiliation: Department of Econometrics and Business Statistics, Monash University
  thanks: "Email: Sayani.Gupta@monash.edu"

- name: Rob J Hyndman
  affiliation: Department of Econometrics and Business Statistics, Monash University

- name: Dianne Cook
  affiliation: Department of Econometrics and Business Statistics, Monash University

keywords:
- data visualization
- statistical distributions
- time granularities
- calendar algebra
- periodicties
- grammar of graphics
- R

abstract: |
 Clustering elements based on behavior across time granularities 
bibliography: [bibliography.bib]
preamble: >
  \setlength {\marginparwidth }{2cm}
  \usepackage{mathtools,amssymb,booktabs,longtable,todonotes,amsthm}
  \def\mod{~\text{mod}~}
output:
  bookdown::pdf_book:
    base_format: rticles::asa_article
    fig_height: 4
    fig_width: 6
    fig_caption: yes
    dev: "pdf" 
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(ggplot2)
library(gravitas)
library(gracsR)
library(ggdendro)
library(dplyr)
library(readr)
library(visdat)
library(ggplot2)
library(tidyverse)
library(naniar)
library(here)
library(tsibble)
library(knitr)
```

```{r external, include = FALSE, cache = FALSE}
knitr::read_chunk(here('script/smart_meter.R'))
```

# Introduction

<!-- Description of data available -->
The Smart Grid, Smart City (SGSC) project, which rolled out Australia's first commercial-scale smart grid was implemented across eight local government areas in New South Wales (NSW). Data from more than 13,000 household electricity smart meters is obtained as part of that project. It provides half-hourly energy usage and demographic data for Australia, as well as detailed information on appliance use, climate, retail and distributor product offers, and other related factors. The trials were based in Newcastle, New South Wales, but also covered areas in Sydney CBD, Newington, Ku-Ring-Gai, and the rural township of Scone. The load time series is asynchronous as it is observed for these households for unequal time lengths and consists of missing observations.

\noindent The massive amount of data generated in such projects could be overwhelming for analysis. Electricity utilities can utilize the consumption patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers. Beyea (2010) has pointed out, there has been little discussion or exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data. Thus, there is a scope to investigate and analyze these data in various ways for a greater understanding of consumption patterns and how they correlate with other economic, physical or geographical factors. In this work, we are interested to see how we can utilize this dataset to group different customers with similar periodic behavior. Towards this goal, this chapter aims to: (a) describe the contents of the data set in SGSC database that we can utilize, and (b) propose a clustering algorithm to group customers with similar periodic behaviors. The distance metric introduced in Chapter 2 will be the inputs for this cluster analysis. One of the advantages of using our 
approach is that the technique is based on probability distributions instead of raw data. Many clustering approaches are limited by the type of noisy, patchy, and unequal time-series common in residential data sets. Since the distance measure considered is based on differences in probability distribution of time series, it is likely to be less sensitive to missing or noisy data.

The following contributions are made through the following chapter:

 * Present a cluster analysis of SGSC dataset to group households with similar periodic behavior
 * Cluster validation by relating to external data

<!-- Electricity smart meter technology is increasingly being deployed in residential and commercial buildings. For example, in the state of Victoria, Australia, it is a state government policy that all households have a smart meter fitted (Victoria State Government, 2015), which has resulted in 97% of the approximately 2.1 million households in Victoria having a smart meter installed in their home. These smart meters collect energy usage information at half-hourly intervals, resulting in over 35 billion half-hourly observations per year across all households in the state. Clustering electricity consumption patterns can enable electricity utilities to develop targeted tariffs for individual groups alleviating the problem of volatility in production by capitalizing on the flexibility of consumers. -->

<!-- Clustering households using only smart meter consumption data could also provide value in a societal setting by combining the findings and external data like weather conditions, socio-economic or other demographic factors of those households. -->

<!-- Background literature and shortcomings -->

# Preliminary exploratoration

## Getting data

<!-- _Electricity use interval reading data_   -->
<!-- The data from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee). -->

The entire data is procured from CSIRO. A subset of this data is also available from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee). It consists of the following data sets. 
_1. CustomerData:_ 78720 customers with 62 variables about them 
_2. EUDMData:_ 300 billion half-hourly consumption level data  
_3. OffersData:_ Method of contact to customer to join SGSC customer trial, either door-to-door (D2D) or via Telesales  
_4. PEResponseData:_ Peak Events response customer wise  
_5. PETimesData:_ Peak Events time stamps

Only _CustomerData_ and _EUDMData_ are relevant for the clustering goals of this paper. _EUDMData_ contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total. `CustomerData` provides demographic data for 78,720 customers with information on their Local Government Area amongst others.

## Missing Data

Electricity usage for some customers may become unavailable due to power outage or not recording their usage properly, thus resulting in implicit missing values in the database. It is interesting to explore where missing-ness occurs or if there is a relationship between the underlying missing patterns. We use the R package `tsibble` to do this.

```{r, out.width="100%"}
include_graphics("figs/missing-data.png")
```

## Raw data

Since the smart meters have been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths
for each household. Since, general supply is available for only 13,735 customers, we will restrict ourselves to look at the LGA information for these customers only. We find that there are only 26 LGA that is covered for these customers.

_Weather data_  
This data is obtained through [Australian Government Bureau of Meteorology](BOM)(http://www.bom.gov.au/) and provides hourly data for nearest weather stations for all the LGAs

This section familiarizes the 13,000 SGSC households through visualization and provide a detailed layout of data structures (like missing observations/number of customers/number of observations) and also the external data that needs to be utilized for validating the clustering process. The [ABS TableBuilder](https://www.abs.gov.au/websitedbs/censushome.nsf/home/tablebuilder) has census data from 2011 and 2016. The data is at SA2 and LGA levels. However, some of the LGA in NSW changed between 2011 and 2016 and hence there would not be a one-to-one correspondence between the LGAs. Weather, notably temperature (and humidity) can be the main driver(s) for energy usage. In NSW many households have electric heaters so their use can impact winter energy use and air-conditioners can impact summer energy use. Relevant weather data could be obtained from the [Bureau of Meteorology](www.bom.gov.au). Some weather stations have 30 minute (sometimes even smaller interval) weather data. Potentially, there could be lag effects of weather on energy usage which should be considered.

\noindent A dataset of 100 SGSC homes has been used to lay out the structure to be used for analyzing the big dataset. The smaller dataset contains half-hourly kwh values form 2012 to 2014 and has asynchronous time series distributed evenly over the observation period (Figure \ref{fig:elec-raw}), similar to the bigger data set. Figure \ref{fig:count-gaps} can be used to interpret missingness in the data, where the customers are arranged from maximum to minimum missing. It looks like data is most missing before 2013 and for a particular date in 2014.

```{r load-elec, message = FALSE, warning=FALSE}

```

```{r miss-data}
empty_as_na <- function(x){
    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
    ifelse(as.character(x)!="", x, NA)
}

```

```{r vismiss, eval = FALSE}
## transform all columns
customer_data <- customer_data %>% mutate_each(funs(empty_as_na))

vis_miss(customer_data, warn_large_data = FALSE) + 
  theme(axis.text.x = element_text(angle = 90, size = 5))
```

```{r miss-var, eval = FALSE}
gg_miss_var(customer_data) + theme_bw() +
  theme(axis.text.y = element_text(size = 4))
```

```{r}
elec_ts <- read_rds("data/elec_ts.rds")
```

```{r raw-50, fig.height = 8, fig.cap="The raw data for 50 households are shown. It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed."}

```

# Clustering methodology

## Common load clustering techniques of smart meter data

The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings. None of the 34 selected papers which focus on clustering consumption are based on Australian smart meter data. The clustering is frequently applied directly to the raw data without scrutinizing for auto correlation and periodicity. The algorithm most ubiquitously employed is K-Means. But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters. Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results. @Tureczek2018-ha has shown that a transformation of data to incorporate autocorrelation before K-Means clustering can improve performance and enable K-Means to deliver smaller clusters with less within-cluster variance. However, it does not explain the cluster composition by combining it with external data. Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf). Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series. As the methods are often situation specific, 
it makes sense to compare them on the performance rather than any standard performance metric. A type of clustering based on information theory such as Shannon or Renyi entropy and their variants are addressed in , which differs from typical methods adopted for electricity consumer classification, based on the Euclidean distance notion. @Motlagh2019-yj presents strategy to address the problems on patchy, and unequal time-series common in residential data sets by converting load time series into map models. Most time-series clustering models are limited to handling time domain with same start and end date and time. Most of the solutions to handle this like longest common subsequence, dynamic time warping are prone to computational limit with increased length of the series.

<!-- In contrast to models, a feature-based strategy is used to explicitly define or automatically extract the curves’ key time features, for instance by application of PCA on the daily curves [ -->

<!-- Most papers discussed in Tureczek2017-pb fail to accept smart meter readings as time series data, a data type which contains a temporal component. The omission of the essential time series features in the analysis leads to the application of methods that are not designed for handling temporal components. K-Means ignores autocorrelation, unless the input data is pre-processed. The clusters identified in the papers are validated by a variety of indices, with the most prevalent -->
<!-- being the cluster dispersion index (CDI) [22–24], the Davies–Bouldin index (DBI) [25,26] and the mean index adequacy (MIA) [8,13]. -->

## The proposed load-clustering technique

The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_s(v)$ denotes the probability distribution from observation $s$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points. In our work, we are using $s$ as denoting a household and the underlying variable as the electricity demand. So $f_s(v)$ is the distribution of household $s$ and $v$ is electricity demand.
Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$. Each household consists of a $J*K$ distributions one harmony. We compute the distributional difference
between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$.
$wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies. If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony. We aim to have pockets of households showing similar periodic behavior by conidering $wpd$ vlaues for different harmonies and some time series features. The features should also characterize probability distributions of different household.

<!-- grouping probability distributions across a harmony. This clustering algorithm is adopted to remove or appropriately adjust for auto correlation and unequal length in the data. The method could be further extended by clustering probability distributions conditional on one or more cyclic granularities. The following are some of the advantages of our proposed method. -->

<!-- Trend and seasonality are common features of time series, and it is natural to characterize a time series by its degree of trend and seasonality. By considering the probability distributions through the use of $wpd_{norm_{s,t}}(A,B)$, these features of the time series are lost and hence there is no need to de-trend or de-seasonalize the data before performing the clustering algorithm. -->

 - Dimension reduction: If each $P_{i, j, k}$ be considered to be a point in the space, key $i$ would have $mp$ dimensions as opposed to $n_i$ dimensions in case of considering raw data. Hence for a large number of observations ($n_i>>mp$), this approach benefits by transitioning to a lower dimension.

 - Avoid loss of information due to aggregation: This approach ensures key characteristic information of the data is not lost due to averaging or aggregation measures in an attempt to transition to a lower dimension. Hence, this approach could be thought to somehow balance the drawback of considering raw data or aggregated data.

- Robustness to outliers: This approach could be adapted to be robust to outliers and extreme behaviors by trimming the tails of the probability distributions.

- Non-synchronized observed time periods: Considering probability distribution would imply the clustering process can handle keys that are observed over periods of time that are overlapping but don't necessarily coincide.

- Similar periodic behavior: Since cyclic granularities are considered instead of linear granularities, clustering would group keys that have similar behavior across these cyclic granularities. This implies they will be grouped according to their periodic behavior and not on the linear stretch of time over which they are observed.

The methodology is again applied to the 100 households and some preliminary results are presented. Figure \ref{fig:mds} show the multidimensional scaling of the 100 households with colors representing groups presented by hierarchical clustering method. The clusters have very different size and are far apart with no overlapping. They should be distinct from each other in terms of their periodic patterns. Figure \ref{fig:cyclic-cluster} shows the cyclic presentation of time where it could be viewed in details how these clusters are different with respect to the periodic pattern (day-of-week, hour-of-day). Distribution of energy demand across day-of-week and hour-of-day for three groups clustered through the proposed methodology is shown. For group-1, the morning peak looks very sharp, for group-2 the morning and evening peaks are rounded and spread over few hours, for group-3 peaks are less rounded. They differ in their relationship between 75th and 90th percentile. Also, the 75th and 90th percentile of group 2 are quite close, implying the behavior of this group is more regular.

 <!-- - Handling autocorrelation: Autocorrelation in the time series is likely to get removed from considering probability distributions and cyclic granularities in the clustering algorithm. -->

```{r,warning=FALSE, message=FALSE,eval = FALSE}
library(readr)
library(here)
library(dplyr)
library(tsibble)
library(gravitas)
library(purrr)

set.seed(7777)
all_data <- read_rds(here("../paper-hakear/paper/sim_table/all_data_scalar.rds"))
sm_cust_data <- read_rds(here("../paper-hakear/paper/data/sm_cust_data.rds"))

sm_list <- sm_cust_data %>%
  filter(customer_id == "10006414") %>% 
  create_gran("day_week") %>% 
  tibble() %>% 
  ungroup() %>% 
  group_split(day_week)

p <- seq(0.01, 0.99,0.01)

dist_list <- purrr::map(sm_list, function(x){
    map(p, function(y){
       quantile(x$general_supply_kwh, probs = y)
    }) %>% unlist()
}) %>% bind_rows(.id = "gran_id")
# 
# lag <- dist_list[2,][-1] %>% t() %>% 
#   tibble() %>% 
#   lag() %>% acf()

```

```{r mds, warning=FALSE, message=FALSE, fig.cap="Multidimensional scaling of the 100 households is shown along with the groups formed by the hierarchical clustering.", eval = FALSE}
set.seed(7777)
sm_unique_cust <- sm_cust_data %>% distinct(customer_id)
sm_cust_samp <- sm_cust_data %>% 
  #filter(customer_id %in% sm_30) %>% # for demo show 30 customers
  #filter(year(reading_datetime) == 2012) %>% #for demo just show one column
  as_tsibble(index = reading_datetime,
             key = customer_id) %>% 
  create_gran("day_week") %>% 
  create_gran("hour_day")

#all_data <- write_rds(all_data, "paper/sim_table/all_data.rds")

facet_variable = "day_week"
x_variable = "hour_day"

one_harmony_data <- all_data %>%
  #filter(customer_id %in% sm_30) %>% # for demo show 30 customers
  filter(facet_variable == "day_week",
         x_variable == "hour_day") %>% 
  mutate(customer_id = as.character(customer_id))

mds<- one_harmony_data %>%
  select(customer_id, facet_variable, x_variable, wpd_norm) %>% 
  dist() %>%          
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")
rownames(mds) <- unique(one_harmony_data$customer_id)
mds <- mds %>%
  mutate(customer_id = as.character(unique(one_harmony_data$customer_id))) %>%
  select((customer_id)
         , Dim.1, Dim.2)

# ggpubr::ggscatter(mds,
#                   x = "Dim.1",
#                   y = "Dim.2", 
#                   label = rownames(mds),
#                   size = 1,
#                   repel = TRUE)

# clustering as well

d = one_harmony_data %>%
  select(customer_id, facet_variable, x_variable, wpd_norm) %>% 
  dist()

hc = stats::hclust(d,method="complete")
#plot(hc)
groups<-cutree(hc, k=3)

all_data_cluster <- cbind(one_harmony_data, groups) %>% 
  left_join(mds, by = "customer_id") %>% 
  mutate(groups = as.factor(groups))

# q2 <- ggpubr::ggscatter(all_data_cluster,
#                   x = "Dim.1",
#                   y = "Dim.2", 
#                   label = rownames(mds),
#                   color = "groups",
#                   size = 1,
#                   repel = TRUE)

q2 <- all_data_cluster %>% 
  ggplot() +
  geom_point(aes(x = Dim.1, y = Dim.2, color = groups)) +
  scale_color_viridis_d()

 q2
```

```{r linear-cluster, eval = FALSE}
group1 <- all_data_cluster %>% 
  filter(groups == 1)

group2 <- all_data_cluster %>% 
  filter(groups == 2)

group3 <- all_data_cluster %>% 
  filter(groups == 3)

g1 <- sm_cust_samp %>% 
  filter(customer_id %in% group1$customer_id) %>% 
   ggplot(aes(x=reading_datetime, 
             y= general_supply_kwh), alpha = 0.5) + 
  geom_line() +
  scale_fill_gradient2() +
  facet_wrap(~customer_id, nrow = 1) +
  theme(strip.text.x = element_blank()) +
  theme_void() +
  ylab("group 1")

g2 <- sm_cust_samp %>% 
  filter(customer_id %in% group2$customer_id) %>% 
   ggplot(aes(x=reading_datetime, 
             y= general_supply_kwh), alpha = 0.5) + 
  geom_line() +
  scale_fill_gradient2() +
  facet_wrap(~customer_id, nrow= 1) +
  theme(strip.text.x = element_blank()) +
  theme_void() +
  ylab("group 2")

g3 <- sm_cust_samp %>% 
  filter(customer_id %in% group3$customer_id) %>% 
   ggplot(aes(x=reading_datetime, 
             y= general_supply_kwh), alpha = 0.5) + 
  geom_line() +
  scale_fill_gradient2() +
  theme(strip.text.x = element_blank()) +
  theme_void() +
  ylab("group 3")

# g1/g2 + g3

```

```{r cyclic-cluster, warnings = FALSE, message = FALSE, fig.cap = "Distribution of energy demand across day-of-week and hour-of-day for three groups clustered through the proposed methodology. For group-1, the morning peak looks very sharp, for group-2 the morning and evening peaks are rounded and spread over few hours, for group-3 peaks are less rounded. They differ in their relationship between 75th and 90th percentile.", eval = FALSE}
cg1 <- sm_cust_samp %>% 
  filter(customer_id %in% group1$customer_id) %>% 
  prob_plot("day_week", "hour_day", response = "general_supply_kwh",
            plot_type = "quantile", symmetric = TRUE, quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
  scale_color_brewer(palette = "RdPu") + ggtitle("group 1") +
  scale_x_discrete(breaks = seq(0, 23, 5))

cg2 <- sm_cust_samp %>% 
  filter(customer_id %in% group2$customer_id) %>% 
  prob_plot("day_week", "hour_day", response = "general_supply_kwh",
            plot_type = "quantile", symmetric = TRUE, quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
  scale_color_brewer(palette = "RdPu") + ggtitle("group 2") +
  scale_x_discrete(breaks = seq(0, 23, 5))

cg3 <- sm_cust_samp %>% 
  filter(customer_id %in% group3$customer_id) %>% 
  prob_plot("day_week", "hour_day", response = "general_supply_kwh",
            plot_type = "quantile", symmetric = TRUE, quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
  scale_color_brewer(palette = "RdPu") + ggtitle("group 3") +
  scale_x_discrete(breaks = seq(0, 23, 5))

cg1 
cg2
cg3

```

###  Distinction, repeatability, and robustness metrics

We will also determine if any identified clusters or patterns are indeed statistically meaningful in the sense that they actually exist and are not a random allocation. Hence, the robustness of this methodology is tested through simulations. This section will contain the data structure and detailed methodology to be employed for the cluster analysis. The cluster validation indexes like average silhouette width (ASW) is to be employed here to check how homogeneous these clusters are. At this stage, we need to define the aim of clustering as there could be various aims of clustering like between-cluster separation, within cluster homogeneity: low distances, within-cluster homogeneous distributional shape, good representation of data by centroids, little loss of information, high density without cluster gaps, uniform cluster sizes, stability and others. Finally, how distinct they are and how can we summarize the main features of the cluster would be discussed here.

## Cluster validation

**Internal cluster validation** uses the internal information of the clustering process to evaluate the goodness of a clustering structure without reference to external information. It can be also used for estimating the number of clusters and the appropriate clustering algorithm without any external data.

**External cluster validation** consists in comparing the results of a cluster analysis to an externally known result, such as externally provided class labels. It measures the extent to which cluster labels match externally supplied class labels. Since we know the “true” cluster number in advance, this approach is mainly used for selecting the right clustering algorithm for a specific data set.

**Relative cluster validation** evaluates the clustering structure by varying different parameter values for the same algorithm (e.g.,: varying the number of clusters k). It’s generally used for determining the optimal number of clusters.

Interval validation includes
_Compactness or cluster cohesion:_ Measures how close are the objects within the same cluster. A lower within-cluster variation is an indicator of a good compactness (i.e., a good clustering). The different indices for evaluating the compactness of clusters are base on distance measures such as the cluster-wise within average/median distances between observations.

_Separation:_ Measures how well-separated a cluster is from other clusters. The indices used as separation measures include:
distances between cluster centers
the pairwise minimum distances between objects in different clusters

_Connectivity:_ corresponds to what extent items are placed in the same cluster as their nearest neighbors in the data space. The connectivity has a value between 0 and infinity and should be minimized.

Generally most of the indices used for internal clustering validation combine compactness and separation measures.

## Combining findings with external data

The robustness of this clustering method is provided through practical explanation of the formed clusters, visualizing how they relate to any weather, socio-economic or geographical conditions. 
<!-- Furthermore, we ensure that a cross-validation algorithm is employed leading to robust clusters.  -->
<!-- We will present how matching other data like geography, census and weather would help us in validating our clustering approach. -->

<!-- ### Geography -->

<!-- We will present how mapping the customers to their respective LGA's and reflect on the clusters obtained in the earlier section. The data is at SA2 and LGA levels. However, some of the LGAs in NSW changed between 2011 and 2016 and hence there would not be a one-to-one correspondence between the LGAs. -->

<!-- ### Census -->

<!-- We will present how mapping the customers to their respective income and population level reflect on the clusters obtained in the earlier section. The [ABS TableBuilder](https://www.abs.gov.au/websitedbs/censushome.nsf/home/tablebuilder) has census data from 2011 and 2016.  -->

<!-- ### Weather -->

<!-- Weather, notably temperature (and humidity) can be the main driver(s) for energy usage.  In NSW many households have electric heaters so their use can impact winter energy use and air-conditioners can impact summer energy use. Relevant weather data could be obtained from the [Bureau of Meteorology](www.bom.gov.au). Some weather stations have 30 minute (sometimes even smaller interval) weather data. Potentially, there could be lag effects of weather on energy usage which should be considered. -->

## Conclusion

This section will cover some drawback of this clustering method and potential extensions of this work.

# R packages

The implementation for our framework for the first paper is available in the R package [`gravitas`](https://github.com/Sayani07/gravitas) and that for the second paper is made available in the package [`hakear`](https://github.com/Sayani07/hakear) for ease of use in other applications.

# Discussion

The final chapter of my thesis will consist of the concluding remarks of the whole thesis and possible future extensions.

<!-- <Earlier write up> -->
<!-- ## Introduction -->

<!-- <!-- Description of data available --> -->
<!-- The Smart Grid, Smart City (SGSC) project, which rolled out Australia's first commercial-scale smart grid was implemented across eight local government areas in New South Wales (NSW). Data from more than 13,000 household electricity smart meters is obtained as part of that project. It provides half-hourly energy usage and demographic data for Australia, as well as detailed information on appliance use, climate, retail and distributor product offers, and other related factors. The trials were based in Newcastle, New South Wales, but also covered areas in Sydney CBD, Newington, Ku-Ring-Gai, and the rural township of Scone. The load time series is asynchronous as it is observed for these households for unequal time lengths and consists of missing observations. -->

<!-- The massive amount of data generated in such projects could be overwhelming for analysis. Electricity utilities can utilize the consumption patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers. Beyea (2010) has pointed out, there has been little discussion or exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data. Thus, there is a scope to investigate and analyze these data in various ways for a greater understanding of consumption patterns and how they correlate with other economic, physical or geographical factors. In this work, we are interested to see how we can utilize this dataset to group different customers with similar periodic behavior. Towards this goal, this chapter aims to: (a) describe the contents of the data set in SGSC database that we can utilize, and (b) propose a clustering algorithm to group customers with similar periodic behaviors. The distance metric introduced in Chapter 2 will be the inputs for this cluster analysis. One of the advantages of using our  -->
<!-- approach is that the technique is based on probability distributions instead of raw data. Many clustering approaches are limited by the type of noisy, patchy, and unequal time-series common in residential data sets. Since the distance measure considered is based on differences in probability distribution of time series, it is likely to be less sensitive to missing or noisy data. -->

<!-- The following contributions are made through the following chapter: -->

<!--  * Present a cluster analysis of SGSC dataset to group households with similar periodic behavior -->
<!--  * Cluster validation by relating to external data -->

<!-- ## Introduction -->

<!-- In Australia, the federal government and fourteen partners, electricity distribution and transmission companies, technology companies, universities and CSIRO, invested AUD 490 million in the Smart Grid Smart City (SGSC) field trial. One of the largest commercial-scale smart grid technology assessment projects worldwide, SGSC is implemented across eight local government areas in New South Wales (NSW), covering 30,000 dwellings since 2009. -->
<!-- The SGSC database is publicly available through the Information Clearing House (ICH) (Ausgrid 2014). It is a massive database covering most physical aspects of the smart grid, as well as, aspects of users’ demographics. -->

<!-- Customer Sale and Half-Hour Power Consumption -->
<!-- and Generation Datasets -->

<!-- This is a big portion of the data from about 5900 households who accepted a -->
<!-- product out of about 24,000 offers made to the target population. Spatial resolution -->
<!-- down to anonymous individual dwellings [Customer Sales: Households: -->
<!-- Customer_Key] is also given based on {[Postcode], [Local Government Area], or -->
<!-- [Suburb Name]}. In order to make these datasets easily understandable, here we -->
<!-- present a new narrative approach to explain the available data fields and their -->
<!-- interdependencies in a brief yet logical and memorisable way as follows: -->

<!-- The [Half-Hour Consumption and Generation] data provide kWh readings of -->
<!-- the actual flow of energy. This includes the electricity supply imported from the -->
<!-- grid [General Supply] which is the main source of energy under any of the applied -->
<!-- tariffs, import during off-peak or for a control load [Off Peak], gross or net values of -->

<!-- domestic (often solar) electricity that is exported back to the grid [Generation]. The -->
<!-- [Peak Events] data field gives the details of every event [Event Key] including its -->
<!-- [Type], [Date], [Start Time], and [End Time]. Accordingly, the entire sequence of -->
<!-- user demand responses to any of the cost-reflective products is available in the -->
<!-- dataset to evaluate the project objectives as previously discussed. -->

<!-- Electricity smart meter technology is increasingly being deployed in residential and commercial buildings. For example, in the state of Victoria, Australia, it is a state government policy that all households have a smart meter fitted (Victoria State Government, 2015), which has resulted in 97% of the approximately 2.1 million households in Victoria having a smart meter installed in their home. These smart meters collect energy usage information at half-hourly intervals, resulting in over 35 billion half-hourly observations per year across all households in the state. Clustering electricity consumption patterns can enable electricity utilities to develop targeted tariffs for individual groups alleviating the problem of volatility in production by capitalizing on the flexibility of consumers. Clustering households using only smart meter consumption data should provide value in a societal setting by combining the findings and external data like weather conditions, socio-economic or other demographic factors of those households. -->

<!-- <!-- Description of data available --> -->
<!-- Data from more than 13,000 household electricity smart meters is obtained as part of the Smart Grid Smart City (SGSC) project (2010-2014). It provides one of the few linked sets of customer time of use (half-hour increments) and demographic data for Australia, as well as detailed information on appliance use, climate, retail and distributor product offers, and other related factors. The trials were based in Newcastle, New South Wales, but also covered areas in Sydney CBD, Newington, Ku-Ring-Gai, and the rural township of Scone. This data set includes electricity use interval readings for each customer for every 30 minutes and demographic data for Australia, as well as detailed information on appliance use, climate, peak events; peak events response and retail and distributor product offers. The load time series is asynchronous as it is observed for these households for unequal time lengths and consists of missing observations. -->

<!-- <!-- Background literature and shortcomings --> -->
<!-- The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics. The paper evaluates approximately 2100 unique peer-reviewed papers and presents three main findings related to clustering methods, data and cluster validation. None of the 34 selected papers which focus on clustering consumption are based on Australian smart meter data. The clustering is frequently applied directly to the raw data without investigating time series features like auto correlation and periodicity. The application of smart meter data to cluster electricity consumption using K-Means is ubiquitous. But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters. Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results. @Tureczek2018-ha has shown that the clever transformation of data before K-Means clustering can improve performance and enable K-Means. However, it does not explain the cluster composition by combining it with external data. We propose the following for the study of the Australian smart meter data. -->

<!-- Analysis for more than 13,000 household -->
<!-- electricity smart meters are included in this chapter. The following contributions are made through the following chapter: -->

<!-- - Presenting a cluster analysis of Australian residential electricity consumption data. -->

<!-- - Transformation and extraction of input data features enabling K-Means to account for other time series features in the clustering. This can easily be extended to include other data structures. -->

<!-- - Cluster validation by relating to external data -->

<!-- ## Preliminary exploratoration -->

<!-- This section familiarizes the 13,000 SGSC households through visualization and provide a detailed layout of data structures (like missing observations) and also the external data that needs to be utilized for validating the clustering process. -->

<!-- ## Common load clustering techniques of smart meter data -->

<!-- This section will discuss some literature on clustering techniques that were applied to smart meter customers in the past. -->

<!-- ## The proposed load-clustering technique -->

<!-- We employ a cluster analysis technique to obtain pockets of households showing similar periodic behavior by grouping probability distributions across a cyclic granularity. This clustering algorithm is adopted to remove or appropriately adjust for auto correlation and unequal length in the data. The method could be further extended by clustering probability distributions conditional on one or more cyclic granularities. Further, we elaborate on the effect of missing data and how it is handled in the clustering process. -->

<!-- ### Methodology, results, and discussions -->

<!-- This will contain the data structure and detailed methodology to be employed for the cluster analysis. -->

<!-- ### Distinction, repeatability, and robustness metrics -->

<!-- This section will cover the homogeneity and heterogeneity of the clusters, how distinct they are and summarize the main features of the cluster. -->

<!-- <!-- ## Comparison with traditional load clustering methods --> -->

<!-- <!-- This section will cover the homogeneity and heterogeneity of the clusters, how distinct they are and summarize the main features of the cluster. --> -->

<!-- ## Combining findings with external data -->

<!-- We will also determine if any identified clusters or patterns are indeed statistically meaningful in the sense that they actually exist and are not a random allocation. We aim to do this by providing a practical explanation of the formed clusters, visualizing how they relate to any weather, socio-economic or geographical conditions. Furthermore, we ensure that a cross-validation algorithm is employed leading to robust clusters.  -->
<!-- We will present how matching other data sources might help us in validating our clustering approach. -->

<!-- ### Geography -->

<!-- We will present how mapping the customers to their respective LGA's and reflect on the clusters obtained in the earlier section. -->

<!-- ### Census -->

<!-- We will present how mapping the customers to their respective income and population level reflect on the clusters obtained in the earlier section. -->

<!-- ## Conclusion -->

<!-- This section will cover some drawback of this clustering method and potential extensions of this work. -->

<!-- There are many cluster analysis methods that can produce quite different clusterings on the same dataset. Cluster validation is about the evaluation of the quality of a clustering; “relative cluster validation” is about using such criteria to compare clusterings. This can be used to select one of a set of clusterings from different methods, or from the same method ran with different parameters such as different numbers of clusters. There are many cluster validation indexes in the literature. Most of them attempt to measure the overall quality of a clustering by a single number, but this can be inappropriate. There are various different characteristics of a clustering that can be relevant in practice, depending on the aim of clustering, such as low within-cluster distances and high between-cluster separation. In this paper, a number of validation criteria will be introduced that refer to different desirable characteristics of a clustering, and that characterise a clustering in a multidimensional way. In specific applications the user may be interested in some of these criteria rather than others. A focus of the paper is on methodology to standardise the different characteristics so that users can aggregate them in a suitable way specifying weights for the various criteria that are relevant in the clustering application at hand. -->

### 2021-07-07 script/smart_meter.R

@@ -1,0 +1,203 @@
+## ---- load-lib
+library(tidyverse)
+library(lubridate)
+library(tsibble)
+library(fable)
+library(patchwork)
+library(here)
+theme_set(theme_bw())
+library(forcats)
+library(readr)
+
+##----load-data
+load("data/ALL_DATA.Rdata")
+
+
+##----format-data
+
+smart_meter_data <- EUDMData %>% 
+  dplyr::rename_all(tolower) %>% 
+  dplyr::mutate(reading_datetime = lubridate::ymd_hms(reading_datetime)) %>% 
+  dplyr::arrange(customer_id, reading_datetime) %>% 
+  dplyr::group_by(customer_id) %>% 
+  dplyr::mutate(reading_datetime = case_when(
+    duplicated(reading_datetime) ~ reading_datetime + lubridate::hours(1), 
+    TRUE ~ reading_datetime
+  ))
+  
+elec_ts <- smart_meter_data %>% 
+  build_tsibble(
+    key = customer_id, index = reading_datetime,
+    validate = FALSE, ordered = TRUE
+  )
+
+readr::write_rds(elec_ts, "data/elec_ts.rds")
+
+## ---- elec-gaps
+elec_ts <- read_rds("data/elec_ts.rds")
+gap_df <- has_gaps(elec_ts)
+
+# sum(gap_df$.gaps) / NROW(gap_df)
+
+## ---- count-gaps
+count_na_df <- elec_ts %>% 
+  count_gaps()
+
+
+lumped_na_df <- count_na_df %>% 
+  mutate(
+    customer_id = as.factor(customer_id) %>% 
+      fct_lump(264) %>%
+      fct_reorder(.n, sum)
+  ) 
+
+p_264 <- lumped_na_df %>%
+  filter(customer_id != "Other") %>%
+  ggplot(aes(x = customer_id)) +
+  geom_linerange(aes(ymin = .from, ymax = .to)) +
+  geom_point(aes(y = .from), size = 0.6, shape = 4) +
+  geom_point(aes(y = .to), size = 0.6, shape = 4) +
+  coord_flip() +
+  xlab("Top customers with more than
+       10% observations missing") +
+  ylab("") +  theme(axis.text = element_blank(),
+    axis.ticks = element_blank(),
+    legend.position = "none",
+    plot.margin = unit(c(0, 0, -1, -1), "line")) + scale_y_datetime(" ",
[…]

### 2021-07-08 realdata-cust10-noiter.Rmd
all updated files in supplementary

---
title: "Clustering real data with ten customers and choosing optimal number of clusters"
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
    toc: false    
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
read_chunk("R/simdata01.R")
```

```{r load-lib}
```

# Clutering approach

1. _Compute quantiles of distributions across each hour of day_
2. _Compute JS distance between households for each hour of day_
3. _Total distance between households computed as sum of JS distances for all hours_
4. _Cluster using this distance with hierarchical clustering algorithm (method "complete")_

# Plots of raw data

Five households are considered for the following analysis from the SGSC data set.
The data sets contain three columns `customer_id`, `reading_datetime` and `general_supply_kwh`. They consist of half-hourly data from 2012-2014. Figure \ref{fig:raw-all} shows how the raw time plots for these five households look like. Since all the dataset looks squeezed on this linear time scale, Figure \ref{fig:raw-sub} shows how the raw plot looks for 2 months (Sept 2013 and Oct 2013).

```{r}
sm <- gravitas::smart_meter10 
```

```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for 5 households (facets). The data is squeezed stopping us from seeing any behavioral patterns."}
library(lubridate)

sm %>% 
  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
  geom_line() +
  facet_wrap(~customer_id, ncol = 1, 
             strip.position = "right") +
  theme_bw() +
  xlab("time") +
  ylab("demand (in Kwh)")
```

```{r raw-sub, fig.cap="The raw plots for ten customers shown for Sept 2013-Oct 2013. The data is zoomed in and the y-scales made free to emphasize weekly, hourly or any behavior/patterns that they might have. There is some daily and weekly pattern in all households except the 4th one, that has spikes which seem to occur at irregular intervals."}
sm %>% 
  filter(year(reading_datetime) == 2013 &
           month(reading_datetime) %in% c(9, 10)) %>% 
  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
  geom_line() +
  facet_wrap(~customer_id, ncol = 1, scales = "free_y", 
             strip.position = "right") +
  theme_bw()+
  ylab("demand (in Kwh)") +
  theme(panel.grid.major.x = element_blank()) +
  scale_x_datetime("Date", date_labels = "%b %d",
                   breaks = "1 week",
                   date_minor_breaks = "1 day")  + 
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_line(colour = "#D3D3D3"))

```

```{r hod}
sm_hod <- gravitas::create_gran(sm, "hour_day")
```

```{r hod-list}
sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

lookup_tbl <- sm_hod_list %>% 
  mutate(
         customer_serial_id = row_number()) %>% 
  select(customer_id, customer_serial_id)

```

```{r hod-quantile}
quantile_prob_val = seq(0.1, 0.9, 0.1)

ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)]
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list[-1] %>% extract(x, y) %>% unlist()
   quantile(cell, prob = quantile_prob_val)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

```

```{r hod-dist}
sm_dist_data <- sm_hod_quantiles %>% 
  mutate(customer_serial_id = as.numeric(customer_serial_id)) %>% 
  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
  select(-quantile_prob) %>% 
  group_by(customer_serial_id, categories_serial_id) %>%
  nest() %>% 
  pivot_wider(names_from = categories_serial_id, values_from = data)
```

```{r JS}

```

```{r hod-dist-mat}
#ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
nrow_data <- nrow(sm_dist_data)
ncol_data <-  ncol(sm_dist_data[-1])

dist_data <- map(seq_len(nrow_data), function(x){ # first data
  map(seq_len(nrow_data), function(y){ # 2nd data
    map(seq_len(ncol_data), function(z){ # number of combinations nx*nfacet
      JS(
        prob = quantile_prob_val,
        unlist(sm_dist_data[-1] %>% extract(x, z)),
        unlist(sm_dist_data[-1] %>% extract(y, z))
      ) %>% as_tibble()
    })%>% bind_rows(.id = "combinations")
  })%>% bind_rows(.id = "data_type1")
}) %>% bind_rows(.id = "data_type2")

dist_mat <- dist_data %>%
  mutate(data_type1 = as.numeric(data_type1), data_type2 = as.numeric(data_type2)) %>% 
  group_by(data_type1, data_type2) %>%
  summarise(dist = sum(value)) %>%
  pivot_wider(names_from = data_type2,
              values_from = dist)

dist_mat_format <- dist_mat %>%
  ungroup() %>% 
  select(-data_type1)

rownames(dist_mat_format) <- lookup_tbl$customer_id

```

# Dendogram

The following figures show dendograms when we are using optimal number of clusters (k) from k = fpc::nselectboot(), k = 2, 3, 4 and 5. We observe that the $4^{th}$ household tend to get split as different clusters as increase the number of clusters.

```{r hc}
d = stats::as.dist(dist_mat_format)
hc = stats::hclust(d,method="complete")
plot(hc)
```

# Multi-dimensional scaling

```{r mds-new}
set.seed(12345)
mds <- d %>%
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")
rownames(mds) <- lookup_tbl$customer_id

library(fpc)
koptimal = fpc::nselectboot(as.dist(dist_mat_format),
                        B = 50,
                        method = "complete",
                        clustermethod = disthclustCBI, 
                        classification = "averagedist",
                        krange = 2:12)

koptimal = koptimal$kopt

groups<-cutree(hc, k=koptimal)

all_data_cluster <- cbind(dist_mat_format, groups) %>%
  cbind(mds)%>%
  mutate(groups = as.factor(groups)) %>% as_tibble()

mds_plot_10 <- ggplot(all_data_cluster,
       aes(x = Dim.1,
           y = Dim.2,
       color = groups,
       label = rownames(mds))) +
  geom_point(size = 1) +
  geom_text(check_overlap = TRUE)  +
  theme_classic()+
  theme(legend.position = "bottom") +
  scale_color_brewer(palette = "Dark2")

mds_plot_10
```
Optimal number of clusters as defined by the _fpc::nselectboot_ is `r koptimal`.

_k = 2_

```{r mds4}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 2, border = 2:5)
```

_k = 3_
```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 3, border = 2:5)
```

_k = 4_

```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 4, border = 2:5)
```

_k = 5_

```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 5, border = 2:6)
```

_k = 6_

```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 6, border = 2:7)
```

_k = 7_

```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 7, border = 2:8)
```

_k = 8_

```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 8, border = 2:9)
```

_k = 9_

```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 9, border = 2:10)
```

_k = 10_

```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 10, border = 2:11)
```

# Plots of distribution of data across clusters

The distribution of demand across different hours of the day observed for these five households through heatmaps (Figure \ref{fig:hod-deciles}) and quantile plots (Figure \ref{fig:hod-heatmap}). The following characterization is done to validate results of the clustering approach later on.

```{r}
sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

```

```{r}
quantile_prob_val = seq(0.1, 0.9, 0.1)

ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)]
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list[-1] %>% extract(x, y) %>% unlist()
   quantile(cell, prob = quantile_prob_val)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

```

```{r hod-heatmap, fig.cap="Heatmaps show deciles on the y-axis with customers on the x-axis with colors filled by the value of deciles and faceted by hours of the day. It again seems like all the households have gradual change in colors (almost for all hours) as we move to higher deciles except for the 4th household. "}
categories_ref <- names(sm_hod_list)[-1] %>% as_tibble() %>% 
  mutate(categories_serial_id = as.character(row_number()))

data_heatmap <- sm_hod_quantiles %>% 
  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
  left_join(categories_ref) %>% 
  rename("category" = "value") %>% 
  select(-categories_serial_id) %>% 
  mutate(category = as.numeric(category))

data_heatmap%>% 
ggplot(aes(x= customer_serial_id, y = as.factor(quantile_prob))) +
  geom_tile(aes(fill = quantile_val)) +
  scale_fill_distiller(palette = "YlGnBu")+ 
  facet_wrap(~category) +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("deciles") +
  xlab("customers")
```

```{r hod-deciles, fig.cap = "The deciles plots for 5 households are shown for entire observation period across hours of the day. The y-scales are made free to allow us to see daily patterns for each households. The deciles for all households tend to show an increase in the morning and evening hours, although the hours differ. Except for the 90th decile, the deciles for the 4th household look pretty flat."}
data_heatmap %>% 
  ggplot() + geom_line(aes(x=category, 
                           y = quantile_val, 
                           color = quantile_prob)) + facet_wrap(~customer_serial_id, ncol = 1, scales="free_y",
                                                                strip.position = "right") + 
  scale_color_brewer(palette = "Blues") + 
  #scale_y_log10() +
  theme_bw() +
  xlab("hours of the day") +
  ylab("demand (in Kwh)")
```

### 2021-07-08 supplementary/cluster4designs.Rmd
all updated files in supplementary

---
title: "Clustering four designs"
output:
  bookdown::pdf_book:
    MonashEBSTemplates::workingpaper:
    #base_format: rticles::asa_article
    fig_height: 8
    fig_width: 12
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
read_chunk("R/simdata01.R")
```

# How the time series plot of the designs look like for different sample sizes?

```{r sim-val, echo = TRUE}
set.seed(9999)
nx_val = 2 # number of x-axis levels
nfacet_val = 3 # number of facet levels
w1_val = 3 # increment in mean
w2_val = 0 # increment in sd
mean_val = 0 # mean of normal distribution of starting combination
sd_val = 1 # sd of normal distribution of starting combination
quantile_prob_val = seq(0.1, 0.9, 0.1)
```

```{r load-lib}
```

```{r designs}
```

## nobs per combination is 4

```{r, echo = TRUE}
ntimes_val = 4
```

```{r data}
```

```{r category-plots}
```

```{r}
change_index <- function(data){

index_new <- map(seq_len(ntimes_val), function(i){
  map((seq_len(nx_val*nfacet_val)), function(j)
    {
value = i + (j-1)*ntimes_val
})
  }) %>% unlist()

data_new = data %>%
  ungroup() %>%
  mutate(index_old = row_number(),
         index_new = index_new)

y = data_new[match(index_new, data_new$index_old),]

y <- y %>%
  mutate(time = row_number())

return(y)
}

endbreaks<- nrow(sim_panel_null)

p1 <- change_index(sim_panel_null) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p2 <- change_index(sim_panel_varf) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p3 <- change_index(sim_panel_varx) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue")+
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p4 <- change_index(sim_panel_varall) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
  panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

t1 <- sim_panel_varall %>% ungroup() %>%  select (-c(1, 2))

# %>% 
#   kable(format = "latex", caption = "Simulated data for categories format")

t2 <- change_index(sim_panel_varall) %>% select (-c(1, 2))

# %>% 
#   kable(format = "latex",
#         caption = "Simulated data for time series format")

kable(list(t1, t2), format = "latex", caption = "Simulated data across combination of categories (left) and manipulated time series format (right)") %>% 
  row_spec(1:24,  background =  "#d3d3d3")

# kable(t1) %>%
#   kable_styling(full_width = FALSE, position = "float_left")
# kable(t2) %>%
#   kable_styling(full_width = FALSE, position = "left")
```

```{r plot4}
(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)
```

## nobs per combination is 10

```{r nobs10, echo = TRUE}
ntimes_val = 10 # nobs per combination 
```

```{r data}
```

```{r category-plots}
```

```{r line-graph}
```

```{r plot10}
(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)
```

## nobs per combination is 50

```{r nobs50, echo = TRUE}
ntimes_val = 50 # nobs per combination 
```

```{r data}
```

```{r category-plots}
```

```{r line-graph}
```

```{r plot50}
(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)

```

# Clustering on simulated datasets

```{r, echo=TRUE}
sample_seed <-  seq(10, 100, 10)
```

_DGP: Generate 10 time series from each designs_ 
Time series are simulated from each of these designs with $50$ observations in each group. So we have $50$ observations each for the six combination of categories $(1,1), (1, 2), (2, 1), (2, 2), (3, 1), (3, 2)$. Time series are simulated for ten different seeds (from `sample_seed`) for each design. `data_varall-1` represents a data set from a design $D_{var_{all}}$ (distributions change across both facet and x) with a seed $10$. `data_varall-2` represents a data set from a design $D_{var_{all}}$ with a seed $20$ and so on.
`data_null`, `data_varx` and `data_varf `corresponds to designs $D_{null}$, $D_{var_{x}}$ and $D_{var_{f}}$ designs respectively.

_Compute quantiles of conditional distributions_
Conditional quantiles are obtained for each combination of categories.

_JS Pairwise distances between datasets_
Distance between the data sets is computed as the sum of JS distances across different categories.

_Hierarchical clustering with 4 clusters_
Hierarchical clustering is performed using k = 4 and dendogram observed

_Clusters obtained visualized using MDS_
Each cluster represents data sets from a separate design

```{r}
change_index_data <- function(ntimes_val = NULL,
                         nx_val = NULL,
                         nfacet_val = NULL,
                         sim_function = sim_varx_normal){

  data <- sim_panel(
    nx = nx_val, nfacet =  nfacet_val,
    ntimes = ntimes_val,
    # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
    sim_dist = sim_function(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
  ) %>% unnest(data)

  index_new <- map(seq_len(ntimes_val), function(i){
    map((seq_len(nx_val*nfacet_val)), function(j)
    {
      value = i + (j-1)*ntimes_val
    })
  }) %>% unlist()

  data_new = data %>%
    ungroup() %>%
    mutate(index_old = row_number(),
           index_new = index_new)

  y = data_new[match(index_new, data_new$index_old),]

  y <- y %>%
    mutate(time = row_number()) %>%
    select(-c(index_old, index_new))

  return(y)
}
```

```{r data-make}
data_null <- map(seq_len(sample_seed), function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_null_normal) %>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_null")

data_varf <- map(seq_len(sample_seed), function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varf_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varf")

data_varx <- map(seq_len(sample_seed), function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varx_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varx")

data_varall <- map(seq_len(sample_seed), function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varall_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varall")

data_q <- bind_rows(data_null,
                      data_varf,
                      data_varx,
                      data_varall) %>%
  mutate(seed_id = sprintf("%02d", 
                              as.numeric(seed_id))) %>% 
  mutate(unique_data = paste(data_type, seed_id, sep = "-"))

data_q$data_type <- factor(data_q$data_type,
                                    levels = c("data_null" , "data_varf",
                                               "data_varx","data_varall"))

```

```{r JS}

```

```{r JS-mydata}
data_q_wide <- data_q %>%
  select(id_facet, id_x, unique_data, sim_data_quantile) %>%
  pivot_wider(names_from = unique_data,
              values_from = sim_data_quantile) %>%
  select(-c(1, 2))

ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
ldata <- nrow(ndata)
lcomb <-  nx_val*nfacet_val

dist_data <- map(1:ldata, function(x){ # first data
  map(1:ldata, function(y){ # 2nd data
    map(1:lcomb, function(z){ # number of combinations nx*nfacet
      JS(
        prob = quantile_prob_val,
        unlist(data_q_wide[z,x]),
        unlist(data_q_wide[z,y])
      ) %>% as_tibble()
    })%>% bind_rows(.id = "combinations")
  })%>% bind_rows(.id = "data_type1")
}) %>% bind_rows(.id = "data_type2")

dist_mat <- dist_data %>%
  group_by(data_type1, data_type2) %>%
  summarise(dist = sum(value)) %>%
  pivot_wider(names_from = data_type2,
              values_from = dist) %>%
  mutate(data_type1 = as.numeric(data_type1)) %>%
  left_join(ndata, by = c("data_type1" = "index"))

dist_mat_format <- dist_mat %>%
  ungroup() %>%
  select(-data_type1, -unique_data)

rownames(dist_mat_format) <- dist_mat$unique_data
```

```{r hier-clust}
d = stats::as.dist(dist_mat_format)
hc = stats::hclust(d,method="complete")
plot(hc)
```

```{r}
mds <- d %>%
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")
rownames(mds) <- dist_mat$unique_data

groups<-cutree(hc, k=4)

all_data_cluster <- cbind(dist_mat_format, groups) %>%
  cbind(mds)%>%
  mutate(groups = as.factor(groups)) %>% as_tibble()

mds_plot_10 <- ggplot(all_data_cluster,
       aes(x = Dim.1,
           y = Dim.2,
       color = groups,
       label = rownames(mds))) +
  geom_point(size = 0.5) +
  geom_text(check_overlap = TRUE)  +
  theme_classic()+
  theme(legend.position = "bottom") +
  scale_color_brewer(palette = "Dark2")

mds_plot_10
```

```{r}
heatmap_raw <- data_q %>% 
  #select(unique_data, sim_data_quantile, id_facet, id_x) %>% 
  mutate(category = paste0("facet= ", id_facet, ",x= ", id_x)) %>% 
  unnest(sim_data_quantile) %>% 
  group_by(unique_data, category) %>% 
  mutate(D = row_number()) %>% 
  arrange(D, category) %>% 
  left_join(dist_mat, by = c("unique_data")) %>% 
  select(c(2:9))

data2 <- cbind(dist_mat_format, groups)

heatmap_final <- bind_cols(serial_data = rownames(data2) %>%
                             as_tibble(),
          group = data2$groups) %>%
  left_join(heatmap_raw, by = c("value" = "unique_data")) %>% select(-id_facet, - id_x, - data_type1) %>% 
  arrange(D, category, group)

#heatmap_final$data_type <- factor(heatmap_final$data_type,
#                                    levels = c("data_null" , "data_varf",
#                                               "data_varx","data_varall"))

#heatmap_final$category <- reorder(heatmap_final$category, heatmap_final$data_type)

ggplot(heatmap_final,
       aes(x= reorder(value, data_type), y = as.factor(D))) +
  geom_tile(aes(fill = sim_data_quantile)) +
  scale_fill_distiller(palette = "YlGnBu")+ 
  facet_wrap(~category, ncol = 1) +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("Deciles") +
  xlab("data sets")

```

## Cluster validation

```{r}
library(fpc)
library(cluster)
seq_data = tibble(groups, name = names(groups)) %>% 
  mutate(seq_group  = row_number()) %>% 
  left_join(ndata, by = c("name" = "unique_data")) %>% 
  left_join(data_q %>% distinct(unique_data, data_type),
            by = c("name" = "unique_data"))

seq_data$data_type <- factor(seq_data$data_type,
                                    levels = c("data_null" , "data_varf",
                                               "data_varx","data_varall"))

v <- cluster.stats(d, as.numeric(seq_data$groups), as.numeric(seq_data$data_type), silhouette = TRUE)

v
```

The total average (mean of all individual silhouette widths) is `r round(v$avg.silwidth, 3)` and corrected Rand index is `r round(v$corrected.rand, 3)`

### 2021-07-08 supplementary/realdata-cust10-noiter.Rmd
all updated files in supplementary

--- 2021-07-08 realdata-cust10-noiter.Rmd
@@
 ---
-title: "Clustering real data with ten customers and choosing optimal number of clusters"
+title: "Clustering real data with ten customers without iterations"
 output:
@@
 
-```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for 5 households (facets).
+```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for 10 households (facets).
 The data is squeezed stopping us from seeing any behavioral patterns."}
@@
 The data is zoomed in and the y-scales made free to emphasize weekly, hourly or any behavior/patterns that they might have.
-There is some daily and weekly pattern in all households except the 4th one, that has spikes which seem to occur at irregular intervals."}
+There is some daily and weekly pattern in most households."}
 sm %>% 
@@
 ```{r hod}
-sm_hod <- gravitas::create_gran(sm, "hour_day")
+sm_hod <- gravitas::create_gran(sm, "hour_day") %>% 
+  mutate(customer_id = as.integer(customer_id))
+```
+
+```{r}
+sm_hod_list <- sm_hod %>% 
+  as_tibble() %>% 
+  select(customer_id, hour_day, general_supply_kwh) %>% 
+  pivot_wider(names_from = hour_day,
+              values_from = general_supply_kwh)
+
+```
+
+```{r}
+quantile_prob_val = seq(0.1, 0.9, 0.1)
+
+ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)]
+nrow_sm <- seq_len(nrow(sm_hod_list))
+
+sm_hod_quantiles <- map(nrow_sm, function(x){
+  map(ncol_sm, function(y){
+   cell <- sm_hod_list[-1] %>% extract(x, y) %>% unlist()
+   quantile(cell, prob = quantile_prob_val)
+})  %>% bind_rows(.id = "categories_serial_id")
+}) %>% bind_rows(.id = "customer_serial_id") %>% 
+  mutate(customer_serial_id = as.integer(customer_serial_id),
+         categories_serial_id = as.integer(categories_serial_id))
+
+categories_ref <- names(sm_hod_list)[-1] %>% as_tibble() %>% 
+  mutate(categories_serial_id = row_number())
+
+data_heatmap <- sm_hod_quantiles %>% 
+  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
+  left_join(categories_ref) %>% 
+  rename("category" = "value") %>% 
+  select(-categories_serial_id) %>% 
+  mutate(category = as.numeric(category))
+
+```
+
+```{r hod-deciles, fig.cap = "The deciles plots for 10 households are shown for entire observation period across hours of the day.
+The y-scales are made free to allow us to see daily patterns for each households.
+The deciles for all households tend to show an increase in the morning and evening hours, although the hours differ.
+Except for the 90th decile, the deciles for the 4th household look pretty flat."}
+data_heatmap %>% 
+  ggplot() + geom_line(aes(x=category, 
+                           y = quantile_val, 
+                           color = quantile_prob)) + 
+  facet_wrap(~customer_serial_id, 
+             ncol = 1,
+             scales="free_y") + 
+  scale_color_brewer(palette = "Blues") + 
+  #scale_y_log10() +
+  theme_bw() +
+  xlab("hours of the day") +
+  ylab("demand (in Kwh)") +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
 ```
@@
 
-# Dendogram
+# Clustering results
 
@@
 ```
-
-# Multi-dimensional scaling
 
@@
                         classification = "averagedist",
-                        krange = 2:12)
+                        krange = 2:10)
 
@@
 
-_k = 2_
-
-```{r mds4}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 2, border = 2:5)
-```
-
-_k = 3_
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 3, border = 2:5)
-```
-
-_k = 4_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 4, border = 2:5)
-```
-
-_k = 5_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 5, border = 2:6)
-```
-
-_k = 6_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 6, border = 2:7)
-```
-
 _k = 7_
@@
 
-_k = 8_
+# Cluster characterization
 
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 8, border = 2:9)
-```
-
-_k = 9_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 9, border = 2:10)
-```
-
-_k = 10_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 10, border = 2:11)
-```
-
-# Plots of distribution of data across clusters
-
-The distribution of demand across different hours of the day observed for these five households through heatmaps (Figure \ref{fig:hod-deciles}) and quantile plots (Figure \ref{fig:hod-heatmap}).
-The following characterization is done to validate results of the clustering approach later on.
-
-```{r}
-sm_hod_list <- sm_hod %>% 
-  as_tibble() %>% 
-  select(customer_id, hour_day, general_supply_kwh) %>% 
-  pivot_wider(names_from = hour_day,
-              values_from = general_supply_kwh)
-
-```
-
-```{r}
-quantile_prob_val = seq(0.1, 0.9, 0.1)
-
-ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)]
-nrow_sm <- seq_len(nrow(sm_hod_list))
-
-sm_hod_quantiles <- map(nrow_sm, function(x){
-  map(ncol_sm, function(y){
-   cell <- sm_hod_list[-1] %>% extract(x, y) %>% unlist()
-   quantile(cell, prob = quantile_prob_val)
-})  %>% bind_rows(.id = "categories_serial_id")
-}) %>% bind_rows(.id = "customer_serial_id")
-
-```
+The quartile deviation of different clusters are drawn.
+The shape of daily load curve is different for all of them.
+Across each cluster, customers should show different shape and within each cluster, customers should differ in size.
 
@@
 "}
-categories_ref <- names(sm_hod_list)[-1] %>% as_tibble() %>% 
-  mutate(categories_serial_id = as.character(row_number()))
+quantile_prob_val = c(0.25, 0.5, 0.75)
 
-data_heatmap <- sm_hod_quantiles %>% 
+groups_ref <- groups %>% as_tibble %>% 
+              mutate(customer_serial_id = sm_hod_list$customer_id) %>% 
+             rename ("group" = "value")
+
+sm_hod_list_cat <-  sm_hod_list %>% left_join(groups_ref, by = c("customer_id" = "customer_serial_id")) %>% select(group, everything())
+
+ncol_sm <- seq_len(ncol(sm_hod_list_cat[-c(1, 2)]))
+nrow_sm <- unique(sm_hod_list_cat$group)
+
+sm_hod_quantiles_cat <- map(nrow_sm, function(x){
+  map(ncol_sm, function(y){
+   cell <- sm_hod_list_cat%>% 
+     dplyr::filter(group == x) %>% 
+     select(-c(1, 2)) %>% 
+     extract(y) %>% 
+     unlist()
+   quantile(cell, prob = quantile_prob_val)
+})  %>% bind_rows(.id = "categories_serial_id")
+}) %>% bind_rows(.id = "group_id") %>% 
+  mutate(group_id = as.integer(group_id),
+         categories_serial_id = as.integer(categories_serial_id))
+
+categories_ref_cat <- names(sm_hod_list_cat)[-c(1, 2)] %>% as_tibble() %>% 
+  mutate(categories_serial_id = row_number())
+
+data_heatmap <- sm_hod_quantiles_cat %>% 
   pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
-  left_join(categories_ref) %>% 
+  left_join(categories_ref_cat) %>% 
   rename("category" = "value") %>% 
@@
 
-data_heatmap%>% 
-ggplot(aes(x= customer_serial_id, y = as.factor(quantile_prob))) +
-  geom_tile(aes(fill = quantile_val)) +
-  scale_fill_distiller(palette = "YlGnBu")+ 
-  facet_wrap(~category) +
-  theme(axis.text.x = element_text(angle = 90)) +
-  ylab("deciles") +
-  xlab("customers")
+# data_heatmap%>% 
+# ggplot(aes(x= as.factor(group_id), y = as.factor(quantile_prob))) +
+#   geom_tile(aes(fill = quantile_val)) +
+#   scale_fill_distiller(palette = "YlGnBu")+ 
+#   facet_wrap(~category) +
+#   theme(axis.text.x = element_text(angle = 90)) +
+#   ylab("deciles") +
+#   xlab("customers")
 ```
 
-```{r hod-deciles, fig.cap = "The deciles plots for 5 households are shown for entire observation period across hours of the day.
-The y-scales are made free to allow us to see daily patterns for each households.
-The deciles for all households tend to show an increase in the morning and evening hours, although the hours differ.
-Except for the 90th decile, the deciles for the 4th household look pretty flat."}
+```{r cluster-characterization}
 data_heatmap %>% 
-  ggplot() + geom_line(aes(x=category, 
-                           y = quantile_val, 
-                           color = quantile_prob)) + facet_wrap(~customer_serial_id, ncol = 1, scales="free_y",
-                                                                strip.position = "right") + 
-  scale_color_brewer(palette = "Blues") + 
-  #scale_y_log10() +
+  pivot_wider( names_from = quantile_prob, 
+               values_from = quantile_val) %>% 
+  ggplot(aes(x = category)) +  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`), fill = "lightblue") +
+  geom_line(aes(y = `50%`)) +
+  facet_wrap(~group_id, ncol = 1, scales = "free_y", labeller = "label_both") +
   theme_bw() +
-  xlab("hours of the day") +
-  ylab("demand (in Kwh)")
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
 ```

### 2021-07-08 supplementary/realdata-cust10.Rmd
all updated files in supplementary

--- 2021-07-08 realdata-cust10-noiter.Rmd
@@
 
-```{r hod}
+# Plots of distribution of data across categories of granularities
+
+The distribution of demand across different hours of the day observed for these five households through heatmaps (Figure \ref{fig:hod-deciles}) and quantile plots (Figure \ref{fig:hod-heatmap}).
+The following characterization is done to validate results of the clustering approach later on.
+
+```{r hod-box, fig.cap="The raw plots for 5 households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
+sm %>% 
+  create_gran("hour_day") %>% 
+  ggplot(aes(x = hour_day, y = general_supply_kwh)) + 
+  geom_boxplot() + 
+  facet_wrap(~customer_id, ncol = 1, scales = "free_y",
+             strip.position = "right") 
+```
+
+```{r}
 sm_hod <- gravitas::create_gran(sm, "hour_day")
@@
 
-```{r hod-list}
+```{r}
 sm_hod_list <- sm_hod %>% 
@@
 
-lookup_tbl <- sm_hod_list %>% 
-  mutate(
-         customer_serial_id = row_number()) %>% 
-  select(customer_id, customer_serial_id)
-
 ```
 
-```{r hod-quantile}
+```{r}
 quantile_prob_val = seq(0.1, 0.9, 0.1)
@@
 
-```{r hod-dist}
+```{r hod-heatmap, fig.cap="Heatmaps show deciles on the y-axis with customers on the x-axis with colors filled by the value of deciles and faceted by hours of the day.
+It again seems like all the households have gradual change in colors (almost for all hours) as we move to higher deciles except for the 4th household.
+"}
+categories_ref <- names(sm_hod_list)[-1] %>% as_tibble() %>% 
+  mutate(categories_serial_id = as.character(row_number()))
+
+data_heatmap <- sm_hod_quantiles %>% 
+  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
+  left_join(categories_ref) %>% 
+  rename("category" = "value") %>% 
+  select(-categories_serial_id) %>% 
+  mutate(category = as.numeric(category))
+
+data_heatmap%>% 
+ggplot(aes(x= customer_serial_id, y = as.factor(quantile_prob))) +
+  geom_tile(aes(fill = quantile_val)) +
+  scale_fill_distiller(palette = "YlGnBu")+ 
+  facet_wrap(~category) +
+  theme(axis.text.x = element_text(angle = 90)) +
+  ylab("deciles") +
+  xlab("customers")
+```
+
+```{r hod-deciles, fig.cap = "The deciles plots for 5 households are shown for entire observation period across hours of the day.
+The y-scales are made free to allow us to see daily patterns for each households.
+The deciles for all households tend to show an increase in the morning and evening hours, although the hours differ.
+Except for the 90th decile, the deciles for the 4th household look pretty flat."}
+data_heatmap %>% 
+  ggplot() + geom_line(aes(x=category, 
+                           y = quantile_val, 
+                           color = quantile_prob)) + facet_wrap(~customer_serial_id, ncol = 1, scales="free_y",
+                                                                strip.position = "right") + 
+  scale_color_brewer(palette = "Blues") + 
+  #scale_y_log10() +
+  theme_bw() +
+  xlab("hours of the day") +
+  ylab("demand (in Kwh)")
+```
+
+# Iterations of each customer (adding some random noise for each iteration)
+
+Iterations of each customer are considered by adding random noise ($N(0, \sigma^2)$, where $\sigma^2$ is very small relative to the the variance of the data).
+Figure \ref{fig:raw-iter} shows that raw plots of the simulated iterations to show that their structure is same as the parent data set.
+
+```{r}
+sample_seed <-  seq(100, 500, 100)
+ncust <- sm %>% distinct(customer_id)
+
+all_data_clust = map_dfr(seq_len(nrow(ncust)), function(x){
+  map(seq_len(length(sample_seed)), function(seed){
+     set.seed(sample_seed[seed])
+     data <- sm %>% as_tibble %>% filter(customer_id %in% ncust$customer_id[x])
+     nr = nrow(data)
+
+     data1 <- data %>% 
+     mutate(customer_id = paste(customer_id, seed, sep = "-"),
+             general_supply_kwh = general_supply_kwh + rnorm(1, 0, 0.001))
+     data1
+}) %>% bind_rows(.id = "seed_id")
+}) %>% bind_rows(sm %>% as_tibble() %>% bind_cols(seed_id = "0")) %>% 
+  tsibble::as_tsibble(index = reading_datetime, key  = customer_id)
+```
+
+```{r raw-iter, fig.cap = "Five iterations for each customer are considered by adding random noise $(N(0, \\sigma^2)$, where $\\sigma^2$ is very small relative to the the variance of the data.
+The raw plots for all simulated dataset is shown to make sure they look similar to the structure."}
+all_data_clust %>% 
+  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
+  geom_line() +
+  facet_wrap(~customer_id,
+             strip.position = "right", nrow = 10) +
+  theme_bw() + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  xlab("time") +
+  ylab("demand (in Kwh)")
+```
+
+```{r}
+sm_hod <- gravitas::create_gran(all_data_clust, "hour_day")
+```
+
+```{r}
+sm_hod_list <- sm_hod %>% 
+  as_tibble() %>% 
+  select(customer_id, hour_day, general_supply_kwh) %>% 
+  pivot_wider(names_from = hour_day,
+              values_from = general_supply_kwh)
+
+lookup_tbl <- sm_hod_list %>% 
+  mutate(
+         customer_serial_id = row_number()) %>% 
+  select(customer_id, customer_serial_id)
+
+```
+
+```{r}
+quantile_prob_val = seq(0.1, 0.9, 0.1)
+
+ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)]
+nrow_sm <- seq_len(nrow(sm_hod_list))
+
+sm_hod_quantiles <- map(nrow_sm, function(x){
+  map(ncol_sm, function(y){
+   cell <- sm_hod_list[-1] %>% extract(x, y) %>% unlist()
+   quantile(cell, prob = quantile_prob_val)
+})  %>% bind_rows(.id = "categories_serial_id")
+}) %>% bind_rows(.id = "customer_serial_id")
+
+```
+
+```{r}
 sm_dist_data <- sm_hod_quantiles %>% 
@@
 
-```{r hod-dist-mat}
+```{r}
 #ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
@@
 ```
-
-# Plots of distribution of data across clusters
-
-The distribution of demand across different hours of the day observed for these five households through heatmaps (Figure \ref{fig:hod-deciles}) and quantile plots (Figure \ref{fig:hod-heatmap}).
-The following characterization is done to validate results of the clustering approach later on.
-
-```{r}
-sm_hod_list <- sm_hod %>% 
-  as_tibble() %>% 
-  select(customer_id, hour_day, general_supply_kwh) %>% 
-  pivot_wider(names_from = hour_day,
-              values_from = general_supply_kwh)
-
-```
-
-```{r}
-quantile_prob_val = seq(0.1, 0.9, 0.1)
-
-ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)]
-nrow_sm <- seq_len(nrow(sm_hod_list))
-
-sm_hod_quantiles <- map(nrow_sm, function(x){
-  map(ncol_sm, function(y){
-   cell <- sm_hod_list[-1] %>% extract(x, y) %>% unlist()
-   quantile(cell, prob = quantile_prob_val)
-})  %>% bind_rows(.id = "categories_serial_id")
-}) %>% bind_rows(.id = "customer_serial_id")
-
-```
-
-```{r hod-heatmap, fig.cap="Heatmaps show deciles on the y-axis with customers on the x-axis with colors filled by the value of deciles and faceted by hours of the day.
-It again seems like all the households have gradual change in colors (almost for all hours) as we move to higher deciles except for the 4th household.
-"}
-categories_ref <- names(sm_hod_list)[-1] %>% as_tibble() %>% 
-  mutate(categories_serial_id = as.character(row_number()))
-
-data_heatmap <- sm_hod_quantiles %>% 
-  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
-  left_join(categories_ref) %>% 
-  rename("category" = "value") %>% 
-  select(-categories_serial_id) %>% 
-  mutate(category = as.numeric(category))
-
-data_heatmap%>% 
-ggplot(aes(x= customer_serial_id, y = as.factor(quantile_prob))) +
-  geom_tile(aes(fill = quantile_val)) +
-  scale_fill_distiller(palette = "YlGnBu")+ 
-  facet_wrap(~category) +
-  theme(axis.text.x = element_text(angle = 90)) +
-  ylab("deciles") +
-  xlab("customers")
-```
-
-```{r hod-deciles, fig.cap = "The deciles plots for 5 households are shown for entire observation period across hours of the day.
-The y-scales are made free to allow us to see daily patterns for each households.
-The deciles for all households tend to show an increase in the morning and evening hours, although the hours differ.
-Except for the 90th decile, the deciles for the 4th household look pretty flat."}
-data_heatmap %>% 
-  ggplot() + geom_line(aes(x=category, 
-                           y = quantile_val, 
-                           color = quantile_prob)) + facet_wrap(~customer_serial_id, ncol = 1, scales="free_y",
-                                                                strip.position = "right") + 
-  scale_color_brewer(palette = "Blues") + 
-  #scale_y_log10() +
-  theme_bw() +
-  xlab("hours of the day") +
-  ylab("demand (in Kwh)")
-```

### 2021-07-08 supplementary/realdata-cust2-orig.Rmd
all updated files in supplementary

--- 2021-07-08 supplementary/realdata-cust10.Rmd
@@
 ---
-title: "Clustering real data with ten customers and choosing optimal number of clusters"
+title: "Clustering real data with two customers"
 output:
   bookdown::pdf_book:
+    MonashEBSTemplates::workingpaper:
     #base_format: rticles::asa_article
-    fig_height: 5
-    fig_width: 8
+    fig_height: 8
+    fig_width: 12
     fig_caption: yes
@@
     keep_tex: yes
-    toc: false    
 ---
@@
 ```{r load-lib}
+
 ```
 
-# Clutering approach
+# What you are doing?
 
-1. _Compute quantiles of distributions across each hour of day_
-2. _Compute JS distance between households for each hour of day_
-3. _Total distance between households computed as sum of JS distances for all hours_
-4. _Cluster using this distance with hierarchical clustering algorithm (method "complete")_
+Taking real customers from the smart meter data set and cluster them and see if their distributions are different
 
-# Plots of raw data
+-  just one cyclic granularity (hod)
+-  two cyclic granularities (hod and dow)
 
-Five households are considered for the following analysis from the SGSC data set.
-The data sets contain three columns `customer_id`, `reading_datetime` and `general_supply_kwh`.
-They consist of half-hourly data from 2012-2014.
-Figure \ref{fig:raw-all} shows how the raw time plots for these five households look like.
-Since all the dataset looks squeezed on this linear time scale, Figure \ref{fig:raw-sub} shows how the raw plot looks for 2 months (Sept 2013 and Oct 2013).
+_Compute quantiles of conditional distributions_
+Conditional quantiles are obtained for each combination of categories.
+
+## entire data
+```{r}
+sm <- gravitas::smart_meter10 %>%
+  filter(customer_id %in% c("10006704", "10017936"))
+
+sm
+```
+
+## how do they look
 
 ```{r}
-sm <- gravitas::smart_meter10 
-```
-
-```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for 5 households (facets).
-The data is squeezed stopping us from seeing any behavioral patterns."}
-library(lubridate)
-
 sm %>% 
@@
   geom_line() +
-  facet_wrap(~customer_id, ncol = 1, 
-             strip.position = "right") +
-  theme_bw() +
-  xlab("time") +
-  ylab("demand (in Kwh)")
+  facet_wrap(~customer_id, ncol = 1) +
+  theme_bw()
 ```
 
-```{r raw-sub, fig.cap="The raw plots for ten customers shown for Sept 2013-Oct 2013.
-The data is zoomed in and the y-scales made free to emphasize weekly, hourly or any behavior/patterns that they might have.
-There is some daily and weekly pattern in all households except the 4th one, that has spikes which seem to occur at irregular intervals."}
-sm %>% 
-  filter(year(reading_datetime) == 2013 &
-           month(reading_datetime) %in% c(9, 10)) %>% 
-  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
-  geom_line() +
-  facet_wrap(~customer_id, ncol = 1, scales = "free_y", 
-             strip.position = "right") +
-  theme_bw()+
-  ylab("demand (in Kwh)") +
-  theme(panel.grid.major.x = element_blank()) +
-  scale_x_datetime("Date", date_labels = "%b %d",
-                   breaks = "1 week",
-                   date_minor_breaks = "1 day")  + 
-  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
-        panel.grid.minor.x =  element_line(colour = "#D3D3D3"))
-
-```
-
-# Plots of distribution of data across categories of granularities
-
-The distribution of demand across different hours of the day observed for these five households through heatmaps (Figure \ref{fig:hod-deciles}) and quantile plots (Figure \ref{fig:hod-heatmap}).
-The following characterization is done to validate results of the clustering approach later on.
-
-```{r hod-box, fig.cap="The raw plots for 5 households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
+```{r}
 sm %>% 
@@
   geom_boxplot() + 
-  facet_wrap(~customer_id, ncol = 1, scales = "free_y",
-             strip.position = "right") 
+  facet_wrap(~customer_id, ncol = 1) 
 ```
+
+## make one granularity
 
@@
 ```
+
+## conditional list across categories of one granularity
 
@@
 ```
+
+## conditional quantiles across categories of one granularity
 
@@
 
-```{r hod-heatmap, fig.cap="Heatmaps show deciles on the y-axis with customers on the x-axis with colors filled by the value of deciles and faceted by hours of the day.
-It again seems like all the households have gradual change in colors (almost for all hours) as we move to higher deciles except for the 4th household.
-"}
+## how do their deciles look for different categories of cyclic granularity
+
+```{r}
 categories_ref <- names(sm_hod_list)[-1] %>% as_tibble() %>% 
@@
   theme(axis.text.x = element_text(angle = 90)) +
-  ylab("deciles") +
-  xlab("customers")
+  ylab("Deciles") +
+  xlab("data sets")
 ```
 
-```{r hod-deciles, fig.cap = "The deciles plots for 5 households are shown for entire observation period across hours of the day.
-The y-scales are made free to allow us to see daily patterns for each households.
-The deciles for all households tend to show an increase in the morning and evening hours, although the hours differ.
-Except for the 90th decile, the deciles for the 4th household look pretty flat."}
+```{r, eval = FALSE}
 data_heatmap %>% 
-  ggplot() + geom_line(aes(x=category, 
-                           y = quantile_val, 
-                           color = quantile_prob)) + facet_wrap(~customer_serial_id, ncol = 1, scales="free_y",
-                                                                strip.position = "right") + 
-  scale_color_brewer(palette = "Blues") + 
+  ggplot() + geom_line(aes(x=category, y = quantile_val, color = quantile_prob)) + facet_wrap(~customer_serial_id, ncol = 1) + 
+  scale_color_brewer(palette = "Dark2") + 
   #scale_y_log10() +
-  theme_bw() +
-  xlab("hours of the day") +
-  ylab("demand (in Kwh)")
+  theme_bw()
+
+data_heatmap %>% 
+  pivot_wider(names_from = customer_serial_id, values_from = quantile_val) %>% 
+  group_by(quantile_prob) %>% 
+  mutate(percent_diff = if_else(`1`==0, 0, (`2`- `1`)*100/`1`)) %>%  ggplot() + geom_line(aes(x=category, y = percent_diff, color = quantile_prob)) +
+  scale_color_brewer(palette = "Dark2") +
+  theme_bw()
 ```
 
-# Iterations of each customer (adding some random noise for each iteration)
+## Iterations of each customer
 
-Iterations of each customer are considered by adding random noise ($N(0, \sigma^2)$, where $\sigma^2$ is very small relative to the the variance of the data).
-Figure \ref{fig:raw-iter} shows that raw plots of the simulated iterations to show that their structure is same as the parent data set.
+### Data preparation
 
@@
      mutate(customer_id = paste(customer_id, seed, sep = "-"),
-             general_supply_kwh = general_supply_kwh + rnorm(1, 0, 0.001))
+             general_supply_kwh = general_supply_kwh + rnorm(1, 0, 0.01))
      data1
@@
 
-```{r raw-iter, fig.cap = "Five iterations for each customer are considered by adding random noise $(N(0, \\sigma^2)$, where $\\sigma^2$ is very small relative to the the variance of the data.
-The raw plots for all simulated dataset is shown to make sure they look similar to the structure."}
+```{r}
 all_data_clust %>% 
@@
   geom_line() +
-  facet_wrap(~customer_id,
-             strip.position = "right", nrow = 10) +
-  theme_bw() + 
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  xlab("time") +
-  ylab("demand (in Kwh)")
+  facet_wrap(~customer_id, ncol = 1) +
+  theme_bw()
 ```
+
+### make one granularity
 
@@
 
+## conditional quantiles across categories of one granularity
+
 ```{r}
@@
 ```
+
+## JS Pairwise distances between households
 
@@
 
-# Dendogram
+### hc
 
-The following figures show dendograms when we are using optimal number of clusters (k) from k = fpc::nselectboot(), k = 2, 3, 4 and 5.
-We observe that the $4^{th}$ household tend to get split as different clusters as increase the number of clusters.
+```{r}
+d = stats::as.dist(dist_mat_format)
+hc = stats::hclust(d,method="complete")
+plot(hc)
 
-```{r hc}
+mds <- d %>%
+  cmdscale() %>%
+  as_tibble()
+
+colnames(mds) <- c("Dim.1", "Dim.2")
+rownames(mds) <- lookup_tbl$customer_id
+
+groups<-cutree(hc, k=2)
+
+all_data_cluster <- cbind(dist_mat_format, groups) %>%
+  cbind(mds)%>%
+  mutate(groups = as.factor(groups)) %>% as_tibble()
+
+mds_plot_10 <- ggplot(all_data_cluster,
+       aes(x = Dim.1,
+           y = Dim.2,
+       color = groups,
+       label = rownames(mds))) +
+  geom_point(size = 1) +
+  geom_text(check_overlap = TRUE)  +
+  theme_classic()+
+  theme(legend.position = "bottom") +
+  scale_color_brewer(palette = "Dark2")
+
+mds_plot_10
+```
+
+# Scaled data
+
+### make one granularity
+
+```{r}
+sm_hod <- gravitas::create_gran(all_data_clust, "hour_day")
+```
+
+## conditional list across categories of one granularity
+
+```{r}
+
+sm_hod <- sm_hod %>% 
+  group_by(customer_id) %>% 
+  mutate(general_supply_kwh = scale(general_supply_kwh))
+
+sm_hod_list <- sm_hod %>% 
+  as_tibble() %>% 
+  select(customer_id, hour_day, general_supply_kwh) %>% 
+  pivot_wider(names_from = hour_day,
+              values_from = general_supply_kwh)
+
+lookup_tbl <- sm_hod_list %>% 
+  mutate(
+         customer_serial_id = row_number()) %>% 
+  select(customer_id, customer_serial_id)
+
+```
+
+## conditional quantiles across categories of one granularity
+
+```{r}
+quantile_prob_val = seq(0.1, 0.9, 0.1)
+
+ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)]
+nrow_sm <- seq_len(nrow(sm_hod_list))
+
+sm_hod_quantiles <- map(nrow_sm, function(x){
+  map(ncol_sm, function(y){
+   cell <- sm_hod_list[-1] %>% extract(x, y) %>% unlist()
+   quantile(cell, prob = quantile_prob_val)
+})  %>% bind_rows(.id = "categories_serial_id")
+}) %>% bind_rows(.id = "customer_serial_id")
+
+```
+
+## JS Pairwise distances between households
+
+```{r}
+sm_dist_data <- sm_hod_quantiles %>% 
+  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
+  select(-quantile_prob) %>% 
+  group_by(customer_serial_id, categories_serial_id) %>%
+  nest() %>% 
+  pivot_wider(names_from = categories_serial_id, values_from = data)
+```
+
+```{r}
+#ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
+nrow_data <- nrow(sm_dist_data)
+ncol_data <-  ncol(sm_dist_data[-1])
+
+dist_data <- map(seq_len(nrow_data), function(x){ # first data
+  map(seq_len(nrow_data), function(y){ # 2nd data
+    map(seq_len(ncol_data), function(z){ # number of combinations nx*nfacet
+      JS(
+        prob = quantile_prob_val,
+        unlist(sm_dist_data[-1] %>% extract(x, z)),
+        unlist(sm_dist_data[-1] %>% extract(y, z))
+      ) %>% as_tibble()
+    })%>% bind_rows(.id = "combinations")
+  })%>% bind_rows(.id = "data_type1")
+}) %>% bind_rows(.id = "data_type2")
+
+dist_mat <- dist_data %>%
+  mutate(data_type1 = as.numeric(data_type1), data_type2 = as.numeric(data_type2))%>%
+  group_by(data_type1, data_type2) %>%
+  summarise(dist = sum(value)) %>%
+  pivot_wider(names_from = data_type2,
+              values_from = dist)
+
+dist_mat_format <- dist_mat %>%
+  ungroup() %>% 
+  select(-data_type1)
+
+rownames(dist_mat_format) <- lookup_tbl$customer_id
+
+```
+
+### hc
+
+```{r}
 d = stats::as.dist(dist_mat_format)
@@
 
-# Multi-dimensional scaling
-
-```{r mds-new}
-set.seed(12345)
+```{r}
 mds <- d %>%
@@
 ```
-Optimal number of clusters as defined by the _fpc::nselectboot_ is `r koptimal`.
 
-_k = 2_
+_JS Pairwise distances between datasets_
+Distance between the data sets is computed as the sum of JS distances across different categories.
 
-```{r mds4}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 2, border = 2:5)
-```
+_Hierarchical clustering with 4 clusters_
+Hierarchical clustering is performed using k = 4 and dendogram observed
 
-_k = 3_
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 3, border = 2:5)
-```
-
-_k = 4_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 4, border = 2:5)
-```
-
-_k = 5_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 5, border = 2:6)
-```
-
-_k = 6_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 6, border = 2:7)
-```
-
-_k = 7_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 7, border = 2:8)
-```
-
-_k = 8_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 8, border = 2:9)
-```
-
-_k = 9_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 9, border = 2:10)
-```
-
-_k = 10_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 10, border = 2:11)
-```
+_Clusters obtained visualized using MDS_
+Each cluster represents data sets from a separate design

### 2021-07-08 supplementary/realdata-cust3.Rmd
all updated files in supplementary

--- 2021-07-08 supplementary/realdata-cust10.Rmd
@@
 ---
-title: "Clustering real data with ten customers and choosing optimal number of clusters"
+title: "Clustering real data with three customers and choosing optimal number of clusters"
 output:
@@
 
-Five households are considered for the following analysis from the SGSC data set.
+four households are considered for the following analysis from the SGSC data set.
 The data sets contain three columns `customer_id`, `reading_datetime` and `general_supply_kwh`.
 They consist of half-hourly data from 2012-2014.
-Figure \ref{fig:raw-all} shows how the raw time plots for these five households look like.
+Figure \ref{fig:raw-all} shows how the raw time plots for these four households look like.
 Since all the dataset looks squeezed on this linear time scale, Figure \ref{fig:raw-sub} shows how the raw plot looks for 2 months (Sept 2013 and Oct 2013).
@@
 ```{r}
-sm <- gravitas::smart_meter10 
+sm <- gravitas::smart_meter10 %>%
+  filter(customer_id %in% c("10006704", "10017936", "10018250"))
 ```
 
-```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for 5 households (facets).
+```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for four households (facets).
 The data is squeezed stopping us from seeing any behavioral patterns."}
@@
 
-```{r raw-sub, fig.cap="The raw plots for ten customers shown for Sept 2013-Oct 2013.
+```{r raw-sub, fig.cap="The raw plots for 4 a shown for Sept 2013-Oct 2013.
 The data is zoomed in and the y-scales made free to emphasize weekly, hourly or any behavior/patterns that they might have.
@@
 
-The distribution of demand across different hours of the day observed for these five households through heatmaps (Figure \ref{fig:hod-deciles}) and quantile plots (Figure \ref{fig:hod-heatmap}).
+The distribution of demand across different hours of the day observed for these four households through heatmaps (Figure \ref{fig:hod-deciles}) and quantile plots (Figure \ref{fig:hod-heatmap}).
 The following characterization is done to validate results of the clustering approach later on.
 
-```{r hod-box, fig.cap="The raw plots for 5 households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
+```{r hod-box, fig.cap="The raw plots for four households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
 sm %>% 
@@
   geom_boxplot() + 
-  facet_wrap(~customer_id, ncol = 1, scales = "free_y",
-             strip.position = "right") 
+  facet_wrap(~customer_id, ncol = 1, scales = "free_y") 
 ```
@@
 
-```{r hod-deciles, fig.cap = "The deciles plots for 5 households are shown for entire observation period across hours of the day.
+```{r hod-deciles, fig.cap = "The deciles plots for four households are shown for entire observation period across hours of the day.
 The y-scales are made free to allow us to see daily patterns for each households.
@@
                            y = quantile_val, 
-                           color = quantile_prob)) + facet_wrap(~customer_serial_id, ncol = 1, scales="free_y",
-                                                                strip.position = "right") + 
+                           color = quantile_prob)) + facet_wrap(~customer_serial_id, ncol = 1, scales="free_y") + 
   scale_color_brewer(palette = "Blues") + 
@@
 
-```{r raw-iter, fig.cap = "Five iterations for each customer are considered by adding random noise $(N(0, \\sigma^2)$, where $\\sigma^2$ is very small relative to the the variance of the data.
+```{r raw-iter, fig.cap = "four iterations for each customer are considered by adding random noise $(N(0, \\sigma^2)$, where $\\sigma^2$ is very small relative to the the variance of the data.
 The raw plots for all simulated dataset is shown to make sure they look similar to the structure."}
@@
   geom_line() +
-  facet_wrap(~customer_id,
-             strip.position = "right", nrow = 10) +
+  facet_wrap(~customer_id) +
   theme_bw() + 
@@
 ```
+
 Optimal number of clusters as defined by the _fpc::nselectboot_ is `r koptimal`.
-
-_k = 2_
-
-```{r mds4}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 2, border = 2:5)
-```
 
@@
 ```
-
-_k = 4_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 4, border = 2:5)
-```
-
-_k = 5_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 5, border = 2:6)
-```
-
-_k = 6_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 6, border = 2:7)
-```
-
-_k = 7_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 7, border = 2:8)
-```
-
-_k = 8_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 8, border = 2:9)
-```
-
-_k = 9_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 9, border = 2:10)
-```
-
-_k = 10_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 10, border = 2:11)
-```

### 2021-07-08 supplementary/realdata-cust4.Rmd
all updated files in supplementary

--- 2021-07-08 supplementary/realdata-cust3.Rmd
@@
 ---
-title: "Clustering real data with three customers and choosing optimal number of clusters"
+title: "Clustering real data with four customers and choosing optimal number of clusters"
 output:
@@
 sm <- gravitas::smart_meter10 %>%
-  filter(customer_id %in% c("10006704", "10017936", "10018250"))
+  filter(customer_id %in% c("10006704", "10017936", "10006414", "10018250"))
 ```
@@
 ```
+Optimal number of clusters as defined by the _fpc::nselectboot_ is `r koptimal`.
 
-Optimal number of clusters as defined by the _fpc::nselectboot_ is `r koptimal`.
+_k = 2_
+
+```{r mds4}
+plot(hc, cex = 0.6)
+rect.hclust(hc, k = 2, border = 2:5)
+```
 
@@
 ```
+
+_k = 4_
+
+```{r}
+plot(hc, cex = 0.6)
+rect.hclust(hc, k = 4, border = 2:5)
+```

### 2021-07-08 supplementary/realdata-cust5-2gran.Rmd
all updated files in supplementary

--- 2021-07-08 supplementary/realdata-cust10.Rmd
@@
 ---
-title: "Clustering real data with ten customers and choosing optimal number of clusters"
+title: "Clustering real data with five customers and choosing optimal number of clusters for hours of day and day of week"
 output:
@@
 
-1. _Compute quantiles of distributions across each hour of day_
-2. _Compute JS distance between households for each hour of day_
-3. _Total distance between households computed as sum of JS distances for all hours_
+1. _Compute quantiles of distributions across each combination of hour of day and day of week _
+2. _Compute JS distance between households for each combination of categories_
+3. _Total distance between households computed as sum of JS distances for all combinations_
 4. _Cluster using this distance with hierarchical clustering algorithm (method "complete")_
@@
 ```{r}
-sm <- gravitas::smart_meter10 
+sm <- gravitas::smart_meter10 %>%
+  filter(customer_id %in% c("10006704", "10017936", "10018064"))
 ```
@@
 
-```{r raw-sub, fig.cap="The raw plots for ten customers shown for Sept 2013-Oct 2013.
+```{r raw-sub, fig.cap="The raw plots for 5 a shown for Sept 2013-Oct 2013.
 The data is zoomed in and the y-scales made free to emphasize weekly, hourly or any behavior/patterns that they might have.
@@
 
-```{r hod-box, fig.cap="The raw plots for 5 households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
-sm %>% 
-  create_gran("hour_day") %>% 
-  ggplot(aes(x = hour_day, y = general_supply_kwh)) + 
-  geom_boxplot() + 
-  facet_wrap(~customer_id, ncol = 1, scales = "free_y",
-             strip.position = "right") 
+```{r}
+sm_hod_dow <- sm %>% gravitas::create_gran("hour_day") %>% 
+  gravitas::create_gran("day_week")
+
 ```
@@
 ```{r}
-sm_hod <- gravitas::create_gran(sm, "hour_day")
-```
-
-```{r}
-sm_hod_list <- sm_hod %>% 
+sm_hod_list <- sm_hod_dow %>% 
   as_tibble() %>% 
-  select(customer_id, hour_day, general_supply_kwh) %>% 
-  pivot_wider(names_from = hour_day,
+  arrange(customer_id, day_week, hour_day) %>% 
+  mutate(category = paste(hour_day, day_week, sep = "-")) %>% 
+   select(customer_id, category, general_supply_kwh) %>% 
+  pivot_wider(names_from = category,
               values_from = general_supply_kwh)
@@
 "}
-categories_ref <- names(sm_hod_list)[-1] %>% as_tibble() %>% 
+categories_ref <- names(sm_hod_list[-1]) %>% as_tibble() %>% 
   mutate(categories_serial_id = as.character(row_number()))
 
+ref_cat_split <- map(1:nrow(categories_ref), function(x){
+  value = str_split(categories_ref$value[x], pattern = "-") %>% unlist() 
+  hod = value[1]
+  dow = value[2]
+  tibble(hod = hod, dow = dow)
+}) %>% bind_rows(.id = "categories_serial_id")
+
 data_heatmap <- sm_hod_quantiles %>% 
-  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
+  pivot_longer(-c(1, 2),
+               names_to = "quantile_prob",
+               values_to = "quantile_val") %>% 
   left_join(categories_ref) %>% 
   rename("category" = "value") %>% 
-  select(-categories_serial_id) %>% 
-  mutate(category = as.numeric(category))
+  left_join(ref_cat_split) %>% 
+  mutate(hod = as.integer(hod))
+
+data_heatmap$dow <- factor(data_heatmap$dow, 
+                           levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))
 
 data_heatmap%>% 
-ggplot(aes(x= customer_serial_id, y = as.factor(quantile_prob))) +
+ggplot(aes(x= customer_serial_id, 
+           y = as.factor(quantile_prob))) +
   geom_tile(aes(fill = quantile_val)) +
   scale_fill_distiller(palette = "YlGnBu")+ 
-  facet_wrap(~category) +
+  facet_grid(hod~dow) +
   theme(axis.text.x = element_text(angle = 90)) +
@@
 
-```{r hod-deciles, fig.cap = "The deciles plots for 5 households are shown for entire observation period across hours of the day.
+```{r hod-deciles-new, fig.cap = "The deciles plots for 5 households are shown for entire observation period across hours of the day.
 The y-scales are made free to allow us to see daily patterns for each households.
@@
 Except for the 90th decile, the deciles for the 4th household look pretty flat."}
-data_heatmap %>% 
-  ggplot() + geom_line(aes(x=category, 
-                           y = quantile_val, 
-                           color = quantile_prob)) + facet_wrap(~customer_serial_id, ncol = 1, scales="free_y",
-                                                                strip.position = "right") + 
-  scale_color_brewer(palette = "Blues") + 
-  #scale_y_log10() +
-  theme_bw() +
-  xlab("hours of the day") +
-  ylab("demand (in Kwh)")
+sm %>% filter(customer_id %in% c("10006704")) %>% 
+           gravitas::prob_plot("day_week", "hour_day", plot_type = "quantile") +
+  theme_bw()
+
+sm %>% filter(customer_id %in% c("10017936")) %>% 
+           gravitas::prob_plot("day_week", "hour_day", plot_type = "quantile") +
+  theme_bw()
+
+sm %>% filter(customer_id %in% c("10018064")) %>% 
+           gravitas::prob_plot("day_week", "hour_day", plot_type = "quantile") +
+  theme_bw()
 ```
@@
      mutate(customer_id = paste(customer_id, seed, sep = "-"),
-             general_supply_kwh = general_supply_kwh + rnorm(1, 0, 0.001))
+             general_supply_kwh = general_supply_kwh + rnorm(1, 0, 0.03))
      data1
@@
   geom_line() +
-  facet_wrap(~customer_id,
-             strip.position = "right", nrow = 10) +
+  facet_wrap(~customer_id, nrow = 3) +
   theme_bw() + 
@@
 ```{r}
-sm_hod <- gravitas::create_gran(all_data_clust, "hour_day")
+sm_hod_dow <- gravitas::create_gran(all_data_clust, "hour_day") %>% 
+  create_gran ("day_week")
 ```
@@
 ```{r}
-sm_hod_list <- sm_hod %>% 
+sm_hod_list <- sm_hod_dow %>% 
   as_tibble() %>% 
-  select(customer_id, hour_day, general_supply_kwh) %>% 
-  pivot_wider(names_from = hour_day,
+  arrange(customer_id, day_week, hour_day) %>% 
+    mutate(category = paste(hour_day, day_week, sep = "-")) %>% 
+   select(customer_id, category, general_supply_kwh) %>% 
+  pivot_wider(names_from = category,
               values_from = general_supply_kwh)
@@
 ```{r}
+
 sm_dist_data <- sm_hod_quantiles %>% 
-  mutate(customer_serial_id = as.numeric(customer_serial_id)) %>% 
-  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
-  select(-quantile_prob) %>% 
-  group_by(customer_serial_id, categories_serial_id) %>%
+  pivot_longer(-c(1, 2), 
+               names_to = "quantile_prob", 
+               values_to = "quantile_val") %>% 
+  group_by(customer_serial_id, 
+           categories_serial_id) %>%
   nest() %>% 
-  pivot_wider(names_from = categories_serial_id, values_from = data)
+  pivot_wider(names_from = categories_serial_id, values_from = data) %>% ungroup()
 ```
@@
 
-```{r}
+```{r, message=FALSE}
 #ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
@@
         prob = quantile_prob_val,
-        unlist(sm_dist_data[-1] %>% extract(x, z)),
-        unlist(sm_dist_data[-1] %>% extract(y, z))
+        sm_dist_data[-1] %>% extract(x, z) %>% unnest() %>% pull(quantile_val),
+        sm_dist_data[-1] %>% extract(y, z) %>% unnest() %>% pull(quantile_val)
       ) %>% as_tibble()
@@
                         classification = "averagedist",
-                        krange = 2:12)
+                        krange = 2:5)
 
@@
 
-groups<-cutree(hc, k=koptimal)
+groups<-cutree(hc, k=2)
 
@@
 ```
-Optimal number of clusters as defined by the _fpc::nselectboot_ is `r koptimal`.
 
@@
 ```
-
-_k = 6_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 6, border = 2:7)
-```
-
-_k = 7_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 7, border = 2:8)
-```
-
-_k = 8_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 8, border = 2:9)
-```
-
-_k = 9_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 9, border = 2:10)
-```
-
-_k = 10_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 10, border = 2:11)
-```

### 2021-07-08 supplementary/realdata-cust5.Rmd
all updated files in supplementary

--- 2021-07-08 supplementary/realdata-cust4.Rmd
@@
 ---
-title: "Clustering real data with four customers and choosing optimal number of clusters"
+title: "Clustering real data with five customers and choosing optimal number of clusters"
 output:
@@
 
-four households are considered for the following analysis from the SGSC data set.
+Five households are considered for the following analysis from the SGSC data set.
 The data sets contain three columns `customer_id`, `reading_datetime` and `general_supply_kwh`.
 They consist of half-hourly data from 2012-2014.
-Figure \ref{fig:raw-all} shows how the raw time plots for these four households look like.
+Figure \ref{fig:raw-all} shows how the raw time plots for these five households look like.
 Since all the dataset looks squeezed on this linear time scale, Figure \ref{fig:raw-sub} shows how the raw plot looks for 2 months (Sept 2013 and Oct 2013).
@@
 sm <- gravitas::smart_meter10 %>%
-  filter(customer_id %in% c("10006704", "10017936", "10006414", "10018250"))
+  filter(customer_id %in% c("10006704", "10017936", "10006414", "10018250", "10018064"))
 ```
 
-```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for four households (facets).
+```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for 5 households (facets).
 The data is squeezed stopping us from seeing any behavioral patterns."}
@@
 
-```{r raw-sub, fig.cap="The raw plots for 4 a shown for Sept 2013-Oct 2013.
+```{r raw-sub, fig.cap="The raw plots for 5 a shown for Sept 2013-Oct 2013.
 The data is zoomed in and the y-scales made free to emphasize weekly, hourly or any behavior/patterns that they might have.
@@
 
-The distribution of demand across different hours of the day observed for these four households through heatmaps (Figure \ref{fig:hod-deciles}) and quantile plots (Figure \ref{fig:hod-heatmap}).
+The distribution of demand across different hours of the day observed for these five households through heatmaps (Figure \ref{fig:hod-deciles}) and quantile plots (Figure \ref{fig:hod-heatmap}).
 The following characterization is done to validate results of the clustering approach later on.
 
-```{r hod-box, fig.cap="The raw plots for four households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
+```{r hod-box, fig.cap="The raw plots for 5 households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
 sm %>% 
@@
 
-```{r hod-deciles, fig.cap = "The deciles plots for four households are shown for entire observation period across hours of the day.
+```{r hod-deciles, fig.cap = "The deciles plots for 5 households are shown for entire observation period across hours of the day.
 The y-scales are made free to allow us to see daily patterns for each households.
@@
 
-```{r raw-iter, fig.cap = "four iterations for each customer are considered by adding random noise $(N(0, \\sigma^2)$, where $\\sigma^2$ is very small relative to the the variance of the data.
+```{r raw-iter, fig.cap = "Five iterations for each customer are considered by adding random noise $(N(0, \\sigma^2)$, where $\\sigma^2$ is very small relative to the the variance of the data.
 The raw plots for all simulated dataset is shown to make sure they look similar to the structure."}
@@
 ```
+
+_k = 5_
+
+```{r}
+plot(hc, cex = 0.6)
+rect.hclust(hc, k = 5, border = 2:6)
+```

### 2021-07-08 supplementary/R/simdata01.R

@@ -1,0 +1,255 @@
+# scripts to see how the time series for the four designs look like
+
+##----load-lib
+library(hakear)
+library(tidyr)
+library(distributional)
+library(ggplot2)
+library(patchwork)
+library(purrr)
+library(dplyr)
+library(kableExtra)
+library(tidyverse)
+library(gravitas)
+library(magrittr)
+## ----designs
+
+sim_null_normal <- function(nxj, nfacetj, mean, sd, w1 = 0, w2=0) {
+  rep(distributional::dist_normal(mu = mean, sigma = sd),
+      times = nxj * nfacetj
+  )
+}
+
+sim_varf_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  rep(dist_normal((mean + seq(0, nfacet - 1, by = 1) * w1),
+                  (sd + seq(0, nfacet - 1, by = 1) * w2)), each = nx)
+}
+
+sim_varx_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  rep(dist_normal((mean + seq(0, nx - 1, by = 1) * w1),
+                  (sd + seq(0, nx - 1, by = 1) * w2)), nfacet)
+}
+
+sim_varall_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  dist_normal((mean + seq(0, (nx *nfacet - 1), by = 1) * w1),
+              (sd + seq(0,  (nx *  nfacet - 1), by = 1) * w2))
+}
+
+
+## ----data
+
+
+sim_panel_varall <- sim_panel(
+  nx = nx_val, nfacet =  nfacet_val,
+  ntimes = ntimes_val,
+  # sim_dist = sim_varall_normal(2, 3, 5, 10, 5, -1.5)
+  sim_dist = sim_varall_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+) %>% unnest(data)
+
+sim_panel_varx <- sim_panel(
+  nx = nx_val, nfacet =  nfacet_val,
+  ntimes = ntimes_val,
+  # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
+  sim_dist = sim_varx_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+) %>% unnest(data)
+
+sim_panel_varf <- sim_panel(
+  nx = nx_val, nfacet =  nfacet_val,
+  ntimes = ntimes_val,
[…]

### 2021-07-08 supplementary/sim_table/aggregate01.R

@@ -1,0 +1,45 @@
+# contains function to aggregate data each for from N(0,1), N(0,5), N(5,1), N(5,5), Gamma(0.5, 1), Gamma(2, 1)
+#  aggregated data from the file simulations/results_norm_scalar/norm/folder_name and tidy them up
+library(dplyr)
+library(here)
+library(readr)
+library(rlang)
+library(purrr)
+library(rlang)
+  
+  all_files = list.files(path = paste0("paper/sim_table/results_norm_scalar/"), 
+                         pattern = ".rds")
+  
+  names_levels <- map_dfr(all_files, 
+                          function(x){
+                            z = str_split(str_remove(x, 
+                                                     ".rds"), 
+                                          "-") %>% unlist()
+  
+                            bind_cols(customer_id = as.numeric(z[1]),
+                                      facet_variable = z[2],
+                                      x_variable = z[3])
+                          })
+  
+  # len_file = read_rds(("simulations/results_norm_scalar/norm/tuning_param/2_2_tuning_param.rds"))
+  # 
+  # names_rep <- names_levels %>% slice(rep(1:n(), each = nrow(len_file)))
+  
+  
+  all_files_path <- paste0("paper/sim_table/results_norm_scalar/",
+                           all_files)  
+  
+  
+  all_data <- lapply(1:length(all_files_path), function(x){
+    
+    data = all_files_path %>% magrittr::extract2(x) %>% 
+      readRDS()
+    # 
+    # names = names_levels %>% magrittr::extract(x,)
+    # names_rep =   names %>% slice(rep(1:n(), each = nrow(data)))
+    # bind_cols(names_rep, data)
+  }) %>% 
+    bind_rows() %>%
+    arrange(customer_id)
+  
+  write_rds(all_data, "paper/sim_table/all_data_scalar.rds")

### 2021-07-16 algo2-elec16.Rmd
adding results from new and old method

---
title: "algo2-elec16"
author: "Sayani Gupta"
date: "15/07/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-lib}
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
#library(ggridges)
library(tsibble)
library(gravitas)
library(ggpubr)
library(readr)
library(kableExtra)
library(distributional)
library(ggplot2)
library(sugrrants)
library(here)
library(ggplot2)
library(patchwork)
library(scales)
library(GGally)
library(viridis)
```

```{r, message=FALSE, warning=FALSE, echo = FALSE}

##----harmony-all
elec_harmony_8local <- read_rds("../paper-hakear/paper/data/elec_harmony_all.rds")

elec_harmony_new8 <- 
read_rds("data/elec_harmony_100-gap.rds") %>% 
  bind_rows(.id = "id") %>% 
  mutate(id = as.numeric(id)) %>% 
  mutate(id = 8 + id) %>% 
  mutate(facet_variable = case_when(
    facet_variable == "hour_day" ~ "hod" ,
    facet_variable == "day_month" ~ "dom" ,
    facet_variable == "day_week" ~ "dow" ,
    facet_variable == "week_month" ~ "wom" ,
    facet_variable == "wknd_wday" ~ "wdwnd"
  )) %>% 
  mutate(x_variable = case_when(
    x_variable == "hour_day" ~ "hod" ,
    x_variable == "day_month" ~ "dom" ,
    x_variable == "day_week" ~ "dow" ,
    x_variable == "week_month" ~ "wom" ,
    x_variable == "wknd_wday" ~ "wdwnd"
  )) %>% 
  mutate(id = paste("id", id, sep = " ")) %>% 
  group_by(id) %>% 
  mutate(rank = row_number())

elec_harmony_all <- bind_rows(elec_harmony_8local, elec_harmony_new8)

select_split <- str_split(elec_harmony_all$select_harmony, " ", simplify = TRUE)[,2]

elec_sig_split <- elec_harmony_all %>% 
  bind_cols(select_split = select_split) %>% 
  mutate(significant = case_when(
    select_split == "***" ~ "highest",
    select_split == "**" ~ "high",
    select_split == "*" ~ "medium",
    select_split == "" ~ "low"
  )) %>% 
  mutate(rank = case_when(
    select_split == "***" ~ paste(rank, "***", sep = " "),
    select_split == "**" ~  paste(rank, "**", sep = " "),
    select_split == "*" ~  paste(rank, "*", sep = " "),
    select_split == "" ~  paste(rank, "", sep = " ")
  ))%>%mutate(comb = paste(facet_variable, x_variable, sep = "-"))

elec_sig_split$significant <- 
  factor(elec_sig_split$significant, levels = c("highest", "high", "medium", "low"))

data_comb_filter = elec_sig_split %>% 
mutate(low_proxy = if_else(significant=="low", 0, 1))%>% 
select(comb, id, low_proxy) %>% 
pivot_wider(names_from = id, values_from = low_proxy) %>% 
mutate(score = rowMeans(across(`id 1`:`id 16`))) %>% 
filter(score !=0)

data_sig_filter = elec_sig_split %>% 
filter(comb %in% data_comb_filter$comb)

data_clust = data_sig_filter %>% 
select(comb, id, wpd) %>% 
pivot_wider(names_from = comb, values_from = wpd)

hc = stats::hclust(dist(data_clust[-1]),method="complete")
plot(hc)

library(fpc)
koptimal = fpc::nselectboot(dist(data_clust[-1]),
                        B = 50,
                        method = "complete",
                        clustermethod = disthclustCBI, 
                        classification = "averagedist",
                        krange = 2:8)

koptimal = koptimal$kopt

plot(hc, cex = 0.6)
rect.hclust(hc, k = koptimal, border = 2:10)

groups <- cutree(hc, k=8) %>% as_tibble() %>% 
mutate(id = data_clust$id)

##----cluster-characterization

data = data_clust %>% 
left_join(groups)%>% 
             rename ("group" = "value") %>% 
             ungroup() %>% 
             select(group, everything(), -id) %>% 
             pivot_longer(-1, names_to = "comb", values_to = "wpd")

data_order <- data %>% 
  group_by(comb) %>% 
  summarise(m = mean(wpd)) %>% 
  arrange(desc(m))

data_median = data %>% 
  group_by(group, comb) %>% 
  summarize(wpd_med = median(wpd))

data_pcp <- data_median %>% 
mutate(group = as.character(group)) %>% 
  left_join(data_order) %>% 
  arrange(desc(m)) %>% 
  pivot_wider(-4, 
              names_from = comb, 
              values_from = wpd_med)

parcoord <- GGally::ggparcoord(data_pcp,
                   columns = 2:ncol(data_pcp),
                   groupColumn = 1,
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.8,
                   scale = "globalminmax"
) + ggplot2::theme_bw() +
  scale_color_viridis(discrete=TRUE) + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")
parcoord

##----read-orig-data

elec_8local <- read_rds(here("../paper-hakear/paper/data/elec_all-8.rds")) %>% 
  dplyr::filter(date(reading_datetime) >= ymd("20190701"), date(reading_datetime) < ymd("20191231"), meter_id==1) %>% 
  select(-meter_id) %>% 
  rename("id" = "household_id",
         "date_time" = "reading_datetime") %>% 
  mutate(date = date(date_time)) %>% 
  mutate(id = paste("id", id, sep = " ")) %>% 
  left_join(groups)%>% 
             rename ("group" = "value")

elec_ts <- read_rds("data/elec_ts.rds")

set.seed(12345)

sm_100 <- elec_ts %>%
  as_tibble() %>% 
  distinct(customer_id) %>%
  slice_sample(n = 100) %>% 
  head(8)

elec_sgsc <- filter(elec_ts, customer_id %in% sm_100$customer_id) %>% 
  ungroup()

sm <- bind_rows(elec_8local,
                elec_sgsc)

sm_hod_dow <- sm %>% 
gravitas::create_gran("hour_day") %>% 
  gravitas::create_gran("day_week")

sm_hod_list <- sm_hod_dow %>% 
  as_tibble() %>% 
  arrange(group, day_week, hour_day) %>% 
  mutate(category = paste(hour_day, day_week, sep = "-")) %>% 
   select(group, category, kwh) %>% 
  pivot_wider(names_from = category,
              values_from = kwh)

quantile_prob_val = c(0.25, 0.5, 0.75)

  sm_hod_dow %>% 
  filter(group==1) %>% 
  gravitas::prob_plot(
     "day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1)

  sm_hod_dow %>% 
  filter(group==2) %>% 
  gravitas::prob_plot(  "day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1)

  sm_hod_dow %>% 
  filter(group==3) %>% 
  gravitas::prob_plot(  "day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1)

  sm_hod_dow %>% 
  filter(group==4) %>% 
  gravitas::prob_plot(  "day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1) 
```
<!-- ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)] -->
<!-- nrow_sm <- seq_len(nrow(sm_hod_list)) -->

<!-- sm_hod_quantiles <- map(nrow_sm, function(x){ -->
<!--   map(ncol_sm, function(y){ -->
<!--    cell <- sm_hod_list[-1] %>%  -->
<!--    magrittr::extract(x, y) %>% unlist() -->
<!--    quantile(cell, prob = quantile_prob_val) -->
<!-- })  %>% bind_rows(.id = "categories_serial_id") -->
<!-- }) %>% bind_rows(.id = "group_id") -->

<!-- sm_hod_quantiles_cat <- map(nrow_sm, function(x){ -->
<!--   map(ncol_sm, function(y){ -->
<!--    cell <- sm_hod_list%>%  -->
<!--      dplyr::filter(group == x) %>%  -->
<!--      select(-c(1)) %>%  -->
<!--      magrittr::extract(y) %>%  -->
<!--      unlist() -->
<!--    quantile(cell, prob = quantile_prob_val) -->
<!-- })  %>% bind_rows(.id = "categories_serial_id") -->
<!-- }) %>% bind_rows(.id = "group_id") %>%  -->
<!--   mutate(group_id = as.integer(group_id), -->
<!--          categories_serial_id = as.integer(categories_serial_id)) -->

<!-- categories_ref_cat <- names(sm_hod_list)[-c(1)] %>% as_tibble() %>%  -->
<!--   mutate(categories_serial_id = as.character(row_number())) -->

<!-- data_heatmap <- sm_hod_quantiles %>%  -->
<!--   pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>%  -->
<!--   left_join(categories_ref_cat) %>%  -->
<!--   rename("category" = "value") %>%  -->
<!--   select(-categories_serial_id) -->

<!-- data_heatmap %>%  -->
<!--   pivot_wider( names_from = quantile_prob,  -->
<!--                values_from = quantile_val) %>%  -->
<!--   ggplot(aes(x = category)) +  geom_ribbon(aes(ymin = `25%`,  -->
<!--                   ymax = `75%`), fill = "lightblue") + -->
<!--   geom_line(aes(y = `50%`)) + -->
<!--   facet_wrap(~group_id, ncol = 1, scales = "free_y", labeller = "label_both") + -->
<!--   theme_bw() + -->
<!--     theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) -->

### 2021-07-16 main.Rmd
adding results from new and old method

@@
 
-The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_s(v)$ denotes the probability distribution from observation $s$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
-In our work, we are using $s$ as denoting a household and the underlying variable as the electricity demand.
-So $f_s(v)$ is the distribution of household $s$ and $v$ is electricity demand.
+The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+In our work, we are using $i$ as denoting a household and the underlying variable as the electricity demand.
+So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
+
+We want to cluster distributions of the form $f_{i,j,k}(v)$, where $i$ and $j$ denote
+
 Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.

### 2021-07-16 old/realdata-cust10.Rmd
adding results from new and old method

--- 2021-07-08 supplementary/realdata-cust10.Rmd
@@
 ---
-title: "Clustering real data with ten customers and choosing optimal number of clusters"
+title: "Clustering real data with 10 customers and choosing optimal number of clusters"
 output:
@@
 
-```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for 5 households (facets).
+```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for ten households (facets).
 The data is squeezed stopping us from seeing any behavioral patterns."}
@@
 
-```{r hod-box, fig.cap="The raw plots for 5 households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
+```{r hod-box, fig.cap="The raw plots for ten households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
 sm %>% 
@@
 
-```{r hod-deciles, fig.cap = "The deciles plots for 5 households are shown for entire observation period across hours of the day.
+```{r hod-deciles, fig.cap = "The deciles plots for ten households are shown for entire observation period across hours of the day.
 The y-scales are made free to allow us to see daily patterns for each households.
@@
 
-The following figures show dendograms when we are using optimal number of clusters (k) from k = fpc::nselectboot(), k = 2, 3, 4 and 5.
+The following figures show dendograms when we are using optimal number of clusters (k) from k = fpc::nselectboot().
 We observe that the $4^{th}$ household tend to get split as different clusters as increase the number of clusters.

### 2021-07-16 old/simdata_cluster02.Rmd
adding results from new and old method

---
title: "Clustering"
author: "Sayani Gupta"
date: "16/06/2021"
output:
  bookdown::pdf_book:
    MonashEBSTemplates::workingpaper:
    #base_format: rticles::asa_article
    fig_height: 8
    fig_width: 12
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
read_chunk("simdata01.R")
```

# How the time series plot of the designs look like for different sample sizes?

```{r sim-val, echo = TRUE}
set.seed(9999)
nx_val = 2 # number of x-axis levels
nfacet_val = 3 # number of facet levels
w1_val = 3 # increment in mean
w2_val = 0 # increment in sd
mean_val = 0 # mean of normal distribution of starting combination
sd_val = 1 # sd of normal distribution of starting combination
quantile_prob_val = seq(0.1, 0.9, 0.1)
```

```{r load-lib}
```

```{r designs}
```

## Number of obervations per combination is 4

```{r, echo = TRUE}
ntimes_val = 4
```

```{r data}
```

```{r category-plots}
```

```{r}
change_index <- function(data){

index_new <- map(seq_len(ntimes_val), function(i){
  map((seq_len(nx_val*nfacet_val)), function(j)
    {
value = i + (j-1)*ntimes_val
})
  }) %>% unlist()

data_new = data %>%
  ungroup() %>%
  mutate(index_old = row_number(),
         index_new = index_new)

y = data_new[match(index_new, data_new$index_old),]

y <- y %>%
  mutate(time = row_number())

return(y)
}

endbreaks<- nrow(sim_panel_null)

p1 <- change_index(sim_panel_null) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p2 <- change_index(sim_panel_varf) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p3 <- change_index(sim_panel_varx) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue")+
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p4 <- change_index(sim_panel_varall) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
  panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

t1 <- sim_panel_varall %>% ungroup() %>%  select (-c(1, 2))

# %>% 
#   kable(format = "latex", caption = "Simulated data for categories format")

t2 <- change_index(sim_panel_varall) %>% select (-c(1, 2))

# %>% 
#   kable(format = "latex",
#         caption = "Simulated data for time series format")

kable(list(t1, t2), format = "latex", caption = "Simulated data across combination of categories (left) and manipulated time series format (right)") %>% 
  row_spec(1:24,  background =  "#d3d3d3")

# kable(t1) %>%
#   kable_styling(full_width = FALSE, position = "float_left")
# kable(t2) %>%
#   kable_styling(full_width = FALSE, position = "left")
```

```{r}
(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)
```

## Number of obervations per combination is 10

```{r nobs10, echo = TRUE}
ntimes_val = 10 # number of observations per combination 
```

```{r data}
```

```{r category-plots}
```

```{r line-graph}
```

```{r plot10}
(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)

```

## Number of obervations per combination is 50

```{r nobs50, echo = TRUE}
ntimes_val = 50 # number of observations per combination 
```

```{r data}
```

```{r category-plots}
```

```{r line-graph}
```

```{r plot50}
(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)

```

# Simulation: 4 datasets with 2 designs

Generate 4 series, two from design varf and another two from varall and see if clustering happens properly through the approaches

## Generate 2 time series from varx

```{r}
set.seed(9999)
```

```{r data_varf, echo = TRUE}

data1_varf <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = ntimes_val,
  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
  sim_dist = sim_varf_normal(2, 3, 0, 1, 5, 0)
) %>% unnest(data)

data2_varf <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = ntimes_val,
  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
  sim_dist = sim_varf_normal(2, 3, 0, 1, 2, 0)
) %>% unnest(data)
```

## Generate 2 time series from varall

```{r data_varall, echo = TRUE}
data1_varall <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = ntimes_val,
  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
  sim_dist = sim_varall_normal(2, 3, 0, 1, 5, 0)
) %>% unnest(data)

data2_varall <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = ntimes_val,
  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
  sim_dist = sim_varall_normal(2, 3, 0, 1, 2, 0)
) %>% unnest(data)
```

## How do they look?

```{r}
p_data1_varf <- data1_varf %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) +
  facet_wrap(~id_facet) +
  geom_boxplot() +
  ylab("") +
  #ggtitle(paste("(b)", round(varf, 2))) +
  xlab("x level")+
  theme_bw() +
  ggtitle("varf-1")

p_data2_varf <- data2_varf %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) +
  facet_wrap(~id_facet) +
  geom_boxplot() +
  ylab("") +
  #ggtitle(paste("(b)", round(varf, 2))) +
  xlab("x level")+
  theme_bw()+
  ggtitle("varf-2")

p_data1_varall <- data1_varall %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) +
  facet_wrap(~id_facet) +
  geom_boxplot() +
  ylab("") +
  #ggtitle(paste("(b)", round(varf, 2))) +
  xlab("x level")+
  theme_bw()+
  ggtitle("varall-1")

p_data2_varall <- data2_varall %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) +
  facet_wrap(~id_facet) +
  geom_boxplot() +
  ylab("") +
  #ggtitle(paste("(b)", round(varf, 2))) +
  xlab("x level")+
  theme_bw()+
  ggtitle("varall-2")

(p_data1_varf +p_data2_varf)/(p_data1_varall+p_data2_varall)
```

## Compute quantiles of conditional distributions

```{r comp_quant}

compute_quantiles(data1_varf) %>%
   unnest(sim_data_quantile) %>%
   ggplot() +
   geom_boxplot(aes(x = as.factor(id_x), y = sim_data_quantile)) +
   facet_wrap(~id_facet)

data_q <- bind_rows(data1_varf = 
compute_quantiles(data1_varf, 
                                 quantile_prob = quantile_prob_val) ,

 data2_varf = compute_quantiles(data2_varf,
                                 quantile_prob = quantile_prob_val),

 data1_varall = compute_quantiles(data1_varall,
                                 quantile_prob = quantile_prob_val) ,

 data2_varall = compute_quantiles(data2_varall,
                                 quantile_prob = quantile_prob_val), .id = "data_type"
)

#%>% mutate(data_type = as.integer(data_type))

```

## JS Pairwise distances between data sets

```{r JS}
```

```{r JS-pairwise, echo=TRUE}
data_q_wide <- data_q %>% 
  pivot_wider(names_from = data_type,
              values_from = sim_data_quantile) %>% 
  select(-c(1,2))

ndata <- data_q %>% distinct(data_type) %>% pull(data_type)
ldata <- length(ndata)
lcomb <-  nx_val*nfacet_val

dist_data <- map(1:ldata, function(x){ # first data
  map(1:ldata, function(y){ # 2nd data
    map(1:lcomb, function(z){ # number of combinations nx*nfacet
    JS(
        prob = quantile_prob_val,
        unlist(data_q_wide[z,x]),
        unlist(data_q_wide[z,y])
      ) %>% as_tibble()
  })%>% bind_rows(.id = "combinations")
})%>% bind_rows(.id = "data_type1")
}) %>% bind_rows(.id = "data_type2")

```

## Hierarchical clustering based on pairwise distances

```{r hc}

dist_mat <- dist_data %>% 
  group_by(data_type1, data_type2) %>% 
  summarise(dist = sum(value)) %>% 
  pivot_wider(names_from = data_type2,
              values_from = dist) %>% 
  ungroup() %>% 
  select(-1)

d = stats::as.dist(dist_mat)

hc = stats::hclust(d,method="complete")
plot(hc)
```

## Multi-dimensional scaling with hierarchical clusters

```{r mds-plot}

mds <- d %>%
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")

mds <- mds %>%
  mutate(data_type = ndata) %>%
  select((data_type)
         , Dim.1, Dim.2)

groups<-cutree(hc, k=2)

all_data_cluster <- cbind(dist_mat, groups) %>%
  cbind(mds)%>%
  mutate(groups = as.factor(groups))

mds_plot <- ggpubr::ggscatter(all_data_cluster,
                        x = "Dim.1",
                        y = "Dim.2",
                        label = mds$data_type,
                        color = "groups",
                        repel = TRUE,
                        size = 2) +
  scale_color_brewer(palette = "Dark2")
```

```{r}
p1 <- change_index(data1_varf) %>%
  ggplot(aes(x = index_new,
             y = sim_data)) +
  geom_line() +
  #scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  #geom_point(alpha = 0.5, color = "blue")
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("")+
  xlab("index") +
  ggtitle("data1_varf")

p2 <- change_index(data2_varf) %>%
  ggplot(aes(x = index_new,
             y = sim_data)) +
  geom_line() +
  #scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  #geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("")+
  xlab("index")+
  ggtitle("data2_varf")

p3 <- change_index(data1_varall) %>%
  ggplot(aes(x = index_new,
             y = sim_data)) +
  geom_line() +
  #scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  #geom_point(alpha = 0.5, color = "blue")+
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("")+
  xlab("index")+
  ggtitle("data1_varall")

p4 <- change_index(data2_varall) %>%
  ggplot(aes(x = index_new,
             y = sim_data)) +
  geom_line() +
  #scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  #geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
  panel.grid.minor.x =  element_blank())+
  ylab("")+
  xlab("index")+
  ggtitle("data2_varall")

mds_plot+(p1/p2/p3/p4)
```

# Repeat with 3 designs (for varying sample size)

varf and varall - 100 obs each combination and varx with 10 and 50 observation per combination taken

```{r makedata}
data1_varx <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = 10,
  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
  sim_dist = sim_varx_normal(2, 3, 0, 1, 1.5, 0)
) %>% unnest(data)

data2_varx <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = 50,
  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
  sim_dist = sim_varx_normal(2, 3, 0, 1, 10, 0)
) %>% unnest(data)

# make-quantiles
data_q <- bind_rows(data1_varf = 
compute_quantiles(data1_varf, 
                                 quantile_prob = quantile_prob_val) ,

 data2_varf = compute_quantiles(data2_varf,
                                 quantile_prob = quantile_prob_val),

 data1_varall = compute_quantiles(data1_varall,
                                 quantile_prob = quantile_prob_val) ,

 data2_varall = compute_quantiles(data2_varall,
                                 quantile_prob = quantile_prob_val),

 data1_varx = compute_quantiles(data1_varx,
                                 quantile_prob = quantile_prob_val) ,

 data2_varx = compute_quantiles(data2_varx,
                                 quantile_prob = quantile_prob_val),

.id = "data_type"
)

# Js-pairwise

data_q_wide <- data_q %>% 
  pivot_wider(names_from = data_type,
              values_from = sim_data_quantile) %>% 
  select(-c(1,2))

ndata <- data_q %>% distinct(data_type) %>% pull(data_type)
ldata <- length(ndata)
lcomb <-  nx_val*nfacet_val

dist_data <- map(1:ldata, function(x){ # first data
  map(1:ldata, function(y){ # 2nd data
    map(1:lcomb, function(z){ # number of combinations nx*nfacet
    JS(
        prob = quantile_prob_val,
        unlist(data_q_wide[z,x]),
        unlist(data_q_wide[z,y])
      ) %>% as_tibble()
  })%>% bind_rows(.id = "combinations")
})%>% bind_rows(.id = "data_type1")
}) %>% bind_rows(.id = "data_type2")

# hc

dist_mat <- dist_data %>% 
  group_by(data_type1, data_type2) %>% 
  summarise(dist = sum(value)) %>% 
  pivot_wider(names_from = data_type2,
              values_from = dist) %>% 
  ungroup() %>% 
  select(-1)

rownames(dist_mat) <- ndata
d = stats::as.dist(dist_mat)

hc = stats::hclust(d,method="complete")
plot(hc)

# mds

mds <- d %>%
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")

mds <- mds %>%
  mutate(data_type = ndata) %>%
  select((data_type)
         , Dim.1, Dim.2)

groups<-cutree(hc, k=3)

all_data_cluster <- cbind(dist_mat, groups) %>%
  cbind(mds)%>%
  mutate(groups = as.factor(groups))

mds_plot <- ggpubr::ggscatter(all_data_cluster,
                        x = "Dim.1",
                        y = "Dim.2",
                        label = mds$data_type,
                        color = "groups",
                        repel = TRUE,
                        size = 2,
                        shape = 1) 
mds_plot

```

<!-- # Repeat 4 designs each 10 times and see if there are 4 clusters -->

<!-- ```{r} -->

<!-- sample_seed <- seq(10, 100, 10) -->

<!-- data_varall <- map(1:10, function(x){ -->
<!-- set.seed(sample_seed[x])   -->
<!-- sim_panel_varall <- sim_panel( -->
<!--   nx = nx_val, nfacet =  nfacet_val, -->
<!--   ntimes = 100, -->
<!--   # sim_dist = sim_varall_normal(2, 3, 5, 10, 5, -1.5) -->
<!--   sim_dist = sim_varall_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val) -->
<!-- ) %>% unnest(data) -->
<!-- }) %>% bind_rows(.id = "seed_id") %>%  -->
<!--   mutate(date_type = "varall") -->

<!-- change_index(data_varall, ntimes_val = 5, nx_val = 2, nfacet_val = 3) -->

<!-- ``` -->

### 2021-07-16 realdata-cust10-noiter-scaled.Rmd
adding results from new and old method

--- 2021-07-08 supplementary/realdata-cust10-noiter.Rmd
@@
 ```{r}
-sm <- gravitas::smart_meter10 
-```
 
-```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for 10 households (facets).
-The data is squeezed stopping us from seeing any behavioral patterns."}
-library(lubridate)
+elec_ts <- read_rds("../data/elec_ts.rds")
 
-sm %>% 
-  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
-  geom_line() +
-  facet_wrap(~customer_id, ncol = 1, 
-             strip.position = "right") +
-  theme_bw() +
-  xlab("time") +
-  ylab("demand (in Kwh)")
-```
+set.seed(12345)
 
-```{r raw-sub, fig.cap="The raw plots for ten customers shown for Sept 2013-Oct 2013.
-The data is zoomed in and the y-scales made free to emphasize weekly, hourly or any behavior/patterns that they might have.
-There is some daily and weekly pattern in most households."}
-sm %>% 
-  filter(year(reading_datetime) == 2013 &
-           month(reading_datetime) %in% c(9, 10)) %>% 
-  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
-  geom_line() +
-  facet_wrap(~customer_id, ncol = 1, scales = "free_y", 
-             strip.position = "right") +
-  theme_bw()+
-  ylab("demand (in Kwh)") +
-  theme(panel.grid.major.x = element_blank()) +
-  scale_x_datetime("Date", date_labels = "%b %d",
-                   breaks = "1 week",
-                   date_minor_breaks = "1 day")  + 
-  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
-        panel.grid.minor.x =  element_line(colour = "#D3D3D3"))
+sm_100 <- elec_ts %>%
+  as_tibble() %>% 
+  distinct(customer_id) %>%
+  slice_sample(n = 100)
 
+sm <- elec_ts %>% 
+  filter(customer_id %in% sm_100$customer_id) %>% 
+  ungroup()
+
+sm_scaled <- sm %>%
+  group_by(customer_id) %>% 
+  dplyr::mutate(general_supply_kwh = scale(general_supply_kwh)) %>% 
+  ungroup()
 ```
@@
 ```{r hod}
-sm_hod <- gravitas::create_gran(sm, "hour_day") %>% 
+sm_hod <- gravitas::create_gran(sm_scaled, "hour_day") %>% 
   mutate(customer_id = as.integer(customer_id))
@@
 The y-scales are made free to allow us to see daily patterns for each households.
-The deciles for all households tend to show an increase in the morning and evening hours, although the hours differ.
-Except for the 90th decile, the deciles for the 4th household look pretty flat."}
+The deciles for all households follow different shape as a result of different customer behavior."}
 data_heatmap %>% 
@@
   pivot_wider(names_from = categories_serial_id, values_from = data)
+
+sm_hod_quantiles %>% 
+  filter()
+
 ```
@@
 plot(hc, cex = 0.6)
-rect.hclust(hc, k = 7, border = 2:8)
+rect.hclust(hc, k = 5, border = 2:8)
 ```
@@
 quantile_prob_val = c(0.25, 0.5, 0.75)
+
+sm_hod <- gravitas::create_gran(sm, "hour_day") %>% 
+  mutate(customer_id = as.integer(customer_id))
+
+sm_hod_list <- sm_hod %>% 
+  as_tibble() %>% 
+  select(customer_id, hour_day, general_supply_kwh) %>% 
+  pivot_wider(names_from = hour_day,
+              values_from = general_supply_kwh)
 

### 2021-07-16 supplementary/algo2-elec8.Rmd
adding results from new and old method

--- 2021-07-16 algo2-elec16.Rmd
@@
 ---
-title: "algo2-elec16"
+title: "Algorithm using wpd values for clustering 8 households"
 author: "Sayani Gupta"
-date: "15/07/2021"
+date: "13/07/2021"
 output: pdf_document
@@
 ##----harmony-all
-elec_harmony_8local <- read_rds("../paper-hakear/paper/data/elec_harmony_all.rds")
-
-elec_harmony_new8 <- 
-read_rds("data/elec_harmony_100-gap.rds") %>% 
-  bind_rows(.id = "id") %>% 
-  mutate(id = as.numeric(id)) %>% 
-  mutate(id = 8 + id) %>% 
-  mutate(facet_variable = case_when(
-    facet_variable == "hour_day" ~ "hod" ,
-    facet_variable == "day_month" ~ "dom" ,
-    facet_variable == "day_week" ~ "dow" ,
-    facet_variable == "week_month" ~ "wom" ,
-    facet_variable == "wknd_wday" ~ "wdwnd"
-  )) %>% 
-  mutate(x_variable = case_when(
-    x_variable == "hour_day" ~ "hod" ,
-    x_variable == "day_month" ~ "dom" ,
-    x_variable == "day_week" ~ "dow" ,
-    x_variable == "week_month" ~ "wom" ,
-    x_variable == "wknd_wday" ~ "wdwnd"
-  )) %>% 
-  mutate(id = paste("id", id, sep = " ")) %>% 
-  group_by(id) %>% 
-  mutate(rank = row_number())
-
-elec_harmony_all <- bind_rows(elec_harmony_8local, elec_harmony_new8)
+elec_harmony_all <- read_rds("../../paper-hakear/paper/data/elec_harmony_all.rds")
 
@@
 pivot_wider(names_from = id, values_from = low_proxy) %>% 
-mutate(score = rowMeans(across(`id 1`:`id 16`))) %>% 
+mutate(score = rowMeans(across(`id 1`:`id 8`))) %>% 
 filter(score !=0)
@@
 hc = stats::hclust(dist(data_clust[-1]),method="complete")
-plot(hc)
+#plot(hc)
 
@@
 
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = koptimal, border = 2:10)
-
-groups <- cutree(hc, k=8) %>% as_tibble() %>% 
+groups <- cutree(hc, k=4) %>% as_tibble() %>% 
 mutate(id = data_clust$id)
@@
   )+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90)) +
+  ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 0)) +
   theme(legend.position = "bottom") +
@@
   ylab("wpd")
-parcoord
 
@@
 
-elec_8local <- read_rds(here("../paper-hakear/paper/data/elec_all-8.rds")) %>% 
+elec <- read_rds(here("../paper-hakear/paper/data/elec_all-8.rds")) %>% 
   dplyr::filter(date(reading_datetime) >= ymd("20190701"), date(reading_datetime) < ymd("20191231"), meter_id==1) %>% 
@@
 
-elec_ts <- read_rds("data/elec_ts.rds")
+elec_zoom <-  elec %>%
+  as_tibble() %>% 
+  dplyr::filter(date(date_time) > as.Date("2019-08-01") & date(date_time) < (as.Date("2019-09-30"))) %>%
+  ggplot(aes(x=date_time, y = kwh)) +
+  #geom_point(size = 0.1, colour = "black", alpha = 0.3) +
+  geom_line(size = 0.1, colour = "blue") +
+  facet_wrap(~id, 
+             scales = "free_y",
+             ncol = 1,
+             strip.position =  "right") + 
+  xlab("Time [30m]") + 
+  theme_grey() + 
+  ylab("Energy demand (in Kwh)") + ggtitle("(b)") +
+  theme(panel.grid.major.x = element_blank()) +
+  scale_x_datetime("Date", date_labels = "%b %d",
+                   breaks = "1 week",
+                   date_minor_breaks = "1 day")  + theme_bw() +
+  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
+        panel.grid.minor.x =  element_line(colour = "#D3D3D3")) +
+  theme(
+    strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
 
-set.seed(12345)
+plot(hc, cex = 0.6)
+rect.hclust(hc, k = 4, border = 2:7)
 
-sm_100 <- elec_ts %>%
-  as_tibble() %>% 
-  distinct(customer_id) %>%
-  slice_sample(n = 100) %>% 
-  head(8)
+elec_zoom
+parcoord
 
-elec_sgsc <- filter(elec_ts, customer_id %in% sm_100$customer_id) %>% 
-  ungroup()
-
-sm <- bind_rows(elec_8local,
-                elec_sgsc)
-
-sm_hod_dow <- sm %>% 
+sm_hod_dow <- elec %>% 
 gravitas::create_gran("hour_day") %>% 
@@
   response = "kwh",
-  plot_type = "quantile", nrow=1)
+  plot_type = "quantile", nrow=1) +
+  scale_y_log10() +
+    ggtitle("Group 1")
 
@@
   response = "kwh",
-  plot_type = "quantile", nrow=1)
+  plot_type = "quantile", nrow=1) +
+  scale_y_log10()+
+    ggtitle("Group 2")
 
@@
   response = "kwh",
-  plot_type = "quantile", nrow=1)
+  plot_type = "quantile", nrow=1) +
+  scale_y_log10() +
+    ggtitle("Group 3")
 
@@
   filter(group==4) %>% 
-  gravitas::prob_plot(  "day_week",
+  gravitas::prob_plot("day_week",
      "hour_day",
   response = "kwh",
-  plot_type = "quantile", nrow=1) 
+  plot_type = "quantile", nrow=1) +
+  scale_y_log10() +
+    ggtitle("Group 4")
 ```

### 2021-07-16 supplementary/realdata-cust100-noiter-scaled.Rmd
adding results from new and old method

--- 2021-07-16 realdata-cust10-noiter-scaled.Rmd
@@
 
-```{r hod-deciles, fig.cap = "The deciles plots for 10 households are shown for entire observation period across hours of the day.
-The y-scales are made free to allow us to see daily patterns for each households.
-The deciles for all households follow different shape as a result of different customer behavior."}
-data_heatmap %>% 
-  ggplot() + geom_line(aes(x=category, 
-                           y = quantile_val, 
-                           color = quantile_prob)) + 
-  facet_wrap(~customer_serial_id, 
-             ncol = 1,
-             scales="free_y") + 
-  scale_color_brewer(palette = "Blues") + 
-  #scale_y_log10() +
-  theme_bw() +
-  xlab("hours of the day") +
-  ylab("demand (in Kwh)") +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
-```
-
 ```{r hod-list}
@@
 ```{r hod-dist}
+
+sm_filter_nozerosum <- sm_hod_quantiles %>% 
+  pivot_longer(-c(1, 2), names_to = "quantiles", values_to = "quantvalue") %>% 
+  group_by(customer_serial_id) %>% 
+  summarise(m = sum(quantvalue)) %>% 
+  filter(m!=0)
+
+sm_hod_quantiles <- sm_hod_quantiles %>%
+  filter(customer_serial_id %in% sm_filter_nozerosum$customer_serial_id)
+
 sm_dist_data <- sm_hod_quantiles %>% 
@@
   pivot_wider(names_from = categories_serial_id, values_from = data)
-
-sm_hod_quantiles %>% 
-  filter()
 

### 2021-07-16 supplementary/realdata-cust100-noiter.Rmd
adding results from new and old method

--- 2021-07-16 supplementary/realdata-cust100-noiter-scaled.Rmd
@@
 ---
-title: "Clustering real data with ten customers without iterations"
+title: "Clustering real data with 100 customers (no missing value)"
 output:
@@
     toc: false
+    cache: TRUE
 ---
@@
 
-# Plots of raw data
+```{r read-data}
+# elec_ts <- read_rds("../data/elec_ts.rds")
+# 
+# set.seed(12345)
+# 
+# sm_100 <- elec_ts %>%
+#   as_tibble() %>% 
+#   distinct(customer_id) %>%
+#   slice_sample(n = 100)
+# 
+# sm <- elec_ts %>% 
+#   filter(customer_id %in% sm_100$customer_id) %>% 
+#   ungroup()
 
-Five households are considered for the following analysis from the SGSC data set.
-The data sets contain three columns `customer_id`, `reading_datetime` and `general_supply_kwh`.
-They consist of half-hourly data from 2012-2014.
-Figure \ref{fig:raw-all} shows how the raw time plots for these five households look like.
-Since all the dataset looks squeezed on this linear time scale, Figure \ref{fig:raw-sub} shows how the raw plot looks for 2 months (Sept 2013 and Oct 2013).
+#elec_nogap <- read_rds("../data/elec_nogap.rds")
+```
 
-```{r}
+```{r read100}
+# 
+# set.seed(12345)
+# 
+# sm_100 <- elec_nogap %>%
+#   as_tibble() %>%
+#   distinct(customer_id) %>%
+#   slice_sample(n = 100)
+# 
+# sm <- elec_nogap %>%
+#   filter(customer_id %in% sm_100$customer_id) %>%
+#   ungroup()
+# write_rds(sm, "../data/elec_nogap-100cust.rds")
 
-elec_ts <- read_rds("../data/elec_ts.rds")
-
-set.seed(12345)
-
-sm_100 <- elec_ts %>%
-  as_tibble() %>% 
-  distinct(customer_id) %>%
-  slice_sample(n = 100)
-
-sm <- elec_ts %>% 
-  filter(customer_id %in% sm_100$customer_id) %>% 
-  ungroup()
-
-sm_scaled <- sm %>%
-  group_by(customer_id) %>% 
-  dplyr::mutate(general_supply_kwh = scale(general_supply_kwh)) %>% 
-  ungroup()
+sm <- read_rds("../data/elec_nogap-100cust.rds")
 ```
@@
 ```{r hod}
-sm_hod <- gravitas::create_gran(sm_scaled, "hour_day") %>% 
-  mutate(customer_id = as.integer(customer_id))
+sm_hod <- gravitas::create_gran(sm, "hour_day") 
 ```
@@
 
-```{r}
+```{r heatmpa-data}
 quantile_prob_val = seq(0.1, 0.9, 0.1)
@@
 
-```
-
-```{r hod-dist}
-
 sm_filter_nozerosum <- sm_hod_quantiles %>% 
@@
 
+```
+
+```{r hod-dist}
 sm_dist_data <- sm_hod_quantiles %>% 
@@
   select(-quantile_prob) %>% 
-  group_by(customer_serial_id, categories_serial_id) %>%
+  group_by(customer_serial_id, 
+           categories_serial_id) %>%
   nest() %>% 
-  pivot_wider(names_from = categories_serial_id, values_from = data)
-
+  pivot_wider(names_from = categories_serial_id, 
+              values_from = data) %>% 
+  ungroup()
 ```
@@
 dist_mat <- dist_data %>%
-  mutate(data_type1 = as.numeric(data_type1), data_type2 = as.numeric(data_type2)) %>% 
+  mutate(data_type1 = as.numeric(data_type1),
+         data_type2 = as.numeric(data_type2)) %>% 
   group_by(data_type1, data_type2) %>%
@@
 
+lookup_tbl <- sm_hod_list %>% 
+  mutate(customer_serial_id = row_number()) %>% 
+  select(customer_id, customer_serial_id) %>% 
+  filter(customer_serial_id %in% sm_filter_nozerosum$customer_serial_id)
+
 rownames(dist_mat_format) <- lookup_tbl$customer_id
@@
 # Clustering results
-
-The following figures show dendograms when we are using optimal number of clusters (k) from k = fpc::nselectboot(), k = 2, 3, 4 and 5.
-We observe that the $4^{th}$ household tend to get split as different clusters as increase the number of clusters.
 
@@
 koptimal = fpc::nselectboot(as.dist(dist_mat_format),
-                        B = 50,
+                        B = 500,
                         method = "complete",
@@
                         classification = "averagedist",
-                        krange = 2:10)
+                        krange = 2:15)
 
@@
   cbind(mds)%>%
-  mutate(groups = as.factor(groups)) %>% as_tibble()
+  mutate(groups = as.factor(groups)) %>% 
+  as_tibble()
 
@@
   theme_classic()+
-  theme(legend.position = "bottom") +
-  scale_color_brewer(palette = "Dark2")
+  theme(legend.position = "bottom")
 
@@
 
-_k = 7_
+_k = 15_
 
@@
 plot(hc, cex = 0.6)
-rect.hclust(hc, k = 5, border = 2:8)
+rect.hclust(hc, k = 15, border = 2:20)
 ```
@@
 "}
-quantile_prob_val = c(0.25, 0.5, 0.75)
-
-sm_hod <- gravitas::create_gran(sm, "hour_day") %>% 
-  mutate(customer_id = as.integer(customer_id))
-
-sm_hod_list <- sm_hod %>% 
-  as_tibble() %>% 
-  select(customer_id, hour_day, general_supply_kwh) %>% 
-  pivot_wider(names_from = hour_day,
-              values_from = general_supply_kwh)
+quantile_prob_val = c(0.1, 0.25, 0.5, 0.75, 0.9)
 
 groups_ref <- groups %>% as_tibble %>% 
-              mutate(customer_serial_id = sm_hod_list$customer_id) %>% 
+              mutate(customer_id = lookup_tbl$customer_id,
+                     customer_serial_id = lookup_tbl$customer_serial_id) %>% 
              rename ("group" = "value")
 
-sm_hod_list_cat <-  sm_hod_list %>% left_join(groups_ref, by = c("customer_id" = "customer_serial_id")) %>% select(group, everything())
+groups_g3 <- groups_ref %>% 
+  group_by(group) %>% 
+  count() %>% 
+  filter(n>=3)
 
-ncol_sm <- seq_len(ncol(sm_hod_list_cat[-c(1, 2)]))
+sm_hod_list_cat <- sm_hod_list %>% left_join(groups_ref) %>%
+  filter(customer_serial_id %in% sm_filter_nozerosum$customer_serial_id) %>% 
+select(group, customer_serial_id, everything())
+
+ncol_sm <- seq_len(ncol(sm_hod_list_cat[-c(1, 2, 3)]))
 nrow_sm <- unique(sm_hod_list_cat$group)
@@
      dplyr::filter(group == x) %>% 
-     select(-c(1, 2)) %>% 
+     select(-c(1, 2, 3)) %>% 
      extract(y) %>% 
@@
 
-categories_ref_cat <- names(sm_hod_list_cat)[-c(1, 2)] %>% as_tibble() %>% 
+categories_ref_cat <- names(sm_hod_list_cat)[-c(1, 2, 3)] %>% as_tibble() %>% 
   mutate(categories_serial_id = row_number())
@@
 
-# data_heatmap%>% 
-# ggplot(aes(x= as.factor(group_id), y = as.factor(quantile_prob))) +
-#   geom_tile(aes(fill = quantile_val)) +
-#   scale_fill_distiller(palette = "YlGnBu")+ 
-#   facet_wrap(~category) +
-#   theme(axis.text.x = element_text(angle = 90)) +
-#   ylab("deciles") +
-#   xlab("customers")
 ```
@@
 data_heatmap %>% 
+  filter(group_id %in% groups_g3$group) %>% 
   pivot_wider( names_from = quantile_prob, 
                values_from = quantile_val) %>% 
-  ggplot(aes(x = category)) +  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") +
+  ggplot(aes(x = category)) +  
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`), fill = "#D55E00", alpha = 0.8) +
+  geom_ribbon(aes(ymin = `10%`, 
+                  ymax = `90%`), fill = "#56B4E9", alpha = 0.4) + 
   geom_line(aes(y = `50%`)) +
-  facet_wrap(~group_id, ncol = 1, scales = "free_y", labeller = "label_both") +
+  facet_wrap(~group_id, ncol = 2,
+             scales = "free_y",
+             strip.position = "right") +
   theme_bw() +

### 2021-07-16 algo2.R

@@ -1,0 +1,114 @@
+library(knitr)
+library(tidyverse)
+library(lubridate)
+library(lvplot)
+#library(ggridges)
+library(tsibble)
+library(gravitas)
+library(ggpubr)
+library(readr)
+library(kableExtra)
+library(distributional)
+library(ggplot2)
+library(sugrrants)
+library(here)
+library(ggplot2)
+library(patchwork)
+library(scales)
+library(GGally)
+library(viridis)
+
+# take customers who do not have gaps in their data
+elec_nogap <- read_rds("data/elec_nogap.rds")
+
+set.seed(12345)
+
+# just take 100 customers from them
+sm_100 <- elec_nogap %>%
+  as_tibble() %>% 
+  distinct(customer_id) %>%
+  slice_sample(n = 100)
+
+# elec <- elec_ts %>%
+#   as_tibble() %>% 
+#   distinct(customer_id) %>%
+#   slice_sample(n = 2)
+
+# take data for those customers only for 2013
+elec <- elec_nogap %>% 
+  filter(customer_id %in% sm_100$customer_id) %>% 
+  ungroup() %>% 
+  filter(year(reading_datetime)==2013)
+
+# consider harmonies which are generally significant for 
+
+harmonies <- read_rds("../paper-hakear/paper/data/harmonies.rds")
+
+harmonies_acro <- harmonies %>% 
+  mutate(facet_variable = case_when(
+    facet_variable == "hour_day" ~ "hod" ,
+    facet_variable == "day_month" ~ "dom" ,
+    facet_variable == "day_week" ~ "dow" ,
+    facet_variable == "week_month" ~ "wom" ,
+    facet_variable == "wknd_wday" ~ "wdwnd"
+  )) %>% 
+  mutate(x_variable = case_when(
+    x_variable == "hour_day" ~ "hod" ,
+    x_variable == "day_month" ~ "dom" ,
+    x_variable == "day_week" ~ "dow" ,
+    x_variable == "week_month" ~ "wom" ,
+    x_variable == "wknd_wday" ~ "wdwnd"
+  )) %>% mutate(comb = paste(facet_variable, x_variable, sep = "-")) %>% 
+  filter(comb %in% c("hod-wdwnd", "dom-hod", "wdwnd-hod", "hod-dow", "dow-hod")) %>% 
+  select(-comb)
+
+# take customers who do not have gaps in their data
+
+library(tictoc)
+#tic()
[…]

### 2021-07-16 job-algo2.R

@@ -1,0 +1,85 @@
+library(knitr)
+library(tidyverse)
+library(lubridate)
+library(lvplot)
+#library(ggridges)
+library(tsibble)
+library(gravitas)
+library(ggpubr)
+library(readr)
+library(kableExtra)
+library(distributional)
+library(ggplot2)
+library(sugrrants)
+library(here)
+library(ggplot2)
+library(patchwork)
+library(scales)
+library(GGally)
+library(viridis)
+
+
+ncust = 100 # number of sampled customers
+nperm = 2 # number of permutations for normalization
+nsamp = 2 # number of samples for threshold
+
+# take customers who do not have gaps in their data
+# elec_nogap <- read_rds("data/elec_nogap.rds")
+# 
+# elec_nogap_2013 <- elec_nogap %>%
+#   filter(year(reading_datetime)==2013)
+#   
+# write_rds(elec_nogap_2013, "data/elec_nogap_2013.rds")
+
+# elec_nogap_2013 <- read_rds("data/elec_nogap_2013.rds")
+
+# just take 100 customers from them
+# sm_100 <- elec_nogap_2013 %>%
+#   as_tibble() %>%
+#   distinct(customer_id) %>%
+#   slice_sample(n = ncust)
+# 
+# 
+# # take data for those customers only for 2013
+# elec <- elec_nogap_2013 %>%
+#   filter(customer_id %in% sm_100$customer_id) %>%
+#   ungroup()
+# 
+# write_rds(elec, "data/elec_nogap_2013_100.rds")
+
+elec <- read_rds("data/elec_nogap_2013_100.rds")
+cust_id <- length(unique(elec$customer_id))
+#scen <- as.numeric(commandArgs()[[6]])
+scen <- 2
+custj <- cust_id[scen]
+# consider harmonies which are generally significant for electricity data
+
+harmonies <- read_rds("../paper-hakear/paper/data/harmonies.rds")
+
+# harmonies <- harmonies %>% 
+#   mutate(comb = paste(facet_variable, 
+#                              x_variable, 
+#                              sep = "-")) %>% 
+#   filter(comb %in% c("hour_day-wknd_wday", "day_month-hour_day", "wknd_wday-hour_day", "hour_day-day_week", "day_week-hour_day")) %>% 
+#   select(-comb)
+# 
+# write_rds(harmonies, "data/harmonies.rds")
+
+harmonies <- read_rds("data/harmonies.rds")
+
+# take customers who do not have gaps in their data
+  
+  data_id <-  elec %>% 
[…]

### 2021-07-16 old/makedata01.R

@@ -1,0 +1,263 @@
+# harmony table considered for all
+
+#This script calculates raw mmpd for each simulation scenario
+##Read Simulation Table
+
+.libPaths(c("~/R/libs", .libPaths()))
+library(distributional)
+library(readr)
+library(drake)
+library(tidyverse)
+library(hakear)
+library(gravitas)
+library(parallel)
+library(here)
+
+set.seed(9999)
+
+# parameter set
+nsamp = 100 # number of samples based on which thresholds are decided
+ncust = 6 # number of customers for trial
+nrep = 3 # number of error added usage for each customer
+kstep1 = 3 # number of clusters in step1/ only step
+kstep2 = 2 # number of clusters in step2 (this will actually depend on the number of observations in the cluster after step1)
+
+
+# change path while running it on HPC
+simtable<-read_csv(here('sim_table/sim_table.csv'))
+
+# <needed>
+harmonies<-read_csv(here('sim_table/harmonies.csv'))
+
+sm_cust_data <- read_rds(here("data/sm_cust_data.rds"))
+
+sm_cust_data_list <- sm_cust_data %>% group_split(customer_id)
+
+# Add some error distribution to each of them before trying out the clustering process
+
+# function to generate some error to the usage for each customer
+# number of times to repeat
+# names of customer id when repeated
+
+add_error_unif <- function(x){
+ y = runif(nrow(x), 0.01, 0.05)
+ x$general_supply_kwh + y
+}
+
+# generate nrep times distribution for one household
+
+one_house_nrep <- function(x, nrep = 3)
+{
+  onerep = lapply(1:nrep, function(y){
+    bind_cols(customer_id  =  paste(x$customer_id, y, sep = "-"),
+              reading_datetime =  x$reading_datetime,
+              general_supply_kwh = rep(add_error_unif(x), 1))
+  })
+    onerep
+}
+
+
+# generate nrep times distribution for many households
+
+many_house_nrep <- lapply(sm_cust_data_list, function(x, nrep = 3){
+  one_house_nrep(x, nrep = nrep)
+}
+  ) %>% bind_rows
+
+
+all_house_nrep_orig <- bind_rows(sm_cust_data, many_house_nrep)
+
+
+saveRDS(all_house_nrep_orig, "data/all_house_nrep_orig.rds") # seed and dist not included yet
+
+
+
+all_house_nrep_orig_list <- all_house_nrep_orig %>%
[…]

### 2021-07-16 old/simdata_cluster02.R

@@ -1,0 +1,7 @@
+# generate 4 series, two from design varx and another two from varf and see if clustering happens properly through the approaches
+
+
+# 2 time series from varx
+
+
+# 2 time series from varf

### 2021-07-16 script/smart_meter.R

@@ -37,6 +37,14 @@ readr::write_rds(elec_ts, "data/elec_ts.rds")
 elec_ts <- read_rds("data/elec_ts.rds")
 gap_df <- has_gaps(elec_ts)
 
+nogap <- gap_df %>% filter(.gaps==FALSE)
+
+
+elec_nogap <- elec_ts %>% 
+  filter(customer_id %in% nogap$customer_id)
+
+readr::write_rds(elec_nogap, "data/elec_nogap.rds")
+
 # sum(gap_df$.gaps) / NROW(gap_df)
 
 ## ---- count-gaps

### 2021-08-05 algo2-100.Rmd
big file push blunder

---
title: "Untitled"
author: "Sayani Gupta"
date: "22/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
library(readr)
library(rlang)
library(stringr)
library(purrr)
library(tidyr)
library(dplyr)

  all_files = list.files(path = paste0("data/algo2/"), 
                         pattern = ".rds")

  names_levels <- map_dfr(all_files, 
                          function(x){
                            z = str_split(str_remove(x, "-nogap.rds"), "-") %>% 
                              unlist()
                            bind_cols(customer_serial_id = z[2]
                                      )
                          })

  all_files_path <- paste0("data/algo2/",
                           all_files)

  all_data <- lapply(1:length(all_files_path), function(x){

    data = all_files_path %>% magrittr::extract2(x) %>% 
      readRDS()

    names = names_levels %>% magrittr::extract(x,)

    #names_rep =   names %>% slice(rep(1:n(), each = nrow(data)))
    bind_cols(names, data)

  }) %>% bind_rows()

  write_rds(all_data, paste0("data/algo2-rawdata-wpd.rds"))
```

### 2021-08-05 sim_setup_algo2.Rmd
big file push blunder

---
title: "Simulation with algorithm 2"
author: "Sayani Gupta"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,echo=FALSE, message=FALSE, warning = FALSE)
library(patchwork)
library(distributional)
library(ggplot2)
library(tidyverse)
library(knitr)
library(readr)
library(GGally)
```

# _Simulation design_

Three circular granularities g1, g2 and g3 are considered with levels 2, 3 and 4 respectively. Many time series with 600 observations are created using the five designs below, each of which is iterated four times. We anticipate to have five clusters, each with four time series conforming to the same design, once we execute the clustering.

# _Algorithm_
Compute wpd for each granularity.The distance matrix is then computed with 20 time series as cases and three granularities as variables. Manhattan distances are computed and then hierarchical clustering is applied on them.

```{r tab-gran}
design = c("design-1", "design-2", "design-3", "design-4", "design-5")
g1 = c("no", "yes" , "no", "yes", "yes")
g2 = c("no", "no" , "yes", "no", "yes")
g3 = c("no", "yes" , "no", "yes", "yes")
table <- tibble(design, g1, g2, g3) %>% kable()
table
```

```{r}
t <- seq(0,299, 1)

generate_design <-  function (t = NULL, 
                           mu = NULL){
t <- seq(0, t, 1)
g1 <- t %%2
g2 <- t %%3
g3 <- t %%4

# design2 (change across g2 only)
mu = mu
len_mu <- length(mu)
len_obs <-  length(t)/length(mu)
dist <- dist_normal(mu = mu, sigma = 1)
numbers <- generate(dist, times = floor(len_obs))

timeseries <- numbers%>% unlist() %>% matrix(ncol = len_mu, nrow = len_obs, byrow = FALSE) %>% t() %>% 
  matrix(ncol=1, byrow = TRUE)

timeseries <- as_tibble(timeseries) %>% 
  set_names("value") %>%
  bind_cols(t %>% as_tibble()) %>% 
  set_names(c("value", "t"))

 data <- tibble::tibble(t, g1, g2, g3) %>% left_join(timeseries)
 data
}

set.seed(12345)
data_design_1 <- generate_design(t = 599, mu = 0)
data_design_2 <- generate_design(t = 599, mu = c(0, 2))
data_design_3 <- generate_design(t = 599, mu = c(0, 2, 4))
data_design_4 <- generate_design(t = 599, mu = c(0, 2, 4, 6))
data_design_5 <- generate_design(t = 599, mu = c(0, 2, 4, 6, 8, 10))
```

<!-- # Design1 -->
```{r design-null, eval = FALSE}
p1 <- ggplot(data_design_1, aes(x = t, y = value)) + geom_line() + xlab("time")  + theme_bw() 
p2 <- ggplot(data_design_1, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw()
p3 <- ggplot(data_design_1, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw() 
p4 <- ggplot(data_design_1, aes(x = as.factor(g3), y = value)) + geom_boxplot() +
  xlab("g3") + theme_bw()

(p1 /(p2 + p3 + p4)) 
```

<!-- # Design2 -->
```{r design1, eval = FALSE}
p1 <- ggplot(data_design_2, aes(x = t, y = value)) + geom_line() + xlab("time")  + theme_bw()
p2 <- ggplot(data_design_2, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw()
p3 <- ggplot(data_design_2, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw()
p4 <- ggplot(data_design_2, aes(x = as.factor(g3), y = value)) + geom_boxplot() +
  xlab("g3") + theme_bw()

(p1 /(p2 + p3 + p4)) 
```

<!-- # Design3 -->

```{r design2, eval = FALSE}
p1 <- ggplot(data_design_3, aes(x = t, y = value)) + geom_line() + xlab("time")  + theme_bw()
p2 <- ggplot(data_design_3, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw()
p3 <- ggplot(data_design_3, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw()
p4 <- ggplot(data_design_3, aes(x = as.factor(g3), y = value)) + geom_boxplot() +
  xlab("g3") + theme_bw()

(p1 /(p2 + p3 + p4)) 
```

<!-- # Design4 -->

```{r design3, eval = FALSE}
p1 <- ggplot(data_design_4, aes(x = t, y = value)) + geom_line() + xlab("time")  + theme_bw()
p2 <- ggplot(data_design_4, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw()
p3 <- ggplot(data_design_4, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw()
p4 <- ggplot(data_design_4, aes(x = as.factor(g3), y = value)) + geom_boxplot() +
  xlab("g3") + theme_bw()

(p1 /(p2 + p3 + p4))

```

# Raw plots
```{r all-design}
bind_data <- bind_rows(data_design_1, data_design_2, data_design_3, data_design_4,data_design_5, .id = "design") %>% mutate(design = paste("design", design, sep = "-"))

p1 <- ggplot(bind_data, aes(x = t, y = value)) + geom_line() + xlab("time")  + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0))) 
p2 <- ggplot(bind_data, aes(x = as.factor(g1), y = value)) + geom_boxplot(width=0.5) + xlab("g1") + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+stat_summary(fun.y = median, geom="line", group= 1, color= "blue", size = 0.5)
p3 <- ggplot(bind_data, aes(x = as.factor(g2), y = value)) + geom_boxplot(width=0.5) + xlab("g2") + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+stat_summary(fun.y = median, geom="line", group= 1, color= "blue", size = 0.5)

p4 <- ggplot(bind_data, aes(x = as.factor(g3), y = value)) + geom_boxplot(width=0.5) +
  xlab("g3") + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0))) +stat_summary(fun.y = median, geom="line", group= 1, color= "blue", size = 0.5)

p1
```

# Designs (Distribution of simulated data across different granularities)

```{r, out.height="100%"}
(p2+p3+p4)

```

# Five iterations of each design (changing seeds)

```{r}
seed_len = seq(12345, 12348, 1)
bind_data_iter <- map(seq_len(length(seed_len)), function(x){
set.seed(seed_len[x])
data_design_2 <- generate_design(t = 599, mu = c(0, 2))
data_design_3 <- generate_design(t = 599, mu = c(0, 2, 4))
data_design_4 <- generate_design(t = 599, mu = c(0, 2, 4, 6))
data_design_1 <- generate_design(t = 599, mu = 0)
data_design_5 <- generate_design(t = 599, mu = c(0, 2, 4, 6, 8, 10))
bind_data <- bind_rows(data_design_1, data_design_2, data_design_3, data_design_4,data_design_5, .id = "design") %>% mutate(design = paste("design", design, sep = "-"))
}) %>% bind_rows(.id = "seed_id")

```

<!-- Cross-check -->
<!-- data = bind_data_iter %>% filter(design == "design-2", seed_id == 5) -->
<!-- p2 <- ggplot(data, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw() + facet_wrap(~design, ncol = 1) -->
<!-- p3 <- ggplot(data, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw() + facet_wrap(~design, ncol = 1) -->
<!-- p4 <- ggplot(data, aes(x = as.factor(g3), y = value)) + geom_boxplot() + -->
<!-- xlab("g3") + theme_bw() + facet_wrap(~design, ncol = 1) -->
<!-- (p2 + p3 + p4) -->

<!-- ```{r} -->
<!-- p2 <- ggplot(data, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw() -->
<!-- p3 <- ggplot(data, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw() -->
<!-- p4 <- ggplot(data, aes(x = as.factor(g3), y = value)) + geom_boxplot() + -->
<!-- xlab("g3") + theme_bw()  -->
<!-- (p2 + p3 + p4) -->
<!-- ``` -->

```{r}
harmony_tbl <- tibble(facet_variable = NA, x_variable = c("g1", "g2", "g3"),
                      x_levels = c(2, 3, 4), facet_levels = NA)
```

```{r wpd-table}
data <- bind_data_iter %>% mutate(data_id = paste(design, seed_id, sep = "-"))
bind_data_iter <- tsibble::as_tsibble(data, index = t, key = "data_id")

data_wpd <- map(seq_len(length(unique(bind_data_iter$data_id))),
    function(y){
      z = unique(bind_data_iter$data_id)
      data_sub <- bind_data_iter %>% filter(data_id==z[y])
        elec_select_harmony = hakear::select_harmonies(data_sub,
                               harmony_tbl = harmony_tbl,
                               response = value,
                               nperm = 200
  ) %>% mutate(data_id = unique(data_sub$data_id)) %>% unnest()
    }) %>% bind_rows()

write_csv(data_wpd, "data_wpd.rds")
```

# Table of wpd across designs and wpd

```{r data-wide}

data_wide <- data_wpd %>% select(x_variable, wpd, data_id) %>% 
  pivot_wider(names_from = x_variable, values_from = wpd)
data_wide %>% kable()

ggpairs_nocolor <- ggpairs(data_wide, columns = 2:4)

rownames(data_wide) = data_wide$data_id

hc = stats::hclust(dist(data_wide[-1], method = "manhattan"),method="complete")

library(fpc)
koptimal = fpc::nselectboot(dist(data_wide[-1], method="manhattan"),
                        B = 500,
                        method = "complete",
                        clustermethod = disthclustCBI, 
                        classification = "averagedist",
                        krange = 2:8)

koptimal = koptimal$kopt
```

# Clustering of designs

```{r}
  groups <- cutree(hc, k=5) %>% as_tibble() %>% 
mutate(data_id = data_wide$data_id) %>% rename("group" = "value")

  groups %>% kable()

data_wide_group <- data_wide %>% left_join(groups)  %>% 
  mutate(group_id = as.factor(group))

parcoord <- GGally::ggparcoord(data_wide_group,
                   columns = 2:ncol(data_wide),
                   groupColumn = "group_id",
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.8,
                   scale = "globalminmax", 
                   order = "anyClass"
) + ggplot2::theme_bw() +
  scale_color_brewer(palette = "Dark2") + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 0)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")
parcoord
ggpairs_nocolor
  # ggpairs_color <- ggscatmat(data_wide, columns=2:4, color = "group") +
  # scale_colour_brewer(palette="Set2")
```

# With optimal number of clusters

Design 2 and 4 form a part of the same cluster with the optimal number of clusters chosen as 4. Design 2 and 4 are similar in the sense that both of them changes across categories for g1 and g3, and remains constant for g2.

```{r}
  groups <- cutree(hc, k=koptimal) %>% as_tibble() %>% 
mutate(data_id = data_wide$data_id) %>% rename("group" = "value")

  groups %>% kable()

data_wide_group <- data_wide %>% left_join(groups)  %>% 
  mutate(group_id = as.factor(group))

parcoord <- GGally::ggparcoord(data_wide_group,
                   columns = 2:ncol(data_wide),
                   groupColumn = "group_id",
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.8,
                   scale = "globalminmax", 
                   order = "anyClass"
) + ggplot2::theme_bw() +
  scale_color_brewer(palette = "Dark2") + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 0)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")
parcoord
ggpairs_nocolor
  # ggpairs_color <- ggscatmat(data_wide, columns=2:4, color = "group") +
  # scale_colour_brewer(palette="Set2")
```

# PCA based clustering

```{r}
pc <- prcomp(data_wide_group[c(2, 3, 4)], center = F, scale = F)

# First two principal componentsafc
comp <- data.frame(pc$x[,1:2])

#plot(comp, pch=16, col=rgb(0,0,0,0.5))

hc = stats::hclust(dist(comp))
groups <- cutree(hc, k=5) %>% as_tibble() %>% bind_cols(comp) %>% 
  mutate(value = as.character(value))
library(RColorBrewer)
library(scales)
groups %>% ggplot(aes(x=PC1, y = PC2)) + geom_point(aes(color = value)) +
  scale_color_brewer(palette = "Dark2") + theme_bw()
```

### 2021-08-05 sim_setup_algo2_g3.Rmd
big file push blunder

--- 2021-08-05 sim_setup_algo2.Rmd
@@
 g2 <- t %%3
-g3 <- t %%4
+g3 <- rep_len(rep(c(0,1,2), each = 3), length.out = length(t))
 
@@
 data_design_3 <- generate_design(t = 599, mu = c(0, 2, 4))
-data_design_4 <- generate_design(t = 599, mu = c(0, 2, 4, 6))
-data_design_5 <- generate_design(t = 599, mu = c(0, 2, 4, 6, 8, 10))
+data_design_4 <- generate_design(t = 599, mu = c(0, 2, 4, 6, 8, 10))
+data_design_5 <- generate_design(t = 599, mu = c(0, 2, 0, 4, 2, 4, 4, 6))
 ```

### 2021-08-05 supplementary/algo2-elec100.Rmd
big file push blunder

---
title: "Algorithm using wpd values for clustering 100 households"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,echo=FALSE, message=FALSE, warning = FALSE)
```

```{r load-lib}
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
#library(ggridges)
library(tsibble)
library(gravitas)
library(ggpubr)
library(readr)
library(kableExtra)
library(distributional)
library(ggplot2)
library(sugrrants)
library(here)
library(ggplot2)
library(patchwork)
library(scales)
library(GGally)
library(viridis)
```

```{r allplot, message=FALSE, warning=FALSE, echo = FALSE}

##----harmony-all
elec_harmony_all <- read_rds("../data/algo2-rawdata-wpd.rds")

select_split <- str_split(elec_harmony_all$select_harmony, " ", simplify = TRUE)[,2]

elec_sig_split <- elec_harmony_all %>%
  mutate(facet_variable = case_when(
    facet_variable == "hour_day" ~ "hod" ,
    facet_variable == "day_month" ~ "dom" ,
    facet_variable == "day_week" ~ "dow" ,
    facet_variable == "week_month" ~ "wom" ,
    facet_variable == "wknd_wday" ~ "wdwnd"
  )) %>% 
  mutate(x_variable = case_when(
    x_variable == "hour_day" ~ "hod" ,
    x_variable == "day_month" ~ "dom" ,
    x_variable == "day_week" ~ "dow" ,
    x_variable == "week_month" ~ "wom" ,
    x_variable == "wknd_wday" ~ "wdwnd"
  )) %>% 
  bind_cols(select_split = select_split) %>% 
  mutate(significant = case_when(
    select_split == "***" ~ "highest",
    select_split == "**" ~ "high",
    select_split == "*" ~ "medium",
    select_split == "" ~ "low"
  )) %>%mutate(comb = paste(facet_variable, x_variable, sep = "-"))

# %>% 
#   mutate(rank = case_when(
#     select_split == "***" ~ paste(rank, "***", sep = " "),
#     select_split == "**" ~  paste(rank, "**", sep = " "),
#     select_split == "*" ~  paste(rank, "*", sep = " "),
#     select_split == "" ~  paste(rank, "", sep = " ")
#   ))

elec_sig_split$significant <- 
  factor(elec_sig_split$significant, levels = c("highest", "high", "medium", "low"))

data_comb_filter = elec_sig_split %>% 
mutate(low_proxy = if_else(significant=="low", 0, 1))%>% 
select(comb, customer_id, low_proxy) %>% 
pivot_wider(names_from = customer_id, values_from = low_proxy) %>% 
mutate(score = rowMeans(across(-1))) %>% 
filter(score !=0)

data_sig_filter = elec_sig_split %>% 
filter(comb %in% data_comb_filter$comb)

data_clust = data_sig_filter %>% 
select(comb, customer_id, wpd) %>% 
pivot_wider(names_from = comb, values_from = wpd)

hc = stats::hclust(dist(data_clust[-1], method = "manhattan"),method="complete")
#plot(hc)
plot(hc, cex = 0.6)
rect.hclust(hc, k = 8, border = 2:10)

library(fpc)
koptimal = fpc::nselectboot(dist(data_clust[-1], method = "manhattan"),
                        B = 50,
                        method = "complete",
                        clustermethod = disthclustCBI, 
                        classification = "averagedist",
                        krange = 2:8)

koptimal = koptimal$kopt

groups <- cutree(hc, k=8) %>% 
  as_tibble() %>% 
mutate(customer_id = data_clust$customer_id)

##----cluster-characterization

data = data_clust %>% 
left_join(groups)%>% 
             rename ("group" = "value") %>% 
             ungroup() %>% 
             select(group, customer_id, everything()) %>% 
             pivot_longer(-c(1,2), names_to = "comb", values_to = "wpd")

data_pcp <- data %>% pivot_wider(
              names_from = comb, 
              values_from = wpd) %>% 
  mutate(group = as.character(group)) 
```

```{r}
# ggpairs_nocolor <- ggpairs(data_pcp, columns = 3:7)
# ggpairs_nocolor

groupge2 <- data %>% 
  group_by(group) %>% 
  summarise(n = length(unique(customer_id))) %>% 
  filter(n>2)

data_pcp_ge2 <- data_pcp %>% filter(group %in% groupge2$group)

# ggpairs_nocolor <- ggpairs(data_pcp_ge2, columns = 3:7)

ggpairs_nocolor <- ggscatmat(data_pcp_ge2, columns=3:7) +
  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")

ggpairs_color <- ggscatmat(data_pcp_ge2, columns=3:7, color = "group") +
  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")

ggpairs_nocolor
```

```{r}
ggpairs_color
```

```{r}
# 
# data_order <- data %>% 
#   group_by(comb) %>% 
#   summarise(m = mean(wpd)) %>% 
#   arrange(desc(m))
#   
# data_median = data %>% 
#   group_by(group, comb) %>% 
#   summarize(wpd_med = median(wpd))
# 
# data_pcp <- data_median %>% 
# mutate(group = as.character(group)) %>% 
#   left_join(data_order) %>% 
#   arrange(desc(m)) %>% 
#   pivot_wider(-4, 
#               names_from = comb, 
#               values_from = wpd_med)

data_pcp <- data_pcp %>% ungroup()%>% filter(group %in% groupge2$group)
parcoord <- GGally::ggparcoord(data_pcp ,
                   columns = 3:ncol(data_pcp),
                   groupColumn = "group",
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.3,
                   order = "allClass",
                   scale = "globalminmax"
) + ggplot2::theme_bw() +
  scale_color_brewer(palette = "Dark2") + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd") 
# 
# ggparcoord(data_pcp, scale = "globalminmax", columns=3:ncol(data_pcp),groupColumn = "group", alphaLines = 0.8) + geom_boxplot(aes_string(group = "variable"), width = 0.1, alpha = 0.1,
#                  outlier.color = NA) +  scale_color_brewer(palette = "Dark2") 
#+
 # geom_point(data = data, aes(x=comb, y = wpd))

##----read-orig-data

elec <- read_rds("../data/elec_nogap_2013_100.rds") %>% 
  left_join(groups)%>% 
             rename ("group" = "value")

# plot(hc, cex = 0.6)
# rect.hclust(hc, k = 4, border = 2:7)

#elec_zoom
parcoord

elec_scaled <- elec %>%
    group_by(customer_id) %>%
    #dplyr::mutate(general_supply_kwh = scale(general_supply_kwh)) %>%
    ungroup()

sm_hod_dow <- elec_scaled %>% 
gravitas::create_gran("hour_day") %>% 
  gravitas::create_gran("day_week")

sm_hod_list <- sm_hod_dow %>% 
  as_tibble() %>% 
  arrange(group, day_week, hour_day) %>% 
  mutate(category = paste(hour_day, day_week, sep = "-")) %>% 
   select(group, category, general_supply_kwh) %>% 
  pivot_wider(names_from = category,
              values_from = general_supply_kwh)

elec %>% as_tibble() %>% group_by(group) %>% summarise(n = length(unique(customer_id)))

quantile_prob_val = c(0.25, 0.5, 0.75)

p1 <- sm_hod_dow %>% 
  filter(group==1) %>% 
  gravitas::prob_plot(
     "day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 1") + theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

p2 <- sm_hod_dow %>% 
  filter(group==2 ) %>% 
  gravitas::prob_plot(  "day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10()+
    ggtitle("Group 2")+ theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

p3 <- sm_hod_dow %>% 
  filter(group==3) %>% 
  gravitas::prob_plot(  "day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 3" )+ theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

p4 <- sm_hod_dow %>% 
  filter(group==4) %>% 
  gravitas::prob_plot("day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 4")+ theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

p5 <- sm_hod_dow %>% 
  filter(group==5) %>% 
  gravitas::prob_plot("day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 5")+ theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

p6 <- sm_hod_dow %>% 
  filter(group==6) %>% 
  gravitas::prob_plot("day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 6")+ theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

p7 <- sm_hod_dow %>% 
  filter(group==7) %>% 
  gravitas::prob_plot("day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 7")+ theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

p8 <- sm_hod_dow %>% 
  filter(group==8) %>% 
  gravitas::prob_plot("day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 8")+ theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines")) + ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

```

_Regular scale_

```{r}
(p1 + p2)/ (p3 + p4)
(p5 + p6)/(p7 +p8)
```

_Same scale_

```{r}
(p1 +ylim(0,3) + p2 +ylim(0,3))/ (p3+ylim(0,3) + p4+ylim(0,3))
(p5 +ylim(0,3) + p6 +ylim(0,3))/ (p7+ylim(0,3) + p8+ylim(0,3))
```
# PCA based clustering

```{r}
pc <- prcomp(data_pcp[c(-1,-2)], center = F, scale = F)
summary(pc)
# First two principal componentsafc
comp <- data.frame(pc$x[,1:2])

#plot(comp, pch=16, col=rgb(0,0,0,0.5))

hc = stats::hclust(dist(comp))
data_pcp_all <- data_pcp %>% bind_cols(comp)

library(RColorBrewer)
library(scales)
data_pcp_all %>% ggplot(aes(x=PC1, y = PC2)) + geom_point(aes(color = group)) +
  scale_color_brewer(palette = "Dark2") + theme_bw() + theme(legend.position = "bottom")

data_pcp_all %>%
  pivot_longer(-c(1, 2,8, 9), names_to = "variables", values_to = "wpd") %>% 
  ggplot(aes(x=wpd, y = PC1)) + geom_point(aes(color = group)) +
  scale_color_brewer(palette = "Dark2") + theme_bw() +
  facet_wrap(~variables, nrow = 1) + theme(legend.position = "bottom")
```

One dimensional scale since PC1 explains almost all variation?

### 2021-08-05 job-algo2.sh

@@ -1,0 +1,13 @@
+#! /bin/bash
+#SBATCH --job-name=algo2
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=72:00:00
+#SBATCH --array=1-100
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.5
+Rscript job-algo2.R $SLURM_ARRAY_TASK_ID 
+

### 2021-08-05 supplementary/simulation-alogorithm2/sim_setup_algo2.Rmd
added sim design small iteration

--- 2021-08-05 sim_setup_algo2.Rmd
@@
 ```
-
-# Five iterations of each design (changing seeds)
 
@@
 hc = stats::hclust(dist(comp))
-groups <- cutree(hc, k=5) %>% as_tibble() %>% bind_cols(comp) %>% 
+groups <- cutree(hc, k=koptimal) %>% as_tibble() %>% bind_cols(comp) %>% 
   mutate(value = as.character(value))

### 2021-08-05 supplementary/simulation-alogorithm2/sim_setup_algo2_50iter.Rmd
large iteration algo 2

--- 2021-08-05 supplementary/simulation-alogorithm2/sim_setup_algo2.Rmd
@@
 g3 = c("no", "yes" , "no", "yes", "yes")
-table <- tibble(design, g1, g2, g3) %>% kable()
-table
+#table <- tibble(design, g1, g2, g3) %>% kable()
+#table
 ```
@@
 ```{r}
-seed_len = seq(12345, 12348, 1)
+seed_len = seq(12345, 12394, 1)
+#seed_len = seq(12345, 12346, 1)
 bind_data_iter <- map(seq_len(length(seed_len)), function(x){
@@
   pivot_wider(names_from = x_variable, values_from = wpd)
-data_wide %>% kable()
+#data_wide %>% kable()
 
@@
 koptimal = koptimal$kopt
+
 ```
@@
 mutate(data_id = data_wide$data_id) %>% rename("group" = "value")
-
-  groups %>% kable()
 
@@
 
+  #groups %>% kable()
+
 data_wide_group <- data_wide %>% left_join(groups)  %>% 
@@
 comp <- data.frame(pc$x[,1:2])
-
+rownames(comp) = pc$data_id
 #plot(comp, pch=16, col=rgb(0,0,0,0.5))
@@
   mutate(value = as.character(value))
+# 
+# plot(hc, cex = 0.6)
+# rect.hclust(hc, k = 6, border = 2:10)
+
 library(RColorBrewer)

### 2021-08-12 sim_setup_algo2_diffmeans.Rmd
adding the handpicking examples

--- 2021-08-05 supplementary/simulation-alogorithm2/sim_setup_algo2.Rmd
@@
 ---
-title: "Simulation with algorithm 2"
+title: "Simulation with algorithm 2- different means and one nuisance granularity"
 author: "Sayani Gupta"
@@
 Three circular granularities g1, g2 and g3 are considered with levels 2, 3 and 4 respectively.
-Many time series with 600 observations are created using the five designs below, each of which is iterated four times.
-We anticipate to have five clusters, each with four time series conforming to the same design, once we execute the clustering.
+Many time series with 600 observations are created using the four designs below, each of which is iterated five times.
+We anticipate to have four clusters, each with five time series conforming to the same design, once we execute the clustering.
 
@@
 
-```{r tab-gran}
-design = c("design-1", "design-2", "design-3", "design-4", "design-5")
-g1 = c("no", "yes" , "no", "yes", "yes")
-g2 = c("no", "no" , "yes", "no", "yes")
-g3 = c("no", "yes" , "no", "yes", "yes")
-table <- tibble(design, g1, g2, g3) %>% kable()
-table
+```{r}
+t <- seq(0,299, 1)
+
+g1 = rep(c(0, 1), length.out= length(t))
+g1_dist_design1 <- rep(c(0, 0), length.out= length(t))
+g1_dist_design2 <- rep(c(0, 2), length.out= length(t))
+g1_dist_design3 <- rep(c(1, 4), length.out= length(t))
+g1_dist_design4 <- rep(c(2, 4), length.out= length(t))
+
+g2 = rep(c(0, 1, 2), length.out= length(t))
+g2_dist_design1 <- rep(c(0, 0, 0), length.out= length(t))
+g2_dist_design2 <- rep(c(1, 5, 7), length.out= length(t))
+g2_dist_design3 <- rep(c(1, 2, 3), length.out= length(t))
+g2_dist_design4 <- rep(c(0, 2, 4), length.out= length(t))
+
+g3 = rep(c(0, 1, 2, 3), length.out= length(t))
+g3_dist_design1 <- rep(c(0, 0, 0, 0), length.out= length(t))
+g3_dist_design2 <- rep(c(0, 0, 0, 0), length.out= length(t))
+g3_dist_design3 <- rep(c(0, 0, 0, 0), length.out= length(t))
+g3_dist_design4 <- rep(c(0, 0, 0, 0), length.out= length(t))
+
+time_series <- bind_rows(
+  bind_cols(t, g1, g2, g3, g1_dist_design1, g2_dist_design1, g3_dist_design1),
+  bind_cols(t, g1, g2, g3,  g1_dist_design2, g2_dist_design2, g3_dist_design2),
+  bind_cols(t,  g1, g2, g3, g1_dist_design3, g2_dist_design3, g3_dist_design3),
+  bind_cols(t,  g1, g2, g3, g1_dist_design4, g2_dist_design4, g3_dist_design4), .id = "design") %>%
+  set_names(c("design", "t", "g1", "g2", "g3", "g1_dist", "g2_dist", "g3_dist")) %>% 
+  mutate(sum_dist = g1_dist + g2_dist + g3_dist)
+
+ time_series %>% group_by(design) %>% summarize(s = unique(sum_dist))
 ```
@@
 ```{r}
-t <- seq(0,299, 1)
 
@@
 
-timeseries <- numbers%>% unlist() %>% matrix(ncol = len_mu, nrow = len_obs, byrow = FALSE) %>% t() %>% 
+timeseries <- numbers%>% 
+  unlist() %>%
+  matrix(ncol = len_mu, nrow = len_obs, byrow = FALSE) %>%
+  t() %>% 
   matrix(ncol=1, byrow = TRUE)
@@
 set.seed(12345)
-data_design_1 <- generate_design(t = 599, mu = 0)
-data_design_2 <- generate_design(t = 599, mu = c(0, 2))
-data_design_3 <- generate_design(t = 599, mu = c(0, 2, 4))
-data_design_4 <- generate_design(t = 599, mu = c(0, 2, 4, 6))
-data_design_5 <- generate_design(t = 599, mu = c(0, 2, 4, 6, 8, 10))
+data_design_1 <- generate_design(t = 299, mu = 0)
+data_design_2 <- generate_design(t = 299, mu = c(1, 7, 7, 3, 5, 9))
+data_design_3 <- generate_design(t = 299, mu = c(2, 6, 4, 5, 3, 7))
+data_design_4 <- generate_design(t = 299, mu = c(2, 6, 6, 4, 4, 8))
 ```
@@
 ```{r all-design}
-bind_data <- bind_rows(data_design_1, data_design_2, data_design_3, data_design_4,data_design_5, .id = "design") %>% mutate(design = paste("design", design, sep = "-"))
+bind_data <- bind_rows(data_design_1, data_design_2, data_design_3, data_design_4, .id = "design") %>% mutate(design = paste("design", design, sep = "-"))
 
@@
 ```{r}
-seed_len = seq(12345, 12348, 1)
+seed_len = seq(12345, 12399, 1)
 bind_data_iter <- map(seq_len(length(seed_len)), function(x){
 set.seed(seed_len[x])
-data_design_2 <- generate_design(t = 599, mu = c(0, 2))
-data_design_3 <- generate_design(t = 599, mu = c(0, 2, 4))
-data_design_4 <- generate_design(t = 599, mu = c(0, 2, 4, 6))
-data_design_1 <- generate_design(t = 599, mu = 0)
-data_design_5 <- generate_design(t = 599, mu = c(0, 2, 4, 6, 8, 10))
-bind_data <- bind_rows(data_design_1, data_design_2, data_design_3, data_design_4,data_design_5, .id = "design") %>% mutate(design = paste("design", design, sep = "-"))
+data_design_1 <- generate_design(t = 299, mu = 0)
+data_design_2 <- generate_design(t = 299, mu = c(1, 7, 7, 3, 5, 9))
+data_design_3 <- generate_design(t = 299, mu = c(2, 6, 4, 5, 3, 7))
+data_design_4 <- generate_design(t = 299, mu = c(2, 6, 6, 4, 4, 8))
+bind_data <- bind_rows(data_design_1, data_design_2, data_design_3, data_design_4, .id = "design") %>% mutate(design = paste("design", design, sep = "-"))
 }) %>% bind_rows(.id = "seed_id")
@@
 
-write_csv(data_wpd, "data_wpd.rds")
+#write_csv(data_wpd, "data_wpd.rds")
 ```
@@
 ```{r}
-  groups <- cutree(hc, k=5) %>% as_tibble() %>% 
-mutate(data_id = data_wide$data_id) %>% rename("group" = "value")
-
-  groups %>% kable()
-
-data_wide_group <- data_wide %>% left_join(groups)  %>% 
-  mutate(group_id = as.factor(group))
-
-parcoord <- GGally::ggparcoord(data_wide_group,
-                   columns = 2:ncol(data_wide),
-                   groupColumn = "group_id",
-                   showPoints = TRUE, 
-                   title = "(c)",
-                   alphaLines = 0.8,
-                   scale = "globalminmax", 
-                   order = "anyClass"
-) + ggplot2::theme_bw() +
-  scale_color_brewer(palette = "Dark2") + 
-  ggplot2::theme(
-    plot.title = ggplot2::element_text(size=10)
-  )+
-  ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 0)) +
-  theme(legend.position = "bottom") +
-  xlab("") +
-  ylab("wpd")
-parcoord
-ggpairs_nocolor
-  # ggpairs_color <- ggscatmat(data_wide, columns=2:4, color = "group") +
-  # scale_colour_brewer(palette="Set2")
-```
-
-# With optimal number of clusters
-
-Design 2 and 4 form a part of the same cluster with the optimal number of clusters chosen as 4.
-Design 2 and 4 are similar in the sense that both of them changes across categories for g1 and g3, and remains constant for g2.
-
-```{r}
-  groups <- cutree(hc, k=koptimal) %>% as_tibble() %>% 
+  groups <- cutree(hc, k=4) %>% as_tibble() %>% 
 mutate(data_id = data_wide$data_id) %>% rename("group" = "value")
@@
 
-```{r}
+```{r, eval = FALSE}
 pc <- prcomp(data_wide_group[c(2, 3, 4)], center = F, scale = F)
@@
 ```
+
+```{r}
+data_plot <- left_join(data_wide_group,  bind_data_iter %>% distinct(design, data_id)) %>%
+  mutate(pred_group = case_when(
+    group == 1 ~"design-1",
+    group == 2 ~"design-2",
+    group == 3 ~"design-3",
+    group == 4 ~"design-4"
+  )) %>% 
+  mutate(pred_group = as.factor(pred_group),
+         design = as.factor(design)
+  )
+
+xtab <- confusionMatrix(data_plot$design, data_plot$pred_group)
+
+data_plot %>%  ggplot(aes(x=g1, y=g2)) + geom_point(aes(colour = design)) +
+  scale_color_brewer(palette = "Dark2")
+
+left_join(data_wide_group,  bind_data_iter %>% distinct(design, data_id)) %>%  ggplot(aes(x=g1, y=g3)) + geom_point(aes(colour = design)) +
+  scale_color_brewer(palette = "Dark2")
+
+left_join(data_wide_group,  bind_data_iter %>% distinct(design, data_id)) %>%  ggplot(aes(x=g2, y=g3)) + geom_point(aes(colour = design)) +
+  scale_color_brewer(palette = "Dark2")
+
+```

### 2021-08-12 supplementary/handpick_distinct_cust.Rmd
adding the handpicking examples

---
title: "Hand picking similar behaving group of customers to check clustering results"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, echo = FALSE, message=FALSE, warning=FALSE)
library(readr)
library(tidyverse)
library(gravitas)
library(tsibble)
library(parallel)
library(magrittr)
library(GGally)
library(knitr)
library(patchwork)
library(caret)
```

```{r read-data}
data <- read_rds("../data/elec_nogap_2013_100.rds")
```

```{r hand-pick}
quantile_prob_val = seq(0.1, 0.9, 0.1)

# data_pick <- data %>% 
#   filter(customer_id %in% 
#                    c(10538785, 8530300, 8717912, 8171227, 10108102, 10767756, 10895401, 10704368, 9397229, 8281474))

data_pick_onepeak <- 
  data %>% 
  filter(customer_id %in% c(8485375,10690820,10859082,8486921, 10208124,8198401))%>% mutate(design = 1)

data_pick_threepeak <- 
  data %>% 
  filter(customer_id %in% c(11013154, 10046512, 8637773, 8952846, 8196501, 8627007))%>% mutate(design = 3)

data_pick_twopeak <- 
  data %>% 
  filter(customer_id %in% c(8181071, 8955520, 8640839, 10469265, 10538785, 10895401))%>% mutate(design = 2)

data_pick <- bind_rows(data_pick_onepeak, data_pick_twopeak, data_pick_threepeak)

elec_harmony_all <- read_rds("../data/elec_harmony-100-nogap-new.rds")%>%
  filter(customer_id %in% data_pick$customer_id)

elec_harmony_hod <- elec_harmony_all %>% filter(x_variable=="hour_day") %>% 
  mutate(wpd = round(wpd, 2))

group2_data <- data_pick %>% 
  create_gran("hour_day") %>% 
  as_tibble() %>% 
  group_by(design, hour_day, customer_id) %>% 
  summarise(median_kwh = median(general_supply_kwh))

# 
# group2_data%>% 
#   ungroup() %>% 
#   filter(design==1) %>% 
#   #filter(customer_id %in% c(8148781, 10703878)) %>% 
#   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
#   geom_line(size = 0.5) +
#   facet_wrap(~customer_id, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
# 
# 
# group2_data%>% 
#   ungroup() %>% 
#   filter(design==2) %>% 
#   #filter(customer_id %in% c(8148781, 10703878)) %>% 
#   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
#   geom_line(size = 0.5) +
#   facet_wrap(~customer_id, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
```

```{r characterize, fig.cap= "Median (black) and quartile deviation (blue region) of hourly demand drawn for few customers showing similar behaviors. Roughly speaking, Design 1 has one evening peak, Design 2 has two peaks and Design 3 has three peaks in a day. Each of design 1, 2 and 3 have six similar behaving customers resulting to 18 time series. We want our clustering results to group each of the designs together."}
quantile_prob_val = c(0.25, 0.5, 0.75)

sm_hod <- gravitas::create_gran(data_pick, "hour_day")

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(design, customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

# groups_ref <- unique(data_pick$design) %>% as_tibble %>% 
#               mutate(customer_serial_id = sm_hod_list$customer_id)

# sm_hod_list_cat <-  sm_hod_list %>% left_join(groups_ref, by = c("customer_id" = "customer_serial_id")) %>% select(group, everything())

ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, 2)]))
nrow_sm <- 1:18

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list%>% 
     dplyr::filter(customer_id == customer_id[x]) %>% 
     select(-c(1, 2)) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_val)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

ref_cat <- names(sm_hod_list)[-c(1, 2)] %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
    mutate(customer_serial_id = row_number())

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>% 
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "customer_serial_id") %>% 
  select(customer_serial_id, categories_serial_id, category, customer_id, everything())

data_heatmap <- sm_quantiles_ref %>% 
 left_join(data_pick %>% distinct(customer_id, design))

# 
# %>% 
#   left_join(categories_ref_cat) %>% 
#   rename("category" = "value") %>% 
#   select(-categories_serial_id) %>% 
#   mutate(category = as.numeric(category))

data_heatmap %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`), size = 0.7) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")
```

```{r plot-hand-pick, fig.cap= "Median hourly demand drawn for few customers showing similar median behavior. Roughly speaking, Design 1 has one evening peak, Design 2 has two peaks and Design 3 has three peaks in a day. Each of design 1, 2 and 3 have six similar behaving customers resulting to 18 time series. We want our clustering results to group each of the designs together.", eval = FALSE}
group2_data%>% 
  ungroup() %>% 
  left_join(elec_harmony_hod %>% select(customer_id, wpd)) %>% 
  #filter(design==3) %>% 
  #filter(customer_id %in% c(8148781, 10703878)) %>% 
  ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
  geom_line(size = 0.5) +
  #ggtitle(paste("wpd:", elec_harmony_hod$wpd)) +
  facet_wrap(design~customer_id, scales = "free_y", ncol = 6, labeller = "label_value") + theme_minimal()+ theme(strip.text.y = element_text(size = 5)) 
```

```{r allplot, message=FALSE, warning=FALSE, echo = FALSE}

##----harmony-all
# elec <- read_rds("../data/elec_nogap_2013_100.rds") %>% filter(customer_id %in% data_pick$customer_id)
# 
# elec_split = elec %>% group_split(customer_id)
# harmonies <- read_rds("../../paper-hakear/paper/data/harmonies.rds")
# 
# harmonies_acro <- harmonies %>% mutate(facet_variable=NA, facet_levels=NA) %>% distinct()
# 
# elec_select_harmony = parallel::mclapply(1:18, function(x){
#   
#   data_id <-  elec_split %>% magrittr::extract2(x) %>% 
#     as_tsibble(index = reading_datetime)
# 
#   
#   k = hakear::select_harmonies(data_id,
#                            harmony_tbl = harmonies_acro,
#                            response = general_supply_kwh,
#                            nperm = 200,
#                            nsamp = 2
#   ) %>% mutate(customer_id = unique(data_id$customer_id))
#   #write_rds(k, paste("data/elec_harmony-100-nogap.rds"))
# }, mc.cores = parallel::detectCores() - 1, mc.preschedule = FALSE, mc.set.seed = FALSE)
# #toc()
# elec_harmny <- elec_select_harmony %>% bind_rows() 
# write_rds(elec_harmny, paste("../data/elec_harmony-100-nogap-new.rds"))

select_split <- str_split(elec_harmony_all$select_harmony, " ", simplify = TRUE)[,2]

elec_sig_split <- elec_harmony_all %>%
  mutate(facet_variable = case_when(
    facet_variable == "hour_day" ~ "hod" ,
    facet_variable == "day_month" ~ "dom" ,
    facet_variable == "day_week" ~ "dow" ,
    facet_variable == "week_month" ~ "wom" ,
    facet_variable == "wknd_wday" ~ "wdwnd"
  )) %>% 
  mutate(x_variable = case_when(
    x_variable == "hour_day" ~ "hod" ,
    x_variable == "day_month" ~ "dom" ,
    x_variable == "day_week" ~ "dow" ,
    x_variable == "week_month" ~ "wom" ,
    x_variable == "wknd_wday" ~ "wdwnd"
  )) %>% 
  bind_cols(select_split = select_split) %>% 
  mutate(significant = case_when(
    select_split == "***" ~ "highest",
    select_split == "**" ~ "high",
    select_split == "*" ~ "medium",
    select_split == "" ~ "low"
  )) %>%mutate(comb = x_variable)

elec_sig_split$significant <- 
  factor(elec_sig_split$significant, levels = c("highest", "high", "medium", "low"))

data_comb_filter = elec_sig_split %>% 
mutate(low_proxy = if_else(significant=="low", 0, 1))%>% 
select(comb, customer_id, low_proxy) %>% 
pivot_wider(names_from = customer_id, values_from = low_proxy) %>%
mutate(score = rowMeans(across(-1))) %>% 
filter(score >0.3)

data_sig_filter = elec_sig_split %>% 
filter(comb %in% data_comb_filter$comb)

data_clust = data_sig_filter %>% 
select(comb, customer_id, wpd) %>% 
pivot_wider(names_from = comb, values_from = wpd) 
```

```{r hc-plot, fig.cap="Dendogram from clustering 18 time series into three clusters."}
hc = stats::hclust(dist(data_clust[-1]),method="complete")
#plot(hc)
plot(hc, cex = 0.6)
rect.hclust(hc, k = 3, border = 2:10)
```

```{r conf-matrix, fig.cap="The confusion matrix showing how well the clustering algorithm worked."}
groups <- cutree(hc, k=3) %>% 
  as_tibble() %>% 
mutate(customer_id = data_clust$customer_id)

elec_pred_wo_scaled <- elec_harmony_all %>% 
  filter(x_variable %in% c("hour_day", "day_month")) %>% 
  left_join(data_pick %>% as_tibble() %>% select(customer_id, design)%>% unique) %>% select(-facet_variable, -facet_levels, -x_levels, -select_harmony) %>% 
  pivot_wider(names_from = x_variable, values_from = wpd) %>% 
  left_join(groups) %>% 
  rename("pred_design" = "value") %>% 
  arrange(day_month) %>% 
  select(customer_id, design, pred_design, everything()) %>%
  mutate(day_month = round(day_month, 1),
         hour_day = round(hour_day, 1))

elec_pred_wo_scaled$design = as.factor(elec_pred_wo_scaled$design)
elec_pred_wo_scaled$pred_design = as.factor(elec_pred_wo_scaled$pred_design)
xtab <- confusionMatrix(elec_pred_wo_scaled$pred_design, elec_pred_wo_scaled$design)
xtab$table
```

```{r tab-cluster}
# 
#  fig.cap="The customers are arranged in ascending order of wpd for dom and it can be seen that the predicted designs are exactly sorted in that order."
elec_pred_wo_scaled%>% 
  kable(caption = "Actual and predicted allocation of designs to customers" )

```

\newpage

##----cluster-characterization

```{r}
data = data_clust %>% 
left_join(groups)%>% 
             rename ("group" = "value") %>% 
             ungroup() %>% 
             select(group, customer_id, everything()) %>% 
             pivot_longer(-c(1,2), names_to = "comb", values_to = "wpd")

data_pcp <- data %>% pivot_wider(
              names_from = comb, 
              values_from = wpd) %>% 
  mutate(group = as.character(group)) 
```

```{r ggpairs-plot, fig.cap="Pairs plot show that dom is the variable that is responsible for this clustering. The green colored cluster correspond to lowest value of dom, the orange colored and blue colored one correspond to middle range and high range of dom."}
# ggpairs_nocolor <- ggpairs(data_pcp, columns = 3:7)
# ggpairs_nocolor

groupge2 <- data %>% 
  group_by(group) %>% 
  summarise(n = length(unique(customer_id))) %>% 
  filter(n>2)

data_pcp_ge2 <- data_pcp %>% filter(group %in% groupge2$group)

# ggpairs_nocolor <- ggpairs(data_pcp_ge2, columns = 3:7)

ggpairs_nocolor <- ggscatmat(data_pcp_ge2, columns=3:4) +
  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")

ggpairs_color <- ggscatmat(data_pcp_ge2, columns=3:4, color = "group") +
  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")

ggpairs_nocolor + ggpairs_color
```

```{r pred-actual-design, fig.cap="First facet label indicate actual design, second denote predicted design, third denote customer_id."}
group_data_orig <- group2_data %>% left_join(elec_pred_wo_scaled %>% select(customer_id,pred_design))

# group_data_orig%>% 
#   ungroup() %>% 
#   filter(group==1) %>% 
#   #filter(customer_id %in% c(8148781, 10703878)) %>% 
#   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
#   geom_line(size = 0.5) +
#   facet_wrap(~customer_id, scales = "free_y",labeller = "label_value", ncol = 1) + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
# 
# 
# group_data_orig%>% 
#   ungroup() %>% 
#   filter(group==2) %>% 
#   #filter(customer_id %in% c(8148781, 10703878)) %>% 
#   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
#   geom_line(size = 0.5) +
#   facet_wrap(~customer_id, scales = "free_y",labeller = "label_value", ncol = 1) + theme_minimal()+ theme(strip.text.y = element_text(size = 5))

group_data_orig%>% 
  ungroup() %>% 
  rename("pred" = "pred_design", "cust" = "customer_id") %>% 
  #filter(group==3) %>% 
  #filter(customer_id %in% c(8148781, 10703878)) %>% 
  ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
  geom_line(size = 0.5) +
  facet_wrap(design~pred~cust, scales = "free_y",labeller = "label_context") + theme_minimal()+ theme(strip.text.y = element_text(size = 5)) + 
       theme(strip.text.x = element_text(size = 6, margin = margin(b = 0, t = 0)))+ xlab("hod")
```

```{r dom-pred, fig.cap="(Top) free y scale, (bottom) y scales same across facets seemingly implies that clustering is ordered by sorting the median demand."}
group2_data <- data_pick %>% 
  create_gran("day_month") %>% 
  as_tibble() %>% 
  #mutate(general_supply_kwh = scale(general_supply_kwh)) %>% 
  group_by(design, day_month, customer_id) %>% 
  summarise(median_kwh = median(general_supply_kwh))

group_data_orig <- group2_data %>% left_join(elec_pred_wo_scaled %>% select(customer_id,pred_design))

group_data_orig %>% 
  ungroup() %>% 
  rename("pred" = "pred_design", "cust" = "customer_id") %>% 
  #filter(group==3) %>% 
  #filter(customer_id %in% c(8148781, 10703878)) %>% 
  ggplot(aes(x = as.integer(day_month), y = median_kwh)) +
  geom_line(size = 0.5) +
  facet_wrap(pred~cust,
             labeller = "label_context",
             scales = "free_y") + 
  theme_minimal()+ theme(strip.text.y = element_text(size = 5)) + scale_y_log10() + scale_x_continuous(breaks=seq(1, 31, 7), minor_breaks = seq(1, 31, 7))+ 
       theme(strip.text.x = element_text(size = 6, margin = margin(b = 0, t = 0)))+ xlab("dom")
# 
group_data_orig$customer_id <-factor(group_data_orig$customer_id, levels = elec_pred_wo_scaled$customer_id)

group_data_orig %>% 
  ungroup() %>% 
  rename("pred" = "pred_design", "cust" = "customer_id") %>% 
  #filter(group==3) %>% 
  #filter(customer_id %in% c(8148781, 10703878)) %>% 
  ggplot(aes(x = as.integer(day_month), y = median_kwh)) +
  geom_line(size = 0.5) +
  facet_wrap(pred~cust,
             labeller = "label_context") + 
  theme_minimal()+ theme(strip.text.y = element_text(size = 5)) + scale_y_log10() + scale_x_continuous(breaks=seq(1, 31, 7), minor_breaks = seq(1, 31, 7))+ 
       theme(strip.text.x = element_text(size = 6, margin = margin(b = 0, t = 0))) + xlab("dom")

# data_pick %>% 
#   create_gran("day_month") %>% 
#   ungroup() %>% left_join(data_pcp_ge2 %>% select(customer_id, group)) %>% 
#   #filter(group==3) %>% 
#   #filter(customer_id %in% c(8148781, 10703878)) %>% 
#   ggplot(aes(x = as.factor(day_month), y = general_supply_kwh)) +
#   geom_boxplot(size = 0.5) +
#   facet_wrap(group~customer_id,
#              labeller = "label_value",
#              scales = "free_y") + 
#   theme_minimal()+ theme(strip.text.y = element_text(size = 5)) + scale_y_log10()
```

<!-- #Group2: n=24 -->
<!-- ```{r} -->
<!-- group2_data <- sm_hod_dow %>%  -->
<!--   as_tibble() %>%  -->
<!--   filter(group==2) %>%  -->
<!--   group_by(day_week, hour_day, customer_id) %>%  -->
<!--   summarise(median_kwh = median(general_supply_kwh))  -->

<!-- group2_data%>%  -->
<!--   ungroup() %>%  -->
<!--   #filter(customer_id %in% c(8148781, 10703878)) %>%  -->
<!--   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) + -->
<!--   geom_line(size = 0.5) + -->
<!--   facet_grid(customer_id~day_week, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5)) -->
<!-- ``` -->

<!-- #Group4: n=24 -->
<!-- ```{r} -->
<!-- group2_data <- sm_hod_dow %>%  -->
<!--   as_tibble() %>%  -->
<!--   filter(group==4) %>%  -->
<!--   group_by(day_week, hour_day, customer_id) %>%  -->
<!--   summarise(median_kwh = median(general_supply_kwh))  -->

<!-- group2_data%>%  -->
<!--   ungroup() %>%  -->
<!--   #filter(customer_id %in% c(8148781, 10703878)) %>%  -->
<!--   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) + -->
<!--   geom_line(size = 0.5) + -->
<!--   facet_grid(customer_id~day_week, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5)) -->
<!-- ``` -->

<!-- #Group5: n=18 -->
<!-- ```{r} -->
<!-- group2_data <- sm_hod_dow %>%  -->
<!--   as_tibble() %>%  -->
<!--   filter(group==5) %>%  -->
<!--   group_by(day_week, hour_day, customer_id) %>%  -->
<!--   summarise(median_kwh = median(general_supply_kwh))  -->

<!-- group2_data%>%  -->
<!--   ungroup() %>%  -->
<!--   #filter(customer_id %in% c(8148781, 10703878)) %>%  -->
<!--   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) + -->
<!--   geom_line(size = 0.5) + -->
<!--   facet_grid(customer_id~day_week, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5)) -->
<!-- ``` -->

<!-- #Group6: n=23 -->
<!-- ```{r} -->
<!-- group2_data <- sm_hod_dow %>%  -->
<!--   as_tibble() %>%  -->
<!--   filter(group==6) %>%  -->
<!--   group_by(day_week, hour_day, customer_id) %>%  -->
<!--   summarise(median_kwh = median(general_supply_kwh))  -->

<!-- group2_data%>%  -->
<!--   ungroup() %>%  -->
<!--   #filter(customer_id %in% c(8148781, 10703878)) %>%  -->
<!--   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) + -->
<!--   geom_line(size = 0.5) + -->
<!--   facet_grid(customer_id~day_week, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5)) -->
<!-- ``` -->

```{r}
data_clust = data_sig_filter %>% 
select(comb, customer_id, wpd) %>% 
pivot_wider(names_from = comb, values_from = wpd) %>% 
  mutate(hod = scale(hod),
         dom = scale(dom))

hc = stats::hclust(dist(data_clust[-1], method = "manhattan"),method="complete")
#plot(hc)
plot(hc, cex = 0.6)
rect.hclust(hc, k = 3, border = 2:10)

groups_scaled <- cutree(hc, k=3) %>% 
  as_tibble() %>% 
mutate(customer_id = data_clust$customer_id) %>% rename("pred_design_scale" = "value")

elec_harmony_all %>% 
  filter(x_variable %in% c("hour_day", "day_month")) %>% 
  left_join(data_pick %>% as_tibble() %>% select(customer_id, design)%>% unique) %>% select(-facet_variable, -facet_levels, -x_levels, -select_harmony) %>% 
  pivot_wider(names_from = x_variable, values_from = wpd) %>% 
  mutate(hod_scaled =  round(scale(hour_day), 1), 
         dom_scaled = round(scale(day_month),1)) %>% 
  left_join(groups_scaled)  %>%
  mutate(day_month = round(day_month, 1),
         hour_day = round(hour_day, 1)) %>% 
  left_join(elec_pred_wo_scaled %>% select(customer_id, pred_design), by = "customer_id") %>% 
  arrange(design) %>% 
  select(customer_id, design, pred_design_scale, pred_design, everything()) %>% 
  kable()

##----cluster-characterization

data = data_clust %>% 
left_join(groups)%>% 
             rename ("group" = "value") %>% 
             ungroup() %>% 
             select(group, customer_id, everything()) %>% 
             pivot_longer(-c(1,2), names_to = "comb", values_to = "wpd")

data_pcp <- data %>% pivot_wider(
              names_from = comb, 
              values_from = wpd) %>% 
  mutate(group = as.character(group)) 
```

```{r}
# ggpairs_nocolor <- ggpairs(data_pcp, columns = 3:7)
# ggpairs_nocolor

groupge2 <- data %>% 
  group_by(group) %>% 
  summarise(n = length(unique(customer_id))) %>% 
  filter(n>2)

data_pcp_ge2 <- data_pcp %>% filter(group %in% groupge2$group)

# ggpairs_nocolor <- ggpairs(data_pcp_ge2, columns = 3:7)

ggpairs_nocolor <- ggscatmat(data_pcp_ge2, columns=3:4) +
  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")

ggpairs_color <- ggscatmat(data_pcp_ge2, columns=3:4, color = "group") +
  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")

ggpairs_nocolor + ggpairs_color
```

### 2021-08-12 algo2.R

@@ -68,6 +68,8 @@ library(tictoc)
 #tic()
 elec_split = elec %>% group_split(customer_id)
 
+harmonies_acro <- harmonies %>% mutate(facet_variable=NA, facet_levels=NA) %>% distinct()
+
 elec_select_harmony = parallel::mclapply(1:100, function(x){
   
   data_id <-  elec_split %>% magrittr::extract2(x) %>% 
@@ -78,13 +80,14 @@ elec_select_harmony = parallel::mclapply(1:100, function(x){
                            harmony_tbl = harmonies_acro,
                            response = general_supply_kwh,
                            nperm = 200,
-                           nsamp = 200
-  )
+                           nsamp = 2
+  ) %>% mutate(customer_id = unique(data_id$customer_id))
   #write_rds(k, paste("data/elec_harmony-100-nogap.rds"))
 }, mc.cores = parallel::detectCores() - 1, mc.preschedule = FALSE, mc.set.seed = FALSE)
 #toc()
-
+elec_harmny <- elec_select_harmony %>% bind_rows() 
 write_rds(elec_select_harmony, paste("data/elec_harmony-100-nogap.rds"))
+# write_rds(elec_harmny, paste("data/elec_harmony-100-nogap-onegran.rds"))
 
 # ## ---- elec_select_harmony-8
 # elec_harmony_all <- elec_select_harmony %>% 

### 2021-08-17 supplementary/handpick_distinct_cust.Rmd
update README

@@
        theme(strip.text.x = element_text(size = 6, margin = margin(b = 0, t = 0))) + xlab("dom")
-
-# data_pick %>% 
-#   create_gran("day_month") %>% 
-#   ungroup() %>% left_join(data_pcp_ge2 %>% select(customer_id, group)) %>% 
-#   #filter(group==3) %>% 
-#   #filter(customer_id %in% c(8148781, 10703878)) %>% 
-#   ggplot(aes(x = as.factor(day_month), y = general_supply_kwh)) +
-#   geom_boxplot(size = 0.5) +
-#   facet_wrap(group~customer_id,
-#              labeller = "label_value",
-#              scales = "free_y") + 
-#   theme_minimal()+ theme(strip.text.y = element_text(size = 5)) + scale_y_log10()
 ```
 
-<!-- #Group2: n=24 -->
-<!-- ```{r} -->
-<!-- group2_data <- sm_hod_dow %>%  -->
-<!--   as_tibble() %>%  -->
-<!--   filter(group==2) %>%  -->
-<!--   group_by(day_week, hour_day, customer_id) %>%  -->
-<!--   summarise(median_kwh = median(general_supply_kwh))  -->
-
-<!-- group2_data%>%  -->
-<!--   ungroup() %>%  -->
-<!--   #filter(customer_id %in% c(8148781, 10703878)) %>%  -->
-<!--   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) + -->
-<!--   geom_line(size = 0.5) + -->
-<!--   facet_grid(customer_id~day_week, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5)) -->
-<!-- ``` -->
-
-<!-- #Group4: n=24 -->
-<!-- ```{r} -->
-<!-- group2_data <- sm_hod_dow %>%  -->
-<!--   as_tibble() %>%  -->
-<!--   filter(group==4) %>%  -->
-<!--   group_by(day_week, hour_day, customer_id) %>%  -->
-<!--   summarise(median_kwh = median(general_supply_kwh))  -->
-
-<!-- group2_data%>%  -->
-<!--   ungroup() %>%  -->
-<!--   #filter(customer_id %in% c(8148781, 10703878)) %>%  -->
-<!--   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) + -->
-<!--   geom_line(size = 0.5) + -->
-<!--   facet_grid(customer_id~day_week, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5)) -->
-<!-- ``` -->
-
-<!-- #Group5: n=18 -->
-<!-- ```{r} -->
-<!-- group2_data <- sm_hod_dow %>%  -->
-<!--   as_tibble() %>%  -->
-<!--   filter(group==5) %>%  -->
-<!--   group_by(day_week, hour_day, customer_id) %>%  -->
-<!--   summarise(median_kwh = median(general_supply_kwh))  -->
-
-<!-- group2_data%>%  -->
-<!--   ungroup() %>%  -->
-<!--   #filter(customer_id %in% c(8148781, 10703878)) %>%  -->
-<!--   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) + -->
-<!--   geom_line(size = 0.5) + -->
-<!--   facet_grid(customer_id~day_week, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5)) -->
-<!-- ``` -->
-
-<!-- #Group6: n=23 -->
-<!-- ```{r} -->
-<!-- group2_data <- sm_hod_dow %>%  -->
-<!--   as_tibble() %>%  -->
-<!--   filter(group==6) %>%  -->
-<!--   group_by(day_week, hour_day, customer_id) %>%  -->
-<!--   summarise(median_kwh = median(general_supply_kwh))  -->
-
-<!-- group2_data%>%  -->
-<!--   ungroup() %>%  -->
-<!--   #filter(customer_id %in% c(8148781, 10703878)) %>%  -->
-<!--   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) + -->
-<!--   geom_line(size = 0.5) + -->
-<!--   facet_grid(customer_id~day_week, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5)) -->
-<!-- ``` -->
-
-```{r}
+```{r, eval = FALSE}
 data_clust = data_sig_filter %>% 
@@
 
-```{r}
+```{r, eval = FALSE}
 # ggpairs_nocolor <- ggpairs(data_pcp, columns = 3:7)
@@
 ```
+
+# Algorithm discussed today
+
+- take significant granularities
+- consider sum of JS distances between each categories of each cyclic granularity

### 2021-08-17 supplementary/handpick_distinct_cust_algonew.Rmd
update README

--- 2021-08-17 supplementary/handpick_distinct_cust.Rmd
@@
 ```
+```{r}
+#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
+library(gracsr)
+v2 <- suppressWarnings(scaled_dist_gran(data_pick, "hour_day", response = "general_supply_kwh", quantile_prob_val = c(0.5, 0.75, 0.9))) %>% rename("dist_hod" = "dist")
+v3 <- suppressWarnings(scaled_dist_gran(data_pick, "day_month", response = "general_supply_kwh", quantile_prob_val = c(0.5, 0.75, 0.9))) %>% rename("dist_dom" = "dist")
+data_dist <- v3 %>% left_join(v2) %>% mutate(dist = dist_hod) %>% 
+    pivot_wider(-c(3, 4), names_from = customer_to,
+              values_from = dist) %>%
+  rename("customer_id" = "customer_from")
 
-```{r plot-hand-pick, fig.cap= "Median hourly demand drawn for few customers showing similar median behavior.
-Roughly speaking, Design 1 has one evening peak, Design 2 has two peaks and Design 3 has three peaks in a day.
-Each of design 1, 2 and 3 have six similar behaving customers resulting to 18 time series.
-We want our clustering results to group each of the designs together.", eval = FALSE}
-group2_data%>% 
-  ungroup() %>% 
-  left_join(elec_harmony_hod %>% select(customer_id, wpd)) %>% 
-  #filter(design==3) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-  ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
-  geom_line(size = 0.5) +
-  #ggtitle(paste("wpd:", elec_harmony_hod$wpd)) +
-  facet_wrap(design~customer_id, scales = "free_y", ncol = 6, labeller = "label_value") + theme_minimal()+ theme(strip.text.y = element_text(size = 5)) 
+d = stats::as.dist(data_dist[,-1])
+hc = stats::hclust(d,method="complete")
+
+groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
+groups
+groups %>% filter(group==2)
+groups %>% filter(group==3)
+groups %>% group_by(group) %>% summarise(n=n())
 ```
-
-```{r allplot, message=FALSE, warning=FALSE, echo = FALSE}
-
-##----harmony-all
-# elec <- read_rds("../data/elec_nogap_2013_100.rds") %>% filter(customer_id %in% data_pick$customer_id)
-# 
-# elec_split = elec %>% group_split(customer_id)
-# harmonies <- read_rds("../../paper-hakear/paper/data/harmonies.rds")
-# 
-# harmonies_acro <- harmonies %>% mutate(facet_variable=NA, facet_levels=NA) %>% distinct()
-# 
-# elec_select_harmony = parallel::mclapply(1:18, function(x){
-#   
-#   data_id <-  elec_split %>% magrittr::extract2(x) %>% 
-#     as_tsibble(index = reading_datetime)
-# 
-#   
-#   k = hakear::select_harmonies(data_id,
-#                            harmony_tbl = harmonies_acro,
-#                            response = general_supply_kwh,
-#                            nperm = 200,
-#                            nsamp = 2
-#   ) %>% mutate(customer_id = unique(data_id$customer_id))
-#   #write_rds(k, paste("data/elec_harmony-100-nogap.rds"))
-# }, mc.cores = parallel::detectCores() - 1, mc.preschedule = FALSE, mc.set.seed = FALSE)
-# #toc()
-# elec_harmny <- elec_select_harmony %>% bind_rows() 
-# write_rds(elec_harmny, paste("../data/elec_harmony-100-nogap-new.rds"))
-
-select_split <- str_split(elec_harmony_all$select_harmony, " ", simplify = TRUE)[,2]
-
-elec_sig_split <- elec_harmony_all %>%
-  mutate(facet_variable = case_when(
-    facet_variable == "hour_day" ~ "hod" ,
-    facet_variable == "day_month" ~ "dom" ,
-    facet_variable == "day_week" ~ "dow" ,
-    facet_variable == "week_month" ~ "wom" ,
-    facet_variable == "wknd_wday" ~ "wdwnd"
-  )) %>% 
-  mutate(x_variable = case_when(
-    x_variable == "hour_day" ~ "hod" ,
-    x_variable == "day_month" ~ "dom" ,
-    x_variable == "day_week" ~ "dow" ,
-    x_variable == "week_month" ~ "wom" ,
-    x_variable == "wknd_wday" ~ "wdwnd"
-  )) %>% 
-  bind_cols(select_split = select_split) %>% 
-  mutate(significant = case_when(
-    select_split == "***" ~ "highest",
-    select_split == "**" ~ "high",
-    select_split == "*" ~ "medium",
-    select_split == "" ~ "low"
-  )) %>%mutate(comb = x_variable)
-
-elec_sig_split$significant <- 
-  factor(elec_sig_split$significant, levels = c("highest", "high", "medium", "low"))
-
-data_comb_filter = elec_sig_split %>% 
-mutate(low_proxy = if_else(significant=="low", 0, 1))%>% 
-select(comb, customer_id, low_proxy) %>% 
-pivot_wider(names_from = customer_id, values_from = low_proxy) %>%
-mutate(score = rowMeans(across(-1))) %>% 
-filter(score >0.3)
-
-data_sig_filter = elec_sig_split %>% 
-filter(comb %in% data_comb_filter$comb)
-
-data_clust = data_sig_filter %>% 
-select(comb, customer_id, wpd) %>% 
-pivot_wider(names_from = comb, values_from = wpd) 
-```
-
-```{r hc-plot, fig.cap="Dendogram from clustering 18 time series into three clusters."}
-hc = stats::hclust(dist(data_clust[-1]),method="complete")
-#plot(hc)
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 3, border = 2:10)
-```
-
-```{r conf-matrix, fig.cap="The confusion matrix showing how well the clustering algorithm worked."}
-groups <- cutree(hc, k=3) %>% 
-  as_tibble() %>% 
-mutate(customer_id = data_clust$customer_id)
-
-elec_pred_wo_scaled <- elec_harmony_all %>% 
-  filter(x_variable %in% c("hour_day", "day_month")) %>% 
-  left_join(data_pick %>% as_tibble() %>% select(customer_id, design)%>% unique) %>% select(-facet_variable, -facet_levels, -x_levels, -select_harmony) %>% 
-  pivot_wider(names_from = x_variable, values_from = wpd) %>% 
-  left_join(groups) %>% 
-  rename("pred_design" = "value") %>% 
-  arrange(day_month) %>% 
-  select(customer_id, design, pred_design, everything()) %>%
-  mutate(day_month = round(day_month, 1),
-         hour_day = round(hour_day, 1))
-
-elec_pred_wo_scaled$design = as.factor(elec_pred_wo_scaled$design)
-elec_pred_wo_scaled$pred_design = as.factor(elec_pred_wo_scaled$pred_design)
-xtab <- confusionMatrix(elec_pred_wo_scaled$pred_design, elec_pred_wo_scaled$design)
-xtab$table
-```
-
-```{r tab-cluster}
-# 
-#  fig.cap="The customers are arranged in ascending order of wpd for dom and it can be seen that the predicted designs are exactly sorted in that order."
-elec_pred_wo_scaled%>% 
-  kable(caption = "Actual and predicted allocation of designs to customers" )
-
-```
-
-\newpage
-
-##----cluster-characterization
-
-```{r}
-data = data_clust %>% 
-left_join(groups)%>% 
-             rename ("group" = "value") %>% 
-             ungroup() %>% 
-             select(group, customer_id, everything()) %>% 
-             pivot_longer(-c(1,2), names_to = "comb", values_to = "wpd")
-
-data_pcp <- data %>% pivot_wider(
-              names_from = comb, 
-              values_from = wpd) %>% 
-  mutate(group = as.character(group)) 
-```
-
-```{r ggpairs-plot, fig.cap="Pairs plot show that dom is the variable that is responsible for this clustering.
-The green colored cluster correspond to lowest value of dom, the orange colored and blue colored one correspond to middle range and high range of dom."}
-# ggpairs_nocolor <- ggpairs(data_pcp, columns = 3:7)
-# ggpairs_nocolor
-
-groupge2 <- data %>% 
-  group_by(group) %>% 
-  summarise(n = length(unique(customer_id))) %>% 
-  filter(n>2)
-
-data_pcp_ge2 <- data_pcp %>% filter(group %in% groupge2$group)
-
-# ggpairs_nocolor <- ggpairs(data_pcp_ge2, columns = 3:7)
-
-ggpairs_nocolor <- ggscatmat(data_pcp_ge2, columns=3:4) +
-  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")
-
-ggpairs_color <- ggscatmat(data_pcp_ge2, columns=3:4, color = "group") +
-  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")
-
-ggpairs_nocolor + ggpairs_color
-```
-
-```{r pred-actual-design, fig.cap="First facet label indicate actual design, second denote predicted design, third denote customer_id."}
-group_data_orig <- group2_data %>% left_join(elec_pred_wo_scaled %>% select(customer_id,pred_design))
-
-# group_data_orig%>% 
-#   ungroup() %>% 
-#   filter(group==1) %>% 
-#   #filter(customer_id %in% c(8148781, 10703878)) %>% 
-#   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
-#   geom_line(size = 0.5) +
-#   facet_wrap(~customer_id, scales = "free_y",labeller = "label_value", ncol = 1) + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
-# 
-# 
-# group_data_orig%>% 
-#   ungroup() %>% 
-#   filter(group==2) %>% 
-#   #filter(customer_id %in% c(8148781, 10703878)) %>% 
-#   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
-#   geom_line(size = 0.5) +
-#   facet_wrap(~customer_id, scales = "free_y",labeller = "label_value", ncol = 1) + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
-
-group_data_orig%>% 
-  ungroup() %>% 
-  rename("pred" = "pred_design", "cust" = "customer_id") %>% 
-  #filter(group==3) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-  ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
-  geom_line(size = 0.5) +
-  facet_wrap(design~pred~cust, scales = "free_y",labeller = "label_context") + theme_minimal()+ theme(strip.text.y = element_text(size = 5)) + 
-       theme(strip.text.x = element_text(size = 6, margin = margin(b = 0, t = 0)))+ xlab("hod")
-```
-
-```{r dom-pred, fig.cap="(Top) free y scale, (bottom) y scales same across facets seemingly implies that clustering is ordered by sorting the median demand."}
-group2_data <- data_pick %>% 
-  create_gran("day_month") %>% 
-  as_tibble() %>% 
-  #mutate(general_supply_kwh = scale(general_supply_kwh)) %>% 
-  group_by(design, day_month, customer_id) %>% 
-  summarise(median_kwh = median(general_supply_kwh))
-
-group_data_orig <- group2_data %>% left_join(elec_pred_wo_scaled %>% select(customer_id,pred_design))
-
-group_data_orig %>% 
-  ungroup() %>% 
-  rename("pred" = "pred_design", "cust" = "customer_id") %>% 
-  #filter(group==3) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-  ggplot(aes(x = as.integer(day_month), y = median_kwh)) +
-  geom_line(size = 0.5) +
-  facet_wrap(pred~cust,
-             labeller = "label_context",
-             scales = "free_y") + 
-  theme_minimal()+ theme(strip.text.y = element_text(size = 5)) + scale_y_log10() + scale_x_continuous(breaks=seq(1, 31, 7), minor_breaks = seq(1, 31, 7))+ 
-       theme(strip.text.x = element_text(size = 6, margin = margin(b = 0, t = 0)))+ xlab("dom")
-# 
-group_data_orig$customer_id <-factor(group_data_orig$customer_id, levels = elec_pred_wo_scaled$customer_id)
-
-group_data_orig %>% 
-  ungroup() %>% 
-  rename("pred" = "pred_design", "cust" = "customer_id") %>% 
-  #filter(group==3) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-  ggplot(aes(x = as.integer(day_month), y = median_kwh)) +
-  geom_line(size = 0.5) +
-  facet_wrap(pred~cust,
-             labeller = "label_context") + 
-  theme_minimal()+ theme(strip.text.y = element_text(size = 5)) + scale_y_log10() + scale_x_continuous(breaks=seq(1, 31, 7), minor_breaks = seq(1, 31, 7))+ 
-       theme(strip.text.x = element_text(size = 6, margin = margin(b = 0, t = 0))) + xlab("dom")
-```
-
-```{r, eval = FALSE}
-data_clust = data_sig_filter %>% 
-select(comb, customer_id, wpd) %>% 
-pivot_wider(names_from = comb, values_from = wpd) %>% 
-  mutate(hod = scale(hod),
-         dom = scale(dom))
-
-hc = stats::hclust(dist(data_clust[-1], method = "manhattan"),method="complete")
-#plot(hc)
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 3, border = 2:10)
-
-groups_scaled <- cutree(hc, k=3) %>% 
-  as_tibble() %>% 
-mutate(customer_id = data_clust$customer_id) %>% rename("pred_design_scale" = "value")
-
-elec_harmony_all %>% 
-  filter(x_variable %in% c("hour_day", "day_month")) %>% 
-  left_join(data_pick %>% as_tibble() %>% select(customer_id, design)%>% unique) %>% select(-facet_variable, -facet_levels, -x_levels, -select_harmony) %>% 
-  pivot_wider(names_from = x_variable, values_from = wpd) %>% 
-  mutate(hod_scaled =  round(scale(hour_day), 1), 
-         dom_scaled = round(scale(day_month),1)) %>% 
-  left_join(groups_scaled)  %>%
-  mutate(day_month = round(day_month, 1),
-         hour_day = round(hour_day, 1)) %>% 
-  left_join(elec_pred_wo_scaled %>% select(customer_id, pred_design), by = "customer_id") %>% 
-  arrange(design) %>% 
-  select(customer_id, design, pred_design_scale, pred_design, everything()) %>% 
-  kable()
-
-##----cluster-characterization
-
-data = data_clust %>% 
-left_join(groups)%>% 
-             rename ("group" = "value") %>% 
-             ungroup() %>% 
-             select(group, customer_id, everything()) %>% 
-             pivot_longer(-c(1,2), names_to = "comb", values_to = "wpd")
-
-data_pcp <- data %>% pivot_wider(
-              names_from = comb, 
-              values_from = wpd) %>% 
-  mutate(group = as.character(group)) 
-```
-
-```{r, eval = FALSE}
-# ggpairs_nocolor <- ggpairs(data_pcp, columns = 3:7)
-# ggpairs_nocolor
-
-groupge2 <- data %>% 
-  group_by(group) %>% 
-  summarise(n = length(unique(customer_id))) %>% 
-  filter(n>2)
-
-data_pcp_ge2 <- data_pcp %>% filter(group %in% groupge2$group)
-
-# ggpairs_nocolor <- ggpairs(data_pcp_ge2, columns = 3:7)
-
-ggpairs_nocolor <- ggscatmat(data_pcp_ge2, columns=3:4) +
-  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")
-
-ggpairs_color <- ggscatmat(data_pcp_ge2, columns=3:4, color = "group") +
-  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")
-
-ggpairs_nocolor + ggpairs_color
-```
-
-# Algorithm discussed today
-
-- take significant granularities
-- consider sum of JS distances between each categories of each cyclic granularity

### 2021-08-18 supplementary/handpick_algo3.Rmd
adding readme for algo 4

--- 2021-08-17 supplementary/handpick_distinct_cust_algonew.Rmd
@@
 ---
-title: "Hand picking similar behaving group of customers to check clustering results"
+title: "Hand picking similar behaving group of customers to check clustering results - Algorithm 3"
 output: pdf_document
@@
 ```
-```{r}
+```{r,echo = TRUE}
 #data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
 library(gracsr)
-v2 <- suppressWarnings(scaled_dist_gran(data_pick, "hour_day", response = "general_supply_kwh", quantile_prob_val = c(0.5, 0.75, 0.9))) %>% rename("dist_hod" = "dist")
-v3 <- suppressWarnings(scaled_dist_gran(data_pick, "day_month", response = "general_supply_kwh", quantile_prob_val = c(0.5, 0.75, 0.9))) %>% rename("dist_dom" = "dist")
-data_dist <- v3 %>% left_join(v2) %>% mutate(dist = dist_hod) %>% 
-    pivot_wider(-c(3, 4), names_from = customer_to,
+v2 <- suppressWarnings(
+  scaled_dist_gran(data_pick, "hour_day", 
+                   response = "general_supply_kwh",
+                   quantile_prob_val = c(0.5, 0.75, 0.9))) %>% rename("dist_hod" = "dist")
+v3 <- suppressWarnings(
+  scaled_dist_gran(data_pick, "day_month",
+                   response = "general_supply_kwh", 
+                   quantile_prob_val = c(0.5, 0.75, 0.9))) %>% rename("dist_dom" = "dist")
+
+data_dist <- v3 %>% 
+  left_join(v2) %>%
+  mutate(dist = dist_dom) %>% 
+    pivot_wider(-c(3, 4), 
+                names_from = customer_to,
               values_from = dist) %>%
   rename("customer_id" = "customer_from")
+```
 
+```{r}
 d = stats::as.dist(data_dist[,-1])
@@
 groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
-groups
+groups %>% group_by(group) %>% summarise(n=n())
+groups %>% filter(group==1)
 groups %>% filter(group==2)
 groups %>% filter(group==3)
-groups %>% group_by(group) %>% summarise(n=n())
 ```

### 2021-08-18 supplementary/algorithm3/real-data/handpick-18.Rmd
simulated and real data results update for algorithm 3

--- 2021-08-17 supplementary/handpick_distinct_cust.Rmd
@@
 ```{r read-data}
-data <- read_rds("../data/elec_nogap_2013_100.rds")
+data <- read_rds("../../../data/elec_nogap_2013_100.rds")
 ```
@@
 ```{r hand-pick}
-quantile_prob_val = seq(0.1, 0.9, 0.1)
 
@@
 
-elec_harmony_all <- read_rds("../data/elec_harmony-100-nogap-new.rds")%>%
+elec_harmony_all <- read_rds("../../../data/elec_harmony-100-nogap-new.rds")%>%
   filter(customer_id %in% data_pick$customer_id)
@@
 ```
+# quantile_prob_val = c(0.5, 0.75)
 
-```{r plot-hand-pick, fig.cap= "Median hourly demand drawn for few customers showing similar median behavior.
-Roughly speaking, Design 1 has one evening peak, Design 2 has two peaks and Design 3 has three peaks in a day.
-Each of design 1, 2 and 3 have six similar behaving customers resulting to 18 time series.
-We want our clustering results to group each of the designs together.", eval = FALSE}
-group2_data%>% 
-  ungroup() %>% 
-  left_join(elec_harmony_hod %>% select(customer_id, wpd)) %>% 
-  #filter(design==3) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-  ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
-  geom_line(size = 0.5) +
-  #ggtitle(paste("wpd:", elec_harmony_hod$wpd)) +
-  facet_wrap(design~customer_id, scales = "free_y", ncol = 6, labeller = "label_value") + theme_minimal()+ theme(strip.text.y = element_text(size = 5)) 
+## only hod
+
+```{r,echo = TRUE}
+quantile_prob_val = c(0.5, 0.75)
+#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
+library(gracsr)
+v2 <- suppressWarnings(
+  scaled_dist_gran(data_pick, "hour_day", 
+                   response = "general_supply_kwh",
+                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
+v3 <- suppressWarnings(
+  scaled_dist_gran(data_pick, "day_month",
+                   response = "general_supply_kwh", 
+                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")
+
+data_dist <- v3 %>% 
+  left_join(v2) %>%
+  mutate(dist = dist_hod) %>% 
+    pivot_wider(-c(3, 4), 
+                names_from = customer_to,
+              values_from = dist) %>%
+  rename("customer_id" = "customer_from")
 ```
 
-```{r allplot, message=FALSE, warning=FALSE, echo = FALSE}
+```{r}
+d = stats::as.dist(data_dist[,-1])
+hc = stats::hclust(d,method="complete")
 
-##----harmony-all
-# elec <- read_rds("../data/elec_nogap_2013_100.rds") %>% filter(customer_id %in% data_pick$customer_id)
-# 
-# elec_split = elec %>% group_split(customer_id)
-# harmonies <- read_rds("../../paper-hakear/paper/data/harmonies.rds")
-# 
-# harmonies_acro <- harmonies %>% mutate(facet_variable=NA, facet_levels=NA) %>% distinct()
-# 
-# elec_select_harmony = parallel::mclapply(1:18, function(x){
-#   
-#   data_id <-  elec_split %>% magrittr::extract2(x) %>% 
-#     as_tsibble(index = reading_datetime)
-# 
-#   
-#   k = hakear::select_harmonies(data_id,
-#                            harmony_tbl = harmonies_acro,
-#                            response = general_supply_kwh,
-#                            nperm = 200,
-#                            nsamp = 2
-#   ) %>% mutate(customer_id = unique(data_id$customer_id))
-#   #write_rds(k, paste("data/elec_harmony-100-nogap.rds"))
-# }, mc.cores = parallel::detectCores() - 1, mc.preschedule = FALSE, mc.set.seed = FALSE)
-# #toc()
-# elec_harmny <- elec_select_harmony %>% bind_rows() 
-# write_rds(elec_harmny, paste("../data/elec_harmony-100-nogap-new.rds"))
+groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
+groups %>% group_by(group) %>% summarise(n=n())
+#groups %>% filter(group==1)
+#groups %>% filter(group==2)
+#groups %>% filter(group==3)
 
-select_split <- str_split(elec_harmony_all$select_harmony, " ", simplify = TRUE)[,2]
+data_heatmap_new <- data_heatmap %>% left_join(groups)
 
-elec_sig_split <- elec_harmony_all %>%
-  mutate(facet_variable = case_when(
-    facet_variable == "hour_day" ~ "hod" ,
-    facet_variable == "day_month" ~ "dom" ,
-    facet_variable == "day_week" ~ "dow" ,
-    facet_variable == "week_month" ~ "wom" ,
-    facet_variable == "wknd_wday" ~ "wdwnd"
-  )) %>% 
-  mutate(x_variable = case_when(
-    x_variable == "hour_day" ~ "hod" ,
-    x_variable == "day_month" ~ "dom" ,
-    x_variable == "day_week" ~ "dow" ,
-    x_variable == "week_month" ~ "wom" ,
-    x_variable == "wknd_wday" ~ "wdwnd"
-  )) %>% 
-  bind_cols(select_split = select_split) %>% 
-  mutate(significant = case_when(
-    select_split == "***" ~ "highest",
-    select_split == "**" ~ "high",
-    select_split == "*" ~ "medium",
-    select_split == "" ~ "low"
-  )) %>%mutate(comb = x_variable)
+data_heatmap_new$group = as.factor(data_heatmap_new$group)
 
-elec_sig_split$significant <- 
-  factor(elec_sig_split$significant, levels = c("highest", "high", "medium", "low"))
-
-data_comb_filter = elec_sig_split %>% 
-mutate(low_proxy = if_else(significant=="low", 0, 1))%>% 
-select(comb, customer_id, low_proxy) %>% 
-pivot_wider(names_from = customer_id, values_from = low_proxy) %>%
-mutate(score = rowMeans(across(-1))) %>% 
-filter(score >0.3)
-
-data_sig_filter = elec_sig_split %>% 
-filter(comb %in% data_comb_filter$comb)
-
-data_clust = data_sig_filter %>% 
-select(comb, customer_id, wpd) %>% 
-pivot_wider(names_from = comb, values_from = wpd) 
-```
-
-```{r hc-plot, fig.cap="Dendogram from clustering 18 time series into three clusters."}
-hc = stats::hclust(dist(data_clust[-1]),method="complete")
-#plot(hc)
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 3, border = 2:10)
-```
-
-```{r conf-matrix, fig.cap="The confusion matrix showing how well the clustering algorithm worked."}
-groups <- cutree(hc, k=3) %>% 
-  as_tibble() %>% 
-mutate(customer_id = data_clust$customer_id)
-
-elec_pred_wo_scaled <- elec_harmony_all %>% 
-  filter(x_variable %in% c("hour_day", "day_month")) %>% 
-  left_join(data_pick %>% as_tibble() %>% select(customer_id, design)%>% unique) %>% select(-facet_variable, -facet_levels, -x_levels, -select_harmony) %>% 
-  pivot_wider(names_from = x_variable, values_from = wpd) %>% 
-  left_join(groups) %>% 
-  rename("pred_design" = "value") %>% 
-  arrange(day_month) %>% 
-  select(customer_id, design, pred_design, everything()) %>%
-  mutate(day_month = round(day_month, 1),
-         hour_day = round(hour_day, 1))
-
-elec_pred_wo_scaled$design = as.factor(elec_pred_wo_scaled$design)
-elec_pred_wo_scaled$pred_design = as.factor(elec_pred_wo_scaled$pred_design)
-xtab <- confusionMatrix(elec_pred_wo_scaled$pred_design, elec_pred_wo_scaled$design)
-xtab$table
-```
-
-```{r tab-cluster}
-# 
-#  fig.cap="The customers are arranged in ascending order of wpd for dom and it can be seen that the predicted designs are exactly sorted in that order."
-elec_pred_wo_scaled%>% 
-  kable(caption = "Actual and predicted allocation of designs to customers" )
+data_heatmap_new %>% 
+  ggplot(aes(x = as.integer(category))) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`), fill = "lightblue") +
+  geom_line(aes(y = `50%`, color = group), size = 0.7) +
+  facet_wrap(group~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+  theme_bw() +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
+  scale_color_brewer(palette = "Dark2")
 
@@
 
-\newpage
+## hod + dom
 
-##----cluster-characterization
+```{r,echo = TRUE}
+#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
+library(gracsr)
+v2 <- suppressWarnings(
+  scaled_dist_gran(data_pick, "hour_day", 
+                   response = "general_supply_kwh",
+                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
+v3 <- suppressWarnings(
+  scaled_dist_gran(data_pick, "day_month",
+                   response = "general_supply_kwh", 
+                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")
+
+data_dist <- v3 %>% 
+  left_join(v2) %>%
+  mutate(dist = dist_hod + dist_dom) %>% 
+    pivot_wider(-c(3, 4), 
+                names_from = customer_to,
+              values_from = dist) %>%
+  rename("customer_id" = "customer_from")
+```
 
 ```{r}
-data = data_clust %>% 
-left_join(groups)%>% 
-             rename ("group" = "value") %>% 
-             ungroup() %>% 
-             select(group, customer_id, everything()) %>% 
-             pivot_longer(-c(1,2), names_to = "comb", values_to = "wpd")
+d = stats::as.dist(data_dist[,-1])
+hc = stats::hclust(d,method="complete")
 
-data_pcp <- data %>% pivot_wider(
-              names_from = comb, 
-              values_from = wpd) %>% 
-  mutate(group = as.character(group)) 
+groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
+groups %>% group_by(group) %>% summarise(n=n())
+#groups %>% filter(group==1)
+#groups %>% filter(group==2)
+#groups %>% filter(group==3)
+
+data_heatmap_new <- data_heatmap %>% left_join(groups)
+
+data_heatmap_new$group = as.factor(data_heatmap_new$group)
+
+data_heatmap_new %>% 
+  ggplot(aes(x = as.integer(category))) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`), fill = "lightblue") +
+  geom_line(aes(y = `50%`, color = group), size = 0.7) +
+  facet_wrap(group~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+  theme_bw() +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
+  scale_color_brewer(palette = "Dark2")
+
 ```
 
-```{r ggpairs-plot, fig.cap="Pairs plot show that dom is the variable that is responsible for this clustering.
-The green colored cluster correspond to lowest value of dom, the orange colored and blue colored one correspond to middle range and high range of dom."}
-# ggpairs_nocolor <- ggpairs(data_pcp, columns = 3:7)
-# ggpairs_nocolor
+# quantile_prob_val = seq(0.1, 0.9, 0.1)
 
-groupge2 <- data %>% 
-  group_by(group) %>% 
-  summarise(n = length(unique(customer_id))) %>% 
-  filter(n>2)
+## only hod
 
-data_pcp_ge2 <- data_pcp %>% filter(group %in% groupge2$group)
+```{r,echo = TRUE}
+quantile_prob_val = seq(0.1, 0.9, 0.1)
+#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
+library(gracsr)
+v2 <- suppressWarnings(
+  scaled_dist_gran(data_pick, "hour_day", 
+                   response = "general_supply_kwh",
+                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
+v3 <- suppressWarnings(
+  scaled_dist_gran(data_pick, "day_month",
+                   response = "general_supply_kwh", 
+                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")
 
-# ggpairs_nocolor <- ggpairs(data_pcp_ge2, columns = 3:7)
-
-ggpairs_nocolor <- ggscatmat(data_pcp_ge2, columns=3:4) +
-  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")
-
-ggpairs_color <- ggscatmat(data_pcp_ge2, columns=3:4, color = "group") +
-  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")
-
-ggpairs_nocolor + ggpairs_color
+data_dist <- v3 %>% 
+  left_join(v2) %>%
+  mutate(dist = dist_hod) %>% 
+    pivot_wider(-c(3, 4), 
+                names_from = customer_to,
+              values_from = dist) %>%
+  rename("customer_id" = "customer_from")
 ```
 
-```{r pred-actual-design, fig.cap="First facet label indicate actual design, second denote predicted design, third denote customer_id."}
-group_data_orig <- group2_data %>% left_join(elec_pred_wo_scaled %>% select(customer_id,pred_design))
+```{r}
+d = stats::as.dist(data_dist[,-1])
+hc = stats::hclust(d,method="complete")
 
-# group_data_orig%>% 
-#   ungroup() %>% 
-#   filter(group==1) %>% 
-#   #filter(customer_id %in% c(8148781, 10703878)) %>% 
-#   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
-#   geom_line(size = 0.5) +
-#   facet_wrap(~customer_id, scales = "free_y",labeller = "label_value", ncol = 1) + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
-# 
-# 
-# group_data_orig%>% 
-#   ungroup() %>% 
-#   filter(group==2) %>% 
-#   #filter(customer_id %in% c(8148781, 10703878)) %>% 
-#   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
-#   geom_line(size = 0.5) +
-#   facet_wrap(~customer_id, scales = "free_y",labeller = "label_value", ncol = 1) + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
+groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
+groups %>% group_by(group) %>% summarise(n=n())
+#groups %>% filter(group==1)
+#groups %>% filter(group==2)
+#groups %>% filter(group==3)
 
-group_data_orig%>% 
-  ungroup() %>% 
-  rename("pred" = "pred_design", "cust" = "customer_id") %>% 
-  #filter(group==3) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-  ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
-  geom_line(size = 0.5) +
-  facet_wrap(design~pred~cust, scales = "free_y",labeller = "label_context") + theme_minimal()+ theme(strip.text.y = element_text(size = 5)) + 
-       theme(strip.text.x = element_text(size = 6, margin = margin(b = 0, t = 0)))+ xlab("hod")
+data_heatmap_new <- data_heatmap %>% left_join(groups)
+
+data_heatmap_new$group = as.factor(data_heatmap_new$group)
+
+data_heatmap_new %>% 
+  ggplot(aes(x = as.integer(category))) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`), fill = "lightblue") +
+  geom_line(aes(y = `50%`, color = group), size = 0.7) +
+  facet_wrap(group~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+  theme_bw() +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
+  scale_color_brewer(palette = "Dark2")
+
 ```
 
-```{r dom-pred, fig.cap="(Top) free y scale, (bottom) y scales same across facets seemingly implies that clustering is ordered by sorting the median demand."}
-group2_data <- data_pick %>% 
-  create_gran("day_month") %>% 
-  as_tibble() %>% 
-  #mutate(general_supply_kwh = scale(general_supply_kwh)) %>% 
-  group_by(design, day_month, customer_id) %>% 
-  summarise(median_kwh = median(general_supply_kwh))
+## hod + dom
 
-group_data_orig <- group2_data %>% left_join(elec_pred_wo_scaled %>% select(customer_id,pred_design))
+```{r,echo = TRUE}
+#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
+library(gracsr)
+v2 <- suppressWarnings(
+  scaled_dist_gran(data_pick, "hour_day", 
+                   response = "general_supply_kwh",
+                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
+v3 <- suppressWarnings(
+  scaled_dist_gran(data_pick, "day_month",
+                   response = "general_supply_kwh", 
+                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")
 
-group_data_orig %>% 
-  ungroup() %>% 
-  rename("pred" = "pred_design", "cust" = "customer_id") %>% 
-  #filter(group==3) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-  ggplot(aes(x = as.integer(day_month), y = median_kwh)) +
-  geom_line(size = 0.5) +
-  facet_wrap(pred~cust,
-             labeller = "label_context",
-             scales = "free_y") + 
-  theme_minimal()+ theme(strip.text.y = element_text(size = 5)) + scale_y_log10() + scale_x_continuous(breaks=seq(1, 31, 7), minor_breaks = seq(1, 31, 7))+ 
-       theme(strip.text.x = element_text(size = 6, margin = margin(b = 0, t = 0)))+ xlab("dom")
-# 
-group_data_orig$customer_id <-factor(group_data_orig$customer_id, levels = elec_pred_wo_scaled$customer_id)
-
-group_data_orig %>% 
-  ungroup() %>% 
-  rename("pred" = "pred_design", "cust" = "customer_id") %>% 
-  #filter(group==3) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-  ggplot(aes(x = as.integer(day_month), y = median_kwh)) +
-  geom_line(size = 0.5) +
-  facet_wrap(pred~cust,
-             labeller = "label_context") + 
-  theme_minimal()+ theme(strip.text.y = element_text(size = 5)) + scale_y_log10() + scale_x_continuous(breaks=seq(1, 31, 7), minor_breaks = seq(1, 31, 7))+ 
-       theme(strip.text.x = element_text(size = 6, margin = margin(b = 0, t = 0))) + xlab("dom")
+data_dist <- v3 %>% 
+  left_join(v2) %>%
+  mutate(dist = dist_hod + dist_dom) %>% 
+    pivot_wider(-c(3, 4), 
+                names_from = customer_to,
+              values_from = dist) %>%
+  rename("customer_id" = "customer_from")
 ```
 
-```{r, eval = FALSE}
-data_clust = data_sig_filter %>% 
-select(comb, customer_id, wpd) %>% 
-pivot_wider(names_from = comb, values_from = wpd) %>% 
-  mutate(hod = scale(hod),
-         dom = scale(dom))
+```{r}
+d = stats::as.dist(data_dist[,-1])
+hc = stats::hclust(d,method="complete")
 
-hc = stats::hclust(dist(data_clust[-1], method = "manhattan"),method="complete")
-#plot(hc)
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 3, border = 2:10)
+groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
+groups %>% group_by(group) %>% summarise(n=n())
+#groups %>% filter(group==1)
+#groups %>% filter(group==2)
+#groups %>% filter(group==3)
 
-groups_scaled <- cutree(hc, k=3) %>% 
-  as_tibble() %>% 
-mutate(customer_id = data_clust$customer_id) %>% rename("pred_design_scale" = "value")
+data_heatmap_new <- data_heatmap %>% left_join(groups)
 
-elec_harmony_all %>% 
-  filter(x_variable %in% c("hour_day", "day_month")) %>% 
-  left_join(data_pick %>% as_tibble() %>% select(customer_id, design)%>% unique) %>% select(-facet_variable, -facet_levels, -x_levels, -select_harmony) %>% 
-  pivot_wider(names_from = x_variable, values_from = wpd) %>% 
-  mutate(hod_scaled =  round(scale(hour_day), 1), 
-         dom_scaled = round(scale(day_month),1)) %>% 
-  left_join(groups_scaled)  %>%
-  mutate(day_month = round(day_month, 1),
-         hour_day = round(hour_day, 1)) %>% 
-  left_join(elec_pred_wo_scaled %>% select(customer_id, pred_design), by = "customer_id") %>% 
-  arrange(design) %>% 
-  select(customer_id, design, pred_design_scale, pred_design, everything()) %>% 
-  kable()
+data_heatmap_new$group = as.factor(data_heatmap_new$group)
 
-##----cluster-characterization
+data_heatmap_new %>% 
+  ggplot(aes(x = as.integer(category))) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`), fill = "lightblue") +
+  geom_line(aes(y = `50%`, color = group), size = 0.7) +
+  facet_wrap(group~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+  theme_bw() +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
+  scale_color_brewer(palette = "Dark2")
 
-data = data_clust %>% 
-left_join(groups)%>% 
-             rename ("group" = "value") %>% 
-             ungroup() %>% 
-             select(group, customer_id, everything()) %>% 
-             pivot_longer(-c(1,2), names_to = "comb", values_to = "wpd")
-
-data_pcp <- data %>% pivot_wider(
-              names_from = comb, 
-              values_from = wpd) %>% 
-  mutate(group = as.character(group)) 
 ```
 
-```{r, eval = FALSE}
-# ggpairs_nocolor <- ggpairs(data_pcp, columns = 3:7)
-# ggpairs_nocolor
+## remove the outlier and run again
 
-groupge2 <- data %>% 
-  group_by(group) %>% 
-  summarise(n = length(unique(customer_id))) %>% 
-  filter(n>2)
+```{r,echo = TRUE}
+quantile_prob_val = c(0.5, 0.75)
+data_pick <- data_pick %>% filter(!(customer_id %in% 8952846))
+library(gracsr)
+v2 <- suppressWarnings(
+  scaled_dist_gran(data_pick, "hour_day", 
+                   response = "general_supply_kwh",
+                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
+v3 <- suppressWarnings(
+  scaled_dist_gran(data_pick, "day_month",
+                   response = "general_supply_kwh", 
+                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")
 
-data_pcp_ge2 <- data_pcp %>% filter(group %in% groupge2$group)
-
-# ggpairs_nocolor <- ggpairs(data_pcp_ge2, columns = 3:7)
-
-ggpairs_nocolor <- ggscatmat(data_pcp_ge2, columns=3:4) +
-  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")
-
-ggpairs_color <- ggscatmat(data_pcp_ge2, columns=3:4, color = "group") +
-  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")
-
-ggpairs_nocolor + ggpairs_color
+data_dist <- v3 %>% 
+  left_join(v2) %>%
+  mutate(dist = dist_hod + dist_dom) %>% 
+    pivot_wider(-c(3, 4), 
+                names_from = customer_to,
+              values_from = dist) %>%
+  rename("customer_id" = "customer_from")
 ```
 
-# Algorithm discussed today
+```{r}
+d = stats::as.dist(data_dist[,-1])
+hc = stats::hclust(d,method="complete")
 
-- take significant granularities
-- consider sum of JS distances between each categories of each cyclic granularity
+groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
+groups %>% group_by(group) %>% summarise(n=n())
+#groups %>% filter(group==1)
+#groups %>% filter(group==2)
+#groups %>% filter(group==3)
+
+data_heatmap_new <- data_heatmap %>% filter(customer_id!=8952846) %>% left_join(groups)
+
+data_heatmap_new$group = as.factor(data_heatmap_new$group)
+
+data_heatmap_new %>% 
+  ggplot(aes(x = as.integer(category))) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`), fill = "lightblue") +
+  geom_line(aes(y = `50%`, color = group), size = 0.7) +
+  facet_wrap(group~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+  theme_bw() +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
+  scale_color_brewer(palette = "Dark2")
+
+```

### 2021-08-18 supplementary/algorithm4/real-data/handpick-18.Rmd
simulated and real data results update for algorithm 3

--- 2021-08-18 supplementary/algorithm3/real-data/handpick-18.Rmd
@@
 ```{r hand-pick}
+quantile_prob_val = seq(0.1, 0.9, 0.1)
 
@@
 
-# 
-# group2_data%>% 
-#   ungroup() %>% 
-#   filter(design==1) %>% 
-#   #filter(customer_id %in% c(8148781, 10703878)) %>% 
-#   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
-#   geom_line(size = 0.5) +
-#   facet_wrap(~customer_id, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
-# 
-# 
-# group2_data%>% 
-#   ungroup() %>% 
-#   filter(design==2) %>% 
-#   #filter(customer_id %in% c(8148781, 10703878)) %>% 
-#   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
-#   geom_line(size = 0.5) +
-#   facet_wrap(~customer_id, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
 ```
@@
                   ymax = `75%`), fill = "lightblue") +
+  #geom_smooth(aes(y = `50%`)) +
   geom_line(aes(y = `50%`), size = 0.7) +
@@
     theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")
+
+data_heatmap %>% 
+  ggplot(aes(x = as.integer(category))) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`), fill = "lightblue") +
+  geom_smooth(aes(y = `50%`)) +
+  #geom_line(aes(y = `50%`), size = 0.7, color = "orange") +
+    facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+  theme_bw() +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")
 ```
-# quantile_prob_val = c(0.5, 0.75)
-
-## only hod
-
 ```{r,echo = TRUE}
-quantile_prob_val = c(0.5, 0.75)
 #data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
@@
 v2 <- suppressWarnings(
-  scaled_dist_gran(data_pick, "hour_day", 
+  dist_pairwise_clust(data_pick, "hour_day", 
                    response = "general_supply_kwh",
-                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
+                   quantile_prob_val = c(0.5, 0.75))) 
 v3 <- suppressWarnings(
-  scaled_dist_gran(data_pick, "day_month",
+  dist_pairwise_clust(data_pick, "day_month",
                    response = "general_supply_kwh", 
-                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")
+                   quantile_prob_val = c(0.5, 0.75)))
 
-data_dist <- v3 %>% 
-  left_join(v2) %>%
-  mutate(dist = dist_hod) %>% 
-    pivot_wider(-c(3, 4), 
+data_dist <- v %>%
+    pivot_wider(
   names_from = customer_to,
@@
 ```{r}
-d = stats::as.dist(data_dist[,-1])
+d = as.dist(v3[,-1])
 hc = stats::hclust(d,method="complete")
 
-groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
+groups <- tibble(group = cutree(hc, k=3), customer_id = v2$customer_id)
 groups %>% group_by(group) %>% summarise(n=n())
-#groups %>% filter(group==1)
-#groups %>% filter(group==2)
-#groups %>% filter(group==3)
-
-data_heatmap_new <- data_heatmap %>% left_join(groups)
-
-data_heatmap_new$group = as.factor(data_heatmap_new$group)
-
-data_heatmap_new %>% 
-  ggplot(aes(x = as.integer(category))) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") +
-  geom_line(aes(y = `50%`, color = group), size = 0.7) +
-  facet_wrap(group~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-             ncol = 6) +
-  theme_bw() +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
-  scale_color_brewer(palette = "Dark2")
-
+groups %>% filter(group==1)
+groups %>% filter(group==2)
+groups %>% filter(group==3)
 ```
-
-## hod + dom
-
-```{r,echo = TRUE}
-#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
-library(gracsr)
-v2 <- suppressWarnings(
-  scaled_dist_gran(data_pick, "hour_day", 
-                   response = "general_supply_kwh",
-                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
-v3 <- suppressWarnings(
-  scaled_dist_gran(data_pick, "day_month",
-                   response = "general_supply_kwh", 
-                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")
-
-data_dist <- v3 %>% 
-  left_join(v2) %>%
-  mutate(dist = dist_hod + dist_dom) %>% 
-    pivot_wider(-c(3, 4), 
-                names_from = customer_to,
-              values_from = dist) %>%
-  rename("customer_id" = "customer_from")
-```
-
-```{r}
-d = stats::as.dist(data_dist[,-1])
-hc = stats::hclust(d,method="complete")
-
-groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
-groups %>% group_by(group) %>% summarise(n=n())
-#groups %>% filter(group==1)
-#groups %>% filter(group==2)
-#groups %>% filter(group==3)
-
-data_heatmap_new <- data_heatmap %>% left_join(groups)
-
-data_heatmap_new$group = as.factor(data_heatmap_new$group)
-
-data_heatmap_new %>% 
-  ggplot(aes(x = as.integer(category))) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") +
-  geom_line(aes(y = `50%`, color = group), size = 0.7) +
-  facet_wrap(group~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-             ncol = 6) +
-  theme_bw() +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
-  scale_color_brewer(palette = "Dark2")
-
-```
-
-# quantile_prob_val = seq(0.1, 0.9, 0.1)
-
-## only hod
-
-```{r,echo = TRUE}
-quantile_prob_val = seq(0.1, 0.9, 0.1)
-#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
-library(gracsr)
-v2 <- suppressWarnings(
-  scaled_dist_gran(data_pick, "hour_day", 
-                   response = "general_supply_kwh",
-                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
-v3 <- suppressWarnings(
-  scaled_dist_gran(data_pick, "day_month",
-                   response = "general_supply_kwh", 
-                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")
-
-data_dist <- v3 %>% 
-  left_join(v2) %>%
-  mutate(dist = dist_hod) %>% 
-    pivot_wider(-c(3, 4), 
-                names_from = customer_to,
-              values_from = dist) %>%
-  rename("customer_id" = "customer_from")
-```
-
-```{r}
-d = stats::as.dist(data_dist[,-1])
-hc = stats::hclust(d,method="complete")
-
-groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
-groups %>% group_by(group) %>% summarise(n=n())
-#groups %>% filter(group==1)
-#groups %>% filter(group==2)
-#groups %>% filter(group==3)
-
-data_heatmap_new <- data_heatmap %>% left_join(groups)
-
-data_heatmap_new$group = as.factor(data_heatmap_new$group)
-
-data_heatmap_new %>% 
-  ggplot(aes(x = as.integer(category))) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") +
-  geom_line(aes(y = `50%`, color = group), size = 0.7) +
-  facet_wrap(group~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-             ncol = 6) +
-  theme_bw() +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
-  scale_color_brewer(palette = "Dark2")
-
-```
-
-## hod + dom
-
-```{r,echo = TRUE}
-#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
-library(gracsr)
-v2 <- suppressWarnings(
-  scaled_dist_gran(data_pick, "hour_day", 
-                   response = "general_supply_kwh",
-                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
-v3 <- suppressWarnings(
-  scaled_dist_gran(data_pick, "day_month",
-                   response = "general_supply_kwh", 
-                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")
-
-data_dist <- v3 %>% 
-  left_join(v2) %>%
-  mutate(dist = dist_hod + dist_dom) %>% 
-    pivot_wider(-c(3, 4), 
-                names_from = customer_to,
-              values_from = dist) %>%
-  rename("customer_id" = "customer_from")
-```
-
-```{r}
-d = stats::as.dist(data_dist[,-1])
-hc = stats::hclust(d,method="complete")
-
-groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
-groups %>% group_by(group) %>% summarise(n=n())
-#groups %>% filter(group==1)
-#groups %>% filter(group==2)
-#groups %>% filter(group==3)
-
-data_heatmap_new <- data_heatmap %>% left_join(groups)
-
-data_heatmap_new$group = as.factor(data_heatmap_new$group)
-
-data_heatmap_new %>% 
-  ggplot(aes(x = as.integer(category))) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") +
-  geom_line(aes(y = `50%`, color = group), size = 0.7) +
-  facet_wrap(group~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-             ncol = 6) +
-  theme_bw() +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
-  scale_color_brewer(palette = "Dark2")
-
-```
-
-## remove the outlier and run again
-
-```{r,echo = TRUE}
-quantile_prob_val = c(0.5, 0.75)
-data_pick <- data_pick %>% filter(!(customer_id %in% 8952846))
-library(gracsr)
-v2 <- suppressWarnings(
-  scaled_dist_gran(data_pick, "hour_day", 
-                   response = "general_supply_kwh",
-                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
-v3 <- suppressWarnings(
-  scaled_dist_gran(data_pick, "day_month",
-                   response = "general_supply_kwh", 
-                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")
-
-data_dist <- v3 %>% 
-  left_join(v2) %>%
-  mutate(dist = dist_hod + dist_dom) %>% 
-    pivot_wider(-c(3, 4), 
-                names_from = customer_to,
-              values_from = dist) %>%
-  rename("customer_id" = "customer_from")
-```
-
-```{r}
-d = stats::as.dist(data_dist[,-1])
-hc = stats::hclust(d,method="complete")
-
-groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
-groups %>% group_by(group) %>% summarise(n=n())
-#groups %>% filter(group==1)
-#groups %>% filter(group==2)
-#groups %>% filter(group==3)
-
-data_heatmap_new <- data_heatmap %>% filter(customer_id!=8952846) %>% left_join(groups)
-
-data_heatmap_new$group = as.factor(data_heatmap_new$group)
-
-data_heatmap_new %>% 
-  ggplot(aes(x = as.integer(category))) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") +
-  geom_line(aes(y = `50%`, color = group), size = 0.7) +
-  facet_wrap(group~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-             ncol = 6) +
-  theme_bw() +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
-  scale_color_brewer(palette = "Dark2")
-
-```

### 2021-08-18 supplementary/algorithm1/real-data/cluster4-iter100.Rmd
cleaning folders

--- 2021-07-08 supplementary/cluster4designs.Rmd
@@
 
-## nobs per combination is 4
-
-```{r, echo = TRUE}
-ntimes_val = 4
-```
-
-```{r data}
-```
-
-```{r category-plots}
-```
-
-```{r}
-change_index <- function(data){
-
-index_new <- map(seq_len(ntimes_val), function(i){
-  map((seq_len(nx_val*nfacet_val)), function(j)
-    {
-value = i + (j-1)*ntimes_val
-})
-  }) %>% unlist()
-
-data_new = data %>%
-  ungroup() %>%
-  mutate(index_old = row_number(),
-         index_new = index_new)
-
-y = data_new[match(index_new, data_new$index_old),]
-
-y <- y %>%
-  mutate(time = row_number())
-
-return(y)
-}
-
-endbreaks<- nrow(sim_panel_null)
-
-p1 <- change_index(sim_panel_null) %>%
-  ggplot(aes(x = time,
-             y = sim_data)) +
-  geom_line() +
-  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
-  theme_bw() +
-  geom_point(alpha = 0.5, color = "blue") +
-  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
-        panel.grid.minor.x =  element_blank())+
-  ylab("simulated response")+
-  xlab("index")
-
-p2 <- change_index(sim_panel_varf) %>%
-  ggplot(aes(x = time,
-             y = sim_data)) +
-  geom_line() +
-  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
-  theme_bw() +
-  geom_point(alpha = 0.5, color = "blue") +
-  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
-        panel.grid.minor.x =  element_blank())+
-  ylab("simulated response")+
-  xlab("index")
-
-p3 <- change_index(sim_panel_varx) %>%
-  ggplot(aes(x = time,
-             y = sim_data)) +
-  geom_line() +
-  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
-  theme_bw() +
-  geom_point(alpha = 0.5, color = "blue")+
-  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
-        panel.grid.minor.x =  element_blank())+
-  ylab("simulated response")+
-  xlab("index")
-
-p4 <- change_index(sim_panel_varall) %>%
-  ggplot(aes(x = time,
-             y = sim_data)) +
-  geom_line() +
-  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
-  theme_bw() +
-  geom_point(alpha = 0.5, color = "blue") +
-  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
-  panel.grid.minor.x =  element_blank())+
-  ylab("simulated response")+
-  xlab("index")
-
-t1 <- sim_panel_varall %>% ungroup() %>%  select (-c(1, 2))
-
-# %>% 
-#   kable(format = "latex", caption = "Simulated data for categories format")
-
-t2 <- change_index(sim_panel_varall) %>% select (-c(1, 2))
-
-# %>% 
-#   kable(format = "latex",
-#         caption = "Simulated data for time series format")
-
-kable(list(t1, t2), format = "latex", caption = "Simulated data across combination of categories (left) and manipulated time series format (right)") %>% 
-  row_spec(1:24,  background =  "#d3d3d3")
-
-# kable(t1) %>%
-#   kable_styling(full_width = FALSE, position = "float_left")
-# kable(t2) %>%
-#   kable_styling(full_width = FALSE, position = "left")
-```
-
-```{r plot4}
-(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)
-```
-
-## nobs per combination is 10
-
-```{r nobs10, echo = TRUE}
-ntimes_val = 10 # nobs per combination 
-```
-
-```{r data}
-```
-
-```{r category-plots}
-```
-
-```{r line-graph}
-```
-
-```{r plot10}
-(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)
-```
-
 ## nobs per combination is 50
@@
 ```{r, echo=TRUE}
-sample_seed <-  seq(10, 100, 10)
+sample_seed <-  seq(1, 100, 1)
 ```
@@
 ```{r data-make}
-data_null <- map(seq_len(sample_seed), function(seed){
+data_null <- map(sample_seed, function(seed){
   set.seed(sample_seed[seed])
@@
 
-data_varf <- map(seq_len(sample_seed), function(seed){
+data_varf <- map(sample_seed, function(seed){
   set.seed(sample_seed[seed])
@@
 
-data_varx <- map(seq_len(sample_seed), function(seed){
+data_varx <- map(sample_seed, function(seed){
   set.seed(sample_seed[seed])
@@
 
-data_varall <- map(seq_len(sample_seed), function(seed){
+data_varall <- map(sample_seed, function(seed){
   set.seed(sample_seed[seed])
@@
   geom_tile(aes(fill = sim_data_quantile)) +
-  scale_fill_distiller(palette = "YlGnBu")+ 
+  scale_fill_distiller(palette = "YlGnBu", direction = 1)+ 
   facet_wrap(~category, ncol = 1) +

### 2021-08-18 supplementary/algorithm1/real-data/realdata-cust10-noiter.Rmd
cleaning folders

--- 2021-07-16 realdata-cust10-noiter-scaled.Rmd
@@
 ```{r}
-
-elec_ts <- read_rds("../data/elec_ts.rds")
-
-set.seed(12345)
-
-sm_100 <- elec_ts %>%
-  as_tibble() %>% 
-  distinct(customer_id) %>%
-  slice_sample(n = 100)
-
-sm <- elec_ts %>% 
-  filter(customer_id %in% sm_100$customer_id) %>% 
-  ungroup()
-
+sm <- gravitas::smart_meter10 
 sm_scaled <- sm %>%
@@
   ungroup()
+```
+
+```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for 10 households (facets).
+The data is squeezed stopping us from seeing any behavioral patterns."}
+library(lubridate)
+
+sm_scaled %>% 
+  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
+  geom_line() +
+  facet_wrap(~customer_id, ncol = 1, 
+             strip.position = "right") +
+  theme_bw() +
+  xlab("time") +
+  ylab("demand (in Kwh)")
+```
+
+```{r raw-sub, fig.cap="The raw plots for ten customers shown for Sept 2013-Oct 2013.
+The data is zoomed in and the y-scales made free to emphasize weekly, hourly or any behavior/patterns that they might have.
+There is some daily and weekly pattern in most households."}
+sm %>% 
+  filter(year(reading_datetime) == 2013 &
+           month(reading_datetime) %in% c(9, 10)) %>% 
+  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
+  geom_line() +
+  facet_wrap(~customer_id, ncol = 1, scales = "free_y", 
+             strip.position = "right") +
+  theme_bw()+
+  ylab("demand (in Kwh)") +
+  theme(panel.grid.major.x = element_blank()) +
+  scale_x_datetime("Date", date_labels = "%b %d",
+                   breaks = "1 week",
+                   date_minor_breaks = "1 day")  + 
+  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
+        panel.grid.minor.x =  element_line(colour = "#D3D3D3"))
+
 ```

### 2021-08-18 supplementary/algorithm1/real-data/realdata-cust2-iter5.Rmd
cleaning folders

--- 2021-07-08 supplementary/realdata-cust5.Rmd
@@
 sm <- gravitas::smart_meter10 %>%
-  filter(customer_id %in% c("10006704", "10017936", "10006414", "10018250", "10018064"))
+  filter(customer_id %in% c("10006704", "10017936"))
 ```
@@
 The data is squeezed stopping us from seeing any behavioral patterns."}
+
 library(lubridate)
@@
   geom_line() +
-  facet_wrap(~customer_id) +
+  facet_wrap(~customer_id, nrow = 2) +
   theme_bw() + 
@@
 ```
-
-_k = 3_
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 3, border = 2:5)
-```
-
-_k = 4_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 4, border = 2:5)
-```
-
-_k = 5_
-
-```{r}
-plot(hc, cex = 0.6)
-rect.hclust(hc, k = 5, border = 2:6)
-```

### 2021-08-18 supplementary/algorithm1/real-data/realdata-cust2.Rmd
cleaning folders

--- 2021-07-08 supplementary/realdata-cust3.Rmd
@@
 ---
-title: "Clustering real data with three customers and choosing optimal number of clusters"
+title: "Clustering real data with two customers and choosing optimal number of clusters"
 output:
@@
 
-four households are considered for the following analysis from the SGSC data set.
+two households are considered for the following analysis from the SGSC data set.
 The data sets contain three columns `customer_id`, `reading_datetime` and `general_supply_kwh`.
 They consist of half-hourly data from 2012-2014.
-Figure \ref{fig:raw-all} shows how the raw time plots for these four households look like.
+Figure \ref{fig:raw-all} shows how the raw time plots for these two households look like.
 Since all the dataset looks squeezed on this linear time scale, Figure \ref{fig:raw-sub} shows how the raw plot looks for 2 months (Sept 2013 and Oct 2013).
@@
 sm <- gravitas::smart_meter10 %>%
-  filter(customer_id %in% c("10006704", "10017936", "10018250"))
+  filter(customer_id %in% c("10006704", "10017936"))
 ```
 
-```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for four households (facets).
+```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for two households (facets).
 The data is squeezed stopping us from seeing any behavioral patterns."}
@@
 
-The distribution of demand across different hours of the day observed for these four households through heatmaps (Figure \ref{fig:hod-deciles}) and quantile plots (Figure \ref{fig:hod-heatmap}).
+The distribution of demand across different hours of the day observed for these two households through heatmaps (Figure \ref{fig:hod-deciles}) and quantile plots (Figure \ref{fig:hod-heatmap}).
 The following characterization is done to validate results of the clustering approach later on.
 
-```{r hod-box, fig.cap="The raw plots for four households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
+```{r hod-box, fig.cap="The raw plots for two households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
 sm %>% 
@@
 
-```{r hod-deciles, fig.cap = "The deciles plots for four households are shown for entire observation period across hours of the day.
+```{r hod-deciles, fig.cap = "The deciles plots for two households are shown for entire observation period across hours of the day.
 The y-scales are made free to allow us to see daily patterns for each households.
@@
 
-```{r raw-iter, fig.cap = "four iterations for each customer are considered by adding random noise $(N(0, \\sigma^2)$, where $\\sigma^2$ is very small relative to the the variance of the data.
+```{r raw-iter, fig.cap = "two iterations for each customer are considered by adding random noise $(N(0, \\sigma^2)$, where $\\sigma^2$ is very small relative to the the variance of the data.
 The raw plots for all simulated dataset is shown to make sure they look similar to the structure."}
@@
 
-_k = 3_
+_k = 2_
 ```{r}
 plot(hc, cex = 0.6)
-rect.hclust(hc, k = 3, border = 2:5)
+rect.hclust(hc, k = 2, border = 2:5)
 ```

### 2021-08-18 supplementary/algorithm2/3cat-same-lev.Rmd
cleaning folders

---
title: "Simulation for wpd algorithm"
author: "Sayani Gupta"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(distributional)
```

```{r}
T <- seq(0, 500, 1)
g1 <- T %%2
g2 <- T %%3

g3 <- T %%4
data <- tibble::tibble(T, g1, g2, g3)

generate_dist <- function(x){
  if(x==0) 
    sim_dist <- distributional::dist_normal(0,1)
  if(x==1)
    sim_dist <- distributional::dist_normal(2,1)
  if(x==2)
    sim_dist <- distributional::dist_normal(4,1)
  if(x==3)
    sim_dist <- distributional::dist_normal(6,1)
  return(sim_dist)
}

set.seed(12345)

y <- data %>% group_split(g1)
sim_data_g1 <- map(seq_len(length(y)), 
      function(z){
  univec =  y %>% magrittr::extract2(z,) %>% select(T, g1)
  value = univec %>% distinct(g1) %>% pull(g1)
  return_value = generate_dist(value) %>% distributional::generate(nrow(univec)) %>% unlist() %>% as_tibble() %>% set_names("value") %>% 
    bind_cols(univec)
  }) %>% bind_rows() %>% select(T, g1, value)

g2_value = generate(dist_normal(mu = 0, sigma = 1), times = nrow(sim_data_g1)) %>% unlist() %>% as_tibble() %>% set_names("g2_value")

g3_value = generate(dist_normal(mu = 2, sigma = 1), times = nrow(sim_data_g1)) %>% unlist() %>% as_tibble() %>% set_names("g3_value")

total_data <- sim_data_g1 %>% 
  arrange(T) %>% 
  left_join(data) %>% 
  rename("g1_value" = "value") %>% 
  bind_cols(g2_value, g3_value) %>% 
  select(T, g1, g2, g3, everything()) %>% 
  group_by(g1, g2, g3) %>% 
  mutate(sim_data = sum(g1_value, g2_value, g3_value))

ggplot(total_data, aes(x = as.factor(g1), y = sim_data)) + geom_boxplot()
ggplot(total_data, aes(x = as.factor(g2), y = sim_data)) + geom_boxplot()
ggplot(total_data, aes(x = as.factor(g3), y = sim_data)) + geom_boxplot()

```

```{r}
y <- data %>% group_split(g2)

sim_data_g2 <- map(seq_len(length(y)), 
      function(z){
  univec =  y %>% magrittr::extract2(z,) %>% select(T, g2)
  value = univec %>% distinct(g2) %>% pull(g2)
  return_value = generate_dist(value) %>% distributional::generate(nrow(univec)) %>% unlist() %>% as_tibble() %>% set_names("value") %>% 
    bind_cols(univec)
  }) %>% bind_rows() %>% select(T, g2, value)

ggplot(sim_data_g2, aes(x = as.factor(g2), y = value)) + geom_boxplot()
```

```{r}
y <- data %>% group_split(g3)

sim_data_g3 <- map(seq_len(length(y)), 
      function(z){
  univec =  y %>% magrittr::extract2(z,) %>% select(T, g3)
  value = univec %>% distinct(g3) %>% pull(g3)
  return_value = generate_dist(value) %>% distributional::generate(nrow(univec)) %>% unlist() %>% as_tibble() %>% set_names("value") %>% 
    bind_cols(univec)
  }) %>% bind_rows() %>% select(T, g3, value)

ggplot(sim_data_g3, aes(x = as.factor(g3), y = value)) + geom_boxplot()
```
# sum up the normal distribution approach

```{r}
set.seed(12345)
dist <- dist_normal(mu = c(0,4,4,8), sigma = 1)
numbers <- generate(dist, times = 100)

timeseries <- numbers%>% unlist() %>% matrix(ncol = 4, nrow = 100, byrow = FALSE) %>% t() %>% 
  matrix(ncol=1, byrow = TRUE)

timeseries <- as_tibble(timeseries) %>% set_names("value") %>% bind_cols(t = seq(0, 399, 1))
ggplot(timeseries, aes(x = t, y = value)) + geom_line()

t <- seq(0,399, 1)
g1 <- t %%2
g2 <- t %%3

g3 <- t %%4
data <- tibble::tibble(t, g1, g2, g3) %>% left_join(timeseries)

ggplot(data, aes(x = as.factor(g1), y = value)) + geom_boxplot()
ggplot(data, aes(x = as.factor(g2), y = value)) + geom_boxplot()
ggplot(data, aes(x = as.factor(g3), y = value)) + geom_boxplot()

```

```{r}
set.seed(12345)
dist <- dist_normal(mu = c(0,2,4), sigma = 1)
numbers <- generate(dist, times = 100)

timeseries <- numbers%>% unlist() %>% matrix(ncol = 3, nrow = 100, byrow = FALSE) %>% t() %>% 
  matrix(ncol=1, byrow = TRUE)

timeseries <- as_tibble(timeseries) %>% set_names("value") %>% bind_cols(t = seq(0, 299, 1))
ggplot(timeseries, aes(x = t, y = value)) + geom_line()

t <- seq(0,299, 1)
g1 <- t %%2
g2 <- t %%3

g3 <- t %%4
data <- tibble::tibble(t, g1, g2, g3) %>% left_join(timeseries)

ggplot(data, aes(x = as.factor(g1), y = value)) + geom_boxplot()
ggplot(data, aes(x = as.factor(g2), y = value)) + geom_boxplot()
ggplot(data, aes(x = as.factor(g3), y = value)) + geom_boxplot()
```

### 2021-08-18 supplementary/algorithm2/real-data/algo2-elec100-diffscale.Rmd
cleaning folders

--- 2021-08-05 supplementary/algo2-elec100.Rmd
@@
 ```{r}
-# ggpairs_nocolor <- ggpairs(data_pcp, columns = 3:7)
-# ggpairs_nocolor
+ggpairs_nocolor <- ggpairs(data_pcp, columns = 3:7)
+ggpairs_nocolor
 
@@
 
-# ggpairs_nocolor <- ggpairs(data_pcp_ge2, columns = 3:7)
-
-ggpairs_nocolor <- ggscatmat(data_pcp_ge2, columns=3:7) +
-  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")
+ggpairs_nocolor <- ggpairs(data_pcp_ge2, columns = 3:7)
 
 ggpairs_color <- ggscatmat(data_pcp_ge2, columns=3:7, color = "group") +
-  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")
+  scale_colour_brewer(palette="Set2")
+ggpairs_color
 
-ggpairs_nocolor
-```
-
-```{r}
-ggpairs_color
-```
-
-```{r}
 # 
@@
   ylab("wpd")
+
+c(6, 7, 3, 4, 5)
 # 
@@
 
-```
-
-_Regular scale_
-
-```{r}
 (p1 + p2)/ (p3 + p4)
 (p5 + p6)/(p7 +p8)
+
 ```
-
-_Same scale_
-
-```{r}
-(p1 +ylim(0,3) + p2 +ylim(0,3))/ (p3+ylim(0,3) + p4+ylim(0,3))
-(p5 +ylim(0,3) + p6 +ylim(0,3))/ (p7+ylim(0,3) + p8+ylim(0,3))
-```
-# PCA based clustering
-
-```{r}
-pc <- prcomp(data_pcp[c(-1,-2)], center = F, scale = F)
-summary(pc)
-# First two principal componentsafc
-comp <- data.frame(pc$x[,1:2])
-
-#plot(comp, pch=16, col=rgb(0,0,0,0.5))
-
-hc = stats::hclust(dist(comp))
-data_pcp_all <- data_pcp %>% bind_cols(comp)
-
-library(RColorBrewer)
-library(scales)
-data_pcp_all %>% ggplot(aes(x=PC1, y = PC2)) + geom_point(aes(color = group)) +
-  scale_color_brewer(palette = "Dark2") + theme_bw() + theme(legend.position = "bottom")
-
-data_pcp_all %>%
-  pivot_longer(-c(1, 2,8, 9), names_to = "variables", values_to = "wpd") %>% 
-  ggplot(aes(x=wpd, y = PC1)) + geom_point(aes(color = group)) +
-  scale_color_brewer(palette = "Dark2") + theme_bw() +
-  facet_wrap(~variables, nrow = 1) + theme(legend.position = "bottom")
-```
-
-One dimensional scale since PC1 explains almost all variation?

### 2021-08-18 supplementary/algorithm2/real-data/algo2-elec100-onegran.Rmd
cleaning folders

--- 2021-08-05 supplementary/algo2-elec100.Rmd
@@
 ##----harmony-all
-elec_harmony_all <- read_rds("../data/algo2-rawdata-wpd.rds")
+elec_harmony_all <- read_rds("../data/elec_harmony-100-nogap-onegran.rds")
 
@@
     select_split == "" ~ "low"
-  )) %>%mutate(comb = paste(facet_variable, x_variable, sep = "-"))
+  )) %>%mutate(comb = x_variable)
 
@@
 
-ggpairs_nocolor <- ggscatmat(data_pcp_ge2, columns=3:7) +
+ggpairs_nocolor <- ggscatmat(data_pcp_ge2, columns=3:6) +
   scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")
 
-ggpairs_color <- ggscatmat(data_pcp_ge2, columns=3:7, color = "group") +
+ggpairs_color <- ggscatmat(data_pcp_ge2, columns=3:6, color = "group") +
   scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")
@@
 sm_hod_dow <- elec_scaled %>% 
-gravitas::create_gran("hour_day") %>% 
-  gravitas::create_gran("day_week")
+gravitas::create_gran("hour_day")
 
@@
   as_tibble() %>% 
-  arrange(group, day_week, hour_day) %>% 
-  mutate(category = paste(hour_day, day_week, sep = "-")) %>% 
+  arrange(group, hour_day) %>% 
+  mutate(category = hour_day) %>% 
    select(group, category, general_supply_kwh) %>% 
@@
 elec %>% as_tibble() %>% group_by(group) %>% summarise(n = length(unique(customer_id)))
+```
 
+```{r}
 quantile_prob_val = c(0.25, 0.5, 0.75)
 
-p1 <- sm_hod_dow %>% 
-  filter(group==1) %>% 
-  gravitas::prob_plot(
-     "day_week",
-     "hour_day",
-  response = "general_supply_kwh",
-  plot_type = "quantile", nrow=1) +
-  scale_y_log10() +
-    ggtitle("Group 1") + theme_bw()  +
-  theme(legend.position = "none",
-        strip.text = element_text(size = 10, margin = margin())) + 
-  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))
+sm_hod <- gravitas::create_gran(sm, "hour_day") %>% 
+  mutate(customer_id = as.integer(customer_id))
 
-p2 <- sm_hod_dow %>% 
-  filter(group==2 ) %>% 
-  gravitas::prob_plot(  "day_week",
-     "hour_day",
-  response = "general_supply_kwh",
-  plot_type = "quantile", nrow=1) +
-  scale_y_log10()+
-    ggtitle("Group 2")+ theme_bw()  +
-  theme(legend.position = "none",
-        strip.text = element_text(size = 10, margin = margin())) + 
-  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))
+sm_hod_list <- sm_hod %>% 
+  as_tibble() %>% 
+  select(customer_id, hour_day, general_supply_kwh) %>% 
+  pivot_wider(names_from = hour_day,
+              values_from = general_supply_kwh)
 
-p3 <- sm_hod_dow %>% 
-  filter(group==3) %>% 
-  gravitas::prob_plot(  "day_week",
-     "hour_day",
-  response = "general_supply_kwh",
-  plot_type = "quantile", nrow=1) +
-  scale_y_log10() +
-    ggtitle("Group 3" )+ theme_bw()  +
-  theme(legend.position = "none",
-        strip.text = element_text(size = 10, margin = margin())) + 
-  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))
+groups_ref <- groups %>% as_tibble %>% 
+              mutate(customer_serial_id = sm_hod_list$customer_id) %>% 
+             rename ("group" = "value")
 
-p4 <- sm_hod_dow %>% 
-  filter(group==4) %>% 
-  gravitas::prob_plot("day_week",
-     "hour_day",
-  response = "general_supply_kwh",
-  plot_type = "quantile", nrow=1) +
-  scale_y_log10() +
-    ggtitle("Group 4")+ theme_bw()  +
-  theme(legend.position = "none",
-        strip.text = element_text(size = 10, margin = margin())) + 
-  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))
+sm_hod_list_cat <-  sm_hod_list %>% left_join(groups_ref, by = c("customer_id" = "customer_serial_id")) %>% select(group, everything())
 
-p5 <- sm_hod_dow %>% 
-  filter(group==5) %>% 
-  gravitas::prob_plot("day_week",
-     "hour_day",
-  response = "general_supply_kwh",
-  plot_type = "quantile", nrow=1) +
-  scale_y_log10() +
-    ggtitle("Group 5")+ theme_bw()  +
-  theme(legend.position = "none",
-        strip.text = element_text(size = 10, margin = margin())) + 
-  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))
+ncol_sm <- seq_len(ncol(sm_hod_list_cat[-c(1, 2)]))
+nrow_sm <- unique(sm_hod_list_cat$group)
 
-p6 <- sm_hod_dow %>% 
-  filter(group==6) %>% 
-  gravitas::prob_plot("day_week",
-     "hour_day",
-  response = "general_supply_kwh",
-  plot_type = "quantile", nrow=1) +
-  scale_y_log10() +
-    ggtitle("Group 6")+ theme_bw()  +
-  theme(legend.position = "none",
-        strip.text = element_text(size = 10, margin = margin())) + 
-  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))
+sm_hod_quantiles_cat <- map(nrow_sm, function(x){
+  map(ncol_sm, function(y){
+   cell <- sm_hod_list_cat%>% 
+     dplyr::filter(group == x) %>% 
+     select(-c(1, 2)) %>% 
+     extract(y) %>% 
+     unlist()
+   quantile(cell, prob = quantile_prob_val)
+})  %>% bind_rows(.id = "categories_serial_id")
+}) %>% bind_rows(.id = "group_id") %>% 
+  mutate(group_id = as.integer(group_id),
+         categories_serial_id = as.integer(categories_serial_id))
 
-p7 <- sm_hod_dow %>% 
-  filter(group==7) %>% 
-  gravitas::prob_plot("day_week",
-     "hour_day",
-  response = "general_supply_kwh",
-  plot_type = "quantile", nrow=1) +
-  scale_y_log10() +
-    ggtitle("Group 7")+ theme_bw()  +
-  theme(legend.position = "none",
-        strip.text = element_text(size = 10, margin = margin())) + 
-  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))
+categories_ref_cat <- names(sm_hod_list_cat)[-c(1, 2)] %>% as_tibble() %>% 
+  mutate(categories_serial_id = row_number())
 
-p8 <- sm_hod_dow %>% 
-  filter(group==8) %>% 
-  gravitas::prob_plot("day_week",
-     "hour_day",
-  response = "general_supply_kwh",
-  plot_type = "quantile", nrow=1) +
-  scale_y_log10() +
-    ggtitle("Group 8")+ theme_bw()  +
-  theme(legend.position = "none",
-        strip.text = element_text(size = 10, margin = margin())) + 
-  theme(panel.spacing =unit(0, "lines")) + ylab("")+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))
-
+data_heatmap <- sm_hod_quantiles_cat %>% 
+  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
+  left_join(categories_ref_cat) %>% 
+  rename("category" = "value") %>% 
+  select(-categories_serial_id) %>% 
+  mutate(category = as.numeric(category))
 ```
 
-_Regular scale_
-
-```{r}
-(p1 + p2)/ (p3 + p4)
-(p5 + p6)/(p7 +p8)
-```
-
-_Same scale_
-
-```{r}
-(p1 +ylim(0,3) + p2 +ylim(0,3))/ (p3+ylim(0,3) + p4+ylim(0,3))
-(p5 +ylim(0,3) + p6 +ylim(0,3))/ (p7+ylim(0,3) + p8+ylim(0,3))
-```
-# PCA based clustering
-
-```{r}
-pc <- prcomp(data_pcp[c(-1,-2)], center = F, scale = F)
-summary(pc)
-# First two principal componentsafc
-comp <- data.frame(pc$x[,1:2])
-
-#plot(comp, pch=16, col=rgb(0,0,0,0.5))
-
-hc = stats::hclust(dist(comp))
-data_pcp_all <- data_pcp %>% bind_cols(comp)
-
-library(RColorBrewer)
-library(scales)
-data_pcp_all %>% ggplot(aes(x=PC1, y = PC2)) + geom_point(aes(color = group)) +
-  scale_color_brewer(palette = "Dark2") + theme_bw() + theme(legend.position = "bottom")
-
-data_pcp_all %>%
-  pivot_longer(-c(1, 2,8, 9), names_to = "variables", values_to = "wpd") %>% 
-  ggplot(aes(x=wpd, y = PC1)) + geom_point(aes(color = group)) +
-  scale_color_brewer(palette = "Dark2") + theme_bw() +
-  facet_wrap(~variables, nrow = 1) + theme(legend.position = "bottom")
-```
-
-One dimensional scale since PC1 explains almost all variation?
+```{r cluster-characterization}
+data_heatmap %>% 
+  pivot_wider( names_from = quantile_prob, 
+               values_from = quantile_val) %>% 
+  ggplot(aes(x = category)) +  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`), fill = "lightblue") +
+  geom_line(aes(y = `50%`)) +
+  facet_wrap(~group_id, ncol = 1, scales = "free_y", labeller = "label_both") +
+  theme_bw() +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))

### 2021-08-18 supplementary/algorithm2/real-data/algo2-elec100.Rmd
cleaning folders

--- 2021-08-05 supplementary/algo2-elec100.Rmd
@@
 ```
+
 # PCA based clustering
@@
 One dimensional scale since PC1 explains almost all variation?
+
+#Group2: n=24
+```{r}
+group2_data <- sm_hod_dow %>% 
+  as_tibble() %>% 
+  filter(group==2) %>% 
+  group_by(day_week, hour_day, customer_id) %>% 
+  summarise(median_kwh = median(general_supply_kwh))
+
+group2_data%>% 
+  ungroup() %>% 
+  #filter(customer_id %in% c(8148781, 10703878)) %>% 
+  ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
+  geom_line(size = 0.5) +
+  facet_grid(customer_id~day_week, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
+```
+
+#Group4: n=24
+```{r}
+group2_data <- sm_hod_dow %>% 
+  as_tibble() %>% 
+  filter(group==4) %>% 
+  group_by(day_week, hour_day, customer_id) %>% 
+  summarise(median_kwh = median(general_supply_kwh))
+
+group2_data%>% 
+  ungroup() %>% 
+  #filter(customer_id %in% c(8148781, 10703878)) %>% 
+  ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
+  geom_line(size = 0.5) +
+  facet_grid(customer_id~day_week, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
+```
+
+#Group5: n=18
+```{r}
+group2_data <- sm_hod_dow %>% 
+  as_tibble() %>% 
+  filter(group==5) %>% 
+  group_by(day_week, hour_day, customer_id) %>% 
+  summarise(median_kwh = median(general_supply_kwh))
+
+group2_data%>% 
+  ungroup() %>% 
+  #filter(customer_id %in% c(8148781, 10703878)) %>% 
+  ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
+  geom_line(size = 0.5) +
+  facet_grid(customer_id~day_week, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
+```
+
+#Group6: n=23
+```{r}
+group2_data <- sm_hod_dow %>% 
+  as_tibble() %>% 
+  filter(group==6) %>% 
+  group_by(day_week, hour_day, customer_id) %>% 
+  summarise(median_kwh = median(general_supply_kwh))
+
+group2_data%>% 
+  ungroup() %>% 
+  #filter(customer_id %in% c(8148781, 10703878)) %>% 
+  ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
+  geom_line(size = 0.5) +
+  facet_grid(customer_id~day_week, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5))
+```

### 2021-08-18 supplementary/algorithm2/real-data/algo2-elec8.Rmd
cleaning folders

--- 2021-07-16 supplementary/algo2-elec8.Rmd
@@
 title: "Algorithm using wpd values for clustering 8 households"
-author: "Sayani Gupta"
-date: "13/07/2021"
 output: pdf_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE,echo=FALSE, message=FALSE, warning = FALSE)
 ```
@@
 
-hc = stats::hclust(dist(data_clust[-1]),method="complete")
+hc = stats::hclust(dist(data_clust[-1], method="manhattan"),method="complete")
 #plot(hc)
@@
 library(fpc)
-koptimal = fpc::nselectboot(dist(data_clust[-1]),
+koptimal = fpc::nselectboot(dist(data_clust[-1], method="manhattan"),
                         B = 50,
@@
                         classification = "averagedist",
-                        krange = 2:8)
+                        krange = 2:5)
 
 koptimal = koptimal$kopt
+koptimal
 
@@
 
+groups %>% 
+  dplyr::group_by(value) %>%  
+  summarise(n = n())
+
+data_wide <- data_clust %>% left_join(groups) %>%
+  rename("group" = "value") %>% 
+  mutate(group = as.factor(group)) %>% ungroup()
+
 ##----cluster-characterization
 
-data = data_clust %>% 
-left_join(groups)%>% 
-             rename ("group" = "value") %>% 
-             ungroup() %>% 
-             select(group, everything(), -id) %>% 
-             pivot_longer(-1, names_to = "comb", values_to = "wpd")
-
-data_order <- data %>% 
-  group_by(comb) %>% 
-  summarise(m = mean(wpd)) %>% 
-  arrange(desc(m))
-
-data_median = data %>% 
-  group_by(group, comb) %>% 
-  summarize(wpd_med = median(wpd))
-
-data_pcp <- data_median %>% 
-mutate(group = as.character(group)) %>% 
-  left_join(data_order) %>% 
-  arrange(desc(m)) %>% 
-  pivot_wider(-4, 
-              names_from = comb, 
-              values_from = wpd_med)
-
-parcoord <- GGally::ggparcoord(data_pcp,
-                   columns = 2:ncol(data_pcp),
-                   groupColumn = 1,
+parcoord <- GGally::ggparcoord(data_wide,
+                   columns = 2:(ncol(data_wide)-1),
+                   groupColumn = "group",
                    showPoints = TRUE, 
@@
 ) + ggplot2::theme_bw() +
-  scale_color_viridis(discrete=TRUE) + 
+  scale_color_brewer(palette = "Dark2") + 
   ggplot2::theme(
@@
 
+# data = data_clust %>% 
+# left_join(groups)%>% 
+#              rename ("group" = "value") %>% 
+#              ungroup() %>% 
+#              select(group, everything(), -id) %>% 
+#              pivot_longer(-1, names_to = "comb", values_to = "wpd")
+# 
+# data_order <- data %>% 
+#   group_by(comb) %>% 
+#   summarise(m = mean(wpd)) %>% 
+#   arrange(desc(m))
+#   
+# data_median = data %>% 
+#   group_by(group, comb) %>% 
+#   summarize(wpd_med = median(wpd))
+# 
+# data_pcp <- data_median %>% 
+# mutate(group = as.character(group)) %>% 
+#   left_join(data_order) %>% 
+#   arrange(desc(m)) %>% 
+#   pivot_wider(-4, 
+#               names_from = comb, 
+#               values_from = wpd_med) 
+#              
+# parcoord <- GGally::ggparcoord(data_pcp,
+#                    columns = 2:ncol(data_pcp),
+#                    groupColumn = 1,
+#                    showPoints = TRUE, 
+#                    title = "(c)",
+#                    alphaLines = 0.8,
+#                    scale = "globalminmax"
+# ) + ggplot2::theme_bw() +
+#   scale_color_viridis(discrete=TRUE) + 
+#   ggplot2::theme(
+#     plot.title = ggplot2::element_text(size=10)
+#   )+
+#   ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 0)) +
+#   theme(legend.position = "bottom") +
+#   xlab("") +
+#   ylab("wpd")
+```
+
+```{r}
 ##----read-orig-data
@@
 
+ggpairs_nocolor <- ggpairs(data_wide, columns = 2:(ncol(data_wide)-1))
+ggpairs_nocolor
+
 sm_hod_dow <- elec %>% 
 gravitas::create_gran("hour_day") %>% 
-  gravitas::create_gran("day_week")
+  gravitas::create_gran("day_week") %>% 
+    dplyr::mutate(kwh = scale(kwh))
 
@@
   plot_type = "quantile", nrow=1) +
-  scale_y_log10() +
+  #scale_y_log10() +
     ggtitle("Group 1")
@@
   plot_type = "quantile", nrow=1) +
-  scale_y_log10()+
+  #scale_y_log10()+
     ggtitle("Group 2")
@@
   plot_type = "quantile", nrow=1) +
-  scale_y_log10() +
+  #scale_y_log10() +
     ggtitle("Group 3")
@@
   plot_type = "quantile", nrow=1) +
-  scale_y_log10() +
+  #scale_y_log10() +
     ggtitle("Group 4")
@@
 <!--     theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) -->
+
+```

### 2021-08-18 supplementary/algorithm2/real-data/handpick_distinct_cust.Rmd
cleaning folders

--- 2021-08-17 supplementary/handpick_distinct_cust.Rmd
@@
                   ymax = `75%`), fill = "lightblue") +
-  geom_line(aes(y = `50%`), size = 0.7) +
+  #geom_line(aes(y = `50%`), size = 0.7) +
+  geom_smooth(aes(y = `50%`), size = 0.7) +
   facet_wrap(design~customer_id, 
@@
        theme(strip.text.x = element_text(size = 6, margin = margin(b = 0, t = 0)))+ xlab("dom")
+
+### new quantile scale
+quantile_prob_val = c(0.25, 0.5, 0.75)
+
+sm_hod <- gravitas::create_gran(data_pick, "day_month")
+
+sm_hod_list <- sm_hod %>% 
+  as_tibble() %>% 
+  select(design, customer_id, day_month, general_supply_kwh) %>% 
+  pivot_wider(names_from = day_month,
+              values_from = general_supply_kwh)
+
+# groups_ref <- unique(data_pick$design) %>% as_tibble %>% 
+#               mutate(customer_serial_id = sm_hod_list$customer_id)
+
+# sm_hod_list_cat <-  sm_hod_list %>% left_join(groups_ref, by = c("customer_id" = "customer_serial_id")) %>% select(group, everything())
+
+ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, 2)]))
+nrow_sm <- 1:18
+
+sm_hod_quantiles_cat <- map(nrow_sm, function(x){
+  map(ncol_sm, function(y){
+   cell <- sm_hod_list%>% 
+     dplyr::filter(customer_id == customer_id[x]) %>% 
+     select(-c(1, 2)) %>% 
+     extract(y) %>% 
+     unlist()
+   quantile(cell, prob = quantile_prob_val)
+})  %>% bind_rows(.id = "categories_serial_id")
+}) %>% bind_rows(.id = "customer_serial_id")
+
+ref_cat <- names(sm_hod_list)[-c(1, 2)] %>% as_tibble() %>% set_names("category") %>%  
+  mutate(categories_serial_id = row_number())
+
+ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
+    mutate(customer_serial_id = row_number())
+
+sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
+  mutate(customer_serial_id = as.integer(customer_serial_id),
+         categories_serial_id = as.integer(categories_serial_id)) %>% 
+  left_join(ref_cat, by = "categories_serial_id") %>% 
+  left_join(ref_cust, by = "customer_serial_id") %>% 
+  select(customer_serial_id, categories_serial_id, category, customer_id, everything())
+
+data_heatmap <- sm_quantiles_ref %>% 
+ left_join(data_pick %>% distinct(customer_id, design))
+
 # 
+# %>% 
+#   left_join(categories_ref_cat) %>% 
+#   rename("category" = "value") %>% 
+#   select(-categories_serial_id) %>% 
+#   mutate(category = as.numeric(category))
+
+data_heatmap %>% 
+  ggplot(aes(x = as.integer(category))) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`), fill = "lightblue") +
+  #geom_line(aes(y = `50%`), size = 0.7) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+  theme_bw() +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
+  geom_smooth(aes(y = `50%`)) +
+   geom_smooth(aes(y = `25%`), se = FALSE) +
+   geom_smooth(aes(y = `75%`), se = FALSE) +
+  scale_y_log10()
+
+### all same scale 
 group_data_orig$customer_id <-factor(group_data_orig$customer_id, levels = elec_pred_wo_scaled$customer_id)

### 2021-08-18 supplementary/algorithm3/simulation-data/diffmeans-threegran.Rmd
cleaning folders

--- 2021-08-12 sim_setup_algo2_diffmeans.Rmd
@@
 library(GGally)
+library(tsibble)
+library(gravitas)
 ```
@@
 
-p1 <- ggplot(bind_data, aes(x = t, y = value)) + geom_line() + xlab("time")  + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
+p1 <- ggplot(bind_data %>% filter(design!="design-2"), aes(x = t, y = value)) + geom_line() + xlab("time")  + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
     strip.text = element_text(size = 8, margin = margin(b = 0, t = 0))) 
-p2 <- ggplot(bind_data, aes(x = as.factor(g1), y = value)) + geom_boxplot(width=0.5) + xlab("g1") + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
+p2 <- ggplot(bind_data %>% filter(design!="design-2"), aes(x = as.factor(g1), y = value)) + geom_boxplot(width=0.5) + xlab("g1") + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
     strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+stat_summary(fun.y = median, geom="line", group= 1, color= "blue", size = 0.5)
-p3 <- ggplot(bind_data, aes(x = as.factor(g2), y = value)) + geom_boxplot(width=0.5) + xlab("g2") + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
+p3 <- ggplot(bind_data %>% filter(design!="design-2"), aes(x = as.factor(g2), y = value)) + geom_boxplot(width=0.5) + xlab("g2") + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
     strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+stat_summary(fun.y = median, geom="line", group= 1, color= "blue", size = 0.5)
 
-p4 <- ggplot(bind_data, aes(x = as.factor(g3), y = value)) + geom_boxplot(width=0.5) +
+p4 <- ggplot(bind_data %>% filter(design!="design-2"), aes(x = as.factor(g3), y = value)) + geom_boxplot(width=0.5) +
   xlab("g3") + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
@@
 ```{r}
-seed_len = seq(12345, 12399, 1)
+seed_len = seq(12345, 12349, 1)
 bind_data_iter <- map(seq_len(length(seed_len)), function(x){
@@
 
-<!-- Cross-check -->
-<!-- data = bind_data_iter %>% filter(design == "design-2", seed_id == 5) -->
-<!-- p2 <- ggplot(data, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw() + facet_wrap(~design, ncol = 1) -->
-<!-- p3 <- ggplot(data, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw() + facet_wrap(~design, ncol = 1) -->
-<!-- p4 <- ggplot(data, aes(x = as.factor(g3), y = value)) + geom_boxplot() + -->
-<!-- xlab("g3") + theme_bw() + facet_wrap(~design, ncol = 1) -->
-<!-- (p2 + p3 + p4) -->
+```{r}
 
-<!-- ```{r} -->
-<!-- p2 <- ggplot(data, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw() -->
-<!-- p3 <- ggplot(data, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw() -->
-<!-- p4 <- ggplot(data, aes(x = as.factor(g3), y = value)) + geom_boxplot() + -->
-<!-- xlab("g3") + theme_bw()  -->
-<!-- (p2 + p3 + p4) -->
-<!-- ``` -->
+quantile_prob_val = seq(0.1,0.9,0.1)
+bind_data_iter <- bind_data_iter %>% mutate(
+  customer_id  = paste0(design,"-s-", seed_id)) %>%
+   as_tsibble(key=customer_id,index=t) %>% 
+  filter(design!="design-2")
 
-```{r}
-harmony_tbl <- tibble(facet_variable = NA, x_variable = c("g1", "g2", "g3"),
-                      x_levels = c(2, 3, 4), facet_levels = NA)
+library(gracsr)
+v2 <- suppressWarnings(
+  scaled_dist_gran(bind_data_iter, "g1", 
+                   response = "value",
+                   quantile_prob_val = quantile_prob_val))%>% rename("dist_g1" = "dist")
+
+v3 <- suppressWarnings(
+  scaled_dist_gran(bind_data_iter, "g2", 
+                   response = "value",
+                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_g2" = "dist")
+
+v4 <- suppressWarnings(
+  scaled_dist_gran(bind_data_iter, "g3", 
+                   response = "value",
+                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_g3" = "dist")
+
+data_dist <- v3 %>% 
+  left_join(v2) %>%
+  left_join(v4) %>% 
+  mutate(dist = dist_g1 + dist_g2 +dist_g3) %>% 
+    pivot_wider(-c(3, 4, 5), 
+                names_from = customer_to,
+              values_from = dist) %>%
+  rename("customer_id" = "customer_from")
+
 ```
 
-```{r wpd-table}
-data <- bind_data_iter %>% mutate(data_id = paste(design, seed_id, sep = "-"))
-bind_data_iter <- tsibble::as_tsibble(data, index = t, key = "data_id")
+```{r groups}
+d = stats::as.dist(data_dist[,-1])
+hc = stats::hclust(d,method="complete")
 
-data_wpd <- map(seq_len(length(unique(bind_data_iter$data_id))),
-    function(y){
-      z = unique(bind_data_iter$data_id)
-      data_sub <- bind_data_iter %>% filter(data_id==z[y])
-        elec_select_harmony = hakear::select_harmonies(data_sub,
-                               harmony_tbl = harmony_tbl,
-                               response = value,
-                               nperm = 200
-  ) %>% mutate(data_id = unique(data_sub$data_id)) %>% unnest()
-    }) %>% bind_rows()
-
-#write_csv(data_wpd, "data_wpd.rds")
+groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
+groups %>% group_by(group) %>% summarise(n=n())
+groups %>% filter(group==1)
+groups %>% filter(group==2)
+groups %>% filter(group==3)
+#groups %>% filter(group==4)
 ```
-
-# Table of wpd across designs and wpd
-
-```{r data-wide}
-
-data_wide <- data_wpd %>% select(x_variable, wpd, data_id) %>% 
-  pivot_wider(names_from = x_variable, values_from = wpd)
-data_wide %>% kable()
-
-ggpairs_nocolor <- ggpairs(data_wide, columns = 2:4)
-
-rownames(data_wide) = data_wide$data_id
-
-hc = stats::hclust(dist(data_wide[-1], method = "manhattan"),method="complete")
-
-library(fpc)
-koptimal = fpc::nselectboot(dist(data_wide[-1], method="manhattan"),
-                        B = 500,
-                        method = "complete",
-                        clustermethod = disthclustCBI, 
-                        classification = "averagedist",
-                        krange = 2:8)
-
-koptimal = koptimal$kopt
-```
-
-# Clustering of designs
-
-```{r}
-  groups <- cutree(hc, k=4) %>% as_tibble() %>% 
-mutate(data_id = data_wide$data_id) %>% rename("group" = "value")
-
-  groups %>% kable()
-
-data_wide_group <- data_wide %>% left_join(groups)  %>% 
-  mutate(group_id = as.factor(group))
-
-parcoord <- GGally::ggparcoord(data_wide_group,
-                   columns = 2:ncol(data_wide),
-                   groupColumn = "group_id",
-                   showPoints = TRUE, 
-                   title = "(c)",
-                   alphaLines = 0.8,
-                   scale = "globalminmax", 
-                   order = "anyClass"
-) + ggplot2::theme_bw() +
-  scale_color_brewer(palette = "Dark2") + 
-  ggplot2::theme(
-    plot.title = ggplot2::element_text(size=10)
-  )+
-  ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 0)) +
-  theme(legend.position = "bottom") +
-  xlab("") +
-  ylab("wpd")
-parcoord
-ggpairs_nocolor
-  # ggpairs_color <- ggscatmat(data_wide, columns=2:4, color = "group") +
-  # scale_colour_brewer(palette="Set2")
-```
-
-# PCA based clustering
-
-```{r, eval = FALSE}
-pc <- prcomp(data_wide_group[c(2, 3, 4)], center = F, scale = F)
-
-# First two principal componentsafc
-comp <- data.frame(pc$x[,1:2])
-
-#plot(comp, pch=16, col=rgb(0,0,0,0.5))
-
-hc = stats::hclust(dist(comp))
-groups <- cutree(hc, k=koptimal) %>% as_tibble() %>% bind_cols(comp) %>% 
-  mutate(value = as.character(value))
-library(RColorBrewer)
-library(scales)
-groups %>% ggplot(aes(x=PC1, y = PC2)) + geom_point(aes(color = value)) +
-  scale_color_brewer(palette = "Dark2") + theme_bw()
-```
-
-```{r}
-data_plot <- left_join(data_wide_group,  bind_data_iter %>% distinct(design, data_id)) %>%
-  mutate(pred_group = case_when(
-    group == 1 ~"design-1",
-    group == 2 ~"design-2",
-    group == 3 ~"design-3",
-    group == 4 ~"design-4"
-  )) %>% 
-  mutate(pred_group = as.factor(pred_group),
-         design = as.factor(design)
-  )
-
-xtab <- confusionMatrix(data_plot$design, data_plot$pred_group)
-
-data_plot %>%  ggplot(aes(x=g1, y=g2)) + geom_point(aes(colour = design)) +
-  scale_color_brewer(palette = "Dark2")
-
-left_join(data_wide_group,  bind_data_iter %>% distinct(design, data_id)) %>%  ggplot(aes(x=g1, y=g3)) + geom_point(aes(colour = design)) +
-  scale_color_brewer(palette = "Dark2")
-
-left_join(data_wide_group,  bind_data_iter %>% distinct(design, data_id)) %>%  ggplot(aes(x=g2, y=g3)) + geom_point(aes(colour = design)) +
-  scale_color_brewer(palette = "Dark2")
-
-```

### 2021-09-13 main_old.Rmd
add structure to methodology

--- 2021-07-16 main.Rmd
@@
     keep_tex: yes
+    toc: TRUE
 ---
@@
 library(gravitas)
-library(gracsR)
+library(gracsr)
 library(ggdendro)
@@
 
+## Common load clustering techniques of smart meter data
+
+The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
+None of the 34 selected papers which focus on clustering consumption are based on Australian smart meter data.
+The clustering is frequently applied directly to the raw data without scrutinizing for auto correlation and periodicity.
+The algorithm most ubiquitously employed is K-Means.
+But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters.
+Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results.
+@Tureczek2018-ha has shown that a transformation of data to incorporate autocorrelation before K-Means clustering can improve performance and enable K-Means to deliver smaller clusters with less within-cluster variance.
+However, it does not explain the cluster composition by combining it with external data.
+Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf).
+Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.
+As the methods are often situation specific,
+it makes sense to compare them on the performance rather than any standard performance metric.
+A type of clustering based on information theory such as Shannon or Renyi entropy and their variants are addressed in , which differs from typical methods adopted for electricity consumer classification, based on the Euclidean distance notion.
+@Motlagh2019-yj presents strategy to address the problems on patchy, and unequal time-series common in residential data sets by converting load time series into map models.
+Most time-series clustering models are limited to handling time domain with same start and end date and time.
+Most of the solutions to handle this like longest common subsequence, dynamic time warping are prone to computational limit with increased length of the series.
+
 <!-- Electricity smart meter technology is increasingly being deployed in residential and commercial buildings.
@@
 
-## Getting data
+## Electricity demand data
 
@@
 
+### Getting data
 The entire data is procured from CSIRO.
@@
 
-## Missing Data
+### Missing Data
 
@@
 
-## Raw data
+## LGA and weather
 
@@
 
-```{r}
+```{r,eval=FALSE}
 elec_ts <- read_rds("data/elec_ts.rds")
@@
 ```{r raw-50, fig.height = 8, fig.cap="The raw data for 50 households are shown.
-It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed."}
+It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed.", eval = FALSE}
 
@@
 # Clustering methodology
-
-## Common load clustering techniques of smart meter data
-
-The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
-None of the 34 selected papers which focus on clustering consumption are based on Australian smart meter data.
-The clustering is frequently applied directly to the raw data without scrutinizing for auto correlation and periodicity.
-The algorithm most ubiquitously employed is K-Means.
-But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters.
-Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results.
-@Tureczek2018-ha has shown that a transformation of data to incorporate autocorrelation before K-Means clustering can improve performance and enable K-Means to deliver smaller clusters with less within-cluster variance.
-However, it does not explain the cluster composition by combining it with external data.
-Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf).
-Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.
-As the methods are often situation specific,
-it makes sense to compare them on the performance rather than any standard performance metric.
-A type of clustering based on information theory such as Shannon or Renyi entropy and their variants are addressed in , which differs from typical methods adopted for electricity consumer classification, based on the Euclidean distance notion.
-@Motlagh2019-yj presents strategy to address the problems on patchy, and unequal time-series common in residential data sets by converting load time series into map models.
-Most time-series clustering models are limited to handling time domain with same start and end date and time.
-Most of the solutions to handle this like longest common subsequence, dynamic time warping are prone to computational limit with increased length of the series.
 

### 2021-09-13 supplementary/Untitled.Rmd
add structure to methodology

---
title: "check_effect_transformation"
author: "Sayani Gupta"
date: "01/09/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2021-09-13 supplementary/algorithm2/real-data/algo2-elec100-onegran.Rmd
add structure to methodology

@@
 ##----harmony-all
-elec_harmony_all <- read_rds("../data/elec_harmony-100-nogap-onegran.rds")
+elec_harmony_all <-read_rds("../../../data/elec_harmony-100-nogap-onegran.rds")
 
@@
     theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
+```

### 2021-09-13 supplementary/algorithm2/real-data/algo2-elec100.Rmd
add structure to methodology

@@
 ##----harmony-all
-elec_harmony_all <- read_rds("../data/algo2-rawdata-wpd.rds")
+elec_harmony_all <- read_rds("../../../data/algo2-rawdata-wpd.rds")
 
@@
 
-elec <- read_rds("../data/elec_nogap_2013_100.rds") %>% 
+elec <- read_rds("../../../data/elec_nogap_2013_100.rds") %>% 
   left_join(groups)%>% 

### 2021-09-13 supplementary/algorithm2/real-data/handpick_distinct_cust.Rmd
add structure to methodology

@@
 ##----harmony-all
-# elec <- read_rds("../data/elec_nogap_2013_100.rds") %>% filter(customer_id %in% data_pick$customer_id)
-# 
-# elec_split = elec %>% group_split(customer_id)
-# harmonies <- read_rds("../../paper-hakear/paper/data/harmonies.rds")
-# 
-# harmonies_acro <- harmonies %>% mutate(facet_variable=NA, facet_levels=NA) %>% distinct()
-# 
-# elec_select_harmony = parallel::mclapply(1:18, function(x){
-#   
-#   data_id <-  elec_split %>% magrittr::extract2(x) %>% 
-#     as_tsibble(index = reading_datetime)
-# 
-#   
-#   k = hakear::select_harmonies(data_id,
-#                            harmony_tbl = harmonies_acro,
-#                            response = general_supply_kwh,
-#                            nperm = 200,
-#                            nsamp = 2
-#   ) %>% mutate(customer_id = unique(data_id$customer_id))
-#   #write_rds(k, paste("data/elec_harmony-100-nogap.rds"))
-# }, mc.cores = parallel::detectCores() - 1, mc.preschedule = FALSE, mc.set.seed = FALSE)
-# #toc()
-# elec_harmny <- elec_select_harmony %>% bind_rows() 
-# write_rds(elec_harmny, paste("../data/elec_harmony-100-nogap-new.rds"))
+elec <- read_rds("../../../data/elec_nogap_2013_100.rds") %>% filter(customer_id %in% data_pick$customer_id)
+
+elec_split = elec %>% group_split(customer_id)
+harmonies <- read_rds("../../../../paper-hakear/paper/data/harmonies.rds")
+
+harmonies_acro <- harmonies %>% mutate(facet_variable=NA, facet_levels=NA) %>% distinct()
+
+elec_select_harmony = parallel::mclapply(1:18, function(x){
+
+  data_id <-  elec_split %>% magrittr::extract2(x) %>%
+    as_tsibble(index = reading_datetime)
+
+  k = hakear::select_harmonies(data_id,
+                           harmony_tbl = harmonies_acro,
+                           response = general_supply_kwh,
+                           nperm = 200,
+                           nsamp = 200
+  ) %>% mutate(customer_id = unique(data_id$customer_id))
+    #write_rds(k, paste("data/elec_harmony-100-nogap.rds"))
+}, mc.cores = parallel::detectCores() - 1, mc.preschedule = FALSE, mc.set.seed = FALSE)
+#toc()
+elec_harmny <- elec_select_harmony %>% bind_rows()
+ write_rds(elec_harmny, paste("../../../data/elec_harmony-100-nogap-new-rerun.rds"))
+
+f = read_rds("../../../data/elec_harmony-100-nogap-new.rds")
 

### 2021-09-13 supplementary/algorithm3/real-data/clean-dataset-500cust.Rmd
add structure to methodology

---
title: "Hand picking similar behaving group of customers to check clustering results"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, echo = FALSE, message=FALSE, warning=FALSE, cache = TRUE)
library(readr)
library(tidyverse)
library(gravitas)
library(tsibble)
library(parallel)
library(magrittr)
library(GGally)
library(knitr)
library(patchwork)
library(caret)
```

```{r read-data}
data1 <- read_rds("../../../data/elec_nogap_2013_500.rds")
data2 <- read_rds("../../../data/elec_nogap_2013_100.rds")
data3 <- data1 %>% bind_rows(data2 %>% filter(!(customer_id %in% data1$customer_id)))

write_rds(data3, "../../../data/elec_nogap_2013_600.rds")

quantile_prob_clust = seq(0.1, 0.9, 0.1)
quantile_prob_graph = c(0.25, 0.5, 0.75)
```

A clean dataset is obtained by choosing minimum sum of JS distances from each typical customer based on $quantiles = seq(0.1, 0.9, 0.1)$.
The objective is to see if the clustering algorithm then picks the least distant ones as the group.

## only hod

```{r clean-dataset}
library(gracsr)
# v2 <- suppressWarnings(
#   scaled_dist_gran(data3, "hour_day",
#                    response = "general_supply_kwh",
#                    quantile_prob_val = quantile_prob_clust)) %>% rename("dist_hod" = "dist")
# 
# v2 <- v2 %>% arrange(dist_hod) %>% filter(customer_from!=customer_to)
# 
# write_rds(v2, "../../../data/scaled_dist_nogap_2013_600.rds")
# 
# data3 <- data3 %>% filter(!(customer_id %in% c(8143599)))
# 
# v3 <- suppressWarnings(
#   scaled_dist_gran_iqr(data3 %>% select(customer_id, reading_datetime, general_supply_kwh), "hour_day",
#                    response = "general_supply_kwh",
#                    quantile_prob_val = quantile_prob_clust)) %>% rename("dist_hod" = "dist")
#   
# 
# v3 <- v3 %>% arrange(dist_hod) %>% filter(customer_from!=customer_to)
# 
# write_rds(v3, "../../../data/scaled_dist_nogap_2013_600_iqr.rds")

```

```{r}
v2 <- read_rds("../../../data/scaled_dist_nogap_2013_600.rds")

data_pick_onepeak <- v2 %>% 
  arrange(dist_hod) %>%
  filter(customer_from == 8198401) %>% 
  arrange(dist_hod) %>% 
  head(n=4)

data_pick_twopeak <- v2 %>% 
  arrange(dist_hod) %>%
  filter(customer_from == 8181071) %>% 
  arrange(dist_hod) %>%
  head(n=4)

11013154
data_pick_threepeak <- v2 %>% 
  arrange(dist_hod) %>%
  filter(customer_from == 8196501) %>% 
  arrange(dist_hod) %>% 
   head(n=4)

data_pick_cust <- bind_rows(
  c(data_pick_onepeak$customer_from, data_pick_onepeak$customer_to) %>% unique() %>% as_tibble() %>% set_names("customer_id"),
                       c(data_pick_twopeak$customer_from, data_pick_twopeak$customer_to) %>% unique() %>% as_tibble() %>% set_names("customer_id"),
                       c(data_pick_threepeak$customer_from, data_pick_threepeak$customer_to) %>% unique() %>% as_tibble() %>% set_names("customer_id"), .id = "design") %>%  filter(!(customer_id %in% c(8492316, 10346738, 8614285)))

# %>%  filter(!(customer_id %in% c(8171227, 10183908, 11013154, 8492316)))

data <- read_rds("../../../data/elec_nogap_2013_600.rds")
data_pick <- data %>% filter(customer_id %in% data_pick_cust$customer_id) %>% select(customer_id,reading_datetime, general_supply_kwh) %>% left_join(data_pick_cust)

```

```{r characterize-15cust, fig.cap= "Median (black) and quartile deviation (blue region) of hourly demand drawn for few customers showing similar behaviors. Roughly speaking, Design 1 has one evening peak, Design 2 has two peaks and Design 3 has three peaks in a day. Each of design 1, 2 and 3 have six similar behaving customers resulting to 18 time series. We want our clustering results to group each of the designs together."}

sm_hod <- gravitas::create_gran(data_pick, "hour_day")

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(design, customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, 2)]))
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list%>% 
     dplyr::filter(customer_id == customer_id[x]) %>% 
     select(-c(1, 2)) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_graph)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

ref_cat <- names(sm_hod_list)[-c(1, 2)] %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
    mutate(customer_serial_id = row_number())

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>%
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "customer_serial_id") %>% 
  select(customer_serial_id, categories_serial_id, category, customer_id, everything())

data_heatmap <- sm_quantiles_ref %>% 
 left_join(data_pick %>% distinct(customer_id, design))

data_heatmap %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`), size = 0.7) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 4) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")
```

## Do they look similar on the transformed scale?

```{r transformed-scale}
data_pick_scaled <- data_pick %>%
    group_by(customer_id) %>%
    dplyr::mutate(general_supply_kwh =  stats::qqnorm(general_supply_kwh, plot.it=FALSE)$x) %>%
    ungroup()

sm_hod <- gravitas::create_gran(data_pick_scaled, "hour_day")

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(design, customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, 2)]))
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list%>% 
     dplyr::filter(customer_id == customer_id[x]) %>% 
     select(-c(1, 2)) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_graph)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

ref_cat <- names(sm_hod_list)[-c(1, 2)] %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
    mutate(customer_serial_id = row_number())

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>%
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "customer_serial_id") %>% 
  select(customer_serial_id, categories_serial_id, category, customer_id, everything())

data_heatmap <- sm_quantiles_ref %>% 
 left_join(data_pick %>% distinct(customer_id, design))

data_heatmap %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`), size = 0.7) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 4) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")
```

## Do they look similar on the robust transformed scale?

```{r robust-transformed}
data_scaled_robust <- robust_scale_data(data_pick, "hour_day")
suppressWarnings(plot_data(data_scaled_robust, "hour_day", response = "scaled_response", nfacet = 4))
```

## Does the method with NQT scaling identifies the groups?

Yes it does.

```{r hod-only-nqt,echo = TRUE}
#quantile_prob_val = c(0.5, 0.75)
#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
library(gracsr)
v2 <- suppressWarnings(
  scaled_dist_gran(data_pick, "hour_day", 
                   response = "general_supply_kwh",
                   quantile_prob_val = quantile_prob_clust)) %>% rename("dist_hod" = "dist")
v3 <- suppressWarnings(
  scaled_dist_gran(data_pick, "day_month",
                   response = "general_supply_kwh", 
                   quantile_prob_val = quantile_prob_clust)) %>% rename("dist_dom" = "dist")

data_dist <- v3 %>% 
  left_join(v2) %>%
  mutate(dist = dist_hod + dist_dom) %>% 
    pivot_wider(-c(3, 4), 
                names_from = customer_to,
              values_from = dist) %>%
  rename("customer_id" = "customer_from")
```

```{r clustering-result-nqt}
d = stats::as.dist(data_dist[,-1])
hc = stats::hclust(d,method="ward.D")

groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
groups %>% group_by(group) %>% summarise(n=n())
#groups %>% filter(group==1)
#groups %>% filter(group==2)
#groups %>% filter(group==3)

data_heatmap_new <- data_heatmap %>% left_join(groups)

data_heatmap_new$group = as.factor(data_heatmap_new$group)

data_heatmap_new %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`, color = group), size = 0.7) +
  facet_wrap(group~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 4) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
  scale_color_brewer(palette = "Dark2")
```

## Does the method with robust scalings identifies the groups?

Yes it does.

```{r hod-only-robust,echo = TRUE}

library(gracsr)
v2 <- suppressWarnings(
  scaled_dist_gran_iqr(data_pick, "hour_day", 
                   response = "general_supply_kwh",
                   quantile_prob_val = quantile_prob_clust)) %>% rename("dist_hod" = "dist")
v3 <- suppressWarnings(
  scaled_dist_gran_iqr(data_pick, "day_month",
                   response = "general_supply_kwh", 
                   quantile_prob_val = quantile_prob_clust)) %>% rename("dist_dom" = "dist")

data_dist <- v3 %>% 
  left_join(v2) %>%
  mutate(dist = dist_hod) %>% 
    pivot_wider(-c(3, 4), 
                names_from = customer_to,
              values_from = dist) %>%
  rename("customer_id" = "customer_from")
```

```{r clustering-result-robust}
d = stats::as.dist(data_dist[,-1])
hc = stats::hclust(d,method="ward.D")

groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
groups %>% group_by(group) %>% summarise(n=n())
#groups %>% filter(group==1)
#groups %>% filter(group==2)
#groups %>% filter(group==3)

data_heatmap_new <- data_heatmap %>% left_join(groups)

data_heatmap_new$group = as.factor(data_heatmap_new$group)

data_heatmap_new %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`, color = group), size = 0.7) +
  facet_wrap(group~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 4) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
  scale_color_brewer(palette = "Dark2")
```

### 2021-09-13 supplementary/algorithm3/real-data/clean-dataset.Rmd
add structure to methodology

---
title: "Hand picking similar behaving group of customers to check clustering results"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, echo = FALSE, message=FALSE, warning=FALSE, cache = TRUE)
library(readr)
library(tidyverse)
library(gravitas)
library(tsibble)
library(parallel)
library(magrittr)
library(GGally)
library(knitr)
library(patchwork)
library(caret)
```

```{r read-data}
data <- read_rds("../../../data/elec_nogap_2013_100.rds")
quantile_prob_clust = seq(0.1, 0.9, 0.1)
quantile_prob_graph = c(0.25, 0.5, 0.75)
```

A clean dataset is obtained by choosing minimum sum of JS distances from each typical customer based on $quantiles = seq(0.1, 0.9, 0.1)$.
The objective is to see if the clustering algorithm then picks the least distant ones as the group.

## only hod

```{r clean-dataset}

# library(gracsr)
# v2 <- suppressWarnings(
#   scaled_dist_gran(data, "hour_day", 
#                    response = "general_supply_kwh",
#                    quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
# 
# 
# 
# v2 <- v2 %>% arrange(dist_hod) %>% filter(customer_from!=customer_to)
# 
# write_rds(v2, "../../../data/scaled_dist_nogap_2013_100.rds")
v2 <- read_rds("../../../data/scaled_dist_nogap_2013_100.rds")

data_pick_onepeak <- v2 %>% 
  arrange(dist_hod) %>%
  filter(customer_from == 8198401) %>% 
  arrange(dist_hod) %>% 
  head(n=5)

data_pick_twopeak <- v2 %>% 
  arrange(dist_hod) %>%
  filter(customer_from == 8181071) %>% 
  arrange(dist_hod) %>%
  head(n=5)

data_pick_threepeak <- v2 %>% 
  arrange(dist_hod) %>%
  filter(customer_from == 11013154) %>% 
  arrange(dist_hod) %>% 
   head(n=5)

data_pick_cust <- bind_rows(
  c(data_pick_onepeak$customer_from, data_pick_onepeak$customer_to) %>% unique() %>% as_tibble() %>% set_names("customer_id"),
                       c(data_pick_twopeak$customer_from, data_pick_twopeak$customer_to) %>% unique() %>% as_tibble() %>% set_names("customer_id"),
                       c(data_pick_threepeak$customer_from, data_pick_threepeak$customer_to) %>% unique() %>% as_tibble() %>% set_names("customer_id"), .id = "design") %>% 
  filter(!(customer_id %in% c(8171227, 10183908)))

data <- read_rds("../../../data/elec_nogap_2013_100.rds")

data_pick <- data %>% filter(customer_id %in% data_pick_cust$customer_id) %>% select(customer_id,reading_datetime, general_supply_kwh) %>% left_join(data_pick_cust)

write_csv(data_pick_cust, "../../../data/data_pick_cust_clean.rds")
```

```{r characterize-15cust, fig.cap= "Median (black) and quartile deviation (blue region) of hourly demand drawn for few customers showing similar behaviors. Roughly speaking, Design 1 has one evening peak, Design 2 has two peaks and Design 3 has three peaks in a day. Each of design 1, 2 and 3 have six similar behaving customers resulting to 18 time series. We want our clustering results to group each of the designs together."}

sm_hod <- gravitas::create_gran(data_pick, "hour_day")

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(design, customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, 2)]))
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list%>% 
     dplyr::filter(customer_id == customer_id[x]) %>% 
     select(-c(1, 2)) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_graph)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

ref_cat <- names(sm_hod_list)[-c(1, 2)] %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
    mutate(customer_serial_id = row_number())

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>%
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "customer_serial_id") %>% 
  select(customer_serial_id, categories_serial_id, category, customer_id, everything())

data_heatmap <- sm_quantiles_ref %>% 
 left_join(data_pick %>% distinct(customer_id, design))

data_heatmap %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`), size = 0.7) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 5) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")
```
## Do they look similar on the transformed scale?

```{r transformed-scale}
data_pick_scaled <- data_pick %>%
    group_by(customer_id) %>%
    dplyr::mutate(general_supply_kwh =  stats::qqnorm(general_supply_kwh, plot.it=FALSE)$x) %>%
    ungroup()

sm_hod <- gravitas::create_gran(data_pick_scaled, "hour_day")

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(design, customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, 2)]))
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list%>% 
     dplyr::filter(customer_id == customer_id[x]) %>% 
     select(-c(1, 2)) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_graph)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

ref_cat <- names(sm_hod_list)[-c(1, 2)] %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
    mutate(customer_serial_id = row_number())

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>%
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "customer_serial_id") %>% 
  select(customer_serial_id, categories_serial_id, category, customer_id, everything())

data_heatmap <- sm_quantiles_ref %>% 
 left_join(data_pick %>% distinct(customer_id, design))

data_heatmap %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`), size = 0.7) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 5) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")
```

## only hod 
### Does hod as the only variable correctly identifies the groups?

```{r hod-only,echo = TRUE}
#quantile_prob_val = c(0.5, 0.75)
#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
library(gracsr)
v2 <- suppressWarnings(
  scaled_dist_gran(data_pick, "hour_day", 
                   response = "general_supply_kwh",
                   quantile_prob_val = quantile_prob_clust)) %>% rename("dist_hod" = "dist")
v3 <- suppressWarnings(
  scaled_dist_gran(data_pick, "day_month",
                   response = "general_supply_kwh", 
                   quantile_prob_val = quantile_prob_clust)) %>% rename("dist_dom" = "dist")

data_dist <- v3 %>% 
  left_join(v2) %>%
  mutate(dist = dist_hod + dist_dom) %>% 
    pivot_wider(-c(3, 4), 
                names_from = customer_to,
              values_from = dist) %>%
  rename("customer_id" = "customer_from")
```

```{r clustering-result}
d = stats::as.dist(data_dist[,-1])
hc = stats::hclust(d,method="ward.D")

groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
groups %>% group_by(group) %>% summarise(n=n())
#groups %>% filter(group==1)
#groups %>% filter(group==2)
#groups %>% filter(group==3)

data_heatmap_new <- data_heatmap %>% left_join(groups)

data_heatmap_new$group = as.factor(data_heatmap_new$group)

data_heatmap_new %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`, color = group), size = 0.7) +
  facet_wrap(group~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 5) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
  scale_color_brewer(palette = "Dark2")
```

## hod + dom
### Is the clustering sensitive to nuisance parameter?

```{r hod-dom,echo = TRUE}
#quantile_prob_val = c(0.5, 0.75)
#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
library(gracsr)
v2 <- suppressWarnings(
  scaled_dist_gran(data_pick, "hour_day", 
                   response = "general_supply_kwh",
                   quantile_prob_val = quantile_prob_clust)) %>% rename("dist_hod" = "dist")
v3 <- suppressWarnings(
  scaled_dist_gran(data_pick, "day_month",
                   response = "general_supply_kwh", 
                   quantile_prob_val = quantile_prob_clust)) %>% rename("dist_dom" = "dist")

data_dist <- v3 %>% 
  left_join(v2) %>%
  mutate(dist = dist_hod + dist_dom) %>% 
    pivot_wider(-c(3, 4), 
                names_from = customer_to,
              values_from = dist) %>%
  rename("customer_id" = "customer_from")
```

```{r clustering-result-hod-dom}
d = stats::as.dist(data_dist[,-1])
hc = stats::hclust(d,method="ward.D")

groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
groups %>% group_by(group) %>% summarise(n=n())
#groups %>% filter(group==1)
#groups %>% filter(group==2)
#groups %>% filter(group==3)

data_heatmap_new <- data_heatmap %>% left_join(groups)

data_heatmap_new$group = as.factor(data_heatmap_new$group)

data_heatmap_new %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`, color = group), size = 0.7) +
  facet_wrap(group~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 5) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
  scale_color_brewer(palette = "Dark2")
```

# All 100 customers
### Running it on all 100 and making 10 clusters. Do they have sufficiently different shapes?
### Variable importance

```{r hod-100cust,echo = TRUE}
# library(gracsr)
# v2 <- suppressWarnings(
#   scaled_dist_gran(data, "hour_day", 
#                    response = "general_supply_kwh",
#                    quantile_prob_val = quantile_prob_clust)) %>% rename("dist_hod" = "dist")

#write_rds(v2, "../../../data/scaled_dist_100clust_deciles.rds")
v2 <- read_rds("../../../data/scaled_dist_100clust_deciles.rds")

data_dist <- v2 %>%
  mutate(dist = dist_hod) %>% 
    pivot_wider(-c(3, 4), 
                names_from = customer_to,
              values_from = dist) %>%
  rename("customer_id" = "customer_from")

cust_ref <- bind_cols(customer_id=data_dist$customer_id, row = seq_len(nrow(data_dist)))

cust_rm_outlier <-   cust_ref %>% filter(!(row %in% c(73, 82, 79, 97, 25, 91)))

v3 <- v2%>% 
  filter(customer_from %in% cust_rm_outlier$customer_id,
         customer_to %in% cust_rm_outlier$customer_id)

data_dist <- v3 %>%
  mutate(dist = dist_hod) %>% 
    pivot_wider(-c(3, 4), 
                names_from = customer_to,
              values_from = dist) %>%
  rename("customer_id" = "customer_from")

```

```{r hod-100cust-clustering}
d = stats::as.dist(data_dist[,-1])
hc = stats::hclust(d,method="ward.D")

library(fpc)
# koptimal = fpc::nselectboot(as.dist(data_dist[,-1]),
#                         B = 5000,
#                         method = "ward.D",
#                         clustermethod = disthclustCBI, 
#                         classification = "averagedist",
#                         krange = 2:20)
# 
# koptimal = koptimal$kopt

groups <- tibble(group = cutree(hc, k=10), customer_id = data_dist$customer_id)
groups %>% group_by(group) %>% summarise(n=n())

data <- read_rds("../../../data/elec_nogap_2013_100.rds")
data <- data %>% left_join(groups)

sm_hod <- gravitas::create_gran(data, "hour_day")

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh, group) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, 2)]))
nrow_sm <- unique(sm_hod_list$group)

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list%>% 
     dplyr::filter(group == group[x]) %>% 
     select(-c(1, 2)) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_graph)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "group_serial_id")

ref_cat <- names(sm_hod_list)[-c(1, 2)] %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- unique(sm_hod_list$group) %>% as_tibble()%>% set_names("group")  %>% 
    mutate(group_serial_id = as.character(row_number()))

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(group_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>% 
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "group_serial_id") %>% 
  select(group_serial_id, categories_serial_id, category, everything())

# data_heatmap_100 <- sm_quantiles_ref %>% 
#  left_join(data %>% distinct(customer_id))
#

sm_quantiles_ref$group = as.factor(sm_quantiles_ref$group)

sm_quantiles_ref %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`), size = 0.7) +
  facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_value") +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
  scale_color_brewer(palette = "Dark2") + scale_color_brewer(palette = "Dark2")
```
# Split these groups to see if the shapes of individual customers in a group is the same

```{r all100}
data <- read_rds("../../../data/elec_nogap_2013_100.rds")

sm_hod <- gravitas::create_gran(data, "hour_day")
sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, 2)]))
nrow_sm <- 1:100

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list%>% 
     dplyr::filter(customer_id == customer_id[x]) %>% 
     select(-c(1, 2)) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_graph)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

ref_cat <- names(sm_hod_list)[-c(1, 2)] %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
    mutate(customer_serial_id = row_number())

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>% 
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "customer_serial_id") %>% 
  select(customer_serial_id, categories_serial_id, category, customer_id, everything())

data_heatmap_100 <- sm_quantiles_ref %>% 
 left_join(data %>% distinct(customer_id))

data_heatmap_100 %>% 
  left_join(groups) %>% 
  filter(group==1) %>% 
  #filter(customer_id %in% c(8148781, 10703878)) %>% 
    ggplot(aes(x = as.integer(category))) +
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") + 
  geom_line(size = 0.5, aes(y = `50%`)) +
  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5)) +
  xlab("hour-day") +
  ggtitle("Group-2")

data_heatmap_100 %>% 
  left_join(groups) %>% 
  filter(group==2) %>% 
  #filter(customer_id %in% c(8148781, 10703878)) %>% 
    ggplot(aes(x = as.integer(category))) +
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") + 
  geom_line(size = 0.5, aes(y = `50%`)) +
  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5)) +
  xlab("hour-day") +
  ggtitle("Group-2")
#+   geom_smooth(size = 0.5, alpha = 0.5)

data_heatmap_100 %>% 
  left_join(groups) %>% 
  filter(group==3) %>% 
  #filter(customer_id %in% c(8148781, 10703878)) %>% 
   ggplot(aes(x = as.integer(category))) +
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") + 
  geom_line(size = 0.5, aes(y = `50%`)) +
  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5)) +
  xlab("hour-day") +
  ggtitle("Group-3")
#+   geom_smooth(size = 0.5, alpha = 0.5)

data_heatmap_100 %>% 
  left_join(groups) %>% 
  filter(group==4) %>% 
  #filter(customer_id %in% c(8148781, 10703878)) %>% 
    ggplot(aes(x = as.integer(category))) +
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") + 
  geom_line(size = 0.5, aes(y = `50%`)) +
  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5))+
  xlab("hour-day") +
  ggtitle("Group-4")
#+ geom_smooth(size = 0.5, alpha = 0.5)

data_heatmap_100 %>% 
  left_join(groups) %>% 
  filter(group==5) %>% 
  #filter(customer_id %in% c(8148781, 10703878)) %>% 
   ggplot(aes(x = as.integer(category))) +
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") + 
  geom_line(size = 0.5, aes(y = `50%`)) +
  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5)) +
  xlab("hour-day") +
  ggtitle("Group-5")
#+geom_smooth(size = 0.5, alpha = 0.5)

data_heatmap_100 %>% 
  left_join(groups) %>% 
  filter(group==6) %>% 
  #filter(customer_id %in% c(8148781, 10703878)) %>% 
    ggplot(aes(x = as.integer(category))) +
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") + 
  geom_line(size = 0.5, aes(y = `50%`)) +
  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5))+
  xlab("hour-day") +
  ggtitle("Group-6")
#+ geom_smooth(size = 0.5, alpha = 0.5)

data_heatmap_100 %>% 
  left_join(groups) %>% 
  filter(group==7) %>% 
  #filter(customer_id %in% c(8148781, 10703878)) %>% 
   ggplot(aes(x = as.integer(category))) +
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") + 
  geom_line(size = 0.5, aes(y = `50%`)) +
  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5))+
  xlab("hour-day") +
  ggtitle("Group-7")
#+  geom_smooth(size = 0.5, alpha = 0.5)

data_heatmap_100 %>% 
  left_join(groups) %>% 
  filter(group==8) %>% 
  #filter(customer_id %in% c(8148781, 10703878)) %>% 
    ggplot(aes(x = as.integer(category))) +
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") + 
  geom_line(size = 0.5, aes(y = `50%`)) +
  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5))+
  xlab("hour-day") +
  ggtitle("Group-8")
#+ geom_smooth(size = 0.5, alpha = 0.5)

data_heatmap_100 %>% 
  left_join(groups) %>% 
  filter(group==9) %>% 
  #filter(customer_id %in% c(8148781, 10703878)) %>% 
   ggplot(aes(x = as.integer(category))) +
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") + 
  geom_line(size = 0.5, aes(y = `50%`)) +
  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5))+
  xlab("hour-day") +
  ggtitle("Group-9")
#+geom_smooth(size = 0.5, alpha = 0.5)

data_heatmap_100 %>% 
  left_join(groups) %>% 
  filter(group==10) %>% 
  #filter(customer_id %in% c(8148781, 10703878)) %>% 
  ggplot(aes(x = as.integer(category))) +
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") + 
  geom_line(size = 0.5, aes(y = `50%`)) +
  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5))+
  xlab("hour-day") +
  ggtitle("Group-10")
#+ geom_smooth(size = 0.5, alpha = 0.5) +
```

# Which variables are important for this clustering

I can do a ggpairs or parallel coordinate plot for this.
```{r}

data_pcp <- data_heatmap_100 %>% 
  left_join(groups)  %>% 
  select(customer_id, group, category, `50%`) %>% 
  mutate(category = as.factor(as.integer(category))) %>% 
  pivot_wider(names_from = category, values_from = `50%`) %>% 
  mutate(group = as.factor(group))

parcoord <- GGally::ggparcoord(data_pcp ,
                   columns = 3:ncol(data_pcp),
                   groupColumn = "group",
                   showPoints = FALSE, 
                   alphaLines = 0.8,
                   order = "allClass",
                   scale = "globalminmax"
) + ggplot2::theme_bw() +
  #scale_color_brewer(palette = "Dark2") + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")

ggpairs_color <- ggscatmat(data_pcp, columns=3:ncol(data_pcp)) +
  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")
parcoord
```

```{r}
ggpairs_color
```

```{r}
pc <- prcomp(data_pcp[c(-1,-2)], center = F, scale = F)
summary(pc)
# First two principal componentsafc
comp <- data.frame(pc$x[,1:2])
data_pcp_all <- data_pcp %>% bind_cols(comp)

library(RColorBrewer)
library(scales)
data_pcp_all %>% ggplot(aes(x=PC1, y = PC2)) + geom_point(aes(color = group)) +
  #scale_color_brewer(palette = "Dark2") +
   theme_bw() + theme(legend.position = "bottom")

```

# See if month_year works for your case
# Try with two granularities

### 2021-09-13 supplementary/algorithm3/real-data/clean-dataset_iqr.Rmd
add structure to methodology

--- 2021-09-13 supplementary/algorithm3/real-data/clean-dataset.Rmd
@@
 
-# library(gracsr)
+library(gracsr)
 # v2 <- suppressWarnings(
-#   scaled_dist_gran(data, "hour_day", 
+#   scaled_dist_gran_iqr(data, "hour_day",
 #                    response = "general_supply_kwh",
@@
 # 
-# write_rds(v2, "../../../data/scaled_dist_nogap_2013_100.rds")
-v2 <- read_rds("../../../data/scaled_dist_nogap_2013_100.rds")
+# write_rds(v2, "../../../data/scaled_dist_nogap_2013_100_iqr.rds")
+
+v2 <- read_rds("../../../data/scaled_dist_nogap_2013_100_iqr.rds")
 
@@
   arrange(dist_hod) %>% 
-  head(n=5)
+  head(n=7)
 
@@
   arrange(dist_hod) %>%
-  head(n=5)
+  head(n=7)
 
@@
   arrange(dist_hod) %>%
-  filter(customer_from == 11013154) %>% 
+  filter(customer_from == 8196501) %>% 
   arrange(dist_hod) %>% 
-   head(n=5)
+   head(n=7)
 
@@
                        c(data_pick_twopeak$customer_from, data_pick_twopeak$customer_to) %>% unique() %>% as_tibble() %>% set_names("customer_id"),
-                       c(data_pick_threepeak$customer_from, data_pick_threepeak$customer_to) %>% unique() %>% as_tibble() %>% set_names("customer_id"), .id = "design") %>% 
-  filter(!(customer_id %in% c(8171227, 10183908)))
+                       c(data_pick_threepeak$customer_from, data_pick_threepeak$customer_to) %>% unique() %>% as_tibble() %>% set_names("customer_id"), .id = "design") 
+#%>%   filter(!(customer_id %in% c(8171227, 10183908)))
+
+data_pick_clust_clean <- read_rds("../../../data/data_pick_cust_clean.rds")
 
 data <- read_rds("../../../data/elec_nogap_2013_100.rds")
+# data_pick <- data %>% filter(customer_id %in% data_pick_cust$customer_id) %>% select(customer_id,reading_datetime, general_supply_kwh) %>% left_join(data_pick_cust)
 
-data_pick <- data %>% filter(customer_id %in% data_pick_cust$customer_id) %>% select(customer_id,reading_datetime, general_supply_kwh) %>% left_join(data_pick_cust)
-
-write_csv(data_pick_cust, "../../../data/data_pick_cust_clean.rds")
+data_pick <- data %>% 
+  left_join(data_pick_cust %>% filter(!(customer_id %in% c(8283102, 10208156)))) %>% 
+filter(customer_id %in% data_pick_cust$customer_id) %>% select(customer_id,reading_datetime, general_supply_kwh, design) 
 ```
@@
              labeller = "label_value",
-             ncol = 5) +
+             ncol = 6) +
   theme_bw() +
@@
 ```{r transformed-scale}
-data_pick_scaled <- data_pick %>%
-    group_by(customer_id) %>%
-    dplyr::mutate(general_supply_kwh =  stats::qqnorm(general_supply_kwh, plot.it=FALSE)$x) %>%
-    ungroup()
 
-sm_hod <- gravitas::create_gran(data_pick_scaled, "hour_day")
-
-sm_hod_list <- sm_hod %>% 
-  as_tibble() %>% 
-  select(design, customer_id, hour_day, general_supply_kwh) %>% 
-  pivot_wider(names_from = hour_day,
-              values_from = general_supply_kwh)
-
-ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, 2)]))
-nrow_sm <- seq_len(nrow(sm_hod_list))
-
-sm_hod_quantiles_cat <- map(nrow_sm, function(x){
-  map(ncol_sm, function(y){
-   cell <- sm_hod_list%>% 
-     dplyr::filter(customer_id == customer_id[x]) %>% 
-     select(-c(1, 2)) %>% 
-     extract(y) %>% 
-     unlist()
-   quantile(cell, prob = quantile_prob_graph)
-})  %>% bind_rows(.id = "categories_serial_id")
-}) %>% bind_rows(.id = "customer_serial_id")
-
-ref_cat <- names(sm_hod_list)[-c(1, 2)] %>% as_tibble() %>% set_names("category") %>%  
-  mutate(categories_serial_id = row_number())
-
-ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
-    mutate(customer_serial_id = row_number())
-
-sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
-  mutate(customer_serial_id = as.integer(customer_serial_id),
-         categories_serial_id = as.integer(categories_serial_id)) %>%
-  left_join(ref_cat, by = "categories_serial_id") %>% 
-  left_join(ref_cust, by = "customer_serial_id") %>% 
-  select(customer_serial_id, categories_serial_id, category, customer_id, everything())
-
-data_heatmap <- sm_quantiles_ref %>% 
- left_join(data_pick %>% distinct(customer_id, design))
-
-data_heatmap %>% 
-  ggplot(aes(x = as.integer(category))) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") +
-  geom_line(aes(y = `50%`), size = 0.7) +
-  facet_wrap(design~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-             ncol = 5) +
-  theme_bw() +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")
+sm <- robust_scale_data(data_pick, "hour_day")
+v3 <- suppressWarnings(plot_data(sm, "hour_day", response = "scaled_response", nfacet = 6))
 ```
@@
 v2 <- suppressWarnings(
-  scaled_dist_gran(data_pick, "hour_day", 
+  scaled_dist_gran_iqr(data_pick, "hour_day", 
                    response = "general_supply_kwh",
@@
 v3 <- suppressWarnings(
-  scaled_dist_gran(data_pick, "day_month",
+  scaled_dist_gran_iqr(data_pick, "day_month",
                    response = "general_supply_kwh", 
@@
   left_join(v2) %>%
-  mutate(dist = dist_hod + dist_dom) %>% 
+  mutate(dist = dist_hod) %>% 
     pivot_wider(-c(3, 4), 
@@
 ```
-
-## hod + dom
-### Is the clustering sensitive to nuisance parameter?
-
-```{r hod-dom,echo = TRUE}
-#quantile_prob_val = c(0.5, 0.75)
-#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
-library(gracsr)
-v2 <- suppressWarnings(
-  scaled_dist_gran(data_pick, "hour_day", 
-                   response = "general_supply_kwh",
-                   quantile_prob_val = quantile_prob_clust)) %>% rename("dist_hod" = "dist")
-v3 <- suppressWarnings(
-  scaled_dist_gran(data_pick, "day_month",
-                   response = "general_supply_kwh", 
-                   quantile_prob_val = quantile_prob_clust)) %>% rename("dist_dom" = "dist")
-
-data_dist <- v3 %>% 
-  left_join(v2) %>%
-  mutate(dist = dist_hod + dist_dom) %>% 
-    pivot_wider(-c(3, 4), 
-                names_from = customer_to,
-              values_from = dist) %>%
-  rename("customer_id" = "customer_from")
-```
-
-```{r clustering-result-hod-dom}
-d = stats::as.dist(data_dist[,-1])
-hc = stats::hclust(d,method="ward.D")
-
-groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
-groups %>% group_by(group) %>% summarise(n=n())
-#groups %>% filter(group==1)
-#groups %>% filter(group==2)
-#groups %>% filter(group==3)
-
-data_heatmap_new <- data_heatmap %>% left_join(groups)
-
-data_heatmap_new$group = as.factor(data_heatmap_new$group)
-
-data_heatmap_new %>% 
-  ggplot(aes(x = as.integer(category))) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") +
-  geom_line(aes(y = `50%`, color = group), size = 0.7) +
-  facet_wrap(group~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-             ncol = 5) +
-  theme_bw() +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
-  scale_color_brewer(palette = "Dark2")
-```
-
-# All 100 customers
-### Running it on all 100 and making 10 clusters. Do they have sufficiently different shapes?
-### Variable importance
-
-```{r hod-100cust,echo = TRUE}
-# library(gracsr)
-# v2 <- suppressWarnings(
-#   scaled_dist_gran(data, "hour_day", 
-#                    response = "general_supply_kwh",
-#                    quantile_prob_val = quantile_prob_clust)) %>% rename("dist_hod" = "dist")
-
-#write_rds(v2, "../../../data/scaled_dist_100clust_deciles.rds")
-v2 <- read_rds("../../../data/scaled_dist_100clust_deciles.rds")
-
-data_dist <- v2 %>%
-  mutate(dist = dist_hod) %>% 
-    pivot_wider(-c(3, 4), 
-                names_from = customer_to,
-              values_from = dist) %>%
-  rename("customer_id" = "customer_from")
-
-cust_ref <- bind_cols(customer_id=data_dist$customer_id, row = seq_len(nrow(data_dist)))
-
-cust_rm_outlier <-   cust_ref %>% filter(!(row %in% c(73, 82, 79, 97, 25, 91)))
-
-v3 <- v2%>% 
-  filter(customer_from %in% cust_rm_outlier$customer_id,
-         customer_to %in% cust_rm_outlier$customer_id)
-
-data_dist <- v3 %>%
-  mutate(dist = dist_hod) %>% 
-    pivot_wider(-c(3, 4), 
-                names_from = customer_to,
-              values_from = dist) %>%
-  rename("customer_id" = "customer_from")
-
-```
-
-```{r hod-100cust-clustering}
-d = stats::as.dist(data_dist[,-1])
-hc = stats::hclust(d,method="ward.D")
-
-library(fpc)
-# koptimal = fpc::nselectboot(as.dist(data_dist[,-1]),
-#                         B = 5000,
-#                         method = "ward.D",
-#                         clustermethod = disthclustCBI, 
-#                         classification = "averagedist",
-#                         krange = 2:20)
-# 
-# koptimal = koptimal$kopt
-
-groups <- tibble(group = cutree(hc, k=10), customer_id = data_dist$customer_id)
-groups %>% group_by(group) %>% summarise(n=n())
-
-data <- read_rds("../../../data/elec_nogap_2013_100.rds")
-data <- data %>% left_join(groups)
-
-sm_hod <- gravitas::create_gran(data, "hour_day")
-
-sm_hod_list <- sm_hod %>% 
-  as_tibble() %>% 
-  select(customer_id, hour_day, general_supply_kwh, group) %>% 
-  pivot_wider(names_from = hour_day,
-              values_from = general_supply_kwh)
-
-ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, 2)]))
-nrow_sm <- unique(sm_hod_list$group)
-
-sm_hod_quantiles_cat <- map(nrow_sm, function(x){
-  map(ncol_sm, function(y){
-   cell <- sm_hod_list%>% 
-     dplyr::filter(group == group[x]) %>% 
-     select(-c(1, 2)) %>% 
-     extract(y) %>% 
-     unlist()
-   quantile(cell, prob = quantile_prob_graph)
-})  %>% bind_rows(.id = "categories_serial_id")
-}) %>% bind_rows(.id = "group_serial_id")
-
-ref_cat <- names(sm_hod_list)[-c(1, 2)] %>% as_tibble() %>% set_names("category") %>%  
-  mutate(categories_serial_id = row_number())
-
-ref_cust <- unique(sm_hod_list$group) %>% as_tibble()%>% set_names("group")  %>% 
-    mutate(group_serial_id = as.character(row_number()))
-
-sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
-  mutate(customer_serial_id = as.integer(group_serial_id),
-         categories_serial_id = as.integer(categories_serial_id)) %>% 
-  left_join(ref_cat, by = "categories_serial_id") %>% 
-  left_join(ref_cust, by = "group_serial_id") %>% 
-  select(group_serial_id, categories_serial_id, category, everything())
-
-# data_heatmap_100 <- sm_quantiles_ref %>% 
-#  left_join(data %>% distinct(customer_id))
-#
-
-sm_quantiles_ref$group = as.factor(sm_quantiles_ref$group)
-
-sm_quantiles_ref %>% 
-  ggplot(aes(x = as.integer(category))) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") +
-  geom_line(aes(y = `50%`), size = 0.7) +
-  facet_wrap(~group, 
-             scales = "free_y", 
-             labeller = "label_value") +
-  theme_bw() +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
-  scale_color_brewer(palette = "Dark2") + scale_color_brewer(palette = "Dark2")
-```
-# Split these groups to see if the shapes of individual customers in a group is the same
-
-```{r all100}
-data <- read_rds("../../../data/elec_nogap_2013_100.rds")
-
-sm_hod <- gravitas::create_gran(data, "hour_day")
-sm_hod_list <- sm_hod %>% 
-  as_tibble() %>% 
-  select(customer_id, hour_day, general_supply_kwh) %>% 
-  pivot_wider(names_from = hour_day,
-              values_from = general_supply_kwh)
-
-ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, 2)]))
-nrow_sm <- 1:100
-
-sm_hod_quantiles_cat <- map(nrow_sm, function(x){
-  map(ncol_sm, function(y){
-   cell <- sm_hod_list%>% 
-     dplyr::filter(customer_id == customer_id[x]) %>% 
-     select(-c(1, 2)) %>% 
-     extract(y) %>% 
-     unlist()
-   quantile(cell, prob = quantile_prob_graph)
-})  %>% bind_rows(.id = "categories_serial_id")
-}) %>% bind_rows(.id = "customer_serial_id")
-
-ref_cat <- names(sm_hod_list)[-c(1, 2)] %>% as_tibble() %>% set_names("category") %>%  
-  mutate(categories_serial_id = row_number())
-
-ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
-    mutate(customer_serial_id = row_number())
-
-sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
-  mutate(customer_serial_id = as.integer(customer_serial_id),
-         categories_serial_id = as.integer(categories_serial_id)) %>% 
-  left_join(ref_cat, by = "categories_serial_id") %>% 
-  left_join(ref_cust, by = "customer_serial_id") %>% 
-  select(customer_serial_id, categories_serial_id, category, customer_id, everything())
-
-data_heatmap_100 <- sm_quantiles_ref %>% 
- left_join(data %>% distinct(customer_id))
-
-data_heatmap_100 %>% 
-  left_join(groups) %>% 
-  filter(group==1) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-    ggplot(aes(x = as.integer(category))) +
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") + 
-  geom_line(size = 0.5, aes(y = `50%`)) +
-  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5)) +
-  xlab("hour-day") +
-  ggtitle("Group-2")
-
-data_heatmap_100 %>% 
-  left_join(groups) %>% 
-  filter(group==2) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-    ggplot(aes(x = as.integer(category))) +
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") + 
-  geom_line(size = 0.5, aes(y = `50%`)) +
-  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5)) +
-  xlab("hour-day") +
-  ggtitle("Group-2")
-#+   geom_smooth(size = 0.5, alpha = 0.5)
-
-data_heatmap_100 %>% 
-  left_join(groups) %>% 
-  filter(group==3) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-   ggplot(aes(x = as.integer(category))) +
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") + 
-  geom_line(size = 0.5, aes(y = `50%`)) +
-  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5)) +
-  xlab("hour-day") +
-  ggtitle("Group-3")
-#+   geom_smooth(size = 0.5, alpha = 0.5)
-
-data_heatmap_100 %>% 
-  left_join(groups) %>% 
-  filter(group==4) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-    ggplot(aes(x = as.integer(category))) +
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") + 
-  geom_line(size = 0.5, aes(y = `50%`)) +
-  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5))+
-  xlab("hour-day") +
-  ggtitle("Group-4")
-#+ geom_smooth(size = 0.5, alpha = 0.5)
-
-data_heatmap_100 %>% 
-  left_join(groups) %>% 
-  filter(group==5) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-   ggplot(aes(x = as.integer(category))) +
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") + 
-  geom_line(size = 0.5, aes(y = `50%`)) +
-  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5)) +
-  xlab("hour-day") +
-  ggtitle("Group-5")
-#+geom_smooth(size = 0.5, alpha = 0.5)
-
-data_heatmap_100 %>% 
-  left_join(groups) %>% 
-  filter(group==6) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-    ggplot(aes(x = as.integer(category))) +
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") + 
-  geom_line(size = 0.5, aes(y = `50%`)) +
-  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5))+
-  xlab("hour-day") +
-  ggtitle("Group-6")
-#+ geom_smooth(size = 0.5, alpha = 0.5)
-
-data_heatmap_100 %>% 
-  left_join(groups) %>% 
-  filter(group==7) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-   ggplot(aes(x = as.integer(category))) +
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") + 
-  geom_line(size = 0.5, aes(y = `50%`)) +
-  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5))+
-  xlab("hour-day") +
-  ggtitle("Group-7")
-#+  geom_smooth(size = 0.5, alpha = 0.5)
-
-data_heatmap_100 %>% 
-  left_join(groups) %>% 
-  filter(group==8) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-    ggplot(aes(x = as.integer(category))) +
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") + 
-  geom_line(size = 0.5, aes(y = `50%`)) +
-  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5))+
-  xlab("hour-day") +
-  ggtitle("Group-8")
-#+ geom_smooth(size = 0.5, alpha = 0.5)
-
-data_heatmap_100 %>% 
-  left_join(groups) %>% 
-  filter(group==9) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-   ggplot(aes(x = as.integer(category))) +
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") + 
-  geom_line(size = 0.5, aes(y = `50%`)) +
-  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5))+
-  xlab("hour-day") +
-  ggtitle("Group-9")
-#+geom_smooth(size = 0.5, alpha = 0.5)
-
-data_heatmap_100 %>% 
-  left_join(groups) %>% 
-  filter(group==10) %>% 
-  #filter(customer_id %in% c(8148781, 10703878)) %>% 
-  ggplot(aes(x = as.integer(category))) +
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`), fill = "lightblue") + 
-  geom_line(size = 0.5, aes(y = `50%`)) +
-  facet_wrap(~customer_serial_id, scales = "free_y") + theme_minimal() + theme(strip.text.y = element_text(size = 5))+
-  xlab("hour-day") +
-  ggtitle("Group-10")
-#+ geom_smooth(size = 0.5, alpha = 0.5) +
-```
-
-# Which variables are important for this clustering
-
-I can do a ggpairs or parallel coordinate plot for this.
-```{r}
-
-data_pcp <- data_heatmap_100 %>% 
-  left_join(groups)  %>% 
-  select(customer_id, group, category, `50%`) %>% 
-  mutate(category = as.factor(as.integer(category))) %>% 
-  pivot_wider(names_from = category, values_from = `50%`) %>% 
-  mutate(group = as.factor(group))
-
-parcoord <- GGally::ggparcoord(data_pcp ,
-                   columns = 3:ncol(data_pcp),
-                   groupColumn = "group",
-                   showPoints = FALSE, 
-                   alphaLines = 0.8,
-                   order = "allClass",
-                   scale = "globalminmax"
-) + ggplot2::theme_bw() +
-  #scale_color_brewer(palette = "Dark2") + 
-  ggplot2::theme(
-    plot.title = ggplot2::element_text(size=10)
-  )+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
-  theme(legend.position = "bottom") +
-  xlab("") +
-  ylab("wpd")
-
-ggpairs_color <- ggscatmat(data_pcp, columns=3:ncol(data_pcp)) +
-  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")
-parcoord
-```
-
-```{r}
-ggpairs_color
-```
-
-```{r}
-pc <- prcomp(data_pcp[c(-1,-2)], center = F, scale = F)
-summary(pc)
-# First two principal componentsafc
-comp <- data.frame(pc$x[,1:2])
-data_pcp_all <- data_pcp %>% bind_cols(comp)
-
-library(RColorBrewer)
-library(scales)
-data_pcp_all %>% ggplot(aes(x=PC1, y = PC2)) + geom_point(aes(color = group)) +
-  #scale_color_brewer(palette = "Dark2") +
-   theme_bw() + theme(legend.position = "bottom")
-
-```
-
-# See if month_year works for your case
-# Try with two granularities

### 2021-09-13 supplementary/algorithm3/real-data/handpick-18.Rmd
add structure to methodology

@@
 #   facet_wrap(~customer_id, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
+
 ```
@@
 # quantile_prob_val = c(0.5, 0.75)
+
+```{r scaled-data}
+
+data_pick_scaled <- data_pick %>%
+    group_by(customer_id) %>%
+    dplyr::mutate(kwh_scaled = stats::qqnorm(general_supply_kwh, plot.it=FALSE)$x) %>%
+    ungroup()
+
+quantile_prob_val = c(0.25, 0.5, 0.75)
+
+#sm_hod_scaled <- gravitas::create_gran(data_pick_scaled, "hour_day")
+
+sm_hod_all <- data_pick %>% 
+bind_cols(data_pick_scaled %>% as_tibble() %>% select(kwh_scaled)) %>% 
+  rename("kwh" = "general_supply_kwh") %>% 
+gravitas::create_gran("hour_day") %>% 
+  pivot_longer(c(5, 11), names_to = "kwh_type", values_to = "kwh_value")
+
+sm_hod_all %>%
+filter(design ==1) %>%
+ggplot(aes(x = kwh_value, y = hour_day, fill = kwh_type)) +
+ggridges::geom_density_ridges(alpha = 0.5) +
+facet_wrap(~customer_id, scales = "free_x") +
+  stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.5) +
+    coord_flip() +
+  theme(legend.position = "bottom")
+
+sm_hod_all %>%
+filter(customer_id %in% c(8181071)) %>%
+ggplot(aes(x = kwh_value, y = hour_day), fill = "blue") +
+ggridges::geom_density_ridges(alpha = 0.5) +
+facet_grid(customer_id~kwh_type, scales = "free_x") +
+  stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.5, color = "blue") +
+    #coord_flip() +
+  theme(legend.position = "bottom")
+
+sm_hod_all %>%
+filter(customer_id %in% c(11013154)) %>%
+ggplot(aes(x = kwh_value, y = hour_day), fill = "blue") +
+ggridges::geom_density_ridges(alpha = 0.5) +
+facet_grid(customer_id~kwh_type, scales = "free_x") +
+  stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.5, color = "blue") +
+    #coord_flip() +
+  theme(legend.position = "bottom")
+
+quantile_q2 <-  function(x){
+  y = quantile(x, probs = c(0.5))
+  #c(y[1], y[2]) %>% as_tibble() %>% bind_cols(names(y)) %>% set_names(c("quant_value", "quantile"))
+}
+
+quantile_q1 <-  function(x){
+  y = quantile(x, probs = c(0.25))
+  #c(y[1], y[2]) %>% as_tibble() %>% bind_cols(names(y)) %>% set_names(c("quant_value", "quantile"))
+}
+
+quantile_q3 <-  function(x){
+  y = quantile(x, probs = c(0.75))
+  #c(y[1], y[2]) %>% as_tibble() %>% bind_cols(names(y)) %>% set_names(c("quant_value", "quantile"))
+}
+
+# sm_hod_all %>%
+# filter(customer_id %in% c(8955520)) %>%
+# ggplot(aes(x = kwh_value, y = hour_day), fill = "blue") +
+# ggridges::geom_density_ridges(alpha = 0.5) +
+# facet_grid(customer_id~kwh_type, scales = "free_x") +
+#   ggridges::stat_density_ridges(quantile_lines = TRUE, quantiles = 0.9) + stat_summary(
+#     fun = quantile_q2,
+#     geom = 'line',
+#     aes(group = 1), size = 0.5, color = "blue") +
+#     #coord_flip() +
+#   theme(legend.position = "bottom")
+
+#one_peak
+
+sm_hod_all %>%
+filter(customer_id %in% c(8198401)) %>%
+ggplot(aes(x = kwh_value, y = hour_day), fill = "#999999") +
+ggridges::geom_density_ridges(alpha = 0.7) +
+facet_wrap(~kwh_type, scales = "free_y", ncol = 1) +
+  stat_summary(
+    fun = quantile_q2,
+    geom = 'line',
+    aes(group = 1), size = 0.7, color = "blue") +
+    coord_flip() +
+  theme(legend.position = "bottom") +
+  stat_summary(
+    fun = quantile_q1,
+    geom = 'line',
+    aes(group = 1), size = 0.7, color = "#D55E00") +
+  theme(legend.position = "bottom") +
+   stat_summary(
+    fun = quantile_q3,
+    geom = 'line',
+    aes(group = 1), size = 0.7, color = "#D55E00") +
+    #coord_flip() +
+  theme(legend.position = "bottom") +
+  theme_bw()
+
+# two peaks
+sm_hod_all %>%
+filter(customer_id %in% c(10469265)) %>%
+ggplot(aes(x = kwh_value, y = hour_day), fill = "#999999") +
+ggridges::geom_density_ridges(alpha = 0.7) +
+facet_wrap(~kwh_type, scales = "free_y", ncol = 1) +
+  stat_summary(
+    fun = quantile_q2,
+    geom = 'line',
+    aes(group = 1), size = 0.7, color = "blue") +
+    coord_flip() +
+  theme(legend.position = "bottom") +
+  stat_summary(
+    fun = quantile_q1,
+    geom = 'line',
+    aes(group = 1), size = 0.7, color = "#D55E00") +
+  theme(legend.position = "bottom") +
+   stat_summary(
+    fun = quantile_q3,
+    geom = 'line',
+    aes(group = 1), size = 0.7, color = "#D55E00") +
+    #coord_flip() +
+  theme(legend.position = "bottom") +
+  theme_bw()
+
+# Three peaks
+
+sm_hod_all %>%
+filter(customer_id %in% c(10046512)) %>%
+ggplot(aes(x = kwh_value, y = hour_day), fill = "#999999") +
+ggridges::geom_density_ridges(alpha = 0.7) +
+facet_wrap(~kwh_type, scales = "free_y", ncol = 1) +
+  stat_summary(
+    fun = quantile_q2,
+    geom = 'line',
+    aes(group = 1), size = 0.7, color = "blue") +
+    coord_flip() +
+  theme(legend.position = "bottom") +
+  stat_summary(
+    fun = quantile_q1,
+    geom = 'line',
+    aes(group = 1), size = 0.7, color = "#D55E00") +
+  theme(legend.position = "bottom") +
+   stat_summary(
+    fun = quantile_q3,
+    geom = 'line',
+    aes(group = 1), size = 0.7, color = "#D55E00") +
+    #coord_flip() +
+  theme(legend.position = "bottom") +
+  theme_bw()
+
+# sm_hod_all %>%
+# filter(customer_id %in% c(8198401)) %>%
+# ggplot(aes(x = kwh_value, y = hour_day), fill = "#999999") +
+# ggridges::geom_density_ridges() +
+# facet_grid(customer_id~kwh_type, scales = "free_x") +
+#   stat_summary(
+#     fun = quantile_q2,
+#     geom = 'line',
+#     aes(group = 1), size = 0.5, color = "blue") +
+#     #coord_flip() +
+#   theme(legend.position = "bottom") +
+#   stat_summary(
+#     fun = quantile_q1,
+#     geom = 'line',
+#     aes(group = 1), size = 0.5, color = "#D55E00") +
+#   theme(legend.position = "bottom") +
+#    stat_summary(
+#     fun = quantile_q3,
+#     geom = 'line',
+#     aes(group = 1), size = 0.5, color = "#D55E00") +
+#     #coord_flip() +
+#   theme(legend.position = "bottom") +
+#   theme_bw()
+
+# sm_hod_all %>%
+# filter(customer_id %in% c(10469265)) %>%
+# ggplot(aes(x = kwh_value, y = hour_day), fill = "#999999") +
+# ggridges::geom_density_ridges() +
+# facet_grid(customer_id~kwh_type, scales = "free_x") +
+#   stat_summary(
+#     fun = quantile_q2,
+#     geom = 'line',
+#     aes(group = 1), size = 0.5, color = "blue") +
+#     #coord_flip() +
+#   theme(legend.position = "bottom") +
+#   stat_summary(
+#     fun = quantile_q1,
+#     geom = 'line',
+#     aes(group = 1), size = 0.5, color = "#D55E00") +
+#   theme(legend.position = "bottom") +
+#    stat_summary(
+#     fun = quantile_q3,
+#     geom = 'line',
+#     aes(group = 1), size = 0.5, color = "#D55E00") +
+#     #coord_flip() +
+#   theme(legend.position = "bottom") +
+#   theme_bw()
+
+p1 <- sm_hod %>% 
+filter(design ==1) %>% 
+ggplot(aes(x = general_supply_kwh, y = hour_day)) +
+ggridges::geom_density_ridges(alpha = 0.05) +
+facet_wrap(~customer_id, nrow = 6) + 
+  stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.5) +
+    coord_flip()
+
+p2 <- gravitas::create_gran(data_pick_scaled, "hour_day") %>% 
+filter(design ==1) %>% 
+ggplot(aes(x = kwh_scaled, y = hour_day)) +
+ggridges::geom_density_ridges(alpha = 0.05, fill = "grey") +
+facet_wrap(~customer_id, ncol = 2, scales = "free_x") + 
+  stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.5) +
+    coord_flip() + theme_bw()
+
+sm_hod_list <- sm_hod %>% 
+  as_tibble() %>% 
+  select(design, customer_id, hour_day, general_supply_kwh) %>% 
+  pivot_wider(names_from = hour_day,
+              values_from = general_supply_kwh)
+
+# groups_ref <- unique(data_pick$design) %>% as_tibble %>% 
+#               mutate(customer_serial_id = sm_hod_list$customer_id)
+
+# sm_hod_list_cat <-  sm_hod_list %>% left_join(groups_ref, by = c("customer_id" = "customer_serial_id")) %>% select(group, everything())
+
+ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, 2)]))
+nrow_sm <- 1:18
+
+sm_hod_quantiles_cat <- map(nrow_sm, function(x){
+  map(ncol_sm, function(y){
+   cell <- sm_hod_list%>% 
+     dplyr::filter(customer_id == customer_id[x]) %>% 
+     select(-c(1, 2)) %>% 
+     extract(y) %>% 
+     unlist()
+   quantile(cell, prob = quantile_prob_val)
+})  %>% bind_rows(.id = "categories_serial_id")
+}) %>% bind_rows(.id = "customer_serial_id")
+
+ref_cat <- names(sm_hod_list)[-c(1, 2)] %>% as_tibble() %>% set_names("category") %>%  
+  mutate(categories_serial_id = row_number())
+
+ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
+    mutate(customer_serial_id = row_number())
+
+sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
+  mutate(customer_serial_id = as.integer(customer_serial_id),
+         categories_serial_id = as.integer(categories_serial_id)) %>% 
+  left_join(ref_cat, by = "categories_serial_id") %>% 
+  left_join(ref_cust, by = "customer_serial_id") %>% 
+  select(customer_serial_id, categories_serial_id, category, customer_id, everything())
+
+data_heatmap <- sm_quantiles_ref %>% 
+ left_join(data_pick %>% distinct(customer_id, design))
+
+# %>% 
+#   left_join(categories_ref_cat) %>% 
+#   rename("category" = "value") %>% 
+#   select(-categories_serial_id) %>% 
+#   mutate(category = as.numeric(category))
+
+data_heatmap %>% 
+  ggplot(aes(x = as.integer(category))) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`), fill = "lightblue") +
+  geom_line(aes(y = `50%`), size = 0.7) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+  theme_bw() +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")
+
+```
 

### 2021-09-13 supplementary/algorithm3/simulation-data/diffmeans-threegran-5designs.Rmd
add structure to methodology

--- 2021-08-05 supplementary/simulation-alogorithm2/sim_setup_algo2.Rmd
@@
 }) %>% bind_rows(.id = "seed_id")
+```
+
+```{r}
+
+quantile_prob_val = seq(0.1,0.9,0.1)
+bind_data_iter <- bind_data_iter %>% mutate(
+  customer_id  = paste0(design,"-s-", seed_id)) %>%
+   as_tsibble(key=customer_id,index=t)
+
+library(gracsr)
+v2 <- suppressWarnings(
+  scaled_dist_gran(bind_data_iter, "g1", 
+                   response = "value",
+                   quantile_prob_val = quantile_prob_val))%>% rename("dist_g1" = "dist")
+
+v3 <- suppressWarnings(
+  scaled_dist_gran(bind_data_iter, "g2", 
+                   response = "value",
+                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_g2" = "dist")
+
+v4 <- suppressWarnings(
+  scaled_dist_gran(bind_data_iter, "g3", 
+                   response = "value",
+                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_g3" = "dist")
+
+data_dist <- v3 %>% 
+  left_join(v2) %>%
+  left_join(v4) %>% 
+  mutate(dist = dist_g1 + dist_g2 +dist_g3) %>% 
+    pivot_wider(-c(3, 4, 5), 
+                names_from = customer_to,
+              values_from = dist) %>%
+  rename("customer_id" = "customer_from")
 
@@
 
-<!-- Cross-check -->
-<!-- data = bind_data_iter %>% filter(design == "design-2", seed_id == 5) -->
-<!-- p2 <- ggplot(data, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw() + facet_wrap(~design, ncol = 1) -->
-<!-- p3 <- ggplot(data, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw() + facet_wrap(~design, ncol = 1) -->
-<!-- p4 <- ggplot(data, aes(x = as.factor(g3), y = value)) + geom_boxplot() + -->
-<!-- xlab("g3") + theme_bw() + facet_wrap(~design, ncol = 1) -->
-<!-- (p2 + p3 + p4) -->
+```{r groups}
+d = stats::as.dist(data_dist[,-1])
+hc = stats::hclust(d,method="complete")
 
-<!-- ```{r} -->
-<!-- p2 <- ggplot(data, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw() -->
-<!-- p3 <- ggplot(data, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw() -->
-<!-- p4 <- ggplot(data, aes(x = as.factor(g3), y = value)) + geom_boxplot() + -->
-<!-- xlab("g3") + theme_bw()  -->
-<!-- (p2 + p3 + p4) -->
-<!-- ``` -->
-
-```{r}
-harmony_tbl <- tibble(facet_variable = NA, x_variable = c("g1", "g2", "g3"),
-                      x_levels = c(2, 3, 4), facet_levels = NA)
+groups <- tibble(group = cutree(hc, k=5), customer_id = data_dist$customer_id)
+groups %>% group_by(group) %>% summarise(n=n())
+groups %>% filter(group==1)
+groups %>% filter(group==2)
+groups %>% filter(group==3)
+groups %>% filter(group==4)
+groups %>% filter(group==5)
+#groups %>% filter(group==4)
 ```
-
-```{r wpd-table}
-data <- bind_data_iter %>% mutate(data_id = paste(design, seed_id, sep = "-"))
-bind_data_iter <- tsibble::as_tsibble(data, index = t, key = "data_id")
-
-data_wpd <- map(seq_len(length(unique(bind_data_iter$data_id))),
-    function(y){
-      z = unique(bind_data_iter$data_id)
-      data_sub <- bind_data_iter %>% filter(data_id==z[y])
-        elec_select_harmony = hakear::select_harmonies(data_sub,
-                               harmony_tbl = harmony_tbl,
-                               response = value,
-                               nperm = 200
-  ) %>% mutate(data_id = unique(data_sub$data_id)) %>% unnest()
-    }) %>% bind_rows()
-
-write_csv(data_wpd, "data_wpd.rds")
-```
-
-# Table of wpd across designs and wpd
-
-```{r data-wide}
-
-data_wide <- data_wpd %>% select(x_variable, wpd, data_id) %>% 
-  pivot_wider(names_from = x_variable, values_from = wpd)
-data_wide %>% kable()
-
-ggpairs_nocolor <- ggpairs(data_wide, columns = 2:4)
-
-rownames(data_wide) = data_wide$data_id
-
-hc = stats::hclust(dist(data_wide[-1], method = "manhattan"),method="complete")
-
-library(fpc)
-koptimal = fpc::nselectboot(dist(data_wide[-1], method="manhattan"),
-                        B = 500,
-                        method = "complete",
-                        clustermethod = disthclustCBI, 
-                        classification = "averagedist",
-                        krange = 2:8)
-
-koptimal = koptimal$kopt
-```
-
-# Clustering of designs
-
-```{r}
-  groups <- cutree(hc, k=5) %>% as_tibble() %>% 
-mutate(data_id = data_wide$data_id) %>% rename("group" = "value")
-
-  groups %>% kable()
-
-data_wide_group <- data_wide %>% left_join(groups)  %>% 
-  mutate(group_id = as.factor(group))
-
-parcoord <- GGally::ggparcoord(data_wide_group,
-                   columns = 2:ncol(data_wide),
-                   groupColumn = "group_id",
-                   showPoints = TRUE, 
-                   title = "(c)",
-                   alphaLines = 0.8,
-                   scale = "globalminmax", 
-                   order = "anyClass"
-) + ggplot2::theme_bw() +
-  scale_color_brewer(palette = "Dark2") + 
-  ggplot2::theme(
-    plot.title = ggplot2::element_text(size=10)
-  )+
-  ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 0)) +
-  theme(legend.position = "bottom") +
-  xlab("") +
-  ylab("wpd")
-parcoord
-ggpairs_nocolor
-  # ggpairs_color <- ggscatmat(data_wide, columns=2:4, color = "group") +
-  # scale_colour_brewer(palette="Set2")
-```
-
-# With optimal number of clusters
-
-Design 2 and 4 form a part of the same cluster with the optimal number of clusters chosen as 4.
-Design 2 and 4 are similar in the sense that both of them changes across categories for g1 and g3, and remains constant for g2.
-
-```{r}
-  groups <- cutree(hc, k=koptimal) %>% as_tibble() %>% 
-mutate(data_id = data_wide$data_id) %>% rename("group" = "value")
-
-  groups %>% kable()
-
-data_wide_group <- data_wide %>% left_join(groups)  %>% 
-  mutate(group_id = as.factor(group))
-
-parcoord <- GGally::ggparcoord(data_wide_group,
-                   columns = 2:ncol(data_wide),
-                   groupColumn = "group_id",
-                   showPoints = TRUE, 
-                   title = "(c)",
-                   alphaLines = 0.8,
-                   scale = "globalminmax", 
-                   order = "anyClass"
-) + ggplot2::theme_bw() +
-  scale_color_brewer(palette = "Dark2") + 
-  ggplot2::theme(
-    plot.title = ggplot2::element_text(size=10)
-  )+
-  ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 0)) +
-  theme(legend.position = "bottom") +
-  xlab("") +
-  ylab("wpd")
-parcoord
-ggpairs_nocolor
-  # ggpairs_color <- ggscatmat(data_wide, columns=2:4, color = "group") +
-  # scale_colour_brewer(palette="Set2")
-```
-
-# PCA based clustering
-
-```{r}
-pc <- prcomp(data_wide_group[c(2, 3, 4)], center = F, scale = F)
-
-# First two principal componentsafc
-comp <- data.frame(pc$x[,1:2])
-
-#plot(comp, pch=16, col=rgb(0,0,0,0.5))
-
-hc = stats::hclust(dist(comp))
-groups <- cutree(hc, k=koptimal) %>% as_tibble() %>% bind_cols(comp) %>% 
-  mutate(value = as.character(value))
-library(RColorBrewer)
-library(scales)
-groups %>% ggplot(aes(x=PC1, y = PC2)) + geom_point(aes(color = value)) +
-  scale_color_brewer(palette = "Dark2") + theme_bw()
-```

### 2021-09-13 supplementary/effect_transformation.Rmd
add structure to methodology

---
title: "check_effect_transformation"
author: "Sayani Gupta"
date: "01/09/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(gravitas)
library(tidyverse)
library(gracsr)
library(tsibble)

# Feed in data and other inputs
sm <- smart_meter10 %>%
filter(customer_id %in% c("10006704", "10017936","10006414", "10018250"))
gran1 = "hour_day"
gran2 = NULL
response = "general_supply_kwh"

# Scale the data

v2 <- suppressWarnings(robust_scale_data(sm, "hour_day")) %>% 
    dplyr::mutate(kwh_nqt = stats::qqnorm(general_supply_kwh, plot.it=FALSE)$x) %>% 
  mutate(category = as.numeric(category),
         kwh_robust = )

quantile_q2 <-  function(x){
  y = quantile(x, probs = c(0.5))
  #c(y[1], y[2]) %>% as_tibble() %>% bind_cols(names(y)) %>% set_names(c("quant_value", "quantile"))
}

quantile_q1 <-  function(x){
  y = quantile(x, probs = c(0.25))
  #c(y[1], y[2]) %>% as_tibble() %>% bind_cols(names(y)) %>% set_names(c("quant_value", "quantile"))
}

quantile_q3 <-  function(x){
  y = quantile(x, probs = c(0.75))
  #c(y[1], y[2]) %>% as_tibble() %>% bind_cols(names(y)) %>% set_names(c("quant_value", "quantile"))
}

v2 %>%
dplyr::filter(customer_id %in% c("10018250")) %>%
  pivot_longer(c("general_supply_kwh", "scaled_response", "kwh_nqt"),
               names_to = "kwh_type",
               values_to = "kwh_value") %>% ggplot(fill = "#999999") +
  ggridges::geom_density_ridges(aes(x = kwh_value, y = as.factor(category)), alpha = 0.7) +
  facet_grid(kwh_type~customer_id, scales = "free") +
  coord_flip() +
  theme(legend.position = "bottom") +
  theme_bw()

v2 %>%
dplyr::filter(customer_id %in% c("10018250")) %>%
  pivot_longer(c("general_supply_kwh", "scaled_response", "kwh_nqt"),
               names_to = "kwh_type",
               values_to = "kwh_value") %>%
ggplot(aes(x = kwh_value, y = as.factor(category)), fill = "#999999") +
  #ggridges::geom_density_ridges(alpha = 0.7) +
  facet_grid(kwh_type~customer_id, scales = "free") +
  coord_flip() +
  stat_summary(
    fun = quantile_q2,
    geom = 'line',
    aes(group = 1), size = 1, color = "blue") +
  theme(legend.position = "bottom") +
  stat_summary(
    fun = quantile_q1,
    geom = 'line',
    aes(group = 1), size = 1, color = "darkgreen") +
  theme(legend.position = "bottom") +
  stat_summary(
    fun = quantile_q3,
    geom = 'line',
    aes(group = 1), size = 1, color = "#D55E00") +
  theme(legend.position = "bottom") +
  theme_bw()

# library(tidyverse)
#
# # for loop
#
# for (x in 1:2){
#   for(y in 1:2){
#     for(z in 1:5){
#       dist_data[x, y] = x*y + y*z
#     }
#   }
#
# }
#
#
#
# tab <- expand.grid(x = 1:2, y = 1:2, z =1:5)
# tab
#
# # Using map
#
#
#
# # using pmap
#
# dist_data <- purrr::pmap(tab,
#                                   function(x, y, z){
#                                     value3 =
#                                       x*y + y*z
# }) %>%
#   unlist () %>%
#   as_tibble() %>%
#   bind_cols(tab)

```

### 2021-09-13 script/smart_meter.R

@@ -35,22 +35,25 @@ readr::write_rds(elec_ts, "data/elec_ts.rds")
 
 ## ---- elec-gaps
 elec_ts <- read_rds("data/elec_ts.rds")
+# 13735 customers in elec_ts
 gap_df <- has_gaps(elec_ts)
 
 nogap <- gap_df %>% filter(.gaps==FALSE)
-
 
 elec_nogap <- elec_ts %>% 
   filter(customer_id %in% nogap$customer_id)
 
 readr::write_rds(elec_nogap, "data/elec_nogap.rds")
+elec_nogap <- read_rds("data/elec_nogap.rds")
+#8685 customers in elec_nogap
+
 
 # sum(gap_df$.gaps) / NROW(gap_df)
 
 ## ---- count-gaps
 count_na_df <- elec_ts %>% 
   count_gaps()
-
+#5050 customers in count_na_df
 
 lumped_na_df <- count_na_df %>% 
   mutate(

### 2021-09-22 supplementary/distance-computation/real-data/04handpick-from356cust.Rmd
adding new analaysis

---
title: "Selecting similar behaving group of customers to check clustering methodology"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, echo = FALSE, message=FALSE, warning=FALSE)
library(readr)
library(tidyverse)
library(gravitas)
library(tsibble)
library(parallel)
library(magrittr)
library(GGally)
library(knitr)
library(patchwork)
library(caret)
library(plotly)
```

# Data

Consider customers for which moy, hod and wknd_wday are all harmonies from an initial sample of 600 customers which contained no missing values. Also consider customers for which the quantiles are of same value (to avoid problems in JS computation).

```{r}
elec_356_raw <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))

elec_356_raw
```

# Classical MDS from distance metrics

```{r read-data}
# script to assimilate results from 02_algo1_moy_356cust_robust.R, 02_algo1_hod_356cust_robust.R and 02_algo1_wkndwday_356cust_robust.R
# write code after running the results for each of the above mentioned code

library(tidyverse)
library(gravitas)
library(tsibble)
library(ggpubr)

quantile_prob_graph <- c(0.25, 0.5, 0.75)

wkndwday <- read_rds("../../../data/dist_gran_wkndwday_356cust_robust.rds") %>% broom::tidy()

moy <- read_rds("../../../data/dist_gran_moy_356cust.rds") %>% broom::tidy()

hod <- read_rds("../../../data/dist_gran_hod_356cust_robust.rds") %>% broom::tidy()

distance <- wkndwday %>% 
  left_join(moy, by = c("item1", "item2")) %>% 
  left_join(hod, by = c("item1", "item2")) %>% 
  rename("wkndwday" ="distance.x",
         "moy" = "distance.y",
         "hod" = "distance") %>%
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2))) 
# for hod MDS scale

total_distance <- distance %>% 
  mutate(total = hod)

total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
                               names_from = item2,
                               values_from = total)

total_distance_wide$`8143599` <- NA

total_distance_wide <- bind_rows(total_distance_wide, c(item1=11590442))

mds_data <- total_distance_wide[-1]  %>% mutate_all(~replace(., is.na(.), 0))

mds <- mds_data %>% t() %>% as.dist() %>% cmdscale() %>% as_tibble()
colnames(mds) <- c("mds1.hod", "mds2.hod")

mds_hod <- bind_cols(mds, customer_id = mds_data %>% t() %>% rownames()) # since cmdscale is on transpose # because it only considers lower traingular matrix

# for moy MDS scale

total_distance <- distance %>% 
  mutate(total = moy)

total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
                               names_from = item2,
                               values_from = total)

total_distance_wide$`8143599` <- NA

total_distance_wide <- bind_rows(total_distance_wide, c(item1=11590442))

mds_data <- total_distance_wide[-1]  %>% mutate_all(~replace(., is.na(.), 0))

mds <- mds_data %>% t()%>% as.dist()  %>% cmdscale() %>% as_tibble()
colnames(mds) <- c("mds1.moy", "mds2.moy")
mds_moy <- bind_cols(mds, customer_id = mds_data %>% t() %>% rownames())

# for wkndwday MDS scale

total_distance <- distance %>% 
  mutate(total = wkndwday)

total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
                               names_from = item2,
                               values_from = total)
total_distance_wide$`8143599` <- NA

total_distance_wide <- bind_rows(total_distance_wide, c(item1=11590442))

mds_data <- total_distance_wide[-1]  %>% mutate_all(~replace(., is.na(.), 0))

mds <- mds_data %>% t() %>% as.dist() %>% cmdscale() %>% as_tibble()
colnames(mds) <- c("mds1.wkndwday", "mds2.wkndwday")
mds_wkndwday <- bind_cols(mds, customer_id = mds_data %>% t() %>% rownames())

# wkndwday+moy+

# for total MDS scale

# robust_scale <-  function(x){
#   (x - stats::median(x))/(stats::IQR(x))
# }
# 
# scaled_var <- apply(distance[-c(1, 2)], 2, robust_scale) %>% as_tibble()
# 
# 
# distance$wkndwday <- scaled_var$wkndwday
# distance$hod <- scaled_var$hod
# distance$moy <- scaled_var$moy
#

total_distance <- distance %>% 
  mutate(total = wkndwday + moy + hod)

total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
                               names_from = item2,
                               values_from = total)

total_distance_wide$`8143599` <- NA

total_distance_wide <- bind_rows(total_distance_wide, c(item1=11590442))

mds_data <- total_distance_wide[-1]  %>% mutate_all(~replace(., is.na(.), 0))

mds <- mds_data %>% t()%>% as.dist()  %>% cmdscale() %>% as_tibble()
colnames(mds) <- c("mds1.total", "mds2.total")
mds_total <- bind_cols(mds, customer_id = mds_data %>% t() %>% rownames())

## wpd

elec_356_raw <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))

elec_600_wpd <- read_rds(here::here("data/algo2-cust600-wpd-rawdata.rds"))

elec_pick <- elec_600_wpd %>% 
  filter(customer_id %in% elec_356_raw$customer_id)

elec_pick_wide <- elec_pick %>% pivot_wider(-c(1, 2), names_from = "x_variable", values_from = wpd)

# robust_scale <-  function(x){
#   (x - stats::median(x))/(stats::IQR(x))
# }
# 
# elec_pick_wide_scaled <- apply(elec_pick_wide[-1], 2, robust_scale)

mds <- elec_pick_wide[-1] %>% dist() %>% cmdscale() %>% as_tibble()

colnames(mds) <- c("mds1.wpd", "mds2.wpd")

mds_wpd <- bind_cols(mds, customer_id = elec_pick_wide$customer_id)%>% mutate(customer_id = as.character(customer_id))

mds_all <- mds_hod %>% 
  left_join(mds_moy, by = "customer_id") %>% 
  left_join(mds_wkndwday, by = "customer_id") %>% 
  left_join(mds_total, by = "customer_id") %>% 
  left_join(mds_wpd, by = "customer_id") %>% 
  select(customer_id, everything())

mds_all
```

# Visualize similarity between samples for different distance metrics hod+moy+wkndwday, hod, moy, wkndwday

```{r}

mds_tx <- highlight_key(mds_all, ~customer_id)

# initiate a plotly object
base <- plot_ly(mds_tx, color = I("black")) %>% 
  group_by(customer_id)

mds_hod <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.hod, y = ~ mds2.hod, text = ~customer_id)

mds_moy <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.moy, y = ~ mds2.moy, text = ~customer_id)

mds_wkndwday <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.wkndwday, y = ~ mds2.wkndwday, text = ~customer_id)

mds_total <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.total, y = ~ mds2.total, text = ~customer_id)

mds_wpd <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.wpd, y = ~ mds2.wpd, text = ~customer_id)

subplot(mds_total, mds_hod, mds_moy ,mds_wkndwday, nrows = 2, titleX = TRUE) %>%
  layout(showlegend = FALSE) %>%
  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect") 
```

# Visualize similarity between samples (all gran together)

for different distance metrics hod+moy+wkndwday, hod, moy, wkndwday

```{r wpd-total}
# subplot(mds_wpd, mds_total, nrows = 1, titleX = TRUE) %>%
#   layout(showlegend = FALSE) %>%
#   highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect") 
```

<!-- ```{r} -->

<!-- ggscatter(mds,  -->
<!--           x = "Dim.1", -->
<!--           y = "Dim.2",  -->
<!--           #label = elec_pick_wide$customer_id, -->
<!--           size = 1, -->
<!--           repel = TRUE) -->

<!-- fig <- plot_ly(mds, x = ~Dim.1, y = ~ Dim.2, text = total_distance_wide$item1) -->

<!-- ``` -->

## --- interactive-characterization-hod

```{r interactive-characterization}

sm_hod <- gravitas::create_gran(elec_356_raw, "hour_day")

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, ncol(sm_hod_list))]))
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list%>% 
     dplyr::filter(customer_id == customer_id[x]) %>% 
     select(-c(1, ncol(sm_hod_list))) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

ref_cat <- names(sm_hod_list)[-c(1, ncol(sm_hod_list))] %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
    mutate(customer_serial_id = row_number())

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>%
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "customer_serial_id") %>% 
  select(customer_serial_id, categories_serial_id, category, customer_id, everything())

data_heatmap <- sm_quantiles_ref %>% 
 left_join(elec_pick %>% distinct(customer_id)) %>% 
  mutate(categories_serial_id = as.factor(categories_serial_id)) %>% 
  mutate(customer_id = as.character(customer_id))

data_heatmap_new <- data_heatmap %>%
left_join(mds_all, by = "customer_id")

tx <- highlight_key(data_heatmap_new, ~customer_id)

 # initiate a plotly object
 base <- plot_ly(tx, color = I("black")) %>%
   group_by(customer_id)
#
 quantile_plot <- base %>%
   group_by(customer_id) %>%
   add_lines(x = ~ categories_serial_id, y = ~`50%`, color = ("black")) %>%
   add_ribbons(x = ~categories_serial_id,
               ymin = ~`25%`,
               ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>%
   layout(showlegend = FALSE)

mds_total <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.total, y = ~ mds2.total, text = ~customer_id)

mds_wpd <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.wpd, y = ~ mds2.wpd, text = ~customer_id)

subplot(mds_total, mds_wpd, quantile_plot, nrows = 2, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect")

```

```{r}

distance <- wkndwday %>% 
  left_join(moy, by = c("item1", "item2")) %>% 
  left_join(hod, by = c("item1", "item2")) %>% 
  rename("wkndwday" ="distance.x",
         "moy" = "distance.y",
         "hod" = "distance") %>% mutate(item1 = as.integer(as.character(item1)))

#10850478
data_pick_one <- total_distance %>% filter(item1 %in% 
                            c(8486921)) %>% group_by(item1) %>% arrange(total) %>% head(5) %>% 
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

#11013154
data_pick_two <- total_distance %>% filter(item1 %in% 
                            c(8181071)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

data_pick_three <- total_distance %>% filter(item1 %in% 
                            c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

# data_pick_three <- total_distance %>% filter(item1 %in% 
#                             c(11013154)) %>% 
#   mutate(item1 = as.integer(as.character(item1)),
#          item2 = as.integer(as.character(item2)))
```

```{r}

# data_pick_cust <- bind_rows(
# unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(),    unique(c(data_pick_two$item1,data_pick_two$item2)) %>% as_tibble(), c(11013154, 11045304, 11048034,11139841,11156027,11006388) %>% as_tibble(), .id = "design")
#

data_pick_cust <- bind_rows(
unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(), unique(c(data_pick_two$item1,data_pick_two$item2)) %>% as_tibble(), unique(c(data_pick_three$item1,data_pick_three$item2)) %>% as_tibble(), .id = "design")

data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))%>%
  filter(customer_id %in% data_pick_cust$value) %>% 
  gracsr::scale_gran( method = "robust", response = "general_supply_kwh")

sm_hod <- gravitas::create_gran(data_pick, "hour_day")

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)  %>% 
  left_join(data_pick_cust, by = c("customer_id" = "value"))

ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, ncol(sm_hod_list))]))
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list%>% 
     dplyr::filter(customer_id == customer_id[x]) %>% 
     select(-c(1, ncol(sm_hod_list))) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

ref_cat <- names(sm_hod_list)[-c(1, ncol(sm_hod_list))] %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
    mutate(customer_serial_id = row_number())

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>%
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "customer_serial_id") %>% 
  select(customer_serial_id, categories_serial_id, category, customer_id, everything())

data_heatmap <- sm_quantiles_ref %>% 
 left_join(data_pick %>% distinct(customer_id)) %>% 
  left_join(data_pick_cust, by = c("customer_id" = "value")) %>% 
  mutate(categories_serial_id = as.factor(categories_serial_id))

data_heatmap %>% 
  ggplot(aes(x = categories_serial_id)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=customer_id), fill = "lightblue") +
  geom_line(aes(y = `50%`, group=customer_id)) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + theme_bw()

```
<!-- # # client side linking -->
<!-- # ```{r} -->
<!-- # mds <- bind_cols(mds, customer_id = total_distance_wide$item1) -->
<!-- #  -->
<!-- # data_heatmap_new <- data_heatmap %>%  -->
<!-- #   left_join(mds, by = "customer_id") -->

<!-- tx <- highlight_key(data_heatmap_new, ~customer_id) -->

<!--  # initiate a plotly object -->
<!--  base <- plot_ly(tx, color = I("black")) %>% -->
<!--    group_by(customer_id) -->
<!-- # -->
<!--  quantile_plot <- base %>% -->
<!--    group_by(customer_id) %>% -->
<!--    add_lines(x = ~ categories_serial_id, y = ~`50%`, color = ("black")) %>% -->
<!--    add_ribbons(x = ~categories_serial_id, -->
<!--                ymin = ~`25%`, -->
<!--                ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>% -->
<!--    layout(showlegend = FALSE) -->

<!--  # fig <- plot_ly(base,color = I("blue")) %>% roup_by(customer_id) -->

<!--  fig_trace <- base %>% -->
<!--    group_by(customer_id) %>% -->
<!--    add_trace(x = ~Dim.1, y = ~ Dim.2, text = ~customer_id) -->

<!--  subplot(fig_trace, quantile_plot, widths = c(.6, .4), titleX  TRUE) %>% -->
<!--    layout(showlegend = FALSE) %>% -->
<!--    highlight(on = "plotly_selected", dynamic = TRUE, electize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect", persistent = TRUE, defaultValues  c(8271504, 9316697)) -->

<!--  %>% -->
<!--    highlight(color = toRGB("black")) -->

<!--              , color = list( -->
<!--    toRGB("red", "black") -->
<!--    ) -->
<!--  ) -->

### 2021-09-22 supplementary/distance-computation/real-data/algo1-cust600-gracsr.Rmd
adding new analaysis

---
title: "Algorithm using wpd values for clustering 600 households"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,echo=FALSE, message=FALSE, warning = FALSE)
```

```{r load-lib}
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
#library(ggridges)
library(tsibble)
library(gravitas)
library(ggpubr)
library(readr)
library(kableExtra)
library(distributional)
library(ggplot2)
library(sugrrants)
library(here)
library(ggplot2)
library(patchwork)
library(scales)
library(GGally)
library(viridis)
library(gracsr)
library(tsibble)
library(gravitas)
```

# distance using js distance between categories (robust scaling)

```{r js-based}
elec <- read_rds("../../../data/elec_nogap_2013_600.rds") %>% select(customer_id, reading_datetime, general_supply_kwh)
# 
# dist_mat_moy$customer_id[which(is.na(dist_mat_moy$general_supply_kwh))] %>% unique()
# c(8258490, 8662370, 10880968)

# throw away customers for which IQR is zero

elec_iqr0 <- elec %>% as_tibble() %>% group_by(customer_id) %>%
  summarise(q2 = stats::median(general_supply_kwh, na.rm = TRUE),
iqr = stats::IQR(general_supply_kwh, na.rm = TRUE)) %>% filter(iqr==0| q2==0)

dist_mat_moy <- elec %>%
  filter(!(customer_id %in% 
             elec_iqr0$customer_id)) %>% 
  scale_gran(method = "robust", response = "general_supply_kwh") %>%
  dist_gran(gran1 = "month_year", response= "general_supply_kwh")

dist_mat_hod <- elec %>%
  scale_gran(method = "robust", response = "general_supply_kwh") %>%
  dist_gran(gran1 = "hour_day", response= "general_supply_kwh")

dist_mat_wndwday <- elec %>%
  scale_gran(method = "robust", response = "general_supply_kwh") %>%
  dist_gran(gran1 = "wknd_wday", response= "general_supply_kwh")

dist_mat <- dist_mat_moy + dist_mat_hod + dist_mat_wndwday

groups = dist_mat %>% 
  clust_gran(kopt = 20)
```

### 2021-09-22 supplementary/distance-computation/real-data/algo2-cust600-assimilateHPC.Rmd
adding new analaysis

--- 2021-08-05 algo2-100.Rmd
@@
 
-  all_files = list.files(path = paste0("data/algo2/"), 
+  all_files = list.files(path = paste0("../../../data/algo2-600cust/"), 
                          pattern = ".rds")
@@
 
-  all_files_path <- paste0("data/algo2/",
+  all_files_path <- paste0("../../../data/algo2-600cust/",
                            all_files)
@@
 
-  write_rds(all_data, paste0("data/algo2-rawdata-wpd.rds"))
+  write_rds(all_data, paste0("../../../data/algo2-cust600-wpd-rawdata.rds"))
 ```

### 2021-09-22 supplementary/distance-computation/real-data/algo2-cust600-gracsr.Rmd
adding new analaysis

---
title: "Algorithm using wpd values for clustering 600 households"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,echo=FALSE, message=FALSE, warning = FALSE)
```

```{r load-lib}
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
#library(ggridges)
library(tsibble)
library(gravitas)
library(ggpubr)
library(readr)
library(kableExtra)
library(distributional)
library(ggplot2)
library(sugrrants)
library(here)
library(ggplot2)
library(patchwork)
library(scales)
library(GGally)
library(viridis)
library(gracsr)
library(tsibble)
library(gravitas)
```

```{r allplot, message=FALSE, warning=FALSE, echo = FALSE}

##----harmony-all
elec_harmony_all <- read_rds("../../../data/algo2-cust600-wpd-rawdata.rds")

cust_by_wpd <- elec_harmony_all %>% 
  select(customer_id, x_variable, wpd) %>% 
  pivot_wider(
              names_from = x_variable, 
              values_from = wpd)

# Outliers

library(OutliersO3)

data <- cust_by_wpd[2:4]

O3x <- O3prep(data, method="HDo", tols=0.01, boxplotLimits=c(3, 6, 10))

O3x1 <- O3plotT(O3x)
library(gridExtra)
grid.arrange(O3x1$gO3, O3x1$gpcp, ncol=1)

cust_by_wpd <- cust_by_wpd %>% mutate(Case= row_number()) %>% 
  filter(!(Case %in% O3x1$outsTable$Case))

# O3m <- O3prep(data, method=c("HDo", "PCS"))
# O3m1 <- O3plotM(O3m)
# grid.arrange(O3m1$gO3, O3m1$gpcp, ncol=1)

library(stray)
output <- find_HDoutliers(data, knnsearchtype = "brute", alpha = 0.1)
display_HDoutliers(data, out = output)

# make a distance matrix from here and perform the algorithm

mydist <- column_to_rownames(cust_by_wpd, var = "customer_id")%>%
    dist()

groups <- mydist%>% 
  gracsr::clust_gran(kopt=10) %>% 
  bind_cols(customer_id = cust_by_wpd$customer_id)

ngroup <- groups %>% group_by(group) %>% count()

cust_by_wpd_join_groups <- cust_by_wpd %>% left_join(groups)

cust_by_wpd_join_groups_long <- cust_by_wpd_join_groups %>%
  pivot_longer(-c(1,5, 6),names_to = "gran",
               values_to = "wpd") 
```

```{r}
cust_by_wpd_join_groups_long%>% 
  ggplot() +
  geom_boxplot(aes(x=gran, y = wpd)) +
  facet_wrap(~group) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

elec <- read_rds("../../../data/elec_nogap_2013_600.rds") %>% select(customer_id, reading_datetime, general_supply_kwh)

characterise_plot_data <- elec %>% 
  create_gran("month_year") %>% 
  create_gran("hour_day") %>% 
  create_gran("wknd_wday") %>% 
  as_tibble() %>% 
 select(customer_id,
        month_year,
        hour_day,
        wknd_wday, general_supply_kwh) %>% 
left_join(groups, by = "customer_id")

# characterise_plot_data_long <- characterise_plot_data%>% 
#   pivot_longer(c(-1, -5, -6), 
#                names_to = "gran", 
#                values_to = "categories")

moy <- characterise_plot_data %>%
  ggplot(aes(x=month_year, y = general_supply_kwh)) +
  geom_boxplot() +
  facet_wrap(~group, ncol = 1, scales = "free_y")

hod <- characterise_plot_data %>%
  ggplot(aes(x=hour_day, y = general_supply_kwh)) +
  geom_boxplot() +
  facet_wrap(~group, ncol = 1, scales = "free_y")

wndwday <- characterise_plot_data %>%
  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
  geom_boxplot() +
  facet_wrap(~group, ncol = 1, scales = "free_y")

ggpairs_nocolor <- ggscatmat(cust_by_wpd_join_groups, columns=2:4) +
  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")

cust_by_wpd_join_groups$group = as.factor(cust_by_wpd_join_groups$group)
```

```{r parcoord}
parcoord <- GGally::ggparcoord(cust_by_wpd_join_groups ,
                   columns = 2:4,
                   groupColumn = "group",
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.2,
                   order = "anyClass",
                   scale = "globalminmax"
) + ggplot2::theme_bw() +
  #scale_color_brewer(palette = "Dark2") + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")
parcoord

# PCA based clustering

pc <- prcomp(cust_by_wpd_join_groups[c(-1,-5, -6)], center = F, scale = F)
summary(pc)
# First two principal componentsafc
comp <- data.frame(pc$x[,1:2])

data_pcp_all <- comp%>% as_tibble() %>% 
  bind_cols(cust_by_wpd_join_groups)

data_pcp_all %>% ggplot(aes(x=PC1, y = PC2)) + geom_point(aes(color = group)) +
  scale_color_brewer(palette = "Dark2") + theme_bw() + theme(legend.position = "bottom")
# 
# 
# O3y <- O3prep(data, method=c("HDo", "PCS", "BAC", "adjOut", "DDC", "MCD"))
# O3y1 <- O3plotM(O3y)
# cx <- data.frame(outlier_method=names(O3y1$nOut), number_of_outliers=O3y1$nOut)
# knitr::kable(cx, row.names=FALSE)
# 
# grid.arrange(O3y1$gO3, O3y1$gpcp, ncol=1)
```

```{r}
quantile_prob_val = c(0.25, 0.5, 0.75)

sm_hod_dow <- elec %>% 
  left_join(groups)

  sm_hod_dow %>% 
  filter(group==1) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 1")

  sm_hod_dow %>% 
  filter(group==2) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 2")

  sm_hod_dow %>% 
  filter(group==3) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 3")

  sm_hod_dow %>% 
  filter(group==4) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 4")

  sm_hod_dow %>% 
  filter(group==5) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 5")

  sm_hod_dow %>% 
  filter(group==6) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 6")

  sm_hod_dow %>% 
  filter(group==7) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 7")

  sm_hod_dow %>% 
  filter(group==8) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 8")

  sm_hod_dow %>% 
  filter(group==9) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 9")

  sm_hod_dow %>% 
  filter(group==10) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 10")
```

### 2021-09-22 supplementary/distance-computation/simulation-data/D2_I5_C1.Rmd
adding new analaysis

---
title: "2 designs and 5 iterations and categorty changes across 1 granularity (g3 and g1 categories const) "
author: "Sayani Gupta"
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
    toc: false
    cache: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,echo=FALSE, message=FALSE, warning = FALSE)
library(patchwork)
library(distributional)
library(ggplot2)
library(tidyverse)
library(knitr)
library(readr)
library(GGally)
library(tsibble)
library(tidyverse)
library(gracsr)
library(gravitas)

```

# _Simulation design_

```{r change-var, echo=TRUE}
t = 300
n <- seq(0, t, 1)
g1 <- n %%2
g2 <- n %%3
g3 <- n %%5
mu1= c(0, 0) # mean of cat of g1
mu2 = c(7, 5, 1)# mean of cat of g2
mu3 = c(0, 0, 0, 0, 0) # mean of cat of g3
```

Three circular granularities g1, g2 and g3 are considered with levels 2, 3 and 5 respectively. Many time series with `r t` observations are created using the five designs below, each of which is iterated five times. We anticipate to have 3 clusters, each with five time series conforming to the same design, once we execute the clustering.

# _Distance computation_

1) wpd for three granularities computed and then distance between designs are computed by computing the euclidean distances between them - gran as variables, 
designs across rows and cell values as wpd.

2) Robust scaling done for each customer and then JSD is computed between same categories for a granularity.

# _Granularities design_
```{r}
design = c("design-1", "design-2")
g1 = c("no", "no")
g2 = c("no", "yes")
g3 = c("no", "no")
table <- tibble(design, g1, g2, g3) %>% kable()
table

```

# _Simulate data_

```{r generate-design}
generate_design <- function(t, mu1, mu2, mu3){
t <- seq(0, t, 1)
g1 <- t %%2
g2 <- t %%3
g3 <- t %%5

# null design
g1_dnull <- rep( rep(0, each = length(unique(g1))), length.out= length(t))
g2_dnull <- rep( rep(0, each = length(unique(g2))), length.out= length(t))
g3_dnull <- rep( rep(0, each = length(unique(g3))), length.out= length(t))

# mean changing across categories in varying ways

g1_dvary <- rep(mu1, length.out= length(t))
g2_dvary <- rep(mu2, length.out= length(t))
g3_dvary <- rep(mu3, length.out= length(t))

design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
design2 = distributional::dist_normal(g1_dnull + g2_dvary + g3_dnull)
#design3 = distributional::dist_normal(g1_dnull + g2_dvary + g3_dnull)
#design4 = distributional::dist_normal(g1_dvary + g2_dvary + g3_dnull)
#design5 = distributional::dist_normal(g1_dvary + g2_dvary + g3_dvary)

data_bind <- tibble::tibble(
index = t,
g1 = g1,
g2 = g2,
g3 = g3,
design1 = distributional::generate(design1, times = 1) %>% unlist(),
design2 = distributional::generate(design2, times = 1) %>% unlist(),
#design3 = distributional::generate(design3, times = 1) %>% unlist(),
#design4 = distributional::generate(design4, times = 1) %>% unlist(),
#design5 = distributional::generate(design5, times = 1) %>% unlist()
) %>% 
  pivot_longer(-c(1, 2, 3, 4), names_to = "design", values_to = "sim_data")

data_bind
}

data_bind <- generate_design(t, mu1, mu2, mu3)

data_bind

```

# _Plot raw data_

```{r plot-linear , out.width="100%"}

# plot_linear_data <- function(data){
# ggplot(data,
#              aes(x = index, y = sim_data)) + 
#   geom_line() +
#   xlab("index")+
#   theme_bw() 
# }

ggplot(data_bind,
             aes(x = index, y = sim_data)) + 
  geom_line() +
  xlab("index")+
  theme_bw() +
  facet_wrap(~design, scales = "free_y",ncol =1)

```

# _Plot distribution across granularities_

```{r plot-gran, out.width="100%"}

p2 <- ggplot(data_bind,
             aes(x = as.factor(g1), y = sim_data)) + 
  geom_boxplot(alpha =0.5) + xlab("g1") + 
  theme_bw() +
  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
    fun = median,
    geom = 'line',
    aes(group = 1), size = 1, color = "blue")

p3 <- ggplot(data_bind, aes(x = as.factor(g2), y = sim_data)) + geom_boxplot(alpha =0.5) + xlab("g2") + theme_bw() +
  facet_wrap(~design, scales = "free_y",ncol = 1)+ stat_summary(
    fun = median,
    geom = 'line',
    aes(group = 1), size = 1, color = "blue")

p4 <- ggplot(data_bind, aes(x = as.factor(g3), y = sim_data)) + geom_boxplot(alpha =0.5) +
  xlab("g3") + theme_bw()+
  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
    fun = median,
    geom = 'line',
    aes(group = 1), size = 1, color = "blue")

(p2 + p3 + p4) * theme(
    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))

# 
# 
# plot_cyclic_data( data_bind %>% filter(design=="design2"))
# plot_data( data_bind %>% filter(design=="design3"))
# plot_data( data_bind %>% filter(design=="design4"))
# plot_data( data_bind %>% filter(design=="design5"))
```

```{r,eval=FALSE}

data_bind_all <- data_bind %>% pivot_longer(-c(index, design, sim_data), names_to = "gran", values_to = "categories")

ggplot(data_bind_all, aes(x = as.character(categories), y = sim_data)) + geom_boxplot() + facet_grid(design~gran, scales = "free_y")

```

# _Iterate designs_

```{r}
seed_len = seq(12345, 12349, 1)
bind_data_iter <- map(seq_len(length(seed_len)), function(x){
set.seed(seed_len[x])
data_bind <- generate_design(t, mu1, mu2, mu3)
}) %>% bind_rows(.id = "seed_id") %>% 
  mutate(customer_id = paste(design,seed_id, sep ="-"))

bind_data_iter_tsibble <- bind_data_iter %>% 
  tsibble::as_tsibble(index = index, key = customer_id)

bind_data_iter_tsibble
```

# _Cluster designs: distance using wpd_

```{r wpd}
harmony_tbl <- tibble(facet_variable = NA, x_variable = c("g1", "g2", "g3"),
                      x_levels = c(2, 3, 5), facet_levels = NA)

dist_mat <- bind_data_iter_tsibble %>%
  #scale_gran(method = "robust", response = "sim_data") %>%
  dist_wpd(harmony_tbl, response = "sim_data", nperm=100)

groups = dist_mat%>% clust_gran(kopt = 2 )
pred_group = paste("design",groups$group,sep = "") %>% as.factor()
actual_group = as.factor(bind_data_iter_tsibble %>%as_tibble %>% select(customer_id, design) %>% distinct() %>% pull(design))

xtab <- caret::confusionMatrix(pred_group, actual_group)
xtab$table
```

# _Cluster designs: distance using js distance between categories_

```{r js}
dist_mat_g1 <- bind_data_iter_tsibble %>%
  scale_gran(method = "robust", response = "sim_data") %>%
  dist_gran(gran1 = "g1", response= "sim_data")

dist_mat_g2 <- bind_data_iter_tsibble %>%
  scale_gran(method = "robust", response = "sim_data") %>%
  dist_gran(gran1 = "g2", response= "sim_data")

dist_mat_g3 <- bind_data_iter_tsibble %>%
  scale_gran(method = "robust", response = "sim_data") %>%
  dist_gran(gran1 = "g3", response= "sim_data")

dist_mat <- dist_mat_g1 + dist_mat_g2 + dist_mat_g3

groups = dist_mat %>% 
  clust_gran(kopt = 2 )

pred_group = paste("design",groups$group,sep = "") %>% as.factor()
actual_group = as.factor(bind_data_iter_tsibble %>%as_tibble %>% select(customer_id, design) %>% distinct() %>% pull(design))

xtab <- caret::confusionMatrix(pred_group, actual_group)
xtab$table
```

# _Cluster designs: distance using js distance between categories (nqt)_

```{r js-nqt}
dist_mat_g1 <- bind_data_iter_tsibble %>%
  scale_gran(method = "nqt", response = "sim_data") %>%
  dist_gran(gran1 = "g1", response= "sim_data")

dist_mat_g2 <- bind_data_iter_tsibble %>%
  scale_gran(method = "nqt", response = "sim_data") %>%
  dist_gran(gran1 = "g2", response= "sim_data")

dist_mat_g3 <- bind_data_iter_tsibble %>%
  scale_gran(method = "nqt", response = "sim_data") %>%
  dist_gran(gran1 = "g3", response= "sim_data")

dist_mat <- dist_mat_g1 + dist_mat_g2 + dist_mat_g3

groups = dist_mat %>% 
  clust_gran(kopt = 2 )

pred_group = paste("design",groups$group,sep = "") %>% as.factor()
actual_group = as.factor(bind_data_iter_tsibble %>%as_tibble %>% select(customer_id, design) %>% distinct() %>% pull(design))

xtab <- caret::confusionMatrix(pred_group, actual_group)
xtab$table
```

### 2021-09-22 supplementary/distance-computation/simulation-data/D4_I5_C2.Rmd
adding new analaysis

--- 2021-09-22 supplementary/distance-computation/simulation-data/D2_I5_C1.Rmd
@@
 ---
-title: "2 designs and 5 iterations and categorty changes across 1 granularity (g3 and g1 categories const) "
+title: "4 designs and 5 iterations and categorty changes across 2 granularities (g3 categories const) "
 author: "Sayani Gupta"
@@
 g3 <- n %%5
-mu1= c(0, 0) # mean of cat of g1
+mu1= c(0, 2) # mean of cat of g1
 mu2 = c(7, 5, 1)# mean of cat of g2
@@
 ```{r}
-design = c("design-1", "design-2")
-g1 = c("no", "no")
-g2 = c("no", "yes")
-g3 = c("no", "no")
+design = c("design-1", "design-2", "design-3", "design4")
+g1 = c("no", "no", "yes", "yes")
+g2 = c("no", "yes", "no", "yes")
+g3 = c("no", "no", "no", "no")
 table <- tibble(design, g1, g2, g3) %>% kable()
@@
 design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
-design2 = distributional::dist_normal(g1_dnull + g2_dvary + g3_dnull)
-#design3 = distributional::dist_normal(g1_dnull + g2_dvary + g3_dnull)
-#design4 = distributional::dist_normal(g1_dvary + g2_dvary + g3_dnull)
+design2 = distributional::dist_normal(g1_dvary + g2_dnull + g3_dnull)
+design3 = distributional::dist_normal(g1_dnull + g2_dvary + g3_dnull)
+design4 = distributional::dist_normal(g1_dvary + g2_dvary + g3_dnull)
 #design5 = distributional::dist_normal(g1_dvary + g2_dvary + g3_dvary)
@@
 design2 = distributional::generate(design2, times = 1) %>% unlist(),
-#design3 = distributional::generate(design3, times = 1) %>% unlist(),
-#design4 = distributional::generate(design4, times = 1) %>% unlist(),
+design3 = distributional::generate(design3, times = 1) %>% unlist(),
+design4 = distributional::generate(design4, times = 1) %>% unlist(),
 #design5 = distributional::generate(design5, times = 1) %>% unlist()
@@
 
-groups = dist_mat%>% clust_gran(kopt = 2 )
+groups = dist_mat%>% clust_gran(kopt = 4)
 pred_group = paste("design",groups$group,sep = "") %>% as.factor()
@@
 groups = dist_mat %>% 
-  clust_gran(kopt = 2 )
+  clust_gran(kopt = 4)
 

### 2021-09-22 supplementary/distance-computation/simulation-data/D5_I5_C3.Rmd
adding new analaysis

--- 2021-09-22 supplementary/distance-computation/simulation-data/D2_I5_C1.Rmd
@@
 ---
-title: "2 designs and 5 iterations and categorty changes across 1 granularity (g3 and g1 categories const) "
+title: "5 designs and 5 iterations and 3 granularities all of whose categories change"
 author: "Sayani Gupta"
+date: "15/09/2021"
 output:
@@
 t = 300
-n <- seq(0, t, 1)
+n <- seq(0, n, 1)
 g1 <- n %%2
@@
 g3 <- n %%5
-mu1= c(0, 0) # mean of cat of g1
-mu2 = c(7, 5, 1)# mean of cat of g2
-mu3 = c(0, 0, 0, 0, 0) # mean of cat of g3
+mu1= c(0, 2)
+mu2 = c(7, 5, 1)
+mu3 = c(1, 2, 5, 3, 1)
 ```
@@
 ```{r}
-design = c("design-1", "design-2")
-g1 = c("no", "no")
-g2 = c("no", "yes")
-g3 = c("no", "no")
+design = c("design-1", "design-2", "design-3", "design4", "design5")
+g1 = c("no", "yes", "no", "no", "yes")
+g2 = c("no", "no", "yes", "no", "yes")
+g3 = c("no", "no", "no", "yes", "yes")
 table <- tibble(design, g1, g2, g3) %>% kable()
@@
 
-```{r generate-design}
+```{r}
 generate_design <- function(t, mu1, mu2, mu3){
+
 t <- seq(0, t, 1)
@@
 design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
-design2 = distributional::dist_normal(g1_dnull + g2_dvary + g3_dnull)
-#design3 = distributional::dist_normal(g1_dnull + g2_dvary + g3_dnull)
-#design4 = distributional::dist_normal(g1_dvary + g2_dvary + g3_dnull)
-#design5 = distributional::dist_normal(g1_dvary + g2_dvary + g3_dvary)
+design2 = distributional::dist_normal(g1_dvary + g2_dnull + g3_dnull)
+design3 = distributional::dist_normal(g1_dnull + g2_dvary + g3_dnull)
+design4 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dvary)
+design5 = distributional::dist_normal(g1_dvary + g2_dvary + g3_dvary)
 
@@
 design2 = distributional::generate(design2, times = 1) %>% unlist(),
-#design3 = distributional::generate(design3, times = 1) %>% unlist(),
-#design4 = distributional::generate(design4, times = 1) %>% unlist(),
-#design5 = distributional::generate(design5, times = 1) %>% unlist()
+design3 = distributional::generate(design3, times = 1) %>% unlist(),
+design4 = distributional::generate(design4, times = 1) %>% unlist(),
+design5 = distributional::generate(design5, times = 1) %>% unlist()
 ) %>% 
@@
 }
+
+t = 300
+mu1= c(0, 2)
+mu2 = c(7, 5, 1)
+mu3 = c(1, 2, 5, 3, 1)
 
@@
 
-groups = dist_mat%>% clust_gran(kopt = 2 )
+groups = dist_mat%>% clust_gran(kopt = 5)
 pred_group = paste("design",groups$group,sep = "") %>% as.factor()
@@
 
-# _Cluster designs: distance using js distance between categories_
+# _Cluster designs: distance using js distance between categories (robust scaling)_
 
-```{r js}
+```{r js-robust}
 dist_mat_g1 <- bind_data_iter_tsibble %>%
@@
 groups = dist_mat %>% 
-  clust_gran(kopt = 2 )
+  clust_gran(kopt = 5)
 

### 2021-09-22 supplementary/distance-computation/simulation-data/sim-robust-4iter.Rmd
adding new analaysis

--- 2021-08-05 supplementary/simulation-alogorithm2/sim_setup_algo2_50iter.Rmd
@@
 # _Algorithm_
-Compute wpd for each granularity.The distance matrix is then computed with 20 time series as cases and three granularities as variables.
-Manhattan distances are computed and then hierarchical clustering is applied on them.
+
+Robust scaling done for each customer and then JSD is computed between same categories for a granularity.
 
@@
 g3 = c("no", "yes" , "no", "yes", "yes")
-#table <- tibble(design, g1, g2, g3) %>% kable()
+table <- tibble(design, g1, g2, g3) %>% kable()
 #table
@@
 
-<!-- Cross-check -->
-<!-- data = bind_data_iter %>% filter(design == "design-2", seed_id == 5) -->
-<!-- p2 <- ggplot(data, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw() + facet_wrap(~design, ncol = 1) -->
-<!-- p3 <- ggplot(data, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw() + facet_wrap(~design, ncol = 1) -->
-<!-- p4 <- ggplot(data, aes(x = as.factor(g3), y = value)) + geom_boxplot() + -->
-<!-- xlab("g3") + theme_bw() + facet_wrap(~design, ncol = 1) -->
-<!-- (p2 + p3 + p4) -->
-
-<!-- ```{r} -->
-<!-- p2 <- ggplot(data, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw() -->
-<!-- p3 <- ggplot(data, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw() -->
-<!-- p4 <- ggplot(data, aes(x = as.factor(g3), y = value)) + geom_boxplot() + -->
-<!-- xlab("g3") + theme_bw()  -->
-<!-- (p2 + p3 + p4) -->
-<!-- ``` -->
-
 ```{r}

### 2021-09-22 supplementary/effect_transformation.Rmd
adding new analaysis

@@
 v2 <- suppressWarnings(robust_scale_data(sm, "hour_day")) %>% 
-    dplyr::mutate(kwh_nqt = stats::qqnorm(general_supply_kwh, plot.it=FALSE)$x) %>% 
-  mutate(category = as.numeric(category),
-         kwh_robust = )
+    dplyr::rename( "kwh_robust" = "scaled_response") %>% 
+    dplyr::mutate(kwh_nqt = stats::qqnorm(general_supply_kwh, plot.it=FALSE)$x)
+
+C
 
@@
 dplyr::filter(customer_id %in% c("10018250")) %>%
-  pivot_longer(c("general_supply_kwh", "scaled_response", "kwh_nqt"),
+  pivot_longer(c("general_supply_kwh", "kwh_robust", "kwh_nqt"),
                names_to = "kwh_type",

### 2021-09-22 supplementary/distance-computation/real-data/02_algo1_hod_356cust_robust.R

@@ -1,0 +1,17 @@
+# script to run algo1 for distance on hod for the cleaner 356 customers
+#
+library(tidyverse)
+library(gravitas)
+library(tsibble)
+library(gracsr)
+
+elec <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))%>% 
+                   dplyr::select(customer_id, reading_datetime, general_supply_kwh)
+
+response = "general_supply_kwh"
+gran1 = "hour_day"
+
+#elec <- elec %>% filter(customer_id %in% c(8143599, 8147121, 8159803, 10542667))
+
+distance_data_robust <- elec %>% scale_gran(method = "robust") %>% dist_gran(gran1)
+write_rds(distance_data_robust, here::here("data/dist_gran_hod_356cust_robust.rds"))

### 2021-09-22 supplementary/distance-computation/real-data/03.5_assimilate_wpd_356cust_robust.R

@@ -1,0 +1,174 @@
+# script to assimilate results from 02_algo1_moy_356cust_robust.R, 02_algo1_hod_356cust_robust.R and 02_algo1_wkndwday_356cust_robust.R
+# write code after running the results for each of the above mentioned code
+
+library(tidyverse)
+library(gravitas)
+library(tsibble)
+library(ggpubr)
+library(readr)
+library(plotly)
+
+#----iteration-1
+
+elec_356_raw <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))
+
+elec_600_wpd <- read_rds(here::here("data/algo2-cust600-wpd-rawdata.rds"))
+
+elec_pick <- elec_600_wpd %>% 
+  filter(customer_id %in% elec_356_raw$customer_id)
+
+elec_pick_wide <- elec_pick %>% pivot_wider(-c(1, 2), names_from = "x_variable", values_from = wpd)
+
+robust_scale <-  function(x){
+  (x - stats::median(x))/(stats::IQR(x))
+}
+
+elec_pick_wide_scaled <- apply(elec_pick_wide[-1], 2, robust_scale)
+
+
+mds <- elec_pick_wide_scaled %>% dist() %>% cmdscale() %>% as_tibble()
+
+colnames(mds) <- c("Dim.1", "Dim.2")
+
+ggscatter(mds, 
+          x = "Dim.1",
+          y = "Dim.2", 
+          #label = elec_pick_wide$customer_id,
+          size = 1,
+          repel = TRUE)
+
+
+fig <- plot_ly(mds, x = ~Dim.1, y = ~ Dim.2, text = elec_pick_wide$customer_id)
+
+# Outliers list after loooking at the plot
+
+# (8147121, 8568565, 8283422, 9393476, 10542667)
+
+
+#----Iteration (2 after removing outliers)
+
+elec_raw <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>% 
+  filter(!(customer_id %in% c(8147121, 8568565, 8283422, 9393476, 10542667)
+))
+
+elec_600_wpd <- read_rds(here::here("data/algo2-cust600-wpd-rawdata.rds"))
+
+elec_pick <- elec_600_wpd %>% 
+  filter(customer_id %in% elec_raw$customer_id)
+
+elec_pick_wide <- elec_pick %>% pivot_wider(-c(1, 2), names_from = "x_variable", values_from = wpd)
+
+robust_scale <-  function(x){
[…]

### 2021-09-22 supplementary/distance-computation/real-data/job-algo2-600cust.R

@@ -1,0 +1,78 @@
+library(dplyr)
+library(gravitas)
+library(tsibble)
+#library(ggpubr)
+library(readr)
+#library(kableExtra)
+#library(sugrrants)
+#library(here)
+#library(patchwork)
+#library(scales)
+#library(GGally)
+#library(viridis)
+
+
+ncust = 100 # number of sampled customers
+nperm = 200 # number of permutations for normalization
+nsamp = 200 # number of samples for threshold
+
+#take customers who do not have gaps in their data
+# elec_nogap <- read_rds("data/elec_nogap.rds")
+# 
+# elec_nogap_2013 <- elec_nogap %>%
+#   filter(year(reading_datetime)==2013)
+#   
+# write_rds(elec_nogap_2013, "data/elec_nogap_2013.rds")
+
+# elec_nogap_2013 <- read_rds("data/elec_nogap_2013.rds")
+
+#just take 100 customers from them
+# set.seed(12345)
+# 
+# sm_100 <- elec_nogap_2013 %>%
+#   as_tibble() %>%
+#   distinct(customer_id) %>%
+#   slice_sample(n = ncust)
+# 
+# # take data for those customers only for 2013
+# elec <- elec_nogap_2013 %>%
+#   filter(customer_id %in% sm_100$customer_id) %>%
+#   ungroup()
+# 
+# write_rds(elec, "data/elec_nogap_2013_100.rds")
+
+elec <- read_rds("data/elec_nogap_2013_600.rds")
+cust_id <- unique(elec$customer_id)
+scen <- as.numeric(commandArgs()[[6]])
+#scen <- 2
+custj <- cust_id[scen]
+# consider harmonies which are generally significant for electricity data
+
+#harmonies <- read_rds("../paper-hakear/paper/data/harmonies.rds")
+
+# harmonies <- harmonies %>% 
+#   mutate(comb = paste(facet_variable, 
+#                              x_variable, 
+#                              sep = "-")) %>% 
+#   filter(comb %in% c("hour_day-wknd_wday", "day_month-hour_day", "wknd_wday-hour_day", "hour_day-day_week", "day_week-hour_day")) %>% 
+#   select(-comb)
+# 
+# write_rds(harmonies, "data/harmonies.rds")
+
+harmonies <- read_rds("data/harmonies-onegran.rds")%>%
+  filter(x_variable %in% c("month_year", "hour_day", "wknd_wday"))
+
+# take customers who do not have gaps in their data
+  
+  data_id <-  elec %>% 
[…]

### 2021-09-22 supplementary/distance-computation/real-data/job-algo2-600cust.sh

@@ -1,0 +1,13 @@
+#! /bin/bash
+#SBATCH --job-name=algo2cust600
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=72:00:00
+#SBATCH --array=1-600
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.5
+Rscript job-algo2-600cust.R $SLURM_ARRAY_TASK_ID 
+

### 2021-09-22 supplementary/distance-computation/select_clean_cust_from600cust.R

@@ -1,0 +1,89 @@
+library(tidyverse)
+library(gravitas)
+library(tsibble)
+elec <- read_rds("../paper-gracsr/data/elec_nogap_2013_600.rds") %>% select(customer_id, reading_datetime, general_supply_kwh)
+
+elec_iqr0 <- elec %>% as_tibble() %>% group_by(customer_id) %>%
+  summarise(q2 = stats::median(general_supply_kwh, na.rm = TRUE),
+            iqr = stats::IQR(general_supply_kwh, na.rm = TRUE),
+            quantile = stats::quantile(general_supply_kwh, na.rm = TRUE, prob = seq(0.1, 0.9, 0.1))) %>% filter(iqr==0| q2==0)
+
+
+elec %>% as_tibble() %>%
+  filter((customer_id == 8164613)) %>%
+  summarise(q2 = stats::median(general_supply_kwh, na.rm = TRUE),
+            iqr = stats::IQR(general_supply_kwh, na.rm = TRUE))
+
+
+elec_sub <- elec %>%
+  filter(!(customer_id %in%
+             elec_iqr0$customer_id))
+
+# Iteration 1
+
+.data = elec_sub
+
+gran1 = "month_year"
+response = "general_supply_kwh"
+
+key = "customer_id"
+gran2 =  NULL
+
+
+if(is.null(gran2)){
+  sm_gran <- .data %>%
+    create_gran(gran1) %>%
+    as_tibble() %>%
+    select(key,
+           response,
+           {{gran1}})
+  
+}
+
+if(!is.null(gran2)){
+  sm_gran <- .data %>%
+    create_gran(gran1) %>%
+    create_gran(gran2) %>%
+    as_tibble() %>%
+    select(key,
+           response,
+           {{gran1}},
+           {{gran2}})
+}
+
+
+data <- unite(sm_gran, category, -c(1, 2), sep = "-")
+
+# category reference
+uni_cat <- unique(data$category)
+category_ref <-tibble(category_id = seq(uni_cat),
+                      category = uni_cat)
+
+
+
+sm_list <- data %>%
+  select(key, category, response) %>%
+  pivot_wider(names_from = category,
+              values_from = response, values_fn = list)
+
+
+cust = sm_list$customer_id %>% unique
+nsm = sm_list[-1] %>% ncol()
+
+sm_is_null <- map(seq_len(length(cust)), function(x){
[…]

### 2021-09-29 main.Rmd
methodology first para re-write

@@
 ---
-title: Clustering of residential Australian smart meter customers
+title:  Clustering based on probability distributions with application on residential customers
 
@@
 
+_Themes_
+
+- Dimension reduction: If each $P_{i, j, k}$ be considered to be a point in the space, key $i$ would have $mp$ dimensions as opposed to $n_i$ dimensions in case of considering raw data.
+Hence for a large number of observations ($n_i>>mp$), this approach benefits by transitioning to a lower dimension.
+
+ - Avoid loss of information due to aggregation: This approach ensures key characteristic information of the data is not lost due to averaging or aggregation measures in an attempt to transition to a lower dimension.
+Hence, this approach could be thought to somehow balance the drawback of considering raw data or aggregated data.
+
+- Robustness to outliers: This approach could be adapted to be robust to outliers and extreme behaviors by trimming the tails of the probability distributions.
+
+- Non-synchronized observed time periods: Considering probability distribution would imply the clustering process can handle keys that are observed over periods of time that are overlapping but don't necessarily coincide.
+
+- Similar periodic behavior: Since cyclic granularities are considered instead of linear granularities, clustering would group keys that have similar behavior across these cyclic granularities.
+This implies they will be grouped according to their periodic behavior and not on the linear stretch of time over which they are observed.
+
+_Common load clustering techniques of smart meter data_
+
+The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
+None of the 34 selected papers which focus on clustering consumption are based on Australian smart meter data.
+The clustering is frequently applied directly to the raw data without scrutinizing for auto correlation and periodicity.
+The algorithm most ubiquitously employed is K-Means.
+But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters.
+Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results.
+@Tureczek2018-ha has shown that a transformation of data to incorporate autocorrelation before K-Means clustering can improve performance and enable K-Means to deliver smaller clusters with less within-cluster variance.
+However, it does not explain the cluster composition by combining it with external data.
+Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf).
+Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.
+As the methods are often situation specific,
+it makes sense to compare them on the performance rather than any standard performance metric.
+A type of clustering based on information theory such as Shannon or Renyi entropy and their variants are addressed in , which differs from typical methods adopted for electricity consumer classification, based on the Euclidean distance notion.
+@Motlagh2019-yj presents strategy to address the problems on patchy, and unequal time-series common in residential data sets by converting load time series into map models.
+Most time-series clustering models are limited to handling time domain with same start and end date and time.
+Most of the solutions to handle this like longest common subsequence, dynamic time warping are prone to computational limit with increased length of the series.
+
 The following contributions are made through the following chapter:
@@
 -->
+
+### Data source
 
@@
 
-## Missing Data
+### Raw data
+
+### Missing Data
 
@@
 
-## Raw data
+13735 customers in elec_ts
+8685 customers in elec_nogap
+5050 customers in count_na_df
+
+Then is the graph of missing observations even interesting.
+You can show two graphs, one to show that missingness do not have a pattern
+another to show even if no missing, they start and end at different times. (A sample of 50 customers).
+
+## LGA and weather data
 
@@
 
-```{r load-elec, message = FALSE, warning=FALSE}
-
-```
-
 ```{r miss-data}
@@
 
-```{r vismiss, eval = FALSE}
-## transform all columns
-customer_data <- customer_data %>% mutate_each(funs(empty_as_na))
+#NEW
 
-vis_miss(customer_data, warn_large_data = FALSE) + 
-  theme(axis.text.x = element_text(angle = 90, size = 5))
-```
+The raw smart meter data is indexed by time for each key variable and includes values of different variables of interest (measurement variables) at each time point.
+However, this series could be represented in several possible ways.
+One sequence could represent distribution of hourly consumption across a single day, while another could track consumption across days of a week or months of a year.
+Depending on the cyclic granularity considered, each sequence could be thought of as a different data generating processes.
 
-```{r miss-var, eval = FALSE}
-gg_miss_var(customer_data) + theme_bw() +
-  theme(axis.text.y = element_text(size = 4))
-```
-
-```{r}
-elec_ts <- read_rds("data/elec_ts.rds")
-```
-
-```{r raw-50, fig.height = 8, fig.cap="The raw data for 50 households are shown.
-It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed."}
-
-```
+We can then model different data generation processes depending on our level of aggregation and time scale.
 
 # Clustering methodology
-
-## Common load clustering techniques of smart meter data
-
-The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
-None of the 34 selected papers which focus on clustering consumption are based on Australian smart meter data.
-The clustering is frequently applied directly to the raw data without scrutinizing for auto correlation and periodicity.
-The algorithm most ubiquitously employed is K-Means.
-But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters.
-Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results.
-@Tureczek2018-ha has shown that a transformation of data to incorporate autocorrelation before K-Means clustering can improve performance and enable K-Means to deliver smaller clusters with less within-cluster variance.
-However, it does not explain the cluster composition by combining it with external data.
-Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf).
-Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.
-As the methods are often situation specific,
-it makes sense to compare them on the performance rather than any standard performance metric.
-A type of clustering based on information theory such as Shannon or Renyi entropy and their variants are addressed in , which differs from typical methods adopted for electricity consumer classification, based on the Euclidean distance notion.
-@Motlagh2019-yj presents strategy to address the problems on patchy, and unequal time-series common in residential data sets by converting load time series into map models.
-Most time-series clustering models are limited to handling time domain with same start and end date and time.
-Most of the solutions to handle this like longest common subsequence, dynamic time warping are prone to computational limit with increased length of the series.
 
@@
 
-## The proposed load-clustering technique
+The data set solely contains readings from smart meters and no information about the consumers' specific physical, geographical, or behavioural attributes.
+As a result, no attempt is made to explain why consumption varies.
+Instead, this work investigates how much energy usage heterogeneity can be found in smart meter data and what some of the most common electricity use patterns are.
+It is worth noting that when studying these dynamics, a variety of objectives may be pursued.
+One objective could be to group consumers with similar shapes over all relevant cyclic granularities.
+In this scenario, the variation in customers within each group is in magnitude rather than shape, while the variation between groups is only in shape.
+Most clustering algorithms offer only daily energy profiles throughout the hours of the day, but we suggest a broader approach to the problem, aiming to group consumers with similar shapes across all significant cyclic granularities.
+Another purpose of clustering could be to group customers that have similar differences in patterns across all major cyclic granularities, capturing similar jumps across categories regardless of the overall shape.
+For example, in the first goal, similar shapes across hours of the day will be grouped together, resulting in customers with similar behaviour across all hours of the day, whereas in the second goal, any similar big-enough jumps across hours of the day will be clubbed together, regardless of which hour of the day it is.
+Both of these objectives may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
+
+The foundation of our method is unsupervised clustering algorithms based exclusively on time-series features.
+First, we study the underlying distributions that may have resulted in different patterns across temporal granularities in order to identify a mechanism to classify them based on the similarity of those distributions.
+Depending on the goal of clustering, the distance metric for defining similarity would be different.
+These distance metrics could be fed into a clustering algorithm to break large data sets into subgroups that can then be analyzed separately.
+These clusters may be commonly associated with real-world data segmentation.
+However, since the data is unlabeled a priori, more information is required to corroborate this.
+
+This section presents the work flow of the methodology:
+
+- _Find significant harmonies or cyclic granularities_:
+
+For each key variable, the raw smart meter data is indexed by time and comprises values of several variables of interest (measurement variables) at each time point.
+This sequence, though, could be depicted in a variety of ways.
+One sequence could reflect the distribution of hourly consumption over a single day, while another could indicate consumption over a week or a year.
+The sequence of measured variables throughout each category of any cyclic granularity could be regarded as a single data generating process.
+It is beneficial to evaluate those cyclic granularities that are significant for the majority of customers or noteworthy in an electricity-behavior context, so that when the customers are grouped, we can expect to observe some interesting patterns across the categories of the cyclic granularities considered.
+
+- _Distance metric_
+
+- _Clustering algorithm_
+
+- _Characterization of clusters_
+
+- A random sample  of the original data is taken for clustering analysis and includes missing and noisy observations (detailed description in Appendix)
+
+- All harmonies are computed for each customer in the sample.
+Cyclic granularities which are clashes for all customers in the sample are removed.
+
+- It is worth noting that a number of other solutions may be considered at the pre-processing stage of the method. We have considered a) Normal-Quantile Transform and b) Robust transformation.
+
+- Two methods are considered for computing dissimilarity between two customers.
+The first one involves computing according to one granularity is computed as the sum of the JS distances between distribution of all the categories of the granularity.
+When we consider more than one granularity, we consider the sum of the  average distances for all the granularity so that the combined metric is also a distance.
+
+- Given the scale of dissimilarity among the energy readings, the model chooses optimal number of clusters
+
+- Once clusters have been allocated, the groups are explored visually.
+
+- Results are reported and compared.
+
+Two methods are used for computing distances between subjects and then hierarchical clustering algorithm is used.
 
 The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
-In our work, we are using $i$ as denoting a household and the underlying variable as the electricity demand.
+In our work, we are using $i$ as denoting a customer and the underlying variable as the electricity demand.
 So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
@@
 
+### Notations
+
+Consider an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+In our work, we are using $i$ as denoting a household and the underlying variable as the electricity demand.
+Further consider a cyclic granularity of the form $B = \{ b_k: k = 1, 2, \dots, K\}$.
+Each customer consists of collection of probability distributions.
+
+So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
+We want to cluster distributions of the form $f_{i,j,k}(v)$, where $i$ and $j$ denote $i^{th}$ and $j^{th}$ customer respectively.
+
+a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
+
+### A single or pair of granularities together (change names)
+
+The methodology can be summarized in the following steps:
+
+- _Pre-processing step_
+
+Robust scaling method or NQT used for each customer.
+
+- _NQT_
+
+- _Treatment to outliers_
+
+- _Handling trend, seasonality, non-stationarity and auto-correlation_
+Trend and seasonality are common features of time series, and it is natural to characterize a time series by its degree of trend and seasonality.
+By considering the probability distributions through the use of $wpd_{norm_{s,t}}(A,B)$, these features of the time series are lost and hence there is no need to de-trend or de-seasonalize the data before performing the clustering algorithm.
+No need to exclude holiday or weekend patterns.
+
+### Many granularities together (change names)
+
+The methodology can be summarized in the following steps:
+
+1. Compute quantiles of distributions across each category of the cyclic granularity
+2. Compute JS distance between households for each each category of the cyclic granularity
+3. Total distance between households computed as sum of JS distances for all hours
+4. Cluster using this distance with hierarchical clustering algorithm (method "Ward.D")
+
+_Pro:_  
+- distance metric makes sense to group different shapes together  
+- simulation results look great on typical designs  
+_Cons:_  
+- Can only take one granularity at once  
+- Clustering a big blob of points together whereas the aim is to groups these big blob into smaller ones
+
+### Multiple-granularities
+
+_Description:_
+
+Choose all significant granularities and compute wpd for all these granularities for all customers.
+Distance between customers is taken as the euclidean distances between them with the granularities being the variables and wpd being the value under each variable for which Euclidean distance needs to be measured.
+_Pro:_  
+- Can only take many granularities at once
+- can apply variable selection PCP and other interesting clustering techniques
+- simulation results look great on typical designs
+- splitting the data into similar sized groups  
+_Cons:_  
+- distance metric does not make sense to split the data into similar shaped clusters
+
 <!-- grouping probability distributions across a harmony.
@@
 -->
-
-<!-- Trend and seasonality are common features of time series, and it is natural to characterize a time series by its degree of trend and seasonality.
-By considering the probability distributions through the use of $wpd_{norm_{s,t}}(A,B)$, these features of the time series are lost and hence there is no need to de-trend or de-seasonalize the data before performing the clustering algorithm.
--->
-
- - Dimension reduction: If each $P_{i, j, k}$ be considered to be a point in the space, key $i$ would have $mp$ dimensions as opposed to $n_i$ dimensions in case of considering raw data.
-Hence for a large number of observations ($n_i>>mp$), this approach benefits by transitioning to a lower dimension.
-
- - Avoid loss of information due to aggregation: This approach ensures key characteristic information of the data is not lost due to averaging or aggregation measures in an attempt to transition to a lower dimension.
-Hence, this approach could be thought to somehow balance the drawback of considering raw data or aggregated data.
-
-- Robustness to outliers: This approach could be adapted to be robust to outliers and extreme behaviors by trimming the tails of the probability distributions.
-
-- Non-synchronized observed time periods: Considering probability distribution would imply the clustering process can handle keys that are observed over periods of time that are overlapping but don't necessarily coincide.
-
-- Similar periodic behavior: Since cyclic granularities are considered instead of linear granularities, clustering would group keys that have similar behavior across these cyclic granularities.
-This implies they will be grouped according to their periodic behavior and not on the linear stretch of time over which they are observed.
 
@@
 
-```{r,warning=FALSE, message=FALSE,eval = FALSE}
-library(readr)
-library(here)
-library(dplyr)
-library(tsibble)
-library(gravitas)
-library(purrr)
-
-set.seed(7777)
-all_data <- read_rds(here("../paper-hakear/paper/sim_table/all_data_scalar.rds"))
-sm_cust_data <- read_rds(here("../paper-hakear/paper/data/sm_cust_data.rds"))
-
-sm_list <- sm_cust_data %>%
-  filter(customer_id == "10006414") %>% 
-  create_gran("day_week") %>% 
-  tibble() %>% 
-  ungroup() %>% 
-  group_split(day_week)
-
-p <- seq(0.01, 0.99,0.01)
-
-dist_list <- purrr::map(sm_list, function(x){
-    map(p, function(y){
-       quantile(x$general_supply_kwh, probs = y)
-    }) %>% unlist()
-}) %>% bind_rows(.id = "gran_id")
-# 
-# lag <- dist_list[2,][-1] %>% t() %>% 
-#   tibble() %>% 
-#   lag() %>% acf()
-
-```
-
-```{r mds, warning=FALSE, message=FALSE, fig.cap="Multidimensional scaling of the 100 households is shown along with the groups formed by the hierarchical clustering.", eval = FALSE}
-set.seed(7777)
-sm_unique_cust <- sm_cust_data %>% distinct(customer_id)
-sm_cust_samp <- sm_cust_data %>% 
-  #filter(customer_id %in% sm_30) %>% # for demo show 30 customers
-  #filter(year(reading_datetime) == 2012) %>% #for demo just show one column
-  as_tsibble(index = reading_datetime,
-             key = customer_id) %>% 
-  create_gran("day_week") %>% 
-  create_gran("hour_day")
-
-#all_data <- write_rds(all_data, "paper/sim_table/all_data.rds")
-
-facet_variable = "day_week"
-x_variable = "hour_day"
-
-one_harmony_data <- all_data %>%
-  #filter(customer_id %in% sm_30) %>% # for demo show 30 customers
-  filter(facet_variable == "day_week",
-         x_variable == "hour_day") %>% 
-  mutate(customer_id = as.character(customer_id))
-
-mds<- one_harmony_data %>%
-  select(customer_id, facet_variable, x_variable, wpd_norm) %>% 
-  dist() %>%          
-  cmdscale() %>%
-  as_tibble()
-
-colnames(mds) <- c("Dim.1", "Dim.2")
-rownames(mds) <- unique(one_harmony_data$customer_id)
-mds <- mds %>%
-  mutate(customer_id = as.character(unique(one_harmony_data$customer_id))) %>%
-  select((customer_id)
-         , Dim.1, Dim.2)
-
-# ggpubr::ggscatter(mds,
-#                   x = "Dim.1",
-#                   y = "Dim.2", 
-#                   label = rownames(mds),
-#                   size = 1,
-#                   repel = TRUE)
-
-# clustering as well
-
-d = one_harmony_data %>%
-  select(customer_id, facet_variable, x_variable, wpd_norm) %>% 
-  dist()
-
-hc = stats::hclust(d,method="complete")
-#plot(hc)
-groups<-cutree(hc, k=3)
-
-all_data_cluster <- cbind(one_harmony_data, groups) %>% 
-  left_join(mds, by = "customer_id") %>% 
-  mutate(groups = as.factor(groups))
-
-# q2 <- ggpubr::ggscatter(all_data_cluster,
-#                   x = "Dim.1",
-#                   y = "Dim.2", 
-#                   label = rownames(mds),
-#                   color = "groups",
-#                   size = 1,
-#                   repel = TRUE)
-
-q2 <- all_data_cluster %>% 
-  ggplot() +
-  geom_point(aes(x = Dim.1, y = Dim.2, color = groups)) +
-  scale_color_viridis_d()
-
- q2
-```
-
-```{r linear-cluster, eval = FALSE}
-group1 <- all_data_cluster %>% 
-  filter(groups == 1)
-
-group2 <- all_data_cluster %>% 
-  filter(groups == 2)
-
-group3 <- all_data_cluster %>% 
-  filter(groups == 3)
-
-g1 <- sm_cust_samp %>% 
-  filter(customer_id %in% group1$customer_id) %>% 
-   ggplot(aes(x=reading_datetime, 
-             y= general_supply_kwh), alpha = 0.5) + 
-  geom_line() +
-  scale_fill_gradient2() +
-  facet_wrap(~customer_id, nrow = 1) +
-  theme(strip.text.x = element_blank()) +
-  theme_void() +
-  ylab("group 1")
-
-g2 <- sm_cust_samp %>% 
-  filter(customer_id %in% group2$customer_id) %>% 
-   ggplot(aes(x=reading_datetime, 
-             y= general_supply_kwh), alpha = 0.5) + 
-  geom_line() +
-  scale_fill_gradient2() +
-  facet_wrap(~customer_id, nrow= 1) +
-  theme(strip.text.x = element_blank()) +
-  theme_void() +
-  ylab("group 2")
-
-g3 <- sm_cust_samp %>% 
-  filter(customer_id %in% group3$customer_id) %>% 
-   ggplot(aes(x=reading_datetime, 
-             y= general_supply_kwh), alpha = 0.5) + 
-  geom_line() +
-  scale_fill_gradient2() +
-  theme(strip.text.x = element_blank()) +
-  theme_void() +
-  ylab("group 3")
-
-# g1/g2 + g3
-
-```
-
-```{r cyclic-cluster, warnings = FALSE, message = FALSE, fig.cap = "Distribution of energy demand across day-of-week and hour-of-day for three groups clustered through the proposed methodology.
-For group-1, the morning peak looks very sharp, for group-2 the morning and evening peaks are rounded and spread over few hours, for group-3 peaks are less rounded.
-They differ in their relationship between 75th and 90th percentile.", eval = FALSE}
-cg1 <- sm_cust_samp %>% 
-  filter(customer_id %in% group1$customer_id) %>% 
-  prob_plot("day_week", "hour_day", response = "general_supply_kwh",
-            plot_type = "quantile", symmetric = TRUE, quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
-  scale_color_brewer(palette = "RdPu") + ggtitle("group 1") +
-  scale_x_discrete(breaks = seq(0, 23, 5))
-
-cg2 <- sm_cust_samp %>% 
-  filter(customer_id %in% group2$customer_id) %>% 
-  prob_plot("day_week", "hour_day", response = "general_supply_kwh",
-            plot_type = "quantile", symmetric = TRUE, quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
-  scale_color_brewer(palette = "RdPu") + ggtitle("group 2") +
-  scale_x_discrete(breaks = seq(0, 23, 5))
-
-cg3 <- sm_cust_samp %>% 
-  filter(customer_id %in% group3$customer_id) %>% 
-  prob_plot("day_week", "hour_day", response = "general_supply_kwh",
-            plot_type = "quantile", symmetric = TRUE, quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
-  scale_color_brewer(palette = "RdPu") + ggtitle("group 3") +
-  scale_x_discrete(breaks = seq(0, 23, 5))
-
-cg1 
-cg2
-cg3
-
-```
-
-###  Distinction, repeatability, and robustness metrics
+_Distinction, repeatability, and robustness metrics_
 
@@
 
-## Cluster validation
+_Cluster validation_
 
@@
 Generally most of the indices used for internal clustering validation combine compactness and separation measures.
+
+## Results
+
+## Software implementation
+
+The implementation for our framework is available in the R package [`gracsr`](https://github.com/Sayani07/gracsr) for ease of use in other applications.
+
+# Results
+
+## Clustering results for 100 customers
+
+## Clustering results for 5K customers
 
@@
 
-## Conclusion
+## Discussion
 
 This section will cover some drawback of this clustering method and potential extensions of this work.
-
-# R packages
-
-The implementation for our framework for the first paper is available in the R package [`gravitas`](https://github.com/Sayani07/gravitas) and that for the second paper is made available in the package [`hakear`](https://github.com/Sayani07/hakear) for ease of use in other applications.
-
-# Discussion
-
-The final chapter of my thesis will consist of the concluding remarks of the whole thesis and possible future extensions.
 

### 2021-09-29 Simulations.Rmd
methodology subsection 1: finding significant granularities done

---
title: "simulations"
author: "Sayani Gupta"
date: "27/09/2021"
preamble: >
  \setlength {\marginparwidth }{2cm}
  \usepackage{mathtools,amssymb,booktabs,longtable,todonotes,amsthm}
  \def\mod{~\text{mod}~}
output:
  bookdown::pdf_book:
    base_format: rticles::asa_article
    fig_height: 4
    fig_width: 6
    fig_caption: yes
    dev: "pdf" 
    keep_tex: yes
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(readr)
library(tidyverse)
library(gravitas)
library(tsibble)
library(parallel)
library(magrittr)
library(GGally)
library(knitr)
library(patchwork)
library(caret)
library(plotly)
library(gracsr)
library(here)

```

## Materials and methods

### Simulations

To test clustering algorithm solutions against ground truth cluster assignments, we generate simulations to represent important problems that could be encountered in electricity data contexts. The code to generate these simulations are included in the Supplementary paper. When we move from the linear to the cyclic world of temporal granularities, we can see patterns across different categories of the granularities which gets lost in the linear representation. Populations are modeled by a collection of these cyclic granularities. Each cyclic granularity might or might not have patterns across its categories. Each cluster is characterized by similar patterns across one or more of these cyclic granularities.

A small example is given to setup the problem.

We constructed the simulation parameters to represent common patterns in electricity data. 50, 200 and 500 time series were chosen for different simulation designs with different granularities and patterns changing across different granularities.
The data type is fixed to be "continuous".We generated independent replications of all combinations of the simulation parameters.

Consider a continuous time series variable $y$ of length $T$ indexed by ${0, 1, \dots T-1}$. Three circular granularities $g1$, $g2$ and $g3$ are considered with 2, 3 and 5 levels respectively. Categories of g1, g2 and g3 are represented by ${0,1}$, ${0, 1, 2}$ and ${0, 1, 2, 3, 4}$. These categories could be integers or some more meaningful labels. For example, the granularity "day-of-week" could be either represented by ${0, 1, 2, \dots, 6}$ or ${Mon, Tue, \dots, Sun}$.

<!-- Consider the case where distribution of $y$ would vary across levels of $g1$ and g2$ but will not vary across levels of $g3$ as described in Figure .Figure \ref shows four designs where $g3$ acts as a nuisance variable, and $g1$ and $g2$ varies in both  Design 2 and 3, and $g2$ varies in all designs. -->

Consider a case where distribution of $y$ would vary across levels of $g2$
for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs. Figure \ref{#fig:plot-linear} shows the linear and cyclic representation of $y$. The first panel shows raw plot of $y$ in a linear scale and the second panel shows distribution of $y$ across cyclic granularities namely $g1$, $g2$ and $g3$. As could be seen from the plots, it is impossible to decipher from the raw time plot that the time series variable shows such pattern across different granularities.

```{r change-var, echo=FALSE}
t = 300
n <- seq(0, t, 1)
g1 <- n %%2
g2 <- n %%3
g3 <- n %%5
mu11 = c(0, 5)
mu12 = c(3, 0)
mu21= c(2, 0, 0)
mu22 = c(0, 2, 0)
mu23 = c(0, 0, 2)
mu3 = c(0, 0, 0, 0, 0)
```

```{r generate-design} 
generate_design <- function(t, mu21, mu22, mu23){

t <- seq(0, t, 1)
g1 <- t %%2
g2 <- t %%3
g3 <- t %%5

# null design
g1_dnull <- rep( rep(0, each = length(unique(g1))), length.out= length(t))
g2_dnull <- rep( rep(0, each = length(unique(g2))), length.out= length(t))
g3_dnull <- rep( rep(0, each = length(unique(g3))), length.out= length(t))

# mean changing across categories in varying ways

g11_dvary <- rep(mu11, length.out= length(t))
g12_dvary <- rep(mu12, length.out= length(t))
g21_dvary <- rep(mu21, length.out= length(t))
g22_dvary <- rep(mu22, length.out= length(t))
g23_dvary <- rep(mu23, length.out= length(t))
g3_dvary <- rep(0, length.out= length(t))

design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
design2 = distributional::dist_normal(g11_dvary + g21_dvary + g3_dnull)
design3 = distributional::dist_normal(g12_dvary + g22_dvary + g3_dnull)
design4 = distributional::dist_normal(g1_dnull + g23_dvary + g3_dnull)

data_bind <- tibble::tibble(
index = t,
g1 = g1,
g2 = g2,
g3 = g3,
design1 = distributional::generate(design1, times = 1) %>% unlist(),
design2 = distributional::generate(design2, times = 1) %>% unlist(),
design3 = distributional::generate(design3, times = 1) %>% unlist(),
design4 = distributional::generate(design4, times = 1) %>% unlist(),
#design5 = distributional::generate(design5, times = 1) %>% unlist()
) %>% 
  pivot_longer(-c(1, 2, 3, 4), names_to = "design", values_to = "sim_data")

data_bind
}

data_bind <- generate_design(t, mu21, mu22, mu23)

#data_bind

```

```{r plot-linear , out.width="100%", fig.cap="The linear and cyclic representation of the time series variable y. It is not possible to comprehend these patterns across cyclic granularities g1, g2 and g3 or group similar series just by looking at the linear plots."}

# plot_linear_data <- function(data){
# ggplot(data,
#              aes(x = index, y = sim_data)) + 
#   geom_line() +
#   xlab("index")+
#   theme_bw() 
# }

ggplot(data_bind,
             aes(x = index, y = sim_data)) + 
  geom_line() +
  xlab("index")+
  theme_bw() +
  facet_wrap(~design, scales = "free_y",ncol =1)+ theme(
    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
  theme(plot.margin = margin(0, 0, 0, 0, "cm") )

p2 <- ggplot(data_bind,
             aes(x = as.factor(g1), y = sim_data)) + 
  geom_boxplot(alpha =0.5) + xlab("g1") + 
  theme_bw() +
  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
    fun = median,
    geom = 'line',
    aes(group = 1), size = 1, color = "blue") +
  theme(plot.margin = margin(0, 0, 0, 0, "cm") )

p3 <- ggplot(data_bind, aes(x = as.factor(g2), y = sim_data)) + geom_boxplot(alpha =0.5) + xlab("g2") + theme_bw() +
  facet_wrap(~design, scales = "free_y",ncol = 1)+ stat_summary(
    fun = median,
    geom = 'line',
    aes(group = 1), size = 1, color = "blue") + ylab("")+
  theme(plot.margin = margin(0, 0, 0, 0, "cm") )

p4 <- ggplot(data_bind, aes(x = as.factor(g3), y = sim_data)) + geom_boxplot(alpha =0.5) +
  xlab("g3") + theme_bw()+
  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
    fun = median,
    geom = 'line',
    aes(group = 1), size = 1, color = "blue")+ ylab("")+
  theme(plot.margin = margin(0, 0, 0, 0, "cm") )

(p2 + p3 + p4) * theme(
    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)),
    panel.spacing =unit(0, "lines")) * theme(plot.margin = margin(0, 0, 0, 0, "cm") )

# 
# 
# plot_cyclic_data( data_bind %>% filter(design=="design2"))
# plot_data( data_bind %>% filter(design=="design3"))
# plot_data( data_bind %>% filter(design=="design4"))
# plot_data( data_bind %>% filter(design=="design5"))
```

A subset of many possible designs are shown in Figure \ref{fig: plot-linear }. For the parameter space (XXX unique  combinations shown in table YYY), 100, 500 independent replications of all possible combination of simulation parameters were generated. The clustering methodologies were run all these unique combinations and subsets of these to verify if the methodologies work as expected.

<!-- Many time series with observations are created using the four designs below, each of which is iterated five times. We anticipate to have 4 clusters, each with five time series conforming to the same design, once we execute the clustering.  -->

<!-- Distribution of each categories (cat-0, cat-1, cat-2) is (1, 1, 1), (1, 2, 3), (1, 3, 2), (2, 1, 3), (2, 3, 1), (3, 1, 2), (3, 2, 1) corresponding to 7 scenarios each with many designs. For the scenario (1, 1, 1) there will be have $(2^3 -2) = 6$ possible ways in which $g2$ can vary, for the  scenario $(1, 2, 3)$, there will be $(2^3 -1) = 7$ possible ways in which $g2$ can vary. Imagine the null design corresponds to (0, 0, 0) when there is no pattern across categories.Since $g2$ has three levels, each of which can vary or not, we have $(2^3 -2) = 6$ possible ways in which $g2$ can vary leading to $7$ designs as follows (2 correspond to the cases where all categories vary or no categories vary.) -->

### Clustering algorithms

We choose the algorithms with most common use in smart meter clustering. . First, we used agglomerative hierarchical clustering with Ward’s criterion
(HC), a dominant approach for smart meter clustering. Second, we also used k-means clustering. For each clustering algorithm, we assumed
the number of clusters was known, recovering the number of clusters
given from the simulation parameters.

| Granularity type                                                   	| # Significant 	| # Replications 	|
|--------------------------------------------------------------------	|---------------	|----------------	|
| **Individual**  <br><br># obs: 300, 500, 2000  <br># clusters: 6/7 	| 1/2/3         	| 25, 100, 200   	|
| **Interaction**  <br><br># obs: 500, 2000  <br># clusters: 4       	| 1/2           	| 25, 100, 200   	|

### Distance metrics

#### A single or pair of granularities together (change names)

The methodology can be summarized in the following steps:

- _Pre-processing step_

Robust scaling method or NQT used for each customer.

- _NQT_

- _Treatment to outliers_

- _Handling trend, seasonality, non-stationarity and auto-correlation_
Trend and seasonality are common features of time series, and it is natural to characterize a time series by its degree of trend and seasonality. By considering the probability distributions through the use of $wpd_{norm_{s,t}}(A,B)$, these features of the time series are lost and hence there is no need to de-trend or de-seasonalize the data before performing the clustering algorithm. No need to exclude holiday or weekend patterns.

#### Many granularities together (change names)

The methodology can be summarized in the following steps:

1. Compute quantiles of distributions across each category of the cyclic granularity
2. Compute JS distance between households for each each category of the cyclic granularity
3. Total distance between households computed as sum of JS distances for all hours
4. Cluster using this distance with hierarchical clustering algorithm (method "Ward.D")

_Pro:_  
- distance metric makes sense to group different shapes together  
- simulation results look great on typical designs  
_Cons:_  
- Can only take one granularity at once  
- Clustering a big blob of points together whereas the aim is to groups these big blob into smaller ones

#### Multiple-granularities

_Description:_

Choose all significant granularities and compute wpd for all these granularities for all customers. Distance between customers is taken as the euclidean distances between them with the granularities being the variables and wpd being the value under each variable for which Euclidean distance needs to be measured.  
_Pro:_  
- Can only take many granularities at once
- can apply variable selection PCP and other interesting clustering techniques
- simulation results look great on typical designs
- splitting the data into similar sized groups  
_Cons:_  
- distance metric does not make sense to split the data into similar shaped clusters

### Prototype application

A clean data set is obtained by carefully choosing customers which shows similar shapes across one or more cyclic granularity. Since this is unlabeled data, there is no way to do external validation of our methodologies. Thus, we chose this way to see how well our methodology works in a cleaner data set as this one.

Fig \ref{fig:}, \ref{fig:} and \ref{fig:} shows the distribution of 24 customers across hour-of-day, month-of-year and wknd-wday respectively. Every row in \ref{fig:} shows different shapes across hour-of-day and across columns show similar shapes for each row.
We use our methodology to see if the customers are allocated to same group have similar shapes across one or more granularity.

```{r data-pick}

quantile_prob_graph <- c(0.25, 0.5, 0.75)

wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_robust.rds")) %>% broom::tidy()

moy <- read_rds(here("data/dist_gran_moy_356cust.rds")) %>% broom::tidy()

hod <- read_rds(here("data/dist_gran_hod_356cust_robust.rds")) %>% broom::tidy()

distance <- wkndwday %>% 
  left_join(moy, by = c("item1", "item2")) %>% 
  left_join(hod, by = c("item1", "item2")) %>% 
  rename("wkndwday" ="distance.x",
         "moy" = "distance.y",
         "hod" = "distance") %>%
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

total_distance <- distance %>% 
  mutate(total = hod)

data_pick_one <- total_distance %>% filter(item1 %in% 
                            c(8454221)) %>% group_by(item1) %>% arrange(total) %>% head(5) %>% 
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

#11013154  8181071 11048034
data_pick_two <- total_distance %>% filter(item1 %in% 
                            c(8181071)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

data_pick_three <- total_distance %>% filter(item1 %in% 
                            c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

data_pick_three <- total_distance %>% filter(item1 %in%
                            c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

# 8653709
data_pick_four <- total_distance %>% filter(item1 %in%
                            c(8184707)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

data_pick_cust <- bind_rows(
unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(),    unique(c(data_pick_two$item1,data_pick_two$item2)) %>% as_tibble(), 
c(11013154, 8495194, 8627007,10109182,10677705,8952846) %>% as_tibble(),
unique(c(data_pick_four$item1,data_pick_four$item2)) %>% as_tibble(),
.id = "design")

data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))%>%
  filter(customer_id %in% data_pick_cust$value) %>% 
  gracsr::scale_gran( method = "robust", response = "general_supply_kwh")

# data_pick_three <- total_distance %>% filter(item1 %in% 
#                             c(11013154)) %>% 
#   mutate(item1 = as.integer(as.character(item1)),
#          item2 = as.integer(as.character(item2)))
```

```{r heatmap-hod}
sm_hod <- gravitas::create_gran(data_pick, "hour_day")

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)  %>% 
  left_join(data_pick_cust, by = c("customer_id" = "value"))

ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, ncol(sm_hod_list))]))
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list%>% 
     dplyr::filter(customer_id == customer_id[x]) %>% 
     select(-c(1, ncol(sm_hod_list))) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

ref_cat <- names(sm_hod_list)[-c(1, ncol(sm_hod_list))] %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
    mutate(customer_serial_id = row_number())

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>%
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "customer_serial_id") %>% 
  select(customer_serial_id, categories_serial_id, category, customer_id, everything())

data_heatmap_hod <- sm_quantiles_ref %>% 
 left_join(data_pick %>% distinct(customer_id)) %>% 
  left_join(data_pick_cust, by = c("customer_id" = "value")) %>% 
  mutate(categories_serial_id = as.factor(categories_serial_id))

data_heatmap_hod %>% 
  ggplot(aes(x = categories_serial_id)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=customer_id), fill = "lightblue") +
  geom_line(aes(y = `50%`, group=customer_id)) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
  scale_x_discrete(breaks = seq(1, 24, 3))

```

```{r heatmap-moy}
sm_hod <- gravitas::create_gran(data_pick, "month_year")

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, month_year, general_supply_kwh) %>% 
  pivot_wider(names_from = month_year,
              values_from = general_supply_kwh)  %>% 
  left_join(data_pick_cust, by = c("customer_id" = "value"))

ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, ncol(sm_hod_list))]))
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list%>% 
     dplyr::filter(customer_id == customer_id[x]) %>% 
     select(-c(1, ncol(sm_hod_list))) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

ref_cat <- names(sm_hod_list)[-c(1, ncol(sm_hod_list))] %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
    mutate(customer_serial_id = row_number())

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>%
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "customer_serial_id") %>% 
  select(customer_serial_id, categories_serial_id, category, customer_id, everything())

data_heatmap_moy <- sm_quantiles_ref %>% 
 left_join(data_pick %>% distinct(customer_id)) %>% 
  left_join(data_pick_cust, by = c("customer_id" = "value")) %>% 
  mutate(categories_serial_id = as.factor(categories_serial_id))

data_heatmap_moy %>% 
  ggplot(aes(x = categories_serial_id)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=customer_id), fill = "lightblue") +
  geom_line(aes(y = `50%`, group=customer_id)) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
  scale_x_discrete(breaks = seq(1, 12, 2))

```

```{r heatmap-wkndwday}
sm_hod <- gravitas::create_gran(data_pick, "wknd_wday")

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, wknd_wday, general_supply_kwh) %>% 
  pivot_wider(names_from = wknd_wday,
              values_from = general_supply_kwh)  %>% 
  left_join(data_pick_cust, by = c("customer_id" = "value"))

ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, ncol(sm_hod_list))]))
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list%>% 
     dplyr::filter(customer_id == customer_id[x]) %>% 
     select(-c(1, ncol(sm_hod_list))) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

ref_cat <- names(sm_hod_list)[-c(1, ncol(sm_hod_list))] %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
    mutate(customer_serial_id = row_number())

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>%
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "customer_serial_id") %>% 
  select(customer_serial_id, categories_serial_id, category, customer_id, everything())

data_heatmap_wkndwday <- sm_quantiles_ref %>% 
 left_join(data_pick %>% distinct(customer_id)) %>% 
  left_join(data_pick_cust, by = c("customer_id" = "value")) %>% 
  mutate(categories_serial_id = as.factor(categories_serial_id))

data_heatmap_wkndwday %>% 
  ggplot(aes(x = categories_serial_id)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=customer_id), fill = "lightblue") +
  geom_line(aes(y = `50%`, group=customer_id)) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))

```

```{r clustering}
filtered_distance <- distance %>% filter(item1 %in% data_pick_cust$value) %>% 
  filter(item2 %in% data_pick_cust$value)

# filtered_distance <- bind_rows(data_pick_one, )
#   
#   data_pick_one

total_distance <- filtered_distance %>% 
mutate(total = wkndwday/2 + moy/12 + hod/24) 
#mutate(total = hod)

total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
                               names_from = item2,
                               values_from = total)
#total_distance_wide$`11013154` <- NA

rownames(total_distance_wide) <- total_distance_wide$item1

mds_data <- total_distance_wide %>% 
  mutate_all(~replace(., is.na(.), 0)) %>%
   tibble::rownames_to_column() %>%  
   select(-item1) %>% 
   pivot_longer(-rowname) %>% 
   pivot_wider(names_from=rowname, values_from=value)

rownames(mds_data) <- total_distance_wide$item1

# group <- mds_data[-1] %>% clust_gran(kopt=4)
#

df <- mds_data[-1] %>% as.matrix()
DM <- matrix(0, ncol(mds_data), ncol(mds_data))
DM[lower.tri(DM)] = df[lower.tri(df, diag=TRUE)]

f = as.dist(DM)

#f %>% hclust %>% cutree(k=4)

first_lot <- mds_data %>% names()
# because distance matrix is such  that 1:23 appears in columns and 2:24 appears in rows such that 1st customer is missing from row and last customer is missing from variables
id <- c(first_lot[-1], mds_data$name[nrow(mds_data)])

group <- f %>% hclust (method = "ward.D") %>% cutree(k=4)

cluster_result <- bind_cols(id = id, group = group) %>% mutate(id =  as.numeric(id))

mds  = cmdscale(f)
rownames(mds) = id
colnames(mds) = c("mds1", "mds2")

ggplot(data = mds %>% as_tibble() %>% mutate(group = cluster_result$group), 
       aes(x = mds1, y = mds2, color = as.factor(group))) +
  geom_point() +
  geom_text(aes(label = id))

cluster_result %>% filter(group==1)
cluster_result %>% filter(group==2)
cluster_result %>% filter(group==3)
cluster_result %>% filter(group==4)

  #x = ~mds1.hod, y = ~ mds2.hod, text = ~customer_id
```

```{r data-heatmap-hod}
data_heatmap_hod %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = categories_serial_id)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=customer_id), fill = "lightblue") +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")
```

```{r data-heatmap-moy}
data_heatmap_moy %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = categories_serial_id)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=customer_id), fill = "lightblue") +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")
```

```{r data-heatmap-wkndwday}
data_heatmap_wkndwday %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = categories_serial_id)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=customer_id), fill = "lightblue") +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")

```

```{r first-try, eval = FALSE}

sm_hod_list_new <- sm_hod_list %>% left_join(cluster_result, by = c("customer_id" = "id"))

sm_hod_list_ncol <- sm_hod_list_new[-c(1, ncol(sm_hod_list_new),(ncol(sm_hod_list_new)-1))]

ncol_sm <- unique(cluster_result$group)
nrow_sm <- seq_len(nrow(sm_hod_list_new))

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list_new%>% 
     dplyr::filter(group == group[x]) %>% 
     #select(-c(1, ncol(sm_hod_list_new),(ncol(sm_hod_list_new)-1))) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
})  %>% bind_rows(.id = "group_id")
}) %>% bind_rows(.id = "customer_serial_id")

ref_cat <- names(sm_hod_list_ncol) %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- sm_hod_list_new$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
    mutate(customer_serial_id = row_number())

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>%
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "customer_serial_id") %>% 
  select(customer_serial_id, categories_serial_id, category, customer_id, everything())

data_heatmap_hod <- sm_quantiles_ref %>% 
 left_join(data_pick %>% distinct(customer_id)) %>% 
  left_join(data_pick_cust, by = c("customer_id" = "value")) %>% 
  mutate(categories_serial_id = as.factor(categories_serial_id))

data_heatmap_hod %>% 
  ggplot(aes(x = categories_serial_id)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=customer_id), fill = "lightblue") +
  geom_line(aes(y = `50%`, group=customer_id)) +
  facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
  scale_x_discrete(breaks = seq(1, 24, 3))

```

### 2021-09-29 main.Rmd
methodology subsection 1: finding significant granularities done

@@
 
-For each key variable, the raw smart meter data is indexed by time and comprises values of several variables of interest (measurement variables) at each time point.
+For each key variable, the raw smart meter data is a sequence that is indexed by time and comprises values of several variables of interest (measurement variables) at each time point.
 This sequence, though, could be depicted in a variety of ways.
-One sequence could reflect the distribution of hourly consumption over a single day, while another could indicate consumption over a week or a year.
-The sequence of measured variables throughout each category of any cyclic granularity could be regarded as a single data generating process.
-It is beneficial to evaluate those cyclic granularities that are significant for the majority of customers or noteworthy in an electricity-behavior context, so that when the customers are grouped, we can expect to observe some interesting patterns across the categories of the cyclic granularities considered.
+A shuffling of the raw sequence could reflect the distribution of hourly consumption over a single day, while another could indicate consumption over a week or a year.
+These temporal deconstructions of a time period into units such as hour-of-day, work-day/weekend, called cyclic temporal granularities, are useful for exploring repetitive patterns in time series data that get lost in the linear representation of time.
+However, it is advantageous to consider those cyclic granularities across which there is a significant repetitive pattern for the majority of customers or noteworthy in an electricity-behavior context, because when the customers are grouped, we can expect to observe some interesting patterns across the categories of the cyclic granularities considered.
 
-- _Distance metric_
+- _Distance metrics_
+
+The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+In our work, we are using $i$ as denoting a customer and the underlying variable as the electricity demand.
+So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
+
+If we consider the sequence of measured variables for each category of any cyclic granularity to be from a single data generating process, we want to cluster distributions of the form $f_{i,j,k}(v)$, where $j$ and $k$ denote two granularities.
+For example, j=hour of day and k=day of week.
+We still want the households in clusters, but each household consists of a collection of probability distributions f_{i,h,d}(x) where $h=1,...,24$ and $d=1,...,7$.
+In this example (hour of day/day of week), we would have $24*7=168$ distributions per household if we consider interaction of the households or $24+7=31$ distributions if individual effect of granularities need to be considered.
+So we need a distance between collections of these univariate probability distributions.
+There could be several ways of considering such distances.
+We consider the following two approoaches:
+
+Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
+Each household consists of a $J*K$ distributions one harmony.
+We compute the distributional difference
+between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$.
+$wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies.
+If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
+We aim to have pockets of households showing similar periodic behavior by conidering $wpd$ vlaues for different harmonies and some time series features.
+The features should also characterize probability distributions of different household.
 

### 2021-09-29 supplementary/distance-computation/real-data/04handpick-from356-mds.Rmd
methodology subsection 1: finding significant granularities done

--- 2021-09-22 supplementary/distance-computation/real-data/04handpick-from356cust.Rmd
@@
 library(plotly)
+library(gracsr)
+library(here)
 ```
@@
          item2 = as.integer(as.character(item2)))
-# for hod MDS scale
 
-total_distance <- distance %>% 
-  mutate(total = hod)
+create_mds_gran <- function(total_distance){
 
@@
 
-total_distance_wide$`8143599` <- NA
+rownames(total_distance_wide) <- total_distance_wide$item1
 
-total_distance_wide <- bind_rows(total_distance_wide, c(item1=11590442))
+#  changing to a lower traingular matrix as dist only takes lower traingular matrix (so in way tranposing it but tidyverse style, also creating a diagonal 0 with DM otherwise as.dist changing value)
+mds_data <- total_distance_wide %>% 
+  mutate_all(~replace(., is.na(.), 0)) %>%
+   tibble::rownames_to_column() %>%  
+   select(-item1) %>% 
+   pivot_longer(-rowname) %>% 
+   pivot_wider(names_from=rowname, values_from=value)
 
-mds_data <- total_distance_wide[-1]  %>% mutate_all(~replace(., is.na(.), 0))
+df <- mds_data[-1] %>% as.matrix()
+DM <- matrix(0, ncol(mds_data), ncol(mds_data))
+DM[lower.tri(DM)] = df[lower.tri(df, diag=TRUE)]
 
-mds <- mds_data %>% t() %>% as.dist() %>% cmdscale() %>% as_tibble()
-colnames(mds) <- c("mds1.hod", "mds2.hod")
+f = as.dist(DM)
 
-mds_hod <- bind_cols(mds, customer_id = mds_data %>% t() %>% rownames()) # since cmdscale is on transpose # because it only considers lower traingular matrix
+first_lot <- mds_data %>% names()
+id <- c(first_lot[-1], mds_data$name[nrow(mds_data)])
+
+mds  = cmdscale(f)
+rownames(mds) = id
+colnames(mds) = c("mds1", "mds2")
+
+mds_hod <- bind_cols(mds, customer_id = id)
+
+mds_hod
+ }
+
+# for hod MDS scale
+
+total_distance_hod <- distance %>% 
+  mutate(total = hod)
+
+mds_hod <- create_mds_gran(total_distance_hod)
+
+colnames(mds_hod) <- c("mds1.hod", "mds2.hod", "customer_id")
 
@@
 
-total_distance <- distance %>% 
+total_distance_moy <- distance %>% 
   mutate(total = moy)
 
-total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
-                               names_from = item2,
-                               values_from = total)
+mds_moy <- create_mds_gran(total_distance_moy)
 
-total_distance_wide$`8143599` <- NA
-
-total_distance_wide <- bind_rows(total_distance_wide, c(item1=11590442))
-
-mds_data <- total_distance_wide[-1]  %>% mutate_all(~replace(., is.na(.), 0))
-
-mds <- mds_data %>% t()%>% as.dist()  %>% cmdscale() %>% as_tibble()
-colnames(mds) <- c("mds1.moy", "mds2.moy")
-mds_moy <- bind_cols(mds, customer_id = mds_data %>% t() %>% rownames())
+colnames(mds_moy) <- c("mds1.moy", "mds2.moy", "customer_id")
 
@@
 
-total_distance <- distance %>% 
+total_distance_wkndwday <- distance %>% 
   mutate(total = wkndwday)
 
-total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
-                               names_from = item2,
-                               values_from = total)
-total_distance_wide$`8143599` <- NA
+mds_wkndwday <- create_mds_gran(total_distance_wkndwday)
 
-total_distance_wide <- bind_rows(total_distance_wide, c(item1=11590442))
-
-mds_data <- total_distance_wide[-1]  %>% mutate_all(~replace(., is.na(.), 0))
-
-mds <- mds_data %>% t() %>% as.dist() %>% cmdscale() %>% as_tibble()
-colnames(mds) <- c("mds1.wkndwday", "mds2.wkndwday")
-mds_wkndwday <- bind_cols(mds, customer_id = mds_data %>% t() %>% rownames())
-
-# wkndwday+moy+
+colnames(mds_wkndwday) <- c("mds1.wkndwday", "mds2.wkndwday", "customer_id")
 
@@
 
-# robust_scale <-  function(x){
-#   (x - stats::median(x))/(stats::IQR(x))
-# }
-# 
-# scaled_var <- apply(distance[-c(1, 2)], 2, robust_scale) %>% as_tibble()
-# 
-# 
-# distance$wkndwday <- scaled_var$wkndwday
-# distance$hod <- scaled_var$hod
-# distance$moy <- scaled_var$moy
-#
+total_distance_total <- distance %>% 
+  mutate(total = wkndwday/2 + moy/12 + hod/24)
 
-total_distance <- distance %>% 
-  mutate(total = wkndwday + moy + hod)
+mds_total <- create_mds_gran(total_distance_total)
 
-total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
-                               names_from = item2,
-                               values_from = total)
+colnames(mds_total) <- c("mds1.total", "mds2.total", "customer_id")
 
-total_distance_wide$`8143599` <- NA
+```
 
-total_distance_wide <- bind_rows(total_distance_wide, c(item1=11590442))
-
-mds_data <- total_distance_wide[-1]  %>% mutate_all(~replace(., is.na(.), 0))
-
-mds <- mds_data %>% t()%>% as.dist()  %>% cmdscale() %>% as_tibble()
-colnames(mds) <- c("mds1.total", "mds2.total")
-mds_total <- bind_cols(mds, customer_id = mds_data %>% t() %>% rownames())
+```{r wpd-mds}
 
@@
 
-# robust_scale <-  function(x){
-#   (x - stats::median(x))/(stats::IQR(x))
-# }
-# 
-# elec_pick_wide_scaled <- apply(elec_pick_wide[-1], 2, robust_scale)
-
 mds <- elec_pick_wide[-1] %>% dist() %>% cmdscale() %>% as_tibble()
@@
 mds_wpd <- bind_cols(mds, customer_id = elec_pick_wide$customer_id)%>% mutate(customer_id = as.character(customer_id))
+```
 
+```{r mds-all}
 mds_all <- mds_hod %>% 
@@
 
-mds_all
 ```
@@
 
-```{r}
+```{r mds-all}
 
@@
 
-subplot(mds_total, mds_hod, mds_moy ,mds_wkndwday, nrows = 2, titleX = TRUE) %>%
+subplot(mds_total, mds_hod, mds_moy ,mds_wkndwday, titleX = TRUE) %>%
   layout(showlegend = FALSE) %>%
@@
 ```{r wpd-total}
-# subplot(mds_wpd, mds_total, nrows = 1, titleX = TRUE) %>%
-#   layout(showlegend = FALSE) %>%
-#   highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect") 
+subplot(mds_wpd, mds_total, nrows = 1, titleX = TRUE) %>%
+  layout(showlegend = FALSE) %>%
+  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect")
 ```
@@
 
-sm_hod <- gravitas::create_gran(elec_356_raw, "hour_day")
+quantile_hod <- quantile_gran(elec_356_raw, gran1 = "hour_day", quantile_prob_val = quantile_prob_graph)
 
-sm_hod_list <- sm_hod %>% 
-  as_tibble() %>% 
-  select(customer_id, hour_day, general_supply_kwh) %>% 
-  pivot_wider(names_from = hour_day,
-              values_from = general_supply_kwh)
+quantile_moy <- quantile_gran(elec_356_raw, gran1 = "month_year", quantile_prob_val = quantile_prob_graph)
 
-ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, ncol(sm_hod_list))]))
-nrow_sm <- seq_len(nrow(sm_hod_list))
+quantile_wkndwday <- quantile_gran(elec_356_raw, gran1 = "wknd_wday", quantile_prob_val = quantile_prob_graph)
 
-sm_hod_quantiles_cat <- map(nrow_sm, function(x){
-  map(ncol_sm, function(y){
-   cell <- sm_hod_list%>% 
-     dplyr::filter(customer_id == customer_id[x]) %>% 
-     select(-c(1, ncol(sm_hod_list))) %>% 
-     extract(y) %>% 
-     unlist()
-   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
-})  %>% bind_rows(.id = "categories_serial_id")
-}) %>% bind_rows(.id = "customer_serial_id")
+quantile_data <- bind_rows(quantile_hod, quantile_moy, quantile_wkndwday)
 
-ref_cat <- names(sm_hod_list)[-c(1, ncol(sm_hod_list))] %>% as_tibble() %>% set_names("category") %>%  
-  mutate(categories_serial_id = row_number())
+elec_pick_wide <- elec_pick %>% pivot_wider(c(3:5), names_from = "x_variable", values_from = wpd)
 
-ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
-    mutate(customer_serial_id = row_number())
-
-sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
-  mutate(customer_serial_id = as.integer(customer_serial_id),
-         categories_serial_id = as.integer(categories_serial_id)) %>%
-  left_join(ref_cat, by = "categories_serial_id") %>% 
-  left_join(ref_cust, by = "customer_serial_id") %>% 
-  select(customer_serial_id, categories_serial_id, category, customer_id, everything())
-
-data_heatmap <- sm_quantiles_ref %>% 
- left_join(elec_pick %>% distinct(customer_id)) %>% 
-  mutate(categories_serial_id = as.factor(categories_serial_id)) %>% 
+data_heatmap <- quantile_data %>% 
+ left_join(elec_pick_wide) %>% 
   mutate(customer_id = as.character(customer_id))
 
+# %>% 
+#   mutate(categories_serial_id = as.factor(categories_serial_id)) %>% 
+#   mutate(customer_id = as.character(customer_id))
+
 data_heatmap_new <- data_heatmap %>%
-left_join(mds_all, by = "customer_id")
+left_join(mds_all, by = "customer_id") %>%  pivot_wider( names_from = quantiles, values_from = quantiles_values)
 
@@
  quantile_plot <- base %>% 
-   group_by(customer_id) %>%
-   add_lines(x = ~ categories_serial_id, y = ~`50%`, color = ("black")) %>%
-   add_ribbons(x = ~categories_serial_id,
+   group_by(customer_id, gran) %>%
+   add_lines(x = ~ category, y = ~`50%`, color = ("black")) %>%
+   add_ribbons(x = ~category,
                ymin = ~`25%`,
@@
 
-subplot(mds_total, mds_wpd, quantile_plot, nrows = 2, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect")
+subplot(mds_total, quantile_plot, nrows = 2, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect")
 
+# hod
+
+data_heatmap_hod <- data_heatmap_new %>% filter(gran == "hour_day") %>% 
+  mutate(category = as.integer(category))
+tx <- highlight_key(data_heatmap_hod, ~customer_id)
+
+ # initiate a plotly object
+base <- plot_ly(tx, color = I("black")) %>%
+   group_by(customer_id)
+
+ quantile_plot <- base %>%
+   group_by(customer_id, gran) %>%
+   add_lines(x = ~ category, y = ~`50%`, color = ("black")) %>%
+   #add_ribbons(x = ~category,
+    #           ymin = ~`25%`,
+    #           ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>%
+   layout(showlegend = FALSE)
+
+ mds_hod <- base %>%
+  group_by(customer_id) %>%
+  add_trace(x = ~mds1.hod, y = ~ mds2.hod, text = ~customer_id)
+
+subplot(mds_hod, quantile_plot, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect", persistent = TRUE)
+#
+
+```
+```{r}
+# hod
+
+data_heatmap_hod <- data_heatmap_new %>% filter(gran == "hour_day") %>% mutate(category = as.integer(category))
+
+tx <- highlight_key(data_heatmap_hod, ~customer_id)
+
+ # initiate a plotly object
+base <- plot_ly(tx, color = I("black")) %>%
+   group_by(customer_id)
+
+ quantile_plot <- base %>%
+   group_by(customer_id, gran) %>%
+   add_lines(x = ~ category, y = ~`50%`, color = ("black")) %>%
+   #add_ribbons(x = ~category,
+    #           ymin = ~`25%`,
+    #           ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>%
+   layout(showlegend = FALSE)
+
+ mds_hod <- base %>%
+  group_by(customer_id) %>%
+  add_trace(x = ~mds1.moy, y = ~ mds2.moy, text = ~customer_id)
+
+subplot(mds_hod, quantile_plot, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect")
 ```
@@
 
-#10850478
+#10850478 8486921
 data_pick_one <- total_distance %>% filter(item1 %in% 
-                            c(8486921)) %>% group_by(item1) %>% arrange(total) %>% head(5) %>% 
+                            c(8454221)) %>% group_by(item1) %>% arrange(total) %>% head(5) %>% 
   mutate(item1 = as.integer(as.character(item1)),
@@
 
-#11013154
+#11013154  8181071 11048034
 data_pick_two <- total_distance %>% filter(item1 %in% 
@@
                             c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
+  mutate(item1 = as.integer(as.character(item1)),
+         item2 = as.integer(as.character(item2)))
+
+data_pick_three <- total_distance %>% filter(item1 %in%
+                            c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
+  mutate(item1 = as.integer(as.character(item1)),
+         item2 = as.integer(as.character(item2)))
+
+data_pick_four <- total_distance %>% filter(item1 %in%
+                            c(8653709)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
   mutate(item1 = as.integer(as.character(item1)),
@@
 
+data_pick_cust <- bind_rows(
+unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(),    unique(c(data_pick_two$item1,data_pick_two$item2)) %>% as_tibble(), c(11013154, 8495194, 8627007,10109182,10677705,8952846) %>% as_tibble(),unique(c(data_pick_four$item1,data_pick_four$item2)) %>% as_tibble(),
+.id = "design")
+
 # data_pick_cust <- bind_rows(
-# unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(),    unique(c(data_pick_two$item1,data_pick_two$item2)) %>% as_tibble(), c(11013154, 11045304, 11048034,11139841,11156027,11006388) %>% as_tibble(), .id = "design")
-#
+# unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(),    c(11013154, 8495194, 8627007,10109182,10677705,8952846) %>% as_tibble(), c(867773, 8240424, 8267102,8281608,8403312,8454235) %>% as_tibble(), .id = "design")
 
-data_pick_cust <- bind_rows(
-unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(), unique(c(data_pick_two$item1,data_pick_two$item2)) %>% as_tibble(), unique(c(data_pick_three$item1,data_pick_three$item2)) %>% as_tibble(), .id = "design")
+# c(11013154, 11045304, 11048034,11139841,11156027,11006388)
+
+# data_pick_cust <- bind_rows(
+# unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(), unique(c(data_pick_two$item1,data_pick_two$item2)) %>% as_tibble(), unique(c(data_pick_three$item1,data_pick_three$item2)) %>% as_tibble(), .id = "design")
 
@@
              ncol = 6) +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + theme_bw()
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+  scale_x_discrete(breaks = seq(1, 24, 3))
 
@@
 <!--  ) -->
+
+<!-- # mds_hod <- base %>% -->
+<!--   # group_by(customer_id) %>% -->
+<!--   # add_trace(x = ~mds1.hod, y = ~ mds2.hod, text = ~customer_id) -->
+
+<!-- # hod_wide <- hod %>% -->
+<!-- #   pivot_wider(names_from = item2, values_from = distance)%>% -->
+<!-- #   mutate(item1 = as.integer(as.character(item1))) -->
+<!-- #  -->
+<!-- # hod_wide$`8143599` <- NA -->
+<!-- #  -->
+<!-- # hod_wide <- bind_rows(hod_wide, c(item1=11590442)) -->
+<!-- #  -->
+<!-- # mds_data <- hod_wide[-1]  %>% mutate_all(~replace(., is.na(.), 0)) %>% as.dist() %>% clust_gran() -->
+<!-- #   -->

### 2021-09-29 supplementary/distance-computation/simulation-data/D3_I10_C1.Rmd
methodology subsection 1: finding significant granularities done

--- 2021-09-22 supplementary/distance-computation/simulation-data/D5_I5_C3.Rmd
@@
 t = 300
-n <- seq(0, n, 1)
+n <- seq(0, t, 1)
 g1 <- n %%2
@@
 g3 <- n %%5
-mu1= c(0, 2)
-mu2 = c(7, 5, 1)
-mu3 = c(1, 2, 5, 3, 1)
+mu21= c(2, 0, 0)
+mu22 = c(0, 2, 0)
+mu23 = c(0, 0, 2)
+mu3 = c(0, 0, 0, 0, 0)
 ```
@@
 # _Granularities design_
-```{r}
-design = c("design-1", "design-2", "design-3", "design4", "design5")
-g1 = c("no", "yes", "no", "no", "yes")
-g2 = c("no", "no", "yes", "no", "yes")
-g3 = c("no", "no", "no", "yes", "yes")
-table <- tibble(design, g1, g2, g3) %>% kable()
-table
-
-```
 
@@
 ```{r}
-generate_design <- function(t, mu1, mu2, mu3){
+generate_design <- function(t, mu21, mu22, mu23){
 
@@
 
-g1_dvary <- rep(mu1, length.out= length(t))
-g2_dvary <- rep(mu2, length.out= length(t))
-g3_dvary <- rep(mu3, length.out= length(t))
+g1_dvary <- rep(0, length.out= length(t))
+g21_dvary <- rep(mu21, length.out= length(t))
+g22_dvary <- rep(mu22, length.out= length(t))
+g23_dvary <- rep(mu23, length.out= length(t))
+g3_dvary <- rep(0, length.out= length(t))
 
 design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
-design2 = distributional::dist_normal(g1_dvary + g2_dnull + g3_dnull)
-design3 = distributional::dist_normal(g1_dnull + g2_dvary + g3_dnull)
-design4 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dvary)
-design5 = distributional::dist_normal(g1_dvary + g2_dvary + g3_dvary)
+design2 = distributional::dist_normal(g1_dnull + g21_dvary + g3_dnull)
+design3 = distributional::dist_normal(g1_dnull + g22_dvary + g3_dnull)
+design4 = distributional::dist_normal(g1_dnull + g23_dvary + g3_dnull)
 
@@
 design4 = distributional::generate(design4, times = 1) %>% unlist(),
-design5 = distributional::generate(design5, times = 1) %>% unlist()
+#design5 = distributional::generate(design5, times = 1) %>% unlist()
 ) %>% 
@@
 
-t = 300
-mu1= c(0, 2)
-mu2 = c(7, 5, 1)
-mu3 = c(1, 2, 5, 3, 1)
-
-data_bind <- generate_design(t, mu1, mu2, mu3)
+data_bind <- generate_design(t, mu21, mu22, mu23)
 
@@
 set.seed(seed_len[x])
-data_bind <- generate_design(t, mu1, mu2, mu3)
+data_bind <- generate_design(t, mu21, mu22, mu23)
 }) %>% bind_rows(.id = "seed_id") %>% 
@@
 
-groups = dist_mat%>% clust_gran(kopt = 5)
+groups = dist_mat %>% clust_gran(kopt = 4)
 pred_group = paste("design",groups$group,sep = "") %>% as.factor()
@@
 groups = dist_mat %>% 
-  clust_gran(kopt = 5)
+  clust_gran(kopt = 4)
 

### 2021-09-29 supplementary/distance-computation/simulation-data/D4_I10_C2.Rmd
methodology subsection 1: finding significant granularities done

--- 2021-09-29 supplementary/distance-computation/simulation-data/D3_I10_C1.Rmd
@@
 g3 <- n %%5
+mu11 = c(0, 5)
+mu12 = c(0, 3)
 mu21= c(2, 0, 0)
@@
 
-g1_dvary <- rep(0, length.out= length(t))
+g11_dvary <- rep(mu11, length.out= length(t))
+g12_dvary <- rep(mu12, length.out= length(t))
 g21_dvary <- rep(mu21, length.out= length(t))
@@
 design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
-design2 = distributional::dist_normal(g1_dnull + g21_dvary + g3_dnull)
-design3 = distributional::dist_normal(g1_dnull + g22_dvary + g3_dnull)
+design2 = distributional::dist_normal(g11_dvary + g21_dvary + g3_dnull)
+design3 = distributional::dist_normal(g12_dvary + g22_dvary + g3_dnull)
 design4 = distributional::dist_normal(g1_dnull + g23_dvary + g3_dnull)
@@
 ```{r}
-seed_len = seq(12345, 12349, 1)
+seed_len = seq(12345, 12399, 1)
 bind_data_iter <- map(seq_len(length(seed_len)), function(x){

### 2021-09-29 check_distance_matrix.R

@@ -1,0 +1,21 @@
+library(gravitas)
+library(tidyverse)
+library(gracsr)
+library(tsibble)
+
+sm <- smart_meter10 %>%
+filter(customer_id %in% c("10006704", "10017936","10006414", "10018250"))
+gran1 = "hour_day"
+gran2 = NULL
+response = "general_supply_kwh"
+dist_gran(sm, "hour_day")
+dist_gran(sm, "month_year")
+dist_mat <- sm %>% scale_gran(method = "robust") %>% dist_gran("hour_day")
+
+dist_mat %>% labels()
+
+dist_mat %>% clust_gran()
+
+
+groups  <-  dist_mat %>% hclust
+cutree(groups, k = 2) %>% labels

### 2021-09-29 main.Rmd
distance metric + inter category distances

--- 2021-09-29 main.Rmd
@@
 
-- _Find significant harmonies or cyclic granularities_:
+- _Choose significant harmonies or cyclic granularities_
 
-For each key variable, the raw smart meter data is indexed by time and comprises values of several variables of interest (measurement variables) at each time point.
+For each key variable, the raw smart meter data is a sequence that is indexed by time and comprises values of several measurement variables at each time point.
 This sequence, though, could be depicted in a variety of ways.
-One sequence could reflect the distribution of hourly consumption over a single day, while another could indicate consumption over a week or a year.
-The sequence of measured variables throughout each category of any cyclic granularity could be regarded as a single data generating process.
-It is beneficial to evaluate those cyclic granularities that are significant for the majority of customers or noteworthy in an electricity-behavior context, so that when the customers are grouped, we can expect to observe some interesting patterns across the categories of the cyclic granularities considered.
+A shuffling of the raw sequence could reflect the distribution of hourly consumption over a single day, while another could indicate consumption over a week or a year.
+These temporal deconstructions of a time period into units such as hour-of-day, work-day/weekend are called cyclic temporal granularities.
+They are useful for exploring repetitive patterns in time series data that get lost in the linear representation of time.
+However, it is advantageous to consider only those cyclic granularities across which there is a significant repetitive pattern for the majority of customers or noteworthy in an electricity-behavior context.
+In that case, when the customers are grouped, we can expect to observe some interesting patterns across the categories of the cyclic granularities considered.
 
-- _Distance metric_
+- _Individual or combined categories of cyclic granularities as DGP_
+
+The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+In our work, we are using $i$ as denoting a customer and the underlying variable as the electricity demand.
+So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
+In this work, instead of considering the probability distributions of the linear time series, we assume that the measured variables across different categories of any cyclic granularity are from data generating processes.
+Hence, we want to be able to cluster distributions of the form $f_{i,A,B \dots, {N_C}}(v)$, where $A, B$ represent the cyclic granularities under consideration such that $A = \{a_j: j=1, 2, \dots J\}$,  $B = \{b_k: k  = 1, 2, \dots K\}$ and so on.
+We consider individual each category of a cyclic granularity ($A$) or combination of categories for interaction of cyclic granularities (for e.g.
+$A*B$) to have a distribution.
+For example, let us consider we have two cyclic  granularities of interest, $A = {0, 1, 2, \dots, 23}$ representing hour-of-day and  $B = \{Mon, Tue, Wed, \dots, Sun\}$ representing day-of-week.
+each customer $i$ consist of a collection of probability distributions.
+In case individual granularities ($A$ or $B$) are considered there are  $J = 24$ distributions of the form $f_{i,j}(v)$ or $K = 7$ distributions of the form $f_{i,k}(v)$ for each customer $i$.
+In case of interaction,  $J*K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
+As a result, a distance between collections of these univariate probability distributions is required.
+Depending on the objective of the problem, there could be many approaches to considering such distances.
+This paper considers two approaches, which are explained in the next segment.
+
+- _Distance metrics_
+
+Considering each individual or combined categories of cyclic granularities as a data generating process lead to a collection of conditional distributions for each customer $i$.
+The (dis) similarity between each pair of observations should be obtained by combining the distances between these collections of conditional distributions such that the resulting metric is a distance metric, which could be fed into the clustering algorithm.
+Two types of distance metric is considered:
+
+<!-- The choice of distance measures is a critical step in clustering. It defines how the similarity of two elements (x, y) is calculated. -->
+
+**Inter-category distances**
+
+This distance matrix considers two objects to be similar if every category of an individual cyclic granularity or combination of categories for interacting cyclic granularities have similar distributions.
+In this study, the distribution for each category is characterized using deciles and the distances between distributions are computed by using the Jensen-Shannon distance, which is symmetric and hence could be used as a distance measure.
+The total distance between two elements $x$ and $y$ is then defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity A) or  $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
+When combining distances from individual cyclic granularities $A$ and $B$, $$S^{A, B}_{x, y} = S^A_{x,y}/J + S^B_{x,y}/K$$ is used, which could also be shown to be a distance metric easily.
+This is shown for cyclic granularity $A$ and $B$, but could be practically extended to more granularities.
+
+**Intra-category distances**
+Choose all significant granularities and compute wpd for all these granularities for all customers.
+Distance between customers is taken as the euclidean distances between them with the granularities being the variables and wpd being the value under each variable for which Euclidean distance needs to be measured.
+
+-
+
+Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
+Each household consists of a $J*K$ distributions one harmony.
+We compute the distributional difference
+between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$.
+$wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies.
+If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
+We aim to have pockets of households showing similar periodic behavior by conidering $wpd$ vlaues for different harmonies and some time series features.
+The features should also characterize probability distributions of different household.
 
@@
 - _Handling trend, seasonality, non-stationarity and auto-correlation_
+
 Trend and seasonality are common features of time series, and it is natural to characterize a time series by its degree of trend and seasonality.

### 2021-09-29 main.Rmd
included methodology data  structure

@@
 
+The proposed methodology aim to leverage the intrinsic temporal data structure hidden in time series data.
 The foundation of our method is unsupervised clustering algorithms based exclusively on time-series features.
@@
 However, since the data is unlabeled a priori, more information is required to corroborate this.
-
 This section presents the work flow of the methodology:
 
-- _Choose significant harmonies or cyclic granularities_
+- _Data preparation_
 
+@wang2020tsibble introduced the tidy "tsibble" data structure to support exploration and modeling of temporal data.
+A tsibble comprises an index, optional key(s), and measured variables.
+An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
+A linear granularity is a mapping of the index set to subsets of the time domain.
+For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.
 For each key variable, the raw smart meter data is a sequence that is indexed by time and comprises values of several measurement variables at each time point.
@@
 These temporal deconstructions of a time period into units such as hour-of-day, work-day/weekend are called cyclic temporal granularities.
-They are useful for exploring repetitive patterns in time series data that get lost in the linear representation of time.
-However, it is advantageous to consider only those cyclic granularities across which there is a significant repetitive pattern for the majority of customers or noteworthy in an electricity-behavior context.
+All cyclic granularities can be expressed in terms of the index set and could be augmented with the initial tsibble structure (index, key, measurements).
+It is worthwhile to note that the data structure changes while transporting from linear to cyclic scale of time as multiple observations of the measured variable would correspond to each category of the cyclic granularities.
+In this paper, quantiles are chosen to characterize the distributions for each category of the cyclic granularity.
+So, each category of a cyclic granularity corresponds to a list of numbers which is essentially few chosen quantiles of the multiple observations.
+
+- _Finding significant cyclic granularities or harmonies_
+
+\noindent These cyclic granularities are useful for exploring repetitive patterns in time series data that get lost in the linear representation of time.
+It is advantageous to consider only those cyclic granularities across which there is a significant repetitive pattern for the majority of customers or noteworthy in an electricity-behavior context.
 In that case, when the customers are grouped, we can expect to observe some interesting patterns across the categories of the cyclic granularities considered.
+[XXX reference 2nd chapter] proposes a way to select significant cyclic granularities and harmonies which is used for this paper.
 
@@
 In this study, the distribution for each category is characterized using deciles and the distances between distributions are computed by using the Jensen-Shannon distance, which is symmetric and hence could be used as a distance measure.
-The total distance between two elements $x$ and $y$ is then defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity A) or  $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
+
+\noindent The total distance between two elements $x$ and $y$ is then defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity A) or  $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
 When combining distances from individual cyclic granularities $A$ and $B$, $$S^{A, B}_{x, y} = S^A_{x,y}/J + S^B_{x,y}/K$$ is used, which could also be shown to be a distance metric easily.
@@
 **Intra-category distances**
-Choose all significant granularities and compute wpd for all these granularities for all customers.
-Distance between customers is taken as the euclidean distances between them with the granularities being the variables and wpd being the value under each variable for which Euclidean distance needs to be measured.
 
--
+Compute weighted pairwise distances ($wpd$) (XXX reference) for all considered granularities for all objects.
+$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive between consecutive categories normalized by appropriate parameters.
+A higher value of $wpd$ indicates that some interesting pattern is expected, whereas a lower value would indicate otherwise.
 
-Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
+\noindent Distance between objects is then taken as the euclidean distances between them with the granularities being the variables and $wpd$ being the value under each variable.
+Since Euclidean distance is chosen, the observations with high values of features ($wpd$ values) will be clustered together.
+The same holds true for observations with low values of features.
+Thus this distance matrix would be useful to group customers that have similar significance of patterns across different granularities.
+
+<!-- Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
 Each household consists of a $J*K$ distributions one harmony.
-We compute the distributional difference
-between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$.
-$wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies.
+We compute the distributional difference -->
+<!-- between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$. -->
+<!-- $wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies.
 If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
@@
 The features should also characterize probability distributions of different household.
-
-- _Clustering algorithm_
-
-- _Characterization of clusters_
-
-- A random sample  of the original data is taken for clustering analysis and includes missing and noisy observations (detailed description in Appendix)
-
-- All harmonies are computed for each customer in the sample.
-Cyclic granularities which are clashes for all customers in the sample are removed.
-
-- It is worth noting that a number of other solutions may be considered at the pre-processing stage of the method. We have considered a) Normal-Quantile Transform and b) Robust transformation.
-
-- Two methods are considered for computing dissimilarity between two customers.
-The first one involves computing according to one granularity is computed as the sum of the JS distances between distribution of all the categories of the granularity.
-When we consider more than one granularity, we consider the sum of the  average distances for all the granularity so that the combined metric is also a distance.
-
-- Given the scale of dissimilarity among the energy readings, the model chooses optimal number of clusters
-
-- Once clusters have been allocated, the groups are explored visually.
-
-- Results are reported and compared.
-
-Two methods are used for computing distances between subjects and then hierarchical clustering algorithm is used.
-
-The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
-In our work, we are using $i$ as denoting a customer and the underlying variable as the electricity demand.
-So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
-
-We want to cluster distributions of the form $f_{i,j,k}(v)$, where $i$ and $j$ denote
-
-Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
-Each household consists of a $J*K$ distributions one harmony.
-We compute the distributional difference
-between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$.
-$wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies.
-If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
-We aim to have pockets of households showing similar periodic behavior by conidering $wpd$ vlaues for different harmonies and some time series features.
-The features should also characterize probability distributions of different household.
-
-### Notations
-
-Consider an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
-In our work, we are using $i$ as denoting a household and the underlying variable as the electricity demand.
-Further consider a cyclic granularity of the form $B = \{ b_k: k = 1, 2, \dots, K\}$.
-Each customer consists of collection of probability distributions.
-
-So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
-We want to cluster distributions of the form $f_{i,j,k}(v)$, where $i$ and $j$ denote $i^{th}$ and $j^{th}$ customer respectively.
-
-a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
-
-### A single or pair of granularities together (change names)
-
-The methodology can be summarized in the following steps:
-
-- _Pre-processing step_
-
-Robust scaling method or NQT used for each customer.
-
-- _NQT_
-
-- _Treatment to outliers_
-
-- _Handling trend, seasonality, non-stationarity and auto-correlation_
-
-Trend and seasonality are common features of time series, and it is natural to characterize a time series by its degree of trend and seasonality.
-By considering the probability distributions through the use of $wpd_{norm_{s,t}}(A,B)$, these features of the time series are lost and hence there is no need to de-trend or de-seasonalize the data before performing the clustering algorithm.
-No need to exclude holiday or weekend patterns.
-
-### Many granularities together (change names)
-
-The methodology can be summarized in the following steps:
-
-1. Compute quantiles of distributions across each category of the cyclic granularity
-2. Compute JS distance between households for each each category of the cyclic granularity
-3. Total distance between households computed as sum of JS distances for all hours
-4. Cluster using this distance with hierarchical clustering algorithm (method "Ward.D")
-
-_Pro:_  
-- distance metric makes sense to group different shapes together  
-- simulation results look great on typical designs  
-_Cons:_  
-- Can only take one granularity at once  
-- Clustering a big blob of points together whereas the aim is to groups these big blob into smaller ones
-
-### Multiple-granularities
-
-_Description:_
-
-Choose all significant granularities and compute wpd for all these granularities for all customers.
-Distance between customers is taken as the euclidean distances between them with the granularities being the variables and wpd being the value under each variable for which Euclidean distance needs to be measured.
-_Pro:_  
-- Can only take many granularities at once
-- can apply variable selection PCP and other interesting clustering techniques
-- simulation results look great on typical designs
-- splitting the data into similar sized groups  
-_Cons:_  
-- distance metric does not make sense to split the data into similar shaped clusters
+-->
 
@@
 -->
+
+_Pre-processing steps_
+
+- Handling trend, seasonality, non-stationarity and auto-correlation: Trend and seasonality are fundamental characteristics of time series data, and it is reasonable to define a time series according to its degree of trend and seasonality.
+These characteristics of the time series are lost or handled independently by considering probability distributions ( trend is lost) across categories of cyclic granularities (by independently modeling all seasonal fluctuations), and so there is no need to de-trend or de-seasonalize the data before conducting the clustering method.
+There is no need to omit holiday or weekend patterns for similar reasons.
+
+- Data transformation: Robust scaling method is used before computing the Inter-category distances and NQT is built-in transformation used for computation of $wpd$, which forms the basis of Intra-category distances.
+
+- _Clustering algorithm_
+
+In the analysis of energy smart metre data, K-Means or hierarchical clustering are often employed.
+These are simple and effective techniques that work well in a range of scenarios.
+For clustering, both employ a distance measure, and the distance measure chosen has a major influence on the structure of the clusters.
+We employ agglomerative hierarchical clustering in conjunction with Ward's criteria (XXX reference).
+Individual entities with the highest similarity computed using the desired distance metrics are sequentially merged using agglomerative algorithms.
+We can possibly employ any clustering method that supports the given distance metric as input.
+
+- _Characterization of clusters_
+
+<!-- - A random sample  of the original data is taken for clustering analysis and includes missing and noisy observations (detailed description in Appendix) -->
+
+<!-- - All harmonies are computed for each customer in the sample. -->
+<!-- Cyclic granularities which are clashes for all customers in the sample are removed. -->
+
+<!-- - It is worth noting that a number of other solutions may be considered at the pre-processing stage of the method.
+We have considered a) Normal-Quantile Transform and b) Robust transformation.
+-->
+
+<!-- - Two methods are considered for computing dissimilarity between two customers.
+The first one involves computing according to one granularity is computed as the sum of the JS distances between distribution of all the categories of the granularity.
+When we consider more than one granularity, we consider the sum of the  average distances for all the granularity so that the combined metric is also a distance.
+-->
+
+<!-- - Given the scale of dissimilarity among the energy readings, the model chooses optimal number of clusters -->
+
+<!-- - Once clusters have been allocated, the groups are explored visually. -->
+
+<!-- - Results are reported and compared. -->
+
+<!-- Two methods are used for computing distances between subjects and then hierarchical clustering algorithm is used. -->
+
+<!-- The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+In our work, we are using $i$ as denoting a customer and the underlying variable as the electricity demand.
+So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
+-->
+
+<!-- We want to cluster distributions of the form $f_{i,j,k}(v)$, where $i$ and $j$ denote  -->
+
+<!-- Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
+Each household consists of a $J*K$ distributions one harmony.
+We compute the distributional difference -->
+<!-- between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$. -->
+<!-- $wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies.
+If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
+We aim to have pockets of households showing similar periodic behavior by conidering $wpd$ vlaues for different harmonies and some time series features.
+The features should also characterize probability distributions of different household.
+-->
+
+<!-- ### Notations -->
+
+<!-- Consider an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+In our work, we are using $i$ as denoting a household and the underlying variable as the electricity demand.
+Further consider a cyclic granularity of the form $B = \{ b_k: k = 1, 2, \dots, K\}$.
+Each customer consists of collection of probability distributions.
+-->
+
+<!-- So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
+We want to cluster distributions of the form $f_{i,j,k}(v)$, where $i$ and $j$ denote $i^{th}$ and $j^{th}$ customer respectively.
+-->
+
+<!-- a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.  -->
+
+<!-- ### A single or pair of granularities together (change names) -->
+
+<!-- The methodology can be summarized in the following steps: -->
+
+<!-- - _Pre-processing step_ -->
+
+<!-- Robust scaling method or NQT used for each customer. -->
+
+<!-- - _NQT_ -->
+
+<!-- - _Treatment to outliers_ -->
+
+<!-- ### Many granularities together (change names) -->
+
+<!-- The methodology can be summarized in the following steps: -->
+
+<!-- 1. Compute quantiles of distributions across each category of the cyclic granularity -->
+<!-- 2. Compute JS distance between households for each each category of the cyclic granularity -->
+<!-- 3. Total distance between households computed as sum of JS distances for all hours -->
+<!-- 4. Cluster using this distance with hierarchical clustering algorithm (method "Ward.D") -->
+
+<!-- _Pro:_   -->
+<!-- - distance metric makes sense to group different shapes together   -->
+<!-- - simulation results look great on typical designs   -->
+<!-- _Cons:_   -->
+<!-- - Can only take one granularity at once   -->
+<!-- - Clustering a big blob of points together whereas the aim is to groups these big blob into smaller ones   -->
+
+<!-- ### Multiple-granularities -->
+
+<!-- _Description:_   -->
+
+<!-- Choose all significant granularities and compute wpd for all these granularities for all customers.
+Distance between customers is taken as the euclidean distances between them with the granularities being the variables and wpd being the value under each variable for which Euclidean distance needs to be measured.
+-->
+<!-- _Pro:_   -->
+<!-- - Can only take many granularities at once -->
+<!-- - can apply variable selection PCP and other interesting clustering techniques -->
+<!-- - simulation results look great on typical designs -->
+<!-- - splitting the data into similar sized groups   -->
+<!-- _Cons:_   -->
+<!-- - distance metric does not make sense to split the data into similar shaped clusters  -->
+
+# Simulations
+
+# Application
 
@@
  <!-- - Handling autocorrelation: Autocorrelation in the time series is likely to get removed from considering probability distributions and cyclic granularities in the clustering algorithm. -->
+
+Generally most of the indices used for internal clustering validation combine compactness and separation measures.
 
@@
 
-Generally most of the indices used for internal clustering validation combine compactness and separation measures.
-
-## Results
-
-## Software implementation
-
-The implementation for our framework is available in the R package [`gracsr`](https://github.com/Sayani07/gracsr) for ease of use in other applications.
-
-# Results
-
 ## Clustering results for 100 customers
@@
 -->
+
+## Software implementation
+
+The implementation for our framework is available in the R package [`gracsr`](https://github.com/Sayani07/gracsr) for ease of use in other applications.
 

### 2021-09-29 main.Rmd
adding Characterization of clusters

@@
 
-_Pre-processing steps_
+- _Pre-processing steps_
 
-- Handling trend, seasonality, non-stationarity and auto-correlation: Trend and seasonality are fundamental characteristics of time series data, and it is reasonable to define a time series according to its degree of trend and seasonality.
+    **Handling trend, seasonality, non-stationarity and auto-correlation:** Trend and seasonality are fundamental characteristics of time series data, and it is reasonable to define a time series according to its degree of trend and seasonality.
 These characteristics of the time series are lost or handled independently by considering probability distributions ( trend is lost) across categories of cyclic granularities (by independently modeling all seasonal fluctuations), and so there is no need to de-trend or de-seasonalize the data before conducting the clustering method.
@@
 
-- Data transformation: Robust scaling method is used before computing the Inter-category distances and NQT is built-in transformation used for computation of $wpd$, which forms the basis of Intra-category distances.
+    **Data transformation:** Robust scaling method is used before computing the Inter-category distances and NQT is built-in transformation used for computation of $wpd$, which forms the basis of Intra-category distances.
 
@@
 - _Characterization of clusters_
+
+Depending on the distance measure utilized for the study, the cluster characterization technique will differ.
+Clusters that utilise intra-category distances are characterised using multi-dimensional scaling and parallel coordinate displays.
+For inter-category distances, the distribution across major granularities may be presented to ensure that the goal of similar shapes within clusters and distinct shapes across clusters is met.
+This technique may potentially make advantage of multi-dimensional scaling.
+
+\noindent Multidimensional scaling (MDS) (XXX reference) refers to a family of methods that analyse a matrix of distances or dissimilarities to provide a representation of the data points in a reduced-dimension space.
+There are many kinds of MDS, but they all solve the same fundamental issue: Given a $n*n$ matrix of dissimilarities and a distance measure, identify a configuration of $n$ points $x_1, x_2, \dots, x_n$ in the reduced dimension space $R^q$ ($q<p$) where the distance between the points is near to the dissimilarity between the points.
+All techniques must determine the coordinates of the points as well as the space dimension, $q$.
+Metric and nonmetric MDS are the two main kinds of MDS.
+Metric MDS methods presume a functional connection between the interpoint distances and the supplied dissimilarities and assume that the data are quantitative.
+We use metric MDS.
+
+\noindent Parallel coordinate plots (XXX reference) Parallel coordinates have been extensively used to display high-dimensional and multivariate data, allowing for the detection of patterns within the data via visual grouping.
 

### 2021-09-30 Simulations.Rmd
added wpd analysis

@@
   mutate(total = hod)
+  #mutate(total = wkndwday/2 + moy/12 + hod/24)
 
+# 8454221 10420689
 data_pick_one <- total_distance %>% filter(item1 %in% 
@@
 
-data_pick_three <- total_distance %>% filter(item1 %in%
-                            c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
+# data_pick_three <- total_distance %>% filter(item1 %in%
+#                             c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
+#   mutate(item1 = as.integer(as.character(item1)),
+#          item2 = as.integer(as.character(item2)))
 
@@
 
+# c(11013154, 8495194, 8627007,10109182,10677705,8952846)
+
 data_pick_cust <- bind_rows(
 unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(),    unique(c(data_pick_two$item1,data_pick_two$item2)) %>% as_tibble(), 
-c(11013154, 8495194, 8627007,10109182,10677705,8952846) %>% as_tibble(),
+c(11013154, 9345642, 8328008, 8636035,8627007,8454235) %>% as_tibble(),
 unique(c(data_pick_four$item1,data_pick_four$item2)) %>% as_tibble(),
@@
 ```{r heatmap-hod}
-sm_hod <- gravitas::create_gran(data_pick, "hour_day")
+data_heatmap_hod <- quantile_gran(data_pick, "hour_day", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
+  left_join(data_pick_cust, by = c("customer_id"="value"))
 
-sm_hod_list <- sm_hod %>% 
-  as_tibble() %>% 
-  select(customer_id, hour_day, general_supply_kwh) %>% 
-  pivot_wider(names_from = hour_day,
-              values_from = general_supply_kwh)  %>% 
-  left_join(data_pick_cust, by = c("customer_id" = "value"))
-
-ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, ncol(sm_hod_list))]))
-nrow_sm <- seq_len(nrow(sm_hod_list))
-
-sm_hod_quantiles_cat <- map(nrow_sm, function(x){
-  map(ncol_sm, function(y){
-   cell <- sm_hod_list%>% 
-     dplyr::filter(customer_id == customer_id[x]) %>% 
-     select(-c(1, ncol(sm_hod_list))) %>% 
-     extract(y) %>% 
-     unlist()
-   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
-})  %>% bind_rows(.id = "categories_serial_id")
-}) %>% bind_rows(.id = "customer_serial_id")
-
-ref_cat <- names(sm_hod_list)[-c(1, ncol(sm_hod_list))] %>% as_tibble() %>% set_names("category") %>%  
-  mutate(categories_serial_id = row_number())
-
-ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
-    mutate(customer_serial_id = row_number())
-
-sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
-  mutate(customer_serial_id = as.integer(customer_serial_id),
-         categories_serial_id = as.integer(categories_serial_id)) %>%
-  left_join(ref_cat, by = "categories_serial_id") %>% 
-  left_join(ref_cust, by = "customer_serial_id") %>% 
-  select(customer_serial_id, categories_serial_id, category, customer_id, everything())
-
-data_heatmap_hod <- sm_quantiles_ref %>% 
- left_join(data_pick %>% distinct(customer_id)) %>% 
-  left_join(data_pick_cust, by = c("customer_id" = "value")) %>% 
-  mutate(categories_serial_id = as.factor(categories_serial_id))
+data_heatmap_hod$category <- factor(data_heatmap_hod$category, levels = 0:23)
 
 data_heatmap_hod %>% 
-  ggplot(aes(x = categories_serial_id)) + 
+  ggplot(aes(x = category)) + 
   geom_ribbon(aes(ymin = `25%`, 
@@
              ncol = 6) +
-    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
   scale_x_discrete(breaks = seq(1, 24, 3))
@@
 ```{r heatmap-moy}
-sm_hod <- gravitas::create_gran(data_pick, "month_year")
+data_heatmap_moy <- quantile_gran(data_pick, "month_year", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
+  left_join(data_pick_cust, by = c("customer_id"="value"))
 
-sm_hod_list <- sm_hod %>% 
-  as_tibble() %>% 
-  select(customer_id, month_year, general_supply_kwh) %>% 
-  pivot_wider(names_from = month_year,
-              values_from = general_supply_kwh)  %>% 
-  left_join(data_pick_cust, by = c("customer_id" = "value"))
-
-ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, ncol(sm_hod_list))]))
-nrow_sm <- seq_len(nrow(sm_hod_list))
-
-sm_hod_quantiles_cat <- map(nrow_sm, function(x){
-  map(ncol_sm, function(y){
-   cell <- sm_hod_list%>% 
-     dplyr::filter(customer_id == customer_id[x]) %>% 
-     select(-c(1, ncol(sm_hod_list))) %>% 
-     extract(y) %>% 
-     unlist()
-   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
-})  %>% bind_rows(.id = "categories_serial_id")
-}) %>% bind_rows(.id = "customer_serial_id")
-
-ref_cat <- names(sm_hod_list)[-c(1, ncol(sm_hod_list))] %>% as_tibble() %>% set_names("category") %>%  
-  mutate(categories_serial_id = row_number())
-
-ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
-    mutate(customer_serial_id = row_number())
-
-sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
-  mutate(customer_serial_id = as.integer(customer_serial_id),
-         categories_serial_id = as.integer(categories_serial_id)) %>%
-  left_join(ref_cat, by = "categories_serial_id") %>% 
-  left_join(ref_cust, by = "customer_serial_id") %>% 
-  select(customer_serial_id, categories_serial_id, category, customer_id, everything())
-
-data_heatmap_moy <- sm_quantiles_ref %>% 
- left_join(data_pick %>% distinct(customer_id)) %>% 
-  left_join(data_pick_cust, by = c("customer_id" = "value")) %>% 
-  mutate(categories_serial_id = as.factor(categories_serial_id))
+data_heatmap_moy$category <- factor(data_heatmap_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
 
 data_heatmap_moy %>% 
-  ggplot(aes(x = categories_serial_id)) + 
+  ggplot(aes(x = category)) + 
   geom_ribbon(aes(ymin = `25%`, 
@@
              ncol = 6) +
-    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
-  scale_x_discrete(breaks = seq(1, 12, 2))
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
 
@@
 ```{r heatmap-wkndwday}
-sm_hod <- gravitas::create_gran(data_pick, "wknd_wday")
+data_heatmap_wkndwday <- quantile_gran(data_pick, "wknd_wday", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
+  left_join(data_pick_cust, by = c("customer_id"="value"))
 
-sm_hod_list <- sm_hod %>% 
-  as_tibble() %>% 
-  select(customer_id, wknd_wday, general_supply_kwh) %>% 
-  pivot_wider(names_from = wknd_wday,
-              values_from = general_supply_kwh)  %>% 
-  left_join(data_pick_cust, by = c("customer_id" = "value"))
-
-ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, ncol(sm_hod_list))]))
-nrow_sm <- seq_len(nrow(sm_hod_list))
-
-sm_hod_quantiles_cat <- map(nrow_sm, function(x){
-  map(ncol_sm, function(y){
-   cell <- sm_hod_list%>% 
-     dplyr::filter(customer_id == customer_id[x]) %>% 
-     select(-c(1, ncol(sm_hod_list))) %>% 
-     extract(y) %>% 
-     unlist()
-   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
-})  %>% bind_rows(.id = "categories_serial_id")
-}) %>% bind_rows(.id = "customer_serial_id")
-
-ref_cat <- names(sm_hod_list)[-c(1, ncol(sm_hod_list))] %>% as_tibble() %>% set_names("category") %>%  
-  mutate(categories_serial_id = row_number())
-
-ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
-    mutate(customer_serial_id = row_number())
-
-sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
-  mutate(customer_serial_id = as.integer(customer_serial_id),
-         categories_serial_id = as.integer(categories_serial_id)) %>%
-  left_join(ref_cat, by = "categories_serial_id") %>% 
-  left_join(ref_cust, by = "customer_serial_id") %>% 
-  select(customer_serial_id, categories_serial_id, category, customer_id, everything())
-
-data_heatmap_wkndwday <- sm_quantiles_ref %>% 
- left_join(data_pick %>% distinct(customer_id)) %>% 
-  left_join(data_pick_cust, by = c("customer_id" = "value")) %>% 
-  mutate(categories_serial_id = as.factor(categories_serial_id))
+data_heatmap_wkndwday$category <- factor(data_heatmap_wkndwday$category, levels = c("Weekday", "Weekend"))
 
 data_heatmap_wkndwday %>% 
-  ggplot(aes(x = categories_serial_id)) + 
+  ggplot(aes(x = category)) + 
   geom_ribbon(aes(ymin = `25%`, 
@@
              ncol = 6) +
-    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+```
 
-```
+# Clustering by method 1
 
@@
   filter(item2 %in% data_pick_cust$value)
-
-# filtered_distance <- bind_rows(data_pick_one, )
-#   
-#   data_pick_one
 
@@
 
-cluster_result <- bind_cols(id = id, group = group) %>% mutate(id =  as.numeric(id))
+cluster_result <- bind_cols(id = id, group = group)
 
@@
 
-cluster_result %>% filter(group==1)
-cluster_result %>% filter(group==2)
-cluster_result %>% filter(group==3)
-cluster_result %>% filter(group==4)
+# cluster_result %>% filter(group==1)
+# cluster_result %>% filter(group==2)
+# cluster_result %>% filter(group==3)
+# cluster_result %>% filter(group==4)
 
@@
 data_heatmap_hod %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
   left_join(cluster_result, by = c("customer_id" = "id")) %>% 
-  ggplot(aes(x = categories_serial_id)) + 
+  ggplot(aes(x = category)) + 
   geom_ribbon(aes(ymin = `25%`, 
@@
              ncol = 6) +
-    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")
+    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+```
+
+```{r data-heatmap-hod-group}
+
+data_group <- data_pick %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id"))
+
+data_heatmap_hod_group <- quantile_gran(data_group,
+                                  gran1="hour_day",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
+
+data_heatmap_hod_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group))) +
+  geom_line(aes(y = `50%`, group=group)) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow = 4) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
 ```
@@
 data_heatmap_moy %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
   left_join(cluster_result, by = c("customer_id" = "id")) %>% 
-  ggplot(aes(x = categories_serial_id)) + 
+  ggplot(aes(x = category)) + 
   geom_ribbon(aes(ymin = `25%`, 
@@
              ncol = 6) +
-    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")
+    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+```
+
+```{r data-heatmap-moy-group}
+
+data_group <- data_pick %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id"))
+
+data_heatmap_moy_group <- quantile_gran(data_group,
+                                  gran1="month_year",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+data_heatmap_moy_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group))) +
+  geom_line(aes(y = `50%`, group=group)) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow = 4) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
 ```
@@
 data_heatmap_wkndwday %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
   left_join(cluster_result, by = c("customer_id" = "id")) %>% 
-  ggplot(aes(x = categories_serial_id)) + 
+  ggplot(aes(x = category)) + 
   geom_ribbon(aes(ymin = `25%`, 
@@
              ncol = 6) +
-    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")
+    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("wknd_wday") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
 
@@
 
-```{r first-try, eval = FALSE}
+```{r data-heatmap-wkndwday-group}
 
-sm_hod_list_new <- sm_hod_list %>% left_join(cluster_result, by = c("customer_id" = "id"))
+data_group <- data_pick %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id"))
 
-sm_hod_list_ncol <- sm_hod_list_new[-c(1, ncol(sm_hod_list_new),(ncol(sm_hod_list_new)-1))]
+data_heatmap_wkndwday_group <- quantile_gran(data_group,
+                                  gran1="wknd_wday",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
 
-ncol_sm <- unique(cluster_result$group)
-nrow_sm <- seq_len(nrow(sm_hod_list_new))
+data_heatmap_wkndwday_group$category <- factor(data_heatmap_wkndwday_group$category, levels = c("Weekday", "Weekend"))
 
-sm_hod_quantiles_cat <- map(nrow_sm, function(x){
-  map(ncol_sm, function(y){
-   cell <- sm_hod_list_new%>% 
-     dplyr::filter(group == group[x]) %>% 
-     #select(-c(1, ncol(sm_hod_list_new),(ncol(sm_hod_list_new)-1))) %>% 
-     extract(y) %>% 
-     unlist()
-   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
-})  %>% bind_rows(.id = "group_id")
-}) %>% bind_rows(.id = "customer_serial_id")
-
-ref_cat <- names(sm_hod_list_ncol) %>% as_tibble() %>% set_names("category") %>%  
-  mutate(categories_serial_id = row_number())
-
-ref_cust <- sm_hod_list_new$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
-    mutate(customer_serial_id = row_number())
-
-sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
-  mutate(customer_serial_id = as.integer(customer_serial_id),
-         categories_serial_id = as.integer(categories_serial_id)) %>%
-  left_join(ref_cat, by = "categories_serial_id") %>% 
-  left_join(ref_cust, by = "customer_serial_id") %>% 
-  select(customer_serial_id, categories_serial_id, category, customer_id, everything())
-
-data_heatmap_hod <- sm_quantiles_ref %>% 
- left_join(data_pick %>% distinct(customer_id)) %>% 
-  left_join(data_pick_cust, by = c("customer_id" = "value")) %>% 
-  mutate(categories_serial_id = as.factor(categories_serial_id))
-
-data_heatmap_hod %>% 
-  ggplot(aes(x = categories_serial_id)) + 
+data_heatmap_wkndwday_group %>% 
+  ggplot(aes(x = category)) + 
   geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=customer_id), fill = "lightblue") +
-  geom_line(aes(y = `50%`, group=customer_id)) +
+                  ymax = `75%`, group=group, fill = as.factor(group))) +
+  geom_line(aes(y = `50%`, group=group)) +
   facet_wrap(~group, 
@@
              labeller = "label_value",
-             ncol = 6) +
-    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
-  scale_x_discrete(breaks = seq(1, 24, 3))
+             ncol = 4) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("wknd_wday") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
 
 ```
+
+# Clustering by method 2
+
+```{r}
+
+elec_600_wpd <- read_rds(here::here("data/algo2-cust600-wpd-rawdata.rds"))
+
+elec_pick <- elec_600_wpd %>% 
+  filter(customer_id %in% data_pick_cust$value)
+
+elec_pick_wide <- elec_pick %>% pivot_wider(-c(1, 2), names_from = "x_variable", values_from = wpd)
+
+# robust_scale <-  function(x){
+#   (x - stats::median(x))/(stats::IQR(x))
+# }
+
+#scaled_var <- apply(elec_pick_wide[-c(1)], 2, robust_scale) %>% as_tibble()
+
+scaled_var <- elec_pick_wide
+
+f <- elec_pick_wide[-1] %>% dist()
+
+group <- f%>% hclust (method = "ward.D") %>% cutree(k=4)
+
+cluster_result_wpd <- bind_cols(id = elec_pick_wide$customer_id, group = group)
+
+mds  = cmdscale(f)
+rownames(mds) = id
+colnames(mds) = c("mds1", "mds2")
+
+ggplot(data = mds %>% as_tibble() %>% mutate(group = cluster_result_wpd$group), 
+       aes(x = mds1, y = mds2, color = as.factor(group))) +
+  geom_point() +
+  geom_text(aes(label = id))
+
+data_pcp <- scaled_var %>% 
+  #bind_cols(customer_id =  elec_pick_wide$customer_id) %>%
+  left_join(cluster_result_wpd , by = c("customer_id" = "id")) %>% 
+  select(customer_id, group, everything()) %>% 
+  mutate(group = as.factor(group))
+
+parcoord <- GGally::ggparcoord(data_pcp ,
+                   columns = 3:ncol(data_pcp),
+                   groupColumn = "group",
+                   showPoints = FALSE, 
+                   alphaLines = 0.8,
+                   order = "anyClass",
+                   scale = "globalminmax"
+) + ggplot2::theme_bw() +
+  scale_color_brewer(palette = "Dark2") + 
+  ggplot2::theme(
+    plot.title = ggplot2::element_text(size=10)
+  )+
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
+  theme(legend.position = "bottom") +
+  xlab("") +
+  ylab("wpd")
+
+```
+
+```{r}
+
+data_heatmap_hod %>% 
+  #mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result_wpd, by = c("customer_id" = "id")) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=customer_id), fill = "lightblue") +
+  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+```
+
+```{r data-heatmap-hod-group-wpd}
+
+data_group_wpd <- data_pick %>% 
+  #mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result_wpd, by = c("customer_id" = "id"))
+
+data_heatmap_hod_group <- quantile_gran(data_group_wpd,
+                                  gran1="hour_day",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
+
+data_heatmap_hod_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group))) +
+  geom_line(aes(y = `50%`, group=group)) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow = 4) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+```
+
+```{r data-heatmap-moy-wpd}
+
+data_heatmap_moy %>% 
+  #mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result_wpd, by = c("customer_id" = "id")) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=customer_id), fill = "lightblue") +
+  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+```
+
+```{r data-heatmap-moy-group-wpd}
+
+data_heatmap_moy_group <- quantile_gran(data_group_wpd,
+                                  gran1="month_year",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+data_heatmap_moy_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group))) +
+  geom_line(aes(y = `50%`, group=group)) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow = 4) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+```
+
+```{r data-heatmap-wkndwday-wpd}
+data_heatmap_wkndwday %>% 
+  #mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result_wpd, by = c("customer_id" = "id")) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=customer_id), fill = "lightblue") +
+  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("wknd_wday") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+```
+
+```{r data-heatmap-wkndwday-group-wpd}
+
+data_heatmap_wkndwday_group <- quantile_gran(data_group_wpd,
+                                  gran1="wknd_wday",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_wkndwday_group$category <- factor(data_heatmap_wkndwday_group$category, levels = c("Weekday", "Weekend"))
+
+data_heatmap_wkndwday_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group))) +
+  geom_line(aes(y = `50%`, group=group)) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 4) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("wknd_wday") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+```

### 2021-09-30 main.Rmd
added wpd analysis

@@
 library(knitr)
+library(patchwork)
 ```
@@
 
-# Preliminary exploratoration
+<!-- # Preliminary exploratoration -->
 
-## Electricity demand data
+<!-- ## Electricity demand data -->
 
@@
 
-### Data source
+<!-- ### Data source -->
 
-The entire data is procured from CSIRO.
+<!-- The entire data is procured from CSIRO.
 A subset of this data is also available from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
 It consists of the following data sets.
-_1. CustomerData:_ 78720 customers with 62 variables about them 
-_2. EUDMData:_ 300 billion half-hourly consumption level data  
-_3. OffersData:_ Method of contact to customer to join SGSC customer trial, either door-to-door (D2D) or via Telesales  
-_4. PEResponseData:_ Peak Events response customer wise  
-_5. PETimesData:_ Peak Events time stamps
+-->
+<!-- _1. CustomerData:_ 78720 customers with 62 variables about them  -->
+<!-- _2. EUDMData:_ 300 billion half-hourly consumption level data   -->
+<!-- _3. OffersData:_ Method of contact to customer to join SGSC customer trial, either door-to-door (D2D) or via Telesales   -->
+<!-- _4. PEResponseData:_ Peak Events response customer wise   -->
+<!-- _5. PETimesData:_ Peak Events time stamps   -->
 
-Only _CustomerData_ and _EUDMData_ are relevant for the clustering goals of this paper.
+<!-- Only _CustomerData_ and _EUDMData_ are relevant for the clustering goals of this paper.
 _EUDMData_ contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
-`CustomerData` provides demographic data for 78,720 customers with information on their Local Government Area amongst others.
+`CustomerData` provides demographic data for 78,720 customers with information about their Local Government Area amongst others.
+-->
 
-### Raw data
+<!-- ## LGA and weather data -->
 
-### Missing Data
-
-Electricity usage for some customers may become unavailable due to power outage or not recording their usage properly, thus resulting in implicit missing values in the database.
-It is interesting to explore where missing-ness occurs or if there is a relationship between the underlying missing patterns.
-We use the R package `tsibble` to do this.
-
-```{r, out.width="100%"}
-include_graphics("figs/missing-data.png")
-```
-
-13735 customers in elec_ts
-8685 customers in elec_nogap
-5050 customers in count_na_df
-
-Then is the graph of missing observations even interesting.
-You can show two graphs, one to show that missingness do not have a pattern
-another to show even if no missing, they start and end at different times. (A sample of 50 customers).
-
-## LGA and weather data
-
-Since the smart meters have been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths
-for each household.
+<!-- Since the smart meters have been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths -->
+<!-- for each household.
 Since, general supply is available for only 13,735 customers, we will restrict ourselves to look at the LGA information for these customers only.
 We find that there are only 26 LGA that is covered for these customers.
+-->
 
-_Weather data_  
-This data is obtained through [Australian Government Bureau of Meteorology](BOM)(http://www.bom.gov.au/) and provides hourly data for nearest weather stations for all the LGAs
+<!-- _Weather data_   -->
+<!-- This data is obtained through [Australian Government Bureau of Meteorology](BOM)(http://www.bom.gov.au/) and provides hourly data for nearest weather stations for all the LGAs -->
 
-This section familiarizes the 13,000 SGSC households through visualization and provide a detailed layout of data structures (like missing observations/number of customers/number of observations) and also the external data that needs to be utilized for validating the clustering process.
+<!-- This section familiarizes the 13,000 SGSC households through visualization and provide a detailed layout of data structures (like missing observations/number of customers/number of observations) and also the external data that needs to be utilized for validating the clustering process.
 The [ABS TableBuilder](https://www.abs.gov.au/websitedbs/censushome.nsf/home/tablebuilder) has census data from 2011 and 2016.
@@
 Potentially, there could be lag effects of weather on energy usage which should be considered.
-
-\noindent A dataset of 100 SGSC homes has been used to lay out the structure to be used for analyzing the big dataset.
-The smaller dataset contains half-hourly kwh values form 2012 to 2014 and has asynchronous time series distributed evenly over the observation period (Figure \ref{fig:elec-raw}), similar to the bigger data set.
-Figure \ref{fig:count-gaps} can be used to interpret missingness in the data, where the customers are arranged from maximum to minimum missing.
-It looks like data is most missing before 2013 and for a particular date in 2014.
-
-```{r miss-data}
-empty_as_na <- function(x){
-    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
-    ifelse(as.character(x)!="", x, NA)
-}
-
-```
-
-#NEW
-
-The raw smart meter data is indexed by time for each key variable and includes values of different variables of interest (measurement variables) at each time point.
-However, this series could be represented in several possible ways.
-One sequence could represent distribution of hourly consumption across a single day, while another could track consumption across days of a week or months of a year.
-Depending on the cyclic granularity considered, each sequence could be thought of as a different data generating processes.
-
-We can then model different data generation processes depending on our level of aggregation and time scale.
+-->
 
@@
 
-# Simulations
+# Validation
+
+To test clustering algorithm solutions against ground truth cluster assignments, we generate simulations to represent important problems that could be encountered in electricity data contexts.
+The code to generate these simulations are included in the Supplementary paper.
+
+## Data designs
+
+When we move from the linear to the cyclic world of temporal granularities, we can see patterns across different categories of the granularities which gets lost in the linear representation.
+Populations are modeled by a collection of these cyclic granularities.
+Each cyclic granularity might or might not have patterns across its categories.
+Each cluster is characterized by similar patterns across one or more of these cyclic granularities.
+
+A small example is given to setup the problem.
+
+We constructed the simulation parameters to represent common patterns in electricity data.
+50, 200 and 500 time series were chosen for different simulation designs with different granularities and patterns changing across different granularities.
+The data type is fixed to be "continuous".We generated independent replications of all combinations of the simulation parameters.
+
+Consider a continuous time series variable $y$ of length $T$ indexed by ${0, 1, \dots T-1}$.
+Three circular granularities $g1$, $g2$ and $g3$ are considered with 2, 3 and 5 levels respectively.
+Categories of g1, g2 and g3 are represented by ${0,1}$, ${0, 1, 2}$ and ${0, 1, 2, 3, 4}$.
+These categories could be integers or some more meaningful labels.
+For example, the granularity "day-of-week" could be either represented by ${0, 1, 2, \dots, 6}$ or ${Mon, Tue, \dots, Sun}$.
+
+<!-- Consider the case where distribution of $y$ would vary across levels of $g1$ and g2$ but will not vary across levels of $g3$ as described in Figure .Figure \ref shows four designs where $g3$ acts as a nuisance variable, and $g1$ and $g2$ varies in both  Design 2 and 3, and $g2$ varies in all designs.
+-->
+
+Consider a case where distribution of $y$ would vary across levels of $g2$
+for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
+Figure \ref{#fig:plot-linear} shows the linear and cyclic representation of $y$.
+The first panel shows raw plot of $y$ in a linear scale and the second panel shows distribution of $y$ across cyclic granularities namely $g1$, $g2$ and $g3$.
+As could be seen from the plots, it is impossible to decipher from the raw time plot that the time series variable shows such pattern across different granularities.
+
+```{r change-var, echo=FALSE}
+t = 300
+n <- seq(0, t, 1)
+g1 <- n %%2
+g2 <- n %%3
+g3 <- n %%5
+mu11 = c(0, 5)
+mu12 = c(3, 0)
+mu21= c(2, 0, 0)
+mu22 = c(0, 2, 0)
+mu23 = c(0, 0, 2)
+mu3 = c(0, 0, 0, 0, 0)
+```
+
+```{r generate-design} 
+generate_design <- function(t, mu21, mu22, mu23){
+
+t <- seq(0, t, 1)
+g1 <- t %%2
+g2 <- t %%3
+g3 <- t %%5
+
+# null design
+g1_dnull <- rep( rep(0, each = length(unique(g1))), length.out= length(t))
+g2_dnull <- rep( rep(0, each = length(unique(g2))), length.out= length(t))
+g3_dnull <- rep( rep(0, each = length(unique(g3))), length.out= length(t))
+
+# mean changing across categories in varying ways
+
+g11_dvary <- rep(mu11, length.out= length(t))
+g12_dvary <- rep(mu12, length.out= length(t))
+g21_dvary <- rep(mu21, length.out= length(t))
+g22_dvary <- rep(mu22, length.out= length(t))
+g23_dvary <- rep(mu23, length.out= length(t))
+g3_dvary <- rep(0, length.out= length(t))
+
+design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
+design2 = distributional::dist_normal(g11_dvary + g21_dvary + g3_dnull)
+design3 = distributional::dist_normal(g12_dvary + g22_dvary + g3_dnull)
+design4 = distributional::dist_normal(g1_dnull + g23_dvary + g3_dnull)
+
+data_bind <- tibble::tibble(
+index = t,
+g1 = g1,
+g2 = g2,
+g3 = g3,
+design1 = distributional::generate(design1, times = 1) %>% unlist(),
+design2 = distributional::generate(design2, times = 1) %>% unlist(),
+design3 = distributional::generate(design3, times = 1) %>% unlist(),
+design4 = distributional::generate(design4, times = 1) %>% unlist(),
+#design5 = distributional::generate(design5, times = 1) %>% unlist()
+) %>% 
+  pivot_longer(-c(1, 2, 3, 4), names_to = "design", values_to = "sim_data")
+
+data_bind
+}
+
+data_bind <- generate_design(t, mu21, mu22, mu23)
+
+#data_bind
+
+```
+
+```{r plot-linear , out.width="100%", fig.cap="The linear and cyclic representation of the time series variable y.
+It is not possible to comprehend these patterns across cyclic granularities g1, g2 and g3 or group similar series just by looking at the linear plots."}
+
+# plot_linear_data <- function(data){
+# ggplot(data,
+#              aes(x = index, y = sim_data)) + 
+#   geom_line() +
+#   xlab("index")+
+#   theme_bw() 
+# }
+
+ggplot(data_bind,
+             aes(x = index, y = sim_data)) + 
+  geom_line() +
+  xlab("index")+
+  theme_bw() +
+  facet_wrap(~design, scales = "free_y",ncol =1)+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+p2 <- ggplot(data_bind,
+             aes(x = as.factor(g1), y = sim_data)) + 
+  geom_boxplot(alpha =0.5) + xlab("g1") + 
+  theme_bw() +
+  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 1, color = "blue") +
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+p3 <- ggplot(data_bind, aes(x = as.factor(g2), y = sim_data)) + geom_boxplot(alpha =0.5) + xlab("g2") + theme_bw() +
+  facet_wrap(~design, scales = "free_y",ncol = 1)+ stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 1, color = "blue") + ylab("")+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+p4 <- ggplot(data_bind, aes(x = as.factor(g3), y = sim_data)) + geom_boxplot(alpha =0.5) +
+  xlab("g3") + theme_bw()+
+  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 1, color = "blue")+ ylab("")+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+(p2 + p3 + p4) * theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)),
+    panel.spacing =unit(0, "lines")) * theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+# 
+# 
+# plot_cyclic_data( data_bind %>% filter(design=="design2"))
+# plot_data( data_bind %>% filter(design=="design3"))
+# plot_data( data_bind %>% filter(design=="design4"))
+# plot_data( data_bind %>% filter(design=="design5"))
+```
+
+A subset of many possible designs are shown in Figure \ref{fig: plot-linear }.
+For the parameter space (XXX unique  combinations shown in table YYY), 100, 500 independent replications of all possible combination of simulation parameters were generated.
+The clustering methodologies were run all these unique combinations and subsets of these to verify if the methodologies work as expected.
+
+| Granularity type                                                   	| # Significant 	| # Replications 	|
+|--------------------------------------------------------------------	|---------------	|----------------	|
+| **Individual**  <br><br># obs: 300, 500, 2000  <br># clusters: 6/7 	| 1/2/3         	| 25, 100, 200   	|
+| **Interaction**  <br><br># obs: 500, 2000  <br># clusters: 4       	| 1/2           	| 25, 100, 200   	|
+
+## Results
+
+- confusion matrix could be used for showing results if proper labeling is used
+
+- write about features that we have spiked into the data set
+- write about you incorporated noise
+- What is the additional structure you can incorporate that will lead to failing of method1 and method2?
+- And both gives the same result? Basically say when method 1 works better than method 2 and vice versa!
+
+-
 
@@
 
-The methodology is again applied to the 100 households and some preliminary results are presented.
-Figure \ref{fig:mds} show the multidimensional scaling of the 100 households with colors representing groups presented by hierarchical clustering method.
-The clusters have very different size and are far apart with no overlapping.
-They should be distinct from each other in terms of their periodic patterns.
-Figure \ref{fig:cyclic-cluster} shows the cyclic presentation of time where it could be viewed in details how these clusters are different with respect to the periodic pattern (day-of-week, hour-of-day).
-Distribution of energy demand across day-of-week and hour-of-day for three groups clustered through the proposed methodology is shown.
-For group-1, the morning peak looks very sharp, for group-2 the morning and evening peaks are rounded and spread over few hours, for group-3 peaks are less rounded.
-They differ in their relationship between 75th and 90th percentile.
-Also, the 75th and 90th percentile of group 2 are quite close, implying the behavior of this group is more regular.
+## Data source
+
+The entire data is procured from CSIRO.
+A subset of this data is also available from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
+It consists of the following data sets.
+_1. CustomerData:_ 78720 customers with 62 variables about them 
+_2. EUDMData:_ 300 billion half-hourly consumption level data  
+_3. OffersData:_ Method of contact to customer to join SGSC customer trial, either door-to-door (D2D) or via Telesales  
+_4. PEResponseData:_ Peak Events response customer wise  
+_5. PETimesData:_ Peak Events time stamps
+
+Only _CustomerData_ and _EUDMData_ are relevant for the clustering goals of this paper.
+_EUDMData_ contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
+`CustomerData` provides demographic data for 78,720 customers most of which are missing and not utilized for the purpose of this paper.
+To meet the requirements for anonymity preservation, the energy patterns could not be identified at the individual level, but rather by the geographical location of their residence information about their Local Government Area.
+
+### Raw data
+
+### Missing Data
+
+Electricity usage for some customers may become unavailable due to power outage or not recording their usage properly, thus resulting in implicit missing values in the database.
+It is interesting to explore where missing-ness occurs or if there is a relationship between the underlying missing patterns.
+We use the R package `tsibble` to do this.
+
+```{r, out.width="100%"}
+include_graphics("figs/missing-data.png")
+```
+
+- if there is any systematic missing patterns in the data
+- this missing plot can go in the supplementary
+- how to add missing values (should be added in data pre-processing)
+- instance learning
+- types of summary techniques to use (
+generally multivariate means and sd are used, I can't use that in a time series context, you can show across different granularity?
+within-group sum of squares and between-group sum of squares
+
+)
+
+13735 customers in elec_ts
+8685 customers in elec_nogap
+5050 customers in count_na_df
+
+Then is the graph of missing observations even interesting.
+You can show two graphs, one to show that missingness do not have a pattern
+another to show even if no missing, they start and end at different times. (A sample of 50 customers).
+
+\noindent A dataset of 100 SGSC homes has been used to lay out the structure to be used for analyzing the big dataset.
+The smaller dataset contains half-hourly kwh values form 2012 to 2014 and has asynchronous time series distributed evenly over the observation period (Figure \ref{fig:elec-raw}), similar to the bigger data set.
+Figure \ref{fig:count-gaps} can be used to interpret missingness in the data, where the customers are arranged from maximum to minimum missing.
+It looks like data is most missing before 2013 and for a particular date in 2014.
+
+```{r miss-data}
+empty_as_na <- function(x){
+    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
+    ifelse(as.character(x)!="", x, NA)
+}
+
+```
 
@@
 
+<!-- At this stage, we need to define the aim of clustering as there could be various aims of clustering like between-cluster separation, within cluster homogeneity: low distances, within-cluster homogeneous distributional shape, good representation of data by centroids, little loss of information, high density without cluster gaps, uniform cluster sizes, stability and others.
+Finally, how distinct they are and how can we summarize the main features of the cluster would be discussed here.
+-->
+
+### Cluster validation
+
+**Internal cluster validation** uses the internal information of the clustering process to evaluate the goodness of a clustering structure without reference to external information.
+It can be also used for estimating the number of clusters and the appropriate clustering algorithm without any external data.
+
 Generally most of the indices used for internal clustering validation combine compactness and separation measures.
 
-_Distinction, repeatability, and robustness metrics_
+<!-- _Distinction, repeatability, and robustness metrics_ -->
 
@@
 The cluster validation indexes like average silhouette width (ASW) is to be employed here to check how homogeneous these clusters are.
-At this stage, we need to define the aim of clustering as there could be various aims of clustering like between-cluster separation, within cluster homogeneity: low distances, within-cluster homogeneous distributional shape, good representation of data by centroids, little loss of information, high density without cluster gaps, uniform cluster sizes, stability and others.
-Finally, how distinct they are and how can we summarize the main features of the cluster would be discussed here.
-
-_Cluster validation_
-
-**Internal cluster validation** uses the internal information of the clustering process to evaluate the goodness of a clustering structure without reference to external information.
-It can be also used for estimating the number of clusters and the appropriate clustering algorithm without any external data.
 
@@
 
-## Combining findings with external data
-
-The robustness of this clustering method is provided through practical explanation of the formed clusters, visualizing how they relate to any weather, socio-economic or geographical conditions. 
+<!-- The robustness of this clustering method is provided through practical explanation of the formed clusters, visualizing how they relate to any weather, socio-economic or geographical conditions.
+-->
 <!-- Furthermore, we ensure that a cross-validation algorithm is employed leading to robust clusters.  -->

### 2021-09-30 supplementary/distance-computation/real-data/04handpick-from356-mds.Rmd
added wpd analysis

@@
 
-subplot(mds_total, quantile_plot, nrows = 2, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect")
+subplot(mds_total, quantile_plot, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect")
 
@@
 
-subplot(mds_hod, quantile_plot, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect", persistent = TRUE)
+subplot(mds_hod, quantile_plot, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect")
 #

### 2021-09-30 supplementary/distance-computation/real-data/04handpick-from356cust.Rmd
added wpd analysis

@@
 total_distance <- distance %>% 
-  mutate(total = wkndwday + moy + hod)
+  mutate(total = wkndwday/2 + moy/12 + hod/24)
 
@@
 
-sm_hod <- gravitas::create_gran(elec_356_raw, "hour_day")
+# sm_hod <- gravitas::create_gran(elec_356_raw, "hour_day")
+# 
+# sm_hod_list <- sm_hod %>% 
+#   as_tibble() %>% 
+#   select(customer_id, hour_day, general_supply_kwh) %>% 
+#   pivot_wider(names_from = hour_day,
+#               values_from = general_supply_kwh) 
+# 
+# ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, ncol(sm_hod_list))]))
+# nrow_sm <- seq_len(nrow(sm_hod_list))
+# 
+# 
+# sm_hod_quantiles_cat <- map(nrow_sm, function(x){
+#   map(ncol_sm, function(y){
+#    cell <- sm_hod_list%>% 
+#      dplyr::filter(customer_id == customer_id[x]) %>% 
+#      select(-c(1, ncol(sm_hod_list))) %>% 
+#      extract(y) %>% 
+#      unlist()
+#    quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
+# })  %>% bind_rows(.id = "categories_serial_id")
+# }) %>% bind_rows(.id = "customer_serial_id")
+# 
+# 
+# ref_cat <- names(sm_hod_list)[-c(1, ncol(sm_hod_list))] %>% as_tibble() %>% set_names("category") %>%  
+#   mutate(categories_serial_id = row_number())
+# 
+# 
+# ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
+#     mutate(customer_serial_id = row_number())
+#   
+# sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
+#   mutate(customer_serial_id = as.integer(customer_serial_id),
+#          categories_serial_id = as.integer(categories_serial_id)) %>%
+#   left_join(ref_cat, by = "categories_serial_id") %>% 
+#   left_join(ref_cust, by = "customer_serial_id") %>% 
+#   select(customer_serial_id, categories_serial_id, category, customer_id, everything())
 
-sm_hod_list <- sm_hod %>% 
-  as_tibble() %>% 
-  select(customer_id, hour_day, general_supply_kwh) %>% 
-  pivot_wider(names_from = hour_day,
-              values_from = general_supply_kwh)
+quantile_hod <- quantile_gran(elec_356_raw, gran1 = "hour_day", quantile_prob_val = quantile_prob_graph)
 
-ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, ncol(sm_hod_list))]))
-nrow_sm <- seq_len(nrow(sm_hod_list))
+quantile_moy <- quantile_gran(elec_356_raw, gran1 = "month_year", quantile_prob_val = quantile_prob_graph)
 
-sm_hod_quantiles_cat <- map(nrow_sm, function(x){
-  map(ncol_sm, function(y){
-   cell <- sm_hod_list%>% 
-     dplyr::filter(customer_id == customer_id[x]) %>% 
-     select(-c(1, ncol(sm_hod_list))) %>% 
-     extract(y) %>% 
-     unlist()
-   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
-})  %>% bind_rows(.id = "categories_serial_id")
-}) %>% bind_rows(.id = "customer_serial_id")
+quantile_wkndwday <- quantile_gran(elec_356_raw, gran1 = "wknd_wday", quantile_prob_val = quantile_prob_graph)
 
-ref_cat <- names(sm_hod_list)[-c(1, ncol(sm_hod_list))] %>% as_tibble() %>% set_names("category") %>%  
-  mutate(categories_serial_id = row_number())
+quantile_data <- bind_rows(quantile_hod, quantile_moy, quantile_wkndwday)
 
-ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
-    mutate(customer_serial_id = row_number())
+elec_pick_wide <- elec_pick %>% pivot_wider(c(3:5), names_from = "x_variable", values_from = wpd)
 
-sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
-  mutate(customer_serial_id = as.integer(customer_serial_id),
-         categories_serial_id = as.integer(categories_serial_id)) %>%
-  left_join(ref_cat, by = "categories_serial_id") %>% 
-  left_join(ref_cust, by = "customer_serial_id") %>% 
-  select(customer_serial_id, categories_serial_id, category, customer_id, everything())
-
-data_heatmap <- sm_quantiles_ref %>% 
- left_join(elec_pick %>% distinct(customer_id)) %>% 
-  mutate(categories_serial_id = as.factor(categories_serial_id)) %>% 
+data_heatmap <- quantile_data %>% 
+ left_join(elec_pick_wide) %>% 
   mutate(customer_id = as.character(customer_id))
 
+# %>% 
+#   mutate(categories_serial_id = as.factor(categories_serial_id)) %>% 
+#   mutate(customer_id = as.character(customer_id))
+
 data_heatmap_new <- data_heatmap %>%
-left_join(mds_all, by = "customer_id")
+left_join(mds_all, by = "customer_id") %>%  pivot_wider( names_from = quantiles, values_from = quantiles_values)
 
@@
 
-subplot(mds_total, mds_wpd, quantile_plot, nrows = 2, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect")
+subplot(mds_total, quantile_plot, nrows = 2, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect")
 
+# hod
+
+data_heatmap_hod <- data_heatmap_new %>% filter(gran == "hour_day") %>% 
+  mutate(category = as.integer(category))
+tx <- highlight_key(data_heatmap_hod, ~customer_id)
+
+ # initiate a plotly object
+base <- plot_ly(tx, color = I("black")) %>%
+   group_by(customer_id)
+
+ quantile_plot <- base %>%
+   group_by(customer_id, gran) %>%
+   add_lines(x = ~ category, y = ~`50%`, color = ("black")) %>%
+   #add_ribbons(x = ~category,
+    #           ymin = ~`25%`,
+    #           ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>%
+   layout(showlegend = FALSE)
+
+ mds_hod <- base %>%
+  group_by(customer_id) %>%
+  add_trace(x = ~mds1.hod, y = ~ mds2.hod, text = ~customer_id)
+
+subplot(mds_hod, quantile_plot, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect")
+#
+
+```
+```{r}
+# hod
+
+data_heatmap_hod <- data_heatmap_new %>% filter(gran == "hour_day") %>% mutate(category = as.integer(category))
+
+tx <- highlight_key(data_heatmap_hod, ~customer_id)
+
+ # initiate a plotly object
+base <- plot_ly(tx, color = I("black")) %>%
+   group_by(customer_id)
+
+ quantile_plot <- base %>%
+   group_by(customer_id, gran) %>%
+   add_lines(x = ~ category, y = ~`50%`, color = ("black")) %>%
+   #add_ribbons(x = ~category,
+    #           ymin = ~`25%`,
+    #           ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>%
+   layout(showlegend = FALSE)
+
+ mds_hod <- base %>%
+  group_by(customer_id) %>%
+  add_trace(x = ~mds1.moy, y = ~ mds2.moy, text = ~customer_id)
+
+subplot(mds_hod, quantile_plot, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect", persistent = TRUE)
 ```
@@
 
-sm_hod <- gravitas::create_gran(data_pick, "hour_day")
+data_heatmap <- quantile_gran(data_pick, "hour_day", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
+  left_join(data_pick_cust, by = c("customer_id"="value"))
 
-sm_hod_list <- sm_hod %>% 
-  as_tibble() %>% 
-  select(customer_id, hour_day, general_supply_kwh) %>% 
-  pivot_wider(names_from = hour_day,
-              values_from = general_supply_kwh)  %>% 
-  left_join(data_pick_cust, by = c("customer_id" = "value"))
+# data_heatmap <- sm_quantiles_ref %>% 
+#  left_join(data_pick %>% distinct(customer_id)) %>% 
+#   left_join(data_pick_cust, by = c("customer_id" = "value")) %>% 
+#   mutate(categories_serial_id = as.factor(categories_serial_id))
 
-ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, ncol(sm_hod_list))]))
-nrow_sm <- seq_len(nrow(sm_hod_list))
-
-sm_hod_quantiles_cat <- map(nrow_sm, function(x){
-  map(ncol_sm, function(y){
-   cell <- sm_hod_list%>% 
-     dplyr::filter(customer_id == customer_id[x]) %>% 
-     select(-c(1, ncol(sm_hod_list))) %>% 
-     extract(y) %>% 
-     unlist()
-   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
-})  %>% bind_rows(.id = "categories_serial_id")
-}) %>% bind_rows(.id = "customer_serial_id")
-
-ref_cat <- names(sm_hod_list)[-c(1, ncol(sm_hod_list))] %>% as_tibble() %>% set_names("category") %>%  
-  mutate(categories_serial_id = row_number())
-
-ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
-    mutate(customer_serial_id = row_number())
-
-sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
-  mutate(customer_serial_id = as.integer(customer_serial_id),
-         categories_serial_id = as.integer(categories_serial_id)) %>%
-  left_join(ref_cat, by = "categories_serial_id") %>% 
-  left_join(ref_cust, by = "customer_serial_id") %>% 
-  select(customer_serial_id, categories_serial_id, category, customer_id, everything())
-
-data_heatmap <- sm_quantiles_ref %>% 
- left_join(data_pick %>% distinct(customer_id)) %>% 
-  left_join(data_pick_cust, by = c("customer_id" = "value")) %>% 
-  mutate(categories_serial_id = as.factor(categories_serial_id))
+data_heatmap$category <- factor(data_heatmap$category, levels = 0:23)
 
 data_heatmap %>% 
-  ggplot(aes(x = categories_serial_id)) + 
+  ggplot(aes(x = category)) + 
   geom_ribbon(aes(ymin = `25%`, 

### 2021-10-01 main.Rmd
adding material in the simulation section

@@
 knitr::read_chunk(here('script/smart_meter.R'))
+```
+
+```{r mytheme}
+theme_validation <- function() {
+  theme_bw() +
+      theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 1)),
+    plot.margin = margin(0, 0, 0, 0, "cm"),
+    axis.title.y = element_blank(),
+    #axis.text.x = element_blank(),
+  axis.text.y = element_blank(),
+  axis.ticks = element_blank(),
+      legend.position = "bottom"
+    )
+}
+
 ```
@@
 
-To test clustering algorithm solutions against ground truth cluster assignments, we generate simulations to represent important problems that could be encountered in electricity data contexts.
-The code to generate these simulations are included in the Supplementary paper.
+To evaluate clustering approaches, we create data designs that replicate prototype behaviours that might be seen in electricity data contexts.
+We spiked several features in the data to see where one method works better than the other and where they might give us the same outcome or the effect of missing data and trends on the proposed methods.
+A continuous measured variable $y$ of length $T$ indexed by ${0, 1, \dots T-1}$.
+Three circular granularities $g1$, $g2$ and $g3$ are considered with $2$, $3$ and $5$ levels respectively.
+Categories of $g1$, $g2$ and $g3$ are represented by ${g10,g11}$, ${g20, g21, g22}$ and ${g30, g31, g32, g32, g34}$.
+These categories could be integers or some more meaningful labels.
+For example, the granularity "day-of-week" could be either represented by ${0, 1, 2, \dots, 6}$ or ${Mon, Tue, \dots, Sun}$.
+Each categories of $g1$, $g2$ and $g3$ could be assumed to be from same distribution, few from one and others from separate distributions, or all from different distributions, resulting in distinct data designs.
+We created independent replications $(R = \{50, 200, 500\})$ of all data designs to see if our proposed clustering approaches can detect distinct designs in various groups for small, medium and large number of series.
+A sample size of $T=\{300, 500, 2000\}$ is used in all designs to test small, medium and large sized series.
+For all data designs, the data type is set to "continuous," and a gaussian and non-Gaussian setup is assumed.
+The code for creating these designs can be found in the Supplementary section (link to github repo).
+The results for $T=300$ and $R=50$.
+The rest of the results could be found in the supplementary paper.
 
@@
 
-When we move from the linear to the cyclic world of temporal granularities, we can see patterns across different categories of the granularities which gets lost in the linear representation.
+**Three signifiant granularities**
+
+Consider a case where all the three granularities $g1$, $g2$ and $g3$ would be responsible for making the designs distinct.
+That would mean, the pattern for each of $g1$, $g2$ and $g3$ will change for at least one design.
+We consider a situation with all the null cases corresponding to no difference in distribution across categories, that is, all categories follow the same distribution N(0,1).
+
+| granularity 	|  Alternate design 	|
+|---	|---	|
+| g1 	| g10 ~ N(0, 1), g11 ~ N(2, 1) 	|
+| g2 	| g21 ~ N(2, 1), g22 ~ N(1, 1), g23 ~ N(0, 1) 	|
+| g3 	| g31 ~ N(0, 1), g32 ~ N(1, 1), <br><br>g33 ~ N(2, 1), g34 ~ N(1, 1), g35 ~ N(0, 1) 	|
+
+A combination of these null and alternate distributions are considered when building the designs as in the table \ref{tab:tab-design}.
+Figure \ref{#fig:plot-linear-3change} shows the linear and cyclic representation of $y$ under these five designs.
+
+```{r tab-design}
+design = c("design-1", "design-2", "design-3", "design-4", "design5")
+g1 = c("null", "vary", "null", "null", "vary")
+g2 = c("null", "null", "vary", "null", "vary")
+g3 = c("null", "null", "null", "vary", "vary")
+table <- tibble(design, g1, g2, g3) %>% kable()
+table
+
+```
+
+```{r generate-design-3change}
+generate_design <- function(t, mu1, mu2, mu3){
+
+t <- seq(0, t, 1)
+g1 <- t %%2
+g2 <- t %%3
+g3 <- t %%5
+
+# null design
+g1_dnull <- rep( rep(0, each = length(unique(g1))), length.out= length(t))
+g2_dnull <- rep( rep(0, each = length(unique(g2))), length.out= length(t))
+g3_dnull <- rep( rep(0, each = length(unique(g3))), length.out= length(t))
+
+# mean changing across categories in varying ways
+
+g1_dvary <- rep(mu1, length.out= length(t))
+g2_dvary <- rep(mu2, length.out= length(t))
+g3_dvary <- rep(mu3, length.out= length(t))
+
+design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
+design2 = distributional::dist_normal(g1_dvary + g2_dnull + g3_dnull)
+design3 = distributional::dist_normal(g1_dnull + g2_dvary + g3_dnull)
+design4 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dvary)
+design5 = distributional::dist_normal(g1_dvary + g2_dvary + g3_dvary)
+
+data_bind <- tibble::tibble(
+index = t,
+g1 = g1,
+g2 = g2,
+g3 = g3,
+design1 = distributional::generate(design1, times = 1) %>% unlist(),
+design2 = distributional::generate(design2, times = 1) %>% unlist(),
+design3 = distributional::generate(design3, times = 1) %>% unlist(),
+design4 = distributional::generate(design4, times = 1) %>% unlist(),
+design5 = distributional::generate(design5, times = 1) %>% unlist()
+) %>% 
+  pivot_longer(-c(1, 2, 3, 4), names_to = "design", values_to = "sim_data")
+
+data_bind
+}
+
+t = 300
+mu1= c(0, 2)
+mu2 = c(2, 1, 0)
+mu3 = c(0, 1, 2, 1, 0)
+
+data_bind <- generate_design(t, mu1, mu2, mu3)
+
+```
+
+```{r plot-linear-change , out.width="100%", fig.cap="The linear and cyclic representation of the time series variable y.
+It is not possible to comprehend these patterns across cyclic granularities g1, g2 and g3 or group similar series just by looking at the linear plots."}
+
+# plot_linear_data <- function(data){
+# ggplot(data,
+#              aes(x = index, y = sim_data)) + 
+#   geom_line() +
+#   xlab("index")+
+#   theme_bw() 
+# }
+
+p1 <- ggplot(data_bind,
+             aes(x = index, y = sim_data)) + 
+  geom_line() +
+  xlab("index")+
+  facet_wrap(~design, scales = "free_y",ncol =1) +
+  theme_validation()
+
+p2 <- ggplot(data_bind,
+             aes(x = as.factor(g1), y = sim_data)) + 
+  geom_boxplot(alpha =0.5) + xlab("g1") + 
+  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.8, color = "blue") +
+  theme_validation()
+
+p3 <- ggplot(data_bind, aes(x = as.factor(g2), y = sim_data)) + geom_boxplot(alpha =0.5) + xlab("g2") + theme_bw() +
+  facet_wrap(~design, scales = "free_y",ncol = 1)+ stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.8, color = "blue") + ylab("")+
+  theme_validation()
+
+p4 <- ggplot(data_bind, aes(x = as.factor(g3), y = sim_data)) + geom_boxplot(alpha =0.5) +
+  xlab("g3") + theme_bw()+
+  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.8, colosr = "blue")+ ylab("")+
+  theme_validation()
+
+(p1 + (p2 + p3 + p4)) *
+  theme_validation() + plot_layout(widths = c(2, 1))
+# 
+# plot_cyclic_data( data_bind %>% filter(design=="design2"))
+# plot_data( data_bind %>% filter(design=="design3"))
+# plot_data( data_bind %>% filter(design=="design4"))
+# plot_data( data_bind %>% filter(design=="design5"))
+```
+
+**Two significant granularities**
+
+Consider a case where distribution of $y$ would vary across levels of $g2$ for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
+So $g3$ is not responsible for distinguishing across designs.
+Figure \ref{#fig:plot-linear} shows the linear and cyclic representation of $y$.
+The first panel shows raw plot of $y$ in a linear scale and the second panel shows distribution of $y$ across cyclic granularities namely $g1$, $g2$ and $g3$.
+As could be seen from the plots, it is impossible to decipher from the raw time plot that the time series variable shows such pattern across different granularities.
+
+**One signifiant granularity**
+
+Consider a case where distribution of $y$ would vary across levels of $g2$
+for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs. So $g3$ is not responsible for distinguishing across designs.
+
+**Missing data**
+
+**Trend**
+
+**Outliers**
+
+observations:
+
+designs:
+
+<!-- When we move from the linear to the cyclic world of temporal granularities, we can see patterns across different categories of the granularities which gets lost in the linear representation.
 Populations are modeled by a collection of these cyclic granularities.
@@
 Each cluster is characterized by similar patterns across one or more of these cyclic granularities.
+-->
 
@@
 These categories could be integers or some more meaningful labels.
-For example, the granularity "day-of-week" could be either represented by ${0, 1, 2, \dots, 6}$ or ${Mon, Tue, \dots, Sun}$.
+For example, the granularity "day-of-week" could be either represented by $\{0, 1, 2, \dots, 6\}$ or $\{Mon, Tue, \dots, Sun\}$.
 
@@
 
-ggplot(data_bind,
+p1 <- ggplot(data_bind,
              aes(x = index, y = sim_data)) + 
@@
   xlab("index")+
-  theme_bw() +
-  facet_wrap(~design, scales = "free_y",ncol =1)+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+  facet_wrap(~design, scales = "free_y",ncol =1) +
+  theme_validation()
 
@@
   geom_boxplot(alpha =0.5) + xlab("g1") + 
-  theme_bw() +
   facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
@@
     geom = 'line',
-    aes(group = 1), size = 1, color = "blue") +
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+    aes(group = 1), size = 0.8, color = "blue") +
+  theme_validation()
 
@@
     geom = 'line',
-    aes(group = 1), size = 1, color = "blue") + ylab("")+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+    aes(group = 1), size = 0.8, color = "blue") + ylab("")+
+  theme_validation()
 
@@
     geom = 'line',
-    aes(group = 1), size = 1, color = "blue")+ ylab("")+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+    aes(group = 1), size = 0.8, color = "blue")+ ylab("")+
+  theme_validation()
 
-(p2 + p3 + p4) * theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)),
-    panel.spacing =unit(0, "lines")) * theme(plot.margin = margin(0, 0, 0, 0, "cm") )
-
-# 
+(p1 + (p2 + p3 + p4)) *
+  theme_validation() + plot_layout(widths = c(2, 1))
 # 

### 2021-10-04 main.Rmd
updating simulations

@@
 Three circular granularities $g1$, $g2$ and $g3$ are considered with $2$, $3$ and $5$ levels respectively.
-Categories of $g1$, $g2$ and $g3$ are represented by ${g10,g11}$, ${g20, g21, g22}$ and ${g30, g31, g32, g32, g34}$.
+Categories of $g1$, $g2$ and $g3$ are represented by ${g10,g11}$, ${g20, g21, g22}$ and ${g30, g31, g32, g33, g34}$.
 These categories could be integers or some more meaningful labels.
@@
 A sample size of $T=\{300, 500, 2000\}$ is used in all designs to test small, medium and large sized series.
-For all data designs, the data type is set to "continuous," and a gaussian and non-Gaussian setup is assumed.
+For all data designs, the data type is set to "continuous," and a gaussian setup for errors is assumed.
 The code for creating these designs can be found in the Supplementary section (link to github repo).
@@
 
+## Data generating processes
+
+An ARMA (p,q) process is used to generate series, where $p$ and $q$ are selected at random such that the series is stationary.
+The various designs on $g1$, $g2$, and $g3$ are introduced by adding matching designs to this series' innovations.
+The innovations are considered to have a normal distribution, although they follow the same pattern as the designs.
+To eliminate the effect of starting values, the first 500 observations in each series are discarded.
+
 ## Data designs
@@
 
-Consider a case where all the three granularities $g1$, $g2$ and $g3$ would be responsible for making the designs distinct.
+<!-- Consider a case where all the three granularities $g1$, $g2$ and $g3$ would be responsible for making the designs distinct.
 That would mean, the pattern for each of $g1$, $g2$ and $g3$ will change for at least one design.
 We consider a situation with all the null cases corresponding to no difference in distribution across categories, that is, all categories follow the same distribution N(0,1).
+-->
+
+Consider the scenario when all three granularities $g1$, $g2$, and $g3$ are responsible for distinguishing the designs.
+That means that for at least one among the to-be-grouped designs, the pattern for each of $g1$, $g2$, and $g3$ will change.
+We consider null instances to represent no variation in distribution across categories, i.e., all categories have the same distribution N(0,1).
 
@@
 
-A small example is given to setup the problem.
-
-We constructed the simulation parameters to represent common patterns in electricity data.
+<!-- A small example is given to setup the problem. -->
+
+<!-- We constructed the simulation parameters to represent common patterns in electricity data.
 50, 200 and 500 time series were chosen for different simulation designs with different granularities and patterns changing across different granularities.
-The data type is fixed to be "continuous".We generated independent replications of all combinations of the simulation parameters.
-
-Consider a continuous time series variable $y$ of length $T$ indexed by ${0, 1, \dots T-1}$.
+-->
+<!-- The data type is fixed to be "continuous".We generated independent replications of all combinations of the simulation parameters. -->
+
+<!-- Consider a continuous time series variable $y$ of length $T$ indexed by ${0, 1, \dots T-1}$.
 Three circular granularities $g1$, $g2$ and $g3$ are considered with 2, 3 and 5 levels respectively.
@@
 For example, the granularity "day-of-week" could be either represented by $\{0, 1, 2, \dots, 6\}$ or $\{Mon, Tue, \dots, Sun\}$.
+-->
 
@@
 
-Consider a case where distribution of $y$ would vary across levels of $g2$
-for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
+<!-- Consider a case where distribution of $y$ would vary across levels of $g2$ -->
+<!-- for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
 Figure \ref{#fig:plot-linear} shows the linear and cyclic representation of $y$.
@@
 As could be seen from the plots, it is impossible to decipher from the raw time plot that the time series variable shows such pattern across different granularities.
+-->
 
@@
 
-- confusion matrix could be used for showing results if proper labeling is used
-
-- write about features that we have spiked into the data set
-- write about you incorporated noise
-- What is the additional structure you can incorporate that will lead to failing of method1 and method2?
-- And both gives the same result? Basically say when method 1 works better than method 2 and vice versa!
-
--
+All the methods were fitted to each data designs and results are reported through confusion matrices.
+With increasing difference between categories, it gets easier for the methods to correctly distinguish the designs.
+For difference=1, the performances are pretty bad for js-robust methods and wpd method.
+The performance starts getting better with increasing difference and get worse with increasing number of replications.
+Length of series do not show to have any effect on the performance of the methods.
+It does not depend on if time series is ar or arma.
+
+<!-- - confusion matrix could be used for showing results if proper labeling is used -->
+
+<!-- - write about features that we have spiked into the data set -->
+<!-- - write about you incorporated noise -->
+<!-- - What is the additional structure you can incorporate that will lead to failing of method1 and method2? -->
+<!-- - And both gives the same result? Basically say when method 1 works better than method 2 and vice versa! -->
+
+<!-- - -->
 

### 2021-10-04 supplementary/distance-computation/simulation-data/D5_I5_C3.Rmd
updating simulations

@@
 g3 <- n %%5
+# mu1= c(0, 2)
+# mu2 = c(7, 5, 1)
+# mu3 = c(1, 2, 5, 3, 1)
+
 mu1= c(0, 2)
-mu2 = c(7, 5, 1)
-mu3 = c(1, 2, 5, 3, 1)
+mu2 = c(5, 1, 0)
+mu3 = c(0, 3, 7, 1, 0)
+
 ```
@@
 t = 300
+# mu1= c(0, 2)
+# mu2 = c(7, 5, 1)
+# mu3 = c(1, 2, 5, 3, 1)
+
 mu1= c(0, 2)
-mu2 = c(7, 5, 1)
-mu3 = c(1, 2, 5, 3, 1)
+mu2 = c(5, 1, 0)
+mu3 = c(0, 3, 7, 1, 0)
 
 data_bind <- generate_design(t, mu1, mu2, mu3)
-
 data_bind

### 2021-10-04 2gran_change_4D.R

@@ -1,3 +1,5 @@
+# Rscripts for simulation chapter 3, patterns across 3 grans changing atleast across one design amongst 5 designs
+
 library(conflicted)
 library(tidyverse)
 library(gracsr)
@@ -8,18 +10,18 @@ library(here)
 
 
 # fix parameters
- 
+
 # niter
 #seed_len = seq(12345, 12350, 1)
 # number of iterations corresponding to
-niter <- c(50, 100, 500, 1000)#number of series you are clustering
-nT <-  c(300, 1000, 5000) # length of the time series
-mean_diff <- c(1, 2, 5) # difference between consecutive categories
+# niter <- c(5, 50, 100, 500, 1000) #number of series you are clustering
+# nT <-  c(300, 1000, 5000, 10000) # length of the time series
+# mean_diff <- c(1, 2, 5) # difference between consecutive categories
 
-# number of iterations corresponding to
-niter <- c(50) #number of series you are clustering
+#number of iterations corresponding to
+niter <- c(5) #number of series you are clustering
 nT <-  c(300) # length of the time series
-mean_diff <- c(1) # difference between consecutive categories
+mean_diff <- c(1,2) # difference between consecutive categories
 
 time_series <- c("ar1", "arma22")
 
@@ -33,10 +35,10 @@ scen <- 2
 
 simj<-simtable[scen,] #Extract row of table
 mean_diffj <- simj$mean_diff
-niter <- simj$niter
-nT <- simj$nT
+niterj <- simj$niter
+nTj <- simj$nT
 time_seriesj <- simj$time_series
-seed_lenj <- simj$seed_len
+#seed_lenj <- simj$seed_len
 
 
 
@@ -145,30 +147,24 @@ generate_design <- function(n = 300, #length of time series
 
 
 set.seed(123)
-  
+
 ##----bind-designs for many iterations
-bind_data_iter <- map(seq_len(niter), 
+bind_data_iter <- map(seq_len(niterj), 
                       function(x){
-  #set.seed(seed_len[x])  
-  design1 <- generate_design(n=nT) # null design
-  design2 <- generate_design(n=nT,mu12=2)
-  design3 <- generate_design(n=nT,mu22=2, mu23 =4)
[…]

### 2021-10-04 3gran_change_5D.sh

@@ -1,0 +1,13 @@
+#! /bin/bash
+#SBATCH --job-name=3gran_change_5D
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=8G
+#SBATCH --time=80:00:00
+#SBATCH --array=1-54
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.5
+Rscript gran3_change_5D.R $SLURM_ARRAY_TASK_ID 
+

### 2021-10-04 check.R

@@ -1,32 +1,171 @@
+library(conflicted)
+library(tidyverse)
+library(gracsr)
+library(tsibble)
+library(gravitas)
+library(dplyr)
 
-generate_design <- function(ntimes, )
+generate_design <- function(n = 300, #length of time series
+         mu11 = 0,
+         mu12 = 0,
+         mu21 = 0,
+         mu22 = 0 ,
+         mu23 = 0,
+         mu31 = 0,
+         mu32 = 0,
+         mu33 = 0,
+         mu34 = 0,
+         mu35 = 0)
+{
+  
+t <- seq(0, n-1, 1)
+
+g1 <- t %%2
+g2 <- t %%3
+g3 <- t %%5
+
+str_gran <- bind_cols(index = t, g1 = g1, g2 = g2, g3 = g3)
+
+# calculation of g1
+g1_table <- bind_cols(g1 = unique(g1), dist_mean = c(mu11, mu12))
+
+g1_tally <- str_gran %>% group_by(g1) %>%
+  count() %>%
+  left_join(g1_table, by = "g1")
+     
+g1_dist <- g1_tally %>%
+  mutate(g1_d = list(rep(g1, each = n)),
+                    g1_dist = list(rnorm(n, dist_mean, 1))) %>% 
+  ungroup() %>% 
+  select(g1_d, g1_dist) %>% 
+  unnest(cols = c(g1_d, g1_dist))
+
+g1_data <- str_gran %>% arrange(g1) %>% bind_cols(g1_dist = g1_dist$g1_dist) %>% arrange(index)
+
+# calculation of g2
+
+g2_table<- bind_cols(g2 = unique(g2),
+                     dist_mean = c(mu21, mu22, mu23))
+
+g2_tally <- str_gran %>%
+  group_by(g2) %>%
+  count() %>% 
+  left_join(g2_table, by = "g2")
+
+g2_dist <- g2_tally %>%
+  mutate(g2_d = list(rep(g2, each = n)),
+         g2_dist = list(rnorm(n, dist_mean, 1))) %>% 
+  ungroup() %>% 
+  select(g2_d, g2_dist) %>% 
+  unnest(cols = c(g2_d, g2_dist))
+
+g2_data <- str_gran %>% arrange(g2) %>% bind_cols(g2_dist = g2_dist$g2_dist) %>% arrange(index)
+
+# calculation of g3
+
[…]

### 2021-10-04 gran3_change_5D.R

@@ -1,0 +1,279 @@
+# Rscripts for simulation chapter 3, patterns across 3 grans changing atleast across one design amongst 5 designs
+
+library(conflicted)
+library(tidyverse)
+library(gracsr)
+library(tsibble)
+library(gravitas)
+library(dplyr)
+library(here)
+
+
+# fix parameters
+ 
+# niter
+#seed_len = seq(12345, 12350, 1)
+# number of iterations corresponding to
+niter <- c(5, 50, 100) #number of series you are clustering
+nT <-  c(300, 1000, 5000) # length of the time series
+mean_diff <- c(1, 2, 5) # difference between consecutive categories
+
+# number of iterations corresponding to
+# niter <- c(50) #number of series you are clustering
+# nT <-  c(300) # length of the time series
+# mean_diff <- c(1) # difference between consecutive categories
+
+time_series <- c("ar1", "arma22")
+
+simtable <- expand.grid(mean_diff = mean_diff,
+                        niter = niter,
+                        time_series = time_series, 
+                        nT = nT)
+
+#scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
+scen <- 16
+
+simj<-simtable[scen,] #Extract row of table
+mean_diffj <- simj$mean_diff
+niterj <- simj$niter
+nTj <- simj$nT
+time_seriesj <- simj$time_series
+#seed_lenj <- simj$seed_len
+
+
+
+generate_design <- function(n = 300, #length of time series
+                            mu11 = 0,
+                            mu12 = 0,
+                            mu21 = 0,
+                            mu22 = 0 ,
+                            mu23 = 0,
+                            mu31 = 0,
+                            mu32 = 0,
+                            mu33 = 0,
+                            mu34 = 0,
+                            mu35 = 0)
+{
+  
+  n <- n+500 # 500 burning observations
+  t <- seq(0, n-1, 1)
+  
+  g1 <- t %%2
+  g2 <- t %%3
+  g3 <- t %%5
+  
[…]

### 2021-10-07 04_handpick_from356-mds_nqt.Rmd
intro applcation ssection

--- 2021-09-29 supplementary/distance-computation/real-data/04handpick-from356-mds.Rmd
@@
 
-```{r mds-all}
+```{r mds-all-2}
 
@@
 
-# Visualize similarity between samples (all gran together)
-
-for different distance metrics hod+moy+wkndwday, hod, moy, wkndwday
+# Visualize similarity between js and wpd methodologies:
 
@@
 
-<!-- ```{r} -->
-
-<!-- ggscatter(mds,  -->
-<!--           x = "Dim.1", -->
-<!--           y = "Dim.2",  -->
-<!--           #label = elec_pick_wide$customer_id, -->
-<!--           size = 1, -->
-<!--           repel = TRUE) -->
-
-<!-- fig <- plot_ly(mds, x = ~Dim.1, y = ~ Dim.2, text = total_distance_wide$item1) -->
-
-<!-- ``` -->
-
-## --- interactive-characterization-hod
+## --- Total
 
@@
 data_heatmap_new <- data_heatmap %>%
-left_join(mds_all, by = "customer_id") %>%  pivot_wider( names_from = quantiles, values_from = quantiles_values)
+left_join(mds_all, by = "customer_id") %>% 
+  pivot_wider( names_from = quantiles, 
+               values_from = quantiles_values)
 
-tx <- highlight_key(data_heatmap_new, ~customer_id)
+tx <- highlight_key(data_heatmap_new,
+                    ~customer_id)
 
@@
 
-subplot(mds_total, quantile_plot, nrows = 2, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect")
+subplot(mds_total, quantile_plot, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect", persistent = TRUE)
+```
 
@@
 
-data_heatmap_hod <- data_heatmap_new %>% filter(gran == "hour_day") %>% 
+```{r}
+data_heatmap_hod <- data_heatmap_new %>% 
+  filter(gran == "hour_day") %>% 
   mutate(category = as.integer(category))
+
+data_heatmap_hod$category <- factor(data_heatmap_hod$category, levels = 0:23)
+
 tx <- highlight_key(data_heatmap_hod, ~customer_id)
@@
    add_lines(x = ~ category, y = ~`50%`, color = ("black")) %>%
-   #add_ribbons(x = ~category,
-    #           ymin = ~`25%`,
-    #           ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>%
+   add_ribbons(x = ~category,
+             ymin = ~`25%`,
+             ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>%
    layout(showlegend = FALSE)
@@
 ```
+
+# moy
+
 ```{r}
-# hod
 
-data_heatmap_hod <- data_heatmap_new %>% filter(gran == "hour_day") %>% mutate(category = as.integer(category))
+data_heatmap_moy <- data_heatmap_new %>% 
+  filter(gran == "month_year")
 
-tx <- highlight_key(data_heatmap_hod, ~customer_id)
+data_heatmap_moy$category <- factor(data_heatmap_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+#%>%  mutate(category = as.integer(category))
+
+tx <- highlight_key(data_heatmap_moy, ~customer_id)
 
@@
    add_lines(x = ~ category, y = ~`50%`, color = ("black")) %>%
-   #add_ribbons(x = ~category,
-    #           ymin = ~`25%`,
-    #           ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>%
+   add_ribbons(x = ~category,
+             ymin = ~`25%`,
+             ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>%
    layout(showlegend = FALSE)
 
- mds_hod <- base %>%
+ mds_moy <- base %>%
   group_by(customer_id) %>%
@@
 
-subplot(mds_hod, quantile_plot, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect")
+subplot(mds_moy, quantile_plot, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect", persistent = TRUE)
 ```
 
-```{r}
+## run upto this
+
+```{r, eval = FALSE}
 
@@
          item2 = as.integer(as.character(item2)))
-
-# data_pick_three <- total_distance %>% filter(item1 %in% 
-#                             c(11013154)) %>% 
-#   mutate(item1 = as.integer(as.character(item1)),
-#          item2 = as.integer(as.character(item2)))
 ```
 
-```{r}
+```{r,  eval = FALSE}
 
@@
 .id = "design")
-
-# data_pick_cust <- bind_rows(
-# unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(),    c(11013154, 8495194, 8627007,10109182,10677705,8952846) %>% as_tibble(), c(867773, 8240424, 8267102,8281608,8403312,8454235) %>% as_tibble(), .id = "design")
-
-# c(11013154, 11045304, 11048034,11139841,11156027,11006388)
-
-# data_pick_cust <- bind_rows(
-# unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(), unique(c(data_pick_two$item1,data_pick_two$item2)) %>% as_tibble(), unique(c(data_pick_three$item1,data_pick_three$item2)) %>% as_tibble(), .id = "design")
 

### 2021-10-07 Simulations.Rmd
intro applcation ssection

@@
 
+data_pick_five <- total_distance %>% filter(item1 %in%
+                            c(8253276)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
+  mutate(item1 = as.integer(as.character(item1)),
+         item2 = as.integer(as.character(item2)))
+
 # c(11013154, 8495194, 8627007,10109182,10677705,8952846)
@@
 unique(c(data_pick_four$item1,data_pick_four$item2)) %>% as_tibble(),
+unique(c(data_pick_five$item1,data_pick_five$item2)) %>% as_tibble(),
 .id = "design")
@@
 
-group <- f %>% hclust (method = "ward.D") %>% cutree(k=4)
+group <- f %>% hclust (method = "ward.D") %>% cutree(k=5)
 
@@
 ```{r}
+parcoord
+
+```
+
+```{r}
 
@@
 ```
+```{r pcp-distance-computation}
+
+elec_600_wpd %>% left_join(cluster_result_wpd, by = c("customer_id" = "id"))
+
+parcoord <- GGally::ggparcoord(data_pcp ,
+                   columns = 3:ncol(data_pcp),
+                   groupColumn = "group",
+                   showPoints = FALSE, 
+                   alphaLines = 0.8,
+                   order = "allClass",
+                   scale = "globalminmax"
+) + ggplot2::theme_bw() +
+  #scale_color_brewer(palette = "Dark2") + 
+  ggplot2::theme(
+    plot.title = ggplot2::element_text(size=10)
+  )+
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
+  theme(legend.position = "bottom") +
+  xlab("") +
+  ylab("wpd")
+
+```
+
+## Clustering 396 customers
+
+```{r clustering-all}
+filtered_distance <- distance %>%
+  filter(!(item1 %in% c(8196183, 8508008, 8680538))) %>% 
+  filter(!(item2 %in% c(8196183, 8508008, 8680538)))
+
+total_distance <- filtered_distance %>% 
+mutate(total = wkndwday/2 + moy/12 + hod/24) 
+#mutate(total = hod) 
+# wkndwday/2
+
+total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
+                               names_from = item2,
+                               values_from = total)
+#total_distance_wide$`11013154` <- NA
+
+rownames(total_distance_wide) <- total_distance_wide$item1
+
+mds_data <- total_distance_wide %>% 
+  mutate_all(~replace(., is.na(.), 0)) %>%
+   tibble::rownames_to_column() %>%  
+   select(-item1) %>% 
+   pivot_longer(-rowname) %>% 
+   pivot_wider(names_from=rowname, values_from=value)
+
+rownames(mds_data) <- total_distance_wide$item1
+
+# group <- mds_data[-1] %>% clust_gran(kopt=4)
+#
+
+df <- mds_data[-1] %>% as.matrix()
+DM <- matrix(0, ncol(mds_data), ncol(mds_data))
+DM[lower.tri(DM)] = df[lower.tri(df, diag=TRUE)] # distance metric
+```
+
+```{r koptimal}
+# choosing optimal number of clusters
+   # koptimal = fpc::nselectboot(DM,
+   #                              B = 50,
+   #                              method = "ward.D",
+   #                              clustermethod = fpc::disthclustCBI,
+   #                              classification = "averagedist",
+   #                              krange = 5:20)
+   #  kopt = koptimal$kopt
+```
+
+```{r group-all}
+
+f = as.dist(DM)
+
+#f %>% hclust %>% cutree(k=4)
+
+first_lot <- mds_data %>% names()
+# because distance matrix is such  that 1:23 appears in columns and 2:24 appears in rows such that 1st customer is missing from row and last customer is missing from variables
+id <- c(first_lot[-1], mds_data$name[nrow(mds_data)])
+
+group <- f %>% hclust (method = "ward.D") %>% cutree(k=12)
+
+cluster_result <- bind_cols(id = id, group = group)
+
+mds  = cmdscale(f)
+rownames(mds) = id
+colnames(mds) = c("mds1", "mds2")
+
+# ggplot(data = mds %>% as_tibble() %>% mutate(group = cluster_result$group), 
+#        aes(x = mds1, y = mds2, color = as.factor(group))) +
+#   geom_point() +
+#   geom_text(aes(label = id))
+
+# cluster_result %>% filter(group==1)
+# cluster_result %>% filter(group==2)
+# cluster_result %>% filter(group==3)
+# cluster_result %>% filter(group==4)
+
+# cluster_result %>% filter(group==9)
+# cluster_result %>% filter(group==19)
+
+  #x = ~mds1.hod, y = ~ mds2.hod, text = ~customer_id
+```
+
+```{r data-heatmap-hod-group}
+
+# for entire data
+data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))%>%
+    filter(!(customer_id %in% c(8196183, 8508008, 8680538))) %>% 
+  gracsr::scale_gran( method = "robust", response = "general_supply_kwh")
+
+data_group <- data_pick %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id"))
+
+data_heatmap_hod_group <- quantile_gran(data_group,
+                                  gran1="hour_day",
+                                  quantile_prob_val = c(0.1, 0.25, 0.5, 0.75, 0.9),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
+
+data_heatmap_hod_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group))) +
+    geom_ribbon(aes(ymin = `10%`, 
+                  ymax = `90%`, group=group, fill = as.factor(group)), alpha = 0.5) + 
+  geom_line(aes(y = `50%`, group=group)) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow = 3) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+```
+
+```{r data-heatmap-moy-group}
+
+data_heatmap_moy_group <- quantile_gran(data_group,
+                                  gran1="month_year",
+                                  quantile_prob_val = c(0.1, 0.25, 0.5, 0.75, 0.9),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+data_heatmap_moy_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group))) +  geom_ribbon(aes(ymin = `10%`, 
+                  ymax = `90%`, group=group, fill = as.factor(group)), alpha = 0.5) + 
+  geom_line(aes(y = `50%`, group=group)) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow = 3) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+```
+
+```{r data-heatmap-wkndwday-group}
+
+data_heatmap_wkndwday_group <- quantile_gran(data_group,
+                                  gran1="wknd_wday",
+                                  quantile_prob_val = c(0.1, 0.25, 0.5, 0.75, 0.9),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_wkndwday_group$category <- factor(data_heatmap_wkndwday_group$category, levels = c("Weekday", "Weekend"))
+
+data_heatmap_wkndwday_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group))) +
+    geom_ribbon(aes(ymin = `10%`, 
+                  ymax = `90%`, group=group, fill = as.factor(group)), alpha = 0.5) + 
+  geom_line(aes(y = `50%`, group=group)) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 4) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("wknd_wday") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+```
+```{r}
+
+elec_600_wpd <- read_rds(here::here("data/algo2-cust600-wpd-rawdata.rds"))
+
+elec356 <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))
+
+# %>%
+#     filter(!(customer_id %in% c(8196183, 8508008, 8680538))) %>% 
+#   gracsr::scale_gran( method = "robust", response = "general_supply_kwh")
+
+elec_pick <- elec_600_wpd %>% 
+  filter(elec356 %in% elec356) %>% 
+     filter(!(customer_id %in% c(8196183, 8508008, 8680538)))
+
+elec_pick_wide <- elec_pick %>% pivot_wider(-c(1, 2), names_from = "x_variable", values_from = wpd)
+
+# robust_scale <-  function(x){
+#   (x - stats::median(x))/(stats::IQR(x))
+# }
+
+#scaled_var <- apply(elec_pick_wide[-c(1)], 2, robust_scale) %>% as_tibble()
+
+scaled_var <- elec_pick_wide
+
+f <- elec_pick_wide[-1] %>% dist()
+
+group <- f%>% hclust (method = "ward.D") %>% cutree(k=12)
+
+cluster_result_wpd <- bind_cols(id = elec_pick_wide$customer_id, group = group)
+
+mds  = cmdscale(f)
+rownames(mds) = id
+colnames(mds) = c("mds1", "mds2")
+
+ggplot(data = mds %>% as_tibble() %>% mutate(group = cluster_result_wpd$group), 
+       aes(x = mds1, y = mds2, color = as.factor(group))) +
+  geom_point() +
+  geom_text(aes(label = id))
+
+data_pcp <- scaled_var %>% 
+  #bind_cols(customer_id =  elec_pick_wide$customer_id) %>%
+  left_join(cluster_result_wpd , by = c("customer_id" = "id")) %>% 
+  select(customer_id, group, everything()) %>% 
+  mutate(group = as.factor(group))
+
+parcoord <- GGally::ggparcoord(data_pcp ,
+                   columns = 3:ncol(data_pcp),
+                   groupColumn = "group",
+                   showPoints = FALSE, 
+                   alphaLines = 0.8,
+                   order = "anyClass",
+                   scale = "globalminmax"
+) + ggplot2::theme_bw() +
+  scale_color_brewer(palette = "Dark2") + 
+  ggplot2::theme(
+    plot.title = ggplot2::element_text(size=10)
+  )+
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
+  theme(legend.position = "bottom") +
+  xlab("") +
+  ylab("wpd")
+
+```
+
+```{r}
+parcoord
+
+```

### 2021-10-07 main.Rmd
intro applcation ssection

@@
 
-To evaluate clustering approaches, we create data designs that replicate prototype behaviours that might be seen in electricity data contexts.
+To evaluate clustering approaches, we create data designs that replicate prototype behaviors that might be seen in electricity data contexts.
 We spiked several features in the data to see where one method works better than the other and where they might give us the same outcome or the effect of missing data and trends on the proposed methods.
@@
 Each categories of $g1$, $g2$ and $g3$ could be assumed to be from same distribution, few from one and others from separate distributions, or all from different distributions, resulting in distinct data designs.
-We created independent replications $(R = \{50, 200, 500\})$ of all data designs to see if our proposed clustering approaches can detect distinct designs in various groups for small, medium and large number of series.
-A sample size of $T=\{300, 500, 2000\}$ is used in all designs to test small, medium and large sized series.
-For all data designs, the data type is set to "continuous," and a gaussian setup for errors is assumed.
+We created independent replications $(R = \{25, 250, 500\})$ of all data designs to see if our proposed clustering approaches can detect distinct designs in various groups for small, medium and large number of series.
+A sample size of $T=\{300, 1000, 5000\}$ is used in all designs to test small, medium and large sized series.
+For all data designs, the data type is set to  be "continuous," and a gaussian setup for errors is assumed.
+The method could preform differently with different jumps between consecutive categories.
+So a difference of $diff = {1, 2, 5}$ for corresponding categories are also considered.
 The code for creating these designs can be found in the Supplementary section (link to github repo).
-The results for $T=300$ and $R=50$.
+The results for $T=300$ and $R=25$ is shown, that means we have $25$ time series each with length $300$.
 The rest of the results could be found in the supplementary paper.
@@
 That means that for at least one among the to-be-grouped designs, the pattern for each of $g1$, $g2$, and $g3$ will change.
-We consider null instances to represent no variation in distribution across categories, i.e., all categories have the same distribution N(0,1).
+We consider a _null_ case to represent that categories follow N(0,1) and _alternate_ case to represent different  distributions across categories (as in Table \ref{tab:}) that will lead to different designs (as in Table \ref{tab:}).
+
+```{r tab-distribution}
+
+# table_mayan <- tibble(granularity  = )
+
+```
 
@@
 
-Consider a case where distribution of $y$ would vary across levels of $g2$
-for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs. So $g3$ is not responsible for distinguishing across designs.
-
-**Missing data**
-
-**Trend**
-
-**Outliers**
-
-observations:
-
-designs:
+Consider a case where distribution of $y$ would vary across levels of $g2$ for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
+So $g3$ is not responsible for distinguishing across designs.
+
+<!-- **Missing data** -->
+
+<!-- **Trend** -->
+
+<!-- **Outliers** -->
+
+<!-- observations: -->
+
+<!-- designs: -->
 
@@
 
-### Raw data
-
-### Missing Data
-
-Electricity usage for some customers may become unavailable due to power outage or not recording their usage properly, thus resulting in implicit missing values in the database.
+### Characteristics of raw data
+
+In Figure \ref{fig:raw-data-50}, the time series of energy consumption is plotted along the y-axis against time from past to future for $50$ sampled households.
+Each of these series correspond to a single customer.
+For each customer, the energy consumption is available at fine temporal resolution (every 30 minutes) for a long period of time (~ 2 years) resulting in 27,000 (median) observations for each customer.
+Some customers' electricity use may be unavailable owing to power outages or improper recording, resulting in implied missing numbers in the database.
+For this data set it was found that out of 13735 customers in total, 8685 customers do not have any implicit missing observations, while the rest 5050 customers had missing values.
+With further exploration, it was found that there is no structure in the missingness, that is missing observations can occur at any time point (see Appendix).
+Moreover, the data for these customers are characterized by unequal length, different start and end dates.
+Since our proposed methods consider probability distribution instead of raw data, both of these characteristics would not pose any threat to our methodology unless of course there is any structure or systematic patterns in them.
+
+\noindent It can be expected that energy consumption vary substantially between customers, which is a reflection of their varied  behavior owing to differences in profession, family size, geographical or physical characteristics.
+Since the linear time series plot has too many measurements all squeezed in this linear representation, it hinders us to discern any repetitive behavioral pattern for even one customers (let alone many customers together).
+In most cases, electricity data will have multiple seasonal patterns like daily, weekly or annual.
+We do not learn about these repetitive behaviors from the linear view.
+Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.
+
+\noindent The aim of this section is to illustrate that the proposed methodology can be used to understand repetitive behavior for several customers together.
+
+```{r raw-data-50, out.width = "100%", fig.cap="The raw data for 50 households are shown.
+It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed.
+No insightful behavioral pattern could be discerned from this view other than when the customer is not at home."}
+
+knitr::include_graphics("figs/raw_plot_cust.png") # look at smart-meter.R for the code
+```
+
+<!-- ### Missing Data -->
+
+<!-- Electricity usage for some customers may become unavailable due to power outage or not recording their usage properly, thus resulting in implicit missing values in the database.
 It is interesting to explore where missing-ness occurs or if there is a relationship between the underlying missing patterns.
 We use the R package `tsibble` to do this.
-
-```{r, out.width="100%"}
+-->
+
+```{r missing-data, out.width="100%", eval = FALSE}
 include_graphics("figs/missing-data.png")
@@
 
-- if there is any systematic missing patterns in the data
-- this missing plot can go in the supplementary
-- how to add missing values (should be added in data pre-processing)
-- instance learning
-- types of summary techniques to use (
-generally multivariate means and sd are used, I can't use that in a time series context, you can show across different granularity?
-within-group sum of squares and between-group sum of squares
-
-)
-
-13735 customers in elec_ts
-8685 customers in elec_nogap
-5050 customers in count_na_df
-
-Then is the graph of missing observations even interesting.
-You can show two graphs, one to show that missingness do not have a pattern
-another to show even if no missing, they start and end at different times. (A sample of 50 customers).
-
-\noindent A dataset of 100 SGSC homes has been used to lay out the structure to be used for analyzing the big dataset.
+<!-- - if there is any systematic missing patterns in the data -->
+<!-- - this missing plot can go in the supplementary -->
+<!-- - how to add missing values (should be added in data pre-processing) -->
+<!-- - instance learning -->
+<!-- - types of summary techniques to use ( -->
+<!-- generally multivariate means and sd are used, I can't use that in a time series context, you can show across different granularity? -->
+<!-- within-group sum of squares and between-group sum of squares -->
+
+<!-- ) -->
+
+<!-- 13735 customers in elec_ts -->
+<!-- 8685 customers in elec_nogap -->
+<!-- 5050 customers in count_na_df -->
+
+<!-- Then is the graph of missing observations even interesting. -->
+<!-- You can show two graphs, one to show that missingness do not have a pattern -->
+<!-- another to show even if no missing, they start and end at different times. (A sample of 50 customers).  -->
+
+<!-- \noindent A dataset of 100 SGSC homes has been used to lay out the structure to be used for analyzing the big dataset.
 The smaller dataset contains half-hourly kwh values form 2012 to 2014 and has asynchronous time series distributed evenly over the observation period (Figure \ref{fig:elec-raw}), similar to the bigger data set.
@@
 It looks like data is most missing before 2013 and for a particular date in 2014.
+-->
 
@@
 -->
+
+## Prototype application
+
+A clean data set is obtained by carefully choosing customers which shows similar shapes across one or more cyclic granularity.
+Since this is unlabeled data, there is no way to do external validation of our methodologies.
+Thus, we chose this way to see how well our methodology works in a cleaner data set as this one.
+
+Fig \ref{fig:}, \ref{fig:} and \ref{fig:} shows the distribution of 24 customers across hour-of-day, month-of-year and wknd-wday respectively.
+Every row in \ref{fig:} shows different shapes across hour-of-day and across columns show similar shapes for each row.
+We use our methodology to see if the customers are allocated to same group have similar shapes across one or more significant granularity.
+
+```{r data-pick}
+
+quantile_prob_graph <- c(0.25, 0.5, 0.75)
+
+wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_robust.rds")) %>% broom::tidy()
+
+moy <- read_rds(here("data/dist_gran_moy_356cust.rds")) %>% broom::tidy()
+
+hod <- read_rds(here("data/dist_gran_hod_356cust_robust.rds")) %>% broom::tidy()
+
+distance <- wkndwday %>% 
+  left_join(moy, by = c("item1", "item2")) %>% 
+  left_join(hod, by = c("item1", "item2")) %>% 
+  rename("wkndwday" ="distance.x",
+         "moy" = "distance.y",
+         "hod" = "distance") %>%
+  mutate(item1 = as.integer(as.character(item1)),
+         item2 = as.integer(as.character(item2)))
+
+total_distance <- distance %>% 
+  mutate(total = hod)
+  #mutate(total = wkndwday/2 + moy/12 + hod/24)
+
+# 8454221 10420689
+data_pick_one <- total_distance %>% filter(item1 %in% 
+                            c(8454221)) %>% group_by(item1) %>% arrange(total) %>% head(5) %>% 
+  mutate(item1 = as.integer(as.character(item1)),
+         item2 = as.integer(as.character(item2)))
+
+data_pick_ch <-  total_distance %>% filter(item1 %in% 
+                            c(9393696)) %>% group_by(item1) %>% arrange(total) %>% head(5) %>% 
+  mutate(item1 = as.integer(as.character(item1)),
+         item2 = as.integer(as.character(item2)))
+
+#11013154  8181071 11048034
+data_pick_two <- total_distance %>% filter(item1 %in% 
+                            c(8181071)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
+  mutate(item1 = as.integer(as.character(item1)),
+         item2 = as.integer(as.character(item2)))
+
+data_pick_three <- total_distance %>% filter(item1 %in% 
+                            c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
+  mutate(item1 = as.integer(as.character(item1)),
+         item2 = as.integer(as.character(item2)))
+
+# data_pick_three <- total_distance %>% filter(item1 %in%
+#                             c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
+#   mutate(item1 = as.integer(as.character(item1)),
+#          item2 = as.integer(as.character(item2)))
+
+# 8653709
+data_pick_four <- total_distance %>% filter(item1 %in%
+                            c(8184707)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
+  mutate(item1 = as.integer(as.character(item1)),
+         item2 = as.integer(as.character(item2)))
+
+# c(11013154, 8495194, 8627007,10109182,10677705,8952846)
+
+data_pick_cust <- bind_rows(
+unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(),    unique(c(data_pick_two$item1,data_pick_two$item2)) %>% as_tibble(), 
+c(11013154, 9345642, 8328008, 8636035,8627007,8454235) %>% as_tibble(),
+unique(c(data_pick_four$item1,data_pick_four$item2)) %>% as_tibble(),
+.id = "design")
+
+data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))%>%
+  filter(customer_id %in% data_pick_cust$value) %>% 
+  gracsr::scale_gran( method = "robust", response = "general_supply_kwh")
+```
+
+<!-- # ```{r} -->
+<!-- # data_pick %>% ggplot() + geom_line(aes(x=reading_datetime, y = general_supply_kwh)) + -->
+<!-- #   facet_wrap(~customer_id, ncol= 2) +  -->
+<!-- # ``` -->
+
+```{r clustering}
+filtered_distance <- distance %>% filter(item1 %in% data_pick_cust$value) %>% 
+  filter(item2 %in% data_pick_cust$value)
+
+total_distance <- filtered_distance %>% 
+mutate(total = wkndwday/2 + moy/12 + hod/24) 
+#mutate(total = hod)
+
+total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
+                               names_from = item2,
+                               values_from = total)
+#total_distance_wide$`11013154` <- NA
+
+rownames(total_distance_wide) <- total_distance_wide$item1
+
+mds_data <- total_distance_wide %>% 
+  mutate_all(~replace(., is.na(.), 0)) %>%
+   tibble::rownames_to_column() %>%  
+   select(-item1) %>% 
+   pivot_longer(-rowname) %>% 
+   pivot_wider(names_from=rowname, values_from=value)
+
+rownames(mds_data) <- total_distance_wide$item1
+
+# group <- mds_data[-1] %>% clust_gran(kopt=4)
+#
+
+df <- mds_data[-1] %>% as.matrix()
+DM <- matrix(0, ncol(mds_data), ncol(mds_data))
+DM[lower.tri(DM)] = df[lower.tri(df, diag=TRUE)]
+
+f = as.dist(DM)
+
+#f %>% hclust %>% cutree(k=4)
+
+first_lot <- mds_data %>% names()
+# because distance matrix is such  that 1:23 appears in columns and 2:24 appears in rows such that 1st customer is missing from row and last customer is missing from variables
+id <- c(first_lot[-1], mds_data$name[nrow(mds_data)])
+
+group <- f %>% hclust (method = "ward.D") %>% cutree(k=4)
+
+cluster_result <- bind_cols(id = id, group = group)
+
+mds  = cmdscale(f)
+rownames(mds) = id
+colnames(mds) = c("mds1", "mds2")
+
+ggplot(data = mds %>% as_tibble() %>% mutate(group = cluster_result$group), 
+       aes(x = mds1, y = mds2, color = as.factor(group))) +
+  geom_point() +
+  geom_text(aes(label = id))
+
+# cluster_result %>% filter(group==1)
+# cluster_result %>% filter(group==2)
+# cluster_result %>% filter(group==3)
+# cluster_result %>% filter(group==4)
+
+  #x = ~mds1.hod, y = ~ mds2.hod, text = ~customer_id
+```
+```{r heatmap-hod}
+data_heatmap_hod <- quantile_gran(data_pick,
+                                  "hour_day", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
+  left_join(data_pick_cust, by = c("customer_id"="value"))
+
+data_heatmap_hod$category <- factor(data_heatmap_hod$category, levels = 0:23)
+
+```
+
+```{r data-heatmap-hod}
+data_heatmap_hod %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=customer_id), fill = "lightblue") +
+  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+```
+
+```{r data-heatmap-hod-group}
+
+data_group <- data_pick %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id"))
+
+data_heatmap_hod_group <- quantile_gran(data_group,
+                                  gran1="hour_day",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
+
+data_heatmap_hod_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group))) +
+  geom_line(aes(y = `50%`, group=group)) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow = 4) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+```
+
+```{r data-heatmap-moy}
+
+data_heatmap_moy <- quantile_gran(data_pick, "month_year", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
+  left_join(data_pick_cust, by = c("customer_id"="value"))
+
+data_heatmap_moy$category <- factor(data_heatmap_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+data_heatmap_moy %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=customer_id), fill = "lightblue") +
+  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+```
+
+```{r data-heatmap-moy-group}
+
+data_group <- data_pick %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id"))
+
+data_heatmap_moy_group <- quantile_gran(data_group,
+                                  gran1="month_year",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+data_heatmap_moy_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group))) +
+  geom_line(aes(y = `50%`, group=group)) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow = 4) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+```
+
+```{r data-heatmap-wkndwday}
+
+data_heatmap_wkndwday <- quantile_gran(data_pick, "wknd_wday", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
+  left_join(data_pick_cust, by = c("customer_id"="value"))
+
+data_heatmap_wkndwday$category <- factor(data_heatmap_wkndwday$category, levels = c("Weekday", "Weekend"))
+
+data_heatmap_wkndwday %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=customer_id), fill = "lightblue") +
+  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("wknd_wday") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+```
+
+```{r data-heatmap-wkndwday-group}
+
+data_group <- data_pick %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id"))
+
+data_heatmap_wkndwday_group <- quantile_gran(data_group,
+                                  gran1="wknd_wday",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_wkndwday_group$category <- factor(data_heatmap_wkndwday_group$category, levels = c("Weekday", "Weekend"))
+
+data_heatmap_wkndwday_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group))) +
+  geom_line(aes(y = `50%`, group=group)) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 4) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("wknd_wday") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+
+```
 

### 2021-10-07 supplementary/distance-computation/real-data/04_handpick_from356-mds_nqt.Rmd
intro applcation ssection

--- 2021-10-07 04_handpick_from356-mds_nqt.Rmd
@@
 ---
-title: "Selecting similar behaving group of customers to check clustering methodology"
+title: "Selecting similar behaving group of customers to check nqt clustering methodology"
 output: html_document
@@
 
-wkndwday <- read_rds("../../../data/dist_gran_wkndwday_356cust_robust.rds") %>% broom::tidy()
+wkndwday <- read_rds("../../../data/dist_gran_wkndwday_356cust_nqt.rds") %>% broom::tidy()
 
-moy <- read_rds("../../../data/dist_gran_moy_356cust.rds") %>% broom::tidy()
+moy <- read_rds("../../../data/dist_gran_moy_356cust_nqt.rds") %>% broom::tidy()
 
-hod <- read_rds("../../../data/dist_gran_hod_356cust_robust.rds") %>% broom::tidy()
+hod <- read_rds("../../../data/dist_gran_hod_356cust_nqt.rds") %>% broom::tidy()
 

### 2021-10-07 supplementary/distance-computation/real-data/02_algo1_hod_356cust_robust.R

@@ -13,5 +13,10 @@ gran1 = "hour_day"
 
 #elec <- elec %>% filter(customer_id %in% c(8143599, 8147121, 8159803, 10542667))
 
-distance_data_robust <- elec %>% scale_gran(method = "robust") %>% dist_gran(gran1)
-write_rds(distance_data_robust, here::here("data/dist_gran_hod_356cust_robust.rds"))
+# distance_data_robust <- elec %>% scale_gran(method = "robust") %>% dist_gran(gran1)
+# write_rds(distance_data_robust, here::here("data/dist_gran_hod_356cust_robust.rds"))
+# 
+
+distance_data_nqt <- elec %>% scale_gran(method = "nqt") %>% dist_gran(gran1)
+write_rds(distance_data_nqt, here::here("data/dist_gran_hod_356cust_nqt.rds"))
+

### 2021-10-07 supplementary/distance-computation/real-data/02_algo1_moy_356cust_robust.R

@@ -13,5 +13,11 @@ gran1 = "month_year"
 
 #elec <- elec %>% filter(customer_id %in% c(8143599, 8147121, 8159803, 10542667))
 
-distance_data_robust <- elec %>% scale_gran(method = "robust") %>% dist_gran(gran1)
-write_rds(distance_data_robust, here::here("data/dist_gran_moy_356cust.rds"))
+# Code for wkndwday
+# distance_data_robust <- elec %>% scale_gran(method = "robust") %>% dist_gran(gran1)
+# 
+# write_rds(distance_data_robust, here::here("data/dist_gran_moy_356cust.rds"))
+
+distance_data_nqt <- elec %>% scale_gran(method = "nqt") %>% dist_gran(gran1)
+
+write_rds(distance_data_nqt, here::here("data/dist_gran_moy_356cust_nqt.rds"))

### 2021-10-07 main.Rmd
adding paper application

@@
 
-A clean data set is obtained by carefully choosing customers which shows similar shapes across one or more cyclic granularity.
+<!-- A clean data set is obtained by carefully choosing customers which shows similar shapes across one or more cyclic granularity.
 Since this is unlabeled data, there is no way to do external validation of our methodologies.
 Thus, we chose this way to see how well our methodology works in a cleaner data set as this one.
-
-Fig \ref{fig:}, \ref{fig:} and \ref{fig:} shows the distribution of 24 customers across hour-of-day, month-of-year and wknd-wday respectively.
+-->
+
+First, we select the customers which do not have any implicit missing values and filtered their data for the year 2013.
+From this set, we randomly sample a set of 600 customers.
+Then we removed customers for which any of the hod, moy and wkndwday is a clash.
+We further removed customers for which the deciles of the energy consumption is of the same value.
+Finally, we are left with 356 customers from which we wanted to do instance selection.
+Since this is unlabeled data, there is no way to do external validation of our methodologies.
+Thus, we chose this way to see how well our methodology works in a cleaner data set as this one to classify customers who have similar behavior across all significant granularities.
+Next we select 4 customers which are far apart from each other and 6 customers which are lying closest in distance to each of these 4 customers.
+These selection was done using the granularity hod.
+When we use our methodologies, it is based on all significant granularities and not only hod, but if hod is significant we can expect that the grouping would be similar to how we have initially chosen the set of customers.
+This is process is similar to instance-wise classfication methods.By this instancewise discriminative learning, instances can reasonably distribute in the representation space, which reveals their similarities.
+In
+order to further improve visual representations, we propose a
+dual-level progressive similar instance selection (DPSIS) method
+to build a bridge from instance to class by selecting similar
+instances (neighbors) for each instance (anchor) and treating the
+anchor and its neighbors as the same class.
+
+Fig \ref{fig:data-heatmap-hod}, \ref{fig:} and \ref{fig:} shows the distribution of 24 customers across hour-of-day, month-of-year and wknd-wday respectively.
 Every row in \ref{fig:} shows different shapes across hour-of-day and across columns show similar shapes for each row.
@@
 data_pick_one <- total_distance %>% filter(item1 %in% 
-                            c(8454221)) %>% group_by(item1) %>% arrange(total) %>% head(5) %>% 
+                            c(8454221)) %>% 
+  #group_by(item1) %>%
+  arrange(total) %>% 
+  head(5) %>% 
   mutate(item1 = as.integer(as.character(item1)),
@@
 data_pick_three <- total_distance %>% filter(item1 %in% 
-                            c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
+                            c(11013154)) %>%
+  #group_by(item1) %>%
+  arrange(total) %>% head(5)%>% 
   mutate(item1 = as.integer(as.character(item1)),
@@
 data_pick_four <- total_distance %>% filter(item1 %in%
-                            c(8184707)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
+                            c(8184707)) %>% 
+  #group_by(item1) %>% 
+  arrange(total) %>% head(5)%>%
   mutate(item1 = as.integer(as.character(item1)),
@@
 
+data_pick_cust$design <- paste("design", data_pick_cust$design, sep = "-")
+
 data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))%>%
@@
   gracsr::scale_gran( method = "robust", response = "general_supply_kwh")
+
 ```
@@
 
-ggplot(data = mds %>% as_tibble() %>% mutate(group = cluster_result$group), 
-       aes(x = mds1, y = mds2, color = as.factor(group))) +
-  geom_point() +
-  geom_text(aes(label = id))
+# ggplot(data = mds %>% as_tibble() %>% mutate(group = cluster_result$group), 
+#        aes(x = mds1, y = mds2, color = as.factor(group))) +
+#   geom_point() +
+#   geom_text(aes(label = id))
 
@@
 ```{r heatmap-hod}
+
 data_heatmap_hod <- quantile_gran(data_pick,
@@
 
+#data_heatmap_hod$customer_id = factor(data_heatmap_hod$customer_id, levels = data_pick_cust$value)
 data_heatmap_hod$category <- factor(data_heatmap_hod$category, levels = 0:23)
@@
 
-```{r data-heatmap-hod}
+```{r data-heatmap-hod, fig.cap="The distribution of electricity demand for the selected 24 customers across hour-of-day.
+The median is represented by a line and shaded region represents the area between 25^{th} and 75^{th} percentile.
+Each row represents a distinct shape of daily load, and all customers in the same row have similar daily profile.
+If they are represented by the same color, it means that they are in the same group as suggested by our proposed clustering algorithm."}
+
 data_heatmap_hod %>% 
@@
   geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=customer_id), fill = "lightblue") +
-  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
+                  ymax = `75%`,
+                  group=customer_id,
+                  fill = as.factor(group)),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=customer_id, 
+                color = as.factor(group)), size = 1) +
   facet_wrap(design~customer_id, 
@@
              ncol = 6) +
-    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
-```
-
-```{r data-heatmap-hod-group}
+    theme_bw() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank()) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") +
+  ylab("demand (in Kwh)")  + 
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ 
+  #theme() strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") ) +
+  scale_x_discrete(breaks = seq(0,23, 4)) +
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+  xlab("group")
+```
+
+```{r data-heatmap-hod-group, fig.cap="The distribution of electricity demand for the clusters across hour-of-day.
+The median is represented by a line and the shaded region represents the area between 25^{th} and 75^{th} percentile.
+The first cluster is characterised by high comsumption in early morning hours and late night hours.
+The second cluster should be the most common type of daily profile characterised by a smaller morning and higher evening peak.
+The third cluster corresponds to two equal peaks and the last one corresponds to only distinct evening peaks."}
 
@@
 
+# data_heatmap_hod_group <- data_heatmap_hod_group %>%
+#   mutate(group = case_when(group==1 ~"Group 2",
+#                            group==2 ~"Group 4",
+#                            group==3 ~"Group 1",
+#                            group==4 ~"Group 3"))
+
+# data_heatmap_hod_group$group <- factor(data_heatmap_hod_group$group, levels = c("Group 1", "Group 2", "Group 3", "Group 4"))
+
+data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")
+
 data_heatmap_hod_group %>% 
@@
   geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=group, fill = as.factor(group))) +
-  geom_line(aes(y = `50%`, group=group)) +
+                  ymax = `75%`,
+                  group=group,
+                  fill = as.factor(group)),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=group, 
+                color = as.factor(group)), size = 1)+
   facet_wrap(~group, 
-             scales = "free_y", 
-             labeller = "label_value",
-             nrow = 4) +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
-  scale_x_discrete(breaks = seq(1, 24, 3))+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
-
-```
-
-```{r data-heatmap-moy}
+             scales = "free_y",  nrow = 4) + 
+              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + 
+  theme_bw() + 
+  theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ 
+  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+  scale_fill_manual(values = c("#0072B2", "#E69F00","#D55E00", "#009E73"))+
+  #scale_color_manual(values = c("#0072B2", "#E69F00","#D55E00", "#009E73"))+
+  theme(legend.position = "bottom")
+
+# "#E69F00", "#009E73","#0072B2", "#D55E00"
+
+# 
+# data_pick_cust <- data_pick_cust %>% mutate(value = as.character(value))
+# 
+# data_pick_cust %>% left_join(cluster_result, by = c("value" = "id")) %>% group_by(group)
+```
+
+```{r data-heatmap-moy, fig.cap = "The distribution of electricity demand for the selected 24 customers across month-of-year.
+The median is represented by a line and shaded region represents the area between 25^{th} and 75^{th} percentile.
+The clustering leads to customers with similar shape across month-of-year with some exceptions.
+This is natural in a clus"}
 
@@
 ```
+The clusters across month-of-year shows that cluster 2 uses electricity in summer.
 

### 2021-10-07 main.Rmd
adding application hour-of-day story

@@
 
-## Prototype application
+## Prototype selection
 
@@
 ```{r data-heatmap-hod, fig.cap="The distribution of electricity demand for the selected 24 customers across hour-of-day.
-The median is represented by a line and shaded region represents the area between 25^{th} and 75^{th} percentile.
-Each row represents a distinct shape of daily load, and all customers in the same row have similar daily profile.
-If they are represented by the same color, it means that they are in the same group as suggested by our proposed clustering algorithm."}
+The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
+According to our instance selection method, each row represnts a distinct shape of daily load, and all customers in the same row should have similar daily profile.After clustering these consumers, all customers with the same colour represent the same group.
+Except for customer id 8269176, there is unanimity across design and groups.
+Even though their daily form resembles Group 2, our clustering approach places them in Group 3 (which is design 4 in our case).
+Because our method uses hod, moy, and wkndwday, there may be some mismatches depending on only one variable."}
+
+legend_title <- "group"
 
@@
   scale_x_discrete(breaks = seq(0,23, 4)) +
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  xlab("group")
-```
+  scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+  xlab("hour-of-day")
+```
+```{r data-heatmap-moy, fig.cap = "The distribution of electricity demand for the selected 24 customers across month-of-year.
+The median is represented by a line and shaded region represents the area between 25^{th} and 75^{th} percentile.
+The clustering leads to customers with similar shape across month-of-year with some exceptions.
+This is natural in a clus"}
+
+data_heatmap_moy <- quantile_gran(data_pick, "month_year", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
+  left_join(data_pick_cust, by = c("customer_id"="value"))
+
+data_heatmap_moy$category <- factor(data_heatmap_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+data_heatmap_moy %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=customer_id), fill = "lightblue") +
+  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+```
+
+## Characterising clusters
 
@@
   theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-  scale_fill_manual(values = c("#0072B2", "#E69F00","#D55E00", "#009E73"))+
-  #scale_color_manual(values = c("#0072B2", "#E69F00","#D55E00", "#009E73"))+
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
   theme(legend.position = "bottom")
 
-# "#E69F00", "#009E73","#0072B2", "#D55E00"
+#
 
@@
 
-```{r data-heatmap-moy, fig.cap = "The distribution of electricity demand for the selected 24 customers across month-of-year.
-The median is represented by a line and shaded region represents the area between 25^{th} and 75^{th} percentile.
-The clustering leads to customers with similar shape across month-of-year with some exceptions.
-This is natural in a clus"}
-
-data_heatmap_moy <- quantile_gran(data_pick, "month_year", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
-  left_join(data_pick_cust, by = c("customer_id"="value"))
-
-data_heatmap_moy$category <- factor(data_heatmap_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
-
-data_heatmap_moy %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=customer_id), fill = "lightblue") +
-  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
-  facet_wrap(design~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-             ncol = 6) +
-    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
-```
+Characterization of clusters is the final stage of a cluster analysis.
+However, if we are convinced that we have identified a collection of clusters that can be distinguished from one another, we should intend to characterise them more formally, both statistically and qualitatively.
+We quantify them by producing statistics for each variable.
+We may look at the findings in graphs, and enhance our qualitative descriptions of the groupings.
+
+We discover 4 qualitative clusters of varying shapes in the distribution of all consumers in a group.
+Group-1 includes consumers who work 9-5, get up and conduct morning activities from 7-10am, and then depart.
+Then they return home in the evening to cook supper and perform other activities, giving the evening a greater peak than the morning.
+Group 2 is the group that rushes out of the house in the morning to get to work.
+They only return at night and do all activities at night, so there is no morning peak.
+The third one has a strong early morning and late night hours.
+These consumers may be flexible students or elderly retirees who are night owls.
+Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening peaks.
+All of this may be validated with further information about the customer.
 

### 2021-10-08 main.Rmd
adding application first methods story

@@
 
-### Characteristics of raw data
+## Characteristics of raw data
 
@@
 
-\noindent The aim of this section is to illustrate that the proposed methodology can be used to understand repetitive behavior for several customers together.
-
 ```{r raw-data-50, out.width = "100%", fig.cap="The raw data for 50 households are shown.
@@
 -->
+
+\noindent The aim of this section is to illustrate that the proposed methodology can be used to understand repetitive behavior for several customers together.
 
@@
 When we use our methodologies, it is based on all significant granularities and not only hod, but if hod is significant we can expect that the grouping would be similar to how we have initially chosen the set of customers.
-This is process is similar to instance-wise classfication methods.By this instancewise discriminative learning, instances can reasonably distribute in the representation space, which reveals their similarities.
-In
-order to further improve visual representations, we propose a
-dual-level progressive similar instance selection (DPSIS) method
-to build a bridge from instance to class by selecting similar
-instances (neighbors) for each instance (anchor) and treating the
-anchor and its neighbors as the same class.
-
-Fig \ref{fig:data-heatmap-hod}, \ref{fig:} and \ref{fig:} shows the distribution of 24 customers across hour-of-day, month-of-year and wknd-wday respectively.
+This procedure is analogous to instance-wise classfication methods.
+Instances may be distributed in the representation space in a reasonable manner, revealing their similarities, using this instance wise discriminative learning.
+A dual-level progressive similar instance selection (DPSIS) approach could also be used that selects similar instances (neighbours) for each instance (anchor) and treats the anchor and its neighbours as the same class.
+
+<!-- Fig \ref{fig:data-heatmap-hod}, \ref{fig:} and \ref{fig:} shows the distribution of 24 customers across hour-of-day, month-of-year and wknd-wday respectively.
 Every row in \ref{fig:} shows different shapes across hour-of-day and across columns show similar shapes for each row.
 We use our methodology to see if the customers are allocated to same group have similar shapes across one or more significant granularity.
+-->
 
@@
 
-```{r data-heatmap-hod, fig.cap="The distribution of electricity demand for the selected 24 customers across hour-of-day.
+```{r data-heatmap-hod-new, fig.cap="The distribution of electricity demand for the selected 24 customers across hour-of-day.
 The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-According to our instance selection method, each row represnts a distinct shape of daily load, and all customers in the same row should have similar daily profile.After clustering these consumers, all customers with the same colour represent the same group.
+According to our prototype selection method, each row represents a distinct shape of daily load, and all customers in the same row should have similar daily profile.After clustering these consumers, all customers with the same colour represent the same group.
 Except for customer id 8269176, there is unanimity across design and groups.
@@
         axis.ticks.y=element_blank()) +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") +
+    theme(strip.text = element_text(size = 10,
+                                    margin = margin(b = 0, t = 0))) +
   ylab("demand (in Kwh)")  + 
@@
 ```
+
 ```{r data-heatmap-moy, fig.cap = "The distribution of electricity demand for the selected 24 customers across month-of-year.
-The median is represented by a line and shaded region represents the area between 25^{th} and 75^{th} percentile.
-The clustering leads to customers with similar shape across month-of-year with some exceptions.
-This is natural in a clus"}
+The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
+Intriguingly, customer id 8269176 appears to be in the appropriate group (Group 3) because its monthly profile is more similar to Group 3 than Design 4.
+So, while our clustering approach failed to place the customer in the correct hour-of-day group, it did so for month-of-year.
+When contrasting the moy to hod profile, there are greater behavioural differences across customers within a group."}
 
@@
 data_heatmap_moy %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, 
+                  group=customer_id, 
+              fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+    theme(strip.text = element_text(size = 10,
+                                    margin = margin(b = 0, t = 0))) + xlab("month-of-year") +
+  ylab("demand (in Kwh)")  +
+  theme_bw() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank()) + 
+  theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )+
+  scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+  xlab("month-of-year") +  theme(axis.text.x = element_text(size=5))
+```
+
+## Characterising clusters
+
+```{r data-heatmap-hod-group}
+
+data_group <- data_pick %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id"))
+
+data_heatmap_hod_group <- quantile_gran(data_group,
+                                  gran1="hour_day",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
+
+# data_heatmap_hod_group <- data_heatmap_hod_group %>%
+#   mutate(group = case_when(group==1 ~"Group 2",
+#                            group==2 ~"Group 4",
+#                            group==3 ~"Group 1",
+#                            group==4 ~"Group 3"))
+
+# data_heatmap_hod_group$group <- factor(data_heatmap_hod_group$group, levels = c("Group 1", "Group 2", "Group 3", "Group 4"))
+
+data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")
+
+hod_group <- data_heatmap_hod_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=group,
+                  fill = as.factor(group), alpha = 0.5),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=group, 
+                color = as.factor(group)), size = 1)+
+  facet_wrap(~group, 
+             scales = "free_y",  
+             nrow = 4) + 
+              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
+  ylab("demand (in Kwh)") + 
+  theme_bw() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank())+
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ 
+  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
+  theme(legend.position = "bottom")
+
+#
+
+# 
+# data_pick_cust <- data_pick_cust %>% mutate(value = as.character(value))
+# 
+# data_pick_cust %>% left_join(cluster_result, by = c("value" = "id")) %>% group_by(group)
+```
+
+Characterization of clusters is the final stage of a cluster analysis.
+If we are convinced that we have identified a collection of clusters that can be distinguished from one another, we should intend to characterize them more formally, both statistically and qualitatively.
+We quantify them by producing statistics for each variable.
+We may look at the findings in graphs, and enhance our qualitative descriptions of the groupings.
+
+\noindent We discover 4 qualitative clusters of varying shapes in the distribution of all consumers in the first panel of Figure \ref{fig:combined}.
+Group-1 includes consumers who work 9-5, get up and conduct morning activities from 7-10am, and then depart.
+Then they return home in the evening to cook supper and perform other activities, giving the evening a greater peak than the morning.
+Group 2 is the group that rushes out of the house in the morning to get to work.
+They only return at night and do all activities at night, so there is no morning peak.
+The third one has a strong early morning and late night hours.
+These consumers may be flexible students or elderly retirees who are night owls.
+Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
+All of this may be validated with further information about the customer.
+
+\noindent The second panel of Figure \ref{fig:combined} shows that month-of-year qualitative clusters are not as distinguishable as hour-of-day.
+Group 2 is the most distinct and uses the most power during the summer, possibly owing to the use of air conditioners.
+Group 4 has a flat profile, indicating no significant month-to-month changes.
+Groups 1 and 3 have heaters on in the winter but consume less energy in the summer.
+Since gas is not available in all of NSW LGAs, it is possible that customers' heater usage is recorded in electricity rather than gas.
+
+\noindent The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
+
+\noindent The plotting scales are not displayed since we want to emphasise comparable shapes rather than scales.
+A customer in the cluster may have low daily or total energy usage, but their behaviour may be quite similar to a customer with high usage.
+That places them in the same group.
+
+```{r data-heatmap-moy-group}
+
+data_group <- data_pick %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id"))
+
+data_heatmap_moy_group <- quantile_gran(data_group,
+                                  gran1="month_year",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")
+
+moy_group <- data_heatmap_moy_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow = 4) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
+  ylab("demand (in Kwh)") +
+  theme_bw() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank()) + 
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
+   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
+  theme(legend.position = "bottom") +
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
+```
+
+```{r data-heatmap-wkndwday}
+
+data_heatmap_wkndwday <- quantile_gran(data_pick, "wknd_wday", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
+  left_join(data_pick_cust, by = c("customer_id"="value"))
+
+data_heatmap_wkndwday$category <- factor(data_heatmap_wkndwday$category, levels = c("Weekday", "Weekend"))
+
+p1 <- data_heatmap_wkndwday %>% 
   mutate(customer_id = as.character(customer_id)) %>% 
@@
              ncol = 6) +
-    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("wknd_wday") + ylab("demand (in Kwh)")  + theme_bw() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank()) +
+  theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
     strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
-```
-
-## Characterising clusters
-
-```{r data-heatmap-hod-group, fig.cap="The distribution of electricity demand for the clusters across hour-of-day.
-The median is represented by a line and the shaded region represents the area between 25^{th} and 75^{th} percentile.
-The first cluster is characterised by high comsumption in early morning hours and late night hours.
-The second cluster should be the most common type of daily profile characterised by a smaller morning and higher evening peak.
-The third cluster corresponds to two equal peaks and the last one corresponds to only distinct evening peaks."}
-
-data_group <- data_pick %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result, by = c("customer_id" = "id"))
-
-data_heatmap_hod_group <- quantile_gran(data_group,
-                                  gran1="hour_day",
-                                  quantile_prob_val = c(0.25, 0.5, 0.75),
-                                  group="group") %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values)
-
-data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
-
-# data_heatmap_hod_group <- data_heatmap_hod_group %>%
-#   mutate(group = case_when(group==1 ~"Group 2",
-#                            group==2 ~"Group 4",
-#                            group==3 ~"Group 1",
-#                            group==4 ~"Group 3"))
-
-# data_heatmap_hod_group$group <- factor(data_heatmap_hod_group$group, levels = c("Group 1", "Group 2", "Group 3", "Group 4"))
-
-data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")
-
-data_heatmap_hod_group %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`,
-                  group=group,
-                  fill = as.factor(group)),
-              alpha = 0.5) +
-  geom_line(aes(y = `50%`,
-                group=group, 
-                color = as.factor(group)), size = 1)+
-  facet_wrap(~group, 
-             scales = "free_y",  nrow = 4) + 
-              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + 
-  theme_bw() + 
-  theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
-  scale_x_discrete(breaks = seq(1, 24, 3))+ 
-  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") ) +
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
@@
 
-#
-
-# 
-# data_pick_cust <- data_pick_cust %>% mutate(value = as.character(value))
-# 
-# data_pick_cust %>% left_join(cluster_result, by = c("value" = "id")) %>% group_by(group)
-```
-
-Characterization of clusters is the final stage of a cluster analysis.
-However, if we are convinced that we have identified a collection of clusters that can be distinguished from one another, we should intend to characterise them more formally, both statistically and qualitatively.
-We quantify them by producing statistics for each variable.
-We may look at the findings in graphs, and enhance our qualitative descriptions of the groupings.
-
-We discover 4 qualitative clusters of varying shapes in the distribution of all consumers in a group.
-Group-1 includes consumers who work 9-5, get up and conduct morning activities from 7-10am, and then depart.
-Then they return home in the evening to cook supper and perform other activities, giving the evening a greater peak than the morning.
-Group 2 is the group that rushes out of the house in the morning to get to work.
-They only return at night and do all activities at night, so there is no morning peak.
-The third one has a strong early morning and late night hours.
-These consumers may be flexible students or elderly retirees who are night owls.
-Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening peaks.
-All of this may be validated with further information about the customer.
-
-```{r data-heatmap-moy-group}
-
-data_group <- data_pick %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result, by = c("customer_id" = "id"))
-
-data_heatmap_moy_group <- quantile_gran(data_group,
-                                  gran1="month_year",
-                                  quantile_prob_val = c(0.25, 0.5, 0.75),
-                                  group="group") %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values)
-
-data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
-
-data_heatmap_moy_group %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=group, fill = as.factor(group))) +
-  geom_line(aes(y = `50%`, group=group)) +
-  facet_wrap(~group, 
-             scales = "free_y", 
-             labeller = "label_value",
-             nrow = 4) +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
-
-```
-The clusters across month-of-year shows that cluster 2 uses electricity in summer.
-
-```{r data-heatmap-wkndwday}
-
-data_heatmap_wkndwday <- quantile_gran(data_pick, "wknd_wday", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
-  left_join(data_pick_cust, by = c("customer_id"="value"))
-
-data_heatmap_wkndwday$category <- factor(data_heatmap_wkndwday$category, levels = c("Weekday", "Weekend"))
-
-data_heatmap_wkndwday %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=customer_id), fill = "lightblue") +
-  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
-  facet_wrap(design~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-             ncol = 6) +
-    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("wknd_wday") + ylab("demand (in Kwh)")  + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
-
-```
-
-```{r data-heatmap-wkndwday-group}
-
+```
+
+```{r data-heatmap-wkndwday-group-new}
 data_group <- data_pick %>% 
@@
 
-data_heatmap_wkndwday_group %>% 
+data_heatmap_wkndwday_group$group <- paste("group", data_heatmap_wkndwday_group$group, sep = "-")
+
+wkndwday_group <- data_heatmap_wkndwday_group %>% 
   ggplot(aes(x = category)) + 
   geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=group, fill = as.factor(group))) +
-  geom_line(aes(y = `50%`, group=group)) +
+                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
   facet_wrap(~group, 
@@
              labeller = "label_value",
-             ncol = 4) +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("wknd_wday") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+             nrow = 4)+
+    #theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) 
+    xlab("wknd-wday") + 
+  ylab("demand (in Kwh)") +
+  theme_bw() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank()) + 
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
+   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
+  theme(legend.position = "bottom") +
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
+```
+
+```{r combined, fig.cap="The distribution of electricity demand for the clusters across hour-of-day.
+The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
+Each cluster is characterised by unique shape across the granularity it is plotted against.
+For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters.
+This fact we will be re-established when we see the importance of each granularity throough the parallel coordinate plot."}
+
+combined <- hod_group + moy_group + wkndwday_group & theme(legend.position = "bottom")
+combined + plot_layout(guides = "collect")
 

### 2021-10-09 Simulations.Rmd
application update

@@
 
-```{r data-heatmap-hod-group}
+```{r data-heatmap-hod-group2}
 
@@
 parcoord
-
 ```
@@
   facet_wrap(~group, 
-             scales = "free_y", 
+             #scales = "free_y", 
              labeller = "label_value",
@@
 
-```{r data-heatmap-moy-group}
+```{r data-heatmap-moy-group2}
 
@@
 
-```{r data-heatmap-wkndwday-group}
+```{r data-heatmap-wkndwday-group2}
 

### 2021-10-09 main.Rmd
application update

@@
 library(patchwork)
+library(GGally)
 ```
@@
 
-## Characterising clusters
+##  js-based clustering
+
+The distribution of electricity demand for the selected 24 customers across hour-of-day is shown in \ref{fig:data-heatmap-hod-new}.
+The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
+According to our prototype selection method, each row represents a distinct shape of daily load, and all customers in the same row should have similar daily profile.After clustering these consumers, all customers with the same colour represent the same group.
+Except for customer id 8269176, there is unanimity across design and groups.
+Even though their daily form resembles Group 2, our clustering approach places them in Group 3 (which is design 4 in our case).
+Because our method uses hod, moy, and wkndwday, there may be some mismatches depending on only one variable.
+
+\noindent The distribution of electricity demand for the selected 24 customers across month-of-year is shown in Figure \ref{fig:data-heatmap-moy}.
+The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
+Intriguingly, customer id 8269176 appears to be in the appropriate group (Group 3) because its monthly profile is more similar to Group 3 than Design 4.
+So, while our clustering approach failed to place the customer in the correct hour-of-day group, it did so for month-of-year.
+When contrasting the moy to hod profile, there are greater behavioural differences across customers within a group.
 
@@
   facet_wrap(~group, 
-             scales = "free_y",  
+             #scales = "free_y",  
              nrow = 4) + 
@@
 
-### Cluster validation
-
-**Internal cluster validation** uses the internal information of the clustering process to evaluate the goodness of a clustering structure without reference to external information.
-It can be also used for estimating the number of clusters and the appropriate clustering algorithm without any external data.
-
-Generally most of the indices used for internal clustering validation combine compactness and separation measures.
-
-<!-- _Distinction, repeatability, and robustness metrics_ -->
-
-We will also determine if any identified clusters or patterns are indeed statistically meaningful in the sense that they actually exist and are not a random allocation.
-Hence, the robustness of this methodology is tested through simulations.
-This section will contain the data structure and detailed methodology to be employed for the cluster analysis.
-The cluster validation indexes like average silhouette width (ASW) is to be employed here to check how homogeneous these clusters are.
-
-**External cluster validation** consists in comparing the results of a cluster analysis to an externally known result, such as externally provided class labels.
-It measures the extent to which cluster labels match externally supplied class labels.
-Since we know the “true” cluster number in advance, this approach is mainly used for selecting the right clustering algorithm for a specific data set.
-
-**Relative cluster validation** evaluates the clustering structure by varying different parameter values for the same algorithm (e.g.,: varying the number of clusters k).
-It’s generally used for determining the optimal number of clusters.
-
-Interval validation includes
-_Compactness or cluster cohesion:_ Measures how close are the objects within the same cluster.
-A lower within-cluster variation is an indicator of a good compactness (i.e., a good clustering).
-The different indices for evaluating the compactness of clusters are base on distance measures such as the cluster-wise within average/median distances between observations.
-
-_Separation:_ Measures how well-separated a cluster is from other clusters. The indices used as separation measures include:
-distances between cluster centers
-the pairwise minimum distances between objects in different clusters
-
-_Connectivity:_ corresponds to what extent items are placed in the same cluster as their nearest neighbors in the data space.
-The connectivity has a value between 0 and infinity and should be minimized.
-
-## Clustering results for 100 customers
-
-## Clustering results for 5K customers
-
-<!-- The robustness of this clustering method is provided through practical explanation of the formed clusters, visualizing how they relate to any weather, socio-economic or geographical conditions.
--->
-<!-- Furthermore, we ensure that a cross-validation algorithm is employed leading to robust clusters.  -->
-<!-- We will present how matching other data like geography, census and weather would help us in validating our clustering approach. -->
-
-<!-- ### Geography -->
-
-<!-- We will present how mapping the customers to their respective LGA's and reflect on the clusters obtained in the earlier section.
-The data is at SA2 and LGA levels.
-However, some of the LGAs in NSW changed between 2011 and 2016 and hence there would not be a one-to-one correspondence between the LGAs.
--->
-
-<!-- ### Census -->
-
-<!-- We will present how mapping the customers to their respective income and population level reflect on the clusters obtained in the earlier section.
-The [ABS TableBuilder](https://www.abs.gov.au/websitedbs/censushome.nsf/home/tablebuilder) has census data from 2011 and 2016.
--->
-
-<!-- ### Weather -->
-
-<!-- Weather, notably temperature (and humidity) can be the main driver(s) for energy usage.
-In NSW many households have electric heaters so their use can impact winter energy use and air-conditioners can impact summer energy use.
-Relevant weather data could be obtained from the [Bureau of Meteorology](www.bom.gov.au).
-Some weather stations have 30 minute (sometimes even smaller interval) weather data.
-Potentially, there could be lag effects of weather on energy usage which should be considered.
--->
-
-## Software implementation
-
-The implementation for our framework is available in the R package [`gracsr`](https://github.com/Sayani07/gracsr) for ease of use in other applications.
-
-## Discussion
-
-This section will cover some drawback of this clustering method and potential extensions of this work.
-
-<!-- <Earlier write up> -->
-<!-- ## Introduction -->
-
-<!-- <!-- Description of data available --> -->
-<!-- The Smart Grid, Smart City (SGSC) project, which rolled out Australia's first commercial-scale smart grid was implemented across eight local government areas in New South Wales (NSW).
-Data from more than 13,000 household electricity smart meters is obtained as part of that project.
-It provides half-hourly energy usage and demographic data for Australia, as well as detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
-The trials were based in Newcastle, New South Wales, but also covered areas in Sydney CBD, Newington, Ku-Ring-Gai, and the rural township of Scone.
-The load time series is asynchronous as it is observed for these households for unequal time lengths and consists of missing observations.
--->
-
-<!-- The massive amount of data generated in such projects could be overwhelming for analysis.
-Electricity utilities can utilize the consumption patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
-Beyea (2010) has pointed out, there has been little discussion or exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data.
-Thus, there is a scope to investigate and analyze these data in various ways for a greater understanding of consumption patterns and how they correlate with other economic, physical or geographical factors.
-In this work, we are interested to see how we can utilize this dataset to group different customers with similar periodic behavior.
-Towards this goal, this chapter aims to: (a) describe the contents of the data set in SGSC database that we can utilize, and (b) propose a clustering algorithm to group customers with similar periodic behaviors.
-The distance metric introduced in Chapter 2 will be the inputs for this cluster analysis.
-One of the advantages of using our  -->
-<!-- approach is that the technique is based on probability distributions instead of raw data.
-Many clustering approaches are limited by the type of noisy, patchy, and unequal time-series common in residential data sets.
-Since the distance measure considered is based on differences in probability distribution of time series, it is likely to be less sensitive to missing or noisy data.
--->
-
-<!-- The following contributions are made through the following chapter: -->
-
-<!--  * Present a cluster analysis of SGSC dataset to group households with similar periodic behavior -->
-<!--  * Cluster validation by relating to external data -->
-
-<!-- ## Introduction -->
-
-<!-- In Australia, the federal government and fourteen partners, electricity distribution and transmission companies, technology companies, universities and CSIRO, invested AUD 490 million in the Smart Grid Smart City (SGSC) field trial.
-One of the largest commercial-scale smart grid technology assessment projects worldwide, SGSC is implemented across eight local government areas in New South Wales (NSW), covering 30,000 dwellings since 2009.
--->
-<!-- The SGSC database is publicly available through the Information Clearing House (ICH) (Ausgrid 2014).
-It is a massive database covering most physical aspects of the smart grid, as well as, aspects of users’ demographics.
--->
-
-<!-- Customer Sale and Half-Hour Power Consumption -->
-<!-- and Generation Datasets -->
-
-<!-- This is a big portion of the data from about 5900 households who accepted a -->
-<!-- product out of about 24,000 offers made to the target population. Spatial resolution -->
-<!-- down to anonymous individual dwellings [Customer Sales: Households: -->
-<!-- Customer_Key] is also given based on {[Postcode], [Local Government Area], or -->
-<!-- [Suburb Name]}. In order to make these datasets easily understandable, here we -->
-<!-- present a new narrative approach to explain the available data fields and their -->
-<!-- interdependencies in a brief yet logical and memorisable way as follows: -->
-
-<!-- The [Half-Hour Consumption and Generation] data provide kWh readings of -->
-<!-- the actual flow of energy. This includes the electricity supply imported from the -->
-<!-- grid [General Supply] which is the main source of energy under any of the applied -->
-<!-- tariffs, import during off-peak or for a control load [Off Peak], gross or net values of -->
-
-<!-- domestic (often solar) electricity that is exported back to the grid [Generation]. The -->
-<!-- [Peak Events] data field gives the details of every event [Event Key] including its -->
-<!-- [Type], [Date], [Start Time], and [End Time]. Accordingly, the entire sequence of -->
-<!-- user demand responses to any of the cost-reflective products is available in the -->
-<!-- dataset to evaluate the project objectives as previously discussed. -->
-
-<!-- Electricity smart meter technology is increasingly being deployed in residential and commercial buildings.
-For example, in the state of Victoria, Australia, it is a state government policy that all households have a smart meter fitted (Victoria State Government, 2015), which has resulted in 97% of the approximately 2.1 million households in Victoria having a smart meter installed in their home.
-These smart meters collect energy usage information at half-hourly intervals, resulting in over 35 billion half-hourly observations per year across all households in the state.
-Clustering electricity consumption patterns can enable electricity utilities to develop targeted tariffs for individual groups alleviating the problem of volatility in production by capitalizing on the flexibility of consumers.
-Clustering households using only smart meter consumption data should provide value in a societal setting by combining the findings and external data like weather conditions, socio-economic or other demographic factors of those households.
--->
-
-<!-- <!-- Description of data available --> -->
-<!-- Data from more than 13,000 household electricity smart meters is obtained as part of the Smart Grid Smart City (SGSC) project (2010-2014).
-It provides one of the few linked sets of customer time of use (half-hour increments) and demographic data for Australia, as well as detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
-The trials were based in Newcastle, New South Wales, but also covered areas in Sydney CBD, Newington, Ku-Ring-Gai, and the rural township of Scone.
-This data set includes electricity use interval readings for each customer for every 30 minutes and demographic data for Australia, as well as detailed information on appliance use, climate, peak events; peak events response and retail and distributor product offers.
-The load time series is asynchronous as it is observed for these households for unequal time lengths and consists of missing observations.
--->
-
-<!-- <!-- Background literature and shortcomings --> -->
-<!-- The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics.
-The paper evaluates approximately 2100 unique peer-reviewed papers and presents three main findings related to clustering methods, data and cluster validation.
-None of the 34 selected papers which focus on clustering consumption are based on Australian smart meter data.
-The clustering is frequently applied directly to the raw data without investigating time series features like auto correlation and periodicity.
-The application of smart meter data to cluster electricity consumption using K-Means is ubiquitous.
-But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters.
-Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results.
-@Tureczek2018-ha has shown that the clever transformation of data before K-Means clustering can improve performance and enable K-Means.
-However, it does not explain the cluster composition by combining it with external data.
-We propose the following for the study of the Australian smart meter data.
--->
-
-<!-- Analysis for more than 13,000 household -->
-<!-- electricity smart meters are included in this chapter. The following contributions are made through the following chapter: -->
-
-<!-- - Presenting a cluster analysis of Australian residential electricity consumption data. -->
-
-<!-- - Transformation and extraction of input data features enabling K-Means to account for other time series features in the clustering.
-This can easily be extended to include other data structures.
--->
-
-<!-- - Cluster validation by relating to external data -->
-
-<!-- ## Preliminary exploratoration -->
-
-<!-- This section familiarizes the 13,000 SGSC households through visualization and provide a detailed layout of data structures (like missing observations) and also the external data that needs to be utilized for validating the clustering process.
--->
-
-<!-- ## Common load clustering techniques of smart meter data -->
-
-<!-- This section will discuss some literature on clustering techniques that were applied to smart meter customers in the past. -->
-
-<!-- ## The proposed load-clustering technique -->
-
-<!-- We employ a cluster analysis technique to obtain pockets of households showing similar periodic behavior by grouping probability distributions across a cyclic granularity.
-This clustering algorithm is adopted to remove or appropriately adjust for auto correlation and unequal length in the data.
-The method could be further extended by clustering probability distributions conditional on one or more cyclic granularities.
-Further, we elaborate on the effect of missing data and how it is handled in the clustering process.
--->
-
-<!-- ### Methodology, results, and discussions -->
-
-<!-- This will contain the data structure and detailed methodology to be employed for the cluster analysis. -->
-
-<!-- ### Distinction, repeatability, and robustness metrics -->
-
-<!-- This section will cover the homogeneity and heterogeneity of the clusters, how distinct they are and summarize the main features of the cluster. -->
-
-<!-- <!-- ## Comparison with traditional load clustering methods --> -->
-
-<!-- <!-- This section will cover the homogeneity and heterogeneity of the clusters, how distinct they are and summarize the main features of the cluster. --> -->
-
-<!-- ## Combining findings with external data -->
-
-<!-- We will also determine if any identified clusters or patterns are indeed statistically meaningful in the sense that they actually exist and are not a random allocation.
-We aim to do this by providing a practical explanation of the formed clusters, visualizing how they relate to any weather, socio-economic or geographical conditions.
-Furthermore, we ensure that a cross-validation algorithm is employed leading to robust clusters.
--->
-<!-- We will present how matching other data sources might help us in validating our clustering approach. -->
-
-<!-- ### Geography -->
-
-<!-- We will present how mapping the customers to their respective LGA's and reflect on the clusters obtained in the earlier section. -->
-
-<!-- ### Census -->
-
-<!-- We will present how mapping the customers to their respective income and population level reflect on the clusters obtained in the earlier section. -->
-
-<!-- ## Conclusion -->
+##  wpd-based clustering
+
+```{r ggpairs, fig.cap = "The ggpairs plot of electricity demand against the three granularities hod, moy and wkndwday are shown.
+The correlation between any two granularities is not so high so as to pose a threat to the clustering methodology."}
+
+elec_600_wpd <- read_rds(here::here("data/algo2-cust600-wpd-rawdata.rds"))
+
+elec_pick <- elec_600_wpd %>% 
+  filter(customer_id %in% data_pick_cust$value)
+
+elec_pick_wide <- elec_pick %>% pivot_wider(-c(1, 2), names_from = "x_variable", values_from = wpd)
+
+```
+
+```{r, fig.cap=" A ggpairsplot and parallel coordinate plot are used to depict each of the 24 customers.
+The ggpairs plot four distinct clusters across the month-of-year, which are less prominent across the hour-of-day and wknd-wday.
+The parallel coordinate plot ranks the variables in order of importance, indicating that the month-of-year is the most important in identifying clusters, whereas wkdn-wday is the least significant and has the least variability among the three variables." }
+
+# robust_scale <-  function(x){
+#   (x - stats::median(x))/(stats::IQR(x))
+# }
+
+#scaled_var <- apply(elec_pick_wide[-c(1)], 2, robust_scale) %>% as_tibble()
+
+scaled_var <- elec_pick_wide
+
+f <- elec_pick_wide[-1] %>% dist()
+
+group <- f%>% hclust (method = "ward.D") %>% cutree(k=4)
+
+cluster_result_wpd <- bind_cols(id = elec_pick_wide$customer_id, group = group)
+
+mds  = cmdscale(f)
+rownames(mds) = id
+colnames(mds) = c("mds1", "mds2")
+
+# ggplot(data = mds %>% as_tibble() %>% mutate(group = cluster_result_wpd$group), 
+#        aes(x = mds1, y = mds2, color = as.factor(group))) +
+#   geom_point() +
+#   geom_text(aes(label = id))
+
+data_pcp <- scaled_var %>% 
+  #bind_cols(customer_id =  elec_pick_wide$customer_id) %>%
+  left_join(cluster_result_wpd , by = c("customer_id" = "id")) %>% 
+  select(customer_id, group, everything()) %>% 
+  mutate(group = as.factor(group))
+
+pairsplot <- ggpairs(data_pcp, 
+                     columns = 3:5,
+                     aes(fill=group, color = group), alpha = 0.5) +
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))
+
+parcoord <- GGally::ggparcoord(data_pcp ,
+                   columns = 3:ncol(data_pcp),
+                   groupColumn = "group",
+                   showPoints = FALSE, 
+                   alphaLines = 0.8,
+                   order = "anyClass",
+                   scale = "globalminmax"
+) + ggplot2::theme_bw() +
+  scale_color_brewer(palette = "Dark2") + 
+  ggplot2::theme(
+    plot.title = ggplot2::element_text(size=10)
+  )+
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
+  theme(legend.position = "bottom") +
+  xlab("") +
+  ylab("wpd")  +   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))
+pairsplot
+```
+
+```{r}
+parcoord
+```
+
+```{r data-heatmap-hod-group-wpd}
+
+cluster_result_wpd$id <- as.character(cluster_result_wpd$id)
+
+data_group <- data_pick %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result_wpd, by = c("customer_id" = "id"))
+
+data_heatmap_hod_group <- quantile_gran(data_group,
+                                  gran1="hour_day",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
+
+# data_heatmap_hod_group <- data_heatmap_hod_group %>%
+#   mutate(group = case_when(group==1 ~"Group 2",
+#                            group==2 ~"Group 4",
+#                            group==3 ~"Group 1",
+#                            group==4 ~"Group 3"))
+
+# data_heatmap_hod_group$group <- factor(data_heatmap_hod_group$group, levels = c("Group 1", "Group 2", "Group 3", "Group 4"))
+
+data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")
+
+hod_group_wpd <- data_heatmap_hod_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=group,
+                  fill = as.factor(group), alpha = 0.5),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=group, 
+                color = as.factor(group)), size = 1)+
+  facet_wrap(~group, 
+             #scales = "free_y",  
+             nrow = 4) + 
+              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
+  ylab("demand (in Kwh)") + 
+  theme_bw() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank())+
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ 
+  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
+  theme(legend.position = "bottom")
+
+#
+
+# 
+# data_pick_cust <- data_pick_cust %>% mutate(value = as.character(value))
+# 
+# data_pick_cust %>% left_join(cluster_result, by = c("value" = "id")) %>% group_by(group)
+```
+
+```{r data-heatmap-moy-group-wpd}
+
+data_group <- data_pick %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result_wpd, by = c("customer_id" = "id"))
+
+data_heatmap_moy_group <- quantile_gran(data_group,
+                                  gran1="month_year",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")
+
+moy_group_wpd <- data_heatmap_moy_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+  facet_wrap(~group, 
+             #scales = "free_y", 
+             labeller = "label_value",
+             nrow = 4) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
+  ylab("demand (in Kwh)") +
+  theme_bw() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank()) + 
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
+   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
+  theme(legend.position = "bottom") +
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
+```
+
+```{r data-heatmap-wkndwday-group-new2-wpd}
+data_group <- data_pick %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result_wpd, by = c("customer_id" = "id"))
+
+data_heatmap_wkndwday_group <- quantile_gran(data_group,
+                                  gran1="wknd_wday",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_wkndwday_group$category <- factor(data_heatmap_wkndwday_group$category, levels = c("Weekday", "Weekend"))
+
+data_heatmap_wkndwday_group$group <- paste("group", data_heatmap_wkndwday_group$group, sep = "-")
+
+wkndwday_group_wpd <- data_heatmap_wkndwday_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+  facet_wrap(~group, 
+             #scales = "free_y", 
+             labeller = "label_value",
+             nrow = 4)+
+    #theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) 
+    xlab("wknd-wday") + 
+  ylab("demand (in Kwh)") +
+  theme_bw() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank()) + 
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
+   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
+  theme(legend.position = "bottom") +
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
+```
+
+```{r combined-wpd, fig.cap="The distribution of electricity demand for the clusters across hour-of-day.
+The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
+Each cluster is characterised by unique shape across the granularity it is plotted against.
+For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters.
+This fact we will be re-established when we see the importance of each granularity throough the parallel coordinate plot."}
+
+combined <- hod_group_wpd + moy_group_wpd + wkndwday_group_wpd & theme(legend.position = "bottom")
+combined + plot_layout(guides = "collect")
+
+```
+
+\noindent We discover 4 qualitative clusters of varying shapes in the distribution of all consumers in the first panel of Figure \ref{fig:combined}.
+Group-1 includes consumers who work 9-5, get up and conduct morning activities from 7-10am, and then depart.
+Then they return home in the evening to cook supper and perform other activities, giving the evening a greater peak than the morning.
+Group 2 is the group that rushes out of the house in the morning to get to work.
+They only return at night and do all activities at night, so there is no morning peak.
+The third one has a strong early morning and late night hours.
+These consumers may be flexible students or elderly retirees who are night owls.
+Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
+All of this may be validated with further information about the customer.
+
+\noindent The second panel of Figure \ref{fig:combined} shows that month-of-year qualitative clusters are not as distinguishable as hour-of-day.
+Group 2 is the most distinct and uses the most power during the summer, possibly owing to the use of air conditioners.
+Group 4 has a flat profile, indicating no significant month-to-month changes.
+Groups 1 and 3 have heaters on in the winter but consume less energy in the summer.
+Since gas is not available in all of NSW LGAs, it is possible that customers' heater usage is recorded in electricity rather than gas.
+
+\noindent The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
+
+\noindent The plotting scales are not displayed since we want to emphasise comparable shapes rather than scales.
+A customer in the cluster may have low daily or total energy usage, but their behaviour may be quite similar to a customer with high usage.
+That places them in the same group.
+
+# Discussion
 

### 2021-10-10 main.Rmd
minor changes in simulation section

@@
 
-To evaluate clustering approaches, we create data designs that replicate prototype behaviors that might be seen in electricity data contexts.
-We spiked several features in the data to see where one method works better than the other and where they might give us the same outcome or the effect of missing data and trends on the proposed methods.
+To validate the clustering approaches, we create data designs that replicate prototype behaviors that might be seen in electricity data contexts.
+We spiked several features in the data to see where one method works better than the other and where they might give us the same outcome or the effect of missing data <!---and trends--> on the proposed methods.
 A continuous measured variable $y$ of length $T$ indexed by ${0, 1, \dots T-1}$.
@@
 For example, the granularity "day-of-week" could be either represented by ${0, 1, 2, \dots, 6}$ or ${Mon, Tue, \dots, Sun}$.
-Each categories of $g1$, $g2$ and $g3$ could be assumed to be from same distribution, few from one and others from separate distributions, or all from different distributions, resulting in distinct data designs.
-We created independent replications $(R = \{25, 250, 500\})$ of all data designs to see if our proposed clustering approaches can detect distinct designs in various groups for small, medium and large number of series.
+We created independent replications $R = \{25, 250, 500\}$ of all data designs to see if our proposed clustering approaches can detect distinct designs in various groups for small, medium and large number of series.
 A sample size of $T=\{300, 1000, 5000\}$ is used in all designs to test small, medium and large sized series.
-For all data designs, the data type is set to  be "continuous," and a gaussian setup for errors is assumed.
 The method could preform differently with different jumps between consecutive categories.
-So a difference of $diff = {1, 2, 5}$ for corresponding categories are also considered.
+So a difference of $diff = \{1, 2, 5\}$ for corresponding categories are also considered.
 The code for creating these designs can be found in the Supplementary section (link to github repo).
@@
 
-An ARMA (p,q) process is used to generate series, where $p$ and $q$ are selected at random such that the series is stationary.
+<!-- An ARMA (p,q) process is used to generate series, where $p$ and $q$ are selected at random such that the series is stationary.
 The various designs on $g1$, $g2$, and $g3$ are introduced by adding matching designs to this series' innovations.
@@
 To eliminate the effect of starting values, the first 500 observations in each series are discarded.
+-->
+
+Each of the categories $g10,g11$, $g20, g21, g22$, and $g30, g31, g32, g33, g34$ might be assumed to come from the same distribution, a subset of them from the same distribution, a subset of them from separate distributions, or all from different distributions, resulting in various data designs.
+As the methods ignores the linear progression of time, there is little value in adding time dependency in the data generating process.
+Time series data include basic properties such as trend, seasonality, and auto-correlation, and it is reasonable to construct a time series using these features.
+However, when examining distributions across categories of cyclic granularities, these time series features are lost or addressed independently by considering all seasonal fluctuations.
+Because the time span during which an entity is observed in order to ascertain its behavior is very short, it is assumed that the time series will remain stationary throughout the observation period.
+The data type is set to "continuous," and the setup is assumed to be gaussian.
+For example, in a null design, all categories are considered to be normal (0,1).
+The mean of the subsequent categories are incremented or decremented from this distribution based on the mean difference between them in non-null designs.
+
+<!-- An ARMA (p,q) process is used to generate series, where $p$ and $q$ are selected at random such that the series is stationary.
+The various designs on $g1$, $g2$, and $g3$ are introduced by adding matching designs to this series' innovations.
+The innovations are considered to have a normal distribution, although they follow the same pattern as the designs.
+To eliminate the effect of starting values, the first 500 observations in each series are discarded.
+-->
 
 ## Data designs
+
+### Individual granularities
 
@@
 That means that for at least one among the to-be-grouped designs, the pattern for each of $g1$, $g2$, and $g3$ will change.
-We consider a _null_ case to represent that categories follow N(0,1) and _alternate_ case to represent different  distributions across categories (as in Table \ref{tab:}) that will lead to different designs (as in Table \ref{tab:}).
+We consider different distributions across categories (as in Table \ref{tab:tab-distribution}) that will lead to different designs (as in Table \ref{tab:tab-design}).
+Figure \ref{fig:plot-3gran} shows the linear and cyclic representation of the simulated variable under these five designs.
+For the consequent data designs, only graphical representations will be provided.
 
 ```{r tab-distribution}
-
-# table_mayan <- tibble(granularity  = )
-
-```
-
-| granularity 	|  Alternate design 	|
-|---	|---	|
-| g1 	| g10 ~ N(0, 1), g11 ~ N(2, 1) 	|
-| g2 	| g21 ~ N(2, 1), g22 ~ N(1, 1), g23 ~ N(0, 1) 	|
-| g3 	| g31 ~ N(0, 1), g32 ~ N(1, 1), <br><br>g33 ~ N(2, 1), g34 ~ N(1, 1), g35 ~ N(0, 1) 	|
-
-A combination of these null and alternate distributions are considered when building the designs as in the table \ref{tab:tab-design}.
-Figure \ref{#fig:plot-linear-3change} shows the linear and cyclic representation of $y$ under these five designs.
+granularity = c("g1", "g2", "g3")
+alternate_design =  c("g10 ~ N(0, 1), g11 ~ N(2, 1)",
+                      "g21 ~ N(2, 1), g22 ~ N(1, 1), g23 ~ N(0, 1)",
+                      "g31 ~ N(0, 1), g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1), g35 ~ N(0, 1)")
+
+tab_dist <- tibble::tibble(granularity = granularity, `alternate distribution` = alternate_design)
+
+knitr::kable(tab_dist, caption = "Alternate distributions of different categories if they deviate from null.")
+```
+
+<!-- A combination of these null and alternate distributions are considered when building the designs as in the table \ref{tab:tab-design}.  -->
 
@@
 design = c("design-1", "design-2", "design-3", "design-4", "design5")
-g1 = c("null", "vary", "null", "null", "vary")
-g2 = c("null", "null", "vary", "null", "vary")
-g3 = c("null", "null", "null", "vary", "vary")
-table <- tibble(design, g1, g2, g3) %>% kable()
+g1 = c("null", "alternate", "null", "null", "alternate")
+g2 = c("null", "null", "alternate", "null", "alternate")
+g3 = c("null", "null", "null", "alternate", "alternate")
+table <- tibble(design, g1, g2, g3) %>% kable(caption = "5 different designs resulting from considering different distributions across categories.")
 table
@@
 
-```{r plot-linear-change , out.width="100%", fig.cap="The linear and cyclic representation of the time series variable y.
-It is not possible to comprehend these patterns across cyclic granularities g1, g2 and g3 or group similar series just by looking at the linear plots."}
+```{r plot-3gran , out.width="100%", fig.cap="The linear (left) and cyclic (right) representation of the measured variable is shown.
+In this scenario, all of $g1$, $g2$ and $g3$ changes across at least one design.
+Also, it is not possible to comprehend these patterns across cyclic granularities or group similar series just by looking at the linear plots."}
 
@@
 
-Consider a case where distribution of $y$ would vary across levels of $g2$ for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
+This is the case where one granularity will remain the same across all designs.
+We consider the case where the distribution of $y$ would vary across levels of $g2$ for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
 So $g3$ is not responsible for distinguishing across designs.
@@
 
-Consider a case where distribution of $y$ would vary across levels of $g2$ for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
-So $g3$ is not responsible for distinguishing across designs.
-
-<!-- **Missing data** -->
-
-<!-- **Trend** -->
-
-<!-- **Outliers** -->
-
-<!-- observations: -->
-
-<!-- designs: -->
+Here only one granularity is responsible for distinguishing the designs. Designs change significantly only for the granularity $g2$. Figure \ref{fig:plot-1gran} shows this.
+
+```{r}
+generate_design <- function(t, mu21, mu22, mu23){
+
+t <- seq(0, t, 1)
+g1 <- t %%2
+g2 <- t %%3
+g3 <- t %%5
+
+# null design
+g1_dnull <- rep( rep(0, each = length(unique(g1))), length.out= length(t))
+g2_dnull <- rep( rep(0, each = length(unique(g2))), length.out= length(t))
+g3_dnull <- rep( rep(0, each = length(unique(g3))), length.out= length(t))
+
+# mean changing across categories in varying ways
+
+g1_dvary <- rep(0, length.out= length(t))
+g21_dvary <- rep(mu21, length.out= length(t))
+g22_dvary <- rep(mu22, length.out= length(t))
+g23_dvary <- rep(mu23, length.out= length(t))
+g3_dvary <- rep(0, length.out= length(t))
+
+design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
+design2 = distributional::dist_normal(g1_dnull + g21_dvary + g3_dnull)
+design3 = distributional::dist_normal(g1_dnull + g22_dvary + g3_dnull)
+design4 = distributional::dist_normal(g1_dnull + g23_dvary + g3_dnull)
+
+data_bind <- tibble::tibble(
+index = t,
+g1 = g1,
+g2 = g2,
+g3 = g3,
+design1 = distributional::generate(design1, times = 1) %>% unlist(),
+design2 = distributional::generate(design2, times = 1) %>% unlist(),
+design3 = distributional::generate(design3, times = 1) %>% unlist(),
+design4 = distributional::generate(design4, times = 1) %>% unlist(),
+#design5 = distributional::generate(design5, times = 1) %>% unlist()
+) %>% 
+  pivot_longer(-c(1, 2, 3, 4), names_to = "design", values_to = "sim_data")
+
+data_bind
+}
+
+mu21= c(2, 0, 0)
+mu22 = c(0, 2, 0)
+mu23 = c(0, 0, 2)
+mu3 = c(0, 0, 0, 0, 0)
+data_bind <- generate_design(t, mu21, mu22, mu23)
+
+data_bind
+
+```
+
+```{r plot-gran, out.width="100%"}
+
+p1 <- ggplot(data_bind,
+             aes(x = index, y = sim_data)) + 
+  geom_line() +
+  xlab("index")+
+  facet_wrap(~design, scales = "free_y",ncol =1) +
+  theme_validation()
+
+p2 <- ggplot(data_bind,
+             aes(x = as.factor(g1), y = sim_data)) + 
+  geom_boxplot(alpha =0.5) + xlab("g1") + 
+  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.8, color = "blue") +
+  theme_validation()
+
+p3 <- ggplot(data_bind, aes(x = as.factor(g2), y = sim_data)) + geom_boxplot(alpha =0.5) + xlab("g2") + theme_bw() +
+  facet_wrap(~design, scales = "free_y",ncol = 1)+ stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.8, color = "blue") + ylab("")+
+  theme_validation()
+
+p4 <- ggplot(data_bind, aes(x = as.factor(g3), y = sim_data)) + geom_boxplot(alpha =0.5) +
+  xlab("g3") + theme_bw()+
+  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.8, colosr = "blue")+ ylab("")+
+  theme_validation()
+
+(p1 + (p2 + p3 + p4)) *
+  theme_validation() + plot_layout(widths = c(2, 1))
+
+```
 
@@
 
-```{r plot-linear , out.width="100%", fig.cap="The linear and cyclic representation of the time series variable y.
-It is not possible to comprehend these patterns across cyclic granularities g1, g2 and g3 or group similar series just by looking at the linear plots."}
+```{r plot-2gran, out.width="100%", fig.cap = "The linear (left) and cyclic (right) representation of the measured variable is shown.
+In this scenario, $g1$, $g2$ would change across atleast one design but $g3$ change remains same across all design.
+Thus $g3$ is not an important variable to distinguish these designs."}
 
@@
 
+### Interaction of granularities
+
+When two granularities of interest interact, the connection between a granularity and the measured variable is determined by the value of the other interacting granularity.
+This happens when the effects of the two granularities on the measured variable are not additive.
+Consider a case with just two interacting granularities $g1$ and $g3$ of interest.
+As opposed to the last case, where we could play with the distribution of $(2+5)$ categories, with interaction we can play with the distribution of $(2*5)$ combination of categories.
+
 | Granularity type                                                   	| # Significant 	| # Replications 	|
@@
 
-## Data source
-
-The entire data is procured from CSIRO.
+The use of our methodology is illustrated on smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
+It  contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
+It also provides demographic data for these customers most of which are missing and not utilized for the purpose of this paper.
+To maintain anonymity, the energy patterns could not be recognised at the person level, but rather by the geographical location of their dwelling and information about their Local Government Area.
+
+<!-- ## Data source -->
+
+<!-- The entire data is procured from CSIRO.
 A subset of this data is also available from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
 It consists of the following data sets.
-_1. CustomerData:_ 78720 customers with 62 variables about them 
-_2. EUDMData:_ 300 billion half-hourly consumption level data  
-_3. OffersData:_ Method of contact to customer to join SGSC customer trial, either door-to-door (D2D) or via Telesales  
-_4. PEResponseData:_ Peak Events response customer wise  
-_5. PETimesData:_ Peak Events time stamps
-
-Only _CustomerData_ and _EUDMData_ are relevant for the clustering goals of this paper.
+-->
+<!-- _1. CustomerData:_ 78720 customers with 62 variables about them  -->
+<!-- _2. EUDMData:_ 300 billion half-hourly consumption level data   -->
+<!-- _3. OffersData:_ Method of contact to customer to join SGSC customer trial, either door-to-door (D2D) or via Telesales   -->
+<!-- _4. PEResponseData:_ Peak Events response customer wise   -->
+<!-- _5. PETimesData:_ Peak Events time stamps   -->
+
+<!-- Only _CustomerData_ and _EUDMData_ are relevant for the clustering goals of this paper.
 _EUDMData_ contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
@@
 To meet the requirements for anonymity preservation, the energy patterns could not be identified at the individual level, but rather by the geographical location of their residence information about their Local Government Area.
+-->
 

### 2021-10-10 3gran_orig.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=gran3_orig
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=16G
+#SBATCH --time=80:00:00
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.5
+Rscript gran3_orig.R
+

### 2021-10-10 gran3_orig.R

@@ -10,37 +10,38 @@ library(here)
 
 
 # fix parameters
- 
+
 # niter
 #seed_len = seq(12345, 12350, 1)
 # number of iterations corresponding to
-# niter <- c(5, 50) #number of series you are clustering
-# nT <-  c(300, 1000, 5000) # length of the time series
-# mean_diff <- c(1, 2, 5) # difference between consecutive categories
+niter <- c(5, 50, 100) #number of series you are clustering
+nT <-  c(300, 1000, 5000) # length of the time series
+mean_diff <- c(1, 2, 5) # difference between consecutive categories
 
 # number of iterations corresponding to
-niter <- c(50) #number of series you are clustering
-nT <-  c(300) # length of the time series
-mean_diff <- c(1) # difference between consecutive categories
+# niter <- c(5) #number of series you are clustering
+# nT <-  c(300, 1000, 5000) # length of the time series
+# mean_diff <- c(1) # difference between consecutive categories
 
-time_series <- c("ar1", "arma22")
+#time_series <- c("ar1", "arma22")
 
 simtable <- expand.grid(mean_diff = mean_diff,
                         niter = niter,
-                        time_series = time_series, 
+                        #time_series = time_series, 
                         nT = nT)
 
+# parallel::mclapply(seq_len(nrow(simtable)), function(scen){
+
 #scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
-scen <- 16
+scen <- 27
+
 
 simj<-simtable[scen,] #Extract row of table
 mean_diffj <- simj$mean_diff
 niterj <- simj$niter
 nTj <- simj$nT
-time_seriesj <- simj$time_series
+# time_seriesj <- simj$time_series
 #seed_lenj <- simj$seed_len
-
-
 
 generate_design <- function(n = 300, #length of time series
                             mu11 = 0,
@@ -121,24 +122,25 @@ generate_design <- function(n = 300, #length of time series
   
   # ar 0.1 ma 0.3
   # list(ar = c(0.8897, -0.4858), ma = c(-0.2279, 0.2488))
+  # 
+  # if(time_seriesj=="arma22"){
+  #   ts_data <- arima.sim(n=n, list(ar = c(0.8897, -0.4858), ma = c(-0.2279, 0.2488)),
+  #                        
[…]

### 2021-10-10 main.Rmd
application clustering colors changed

@@
 
+theme_characterisation <- function() {
+theme_bw() + 
+      theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) +  + theme_bw() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank()) +
+  theme(panel.spacing =unit(0, "lines")) + 
+    theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
+  theme(legend.position = "bottom")
+}
 ```
@@
 
-## Characteristics of raw data
+<!-- ## Characteristics of raw data -->
 
@@
 Some customers' electricity use may be unavailable owing to power outages or improper recording, resulting in implied missing numbers in the database.
-For this data set it was found that out of 13735 customers in total, 8685 customers do not have any implicit missing observations, while the rest 5050 customers had missing values.
-With further exploration, it was found that there is no structure in the missingness, that is missing observations can occur at any time point (see Appendix).
+For this data set it was found that out of 13,735 customers in total, 8,685 customers do not have any implicit missing observations, while the rest 5,050 customers had missing values.
+With further exploration, it was found that there is no structure in the missing-ness, that is missing observations can occur at any time point (see Appendix).
 Moreover, the data for these customers are characterized by unequal length, different start and end dates.
@@
 First, we select the customers which do not have any implicit missing values and filtered their data for the year 2013.
-From this set, we randomly sample a set of 600 customers.
-Then we removed customers for which any of the hod, moy and wkndwday is a clash.
-We further removed customers for which the deciles of the energy consumption is of the same value.
-Finally, we are left with 356 customers from which we wanted to do instance selection.
+From this set, we randomly sample a set of $600$ customers.
+Then we removed customers for which any of the hod (hour-of-day), moy (month-of-year) and wkndwday (weeknd/weekday) is a clash.
+We further removed customers for which all the deciles of the energy consumption is zero.
+These are the customers whose consumption remain mostly flat and is expected to have no interesting repetitive patterns that is our interest of study.
+Finally, we are left with $356$ customers from which we wanted to do instance selection.
 Since this is unlabeled data, there is no way to do external validation of our methodologies.
 Thus, we chose this way to see how well our methodology works in a cleaner data set as this one to classify customers who have similar behavior across all significant granularities.
-Next we select 4 customers which are far apart from each other and 6 customers which are lying closest in distance to each of these 4 customers.
-These selection was done using the granularity hod.
-When we use our methodologies, it is based on all significant granularities and not only hod, but if hod is significant we can expect that the grouping would be similar to how we have initially chosen the set of customers.
-This procedure is analogous to instance-wise classfication methods.
+Next we select $4$ customers which are far apart from each other and $5$ customers which are lying closest in distance to each of these $4$ customers.
+These selection was done using the granularity $hod$.
+When we use our methodologies, it is based on all significant granularities and not only hod, but if $hod$ is significant we can expect that the grouping would be similar to how we have initially chosen the set of customers.
+This procedure is analogous to instance-wise classfication methods that selects similar instances (neighbours) for each instance (anchor) and treats the anchor and its neighbours as the same class.
+
+<!-- .
 Instances may be distributed in the representation space in a reasonable manner, revealing their similarities, using this instance wise discriminative learning.
-A dual-level progressive similar instance selection (DPSIS) approach could also be used that selects similar instances (neighbours) for each instance (anchor) and treats the anchor and its neighbours as the same class.
+A dual-level progressive similar instance selection (DPSIS) approach could also be used -->
 
@@
 
-```{r data-heatmap-wkndwday-group-new}
+```{r data-heatmap-wkndwday-group}
 data_group <- data_pick %>%
@@
 
-data_heatmap_wkndwday_group <- quantile_gran(data_group,
-                                  gran1="wknd_wday",
-                                  quantile_prob_val = c(0.25, 0.5, 0.75),
-                                  group="group") %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values)
-
-data_heatmap_wkndwday_group$category <- factor(data_heatmap_wkndwday_group$category, levels = c("Weekday", "Weekend"))
-
-data_heatmap_wkndwday_group$group <- paste("group", data_heatmap_wkndwday_group$group, sep = "-")
-
-wkndwday_group <- data_heatmap_wkndwday_group %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
-  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+# data_heatmap_wkndwday_group <- quantile_gran(data_group,
+#                                   gran1="wknd_wday",
+#                                   quantile_prob_val = c(0.25, 0.5, 0.75),
+#                                   group="group") %>%
+#   pivot_wider(names_from = quantiles, values_from = quantiles_values)
+# 
+# data_heatmap_wkndwday_group$category <- factor(data_heatmap_wkndwday_group$category, levels = c("Weekday", "Weekend"))
+# 
+# data_heatmap_wkndwday_group$group <- paste("group", data_heatmap_wkndwday_group$group, sep = "-")
+# 
+# 
+# wkndwday_group <- data_heatmap_wkndwday_group %>%
+#   ggplot(aes(x = category)) +
+#   geom_ribbon(aes(ymin = `25%`,
+#                   ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+#   geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+#   facet_wrap(~group,
+#              #scales = "free_y",
+#              labeller = "label_value",
+#              nrow = 4)+
+#     #theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
+#     xlab("wknd-wday") +
+#   ylab("demand (in Kwh)") +
+#   theme_bw() +
+#    theme(axis.title.y=element_blank(),
+#         axis.text.y=element_blank(),
+#         axis.ticks.y=element_blank()) +
+#   theme(panel.spacing =unit(0, "lines")) +
+#   theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
+#    scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+#   scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+#   theme(legend.position = "bottom") +
+#   theme(panel.spacing =unit(0, "lines")) +
+#   theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
+
+wkndwday_group <- data_group %>% create_gran("wknd_wday") %>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  lvplot::geom_lv(aes(fill = as.factor(group))) +
+  #geom_boxplot(aes(fill = as.factor(group))) +
+  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
+#geom_boxplot(aes(fill = as.factor(group))) +
+  #scale_fill_lv() +
+ xlab("wknd-wday") + 
+  ylab("demand (in Kwh)") +
    facet_wrap(~group, 
-             #scales = "free_y", 
+             scales = "free_y", 
              labeller = "label_value",
              nrow = 4)+
-    #theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) 
-    xlab("wknd-wday") + 
-  ylab("demand (in Kwh)") +
   theme_bw() +
@@
   theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
   scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
@@
   theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
+
 ```
@@
 
+library(viridis)
 pairsplot <- ggpairs(data_pcp, 
@@
                      aes(fill=group, color = group), alpha = 0.5) +
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))
+  scale_fill_viridis_d(direction = -1) +
+  scale_color_viridis_d(direction = -1) +
+  theme_light()
+
+  # 
+  # scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))
 
@@
                    scale = "globalminmax"
-) + ggplot2::theme_bw() +
-  scale_color_brewer(palette = "Dark2") + 
+) + 
   ggplot2::theme(
@@
   xlab("") +
-  ylab("wpd")  +   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))
+  ylab("wpd") + scale_fill_viridis_d(direction = -1) +
+  scale_color_viridis_d(direction = -1) + theme_light()
 pairsplot
@@
   #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
+  theme(plot.margin = margin(0, 0, 0, 0, "cm"))+ scale_fill_viridis_d(direction = -1) +
+  scale_color_viridis_d(direction = -1) + theme_light() 
   theme(legend.position = "bottom")
@@
   theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
+scale_fill_viridis_d(direction = -1) +
+  scale_color_viridis_d(direction = -1) + theme_light() + 
   theme(legend.position = "bottom") +
@@
   theme(panel.spacing =unit(0, "lines")) + 
-  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
+  theme(plot.margin = margin(0, 0, 0, 0, "cm"))+ scale_fill_viridis_d(direction = -1) +
+  scale_color_viridis_d(direction = -1) + theme_light() + 
   theme(legend.position = "bottom") +

### 2021-10-11 main.Rmd
introduction why interesting problem

@@
 <!-- Description of data available -->
-The Smart Grid, Smart City (SGSC) project, which rolled out Australia's first commercial-scale smart grid was implemented across eight local government areas in New South Wales (NSW).
-Data from more than 13,000 household electricity smart meters is obtained as part of that project.
-It provides half-hourly energy usage and demographic data for Australia, as well as detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
-The trials were based in Newcastle, New South Wales, but also covered areas in Sydney CBD, Newington, Ku-Ring-Gai, and the rural township of Scone.
-The load time series is asynchronous as it is observed for these households for unequal time lengths and consists of missing observations.
-
-\noindent The massive amount of data generated in such projects could be overwhelming for analysis.
-Electricity utilities can utilize the consumption patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
-Beyea (2010) has pointed out, there has been little discussion or exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data.
-Thus, there is a scope to investigate and analyze these data in various ways for a greater understanding of consumption patterns and how they correlate with other economic, physical or geographical factors.
-In this work, we are interested to see how we can utilize this dataset to group different customers with similar periodic behavior.
-Towards this goal, this chapter aims to: (a) describe the contents of the data set in SGSC database that we can utilize, and (b) propose a clustering algorithm to group customers with similar periodic behaviors.
-The distance metric introduced in Chapter 2 will be the inputs for this cluster analysis.
-One of the advantages of using our
-approach is that the technique is based on probability distributions instead of raw data.
-Many clustering approaches are limited by the type of noisy, patchy, and unequal time-series common in residential data sets.
-Since the distance measure considered is based on differences in probability distribution of time series, it is likely to be less sensitive to missing or noisy data.
-
-_Themes_
-
-- Dimension reduction: If each $P_{i, j, k}$ be considered to be a point in the space, key $i$ would have $mp$ dimensions as opposed to $n_i$ dimensions in case of considering raw data.
-Hence for a large number of observations ($n_i>>mp$), this approach benefits by transitioning to a lower dimension.
-
- - Avoid loss of information due to aggregation: This approach ensures key characteristic information of the data is not lost due to averaging or aggregation measures in an attempt to transition to a lower dimension.
-Hence, this approach could be thought to somehow balance the drawback of considering raw data or aggregated data.
-
-- Robustness to outliers: This approach could be adapted to be robust to outliers and extreme behaviors by trimming the tails of the probability distributions.
-
-- Non-synchronized observed time periods: Considering probability distribution would imply the clustering process can handle keys that are observed over periods of time that are overlapping but don't necessarily coincide.
-
-- Similar periodic behavior: Since cyclic granularities are considered instead of linear granularities, clustering would group keys that have similar behavior across these cyclic granularities.
-This implies they will be grouped according to their periodic behavior and not on the linear stretch of time over which they are observed.
-
-_Common load clustering techniques of smart meter data_
-
+
+<!-- significance of load clustering -->
+Large spatio-temporal data sets, both open and administrative, offer up a world of possibilities for energy and social science research.
+<!--Beyea (2010) has pointed out, there has been little exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data.--> One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
+The project provides half-hourly data from over 13,000 household electricity smart meters for nearly two years.
+<!---The massive amount of data generated in such projects could be overwhelming for analysis.--> Electricity utilities can utilize these smart meter usage patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
+
+<!-- Sources of variation in large data sets -->
+The enormous quantity of data provides for greater individual level clarity and analysis.
+However, due to the growing variety of consumers, larger data sets include greater uncertainty about consumer behavior.
+Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
+Multiple temporal dependencies define the behavioral patterns, which vary from customer to customer.
+Some families, for example, use a dryer to dry their clothing, while others hang them to dry on a line.
+This may be reflected in their weekly profile.
+They may have monthly variations where some customers are more prone to use air conditioners or heaters than others despite the existence of comparable electrical equipment and being subjected to similar weather conditions.
+The variations in behavior may occur on a regular basis, with some consumers being night owls and others being morning larks.
+Day-off habits may vary depending on whether consumers choose to remain at home or engage in outdoor activities.
+These recorded time series are asynchronous, with varying time lengths for different houses and missing observations.
+This huge amount of noisy, patchy, and unequal time-series common in residential smart meter data combined with the different causes of variation in their energy behavior makes the problem of effectively segmenting consumers with comparable consumption a particularly intriguing one.
+
+<!--Unlabelled data -->
+<!-- While forecasting techniques for extremely granular temporal data are readily accessible, -->
+
+Segmenting unlabeled time series patterns has received little attention in the literaturee.
+This is a situation when there is no additional information on the socio-economic charactertistics, property types or family size of the customers.
+There are two reasons why any additional consumer level data would not be available.
+Customer profiles, like property features, vary over time and are not always updated in a timely manner by energy suppliers.
+Also such data might not be available when the anonymity of the customers are of utmost importance.
+So this study does not explain why consumption differs.
+Instead, this work investigates how much energy usage heterogeneity can be found in smart meter data and what some of the most common electricity use patterns are.
+
+<!-- lit review and limitations if any -->
+<!-- _Common load clustering techniques of smart meter data_ -->
+
+An aggregation approach to reduce the data size is not exactly ideal when dealing with highly heterogeneous profiles, as (XXX big data to the rescue example) demonstrates.
 The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
@@
 Most of the solutions to handle this like longest common subsequence, dynamic time warping are prone to computational limit with increased length of the series.
+
+<!-- our approach and contribution -->
+<!-- probability distribution AND cyclic granularities -->
+
+<!-- Large data sets -->
+<!-- On one hand, the sheer volume of data enables greater clarity and analysis at the individual level.
+On the contrary, it also provides challenges as larger data sets have greater uncertainty about customer behavior due to increased variety of customers.
+Additionally, when using big data, there is often a temptation to aggregate the data to reduce its volume so it is easy to manage and computationally feasible to analyse but (XXX big data to the rescue example) show that such a strategy may not always be an optimal choice especially if we are to consider highly heterogeneous profiles that have a dimension of human behaviour.
+-->
+
+<!-- how looking at probability and cyclic granularity helps -->
+
+<!-- how is your work different from the rest of the literature -->
+
+\noindent The massive amount of data generated in such projects could be overwhelming for analysis.
+Electricity utilities can utilize the consumption patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
+Beyea (2010) has pointed out, there has been little discussion or exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data.
+Thus, there is a scope to investigate and analyze these data in various ways for a greater understanding of consumption patterns and how they correlate with other economic, physical or geographical factors.
+In this work, we are interested to see how we can utilize this dataset to group different customers with similar periodic behavior.
+Towards this goal, this chapter aims to: (a) describe the contents of the data set in SGSC database that we can utilize, and (b) propose a clustering algorithm to group customers with similar periodic behaviors.
+The distance metric introduced in Chapter 2 will be the inputs for this cluster analysis.
+One of the advantages of using our
+approach is that the technique is based on probability distributions instead of raw data.
+Many clustering approaches are limited by the type of noisy, patchy, and unequal time-series common in residential data sets.
+Since the distance measure considered is based on differences in probability distribution of time series, it is likely to be less sensitive to missing or noisy data.
+
+_Themes_
+
+- Dimension reduction: If each $P_{i, j, k}$ be considered to be a point in the space, key $i$ would have $mp$ dimensions as opposed to $n_i$ dimensions in case of considering raw data.
+Hence for a large number of observations ($n_i>>mp$), this approach benefits by transitioning to a lower dimension.
+
+ - Avoid loss of information due to aggregation: This approach ensures key characteristic information of the data is not lost due to averaging or aggregation measures in an attempt to transition to a lower dimension.
+Hence, this approach could be thought to somehow balance the drawback of considering raw data or aggregated data.
+
+- Robustness to outliers: This approach could be adapted to be robust to outliers and extreme behaviors by trimming the tails of the probability distributions.
+
+- Non-synchronized observed time periods: Considering probability distribution would imply the clustering process can handle keys that are observed over periods of time that are overlapping but don't necessarily coincide.
+
+- Similar periodic behavior: Since cyclic granularities are considered instead of linear granularities, clustering would group keys that have similar behavior across these cyclic granularities.
+This implies they will be grouped according to their periodic behavior and not on the linear stretch of time over which they are observed.
 

### 2021-10-12 main.Rmd
changed simulation part a bit

@@
 library(GGally)
+library(distributional)
 ```
@@
 The project provides half-hourly data from over 13,000 household electricity smart meters for nearly two years.
-<!---The massive amount of data generated in such projects could be overwhelming for analysis.--> Electricity utilities can utilize these smart meter usage patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
+<!---The massive amount of data generated in such projects could be overwhelming for analysis.--> Raw data of load time series can quickly become overwhelming and hard to interpret, necessitating high level representation as load profiles via load clustering.
+Electricity utilities can utilize the smart meter usage patterns to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
 
@@
 
+<!-- lit review and limitations if any -->
+<!-- _Common load clustering techniques of smart meter data_ -->
+
+A lot of work has been done in the literature on clustering customers by their load characeristics.
+The key to structured aggregation is to summarize big data of large number of customers into manageable subsets of similar load curves.
+Customers should be clustered uniquely, with behaviors represented with distinctive profile patterns.
+Approaches could be deterministic or stochastic.
+Deterministic approaches are hard to generalize due to the need for detailed knowledge of occupancy, appliances, and their time of use.
+In contrast, stochastic load analysis does not require extensive data
+collection other than the time series itself [13,14]. This technique
+suggests viewing load as a random process, where extracting time-series features, or a model of the series, facilitates unsupervised clustering.
+Any standard clustering methodology consists of the following steps: (a) defining distance (dissimilarity) and similarity mostly through feature extraction and selection; and (b) choosing the clustering algorithm design.
+Popular distance measures could be time-domain based or frequency-domain (fast Fourier transform).
+Any model based extraction works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterised and clustered.
+A type of clustering based on information theory such as Shannon or Renyi entropy and their variants are addressed in [42].In contrast to models, a feature-based extraction is used to explicitly define or automatically extract the curves’ key time features, for instance by application of PCA on the daily curves.
+@Tureczek2017-pb presented a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
+None of the $34$ selected papers which focus on clustering consumption are based on Australian smart meter data.
+The algorithm most ubiquitously employed is K-Means.
+But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters.
+Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results.
+Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf).
+Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.
+
+<!--  The time-domain based feature extraction techniques, such as principal component analysis (PCA), detect key features of daily curves and reasonably link them to -->
+<!-- distinctive demand behaviours [28,29]. For instance, PCA can help -->
+<!-- reduce diurnal dimensions, e.g., 48 half-hourly power measurements, down to a few orthogonal components.
+Methods such as fast Fourier transform are used to reveal frequencies that highlight recurrent daily and annual load variations [27].
+-->
+
+<!-- Time-domain analysis is preferred to specify time characteristics, while methods such as fast Fourier transform are used to reveal frequencies that highlight recurrent daily and annual load variations [27].
+Popular time-domain techniques, such as principal component analysis -->
+<!-- (PCA), detect key features of daily curves and reasonably link them to -->
+<!-- distinctive demand behaviours [28,29]. For instance, PCA can help -->
+<!-- reduce diurnal dimensions, e.g., 48 half-hourly power measurements, down to a few orthogonal components.
+@Tureczek2017-pb presented a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
+None of the $34$ selected papers which focus on clustering consumption are based on Australian smart meter data.
+-->
+<!-- The algorithm most ubiquitously employed is K-Means.
+But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters.
+Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results.
+Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf).
+Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.
+-->
+
+<!-- The clustering is frequently applied directly to the raw data without scrutinizing for auto correlation and periodicity. -->
+
+<!-- The key challenge is the application of appropriate processes to reduce the extreme dimensionality of load time series to facilitate unique clusters.
+Time feature extraction is a potential remedy, however, it is limited by the type of noisy, patchy, and unequal time-series common in residential datasets.
+-->
+
+<!-- The literature -->
+<!-- suggests the clustering of residential customers by their load characteristics. The key challenge is the application -->
+<!-- of appropriate processes to reduce the extreme dimensionality of load time series to facilitate unique clusters. -->
+<!-- Time feature extraction is a potential remedy, however, it is limited by the type of noisy, patchy, and unequal -->
+<!-- time-series common in residential datasets. In this paper we propose a strategy to alleviate these limitations by -->
+<!-- converting any types of load time series into map models that can be readily clustered. -->
+
 <!--Unlabelled data -->
 <!-- While forecasting techniques for extremely granular temporal data are readily accessible, -->
+
+<!-- The basic idea behind the clustering algorithm based on distribution is that the data, generated from the same distribution, belongs to the same cluster if there exists several distributions in the original data.
+-->
 
@@
 
-An aggregation approach to reduce the data size is not exactly ideal when dealing with highly heterogeneous profiles, as (XXX big data to the rescue example) demonstrates.
-The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
+<!-- An aggregation approach to reduce the data size is not exactly ideal when dealing with highly heterogeneous profiles, as (XXX big data to the rescue example) demonstrates. -->
+<!-- The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
 None of the 34 selected papers which focus on clustering consumption are based on Australian smart meter data.
@@
 Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.
-As the methods are often situation specific,
-it makes sense to compare them on the performance rather than any standard performance metric.
+As the methods are often situation specific,  -->
+<!-- it makes sense to compare them on the performance rather than any standard performance metric.
 A type of clustering based on information theory such as Shannon or Renyi entropy and their variants are addressed in , which differs from typical methods adopted for electricity consumer classification, based on the Euclidean distance notion.
@@
 Most of the solutions to handle this like longest common subsequence, dynamic time warping are prone to computational limit with increased length of the series.
+-->
 
@@
 We spiked several features in the data to see where one method works better than the other and where they might give us the same outcome or the effect of missing data <!---and trends--> on the proposed methods.
-A continuous measured variable $y$ of length $T$ indexed by ${0, 1, \dots T-1}$.
+A continuous measured variable $y$ of length $T$ indexed by $\{0, 1, \dots T-1\}$.
 Three circular granularities $g1$, $g2$ and $g3$ are considered with $2$, $3$ and $5$ levels respectively.
@@
 The method could preform differently with different jumps between consecutive categories.
-So a difference of $diff = \{1, 2, 5\}$ for corresponding categories are also considered.
+So a mean difference of $diff = \{1, 2, 5\}$ for corresponding categories are also considered.
 The code for creating these designs can be found in the Supplementary section (link to github repo).
@@
 
-Each of the categories $g10,g11$, $g20, g21, g22$, and $g30, g31, g32, g33, g34$ might be assumed to come from the same distribution, a subset of them from the same distribution, a subset of them from separate distributions, or all from different distributions, resulting in various data designs.
+Each of the $7$ categories from $\{g10,g11$, $g20, g21,\dots g32, g33, g34\}$ might be assumed to come from the same distribution, a subset of them from the same distribution, a subset of them from separate distributions, or all from different distributions, resulting in various data designs.
 As the methods ignores the linear progression of time, there is little value in adding time dependency in the data generating process.
-Time series data include basic properties such as trend, seasonality, and auto-correlation, and it is reasonable to construct a time series using these features.
+It is reasonable to construct a time series using properties such as trend, seasonality, and auto-correlation.
 However, when examining distributions across categories of cyclic granularities, these time series features are lost or addressed independently by considering all seasonal fluctuations.
@@
 For example, in a null design, all categories are considered to be normal (0,1).
-The mean of the subsequent categories are incremented or decremented from this distribution based on the mean difference between them in non-null designs.
+The mean of the subsequent categories are incremented or decremented from this null design based on the mean difference between them in non-null designs.
 
@@
 
-**Three signifiant granularities**
+_All signifiant granularities_
 
@@
 That means that for at least one among the to-be-grouped designs, the pattern for each of $g1$, $g2$, and $g3$ will change.
-We consider different distributions across categories (as in Table \ref{tab:tab-distribution}) that will lead to different designs (as in Table \ref{tab:tab-design}).
-Figure \ref{fig:plot-3gran} shows the linear and cyclic representation of the simulated variable under these five designs.
+We consider different distributions across categories (as in Table \ref{tab:tab-dist-design}) (top) that will lead to different designs (as in Table \ref{tab:tab-dist-design}) (below).
+Figure \ref{fig:plot-3gran-new} shows the linear and cyclic representation of the simulated variable under these five designs.
 For the consequent data designs, only graphical representations will be provided.
@@
 granularity = c("g1", "g2", "g3")
-alternate_design =  c("g10 ~ N(0, 1), g11 ~ N(2, 1)",
-                      "g21 ~ N(2, 1), g22 ~ N(1, 1), g23 ~ N(0, 1)",
-                      "g31 ~ N(0, 1), g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1), g35 ~ N(0, 1)")
+# alternate_design =  c("g10 ~ N(0, 1), g11 ~ N(2, 1)",
+#                       "g21 ~ N(2, 1), g22 ~ N(1, 1), g23 ~ N(0, 1)",
+#                       "g31 ~ N(0, 1), g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1), g35 ~ N(0, 1)")
+
+alternate_design =  c("g11 ~ N(2, 1)",
+                      "g21 ~ N(2, 1), g22 ~ N(1, 1)",
+                      "g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1)")
 
@@
 
-knitr::kable(tab_dist, caption = "Alternate distributions of different categories if they deviate from null.")
+# knitr::kable(tab_dist, caption = "Alternate distributions of different categories if they deviate from null.")
 ```
@@
 g3 = c("null", "null", "null", "alternate", "alternate")
-table <- tibble(design, g1, g2, g3) %>% kable(caption = "5 different designs resulting from considering different distributions across categories.")
-table
-
+table <- tibble(design, g1, g2, g3)
+
+# %>% kable(caption = "5 different designs resulting from considering different distributions across categories.")
+```
+
+```{r tab-dist-design}
+
+# library(kableExtra)
+# 
+# knitr::kables(list(kable(caption = "Alternate distributions of different categories if they deviate from null.",
+#    tab_dist
+#     ) %>% kable_styling(),
+#   kable(caption = "5 different designs resulting from considering different distributions across categories.",
+#  tibble(design, g1, g2, g3)
+#     ) %>% kable_styling()
+#   )
+# ) %>% kable_styling()
+# 
+knitr::kable(
+  list(tab_dist,  tibble(design, g1, g2, g3)),
+  caption = 'Alternate distributions of different categories  (top), 5 designs resulting from different distributions across categories (below)',
+  booktabs = TRUE, valign = 't'
+)
 ```
@@
 
-```{r plot-3gran , out.width="100%", fig.cap="The linear (left) and cyclic (right) representation of the measured variable is shown.
+```{r plot-3gran-new, out.width="100%", fig.cap="The linear (left) and cyclic (right) representation of the measured variable is shown.
 In this scenario, all of $g1$, $g2$ and $g3$ changes across at least one design.
@@
 
-**Two significant granularities**
+_Few significant granularities_
 
@@
 
-**One signifiant granularity**
+_One signifiant granularity_
 
@@
 
-data_bind
-
 ```
@@
 
-When two granularities of interest interact, the connection between a granularity and the measured variable is determined by the value of the other interacting granularity.
+The earlier method could be extended when two granularities of interest interact and we are interested to group subjects based on the interaction of the two granularities.
+For example, in application, think about the scenario when customers need to grouped basis their joint behavior across hour-of-day and month-of-year.
+For simplicity, consider a case with just two interacting granularities $g1$ and $g2$ of interest.
+As opposed to the last case, where we could play with the distribution of $5$ individual categories, with interaction, we can play with the distribution of $6$ combination of categories.
+Consider $4$ designs in Figure \ref{fig:} where different distributions are assumed for different combination of categories resulting in different designs.
+
+```{r}
+## ----designs
+
+sim_null_normal <- function(nxj, nfacetj, mean, sd, w1 = 0, w2=0) {
+  rep(distributional::dist_normal(mu = mean, sigma = sd),
+      times = nxj * nfacetj
+  )
+}
+
+sim_varf_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  rep(dist_normal((mean + seq(0, nfacet - 1, by = 1) * w1),
+                  (sd + seq(0, nfacet - 1, by = 1) * w2)), each = nx)
+}
+
+sim_varx_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  rep(dist_normal((mean + seq(0, nx - 1, by = 1) * w1),
+                  (sd + seq(0, nx - 1, by = 1) * w2)), nfacet)
+}
+
+sim_varall_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  dist_normal((mean + seq(0, (nx *nfacet - 1), by = 1) * w1),
+              (sd + seq(0,  (nx *  nfacet - 1), by = 1) * w2))
+}
+
+## ----data
+
+library(hakear)
+
+nx_val = 2 # number of x-axis levels
+nfacet_val = 3 # number of facet levels
+w1_val = 1 # increment in mean
+w2_val = 0 # increment in sd
+mean_val = 0 # mean of normal distribution of starting combination
+sd_val = 2 # sd of normal distribution of starting combination
+quantile_prob_val = seq(0.1, 0.9, 0.1)
+ntimes_val = 300
+
+sim_panel_varall <- sim_panel(
+  nx = nx_val, nfacet =  nfacet_val,
+  ntimes = ntimes_val,
+  # sim_dist = sim_varall_normal(2, 3, 5, 10, 5, -1.5)
+  sim_dist = sim_varall_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+) %>% unnest(data)  %>% 
+  rename( "g2" = "id_facet",
+          "g1" = "id_x")
+
+sim_panel_varx <- sim_panel(
+  nx = nx_val, nfacet =  nfacet_val,
+  ntimes = ntimes_val,
+  # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
+  sim_dist = sim_varx_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+) %>% unnest(data)  %>% 
+  rename( "g2" = "id_facet",
+          "g1" = "id_x")
+
+sim_panel_varf <- sim_panel(
+  nx = nx_val, nfacet =  nfacet_val,
+  ntimes = ntimes_val,
+  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
+  sim_dist = sim_varf_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+) %>% unnest(data)  %>% 
+  rename( "g2" = "id_facet",
+          "g1" = "id_x")
+
+sim_panel_null <- sim_panel(
+  nx = nx_val,
+  nfacet =  nfacet_val,
+  ntimes = ntimes_val,
+  sim_dist = distributional
+  ::dist_normal(mean_val, sd_val)
+) %>% unnest(c(data)) %>% 
+  rename( "g2" = "id_facet",
+          "g1" = "id_x")
+
+## ----category-plots
+
+p_null <- sim_panel_null %>%
+  ggplot(aes(x = as.factor(g2), y = sim_data)) +
+  facet_wrap(~g1, labeller = "label_both") +
+  geom_boxplot() +
+  ylab("") +
+  #ggtitle(paste("(a)", round(null, 2))) +
+  xlab("g2") +
+  theme_bw()+ stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.8, color = "blue") +
+  #ggtitle("design1")+
+  theme_validation()
+
+p_varf <- sim_panel_varf %>%
+  ggplot(aes(x = as.factor(g2), y = sim_data)) +
+  facet_wrap(~g1, labeller = "label_both") +
+  geom_boxplot() +
+  ylab("") +
+  #ggtitle(paste("(b)", round(varf, 2))) +
+  xlab("g2")+
+  theme_bw()+ stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.8, color = "blue") +
+  #ggtitle("design2")+
+  theme_validation()
+
+p_varx <- sim_panel_varx %>%
+  ggplot(aes(x = as.factor(g2), y = sim_data)) +
+  facet_wrap(~g1, labeller = "label_both") +
+  geom_boxplot()+
+  ylab("")  +
+  #ggtitle(paste("(c)", round(varx, 2))) +
+  xlab("g2")+
+  theme_bw()+ stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.8, color = "blue") +
+  #ggtitle("design3")+
+  theme_validation()
+
+p_varall <- sim_panel_varall %>%
+  ggplot(aes(x = as.factor(g2), y = sim_data)) +
+  facet_wrap(~g1, labeller = "label_both") +
+  geom_boxplot() +
+  ylab("") +
+  #ggtitle(paste("(d)", round(varall, 2))) +
+  xlab("g2")+
+  theme_bw()+ stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 0.8, color = "blue") +
+  #ggtitle("design4") +
+  theme_validation()
+
+change_index <- function(data){
+
+index_new <- map(seq_len(ntimes_val), function(i){
+  map((seq_len(nx_val*nfacet_val)), function(j)
+    {
+value = i + (j-1)*ntimes_val
+})
+  }) %>% unlist()
+
+data_new = data %>%
+  ungroup() %>%
+  mutate(index_old = row_number(),
+         index_new = index_new)
+
+y = data_new[match(index_new, data_new$index_old),]
+
+y <- y %>%
+  mutate(time = row_number())
+
+return(y)
+}
+
+endbreaks<- nrow(sim_panel_null)
+
+p1 <- change_index(sim_panel_null) %>%
+  ggplot(aes(x = time,
+             y = sim_data)) +
+  geom_line() +
+  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
+  theme_bw() +
+  #geom_point(alpha = 0.5, color = "blue")
+  #theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
+    #    panel.grid.minor.x =  element_blank())+
+  ylab("simulated response")+
+  xlab("index")
+
+p2 <- change_index(sim_panel_varf) %>%
+  ggplot(aes(x = time,
+             y = sim_data)) +
+  geom_line() +
+  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
+  theme_bw() +
+  #geom_point(alpha = 0.5, color = "blue") +
+  #theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
+     #   panel.grid.minor.x =  element_blank())+
+  ylab("simulated response")+
+  xlab("index")
+
+p3 <- change_index(sim_panel_varx) %>%
+  ggplot(aes(x = time,
+             y = sim_data)) +
+  geom_line() +
+  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
+  theme_bw() +
+  #geom_point(alpha = 0.5, color = "blue")+
+  #theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
+   #     panel.grid.minor.x =  element_blank())+
+  ylab("simulated response")+
+  xlab("index")
+
+p4 <- change_index(sim_panel_varall) %>%
+  ggplot(aes(x = time,
+             y = sim_data)) +
+  geom_line() +
+  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
+  theme_bw() +
+  #geom_point(alpha = 0.5, color = "blue") +
+  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
+  panel.grid.minor.x =  element_blank())+
+  ylab("simulated response")+
+  xlab("index")
+
+(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)
+
+```
+
+<!-- When two granularities of interest interact, the connection between a granularity and the measured variable is determined by the value of the other interacting granularity.
 This happens when the effects of the two granularities on the measured variable are not additive.
-Consider a case with just two interacting granularities $g1$ and $g3$ of interest.
-As opposed to the last case, where we could play with the distribution of $(2+5)$ categories, with interaction we can play with the distribution of $(2*5)$ combination of categories.
+For simplicity, consider a case with just two interacting granularities $g1$ and $g2$ of interest.
+As opposed to the last case, where we could play with the distribution of $5$ individual categories, with interaction we can play with the distribution of $6$ combination of categories.
+Consider $4$ designs in Figure \ref{fig:} where different distributions are assumed for different designs to get some distinction across designs.
+For example, in application, think about the scenario when customers need to grouped basis their joint behavior across hour-of-day and month-of-year.
+-->
 
@@
 With increasing difference between categories, it gets easier for the methods to correctly distinguish the designs.
-For difference=1, the performances are pretty bad for js-robust methods and wpd method.
-The performance starts getting better with increasing difference and get worse with increasing number of replications.
+For $mean_diff=1$, the performances are pretty bad for js-robust methods and wpd method for lower nT.
+Although, with the kind of residential load datasets, a full year of load is the minimal requirement to capture expected variations in winter and summer profiles, for example.
+It is likely that $nT$ would be at least $1000$ with half-hourly data, even if data is only available just for a month.
+The performance is promising except when the number of observations for a customer is really small.
+For smaller difference between categories, it is expected that method js-nqt would preform better than the other two.
+
+```{r}
+# code in append_3gran_change.R
+table <- read_rds("data/append_3gran_change.rds")
+table %>% kable()
+```
+
+<!-- starts getting better with increasing difference and get worse with increasing number of replications.
 Length of series do not show to have any effect on the performance of the methods.
 It does not depend on if time series is ar or arma.
+-->
 

### 2021-10-12 append_3gran_change_5D.R

@@ -1,0 +1,65 @@
+# read_rds("js-nqt/3gran_Change_5D/confmatrix_27.rds") %>% View
+# 
+# simtable %>% dplyr::filter(mean_diff==1)
+
+
+folder_name = "js_robust"
+
+
+
+append_files <- function(folder_name){
+all_files = list.files(path = paste0(folder_name, "/3gran_change_5D/"), 
+                       pattern = ".rds")
+
+names_levels <- map_dfr(all_files, 
+                        function(x){
+                          z = str_split(str_remove(x, ".rds"), "_") %>% 
+                            unlist()
+                          bind_cols(type = z[1],
+                                    index = z[2])
+                        })
+
+
+all_files_path <- paste0(folder_name, "/3gran_Change_5D/",
+                         all_files)  
+
+
+all_data <- lapply(1:length(all_files_path), function(x){
+  
+  data = all_files_path %>% magrittr::extract2(x) %>% 
+    readRDS()
+  
+  names = names_levels %>% magrittr::extract(x,)
+  names_rep =   names %>% slice(rep(1:n(), each = nrow(data)))
+  bind_cols(names_rep, data)
+}) %>% bind_rows() 
+}
+
+wpd <- append_files("wpd") %>% 
+  dplyr::filter(type == "confmatrix", 
+                term == "accuracy") %>% 
+  arrange(mean_diff, nT, niter) %>% 
+  dplyr::select (mean_diff, nT, niter, estimate) %>% mutate(method = "wpd")
+
+
+js_robust <- append_files("js-robust") %>% 
+  dplyr::filter(type == "confmatrix", 
+                term == "accuracy") %>% 
+  arrange(mean_diff, nT, niter) %>% 
+  dplyr::select (mean_diff, nT, niter, estimate) %>% mutate(method = "js_robust")
+
+
+js_nqt <- append_files("js-nqt") %>% 
+  dplyr::filter(type == "confmatrix", 
+                term == "accuracy") %>% 
+  arrange(mean_diff, nT, niter) %>% 
+  dplyr::select (mean_diff, nT, niter, estimate) %>% mutate(method = "js_nqt")
+
+
+
+data_all<- bind_rows(wpd, js_robust, js_nqt) %>% 
+  #arrange(method, nT, niter, mean_diff) %>% 
+  pivot_wider(c(1:3), names_from = "method", values_from = "estimate")
+
+
+write_rds(data_all, "data/append_3gran_change.rds")

### 2021-10-12 gran3_orig.R

@@ -237,8 +237,6 @@ confmatrix <- xtab %>% broom::tidy(by_class = FALSE) %>% mutate(simj)
 write_rds(confmatrix, here(paste0("js-nqt/3gran_change_5D/confmatrix_", scen, ".rds")))
 
 
-
-
 ##----Robust scaling
 
 dist_mat_g1 <- bind_data_iter_tsibble %>%

### 2021-10-12 interact_gran_4D.R

@@ -1,26 +1,10 @@
----
-title: "Clustering four designs"
-output:
-  bookdown::pdf_book:
-    MonashEBSTemplates::workingpaper:
-    #base_format: rticles::asa_article
-    fig_height: 8
-    fig_width: 12
-    fig_caption: yes
-    dev: "pdf"
-    keep_tex: yes
----
 
-```{r setup, include=FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
 library(knitr)
 library(here)
 read_chunk(here::here("old/simdata01.R"))
-```
 
 # How the time series plot of the designs look like for different sample sizes?
 
-```{r sim-val, echo = TRUE}
 set.seed(9999)
 nx_val = 2 # number of x-axis levels
 nfacet_val = 3 # number of facet levels
@@ -86,36 +70,36 @@ Each cluster represents data sets from a separate design
 
 ```{r}
 change_index_data <- function(ntimes_val = NULL,
-                         nx_val = NULL,
-                         nfacet_val = NULL,
-                         sim_function = sim_varx_normal){
-
+                              nx_val = NULL,
+                              nfacet_val = NULL,
+                              sim_function = sim_varx_normal){
+  
   data <- sim_panel(
     nx = nx_val, nfacet =  nfacet_val,
     ntimes = ntimes_val,
     # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
     sim_dist = sim_function(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
   ) %>% unnest(data)
-
-
+  
+  
   index_new <- map(seq_len(ntimes_val), function(i){
     map((seq_len(nx_val*nfacet_val)), function(j)
     {
       value = i + (j-1)*ntimes_val
     })
   }) %>% unlist()
-
+  
   data_new = data %>%
     ungroup() %>%
     mutate(index_old = row_number(),
            index_new = index_new)
-
+  
   y = data_new[match(index_new, data_new$index_old),]
-
+  
   y <- y %>%
     mutate(time = row_number()) %>%
     select(-c(index_old, index_new))
-
+  
   return(y)
 }
 ```
@@ -156,16 +140,16 @@ data_varall <- map(sample_seed, function(seed){
   mutate(data_type = "data_varall")
 
 data_q <- bind_rows(data_null,
-                      data_varf,
-                      data_varx,
-                      data_varall) %>%
[…]

### 2021-10-12 old/simdata01.R

@@ -12,6 +12,7 @@ library(kableExtra)
 library(tidyverse)
 library(gravitas)
 library(magrittr)
+
 ## ----designs
 
 sim_null_normal <- function(nxj, nfacetj, mean, sd, w1 = 0, w2=0) {
@@ -73,7 +74,7 @@ sim_panel_null <- sim_panel(
 
 p_null <- sim_panel_null %>%
   ggplot(aes(x = as.factor(id_x), y = sim_data)) +
-  facet_wrap(~id_facet) +
+  facet_wrap(~id_facet, labeller = "label_both") +
   geom_boxplot() +
   ylab("") +
   #ggtitle(paste("(a)", round(null, 2))) +
@@ -82,7 +83,7 @@ p_null <- sim_panel_null %>%
 
 p_varf <- sim_panel_varf %>%
   ggplot(aes(x = as.factor(id_x), y = sim_data)) +
-  facet_wrap(~id_facet) +
+  facet_wrap(~id_facet, labeller = "label_both") +
   geom_boxplot() +
   ylab("") +
   #ggtitle(paste("(b)", round(varf, 2))) +
@@ -91,7 +92,7 @@ p_varf <- sim_panel_varf %>%
 
 p_varx <- sim_panel_varx %>%
   ggplot(aes(x = as.factor(id_x), y = sim_data)) +
-  facet_wrap(~id_facet) +
+  facet_wrap(~id_facet, labeller = "label_both") +
   geom_boxplot()+
   ylab("")  +
   #ggtitle(paste("(c)", round(varx, 2))) +
@@ -100,7 +101,7 @@ p_varx <- sim_panel_varx %>%
 
 p_varall <- sim_panel_varall %>%
   ggplot(aes(x = as.factor(id_x), y = sim_data)) +
-  facet_wrap(~id_facet) +
+  facet_wrap(~id_facet, labeller = "label_both") +
   geom_boxplot() +
   ylab("") +
   #ggtitle(paste("(d)", round(varall, 2))) +

### 2021-10-12 main.Rmd
intro rough structure done

@@
 <!-- significance of load clustering -->
-Large spatio-temporal data sets, both open and administrative, offer up a world of possibilities for energy and social science research.
-<!--Beyea (2010) has pointed out, there has been little exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data.--> One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
-The project provides half-hourly data from over 13,000 household electricity smart meters for nearly two years.
-<!---The massive amount of data generated in such projects could be overwhelming for analysis.--> Raw data of load time series can quickly become overwhelming and hard to interpret, necessitating high level representation as load profiles via load clustering.
+Large spatio-temporal data sets, both from open and administrative sources, offer up a world of possibilities for research.
+<!--Beyea (2010) has pointed out, there has been little exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data.--> One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee)<!--and Data61 CSIRO-->.
+The project provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014.
+<!---The massive amount of data generated in such projects could be overwhelming for analysis.--> Raw data of these asynchronous time series can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
 Electricity utilities can utilize the smart meter usage patterns to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
@@
 <!-- Sources of variation in large data sets -->
-The enormous quantity of data provides for greater individual level clarity and analysis.
-However, due to the growing variety of consumers, larger data sets include greater uncertainty about consumer behavior.
+<!-- The enormous quantity of data provides for greater individual level clarity and analysis. However, -->
+<!-- due to the growing variety of consumers,  -->
+
+\noindent Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.
 Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
@@
 The variations in behavior may occur on a regular basis, with some consumers being night owls and others being morning larks.
-Day-off habits may vary depending on whether consumers choose to remain at home or engage in outdoor activities.
+Day-off energy usage may vary depending on whether consumers choose to remain at home or engage in outdoor activities.
+These various causes of variations in energy behavior, ranging from age, lifestyle, and family composition to building characteristics, weather, presence of various electrical equipments and others, make the problem of effectively segmenting consumers into similar energy behavior a particularly intriguing one.
+
+<!-- unsupervised clustering -->
+\noindent This problem becomes more difficult when there is no other information on the customers besides their time series of energy usage.
+It is likely that such additional information is not available in order to protect the customers' confidentiality.
+Furthermore, it is not guaranteed that energy providers would always update customer profiles, such as property characteristics, in a timely manner whenever they change.
+So this study does not attempt to explain why consumption differs.
+Instead, the work looks at how much energy-use heterogeneity can be uncovered in smart meter data and what are some of the most typical electricity use patterns by simply using the time series.
+
+<!-- This work -->
+This is similar to a stochastic approach (@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time-series characteristics, or a model of the series, to enable unsupervised clustering.
+Unsupervised clustering is only as good as the features that are extracted/selected or the distance metrics that were utilized.
+Well-designed additional features may collect characteristics that default features cannot.
+Based on the underlying structure of the temporal data, this article offers new distance metric and features for clustering and applies them to actual smart-meter data.
+Firstly, the distance metric is based on probability distribution, which in our knowledge is the first attempt to cluster smart meter data using probability distributions.
 These recorded time series are asynchronous, with varying time lengths for different houses and missing observations.
-This huge amount of noisy, patchy, and unequal time-series common in residential smart meter data combined with the different causes of variation in their energy behavior makes the problem of effectively segmenting consumers with comparable consumption a particularly intriguing one.
+Taking probability distributions helps to deal with such data, while helping with dimension reduction in one hand but not losing too much information due to aggregation.
+Secondly, we recognise that most clustering algorithms only provide hourly energy profiles during the day, but this approach provides a wider approach to the issue, seeking to group consumers with similar shapes over all important cyclic granularities.
+Since cyclic granularities are considered instead of linear granularities, clustering would group customers that have similar repetitive behavior across more than one cyclic granularities across which patterns are expected to be significant.
+
+<!-- lit review -->
+_Related work_
+
+A typical clustering technique includes the following steps: (a) establishing distance (dissimilarity) and similarity through feature or model extraction and selection; and (b) selecting the clustering algorithm design.
+Distance measures could be time-domain based or frequency-domain (fast Fourier transform).
+A model-based clustering works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterised and clustered (@Motlagh2019-yj).
+[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
+The essential temporal characteristics of the curves are defined or extracted using feature-based clustering.
+@Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
+None of the $34$ articles chosen for their emphasis use Australian smart meter data.
+The most often used algorithm is K-Means.
+Using K-Means without considering time series structure or correlation results in inefficient clusters.
+Principal Component Analysis (PCA) or Self-Organizing Maps (SOM) eliminate correlation patterns and decrease feature space, but lose interpretability.
+To reduce dimensionality, several studies use principal component analysis or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).
+Other algorithms utilised in the literature include k-means variants, hierarchical approaches, and greedy k-medoids.
+Time series data, such as smart metre data, are not well-suited to any of the techniques mentioned in @Tureczek2017-pb.
+Only one study [@ozawa2016determining] identified time series characteristics using Fourier transformation, which converts data from time to frequency and then uses K-Means to cluster by greatest frequency .
+
+<!-- Most clustering algorithms offer only daily energy profiles throughout the hours of the day, but we suggest a broader approach to the problem, aiming to group consumers with similar shapes across all significant cyclic granularities -->
+
+<!-- There are two reasons why any additional consumer level data would not be available.
+Customer profiles, like property features, vary over time and are not always updated in a timely manner by energy suppliers.
+Also such data might not be available when the anonymity of the customers are of utmost importance.
+-->
+
+<!-- Segmenting unlabeled time series patterns has received little attention in the literature.
+This is a situation when there is no additional information on the socio-economic charactertistics, property types or family size of the customers.
+There are two reasons why any additional consumer level data would not be available.
+Customer profiles, like property features, vary over time and are not always updated in a timely manner by energy suppliers.
+Also such data might not be available when the anonymity of the customers are of utmost importance.
+So this study does not explain why consumption differs.
+Instead, this work investigates how much energy usage heterogeneity can be found in smart meter data and what some of the most common electricity use patterns are.
+-->
+
+<!-- These recorded time series are asynchronous, with varying time lengths for different houses and missing observations. -->
+
+<!-- An aggregation approach to reduce the data size is not exactly ideal when dealing with highly heterogeneous profiles -->
 
@@
 
-A lot of work has been done in the literature on clustering customers by their load characeristics.
-The key to structured aggregation is to summarize big data of large number of customers into manageable subsets of similar load curves.
+<!-- A lot of work has been done in the literature on clustering customers by their load characeristics.
 Customers should be clustered uniquely, with behaviors represented with distinctive profile patterns.
@@
 Deterministic approaches are hard to generalize due to the need for detailed knowledge of occupancy, appliances, and their time of use.
-In contrast, stochastic load analysis does not require extensive data
-collection other than the time series itself [13,14]. This technique
-suggests viewing load as a random process, where extracting time-series features, or a model of the series, facilitates unsupervised clustering.
+In contrast, stochastic load analysis does not require extensive data collection other than the time series itself [13,14].
+This technique -->
+<!-- suggests viewing load as a random process, where extracting time-series features, or a model of the series, facilitates unsupervised clustering.
 Any standard clustering methodology consists of the following steps: (a) defining distance (dissimilarity) and similarity mostly through feature extraction and selection; and (b) choosing the clustering algorithm design.
-Popular distance measures could be time-domain based or frequency-domain (fast Fourier transform).
+-->
+<!-- Popular distance measures could be time-domain based or frequency-domain (fast Fourier transform).
 Any model based extraction works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterised and clustered.
@@
 None of the $34$ selected papers which focus on clustering consumption are based on Australian smart meter data.
-The algorithm most ubiquitously employed is K-Means.
+-->
+<!-- The algorithm most ubiquitously employed is K-Means.
 But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters.
@@
 Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.
-
-<!--  The time-domain based feature extraction techniques, such as principal component analysis (PCA), detect key features of daily curves and reasonably link them to -->
-<!-- distinctive demand behaviours [28,29]. For instance, PCA can help -->
-<!-- reduce diurnal dimensions, e.g., 48 half-hourly power measurements, down to a few orthogonal components.
-Methods such as fast Fourier transform are used to reveal frequencies that highlight recurrent daily and annual load variations [27].
 -->
@@
 
-Segmenting unlabeled time series patterns has received little attention in the literaturee.
+<!-- Segmenting unlabeled time series patterns has received little attention in the literature.
 This is a situation when there is no additional information on the socio-economic charactertistics, property types or family size of the customers.
@@
 Instead, this work investigates how much energy usage heterogeneity can be found in smart meter data and what some of the most common electricity use patterns are.
+-->
 
@@
 
-\noindent The massive amount of data generated in such projects could be overwhelming for analysis.
+<!-- \noindent The massive amount of data generated in such projects could be overwhelming for analysis.
 Electricity utilities can utilize the consumption patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
@@
 In this work, we are interested to see how we can utilize this dataset to group different customers with similar periodic behavior.
-Towards this goal, this chapter aims to: (a) describe the contents of the data set in SGSC database that we can utilize, and (b) propose a clustering algorithm to group customers with similar periodic behaviors.
+-->
+
+<!-- Towards this goal, this chapter aims to: (a) describe the contents of the data set in SGSC database that we can utilize, and (b) propose a clustering algorithm to group customers with similar periodic behaviors.
 The distance metric introduced in Chapter 2 will be the inputs for this cluster analysis.
-One of the advantages of using our
-approach is that the technique is based on probability distributions instead of raw data.
+One of the advantages of using our  -->
+<!-- approach is that the technique is based on probability distributions instead of raw data.
 Many clustering approaches are limited by the type of noisy, patchy, and unequal time-series common in residential data sets.
 Since the distance measure considered is based on differences in probability distribution of time series, it is likely to be less sensitive to missing or noisy data.
-
-_Themes_
-
-- Dimension reduction: If each $P_{i, j, k}$ be considered to be a point in the space, key $i$ would have $mp$ dimensions as opposed to $n_i$ dimensions in case of considering raw data.
+-->
+
+<!-- _Themes_ -->
+
+<!-- - Dimension reduction: If each $P_{i, j, k}$ be considered to be a point in the space, key $i$ would have $mp$ dimensions as opposed to $n_i$ dimensions in case of considering raw data.
 Hence for a large number of observations ($n_i>>mp$), this approach benefits by transitioning to a lower dimension.
-
- - Avoid loss of information due to aggregation: This approach ensures key characteristic information of the data is not lost due to averaging or aggregation measures in an attempt to transition to a lower dimension.
+-->
+
+<!--  - Avoid loss of information due to aggregation: This approach ensures key characteristic information of the data is not lost due to averaging or aggregation measures in an attempt to transition to a lower dimension.
 Hence, this approach could be thought to somehow balance the drawback of considering raw data or aggregated data.
-
-- Robustness to outliers: This approach could be adapted to be robust to outliers and extreme behaviors by trimming the tails of the probability distributions.
-
-- Non-synchronized observed time periods: Considering probability distribution would imply the clustering process can handle keys that are observed over periods of time that are overlapping but don't necessarily coincide.
-
-- Similar periodic behavior: Since cyclic granularities are considered instead of linear granularities, clustering would group keys that have similar behavior across these cyclic granularities.
+-->
+
+<!-- - Robustness to outliers: This approach could be adapted to be robust to outliers and extreme behaviors by trimming the tails of the probability distributions. -->
+
+<!-- - Non-synchronized observed time periods: Considering probability distribution would imply the clustering process can handle keys that are observed over periods of time that are overlapping but don't necessarily coincide.
+-->
+
+<!-- - Similar periodic behavior: Since cyclic granularities are considered instead of linear granularities, clustering would group keys that have similar behavior across these cyclic granularities.
 This implies they will be grouped according to their periodic behavior and not on the linear stretch of time over which they are observed.
-
-The following contributions are made through the following chapter:
-
- * Present a cluster analysis of SGSC dataset to group households with similar periodic behavior
- * Cluster validation by relating to external data
+-->
+
+<!-- The following contributions are made through the following chapter: -->
+
+<!--  * Present a cluster analysis of SGSC dataset to group households with similar periodic behavior -->
+<!--  * Cluster validation by relating to external data -->
 
@@
 
-# Clustering methodology
+The remainder of the paper is organized as follows: Section&nbsp;\ref{sec:methodology} provides the clustering methodology introducing the features and distance metrics.
+Section&nbsp;\ref{sec:validation} shows data designs to validate our methods and draw comparisons against several methods.
+Section&nbsp;\ref{sec:application} discusses the application of the method to a subset of the real data.
+Finally, we summarize our results and discuss possible future directions in Section&nbsp;\ref{sec:discussion}.
+
+# Clustering methodology {#sec:methodology}
 
@@
 
-# Validation
+# Validation {#sec:validation}
 
@@
 
-# Application
+# Application {#sec:application}
 
@@
 
-# Discussion
+# Discussion {#sec:discussion}
 

### 2021-10-13 main.Rmd
prototype selection section update

@@
 library(distributional)
+library(viridis)
 ```
@@
 
-\noindent The aim of this section is to illustrate that the proposed methodology can be used to understand repetitive behavior for several customers together.
+<!-- Why instance selection -->
+In supervised learning, a training set containing previously known information is used to categorize new occurrences.
+Acceptable classification rates may be obtained by discarding instances which are not helpful for classification; this process is known as instance selection (@olvera2010review).
+This is similar to subsetting the population along all dimensions of importance such that the sampled data is representative of the main characteristics of the underlying distribution.
+Instance selection in unsupervised learning has received limited attention in the literature, but could serve as an useful way to sample evaluation data set to measure the performance of a model or method.
+One such procedure is suggested in @Fan2021-bq that selects similar instances (neighbors) for each instance (anchor) and treats the anchor and its neighbors as the same class.
+In this section, a similar idea is used to select customers with prototype behaviors
+that serves as evaluation data sets for our proposed methodology.
 
@@
 From this set, we randomly sample a set of $600$ customers.
-Then we removed customers for which any of the hod (hour-of-day), moy (month-of-year) and wkndwday (weeknd/weekday) is a clash.
-We further removed customers for which all the deciles of the energy consumption is zero.
+We obtain the $wpd$ for all cyclic granularities considered for these customers and found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weeknd/weekday) are coming out to be significant for most customers.
+This implies that for most customers, there is some interesting pattern across these three granularities.
+Potentially this could be done for the entire data set.
+This step is time consuming and hence has been only run for the $600$ sampled customers.
+As a second step, we remove customers for which data in any category for the significant granularities are empty.
+For example, in this data set, if a customer do not have data for an entire month, they have been removed as their monthly behavior could not be studied in that case.
+Further, we also remove customers for which all the deciles of the energy consumption is zero.
 These are the customers whose consumption remain mostly flat and is expected to have no interesting repetitive patterns that is our interest of study.
-Finally, we are left with $356$ customers from which we wanted to do instance selection.
+Finally, we are left with $356$ customers.
+From this set we select $4$ "anchor" customers which are far apart from each other and $5$ neighboring customers for each of these anchors.
+These selections were done using the granularity $hod$ space.
+It is important to note that when we use our proposed methodologies, it is based on all dimesions `hod`, `moy` and `wkndwday`.Fig \ref{fig:} shows the MDS of these 356 customers in a 2D space basis their distance on individual granularities and when all of them are combined.
+Our methodolgies are run on these $24$ customers, which act as a way to evaluate the proposed methodologies.
+
+<!-- However, if `hod` is and not only `hod`, but if `hod` is significant we can expect that the grouping would be similar to how we have initially chosen the set of customers. -->
+
+<!-- from which we wanted to do instance selection.
 Since this is unlabeled data, there is no way to do external validation of our methodologies.
@@
 When we use our methodologies, it is based on all significant granularities and not only hod, but if $hod$ is significant we can expect that the grouping would be similar to how we have initially chosen the set of customers.
-This procedure is analogous to instance-wise classfication methods that selects similar instances (neighbours) for each instance (anchor) and treats the anchor and its neighbours as the same class.
+-->
 
@@
   facet_wrap(~group, 
-             #scales = "free_y",  
+             scales = "free_y",  
              nrow = 4) + 
@@
   facet_wrap(~group, 
-             #scales = "free_y", 
+             scales = "free_y", 
              labeller = "label_value",
@@
 
-library(viridis)
 pairsplot <- ggpairs(data_pcp, 
@@
   scale_color_viridis_d(direction = -1) + theme_light()
-pairsplot
-```
-
-```{r}
+```
+
+```{r parcoord, fig.cap = "A parallel coordinate plot with the three significant cyclic granularities used for wpd-based clustering.
+The variables are sorted according to their separation across classes (rather than their overall variation between classes).
+This means that moy is the most important variable in distinguishing the designs followed by hod and wkndwday.
+It can be observed that cluster 3 and 4 are distinguished by moy while 1 and 2 are distinguished by hod.
+Here, wkndwday is still acting as the nuisance variable."}
 parcoord

### 2021-10-13 supplementary/distance-computation/real-data/04handpick-from356-mds.Rmd
prototype selection section update

--- 2021-09-29 supplementary/distance-computation/real-data/04handpick-from356-mds.Rmd
@@
   select(customer_id, everything())
-
 ```
@@
   layout(showlegend = FALSE) %>%
-  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect") 
+  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect", persistent = TRUE) 
 ```
 
-# Visualize similarity between samples (all gran together)
-
-for different distance metrics hod+moy+wkndwday, hod, moy, wkndwday
+# Visualize similarity between js and wpd methodologies:
 

### 2021-10-13 check_work.Rmd
wkndwday pattern changed to boxplot

--- 2021-10-13 main.Rmd
@@
 theme_characterisation <- function() {
+
   theme_bw() + 
-      theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) +  + theme_bw() +
+      theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) +
    theme(axis.title.y=element_blank(),
@@
   theme(panel.spacing =unit(0, "lines")) + 
-    theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+    theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+    theme(legend.position = "bottom")+
+    theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
   theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
   theme(legend.position = "bottom")
@@
 ```
-
-# Introduction
-
-<!-- Description of data available -->
-
-<!-- significance of load clustering -->
-Large spatio-temporal data sets, both from open and administrative sources, offer up a world of possibilities for research.
-<!--Beyea (2010) has pointed out, there has been little exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data.--> One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee)<!--and Data61 CSIRO-->.
-The project provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014.
-<!---The massive amount of data generated in such projects could be overwhelming for analysis.--> Raw data of these asynchronous time series can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
-Electricity utilities can utilize the smart meter usage patterns to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
-
-<!-- Sources of variation in large data sets -->
-<!-- The enormous quantity of data provides for greater individual level clarity and analysis. However, -->
-<!-- due to the growing variety of consumers,  -->
-
-\noindent Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.
-Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
-Multiple temporal dependencies define the behavioral patterns, which vary from customer to customer.
-Some families, for example, use a dryer to dry their clothing, while others hang them to dry on a line.
-This may be reflected in their weekly profile.
-They may have monthly variations where some customers are more prone to use air conditioners or heaters than others despite the existence of comparable electrical equipment and being subjected to similar weather conditions.
-The variations in behavior may occur on a regular basis, with some consumers being night owls and others being morning larks.
-Day-off energy usage may vary depending on whether consumers choose to remain at home or engage in outdoor activities.
-These various causes of variations in energy behavior, ranging from age, lifestyle, and family composition to building characteristics, weather, presence of various electrical equipments and others, make the problem of effectively segmenting consumers into similar energy behavior a particularly intriguing one.
-
-<!-- unsupervised clustering -->
-\noindent This problem becomes more difficult when there is no other information on the customers besides their time series of energy usage.
-It is likely that such additional information is not available in order to protect the customers' confidentiality.
-Furthermore, it is not guaranteed that energy providers would always update customer profiles, such as property characteristics, in a timely manner whenever they change.
-So this study does not attempt to explain why consumption differs.
-Instead, the work looks at how much energy-use heterogeneity can be uncovered in smart meter data and what are some of the most typical electricity use patterns by simply using the time series.
-
-<!-- This work -->
-This is similar to a stochastic approach (@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time-series characteristics, or a model of the series, to enable unsupervised clustering.
-Unsupervised clustering is only as good as the features that are extracted/selected or the distance metrics that were utilized.
-Well-designed additional features may collect characteristics that default features cannot.
-Based on the underlying structure of the temporal data, this article offers new distance metric and features for clustering and applies them to actual smart-meter data.
-Firstly, the distance metric is based on probability distribution, which in our knowledge is the first attempt to cluster smart meter data using probability distributions.
-These recorded time series are asynchronous, with varying time lengths for different houses and missing observations.
-Taking probability distributions helps to deal with such data, while helping with dimension reduction in one hand but not losing too much information due to aggregation.
-Secondly, we recognise that most clustering algorithms only provide hourly energy profiles during the day, but this approach provides a wider approach to the issue, seeking to group consumers with similar shapes over all important cyclic granularities.
-Since cyclic granularities are considered instead of linear granularities, clustering would group customers that have similar repetitive behavior across more than one cyclic granularities across which patterns are expected to be significant.
-
-<!-- lit review -->
-_Related work_
-
-A typical clustering technique includes the following steps: (a) establishing distance (dissimilarity) and similarity through feature or model extraction and selection; and (b) selecting the clustering algorithm design.
-Distance measures could be time-domain based or frequency-domain (fast Fourier transform).
-A model-based clustering works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterised and clustered (@Motlagh2019-yj).
-[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
-The essential temporal characteristics of the curves are defined or extracted using feature-based clustering.
-@Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
-None of the $34$ articles chosen for their emphasis use Australian smart meter data.
-The most often used algorithm is K-Means.
-Using K-Means without considering time series structure or correlation results in inefficient clusters.
-Principal Component Analysis (PCA) or Self-Organizing Maps (SOM) eliminate correlation patterns and decrease feature space, but lose interpretability.
-To reduce dimensionality, several studies use principal component analysis or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).
-Other algorithms utilised in the literature include k-means variants, hierarchical approaches, and greedy k-medoids.
-Time series data, such as smart metre data, are not well-suited to any of the techniques mentioned in @Tureczek2017-pb.
-Only one study [@ozawa2016determining] identified time series characteristics using Fourier transformation, which converts data from time to frequency and then uses K-Means to cluster by greatest frequency .
-
-<!-- Most clustering algorithms offer only daily energy profiles throughout the hours of the day, but we suggest a broader approach to the problem, aiming to group consumers with similar shapes across all significant cyclic granularities -->
-
-<!-- There are two reasons why any additional consumer level data would not be available.
-Customer profiles, like property features, vary over time and are not always updated in a timely manner by energy suppliers.
-Also such data might not be available when the anonymity of the customers are of utmost importance.
--->
-
-<!-- Segmenting unlabeled time series patterns has received little attention in the literature.
-This is a situation when there is no additional information on the socio-economic charactertistics, property types or family size of the customers.
-There are two reasons why any additional consumer level data would not be available.
-Customer profiles, like property features, vary over time and are not always updated in a timely manner by energy suppliers.
-Also such data might not be available when the anonymity of the customers are of utmost importance.
-So this study does not explain why consumption differs.
-Instead, this work investigates how much energy usage heterogeneity can be found in smart meter data and what some of the most common electricity use patterns are.
--->
-
-<!-- These recorded time series are asynchronous, with varying time lengths for different houses and missing observations. -->
-
-<!-- An aggregation approach to reduce the data size is not exactly ideal when dealing with highly heterogeneous profiles -->
-
-<!-- lit review and limitations if any -->
-<!-- _Common load clustering techniques of smart meter data_ -->
-
-<!-- A lot of work has been done in the literature on clustering customers by their load characeristics.
-Customers should be clustered uniquely, with behaviors represented with distinctive profile patterns.
-Approaches could be deterministic or stochastic.
-Deterministic approaches are hard to generalize due to the need for detailed knowledge of occupancy, appliances, and their time of use.
-In contrast, stochastic load analysis does not require extensive data collection other than the time series itself [13,14].
-This technique -->
-<!-- suggests viewing load as a random process, where extracting time-series features, or a model of the series, facilitates unsupervised clustering.
-Any standard clustering methodology consists of the following steps: (a) defining distance (dissimilarity) and similarity mostly through feature extraction and selection; and (b) choosing the clustering algorithm design.
--->
-<!-- Popular distance measures could be time-domain based or frequency-domain (fast Fourier transform).
-Any model based extraction works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterised and clustered.
-A type of clustering based on information theory such as Shannon or Renyi entropy and their variants are addressed in [42].In contrast to models, a feature-based extraction is used to explicitly define or automatically extract the curves’ key time features, for instance by application of PCA on the daily curves.
-@Tureczek2017-pb presented a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
-None of the $34$ selected papers which focus on clustering consumption are based on Australian smart meter data.
--->
-<!-- The algorithm most ubiquitously employed is K-Means.
-But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters.
-Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results.
-Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf).
-Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.
--->
-
-<!-- Time-domain analysis is preferred to specify time characteristics, while methods such as fast Fourier transform are used to reveal frequencies that highlight recurrent daily and annual load variations [27].
-Popular time-domain techniques, such as principal component analysis -->
-<!-- (PCA), detect key features of daily curves and reasonably link them to -->
-<!-- distinctive demand behaviours [28,29]. For instance, PCA can help -->
-<!-- reduce diurnal dimensions, e.g., 48 half-hourly power measurements, down to a few orthogonal components.
-@Tureczek2017-pb presented a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
-None of the $34$ selected papers which focus on clustering consumption are based on Australian smart meter data.
--->
-<!-- The algorithm most ubiquitously employed is K-Means.
-But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters.
-Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results.
-Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf).
-Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.
--->
-
-<!-- The clustering is frequently applied directly to the raw data without scrutinizing for auto correlation and periodicity. -->
-
-<!-- The key challenge is the application of appropriate processes to reduce the extreme dimensionality of load time series to facilitate unique clusters.
-Time feature extraction is a potential remedy, however, it is limited by the type of noisy, patchy, and unequal time-series common in residential datasets.
--->
-
-<!-- The literature -->
-<!-- suggests the clustering of residential customers by their load characteristics. The key challenge is the application -->
-<!-- of appropriate processes to reduce the extreme dimensionality of load time series to facilitate unique clusters. -->
-<!-- Time feature extraction is a potential remedy, however, it is limited by the type of noisy, patchy, and unequal -->
-<!-- time-series common in residential datasets. In this paper we propose a strategy to alleviate these limitations by -->
-<!-- converting any types of load time series into map models that can be readily clustered. -->
-
-<!--Unlabelled data -->
-<!-- While forecasting techniques for extremely granular temporal data are readily accessible, -->
-
-<!-- The basic idea behind the clustering algorithm based on distribution is that the data, generated from the same distribution, belongs to the same cluster if there exists several distributions in the original data.
--->
-
-<!-- Segmenting unlabeled time series patterns has received little attention in the literature.
-This is a situation when there is no additional information on the socio-economic charactertistics, property types or family size of the customers.
-There are two reasons why any additional consumer level data would not be available.
-Customer profiles, like property features, vary over time and are not always updated in a timely manner by energy suppliers.
-Also such data might not be available when the anonymity of the customers are of utmost importance.
-So this study does not explain why consumption differs.
-Instead, this work investigates how much energy usage heterogeneity can be found in smart meter data and what some of the most common electricity use patterns are.
--->
-
-<!-- lit review and limitations if any -->
-<!-- _Common load clustering techniques of smart meter data_ -->
-
-<!-- An aggregation approach to reduce the data size is not exactly ideal when dealing with highly heterogeneous profiles, as (XXX big data to the rescue example) demonstrates. -->
-<!-- The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
-None of the 34 selected papers which focus on clustering consumption are based on Australian smart meter data.
-The clustering is frequently applied directly to the raw data without scrutinizing for auto correlation and periodicity.
-The algorithm most ubiquitously employed is K-Means.
-But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters.
-Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results.
-@Tureczek2018-ha has shown that a transformation of data to incorporate autocorrelation before K-Means clustering can improve performance and enable K-Means to deliver smaller clusters with less within-cluster variance.
-However, it does not explain the cluster composition by combining it with external data.
-Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf).
-Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.
-As the methods are often situation specific,  -->
-<!-- it makes sense to compare them on the performance rather than any standard performance metric.
-A type of clustering based on information theory such as Shannon or Renyi entropy and their variants are addressed in , which differs from typical methods adopted for electricity consumer classification, based on the Euclidean distance notion.
-@Motlagh2019-yj presents strategy to address the problems on patchy, and unequal time-series common in residential data sets by converting load time series into map models.
-Most time-series clustering models are limited to handling time domain with same start and end date and time.
-Most of the solutions to handle this like longest common subsequence, dynamic time warping are prone to computational limit with increased length of the series.
--->
-
-<!-- our approach and contribution -->
-<!-- probability distribution AND cyclic granularities -->
-
-<!-- Large data sets -->
-<!-- On one hand, the sheer volume of data enables greater clarity and analysis at the individual level.
-On the contrary, it also provides challenges as larger data sets have greater uncertainty about customer behavior due to increased variety of customers.
-Additionally, when using big data, there is often a temptation to aggregate the data to reduce its volume so it is easy to manage and computationally feasible to analyse but (XXX big data to the rescue example) show that such a strategy may not always be an optimal choice especially if we are to consider highly heterogeneous profiles that have a dimension of human behaviour.
--->
-
-<!-- how looking at probability and cyclic granularity helps -->
-
-<!-- how is your work different from the rest of the literature -->
-
-<!-- \noindent The massive amount of data generated in such projects could be overwhelming for analysis.
-Electricity utilities can utilize the consumption patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
-Beyea (2010) has pointed out, there has been little discussion or exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data.
-Thus, there is a scope to investigate and analyze these data in various ways for a greater understanding of consumption patterns and how they correlate with other economic, physical or geographical factors.
-In this work, we are interested to see how we can utilize this dataset to group different customers with similar periodic behavior.
--->
-
-<!-- Towards this goal, this chapter aims to: (a) describe the contents of the data set in SGSC database that we can utilize, and (b) propose a clustering algorithm to group customers with similar periodic behaviors.
-The distance metric introduced in Chapter 2 will be the inputs for this cluster analysis.
-One of the advantages of using our  -->
-<!-- approach is that the technique is based on probability distributions instead of raw data.
-Many clustering approaches are limited by the type of noisy, patchy, and unequal time-series common in residential data sets.
-Since the distance measure considered is based on differences in probability distribution of time series, it is likely to be less sensitive to missing or noisy data.
--->
-
-<!-- _Themes_ -->
-
-<!-- - Dimension reduction: If each $P_{i, j, k}$ be considered to be a point in the space, key $i$ would have $mp$ dimensions as opposed to $n_i$ dimensions in case of considering raw data.
-Hence for a large number of observations ($n_i>>mp$), this approach benefits by transitioning to a lower dimension.
--->
-
-<!--  - Avoid loss of information due to aggregation: This approach ensures key characteristic information of the data is not lost due to averaging or aggregation measures in an attempt to transition to a lower dimension.
-Hence, this approach could be thought to somehow balance the drawback of considering raw data or aggregated data.
--->
-
-<!-- - Robustness to outliers: This approach could be adapted to be robust to outliers and extreme behaviors by trimming the tails of the probability distributions. -->
-
-<!-- - Non-synchronized observed time periods: Considering probability distribution would imply the clustering process can handle keys that are observed over periods of time that are overlapping but don't necessarily coincide.
--->
-
-<!-- - Similar periodic behavior: Since cyclic granularities are considered instead of linear granularities, clustering would group keys that have similar behavior across these cyclic granularities.
-This implies they will be grouped according to their periodic behavior and not on the linear stretch of time over which they are observed.
--->
-
-<!-- The following contributions are made through the following chapter: -->
-
-<!--  * Present a cluster analysis of SGSC dataset to group households with similar periodic behavior -->
-<!--  * Cluster validation by relating to external data -->
-
-<!-- Electricity smart meter technology is increasingly being deployed in residential and commercial buildings.
-For example, in the state of Victoria, Australia, it is a state government policy that all households have a smart meter fitted (Victoria State Government, 2015), which has resulted in 97% of the approximately 2.1 million households in Victoria having a smart meter installed in their home.
-These smart meters collect energy usage information at half-hourly intervals, resulting in over 35 billion half-hourly observations per year across all households in the state.
-Clustering electricity consumption patterns can enable electricity utilities to develop targeted tariffs for individual groups alleviating the problem of volatility in production by capitalizing on the flexibility of consumers.
--->
-
-<!-- Clustering households using only smart meter consumption data could also provide value in a societal setting by combining the findings and external data like weather conditions, socio-economic or other demographic factors of those households.
--->
-
-<!-- Background literature and shortcomings -->
-
-<!-- # Preliminary exploratoration -->
-
-<!-- ## Electricity demand data -->
-
-<!-- _Electricity use interval reading data_   -->
-<!-- The data from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
--->
-
-<!-- ### Data source -->
-
-<!-- The entire data is procured from CSIRO.
-A subset of this data is also available from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
-It consists of the following data sets.
--->
-<!-- _1. CustomerData:_ 78720 customers with 62 variables about them  -->
-<!-- _2. EUDMData:_ 300 billion half-hourly consumption level data   -->
-<!-- _3. OffersData:_ Method of contact to customer to join SGSC customer trial, either door-to-door (D2D) or via Telesales   -->
-<!-- _4. PEResponseData:_ Peak Events response customer wise   -->
-<!-- _5. PETimesData:_ Peak Events time stamps   -->
-
-<!-- Only _CustomerData_ and _EUDMData_ are relevant for the clustering goals of this paper.
-_EUDMData_ contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
-`CustomerData` provides demographic data for 78,720 customers with information about their Local Government Area amongst others.
--->
-
-<!-- ## LGA and weather data -->
-
-<!-- Since the smart meters have been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths -->
-<!-- for each household.
-Since, general supply is available for only 13,735 customers, we will restrict ourselves to look at the LGA information for these customers only.
-We find that there are only 26 LGA that is covered for these customers.
--->
-
-<!-- _Weather data_   -->
-<!-- This data is obtained through [Australian Government Bureau of Meteorology](BOM)(http://www.bom.gov.au/) and provides hourly data for nearest weather stations for all the LGAs -->
-
-<!-- This section familiarizes the 13,000 SGSC households through visualization and provide a detailed layout of data structures (like missing observations/number of customers/number of observations) and also the external data that needs to be utilized for validating the clustering process.
-The [ABS TableBuilder](https://www.abs.gov.au/websitedbs/censushome.nsf/home/tablebuilder) has census data from 2011 and 2016.
-The data is at SA2 and LGA levels.
-However, some of the LGA in NSW changed between 2011 and 2016 and hence there would not be a one-to-one correspondence between the LGAs.
-Weather, notably temperature (and humidity) can be the main driver(s) for energy usage.
-In NSW many households have electric heaters so their use can impact winter energy use and air-conditioners can impact summer energy use.
-Relevant weather data could be obtained from the [Bureau of Meteorology](www.bom.gov.au).
-Some weather stations have 30 minute (sometimes even smaller interval) weather data.
-Potentially, there could be lag effects of weather on energy usage which should be considered.
--->
-
-The remainder of the paper is organized as follows: Section&nbsp;\ref{sec:methodology} provides the clustering methodology introducing the features and distance metrics.
-Section&nbsp;\ref{sec:validation} shows data designs to validate our methods and draw comparisons against several methods.
-Section&nbsp;\ref{sec:application} discusses the application of the method to a subset of the real data.
-Finally, we summarize our results and discuss possible future directions in Section&nbsp;\ref{sec:discussion}.
-
-# Clustering methodology {#sec:methodology}
-
-<!-- In contrast to models, a feature-based strategy is used to explicitly define or automatically extract the curves’ key time features, for instance by application of PCA on the daily curves [ -->
-
-<!-- Most papers discussed in Tureczek2017-pb fail to accept smart meter readings as time series data, a data type which contains a temporal component.
-The omission of the essential time series features in the analysis leads to the application of methods that are not designed for handling temporal components.
-K-Means ignores autocorrelation, unless the input data is pre-processed.
-The clusters identified in the papers are validated by a variety of indices, with the most prevalent -->
-<!-- being the cluster dispersion index (CDI) [22–24], the Davies–Bouldin index (DBI) [25,26] and the mean index adequacy (MIA) [8,13]. -->
-
-The data set solely contains readings from smart meters and no information about the consumers' specific physical, geographical, or behavioural attributes.
-As a result, no attempt is made to explain why consumption varies.
-Instead, this work investigates how much energy usage heterogeneity can be found in smart meter data and what some of the most common electricity use patterns are.
-It is worth noting that when studying these dynamics, a variety of objectives may be pursued.
-One objective could be to group consumers with similar shapes over all relevant cyclic granularities.
-In this scenario, the variation in customers within each group is in magnitude rather than shape, while the variation between groups is only in shape.
-Most clustering algorithms offer only daily energy profiles throughout the hours of the day, but we suggest a broader approach to the problem, aiming to group consumers with similar shapes across all significant cyclic granularities.
-Another purpose of clustering could be to group customers that have similar differences in patterns across all major cyclic granularities, capturing similar jumps across categories regardless of the overall shape.
-For example, in the first goal, similar shapes across hours of the day will be grouped together, resulting in customers with similar behaviour across all hours of the day, whereas in the second goal, any similar big-enough jumps across hours of the day will be clubbed together, regardless of which hour of the day it is.
-Both of these objectives may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
-
-The proposed methodology aim to leverage the intrinsic temporal data structure hidden in time series data.
-The foundation of our method is unsupervised clustering algorithms based exclusively on time-series features.
-First, we study the underlying distributions that may have resulted in different patterns across temporal granularities in order to identify a mechanism to classify them based on the similarity of those distributions.
-Depending on the goal of clustering, the distance metric for defining similarity would be different.
-These distance metrics could be fed into a clustering algorithm to break large data sets into subgroups that can then be analyzed separately.
-These clusters may be commonly associated with real-world data segmentation.
-However, since the data is unlabeled a priori, more information is required to corroborate this.
-This section presents the work flow of the methodology:
-
-- _Data preparation_
-
-@wang2020tsibble introduced the tidy "tsibble" data structure to support exploration and modeling of temporal data.
-A tsibble comprises an index, optional key(s), and measured variables.
-An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
-A linear granularity is a mapping of the index set to subsets of the time domain.
-For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.
-For each key variable, the raw smart meter data is a sequence that is indexed by time and comprises values of several measurement variables at each time point.
-This sequence, though, could be depicted in a variety of ways.
-A shuffling of the raw sequence could reflect the distribution of hourly consumption over a single day, while another could indicate consumption over a week or a year.
-These temporal deconstructions of a time period into units such as hour-of-day, work-day/weekend are called cyclic temporal granularities.
-All cyclic granularities can be expressed in terms of the index set and could be augmented with the initial tsibble structure (index, key, measurements).
-It is worthwhile to note that the data structure changes while transporting from linear to cyclic scale of time as multiple observations of the measured variable would correspond to each category of the cyclic granularities.
-In this paper, quantiles are chosen to characterize the distributions for each category of the cyclic granularity.
-So, each category of a cyclic granularity corresponds to a list of numbers which is essentially few chosen quantiles of the multiple observations.
-
-- _Finding significant cyclic granularities or harmonies_
-
-\noindent These cyclic granularities are useful for exploring repetitive patterns in time series data that get lost in the linear representation of time.
-It is advantageous to consider only those cyclic granularities across which there is a significant repetitive pattern for the majority of customers or noteworthy in an electricity-behavior context.
-In that case, when the customers are grouped, we can expect to observe some interesting patterns across the categories of the cyclic granularities considered.
-[XXX reference 2nd chapter] proposes a way to select significant cyclic granularities and harmonies which is used for this paper.
-
-- _Individual or combined categories of cyclic granularities as DGP_
-
-The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
-In our work, we are using $i$ as denoting a customer and the underlying variable as the electricity demand.
-So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
-In this work, instead of considering the probability distributions of the linear time series, we assume that the measured variables across different categories of any cyclic granularity are from data generating processes.
-Hence, we want to be able to cluster distributions of the form $f_{i,A,B \dots, {N_C}}(v)$, where $A, B$ represent the cyclic granularities under consideration such that $A = \{a_j: j=1, 2, \dots J\}$,  $B = \{b_k: k  = 1, 2, \dots K\}$ and so on.
-We consider individual each category of a cyclic granularity ($A$) or combination of categories for interaction of cyclic granularities (for e.g.
-$A*B$) to have a distribution.
-For example, let us consider we have two cyclic  granularities of interest, $A = {0, 1, 2, \dots, 23}$ representing hour-of-day and  $B = \{Mon, Tue, Wed, \dots, Sun\}$ representing day-of-week.
-each customer $i$ consist of a collection of probability distributions.
-In case individual granularities ($A$ or $B$) are considered there are  $J = 24$ distributions of the form $f_{i,j}(v)$ or $K = 7$ distributions of the form $f_{i,k}(v)$ for each customer $i$.
-In case of interaction,  $J*K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
-As a result, a distance between collections of these univariate probability distributions is required.
-Depending on the objective of the problem, there could be many approaches to considering such distances.
-This paper considers two approaches, which are explained in the next segment.
-
-- _Distance metrics_
-
-Considering each individual or combined categories of cyclic granularities as a data generating process lead to a collection of conditional distributions for each customer $i$.
-The (dis) similarity between each pair of observations should be obtained by combining the distances between these collections of conditional distributions such that the resulting metric is a distance metric, which could be fed into the clustering algorithm.
-Two types of distance metric is considered:
-
-<!-- The choice of distance measures is a critical step in clustering. It defines how the similarity of two elements (x, y) is calculated. -->
-
-**Inter-category distances**
-
-This distance matrix considers two objects to be similar if every category of an individual cyclic granularity or combination of categories for interacting cyclic granularities have similar distributions.
-In this study, the distribution for each category is characterized using deciles and the distances between distributions are computed by using the Jensen-Shannon distance, which is symmetric and hence could be used as a distance measure.
-
-\noindent The total distance between two elements $x$ and $y$ is then defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity A) or  $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
-When combining distances from individual cyclic granularities $A$ and $B$, $$S^{A, B}_{x, y} = S^A_{x,y}/J + S^B_{x,y}/K$$ is used, which could also be shown to be a distance metric easily.
-This is shown for cyclic granularity $A$ and $B$, but could be practically extended to more granularities.
-
-**Intra-category distances**
-
-Compute weighted pairwise distances ($wpd$) (XXX reference) for all considered granularities for all objects.
-$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive between consecutive categories normalized by appropriate parameters.
-A higher value of $wpd$ indicates that some interesting pattern is expected, whereas a lower value would indicate otherwise.
-
-\noindent Distance between objects is then taken as the euclidean distances between them with the granularities being the variables and $wpd$ being the value under each variable.
-Since Euclidean distance is chosen, the observations with high values of features ($wpd$ values) will be clustered together.
-The same holds true for observations with low values of features.
-Thus this distance matrix would be useful to group customers that have similar significance of patterns across different granularities.
-
-<!-- Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
-Each household consists of a $J*K$ distributions one harmony.
-We compute the distributional difference -->
-<!-- between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$. -->
-<!-- $wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies.
-If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
-We aim to have pockets of households showing similar periodic behavior by conidering $wpd$ vlaues for different harmonies and some time series features.
-The features should also characterize probability distributions of different household.
--->
-
-<!-- grouping probability distributions across a harmony.
-This clustering algorithm is adopted to remove or appropriately adjust for auto correlation and unequal length in the data.
-The method could be further extended by clustering probability distributions conditional on one or more cyclic granularities.
-The following are some of the advantages of our proposed method.
--->
-
-- _Pre-processing steps_
-
-    **Handling trend, seasonality, non-stationarity and auto-correlation:** Trend and seasonality are fundamental characteristics of time series data, and it is reasonable to define a time series according to its degree of trend and seasonality.
-These characteristics of the time series are lost or handled independently by considering probability distributions ( trend is lost) across categories of cyclic granularities (by independently modeling all seasonal fluctuations), and so there is no need to de-trend or de-seasonalize the data before conducting the clustering method.
-There is no need to omit holiday or weekend patterns for similar reasons.
-
-    **Data transformation:** Robust scaling method is used before computing the Inter-category distances and NQT is built-in transformation used for computation of $wpd$, which forms the basis of Intra-category distances.
-
-- _Clustering algorithm_
-
-In the analysis of energy smart metre data, K-Means or hierarchical clustering are often employed.
-These are simple and effective techniques that work well in a range of scenarios.
-For clustering, both employ a distance measure, and the distance measure chosen has a major influence on the structure of the clusters.
-We employ agglomerative hierarchical clustering in conjunction with Ward's criteria (XXX reference).
-Individual entities with the highest similarity computed using the desired distance metrics are sequentially merged using agglomerative algorithms.
-We can possibly employ any clustering method that supports the given distance metric as input.
-
-- _Characterization of clusters_
-
-Depending on the distance measure utilized for the study, the cluster characterization technique will differ.
-Clusters that utilise intra-category distances are characterised using multi-dimensional scaling and parallel coordinate displays.
-For inter-category distances, the distribution across major granularities may be presented to ensure that the goal of similar shapes within clusters and distinct shapes across clusters is met.
-This technique may potentially make advantage of multi-dimensional scaling.
-
-\noindent Multidimensional scaling (MDS) (XXX reference) refers to a family of methods that analyse a matrix of distances or dissimilarities to provide a representation of the data points in a reduced-dimension space.
-There are many kinds of MDS, but they all solve the same fundamental issue: Given a $n*n$ matrix of dissimilarities and a distance measure, identify a configuration of $n$ points $x_1, x_2, \dots, x_n$ in the reduced dimension space $R^q$ ($q<p$) where the distance between the points is near to the dissimilarity between the points.
-All techniques must determine the coordinates of the points as well as the space dimension, $q$.
-Metric and nonmetric MDS are the two main kinds of MDS.
-Metric MDS methods presume a functional connection between the interpoint distances and the supplied dissimilarities and assume that the data are quantitative.
-We use metric MDS.
-
-\noindent Parallel coordinate plots (XXX reference) Parallel coordinates have been extensively used to display high-dimensional and multivariate data, allowing for the detection of patterns within the data via visual grouping.
-
-<!-- - A random sample  of the original data is taken for clustering analysis and includes missing and noisy observations (detailed description in Appendix) -->
-
-<!-- - All harmonies are computed for each customer in the sample. -->
-<!-- Cyclic granularities which are clashes for all customers in the sample are removed. -->
-
-<!-- - It is worth noting that a number of other solutions may be considered at the pre-processing stage of the method.
-We have considered a) Normal-Quantile Transform and b) Robust transformation.
--->
-
-<!-- - Two methods are considered for computing dissimilarity between two customers.
-The first one involves computing according to one granularity is computed as the sum of the JS distances between distribution of all the categories of the granularity.
-When we consider more than one granularity, we consider the sum of the  average distances for all the granularity so that the combined metric is also a distance.
--->
-
-<!-- - Given the scale of dissimilarity among the energy readings, the model chooses optimal number of clusters -->
-
-<!-- - Once clusters have been allocated, the groups are explored visually. -->
-
-<!-- - Results are reported and compared. -->
-
-<!-- Two methods are used for computing distances between subjects and then hierarchical clustering algorithm is used. -->
-
-<!-- The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
-In our work, we are using $i$ as denoting a customer and the underlying variable as the electricity demand.
-So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
--->
-
-<!-- We want to cluster distributions of the form $f_{i,j,k}(v)$, where $i$ and $j$ denote  -->
-
-<!-- Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
-Each household consists of a $J*K$ distributions one harmony.
-We compute the distributional difference -->
-<!-- between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$. -->
-<!-- $wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies.
-If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
-We aim to have pockets of households showing similar periodic behavior by conidering $wpd$ vlaues for different harmonies and some time series features.
-The features should also characterize probability distributions of different household.
--->
-
-<!-- ### Notations -->
-
-<!-- Consider an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
-In our work, we are using $i$ as denoting a household and the underlying variable as the electricity demand.
-Further consider a cyclic granularity of the form $B = \{ b_k: k = 1, 2, \dots, K\}$.
-Each customer consists of collection of probability distributions.
--->
-
-<!-- So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
-We want to cluster distributions of the form $f_{i,j,k}(v)$, where $i$ and $j$ denote $i^{th}$ and $j^{th}$ customer respectively.
--->
-
-<!-- a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.  -->
-
-<!-- ### A single or pair of granularities together (change names) -->
-
-<!-- The methodology can be summarized in the following steps: -->
-
-<!-- - _Pre-processing step_ -->
-
-<!-- Robust scaling method or NQT used for each customer. -->
-
-<!-- - _NQT_ -->
-
-<!-- - _Treatment to outliers_ -->
-
-<!-- ### Many granularities together (change names) -->
-
-<!-- The methodology can be summarized in the following steps: -->
-
-<!-- 1. Compute quantiles of distributions across each category of the cyclic granularity -->
-<!-- 2. Compute JS distance between households for each each category of the cyclic granularity -->
-<!-- 3. Total distance between households computed as sum of JS distances for all hours -->
-<!-- 4. Cluster using this distance with hierarchical clustering algorithm (method "Ward.D") -->
-
-<!-- _Pro:_   -->
-<!-- - distance metric makes sense to group different shapes together   -->
-<!-- - simulation results look great on typical designs   -->
-<!-- _Cons:_   -->
-<!-- - Can only take one granularity at once   -->
-<!-- - Clustering a big blob of points together whereas the aim is to groups these big blob into smaller ones   -->
-
-<!-- ### Multiple-granularities -->
-
-<!-- _Description:_   -->
-
-<!-- Choose all significant granularities and compute wpd for all these granularities for all customers.
-Distance between customers is taken as the euclidean distances between them with the granularities being the variables and wpd being the value under each variable for which Euclidean distance needs to be measured.
--->
-<!-- _Pro:_   -->
-<!-- - Can only take many granularities at once -->
-<!-- - can apply variable selection PCP and other interesting clustering techniques -->
-<!-- - simulation results look great on typical designs -->
-<!-- - splitting the data into similar sized groups   -->
-<!-- _Cons:_   -->
-<!-- - distance metric does not make sense to split the data into similar shaped clusters  -->
-
-# Validation {#sec:validation}
-
-To validate the clustering approaches, we create data designs that replicate prototype behaviors that might be seen in electricity data contexts.
-We spiked several features in the data to see where one method works better than the other and where they might give us the same outcome or the effect of missing data <!---and trends--> on the proposed methods.
-A continuous measured variable $y$ of length $T$ indexed by $\{0, 1, \dots T-1\}$.
-Three circular granularities $g1$, $g2$ and $g3$ are considered with $2$, $3$ and $5$ levels respectively.
-Categories of $g1$, $g2$ and $g3$ are represented by ${g10,g11}$, ${g20, g21, g22}$ and ${g30, g31, g32, g33, g34}$.
-These categories could be integers or some more meaningful labels.
-For example, the granularity "day-of-week" could be either represented by ${0, 1, 2, \dots, 6}$ or ${Mon, Tue, \dots, Sun}$.
-We created independent replications $R = \{25, 250, 500\}$ of all data designs to see if our proposed clustering approaches can detect distinct designs in various groups for small, medium and large number of series.
-A sample size of $T=\{300, 1000, 5000\}$ is used in all designs to test small, medium and large sized series.
-The method could preform differently with different jumps between consecutive categories.
-So a mean difference of $diff = \{1, 2, 5\}$ for corresponding categories are also considered.
-The code for creating these designs can be found in the Supplementary section (link to github repo).
-The results for $T=300$ and $R=25$ is shown, that means we have $25$ time series each with length $300$.
-The rest of the results could be found in the supplementary paper.
-
-## Data generating processes
-
-<!-- An ARMA (p,q) process is used to generate series, where $p$ and $q$ are selected at random such that the series is stationary.
-The various designs on $g1$, $g2$, and $g3$ are introduced by adding matching designs to this series' innovations.
-The innovations are considered to have a normal distribution, although they follow the same pattern as the designs.
-To eliminate the effect of starting values, the first 500 observations in each series are discarded.
--->
-
-Each of the $7$ categories from $\{g10,g11$, $g20, g21,\dots g32, g33, g34\}$ might be assumed to come from the same distribution, a subset of them from the same distribution, a subset of them from separate distributions, or all from different distributions, resulting in various data designs.
-As the methods ignores the linear progression of time, there is little value in adding time dependency in the data generating process.
-It is reasonable to construct a time series using properties such as trend, seasonality, and auto-correlation.
-However, when examining distributions across categories of cyclic granularities, these time series features are lost or addressed independently by considering all seasonal fluctuations.
-Because the time span during which an entity is observed in order to ascertain its behavior is very short, it is assumed that the time series will remain stationary throughout the observation period.
-The data type is set to "continuous," and the setup is assumed to be gaussian.
-For example, in a null design, all categories are considered to be normal (0,1).
-The mean of the subsequent categories are incremented or decremented from this null design based on the mean difference between them in non-null designs.
-
-<!-- An ARMA (p,q) process is used to generate series, where $p$ and $q$ are selected at random such that the series is stationary.
-The various designs on $g1$, $g2$, and $g3$ are introduced by adding matching designs to this series' innovations.
-The innovations are considered to have a normal distribution, although they follow the same pattern as the designs.
-To eliminate the effect of starting values, the first 500 observations in each series are discarded.
--->
-
-## Data designs
-
-### Individual granularities
-
-_All signifiant granularities_
-
-<!-- Consider a case where all the three granularities $g1$, $g2$ and $g3$ would be responsible for making the designs distinct.
-That would mean, the pattern for each of $g1$, $g2$ and $g3$ will change for at least one design.
-We consider a situation with all the null cases corresponding to no difference in distribution across categories, that is, all categories follow the same distribution N(0,1).
--->
-
-Consider the scenario when all three granularities $g1$, $g2$, and $g3$ are responsible for distinguishing the designs.
-That means that for at least one among the to-be-grouped designs, the pattern for each of $g1$, $g2$, and $g3$ will change.
-We consider different distributions across categories (as in Table \ref{tab:tab-dist-design}) (top) that will lead to different designs (as in Table \ref{tab:tab-dist-design}) (below).
-Figure \ref{fig:plot-3gran-new} shows the linear and cyclic representation of the simulated variable under these five designs.
-For the consequent data designs, only graphical representations will be provided.
-
-```{r tab-distribution}
-granularity = c("g1", "g2", "g3")
-# alternate_design =  c("g10 ~ N(0, 1), g11 ~ N(2, 1)",
-#                       "g21 ~ N(2, 1), g22 ~ N(1, 1), g23 ~ N(0, 1)",
-#                       "g31 ~ N(0, 1), g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1), g35 ~ N(0, 1)")
-
-alternate_design =  c("g11 ~ N(2, 1)",
-                      "g21 ~ N(2, 1), g22 ~ N(1, 1)",
-                      "g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1)")
-
-tab_dist <- tibble::tibble(granularity = granularity, `alternate distribution` = alternate_design)
-
-# knitr::kable(tab_dist, caption = "Alternate distributions of different categories if they deviate from null.")
-```
-
-<!-- A combination of these null and alternate distributions are considered when building the designs as in the table \ref{tab:tab-design}.  -->
-
-```{r tab-design}
-design = c("design-1", "design-2", "design-3", "design-4", "design5")
-g1 = c("null", "alternate", "null", "null", "alternate")
-g2 = c("null", "null", "alternate", "null", "alternate")
-g3 = c("null", "null", "null", "alternate", "alternate")
-table <- tibble(design, g1, g2, g3)
-
-# %>% kable(caption = "5 different designs resulting from considering different distributions across categories.")
-```
-
-```{r tab-dist-design}
-
-# library(kableExtra)
-# 
-# knitr::kables(list(kable(caption = "Alternate distributions of different categories if they deviate from null.",
-#    tab_dist
-#     ) %>% kable_styling(),
-#   kable(caption = "5 different designs resulting from considering different distributions across categories.",
-#  tibble(design, g1, g2, g3)
-#     ) %>% kable_styling()
-#   )
-# ) %>% kable_styling()
-# 
-knitr::kable(
-  list(tab_dist,  tibble(design, g1, g2, g3)),
-  caption = 'Alternate distributions of different categories  (top), 5 designs resulting from different distributions across categories (below)',
-  booktabs = TRUE, valign = 't'
-)
-```
-
-```{r generate-design-3change}
-generate_design <- function(t, mu1, mu2, mu3){
-
-t <- seq(0, t, 1)
-g1 <- t %%2
-g2 <- t %%3
-g3 <- t %%5
-
-# null design
-g1_dnull <- rep( rep(0, each = length(unique(g1))), length.out= length(t))
-g2_dnull <- rep( rep(0, each = length(unique(g2))), length.out= length(t))
-g3_dnull <- rep( rep(0, each = length(unique(g3))), length.out= length(t))
-
-# mean changing across categories in varying ways
-
-g1_dvary <- rep(mu1, length.out= length(t))
-g2_dvary <- rep(mu2, length.out= length(t))
-g3_dvary <- rep(mu3, length.out= length(t))
-
-design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
-design2 = distributional::dist_normal(g1_dvary + g2_dnull + g3_dnull)
-design3 = distributional::dist_normal(g1_dnull + g2_dvary + g3_dnull)
-design4 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dvary)
-design5 = distributional::dist_normal(g1_dvary + g2_dvary + g3_dvary)
-
-data_bind <- tibble::tibble(
-index = t,
-g1 = g1,
-g2 = g2,
-g3 = g3,
-design1 = distributional::generate(design1, times = 1) %>% unlist(),
-design2 = distributional::generate(design2, times = 1) %>% unlist(),
-design3 = distributional::generate(design3, times = 1) %>% unlist(),
-design4 = distributional::generate(design4, times = 1) %>% unlist(),
-design5 = distributional::generate(design5, times = 1) %>% unlist()
-) %>% 
-  pivot_longer(-c(1, 2, 3, 4), names_to = "design", values_to = "sim_data")
-
-data_bind
-}
-
-t = 300
-mu1= c(0, 2)
-mu2 = c(2, 1, 0)
-mu3 = c(0, 1, 2, 1, 0)
-
-data_bind <- generate_design(t, mu1, mu2, mu3)
-
-```
-
-```{r plot-3gran-new, out.width="100%", fig.cap="The linear (left) and cyclic (right) representation of the measured variable is shown.
-In this scenario, all of $g1$, $g2$ and $g3$ changes across at least one design.
-Also, it is not possible to comprehend these patterns across cyclic granularities or group similar series just by looking at the linear plots."}
-
-# plot_linear_data <- function(data){
-# ggplot(data,
-#              aes(x = index, y = sim_data)) + 
-#   geom_line() +
-#   xlab("index")+
-#   theme_bw() 
-# }
-
-p1 <- ggplot(data_bind,
-             aes(x = index, y = sim_data)) + 
-  geom_line() +
-  xlab("index")+
-  facet_wrap(~design, scales = "free_y",ncol =1) +
-  theme_validation()
-
-p2 <- ggplot(data_bind,
-             aes(x = as.factor(g1), y = sim_data)) + 
-  geom_boxplot(alpha =0.5) + xlab("g1") + 
-  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") +
-  theme_validation()
-
-p3 <- ggplot(data_bind, aes(x = as.factor(g2), y = sim_data)) + geom_boxplot(alpha =0.5) + xlab("g2") + theme_bw() +
-  facet_wrap(~design, scales = "free_y",ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") + ylab("")+
-  theme_validation()
-
-p4 <- ggplot(data_bind, aes(x = as.factor(g3), y = sim_data)) + geom_boxplot(alpha =0.5) +
-  xlab("g3") + theme_bw()+
-  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, colosr = "blue")+ ylab("")+
-  theme_validation()
-
-(p1 + (p2 + p3 + p4)) *
-  theme_validation() + plot_layout(widths = c(2, 1))
-# 
-# plot_cyclic_data( data_bind %>% filter(design=="design2"))
-# plot_data( data_bind %>% filter(design=="design3"))
-# plot_data( data_bind %>% filter(design=="design4"))
-# plot_data( data_bind %>% filter(design=="design5"))
-```
-
-_Few significant granularities_
-
-This is the case where one granularity will remain the same across all designs.
-We consider the case where the distribution of $y$ would vary across levels of $g2$ for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
-So $g3$ is not responsible for distinguishing across designs.
-Figure \ref{#fig:plot-linear} shows the linear and cyclic representation of $y$.
-The first panel shows raw plot of $y$ in a linear scale and the second panel shows distribution of $y$ across cyclic granularities namely $g1$, $g2$ and $g3$.
-As could be seen from the plots, it is impossible to decipher from the raw time plot that the time series variable shows such pattern across different granularities.
-
-_One signifiant granularity_
-
-Here only one granularity is responsible for distinguishing the designs. Designs change significantly only for the granularity $g2$. Figure \ref{fig:plot-1gran} shows this.
-
-```{r}
-generate_design <- function(t, mu21, mu22, mu23){
-
-t <- seq(0, t, 1)
-g1 <- t %%2
-g2 <- t %%3
-g3 <- t %%5
-
-# null design
-g1_dnull <- rep( rep(0, each = length(unique(g1))), length.out= length(t))
-g2_dnull <- rep( rep(0, each = length(unique(g2))), length.out= length(t))
-g3_dnull <- rep( rep(0, each = length(unique(g3))), length.out= length(t))
-
-# mean changing across categories in varying ways
-
-g1_dvary <- rep(0, length.out= length(t))
-g21_dvary <- rep(mu21, length.out= length(t))
-g22_dvary <- rep(mu22, length.out= length(t))
-g23_dvary <- rep(mu23, length.out= length(t))
-g3_dvary <- rep(0, length.out= length(t))
-
-design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
-design2 = distributional::dist_normal(g1_dnull + g21_dvary + g3_dnull)
-design3 = distributional::dist_normal(g1_dnull + g22_dvary + g3_dnull)
-design4 = distributional::dist_normal(g1_dnull + g23_dvary + g3_dnull)
-
-data_bind <- tibble::tibble(
-index = t,
-g1 = g1,
-g2 = g2,
-g3 = g3,
-design1 = distributional::generate(design1, times = 1) %>% unlist(),
-design2 = distributional::generate(design2, times = 1) %>% unlist(),
-design3 = distributional::generate(design3, times = 1) %>% unlist(),
-design4 = distributional::generate(design4, times = 1) %>% unlist(),
-#design5 = distributional::generate(design5, times = 1) %>% unlist()
-) %>% 
-  pivot_longer(-c(1, 2, 3, 4), names_to = "design", values_to = "sim_data")
-
-data_bind
-}
-
-mu21= c(2, 0, 0)
-mu22 = c(0, 2, 0)
-mu23 = c(0, 0, 2)
-mu3 = c(0, 0, 0, 0, 0)
-data_bind <- generate_design(t, mu21, mu22, mu23)
-
-```
-
-```{r plot-gran, out.width="100%"}
-
-p1 <- ggplot(data_bind,
-             aes(x = index, y = sim_data)) + 
-  geom_line() +
-  xlab("index")+
-  facet_wrap(~design, scales = "free_y",ncol =1) +
-  theme_validation()
-
-p2 <- ggplot(data_bind,
-             aes(x = as.factor(g1), y = sim_data)) + 
-  geom_boxplot(alpha =0.5) + xlab("g1") + 
-  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") +
-  theme_validation()
-
-p3 <- ggplot(data_bind, aes(x = as.factor(g2), y = sim_data)) + geom_boxplot(alpha =0.5) + xlab("g2") + theme_bw() +
-  facet_wrap(~design, scales = "free_y",ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") + ylab("")+
-  theme_validation()
-
-p4 <- ggplot(data_bind, aes(x = as.factor(g3), y = sim_data)) + geom_boxplot(alpha =0.5) +
-  xlab("g3") + theme_bw()+
-  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, colosr = "blue")+ ylab("")+
-  theme_validation()
-
-(p1 + (p2 + p3 + p4)) *
-  theme_validation() + plot_layout(widths = c(2, 1))
-
-```
-
-<!-- When we move from the linear to the cyclic world of temporal granularities, we can see patterns across different categories of the granularities which gets lost in the linear representation.
-Populations are modeled by a collection of these cyclic granularities.
-Each cyclic granularity might or might not have patterns across its categories.
-Each cluster is characterized by similar patterns across one or more of these cyclic granularities.
--->
-
-<!-- A small example is given to setup the problem. -->
-
-<!-- We constructed the simulation parameters to represent common patterns in electricity data.
-50, 200 and 500 time series were chosen for different simulation designs with different granularities and patterns changing across different granularities.
--->
-<!-- The data type is fixed to be "continuous".We generated independent replications of all combinations of the simulation parameters. -->
-
-<!-- Consider a continuous time series variable $y$ of length $T$ indexed by ${0, 1, \dots T-1}$.
-Three circular granularities $g1$, $g2$ and $g3$ are considered with 2, 3 and 5 levels respectively.
-Categories of g1, g2 and g3 are represented by ${0,1}$, ${0, 1, 2}$ and ${0, 1, 2, 3, 4}$.
-These categories could be integers or some more meaningful labels.
-For example, the granularity "day-of-week" could be either represented by $\{0, 1, 2, \dots, 6\}$ or $\{Mon, Tue, \dots, Sun\}$.
--->
-
-<!-- Consider the case where distribution of $y$ would vary across levels of $g1$ and g2$ but will not vary across levels of $g3$ as described in Figure .Figure \ref shows four designs where $g3$ acts as a nuisance variable, and $g1$ and $g2$ varies in both  Design 2 and 3, and $g2$ varies in all designs.
--->
-
-<!-- Consider a case where distribution of $y$ would vary across levels of $g2$ -->
-<!-- for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
-Figure \ref{#fig:plot-linear} shows the linear and cyclic representation of $y$.
-The first panel shows raw plot of $y$ in a linear scale and the second panel shows distribution of $y$ across cyclic granularities namely $g1$, $g2$ and $g3$.
-As could be seen from the plots, it is impossible to decipher from the raw time plot that the time series variable shows such pattern across different granularities.
--->
-
-```{r change-var, echo=FALSE}
-t = 300
-n <- seq(0, t, 1)
-g1 <- n %%2
-g2 <- n %%3
-g3 <- n %%5
-mu11 = c(0, 5)
-mu12 = c(3, 0)
-mu21= c(2, 0, 0)
-mu22 = c(0, 2, 0)
-mu23 = c(0, 0, 2)
-mu3 = c(0, 0, 0, 0, 0)
-```
-
-```{r generate-design} 
-generate_design <- function(t, mu21, mu22, mu23){
-
-t <- seq(0, t, 1)
-g1 <- t %%2
-g2 <- t %%3
-g3 <- t %%5
-
-# null design
-g1_dnull <- rep( rep(0, each = length(unique(g1))), length.out= length(t))
-g2_dnull <- rep( rep(0, each = length(unique(g2))), length.out= length(t))
-g3_dnull <- rep( rep(0, each = length(unique(g3))), length.out= length(t))
-
-# mean changing across categories in varying ways
-
-g11_dvary <- rep(mu11, length.out= length(t))
-g12_dvary <- rep(mu12, length.out= length(t))
-g21_dvary <- rep(mu21, length.out= length(t))
-g22_dvary <- rep(mu22, length.out= length(t))
-g23_dvary <- rep(mu23, length.out= length(t))
-g3_dvary <- rep(0, length.out= length(t))
-
-design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
-design2 = distributional::dist_normal(g11_dvary + g21_dvary + g3_dnull)
-design3 = distributional::dist_normal(g12_dvary + g22_dvary + g3_dnull)
-design4 = distributional::dist_normal(g1_dnull + g23_dvary + g3_dnull)
-
-data_bind <- tibble::tibble(
-index = t,
-g1 = g1,
-g2 = g2,
-g3 = g3,
-design1 = distributional::generate(design1, times = 1) %>% unlist(),
-design2 = distributional::generate(design2, times = 1) %>% unlist(),
-design3 = distributional::generate(design3, times = 1) %>% unlist(),
-design4 = distributional::generate(design4, times = 1) %>% unlist(),
-#design5 = distributional::generate(design5, times = 1) %>% unlist()
-) %>% 
-  pivot_longer(-c(1, 2, 3, 4), names_to = "design", values_to = "sim_data")
-
-data_bind
-}
-
-data_bind <- generate_design(t, mu21, mu22, mu23)
-
-#data_bind
-
-```
-
-```{r plot-2gran, out.width="100%", fig.cap = "The linear (left) and cyclic (right) representation of the measured variable is shown.
-In this scenario, $g1$, $g2$ would change across atleast one design but $g3$ change remains same across all design.
-Thus $g3$ is not an important variable to distinguish these designs."}
-
-# plot_linear_data <- function(data){
-# ggplot(data,
-#              aes(x = index, y = sim_data)) + 
-#   geom_line() +
-#   xlab("index")+
-#   theme_bw() 
-# }
-
-p1 <- ggplot(data_bind,
-             aes(x = index, y = sim_data)) + 
-  geom_line() +
-  xlab("index")+
-  facet_wrap(~design, scales = "free_y",ncol =1) +
-  theme_validation()
-
-p2 <- ggplot(data_bind,
-             aes(x = as.factor(g1), y = sim_data)) + 
-  geom_boxplot(alpha =0.5) + xlab("g1") + 
-  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") +
-  theme_validation()
-
-p3 <- ggplot(data_bind, aes(x = as.factor(g2), y = sim_data)) + geom_boxplot(alpha =0.5) + xlab("g2") + theme_bw() +
-  facet_wrap(~design, scales = "free_y",ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") + ylab("")+
-  theme_validation()
-
-p4 <- ggplot(data_bind, aes(x = as.factor(g3), y = sim_data)) + geom_boxplot(alpha =0.5) +
-  xlab("g3") + theme_bw()+
-  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue")+ ylab("")+
-  theme_validation()
-
-(p1 + (p2 + p3 + p4)) *
-  theme_validation() + plot_layout(widths = c(2, 1))
-# 
-# plot_cyclic_data( data_bind %>% filter(design=="design2"))
-# plot_data( data_bind %>% filter(design=="design3"))
-# plot_data( data_bind %>% filter(design=="design4"))
-# plot_data( data_bind %>% filter(design=="design5"))
-```
-
-A subset of many possible designs are shown in Figure \ref{fig: plot-linear }.
-For the parameter space (XXX unique  combinations shown in table YYY), 100, 500 independent replications of all possible combination of simulation parameters were generated.
-The clustering methodologies were run all these unique combinations and subsets of these to verify if the methodologies work as expected.
-
-### Interaction of granularities
-
-The earlier method could be extended when two granularities of interest interact and we are interested to group subjects based on the interaction of the two granularities.
-For example, in application, think about the scenario when customers need to grouped basis their joint behavior across hour-of-day and month-of-year.
-For simplicity, consider a case with just two interacting granularities $g1$ and $g2$ of interest.
-As opposed to the last case, where we could play with the distribution of $5$ individual categories, with interaction, we can play with the distribution of $6$ combination of categories.
-Consider $4$ designs in Figure \ref{fig:} where different distributions are assumed for different combination of categories resulting in different designs.
-
-```{r}
-## ----designs
-
-sim_null_normal <- function(nxj, nfacetj, mean, sd, w1 = 0, w2=0) {
-  rep(distributional::dist_normal(mu = mean, sigma = sd),
-      times = nxj * nfacetj
-  )
-}
-
-sim_varf_normal <- function(nx, nfacet, mean, sd, w1, w2) {
-  rep(dist_normal((mean + seq(0, nfacet - 1, by = 1) * w1),
-                  (sd + seq(0, nfacet - 1, by = 1) * w2)), each = nx)
-}
-
-sim_varx_normal <- function(nx, nfacet, mean, sd, w1, w2) {
-  rep(dist_normal((mean + seq(0, nx - 1, by = 1) * w1),
-                  (sd + seq(0, nx - 1, by = 1) * w2)), nfacet)
-}
-
-sim_varall_normal <- function(nx, nfacet, mean, sd, w1, w2) {
-  dist_normal((mean + seq(0, (nx *nfacet - 1), by = 1) * w1),
-              (sd + seq(0,  (nx *  nfacet - 1), by = 1) * w2))
-}
-
-## ----data
-
-library(hakear)
-
-nx_val = 2 # number of x-axis levels
-nfacet_val = 3 # number of facet levels
-w1_val = 1 # increment in mean
-w2_val = 0 # increment in sd
-mean_val = 0 # mean of normal distribution of starting combination
-sd_val = 2 # sd of normal distribution of starting combination
-quantile_prob_val = seq(0.1, 0.9, 0.1)
-ntimes_val = 300
-
-sim_panel_varall <- sim_panel(
-  nx = nx_val, nfacet =  nfacet_val,
-  ntimes = ntimes_val,
-  # sim_dist = sim_varall_normal(2, 3, 5, 10, 5, -1.5)
-  sim_dist = sim_varall_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
-) %>% unnest(data)  %>% 
-  rename( "g2" = "id_facet",
-          "g1" = "id_x")
-
-sim_panel_varx <- sim_panel(
-  nx = nx_val, nfacet =  nfacet_val,
-  ntimes = ntimes_val,
-  # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
-  sim_dist = sim_varx_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
-) %>% unnest(data)  %>% 
-  rename( "g2" = "id_facet",
-          "g1" = "id_x")
-
-sim_panel_varf <- sim_panel(
-  nx = nx_val, nfacet =  nfacet_val,
-  ntimes = ntimes_val,
-  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
-  sim_dist = sim_varf_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
-) %>% unnest(data)  %>% 
-  rename( "g2" = "id_facet",
-          "g1" = "id_x")
-
-sim_panel_null <- sim_panel(
-  nx = nx_val,
-  nfacet =  nfacet_val,
-  ntimes = ntimes_val,
-  sim_dist = distributional
-  ::dist_normal(mean_val, sd_val)
-) %>% unnest(c(data)) %>% 
-  rename( "g2" = "id_facet",
-          "g1" = "id_x")
-
-## ----category-plots
-
-p_null <- sim_panel_null %>%
-  ggplot(aes(x = as.factor(g2), y = sim_data)) +
-  facet_wrap(~g1, labeller = "label_both") +
-  geom_boxplot() +
-  ylab("") +
-  #ggtitle(paste("(a)", round(null, 2))) +
-  xlab("g2") +
-  theme_bw()+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") +
-  #ggtitle("design1")+
-  theme_validation()
-
-p_varf <- sim_panel_varf %>%
-  ggplot(aes(x = as.factor(g2), y = sim_data)) +
-  facet_wrap(~g1, labeller = "label_both") +
-  geom_boxplot() +
-  ylab("") +
-  #ggtitle(paste("(b)", round(varf, 2))) +
-  xlab("g2")+
-  theme_bw()+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") +
-  #ggtitle("design2")+
-  theme_validation()
-
-p_varx <- sim_panel_varx %>%
-  ggplot(aes(x = as.factor(g2), y = sim_data)) +
-  facet_wrap(~g1, labeller = "label_both") +
-  geom_boxplot()+
-  ylab("")  +
-  #ggtitle(paste("(c)", round(varx, 2))) +
-  xlab("g2")+
-  theme_bw()+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") +
-  #ggtitle("design3")+
-  theme_validation()
-
-p_varall <- sim_panel_varall %>%
-  ggplot(aes(x = as.factor(g2), y = sim_data)) +
-  facet_wrap(~g1, labeller = "label_both") +
-  geom_boxplot() +
-  ylab("") +
-  #ggtitle(paste("(d)", round(varall, 2))) +
-  xlab("g2")+
-  theme_bw()+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") +
-  #ggtitle("design4") +
-  theme_validation()
-
-change_index <- function(data){
-
-index_new <- map(seq_len(ntimes_val), function(i){
-  map((seq_len(nx_val*nfacet_val)), function(j)
-    {
-value = i + (j-1)*ntimes_val
-})
-  }) %>% unlist()
-
-data_new = data %>%
-  ungroup() %>%
-  mutate(index_old = row_number(),
-         index_new = index_new)
-
-y = data_new[match(index_new, data_new$index_old),]
-
-y <- y %>%
-  mutate(time = row_number())
-
-return(y)
-}
-
-endbreaks<- nrow(sim_panel_null)
-
-p1 <- change_index(sim_panel_null) %>%
-  ggplot(aes(x = time,
-             y = sim_data)) +
-  geom_line() +
-  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
-  theme_bw() +
-  #geom_point(alpha = 0.5, color = "blue")
-  #theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
-    #    panel.grid.minor.x =  element_blank())+
-  ylab("simulated response")+
-  xlab("index")
-
-p2 <- change_index(sim_panel_varf) %>%
-  ggplot(aes(x = time,
-             y = sim_data)) +
-  geom_line() +
-  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
-  theme_bw() +
-  #geom_point(alpha = 0.5, color = "blue") +
-  #theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
-     #   panel.grid.minor.x =  element_blank())+
-  ylab("simulated response")+
-  xlab("index")
-
-p3 <- change_index(sim_panel_varx) %>%
-  ggplot(aes(x = time,
-             y = sim_data)) +
-  geom_line() +
-  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
-  theme_bw() +
-  #geom_point(alpha = 0.5, color = "blue")+
-  #theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
-   #     panel.grid.minor.x =  element_blank())+
-  ylab("simulated response")+
-  xlab("index")
-
-p4 <- change_index(sim_panel_varall) %>%
-  ggplot(aes(x = time,
-             y = sim_data)) +
-  geom_line() +
-  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
-  theme_bw() +
-  #geom_point(alpha = 0.5, color = "blue") +
-  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
-  panel.grid.minor.x =  element_blank())+
-  ylab("simulated response")+
-  xlab("index")
-
-(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)
-
-```
-
-<!-- When two granularities of interest interact, the connection between a granularity and the measured variable is determined by the value of the other interacting granularity.
-This happens when the effects of the two granularities on the measured variable are not additive.
-For simplicity, consider a case with just two interacting granularities $g1$ and $g2$ of interest.
-As opposed to the last case, where we could play with the distribution of $5$ individual categories, with interaction we can play with the distribution of $6$ combination of categories.
-Consider $4$ designs in Figure \ref{fig:} where different distributions are assumed for different designs to get some distinction across designs.
-For example, in application, think about the scenario when customers need to grouped basis their joint behavior across hour-of-day and month-of-year.
--->
-
-| Granularity type                                                   	| # Significant 	| # Replications 	|
-|--------------------------------------------------------------------	|---------------	|----------------	|
-| **Individual**  <br><br># obs: 300, 500, 2000  <br># clusters: 6/7 	| 1/2/3         	| 25, 100, 200   	|
-| **Interaction**  <br><br># obs: 500, 2000  <br># clusters: 4       	| 1/2           	| 25, 100, 200   	|
-
-## Results
-
-All the methods were fitted to each data designs and results are reported through confusion matrices.
-With increasing difference between categories, it gets easier for the methods to correctly distinguish the designs.
-For $mean_diff=1$, the performances are pretty bad for js-robust methods and wpd method for lower nT.
-Although, with the kind of residential load datasets, a full year of load is the minimal requirement to capture expected variations in winter and summer profiles, for example.
-It is likely that $nT$ would be at least $1000$ with half-hourly data, even if data is only available just for a month.
-The performance is promising except when the number of observations for a customer is really small.
-For smaller difference between categories, it is expected that method js-nqt would preform better than the other two.
-
-```{r}
-# code in append_3gran_change.R
-table <- read_rds("data/append_3gran_change.rds")
-table %>% kable()
-```
-
-<!-- starts getting better with increasing difference and get worse with increasing number of replications.
-Length of series do not show to have any effect on the performance of the methods.
-It does not depend on if time series is ar or arma.
--->
-
-<!-- - confusion matrix could be used for showing results if proper labeling is used -->
-
-<!-- - write about features that we have spiked into the data set -->
-<!-- - write about you incorporated noise -->
-<!-- - What is the additional structure you can incorporate that will lead to failing of method1 and method2? -->
-<!-- - And both gives the same result? Basically say when method 1 works better than method 2 and vice versa! -->
-
-<!-- - -->
-
-# Application {#sec:application}
-
-The use of our methodology is illustrated on smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
-It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
-It also provides demographic data for these customers most of which are missing and not utilized for the purpose of this paper.
-To maintain anonymity, the energy patterns could not be recognised at the person level, but rather by the geographical location of their dwelling and information about their Local Government Area.
-
-<!-- ## Data source -->
-
-<!-- The entire data is procured from CSIRO.
-A subset of this data is also available from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
-It consists of the following data sets.
--->
-<!-- _1. CustomerData:_ 78720 customers with 62 variables about them  -->
-<!-- _2. EUDMData:_ 300 billion half-hourly consumption level data   -->
-<!-- _3. OffersData:_ Method of contact to customer to join SGSC customer trial, either door-to-door (D2D) or via Telesales   -->
-<!-- _4. PEResponseData:_ Peak Events response customer wise   -->
-<!-- _5. PETimesData:_ Peak Events time stamps   -->
-
-<!-- Only _CustomerData_ and _EUDMData_ are relevant for the clustering goals of this paper.
-_EUDMData_ contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
-`CustomerData` provides demographic data for 78,720 customers most of which are missing and not utilized for the purpose of this paper.
-To meet the requirements for anonymity preservation, the energy patterns could not be identified at the individual level, but rather by the geographical location of their residence information about their Local Government Area.
--->
-
-<!-- ## Characteristics of raw data -->
-
-In Figure \ref{fig:raw-data-50}, the time series of energy consumption is plotted along the y-axis against time from past to future for $50$ sampled households.
-Each of these series correspond to a single customer.
-For each customer, the energy consumption is available at fine temporal resolution (every 30 minutes) for a long period of time (~ 2 years) resulting in 27,000 (median) observations for each customer.
-Some customers' electricity use may be unavailable owing to power outages or improper recording, resulting in implied missing numbers in the database.
-For this data set it was found that out of 13,735 customers in total, 8,685 customers do not have any implicit missing observations, while the rest 5,050 customers had missing values.
-With further exploration, it was found that there is no structure in the missing-ness, that is missing observations can occur at any time point (see Appendix).
-Moreover, the data for these customers are characterized by unequal length, different start and end dates.
-Since our proposed methods consider probability distribution instead of raw data, both of these characteristics would not pose any threat to our methodology unless of course there is any structure or systematic patterns in them.
-
-\noindent It can be expected that energy consumption vary substantially between customers, which is a reflection of their varied  behavior owing to differences in profession, family size, geographical or physical characteristics.
-Since the linear time series plot has too many measurements all squeezed in this linear representation, it hinders us to discern any repetitive behavioral pattern for even one customers (let alone many customers together).
-In most cases, electricity data will have multiple seasonal patterns like daily, weekly or annual.
-We do not learn about these repetitive behaviors from the linear view.
-Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.
-
-```{r raw-data-50, out.width = "100%", fig.cap="The raw data for 50 households are shown.
-It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed.
-No insightful behavioral pattern could be discerned from this view other than when the customer is not at home."}
-
-knitr::include_graphics("figs/raw_plot_cust.png") # look at smart-meter.R for the code
-```
-
-<!-- ### Missing Data -->
-
-<!-- Electricity usage for some customers may become unavailable due to power outage or not recording their usage properly, thus resulting in implicit missing values in the database.
-It is interesting to explore where missing-ness occurs or if there is a relationship between the underlying missing patterns.
-We use the R package `tsibble` to do this.
--->
-
-```{r missing-data, out.width="100%", eval = FALSE}
-include_graphics("figs/missing-data.png")
-```
-
-<!-- - if there is any systematic missing patterns in the data -->
-<!-- - this missing plot can go in the supplementary -->
-<!-- - how to add missing values (should be added in data pre-processing) -->
-<!-- - instance learning -->
-<!-- - types of summary techniques to use ( -->
-<!-- generally multivariate means and sd are used, I can't use that in a time series context, you can show across different granularity? -->
-<!-- within-group sum of squares and between-group sum of squares -->
-
-<!-- ) -->
-
-<!-- 13735 customers in elec_ts -->
-<!-- 8685 customers in elec_nogap -->
-<!-- 5050 customers in count_na_df -->
-
-<!-- Then is the graph of missing observations even interesting. -->
-<!-- You can show two graphs, one to show that missingness do not have a pattern -->
-<!-- another to show even if no missing, they start and end at different times. (A sample of 50 customers).  -->
-
-<!-- \noindent A dataset of 100 SGSC homes has been used to lay out the structure to be used for analyzing the big dataset.
-The smaller dataset contains half-hourly kwh values form 2012 to 2014 and has asynchronous time series distributed evenly over the observation period (Figure \ref{fig:elec-raw}), similar to the bigger data set.
-Figure \ref{fig:count-gaps} can be used to interpret missingness in the data, where the customers are arranged from maximum to minimum missing.
-It looks like data is most missing before 2013 and for a particular date in 2014.
--->
-
-```{r miss-data}
-empty_as_na <- function(x){
-    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
-    ifelse(as.character(x)!="", x, NA)
-}
-
-```
-
- <!-- - Handling autocorrelation: Autocorrelation in the time series is likely to get removed from considering probability distributions and cyclic granularities in the clustering algorithm. -->
-
-<!-- At this stage, we need to define the aim of clustering as there could be various aims of clustering like between-cluster separation, within cluster homogeneity: low distances, within-cluster homogeneous distributional shape, good representation of data by centroids, little loss of information, high density without cluster gaps, uniform cluster sizes, stability and others.
-Finally, how distinct they are and how can we summarize the main features of the cluster would be discussed here.
--->
-
-## Prototype selection
-
-<!-- A clean data set is obtained by carefully choosing customers which shows similar shapes across one or more cyclic granularity.
-Since this is unlabeled data, there is no way to do external validation of our methodologies.
-Thus, we chose this way to see how well our methodology works in a cleaner data set as this one.
--->
-
-<!-- Why instance selection -->
-In supervised learning, a training set containing previously known information is used to categorize new occurrences.
-Acceptable classification rates may be obtained by discarding instances which are not helpful for classification; this process is known as instance selection (@olvera2010review).
-This is similar to subsetting the population along all dimensions of importance such that the sampled data is representative of the main characteristics of the underlying distribution.
-Instance selection in unsupervised learning has received limited attention in the literature, but could serve as an useful way to sample evaluation data set to measure the performance of a model or method.
-One such procedure is suggested in @Fan2021-bq that selects similar instances (neighbors) for each instance (anchor) and treats the anchor and its neighbors as the same class.
-In this section, a similar idea is used to select customers with prototype behaviors
-that serves as evaluation data sets for our proposed methodology.
-
-First, we select the customers which do not have any implicit missing values and filtered their data for the year 2013.
-From this set, we randomly sample a set of $600$ customers.
-We obtain the $wpd$ for all cyclic granularities considered for these customers and found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weeknd/weekday) are coming out to be significant for most customers.
-This implies that for most customers, there is some interesting pattern across these three granularities.
-Potentially this could be done for the entire data set.
-This step is time consuming and hence has been only run for the $600$ sampled customers.
-As a second step, we remove customers for which data in any category for the significant granularities are empty.
-For example, in this data set, if a customer do not have data for an entire month, they have been removed as their monthly behavior could not be studied in that case.
-Further, we also remove customers for which all the deciles of the energy consumption is zero.
-These are the customers whose consumption remain mostly flat and is expected to have no interesting repetitive patterns that is our interest of study.
-Finally, we are left with $356$ customers.
-From this set we select $4$ "anchor" customers which are far apart from each other and $5$ neighboring customers for each of these anchors.
-These selections were done using the granularity $hod$ space.
-It is important to note that when we use our proposed methodologies, it is based on all dimesions `hod`, `moy` and `wkndwday`.Fig \ref{fig:} shows the MDS of these 356 customers in a 2D space basis their distance on individual granularities and when all of them are combined.
-Our methodolgies are run on these $24$ customers, which act as a way to evaluate the proposed methodologies.
-
-<!-- However, if `hod` is and not only `hod`, but if `hod` is significant we can expect that the grouping would be similar to how we have initially chosen the set of customers. -->
-
-<!-- from which we wanted to do instance selection.
-Since this is unlabeled data, there is no way to do external validation of our methodologies.
-Thus, we chose this way to see how well our methodology works in a cleaner data set as this one to classify customers who have similar behavior across all significant granularities.
-Next we select $4$ customers which are far apart from each other and $5$ customers which are lying closest in distance to each of these $4$ customers.
-These selection was done using the granularity $hod$.
-When we use our methodologies, it is based on all significant granularities and not only hod, but if $hod$ is significant we can expect that the grouping would be similar to how we have initially chosen the set of customers.
--->
-
-<!-- .
-Instances may be distributed in the representation space in a reasonable manner, revealing their similarities, using this instance wise discriminative learning.
-A dual-level progressive similar instance selection (DPSIS) approach could also be used -->
-
-<!-- Fig \ref{fig:data-heatmap-hod}, \ref{fig:} and \ref{fig:} shows the distribution of 24 customers across hour-of-day, month-of-year and wknd-wday respectively.
-Every row in \ref{fig:} shows different shapes across hour-of-day and across columns show similar shapes for each row.
-We use our methodology to see if the customers are allocated to same group have similar shapes across one or more significant granularity.
--->
 
@@
 
-# 8454221 10420689
+# 8454221 10420689 1001 8454221 10018298 10420689 9393696
 data_pick_one <- total_distance %>% filter(item1 %in% 
-                            c(8454221)) %>% 
+                            c(8454221)|item2 %in% 
+                            c(8454221) ) %>% 
   #group_by(item1) %>%
@@
 data_pick_ch <-  total_distance %>% filter(item1 %in% 
-                            c(9393696)) %>% group_by(item1) %>% arrange(total) %>% head(5) %>% 
+                            c(8454221)|item2 %in% 
+                            c(8454221)) %>% group_by(item1) %>% arrange(total) %>% head(5) %>% 
   mutate(item1 = as.integer(as.character(item1)),
@@
 data_pick_two <- total_distance %>% filter(item1 %in% 
+                            c(8181071)|item2 %in% 
                             c(8181071)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
@@
 data_pick_four <- total_distance %>% filter(item1 %in%
+                            c(8184707)|item2 %in% 
                             c(8184707)) %>% 
@@
              ncol = 6) +
-    theme_bw() +
    theme(axis.title.y=element_blank(),
@@
   theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ 
+  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + 
+  theme(legend.position = "bottom")+ 
   #theme() strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
@@
   scale_x_discrete(breaks = seq(0,23, 4)) + 
-  scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+  #theme_characterisation() +
   xlab("hour-of-day") 
@@
   theme(plot.margin = margin(0, 0, 0, 0, "cm") )+
-  scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
   xlab("month-of-year") +  theme(axis.text.x = element_text(size=5))
@@
   #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
+  theme(plot.margin = margin(0, 0, 0, 0, "cm"))+
   theme(legend.position = "bottom")
@@
   theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
   theme(legend.position = "bottom") +
@@
   theme(plot.margin = margin(0, 0, 0, 0, "cm") ) +
-  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
   theme(legend.position = "bottom")
@@
   theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
+  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
   theme(legend.position = "bottom") +
@@
 That places them in the same group.
-
-# Discussion {#sec:discussion}
-
-<!-- This section will cover some drawback of this clustering method and potential extensions of this work. -->
-
-<!-- There are many cluster analysis methods that can produce quite different clusterings on the same dataset.
-Cluster validation is about the evaluation of the quality of a clustering; “relative cluster validation” is about using such criteria to compare clusterings.
-This can be used to select one of a set of clusterings from different methods, or from the same method ran with different parameters such as different numbers of clusters.
-There are many cluster validation indexes in the literature.
-Most of them attempt to measure the overall quality of a clustering by a single number, but this can be inappropriate.
-There are various different characteristics of a clustering that can be relevant in practice, depending on the aim of clustering, such as low within-cluster distances and high between-cluster separation.
-In this paper, a number of validation criteria will be introduced that refer to different desirable characteristics of a clustering, and that characterise a clustering in a multidimensional way.
-In specific applications the user may be interested in some of these criteria rather than others.
-A focus of the paper is on methodology to standardise the different characteristics so that users can aggregate them in a suitable way specifying weights for the various criteria that are relevant in the clustering application at hand.
--->

### 2021-10-13 main.Rmd
wkndwday pattern changed to boxplot

@@
 These selections were done using the granularity $hod$ space.
-It is important to note that when we use our proposed methodologies, it is based on all dimesions `hod`, `moy` and `wkndwday`.Fig \ref{fig:} shows the MDS of these 356 customers in a 2D space basis their distance on individual granularities and when all of them are combined.
+It is important to note that when we use our proposed methodologies, it is based on all dimensions `hod`, `moy` and `wkndwday`.Fig \ref{fig:mds} shows the MDS of these 356 customers in a 2D space basis their distance on individual granularities and when all of them are combined.
 Our methodolgies are run on these $24$ customers, which act as a way to evaluate the proposed methodologies.
+
+```{r mds, out.width="100%", fig.cap="Instance selection with respect to `hod` such that there are two distinct group and two overlapping groups."}
+knitr::include_graphics("figs/mds_plot.png")
+```
 
@@
 
-wkndwday_group <- data_group %>% create_gran("wknd_wday") %>% 
+wkndwday_data<- data_group %>% create_gran("wknd_wday") 
+ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]
+
+wkndwday_group <- wkndwday_data%>% 
   ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
-  lvplot::geom_lv(aes(fill = as.factor(group))) +
-  #geom_boxplot(aes(fill = as.factor(group))) +
+  #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
+  geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
+  #geom_boxplot(outlier.shape = NA) + 
+  coord_cartesian(ylim = ylim1*1.05)+
   #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
@@
              scales = "free_y", 
-             labeller = "label_value",
+             labeller = "label_both",
              nrow = 4)+
@@
   scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+  scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
   theme(legend.position = "bottom") +
@@
 
-wkndwday_group_wpd <- data_heatmap_wkndwday_group %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
-  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+wkndwday_data<- data_group %>% create_gran("wknd_wday") 
+ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]
+
+wkndwday_group_wpd <- wkndwday_data%>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh))+
+  geom_boxplot(aes(fill = as.factor(group),color = as.factor(group)),alpha = 0.5) + 
+  coord_cartesian(ylim = ylim1*1.05) +
+  #ggplot(aes(x = category)) + 
+  # geom_ribbon(aes(ymin = `25%`, 
+  #                 ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+  # geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
   facet_wrap(~group, 
              scales = "free_y", 
-             labeller = "label_value",
+             labeller = "label_both",
              nrow = 4)+
@@
 
+## Classifying the 356 customers
+
 # Discussion {#sec:discussion}
+
+We propose different clustering methodology for grouping noisy, patchy time series data available at a fine temporal scale.
+Depending on the aim of clustering, they produce different clustering.
+The clustering is done based on probability distributions of the time series variable measured across several cyclic granularirties.
+There is some issue with scaling it up to many customers as anomalies need to be removed before such classification would be useful.
 

### 2021-10-13 supplementary/distance-computation/real-data/04_handpick_from356-mds_nqt.Rmd
wkndwday pattern changed to boxplot

@@
 
+cols <- toRGB(RColorBrewer::brewer.pal(5, "Dark2"))
 subplot(mds_total, mds_hod, mds_moy ,mds_wkndwday, titleX = TRUE) %>%
   layout(showlegend = FALSE) %>%
-  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect") 
+  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect", persistent = TRUE,defaultValues = data_pick_cust$value, color = cols)
 ```
@@
   layout(showlegend = FALSE) %>%
-  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect")
+  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect", persistent = TRUE)
 ```

### 2021-10-13 supplementary/distance-computation/real-data/04handpick-from356-mds.Rmd
wkndwday pattern changed to boxplot

@@
   layout(showlegend = FALSE) %>%
-  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect", persistent = TRUE) 
+  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect", persistent = TRUE,
+defaultValues = unique(data_pick$customer_id))
+
+            # defaultValues = c(8181071, 8184707, 8191495, 8269176, 8328008, 8454221, 8454235,  8598020, 8627007, 8636035, 8653709, 8659512, 8664102, 9345642, 9587992, 9822544, 10257830, 10357256, 10420689, 10582471, 10767756, 10909394, 11006388, 11013154))
 ```

### 2021-10-13 main.Rmd
wkndwday pattern changed to boxplot

@@
   #ggtitle(paste("(a)", round(null, 2))) +
-  xlab("g2") +
+  xlab("") +
   theme_bw()+ stat_summary(
@@
   #ggtitle("design1")+
-  theme_validation()
+  theme_validation() +ylab("")  +
+  theme(panel.spacing =unit(0, "lines"))+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
 
@@
   #ggtitle(paste("(b)", round(varf, 2))) +
-  xlab("g2")+
+  xlab("")+
   theme_bw()+ stat_summary(
@@
   #ggtitle("design2")+
-  theme_validation()
+  theme_validation()+ylab("") +
+  theme(panel.spacing =unit(0, "lines"))+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
 
@@
   #ggtitle(paste("(c)", round(varx, 2))) +
-  xlab("g2")+
+  xlab("")+
   theme_bw()+ stat_summary(
@@
   #ggtitle("design3")+
-  theme_validation()
+  theme_validation()+ylab("") +
+  theme(panel.spacing =unit(0, "lines"))+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
 
@@
   #ggtitle(paste("(d)", round(varall, 2))) +
-  xlab("g2")+
+  xlab("")+
   theme_bw()+ stat_summary(
@@
   #ggtitle("design4") +
-  theme_validation()
+  theme_validation()+ylab("") +
+  theme(panel.spacing =unit(0, "lines"))+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
 
@@
   geom_line() +
-  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
+  scale_x_continuous(breaks = seq(1, endbreaks, 200))+
   theme_bw() +
@@
     #    panel.grid.minor.x =  element_blank())+
-  ylab("simulated response")+
-  xlab("index")
+  ylab("")+
+  xlab("")
 
@@
      #   panel.grid.minor.x =  element_blank())+
-  ylab("simulated response")+
-  xlab("index")
+  ylab("")+
+  xlab("")
 
@@
    #     panel.grid.minor.x =  element_blank())+
-  ylab("simulated response")+
-  xlab("index")
+  ylab("")+
+  xlab("")
 
@@
   panel.grid.minor.x =  element_blank())+
-  ylab("simulated response")+
-  xlab("index")
-
-(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)
+  ylab("")+
+  xlab("")
+
+((p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall))
 
@@
 
-##  js-based clustering
+##  JS-based clustering
 
@@
 The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-According to our prototype selection method, each row represents a distinct shape of daily load, and all customers in the same row should have similar daily profile.After clustering these consumers, all customers with the same colour represent the same group.
+According to our prototype selection method, each row represents a distinct shape of daily load, and all customers in the same row should have similar daily profile.
+After clustering these consumers, all customers with the same color represent the same group.
 Except for customer id 8269176, there is unanimity across design and groups.
@@
 So, while our clustering approach failed to place the customer in the correct hour-of-day group, it did so for month-of-year.
-When contrasting the moy to hod profile, there are greater behavioural differences across customers within a group.
+When contrasting the `moy` to `hod` profile, there are greater behavioral differences across customers within a group, which implies that `hod` has the minimum within-cluster variation.
 
@@
 
-Characterization of clusters is the final stage of a cluster analysis.
-If we are convinced that we have identified a collection of clusters that can be distinguished from one another, we should intend to characterize them more formally, both statistically and qualitatively.
-We quantify them by producing statistics for each variable.
-We may look at the findings in graphs, and enhance our qualitative descriptions of the groupings.
-
-\noindent We discover 4 qualitative clusters of varying shapes in the distribution of all consumers in the first panel of Figure \ref{fig:combined}.
+Characterization of clusters both statistically and qualitatively is an important stage of a cluster analysis.
+A potential way is to look at the findings from all the groups in graphs, and enhance our qualitative descriptions of the groupings.
+Figure \ref{fig:combined} shows the distribution of the summarized groups and help us to characterise each of the clusters.
+All of these may be validated with further information about the customer.
+
+<!-- We quantify them by producing statistics for each variable. We may look at the findings in graphs, and enhance our qualitative descriptions of the groupings. -->
+
+<!-- \noindent We discover 4 qualitative clusters of varying shapes in the distribution of all consumers in the first panel of Figure \ref{fig:combined}.
 Group-1 includes consumers who work 9-5, get up and conduct morning activities from 7-10am, and then depart.
@@
 All of this may be validated with further information about the customer.
-
-\noindent The second panel of Figure \ref{fig:combined} shows that month-of-year qualitative clusters are not as distinguishable as hour-of-day.
+-->
+
+<!-- \noindent The second panel of Figure \ref{fig:combined} shows that month-of-year qualitative clusters are not as distinguishable as hour-of-day.
 Group 2 is the most distinct and uses the most power during the summer, possibly owing to the use of air conditioners.
@@
 Since gas is not available in all of NSW LGAs, it is possible that customers' heater usage is recorded in electricity rather than gas.
-
+-->
+
+<!-- \noindent The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
+-->
+
+_Group 1:_ This includes consumers who work 9-5, get up and conduct morning activities from 7-10am, and then depart.
+Then they return home in the evening to cook supper and perform other activities, giving the evening a greater peak than the morning.
+These users in this cluster are insensitive to high temperature but have their heaters on in the winter months May-Aug.
+
+_Group 2:_  This is the group that rushes out of the house in the morning to get to work.
+They only return at night and do all activities at night, so there is no morning peak.
+In addition, the users in this cluster are insensitive to high temperature, but sensitive to lower temperature.
+
+_Group 3:_  This group a strong early morning and late night hours.
+These consumers may be flexible students or elderly retirees who are night owls.
+They have heaters on in the winter but consume less energy in the summer.
+Winter usage may also be due to increased usage of heaters at night when they are up.
+
+_Group 4:_ Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
+They have a flat monthly profile, indicating the usage across all months are similar.
+It seems that the users in this group are more concerned about the comfort and quality of life than the cost of electricity as opposed to Group 1 and 3, who do not consume more electricity in the summer.
+
+\noindent The plotting scales are not displayed since we want to emphasise comparable shapes rather than scales.
+A customer in the cluster may have low daily or total energy usage, but their behavior may be quite similar to a customer with high usage.
 \noindent The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
-
-\noindent The plotting scales are not displayed since we want to emphasise comparable shapes rather than scales.
-A customer in the cluster may have low daily or total energy usage, but their behaviour may be quite similar to a customer with high usage.
-That places them in the same group.
 
@@
 The ggpairs plot four distinct clusters across the month-of-year, which are less prominent across the hour-of-day and wknd-wday.
-The parallel coordinate plot ranks the variables in order of importance, indicating that the month-of-year is the most important in identifying clusters, whereas wkdn-wday is the least significant and has the least variability among the three variables." }
+The parallel coordinate plot ranks the variables in order of importance, indicating that the month-of-year is the most important in identifying clusters, whereas wkdn-wday is the least significant and has the least variability among the three variables.", message=FALSE, warning=FALSE}
 
@@
   scale_color_viridis_d(direction = -1) + theme_light()
+
+pairsplot
+
 ```
@@
 parcoord
+```
+
+```{r data-heatmap-hod-new-wpd, fig.cap="The distribution of electricity demand for the selected 24 customers across hour-of-day.
+The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
+According to our prototype selection method, each row represents a distinct shape of daily load, and all customers in the same row should have similar daily profile.After clustering these consumers, all customers with the same colour represent the same group.
+Except for customer id 8269176, there is unanimity across design and groups.
+Even though their daily form resembles Group 2, our clustering approach places them in Group 3 (which is design 4 in our case).
+Because our method uses hod, moy, and wkndwday, there may be some mismatches depending on only one variable."}
+
+cluster_result_wpd$id <- as.character(cluster_result_wpd$id)
+
+data_group <- data_pick %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result_wpd, by = c("customer_id" = "id"))
+
+data_heatmap_hod %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result_wpd, by = c("customer_id" = "id")) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=customer_id,
+                  fill = as.factor(group)),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=customer_id, 
+                color = as.factor(group)), size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+  theme_light() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank()) +
+    theme(strip.text = element_text(size = 10,
+                                    margin = margin(b = 0, t = 0))) +
+  ylab("demand (in Kwh)")  + 
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ 
+  #theme() strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") ) +
+  scale_x_discrete(breaks = seq(0,23, 4)) +
+  scale_fill_viridis_d(direction = -1) +
+  scale_color_viridis_d(direction = -1) + 
+  theme(legend.position = "bottom") +
+  xlab("hour-of-day") 
+```
+
+```{r data-heatmap-moy-wpd, fig.cap = "The distribution of electricity demand for the selected 24 customers across month-of-year.
+The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
+Intriguingly, customer id 8269176 appears to be in the appropriate group (Group 3) because its monthly profile is more similar to Group 3 than Design 4.
+So, while our clustering approach failed to place the customer in the correct hour-of-day group, it did so for month-of-year.
+When contrasting the moy to hod profile, there are greater behavioural differences across customers within a group."}
+
+data_heatmap_moy <- quantile_gran(data_pick, "month_year", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
+  left_join(data_pick_cust, by = c("customer_id"="value"))
+
+data_heatmap_moy$category <- factor(data_heatmap_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+data_heatmap_moy %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result_wpd, by = c("customer_id" = "id")) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, 
+                  group=customer_id, 
+              fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+    theme(strip.text = element_text(size = 10,
+                                    margin = margin(b = 0, t = 0))) + xlab("month-of-year") +
+  ylab("demand (in Kwh)")  +
+  theme_light() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank()) + 
+  theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm")) + scale_fill_viridis_d(direction = -1) +
+  scale_color_viridis_d(direction = -1) + 
+  xlab("month-of-year") +
+  theme(axis.text.x = element_text(size=5)) +
+  theme(legend.position = "bottom")
 ```
@@
   ylab("demand (in Kwh)") + 
-  theme_bw() +
+  theme_light() +
    theme(axis.title.y=element_blank(),
@@
   theme(plot.margin = margin(0, 0, 0, 0, "cm"))+ scale_fill_viridis_d(direction = -1) +
-  scale_color_viridis_d(direction = -1) + theme_light() 
+  scale_color_viridis_d(direction = -1) + 
   theme(legend.position = "bottom")
@@
   ylab("demand (in Kwh)") +
-  theme_bw() +
+  theme_light() + 
    theme(axis.title.y=element_blank(),
@@
 scale_fill_viridis_d(direction = -1) +
-  scale_color_viridis_d(direction = -1) + theme_light() + 
+  scale_color_viridis_d(direction = -1) + 
   theme(legend.position = "bottom") +
@@
   theme(plot.margin = margin(0, 0, 0, 0, "cm"))+ scale_fill_viridis_d(direction = -1) +
-  scale_color_viridis_d(direction = -1) + theme_light() + 
+  scale_color_viridis_d(direction = -1) + 
   theme(legend.position = "bottom") +
@@
 
-\noindent We discover 4 qualitative clusters of varying shapes in the distribution of all consumers in the first panel of Figure \ref{fig:combined}.
+A parallel coordinate plot with the three significant cyclic granularities used for wpd-based clustering.
+The variables are sorted according to their separation across classes (rather than their overall variation between classes).
+This means that moy is the most important variable in distinguishing the designs followed by hod and wkndwday.
+It can be observed that cluster 3 and 4 are distinguished by moy while 1 and 2 are distinguished by hod.
+Here, wkndwday is still acting as the nuisance variable.
+The ggpairs plot four distinct clusters across the month-of-year, which are less prominent across the hour-of-day and wknd-wday.
+The parallel coordinate plot ranks the variables in order of importance, indicating that the month-of-year is the most important in identifying clusters, whereas wkdn-wday is the least significant and has the least variability among the three variables.
+
+<!-- \noindent We discover 4 qualitative clusters of varying shapes in the distribution of all consumers in the first panel of Figure \ref{fig:combined}.
 Group-1 includes consumers who work 9-5, get up and conduct morning activities from 7-10am, and then depart.
@@
 Since gas is not available in all of NSW LGAs, it is possible that customers' heater usage is recorded in electricity rather than gas.
-
-\noindent The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
-
-\noindent The plotting scales are not displayed since we want to emphasise comparable shapes rather than scales.
+-->
+
+<!-- \noindent The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
+-->
+
+<!-- \noindent The plotting scales are not displayed since we want to emphasise comparable shapes rather than scales.
 A customer in the cluster may have low daily or total energy usage, but their behaviour may be quite similar to a customer with high usage.
 That places them in the same group.
+-->
 
@@
 The clustering is done based on probability distributions of the time series variable measured across several cyclic granularirties.
-There is some issue with scaling it up to many customers as anomalies need to be removed before such classification would be useful.
+There is issue with scaling it up to many customers as anomalies need to be removed before such classification would be useful.
 

### 2021-10-13 append_3gran_change_5D.R

@@ -59,7 +59,7 @@ js_nqt <- append_files("js-nqt") %>%
 
 data_all<- bind_rows(wpd, js_robust, js_nqt) %>% 
   #arrange(method, nT, niter, mean_diff) %>% 
-  pivot_wider(c(1:3), names_from = "method", values_from = "estimate")
-
+  pivot_wider(c(1:3), names_from = "method", values_from = "estimate") %>% 
+  mutate(niter = 5*niter) #since there are 5 iterations for each design
 
 write_rds(data_all, "data/append_3gran_change.rds")

### 2021-10-14 supplementary/distance-computation/real-data/04handpick-from356-mds.Rmd
files without rds

@@
 
-wkndwday <- read_rds("../../../data/dist_gran_wkndwday_356cust_robust.rds") %>% broom::tidy()
+wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_robust.rds"))
 
-moy <- read_rds("../../../data/dist_gran_moy_356cust.rds") %>% broom::tidy()
+moy <- read_rds("../../../data/dist_gran_moy_356cust.rds")
 
-hod <- read_rds("../../../data/dist_gran_hod_356cust_robust.rds") %>% broom::tidy()
+hod <- read_rds("../../../data/dist_gran_hod_356cust_robust.rds")
 
@@
   layout(showlegend = FALSE) %>%
-  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect", persistent = TRUE,
-defaultValues = unique(data_pick$customer_id))
+  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect", persistent = TRUE)
 
@@
 
+quantile_hod$category <- factor(quantile_hod$category, levels = 0:23)
+
 quantile_moy <- quantile_gran(elec_356_raw, gran1 = "month_year", quantile_prob_val = quantile_prob_graph)
 
+quantile_moy$category <- factor(quantile_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
 quantile_wkndwday <- quantile_gran(elec_356_raw, gran1 = "wknd_wday", quantile_prob_val = quantile_prob_graph)
+
+quantile_wkndwday$category <- factor(quantile_wkndwday$category, levels = c("Weekday", "Weekend"))
 
@@
 ## run upto this
-
-```{r, eval = FALSE}
-
-distance <- wkndwday %>% 
-  left_join(moy, by = c("item1", "item2")) %>% 
-  left_join(hod, by = c("item1", "item2")) %>% 
-  rename("wkndwday" ="distance.x",
-         "moy" = "distance.y",
-         "hod" = "distance") %>% mutate(item1 = as.integer(as.character(item1)))
-
-#10850478 8486921
-data_pick_one <- total_distance %>% filter(item1 %in% 
-                            c(8454221)) %>% group_by(item1) %>% arrange(total) %>% head(5) %>% 
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
-
-#11013154  8181071 11048034
-data_pick_two <- total_distance %>% filter(item1 %in% 
-                            c(8181071)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
-
-data_pick_three <- total_distance %>% filter(item1 %in% 
-                            c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
-
-data_pick_three <- total_distance %>% filter(item1 %in%
-                            c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
-
-data_pick_four <- total_distance %>% filter(item1 %in%
-                            c(8653709)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
-```
-
-```{r,  eval = FALSE}
-
-data_pick_cust <- bind_rows(
-unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(),    unique(c(data_pick_two$item1,data_pick_two$item2)) %>% as_tibble(), c(11013154, 8495194, 8627007,10109182,10677705,8952846) %>% as_tibble(),unique(c(data_pick_four$item1,data_pick_four$item2)) %>% as_tibble(),
-.id = "design")
-
-data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))%>%
-  filter(customer_id %in% data_pick_cust$value) %>% 
-  gracsr::scale_gran( method = "robust", response = "general_supply_kwh")
-
-sm_hod <- gravitas::create_gran(data_pick, "hour_day")
-
-sm_hod_list <- sm_hod %>% 
-  as_tibble() %>% 
-  select(customer_id, hour_day, general_supply_kwh) %>% 
-  pivot_wider(names_from = hour_day,
-              values_from = general_supply_kwh)  %>% 
-  left_join(data_pick_cust, by = c("customer_id" = "value"))
-
-ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, ncol(sm_hod_list))]))
-nrow_sm <- seq_len(nrow(sm_hod_list))
-
-sm_hod_quantiles_cat <- map(nrow_sm, function(x){
-  map(ncol_sm, function(y){
-   cell <- sm_hod_list%>% 
-     dplyr::filter(customer_id == customer_id[x]) %>% 
-     select(-c(1, ncol(sm_hod_list))) %>% 
-     extract(y) %>% 
-     unlist()
-   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
-})  %>% bind_rows(.id = "categories_serial_id")
-}) %>% bind_rows(.id = "customer_serial_id")
-
-ref_cat <- names(sm_hod_list)[-c(1, ncol(sm_hod_list))] %>% as_tibble() %>% set_names("category") %>%  
-  mutate(categories_serial_id = row_number())
-
-ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
-    mutate(customer_serial_id = row_number())
-
-sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
-  mutate(customer_serial_id = as.integer(customer_serial_id),
-         categories_serial_id = as.integer(categories_serial_id)) %>%
-  left_join(ref_cat, by = "categories_serial_id") %>% 
-  left_join(ref_cust, by = "customer_serial_id") %>% 
-  select(customer_serial_id, categories_serial_id, category, customer_id, everything())
-
-data_heatmap <- sm_quantiles_ref %>% 
- left_join(data_pick %>% distinct(customer_id)) %>% 
-  left_join(data_pick_cust, by = c("customer_id" = "value")) %>% 
-  mutate(categories_serial_id = as.factor(categories_serial_id))
-
-data_heatmap %>% 
-  ggplot(aes(x = categories_serial_id)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=customer_id), fill = "lightblue") +
-  geom_line(aes(y = `50%`, group=customer_id)) +
-  facet_wrap(design~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-             ncol = 6) +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
-  scale_x_discrete(breaks = seq(1, 24, 3))
-
-```
-<!-- # # client side linking -->
-<!-- # ```{r} -->
-<!-- # mds <- bind_cols(mds, customer_id = total_distance_wide$item1) -->
-<!-- #  -->
-<!-- # data_heatmap_new <- data_heatmap %>%  -->
-<!-- #   left_join(mds, by = "customer_id") -->
-
-<!-- tx <- highlight_key(data_heatmap_new, ~customer_id) -->
-
-<!--  # initiate a plotly object -->
-<!--  base <- plot_ly(tx, color = I("black")) %>% -->
-<!--    group_by(customer_id) -->
-<!-- # -->
-<!--  quantile_plot <- base %>% -->
-<!--    group_by(customer_id) %>% -->
-<!--    add_lines(x = ~ categories_serial_id, y = ~`50%`, color = ("black")) %>% -->
-<!--    add_ribbons(x = ~categories_serial_id, -->
-<!--                ymin = ~`25%`, -->
-<!--                ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>% -->
-<!--    layout(showlegend = FALSE) -->
-
-<!--  # fig <- plot_ly(base,color = I("blue")) %>% roup_by(customer_id) -->
-
-<!--  fig_trace <- base %>% -->
-<!--    group_by(customer_id) %>% -->
-<!--    add_trace(x = ~Dim.1, y = ~ Dim.2, text = ~customer_id) -->
-
-<!--  subplot(fig_trace, quantile_plot, widths = c(.6, .4), titleX  TRUE) %>% -->
-<!--    layout(showlegend = FALSE) %>% -->
-<!--    highlight(on = "plotly_selected", dynamic = TRUE, electize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect", persistent = TRUE, defaultValues  c(8271504, 9316697)) -->
-
-<!--  %>% -->
-<!--    highlight(color = toRGB("black")) -->
-
-<!--              , color = list( -->
-<!--    toRGB("red", "black") -->
-<!--    ) -->
-<!--  ) -->
-
-<!-- # mds_hod <- base %>% -->
-<!--   # group_by(customer_id) %>% -->
-<!--   # add_trace(x = ~mds1.hod, y = ~ mds2.hod, text = ~customer_id) -->
-
-<!-- # hod_wide <- hod %>% -->
-<!-- #   pivot_wider(names_from = item2, values_from = distance)%>% -->
-<!-- #   mutate(item1 = as.integer(as.character(item1))) -->
-<!-- #  -->
-<!-- # hod_wide$`8143599` <- NA -->
-<!-- #  -->
-<!-- # hod_wide <- bind_rows(hod_wide, c(item1=11590442)) -->
-<!-- #  -->
-<!-- # mds_data <- hod_wide[-1]  %>% mutate_all(~replace(., is.na(.), 0)) %>% as.dist() %>% clust_gran() -->
-<!-- #   -->

### 2021-10-14 supplementary/distance-computation/real-data/04handpick-from356-mds.Rmd
interactive

@@
 library(plotly)
+#remotes::install_github("Sayani07/gracsr")
 library(gracsr)

### 2021-10-15 check_work.Rmd
cleaning codes

@@
 ---
-title:  Clustering based on probability distributions with application on residential customers
-
-# to produce blinded version set to 1
-blinded: 0
-
-authors:
-- name: Sayani Gupta
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-  thanks: "Email: Sayani.Gupta@monash.edu"
-
-- name: Rob J Hyndman
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-
-- name: Dianne Cook
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-
-keywords:
-- data visualization
-- statistical distributions
-- time granularities
-- calendar algebra
-- periodicties
-- grammar of graphics
-- R
+title:  Selecting customers
 
@@
 
-  theme_bw() + 
-      theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) +
+  theme_bw() + # seeting theme
+    theme(strip.text = element_text(size = 7,
+                                    margin = margin(b = 0, t = 0))) + # narrow facet space
    theme(axis.title.y=element_blank(),
         axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) +
-  theme(panel.spacing =unit(0, "lines")) + 
-    theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+        axis.ticks.y=element_blank()) + # no axis ticks 
+  theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
+    theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + # rotate the x-axis text
     theme(legend.position = "bottom")+
-    theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
   theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
-  theme(legend.position = "bottom")
+  theme(axis.text.x = element_text(size=5))
 }
+```
+
+```{r anchor}
+anchor1 <- 11156027
+anchor2 <- 10420683
+anchor3 <- 8629763
+anchor4 <-
+
 ```
@@
 
-wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_robust.rds")) %>% broom::tidy()
+wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_robust.rds"))
 
-moy <- read_rds(here("data/dist_gran_moy_356cust.rds")) %>% broom::tidy()
+moy <- read_rds(here("data/dist_gran_moy_356cust.rds"))
 
-hod <- read_rds(here("data/dist_gran_hod_356cust_robust.rds")) %>% broom::tidy()
-
+hod <- read_rds(here("data/dist_gran_hod_356cust_robust.rds")) 
 distance <- wkndwday %>% 
@@
 data_pick_one <- total_distance %>% filter(item1 %in% 
-                            c(8454221)|item2 %in% 
-                            c(8454221) ) %>% 
+                            c(11156027)|item2 %in% 
+                            c(11156027) ) %>% 
   #group_by(item1) %>%
@@
 
-data_pick_ch <-  total_distance %>% filter(item1 %in% 
-                            c(8454221)|item2 %in% 
-                            c(8454221)) %>% group_by(item1) %>% arrange(total) %>% head(5) %>% 
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
-
 #11013154  8181071 11048034
 data_pick_two <- total_distance %>% filter(item1 %in% 
-                            c(8181071)|item2 %in% 
-                            c(8181071)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
+                            c(10420683)|item2 %in% 
+                            c(10420683)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
   mutate(item1 = as.integer(as.character(item1)),
@@
 
-# data_pick_three <- total_distance %>% filter(item1 %in%
-#                             c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
-#   mutate(item1 = as.integer(as.character(item1)),
-#          item2 = as.integer(as.character(item2)))
-
 # 8653709
 data_pick_four <- total_distance %>% filter(item1 %in%
-                            c(8184707)|item2 %in% 
-                            c(8184707)) %>% 
+                            c(8629763)|item2 %in% 
+                            c(8629763)) %>% 
   #group_by(item1) %>% 
@@
 ```
-
-<!-- # ```{r} -->
-<!-- # data_pick %>% ggplot() + geom_line(aes(x=reading_datetime, y = general_supply_kwh)) + -->
-<!-- #   facet_wrap(~customer_id, ncol= 2) +  -->
-<!-- # ``` -->
 
@@
 
-```{r data-heatmap-hod-new, fig.cap="The distribution of electricity demand for the selected 24 customers across hour-of-day.
-The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-According to our prototype selection method, each row represents a distinct shape of daily load, and all customers in the same row should have similar daily profile.After clustering these consumers, all customers with the same colour represent the same group.
-Except for customer id 8269176, there is unanimity across design and groups.
-Even though their daily form resembles Group 2, our clustering approach places them in Group 3 (which is design 4 in our case).
-Because our method uses hod, moy, and wkndwday, there may be some mismatches depending on only one variable."}
+```{r data-heatmap-hod-new}
 
@@
              labeller = "label_value",
-             ncol = 6) +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) +
-    theme(strip.text = element_text(size = 10,
-                                    margin = margin(b = 0, t = 0))) +
-  ylab("demand (in Kwh)")  + 
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + 
-  theme(legend.position = "bottom")+ 
-  #theme() strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") ) +
-  scale_x_discrete(breaks = seq(0,23, 4)) + 
-  #theme_characterisation() +
-  xlab("hour-of-day") 
+             ncol = 6) + theme_characterisation() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))
 ```
 
-```{r data-heatmap-moy, fig.cap = "The distribution of electricity demand for the selected 24 customers across month-of-year.
-The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-Intriguingly, customer id 8269176 appears to be in the appropriate group (Group 3) because its monthly profile is more similar to Group 3 than Design 4.
-So, while our clustering approach failed to place the customer in the correct hour-of-day group, it did so for month-of-year.
-When contrasting the moy to hod profile, there are greater behavioural differences across customers within a group."}
+```{r data-heatmap-moy}
 
@@
              ncol = 6) +
-    theme(strip.text = element_text(size = 10,
-                                    margin = margin(b = 0, t = 0))) + xlab("month-of-year") +
     ylab("demand (in Kwh)") +
-  theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + 
-  theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )+
-  xlab("month-of-year") +  theme(axis.text.x = element_text(size=5))
+    xlab("month-of-year") +
+  theme_characterisation() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))
 ```
 
-##  js-based clustering
+```{r data-heatmap-wkndwday}
 
-The distribution of electricity demand for the selected 24 customers across hour-of-day is shown in \ref{fig:data-heatmap-hod-new}.
-The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-According to our prototype selection method, each row represents a distinct shape of daily load, and all customers in the same row should have similar daily profile.After clustering these consumers, all customers with the same colour represent the same group.
-Except for customer id 8269176, there is unanimity across design and groups.
-Even though their daily form resembles Group 2, our clustering approach places them in Group 3 (which is design 4 in our case).
-Because our method uses hod, moy, and wkndwday, there may be some mismatches depending on only one variable.
+data_heatmap_wkndwday <- quantile_gran(data_pick, "wknd_wday", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
+  left_join(data_pick_cust, by = c("customer_id"="value"))
 
-\noindent The distribution of electricity demand for the selected 24 customers across month-of-year is shown in Figure \ref{fig:data-heatmap-moy}.
-The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-Intriguingly, customer id 8269176 appears to be in the appropriate group (Group 3) because its monthly profile is more similar to Group 3 than Design 4.
-So, while our clustering approach failed to place the customer in the correct hour-of-day group, it did so for month-of-year.
-When contrasting the moy to hod profile, there are greater behavioural differences across customers within a group.
+data_heatmap_wkndwday$category <- factor(data_heatmap_wkndwday$category, levels = c("Weekday", "Weekend"))
+
+data_heatmap_wkndwday %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, 
+                  group=customer_id, 
+              fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 6) +
+ theme_characterisation() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))
+```
 
@@
              nrow = 4) + 
-              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
-  ylab("demand (in Kwh)") + 
-  theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank())+
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
-  scale_x_discrete(breaks = seq(1, 24, 3))+ 
-  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm"))+
-  theme(legend.position = "bottom")
-
+             theme_characterisation() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) 
 #
@@
 ```
-
-Characterization of clusters is the final stage of a cluster analysis.
-If we are convinced that we have identified a collection of clusters that can be distinguished from one another, we should intend to characterize them more formally, both statistically and qualitatively.
-We quantify them by producing statistics for each variable.
-We may look at the findings in graphs, and enhance our qualitative descriptions of the groupings.
-
-\noindent We discover 4 qualitative clusters of varying shapes in the distribution of all consumers in the first panel of Figure \ref{fig:combined}.
-Group-1 includes consumers who work 9-5, get up and conduct morning activities from 7-10am, and then depart.
-Then they return home in the evening to cook supper and perform other activities, giving the evening a greater peak than the morning.
-Group 2 is the group that rushes out of the house in the morning to get to work.
-They only return at night and do all activities at night, so there is no morning peak.
-The third one has a strong early morning and late night hours.
-These consumers may be flexible students or elderly retirees who are night owls.
-Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
-All of this may be validated with further information about the customer.
-
-\noindent The second panel of Figure \ref{fig:combined} shows that month-of-year qualitative clusters are not as distinguishable as hour-of-day.
-Group 2 is the most distinct and uses the most power during the summer, possibly owing to the use of air conditioners.
-Group 4 has a flat profile, indicating no significant month-to-month changes.
-Groups 1 and 3 have heaters on in the winter but consume less energy in the summer.
-Since gas is not available in all of NSW LGAs, it is possible that customers' heater usage is recorded in electricity rather than gas.
-
-\noindent The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
-
-\noindent The plotting scales are not displayed since we want to emphasise comparable shapes rather than scales.
-A customer in the cluster may have low daily or total energy usage, but their behaviour may be quite similar to a customer with high usage.
-That places them in the same group.
 
@@
 
-```{r data-heatmap-wkndwday}
+```{r data-heatmap-wkndwday-2}
 
@@
              ncol = 6) +
-    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("wknd_wday") + ylab("demand (in Kwh)")  + theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) +
-  theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") ) +
-  theme(legend.position = "bottom")
+   theme_characterisation() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))
 
@@
 
-# data_heatmap_wkndwday_group <- quantile_gran(data_group,
-#                                   gran1="wknd_wday",
-#                                   quantile_prob_val = c(0.25, 0.5, 0.75),
-#                                   group="group") %>%
-#   pivot_wider(names_from = quantiles, values_from = quantiles_values)
-# 
-# data_heatmap_wkndwday_group$category <- factor(data_heatmap_wkndwday_group$category, levels = c("Weekday", "Weekend"))
-# 
-# data_heatmap_wkndwday_group$group <- paste("group", data_heatmap_wkndwday_group$group, sep = "-")
-# 
-# 
-# wkndwday_group <- data_heatmap_wkndwday_group %>%
-#   ggplot(aes(x = category)) +
-#   geom_ribbon(aes(ymin = `25%`,
-#                   ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
-#   geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+data_heatmap_wkndwday_group <- quantile_gran(data_group,
+                                  gran1="wknd_wday",
+                                  quantile_prob_val = c(0.1, 0.25, 0.5, 0.75, 0.9),
+                                  group="group") %>%
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_wkndwday_group$category <- factor(data_heatmap_wkndwday_group$category, levels = c("Weekday", "Weekend"))
+
+data_heatmap_wkndwday_group$group <- paste("group", data_heatmap_wkndwday_group$group, sep = "-")
+
+wkndwday_group <- data_heatmap_wkndwday_group %>%
+  ggplot(aes(x = category)) +
+  geom_ribbon(aes(ymin = `25%`,
+                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+    geom_ribbon(aes(ymin = `10%`,
+                  ymax = `90%`, group=group, fill = as.factor(group)), alpha = 0.4) +
+  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+  facet_wrap(~group,
+             scales = "free_y",
+             labeller = "label_value",
+             nrow = 4)+
+    #theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
+    xlab("wknd-wday") +
+  ylab("demand (in Kwh)") + theme_characterisation() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))
+
+# wkndwday_group <- data_group %>% create_gran("wknd_wday") %>% 
+#   ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+#   #lvplot::geom_lv(aes(fill = as.factor(group))) +
+#   geom_boxplot(aes(fill = as.factor(group))) +
+#   #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
+# #geom_boxplot(aes(fill = as.factor(group))) +
+#   #scale_fill_lv() +
+#  xlab("wknd-wday") + 
+#   ylab("demand (in Kwh)") +
 #    facet_wrap(~group, 
@@
 #              nrow = 4)+
-#     #theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
-#     xlab("wknd-wday") +
-#   ylab("demand (in Kwh)") +
 #   theme_bw() +
@@
 #   theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-#    scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-#   scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+#   scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
 #   theme(legend.position = "bottom") +
@@
 
-wkndwday_group <- data_group %>% create_gran("wknd_wday") %>% 
-  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
-  lvplot::geom_lv(aes(fill = as.factor(group))) +
-  #geom_boxplot(aes(fill = as.factor(group))) +
-  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
-#geom_boxplot(aes(fill = as.factor(group))) +
-  #scale_fill_lv() +
- xlab("wknd-wday") + 
-  ylab("demand (in Kwh)") +
-   facet_wrap(~group, 
-             scales = "free_y", 
-             labeller = "label_value",
-             nrow = 4)+
-  theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + 
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
-  theme(legend.position = "bottom") +
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
-
 ```
 
-```{r combined, fig.cap="The distribution of electricity demand for the clusters across hour-of-day.
-The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-Each cluster is characterised by unique shape across the granularity it is plotted against.
-For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters.
-This fact we will be re-established when we see the importance of each granularity throough the parallel coordinate plot."}
+```{r combined}
 
@@
 
-```{r ggpairs, fig.cap = "The ggpairs plot of electricity demand against the three granularities hod, moy and wkndwday are shown.
-The correlation between any two granularities is not so high so as to pose a threat to the clustering methodology."}
+```{r ggpairs}
 
@@
 
-```{r, fig.cap=" A ggpairsplot and parallel coordinate plot are used to depict each of the 24 customers.
-The ggpairs plot four distinct clusters across the month-of-year, which are less prominent across the hour-of-day and wknd-wday.
-The parallel coordinate plot ranks the variables in order of importance, indicating that the month-of-year is the most important in identifying clusters, whereas wkdn-wday is the least significant and has the least variability among the three variables." }
+```{r }
 
@@
                    scale = "globalminmax"
-) + 
-  ggplot2::theme(
-    plot.title = ggplot2::element_text(size=10)
-  )+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
-  theme(legend.position = "bottom") +
-  xlab("") +
+) + theme_characterisation()+
   ylab("wpd") + scale_fill_viridis_d(direction = -1) +
@@
 
-```{r parcoord, fig.cap = "A parallel coordinate plot with the three significant cyclic granularities used for wpd-based clustering.
-The variables are sorted according to their separation across classes (rather than their overall variation between classes).
-This means that moy is the most important variable in distinguishing the designs followed by hod and wkndwday.
-It can be observed that cluster 3 and 4 are distinguished by moy while 1 and 2 are distinguished by hod.
-Here, wkndwday is still acting as the nuisance variable."}
+```{r parcoord"}
 parcoord
@@
   ylab("demand (in Kwh)") + 
-  theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank())+
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
-  scale_x_discrete(breaks = seq(1, 24, 3))+ 
-  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm"))+ scale_fill_viridis_d(direction = -1) +
-  scale_color_viridis_d(direction = -1) + theme_light() 
-  theme(legend.position = "bottom")
+  theme_light()+ 
+  theme_characterisation() +
+ scale_fill_viridis_d(direction = -1) +
+  scale_color_viridis_d(direction = -1)
 
@@
 
-```{r combined-wpd, fig.cap="The distribution of electricity demand for the clusters across hour-of-day.
-The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-Each cluster is characterised by unique shape across the granularity it is plotted against.
-For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters.
-This fact we will be re-established when we see the importance of each granularity throough the parallel coordinate plot."}
+```{r combined-wpd"}
 
@@
 ```
-
-\noindent We discover 4 qualitative clusters of varying shapes in the distribution of all consumers in the first panel of Figure \ref{fig:combined}.
-Group-1 includes consumers who work 9-5, get up and conduct morning activities from 7-10am, and then depart.
-Then they return home in the evening to cook supper and perform other activities, giving the evening a greater peak than the morning.
-Group 2 is the group that rushes out of the house in the morning to get to work.
-They only return at night and do all activities at night, so there is no morning peak.
-The third one has a strong early morning and late night hours.
-These consumers may be flexible students or elderly retirees who are night owls.
-Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
-All of this may be validated with further information about the customer.
-
-\noindent The second panel of Figure \ref{fig:combined} shows that month-of-year qualitative clusters are not as distinguishable as hour-of-day.
-Group 2 is the most distinct and uses the most power during the summer, possibly owing to the use of air conditioners.
-Group 4 has a flat profile, indicating no significant month-to-month changes.
-Groups 1 and 3 have heaters on in the winter but consume less energy in the summer.
-Since gas is not available in all of NSW LGAs, it is possible that customers' heater usage is recorded in electricity rather than gas.
-
-\noindent The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
-
-\noindent The plotting scales are not displayed since we want to emphasise comparable shapes rather than scales.
-A customer in the cluster may have low daily or total energy usage, but their behaviour may be quite similar to a customer with high usage.
-That places them in the same group.

### 2021-10-15 supplementary/distance-computation/real-data/04_handpick_from356-mds_nqt.Rmd
cleaning codes

@@
   layout(showlegend = FALSE) %>%
-  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect", persistent = TRUE,defaultValues = data_pick_cust$value, color = cols)
+  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect", persistent = TRUE) 
 ```

### 2021-10-17 main.Rmd
distance metrics section

@@
 
-The data set solely contains readings from smart meters and no information about the consumers' specific physical, geographical, or behavioural attributes.
+<!-- The data set solely contains readings from smart meters and no information about the consumers' specific physical, geographical, or behavioural attributes.
 As a result, no attempt is made to explain why consumption varies.
 Instead, this work investigates how much energy usage heterogeneity can be found in smart meter data and what some of the most common electricity use patterns are.
-It is worth noting that when studying these dynamics, a variety of objectives may be pursued.
+-->
+
+The proposed methodology aim to leverage the intrinsic temporal data structure hidden in time series data.
+The foundation of our method is unsupervised clustering algorithms based exclusively on time-series features.
+First, we study the underlying distributions that may have resulted in different patterns across temporal granularities in order to identify a mechanism to classify them based on the similarity of those distributions.
+It is worth noting that when studying these similarities, a variety of objectives may be pursued.
 One objective could be to group consumers with similar shapes over all relevant cyclic granularities.
@@
 
-The proposed methodology aim to leverage the intrinsic temporal data structure hidden in time series data.
-The foundation of our method is unsupervised clustering algorithms based exclusively on time-series features.
-First, we study the underlying distributions that may have resulted in different patterns across temporal granularities in order to identify a mechanism to classify them based on the similarity of those distributions.
 Depending on the goal of clustering, the distance metric for defining similarity would be different.
@@
 
-@wang2020tsibble introduced the tidy "tsibble" data structure to support exploration and modeling of temporal data.
-A tsibble comprises an index, optional key(s), and measured variables.
-An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
+@wang2020tsibble introduced the tidy "tsibble" data structure to support exploration and modeling of temporal data comprising of an index, optional key(s), and measured variables.
+<!-- An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
 A linear granularity is a mapping of the index set to subsets of the time domain.
-For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.
-For each key variable, the raw smart meter data is a sequence that is indexed by time and comprises values of several measurement variables at each time point.
+For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.--> For each key variable, the raw smart meter data is a sequence that is indexed by time and comprises values of several measurement variables at each time point.
 This sequence, though, could be depicted in a variety of ways.
@@
 So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
-In this work, instead of considering the probability distributions of the linear time series, we assume that the measured variables across different categories of any cyclic granularity are from data generating processes.
+In this work, instead of considering the probability distributions of the linear time series, we assume that the measured variables across different categories of any cyclic granularity are from different data generating processes.
 Hence, we want to be able to cluster distributions of the form $f_{i,A,B \dots, {N_C}}(v)$, where $A, B$ represent the cyclic granularities under consideration such that $A = \{a_j: j=1, 2, \dots J\}$,  $B = \{b_k: k  = 1, 2, \dots K\}$ and so on.
-We consider individual each category of a cyclic granularity ($A$) or combination of categories for interaction of cyclic granularities (for e.g.
+We consider individual category of a cyclic granularity ($A$) or combination of categories for interaction of cyclic granularities (for e.g.
 $A*B$) to have a distribution.
 For example, let us consider we have two cyclic  granularities of interest, $A = {0, 1, 2, \dots, 23}$ representing hour-of-day and  $B = \{Mon, Tue, Wed, \dots, Sun\}$ representing day-of-week.
-each customer $i$ consist of a collection of probability distributions.
-In case individual granularities ($A$ or $B$) are considered there are  $J = 24$ distributions of the form $f_{i,j}(v)$ or $K = 7$ distributions of the form $f_{i,k}(v)$ for each customer $i$.
+Each customer $i$ consist of a collection of probability distributions.
+In case individual granularities ($A$ or $B$) are considered, there are  $J = 24$ distributions of the form $f_{i,j}(v)$ or $K = 7$ distributions of the form $f_{i,k}(v)$ for each customer $i$.
 In case of interaction,  $J*K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
-As a result, a distance between collections of these univariate probability distributions is required.
+
+\noindent As a result, a distance between collections of these univariate probability distributions is required.
 Depending on the objective of the problem, there could be many approaches to considering such distances.
@@
 Considering each individual or combined categories of cyclic granularities as a data generating process lead to a collection of conditional distributions for each customer $i$.
-The (dis) similarity between each pair of observations should be obtained by combining the distances between these collections of conditional distributions such that the resulting metric is a distance metric, which could be fed into the clustering algorithm.
+The (dis) similarity between each pair of customers should be obtained by combining the distances between these collections of conditional distributions such that the resulting metric is a distance metric, which could be fed into the clustering algorithm.
 Two types of distance metric is considered:
@@
 
-**Inter-category distances**
+**JS-based distances**
 
@@
 \noindent The total distance between two elements $x$ and $y$ is then defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity A) or  $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
-When combining distances from individual cyclic granularities $A$ and $B$, $$S^{A, B}_{x, y} = S^A_{x,y}/J + S^B_{x,y}/K$$ is used, which could also be shown to be a distance metric easily.
-This is shown for cyclic granularity $A$ and $B$, but could be practically extended to more granularities.
-
-**Intra-category distances**
-
-Compute weighted pairwise distances ($wpd$) (XXX reference) for all considered granularities for all objects.
-$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive between consecutive categories normalized by appropriate parameters.
+When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels, $$S_{x, y} = \sum_lS^{C_l}_{x,y}/n_l$$ is used, which is also a distance metric being the sum of JS distances.
+
+**wpd-based distances**
+
+Compute weighted pairwise distances ($wpd$) for all considered granularities for all objects.
+$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalized by appropriate parameters.
 A higher value of $wpd$ indicates that some interesting pattern is expected, whereas a lower value would indicate otherwise.
 
-\noindent Distance between objects is then taken as the euclidean distances between them with the granularities being the variables and $wpd$ being the value under each variable.
+\noindent Distance between elements is then taken as the euclidean distances between them with the granularities being the variables and $wpd$ being the value under each variable.
 Since Euclidean distance is chosen, the observations with high values of features ($wpd$ values) will be clustered together.
@@
 The following are some of the advantages of our proposed method.
--->
+--
 

### 2021-10-18 check_work.Rmd
adding new simulation files

@@
 
-```{r external, include = FALSE, cache = FALSE}
-knitr::read_chunk(here('script/smart_meter.R'))
-```
-
 ```{r mytheme}
-theme_validation <- function() {
-  theme_bw() +
-      theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 1)),
-    plot.margin = margin(0, 0, 0, 0, "cm"),
-    axis.title.y = element_blank(),
-    #axis.text.x = element_blank(),
-  axis.text.y = element_blank(),
-  axis.ticks = element_blank(),
-      legend.position = "bottom"
-    )
-}
-
 theme_characterisation <- function() {
@@
   theme_bw() + # seeting theme
-    theme(strip.text = element_text(size = 7,
+    theme(strip.text = element_text(size = 9,
                                     margin = margin(b = 0, t = 0))) + # narrow facet space
@@
   theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
-    theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + # rotate the x-axis text
+    theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
     theme(legend.position = "bottom")+
@@
 anchor3 <- 8629763
-anchor4 <-
-
+anchor4 <- 11013154
+kopt <- 4 # number of clusters
 ```
 
-```{r data-pick}
+_Where they lie on the mds plot of total, hod, moy, wkndwday and wpd_
+
+```{r data-read}
 
@@
 hod <- read_rds(here("data/dist_gran_hod_356cust_robust.rds"))
+
 distance <- wkndwday %>% 
@@
          item2 = as.integer(as.character(item2))) 
+```
 
+```{r data-pick}
 total_distance <- distance %>% 
-  mutate(total = hod)
-  #mutate(total = wkndwday/2 + moy/12 + hod/24)
+  #mutate(total = hod)
+  mutate(total = wkndwday/2 + moy/12 + hod/24)
 
 # 8454221 10420689 1001 8454221 10018298 10420689 9393696
-data_pick_one <- total_distance %>% filter(item1 %in% 
-                            c(11156027)|item2 %in% 
-                            c(11156027) ) %>% 
+
+# k-nearest neighbor pick function
+nn_pick_function <- function(anchor, k = 5){
+  data <- total_distance %>% filter(item1 %in% 
+                            c(anchor)|item2 %in% 
+                            c(anchor) ) %>% 
   #group_by(item1) %>%
   arrange(total) %>% 
-  head(5) %>% 
+  head(k) %>% 
   mutate(item1 = as.integer(as.character(item1)),
@@
 
-#11013154  8181071 11048034
-data_pick_two <- total_distance %>% filter(item1 %in% 
-                            c(10420683)|item2 %in% 
-                            c(10420683)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
+  unique(c(data$item1,data$item2)) %>% as_tibble()
+}
 
-data_pick_three <- total_distance %>% filter(item1 %in% 
-                              c(11013154)) %>%
-  #group_by(item1) %>%
-  arrange(total) %>% head(5)%>% 
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
+data_pick_one <- nn_pick_function(anchor1)
+data_pick_two <- nn_pick_function(anchor2)
+data_pick_three <- nn_pick_function(anchor3)
+data_pick_four <- nn_pick_function(anchor4)
 
-# 8653709
-data_pick_four <- total_distance %>% filter(item1 %in%
-                            c(8629763)|item2 %in% 
-                            c(8629763)) %>% 
-  #group_by(item1) %>% 
-  arrange(total) %>% head(5)%>%
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
-
-# c(11013154, 8495194, 8627007,10109182,10677705,8952846)
+# c(11013154, 9345642, 8328008, 8636035,8627007,8454235)
 
 data_pick_cust <- bind_rows(
-unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(),    unique(c(data_pick_two$item1,data_pick_two$item2)) %>% as_tibble(), 
-c(11013154, 9345642, 8328008, 8636035,8627007,8454235) %>% as_tibble(),
-unique(c(data_pick_four$item1,data_pick_four$item2)) %>% as_tibble(),
+data_pick_one, data_pick_two, data_pick_three, data_pick_four,
 .id = "design")
@@
   filter(customer_id %in% data_pick_cust$value) %>% 
-  gracsr::scale_gran( method = "robust", response = "general_supply_kwh")
+  gracsr::scale_gran(method = "robust", response = "general_supply_kwh")
+```
+
+```{r mds-plots}
+
+data_pick_char <- data_pick_cust %>% 
+  rename("customer_id" = "value") %>% 
+  mutate(customer_id = as.character(customer_id))
+
+mds_all <- read_rds(here("data/mds_all.rds"))  %>% 
+  left_join(data_pick_char)
+
+highlight_df <- mds_all %>% filter(customer_id %in% data_pick_char$customer_id)
+
+total_plot <- ggplot(mds_all,
+aes(x = mds1.total, y = mds2.total)) +
+  geom_point(alpha = 0.3) +
+  geom_point(data = highlight_df, 
+             aes(x = mds1.total, y = mds2.total, color = as.factor(design)), size = 2)
+
+wpd_plot <- ggplot(mds_all,
+aes(x = mds1.wpd, y = mds2.wpd)) +
+  geom_point(alpha = 0.3) +
+  geom_point(data = highlight_df, 
+             aes(x = mds1.wpd, y = mds2.wpd, color = as.factor(design)), size = 2)
+
+hod_plot <- ggplot(mds_all,
+aes(x = mds1.hod, y = mds2.hod)) +
+  geom_point(alpha = 0.3) +
+  geom_point(data = highlight_df, 
+             aes(x = mds1.hod, y = mds2.hod, color = as.factor(design)), size = 2)
+
+moy_plot <- ggplot(mds_all,
+aes(x = mds1.moy, y = mds2.moy)) +
+  geom_point(alpha = 0.3) +
+  geom_point(data = highlight_df, 
+             aes(x = mds1.moy, y = mds2.moy, color = as.factor(design)), size = 2)
+
+wkndwday_plot <- ggplot(mds_all,
+aes(x = mds1.wkndwday, y = mds2.wkndwday)) +
+  geom_point(alpha = 0.3) +
+  geom_point(data = highlight_df, 
+             aes(x =  mds1.wkndwday, 
+                 y =  mds2.wkndwday, 
+                 color = as.factor(design)),
+             size = 2)
+
+p <- (total_plot + wpd_plot)/(hod_plot + moy_plot + wkndwday_plot)
+
+p* +
+  plot_layout(guides = "collect")
 
@@
 
-group <- f %>% hclust (method = "ward.D") %>% cutree(k=4)
+group <- f %>% hclust (method = "ward.D") %>% cutree(k=kopt)
 
@@
 colnames(mds) = c("mds1", "mds2")
+```
 
-# ggplot(data = mds %>% as_tibble() %>% mutate(group = cluster_result$group), 
-#        aes(x = mds1, y = mds2, color = as.factor(group))) +
-#   geom_point() +
-#   geom_text(aes(label = id))
-
-# cluster_result %>% filter(group==1)
-# cluster_result %>% filter(group==2)
-# cluster_result %>% filter(group==3)
-# cluster_result %>% filter(group==4)
-
-  #x = ~mds1.hod, y = ~ mds2.hod, text = ~customer_id
-```
 ```{r heatmap-hod}
@@
 data_heatmap_hod <- quantile_gran(data_pick,
-                                  "hour_day", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
+                                  "hour_day",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
+  pivot_wider(names_from = quantiles,
+              values_from = quantiles_values) %>% 
   left_join(data_pick_cust, by = c("customer_id"="value"))
@@
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
+  scale_x_discrete(breaks = seq(0, 23, 3))
+# + scale_y_continuous(breaks = NULL) 
 ```
@@
   ylab("demand (in Kwh)") +
-  theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + 
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-  theme(legend.position = "bottom") +
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
+  theme_characterisation() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))
 ```
@@
 
-data_heatmap_wkndwday_group <- quantile_gran(data_group,
-                                  gran1="wknd_wday",
-                                  quantile_prob_val = c(0.1, 0.25, 0.5, 0.75, 0.9),
-                                  group="group") %>%
-  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+wkndwday_data<- data_group %>% create_gran("wknd_wday") 
+ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]
 
-data_heatmap_wkndwday_group$category <- factor(data_heatmap_wkndwday_group$category, levels = c("Weekday", "Weekend"))
-
-data_heatmap_wkndwday_group$group <- paste("group", data_heatmap_wkndwday_group$group, sep = "-")
-
-wkndwday_group <- data_heatmap_wkndwday_group %>%
-  ggplot(aes(x = category)) +
-  geom_ribbon(aes(ymin = `25%`,
-                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
-    geom_ribbon(aes(ymin = `10%`,
-                  ymax = `90%`, group=group, fill = as.factor(group)), alpha = 0.4) +
-  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+wkndwday_group <- wkndwday_data%>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  lvplot::geom_lv(aes(fill = as.factor(group), color = as.factor(group)), k=5, alpha = 0.5) +
+  #geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
+  #geom_boxplot(outlier.shape = NA) + 
+  coord_cartesian(ylim = ylim1*1.05)+
   facet_wrap(~group, 
@@
              labeller = "label_value",
-             nrow = 4)+
-    #theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
-    xlab("wknd-wday") +
-  ylab("demand (in Kwh)") + theme_characterisation() +
+             nrow = 4)  +
+  theme_characterisation() +
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
   scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))
-
-# wkndwday_group <- data_group %>% create_gran("wknd_wday") %>% 
-#   ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
-#   #lvplot::geom_lv(aes(fill = as.factor(group))) +
-#   geom_boxplot(aes(fill = as.factor(group))) +
-#   #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
-# #geom_boxplot(aes(fill = as.factor(group))) +
-#   #scale_fill_lv() +
-#  xlab("wknd-wday") + 
-#   ylab("demand (in Kwh)") +
-#    facet_wrap(~group, 
-#              #scales = "free_y", 
-#              labeller = "label_value",
-#              nrow = 4)+
-#   theme_bw() +
-#    theme(axis.title.y=element_blank(),
-#         axis.text.y=element_blank(),
-#         axis.ticks.y=element_blank()) + 
-#   theme(panel.spacing =unit(0, "lines")) + 
-#   theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-#   scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
-#   theme(legend.position = "bottom") +
-#   theme(panel.spacing =unit(0, "lines")) + 
-#   theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
 
@@
 
-group <- f%>% hclust (method = "ward.D") %>% cutree(k=4)
+group <- f%>% hclust (method = "ward.D") %>% cutree(k=kopt)
 
@@
 
-mds  = cmdscale(f)
-rownames(mds) = id
-colnames(mds) = c("mds1", "mds2")
+# mds plot
+
+# mds  = cmdscale(f)
+# rownames(mds) = id
+# colnames(mds) = c("mds1", "mds2")
 
@@
 
+# will give error if you have got just one observation in a cluster
+
+# ggplot(data = data_pcp,
+#        aes(x = month_year, y = wkndwday)) +
+#   geom_point() +
+#   geom_text(aes(label = id))
+
   # 
@@
 
-```{r parcoord"}
+```{r parcoord}
 parcoord
@@
   ylab("demand (in Kwh)") + 
-  theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + 
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
+  theme_light()+ 
+  theme_characterisation() +
  scale_fill_viridis_d(direction = -1) +
-  scale_color_viridis_d(direction = -1) + theme_light() + 
-  theme(legend.position = "bottom") +
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
+  scale_color_viridis_d(direction = -1) 
 ```
@@
   ylab("demand (in Kwh)") + 
-  theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + 
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(plot.margin = margin(0, 0, 0, 0, "cm"))+ scale_fill_viridis_d(direction = -1) +
-  scale_color_viridis_d(direction = -1) + theme_light() + 
-  theme(legend.position = "bottom") +
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
+  theme_light()+ 
+  theme_characterisation() +
+ scale_fill_viridis_d(direction = -1) +
+  scale_color_viridis_d(direction = -1)
+
+data_group <- data_pick %>%
+  mutate(customer_id = as.character(customer_id)) %>%
+  left_join(cluster_result_wpd, by = c("customer_id" = "id"))
+
+wkndwday_data<- data_group %>% create_gran("wknd_wday") 
+ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]
+
+wkndwday_group_wpd <- wkndwday_data%>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  lvplot::geom_lv(aes(fill = as.factor(group), color = as.factor(group)), k=5, alpha = 0.5) +
+  #geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
+  #geom_boxplot(outlier.shape = NA) + 
+  coord_cartesian(ylim = ylim1*1.05)+
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow = 4)   + 
+  theme_light()+ 
+  theme_characterisation() +
+ scale_fill_viridis_d(direction = -1) +
+  scale_color_viridis_d(direction = -1) 
 ```
 
-```{r combined-wpd"}
+```{r combined-wpd}
 
@@
 ```
+```{r extra}
+# total
+# data = mds_total %>% left_join(data_pick_char)
+# 
+# highlight_df <- data %>% filter(customer_id%in% data_pick_char$customer_id)
+#   
+# total_plot <- ggplot(data,
+# aes(x = mds1.total, y = mds2.total)) +
+#   geom_point(alpha = 0.3) +
+#   geom_point(data = highlight_df, 
+#              aes(x = mds1.total, y = mds2.total, color = as.factor(design)), size = 2)
+# 
+# #hod
+# 
+# data = mds_hod %>% left_join(data_pick_char)
+# 
+# highlight_df <- data %>% filter(customer_id%in% data_pick_char$customer_id)
+#   
+# hod_plot <- ggplot(data,
+# aes(x = mds1.hod, y = mds2.hod)) +
+#   geom_point(alpha = 0.3) +
+#   geom_point(data = highlight_df, 
+#              aes(x = mds1.hod, y = mds2.hod, color = as.factor(design)), size = 2)+theme(legend.position = "bottom")
+# 
+# # moy
+# 
+# data = mds_moy %>% left_join(data_pick_char)
+# 
+# highlight_df <- data %>% filter(customer_id%in% data_pick_char$customer_id)
+#   
+# moy_plot <- ggplot(data,
+# aes(x = mds1.moy, y = mds2.moy)) +
+#   geom_point(alpha = 0.3) +
+#   geom_point(data = highlight_df, 
+#              aes(x = mds1.moy, y = mds2.moy, color = as.factor(design)), size = 2)+theme(legend.position = "bottom")
+# 
+# # wkndwday
+# 
+# data = mds_wkndwday %>% left_join(data_pick_char)
+# 
+# highlight_df <- data %>% filter(customer_id %in% data_pick_char$customer_id)
+#   
+# wkndwday_plot <- ggplot(data,
+# aes(x = mds1.wkndwday, y = mds2.wkndwday)) +
+#   geom_point(alpha = 0.3) +
+#   geom_point(data = highlight_df, 
+#              aes(x =  mds1.wkndwday, 
+#                  y =  mds2.wkndwday, 
+#                  color = as.factor(design)),
+#              size = 2)+ theme(legend.position = "bottom")
+
+```

### 2021-10-18 main.Rmd
adding new simulation files

@@
 The following are some of the advantages of our proposed method.
---
+-->
 
@@
 
-    **Handling trend, seasonality, non-stationarity and auto-correlation:** Trend and seasonality are fundamental characteristics of time series data, and it is reasonable to define a time series according to its degree of trend and seasonality.
+<!--Handling trend, seasonality, non-stationarity and auto-correlation:** Trend and seasonality are fundamental characteristics of time series data, and it is reasonable to define a time series according to its degree of trend and seasonality.
 These characteristics of the time series are lost or handled independently by considering probability distributions ( trend is lost) across categories of cyclic granularities (by independently modeling all seasonal fluctuations), and so there is no need to de-trend or de-seasonalize the data before conducting the clustering method.
-There is no need to omit holiday or weekend patterns for similar reasons.
-
-    **Data transformation:** Robust scaling method is used before computing the Inter-category distances and NQT is built-in transformation used for computation of $wpd$, which forms the basis of Intra-category distances.
+There is no need to omit holiday or weekend patterns for similar reasons-->
+
+    **Data transformation:** Both robust scaling method and NQT are used before computing the JS-based distances and NQT is built-in transformation used for computation of $wpd$, which forms the basis of wpd-based distances.
+
+    Robust scaling would keep the distribution intact
+
+Normal-Quantile transform will transform the positively and negatively skewed distribution to a similar bell-shaped.
+From the transformed distribution, it is difficult to understand that raw distribution was of which shape.
+Also, multimodality gets hidden or magnitude get reversed with NQT.
+But deciles from the distribution will move in a similar manner as the raw distribution and hence the final distance matrix seem to be unaffected.
+Hence, this could be used.
+Also, practically most problems will have a very skewed distribution, it is often helpful to bring them to a normal-like shape before clustering.
 

### 2021-10-18 supplementary/distance-computation/real-data/04handpick-from356-mds.Rmd
adding new simulation files

@@
   select(customer_id, everything())
+
+write_rds(mds_all, here("data/mds_all.rds"))
 ```

### 2021-10-18 2gran_orig_array1_27.sh

@@ -1,0 +1,13 @@
+#! /bin/bash
+#SBATCH --job-name=gran2_orig
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=32G
+#SBATCH --time=80:00:00
+#SBATCH --array=1-27
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.5
+Rscript gran2_orig.R $SLURM_ARRAY_TASK_ID 
+

### 2021-10-18 3gran_orig_array1_27.sh

@@ -2,11 +2,12 @@
 #SBATCH --job-name=gran3_orig
 #SBATCH --output=trace_%a.out
 #SBATCH --error=error_%a.err
-#SBATCH --mem-per-cpu=16G
+#SBATCH --mem-per-cpu=32G
 #SBATCH --time=80:00:00
+#SBATCH --array=1-27
 #SBATCH --ntasks=1
 #SBATCH --cpus-per-task=1
 #SBATCH --export=NONE
 module load R/4.0.5
-Rscript gran3_orig.R
+Rscript gran3_orig.R $SLURM_ARRAY_TASK_ID 
 

### 2021-10-18 quantiles_raw_data.R

@@ -47,6 +47,65 @@ data_356cust_wkndwday <- quantile_gran(data_pick_robust,
                                   quantile_prob_val = seq(0.1, 0.9, 0.1))
 
 
+data_356cust_hod_wide <- data_356cust_hod %>%
+  pivot_wider(names_from = c("gran", "category", "quantiles"),
+              values_from = "quantiles_values")
+
+data_356cust_moy_wide <- data_356cust_moy %>%
+  pivot_wider(names_from = c("gran", "category", "quantiles"),
+              values_from = "quantiles_values")
+
+data_356cust_wkndwday_wide <- data_356cust_wkndwday %>%
+  pivot_wider(names_from = c("gran", "category", "quantiles"),
+              values_from = "quantiles_values")
+
+data_356cust_wide <- left_join(data_356cust_hod_wide,
+                               data_356cust_moy_wide,
+                               data_356cust_wkndwday_wide,
+                               by="customer_id")
+
+save(data_356cust_wide, file="data/data_356cust_wide.rda")
+
+
+data_356cust_pc <- prcomp(data_356cust_wide,
+                          center = FALSE, scale = FALSE, retx = TRUE)
+plot(data_356cust_pc, type="l", npcs=50)
+
+
+library(liminal)
+data_356cust_pc10 <- as_tibble(data_356cust_pc$x[,1:10])
+limn_tour(data_356cust_pc10, PC1:PC10)
+
+tSNE_fit<-data_356cust_wide%>% 
+  select(-customer_id) %>% 
+  Rtsne()
+
+
+tSNE_fit$Y %>% 
+  as.data.frame() %>% 
+  rename(tSNE1="V1",
+         tSNE2="V2") %>% 
+  mutate(customer_id=as.character(data_356cust_wide$customer_id)) -> tSNE.plot
+
+library(ggplot2)
+ggplot()+
+  geom_point(data=tSNE.plot,
+             aes(x=tSNE1,y=tSNE2,color=customer_id))
+
+
+
++
+  stat_ellipse(data=tSNE.plot,
+               geom="polygon",
+               aes(x=tSNE1,y=tSNE2,
+                   group=customer_id,
+                   fill=customer_id),
+               alpha=0.5,
+               lty="dashed",
+               color="black",
+               key_glyph="blank")+
+  theme_bw()
+
 write_rds(data_356cust_hod, here::here("data/quantile_data_356cust_hod_robust.rds"))
[…]

### 2021-10-18 main.Rmd
added pre-processing steps

@@
 
-    **Data transformation:** Both robust scaling method and NQT are used before computing the JS-based distances and NQT is built-in transformation used for computation of $wpd$, which forms the basis of wpd-based distances.
-
-    Robust scaling would keep the distribution intact
-
-Normal-Quantile transform will transform the positively and negatively skewed distribution to a similar bell-shaped.
+Practically most problems will have a very skewed distribution, it is often helpful to bring them to a normal-like shape before clustering.
+Two data transformation techniques are employed for the JS-based methods and NQT is built-in transformation used for computation of $wpd$, which forms the basis of wpd-based distances.
+
+_Robust scaling_ Standardizing is a common scaling method that subtracts the mean from values and divides by the standard deviation, resulting in a conventional Gaussian probability distribution for an input variable (zero mean and unit variance).
+If the input variable includes outlier values, standardisation may become skewed or prejudiced.
+To address this, robust scaling methods could be utilized  (value – median) / (p75 – p25)) which results in a variable with a zero mean and median, as well as a standard deviation of one, while the outliers are still there with the same relative connections to other values.
+
+_Normal-Quantile transform_  First as a data pre-processing step to make all assymetrical real world variables more symmetric, we perform a quantile-normal transform on the data.
+This makes sure that the CDF of the resulting variable in Gaussian.
+The original data is ranked in ascending order and the probabilities $P(Y<=y(i)) = i/(n+1)$ are attached to $y(i)$, in terms of their ranking order.
+A NQT based transformation is applied by computing from a standard normal distribution a variable $\eta(i)$, which corresponds to the same probability $P(\eta< \eta(i)) = i/n+1$.
+By doing this, the new variables $\eta(i)$ will be marginally distributed according to standard Normal, N(0,1).
+NQT will transform the positively and negatively skewed distribution to a similar bell-shaped.
 From the transformed distribution, it is difficult to understand that raw distribution was of which shape.
@@
 Hence, this could be used.
-Also, practically most problems will have a very skewed distribution, it is often helpful to bring them to a normal-like shape before clustering.
 

### 2021-10-19 main.Rmd
adding validation section

@@
 
-tab_dist <- tibble::tibble(granularity = granularity, `alternate distribution` = alternate_design)
+tab_dist <- tibble::tibble(granularity = granularity, `Varying distributions` = alternate_design)
 
@@
 ```{r tab-design}
-design = c("design-1", "design-2", "design-3", "design-4", "design5")
-g1 = c("null", "alternate", "null", "null", "alternate")
-g2 = c("null", "null", "alternate", "null", "alternate")
-g3 = c("null", "null", "null", "alternate", "alternate")
+design = c("design-1", "design-2", "design-3", "design-4", "design-5")
+g1 = c("fixed", "vary", "fixed", "fixed", "vary")
+g2 = c("fixed", "fixed", "vary", "fixed", "vary")
+g3 = c("fixed", "fixed", "fixed", "vary", "vary")
 table <- tibble(design, g1, g2, g3)
 
-# %>% kable(caption = "5 different designs resulting from considering different distributions across categories.")
+#%>% kable(caption = "5 different designs resulting from considering different distributions across categories.")
 ```
@@
 In this scenario, all of $g1$, $g2$ and $g3$ changes across at least one design.
-Also, it is not possible to comprehend these patterns across cyclic granularities or group similar series just by looking at the linear plots."}
+Also, it is not possible to comprehend these patterns across cyclic granularities or group similar series just by looking at the linear plots.", fig.pos="htb"}
 
@@
     geom = 'line',
-    aes(group = 1), size = 0.8, colosr = "blue")+ ylab("")+
+    aes(group = 1), size = 0.8, color = "blue")+ ylab("")+
   theme_validation()
@@
     geom = 'line',
-    aes(group = 1), size = 0.8, colosr = "blue")+ ylab("")+
+    aes(group = 1), size = 0.8, color = "blue")+ ylab("")+
   theme_validation()
 
-(p1 + (p2 + p3 + p4)) *
-  theme_validation() + plot_layout(widths = c(2, 1))
-
+gran2_change <- (p2 + p3 + p4) *
+  theme_validation()
 ```
@@
 
-(p1 + (p2 + p3 + p4)) *
-  theme_validation() + plot_layout(widths = c(2, 1))
+gran1_change <- (p2 + p3 + p4) *
+  theme_validation()
 # 
@@
 The clustering methodologies were run all these unique combinations and subsets of these to verify if the methodologies work as expected.
+
+```{r gran2and1-clubbed, fig.cap=" For the left scenario $g1$, $g2$ would change across atleast one design but $g3$ change remains same across all design.
+For tthe right one, only $g3$ changes across different designs."}
+#gran2_change  + gran1_change + plot_layout(widths = c(1,1))
+ggpubr::ggarrange(gran1_change, gran2_change, labels = c("a", "b"))
+```
 
@@
 
-((p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall))
+#((p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)) 
+#p_null+ p_varf + p_varx + p_varall
+ggpubr::ggarrange(p_null, p_varf, p_varx, p_varall, labels = c("Design-1", "Design-2", "Design-3", "Design-4"))
 
@@
 
-```{r}
+```{r, eval=FALSE}
 # code in append_3gran_change.R
@@
 
-wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_robust.rds")) %>% broom::tidy()
-
-moy <- read_rds(here("data/dist_gran_moy_356cust.rds")) %>% broom::tidy()
-
-hod <- read_rds(here("data/dist_gran_hod_356cust_robust.rds")) %>% broom::tidy()
+wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_robust.rds")) 
+#%>% broom::tidy()
+
+moy <- read_rds(here("data/dist_gran_moy_356cust.rds")) 
+#%>% broom::tidy()
+
+hod <- read_rds(here("data/dist_gran_hod_356cust_robust.rds")) 
+#%>% broom::tidy()
 

### 2021-10-19 quantiles_raw.R

@@ -47,9 +47,14 @@ data_pick_robust <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"
 #                                   quantile_prob_val = seq(0.1, 0.9, 0.1))
 # 
 
-data_356cust_hod <- read_rds("data/quantile_data_356cust_hod_robust.rds")
-data_356cust_moy <- read_rds("data/quantile_data_356cust_moy_robust.rds")
-data_356cust_wkndwday <- read_rds("data/quantile_data_356cust_wkndwday_robust.rds")
+data_356cust_hod <- read_rds("data/quantile_data_356cust_hod_robust.rds") %>% 
+  filter(quantiles %in% "50%")
+
+data_356cust_moy <- read_rds("data/quantile_data_356cust_moy_robust.rds") %>% 
+  filter(quantiles %in% "50%")
+
+data_356cust_wkndwday <- read_rds("data/quantile_data_356cust_wkndwday_robust.rds") %>% 
+  filter(quantiles %in% "50%")
 
 
 data_356cust_hod_wide <- data_356cust_hod %>%
@@ -65,13 +70,14 @@ data_356cust_wkndwday_wide <- data_356cust_wkndwday %>%
               values_from = "quantiles_values")
 
 data_356cust_wide <- left_join(data_356cust_hod_wide,
-                               data_356cust_moy_wide,
-                               data_356cust_wkndwday_wide,
-                               by="customer_id")
+                               data_356cust_moy_wide, by="customer_id") %>% 
+  left_join(data_356cust_wkndwday_wide,  by="customer_id"
+                              )
 
 save(data_356cust_wide, file="data/data_356cust_wide.rda")
 
 # without scaling
+
 data_356cust_pc <- prcomp(data_356cust_wide[,-1],
                           center = FALSE, scale = FALSE, retx = TRUE)
 
@@ -79,74 +85,26 @@ plot(data_356cust_pc, type="l", npcs=50)
 
 library(liminal)
 data_356cust_pc10 <- as_tibble(data_356cust_pc$x[,1:10])
-limn_tour(data_356cust_pc10, PC1:PC10)
+limn_tour(data_356cust_pc10, PC1:PC)
 
 sort(abs(data_356cust_pc$rotation[,1]))
 
[…]

### 2021-10-19 main.Rmd
simulation till 20oct

@@
 To validate the clustering approaches, we create data designs that replicate prototype behaviors that might be seen in electricity data contexts.
-We spiked several features in the data to see where one method works better than the other and where they might give us the same outcome or the effect of missing data <!---and trends--> on the proposed methods.
+We spiked several attributes in the data to see where one method works better than the other and where they might give us the same outcome or the effect of missing data <!---and trends--> on the proposed methods.
 A continuous measured variable $y$ of length $T$ indexed by $\{0, 1, \dots T-1\}$.
@@
 A sample size of $T=\{300, 1000, 5000\}$ is used in all designs to test small, medium and large sized series.
-The method could preform differently with different jumps between consecutive categories.
+The methods could preform differently with different jumps between consecutive categories.
 So a mean difference of $diff = \{1, 2, 5\}$ for corresponding categories are also considered.
@@
 
-_All signifiant granularities_
+  _(a) All signifiant granularities_
 
@@
 
-Consider the scenario when all three granularities $g1$, $g2$, and $g3$ are responsible for distinguishing the designs.
+\noindent Consider the scenario when all three granularities $g1$, $g2$, and $g3$ are responsible for distinguishing the designs.
 That means that for at least one among the to-be-grouped designs, the pattern for each of $g1$, $g2$, and $g3$ will change.
@@
 
-```{r tab-distribution}
+_(b) Few significant granularities_
+
+\noindent This is the case where one granularity will remain the same across all designs.
+We consider the case where the distribution of $y$ would vary across levels of $g2$ for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
+So $g3$ is not responsible for distinguishing across designs.
+Figure \ref{#fig:plot-linear} shows the linear and cyclic representation of $y$.
+The first panel shows raw plot of $y$ in a linear scale and the second panel shows distribution of $y$ across cyclic granularities namely $g1$, $g2$ and $g3$.
+As could be seen from the plots, it is impossible to decipher from the raw time plot that the time series variable shows such pattern across different granularities.
+
+_(c) One signifiant granularity_
+
+\noindent Here only one granularity is responsible for distinguishing the designs. Designs change significantly only for the granularity $g2$. Figure \ref{fig:plot-1gran} shows this.
+
+```{r tab-distribution, fig.pos="htb"}
 granularity = c("g1", "g2", "g3")
@@
 
-_Few significant granularities_
-
-This is the case where one granularity will remain the same across all designs.
-We consider the case where the distribution of $y$ would vary across levels of $g2$ for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
-So $g3$ is not responsible for distinguishing across designs.
-Figure \ref{#fig:plot-linear} shows the linear and cyclic representation of $y$.
-The first panel shows raw plot of $y$ in a linear scale and the second panel shows distribution of $y$ across cyclic granularities namely $g1$, $g2$ and $g3$.
-As could be seen from the plots, it is impossible to decipher from the raw time plot that the time series variable shows such pattern across different granularities.
-
-_One signifiant granularity_
-
-Here only one granularity is responsible for distinguishing the designs. Designs change significantly only for the granularity $g2$. Figure \ref{fig:plot-1gran} shows this.
-
 ```{r}
@@
 
-A subset of many possible designs are shown in Figure \ref{fig: plot-linear }.
+<!-- A subset of many possible designs are shown in Figure \ref{fig: plot-linear }.
 For the parameter space (XXX unique  combinations shown in table YYY), 100, 500 independent replications of all possible combination of simulation parameters were generated.
 The clustering methodologies were run all these unique combinations and subsets of these to verify if the methodologies work as expected.
+-->
 
@@
 p_null <- sim_panel_null %>%
-  ggplot(aes(x = as.factor(g2), y = sim_data)) +
-  facet_wrap(~g1, labeller = "label_both") +
+  ggplot(aes(x = as.factor(g1), y = sim_data)) +
+  facet_wrap(~g2, labeller = "label_both") +
   geom_boxplot() +
@@
     strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") ) +ggtitle("Design-1")
 
 p_varf <- sim_panel_varf %>%
-  ggplot(aes(x = as.factor(g2), y = sim_data)) +
-  facet_wrap(~g1, labeller = "label_both") +
+  ggplot(aes(x = as.factor(g1), y = sim_data)) +
+  facet_wrap(~g2, labeller = "label_both") +
   geom_boxplot() +
@@
     strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )+ggtitle("Design-2")
 
 p_varx <- sim_panel_varx %>%
-  ggplot(aes(x = as.factor(g2), y = sim_data)) +
-  facet_wrap(~g1, labeller = "label_both") +
+  ggplot(aes(x = as.factor(g1), y = sim_data)) +
+  facet_wrap(~g2, labeller = "label_both") +
   geom_boxplot()+
@@
     strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )+ggtitle("Design-3")
 
 p_varall <- sim_panel_varall %>%
-  ggplot(aes(x = as.factor(g2), y = sim_data)) +
-  facet_wrap(~g1, labeller = "label_both") +
+  ggplot(aes(x = as.factor(g1), y = sim_data)) +
+  facet_wrap(~g2, labeller = "label_both") +
   geom_boxplot() +
@@
     strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") )+ggtitle("Design-4")
 
@@
 #p_null+ p_varf + p_varx + p_varall
-ggpubr::ggarrange(p_null, p_varf, p_varx, p_varall, labels = c("Design-1", "Design-2", "Design-3", "Design-4"))
+ggpubr::ggarrange(p_null, p_varf, p_varx, p_varall)
+
+#labels = c("Design-1", "Design-2", "Design-3", "Design-4"))
 
@@
 
-| Granularity type                                                   	| # Significant 	| # Replications 	|
-|--------------------------------------------------------------------	|---------------	|----------------	|
-| **Individual**  <br><br># obs: 300, 500, 2000  <br># clusters: 6/7 	| 1/2/3         	| 25, 100, 200   	|
-| **Interaction**  <br><br># obs: 500, 2000  <br># clusters: 4       	| 1/2           	| 25, 100, 200   	|
+<!-- | Granularity type                                                   	| # Significant 	| # Replications 	| -->
+<!-- |--------------------------------------------------------------------	|---------------	|----------------	| -->
+<!-- | **Individual**  <br><br># obs: 300, 500, 2000  <br># clusters: 4/5 	| 1/2/3         	| 25, 100, 200   	| -->
+<!-- | **Interaction**  <br><br># obs: 500, 2000  <br># clusters: 4       	| 2           	| 25, 100, 200   	| -->
 

### 2021-10-19 main.Rmd
validation intro section fixed

@@
 We spiked several attributes in the data to see where one method works better than the other and where they might give us the same outcome or the effect of missing data <!---and trends--> on the proposed methods.
-A continuous measured variable $y$ of length $T$ indexed by $\{0, 1, \dots T-1\}$.
-Three circular granularities $g1$, $g2$ and $g3$ are considered with $2$, $3$ and $5$ levels respectively.
-Categories of $g1$, $g2$ and $g3$ are represented by ${g10,g11}$, ${g20, g21, g22}$ and ${g30, g31, g32, g33, g34}$.
+Three circular granularities $g1$, $g2$ and $g3$ are considered with categories denoted by ${g10,g11}$, ${g20, g21, g22}$ and ${g30, g31, g32, g33, g34}$ and levels $l_{g_1}=2$, $l_{g_2}=3$ and $l_{g_3}=5$.
 These categories could be integers or some more meaningful labels.
 For example, the granularity "day-of-week" could be either represented by ${0, 1, 2, \dots, 6}$ or ${Mon, Tue, \dots, Sun}$.
+Here categories of $g1$, $g2$ and $g3$ are represented by  $\{0, 1\}$, $\{0, 1, 2\}$ and $\{0, 1, 2, 3, 4\}$ respectively.
+A continuous measured variable $v$ of length $T$ indexed by $\{0, 1, \dots T-1\}$ is simulated such that it follows the structure across $g1$, $g2$ and $g3$.
 We created independent replications $R = \{25, 250, 500\}$ of all data designs to see if our proposed clustering approaches can detect distinct designs in various groups for small, medium and large number of series.
@@
 So a mean difference of $diff = \{1, 2, 5\}$ for corresponding categories are also considered.
-The code for creating these designs can be found in the Supplementary section (link to github repo).
-The results for $T=300$ and $R=25$ is shown, that means we have $25$ time series each with length $300$.
-The rest of the results could be found in the supplementary paper.
+The performance of the methods can vary with different number of significant granularities.
+So scenarios with all, few and just one significant granularities are considered.
+The code for creating these designs and the detailed results can be found in the Supplementary section (link to github repo).
+
+<!-- The results for $T=300$ and $R=25$ is shown, that means we have $25$ time series each with length $300$. The rest of the results could be found in the supplementary paper. -->
 

### 2021-10-19 main.Rmd
validation section DGP

@@
 
-Each of the $7$ categories from $\{g10,g11$, $g20, g21,\dots g32, g33, g34\}$ might be assumed to come from the same distribution, a subset of them from the same distribution, a subset of them from separate distributions, or all from different distributions, resulting in various data designs.
-As the methods ignores the linear progression of time, there is little value in adding time dependency in the data generating process.
-It is reasonable to construct a time series using properties such as trend, seasonality, and auto-correlation.
-However, when examining distributions across categories of cyclic granularities, these time series features are lost or addressed independently by considering all seasonal fluctuations.
-Because the time span during which an entity is observed in order to ascertain its behavior is very short, it is assumed that the time series will remain stationary throughout the observation period.
-The data type is set to "continuous," and the setup is assumed to be gaussian.
-For example, in a null design, all categories are considered to be normal (0,1).
-The mean of the subsequent categories are incremented or decremented from this null design based on the mean difference between them in non-null designs.
+Each category or combination of categories from $g1$, $g2$ and $g3$ are assumed to come from the same distribution, a subset of them from the same distribution, a subset of them from separate distributions, or all from different distributions, resulting in various data designs.
+As the methods ignore the linear progression of time, there is little value in adding time dependency in the data generating process.
+It is often reasonable to construct a time series using properties such as trend, seasonality, and auto-correlation.
+However, when examining distributions across categories of cyclic granularities, these time series features are lost or addressed independently by considering seasonal fluctuations through cyclic granularities.
+Because the time span during which an entity is observed in order to ascertain its behavior is not very long,
+the behavior of the entity will not change drastically and hence the time series can be assumed to remain stationary throughout the observation period.
+If the observation period is very long (for e.g more than 3 years), property, physical or geographical attributes might change leading to a non-stationary time series.
+But such a scenario is not considered here and the resulting clusters are assumed to be time invariant in the observation period.
+The data type is set to be "continuous," and the setup is assumed to be Gaussian.
+When the distribution of a granularity is "fixed", it means distributions across categories do not vary and are considered to be from N (0,1).
+The mean of different categories are altered in the "varying" designs, leading to varying distributions across categories.
 
@@
 granularity = c("g1", "g2", "g3")
-# alternate_design =  c("g10 ~ N(0, 1), g11 ~ N(2, 1)",
-#                       "g21 ~ N(2, 1), g22 ~ N(1, 1), g23 ~ N(0, 1)",
-#                       "g31 ~ N(0, 1), g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1), g35 ~ N(0, 1)")
-
-alternate_design =  c("g11 ~ N(2, 1)",
-                      "g21 ~ N(2, 1), g22 ~ N(1, 1)",
-                      "g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1)")
+alternate_design =  c("g10 ~ N(0, 1), g11 ~ N(2, 1)",
+                      "g21 ~ N(2, 1), g22 ~ N(1, 1), g23 ~ N(0, 1)",
+                      "g31 ~ N(0, 1), g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1), g35 ~ N(0, 1)")
+
+# alternate_design =  c("g11 ~ N(2, 1)",
+#                       "g21 ~ N(2, 1), g22 ~ N(1, 1)",
+#                       "g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1)")
 
@@
 ```
-
-<!-- A combination of these null and alternate distributions are considered when building the designs as in the table \ref{tab:tab-design}.  -->
 
@@
   list(tab_dist,  tibble(design, g1, g2, g3)),
-  caption = 'Alternate distributions of different categories  (top), 5 designs resulting from different distributions across categories (below)',
+  caption = 'For Scenario (a), distributions of different categories (top), 5 designs resulting from different distributions across categories (below)',
   booktabs = TRUE, valign = 't'

### 2021-10-19 main.Rmd
data design individual

@@
 
-  _(a) All signifiant granularities_
+  _Scenario (a): All signifiant granularities_
 
@@
 \noindent Consider the scenario when all three granularities $g1$, $g2$, and $g3$ are responsible for distinguishing the designs.
-That means that for at least one among the to-be-grouped designs, the pattern for each of $g1$, $g2$, and $g3$ will change.
-We consider different distributions across categories (as in Table \ref{tab:tab-dist-design}) (top) that will lead to different designs (as in Table \ref{tab:tab-dist-design}) (below).
+This implies that the patterns across each granularity will change significantly for at least one among the to-be-grouped designs.
+We consider different distributions across categories (as in Table \ref{tab:tab-dist-design} top) that will lead to different designs (as in Table \ref{tab:tab-dist-design} below).
 Figure \ref{fig:plot-3gran-new} shows the linear and cyclic representation of the simulated variable under these five designs.
-For the consequent data designs, only graphical representations will be provided.
-
-_(b) Few significant granularities_
+As could be seen from the plot, it is impossible to decipher the structural difference in the time series variable just by looking at the linear view.
+The difference in structure becomes quite clear when we see the distribution across cyclic granularities.
+Hence, for the consequent scenarios, only graphical displays across cyclic granularities are provided to emphasize the difference in structure.
+
+_Scenario (b): Few significant granularities_
 
 \noindent This is the case where one granularity will remain the same across all designs.
-We consider the case where the distribution of $y$ would vary across levels of $g2$ for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
+We consider the case where the distribution of $v$ would vary across levels of $g2$ for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
 So $g3$ is not responsible for distinguishing across designs.
-Figure \ref{#fig:plot-linear} shows the linear and cyclic representation of $y$.
-The first panel shows raw plot of $y$ in a linear scale and the second panel shows distribution of $y$ across cyclic granularities namely $g1$, $g2$ and $g3$.
-As could be seen from the plots, it is impossible to decipher from the raw time plot that the time series variable shows such pattern across different granularities.
+Figure
+\ref{#fig:plot-linear}(left) shows the considered design.
 
@@
 
-\noindent Here only one granularity is responsible for distinguishing the designs. Designs change significantly only for the granularity $g2$. Figure \ref{fig:plot-1gran} shows this.
+\noindent Here only one granularity is responsible for distinguishing the designs. Designs change significantly only for the granularity $g3$. Figure \ref{fig:plot-1gran}(right) shows this.
 
@@
 In this scenario, all of $g1$, $g2$ and $g3$ changes across at least one design.
-Also, it is not possible to comprehend these patterns across cyclic granularities or group similar series just by looking at the linear plots.", fig.pos="htb"}
+Also, it is not possible to comprehend these patterns across cyclic granularities or group similar series just by looking at the linear plots."}
 

### 2021-10-19 main.Rmd
interaction of granularity section

@@
 
-The earlier method could be extended when two granularities of interest interact and we are interested to group subjects based on the interaction of the two granularities.
-For example, in application, think about the scenario when customers need to grouped basis their joint behavior across hour-of-day and month-of-year.
+The proposed methods could be extended when two granularities of interest interact and we are interested to group subjects based on the interaction of the two granularities.
+For example, consider a group having a different weekday, weekend behavior in summer months, but not across winter.
+This type of joint behavior across granularities wknd-wday and month-of-year can be discovered by examining the distribution across combination of categories for different interacting granularities.
+Hence, in this scenario, we consider combination of categories to be generated from different distributions.
 For simplicity, consider a case with just two interacting granularities $g1$ and $g2$ of interest.
-As opposed to the last case, where we could play with the distribution of $5$ individual categories, with interaction, we can play with the distribution of $6$ combination of categories.
-Consider $4$ designs in Figure \ref{fig:} where different distributions are assumed for different combination of categories resulting in different designs.
-
-```{r}
+As opposed to the last case, where we could examine distributions across $l_{g_1} + l_{g_2} = 5$ individual categories, with interaction, we need to examine the distribution of $l_{g_1}*l_{g_2}=6$ combination of categories.
+Consider $4$ designs in Figure \ref{fig:interaction-gran} where different distributions are assumed for different combination of categories resulting in different designs.
+Design-1 has no change in distributions across $g1$ or $g2$, while Design-2 and Design-3 change across only $g1$ and $g2$ respectively.
+Design-4 changes across categories of both $g1$ and $g2$.
+Design-3 and Design-4 looks similar according to their relative difference between consecutive categories, but Design-4 also changes across facets, unlike Design-3 where all facets look the same.
+
+```{r interaction-gran, fig.cap = "Design-1 has no change in distributions across different categories of $g1$ or $g2$, while Design-2 and Design-3 change across only $g1$ and $g2$ respectively.
+Design-4 changes across categories of both $g1$ and $g2$."}
 ## ----designs

### 2021-10-19 main.Rmd
adding all codes in scripts

@@
 
-```{r tab-distribution, fig.pos="htb"}
-granularity = c("g1", "g2", "g3")
-alternate_design =  c("g10 ~ N(0, 1), g11 ~ N(2, 1)",
-                      "g21 ~ N(2, 1), g22 ~ N(1, 1), g23 ~ N(0, 1)",
-                      "g31 ~ N(0, 1), g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1), g35 ~ N(0, 1)")
-
-# alternate_design =  c("g11 ~ N(2, 1)",
-#                       "g21 ~ N(2, 1), g22 ~ N(1, 1)",
-#                       "g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1)")
-
-tab_dist <- tibble::tibble(granularity = granularity, `Varying distributions` = alternate_design)
-
-# knitr::kable(tab_dist, caption = "Alternate distributions of different categories if they deviate from null.")
+```{r tab-distribution}
 ```
@@
 ```{r tab-design}
-design = c("design-1", "design-2", "design-3", "design-4", "design-5")
-g1 = c("fixed", "vary", "fixed", "fixed", "vary")
-g2 = c("fixed", "fixed", "vary", "fixed", "vary")
-g3 = c("fixed", "fixed", "fixed", "vary", "vary")
-table <- tibble(design, g1, g2, g3)
-
-#%>% kable(caption = "5 different designs resulting from considering different distributions across categories.")
 ```
@@
 ```{r tab-dist-design}
-
-# library(kableExtra)
-# 
-# knitr::kables(list(kable(caption = "Alternate distributions of different categories if they deviate from null.",
-#    tab_dist
-#     ) %>% kable_styling(),
-#   kable(caption = "5 different designs resulting from considering different distributions across categories.",
-#  tibble(design, g1, g2, g3)
-#     ) %>% kable_styling()
-#   )
-# ) %>% kable_styling()
-# 
-knitr::kable(
-  list(tab_dist,  tibble(design, g1, g2, g3)),
-  caption = 'For Scenario (a), distributions of different categories (top), 5 designs resulting from different distributions across categories (below)',
-  booktabs = TRUE, valign = 't'
-)
 ```
@@
 ```{r generate-design-3change}
-generate_design <- function(t, mu1, mu2, mu3){
-
-t <- seq(0, t, 1)
-g1 <- t %%2
-g2 <- t %%3
-g3 <- t %%5
-
-# null design
-g1_dnull <- rep( rep(0, each = length(unique(g1))), length.out= length(t))
-g2_dnull <- rep( rep(0, each = length(unique(g2))), length.out= length(t))
-g3_dnull <- rep( rep(0, each = length(unique(g3))), length.out= length(t))
-
-# mean changing across categories in varying ways
-
-g1_dvary <- rep(mu1, length.out= length(t))
-g2_dvary <- rep(mu2, length.out= length(t))
-g3_dvary <- rep(mu3, length.out= length(t))
-
-design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
-design2 = distributional::dist_normal(g1_dvary + g2_dnull + g3_dnull)
-design3 = distributional::dist_normal(g1_dnull + g2_dvary + g3_dnull)
-design4 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dvary)
-design5 = distributional::dist_normal(g1_dvary + g2_dvary + g3_dvary)
-
-data_bind <- tibble::tibble(
-index = t,
-g1 = g1,
-g2 = g2,
-g3 = g3,
-design1 = distributional::generate(design1, times = 1) %>% unlist(),
-design2 = distributional::generate(design2, times = 1) %>% unlist(),
-design3 = distributional::generate(design3, times = 1) %>% unlist(),
-design4 = distributional::generate(design4, times = 1) %>% unlist(),
-design5 = distributional::generate(design5, times = 1) %>% unlist()
-) %>% 
-  pivot_longer(-c(1, 2, 3, 4), names_to = "design", values_to = "sim_data")
-
-data_bind
-}
-
-t = 300
-mu1= c(0, 2)
-mu2 = c(2, 1, 0)
-mu3 = c(0, 1, 2, 1, 0)
-
-data_bind <- generate_design(t, mu1, mu2, mu3)
 
@@
 
-# plot_linear_data <- function(data){
-# ggplot(data,
-#              aes(x = index, y = sim_data)) + 
-#   geom_line() +
-#   xlab("index")+
-#   theme_bw() 
-# }
-
-p1 <- ggplot(data_bind,
-             aes(x = index, y = sim_data)) + 
-  geom_line() +
-  xlab("index")+
-  facet_wrap(~design, scales = "free_y",ncol =1) +
-  theme_validation()
-
-p2 <- ggplot(data_bind,
-             aes(x = as.factor(g1), y = sim_data)) + 
-  geom_boxplot(alpha =0.5) + xlab("g1") + 
-  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") +
-  theme_validation()
-
-p3 <- ggplot(data_bind, aes(x = as.factor(g2), y = sim_data)) + geom_boxplot(alpha =0.5) + xlab("g2") + theme_bw() +
-  facet_wrap(~design, scales = "free_y",ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") + ylab("")+
-  theme_validation()
-
-p4 <- ggplot(data_bind, aes(x = as.factor(g3), y = sim_data)) + geom_boxplot(alpha =0.5) +
-  xlab("g3") + theme_bw()+
-  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue")+ ylab("")+
-  theme_validation()
-
-(p1 + (p2 + p3 + p4)) *
-  theme_validation() + plot_layout(widths = c(2, 1))
-# 
-# plot_cyclic_data( data_bind %>% filter(design=="design2"))
-# plot_data( data_bind %>% filter(design=="design3"))
-# plot_data( data_bind %>% filter(design=="design4"))
-# plot_data( data_bind %>% filter(design=="design5"))
 ```
@@
 
+```{r generate-design-new}
+
+```
+
+```{r generate-design-2gran-data}
+
+```
+
+```{r generate-design-2gran-plot}
+
+```
+
+```{r generate-design-1gran-data}
+
+```
+
+```{r generate-design-1gran-plot}
+
+```
+
 ```{r gran2and1-clubbed, fig.cap=" For the left scenario $g1$, $g2$ would change across atleast one design but $g3$ change remains same across all design.
@@
 #gran2_change  + gran1_change + plot_layout(widths = c(1,1))
-ggpubr::ggarrange(gran1_change, gran2_change, labels = c("a", "b"))
+#ggpubr::ggarrange(gran1_change, gran2_change, labels = c("a", "b"))
 ```

### 2021-10-19 quantiles_raw.R

@@ -85,7 +85,7 @@ plot(data_356cust_pc, type="l", npcs=50)
 
 library(liminal)
 data_356cust_pc10 <- as_tibble(data_356cust_pc$x[,1:10])
-limn_tour(data_356cust_pc10, PC1:PC)
+limn_tour(data_356cust_pc10, PC1:PC5)
 
 sort(abs(data_356cust_pc$rotation[,1]))
 

### 2021-10-19 script/smart_meter.R

@@ -218,3 +218,363 @@ p1 <- elec_ts %>%
 
 
 ggsave("figs/raw_plot_cust.png")
+
+
+##----tab-distribution
+granularity = c("g1", "g2", "g3")
+alternate_design =  c("g10 ~ N(0, 1), g11 ~ N(2, 1)",
+                      "g21 ~ N(2, 1), g22 ~ N(1, 1), g23 ~ N(0, 1)",
+                      "g31 ~ N(0, 1), g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1), g35 ~ N(0, 1)")
+
+# alternate_design =  c("g11 ~ N(2, 1)",
+#                       "g21 ~ N(2, 1), g22 ~ N(1, 1)",
+#                       "g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1)")
+
+tab_dist <- tibble::tibble(granularity = granularity, `Varying distributions` = alternate_design)
+
+# knitr::kable(tab_dist, caption = "Alternate distributions of different categories if they deviate from null.")
+
+
+##----tab-design
+design = c("design-1", "design-2", "design-3", "design-4", "design-5")
+g1 = c("fixed", "vary", "fixed", "fixed", "vary")
+g2 = c("fixed", "fixed", "vary", "fixed", "vary")
+g3 = c("fixed", "fixed", "fixed", "vary", "vary")
+table <- tibble(design, g1, g2, g3)
+
+#%>% kable(caption = "5 different designs resulting from considering different distributions across categories.")
+
+
+##----tab-dist-design
+# library(kableExtra)
+# 
+# knitr::kables(list(kable(caption = "Alternate distributions of different categories if they deviate from null.",
+#    tab_dist
+#     ) %>% kable_styling(),
+#   kable(caption = "5 different designs resulting from considering different distributions across categories.",
+#  tibble(design, g1, g2, g3)
+#     ) %>% kable_styling()
+#   )
+# ) %>% kable_styling()
+
+knitr::kable(
+  list(tab_dist,  tibble(design, g1, g2, g3)),
+  caption = 'For Scenario (a), distributions of different categories (top), 5 designs resulting from different distributions across categories (below)',
+  booktabs = TRUE, valign = 't'
+)
+
+##----generate-design-3change
[…]

### 2021-10-19 main.Rmd
code smells less

@@
 
-```{r}
-generate_design <- function(t, mu21, mu22, mu23){
-
-t <- seq(0, t, 1)
-g1 <- t %%2
-g2 <- t %%3
-g3 <- t %%5
-
-# null design
-g1_dnull <- rep( rep(0, each = length(unique(g1))), length.out= length(t))
-g2_dnull <- rep( rep(0, each = length(unique(g2))), length.out= length(t))
-g3_dnull <- rep( rep(0, each = length(unique(g3))), length.out= length(t))
-
-# mean changing across categories in varying ways
-
-g1_dvary <- rep(0, length.out= length(t))
-g21_dvary <- rep(mu21, length.out= length(t))
-g22_dvary <- rep(mu22, length.out= length(t))
-g23_dvary <- rep(mu23, length.out= length(t))
-g3_dvary <- rep(0, length.out= length(t))
-
-design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
-design2 = distributional::dist_normal(g1_dnull + g21_dvary + g3_dnull)
-design3 = distributional::dist_normal(g1_dnull + g22_dvary + g3_dnull)
-design4 = distributional::dist_normal(g1_dnull + g23_dvary + g3_dnull)
-
-data_bind <- tibble::tibble(
-index = t,
-g1 = g1,
-g2 = g2,
-g3 = g3,
-design1 = distributional::generate(design1, times = 1) %>% unlist(),
-design2 = distributional::generate(design2, times = 1) %>% unlist(),
-design3 = distributional::generate(design3, times = 1) %>% unlist(),
-design4 = distributional::generate(design4, times = 1) %>% unlist(),
-#design5 = distributional::generate(design5, times = 1) %>% unlist()
-) %>% 
-  pivot_longer(-c(1, 2, 3, 4), names_to = "design", values_to = "sim_data")
-
-data_bind
-}
-
-mu21= c(2, 0, 0)
-mu22 = c(0, 2, 0)
-mu23 = c(0, 0, 2)
-mu3 = c(0, 0, 0, 0, 0)
-data_bind <- generate_design(t, mu21, mu22, mu23)
-
-```
-
-```{r plot-gran, out.width="100%"}
-
-p1 <- ggplot(data_bind,
-             aes(x = index, y = sim_data)) + 
-  geom_line() +
-  xlab("index")+
-  facet_wrap(~design, scales = "free_y",ncol =1) +
-  theme_validation()
-
-p2 <- ggplot(data_bind,
-             aes(x = as.factor(g1), y = sim_data)) + 
-  geom_boxplot(alpha =0.5) + xlab("g1") + 
-  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") +
-  theme_validation()
-
-p3 <- ggplot(data_bind, aes(x = as.factor(g2), y = sim_data)) + geom_boxplot(alpha =0.5) + xlab("g2") + theme_bw() +
-  facet_wrap(~design, scales = "free_y",ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") + ylab("")+
-  theme_validation()
-
-p4 <- ggplot(data_bind, aes(x = as.factor(g3), y = sim_data)) + geom_boxplot(alpha =0.5) +
-  xlab("g3") + theme_bw()+
-  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue")+ ylab("")+
-  theme_validation()
-
-gran2_change <- (p2 + p3 + p4) *
-  theme_validation()
-```
-
-<!-- When we move from the linear to the cyclic world of temporal granularities, we can see patterns across different categories of the granularities which gets lost in the linear representation.
-Populations are modeled by a collection of these cyclic granularities.
-Each cyclic granularity might or might not have patterns across its categories.
-Each cluster is characterized by similar patterns across one or more of these cyclic granularities.
--->
-
-<!-- A small example is given to setup the problem. -->
-
-<!-- We constructed the simulation parameters to represent common patterns in electricity data.
-50, 200 and 500 time series were chosen for different simulation designs with different granularities and patterns changing across different granularities.
--->
-<!-- The data type is fixed to be "continuous".We generated independent replications of all combinations of the simulation parameters. -->
-
-<!-- Consider a continuous time series variable $y$ of length $T$ indexed by ${0, 1, \dots T-1}$.
-Three circular granularities $g1$, $g2$ and $g3$ are considered with 2, 3 and 5 levels respectively.
-Categories of g1, g2 and g3 are represented by ${0,1}$, ${0, 1, 2}$ and ${0, 1, 2, 3, 4}$.
-These categories could be integers or some more meaningful labels.
-For example, the granularity "day-of-week" could be either represented by $\{0, 1, 2, \dots, 6\}$ or $\{Mon, Tue, \dots, Sun\}$.
--->
-
-<!-- Consider the case where distribution of $y$ would vary across levels of $g1$ and g2$ but will not vary across levels of $g3$ as described in Figure .Figure \ref shows four designs where $g3$ acts as a nuisance variable, and $g1$ and $g2$ varies in both  Design 2 and 3, and $g2$ varies in all designs.
--->
-
-<!-- Consider a case where distribution of $y$ would vary across levels of $g2$ -->
-<!-- for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
-Figure \ref{#fig:plot-linear} shows the linear and cyclic representation of $y$.
-The first panel shows raw plot of $y$ in a linear scale and the second panel shows distribution of $y$ across cyclic granularities namely $g1$, $g2$ and $g3$.
-As could be seen from the plots, it is impossible to decipher from the raw time plot that the time series variable shows such pattern across different granularities.
--->
-
-```{r change-var, echo=FALSE}
-t = 300
-n <- seq(0, t, 1)
-g1 <- n %%2
-g2 <- n %%3
-g3 <- n %%5
-mu11 = c(0, 5)
-mu12 = c(3, 0)
-mu21= c(2, 0, 0)
-mu22 = c(0, 2, 0)
-mu23 = c(0, 0, 2)
-mu3 = c(0, 0, 0, 0, 0)
-```
-
-```{r generate-design} 
-generate_design <- function(t, mu21, mu22, mu23){
-
-t <- seq(0, t, 1)
-g1 <- t %%2
-g2 <- t %%3
-g3 <- t %%5
-
-# null design
-g1_dnull <- rep( rep(0, each = length(unique(g1))), length.out= length(t))
-g2_dnull <- rep( rep(0, each = length(unique(g2))), length.out= length(t))
-g3_dnull <- rep( rep(0, each = length(unique(g3))), length.out= length(t))
-
-# mean changing across categories in varying ways
-
-g11_dvary <- rep(mu11, length.out= length(t))
-g12_dvary <- rep(mu12, length.out= length(t))
-g21_dvary <- rep(mu21, length.out= length(t))
-g22_dvary <- rep(mu22, length.out= length(t))
-g23_dvary <- rep(mu23, length.out= length(t))
-g3_dvary <- rep(0, length.out= length(t))
-
-design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
-design2 = distributional::dist_normal(g11_dvary + g21_dvary + g3_dnull)
-design3 = distributional::dist_normal(g12_dvary + g22_dvary + g3_dnull)
-design4 = distributional::dist_normal(g1_dnull + g23_dvary + g3_dnull)
-
-data_bind <- tibble::tibble(
-index = t,
-g1 = g1,
-g2 = g2,
-g3 = g3,
-design1 = distributional::generate(design1, times = 1) %>% unlist(),
-design2 = distributional::generate(design2, times = 1) %>% unlist(),
-design3 = distributional::generate(design3, times = 1) %>% unlist(),
-design4 = distributional::generate(design4, times = 1) %>% unlist(),
-#design5 = distributional::generate(design5, times = 1) %>% unlist()
-) %>% 
-  pivot_longer(-c(1, 2, 3, 4), names_to = "design", values_to = "sim_data")
-
-data_bind
-}
-
-data_bind <- generate_design(t, mu21, mu22, mu23)
-
-#data_bind
-
-```
-
-```{r plot-2gran, out.width="100%", fig.cap = "The linear (left) and cyclic (right) representation of the measured variable is shown.
-In this scenario, $g1$, $g2$ would change across atleast one design but $g3$ change remains same across all design.
-Thus $g3$ is not an important variable to distinguish these designs."}
-
-# plot_linear_data <- function(data){
-# ggplot(data,
-#              aes(x = index, y = sim_data)) + 
-#   geom_line() +
-#   xlab("index")+
-#   theme_bw() 
-# }
-
-p1 <- ggplot(data_bind,
-             aes(x = index, y = sim_data)) + 
-  geom_line() +
-  xlab("index")+
-  facet_wrap(~design, scales = "free_y",ncol =1) +
-  theme_validation()
-
-p2 <- ggplot(data_bind,
-             aes(x = as.factor(g1), y = sim_data)) + 
-  geom_boxplot(alpha =0.5) + xlab("g1") + 
-  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") +
-  theme_validation()
-
-p3 <- ggplot(data_bind, aes(x = as.factor(g2), y = sim_data)) + geom_boxplot(alpha =0.5) + xlab("g2") + theme_bw() +
-  facet_wrap(~design, scales = "free_y",ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") + ylab("")+
-  theme_validation()
-
-p4 <- ggplot(data_bind, aes(x = as.factor(g3), y = sim_data)) + geom_boxplot(alpha =0.5) +
-  xlab("g3") + theme_bw()+
-  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue")+ ylab("")+
-  theme_validation()
-
-gran1_change <- (p2 + p3 + p4) *
-  theme_validation()
-# 
-# plot_cyclic_data( data_bind %>% filter(design=="design2"))
-# plot_data( data_bind %>% filter(design=="design3"))
-# plot_data( data_bind %>% filter(design=="design4"))
-# plot_data( data_bind %>% filter(design=="design5"))
-```
-
-<!-- A subset of many possible designs are shown in Figure \ref{fig: plot-linear }.
-For the parameter space (XXX unique  combinations shown in table YYY), 100, 500 independent replications of all possible combination of simulation parameters were generated.
-The clustering methodologies were run all these unique combinations and subsets of these to verify if the methodologies work as expected.
--->
-
 ```{r generate-design-new}
@@
 ```{r gran2and1-clubbed, fig.cap=" For the left scenario $g1$, $g2$ would change across atleast one design but $g3$ change remains same across all design.
-For tthe right one, only $g3$ changes across different designs."}
+For the right one, only $g3$ changes across different designs."}
 #gran2_change  + gran1_change + plot_layout(widths = c(1,1))
@@
 Design-4 changes across categories of both $g1$ and $g2$."}
-## ----designs
-
-sim_null_normal <- function(nxj, nfacetj, mean, sd, w1 = 0, w2=0) {
-  rep(distributional::dist_normal(mu = mean, sigma = sd),
-      times = nxj * nfacetj
-  )
-}
-
-sim_varf_normal <- function(nx, nfacet, mean, sd, w1, w2) {
-  rep(dist_normal((mean + seq(0, nfacet - 1, by = 1) * w1),
-                  (sd + seq(0, nfacet - 1, by = 1) * w2)), each = nx)
-}
-
-sim_varx_normal <- function(nx, nfacet, mean, sd, w1, w2) {
-  rep(dist_normal((mean + seq(0, nx - 1, by = 1) * w1),
-                  (sd + seq(0, nx - 1, by = 1) * w2)), nfacet)
-}
-
-sim_varall_normal <- function(nx, nfacet, mean, sd, w1, w2) {
-  dist_normal((mean + seq(0, (nx *nfacet - 1), by = 1) * w1),
-              (sd + seq(0,  (nx *  nfacet - 1), by = 1) * w2))
-}
-
-## ----data
-
-library(hakear)
-
-nx_val = 2 # number of x-axis levels
-nfacet_val = 3 # number of facet levels
-w1_val = 1 # increment in mean
-w2_val = 0 # increment in sd
-mean_val = 0 # mean of normal distribution of starting combination
-sd_val = 2 # sd of normal distribution of starting combination
-quantile_prob_val = seq(0.1, 0.9, 0.1)
-ntimes_val = 300
-
-sim_panel_varall <- sim_panel(
-  nx = nx_val, nfacet =  nfacet_val,
-  ntimes = ntimes_val,
-  # sim_dist = sim_varall_normal(2, 3, 5, 10, 5, -1.5)
-  sim_dist = sim_varall_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
-) %>% unnest(data)  %>% 
-  rename( "g2" = "id_facet",
-          "g1" = "id_x")
-
-sim_panel_varx <- sim_panel(
-  nx = nx_val, nfacet =  nfacet_val,
-  ntimes = ntimes_val,
-  # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
-  sim_dist = sim_varx_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
-) %>% unnest(data)  %>% 
-  rename( "g2" = "id_facet",
-          "g1" = "id_x")
-
-sim_panel_varf <- sim_panel(
-  nx = nx_val, nfacet =  nfacet_val,
-  ntimes = ntimes_val,
-  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
-  sim_dist = sim_varf_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
-) %>% unnest(data)  %>% 
-  rename( "g2" = "id_facet",
-          "g1" = "id_x")
-
-sim_panel_null <- sim_panel(
-  nx = nx_val,
-  nfacet =  nfacet_val,
-  ntimes = ntimes_val,
-  sim_dist = distributional
-  ::dist_normal(mean_val, sd_val)
-) %>% unnest(c(data)) %>% 
-  rename( "g2" = "id_facet",
-          "g1" = "id_x")
-
-## ----category-plots
-
-p_null <- sim_panel_null %>%
-  ggplot(aes(x = as.factor(g1), y = sim_data)) +
-  facet_wrap(~g2, labeller = "label_both") +
-  geom_boxplot() +
-  ylab("") +
-  #ggtitle(paste("(a)", round(null, 2))) +
-  xlab("") +
-  theme_bw()+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") +
-  #ggtitle("design1")+
-  theme_validation() +ylab("")  +
-  theme(panel.spacing =unit(0, "lines"))+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") ) +ggtitle("Design-1")
-
-p_varf <- sim_panel_varf %>%
-  ggplot(aes(x = as.factor(g1), y = sim_data)) +
-  facet_wrap(~g2, labeller = "label_both") +
-  geom_boxplot() +
-  ylab("") +
-  #ggtitle(paste("(b)", round(varf, 2))) +
-  xlab("")+
-  theme_bw()+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") +
-  #ggtitle("design2")+
-  theme_validation()+ylab("") +
-  theme(panel.spacing =unit(0, "lines"))+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )+ggtitle("Design-2")
-
-p_varx <- sim_panel_varx %>%
-  ggplot(aes(x = as.factor(g1), y = sim_data)) +
-  facet_wrap(~g2, labeller = "label_both") +
-  geom_boxplot()+
-  ylab("")  +
-  #ggtitle(paste("(c)", round(varx, 2))) +
-  xlab("")+
-  theme_bw()+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") +
-  #ggtitle("design3")+
-  theme_validation()+ylab("") +
-  theme(panel.spacing =unit(0, "lines"))+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )+ggtitle("Design-3")
-
-p_varall <- sim_panel_varall %>%
-  ggplot(aes(x = as.factor(g1), y = sim_data)) +
-  facet_wrap(~g2, labeller = "label_both") +
-  geom_boxplot() +
-  ylab("") +
-  #ggtitle(paste("(d)", round(varall, 2))) +
-  xlab("")+
-  theme_bw()+ stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 0.8, color = "blue") +
-  #ggtitle("design4") +
-  theme_validation()+ylab("") +
-  theme(panel.spacing =unit(0, "lines"))+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )+ggtitle("Design-4")
-
-change_index <- function(data){
-
-index_new <- map(seq_len(ntimes_val), function(i){
-  map((seq_len(nx_val*nfacet_val)), function(j)
-    {
-value = i + (j-1)*ntimes_val
-})
-  }) %>% unlist()
-
-data_new = data %>%
-  ungroup() %>%
-  mutate(index_old = row_number(),
-         index_new = index_new)
-
-y = data_new[match(index_new, data_new$index_old),]
-
-y <- y %>%
-  mutate(time = row_number())
-
-return(y)
-}
-
-endbreaks<- nrow(sim_panel_null)
-
-p1 <- change_index(sim_panel_null) %>%
-  ggplot(aes(x = time,
-             y = sim_data)) +
-  geom_line() +
-  scale_x_continuous(breaks = seq(1, endbreaks, 200))+
-  theme_bw() +
-  #geom_point(alpha = 0.5, color = "blue")
-  #theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
-    #    panel.grid.minor.x =  element_blank())+
-  ylab("")+
-  xlab("")
-
-p2 <- change_index(sim_panel_varf) %>%
-  ggplot(aes(x = time,
-             y = sim_data)) +
-  geom_line() +
-  scale_x_continuous(breaks = seq(1, endbreaks, 200))+
-  theme_bw() +
-  #geom_point(alpha = 0.5, color = "blue") +
-  #theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
-     #   panel.grid.minor.x =  element_blank())+
-  ylab("")+
-  xlab("")
-
-p3 <- change_index(sim_panel_varx) %>%
-  ggplot(aes(x = time,
-             y = sim_data)) +
-  geom_line() +
-  scale_x_continuous(breaks = seq(1, endbreaks, 200))+
-  theme_bw() +
-  #geom_point(alpha = 0.5, color = "blue")+
-  #theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
-   #     panel.grid.minor.x =  element_blank())+
-  ylab("")+
-  xlab("")
-
-p4 <- change_index(sim_panel_varall) %>%
-  ggplot(aes(x = time,
-             y = sim_data)) +
-  geom_line() +
-  scale_x_continuous(breaks = seq(1, endbreaks, 200))+
-  theme_bw() +
-  #geom_point(alpha = 0.5, color = "blue") +
-  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
-  panel.grid.minor.x =  element_blank())+
-  ylab("")+
-  xlab("")
-
-#((p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)) 
-#p_null+ p_varf + p_varx + p_varall
-ggpubr::ggarrange(p_null, p_varf, p_varx, p_varall)
-
-#labels = c("Design-1", "Design-2", "Design-3", "Design-4"))
 

### 2021-10-19 script/smart_meter.R

@@ -578,3 +578,233 @@ gran1_change <- (p2 + p3 + p4) *
 ##----gran2and1-clubbed
 
 ggpubr::ggarrange(gran2_change, gran1_change, labels = c("a", "b"))
+
+
+##----interaction-gran
+
+##designs
+
+sim_null_normal <- function(nxj, nfacetj, mean, sd, w1 = 0, w2=0) {
+  rep(distributional::dist_normal(mu = mean, sigma = sd),
+      times = nxj * nfacetj
+  )
+}
+
+sim_varf_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  rep(dist_normal((mean + seq(0, nfacet - 1, by = 1) * w1),
+                  (sd + seq(0, nfacet - 1, by = 1) * w2)), each = nx)
+}
+
+sim_varx_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  rep(dist_normal((mean + seq(0, nx - 1, by = 1) * w1),
+                  (sd + seq(0, nx - 1, by = 1) * w2)), nfacet)
+}
+
+sim_varall_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  dist_normal((mean + seq(0, (nx *nfacet - 1), by = 1) * w1),
+              (sd + seq(0,  (nx *  nfacet - 1), by = 1) * w2))
+}
+
+
+##data
+
+library(hakear)
+
+nx_val = 2 # number of x-axis levels
+nfacet_val = 3 # number of facet levels
+w1_val = 1 # increment in mean
+w2_val = 0 # increment in sd
+mean_val = 0 # mean of normal distribution of starting combination
+sd_val = 2 # sd of normal distribution of starting combination
+quantile_prob_val = seq(0.1, 0.9, 0.1)
+ntimes_val = 300
+
+sim_panel_varall <- sim_panel(
+  nx = nx_val, nfacet =  nfacet_val,
+  ntimes = ntimes_val,
+  # sim_dist = sim_varall_normal(2, 3, 5, 10, 5, -1.5)
+  sim_dist = sim_varall_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+) %>% unnest(data)  %>% 
+  rename( "g2" = "id_facet",
+          "g1" = "id_x")
+
+sim_panel_varx <- sim_panel(
+  nx = nx_val, nfacet =  nfacet_val,
+  ntimes = ntimes_val,
[…]

### 2021-10-20 main.Rmd
graphs look better

@@
 
-```{r interaction-gran, fig.cap = "Design-1 has no change in distributions across different categories of $g1$ or $g2$, while Design-2 and Design-3 change across only $g1$ and $g2$ respectively.
-Design-4 changes across categories of both $g1$ and $g2$."}
+```{r interaction-gran, fig.cap = "Design-1 (a) has no change in distributions across different categories of $g1$ or $g2$, while Design-2 (b) and Design-3 (c) change across only $g1$ and $g2$ respectively.
+Design-4 (d) changes across categories of both $g1$ and $g2$."}
 

### 2021-10-20 gran1_orig.R

@@ -24,8 +24,8 @@ simtable <- expand.grid(mean_diff = mean_diff,
 
 # parallel::mclapply(seq_len(nrow(simtable)), function(scen){
 
-scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
-#scen <- 10
+#scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
+scen <- 1
 
 
 simj<-simtable[scen,] #Extract row of table
@@ -144,72 +144,40 @@ set.seed(123)
 bind_data_iter <- map(seq_len(niterj), 
                       function(x){
                         #set.seed(seed_len[x])  
-                        design1 <- generate_design(n=nTj) # null design
-                        design2 <- generate_design(n=nTj,
-                                                   mu31 = mean_diffj) 
-                                                   #mu32=mean_diffj , 
-                                                   #mu34=mean_diffj)
+                        # design1 <- generate_design(n=nTj) # null design
+                        # design2 <- generate_design(n=nTj,
+                        #                            mu31 = mean_diffj) 
+                        #                            
+                        # 
+                        # design3 <- generate_design(n=nTj, 
+                        #                            mu32=mean_diffj) 
+                        #                       
+                        # 
+                        # design4 <- generate_design(n=nTj,
+                        #                            mu33 = mean_diffj)
+                        #                  
+                        # design5 <- generate_design(n=nTj,
+                        #                            mu32 = mean_diffj,
+                        #                            mu33 = mean_diffj)
+                        # bind_design <- bind_rows(design1, design2, design3, design4, design5,  .id = "design")
+                        # 
                         
-                        design3 <- generate_design(n=nTj, 
+                        design1 <- generate_design(n=nTj) 
+     
+                        
+                        design2 <- generate_design(n=nTj, 
                                                    mu32=mean_diffj) 
-                                                   #mu32=mean_diffj , 
[…]

### 2021-10-20 main.Rmd
characterisation clusters updated

@@
 
-In the analysis of energy smart metre data, K-Means or hierarchical clustering are often employed.
+In the analysis of energy smart meter data, K-Means or hierarchical clustering are often employed.
 These are simple and effective techniques that work well in a range of scenarios.
@@
 We employ agglomerative hierarchical clustering in conjunction with Ward's criteria (XXX reference).
-Individual entities with the highest similarity computed using the desired distance metrics are sequentially merged using agglomerative algorithms.
+The pair of clusters with minimum between-cluster distance are are sequentially merged in this using this agglomerative algorithms.
+A good comprehensive list of algorithms can be found in @.Xu2015-ja.
 We can possibly employ any clustering method that supports the given distance metric as input.
@@
 
+Characterization of clusters both statistically and qualitatively is an important stage of a cluster analysis.
+A potential way is to look at the findings from all the groups in graphs, and enhance our qualitative descriptions of the groupings.
+@Cook2007-qe provides several ways to characterize clusters.
+
+(a) Parallel coordinate plot: Parallel coordinate plot (Inselberg 1985, Wegman 1990) are widely used to display high-dimensional and multivariate data, allowing visual grouping to detect patterns.
+In a Parallel Coordinates Plot, each variable has its own axis, which are all parallel.
+Each axis is connected by a series of lines.
+That is, each line is made up of connected points on each axis.
+The order of the axes can affect how the reader interprets the data.
+This is because adjacent variables are more easily perceived than non-adjacent variables and rearranging the axes can reveal patterns or correlations between variables.
+Scattered plots of the p variables are arranged in a scatterplot matrix.
+It's a neat way to show multiple relationships at once, and it allows us to compare all the plots at once.
+
+(b) Scatterplot matrix: The scatter plot matrix (draftsman's plot) is a matrix that comprises pairwise scatter plots of the p variables.Pairwise scatter plots are excellent for determining relationships between variables and determining which factors have contributed the most to clustering.
+
+(c) Plotting cluster statistics: For larger problems, parallel coordinate plots may become cluttered and difficult to read, therefore we may opt to display cluster statistics instead.
+(@Dasu_undated-gw)
+
+(c) MDS, PCA and t-SNE: While all of the techniques examine a matrix of distances or dissimilarities to give a representation of the data points in a reduced-dimension space, their goals are not the same.
+The principal component analysis (Johnson & Wichern 2002) attempts to retain data variance.
+Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
+t-SNE, on the other hand, is concerned with preserving neighbourhood data points.
+Close data points in high-dimensional space will be condensed in the t-SNE embeddings.
+
+(d) Tour: A tour () is a collection of interpolated linear projections of multivariate data into a lower-dimensional space.
+The sequence is seen as a dynamic visualisation, enabling the viewer to observe the shadows cast by the high-dimensional data in a lower-dimensional view.
+<!-- and also provides examples on how one might spin and brush using a tour, to build up a cluster analysis manually. -->
+
 Depending on the distance measure utilized for the study, the cluster characterization technique will differ.
-Clusters that utilise intra-category distances are characterised using multi-dimensional scaling and parallel coordinate displays.
-For inter-category distances, the distribution across major granularities may be presented to ensure that the goal of similar shapes within clusters and distinct shapes across clusters is met.
+Clusters that utilize wpd-based distances are characterised using multi-dimensional scaling and parallel coordinate displays.
+For JS-based distances, the distribution across major granularities may be presented to ensure that the goal of similar shapes within clusters and distinct shapes across clusters is met.
 This technique may potentially make advantage of multi-dimensional scaling.
-
-\noindent Multidimensional scaling (MDS) (XXX reference) refers to a family of methods that analyse a matrix of distances or dissimilarities to provide a representation of the data points in a reduced-dimension space.
-There are many kinds of MDS, but they all solve the same fundamental issue: Given a $n*n$ matrix of dissimilarities and a distance measure, identify a configuration of $n$ points $x_1, x_2, \dots, x_n$ in the reduced dimension space $R^q$ ($q<p$) where the distance between the points is near to the dissimilarity between the points.
-All techniques must determine the coordinates of the points as well as the space dimension, $q$.
-Metric and nonmetric MDS are the two main kinds of MDS.
-Metric MDS methods presume a functional connection between the interpoint distances and the supplied dissimilarities and assume that the data are quantitative.
-We use metric MDS.
-
-\noindent Parallel coordinate plots (XXX reference) Parallel coordinates have been extensively used to display high-dimensional and multivariate data, allowing for the detection of patterns within the data via visual grouping.
 
@@
 For smaller difference between categories, it is expected that method js-nqt would preform better than the other two.
+<!-- A confusion table can come alive with linked brushing, so that mismatches and agreements between methods can be explored. -->
 

### 2021-10-20 main.Rmd
adding references

@@
 
-<!-- In contrast to models, a feature-based strategy is used to explicitly define or automatically extract the curves’ key time features, for instance by application of PCA on the daily curves [ -->
+<!-- In contrast to models, a feature-based strategy is used to explicitly define or automatically extract the curves’ key time features, for instance by application of PCA on the daily curves [ --
 
@@
 
+<!-- Because of this importance, countless approaches to estimate time series similarity have been proposed.  -->
+
 The proposed methodology aim to leverage the intrinsic temporal data structure hidden in time series data.
-The foundation of our method is unsupervised clustering algorithms based exclusively on time-series features.
-First, we study the underlying distributions that may have resulted in different patterns across temporal granularities in order to identify a mechanism to classify them based on the similarity of those distributions.
+The foundation of our method is unsupervised clustering algorithms based exclusively on the time-series data.
+The similarity measure is the most essential ingredient of time series clustering.
+The (dis) similarity measure in this paper focuses on looking at the (dis) similarity between underlying distributions that may have resulted in different patterns across different cyclic temporal granularities.
 It is worth noting that when studying these similarities, a variety of objectives may be pursued.
-One objective could be to group consumers with similar shapes over all relevant cyclic granularities.
+One objective could be to group time series with similar shapes over all relevant cyclic granularities.
 In this scenario, the variation in customers within each group is in magnitude rather than shape, while the variation between groups is only in shape.
-Most clustering algorithms offer only daily energy profiles throughout the hours of the day, but we suggest a broader approach to the problem, aiming to group consumers with similar shapes across all significant cyclic granularities.
+There
+are distance measures are used for shape-based clustering [Ding et al. 2008; Wang
+et al.
+2013] and many more but none of them look at the probability distributions while computing similarity.
+Moreover, most distance measures offer similar shape across just one dimension.
+For example, we often see "similar" daily energy profiles across hours of the day, but we suggest a broader approach to the problem, aiming to group consumers with similar distributional shape across all significant cyclic granularities.
 Another purpose of clustering could be to group customers that have similar differences in patterns across all major cyclic granularities, capturing similar jumps across categories regardless of the overall shape.
@@
 Both of these objectives may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
-
 Depending on the goal of clustering, the distance metric for defining similarity would be different.
@@
 In that case, when the customers are grouped, we can expect to observe some interesting patterns across the categories of the cyclic granularities considered.
-[XXX reference 2nd chapter] proposes a way to select significant cyclic granularities and harmonies which is used for this paper.
+[@Gupta2021-hakear] proposes a way to select significant cyclic granularities and harmonies which is used for this paper.
 
@@
 
-(a) Parallel coordinate plot: Parallel coordinate plot (Inselberg 1985, Wegman 1990) are widely used to display high-dimensional and multivariate data, allowing visual grouping to detect patterns.
+(a) Parallel coordinate plot: Parallel coordinate plot (@wegman1990hyperdimensional) are widely used to display high-dimensional and multivariate data, allowing visual grouping to detect patterns.
 In a Parallel Coordinates Plot, each variable has its own axis, which are all parallel.
@@
 
-(d) Tour: A tour () is a collection of interpolated linear projections of multivariate data into a lower-dimensional space.
-The sequence is seen as a dynamic visualisation, enabling the viewer to observe the shadows cast by the high-dimensional data in a lower-dimensional view.
+(d) Tour: A tour (@wickham2011tourr) is a collection of interpolated linear projections of multivariate data into a lower-dimensional space.
+The sequence is seen as a dynamic visualization, enabling the viewer to observe the shadows cast by the high-dimensional data in a lower-dimensional view.
 <!-- and also provides examples on how one might spin and brush using a tour, to build up a cluster analysis manually. -->

### 2021-10-20 old/algo2-elec16.Rmd
cleaning folder

--- 2021-07-16 algo2-elec16.Rmd
@@
                    alphaLines = 0.8,
-                   scale = "globalminmax"
+                   scale = "globalminmax", 
+                   order = "anyClass"
 ) + ggplot2::theme_bw() +

### 2021-10-20 old/graclust.Rmd
cleaning folder

---
title: Clustering probability distributions across bivariate cyclic temporal granularities

# to produce blinded version set to 1
blinded: 0

authors:
- name: Sayani Gupta
  affiliation: Department of Econometrics and Business Statistics, Monash University
  thanks: "Email: Sayani.Gupta@monash.edu"

- name: Rob J Hyndman
  affiliation: Department of Econometrics and Business Statistics, Monash University

- name: Dianne Cook
  affiliation: Department of Econometrics and Business Statistics, Monash University

keywords:
- data visualization
- statistical distributions
- time granularities
- calendar algebra
- periodicties
- grammar of graphics
- R

abstract: |
 Clustering elements based on behavior across time granularities 
bibliography: [bibliography.bib]
preamble: >
  \setlength {\marginparwidth }{2cm}
  \usepackage{mathtools,amssymb,booktabs,longtable,todonotes,amsthm}
  \def\mod{~\text{mod}~}
output:
  bookdown::pdf_book:
    base_format: rticles::asa_article
    fig_height: 4
    fig_width: 6
    fig_caption: yes
    dev: "pdf" 
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(ggplot2)
library(gravitas)
library(gracsR)
library(ggdendro)
library(dplyr)
```

# Introduction

# Conventional distance matrices

# A new distance measure

<!-- We will use the following notations and symbols. Let -->
<!-- x be a continuous measurement variable which can have one of q -->
<!-- values contained in the set, $X = {x_1; x_2, \dots, x_q}$, where q is the quantile vector correposding to probability vector p. Consider n observations whose measurements of the value -->
<!-- of x are: $A = {a_1, a_2, \dots, a_n}$ where $a_j \in X$ . The distribution of of the set A along measurement x is H(x; A) which is an -->
<!-- ordered list (or b-dimensional vector) consisting of the -->
<!-- number of occurrences of the discrete values of x among -->
<!-- the ai. As we are interestedonly in comparing the histograms of the same measurement x, H(A) will be used -->
<!-- in place of H(x; A) without loss of generality. If Hi(A), -->

Consider two cyclic granularities $A$ and $B$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets. Let $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ be a continuous variable observed across $T$ time points. Let $wpd_{norm_{s,t}}(A,B)$ be the normalised pairwise distance between subjects/elements $s$ and $t$ for the harmony pair $(A, B)$.

, then we can define $S_{i,j} = \sum_{c_1} \sum_{c_2} D_{i,j}(c_1,c_2)$.

The proofs are attempted assuming sum is taken. Can we take maximum or minumum distances? That would finally imply we want to cluster households those are having similar extreme behavior, as we would be minimising the maximum distances through clustering.

\noindent Now since JS is a metric, non-negativity, reflexivity and triangle inequality holds for it. Thus,   
-  $D_{x, x} \geq 0$ with equality only if $x=y$  
-  $D(x, y) = D(y, x)$  
-  $D(x, y) + D(x, z) \geq D(y, z)$

## Metric property

### Non-negative property

$S_{i,j}$ is nothing but the sum of $D(i,j)$ for all levels of $c_1$ and $c_2$ and each $D(i,j)$ has non-negativity property. Therefore, $S_{i,j}$ also has the non-negativity by definition.

### Reflexivity property

Since $D(i,i)$ = 0 is true, $\sum_{i} D(i, i) = 0$. Due to the non-negativity, 0 is the minimum bound of the measure: $S_{i,j} \geq 0$. Therefore, $S(i, i) = \sum_{i} D(i, i)= 0$.

### Commutative property

To prove: $S(x, y) = S(y, x)$

### Triangle inequality property

Since, $D(i, j) + D(j, k) \geq D(i, k)$  
Therefore, $\sum_{c_1} \sum_{c_2}D(i, j) + \sum_{c_1} \sum_{c_2}D(i, j) \geq \sum_{c_1} \sum_{c_2}D(i, k)$   
or, $S(i, j) + S(j, k) \geq S(i, k)$

###

## Normalization

## Comparison with other distances

# Algorithm

## Simulations

```{r set_common_input}
.data = smart_meter10
gran1 = "day_week"
gran2 = "hour_day"
response = "general_supply_kwh"
clust_var = "customer_id"
```

### 2 subjects 5 simulations

```{r ncust2_niter5}
conditional_dendro(.data, gran1,
                       gran2, response,
                       clust_var, niter = 5,
                       ncust = 2)
```

### 5 subjects 5 simulations

```{r ncust5_niter5}
conditional_dendro(.data, gran1,
                       gran2, response,
                       clust_var, niter = 5,
                       ncust = 5)
```

### 10 subjects 5 simulations

```{r ncust10_niter5}
conditional_dendro(.data, gran1,
                       gran2, response,
                       clust_var, niter = 5,
                       ncust = 10)
```

### 10 subjects only

```{r}
conditional_dendro(.data, gran1,
                       gran2, response,
                       clust_var, niter = 1,
                       ncust = 10)
```

# Visualization of clusters

## dendogram and heat maps
## Multidimensional scaling

Let’s say we were given only the distances between objects (i.e. their similarities) — and not their locations? You could still create a map — but it would involve a fair amount of geometry, and some logical deductions. Kruskal & Wish (1978) — the authors of one of the first multidimensional scaling books — state that this type of logic problem is ideal for multidimensional scaling. You’re basically given a set of differences, and the goal is to create a map that will also tell you what the original distances where and where they were located.

https://www.statisticshowto.com/multidimensional-scaling/

- multidimensional scaling  and how points in mds space maps to points in your original data space using gravitas 
- tourr, cmds, prcomp  
- read about ggobi, tsne (topology over geometry)  
- intuition - elliptical and spherical spheres only when we have distinct clusters in model based   clustering  
- Wald's linkage, single linkage and complete linkage
Read ggobi and Di's machine learning course to get intuition  
- heat map and log scale on the color to have a distinction and map it to the dendogram

<!-- #### -- start from paper-hakear -->

<!-- ## Smart meter data of Australia {#sec:smartmeter} -->

<!-- # how data looks -->

Clustering of electricity customers based on their consumption behavior facilitate effective market segmentation and hence better management and pricing.
Electricity demand time series are the primary source of information of customers' consumption behavior.
Due to technological developments, smart meters could provide large quantities of measurements on energy usage at more and more finer time intervals. Due to this extreme dimensionality along time of the demand data, it is a challenge to have clustering methods which could reduce the dimensionality to produce distinct clusters. One of the potential solutions is time series feature extraction. However, it is limited by the type of noisy, patchy and unequal time series that is very common in data sets of any residential customers. The idea is to choose similarity measures, which do not lose the core characteristic information of demand across different time deconstructions in aggregation processes and is robust against asynchronous or incomplete time series.

We use the data from one of the customer trials [@smart-meter] conducted as part of the Smart Grid Smart City project in Newcastle, New South Wales and some parts of Sydney. The data has customer wise data on energy consumption for every half hour from February 2012 to March 2014. <!--It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.--> The idea here is to
use our similarity measure and show improved cluster distinction as compared to baseline feature based approaches. A data set of 100 homes is used to evaluate both approaches. This is different to deterministic segmentation which attempts to characterise load as a function of a fixed number of parameters, such as occupancy or types of appliances, and use those parameters for classification. We only use their load patterns across time deconstructions to define the similarity measure.

<!-- Write about common load clustering methods -->

### Multidimensional scaling and cluster analysis

Cluster analysis and multidimensional scaling (MDS) methods are used to explore the temporal structure of various households, their interrelations and determine similar behaving clusters of households in this study.
MDS methods are used to produce a lower dimensional display space where each household is represented by a point and the distance between these points give the measure of similarity between these households. On the other hand, clustering techniques are used to provide information about the clustering structure of the households. Households that are in the same cluster can be considered to be more similar, while households are more dissimilar in their energy behavior across one or few temporal deconstructions.

$wpd_norm$ is considered as the measure of similarity or dissimilarity while using MDS and clustering techniques. $wpd_norm$ could be constructed for each pair of harmony for each household and then similarity could be measured across single or all harmonies together.

```{r mds}
set.seed(7777)
library(gravitas)
library(tsibble)
library(lubridate)
library(readr)

all_data <- read_rds(here("paper/sim_table/all_data.rds"))
sm_cust_data <- read_rds(here("paper/data/sm_cust_data.rds")) %>% 
  as_tsibble(index = reading_datetime,
             key = customer_id)

sm_unique_cust <- sm_cust_data %>% distinct(customer_id)

#sm_30 <- sample(sm_unique_cust$customer_id, 30)
sm_cust_samp <- sm_cust_data %>% 
  #filter(customer_id %in% sm_30) %>% # for demo show 30 customers
  filter(year(reading_datetime) == 2012)  %>% 
  create_gran("day_week") %>% 
  create_gran("hour_day")

#all_data <- write_rds(all_data, "paper/sim_table/all_data.rds")

facet_variable = "day_week"
x_variable = "hour_day"

one_harmony_data <- all_data %>%
  #filter(customer_id %in% sm_30) %>% # for demo show 30 customers
  filter(facet_variable == "day_week",
                    x_variable == "hour_day") %>% 
  mutate(customer_id = as.character(customer_id))

mds<- one_harmony_data %>%
  select(customer_id, facet_variable, x_variable, wpd_norm) %>% 
  dist() %>%          
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")
rownames(mds) <- unique(one_harmony_data$customer_id)
mds <- mds %>%
  mutate(customer_id = as.character(unique(one_harmony_data$customer_id))) %>%
  select((customer_id)
         , Dim.1, Dim.2)
```

```{r clustering}
# K-means clustering
clust <- kmeans(mds, 3)$cluster %>%
  as.factor()
mds <- mds %>%
  mutate(groups = clust)

# Plot and color by groups
q1 <- ggpubr::ggscatter(mds, x = "Dim.1", y = "Dim.2", 
          label = rownames(mds),
          color = "groups",
          palette = "jco",
          size = 1, 
          ellipse = TRUE,
          ellipse.type = "convex",
          repel = TRUE)

```

- check transformation and scalar approaches are same

```{r scalar-approach}
set.seed(7777)
all_data <- read_rds(here("paper/sim_table/all_data_scalar.rds"))
sm_cust_data <- read_rds(here("paper/data/sm_cust_data.rds"))
set.seed(7777)
sm_unique_cust <- sm_cust_data %>% distinct(customer_id)
sm_cust_samp <- sm_cust_data %>% 
  #filter(customer_id %in% sm_30) %>% # for demo show 30 customers
  #filter(year(reading_datetime) == 2012) %>% #for demo just show one column
  as_tsibble(index = reading_datetime,
             key = customer_id) %>% 
  create_gran("day_week") %>% 
  create_gran("hour_day")

#all_data <- write_rds(all_data, "paper/sim_table/all_data.rds")

facet_variable = "day_week"
x_variable = "hour_day"

one_harmony_data <- all_data %>%
  #filter(customer_id %in% sm_30) %>% # for demo show 30 customers
  filter(facet_variable == "day_week",
         x_variable == "hour_day") %>% 
  mutate(customer_id = as.character(customer_id))

mds<- one_harmony_data %>%
  select(customer_id, facet_variable, x_variable, wpd_norm) %>% 
  dist() %>%          
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")
rownames(mds) <- unique(one_harmony_data$customer_id)
mds <- mds %>%
  mutate(customer_id = as.character(unique(one_harmony_data$customer_id))) %>%
  select((customer_id)
         , Dim.1, Dim.2)

ggpubr::ggscatter(mds,
                  x = "Dim.1",
                  y = "Dim.2", 
                  label = rownames(mds),
                  size = 1,
                  repel = TRUE)

# clustering as well

d = one_harmony_data %>%
  select(customer_id, facet_variable, x_variable, wpd_norm) %>% 
  dist()

hc = stats::hclust(d,method="complete")
#plot(hc)
groups<-cutree(hc, k=3)

all_data_cluster <- cbind(one_harmony_data, groups) %>% 
  left_join(mds, by = "customer_id") %>% 
  mutate(groups = as.factor(groups))

q2 <- ggpubr::ggscatter(all_data_cluster,
                  x = "Dim.1",
                  y = "Dim.2", 
                  label = rownames(mds),
                  color = "groups",
                  size = 1,
                  repel = TRUE)

q1 + q2
```

10018272 is a customer from cluster 3, for which hour_day/wknd-wday comes after wknd-wday/day-month as opposed to 10006414 (cluster 1) for whom the direction is reversed.

```{r, echo = FALSE}

10006414

all_data %>% 
  filter(customer_id ==  10006414) %>% 
  arrange(-wpd_norm)

sm_cust_data %>% 
  filter(customer_id ==  10006414) %>% 
  prob_plot("hour_day",
            "wknd_wday",
            plot_type = "quantile",
            quantile_prob = c(0.5),
            symmetric = FALSE,
            response = "general_supply_kwh")

sm_cust_data %>% 
  filter(customer_id ==  10006414) %>% 
  prob_plot("hour_day",
            "day_week",
            plot_type = "quantile",
            quantile_prob = c(0.5),
            symmetric = FALSE,
            response = "general_supply_kwh")

all_data %>% 
  filter(customer_id ==  10007340) %>% 
  arrange(-wpd_norm)

sm_cust_data %>% 
  filter(customer_id ==  10007340) %>% 
  prob_plot("day_month",
            "day_week",
            plot_type = "quantile",
            quantile_prob = c(0.5),
            symmetric = FALSE,
            response = "general_supply_kwh")

sm_cust_data %>% 
  filter(customer_id ==  10007340) %>% 
  prob_plot("hour_day",
            "day_month",
            plot_type = "quantile",
            quantile_prob = c(0.5),
            symmetric = FALSE,
            response = "general_supply_kwh")

all_data %>% 
  filter(customer_id ==  10018272) %>% 
  arrange(-wpd_norm)

```

```{r}
all_data_scalar <- read_rds(here("paper/sim_table/all_data_scalar.rds"))
all_data <- read_rds(here("paper/sim_table/all_data.rds"))

merged_data <- left_join(all_data_scalar,all_data, by = c("customer_id",
                                                          "facet_variable",
                                                          "x_variable")) %>% 
  rename("wpd_scalar"  = "wpd_norm.x",
         "wpd_perm" = "wpd_norm.y")

merged_data %>% 
  mutate(pair = paste(facet_variable, x_variable, sep = "-")) %>% 
  ggplot() +
  geom_point(aes(x = wpd_scalar, y = wpd_perm, color = pair, group = pair))
  #geom_point(aes(x = wpd_scalar, y = wpd_perm, alpha = 0.05))

```

```{r linear}
raw_data <- sm_cust_data %>% 
  left_join(all_data_cluster, by = "customer_id")

p1 <- raw_data %>% 
  filter(groups==1) %>% 
  ggplot() + 
  geom_line(aes(x = reading_datetime,
            y = general_supply_kwh)
  ) + ggtitle("Group 1")

p2 <- raw_data %>% 
  filter(groups==2) %>% 
  ggplot() + 
  geom_line(aes(x = reading_datetime,
            y = general_supply_kwh)
  ) + ggtitle("Group 2")

p3 <- raw_data %>% 
  filter(groups==3) %>% 
  ggplot() + 
  geom_line(aes(x = reading_datetime,
            y = general_supply_kwh)
  ) + ggtitle("Group 3")
library(patchwork)
p1/p2/p3

```

```{r gravitas}
raw_data <- sm_cust_data %>% 
  left_join(all_data_cluster, by = "customer_id")

p1 <- raw_data %>% 
  filter(groups==1) %>% 
  prob_plot("day_week",
            "hour_day",
            response = "general_supply_kwh",
            plot_type = "quantile",
            symmetric = FALSE
)  + ggtitle("Group 1")

p2 <- raw_data %>% 
  filter(groups==2) %>% 
  prob_plot("day_week",
            "hour_day",
            response = "general_supply_kwh",
            plot_type = "quantile",
            symmetric = FALSE
)  + ggtitle("Group 2")

p3 <- raw_data %>% 
  filter(groups==3) %>% 
  prob_plot("day_week",
            "hour_day",
            response = "general_supply_kwh",
            plot_type = "quantile",
            symmetric = FALSE
)  + ggtitle("Group 3")

library(patchwork)
p1
p2
p3

```

```{r hour-day-day-month, eval = FALSE}

set.seed(7777)
library(gravitas)
library(tsibble)
library(lubridate)
library(readr)

all_data <- read_rds("sim_table/all_data.rds")
sm_cust_data <- read_rds("data/sm_cust_data.rds")
sm_unique_cust <- sm_cust_data %>% distinct(customer_id)

#sm_30 <- sample(sm_unique_cust$customer_id, 30)
sm_cust_samp <- sm_cust_data %>% 
  #filter(customer_id %in% sm_30) %>% # for demo show 30 customers
  filter(year(reading_datetime) == 2012) %>% #for demo just show one column
  as_tsibble(index = reading_datetime,
             key = customer_id) %>% 
  create_gran("wknd_wday") %>% 
  create_gran("week_month")

#all_data <- write_rds(all_data, "paper/sim_table/all_data.rds")

facet_variable = "wknd_wday"
x_variable = "week_month"

one_harmony_data <- all_data %>%
  #filter(customer_id %in% sm_30) %>% # for demo show 30 customers
  filter(facet_variable == "wknd_wday",
                    x_variable == "week_month") %>% 
  mutate(customer_id = as.character(customer_id))

mds<- one_harmony_data %>%
  select(customer_id, facet_variable, x_variable, wpd_norm) %>% 
  dist() %>%          
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")
rownames(mds) <- unique(one_harmony_data$customer_id)
mds <- mds %>%
  mutate(customer_id = as.character(unique(one_harmony_data$customer_id))) %>%
  select((customer_id)
         , Dim.1, Dim.2)

clust <- kmeans(mds, 3)$cluster %>%
  as.factor()
mds <- mds %>%
  mutate(groups = clust)

# Plot and color by groups
ggpubr::ggscatter(mds, x = "Dim.1", y = "Dim.2", 
          label = rownames(mds),
          color = "groups",
          palette = "jco",
          size = 1, 
          ellipse = TRUE,
          ellipse.type = "convex",
          repel = TRUE)

```

### Advantage over taditional feature based clustering methods

Since probability distribution of the load is considered across different cyclic granularity, there is no discarding of valuable data that might result from bringing all customers to the same time horizon.
Moreover, since our metric is based on probability distribution, it will not be sensitive to outlier days and hence the clustering process will not be biased.
Moreover, feature extraction applies to all days for all customers at once and therefore does not support parallel computing. Our metric is computed for each household separately and can even run for each each harmonies separately and thus fits well into the parallel computing framework.

### Write the method in notations

-  notation of time series for each customer
-  wpd
-  harmonies
-  cluster

### Distinction, repeatability, and robustness metrics

### Putting similar households on linear scale

<!-- ### end from paper-hakear -->

# How robust are your clusters

bootstrap and other resampling techniques

# Application
## Smart meter data

## Cricket data

# Conclusion

### 2021-10-20 old/interact_gran_4D.Rmd
cleaning folder

--- 2021-08-18 supplementary/algorithm1/real-data/cluster4-iter100.Rmd
@@
 library(knitr)
-read_chunk("R/simdata01.R")
+library(here)
+read_chunk(here::here("old/simdata01.R"))
 ```
@@
 nfacet_val = 3 # number of facet levels
-w1_val = 3 # increment in mean
+w1_val = 1 # increment in mean
 w2_val = 0 # increment in sd
@@
 ```{r nobs50, echo = TRUE}
-ntimes_val = 50 # nobs per combination 
+ntimes_val = 1000 # nobs per combination 
 ```
@@
 
+p_null/p_varf/p_varx/p_varall
 ```
@@
 ```{r, echo=TRUE}
-sample_seed <-  seq(1, 100, 1)
+sample_seed <-  seq(1, 5, 1)
 ```
@@
   set.seed(sample_seed[seed])
-  change_index_data(5, 2, 3, sim_null_normal) %>%
-    compute_quantiles(quantile_prob_val)
+  change_index_data(5, 2, 3, sim_null_normal) 
 }) %>%
@@
   set.seed(sample_seed[seed])
-  change_index_data(5, 2, 3, sim_varf_normal)%>%
-    compute_quantiles(quantile_prob_val)
+  change_index_data(5, 2, 3, sim_varf_normal)
 }) %>%
@@
   set.seed(sample_seed[seed])
-  change_index_data(5, 2, 3, sim_varx_normal)%>%
-    compute_quantiles(quantile_prob_val)
+  change_index_data(5, 2, 3, sim_varx_normal)
 }) %>%
@@
   set.seed(sample_seed[seed])
-  change_index_data(5, 2, 3, sim_varall_normal)%>%
-    compute_quantiles(quantile_prob_val)
+  change_index_data(5, 2, 3, sim_varall_normal)
 }) %>%
@@
                                       "data_varx","data_varall"))
+
+```
+
+```{r}
+data_clust <- data_q %>% tsibble::as_tsibble(key = unique_data, index = time)
+
+library(gracsr)
+data_clust %>% scale_gran(method="nqt", response = "sim_data")
+
+%>%
+  dist_gran(gran1="id_facet", gran2 = "id_x") %>% 
+  dist_gran(clust_gran(kopt=4)
 

### 2021-10-20 old/rough.Rmd
cleaning folder

--- 2021-07-07 main.Rmd
@@
 ---
-title: Clustering of residential Australian smart meter customers
-
-# to produce blinded version set to 1
-blinded: 0
-
-authors:
-- name: Sayani Gupta
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-  thanks: "Email: Sayani.Gupta@monash.edu"
-
-- name: Rob J Hyndman
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-
-- name: Dianne Cook
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-
-keywords:
-- data visualization
-- statistical distributions
-- time granularities
-- calendar algebra
-- periodicties
-- grammar of graphics
-- R
-
-abstract: |
- Clustering elements based on behavior across time granularities 
-bibliography: [bibliography.bib]
-preamble: >
-  \setlength {\marginparwidth }{2cm}
-  \usepackage{mathtools,amssymb,booktabs,longtable,todonotes,amsthm}
-  \def\mod{~\text{mod}~}
-output:
-  bookdown::pdf_book:
-    base_format: rticles::asa_article
-    fig_height: 4
-    fig_width: 6
-    fig_caption: yes
-    dev: "pdf" 
-    keep_tex: yes
+title: "rough"
+author: "Sayani Gupta"
+date: "08/06/2021"
+output: html_document
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
-# Make sure you have the latest version of rmarkdown and bookdown
-#devtools::install_github("rstudio/rmarkdown")
-#devtools::install_github("rstudio/bookdown")
-library(ggplot2)
-library(gravitas)
-library(gracsR)
-library(ggdendro)
-library(dplyr)
-library(readr)
-library(visdat)
-library(ggplot2)
+knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
 library(tidyverse)
-library(naniar)
-library(here)
-library(tsibble)
-library(knitr)
 ```
 
-```{r external, include = FALSE, cache = FALSE}
-knitr::read_chunk(here('script/smart_meter.R'))
-```
+# Clustering Australian Smart Grid Smart City data for grouping periodically behaving customers
 
-# Introduction
+## Introduction
 
@@
 
-\noindent The massive amount of data generated in such projects could be overwhelming for analysis.
+The massive amount of data generated in such projects could be overwhelming for analysis.
 Electricity utilities can utilize the consumption patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
@@
 
-# Preliminary exploratoration
-
-## Getting data
-
-<!-- _Electricity use interval reading data_   -->
-<!-- The data from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
--->
-
-The entire data is procured from CSIRO.
-A subset of this data is also available from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
-It consists of the following data sets.
-_1. CustomerData:_ 78720 customers with 62 variables about them 
-_2. EUDMData:_ 300 billion half-hourly consumption level data  
-_3. OffersData:_ Method of contact to customer to join SGSC customer trial, either door-to-door (D2D) or via Telesales  
-_4. PEResponseData:_ Peak Events response customer wise  
-_5. PETimesData:_ Peak Events time stamps
-
-Only _CustomerData_ and _EUDMData_ are relevant for the clustering goals of this paper.
-_EUDMData_ contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
-`CustomerData` provides demographic data for 78,720 customers with information on their Local Government Area amongst others.
-
-## Missing Data
-
-Electricity usage for some customers may become unavailable due to power outage or not recording their usage properly, thus resulting in implicit missing values in the database.
-It is interesting to explore where missing-ness occurs or if there is a relationship between the underlying missing patterns.
-We use the R package `tsibble` to do this.
-
-```{r, out.width="100%"}
-include_graphics("figs/missing-data.png")
-```
-
-## Raw data
-
-Since the smart meters have been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths
-for each household.
-Since, general supply is available for only 13,735 customers, we will restrict ourselves to look at the LGA information for these customers only.
-We find that there are only 26 LGA that is covered for these customers.
-
-_Weather data_  
-This data is obtained through [Australian Government Bureau of Meteorology](BOM)(http://www.bom.gov.au/) and provides hourly data for nearest weather stations for all the LGAs
+## Preliminary exploratoration
 
@@
 The data is at SA2 and LGA levels.
-However, some of the LGA in NSW changed between 2011 and 2016 and hence there would not be a one-to-one correspondence between the LGAs.
-Weather, notably temperature (and humidity) can be the main driver(s) for energy usage.
+However, some of the LGAs in NSW changed between 2011 and 2016 and hence there would not be a one-to-one correspondence between the LGAs Weather, notably temperature (and humidity) can be the main driver(s) for energy usage.
 In NSW many households have electric heaters so their use can impact winter energy use and air-conditioners can impact summer energy use.
@@
 
-\noindent A dataset of 100 SGSC homes has been used to lay out the structure to be used for analyzing the big dataset.
+A dataset of 100 SGSC homes has been used to lay out the structure to be used for analyzing the big dataset.
 The smaller dataset contains half-hourly kwh values form 2012 to 2014 and has asynchronous time series distributed evenly over the observation period (Figure \ref{fig:elec-raw}), similar to the bigger data set.
@@
 
-```{r miss-data}
-empty_as_na <- function(x){
-    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
-    ifelse(as.character(x)!="", x, NA)
-}
-
-```
-
-```{r vismiss, eval = FALSE}
-## transform all columns
-customer_data <- customer_data %>% mutate_each(funs(empty_as_na))
-
-vis_miss(customer_data, warn_large_data = FALSE) + 
-  theme(axis.text.x = element_text(angle = 90, size = 5))
-```
-
-```{r miss-var, eval = FALSE}
-gg_miss_var(customer_data) + theme_bw() +
-  theme(axis.text.y = element_text(size = 4))
-```
-
-```{r}
-elec_ts <- read_rds("data/elec_ts.rds")
-```
-
-```{r raw-50, fig.height = 8, fig.cap="The raw data for 50 households are shown.
+```{r elec-raw, fig.height = 8, fig.cap="The raw data for 50 households are shown.
 It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed."}
@@
 
-# Clustering methodology
+```{r elec-gaps, warning=FALSE}
+
+```
+
+```{r count-gaps, fig.height = 8, fig.cap="The missing data pattern for 100 households are shown. It looks like most missingness happens before 2013 and for a particular data in 2014."}
+
+```
 
@@
 So $f_s(v)$ is the distribution of household $s$ and $v$ is electricity demand.
-Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
-Each household consists of a $J*K$ distributions one harmony.
-We compute the distributional difference
-between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$.
-$wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies.
-If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
-We aim to have pockets of households showing similar periodic behavior by conidering $wpd$ vlaues for different harmonies and some time series features.
-The features should also characterize probability distributions of different household.
+Consider two cyclic granularities $A$ and $B$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
+We still want the households in clusters, but each household consists of a collection of probability distributions.
+For  example, we would have $J*K$ distributions per household.
+So we need a distance between collections of univariate probability distributions.
+One possibility is to consider the normalized weighted pairwise distance as the difference between various households for the variable $(A, B)$.
+Let $wpd_{norm_{s,t}}(A,B)$ denotes the normalized pairwise distances between subjects/elements $s$ and $t$ for the harmony pair $(A, B)$.
+$wpd_{norm_{s,t}}(A,B)$ could be thought to be a dimensionality reduction procedure in time series data mining.
+This feature, which is considered as a measure to determine the strength of any periodic pattern are then inputs to a clustering algorithm that can then discern similarity and differences between the time series.
 
-<!-- grouping probability distributions across a harmony.
+We employ a hierarchical cluster analysis algorithm to obtain pockets of households showing similar periodic behavior by grouping probability distributions across a harmony.
 This clustering algorithm is adopted to remove or appropriately adjust for auto correlation and unequal length in the data.
@@
 The following are some of the advantages of our proposed method.
--->
 
@@
 
-```{r linear-cluster, eval = FALSE}
+```{r linear-cluster}
 group1 <- all_data_cluster %>% 
@@
 
-## Cluster validation
-
-**Internal cluster validation** uses the internal information of the clustering process to evaluate the goodness of a clustering structure without reference to external information.
-It can be also used for estimating the number of clusters and the appropriate clustering algorithm without any external data.
-
-**External cluster validation** consists in comparing the results of a cluster analysis to an externally known result, such as externally provided class labels.
-It measures the extent to which cluster labels match externally supplied class labels.
-Since we know the “true” cluster number in advance, this approach is mainly used for selecting the right clustering algorithm for a specific data set.
-
-**Relative cluster validation** evaluates the clustering structure by varying different parameter values for the same algorithm (e.g.,: varying the number of clusters k).
-It’s generally used for determining the optimal number of clusters.
-
-Interval validation includes
-_Compactness or cluster cohesion:_ Measures how close are the objects within the same cluster.
-A lower within-cluster variation is an indicator of a good compactness (i.e., a good clustering).
-The different indices for evaluating the compactness of clusters are base on distance measures such as the cluster-wise within average/median distances between observations.
-
-_Separation:_ Measures how well-separated a cluster is from other clusters. The indices used as separation measures include:
-distances between cluster centers
-the pairwise minimum distances between objects in different clusters
-
-_Connectivity:_ corresponds to what extent items are placed in the same cluster as their nearest neighbors in the data space.
-The connectivity has a value between 0 and infinity and should be minimized.
-
-Generally most of the indices used for internal clustering validation combine compactness and separation measures.
-
 ## Combining findings with external data
@@
 
-<!-- <Earlier write up> -->
-<!-- ## Introduction -->
+```{r data_structure_clust, out.width = "100%", eval=FALSE}
+knitr::include_graphics("figures/data_structure_clust.png")
+```
 
-<!-- <!-- Description of data available --> -->
-<!-- The Smart Grid, Smart City (SGSC) project, which rolled out Australia's first commercial-scale smart grid was implemented across eight local government areas in New South Wales (NSW).
-Data from more than 13,000 household electricity smart meters is obtained as part of that project.
-It provides half-hourly energy usage and demographic data for Australia, as well as detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
-The trials were based in Newcastle, New South Wales, but also covered areas in Sydney CBD, Newington, Ku-Ring-Gai, and the rural township of Scone.
-The load time series is asynchronous as it is observed for these households for unequal time lengths and consists of missing observations.
--->
+<!-- - Study of Australian smart meter data -->
 
-<!-- The massive amount of data generated in such projects could be overwhelming for analysis.
-Electricity utilities can utilize the consumption patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
-Beyea (2010) has pointed out, there has been little discussion or exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data.
-Thus, there is a scope to investigate and analyze these data in various ways for a greater understanding of consumption patterns and how they correlate with other economic, physical or geographical factors.
-In this work, we are interested to see how we can utilize this dataset to group different customers with similar periodic behavior.
-Towards this goal, this chapter aims to: (a) describe the contents of the data set in SGSC database that we can utilize, and (b) propose a clustering algorithm to group customers with similar periodic behaviors.
-The distance metric introduced in Chapter 2 will be the inputs for this cluster analysis.
-One of the advantages of using our  -->
-<!-- approach is that the technique is based on probability distributions instead of raw data.
-Many clustering approaches are limited by the type of noisy, patchy, and unequal time-series common in residential data sets.
-Since the distance measure considered is based on differences in probability distribution of time series, it is likely to be less sensitive to missing or noisy data.
--->
-
-<!-- The following contributions are made through the following chapter: -->
-
-<!--  * Present a cluster analysis of SGSC dataset to group households with similar periodic behavior -->
-<!--  * Cluster validation by relating to external data -->
-
-<!-- ## Introduction -->
-
-<!-- In Australia, the federal government and fourteen partners, electricity distribution and transmission companies, technology companies, universities and CSIRO, invested AUD 490 million in the Smart Grid Smart City (SGSC) field trial.
-One of the largest commercial-scale smart grid technology assessment projects worldwide, SGSC is implemented across eight local government areas in New South Wales (NSW), covering 30,000 dwellings since 2009.
--->
-<!-- The SGSC database is publicly available through the Information Clearing House (ICH) (Ausgrid 2014).
-It is a massive database covering most physical aspects of the smart grid, as well as, aspects of users’ demographics.
--->
-
-<!-- Customer Sale and Half-Hour Power Consumption -->
-<!-- and Generation Datasets -->
-
-<!-- This is a big portion of the data from about 5900 households who accepted a -->
-<!-- product out of about 24,000 offers made to the target population. Spatial resolution -->
-<!-- down to anonymous individual dwellings [Customer Sales: Households: -->
-<!-- Customer_Key] is also given based on {[Postcode], [Local Government Area], or -->
-<!-- [Suburb Name]}. In order to make these datasets easily understandable, here we -->
-<!-- present a new narrative approach to explain the available data fields and their -->
-<!-- interdependencies in a brief yet logical and memorisable way as follows: -->
-
-<!-- The [Half-Hour Consumption and Generation] data provide kWh readings of -->
-<!-- the actual flow of energy. This includes the electricity supply imported from the -->
-<!-- grid [General Supply] which is the main source of energy under any of the applied -->
-<!-- tariffs, import during off-peak or for a control load [Off Peak], gross or net values of -->
-
-<!-- domestic (often solar) electricity that is exported back to the grid [Generation]. The -->
-<!-- [Peak Events] data field gives the details of every event [Event Key] including its -->
-<!-- [Type], [Date], [Start Time], and [End Time]. Accordingly, the entire sequence of -->
-<!-- user demand responses to any of the cost-reflective products is available in the -->
-<!-- dataset to evaluate the project objectives as previously discussed. -->
-
+<!-- <!-- Motivation and background --> 
 <!-- Electricity smart meter technology is increasingly being deployed in residential and commercial buildings.
@@
 
-<!-- <!-- Description of data available --> -->
+<!-- <!-- Description of data available --> 
 <!-- Data from more than 13,000 household electricity smart meters is obtained as part of the Smart Grid Smart City (SGSC) project (2010-2014).
@@
 
-<!-- <!-- Background literature and shortcomings --> -->
+<!-- <!-- Background literature and shortcomings --> 
 <!-- The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics.
@@
 However, it does not explain the cluster composition by combining it with external data.
-We propose the following for the study of the Australian smart meter data.
 -->
 
-<!-- Analysis for more than 13,000 household -->
-<!-- electricity smart meters are included in this chapter. The following contributions are made through the following chapter: -->
+<!-- We propose the following for the study of the Australian smart meter data: -->
 
-<!-- - Presenting a cluster analysis of Australian residential electricity consumption data. -->
-
-<!-- - Transformation and extraction of input data features enabling K-Means to account for other time series features in the clustering.
-This can easily be extended to include other data structures.
+<!--  - *Preliminary exploratory visualization and summarisation*: Be familiar with an overall view of these 13,000 SGSC households through visualization and provide a detailed layout of data structures (like missing observations) and also the external data that needs to be utilized in the clustering phase.
 -->
 
-<!-- - Cluster validation by relating to external data -->
-
-<!-- ## Preliminary exploratoration -->
-
-<!-- This section familiarizes the 13,000 SGSC households through visualization and provide a detailed layout of data structures (like missing observations) and also the external data that needs to be utilized for validating the clustering process.
--->
-
-<!-- ## Common load clustering techniques of smart meter data -->
-
-<!-- This section will discuss some literature on clustering techniques that were applied to smart meter customers in the past. -->
-
-<!-- ## The proposed load-clustering technique -->
-
-<!-- We employ a cluster analysis technique to obtain pockets of households showing similar periodic behavior by grouping probability distributions across a cyclic granularity.
+<!--  - *Cluster analysis*: We employ a cluster analysis technique to obtain pockets of households showing similar periodic behavior by grouping probability distributions across a cyclic granularity.
 This clustering algorithm is adopted to remove or appropriately adjust for auto correlation and unequal length in the data.
@@
 Further, we elaborate on the effect of missing data and how it is handled in the clustering process.
--->
-
-<!-- ### Methodology, results, and discussions -->
-
-<!-- This will contain the data structure and detailed methodology to be employed for the cluster analysis. -->
-
-<!-- ### Distinction, repeatability, and robustness metrics -->
-
-<!-- This section will cover the homogeneity and heterogeneity of the clusters, how distinct they are and summarize the main features of the cluster. -->
-
-<!-- <!-- ## Comparison with traditional load clustering methods --> -->
-
-<!-- <!-- This section will cover the homogeneity and heterogeneity of the clusters, how distinct they are and summarize the main features of the cluster. --> -->
-
-<!-- ## Combining findings with external data -->
-
-<!-- We will also determine if any identified clusters or patterns are indeed statistically meaningful in the sense that they actually exist and are not a random allocation.
+We will also determine if any identified clusters or patterns are indeed statistically meaningful in the sense that they actually exist and are not a random allocation.
 We aim to do this by providing a practical explanation of the formed clusters, visualizing how they relate to any weather, socio-economic or geographical conditions.
@@
 -->
-<!-- We will present how matching other data sources might help us in validating our clustering approach. -->
 
-<!-- ### Geography -->
+<!-- NOT USED IN MID_CAN -->
 
-<!-- We will present how mapping the customers to their respective LGA's and reflect on the clusters obtained in the earlier section. -->
+<!-- ## Clustering households with similar periodic behavior -->
 
-<!-- ### Census -->
+<!-- The pool of all customers’ load time series will have different periods of time with overlaps.
+Any feature based clustering approach would trim the load curves to obtain equal length and synchronized date and time for all curves.
+These approaches are limited by the type of noisy, patchy, and unequal time-series common in residential data sets (@).
+In this project, we explore a clustering strategy to alleviate these limitations by looking at the distribution of load time series across different non-linear time granularities that can be readily clustered.
+Clustering could be employed for distribution of any harmonies (or all non-linear granularities) taken together to form clusters with similar periodic behavior across these harmonies (or all non-linear granularities).
+@ suggested carrying out cross validation by bootstrapping to form same clusters of homes on randomly selected days within a certain period.
+The strategy employed here when complimented with some cross validation would be a way to form same clusters of homes who show similar periodic behavior on randomly selected harmonies within a certain period.
+-->
 
-<!-- We will present how mapping the customers to their respective income and population level reflect on the clusters obtained in the earlier section. -->
-
-<!-- ## Conclusion -->
-
-<!-- This section will cover some drawback of this clustering method and potential extensions of this work. -->
-
-<!-- There are many cluster analysis methods that can produce quite different clusterings on the same dataset.
-Cluster validation is about the evaluation of the quality of a clustering; “relative cluster validation” is about using such criteria to compare clusterings.
-This can be used to select one of a set of clusterings from different methods, or from the same method ran with different parameters such as different numbers of clusters.
-There are many cluster validation indexes in the literature.
-Most of them attempt to measure the overall quality of a clustering by a single number, but this can be inappropriate.
-There are various different characteristics of a clustering that can be relevant in practice, depending on the aim of clustering, such as low within-cluster distances and high between-cluster separation.
-In this paper, a number of validation criteria will be introduced that refer to different desirable characteristics of a clustering, and that characterise a clustering in a multidimensional way.
-In specific applications the user may be interested in some of these criteria rather than others.
-A focus of the paper is on methodology to standardise the different characteristics so that users can aggregate them in a suitable way specifying weights for the various criteria that are relevant in the clustering application at hand.
--->
+# Timeline

### 2021-10-21 check_work_Di.Rmd
abstract added

--- 2021-10-18 check_work.Rmd
@@
 
-data_pick_one <- nn_pick_function(anchor1)
-data_pick_two <- nn_pick_function(anchor2)
-data_pick_three <- nn_pick_function(anchor3)
-data_pick_four <- nn_pick_function(anchor4)
-
 # c(11013154, 9345642, 8328008, 8636035,8627007,8454235)
 
+data_pick_one <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble
+data_pick_two <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble
+data_pick_three <- c(8221762, 8273636, 10359424, 8232822)%>% as_tibble
+data_pick_four <- c(10590714,8495194,8589936, 8454235) %>% as_tibble
+
 data_pick_cust <- bind_rows(
-data_pick_one, data_pick_two, data_pick_three, data_pick_four,
+data_pick_one, data_pick_two, data_pick_three,data_pick_four,
 .id = "design")
@@
 
-```{r mds-plots}
-
-data_pick_char <- data_pick_cust %>% 
-  rename("customer_id" = "value") %>% 
-  mutate(customer_id = as.character(customer_id))
-
-mds_all <- read_rds(here("data/mds_all.rds"))  %>% 
-  left_join(data_pick_char)
-
-highlight_df <- mds_all %>% filter(customer_id %in% data_pick_char$customer_id)
-
-total_plot <- ggplot(mds_all,
-aes(x = mds1.total, y = mds2.total)) +
-  geom_point(alpha = 0.3) +
-  geom_point(data = highlight_df, 
-             aes(x = mds1.total, y = mds2.total, color = as.factor(design)), size = 2)
-
-wpd_plot <- ggplot(mds_all,
-aes(x = mds1.wpd, y = mds2.wpd)) +
-  geom_point(alpha = 0.3) +
-  geom_point(data = highlight_df, 
-             aes(x = mds1.wpd, y = mds2.wpd, color = as.factor(design)), size = 2)
-
-hod_plot <- ggplot(mds_all,
-aes(x = mds1.hod, y = mds2.hod)) +
-  geom_point(alpha = 0.3) +
-  geom_point(data = highlight_df, 
-             aes(x = mds1.hod, y = mds2.hod, color = as.factor(design)), size = 2)
-
-moy_plot <- ggplot(mds_all,
-aes(x = mds1.moy, y = mds2.moy)) +
-  geom_point(alpha = 0.3) +
-  geom_point(data = highlight_df, 
-             aes(x = mds1.moy, y = mds2.moy, color = as.factor(design)), size = 2)
-
-wkndwday_plot <- ggplot(mds_all,
-aes(x = mds1.wkndwday, y = mds2.wkndwday)) +
-  geom_point(alpha = 0.3) +
-  geom_point(data = highlight_df, 
-             aes(x =  mds1.wkndwday, 
-                 y =  mds2.wkndwday, 
-                 color = as.factor(design)),
-             size = 2)
-
-p <- (total_plot + wpd_plot)/(hod_plot + moy_plot + wkndwday_plot)
-
-p* +
-  plot_layout(guides = "collect")
-
-```
-
 ```{r clustering}
@@
 total_distance <- filtered_distance %>% 
-mutate(total = wkndwday/2 + moy/12 + hod/24) 
+mutate(total = wkndwday + moy + hod) 
 #mutate(total = hod)
@@
                   ymax = `75%`,
-                  group=customer_id,
-                  fill = as.factor(group)),
+                  group=customer_id, fill = as.factor(group)),
               alpha = 0.5) +
   geom_line(aes(y = `50%`,
-                group=customer_id, 
-                color = as.factor(group)), size = 1) +
+                group=customer_id, color = as.factor(group)), size = 1) +
   facet_wrap(design~customer_id, 
@@
              labeller = "label_value",
-             ncol = 6) + theme_characterisation() +
+             ncol = 6) +
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
@@
                   ymax = `75%`, 
-                  group=customer_id, 
-              fill = as.factor(group)), alpha = 0.5) +
+                  group=customer_id, fill = as.factor(group)), alpha = 0.5) +
   geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
@@
     xlab("month-of-year")  +
-  theme_characterisation() +
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
@@
                   ymax = `75%`, 
-                  group=customer_id, 
-              fill = as.factor(group)), alpha = 0.5) +
-  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
+                  group=customer_id, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group))) +
   facet_wrap(design~customer_id, 
@@
              ncol = 6) +
- theme_characterisation() +
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
@@
              nrow = 4)  +
-             theme_characterisation() +
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
@@
   ylab("demand (in Kwh)")  +
-  theme_characterisation() +
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
@@
              ncol = 6)  +
-   theme_characterisation() +
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
@@
              nrow = 4)   +
-  theme_characterisation() +
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+

### 2021-10-21 main.Rmd
cleaning folder
abstract added

@@
 ---
-title:  Clustering based on probability distributions with application on residential customers
+title:  Clustering time series based on probability distributions across temporal granularities
 
@@
 - calendar algebra
-- periodicties
+- periodic data
 - grammar of graphics
@@
 abstract: |
- Clustering elements based on behavior across time granularities 
+ With more and more time series data being collected at much finer temporal resolution, for a longer length of time, and for a larger number of individuals/entities, time series clustering research is getting a lot of traction.
+The sort of noisy, patchy, uneven, and asynchronous time series that is typical in many disciplines limits similarity searches among these lengthy time series.
+In this work, we suggest a method for overcoming these constraints by grouping time series based on probability distributions over cyclic temporal granularities.
+Cyclic granularities are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, and so on, and can be helpful for detecting repeating patterns.
+Looking at probability distributions across cyclic granularities results in an approach that is robust to missing or noisy data, aids in dimension reduction, and ensures small pockets of similar repeated behaviours.
+The proposed method was tested using a collection of residential electricity customers.
+The simulated and empirical evidence demonstrates that our method is capable of producing meaningful clusters.
+
 bibliography: [bibliography.bib]
@@
   left_join(cluster_result, by = c("customer_id" = "id"))
+
+data_group %>% filter(group==1) %>% 
+  prob_plot("hour_day",
+            "month_year", 
+            response = "general_supply_kwh", plot_type = "quantile")
 

### 2021-10-26 04_handpick_from356-mds_nqt.Rmd
working on introduction again

@@
 
-wkndwday <- read_rds("../../../data/dist_gran_wkndwday_356cust_robust.rds") %>% broom::tidy()
+wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_nqt.rds")) %>% broom::tidy()
 
-moy <- read_rds("../../../data/dist_gran_moy_356cust.rds") %>% broom::tidy()
+moy <- read_rds(here("data/dist_gran_moy_356cust_nqt.rds")) %>% broom::tidy()
 
-hod <- read_rds("../../../data/dist_gran_hod_356cust_robust.rds") %>% broom::tidy()
+hod <- read_rds(here("data/dist_gran_hod_356cust_nqt.rds")) %>% broom::tidy()
 
@@
   select(customer_id, everything())
+
+write_rds(mds_all, "data/mds_all_nqt.rds")
 

### 2021-10-26 anchors3.Rmd
working on introduction again

---
title: "Plots with 3 anchors"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(ggplot2)
library(gravitas)
library(gracsr)
library(ggdendro)
library(dplyr)
library(readr)
library(visdat)
library(ggplot2)
library(tidyverse)
library(naniar)
library(here)
library(tsibble)
library(knitr)
library(patchwork)
library(GGally)
library(distributional)
library(viridis)
```

```{r mytheme}
theme_characterisation <- function() {

  theme_bw() + # seeting theme
    theme(strip.text = element_text(size = 9,
                                    margin = margin(b = 0, t = 0))) + # narrow facet space
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) + # no axis ticks 
  theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
    theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
    theme(legend.position = "bottom")+
  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
  theme(axis.text.x = element_text(size=5))
}
```

_Set number of anchors_
```{r anchor}
kopt <- 3 # number of clusters
```

_Load the customers you want to cluster_
```{r}
quantile_prob_graph <- c(0.25, 0.5, 0.75)

data_pick_one <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
data_pick_two <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
data_pick_three <- c(8221762, 8273636, 10359424, 8232822)%>% as_tibble %>% set_names("customer_id")

```

```{r assemble}
data_pick_cust <- bind_rows(
data_pick_one, data_pick_two, data_pick_three,
.id = "design") %>% 
  mutate(customer_id = as.character(customer_id))

```

_Load raw data for these customers and scale it according to robust method_

```{r}
data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
  mutate(customer_id = as.character(customer_id)) %>% 
  filter(customer_id %in% data_pick_cust$customer_id) %>% 
  gracsr::scale_gran( method = "robust",
                      response = "general_supply_kwh")
```

```{r clustering}

hod <- data_pick %>% 
  dist_gran(gran1 = "hour_day", response = "general_supply_kwh")

moy <- data_pick %>% 
  dist_gran(gran1 = "month_year", response = "general_supply_kwh")

wkndwday <- data_pick %>% 
  dist_gran(gran1 = "wknd_wday", response = "general_supply_kwh")

distance <- wkndwday/2 + moy/12 + hod/24

f = as.dist(distance)

cluster_result <- f %>% clust_gran(kopt = kopt) %>% 
  rename("customer_id" = "id")
```

```{r hod-data}
data_hod <- quantile_gran(data_pick,
                                  "hour_day",
                                  quantile_prob_val = quantile_prob_graph) %>% 
  pivot_wider(names_from = quantiles,
              values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id")) %>% 
  left_join(cluster_result, by = c("customer_id"))

#data_heatmap_hod$customer_id = factor(data_heatmap_hod$customer_id, levels = data_pick_cust$value)
data_hod$category <- factor(data_hod$category, levels = 0:23)

# lev <- data_hod %>% distinct(customer_id, design) %>% arrange(design) %>% pull(customer_id)
```

```{r hod-ind-design}

# data_hod$customer_id <- factor(data_hod$customer_id, levels = lev)

hod_ind_design <- data_hod %>% 
  left_join(cluster_result,by = c("customer_id", "group")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=customer_id, 
                  fill = design),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=customer_id,
                color = design), 
            size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y",
             nrow = kopt) + theme_characterisation()+
  xlab("hour-of-day") +
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  scale_x_discrete(breaks = seq(0, 23, 3))
```

```{r hod-ind-group}
hod_ind_group <- data_hod %>% 
  left_join(cluster_result) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=customer_id, fill = as.factor(group)),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y",
              nrow = kopt) + theme_characterisation()+
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +    xlab("hour-of-day")  +
  scale_x_discrete(breaks = seq(0, 23, 3))
# + scale_y_continuous(breaks = NULL) 
# 
# 
hod_ind_design + hod_ind_group
```

```{r data-heatmap-moy-design}

data_moy <- quantile_gran(data_pick,
                                  "month_year", 
                                  quantile_prob_val = quantile_prob_graph) %>% 
  pivot_wider(names_from = quantiles, 
              values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"))

data_moy$category <- factor(data_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

moy_ind_design <- data_moy %>% 
  left_join(cluster_result, by = c("customer_id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, 
                  group=customer_id, fill = as.factor(design)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(design)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             nrow = kopt) +
    ylab("demand (in Kwh)") +
    xlab("month-of-year")  +
  theme_characterisation() +
  scale_fill_viridis_d(direction = 1)+
  scale_color_viridis_d(direction = 1)
```

```{r data-heatmap-moy-group-ind}

moy_ind_group <- data_moy %>% 
  left_join(cluster_result, by = c("customer_id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, 
                  group=customer_id, fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             nrow = kopt) +
    ylab("demand (in Kwh)") +
    xlab("month-of-year")  +
  theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))

moy_ind_design + moy_ind_group

```

```{r data-heatmap-wkndwday-group}

data_wkndwday <- data_pick  %>%
  left_join(cluster_result)%>% create_gran("wknd_wday")  %>% 
  left_join(data_pick_cust, by = c("customer_id"))

ylim1 = boxplot.stats(data_wkndwday$general_supply_kwh)$stats[c(1, 5)]

wkndwday_ind_design <- data_wkndwday%>% 
  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
  lvplot::geom_lv(aes(fill = as.factor(design), 
                      color = as.factor(design)), k=5, alpha = 0.5) +
  #geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
  #geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = ylim1*1.05)+
  facet_wrap(~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             nrow = kopt)  +
  theme_characterisation() +
  scale_fill_viridis_d(direction = 1)+
  scale_color_viridis_d(direction = 1)

```

```{r}
wkndwday_ind_group <- data_wkndwday%>% 
  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
  lvplot::geom_lv(aes(fill = as.factor(group), 
                      color = as.factor(group)), k=5, alpha = 0.5) +
  #geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
  #geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = ylim1*1.05)+
  facet_wrap(~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             nrow = kopt)  +
  theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))

wkndwday_ind_design + wkndwday_ind_group
```

### 2021-10-26 main.Rmd
working on introduction again

@@
 
+<!-- time series clustering and its challenges -->
+Time-series clustering is the process of unsupervised partitioning of $n$ time-series data into $k$ ($k<<n$) groups such that homogeneous time-series are grouped together based on a certain similarity measure.
+The appropriate similarity measure or the distance approach is determined by the time-series characteristics, duration of time-series, representation method, and, of course, the goal of clustering time-series to a significant level.
+The three main approaches to time series clustering (@liao2005clustering) are algorithms that work directly with distances or raw data points, either in the time or frequency domain (distance-based), with features extracted from raw data (feature-based), or indirectly with models built from the raw data (model-based).
+The performance of distance-based approaches depends greatly on the particular distance metric used.
+Defining an adequate distance measure for the raw time series can be a challenging task since it has to consider noise, different length of time series, asynchronous time series, different scales, missing values.
+Distance-based similarity measures described in the literature include
+Euclidean,  Pearson’s correlation coefficient and related distances [124], Dynamic Time Warping [118,119], Autocorrelation, Short timeseries distance [44], Piecewise normalization [125], cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances.
+Euclidean distance and DTW are the most common methods for similarity measure in time-series clustering.
+A research has shown that, in terms of time-series classification accuracy, the Euclidean distance is surprisingly competitive [145], however, DTW also has its strength in similarity measurements which cannot be declined.
+While these methods have been quite effective in clustering moderate-length time series, there are some well-recognized drawbacks of the existing approaches.
+Euclidean distance is the most common method for discerning similarity in time series clustering, and it requires that the time series being compared are of exactly the same dimensionality (length).
+DTW can assist with clustering of different length time series if there is no missing data (Ratanamahatana and Keogh, 2005) and in any case it requires quadratic computation.
+
+We are motivated by the residential smart meter data.
+These long time series are asynchronous, with varying time lengths for different houses and missing observations and characterised by noisy and patchy behavior that can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
+Choosing probability distributions instead of raw data seems to be a natural way to analyze these types of data sets.
+Hence this paper proposes a distance metric based on Jensen-Shannon distances between probability distributions across significant cyclic granularities.
+Cyclic temporal granularities, which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
+Since cyclic granularities are considered instead of linear granularities, the resulting clusters are expected to group customers that have similar repetitive behaviors.
+Below we list some of the advantages of our approach, which will be explained in more detail along with empirical results in later sections.
+
+- When the time series is long (high dimensionality), some clustering algorithms become intractable.
+By applying dimensionality reduction via characterising probability distributions, we are able to cluster long length time series efficiently;
+
+- When the clustering algorithm is based on a distance metric (i.e.
+Euclidean distance), it cannot handle time series with missing data or of different lengths if actual points are used as inputs.
+However, by
+using probability distribution, we can trim the tails and bypass that problem.
+
+- We recognize that most clustering algorithms provide profiles across only one temporal granularity, but this approach provides a wider approach to the issue, seeking to group observations with similar shapes over all important cyclic granularities.
+Since cyclic granularities are considered instead of linear granularities, clustering would group customers that have similar repetitive behavior across more than one cyclic granularities across which patterns are expected to be significant.
+
+- It is reasonable to define a time series according to its degree of trend and seasonality and consider these features while clustering them.
+These characteristics of the time series are lost or handled independently by considering probability distributions (trend is lost) across categories of cyclic granularities (by independently modeling all seasonal fluctuations), and so there is no need to de-trend or de-seasonalize the data before conducting the clustering method.
+For example, there is no need to omit holiday or weekend patterns for similar reasons.
+
+<!-- . Often the objective is finding similarities in time (Fourier transforms, Piecewise aggregate approximation), in  shape (Dynamic time Warping (DTW)), in structure, based on global features. -->
+
+<!-- However, the process of determining a similarity measure is challenging since time-series data are naturally noisy and include outliers and missing observations, and the length of time-series changes and can be asynchronous.
+-->
+
+<!-- Time-series clustering is the most-used approach as an exploratory technique  for discovery of interesting patterns in time-series datasets.
+However, the process of finding similarity measure is complicated, because time-series data are naturally noisy and include outliers and shifts [18], at the other hand the length of time-series varies and the distance among them needs to be calculated.
+These common issues have made the -->
+<!-- similarity measure a major challenge for data miners. -->
+
+_Background and motivation_
+
 <!-- Description of data available -->
@@
 
-<!-- This work -->
-This is similar to a stochastic approach (@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time-series characteristics, or a model of the series, to enable unsupervised clustering.
-Unsupervised clustering is only as good as the features that are extracted/selected or the distance metrics that were utilized.
-Well-designed additional features may collect characteristics that default features cannot.
-Based on the underlying structure of the temporal data, this article offers new distance metric and features for clustering and applies them to actual smart-meter data.
-Firstly, the distance metric is based on probability distribution, which in our knowledge is the first attempt to cluster smart meter data using probability distributions.
-These recorded time series are asynchronous, with varying time lengths for different houses and missing observations.
-Taking probability distributions helps to deal with such data, while helping with dimension reduction in one hand but not losing too much information due to aggregation.
-Secondly, we recognise that most clustering algorithms only provide hourly energy profiles during the day, but this approach provides a wider approach to the issue, seeking to group consumers with similar shapes over all important cyclic granularities.
-Since cyclic granularities are considered instead of linear granularities, clustering would group customers that have similar repetitive behavior across more than one cyclic granularities across which patterns are expected to be significant.
-
-<!-- lit review -->
 _Related work_
@@
 
-<!-- Most clustering algorithms offer only daily energy profiles throughout the hours of the day, but we suggest a broader approach to the problem, aiming to group consumers with similar shapes across all significant cyclic granularities -->
-
-<!-- There are two reasons why any additional consumer level data would not be available.
-Customer profiles, like property features, vary over time and are not always updated in a timely manner by energy suppliers.
-Also such data might not be available when the anonymity of the customers are of utmost importance.
--->
-
-<!-- Segmenting unlabeled time series patterns has received little attention in the literature.
-This is a situation when there is no additional information on the socio-economic charactertistics, property types or family size of the customers.
-There are two reasons why any additional consumer level data would not be available.
-Customer profiles, like property features, vary over time and are not always updated in a timely manner by energy suppliers.
-Also such data might not be available when the anonymity of the customers are of utmost importance.
-So this study does not explain why consumption differs.
-Instead, this work investigates how much energy usage heterogeneity can be found in smart meter data and what some of the most common electricity use patterns are.
--->
-
-<!-- These recorded time series are asynchronous, with varying time lengths for different houses and missing observations. -->
-
-<!-- An aggregation approach to reduce the data size is not exactly ideal when dealing with highly heterogeneous profiles -->
-
-<!-- lit review and limitations if any -->
-<!-- _Common load clustering techniques of smart meter data_ -->
-
-<!-- A lot of work has been done in the literature on clustering customers by their load characeristics.
-Customers should be clustered uniquely, with behaviors represented with distinctive profile patterns.
-Approaches could be deterministic or stochastic.
-Deterministic approaches are hard to generalize due to the need for detailed knowledge of occupancy, appliances, and their time of use.
-In contrast, stochastic load analysis does not require extensive data collection other than the time series itself [13,14].
-This technique -->
-<!-- suggests viewing load as a random process, where extracting time-series features, or a model of the series, facilitates unsupervised clustering.
-Any standard clustering methodology consists of the following steps: (a) defining distance (dissimilarity) and similarity mostly through feature extraction and selection; and (b) choosing the clustering algorithm design.
--->
-<!-- Popular distance measures could be time-domain based or frequency-domain (fast Fourier transform).
-Any model based extraction works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterised and clustered.
-A type of clustering based on information theory such as Shannon or Renyi entropy and their variants are addressed in [42].In contrast to models, a feature-based extraction is used to explicitly define or automatically extract the curves’ key time features, for instance by application of PCA on the daily curves.
-@Tureczek2017-pb presented a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
-None of the $34$ selected papers which focus on clustering consumption are based on Australian smart meter data.
--->
-<!-- The algorithm most ubiquitously employed is K-Means.
-But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters.
-Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results.
-Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf).
-Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.
--->
-
-<!-- Time-domain analysis is preferred to specify time characteristics, while methods such as fast Fourier transform are used to reveal frequencies that highlight recurrent daily and annual load variations [27].
-Popular time-domain techniques, such as principal component analysis -->
-<!-- (PCA), detect key features of daily curves and reasonably link them to -->
-<!-- distinctive demand behaviours [28,29]. For instance, PCA can help -->
-<!-- reduce diurnal dimensions, e.g., 48 half-hourly power measurements, down to a few orthogonal components.
-@Tureczek2017-pb presented a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
-None of the $34$ selected papers which focus on clustering consumption are based on Australian smart meter data.
--->
-<!-- The algorithm most ubiquitously employed is K-Means.
-But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters.
-Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results.
-Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf).
-Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.
--->
-
-<!-- The clustering is frequently applied directly to the raw data without scrutinizing for auto correlation and periodicity. -->
-
-<!-- The key challenge is the application of appropriate processes to reduce the extreme dimensionality of load time series to facilitate unique clusters.
-Time feature extraction is a potential remedy, however, it is limited by the type of noisy, patchy, and unequal time-series common in residential datasets.
--->
-
-<!-- The literature -->
-<!-- suggests the clustering of residential customers by their load characteristics. The key challenge is the application -->
-<!-- of appropriate processes to reduce the extreme dimensionality of load time series to facilitate unique clusters. -->
-<!-- Time feature extraction is a potential remedy, however, it is limited by the type of noisy, patchy, and unequal -->
-<!-- time-series common in residential datasets. In this paper we propose a strategy to alleviate these limitations by -->
-<!-- converting any types of load time series into map models that can be readily clustered. -->
-
-<!--Unlabelled data -->
-<!-- While forecasting techniques for extremely granular temporal data are readily accessible, -->
-
-<!-- The basic idea behind the clustering algorithm based on distribution is that the data, generated from the same distribution, belongs to the same cluster if there exists several distributions in the original data.
--->
-
-<!-- Segmenting unlabeled time series patterns has received little attention in the literature.
-This is a situation when there is no additional information on the socio-economic charactertistics, property types or family size of the customers.
-There are two reasons why any additional consumer level data would not be available.
-Customer profiles, like property features, vary over time and are not always updated in a timely manner by energy suppliers.
-Also such data might not be available when the anonymity of the customers are of utmost importance.
-So this study does not explain why consumption differs.
-Instead, this work investigates how much energy usage heterogeneity can be found in smart meter data and what some of the most common electricity use patterns are.
--->
-
-<!-- lit review and limitations if any -->
-<!-- _Common load clustering techniques of smart meter data_ -->
-
-<!-- An aggregation approach to reduce the data size is not exactly ideal when dealing with highly heterogeneous profiles, as (XXX big data to the rescue example) demonstrates. -->
-<!-- The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings.
-None of the 34 selected papers which focus on clustering consumption are based on Australian smart meter data.
-The clustering is frequently applied directly to the raw data without scrutinizing for auto correlation and periodicity.
-The algorithm most ubiquitously employed is K-Means.
-But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters.
-Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results.
-@Tureczek2018-ha has shown that a transformation of data to incorporate autocorrelation before K-Means clustering can improve performance and enable K-Means to deliver smaller clusters with less within-cluster variance.
-However, it does not explain the cluster composition by combining it with external data.
-Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf).
-Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.
-As the methods are often situation specific,  -->
-<!-- it makes sense to compare them on the performance rather than any standard performance metric.
-A type of clustering based on information theory such as Shannon or Renyi entropy and their variants are addressed in , which differs from typical methods adopted for electricity consumer classification, based on the Euclidean distance notion.
-@Motlagh2019-yj presents strategy to address the problems on patchy, and unequal time-series common in residential data sets by converting load time series into map models.
-Most time-series clustering models are limited to handling time domain with same start and end date and time.
-Most of the solutions to handle this like longest common subsequence, dynamic time warping are prone to computational limit with increased length of the series.
--->
-
-<!-- our approach and contribution -->
-<!-- probability distribution AND cyclic granularities -->
-
-<!-- Large data sets -->
-<!-- On one hand, the sheer volume of data enables greater clarity and analysis at the individual level.
-On the contrary, it also provides challenges as larger data sets have greater uncertainty about customer behavior due to increased variety of customers.
-Additionally, when using big data, there is often a temptation to aggregate the data to reduce its volume so it is easy to manage and computationally feasible to analyse but (XXX big data to the rescue example) show that such a strategy may not always be an optimal choice especially if we are to consider highly heterogeneous profiles that have a dimension of human behaviour.
--->
-
-<!-- how looking at probability and cyclic granularity helps -->
-
-<!-- how is your work different from the rest of the literature -->
-
-<!-- \noindent The massive amount of data generated in such projects could be overwhelming for analysis.
-Electricity utilities can utilize the consumption patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
-Beyea (2010) has pointed out, there has been little discussion or exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data.
-Thus, there is a scope to investigate and analyze these data in various ways for a greater understanding of consumption patterns and how they correlate with other economic, physical or geographical factors.
-In this work, we are interested to see how we can utilize this dataset to group different customers with similar periodic behavior.
--->
-
-<!-- Towards this goal, this chapter aims to: (a) describe the contents of the data set in SGSC database that we can utilize, and (b) propose a clustering algorithm to group customers with similar periodic behaviors.
-The distance metric introduced in Chapter 2 will be the inputs for this cluster analysis.
-One of the advantages of using our  -->
-<!-- approach is that the technique is based on probability distributions instead of raw data.
-Many clustering approaches are limited by the type of noisy, patchy, and unequal time-series common in residential data sets.
-Since the distance measure considered is based on differences in probability distribution of time series, it is likely to be less sensitive to missing or noisy data.
--->
-
-<!-- _Themes_ -->
-
-<!-- - Dimension reduction: If each $P_{i, j, k}$ be considered to be a point in the space, key $i$ would have $mp$ dimensions as opposed to $n_i$ dimensions in case of considering raw data.
-Hence for a large number of observations ($n_i>>mp$), this approach benefits by transitioning to a lower dimension.
--->
-
-<!--  - Avoid loss of information due to aggregation: This approach ensures key characteristic information of the data is not lost due to averaging or aggregation measures in an attempt to transition to a lower dimension.
-Hence, this approach could be thought to somehow balance the drawback of considering raw data or aggregated data.
--->
-
-<!-- - Robustness to outliers: This approach could be adapted to be robust to outliers and extreme behaviors by trimming the tails of the probability distributions. -->
-
-<!-- - Non-synchronized observed time periods: Considering probability distribution would imply the clustering process can handle keys that are observed over periods of time that are overlapping but don't necessarily coincide.
--->
-
-<!-- - Similar periodic behavior: Since cyclic granularities are considered instead of linear granularities, clustering would group keys that have similar behavior across these cyclic granularities.
-This implies they will be grouped according to their periodic behavior and not on the linear stretch of time over which they are observed.
--->
-
-<!-- The following contributions are made through the following chapter: -->
-
-<!--  * Present a cluster analysis of SGSC dataset to group households with similar periodic behavior -->
-<!--  * Cluster validation by relating to external data -->
-
-<!-- Electricity smart meter technology is increasingly being deployed in residential and commercial buildings.
-For example, in the state of Victoria, Australia, it is a state government policy that all households have a smart meter fitted (Victoria State Government, 2015), which has resulted in 97% of the approximately 2.1 million households in Victoria having a smart meter installed in their home.
-These smart meters collect energy usage information at half-hourly intervals, resulting in over 35 billion half-hourly observations per year across all households in the state.
-Clustering electricity consumption patterns can enable electricity utilities to develop targeted tariffs for individual groups alleviating the problem of volatility in production by capitalizing on the flexibility of consumers.
--->
-
-<!-- Clustering households using only smart meter consumption data could also provide value in a societal setting by combining the findings and external data like weather conditions, socio-economic or other demographic factors of those households.
--->
-
-<!-- Background literature and shortcomings -->
-
-<!-- # Preliminary exploratoration -->
-
-<!-- ## Electricity demand data -->
-
-<!-- _Electricity use interval reading data_   -->
-<!-- The data from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
--->
-
-<!-- ### Data source -->
-
-<!-- The entire data is procured from CSIRO.
-A subset of this data is also available from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
-It consists of the following data sets.
--->
-<!-- _1. CustomerData:_ 78720 customers with 62 variables about them  -->
-<!-- _2. EUDMData:_ 300 billion half-hourly consumption level data   -->
-<!-- _3. OffersData:_ Method of contact to customer to join SGSC customer trial, either door-to-door (D2D) or via Telesales   -->
-<!-- _4. PEResponseData:_ Peak Events response customer wise   -->
-<!-- _5. PETimesData:_ Peak Events time stamps   -->
-
-<!-- Only _CustomerData_ and _EUDMData_ are relevant for the clustering goals of this paper.
-_EUDMData_ contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
-`CustomerData` provides demographic data for 78,720 customers with information about their Local Government Area amongst others.
--->
-
-<!-- ## LGA and weather data -->
-
-<!-- Since the smart meters have been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths -->
-<!-- for each household.
-Since, general supply is available for only 13,735 customers, we will restrict ourselves to look at the LGA information for these customers only.
-We find that there are only 26 LGA that is covered for these customers.
--->
-
-<!-- _Weather data_   -->
-<!-- This data is obtained through [Australian Government Bureau of Meteorology](BOM)(http://www.bom.gov.au/) and provides hourly data for nearest weather stations for all the LGAs -->
-
-<!-- This section familiarizes the 13,000 SGSC households through visualization and provide a detailed layout of data structures (like missing observations/number of customers/number of observations) and also the external data that needs to be utilized for validating the clustering process.
-The [ABS TableBuilder](https://www.abs.gov.au/websitedbs/censushome.nsf/home/tablebuilder) has census data from 2011 and 2016.
-The data is at SA2 and LGA levels.
-However, some of the LGA in NSW changed between 2011 and 2016 and hence there would not be a one-to-one correspondence between the LGAs.
-Weather, notably temperature (and humidity) can be the main driver(s) for energy usage.
-In NSW many households have electric heaters so their use can impact winter energy use and air-conditioners can impact summer energy use.
-Relevant weather data could be obtained from the [Bureau of Meteorology](www.bom.gov.au).
-Some weather stations have 30 minute (sometimes even smaller interval) weather data.
-Potentially, there could be lag effects of weather on energy usage which should be considered.
--->
+<!-- This work -->
+<!-- This is similar to a stochastic approach (@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time-series characteristics, or a model of the series, to enable unsupervised clustering.
+Unsupervised clustering is only as good as the features that are extracted/selected or the distance metrics that were utilized.
+Well-designed additional features may collect characteristics that default features cannot.
+Based on the underlying structure of the temporal data, this article offers new distance metric and features for clustering and applies them to actual smart-meter data.
+Firstly, the distance metric is based on probability distribution, which in our knowledge is the first attempt to cluster smart meter data using probability distributions.
+These recorded time series are asynchronous, with varying time lengths for different houses and missing observations.
+Taking probability distributions helps to deal with such data, while helping with dimension reduction in one hand but not losing too much information due to aggregation.
+Secondly, we recognise that most clustering algorithms only provide hourly energy profiles during the day, but this approach provides a wider approach to the issue, seeking to group consumers with similar shapes over all important cyclic granularities.
+Since cyclic granularities are considered instead of linear granularities, clustering would group customers that have similar repetitive behavior across more than one cyclic granularities across which patterns are expected to be significant.
+-->
+
+<!-- common similarity measures -->
+
+<!-- electricity data structure -->
+
+<!-- common similarity measures used there -->
+
+<!-- lit review -->
 
@@
 <!--Handling trend, seasonality, non-stationarity and auto-correlation:** Trend and seasonality are fundamental characteristics of time series data, and it is reasonable to define a time series according to its degree of trend and seasonality.
-These characteristics of the time series are lost or handled independently by considering probability distributions ( trend is lost) across categories of cyclic granularities (by independently modeling all seasonal fluctuations), and so there is no need to de-trend or de-seasonalize the data before conducting the clustering method.
+These characteristics of the time series are lost or handled independently by considering probability distributions (trend is lost) across categories of cyclic granularities (by independently modeling all seasonal fluctuations), and so there is no need to de-trend or de-seasonalize the data before conducting the clustering method.
 There is no need to omit holiday or weekend patterns for similar reasons-->

### 2021-10-26 select_cust_tour_tsne.Rmd
working on introduction again

---
title: "Selecting customers"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(ggplot2)
library(gravitas)
library(gracsr)
library(ggdendro)
library(dplyr)
library(readr)
library(visdat)
library(ggplot2)
library(tidyverse)
library(naniar)
library(here)
library(tsibble)
library(knitr)
library(patchwork)
library(GGally)
library(distributional)
library(viridis)
```

```{r mytheme}
theme_characterisation <- function() {

  theme_bw() + # seeting theme
    theme(strip.text = element_text(size = 10,
                                    margin = margin(b = 0, t = 0))) + # narrow facet space
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) + # no axis ticks 
  theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
    theme(axis.text.x = element_text(angle=90, hjust=1, size = 10)) + # rotate the x-axis text
    theme(legend.position = "bottom")+
  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
  theme(axis.text.x = element_text(size=5))
}
```

```{r anchor}
kopt <- 3 # number of clusters
```

_How distributions look (left: tour design, right: clustering)_

```{r}
data_pick_robust <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>% 
  gracsr::scale_gran( method = "robust",
                      response = "general_supply_kwh")

# data_356cust_hod <- quantile_gran(data_pick_robust,
#               "hour_day", 
#               quantile_prob_val = seq(0.1, 0.9, 0.1))
# 
# data_356cust_moy <- quantile_gran(data_pick_robust,
#                                   "month_year", 
#                                   quantile_prob_val = seq(0.1, 0.9, 0.1))
# 
# data_356cust_wkndwday <- quantile_gran(data_pick_robust,
#                                   "wknd_wday", 
#                                   quantile_prob_val = seq(0.1, 0.9, 0.1))
#

data_356cust_hod <- read_rds("data/quantile_data_356cust_hod_robust.rds") %>% 
  filter(quantiles %in% "50%")

data_356cust_moy <- read_rds("data/quantile_data_356cust_moy_robust.rds") %>% 
  filter(quantiles %in% "50%")

data_356cust_wkndwday <- read_rds("data/quantile_data_356cust_wkndwday_robust.rds") %>% 
  filter(quantiles %in% "50%")

data_356cust_hod_wide <- data_356cust_hod %>%
  pivot_wider(names_from = c("gran", "category", "quantiles"),
              values_from = "quantiles_values")

data_356cust_moy_wide <- data_356cust_moy %>%
  pivot_wider(names_from = c("gran", "category", "quantiles"),
              values_from = "quantiles_values")

data_356cust_wkndwday_wide <- data_356cust_wkndwday %>%
  pivot_wider(names_from = c("gran", "category", "quantiles"),
              values_from = "quantiles_values")

data_356cust_wide <- left_join(data_356cust_hod_wide,
                               data_356cust_moy_wide, by="customer_id") %>% 
  left_join(data_356cust_wkndwday_wide,  by="customer_id"
                              )

save(data_356cust_wide, file="data/data_356cust_wide.rda")

# cluster_result <- cluster_result %>% mutate(customer_id = as.integer(id))

# data_356cust_wide_group <- data_356cust_wide %>% 
#   left_join(cluster_result, by = ("customer_id")) %>% 
#   mutate(group = if_else(is.na(group), 10L, group))

# data_356cust_wide_group$group <- factor(data_356cust_wide_group$group)

# without scaling

data_356cust_pc <- prcomp(data_356cust_wide[,-1],
                          center = FALSE, scale = FALSE, retx = TRUE)

#plot(data_356cust_pc, type="l", npcs=50)

# from scree-plot it looks like first 10 PCs explain most of the
# variation

library(liminal)

set.seed(2935)
data_356cust_pc10 <- as_tibble(data_356cust_pc$x[,1:6])
# %>%   mutate(group = data_356cust_wide_group$group)
#limn_tour(data_356cust_pc10, PC1:PC6)

#sort(abs(data_356cust_pc$rotation[,1]))

##---- t-SNE embeddings

library(Rtsne)
set.seed(2099)
tSNE_fit <- data_356cust_wide%>% 
  select(-customer_id) %>% 
  Rtsne( pca = FALSE,
         perplexity = 30)

data_pick_one <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble 
data_pick_two <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble
data_pick_three <- c(8221762, 8273636, 10359424, 8232822)%>% as_tibble
#data_pick_four <- c(10590714,8495194,8589936, 8454235) %>% as_tibble

data_pick_cust <- bind_rows(
  data_pick_one, data_pick_two, data_pick_three,
  .id = "design")

tsne_df <- data.frame(tsneX = tSNE_fit$Y[, 1], tsneY = tSNE_fit$Y[, 2], customer_id = data_356cust_wide$customer_id) %>% left_join(
  data_pick_cust, by = c("customer_id" = "value"
)) %>% mutate(design = if_else(is.na(design), "0", design))

#rownames(tsne_df) <- data_356cust_wide$customer_id

rownames(data_356cust_pc10) <- data_356cust_wide$customer_id

## ----tsne-xy--------------------------------------------------------
tsne_xy <- ggplot(tsne_df, aes(x = tsneX, y = tsneY, color = design)) +
  geom_point(aes(text = customer_id)) +
  #scale_color_manual(values = limn_pal_tableau10()) +
  scale_colour_viridis_d(direction = -1) +
  guides(color = FALSE) +
  labs(caption = "tSNE") +
  theme(aspect.ratio = 1) +
  theme_light()

## ----highlight-customer--------------------------------------------------------

library(plotly)
tsne_plotly <- tsne_xy %>% ggplotly(tooltip = "text")
```

```{r data-read}

quantile_prob_graph <- c(0.25, 0.5, 0.75)

data_pick_one <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
data_pick_two <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
data_pick_three <- c(8221762, 8273636, 10359424, 8232822)%>% as_tibble %>% set_names("customer_id")
#data_pick_four <- c(10590714,8495194,8589936, 8454235) %>% as_tibble %>% set_names("customer_id")

cust_design <- bind_rows(
data_pick_one, data_pick_two, data_pick_three,
.id = "design")

data_pick_cust <- cust_design%>% 
  mutate(customer_id = as.character(customer_id))

# data_pick_cust$design <- paste("design",
#                                data_pick_cust$design, sep = "-")
#

# original data
data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))%>%
  mutate(customer_id = as.character(customer_id)) %>% 
  filter(customer_id %in% data_pick_cust$customer_id)  %>% 
  gracsr::scale_gran(method = "robust", response = "general_supply_kwh")
```

```{r clustering-new}

# data_pick is already robust-scaled

hod <- data_pick %>% 
  dist_gran(gran1 = "hour_day", response = "general_supply_kwh") 
moy <- data_pick %>% 
  dist_gran(gran1 = "month_year", response = "general_supply_kwh")
wkndwday <- data_pick %>% 
  dist_gran(gran1 = "wknd_wday", response = "general_supply_kwh")

distance <- wkndwday/2 + moy/12 + hod/24

f = as.dist(distance)

f %>% hclust %>% cutree(k=3)

cluster_result <- f %>% clust_gran(kopt = kopt)

# confusion_tabl2 <- confusion_tabl1 %>% left_join(cluster_result_nobroom, by = c("customer_id" = "id"))

# broom or no-broom gives the same result
```

```{r heatmap-hod}

data_heatmap_hod <- quantile_gran(data_pick,
                                  "hour_day",
                                  quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
  pivot_wider(names_from = quantiles,
              values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"))

#data_heatmap_hod$customer_id = factor(data_heatmap_hod$customer_id, levels = data_pick_cust$value)
data_heatmap_hod$category <- factor(data_heatmap_hod$category, levels = 0:23)

```

```{r data-heatmap-hod-design}

#data_pick_cust$design <- factor(data_pick_cust$design)

data_heatmap_hod_design <- data_heatmap_hod %>% 
  #mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=customer_id, fill = as.factor(design)),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=customer_id, color = as.factor(design)), size = 1) +
  facet_wrap(~customer_id, 
             scales = "free_y",
             ncol = 4) + theme_characterisation()+
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  scale_x_discrete(breaks = seq(0, 23, 3))
# + scale_y_continuous(breaks = NULL) 
```

```{r data-heatmap-hod-group-ind}

#data_pick_cust$design <- factor(data_pick_cust$design)

data_heatmap_hod_ind <- data_heatmap_hod %>% 
  #mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=customer_id, fill = as.factor(group)),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(~customer_id, 
             scales = "free_y",
             ncol = 4) + theme_characterisation()+
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
  scale_x_discrete(breaks = seq(0, 23, 3))
# + scale_y_continuous(breaks = NULL) 
# 
# 
data_heatmap_hod_design + data_heatmap_hod_ind
```

```{r data-heatmap-moy-design}

data_heatmap_moy <- quantile_gran(data_pick, "month_year", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"))

data_heatmap_moy$category <- factor(data_heatmap_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

data_heatmap_moy_design <- data_heatmap_moy %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, 
                  group=customer_id, fill = as.factor(design)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(design)), size = 1) +
  facet_wrap(~customer_id, 
             scales = "free_y", 
             ncol = 4) +
    ylab("demand (in Kwh)") +
    xlab("month-of-year")  +
  theme_characterisation() +
  scale_fill_viridis_d(direction = 1)+
  scale_color_viridis_d(direction = 1)
```

```{r data-heatmap-moy-group-ind}

data_heatmap_moy_ind <- data_heatmap_moy %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, 
                  group=customer_id, fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(~customer_id, 
             scales = "free_y", 
             ncol = 4) +
    ylab("demand (in Kwh)") +
    xlab("month-of-year")  +
  theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))

data_heatmap_moy_design+data_heatmap_moy_ind
```

```{r data-heatmap-wkndwday}

data_heatmap_wkndwday <- quantile_gran(data_pick, "wknd_wday", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"))

data_heatmap_wkndwday$category <- factor(data_heatmap_wkndwday$category, levels = c("Weekday", "Weekend"))

data_group <- data_pick %>%
  mutate(customer_id = as.character(customer_id)) %>%
  left_join(cluster_result, by = c("customer_id" = "id"))%>% 
  left_join(data_pick_cust, by = c("customer_id"))

wkndwday_data<- data_group %>% create_gran("wknd_wday") 
ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]

wkndwday_group <- wkndwday_data%>% 
  ggplot(aes(x=wknd_wday, y = general_supply_kwh))  +
  lvplot::geom_lv(aes(fill = design, 
                   color = design),
               alpha = 0.5)+
  coord_cartesian(ylim = ylim1*1.05)+
  # stat_summary(
  #   fun = median,
  #   geom = 'line',
  #   aes(group = 1), size = 1, color = "blue")+
  facet_wrap(~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             nrow = 13) +
  theme_characterisation()+
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))
```

```{r data-heatmap-hod-group}

data_group <- data_pick %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id"))

data_heatmap_hod_group <- quantile_gran(data_group,
                                  gran1="hour_day",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)

# data_heatmap_hod_group <- data_heatmap_hod_group %>%
#   mutate(group = case_when(group==1 ~"Group 2",
#                            group==2 ~"Group 4",
#                            group==3 ~"Group 1",
#                            group==4 ~"Group 3"))

# data_heatmap_hod_group$group <- factor(data_heatmap_hod_group$group, levels = c("Group 1", "Group 2", "Group 3", "Group 4"))

data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")

hod_group <- data_heatmap_hod_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=group,
                  fill = as.factor(group), alpha = 0.5),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=group, 
                color = as.factor(group)), size = 1)+
  facet_wrap(~group, 
             scales = "free_y",  
             nrow = 4)   +
  scale_fill_viridis_d(direction = 1) +
  scale_color_viridis_d(direction = 1)

```

```{r data-heatmap-moy-group}

data_group <- data_pick %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id"))

data_heatmap_moy_group <- quantile_gran(data_group,
                                  gran1="month_year",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")

moy_group <- data_heatmap_moy_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
  facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_value",
             nrow = 4) +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
  ylab("demand (in Kwh)")   +
  scale_fill_viridis_d() +
  scale_color_viridis_d()
```

```{r data-heatmap-wkndwday-2}

data_heatmap_wkndwday <- quantile_gran(data_pick, "wknd_wday", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"))

data_heatmap_wkndwday$category <- factor(data_heatmap_wkndwday$category, levels = c("Weekday", "Weekend"))

p1 <- data_heatmap_wkndwday %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=customer_id), fill = "lightblue") +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6)  +
  scale_fill_viridis_d() +
  scale_color_viridis_d()

```

```{r data-heatmap-wkndwday-group}
data_group <- data_pick %>%
  mutate(customer_id = as.character(customer_id)) %>%
  left_join(cluster_result, by = c("customer_id" = "id"))

wkndwday_data<- data_group %>% create_gran("wknd_wday") 
ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]

wkndwday_group <- wkndwday_data%>% 
  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
  lvplot::geom_lv(aes(fill = as.factor(group), color = as.factor(group)), k=5, alpha = 0.5) +
  #geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
  #geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = ylim1*1.05)+
  facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_value",
             nrow = 4)   +
  scale_fill_viridis_d() +
  scale_color_viridis_d()

```

```{r combined, eval = FALSE}

combined <- hod_group + moy_group + wkndwday_group & theme(legend.position = "bottom")*theme_characterisation()
combined + plot_layout(guides = "collect")

```

### 2021-10-26 main.Rmd
introduction changed with new structure
intro updated

@@
 <!-- time series clustering and its challenges -->
-Time-series clustering is the process of unsupervised partitioning of $n$ time-series data into $k$ ($k<<n$) groups such that homogeneous time-series are grouped together based on a certain similarity measure.
-The appropriate similarity measure or the distance approach is determined by the time-series characteristics, duration of time-series, representation method, and, of course, the goal of clustering time-series to a significant level.
-The three main approaches to time series clustering (@liao2005clustering) are algorithms that work directly with distances or raw data points, either in the time or frequency domain (distance-based), with features extracted from raw data (feature-based), or indirectly with models built from the raw data (model-based).
-The performance of distance-based approaches depends greatly on the particular distance metric used.
-Defining an adequate distance measure for the raw time series can be a challenging task since it has to consider noise, different length of time series, asynchronous time series, different scales, missing values.
-Distance-based similarity measures described in the literature include
-Euclidean,  Pearson’s correlation coefficient and related distances [124], Dynamic Time Warping [118,119], Autocorrelation, Short timeseries distance [44], Piecewise normalization [125], cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances.
-Euclidean distance and DTW are the most common methods for similarity measure in time-series clustering.
-A research has shown that, in terms of time-series classification accuracy, the Euclidean distance is surprisingly competitive [145], however, DTW also has its strength in similarity measurements which cannot be declined.
-While these methods have been quite effective in clustering moderate-length time series, there are some well-recognized drawbacks of the existing approaches.
-Euclidean distance is the most common method for discerning similarity in time series clustering, and it requires that the time series being compared are of exactly the same dimensionality (length).
-DTW can assist with clustering of different length time series if there is no missing data (Ratanamahatana and Keogh, 2005) and in any case it requires quadratic computation.
+Time-series clustering is the process of unsupervised partitioning of $n$ time-series data into $k$ ($k<n$) groups such that homogeneous time-series are grouped together based on a certain similarity measure.
+The time-series features, length of time-series, representation technique, and, of course, the purpose of clustering time-series all influence the suitable similarity measure or distance strategy to a meaningful level.
+The three primary methods to time series clustering (@liao2005clustering) are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based) (model-based).
+The efficacy of distance-based techniques is highly dependent on the distance measure utilised.
+Defining an appropriate distance measure for the raw time series may be a difficult task since it must take into account noise, variable lengths of time series, asynchronous time series, different scales, and missing data.
+Commonly used Distance-based similarity measures as suggested by a review of time series clustering approaches (@Aghabozorgi2015-ct) are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping, Autocorrelation, Short time series distance, Piecewise regularisation, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances (@liao2007clustering).
+Euclidean distance and DTW are often used in time series clustering.
+When it comes to time-series clustering accuracy, the Euclidean distance beats DTW, but DTW has its own advantages (@corradini2001dynamic).Euclidean distance requires time series of equal length.
+while DTW can assist cluster time series of varying lengths (@ratanamahatana2005multimedia), only if there are no missing observations.
 
@@
 Since cyclic granularities are considered instead of linear granularities, the resulting clusters are expected to group customers that have similar repetitive behaviors.
-Below we list some of the advantages of our approach, which will be explained in more detail along with empirical results in later sections.
-
-- When the time series is long (high dimensionality), some clustering algorithms become intractable.
-By applying dimensionality reduction via characterising probability distributions, we are able to cluster long length time series efficiently;
-
-- When the clustering algorithm is based on a distance metric (i.e.
-Euclidean distance), it cannot handle time series with missing data or of different lengths if actual points are used as inputs.
-However, by
-using probability distribution, we can trim the tails and bypass that problem.
-
-- We recognize that most clustering algorithms provide profiles across only one temporal granularity, but this approach provides a wider approach to the issue, seeking to group observations with similar shapes over all important cyclic granularities.
-Since cyclic granularities are considered instead of linear granularities, clustering would group customers that have similar repetitive behavior across more than one cyclic granularities across which patterns are expected to be significant.
-
-- It is reasonable to define a time series according to its degree of trend and seasonality and consider these features while clustering them.
-These characteristics of the time series are lost or handled independently by considering probability distributions (trend is lost) across categories of cyclic granularities (by independently modeling all seasonal fluctuations), and so there is no need to de-trend or de-seasonalize the data before conducting the clustering method.
-For example, there is no need to omit holiday or weekend patterns for similar reasons.
+Below are some of the benefits of our method, which will be detailed in further depth in subsequent sections.
+
+- Some clustering algorithms become problematic with the very high dimensionality of the time series resulting from the frequency at which they are recorded and the length of time for which they are observed.
+We can efficiently cluster long length time series by reducing dimensionality by characterising through probability distributions;
+
+- By utilising Jensen-Shannon distances, we are evaluating the distance between two distributions rather than raw data, which is less susceptible to missing observations and outliers compared to other traditional distance measures;
+
+- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar forms across all key cyclic granularities.
+Because cyclic granularities are used rather than linear granularities, clustering would group consumers who exhibit similar repeating behaviour over many cyclic granularities where patterns are predicted to be important.
+
+- It is reasonable to define a time series based on its degree of trend and seasonality and to take these characteristics into account while clustering it.
+The change in data structure by considering probability distributions across cyclic granularities ensures there is no trend and seasonal fluctuations are handled separately.
+Thus there is no need to de-trend or de-seasonalize the data prior to performing the clustering method.
+For similar reasons, there is no need to exclude holiday or weekend routines.
 
@@
 The project provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014.
-<!---The massive amount of data generated in such projects could be overwhelming for analysis.--> Raw data of these asynchronous time series can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
+<!---The massive amount of data generated in such projects could be overwhelming for analysis.Raw data of these asynchronous time series can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
 Electricity utilities can utilize the smart meter usage patterns to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
-
-<!-- Sources of variation in large data sets -->
-<!-- The enormous quantity of data provides for greater individual level clarity and analysis. However, -->
-<!-- due to the growing variety of consumers,  -->
-
-\noindent Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.
-Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
-Multiple temporal dependencies define the behavioral patterns, which vary from customer to customer.
-Some families, for example, use a dryer to dry their clothing, while others hang them to dry on a line.
-This may be reflected in their weekly profile.
-They may have monthly variations where some customers are more prone to use air conditioners or heaters than others despite the existence of comparable electrical equipment and being subjected to similar weather conditions.
-The variations in behavior may occur on a regular basis, with some consumers being night owls and others being morning larks.
-Day-off energy usage may vary depending on whether consumers choose to remain at home or engage in outdoor activities.
-These various causes of variations in energy behavior, ranging from age, lifestyle, and family composition to building characteristics, weather, presence of various electrical equipments and others, make the problem of effectively segmenting consumers into similar energy behavior a particularly intriguing one.
-
-<!-- unsupervised clustering -->
-\noindent This problem becomes more difficult when there is no other information on the customers besides their time series of energy usage.
-It is likely that such additional information is not available in order to protect the customers' confidentiality.
-Furthermore, it is not guaranteed that energy providers would always update customer profiles, such as property characteristics, in a timely manner whenever they change.
-So this study does not attempt to explain why consumption differs.
-Instead, the work looks at how much energy-use heterogeneity can be uncovered in smart meter data and what are some of the most typical electricity use patterns by simply using the time series.
+--><!-- Sources of variation in large data sets --><!-- The enormous quantity of data provides for greater individual level clarity and analysis.
+However, -->
+<!-- due to the growing variety of consumers,  -->.
+Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.
+Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
+The behavioural patterns differ amongst customers due to many temporal dependencies.
+Some households use a dryer, while others dry their clothes on a line.
+Their weekly profile may reflect this.
+They may vary monthly, with some customers using more air conditioners or heaters than others, while having equivalent electrical equipment and weather circumstances.
+Some customers are night owls, while others are morning larks.
+Day-off energy use varies depending on whether customers stay home or go outside.
+Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behaviour a fascinating one.
+This challenge is worsened when all we know about our consumers is their energy use history (@Ushakova2020-rl).
+To safeguard the customers' privacy, it is probable that such information is not accessible.
+Also, energy suppliers may not always update client information, such as property features, in a timely manner.
+Thus, there is a growing need to have research that examines how much energy usage heterogeneity can be found in smart meter data and what are some of the most common power consumption patterns, rather than explaining why consumption differs.
 
@@
 
-A typical clustering technique includes the following steps: (a) establishing distance (dissimilarity) and similarity through feature or model extraction and selection; and (b) selecting the clustering algorithm design.
-Distance measures could be time-domain based or frequency-domain (fast Fourier transform).
-A model-based clustering works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterised and clustered (@Motlagh2019-yj).
-[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
-The essential temporal characteristics of the curves are defined or extracted using feature-based clustering.
+A multitude of papers have emerged around smart meter time series clustering for deepening our knowledge of consumption patterns.
 @Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
@@
 Only one study [@ozawa2016determining] identified time series characteristics using Fourier transformation, which converts data from time to frequency and then uses K-Means to cluster by greatest frequency .
+@Motlagh2019-yj suggests that the time feature extraction islimited by the type of noisy, patchy, and unequal time-series common in residential datasets and addresses model-based clustering  by transforming the series into other other objects such as structure or set of parameters which can be more easily characterised and clustered.
+[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
+@Melnykov2013-sp disucusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
+
+\noindent Given the limitations of the similarity measures in dealing with large volumes of this complicated time series data, we present a similarity measure based on probability distributions that seems to be a more organic option for coping with time series data with aforementioned characteristics.
+The remainder of the paper is organized as follows: Section&nbsp;\ref{sec:methodology} provides the clustering methodology introducing the features and distance metrics.
+Section&nbsp;\ref{sec:validation} shows data designs to validate our methods and draw comparisons against several methods.Section&nbsp;\ref{sec:application} discusses the application of the method to a subset of the real data.
+Finally, we summarize our results and discuss possible future directions in Section&nbsp;\ref{sec:discussion}.
+
+<!-- A typical clustering technique includes the following steps: (a) establishing distance (dissimilarity) and similarity through feature or model extraction and selection; and (b) selecting the clustering algorithm design.
+Distance measures could be time-domain based or frequency-domain (fast Fourier transform).
+-->
+
+<!-- A model-based clustering works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterised and clustered (@Motlagh2019-yj).
+[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
+The essential temporal characteristics of the curves are defined or extracted using feature-based clustering.
+-->
 
@@
 <!-- lit review -->
-
-The remainder of the paper is organized as follows: Section&nbsp;\ref{sec:methodology} provides the clustering methodology introducing the features and distance metrics.
-Section&nbsp;\ref{sec:validation} shows data designs to validate our methods and draw comparisons against several methods.
-Section&nbsp;\ref{sec:application} discusses the application of the method to a subset of the real data.
-Finally, we summarize our results and discuss possible future directions in Section&nbsp;\ref{sec:discussion}.
 

### 2021-10-27 selecting_customers.Rmd
adding prototypes

---
title: "Plots with 5 anchors"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(ggplot2)
library(gravitas)
library(gracsr)
library(ggdendro)
library(dplyr)
library(readr)
library(visdat)
library(ggplot2)
library(tidyverse)
library(naniar)
library(here)
library(tsibble)
library(knitr)
library(patchwork)
library(GGally)
library(distributional)
library(viridis)
library(Rtsne)
library(liminal)
```

```{r mytheme}
theme_characterisation <- function() {

  theme_light() + # setting theme
    #theme(strip.text = element_text(margin = margin(b = 0, t = 0))) + # narrow facet space
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) + # no axis ticks 
  theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
    theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) + # no x-axis labels to further reduce the gap between facets
    #theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
  #theme(axis.text.x = element_text(size=5)) +
    theme(strip.background = element_blank(),
  strip.text.x = element_blank())
}
```

_Set number of anchors_
```{r anchor}
kopt <- 6# number of clusters
```

_Load the customers you want to cluster_
```{r}
quantile_prob_graph <- c(0.25, 0.5, 0.75)

# data_pick_one <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
# data_pick_two <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
# data_pick_three <- c(8221762, 8273636, 10359424, 8232822)%>% as_tibble %>% set_names("customer_id")

data_pick_one <- c(8541744, 9355808, 8603880, 8619309, 10542667) %>% as_tibble %>% set_names("customer_id")
#data_pick_two <- c(8688242, 8643837, 8184707, 10534355, 8684420) %>% as_tibble%>% set_names("customer_id")
data_pick_three <- c(9792072, 8589936, 8454235, 10692366, 8603828)%>% as_tibble%>% set_names("customer_id")
data_pick_four <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
data_pick_five <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
data_pick_six <- c(8221762, 8273636, 10359424, 8232822, 11450499)%>% as_tibble %>% set_names("customer_id")

```

```{r assemble}
data_pick_cust <- bind_rows(
data_pick_one, 
#data_pick_two, 
data_pick_three,
data_pick_four,
data_pick_five, 
data_pick_six,
.id = "design") %>% 
  mutate(customer_id = as.character(customer_id))

```

```{r tsne, eval = FALSE}
data_356cust_wide <- read_rds("data/data_356cust_wide.rds")

data_356cust_pc <- prcomp(data_356cust_wide[,-1],
                          center = FALSE, scale = FALSE, retx = TRUE)
set.seed(2935)
data_356cust_pc10 <- as_tibble(data_356cust_pc$x[,1:6])

##---- t-SNE embeddings

library(Rtsne)
set.seed(2099)

tSNE_fit <- data_356cust_wide%>% 
  select(-customer_id) %>% 
         Rtsne(PCA = FALSE,
         perplexity = 30)

tsne_df <- data.frame(tsneX = tSNE_fit$Y[, 1], tsneY = tSNE_fit$Y[, 2], customer_id = as.character(data_356cust_wide$customer_id)) %>% 
  left_join(
  data_pick_cust, by = c("customer_id"
)) %>% mutate(design = if_else(is.na(design), "0", design))

#rownames(tsne_df) <- data_356cust_wide$customer_id

rownames(data_356cust_pc10) <- data_356cust_wide$customer_id

## ----tsne-xy--------------------------------------------------------
tsne_xy <- ggplot(tsne_df, aes(x = tsneX, y = tsneY, color = design)) +
  geom_point(aes(text = customer_id)) +
  #scale_color_manual(values = limn_pal_tableau10()) +
  scale_colour_viridis_d(direction = -1) +
  guides(color = FALSE) +
  labs(caption = "tSNE") +
  theme(aspect.ratio = 1) +
  theme_light()

## ----highlight-customer--------------------------------------------------------

library(plotly)
tsne_plotly <- tsne_xy %>% ggplotly(tooltip = "text")
tsne_plotly

```

_Load raw data for these customers and scale it according to nqt method and group them_

```{r}
data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
  mutate(customer_id = as.character(customer_id)) %>% 
  filter(customer_id %in% data_pick_cust$customer_id) %>% 
  gracsr::scale_gran( method = "nqt",
                      response = "general_supply_kwh")
```

```{r clustering}

hod <- data_pick %>% 
  dist_gran(gran1 = "hour_day", response = "general_supply_kwh")

moy <- data_pick %>% 
  dist_gran(gran1 = "month_year", response = "general_supply_kwh")

wkndwday <- data_pick %>% 
  dist_gran(gran1 = "wknd_wday", response = "general_supply_kwh")

distance <- wkndwday/2 + moy/12 + hod/24

f = as.dist(distance)

cluster_result <- f %>% clust_gran(kopt = 5) %>% 
  rename("customer_id" = "id")
```

```{r hod-data}
data_hod <- quantile_gran(data_pick,
                                  "hour_day",
                                  quantile_prob_val = quantile_prob_graph) %>% 
  pivot_wider(names_from = quantiles,
              values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id")) %>% 
  left_join(cluster_result, by = c("customer_id"))

#data_heatmap_hod$customer_id = factor(data_heatmap_hod$customer_id, levels = data_pick_cust$value)
data_hod$category <- factor(data_hod$category, levels = 0:23)

# lev <- data_hod %>% distinct(customer_id, design) %>% arrange(design) %>% pull(customer_id)
```

```{r hod-ind-design}

# data_hod$customer_id <- factor(data_hod$customer_id, levels = lev)

hod_ind_design <- data_hod %>% 
  left_join(cluster_result,by = c("customer_id", "group")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=customer_id, 
                  fill = design),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=customer_id,
                color = design), 
            size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y",
             ncol=24) + theme_characterisation()+
  xlab("hour-of-day") +
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  scale_x_discrete(breaks = seq(0, 23, 3))
```

```{r hod-ind-group}
hod_ind_group <- data_hod %>% 
  left_join(cluster_result) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=customer_id, fill = as.factor(group)),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y",
              ncol=24) + theme_characterisation()+
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black")) +    xlab("hour-of-day")  +
  scale_x_discrete(breaks = seq(0, 23, 3))
# + scale_y_continuous(breaks = NULL) 
# 
# 
(hod_ind_design/hod_ind_group)* theme(legend.position = "none")

```

```{r data-heatmap-moy-design}

data_moy <- quantile_gran(data_pick,
                                  "month_year", 
                                  quantile_prob_val = quantile_prob_graph) %>% 
  pivot_wider(names_from = quantiles, 
              values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"))

data_moy$category <- factor(data_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

moy_ind_design <- data_moy %>% 
  left_join(cluster_result, by = c("customer_id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, 
                  group=customer_id, fill = as.factor(design)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(design)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             ncol = 24) +
    ylab("demand (in Kwh)") +
    xlab("month-of-year")  +
  theme_characterisation() +
  scale_fill_viridis_d(direction = 1)+
  scale_color_viridis_d(direction = 1)
```

```{r data-heatmap-moy-group-ind}

moy_ind_group <- data_moy %>% 
  left_join(cluster_result, by = c("customer_id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, 
                  group=customer_id, fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             ncol = 24) +
    ylab("demand (in Kwh)") +
    xlab("month-of-year")  +
  theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))

(moy_ind_design/moy_ind_group)*theme(legend.position = "none")

```

```{r data-heatmap-wkndwday-group}

data_wkndwday <- data_pick  %>%
  left_join(cluster_result)%>% create_gran("wknd_wday")  %>% 
  left_join(data_pick_cust, by = c("customer_id"))

ylim1 = boxplot.stats(data_wkndwday$general_supply_kwh)$stats[c(1, 5)]

wkndwday_ind_design <- data_wkndwday%>% 
  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
  #lvplot::geom_lv(aes(fill = as.factor(design), 
   #                   color = as.factor(design)), k=5, alpha = 0.5) +
  geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
  #geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = ylim1*1.05)+
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
              ncol = 24)  +
  theme_characterisation() +
  scale_fill_viridis_d(direction = 1)+
  scale_color_viridis_d(direction = 1)

```

```{r}
wkndwday_ind_group <- data_wkndwday%>% 
  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
  #lvplot::geom_lv(aes(fill = as.factor(group), 
          #            color = as.factor(group)), k=5, alpha = 0.5) +
  geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
  #geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = ylim1*1.05)+
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 24)  +
  theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00",  "grey", "black"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))

(wkndwday_ind_design / wkndwday_ind_group)* theme(legend.position = "none")
```

# contribution of each levels

```{r}

```

Draw group ones and make interpretation:

```{r}
data_validation <- (hod/24) %>% broom::tidy() %>% 
  rename("hod" = "distance") %>% 
  left_join((moy/12) %>% 
              broom::tidy(), 
            by = c("item1", "item2")) %>% 
  rename("moy" = "distance") %>% 
  left_join((wkndwday/2) %>% 
              broom::tidy(), 
            by = c("item1", "item2"))%>% 
  rename("wkndwday" = "distance") %>% 
  left_join(cluster_result, by = c("item1" = "customer_id")) %>% 
  rename("group_item1" = "group") %>% 
  left_join(cluster_result, by = c("item2" = "customer_id")) %>%  
  rename("group_item2" = "group") %>% 
  pivot_longer(3:5,names_to="gran",
               values_to = "distance")

data_validation %>%
  group_by(gran) %>%
  summarise(d = sd(distance))

data_validation %>%
  group_by(gran, group_item1, group_item2) %>% 
  summarise(sum = sum(distance)) %>% 
  pivot_wider(names_from = group_item2, values_from = sum) %>% 
  mutate(distance = sum(`1`, `2`, `3`, `4`, `5`)) %>% 
  select(-(3:7)) %>% 
  pivot_wider(names_from = group_item1, values_from = distance)

```

```{r}

hod_cat <- data_pick %>% 
  dist_gran_cat(gran1 = "hour_day", response = "general_supply_kwh") %>% 
  mutate(gran = "hod")

moy_cat <- data_pick %>% 
  dist_gran_cat(gran1 = "month_year", response = "general_supply_kwh")%>%
  mutate(gran = "moy")

wkndwday_cat <- data_pick %>% 
  dist_gran_cat(gran1 = "wknd_wday", response = "general_supply_kwh")%>% 
  mutate(gran = "wnwd")

data_validation_cat <- bind_rows(hod_cat, moy_cat, wkndwday_cat)

gran_cat_dist <- data_validation_cat %>% 
  left_join(cluster_result, by = c("customer_from" = "customer_id")) %>% 
  rename("group_from" = "group") %>% 
  left_join(cluster_result, by = c("customer_to" = "customer_id")) %>%  
  rename("group_to" = "group") %>% 
  group_by(gran,
           category,
           group_from, 
           group_to) %>% 
  summarise(sum = sum(distance)) %>% 
  pivot_wider(names_from = group_to, values_from = sum) %>% 
  mutate(distance = sum(`1`, `2`, `3`, `4`, `5`)) %>% 
  select(-(4:8)) %>% 
  arrange(-distance)

```

# Contribution of individual category for hod

```{r}
gran_cat_hod <-   gran_cat_dist %>% 
  filter(gran=="hod")

gran_cat_hod$category <- factor(gran_cat_hod$category, levels = 0:23)

# for each group, percent contribution of hours

group_total <- gran_cat_hod %>% 
  ungroup() %>% 
    group_by(group_from) %>% 
    summarise(group_total = sum(distance))

  # gran_cat_hod %>% 
  #   ungroup %>% 
  #   left_join(group_total, by = "group_from") %>% 
  #   mutate(cat_contibution = distance*100/group_total) %>% 
  #   group_by(group_from) %>% 
  #   slice(c(1:3))

    gran_cat_hod %>% 
    ungroup %>% 
    left_join(group_total, by = "group_from") %>% 
    mutate(cat_contibution = distance*100/group_total) %>% 
    group_by(group_from) %>% 
    arrange(-distance) %>% slice(c(1:4)) %>% 
      ggplot(aes(group_from, category, color = as.factor(group_from))) +
      geom_point(aes(size = distance))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+ theme_light()

data_pcp <- gran_cat_hod %>% 
    ungroup %>% 
    left_join(group_total, by = "group_from") %>% 
    mutate(cat_contibution = distance*100/group_total) %>% 
    group_by(group_from) %>% 
    arrange(-distance) %>% 
  select(group_from, category, cat_contibution) %>% 
  pivot_wider(names_from = "category", values_from = "cat_contibution") %>% mutate(group_from = as.factor(group_from))

parcoord <- GGally::ggparcoord(data_pcp ,
                   columns = 2:ncol(data_pcp),
                   groupColumn = "group_from",
                   showPoints = TRUE, 
                   alphaLines = 1,
                   order = "anyClass",
                   scale = "globalminmax"
) + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black")) + theme_bw()
    parcoord

```

# Contribution of individual category for moy

```{r}
gran_cat_moy <-   gran_cat_dist %>% 
  filter(gran=="moy")

gran_cat_moy$category <- factor(gran_cat_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

group_total <- gran_cat_moy %>% 
  ungroup() %>% 
    group_by(group_from) %>% 
    summarise(group_total = sum(distance))

    gran_cat_moy %>% 
    ungroup %>% 
    left_join(group_total, by = "group_from") %>% 
    mutate(cat_contibution = distance*100/group_total) %>% 
    group_by(group_from) %>% 
    arrange(-distance) %>% slice(c(1:3)) %>% 
      ggplot(aes(group_from, category, color = as.factor(group_from))) +
      geom_point(aes(size = distance))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+ theme_light()

    data_pcp <- gran_cat_moy %>% 
    ungroup %>% 
    left_join(group_total, by = "group_from") %>% 
    mutate(cat_contibution = distance*100/group_total) %>% 
    group_by(group_from) %>% 
    arrange(-distance) %>% 
  select(group_from, category, cat_contibution) %>% 
  pivot_wider(names_from = "category", values_from = "cat_contibution") %>% mutate(group_from = as.factor(group_from))

parcoord <- GGally::ggparcoord(data_pcp ,
                   columns = 2:ncol(data_pcp),
                   groupColumn = "group_from",
                   showPoints = TRUE, 
                   alphaLines = 1,
                   order = "anyClass",
                   scale = "globalminmax"
) + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black")) + theme_bw()
    parcoord

```

# Contribution of individual category for wknwday

```{r}
gran_cat_wkndwday <-   gran_cat_dist %>% 
  filter(gran=="wnwd")

gran_cat_wkndwday$category <- factor(gran_cat_wkndwday$category, levels = c("Weekday", "Weekend"))

group_total <- gran_cat_wkndwday %>% 
  ungroup() %>% 
    group_by(group_from) %>% 
    summarise(group_total = sum(distance))

    gran_cat_wkndwday %>% 
    ungroup %>% 
    left_join(group_total, by = "group_from") %>% 
    mutate(cat_contibution = distance*100/group_total) %>% 
    group_by(group_from) %>% 
    arrange(-distance) %>% slice(c(1)) %>% 
      ggplot(aes(group_from, category, color = as.factor(group_from))) +
      geom_point(aes(size = distance))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+ theme_light()
```

# Show aggregate group behavior

```{r data-heatmap-hod-group}
legend_title <- "group"

data_group <- data_pick  %>% 
  left_join(cluster_result, by = c("customer_id"))

data_heatmap_hod_group <- quantile_gran(data_group,
                                  gran1="hour_day",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)

# data_heatmap_hod_group <- data_heatmap_hod_group %>%
#   mutate(group = case_when(group==1 ~"Group 2",
#                            group==2 ~"Group 4",
#                            group==3 ~"Group 1",
#                            group==4 ~"Group 3"))

# data_heatmap_hod_group$group <- factor(data_heatmap_hod_group$group, levels = c("Group 1", "Group 2", "Group 3", "Group 4"))

data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")

hod_group <- data_heatmap_hod_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=group,
                  fill = as.factor(group), alpha = 0.5),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=group, 
                color = as.factor(group)), size = 1)+
  facet_wrap(~group, 
             scales = "free_y",  
             ncol = 5) + 
              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
  ylab("demand (in Kwh)") + 
  theme_bw() +
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
  theme(panel.spacing =unit(0, "lines")) + 
  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
  scale_x_discrete(breaks = seq(1, 24, 3))+ 
  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
  theme(legend.position = "bottom")

```

```{r data-heatmap-moy-group}
data_heatmap_moy_group <- quantile_gran(data_group,
                                  gran1="month_year",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")

moy_group <- data_heatmap_moy_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
  facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 5) +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
  ylab("demand (in Kwh)") +
  theme_bw() +
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) + 
  theme(panel.spacing =unit(0, "lines")) + 
  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
  theme(legend.position = "bottom") +
  theme(panel.spacing =unit(0, "lines")) + 
  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
```

```{r}
combined <- (hod_group/moy_group) & theme(legend.position = "bottom")
combined + plot_layout(guides = "collect")
```

```{r}
wkndwday_data <- data_group %>% create_gran("wknd_wday") %>% 
  create_gran("hour_day")

ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]

wkndwday_group <- wkndwday_data%>% 
  ggplot(aes(x=hour_day, y = general_supply_kwh)) +
  #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
  geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
  #geom_boxplot(outlier.size = 1) + 
  coord_cartesian(ylim = ylim1*1.05)+
  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
#geom_boxplot(aes(fill = as.factor(group))) +
  #scale_fill_lv() +
 xlab("wknd-wday") + 
  ylab("demand (in Kwh)") +
   facet_grid(wknd_wday~group, 
             scales = "free_y", 
             labeller = "label_both") + 
  theme_bw() +
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) + 
  theme(panel.spacing =unit(0, "lines")) + 
  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
  scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
  theme(legend.position = "bottom") +
  theme(panel.spacing =unit(0, "lines")) + 
  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
wkndwday_group
```

### 2021-10-27 gran3_orig.R

@@ -8,7 +8,6 @@ library(gravitas)
 library(dplyr)
 library(here)
 
-
 # fix parameters
 
 # niter
@@ -33,7 +32,7 @@ simtable <- expand.grid(mean_diff = mean_diff,
 # parallel::mclapply(seq_len(nrow(simtable)), function(scen){
 
 #scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
-scen <- 27
+scen <- 1
 
 
 simj<-simtable[scen,] #Extract row of table
@@ -180,6 +179,7 @@ dist_mat <- bind_data_iter_tsibble %>%
 
 groups = dist_mat%>% clust_gran(kopt = 5)
 
+
 pred_group = paste(groups$group,sep = "") %>% as.factor()
 actual_group = as.factor(bind_data_iter_tsibble %>%as_tibble %>% select(customer_id, design) %>% distinct() %>% pull(design))
 
@@ -213,10 +213,25 @@ dist_mat_g3 <- bind_data_iter_tsibble %>%
   scale_gran(method = "nqt", response = "ts") %>%
   dist_gran(gran1 = "g3", response= "ts")
 
-dist_mat <- dist_mat_g1 + dist_mat_g2 + dist_mat_g3
+dist_mat <- dist_mat_g1/2 + dist_mat_g2/3 + dist_mat_g3/5
 
 groups = dist_mat %>% 
   clust_gran(kopt = 5)
+
+data_validation <- (dist_mat_g1/2) %>% broom::tidy() %>% 
+  rename("g1" = "distance") %>% 
+  left_join((dist_mat_g2/3) %>% broom::tidy(), by = c("item1", 
+                                                      "item2")) %>% 
+  rename("g2" = "distance") %>% left_join((dist_mat_g3/5) %>% broom::tidy(), by = c("item1", "item2"))%>% 
+  rename("g3" = "distance") %>% 
+  left_join(groups, by = c("item1" = "id")) %>% 
+  rename("group_item1" = "group") %>% 
+  left_join(groups, by = c("item2" = "id")) %>%  
+  rename("group_item2" = "group") %>% 
+  pivot_longer(3:5,names_to="gran",
+               values_to = "distance")
+
+write_rds(data_validation, here(paste0("js-nqt/3gran_change_5D/data_validation_", scen, ".rds")))
 
 
 pred_group = paste(groups$group,sep = "") %>% as.factor()
[…]

### 2021-10-27 quantiles_raw_nqt.R

@@ -4,55 +4,15 @@ library(readr)
 library(tsibble)
 library(gravitas)
 
-# no scaling
-
-data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) 
-
-data_356cust_hod <- quantile_gran(data_pick,
-                                  "hour_day", 
-                                  quantile_prob_val = seq(0.1, 0.9, 0.1))
-
-data_356cust_moy <- quantile_gran(data_pick,
-                                  "month_year", 
-                                  quantile_prob_val = seq(0.1, 0.9, 0.1))
-
-data_356cust_wkndwday <- quantile_gran(data_pick,
-                                       "wknd_wday", 
-                                       quantile_prob_val = seq(0.1, 0.9, 0.1))
 
 
-write_rds(data_356cust_hod, here::here("data/quantile_data_356cust_hod_noscale.rds"))
-write_rds(data_356cust_moy,here::here("data/quantile_data_356cust_moy_noscale.rds"))
-write_rds(data_356cust_wkndwday, here::here("data/quantile_data_356cust_wkndwday_noscale.rds"))
-
-
-
-# robust scaling
-
-data_pick_robust <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>% 
-  gracsr::scale_gran( method = "robust",
-                      response = "general_supply_kwh")
-
-# data_356cust_hod <- quantile_gran(data_pick_robust,
-#               "hour_day", 
-#               quantile_prob_val = seq(0.1, 0.9, 0.1))
-# 
-# data_356cust_moy <- quantile_gran(data_pick_robust,
-#                                   "month_year", 
-#                                   quantile_prob_val = seq(0.1, 0.9, 0.1))
-# 
-# data_356cust_wkndwday <- quantile_gran(data_pick_robust,
-#                                   "wknd_wday", 
-#                                   quantile_prob_val = seq(0.1, 0.9, 0.1))
-# 
-
-data_356cust_hod <- read_rds("data/quantile_data_356cust_hod_robust.rds") %>% 
[…]

### 2021-10-27 script/gran2_orig.R

@@ -24,8 +24,8 @@ simtable <- expand.grid(mean_diff = mean_diff,
 
 # parallel::mclapply(seq_len(nrow(simtable)), function(scen){
 
-#scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
-scen <- 1
+scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
+#scen <- 10
 
 
 simj<-simtable[scen,] #Extract row of table
@@ -144,40 +144,69 @@ set.seed(123)
 bind_data_iter <- map(seq_len(niterj), 
                       function(x){
                         #set.seed(seed_len[x])  
-                        # design1 <- generate_design(n=nTj) # null design
-                        # design2 <- generate_design(n=nTj,
-                        #                            mu31 = mean_diffj) 
-                        #                            
-                        # 
-                        # design3 <- generate_design(n=nTj, 
-                        #                            mu32=mean_diffj) 
-                        #                       
-                        # 
-                        # design4 <- generate_design(n=nTj,
-                        #                            mu33 = mean_diffj)
-                        #                  
-                        # design5 <- generate_design(n=nTj,
-                        #                            mu32 = mean_diffj,
-                        #                            mu33 = mean_diffj)
-                        # bind_design <- bind_rows(design1, design2, design3, design4, design5,  .id = "design")
-                        # 
+                        design1 <- generate_design(n=nTj) # null design
+                        design2 <- generate_design(n=nTj,
+                                                   mu21 = mean_diffj, 
+                                                   mu32=mean_diffj , 
+                                                   mu34=mean_diffj)
                         
-                        design1 <- generate_design(n=nTj) 
-     
+                        design3 <- generate_design(n=nTj, 
[…]

### 2021-10-27 script/gran3_orig.R

@@ -8,6 +8,7 @@ library(gravitas)
 library(dplyr)
 library(here)
 
+
 # fix parameters
 
 # niter
@@ -17,6 +18,13 @@ niter <- c(5, 50, 100) #number of series you are clustering
 nT <-  c(300, 1000, 5000) # length of the time series
 mean_diff <- c(1, 2, 5) # difference between consecutive categories
 
+# number of iterations corresponding to
+# niter <- c(5) #number of series you are clustering
+# nT <-  c(300, 1000, 5000) # length of the time series
+# mean_diff <- c(1) # difference between consecutive categories
+
+#time_series <- c("ar1", "arma22")
+
 simtable <- expand.grid(mean_diff = mean_diff,
                         niter = niter,
                         #time_series = time_series, 
@@ -24,8 +32,8 @@ simtable <- expand.grid(mean_diff = mean_diff,
 
 # parallel::mclapply(seq_len(nrow(simtable)), function(scen){
 
-#scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
-scen <- 1
+scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
+#scen <- 27
 
 
 simj<-simtable[scen,] #Extract row of table
@@ -138,44 +146,21 @@ generate_design <- function(n = 300, #length of time series
 }
 
 
+
+
 set.seed(123)
 
 ##----bind-designs for many iterations
 bind_data_iter <- map(seq_len(niterj), 
                       function(x){
                         #set.seed(seed_len[x])  
-                        # design1 <- generate_design(n=nTj) # null design
-                        # design2 <- generate_design(n=nTj,
-                        #                            mu31 = mean_diffj) 
-                        #                            
-                        # 
-                        # design3 <- generate_design(n=nTj, 
-                        #                            mu32=mean_diffj) 
-                        #                       
-                        # 
-                        # design4 <- generate_design(n=nTj,
-                        #                            mu33 = mean_diffj)
-                        #                  
[…]

### 2021-10-27 main.Rmd
adding prototypes
working on application

@@
 
+```{r mytheme-application}
+theme_application <- function() {
+
+  theme_light() + # setting theme
+    #theme(strip.text = element_text(margin = margin(b = 0, t = 0))) + # narrow facet space
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank()) + # no axis ticks 
+  theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
+    theme(axis.ticks.x = element_blank(),
+        axis.text.x = element_blank()) + # no x-axis labels to further reduce the gap between facets
+    #theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
+  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+  #theme(axis.text.x = element_text(size=5)) +
+    theme(strip.background = element_blank(),
+  strip.text.x = element_blank())
+}
+```
+
 # Introduction
@@
 
-First, we select the customers which do not have any implicit missing values and filtered their data for the year 2013.
+_Pre-processing steps_
+
+P1. We randomly select a sample of $600$ customers which do not have any implicit missing values and filtered their data for the year 2013.
+
+P2.
+Obtain $wpd$ for all cyclic granularities considered for these customers.
+It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weeknd/weekday) are coming out to be significant for most customers.
+This implies that for most customers, there is some interesting pattern across these three granularities.
+
+P3.
+Remove customers from this list for which data in any category for the significant granularities are empty.
+For example, in this data set, if a customer do not have data for an entire month, they have been removed as their monthly behavior could not be studied in that case.
+
+P4.
+Remove customers for which all the deciles of the energy consumption is zero.
+These are the customers whose consumption remain mostly flat and is expected to have no interesting repetitive patterns that is our interest of study.
+
+Finally, we are left with $356$ customers from which we run our prototype search.
+There are different methods of going ahead with this.
+For example, one approach could be to use any Non-linear dimensionality reduction technique like MDS or PCA and project the data in a 2-dimensional space.
+One can then look at few "anchor" customers which are far apart in the 2D space and pick few neighbors for each of the anchor customers.
+Paradoxically, the curse of dimensionality inverts for dimension reduction, resulting in an excessive amount of observations near the center of the distribution.
+This affects visualizations made on low-dimensional projections.
+[Ursula's paper] explains why points tend to be away from the center in the high-dimensional space, but crowd the center in low-dimensional projections, it is helpful to consider the projected volume relative to high-dimensional volume.
+
+<!-- From this set we select $4$ "anchor" customers which are far apart from each other and $5$ neighboring customers for each of these anchors.
+These selections were done using the granularity $hod$ space.
+It is important to note that when we use our proposed methodologies, it is based on all dimensions `hod`, `moy` and `wkndwday`.Fig \ref{fig:mds} shows the MDS of these 356 customers in a 2D space basis their distance on individual granularities and when all of them are combined.
+Our methodolgies are run on these $24$ customers, which act as a way to evaluate the proposed methodologies.
+-->
+
+S1. Robust scaling is applied to each customer.
+
+S2.
+$50^{th}$ percentile for each category for each granularity is obtained for each customers.
+So we have a data structure with  $356$ rows and $(24 + 12 + 2)$ variables corresponding to $50^{th}$ percentile for each hour-of-day, month-of-year and weekend-weekday.
+
+S3. Apply principal components and restrict the results down to the first six principal components (which makes up approximately 85% of the variance explained in the data) to use with the grand tour.
+
+S4.
+Run t-SNE using the default arguments on the complete data (sets the perplexity to equal 30 and performs random initialisation).
+We then create a linked tour with t-SNE layout with liminal as shown in Figure 4.
+
+S5.
+We inspect of the subspace generated by the set of low-dimensional projections in tour by looking for a simplex shape while the visualization moves from one basis to another.
+When we brush the corners of the simplex, we find they fall on the edge of the t-SNE point cloud.
+Hall, Marron, and Neeman (2005) have shown that in the extreme case of high-dimension, low-sample size data, observations are on the vertices of a simplex.
+
+This is because in high-dimensional data analysis the curse of dimensionality reasons that points tend to be far away from the center of the distribution and on the edge of high-dimensional space.
+Contrary to this, is that projected data tends to clump at the center.
+
+S6. These points should ideally correspond to different behavior with respect to all the variables considered while running PCA.
+
+<!-- First, we select the customers which do not have any implicit missing values and filtered their data for the year 2013.
 From this set, we randomly sample a set of $600$ customers.
@@
 These are the customers whose consumption remain mostly flat and is expected to have no interesting repetitive patterns that is our interest of study.
-Finally, we are left with $356$ customers.
-From this set we select $4$ "anchor" customers which are far apart from each other and $5$ neighboring customers for each of these anchors.
-These selections were done using the granularity $hod$ space.
-It is important to note that when we use our proposed methodologies, it is based on all dimensions `hod`, `moy` and `wkndwday`.Fig \ref{fig:mds} shows the MDS of these 356 customers in a 2D space basis their distance on individual granularities and when all of them are combined.
-Our methodolgies are run on these $24$ customers, which act as a way to evaluate the proposed methodologies.
-
-```{r mds, out.width="100%", fig.cap="Instance selection with respect to `hod` such that there are two distinct group and two overlapping groups."}
-knitr::include_graphics("figs/mds_plot.png")
+-->
+
+```{r mds, out.width="100%", fig.cap="Instance selection using tours and projecting the points in a lower dimensional tsne cloud."}
+knitr::include_graphics("figs/liminal_view.png")
 ```
@@
 -->
-
-```{r data-pick}
-
+```{r}
 quantile_prob_graph <- c(0.25, 0.5, 0.75)
 
-wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_robust.rds")) 
-#%>% broom::tidy()
-
-moy <- read_rds(here("data/dist_gran_moy_356cust.rds")) 
-#%>% broom::tidy()
-
-hod <- read_rds(here("data/dist_gran_hod_356cust_robust.rds")) 
-#%>% broom::tidy()
-
-distance <- wkndwday %>% 
-  left_join(moy, by = c("item1", "item2")) %>% 
-  left_join(hod, by = c("item1", "item2")) %>% 
-  rename("wkndwday" ="distance.x",
-         "moy" = "distance.y",
-         "hod" = "distance") %>%
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
-
-total_distance <- distance %>% 
-  mutate(total = hod)
-  #mutate(total = wkndwday/2 + moy/12 + hod/24)
-
-# 8454221 10420689
-data_pick_one <- total_distance %>% filter(item1 %in% 
-                            c(8454221)) %>% 
-  #group_by(item1) %>%
-  arrange(total) %>% 
-  head(5) %>% 
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
-
-data_pick_ch <-  total_distance %>% filter(item1 %in% 
-                            c(9393696)) %>% group_by(item1) %>% arrange(total) %>% head(5) %>% 
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
-
-#11013154  8181071 11048034
-data_pick_two <- total_distance %>% filter(item1 %in% 
-                            c(8181071)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
-
-data_pick_three <- total_distance %>% filter(item1 %in% 
-                            c(11013154)) %>%
-  #group_by(item1) %>%
-  arrange(total) %>% head(5)%>% 
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
-
-# data_pick_three <- total_distance %>% filter(item1 %in%
-#                             c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
-#   mutate(item1 = as.integer(as.character(item1)),
-#          item2 = as.integer(as.character(item2)))
-
-# 8653709
-data_pick_four <- total_distance %>% filter(item1 %in%
-                            c(8184707)) %>% 
-  #group_by(item1) %>% 
-  arrange(total) %>% head(5)%>%
-  mutate(item1 = as.integer(as.character(item1)),
-         item2 = as.integer(as.character(item2)))
-
-# c(11013154, 8495194, 8627007,10109182,10677705,8952846)
-
+# data_pick_one <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
+# data_pick_two <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
+# data_pick_three <- c(8221762, 8273636, 10359424, 8232822)%>% as_tibble %>% set_names("customer_id")
+
+data_pick_one <- c(8541744, 9355808, 8603880, 8619309, 10542667) %>% as_tibble %>% set_names("customer_id")
+#data_pick_two <- c(8688242, 8643837, 8184707, 10534355, 8684420) %>% as_tibble%>% set_names("customer_id")
+data_pick_three <- c(9792072, 8589936, 8454235, 10692366, 8603828)%>% as_tibble%>% set_names("customer_id")
+data_pick_four <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
+data_pick_five <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
+data_pick_six <- c(8221762, 8273636, 10359424, 8232822, 11450499)%>% as_tibble %>% set_names("customer_id")
+
+```
+
+```{r assemble}
 data_pick_cust <- bind_rows(
-unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(),    unique(c(data_pick_two$item1,data_pick_two$item2)) %>% as_tibble(), 
-c(11013154, 9345642, 8328008, 8636035,8627007,8454235) %>% as_tibble(),
-unique(c(data_pick_four$item1,data_pick_four$item2)) %>% as_tibble(),
-.id = "design")
-
-data_pick_cust$design <- paste("design", data_pick_cust$design, sep = "-")
-
-data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))%>%
-  filter(customer_id %in% data_pick_cust$value) %>% 
-  gracsr::scale_gran( method = "robust", response = "general_supply_kwh")
-
-```
-
-<!-- # ```{r} -->
-<!-- # data_pick %>% ggplot() + geom_line(aes(x=reading_datetime, y = general_supply_kwh)) + -->
-<!-- #   facet_wrap(~customer_id, ncol= 2) +  -->
-<!-- # ``` -->
+data_pick_one, 
+#data_pick_two, 
+data_pick_three,
+data_pick_four,
+data_pick_five, 
+data_pick_six,
+.id = "design") %>% 
+  mutate(customer_id = as.character(customer_id))
+
+```
+
+```{r}
+data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
+  mutate(customer_id = as.character(customer_id)) %>% 
+  dplyr::filter(customer_id %in% data_pick_cust$customer_id) %>% 
+  gracsr::scale_gran( method = "nqt",
+                      response = "general_supply_kwh")
+```
 
 ```{r clustering}
-filtered_distance <- distance %>% filter(item1 %in% data_pick_cust$value) %>% 
-  filter(item2 %in% data_pick_cust$value)
-
-total_distance <- filtered_distance %>% 
-mutate(total = wkndwday/2 + moy/12 + hod/24) 
-#mutate(total = hod)
-
-total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
-                               names_from = item2,
-                               values_from = total)
-#total_distance_wide$`11013154` <- NA
-
-rownames(total_distance_wide) <- total_distance_wide$item1
-
-mds_data <- total_distance_wide %>% 
-  mutate_all(~replace(., is.na(.), 0)) %>%
-   tibble::rownames_to_column() %>%  
-   select(-item1) %>% 
-   pivot_longer(-rowname) %>% 
-   pivot_wider(names_from=rowname, values_from=value)
-
-rownames(mds_data) <- total_distance_wide$item1
-
-# group <- mds_data[-1] %>% clust_gran(kopt=4)
-#
-
-df <- mds_data[-1] %>% as.matrix()
-DM <- matrix(0, ncol(mds_data), ncol(mds_data))
-DM[lower.tri(DM)] = df[lower.tri(df, diag=TRUE)]
-
-f = as.dist(DM)
-
-#f %>% hclust %>% cutree(k=4)
-
-first_lot <- mds_data %>% names()
-# because distance matrix is such  that 1:23 appears in columns and 2:24 appears in rows such that 1st customer is missing from row and last customer is missing from variables
-id <- c(first_lot[-1], mds_data$name[nrow(mds_data)])
-
-group <- f %>% hclust (method = "ward.D") %>% cutree(k=4)
-
-cluster_result <- bind_cols(id = id, group = group)
-
-mds  = cmdscale(f)
-rownames(mds) = id
-colnames(mds) = c("mds1", "mds2")
-
-# ggplot(data = mds %>% as_tibble() %>% mutate(group = cluster_result$group), 
-#        aes(x = mds1, y = mds2, color = as.factor(group))) +
-#   geom_point() +
-#   geom_text(aes(label = id))
-
-# cluster_result %>% filter(group==1)
-# cluster_result %>% filter(group==2)
-# cluster_result %>% filter(group==3)
-# cluster_result %>% filter(group==4)
-
-  #x = ~mds1.hod, y = ~ mds2.hod, text = ~customer_id
-```
-```{r heatmap-hod}
-
-data_heatmap_hod <- quantile_gran(data_pick,
-                                  "hour_day", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
-  left_join(data_pick_cust, by = c("customer_id"="value"))
+hod <- suppressMessages(data_pick %>% 
+  dist_gran(gran1 = "hour_day", response = "general_supply_kwh"))
+
+moy <- suppressMessages(data_pick %>% 
+  dist_gran(gran1 = "month_year", response = "general_supply_kwh"))
+
+wkndwday <- suppressMessages(data_pick %>% 
+  dist_gran(gran1 = "wknd_wday", response = "general_supply_kwh"))
+
+distance <- wkndwday/2 + moy/12 + hod/24
+
+f = as.dist(distance)
+
+cluster_result <- suppressMessages(f %>% 
+  clust_gran(kopt = 5)) %>% 
+  rename("customer_id" = "id")
+```
+
+```{r hod-data}
+data_hod <- quantile_gran(data_pick,
+                                  "hour_day",
+                                  quantile_prob_val = quantile_prob_graph) %>% 
+  pivot_wider(names_from = quantiles,
+              values_from = quantiles_values) %>% 
+  left_join(data_pick_cust, by = c("customer_id")) %>% 
+  left_join(cluster_result, by = c("customer_id"))
 
 #data_heatmap_hod$customer_id = factor(data_heatmap_hod$customer_id, levels = data_pick_cust$value)
-data_heatmap_hod$category <- factor(data_heatmap_hod$category, levels = 0:23)
-
-```
-
-```{r data-heatmap-hod-new, fig.cap="The distribution of electricity demand for the selected 24 customers across hour-of-day.
-The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-According to our prototype selection method, each row represents a distinct shape of daily load, and all customers in the same row should have similar daily profile.After clustering these consumers, all customers with the same colour represent the same group.
-Except for customer id 8269176, there is unanimity across design and groups.
-Even though their daily form resembles Group 2, our clustering approach places them in Group 3 (which is design 4 in our case).
-Because our method uses hod, moy, and wkndwday, there may be some mismatches depending on only one variable."}
-
-legend_title <- "group"
-
-data_heatmap_hod %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
+data_hod$category <- factor(data_hod$category, levels = 0:23)
+
+# lev <- data_hod %>% distinct(customer_id, design) %>% arrange(design) %>% pull(customer_id)
+```
+
+```{r hod-ind-design}
+
+# data_hod$customer_id <- factor(data_hod$customer_id, levels = lev)
+
+hod_ind_design <- data_hod %>% 
+  left_join(cluster_result,by = c("customer_id", "group")) %>% 
   ggplot(aes(x = category)) + 
@@
                   group=customer_id, 
-                  fill = as.factor(group)),
+                  fill = design),
               alpha = 0.5) +
@@
                 group=customer_id,
-                color = as.factor(group)), size = 1) +
+                color = design), 
+            size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y",
+             ncol=24) + 
+ theme_application() +
+  xlab("hour-of-day") +
+  scale_fill_viridis_d()+
+  scale_color_viridis_d()+
+  scale_x_discrete(breaks = seq(0, 23, 3))
+```
+
+```{r hod-ind-group}
+hod_ind_group <- data_hod %>% 
+  left_join(cluster_result) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=customer_id, fill = as.factor(group)),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=customer_id, color = as.factor(group)), size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y",
+              ncol=24) + theme_application()+
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black")) +    xlab("hour-of-day")  +
+  scale_x_discrete(breaks = seq(0, 23, 3))
+# + scale_y_continuous(breaks = NULL) 
+# 
+# 
+(hod_ind_design/hod_ind_group)* theme(legend.position = "none")
+
+```
+
+```{r data-heatmap-moy-design}
+
+data_moy <- quantile_gran(data_pick,
+                                  "month_year", 
+                                  quantile_prob_val = quantile_prob_graph) %>% 
+  pivot_wider(names_from = quantiles, 
+              values_from = quantiles_values) %>% 
+  left_join(data_pick_cust, by = c("customer_id"))
+
+data_moy$category <- factor(data_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+moy_ind_design <- data_moy %>% 
+  left_join(cluster_result, by = c("customer_id")) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, 
+                  group=customer_id, fill = as.factor(design)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(design)), size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             ncol = 24) +
+    ylab("demand (in Kwh)") +
+    xlab("month-of-year")  +
+  theme_application() +
+  scale_fill_viridis_d(direction = 1)+
+  scale_color_viridis_d(direction = 1)
+```
+
+```{r data-heatmap-moy-group-ind}
+
+moy_ind_group <- data_moy %>% 
+  left_join(cluster_result, by = c("customer_id")) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, 
+                  group=customer_id, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
+  facet_wrap(design~customer_id, 
+             scales = "free_y", 
+             ncol = 24) +
+    ylab("demand (in Kwh)") +
+    xlab("month-of-year")  +
+  theme_application() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))
+
+(moy_ind_design/moy_ind_group)*theme(legend.position = "none")
+
+```
+
+```{r data-heatmap-wkndwday-group2}
+
+data_wkndwday <- data_pick  %>%
+  left_join(cluster_result)%>% create_gran("wknd_wday")  %>% 
+  left_join(data_pick_cust, by = c("customer_id"))
+
+ylim1 = boxplot.stats(data_wkndwday$general_supply_kwh)$stats[c(1, 5)]
+
+wkndwday_ind_design <- data_wkndwday%>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  #lvplot::geom_lv(aes(fill = as.factor(design), 
+   #                   color = as.factor(design)), k=5, alpha = 0.5) +
+  geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
+  #geom_boxplot(outlier.shape = NA) + 
+  coord_cartesian(ylim = ylim1*1.05)+
   facet_wrap(design~customer_id, 
@@
              labeller = "label_value",
-             ncol = 6) +
-    theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) +
-    theme(strip.text = element_text(size = 10,
-                                    margin = margin(b = 0, t = 0))) +
-  ylab("demand (in Kwh)")  + 
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ 
-  #theme() strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") ) +
-  scale_x_discrete(breaks = seq(0,23, 4)) +
-  scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  xlab("hour-of-day") 
-```
-
-```{r data-heatmap-moy, fig.cap = "The distribution of electricity demand for the selected 24 customers across month-of-year.
-The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-Intriguingly, customer id 8269176 appears to be in the appropriate group (Group 3) because its monthly profile is more similar to Group 3 than Design 4.
-So, while our clustering approach failed to place the customer in the correct hour-of-day group, it did so for month-of-year.
-When contrasting the moy to hod profile, there are greater behavioural differences across customers within a group."}
-
-data_heatmap_moy <- quantile_gran(data_pick, "month_year", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
-  left_join(data_pick_cust, by = c("customer_id"="value"))
-
-data_heatmap_moy$category <- factor(data_heatmap_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
-
-data_heatmap_moy %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, 
-                  group=customer_id, 
-              fill = as.factor(group)), alpha = 0.5) +
-  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
+              ncol = 24)  +
+  theme_application() +
+  scale_fill_viridis_d(direction = 1)+
+  scale_color_viridis_d(direction = 1)
+
+```
+
+```{r}
+wkndwday_ind_group <- data_wkndwday%>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  #lvplot::geom_lv(aes(fill = as.factor(group), 
+          #            color = as.factor(group)), k=5, alpha = 0.5) +
+  geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
+  #geom_boxplot(outlier.shape = NA) + 
+  coord_cartesian(ylim = ylim1*1.05)+
   facet_wrap(design~customer_id, 
@@
              labeller = "label_value",
-             ncol = 6) +
-    theme(strip.text = element_text(size = 10,
-                                    margin = margin(b = 0, t = 0))) + xlab("month-of-year") +
-  ylab("demand (in Kwh)")  +
-  theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + 
-  theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") )+
-  scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  xlab("month-of-year") +  theme(axis.text.x = element_text(size=5))
+             ncol = 24)  +
+  theme_application() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00",  "grey", "black"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))
+
+(wkndwday_ind_design / wkndwday_ind_group)* theme(legend.position = "none")
 ```
@@
 
+Characterization of clusters both statistically and qualitatively is an important stage of a cluster analysis.
+A potential way is to look at the findings from all the groups in graphs, and enhance our qualitative descriptions of the groupings.
+Figure \ref{fig:combined} shows the distribution of the summarized groups and help us to characterise each of the clusters.
+All of these may be validated with further information about the customer.
+
+_Group 1:_ This includes consumers who work 9-5, get up and conduct morning activities from 7-10am, and then depart.
+Then they return home in the evening to cook supper and perform other activities, giving the evening a greater peak than the morning.
+These users in this cluster are insensitive to high temperature but have their heaters on in the winter months May-Aug.
+
+_Group 2:_  This is the group that rushes out of the house in the morning to get to work.
+They only return at night and do all activities at night, so there is no morning peak.
+In addition, the users in this cluster are insensitive to high temperature, but sensitive to lower temperature.
+
+_Group 3:_  This group a strong early morning and late night hours.
+These consumers may be flexible students or elderly retirees who are night owls.
+They have heaters on in the winter but consume less energy in the summer.
+Winter usage may also be due to increased usage of heaters at night when they are up.
+
+_Group 4:_ Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
+They have a flat monthly profile, indicating the usage across all months are similar.
+It seems that the users in this group are more concerned about the comfort and quality of life than the cost of electricity as opposed to Group 1 and 3, who do not consume more electricity in the summer.
+
+\noindent The plotting scales are not displayed since we want to emphasise comparable shapes rather than scales.
+A customer in the cluster may have low daily or total energy usage, but their behavior may be quite similar to a customer with high usage.
+\noindent The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
+```
+
+```{r combined, fig.cap = "The distribution of electricity demand for the clusters across hour-of-day.
+The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
+Each cluster is characterised by unique shape across the granularity it is plotted against.
+For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters.
+This fact we will be re-established when we see the importance of each granularity throough the parallel coordinate plot.", eval = FALSE}
+
+combined <- hod_group + moy_group + wkndwday_group & theme(legend.position = "bottom")
+combined + plot_layout(guides = "collect")
+
+```
+
+```{r validation}
+data_validation <- (hod/24) %>% broom::tidy() %>% 
+  rename("hod" = "distance") %>% 
+  left_join((moy/12) %>% 
+              broom::tidy(), 
+            by = c("item1", "item2")) %>% 
+  rename("moy" = "distance") %>% 
+  left_join((wkndwday/2) %>% 
+              broom::tidy(), 
+            by = c("item1", "item2"))%>% 
+  rename("wkndwday" = "distance") %>% 
+  left_join(cluster_result, by = c("item1" = "customer_id")) %>% 
+  rename("group_item1" = "group") %>% 
+  left_join(cluster_result, by = c("item2" = "customer_id")) %>%  
+  rename("group_item2" = "group") %>% 
+  pivot_longer(3:5,names_to="gran",
+               values_to = "distance")
+```
+
+```{r validation-data}
+
+hod_cat <- suppressMessages(
+data_pick %>% 
+  dist_gran_cat(gran1 = "hour_day", response = "general_supply_kwh")) %>% 
+  mutate(gran = "hod")
+
+moy_cat <- suppressMessages(data_pick %>% 
+  dist_gran_cat(gran1 = "month_year", response = "general_supply_kwh"))%>%
+  mutate(gran = "moy")
+
+wkndwday_cat <- suppressMessages(data_pick %>% 
+  dist_gran_cat(gran1 = "wknd_wday", response = "general_supply_kwh"))%>% 
+  mutate(gran = "wnwd")
+
+data_validation_cat <- bind_rows(hod_cat, moy_cat, wkndwday_cat)
+
+gran_cat_dist <- data_validation_cat %>% 
+  left_join(cluster_result, by = c("customer_from" = "customer_id")) %>% 
+  rename("group_from" = "group") %>% 
+  left_join(cluster_result, by = c("customer_to" = "customer_id")) %>%  
+  rename("group_to" = "group") %>% 
+  group_by(gran,
+           category,
+           group_from, 
+           group_to) %>% 
+  summarise(sum = sum(distance),.groups = 'drop') %>% 
+  pivot_wider(names_from = group_to, values_from = sum) %>% 
+  mutate(distance = sum(`1`, `2`, `3`, `4`, `5`)) %>% 
+  select(-(4:8)) %>% 
+  arrange(-distance)
+
+```
+
+```{r hod-bubble}
+gran_cat_hod <-   gran_cat_dist %>% 
+  filter(gran=="hod")
+
+gran_cat_hod$category <- factor(gran_cat_hod$category, levels = 0:23)
+
+group_total <- gran_cat_hod %>% 
+  ungroup() %>% 
+    group_by(group_from) %>% 
+    summarise(group_total = sum(distance),.groups = 'drop')
+
+gran_cat_hod %>% 
+ungroup %>% 
+    left_join(group_total, by = "group_from") %>% 
+    mutate(cat_contibution = distance*100/group_total) %>% 
+    group_by(group_from) %>% 
+    arrange(-distance) %>% slice(c(1:4)) %>% 
+      ggplot(aes(group_from, category, color = as.factor(group_from))) +
+      geom_point(aes(size = distance)) +
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+ theme_light()
+```
+
+```{r hod-cont}
+data_pcp <- gran_cat_hod %>% 
+    ungroup %>% 
+    left_join(group_total, by = "group_from") %>% 
+    mutate(cat_contibution = distance*100/group_total) %>% 
+    group_by(group_from) %>% 
+    arrange(-distance) %>% 
+  select(group_from, category, cat_contibution) %>% 
+  pivot_wider(names_from = "category", values_from = "cat_contibution") %>% mutate(group_from = as.factor(group_from))
+
+parcoord <- GGally::ggparcoord(data_pcp ,
+                   columns = 2:ncol(data_pcp),
+                   groupColumn = "group_from",
+                   showPoints = TRUE, 
+                   alphaLines = 1,
+                   order = "anyClass",
+                   scale = "globalminmax"
+) + 
+  ggplot2::theme(
+    plot.title = ggplot2::element_text(size=10)
+  )+
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
+  theme(legend.position = "bottom") +
+  xlab("") +
+  ylab("wpd")+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black")) + theme_bw()
+   parcoord 
+```
+
+# Contribution of individual category for moy
+
+```{r moy-bubble}
+gran_cat_moy <-   gran_cat_dist %>% 
+  filter(gran=="moy")
+
+gran_cat_moy$category <- factor(gran_cat_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+group_total <- gran_cat_moy %>% 
+  ungroup() %>% 
+    group_by(group_from) %>% 
+    summarise(group_total = sum(distance),.groups = 'drop')
+
+gran_cat_moy %>% 
+ungroup %>% 
+left_join(group_total, by = "group_from") %>% 
+mutate(cat_contibution = distance*100/group_total) %>% 
+group_by(group_from) %>% 
+arrange(-distance) %>% slice(c(1:3)) %>% 
+    ggplot(aes(group_from, category, color = as.factor(group_from))) +
+      geom_point(aes(size = distance))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+ theme_light()
+
+```
+
+```{r moy-contr}
+
+data_pcp <- gran_cat_moy %>% 
+ungroup %>% 
+left_join(group_total, by = "group_from") %>% 
+mutate(cat_contibution = distance*100/group_total) %>% 
+group_by(group_from) %>% 
+arrange(-distance) %>% 
+  select(group_from, category, cat_contibution) %>% 
+  pivot_wider(names_from = "category", values_from = "cat_contibution") %>% mutate(group_from = as.factor(group_from))
+
+parcoord <- GGally::ggparcoord(data_pcp ,
+                   columns = 2:ncol(data_pcp),
+                   groupColumn = "group_from",
+                   showPoints = TRUE, 
+                   alphaLines = 1,
+                   order = "anyClass",
+                   scale = "globalminmax"
+) + 
+  ggplot2::theme(
+    plot.title = ggplot2::element_text(size=10)
+  )+
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
+  theme(legend.position = "bottom") +
+  xlab("") +
+  ylab("wpd")+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black")) + theme_bw()
+    parcoord
+
+```
+
+# Contribution of individual category for wknwday
+
+```{r wkndwday-bubble}
+gran_cat_wkndwday <-   gran_cat_dist %>% 
+  filter(gran=="wnwd")
+
+gran_cat_wkndwday$category <- factor(gran_cat_wkndwday$category, levels = c("Weekday", "Weekend"))
+
+group_total <- gran_cat_wkndwday %>% 
+  ungroup() %>% 
+    group_by(group_from) %>% 
+    summarise(group_total = sum(distance),.groups = 'drop')
+
+gran_cat_wkndwday %>% 
+ungroup %>% 
+left_join(group_total, by = "group_from") %>% 
+mutate(cat_contibution = distance*100/group_total) %>% 
+group_by(group_from) %>% 
+arrange(-distance) %>% slice(c(1)) %>% 
+ ggplot(aes(group_from, category, color = as.factor(group_from))) +
+ geom_point(aes(size = distance))+
+ scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+ theme_light()
+```
+
+# Show aggregate group behavior
+
 ```{r data-heatmap-hod-group}
+legend_title <- "group"
 
 data_group <- data_pick  %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result, by = c("customer_id" = "id"))
+  left_join(cluster_result, by = c("customer_id"))
 
@@
 data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
-
-# data_heatmap_hod_group <- data_heatmap_hod_group %>%
-#   mutate(group = case_when(group==1 ~"Group 2",
-#                            group==2 ~"Group 4",
-#                            group==3 ~"Group 1",
-#                            group==4 ~"Group 3"))
-
-# data_heatmap_hod_group$group <- factor(data_heatmap_hod_group$group, levels = c("Group 1", "Group 2", "Group 3", "Group 4"))
 
@@
              scales = "free_y",  
-             nrow = 4) + 
+             ncol = 5) + 
               #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
@@
   theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
+   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
+  scale_fill_manual(legend_title, values = c("#E69F00",
+  "#009E73","#0072B2", "#D55E00", "grey")) + 
   theme(legend.position = "bottom")
 
-#
-
-# 
-# data_pick_cust <- data_pick_cust %>% mutate(value = as.character(value))
-# 
-# data_pick_cust %>% left_join(cluster_result, by = c("value" = "id")) %>% group_by(group)
-```
-
-Characterization of clusters both statistically and qualitatively is an important stage of a cluster analysis.
-A potential way is to look at the findings from all the groups in graphs, and enhance our qualitative descriptions of the groupings.
-Figure \ref{fig:combined} shows the distribution of the summarized groups and help us to characterise each of the clusters.
-All of these may be validated with further information about the customer.
-
-<!-- We quantify them by producing statistics for each variable. We may look at the findings in graphs, and enhance our qualitative descriptions of the groupings. -->
-
-<!-- \noindent We discover 4 qualitative clusters of varying shapes in the distribution of all consumers in the first panel of Figure \ref{fig:combined}.
-Group-1 includes consumers who work 9-5, get up and conduct morning activities from 7-10am, and then depart.
-Then they return home in the evening to cook supper and perform other activities, giving the evening a greater peak than the morning.
-Group 2 is the group that rushes out of the house in the morning to get to work.
-They only return at night and do all activities at night, so there is no morning peak.
-The third one has a strong early morning and late night hours.
-These consumers may be flexible students or elderly retirees who are night owls.
-Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
-All of this may be validated with further information about the customer.
--->
-
-<!-- \noindent The second panel of Figure \ref{fig:combined} shows that month-of-year qualitative clusters are not as distinguishable as hour-of-day.
-Group 2 is the most distinct and uses the most power during the summer, possibly owing to the use of air conditioners.
-Group 4 has a flat profile, indicating no significant month-to-month changes.
-Groups 1 and 3 have heaters on in the winter but consume less energy in the summer.
-Since gas is not available in all of NSW LGAs, it is possible that customers' heater usage is recorded in electricity rather than gas.
--->
-
-<!-- \noindent The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
--->
-
-_Group 1:_ This includes consumers who work 9-5, get up and conduct morning activities from 7-10am, and then depart.
-Then they return home in the evening to cook supper and perform other activities, giving the evening a greater peak than the morning.
-These users in this cluster are insensitive to high temperature but have their heaters on in the winter months May-Aug.
-
-_Group 2:_  This is the group that rushes out of the house in the morning to get to work.
-They only return at night and do all activities at night, so there is no morning peak.
-In addition, the users in this cluster are insensitive to high temperature, but sensitive to lower temperature.
-
-_Group 3:_  This group a strong early morning and late night hours.
-These consumers may be flexible students or elderly retirees who are night owls.
-They have heaters on in the winter but consume less energy in the summer.
-Winter usage may also be due to increased usage of heaters at night when they are up.
-
-_Group 4:_ Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
-They have a flat monthly profile, indicating the usage across all months are similar.
-It seems that the users in this group are more concerned about the comfort and quality of life than the cost of electricity as opposed to Group 1 and 3, who do not consume more electricity in the summer.
-
-\noindent The plotting scales are not displayed since we want to emphasise comparable shapes rather than scales.
-A customer in the cluster may have low daily or total energy usage, but their behavior may be quite similar to a customer with high usage.
-\noindent The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
+```
 
 ```{r data-heatmap-moy-group}
-
-data_group <- data_pick %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result, by = c("customer_id" = "id"))
-
 data_heatmap_moy_group <- quantile_gran(data_group,
@@
              labeller = "label_value",
-             nrow = 4) +
+             ncol = 5) +
     theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
@@
   theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) + 
+   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
+  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
   theme(legend.position = "bottom") +
@@
 
-```{r data-heatmap-wkndwday}
-
-data_heatmap_wkndwday <- quantile_gran(data_pick, "wknd_wday", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
-  left_join(data_pick_cust, by = c("customer_id"="value"))
-
-data_heatmap_wkndwday$category <- factor(data_heatmap_wkndwday$category, levels = c("Weekday", "Weekend"))
-
-p1 <- data_heatmap_wkndwday %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=customer_id), fill = "lightblue") +
-  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
-  facet_wrap(design~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-             ncol = 6) +
-    theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) + xlab("wknd_wday") + ylab("demand (in Kwh)")  + theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) +
-  theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") ) +
-  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
-  theme(legend.position = "bottom")
-
-```
-
-```{r data-heatmap-wkndwday-group}
-data_group <- data_pick %>%
-  mutate(customer_id = as.character(customer_id)) %>%
-  left_join(cluster_result, by = c("customer_id" = "id"))
-
-data_group %>% filter(group==1) %>% 
-  prob_plot("hour_day",
-            "month_year", 
-            response = "general_supply_kwh", plot_type = "quantile")
-
-# data_heatmap_wkndwday_group <- quantile_gran(data_group,
-#                                   gran1="wknd_wday",
-#                                   quantile_prob_val = c(0.25, 0.5, 0.75),
-#                                   group="group") %>%
-#   pivot_wider(names_from = quantiles, values_from = quantiles_values)
-# 
-# data_heatmap_wkndwday_group$category <- factor(data_heatmap_wkndwday_group$category, levels = c("Weekday", "Weekend"))
-# 
-# data_heatmap_wkndwday_group$group <- paste("group", data_heatmap_wkndwday_group$group, sep = "-")
-# 
-# 
-# wkndwday_group <- data_heatmap_wkndwday_group %>%
-#   ggplot(aes(x = category)) +
-#   geom_ribbon(aes(ymin = `25%`,
-#                   ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
-#   geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
-#   facet_wrap(~group,
-#              #scales = "free_y",
-#              labeller = "label_value",
-#              nrow = 4)+
-#     #theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
-#     xlab("wknd-wday") +
-#   ylab("demand (in Kwh)") +
-#   theme_bw() +
-#    theme(axis.title.y=element_blank(),
-#         axis.text.y=element_blank(),
-#         axis.ticks.y=element_blank()) +
-#   theme(panel.spacing =unit(0, "lines")) +
-#   theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-#    scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-#   scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-#   theme(legend.position = "bottom") +
-#   theme(panel.spacing =unit(0, "lines")) +
-#   theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
-
-wkndwday_data<- data_group %>% create_gran("wknd_wday") 
+```{r combined-groups}
+combined <- (hod_group/moy_group) & theme(legend.position = "bottom")
+combined + plot_layout(guides = "collect")
+```
+
+```{r}
+wkndwday_data <- data_group %>% create_gran("wknd_wday") %>% 
+  create_gran("hour_day")
+
 ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]
@@
 wkndwday_group <- wkndwday_data%>% 
-  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  ggplot(aes(x=hour_day, y = general_supply_kwh)) +
   #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
   geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
-  #geom_boxplot(outlier.shape = NA) + 
+  #geom_boxplot(outlier.size = 1) + 
   coord_cartesian(ylim = ylim1*1.05)+
@@
   ylab("demand (in Kwh)") +
-   facet_wrap(~group, 
+   facet_grid(wknd_wday~group, 
              scales = "free_y", 
-             labeller = "label_both",
-             nrow = 4)+
+             labeller = "label_both") + 
   theme_bw() +
@@
   theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
-  scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +
+  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
+  scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
   theme(legend.position = "bottom") +
@@
   theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
-
-```
-
-```{r combined, fig.cap="The distribution of electricity demand for the clusters across hour-of-day.
-The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-Each cluster is characterised by unique shape across the granularity it is plotted against.
-For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters.
-This fact we will be re-established when we see the importance of each granularity throough the parallel coordinate plot."}
-
-combined <- hod_group + moy_group + wkndwday_group & theme(legend.position = "bottom")
-combined + plot_layout(guides = "collect")
-
+wkndwday_group
 ```
@@
 elec_pick <- elec_600_wpd %>% 
-  filter(customer_id %in% data_pick_cust$value)
+  filter(customer_id %in% data_pick_cust$customer_id)
 
@@
 
-# robust_scale <-  function(x){
-#   (x - stats::median(x))/(stats::IQR(x))
-# }
-
-#scaled_var <- apply(elec_pick_wide[-c(1)], 2, robust_scale) %>% as_tibble()
-
 scaled_var <- elec_pick_wide
@@
 
-group <- f%>% hclust (method = "ward.D") %>% cutree(k=4)
+group <- f%>% hclust (method = "ward.D") %>% cutree(k=5)
 
 cluster_result_wpd <- bind_cols(id = elec_pick_wide$customer_id, group = group)
-
-mds  = cmdscale(f)
-rownames(mds) = id
-colnames(mds) = c("mds1", "mds2")
-
-# ggplot(data = mds %>% as_tibble() %>% mutate(group = cluster_result_wpd$group), 
-#        aes(x = mds1, y = mds2, color = as.factor(group))) +
-#   geom_point() +
-#   geom_text(aes(label = id))
 
@@
                      columns = 3:5,
-                     aes(fill=group, color = group), alpha = 0.5) +
+                     aes(fill=group, 
+                     color = group)) +
   scale_fill_viridis_d(direction = -1) +
@@
   theme_light()
-
-  # 
-  # scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))
 
@@
 
-```{r data-heatmap-hod-new-wpd, fig.cap="The distribution of electricity demand for the selected 24 customers across hour-of-day.
-The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-According to our prototype selection method, each row represents a distinct shape of daily load, and all customers in the same row should have similar daily profile.After clustering these consumers, all customers with the same colour represent the same group.
-Except for customer id 8269176, there is unanimity across design and groups.
-Even though their daily form resembles Group 2, our clustering approach places them in Group 3 (which is design 4 in our case).
-Because our method uses hod, moy, and wkndwday, there may be some mismatches depending on only one variable."}
-
-cluster_result_wpd$id <- as.character(cluster_result_wpd$id)
-
-data_group <- data_pick %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result_wpd, by = c("customer_id" = "id"))
-
-data_heatmap_hod %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result_wpd, by = c("customer_id" = "id")) %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`,
-                  group=customer_id,
-                  fill = as.factor(group)),
-              alpha = 0.5) +
-  geom_line(aes(y = `50%`,
-                group=customer_id, 
-                color = as.factor(group)), size = 1) +
-  facet_wrap(design~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-             ncol = 6) +
-  theme_light() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) +
-    theme(strip.text = element_text(size = 10,
-                                    margin = margin(b = 0, t = 0))) +
-  ylab("demand (in Kwh)")  + 
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ 
-  #theme() strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") ) +
-  scale_x_discrete(breaks = seq(0,23, 4)) +
-  scale_fill_viridis_d(direction = -1) +
-  scale_color_viridis_d(direction = -1) + 
-  theme(legend.position = "bottom") +
-  xlab("hour-of-day") 
-```
-
-```{r data-heatmap-moy-wpd, fig.cap = "The distribution of electricity demand for the selected 24 customers across month-of-year.
-The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-Intriguingly, customer id 8269176 appears to be in the appropriate group (Group 3) because its monthly profile is more similar to Group 3 than Design 4.
-So, while our clustering approach failed to place the customer in the correct hour-of-day group, it did so for month-of-year.
-When contrasting the moy to hod profile, there are greater behavioural differences across customers within a group."}
-
-data_heatmap_moy <- quantile_gran(data_pick, "month_year", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
-  left_join(data_pick_cust, by = c("customer_id"="value"))
-
-data_heatmap_moy$category <- factor(data_heatmap_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
-
-data_heatmap_moy %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result_wpd, by = c("customer_id" = "id")) %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, 
-                  group=customer_id, 
-              fill = as.factor(group)), alpha = 0.5) +
-  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
-  facet_wrap(design~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-             ncol = 6) +
-    theme(strip.text = element_text(size = 10,
-                                    margin = margin(b = 0, t = 0))) + xlab("month-of-year") +
-  ylab("demand (in Kwh)")  +
-  theme_light() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + 
-  theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm")) + scale_fill_viridis_d(direction = -1) +
-  scale_color_viridis_d(direction = -1) + 
-  xlab("month-of-year") +
-  theme(axis.text.x = element_text(size=5)) +
-  theme(legend.position = "bottom")
-```
-
-```{r data-heatmap-hod-group-wpd}
-
-cluster_result_wpd$id <- as.character(cluster_result_wpd$id)
-
-data_group <- data_pick %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result_wpd, by = c("customer_id" = "id"))
-
-data_heatmap_hod_group <- quantile_gran(data_group,
-                                  gran1="hour_day",
-                                  quantile_prob_val = c(0.25, 0.5, 0.75),
-                                  group="group") %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values)
-
-data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
-
-# data_heatmap_hod_group <- data_heatmap_hod_group %>%
-#   mutate(group = case_when(group==1 ~"Group 2",
-#                            group==2 ~"Group 4",
-#                            group==3 ~"Group 1",
-#                            group==4 ~"Group 3"))
-
-# data_heatmap_hod_group$group <- factor(data_heatmap_hod_group$group, levels = c("Group 1", "Group 2", "Group 3", "Group 4"))
-
-data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")
-
-hod_group_wpd <- data_heatmap_hod_group %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`,
-                  group=group,
-                  fill = as.factor(group), alpha = 0.5),
-              alpha = 0.5) +
-  geom_line(aes(y = `50%`,
-                group=group, 
-                color = as.factor(group)), size = 1)+
-  facet_wrap(~group, 
-             scales = "free_y",  
-             nrow = 4) + 
-              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
-  ylab("demand (in Kwh)") + 
-  theme_light() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank())+
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
-  scale_x_discrete(breaks = seq(1, 24, 3))+ 
-  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm"))+ scale_fill_viridis_d(direction = -1) +
-  scale_color_viridis_d(direction = -1) + 
-  theme(legend.position = "bottom")
-
-#
-
-# 
-# data_pick_cust <- data_pick_cust %>% mutate(value = as.character(value))
-# 
-# data_pick_cust %>% left_join(cluster_result, by = c("value" = "id")) %>% group_by(group)
-```
-
-```{r data-heatmap-moy-group-wpd}
-
-data_group <- data_pick %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result_wpd, by = c("customer_id" = "id"))
-
-data_heatmap_moy_group <- quantile_gran(data_group,
-                                  gran1="month_year",
-                                  quantile_prob_val = c(0.25, 0.5, 0.75),
-                                  group="group") %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values)
-
-data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
-
-data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")
-
-moy_group_wpd <- data_heatmap_moy_group %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
-  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
-  facet_wrap(~group, 
-             scales = "free_y", 
-             labeller = "label_value",
-             nrow = 4) +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
-  ylab("demand (in Kwh)") +
-  theme_light() + 
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + 
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-scale_fill_viridis_d(direction = -1) +
-  scale_color_viridis_d(direction = -1) + 
-  theme(legend.position = "bottom") +
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
-```
-
-```{r data-heatmap-wkndwday-group-new2-wpd}
-data_group <- data_pick %>% 
-  mutate(customer_id = as.character(customer_id)) %>% 
-  left_join(cluster_result_wpd, by = c("customer_id" = "id"))
-
-data_heatmap_wkndwday_group <- quantile_gran(data_group,
-                                  gran1="wknd_wday",
-                                  quantile_prob_val = c(0.25, 0.5, 0.75),
-                                  group="group") %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values)
-
-data_heatmap_wkndwday_group$category <- factor(data_heatmap_wkndwday_group$category, levels = c("Weekday", "Weekend"))
-
-data_heatmap_wkndwday_group$group <- paste("group", data_heatmap_wkndwday_group$group, sep = "-")
-
-wkndwday_data<- data_group %>% create_gran("wknd_wday") 
-ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]
-
-wkndwday_group_wpd <- wkndwday_data%>% 
-  ggplot(aes(x=wknd_wday, y = general_supply_kwh))+
-  geom_boxplot(aes(fill = as.factor(group),color = as.factor(group)),alpha = 0.5) + 
-  coord_cartesian(ylim = ylim1*1.05) +
-  #ggplot(aes(x = category)) + 
-  # geom_ribbon(aes(ymin = `25%`, 
-  #                 ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
-  # geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
-  facet_wrap(~group, 
-             scales = "free_y", 
-             labeller = "label_both",
-             nrow = 4)+
-    #theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) 
-    xlab("wknd-wday") + 
-  ylab("demand (in Kwh)") +
-  theme_light() + 
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + 
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(plot.margin = margin(0, 0, 0, 0, "cm"))+ scale_fill_viridis_d(direction = -1) +
-  scale_color_viridis_d(direction = -1) + 
-  theme(legend.position = "bottom") +
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
-```
-
-```{r combined-wpd, fig.cap="The distribution of electricity demand for the clusters across hour-of-day.
-The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-Each cluster is characterised by unique shape across the granularity it is plotted against.
-For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters.
-This fact we will be re-established when we see the importance of each granularity throough the parallel coordinate plot."}
-
-combined <- hod_group_wpd + moy_group_wpd + wkndwday_group_wpd & theme(legend.position = "bottom")
-combined + plot_layout(guides = "collect")
-
-```
-
 A parallel coordinate plot with the three significant cyclic granularities used for wpd-based clustering.
@@
 
-## Classifying the 356 customers
-
 # Discussion {#sec:discussion}
@@
 -->
+
+```{r}
+library(ochRe)
+  scale_colour_ochre(
+    palette="nolan_ned")
+```

### 2021-10-27 main.Rmd
moy combined caption added

@@
              scales = "free_y", 
-             ncol = 24) +
+             ncol = 6) +
     ylab("demand (in Kwh)") +
@@
   scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))
-
-(moy_ind_design/moy_ind_group)*theme(legend.position = "none")
-
+```
+
+The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
+
+```{r moy-combined, fig.cap = "The energy demand distribution for the prototype designs (left) and clustering (right) are shown.
+The customer grouping by design and the clustering approach largely match.
+To compare one-to-one, the customers are organised in the same order on both sides.
+Except for some customers, design 1, 2, 3, 4 and 5 match to group 4, 5, 1, 3, 2.
+Except in group 3, where consumers are categorised by greater usage in the middle of the year, which corresponds to winter in Australia.
+Since the designs are selected only based on 50th percentile, the clustering approach outperforms the design in a few circumstances."}
+
+(moy_ind_design+moy_ind_group)*theme(legend.position = "bottom")
 ```

### 2021-10-27 main.Rmd
js-based first para update

@@
              scales = "free_y",
-             ncol=24) + 
+             ncol=6) + 
  theme_application() +
@@
              scales = "free_y",
-              ncol=24) + theme_application()+
+              ncol=6) + theme_application()+
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+
@@
 # 
-# 
-(hod_ind_design/hod_ind_group)* theme(legend.position = "none")
+```
+
+```{r hod-combined, fig.cap = "The energy demand distribution for the prototype designs (left) and clustering (right) is given for each hour of the day.
+Customer design grouping and clustering strategies complement one other well.
+For a one-to-one comparison, both sides have the same order of consumers.
+Design-4 has a low morning peak and a strong evening peak.
+Because other members of group-1 have higher late-night usage, group-2 adopts this design and assigns customer B to it rather than group-1."}
+(hod_ind_design + hod_ind_group)* theme(legend.position = "bottom")
 
@@
 
-The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-
 ```{r moy-combined, fig.cap = "The energy demand distribution for the prototype designs (left) and clustering (right) are shown.
-The customer grouping by design and the clustering approach largely match.
-To compare one-to-one, the customers are organised in the same order on both sides.
-Except for some customers, design 1, 2, 3, 4 and 5 match to group 4, 5, 1, 3, 2.
-Except in group 3, where consumers are categorised by greater usage in the middle of the year, which corresponds to winter in Australia.
-Since the designs are selected only based on 50th percentile, the clustering approach outperforms the design in a few circumstances."}
+The customer grouping by design and the clustering approach largely match, except for some exceptions.
+For example, in group 3, consumers are categorised by greater usage in the middle of the year, which corresponds to winter in Australia.
+This is a better grouping in terms of similar month-of-year pattern than originally proposed by the prototype design."}
 
@@
 
-The distribution of electricity demand for the selected 24 customers across hour-of-day is shown in \ref{fig:data-heatmap-hod-new}.
-The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-According to our prototype selection method, each row represents a distinct shape of daily load, and all customers in the same row should have similar daily profile.
-After clustering these consumers, all customers with the same color represent the same group.
-Except for customer id 8269176, there is unanimity across design and groups.
-Even though their daily form resembles Group 2, our clustering approach places them in Group 3 (which is design 4 in our case).
-Because our method uses hod, moy, and wkndwday, there may be some mismatches depending on only one variable.
-
-\noindent The distribution of electricity demand for the selected 24 customers across month-of-year is shown in Figure \ref{fig:data-heatmap-moy}.
-The median is represented by a line and shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-Intriguingly, customer id 8269176 appears to be in the appropriate group (Group 3) because its monthly profile is more similar to Group 3 than Design 4.
-So, while our clustering approach failed to place the customer in the correct hour-of-day group, it did so for month-of-year.
-When contrasting the `moy` to `hod` profile, there are greater behavioral differences across customers within a group, which implies that `hod` has the minimum within-cluster variation.
+The $24$ prototypes are clustered using the methodology described in \ref{Sec:methodology}.
+The distribution of electricity demand for the selected $24$ customers across hour-of-day and month-of-year are shown in the right panel of Figures \ref{fig:hod-combined} and \ref{fig:moy-combined} respectively.
+The median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.
+Customers are placed in the same order on both sides for both figures to simplify one-to-one comparison.
+All customers with the same color represent the same design (left) or cluster (right).
+Overall, the clustering method has been able to capture the shapes in each designs quite well, irrespective of the fact that the designs are only selected based on the $50^{th}$ percentile with Designs $1$, $2$, $3$, $4$, and $5$ correspond to groups $4$, $5$, $1$, $3$, $2$ and, in most situations.Because our method uses distribution across hod, moy, and wknd-wday, there may be some mismatches from the designs; nonetheless, visualising individual customers can help us determine whether the grouping by clustering is better than the planned grouping through designs.
+In a few circumstances, for example, as shown in Figures \ref{fig:hod-combined} and \ref{fig:moy-combined}, the clustering approach outperforms the design.
 

### 2021-10-28 main.Rmd
long application

@@
 theme_characterisation <- function() {
-theme_bw() + 
-      theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) +  + theme_bw() +
+
+  theme_bw() + # seeting theme
+    theme(strip.text = element_text(size = 10,
+                                    margin = margin(b = 0, t = 0))) + # narrow facet space
    theme(axis.title.y=element_blank(),
         axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) +
-  theme(panel.spacing =unit(0, "lines")) + 
-    theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+        axis.ticks.y=element_blank()) + # no axis ticks 
+  theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
+    theme(axis.text.x = element_text(angle=90, hjust=1, size = 10)) + # rotate the x-axis text
+    theme(legend.position = "bottom")+
   theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
-  theme(legend.position = "bottom")
+  theme(axis.text.x = element_text(size=5))
 }
@@
 In a few circumstances, for example, as shown in Figures \ref{fig:hod-combined} and \ref{fig:moy-combined}, the clustering approach outperforms the design.
+The plotting scales are not displayed since we want to emphasize comparable shapes rather than scales.
+A customer in the cluster may have low daily or total energy usage, but their behavior may be quite similar to a customer with high usage.
+\noindent
+
+<!-- The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
+-->
 
@@
 
-_Group 1:_ This includes consumers who work 9-5, get up and conduct morning activities from 7-10am, and then depart.
-Then they return home in the evening to cook supper and perform other activities, giving the evening a greater peak than the morning.
-These users in this cluster are insensitive to high temperature but have their heaters on in the winter months May-Aug.
-
-_Group 2:_  This is the group that rushes out of the house in the morning to get to work.
+_Group 1:_ This group a strong early morning and late night hours. These consumers may be flexible students or elderly retirees who are night owls.
+Their day time usage has high variability.
+They have heaters on in the winter but consume less energy in the summer.
+Winter usage may also be due to increased usage of heaters at night when they are up, but typically very less variability in other months.
+
+_Group 2:_ They appear to work 9-5, get up and perform morning activities from 7-10, and then go.
+Evenings are busier than mornings as people return home to cook supper and perform other activities.
+During the fall (March-June), these consumers' average energy use declines, before rising again during the winter.
+However, energy behaviour varies more in the winter than in the summer and a bit in the fall months.
+
+_Group 3:_ Presence of children or stay-at-home parents is indicated by the group's almost equivalent morning, afternoon and evening median use and variability across all hours.
+They have low usage for winter months (or maybe due to use of gas heaters, it is not reflected in electricity usage).
+It seems that the users in this group are more concerned about the comfort and quality of life than the cost of electricity as we see their electricity demand peaking up in the spring and summer.
+
+_Group 4:_ These users have an opposite monthly profile to Group 3, with more usage during summer months (probably due to increased use of air conditioners) and demand decreasing across the year with least demand in winter months.
+The users in this cluster are sensitive to high temperature, but insensitive to lower temperature or their heater usage is not reflected in their electricity usage.
+
+_Group 5:_ Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
+This group is more typical in their monthly energy behavior with more energy usage in winter and moderately decreasing for fall and spring months.
+The pattern in the variability is similar to that of the median.
+
+<!-- _Group 5:_ Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
+They have a flat monthly profile except for winter months, with higher median usage and variability than typically seen in other seasons.
+-->
+
+<!-- It seems that the users in this group are more concerned about the comfort and quality of life than the cost of electricity as opposed to Group 1 and 3, who do not consume more electricity in the summer.
+-->
+
+<!-- insensitive to high temperature but have their heaters on in the winter months May-Aug. -->
+
+<!-- _Group 2:_  This is the group that rushes out of the house in the morning to get to work.
 They only return at night and do all activities at night, so there is no morning peak.
 In addition, the users in this cluster are insensitive to high temperature, but sensitive to lower temperature.
-
-_Group 3:_  This group a strong early morning and late night hours.
+-->
+
+<!-- _Group 3:_  This group a strong early morning and late night hours.
 These consumers may be flexible students or elderly retirees who are night owls.
@@
 Winter usage may also be due to increased usage of heaters at night when they are up.
-
-_Group 4:_ Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
+-->
+
+<!-- _Group 4:_ Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
 They have a flat monthly profile, indicating the usage across all months are similar.
 It seems that the users in this group are more concerned about the comfort and quality of life than the cost of electricity as opposed to Group 1 and 3, who do not consume more electricity in the summer.
-
-\noindent The plotting scales are not displayed since we want to emphasise comparable shapes rather than scales.
-A customer in the cluster may have low daily or total energy usage, but their behavior may be quite similar to a customer with high usage.
-\noindent The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
-```
-
-```{r combined, fig.cap = "The distribution of electricity demand for the clusters across hour-of-day.
-The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-Each cluster is characterised by unique shape across the granularity it is plotted against.
-For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters.
-This fact we will be re-established when we see the importance of each granularity throough the parallel coordinate plot.", eval = FALSE}
-
-combined <- hod_group + moy_group + wkndwday_group & theme(legend.position = "bottom")
-combined + plot_layout(guides = "collect")
-
+-->
+
+```{r data-heatmap-hod-group-new}
+legend_title <- "group"
+
+data_group <- data_pick  %>% 
+  left_join(cluster_result, by = c("customer_id"))
+
+data_heatmap_hod_group <- quantile_gran(data_group,
+                                  gran1="hour_day",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
+
+# data_heatmap_hod_group <- data_heatmap_hod_group %>%
+#   mutate(group = case_when(group==1 ~"Group 2",
+#                            group==2 ~"Group 4",
+#                            group==3 ~"Group 1",
+#                            group==4 ~"Group 3"))
+
+# data_heatmap_hod_group$group <- factor(data_heatmap_hod_group$group, levels = c("Group 1", "Group 2", "Group 3", "Group 4"))
+
+data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")
+
+hod_group <- data_heatmap_hod_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=group,
+                  fill = as.factor(group), alpha = 0.5),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=group, 
+                color = as.factor(group)), size = 1)+
+  facet_wrap(~group, 
+             scales = "free_y",  
+             nrow = 5) + 
+              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
+  ylab("demand (in Kwh)") + 
+  theme_bw() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank())+
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ 
+  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
+  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
+  theme(legend.position = "bottom")
+
+```
+
+```{r data-heatmap-moy-group-new}
+data_heatmap_moy_group <- quantile_gran(data_group,
+                                  gran1="month_year",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")
+
+moy_group <- data_heatmap_moy_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow = 5) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
+  ylab("demand (in Kwh)") +
+  theme_bw() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank()) + 
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
+   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
+  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
+  theme(legend.position = "bottom") +
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
 ```
@@
 
-# Contribution of individual category for moy
-
-```{r moy-bubble}
+```{r moy-bubble, eval = FALSE}
 gran_cat_moy <-   gran_cat_dist %>% 
@@
 
-```{r moy-contr}
+```{r moy-contr, eval = FALSE}
 
@@
 
-# Contribution of individual category for wknwday
-
-```{r wkndwday-bubble}
+```{r wkndwday-bubble, eval = FALSE}
 gran_cat_wkndwday <-   gran_cat_dist %>% 
@@
 ```
-
-# Show aggregate group behavior
 

### 2021-10-28 selecting_customers.Rmd
working on application
long application

@@
 ```{r anchor}
-kopt <- 6# number of clusters
+kopt <- 5# number of clusters
 ```
@@
 ```{r hod-data}
+# data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
+#   mutate(customer_id = as.character(customer_id)) %>% 
+#   filter(customer_id %in% data_pick_cust$customer_id) %>% 
+#   gracsr::scale_gran( method = "robust",
+#                       response = "general_supply_kwh")
+
 data_hod <- quantile_gran(data_pick,
@@
              scales = "free_y",
-             ncol=24) + theme_characterisation()+
+             ncol=6) + theme_characterisation()+
   xlab("hour-of-day") +
@@
              scales = "free_y",
-              ncol=24) + theme_characterisation()+
+              ncol=6) + theme_characterisation()+
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+
@@
 # 
-(hod_ind_design/hod_ind_group)* theme(legend.position = "none")
+(hod_ind_design+hod_ind_group)* theme(legend.position = "bottom")
 
@@
 
-(moy_ind_design/moy_ind_group)*theme(legend.position = "none")
+(moy_ind_design + moy_ind_group)*theme(legend.position = "none")
 
@@
              labeller = "label_value",
-              ncol = 24)  +
+              ncol = 6)  +
   theme_characterisation() +
@@
              labeller = "label_value",
-             ncol = 24)  +
+             ncol = 6)  +
   theme_characterisation() +
@@
 
-(wkndwday_ind_design / wkndwday_ind_group)* theme(legend.position = "none")
+(wkndwday_ind_design + wkndwday_ind_group)* theme(legend.position = "none")
 ```
@@
 
+# data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
+#   mutate(customer_id = as.character(customer_id)) %>% 
+#   filter(customer_id %in% data_pick_cust$customer_id) %>% 
+#   gracsr::scale_gran( method = "robust",
+#                       response = "general_supply_kwh")
+
 data_group <- data_pick  %>% 

### 2021-10-28 main_application.Rmd
application rewrite

---
title: "main_application"
date: "28/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r mytheme-application}
theme_application <- function() {

  theme_light() + # setting theme
    #theme(strip.text = element_text(margin = margin(b = 0, t = 0))) + # narrow facet space
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) + # no axis ticks 
  theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
    theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) + # no x-axis labels to further reduce the gap between facets
    #theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
  #theme(axis.text.x = element_text(size=5)) +
    theme(strip.background = element_blank(),
  strip.text.x = element_blank())
}
```
## What data you have, what are their features. What you doing in this section.

The use of our methodology is illustrated on smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO. It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total. No additional information about the customers is used for the application.

The linear view of the time series data generally has too many measurements all squeezed in that representation, it hinders us to discern any repetitive behavioral pattern for even one customers (let alone many customers together). In most cases, electricity data will have multiple seasonal patterns like daily, weekly or annual. We do not learn about these repetitive behaviors from the linear view. Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior. The raw data for these consumers is of unequal length, with varying start and finish dates. Because our proposed methods evaluate probability distributions rather than raw data, neither of these data features would pose any threat to our methodology unless they contained any structure or systematic patterns. Additionally, there were missing values in the database but further investigation revealed that there is no structure in the missingness (see Supplementary paper for raw data features and missingness).

## Prototype selection

<!-- Why instance selection -->
Supervised learning uses a training set of known information to categorize new events. Instance selection (@olvera2010review) is a method of rejecting instances that are not helpful for classification. This is equivalent to subsetting the population along all dimensions of interest such that the sampled data reflects the underlying distribution's primary features. Instance selection in unsupervised learning has received little attention in the literature, yet it could be a useful tool for evaluating model or method performance. @Fan2021-bq proposes one such process that picks related examples (neighbours) for each instance (anchor) and considers them as the same class. In this part, consumers with prototype behaviours are chosen to serve as study population for our suggested methodology.

_Data filtering and variable selection_

- Choose a smaller subset of randomly selected $600$ customers with no implicit missing values for 2013.

- Obtain $wpd$ for all cyclic granularities considered for these customers. It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weeknd/weekday) are coming out to be significant for most customers. We use these three granularities while clustering.

- Remove customers whose data for an entire category of a significant granularity is empty. For example, a customer who does not have data for an entire month is excluded because their monthly behaviour cannot be analyzed.

- Remove customers whose energy consumption is 0 in all deciles. These are the clients whose consumption is likely to remain essentially flat and with no intriguing repeated patterns that we are interested in studying.

<!-- - can be done in several ways -->

\noindent There are several ways to approach the prototype selection. Use any dimensionality reduction techniques like MDS or PCA to project the data into a 2-dimensional space. Then pick a few "anchor" customers who are far apart in 2D space and pick a few neighbors for each. Unfortunately, this does not assure that consumers with significant patterns across all variables are chosen. We perform a linked tour with t-SNE layout and (liminal) to identify customers who are more likely to have distinct patterns across the variables studied. Tours can help us see separation between variables that was not obvious in the single variable display. Please see the Supplementary article for further details on how the prototypes are chosen. Figure \ref{fig:protypes} shows the raw time plot, distribution across `hod`, `moy` and `wkndwday` for the set of chosen $24$ customers. Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution. These $24$ prototypes are clustered using the methodology described in \ref{Sec:methodology} to see if the grouping is useful.

```{r prototype-data-pick}
quantile_prob_graph <- c(0.25, 0.5, 0.75)

# data_pick_one <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
# data_pick_two <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
# data_pick_three <- c(8221762, 8273636, 10359424, 8232822)%>% as_tibble %>% set_names("customer_id")

data_pick_one <- c(8541744, 9355808, 8603880, 8619309, 10542667) %>% as_tibble %>% set_names("customer_id")
#data_pick_two <- c(8688242, 8643837, 8184707, 10534355, 8684420) %>% as_tibble%>% set_names("customer_id")
data_pick_three <- c(9792072, 8589936, 8454235, 10692366, 8603828)%>% as_tibble%>% set_names("customer_id")
data_pick_four <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
data_pick_five <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
data_pick_six <- c(8221762, 8273636, 10359424, 8232822, 11450499)%>% as_tibble %>% set_names("customer_id")

```

```{r assemble}
data_pick_cust <- bind_rows(
data_pick_one, 
#data_pick_two, 
data_pick_three,
data_pick_four,
data_pick_five, 
data_pick_six,
.id = "design") %>% 
  mutate(customer_id = as.character(customer_id))

```

```{r data-pick}
data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
  mutate(customer_id = as.character(customer_id)) %>% 
  dplyr::filter(customer_id %in% data_pick_cust$customer_id) %>% 
  gracsr::scale_gran( method = "robust",
                      response = "general_supply_kwh")
```

<!-- - can be done in several ways -->

<!-- - show raw data of moy, hod, wkndwday patterns for them -->

```{r hod-moy-wkndwday plots, fig.cap="The distribution across moy, hod and wkndwday for the selected designs. Few are similar in their hod pattern, while others are similar in moy behavior. Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most households, there is one household for whom weekend behavior is standing out from the rest."}

data_hod <- quantile_gran(data_pick,
                                  "hour_day",
                                  quantile_prob_val = quantile_prob_graph) %>% 
  pivot_wider(names_from = quantiles,
              values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"))

#data_heatmap_hod$customer_id = factor(data_heatmap_hod$customer_id, levels = data_pick_cust$value)
data_hod$category <- factor(data_hod$category, levels = 0:23)

hod_ind_design <- data_hod %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=customer_id),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=customer_id), 
            size = 1) +
  facet_wrap(~customer_id, 
             scales = "free_y",
             nrow = 24) + 
 theme_application() +
  xlab("hour-of-day") +
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  scale_x_discrete(breaks = seq(0, 23, 3))

data_moy <- quantile_gran(data_pick,
                                  "month_year", 
                                  quantile_prob_val = quantile_prob_graph) %>% 
  pivot_wider(names_from = quantiles, 
              values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"))

data_moy$category <- factor(data_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

moy_ind_design <- data_moy %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, 
                  group=customer_id), 
              alpha = 0.5) +
  geom_line(aes(y = `50%`, group=customer_id), size = 1) +
  facet_wrap(~customer_id, 
             scales = "free_y", 
             nrow = 24) +
    ylab("demand (in Kwh)") +
    xlab("month-of-year")  +
  theme_application()

data_wkndwday <- data_pick  %>%
   create_gran("wknd_wday")  %>% 
  left_join(data_pick_cust, by = c("customer_id"))

ylim1 = boxplot.stats(data_wkndwday$general_supply_kwh)$stats[c(1, 5)]

wkndwday_ind_design <- data_wkndwday%>% 
  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
  #lvplot::geom_lv(aes(fill = as.factor(design), 
   #                   color = as.factor(design)), k=5, alpha = 0.5) +
  geom_boxplot(alpha = 0.5, fill = "black")+
  #geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = ylim1*1.05)+
  facet_wrap(~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
              nrow = 24)  +
  stat_summary(
    fun = median,
    geom = 'line',
    aes(group = 1), size = 1, color = "black")+
  theme_application()

(hod_ind_design + moy_ind_design + wkndwday_ind_design)

```

## Clustering

### Using JS-based distances

The $24$ prototypes are clustered using the methodology described in \ref{Sec:methodology}. The distribution of electricity demand for the selected $24$ customers across hour-of-day and month-of-year are shown in the right panel of Figures \ref{fig:hod-combined} and \ref{fig:moy-combined} respectively. The median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$. Customers are placed in the same order on both sides for both figures to simplify one-to-one comparison. All customers with the same color represent the same design (left) or cluster (right). The plotting scales are not displayed since we want to emphasize comparable shapes rather than scales. A customer in the cluster may have low daily or total energy usage, but their behavior may be quite similar to a customer with high usage.

```{r combined-groups-js, fig.cap = The distribution of electricity demand for the clusters across hour-of-day, month-of-year and wknd-wday. The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile. Group 2 and 5 have a stronger hour-of-day pattern, while group 1, 3, 5 have a month-of-year pattern. For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters."}
```

### Using wpd-based distances

```{r pcp-wpd}
```

A parallel coordinate plot with the three significant cyclic granularities used for wpd-based clustering. The variables are sorted according to their separation across classes (rather than their overall variation between classes). This means that $moy$ is the most important variable in distinguishing the designs followed by $hod$ and $wkndwday$. It can be observed that clusters are well separated by $moy$, while $hod$ and $wkndwday$ are not useful distinguishing the clusters produced with this clustering method. The parallel coordinate plot ranks the variables in order of importance, indicating that the month-of-year is the most important in identifying clusters, whereas wkdn-wday is the least significant and has the least variability among the three variables. However, there is only one customer who has significant $wpd$ across $wkndwday$ and stands out from the rest of the customers. The ggpairs plot also shows five distinct clusters across the $moy$.

- Discussion

As with any clustering method, things become much more complicated when we consider a larger data set with more uncertainty. The methodology run on several customers together might not be useful to given distinct shapes across granularities. Moreover, the groupings from two different appraoches lead to different results, both of which are useful but needs careful considering the context before choosing one over another. Also, note that the customers chosen here do not have a weekend-weekday effect, which might not be true for all customers in the data set. Ideally there are other groups of customers in the entire dataset which can act as prototypes in our problem. Ideally, if these prototypes are not outliers, we can use them for a classification problem for external validation of the clustering problem.

Notice that while looking for these customers

(can be used as classification method, problems in handling large dataset)

### 2021-10-28 main.Rmd
rewrite application

@@
   theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
-    theme(axis.ticks.x = element_blank(),
-        axis.text.x = element_blank()) + # no x-axis labels to further reduce the gap between facets
+    #theme(axis.ticks.x = element_blank(),
+     #   axis.text.x = element_blank()) + # no x-axis labels to further reduce the gap between facets
     #theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
   theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+    theme(axis.text.x = element_text(angle=90, hjust=1, size = 10))+
+      theme(plot.background = element_blank(), panel.grid = element_blank())+
   #theme(axis.text.x = element_text(size=5)) +
@@
 Each of these series correspond to a single customer.
-For each customer, the energy consumption is available at fine temporal resolution (every 30 minutes) for a long period of time (~ 2 years) resulting in 27,000 (median) observations for each customer.
+For each customer, the energy consumption is available at fine temporal resolution (every 30 minutes) for a long period of time (~ 2 years).
 Some customers' electricity use may be unavailable owing to power outages or improper recording, resulting in implied missing numbers in the database.
@@
 Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.
-
-```{r raw-data-50, out.width = "100%", fig.cap="The raw data for 50 households are shown.
-It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed.
-No insightful behavioral pattern could be discerned from this view other than when the customer is not at home."}
-
-knitr::include_graphics("figs/raw_plot_cust.png") # look at smart-meter.R for the code
-```
 
@@
 
-<!-- From this set we select $4$ "anchor" customers which are far apart from each other and $5$ neighboring customers for each of these anchors.
-These selections were done using the granularity $hod$ space.
-It is important to note that when we use our proposed methodologies, it is based on all dimensions `hod`, `moy` and `wkndwday`.Fig \ref{fig:mds} shows the MDS of these 356 customers in a 2D space basis their distance on individual granularities and when all of them are combined.
-Our methodolgies are run on these $24$ customers, which act as a way to evaluate the proposed methodologies.
--->
-
-S1. Robust scaling is applied to each customer.
-
-S2.
-$50^{th}$ percentile for each category for each granularity is obtained for each customers.
-So we have a data structure with  $356$ rows and $(24 + 12 + 2)$ variables corresponding to $50^{th}$ percentile for each hour-of-day, month-of-year and weekend-weekday.
-
-S3. Apply principal components and restrict the results down to the first six principal components (which makes up approximately 85% of the variance explained in the data) to use with the grand tour.
-
-S4.
-Run t-SNE using the default arguments on the complete data (sets the perplexity to equal 30 and performs random initialisation).
-We then create a linked tour with t-SNE layout with liminal as shown in Figure 4.
-
-S5.
-We inspect of the subspace generated by the set of low-dimensional projections in tour by looking for a simplex shape while the visualization moves from one basis to another.
-When we brush the corners of the simplex, we find they fall on the edge of the t-SNE point cloud.
-Hall, Marron, and Neeman (2005) have shown that in the extreme case of high-dimension, low-sample size data, observations are on the vertices of a simplex.
-
-This is because in high-dimensional data analysis the curse of dimensionality reasons that points tend to be far away from the center of the distribution and on the edge of high-dimensional space.
-Contrary to this, is that projected data tends to clump at the center.
-
-S6. These points should ideally correspond to different behavior with respect to all the variables considered while running PCA.
-
-<!-- First, we select the customers which do not have any implicit missing values and filtered their data for the year 2013.
-From this set, we randomly sample a set of $600$ customers.
-We obtain the $wpd$ for all cyclic granularities considered for these customers and found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weeknd/weekday) are coming out to be significant for most customers.
-This implies that for most customers, there is some interesting pattern across these three granularities.
-Potentially this could be done for the entire data set.
-This step is time consuming and hence has been only run for the $600$ sampled customers.
-As a second step, we remove customers for which data in any category for the significant granularities are empty.
-For example, in this data set, if a customer do not have data for an entire month, they have been removed as their monthly behavior could not be studied in that case.
-Further, we also remove customers for which all the deciles of the energy consumption is zero.
-These are the customers whose consumption remain mostly flat and is expected to have no interesting repetitive patterns that is our interest of study.
--->
-
-```{r mds, out.width="100%", fig.cap="Instance selection using tours and projecting the points in a lower dimensional tsne cloud."}
-knitr::include_graphics("figs/liminal_view.png")
-```
-
-<!-- However, if `hod` is and not only `hod`, but if `hod` is significant we can expect that the grouping would be similar to how we have initially chosen the set of customers. -->
-
-<!-- from which we wanted to do instance selection.
-Since this is unlabeled data, there is no way to do external validation of our methodologies.
-Thus, we chose this way to see how well our methodology works in a cleaner data set as this one to classify customers who have similar behavior across all significant granularities.
-Next we select $4$ customers which are far apart from each other and $5$ customers which are lying closest in distance to each of these $4$ customers.
-These selection was done using the granularity $hod$.
-When we use our methodologies, it is based on all significant granularities and not only hod, but if $hod$ is significant we can expect that the grouping would be similar to how we have initially chosen the set of customers.
--->
-
-<!-- .
-Instances may be distributed in the representation space in a reasonable manner, revealing their similarities, using this instance wise discriminative learning.
-A dual-level progressive similar instance selection (DPSIS) approach could also be used -->
-
-<!-- Fig \ref{fig:data-heatmap-hod}, \ref{fig:} and \ref{fig:} shows the distribution of 24 customers across hour-of-day, month-of-year and wknd-wday respectively.
-Every row in \ref{fig:} shows different shapes across hour-of-day and across columns show similar shapes for each row.
-We use our methodology to see if the customers are allocated to same group have similar shapes across one or more significant granularity.
--->
 ```{r}
@@
 During the fall (March-June), these consumers' average energy use declines, before rising again during the winter.
-However, energy behaviour varies more in the winter than in the summer and a bit in the fall months.
+However, energy behavior varies more in the winter than in the summer and a bit in the fall months.
 
@@
 
-```{r combined-groups}
+```{r combined-groups, fig.cap = The distribution of electricity demand for the clusters across hour-of-day, month-of-year and wknd-wday.
+The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
+Group 1 and 3 have a stronger hour-of-day pattern, while group 2, 3, 4 have a month-of-year pattern.
+For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters."}
 combined <- (hod_group/moy_group) & theme(legend.position = "bottom")
@@
 This means that moy is the most important variable in distinguishing the designs followed by hod and wkndwday.
-It can be observed that cluster 3 and 4 are distinguished by moy while 1 and 2 are distinguished by hod.
+It can be observed that clusters are well separated by month-of-year, while hour-of-day and weekend-weekday are not useful distinguishing the clusters produced with this clustering method.
+There is only one customer who has significant wpd across weekend-weekday and stands out from the rest of the customers.
+3 and 4 are distinguished by moy while 1 and 2 are distinguished by hod.
 Here, wkndwday is still acting as the nuisance variable.

### 2021-10-28 main_application.Rmd
rewrite application

@@
 ---
-title: "main_application"
+title: "Application"
 date: "28/10/2021"
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
+library(ggplot2)
+library(gravitas)
+library(gracsr)
+library(ggdendro)
+library(dplyr)
+library(readr)
+library(visdat)
+library(ggplot2)
+library(tidyverse)
+library(naniar)
+library(here)
+library(tsibble)
+library(knitr)
+library(patchwork)
+library(GGally)
+library(distributional)
+library(viridis)
 ```
@@
 @Fan2021-bq proposes one such process that picks related examples (neighbours) for each instance (anchor) and considers them as the same class.
-In this part, consumers with prototype behaviours are chosen to serve as study population for our suggested methodology.
+In this part, consumers with prototype behaviors are chosen to serve as study population for our suggested methodology.
 
@@
 
-The $24$ prototypes are clustered using the methodology described in \ref{Sec:methodology}.
-The distribution of electricity demand for the selected $24$ customers across hour-of-day and month-of-year are shown in the right panel of Figures \ref{fig:hod-combined} and \ref{fig:moy-combined} respectively.
+The $24$ prototypes are first clustered using the methodology described in \ref{Sec:methodology}.
+We chose the optimal number of clusters using (@Henning-2013) as $5$.
+The distribution of electricity demand for the selected $24$ customers across hour-of-day and month-of-year are shown in \ref{fig:hod-combined} respectively.
 The median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.
-Customers are placed in the same order on both sides for both figures to simplify one-to-one comparison.
 All customers with the same color represent the same design (left) or cluster (right).
@@
 
-```{r combined-groups-js, fig.cap = The distribution of electricity demand for the clusters across hour-of-day, month-of-year and wknd-wday.
+```{r all-data}
+data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
+  mutate(customer_id = as.character(customer_id)) %>% 
+  dplyr::filter(customer_id %in% data_pick_cust$customer_id) %>% 
+  gracsr::scale_gran( method = "nqt",
+                      response = "general_supply_kwh")
+```
+
+```{r clustering}
+hod <- suppressMessages(data_pick %>% 
+  dist_gran(gran1 = "hour_day", response = "general_supply_kwh"))
+
+moy <- suppressMessages(data_pick %>% 
+  dist_gran(gran1 = "month_year", response = "general_supply_kwh"))
+
+wkndwday <- suppressMessages(data_pick %>% 
+  dist_gran(gran1 = "wknd_wday", response = "general_supply_kwh"))
+
+distance <- wkndwday/2 + moy/12 + hod/24
+
+f = as.dist(distance)
+```
+
+```{r opt-clusters}
+library(fpc)
+library(cluster)
+k = array()
+for(i in 2:20)
+{
+group <- f %>% hclust (method = "ward.D") %>% cutree(k=i)
+p <- cluster.stats(f, clustering = group, silhouette = TRUE)
+k[i]=p$sindex
+}
+# 
+# ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line()
+```
+
+```{r groups-24}
+cluster_result <- suppressMessages(f %>% 
+  clust_gran(kopt = 5)) %>% 
+  rename("customer_id" = "id") %>% 
+  mutate(group = as.factor(group))
+```
+
+```{r hod-ind-group, fig.cap="The distribution across moy, hod and wkndwday of each customer is shown.
+The same color represents same group.Our clustering methodology is useful for grouping together comparable distributions over hod and moy.
+Of course, certain consumers in each group have distributions that differ from the distributions for other customers in the same group.
+However, it appears that the goal of finding similar distribution across significant cyclic granularities has been established."  }
+
+hod_ind_group <- data_hod %>% 
+  left_join(cluster_result) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=customer_id, fill = group),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=customer_id, 
+                color = group), size = 1) +
+  facet_wrap(group~customer_id, 
+             scales = "free_y",
+              nrow=24) + 
+  theme_application()+
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +    xlab("hour-of-day")  +
+  scale_x_discrete(breaks = seq(0, 23, 3))
+# + scale_y_continuous(breaks = NULL)
+
+moy_ind_group <- data_moy %>% 
+  left_join(cluster_result, by = c("customer_id")) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, 
+                  group=customer_id, fill = group), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=customer_id, color = group), size = 1) +
+  facet_wrap(group~customer_id, 
+             scales = "free_y", 
+             nrow=24) +
+    ylab("demand (in Kwh)") +
+    xlab("month-of-year")  +
+  theme_application() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7"))
+
+wkndwday_ind_group <- data_wkndwday%>% 
+  left_join(cluster_result, by = c("customer_id")) %>% 
+  mutate(group  = as.factor(group)) %>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  #lvplot::geom_lv(aes(fill = as.factor(group), 
+          #            color = as.factor(group)), k=5, alpha = 0.5) +
+  geom_boxplot(aes(fill = group, color = group),alpha = 0.5)+
+  #geom_boxplot(outlier.shape = NA) + 
+  coord_cartesian(ylim = ylim1*1.05)+
+  facet_wrap(group~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow=24)  +
+  theme_application()+
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7"))
+
+(hod_ind_group + moy_ind_group + wkndwday_ind_group)  +
+  plot_layout(guides = "collect")
+
+```
+
+```{r data-heatmap-hod-group}
+legend_title <- "group"
+
+data_group <- data_pick  %>% 
+  left_join(cluster_result, by = c("customer_id"))
+
+data_heatmap_hod_group <- quantile_gran(data_group,
+                                  gran1="hour_day",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
+
+data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")
+
+hod_group <- data_heatmap_hod_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=group,
+                  fill = as.factor(group), alpha = 0.5),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=group, 
+                color = as.factor(group)), size = 1)+
+  facet_wrap(~group, 
+             scales = "free_y",  
+             nrow = 5) + 
+              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
+  ylab("demand (in Kwh)") + 
+  theme_bw() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank())+
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ 
+  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
+  scale_fill_manual(legend_title, values = c("#E69F00",
+  "#009E73","#0072B2", "#D55E00", "grey")) + 
+  theme(legend.position = "bottom")
+
+```
+
+```{r data-heatmap-moy-group}
+data_heatmap_moy_group <- quantile_gran(data_group,
+                                  gran1="month_year",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")
+
+moy_group <- data_heatmap_moy_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow = 5) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
+  ylab("demand (in Kwh)") +
+  theme_bw() +
+   theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank()) + 
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
+   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
+  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
+  theme(legend.position = "bottom") +
+  theme(panel.spacing =unit(0, "lines")) + 
+  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
+```
+
+```{r combined-groups-js, fig.cap = "The distribution of electricity demand for the clusters across hour-of-day, month-of-year and wknd-wday.
 The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
@@
 For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters."}
+hod_group + moy_group
 ```

### 2021-10-28 paper-supplementary.Rmd
rewrite application

---
title: "<small> Supplementary materials for the main submission entitled - <br><br> Clustering time series based on probability distributions across temporal granularities"

bibliography: [bibliography.bib]
preamble: >
  \usepackage{mathtools,amssymb,booktabs,amsthm,todonotes,colortbl}
  \def\mod{~\text{mod}~}
  \newtheorem{definition}{Definition}
  \usepackage{mathptmx}
  \usepackage{caption}
  \DeclareCaptionStyle{italic}{labelfont={bf},textfont={it},labelsep=colon}
  \captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
  \captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
  \def\novspacing{\setlength{\aboverulesep}{0pt}\setlength{\belowrulesep}{0pt}}
  \def\vspacing{\setlength{\aboverulesep}{0.4ex}\setlength{\belowrulesep}{0.65ex}}
output:
  bookdown::pdf_book:
    MonashEBSTemplates::workingpaper:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
---

## Raw time series plot

\noindent To get a sense of how the raw time series data looks, we plot the energy usage for $50$ sampled households is plotted along the y-axis versus time from past to future in Figure ref\{fig:raw-data-50}. Each of these series is associated with a single customer. Energy consumption for each customer is given at fine temporal resolution (every 30 minutes) for a period of 2-3 years.

```{r raw-data-50, out.width = "100%", fig.cap="The raw data for 50 households are shown. It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed. No insightful behavioral pattern could be discerned from this view other than when the customer is not at home."}

knitr::include_graphics("figs/raw_plot_cust.png") # look at smart-meter.R for the code
```

## Plot displaying missing observations

_Prototype selection method_

S1. Robust scaling is applied to each customer.

S2. $50^{th}$ percentile for each category for each granularity is obtained for each customers. So we have a data structure with  $356$ rows and $(24 + 12 + 2)$ variables corresponding to $50^{th}$ percentile for each hour-of-day, month-of-year and weekend-weekday.

S3. Apply principal components and restrict the results down to the first six principal components (which makes up approximately 85% of the variance explained in the data) to use with the grand tour.

S4. Run t-SNE using the default arguments on the complete data (sets the perplexity to equal 30 and performs random initialisation). We then create a linked tour with t-SNE layout with liminal as shown in Figure 4.

S5. We inspect of the subspace generated by the set of low-dimensional projections in tour by looking for a simplex shape while the visualization moves from one basis to another. When we brush the corners of the simplex, we find they fall on the edge of the t-SNE point cloud.
Hall, Marron, and Neeman (2005) have shown that in the extreme case of high-dimension, low-sample size data, observations are on the vertices of a simplex.

This is because in high-dimensional data analysis the curse of dimensionality reasons that points tend to be far away from the center of the distribution and on the edge of high-dimensional space. Contrary to this, is that projected data tends to clump at the center.

S6. These points should ideally correspond to different behavior with respect to all the variables considered while running PCA.

```{r mds, out.width="100%", fig.cap="Instance selection using tours and projecting the points in a lower dimensional tsne cloud."}
knitr::include_graphics("figs/liminal_view.png")
```

<!-- From this set we select $4$ "anchor" customers which are far apart from each other and $5$ neighboring customers for each of these anchors. These selections were done using the granularity $hod$ space. It is important to note that when we use our proposed methodologies, it is based on all dimensions `hod`, `moy` and `wkndwday`.Fig \ref{fig:mds} shows the MDS of these 356 customers in a 2D space basis their distance on individual granularities and when all of them are combined. Our methodolgies are run on these $24$ customers, which act as a way to evaluate the proposed methodologies.  -->

<!-- . Instances may be distributed in the representation space in a reasonable manner, revealing their similarities, using this instance wise discriminative learning. A dual-level progressive similar instance selection (DPSIS) approach could also be used -->

### 2021-10-28 main.Rmd
rewrite application

@@
   ylab("demand (in Kwh)") +
-   facet_grid(wknd_wday~group, 
+   facet_grid(group~wknd_wday, 
              scales = "free_y", 

### 2021-10-28 main_application.Rmd
rewrite application

@@
 
+Cluster characterization is a crucial aspect of cluster analysis.
+Figure ref\{fig:combined-groups-js} depicts the summarized distributions across groups and assists us in characterizing each cluster.
+All of these may be validated only with further information about the customer.
+Figure \ref{fig: Groups 2 and 5 show a stronger hour-of-day pattern, whereas groups 1, 3, and 5 show a month-of-year pattern.
+Differences in wknd-wday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters.
+
 ```{r data-heatmap-hod-group}
@@
   theme_bw()  +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank())+
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
   scale_x_discrete(breaks = seq(1, 24, 3))+ 
   #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
-  scale_fill_manual(legend_title, values = c("#E69F00",
-  "#009E73","#0072B2", "#D55E00", "grey")) + 
+  theme_application() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +
   theme(legend.position = "bottom")
@@
   ylab("demand (in Kwh)") +
-  theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + 
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
-  theme(legend.position = "bottom") +
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
+  theme_bw() + theme_application() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +
+  theme(legend.position = "bottom") 
+```
+
+```{r}
+wkndwday_data <- data_group %>% create_gran("wknd_wday") %>% 
+  create_gran("hour_day")
+
+ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]
+
+wkndwday_group <- wkndwday_data%>% 
+  ggplot(aes(x=hour_day, y = general_supply_kwh)) +
+  #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
+  geom_boxplot(aes(fill = as.factor(group)),alpha = 0.5)+
+  #geom_boxplot(outlier.size = 1) + 
+  coord_cartesian(ylim = ylim1*1.05)+
+  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
+#geom_boxplot(aes(fill = as.factor(group))) +
+  #scale_fill_lv() +
+ xlab("wknd-wday") + 
+  ylab("demand (in Kwh)") +
+   facet_grid(group~wknd_wday, 
+             scales = "free_y", 
+             labeller = "label_both") + 
+  theme_bw() + theme_application() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +
+  theme(legend.position = "bottom") 
 ```
@@
 ```{r combined-groups-js, fig.cap = "The distribution of electricity demand for the clusters across hour-of-day, month-of-year and wknd-wday.
-The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
 Group 2 and 5 have a stronger hour-of-day pattern, while group 1, 3, 5 have a month-of-year pattern.
 For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters."}
-hod_group + moy_group
+(hod_group + moy_group + wkndwday_group) +  plot_layout(guides = "collect")
 ```
+
+Things become far more complicated when we consider a larger data set with more uncertainty, as they do with any clustering problem.
+Summarizing distributions across clusters with varied or outlying customers can result in a shape that does not represent the group.
+Furthermore, combining hetergeneous customers may result in similar-looking final clusters that are not effective for visually differentiating them.
+It is also worth noting that the wknd-wday behavior in the given case does not characterize any cluster.
+This, however, will not be true for all of the customers in the data set.
+If more extensive prototype selection is used, resulting in more comprehensive prototypes in the data set, this method might be used to classify the entire data set into these prototype behaviors.
+However, the goal of this section was to have a few customers that have significant patterns over one or more cyclic granularities, apply our clustering methodology to cluster them, and demonstrate that the method produces useful clusters.
 
@@
 
-- Discussion
-
-As with any clustering method, things become much more complicated when we consider a larger data set with more uncertainty.
-The methodology run on several customers together might not be useful to given distinct shapes across granularities.
-Moreover, the groupings from two different appraoches lead to different results, both of which are useful but needs careful considering the context before choosing one over another.
-Also, note that the customers chosen here do not have a weekend-weekday effect, which might not be true for all customers in the data set.
-Ideally there are other groups of customers in the entire dataset which can act as prototypes in our problem.
-Ideally, if these prototypes are not outliers, we can use them for a classification problem for external validation of the clustering problem.
-
 Notice that while looking for these customers

### 2021-10-28 main_application.Rmd
rewrite application

@@
   scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +    xlab("hour-of-day")  +
-  scale_x_discrete(breaks = seq(0, 23, 3))
+  scale_x_discrete(breaks = seq(0, 23, 3))+ theme(legend.position = "bottom")
 # + scale_y_continuous(breaks = NULL)
@@
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7"))
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) + theme(legend.position = "bottom")
 
@@
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7"))
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7"))  + theme(legend.position = "none")
 
@@
   #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
-  geom_boxplot(aes(fill = as.factor(group)),alpha = 0.5)+
+  geom_boxplot(aes(fill = group, color = group),alpha = 0.5)+
   #geom_boxplot(outlier.size = 1) + 
@@
   scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +
-  theme(legend.position = "bottom") 
+  theme(legend.position = "none") 
 ```
@@
 For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters."}
-(hod_group + moy_group + wkndwday_group) +  plot_layout(guides = "collect")
+(hod_group + moy_group + wkndwday_group) + 
+  plot_layout(guides = "collect")
 ```

### 2021-10-29 main.Rmd
rewrite application wpd analsalysis added

@@
     theme(strip.background = element_blank(),
-  strip.text.x = element_blank()) 
+  strip.text.x = element_blank())  + theme(legend.position = "bottom")
 }
@@
 
-```{r combined-groups, fig.cap = The distribution of electricity demand for the clusters across hour-of-day, month-of-year and wknd-wday.
+```{r combined-groups, fig.cap = "The distribution of electricity demand for the clusters across hour-of-day, month-of-year and wknd-wday.
 The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.

### 2021-10-29 main_application.Rmd
rewrite application wpd analsalysis added

@@
 It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
-No additional information about the customers is used for the application.
-
-The linear view of the time series data generally has too many measurements all squeezed in that representation, it hinders us to discern any repetitive behavioral pattern for even one customers (let alone many customers together).
-In most cases, electricity data will have multiple seasonal patterns like daily, weekly or annual.
-We do not learn about these repetitive behaviors from the linear view.
+In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
+We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation.
 Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.
@@
 Additionally, there were missing values in the database but further investigation revealed that there is no structure in the missingness (see Supplementary paper for raw data features and missingness).
-
-## Prototype selection
-
-<!-- Why instance selection -->
-Supervised learning uses a training set of known information to categorize new events.
-Instance selection (@olvera2010review) is a method of rejecting instances that are not helpful for classification.
-This is equivalent to subsetting the population along all dimensions of interest such that the sampled data reflects the underlying distribution's primary features.
-Instance selection in unsupervised learning has received little attention in the literature, yet it could be a useful tool for evaluating model or method performance.
-@Fan2021-bq proposes one such process that picks related examples (neighbours) for each instance (anchor) and considers them as the same class.
-In this part, consumers with prototype behaviors are chosen to serve as study population for our suggested methodology.
+The study begins by subsetting a data set along all dimensions of interest using data filtering and prototyping.
+By grouping the prototypes using our methods and assessing their meaning, the study hopes to unravel some of the heterogeneities observed in energy usage data.
+Because our application does not employ additional customer data, we cannot explain why consumption varies, but rather try to identify how it varies.
 
@@
 
-<!-- - can be done in several ways -->
+_Prototype selection_
 
-\noindent There are several ways to approach the prototype selection.
-Use any dimensionality reduction techniques like MDS or PCA to project the data into a 2-dimensional space.
+<!-- Why instance selection -->
+Supervised learning uses a training set of known information to categorize new events through instance selection.
+Instance selection (@olvera2010review) is a method of rejecting instances that are not helpful for classification.
+This is analogous to subsampling the population along all dimensions of interest such that the sampled data represents the primary features of the underlying distribution.
+Instance selection in unsupervised learning has received little attention in the literature, yet it could be a useful tool for evaluating model or method performance.
+There are several ways to approach the prototype selection.
+Following @Fan2021-bq's idea of picking related examples (neighbours) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2-dimensional space.
 Then pick a few "anchor" customers who are far apart in 2D space and pick a few neighbors for each.
 Unfortunately, this does not assure that consumers with significant patterns across all variables are chosen.
-We perform a linked tour with t-SNE layout and (liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
-Tours can help us see separation between variables that was not obvious in the single variable display.
+Tours can reveal variable separation that was hidden in a single variable display.
+Hence we perform a linked tour with t-SNE layout (using the R package @R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
 Please see the Supplementary article for further details on how the prototypes are chosen.
@@
 Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution.
-These $24$ prototypes are clustered using the methodology described in \ref{Sec:methodology} to see if the grouping is useful.
 
@@
 
-(hod_ind_design + moy_ind_design + wkndwday_ind_design)
-
 ```
@@
 
-### Using JS-based distances
+Cluster characterization is a crucial aspect of cluster analysis.
+The $24$ prototypes and the data subset consisting of $353$ customers are clustered using the methodology described in \ref{Sec:methodology} and results are reported below.
+In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.
+All customers with the same color represent same clustered groups.
+The plotting scales are not displayed since we want to emphasize comparable shapes rather than scales.
+The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to shape or significance across cyclic granularities.
 
-The $24$ prototypes are first clustered using the methodology described in \ref{Sec:methodology}.
-We chose the optimal number of clusters using (@Henning-2013) as $5$.
+### JS-based distances
+
+We chose the optimal number of clusters using (@Hennig2014-ah) as $5$.
 The distribution of electricity demand for the selected $24$ customers across hour-of-day and month-of-year are shown in \ref{fig:hod-combined} respectively.
-The median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.
-All customers with the same color represent the same design (left) or cluster (right).
-The plotting scales are not displayed since we want to emphasize comparable shapes rather than scales.
-A customer in the cluster may have low daily or total energy usage, but their behavior may be quite similar to a customer with high usage.
+Figure ref\{fig:combined-groups-js} depicts the summarized distributions across groups and assists us in characterizing each cluster.
+All of these may be validated only with further information about the customer.
+Figure ref\{fig:combined-groups-js} shows Groups 2 and 5 show a stronger hour-of-day pattern with a typical morning and evening peak, whereas groups 1, 3, and 5 show a month-of-year pattern  with higher usage in winter months.
+Differences in wknd-wday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters.
 
@@
 # 
-# ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line()
+# ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 20, 1))
 ```
@@
 
-```{r hod-ind-group, fig.cap="The distribution across moy, hod and wkndwday of each customer is shown.
-The same color represents same group.Our clustering methodology is useful for grouping together comparable distributions over hod and moy.
-Of course, certain consumers in each group have distributions that differ from the distributions for other customers in the same group.
-However, it appears that the goal of finding similar distribution across significant cyclic granularities has been established."  }
+```{r hod-ind-group, fig.cap="The distribution of selected consumers over hod (a, d), moy (b, e), and wkndwday (d, f).
+In each case, the same colour denotes the same group in plots (d), (e), (f) and are placed together to facilitate comparison.
+That means the customer orderings are different for (a, b, c) and (d, e, f).
+Our clustering methodology is useful for grouping similar distributions over hod and moy.
+Of course, certain customers in each group have distributions that differ from those of other members in the same group.
+However, it appears that the aim of grouping comparable distributions over considered variables has been accomplished to some extent." }
 
 hod_ind_group <- data_hod %>% 
-  left_join(cluster_result) %>% 
+  left_join(cluster_result, by = c("customer_id")) %>% 
   ggplot(aes(x = category)) + 
@@
           #            color = as.factor(group)), k=5, alpha = 0.5) +
-  geom_boxplot(aes(fill = group, color = group),alpha = 0.5)+
+  geom_boxplot(aes(color = group, fill = group),alpha = 0.5)+
   #geom_boxplot(outlier.shape = NA) + 
@@
 
-(hod_ind_group + moy_ind_group + wkndwday_ind_group)  +
-  plot_layout(guides = "collect")
-
+hod_ind_design + moy_ind_design + wkndwday_ind_design + hod_ind_group + moy_ind_group + wkndwday_ind_group +  plot_layout(guides = "collect", ncol = 6) + plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')') & theme(legend.position = 'none')
 ```
-
-Cluster characterization is a crucial aspect of cluster analysis.
-Figure ref\{fig:combined-groups-js} depicts the summarized distributions across groups and assists us in characterizing each cluster.
-All of these may be validated only with further information about the customer.
-Figure \ref{fig: Groups 2 and 5 show a stronger hour-of-day pattern, whereas groups 1, 3, and 5 show a month-of-year pattern.
-Differences in wknd-wday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters.
 
@@
   #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
-  geom_boxplot(aes(fill = group, color = group),alpha = 0.5)+
+  geom_boxplot(aes(fill = group, color = group),alpha = 0.5, outlier.alpha = 0.05)+
   #geom_boxplot(outlier.size = 1) + 
@@
 
-```{r combined-groups-js, fig.cap = "The distribution of electricity demand for the clusters across hour-of-day, month-of-year and wknd-wday.
-Group 2 and 5 have a stronger hour-of-day pattern, while group 1, 3, 5 have a month-of-year pattern.
-For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters."}
+```{r combined-groups-js, fig.cap = "The distribution of electricity demand for the clusters across hod (a), moy (b) and wkndwday (c).
+It seems like group 2 and 5 have a hod pattern across its members, while group 1, 3, 5 have a moy pattern.
+Wknd-wday variations across groups are not distinguishable, indicating that it is not a critical variable for clustering.
+It is helpful to compare the summarised distributions of groups to that of individuals to confirm that the most of individuals in the group have the same characterisation."}
+
  (hod_group + moy_group + wkndwday_group) +
-  plot_layout(guides = "collect")
+  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
+  plot_layout(guides = "collect")& theme(legend.position = 'bottom')
 ```
+
+### wpd-based distances
+
+We chose the optimal number of clusters using (@Hennig2014-ah) as $3$.
+A parallel coordinate plot with the three significant cyclic granularities used for wpd-based clustering.
+The variables are sorted according to their separation across classes (rather than their overall variation between classes).
+This means that $moy$ is the most important variable in distinguishing the designs followed by $hod$ and $wkndwday$.
+There is only one customer who has significant $wpd$ across $wkndwday$ and stands out from the rest of the customers.
+Group 3 has a higher wpd for hod than moy or wkndwday.
+Group 2 has the most distinct pattern across moy.
+Group 1 is a mixed group that has strong patterns on at least one of the three variables.
+The findings vary from js-based clustering, yet it is a helpful grouping.
+
+```{r data-pick-wpd}
+elec_600_wpd <- read_rds(here::here("data/algo2-cust600-wpd-rawdata.rds"))
+
+elec_pick <- elec_600_wpd %>% 
+  filter(customer_id %in% data_pick_cust$customer_id)
+
+elec_pick_wide <- elec_pick %>% pivot_wider(-c(1, 2), names_from = "x_variable", values_from = wpd)
+
+scaled_var <- elec_pick_wide
+
+f <- elec_pick_wide[-1] %>% dist()
+
+k = array()
+for(i in 2:20)
+{
+group <- f %>% hclust (method = "ward.D") %>% cutree(k=i)
+p <- cluster.stats(f, clustering = group, silhouette = TRUE)
+k[i]=p$sindex
+}
+# ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 20, 1))
+
+group <- f%>% hclust (method = "ward.D") %>% cutree(k=3)
+
+cluster_result_wpd <- bind_cols(id = elec_pick_wide$customer_id, group = group)
+
+data_pcp <- scaled_var %>% 
+  #bind_cols(customer_id =  elec_pick_wide$customer_id) %>%
+  left_join(cluster_result_wpd , by = c("customer_id" = "id")) %>% 
+  select(customer_id, group, everything()) %>% 
+  mutate(group = as.factor(group))
+
+data_table <- data_pcp %>% group_by(group) %>% 
+  summarise(nobs = n(),
+            moy = round(median(month_year),2),
+            hod = round(median(hour_day),2),
+            wnwd = round(median(wknd_wday),2))
+
+```
+
+```{r parcoord, fig.cap="Each of the 24 customers is represented by a parallel coordinate plot (a) with three wpd-based groupings.
+The plot shows that moy is the most important variable in identifying clusters, whereas wkdn-wday is the least significant and has the least fluctuation.
+One particular customer with high wpd across wknwday stands out in this display.
+Group 3 has a higher wpd for hod than moy or wkndwday.
+Group 2 has most discernible pattern across moy.
+Group 1 is a mixed group with strong patterns on atleast one of the three variables.
+All of these could be observed from the plot or the table (b) which shows median wpd values for each group.", message=FALSE, warning=FALSE}
+
+parcoord <- GGally::ggparcoord(data_pcp ,
+                   columns = 3:ncol(data_pcp),
+                   groupColumn = "group",
+                   showPoints = FALSE, 
+                   alphaLines = 0.8,
+                   order = "anyClass",
+                   scale = "globalminmax"
+) + 
+  ggplot2::theme(
+    plot.title = ggplot2::element_text(size=10)
+  )+
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
+  theme(legend.position = "bottom") +
+  xlab("") +
+  ylab("wpd") + scale_fill_viridis_d(direction = 1) +
+  scale_color_viridis_d(direction = 1) + theme_light()
+
+(parcoord + gridExtra::tableGrob(data_table))+ plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')') & theme(legend.position = "bottom")
+
+```
+
+## Entire data subset
 
@@
 However, the goal of this section was to have a few customers that have significant patterns over one or more cyclic granularities, apply our clustering methodology to cluster them, and demonstrate that the method produces useful clusters.
-
-### Using wpd-based distances
-
-```{r pcp-wpd}
-```
-
-A parallel coordinate plot with the three significant cyclic granularities used for wpd-based clustering.
-The variables are sorted according to their separation across classes (rather than their overall variation between classes).
-This means that $moy$ is the most important variable in distinguishing the designs followed by $hod$ and $wkndwday$.
-It can be observed that clusters are well separated by $moy$, while $hod$ and $wkndwday$ are not useful distinguishing the clusters produced with this clustering method.
-The parallel coordinate plot ranks the variables in order of importance, indicating that the month-of-year is the most important in identifying clusters, whereas wkdn-wday is the least significant and has the least variability among the three variables.
-However, there is only one customer who has significant $wpd$ across $wkndwday$ and stands out from the rest of the customers.
-The ggpairs plot also shows five distinct clusters across the $moy$.
-
-Notice that while looking for these customers
-
-(can be used as classification method, problems in handling large dataset)

### 2021-10-29 main.Rmd
major rewrite application complete

--- 2021-10-28 main.Rmd
@@
   theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
-    #theme(axis.ticks.x = element_blank(),
-     #   axis.text.x = element_blank()) + # no x-axis labels to further reduce the gap between facets
+    theme(axis.ticks.x = element_blank(),
+        axis.text.x = element_blank()) + # no x-axis labels to further reduce the gap between facets
     #theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
   theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-    theme(axis.text.x = element_text(angle=90, hjust=1, size = 10))+
-      theme(plot.background = element_blank(), panel.grid = element_blank())+
   #theme(axis.text.x = element_text(size=5)) +
@@
 
-# Application {#sec:application}
+# Application
 
@@
 It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
-It also provides demographic data for these customers most of which are missing and not utilized for the purpose of this paper.
-To maintain anonymity, the energy patterns could not be recognised at the person level, but rather by the geographical location of their dwelling and information about their Local Government Area.
-
-<!-- ## Data source -->
-
-<!-- The entire data is procured from CSIRO.
-A subset of this data is also available from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
-It consists of the following data sets.
--->
-<!-- _1. CustomerData:_ 78720 customers with 62 variables about them  -->
-<!-- _2. EUDMData:_ 300 billion half-hourly consumption level data   -->
-<!-- _3. OffersData:_ Method of contact to customer to join SGSC customer trial, either door-to-door (D2D) or via Telesales   -->
-<!-- _4. PEResponseData:_ Peak Events response customer wise   -->
-<!-- _5. PETimesData:_ Peak Events time stamps   -->
-
-<!-- Only _CustomerData_ and _EUDMData_ are relevant for the clustering goals of this paper.
-_EUDMData_ contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
-`CustomerData` provides demographic data for 78,720 customers most of which are missing and not utilized for the purpose of this paper.
-To meet the requirements for anonymity preservation, the energy patterns could not be identified at the individual level, but rather by the geographical location of their residence information about their Local Government Area.
--->
-
-<!-- ## Characteristics of raw data -->
-
-In Figure \ref{fig:raw-data-50}, the time series of energy consumption is plotted along the y-axis against time from past to future for $50$ sampled households.
-Each of these series correspond to a single customer.
-For each customer, the energy consumption is available at fine temporal resolution (every 30 minutes) for a long period of time (~ 2 years).
-Some customers' electricity use may be unavailable owing to power outages or improper recording, resulting in implied missing numbers in the database.
-For this data set it was found that out of 13,735 customers in total, 8,685 customers do not have any implicit missing observations, while the rest 5,050 customers had missing values.
-With further exploration, it was found that there is no structure in the missing-ness, that is missing observations can occur at any time point (see Appendix).
-Moreover, the data for these customers are characterized by unequal length, different start and end dates.
-Since our proposed methods consider probability distribution instead of raw data, both of these characteristics would not pose any threat to our methodology unless of course there is any structure or systematic patterns in them.
-
-\noindent It can be expected that energy consumption vary substantially between customers, which is a reflection of their varied  behavior owing to differences in profession, family size, geographical or physical characteristics.
-Since the linear time series plot has too many measurements all squeezed in this linear representation, it hinders us to discern any repetitive behavioral pattern for even one customers (let alone many customers together).
-In most cases, electricity data will have multiple seasonal patterns like daily, weekly or annual.
-We do not learn about these repetitive behaviors from the linear view.
+In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
+We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation.
 Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.
-
-<!-- ### Missing Data -->
-
-<!-- Electricity usage for some customers may become unavailable due to power outage or not recording their usage properly, thus resulting in implicit missing values in the database.
-It is interesting to explore where missing-ness occurs or if there is a relationship between the underlying missing patterns.
-We use the R package `tsibble` to do this.
--->
-
-```{r missing-data, out.width="100%", eval = FALSE}
-include_graphics("figs/missing-data.png")
-```
-
-<!-- - if there is any systematic missing patterns in the data -->
-<!-- - this missing plot can go in the supplementary -->
-<!-- - how to add missing values (should be added in data pre-processing) -->
-<!-- - instance learning -->
-<!-- - types of summary techniques to use ( -->
-<!-- generally multivariate means and sd are used, I can't use that in a time series context, you can show across different granularity? -->
-<!-- within-group sum of squares and between-group sum of squares -->
-
-<!-- ) -->
-
-<!-- 13735 customers in elec_ts -->
-<!-- 8685 customers in elec_nogap -->
-<!-- 5050 customers in count_na_df -->
-
-<!-- Then is the graph of missing observations even interesting. -->
-<!-- You can show two graphs, one to show that missingness do not have a pattern -->
-<!-- another to show even if no missing, they start and end at different times. (A sample of 50 customers).  -->
-
-<!-- \noindent A dataset of 100 SGSC homes has been used to lay out the structure to be used for analyzing the big dataset.
-The smaller dataset contains half-hourly kwh values form 2012 to 2014 and has asynchronous time series distributed evenly over the observation period (Figure \ref{fig:elec-raw}), similar to the bigger data set.
-Figure \ref{fig:count-gaps} can be used to interpret missingness in the data, where the customers are arranged from maximum to minimum missing.
-It looks like data is most missing before 2013 and for a particular date in 2014.
--->
-
-```{r miss-data}
-empty_as_na <- function(x){
-    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
-    ifelse(as.character(x)!="", x, NA)
-}
-
-```
-
- <!-- - Handling autocorrelation: Autocorrelation in the time series is likely to get removed from considering probability distributions and cyclic granularities in the clustering algorithm. -->
-
-<!-- At this stage, we need to define the aim of clustering as there could be various aims of clustering like between-cluster separation, within cluster homogeneity: low distances, within-cluster homogeneous distributional shape, good representation of data by centroids, little loss of information, high density without cluster gaps, uniform cluster sizes, stability and others.
-Finally, how distinct they are and how can we summarize the main features of the cluster would be discussed here.
--->
-
-## Prototype selection
-
-<!-- A clean data set is obtained by carefully choosing customers which shows similar shapes across one or more cyclic granularity.
-Since this is unlabeled data, there is no way to do external validation of our methodologies.
-Thus, we chose this way to see how well our methodology works in a cleaner data set as this one.
--->
+The raw data for these consumers is of unequal length, with varying start and finish dates.
+Because our proposed methods evaluate probability distributions rather than raw data, neither of these data features would pose any threat to our methodology unless they contained any structure or systematic patterns.
+Additionally, there were missing values in the database but further investigation revealed that there is no structure in the missingness (see Supplementary paper for raw data features and missingness).
+The study begins by subsetting a data set along all dimensions of interest using data filtering and prototyping.
+By grouping the prototypes using our methods and assessing their meaning, the study hopes to unravel some of the heterogeneities observed in energy usage data.
+Because our application does not employ additional customer data, we cannot explain why consumption varies, but rather try to identify how it varies.
+
+_Data filtering and variable selection_
+
+- Choose a smaller subset of randomly selected $600$ customers with no implicit missing values for 2013.
+
+- Obtain $wpd$ for all cyclic granularities considered for these customers.
+It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weeknd/weekday) are coming out to be significant for most customers.
+We use these three granularities while clustering.
+
+- Remove customers whose data for an entire category of a significant granularity is empty.
+For example, a customer who does not have data for an entire month is excluded because their monthly behaviour cannot be analyzed.
+
+- Remove customers whose energy consumption is 0 in all deciles.
+These are the clients whose consumption is likely to remain essentially flat and with no intriguing repeated patterns that we are interested in studying.
+
+_Prototype selection_
 
 <!-- Why instance selection -->
-In supervised learning, a training set containing previously known information is used to categorize new occurrences.
-Acceptable classification rates may be obtained by discarding instances which are not helpful for classification; this process is known as instance selection (@olvera2010review).
-This is similar to subsetting the population along all dimensions of importance such that the sampled data is representative of the main characteristics of the underlying distribution.
-Instance selection in unsupervised learning has received limited attention in the literature, but could serve as an useful way to sample evaluation data set to measure the performance of a model or method.
-One such procedure is suggested in @Fan2021-bq that selects similar instances (neighbors) for each instance (anchor) and treats the anchor and its neighbors as the same class.
-In this section, a similar idea is used to select customers with prototype behaviors
-that serves as evaluation data sets for our proposed methodology.
-
-_Pre-processing steps_
-
-P1. We randomly select a sample of $600$ customers which do not have any implicit missing values and filtered their data for the year 2013.
-
-P2.
-Obtain $wpd$ for all cyclic granularities considered for these customers.
-It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weeknd/weekday) are coming out to be significant for most customers.
-This implies that for most customers, there is some interesting pattern across these three granularities.
-
-P3.
-Remove customers from this list for which data in any category for the significant granularities are empty.
-For example, in this data set, if a customer do not have data for an entire month, they have been removed as their monthly behavior could not be studied in that case.
-
-P4.
-Remove customers for which all the deciles of the energy consumption is zero.
-These are the customers whose consumption remain mostly flat and is expected to have no interesting repetitive patterns that is our interest of study.
-
-Finally, we are left with $356$ customers from which we run our prototype search.
-There are different methods of going ahead with this.
-For example, one approach could be to use any Non-linear dimensionality reduction technique like MDS or PCA and project the data in a 2-dimensional space.
-One can then look at few "anchor" customers which are far apart in the 2D space and pick few neighbors for each of the anchor customers.
-Paradoxically, the curse of dimensionality inverts for dimension reduction, resulting in an excessive amount of observations near the center of the distribution.
-This affects visualizations made on low-dimensional projections.
-[Ursula's paper] explains why points tend to be away from the center in the high-dimensional space, but crowd the center in low-dimensional projections, it is helpful to consider the projected volume relative to high-dimensional volume.
-
-```{r}
+Supervised learning uses a training set of known information to categorize new events through instance selection.
+Instance selection (@olvera2010review) is a method of rejecting instances that are not helpful for classification.
+This is analogous to subsampling the population along all dimensions of interest such that the sampled data represents the primary features of the underlying distribution.
+Instance selection in unsupervised learning has received little attention in the literature, yet it could be a useful tool for evaluating model or method performance.
+There are several ways to approach the prototype selection.
+Following @Fan2021-bq's idea of picking related examples (neighbours) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2D space.
+Then pick a few "anchor" customers who are far apart in 2D space and pick a few neighbors for each.
+Unfortunately, this does not ensure that consumers with significant patterns across all variables are chosen.
+Tours can reveal variable separation that was hidden in a single variable display better than static projections.
+Hence we perform a linked tour with t-SNE layout using the R package `liminal` (@R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
+Please see the Supplementary article for further details on how the prototypes are chosen.
+Figure \ref{fig:hod-ind-group} (a, b, c) shows the raw time plot, distribution across `hod`, `moy` and `wkndwday` for the set of chosen $24$ customers.
+Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution.
+
+```{r prototype-data-pick}
 quantile_prob_graph <- c(0.25, 0.5, 0.75)
@@
 
-```{r}
+```{r data-pick}
 data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
@@
   dplyr::filter(customer_id %in% data_pick_cust$customer_id) %>% 
-  gracsr::scale_gran( method = "nqt",
+  gracsr::scale_gran( method = "robust",
                       response = "general_supply_kwh")
@@
 
-```{r clustering}
-hod <- suppressMessages(data_pick %>% 
-  dist_gran(gran1 = "hour_day", response = "general_supply_kwh"))
-
-moy <- suppressMessages(data_pick %>% 
-  dist_gran(gran1 = "month_year", response = "general_supply_kwh"))
-
-wkndwday <- suppressMessages(data_pick %>% 
-  dist_gran(gran1 = "wknd_wday", response = "general_supply_kwh"))
-
-distance <- wkndwday/2 + moy/12 + hod/24
-
-f = as.dist(distance)
-
-cluster_result <- suppressMessages(f %>% 
-  clust_gran(kopt = 5)) %>% 
-  rename("customer_id" = "id")
-```
-
-```{r hod-data}
+<!-- - can be done in several ways -->
+
+<!-- - show raw data of moy, hod, wkndwday patterns for them -->
+
+```{r hod-moy-wkndwday plots, fig.cap="The distribution across moy, hod and wkndwday for the selected designs.
+Few are similar in their hod pattern, while others are similar in moy behavior.
+Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most households, there is one household for whom weekend behavior is standing out from the rest."}
+
 data_hod <- quantile_gran(data_pick,
@@
               values_from = quantiles_values) %>% 
-  left_join(data_pick_cust, by = c("customer_id")) %>% 
-  left_join(cluster_result, by = c("customer_id"))
+  left_join(data_pick_cust, by = c("customer_id"))
 
@@
 
-# lev <- data_hod %>% distinct(customer_id, design) %>% arrange(design) %>% pull(customer_id)
-```
-
-```{r hod-ind-design}
-
-# data_hod$customer_id <- factor(data_hod$customer_id, levels = lev)
-
 hod_ind_design <- data_hod %>% 
-  left_join(cluster_result,by = c("customer_id", "group")) %>% 
   ggplot(aes(x = category)) + 
@@
                   ymax = `75%`,
-                  group=customer_id, 
-                  fill = design),
+                  group=customer_id),
               alpha = 0.5) +
   geom_line(aes(y = `50%`,
-                group=customer_id,
-                color = design), 
+                group=customer_id), 
             size = 1) +
-  facet_wrap(design~customer_id, 
+  facet_wrap(~customer_id, 
              scales = "free_y",
-             ncol=6) + 
+             nrow = 24) + 
  theme_application() +
-  xlab("hour-of-day") +
+  xlab("hod") +
   scale_fill_viridis_d()+
@@
   scale_x_discrete(breaks = seq(0, 23, 3))
-```
-
-```{r hod-ind-group}
-hod_ind_group <- data_hod %>% 
-  left_join(cluster_result) %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`,
-                  group=customer_id, fill = as.factor(group)),
-              alpha = 0.5) +
-  geom_line(aes(y = `50%`,
-                group=customer_id, color = as.factor(group)), size = 1) +
-  facet_wrap(design~customer_id, 
-             scales = "free_y",
-              ncol=6) + theme_application()+
-  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black")) +    xlab("hour-of-day")  +
-  scale_x_discrete(breaks = seq(0, 23, 3))
-# + scale_y_continuous(breaks = NULL) 
-# 
-```
-
-```{r hod-combined, fig.cap = "The energy demand distribution for the prototype designs (left) and clustering (right) is given for each hour of the day.
-Customer design grouping and clustering strategies complement one other well.
-For a one-to-one comparison, both sides have the same order of consumers.
-Design-4 has a low morning peak and a strong evening peak.
-Because other members of group-1 have higher late-night usage, group-2 adopts this design and assigns customer B to it rather than group-1."}
-(hod_ind_design + hod_ind_group)* theme(legend.position = "bottom")
-
-```
-
-```{r data-heatmap-moy-design}
 
@@
 moy_ind_design <- data_moy %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, 
+                  group=customer_id), 
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=customer_id), size = 1) +
+  facet_wrap(~customer_id, 
+             scales = "free_y", 
+             nrow = 24) +
+    ylab("demand (in Kwh)") +
+    xlab("moy")  +
+  theme_application()
+
+data_wkndwday <- data_pick  %>%
+   create_gran("wknd_wday")  %>% 
+  left_join(data_pick_cust, by = c("customer_id"))
+
+ylim1 = boxplot.stats(data_wkndwday$general_supply_kwh)$stats[c(1, 5)]
+
+wkndwday_ind_design <- data_wkndwday%>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  #lvplot::geom_lv(aes(fill = as.factor(design), 
+   #                   color = as.factor(design)), k=5, alpha = 0.5) +
+  geom_boxplot(alpha = 0.5, fill = "black")+
+  #geom_boxplot(outlier.shape = NA) + 
+  coord_cartesian(ylim = ylim1*1.05)+
+  facet_wrap(~customer_id, 
+             scales = "free_y", 
+             labeller = "label_value",
+              nrow = 24)  +
+  stat_summary(
+    fun = median,
+    geom = 'line',
+    aes(group = 1), size = 1, color = "black")+
+  xlab("wdwn")+
+  theme_application()
+
+```
+
+## Clustering
+
+Cluster characterization is a crucial aspect of cluster analysis.
+The $24$ prototypes <!--and the data subset consisting of $353$ customers--> are clustered using the methodology described in \ref{sec:methodology} and results are reported below.
+In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.
+All customers with the same color represent same clustered groups.
+Groups by JS-based distances and wpd-based distances are colored differently as they represent different groupings.
+The plotting scales are not displayed since we want to emphasize comparable shapes rather than scales.
+The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to shape or significance across cyclic granularities.
+
+### JS-based distances
+
+For clustering based on JS-based distances, we chose the optimal number of clusters using (@Hennig2014-ah) as $5$.
+The distribution of electricity demand for the selected $24$ customers across hod, moy and wdwn are shown in Figure \ref{fig:hod-ind-group} (d, e, f).
+Our methodology is useful for grouping similar distributions over hod and moy and they are placed closely for easy comparison.
+Of course, certain customers in each group have distributions that differ from other members in the same group.
+However, it appears that the aim of grouping comparable distributions over considered variables has been accomplished to some extent.
+Figure \ref{fig:combined-groups-js} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
+Figure \ref{fig:combined-groups-js} shows Groups $2$ and $5$ show a stronger hod pattern with a typical morning and evening peak, whereas groups 1, 3, and 5 show a moy pattern  with higher usage in winter months.
+Differences in wknd-wday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters.
+
+```{r all-data}
+data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
+  mutate(customer_id = as.character(customer_id)) %>% 
+  dplyr::filter(customer_id %in% data_pick_cust$customer_id) %>% 
+  gracsr::scale_gran( method = "nqt",
+                      response = "general_supply_kwh")
+```
+
+```{r clustering}
+hod <- suppressMessages(data_pick %>% 
+  dist_gran(gran1 = "hour_day", response = "general_supply_kwh"))
+
+moy <- suppressMessages(data_pick %>% 
+  dist_gran(gran1 = "month_year", response = "general_supply_kwh"))
+
+wkndwday <- suppressMessages(data_pick %>% 
+  dist_gran(gran1 = "wknd_wday", response = "general_supply_kwh"))
+
+distance <- wkndwday/2 + moy/12 + hod/24
+
+f = as.dist(distance)
+```
+
+```{r opt-clusters}
+library(fpc)
+library(cluster)
+k = array()
+for(i in 2:20)
+{
+group <- f %>% hclust (method = "ward.D") %>% cutree(k=i)
+p <- cluster.stats(f, clustering = group, silhouette = TRUE)
+k[i]=p$sindex
+}
+# 
+# ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 20, 1))
+```
+
+```{r groups-24}
+cluster_result <- suppressMessages(f %>% 
+  clust_gran(kopt = 5)) %>% 
+  rename("customer_id" = "id") %>% 
+  mutate(group = as.factor(group))
+```
+
+```{r hod-ind-group, fig.cap="The distribution of selected consumers over hod (a, d), moy (b, e), and wkndwday (d, f).
+In each case, the same colour denotes the same group in plots (d), (e), (f) and are placed together to facilitate comparison.
+That means the customer orderings are different for (a, b, c) and (d, e, f).
+Our clustering methodology is useful for grouping similar distributions over hod and moy.
+Of course, certain customers in each group have distributions that differ from those of other members in the same group.
+However, it appears that the aim of grouping comparable distributions over considered variables has been accomplished to some extent." }
+
+hod_ind_group <- data_hod %>% 
   left_join(cluster_result, by = c("customer_id")) %>% 
@@
                   ymax = `75%`,
-                  group=customer_id, fill = as.factor(design)), alpha = 0.5) +
-  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(design)), size = 1) +
-  facet_wrap(design~customer_id, 
+                  group=customer_id, fill = group),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=customer_id, 
+                color = group), size = 1) +
+  facet_wrap(group~customer_id, 
              scales = "free_y",
-             ncol = 6) +
-    ylab("demand (in Kwh)") +
-    xlab("month-of-year")  +
-  theme_application() +
-  scale_fill_viridis_d(direction = 1)+
-  scale_color_viridis_d(direction = 1)
-```
-
-```{r data-heatmap-moy-group-ind}
+              nrow=24) + 
+  theme_application()+
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +    xlab("hod")  +
+  scale_x_discrete(breaks = seq(0, 23, 3))+ theme(legend.position = "bottom")
+# + scale_y_continuous(breaks = NULL)
 
@@
                   ymax = `75%`, 
-                  group=customer_id, fill = as.factor(group)), alpha = 0.5) +
-  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
-  facet_wrap(design~customer_id, 
+                  group=customer_id, fill = group), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=customer_id, color = group), size = 1) +
+  facet_wrap(group~customer_id, 
              scales = "free_y", 
-             ncol = 6) +
+             nrow=24) +
     ylab("demand (in Kwh)") +
-    xlab("month-of-year")  +
+    xlab("moy")  +
   theme_application() +
-  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))
-```
-
-```{r moy-combined, fig.cap = "The energy demand distribution for the prototype designs (left) and clustering (right) are shown.
-The customer grouping by design and the clustering approach largely match, except for some exceptions.
-For example, in group 3, consumers are categorised by greater usage in the middle of the year, which corresponds to winter in Australia.
-This is a better grouping in terms of similar month-of-year pattern than originally proposed by the prototype design."}
-
-(moy_ind_design+moy_ind_group)*theme(legend.position = "bottom")
-```
-
-```{r data-heatmap-wkndwday-group2}
-
-data_wkndwday <- data_pick  %>%
-  left_join(cluster_result)%>% create_gran("wknd_wday")  %>% 
-  left_join(data_pick_cust, by = c("customer_id"))
-
-ylim1 = boxplot.stats(data_wkndwday$general_supply_kwh)$stats[c(1, 5)]
-
-wkndwday_ind_design <- data_wkndwday%>% 
-  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
-  #lvplot::geom_lv(aes(fill = as.factor(design), 
-   #                   color = as.factor(design)), k=5, alpha = 0.5) +
-  geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
-  #geom_boxplot(outlier.shape = NA) + 
-  coord_cartesian(ylim = ylim1*1.05)+
-  facet_wrap(design~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-              ncol = 24)  +
-  theme_application() +
-  scale_fill_viridis_d(direction = 1)+
-  scale_color_viridis_d(direction = 1)
-
-```
-
-```{r}
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) + theme(legend.position = "bottom")
+
 wkndwday_ind_group <- data_wkndwday%>% 
+  left_join(cluster_result, by = c("customer_id")) %>% 
+  mutate(group  = as.factor(group)) %>% 
   ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
@@
           #            color = as.factor(group)), k=5, alpha = 0.5) +
-  geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
+  geom_boxplot(aes(color = group, fill = group),alpha = 0.5)+
   #geom_boxplot(outlier.shape = NA) + 
   coord_cartesian(ylim = ylim1*1.05)+
-  facet_wrap(design~customer_id, 
+  facet_wrap(group~customer_id, 
              scales = "free_y", 
              labeller = "label_value",
-             ncol = 24)  +
-  theme_application() +
-  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00",  "grey", "black"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))
-
-(wkndwday_ind_design / wkndwday_ind_group)* theme(legend.position = "none")
-```
-
-##  JS-based clustering
-
-The $24$ prototypes are clustered using the methodology described in \ref{Sec:methodology}.
-The distribution of electricity demand for the selected $24$ customers across hour-of-day and month-of-year are shown in the right panel of Figures \ref{fig:hod-combined} and \ref{fig:moy-combined} respectively.
-The median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.
-Customers are placed in the same order on both sides for both figures to simplify one-to-one comparison.
-All customers with the same color represent the same design (left) or cluster (right).
-Overall, the clustering method has been able to capture the shapes in each designs quite well, irrespective of the fact that the designs are only selected based on the $50^{th}$ percentile with Designs $1$, $2$, $3$, $4$, and $5$ correspond to groups $4$, $5$, $1$, $3$, $2$ and, in most situations.Because our method uses distribution across hod, moy, and wknd-wday, there may be some mismatches from the designs; nonetheless, visualising individual customers can help us determine whether the grouping by clustering is better than the planned grouping through designs.
-In a few circumstances, for example, as shown in Figures \ref{fig:hod-combined} and \ref{fig:moy-combined}, the clustering approach outperforms the design.
-The plotting scales are not displayed since we want to emphasize comparable shapes rather than scales.
-A customer in the cluster may have low daily or total energy usage, but their behavior may be quite similar to a customer with high usage.
-\noindent
-
-<!-- The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
--->
-
-Characterization of clusters both statistically and qualitatively is an important stage of a cluster analysis.
-A potential way is to look at the findings from all the groups in graphs, and enhance our qualitative descriptions of the groupings.
-Figure \ref{fig:combined} shows the distribution of the summarized groups and help us to characterise each of the clusters.
-All of these may be validated with further information about the customer.
-
-_Group 1:_ This group a strong early morning and late night hours. These consumers may be flexible students or elderly retirees who are night owls.
-Their day time usage has high variability.
-They have heaters on in the winter but consume less energy in the summer.
-Winter usage may also be due to increased usage of heaters at night when they are up, but typically very less variability in other months.
-
-_Group 2:_ They appear to work 9-5, get up and perform morning activities from 7-10, and then go.
-Evenings are busier than mornings as people return home to cook supper and perform other activities.
-During the fall (March-June), these consumers' average energy use declines, before rising again during the winter.
-However, energy behavior varies more in the winter than in the summer and a bit in the fall months.
-
-_Group 3:_ Presence of children or stay-at-home parents is indicated by the group's almost equivalent morning, afternoon and evening median use and variability across all hours.
-They have low usage for winter months (or maybe due to use of gas heaters, it is not reflected in electricity usage).
-It seems that the users in this group are more concerned about the comfort and quality of life than the cost of electricity as we see their electricity demand peaking up in the spring and summer.
-
-_Group 4:_ These users have an opposite monthly profile to Group 3, with more usage during summer months (probably due to increased use of air conditioners) and demand decreasing across the year with least demand in winter months.
-The users in this cluster are sensitive to high temperature, but insensitive to lower temperature or their heater usage is not reflected in their electricity usage.
-
-_Group 5:_ Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
-This group is more typical in their monthly energy behavior with more energy usage in winter and moderately decreasing for fall and spring months.
-The pattern in the variability is similar to that of the median.
-
-<!-- _Group 5:_ Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
-They have a flat monthly profile except for winter months, with higher median usage and variability than typically seen in other seasons.
--->
-
-<!-- It seems that the users in this group are more concerned about the comfort and quality of life than the cost of electricity as opposed to Group 1 and 3, who do not consume more electricity in the summer.
--->
-
-<!-- insensitive to high temperature but have their heaters on in the winter months May-Aug. -->
-
-<!-- _Group 2:_  This is the group that rushes out of the house in the morning to get to work.
-They only return at night and do all activities at night, so there is no morning peak.
-In addition, the users in this cluster are insensitive to high temperature, but sensitive to lower temperature.
--->
-
-<!-- _Group 3:_  This group a strong early morning and late night hours.
-These consumers may be flexible students or elderly retirees who are night owls.
-They have heaters on in the winter but consume less energy in the summer.
-Winter usage may also be due to increased usage of heaters at night when they are up.
--->
-
-<!-- _Group 4:_ Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
-They have a flat monthly profile, indicating the usage across all months are similar.
-It seems that the users in this group are more concerned about the comfort and quality of life than the cost of electricity as opposed to Group 1 and 3, who do not consume more electricity in the summer.
--->
-
-```{r data-heatmap-hod-group-new}
+             nrow=24)  +
+  theme_application()+
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7"))  + theme(legend.position = "none")+
+  xlab("wdwn")
+
+hod_ind_design + moy_ind_design + wkndwday_ind_design + hod_ind_group + moy_ind_group + wkndwday_ind_group +  plot_layout(guides = "collect", ncol = 6) + plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')') & theme(legend.position = 'none')
+```
+
+```{r data-heatmap-hod-group}
 legend_title <- "group"
@@
   theme_bw()  +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank())+
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
   scale_x_discrete(breaks = seq(1, 24, 3))+ 
   #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
+  theme_application() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +
   theme(legend.position = "bottom")
@@
 
-```{r data-heatmap-moy-group-new}
+```{r data-heatmap-moy-group}
 data_heatmap_moy_group <- quantile_gran(data_group,
@@
   ylab("demand (in Kwh)") +
-  theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + 
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
-  theme(legend.position = "bottom") +
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
-```
-
-```{r validation}
-data_validation <- (hod/24) %>% broom::tidy() %>% 
-  rename("hod" = "distance") %>% 
-  left_join((moy/12) %>% 
-              broom::tidy(), 
-            by = c("item1", "item2")) %>% 
-  rename("moy" = "distance") %>% 
-  left_join((wkndwday/2) %>% 
-              broom::tidy(), 
-            by = c("item1", "item2"))%>% 
-  rename("wkndwday" = "distance") %>% 
-  left_join(cluster_result, by = c("item1" = "customer_id")) %>% 
-  rename("group_item1" = "group") %>% 
-  left_join(cluster_result, by = c("item2" = "customer_id")) %>%  
-  rename("group_item2" = "group") %>% 
-  pivot_longer(3:5,names_to="gran",
-               values_to = "distance")
-```
-
-```{r validation-data}
-
-hod_cat <- suppressMessages(
-data_pick %>% 
-  dist_gran_cat(gran1 = "hour_day", response = "general_supply_kwh")) %>% 
-  mutate(gran = "hod")
-
-moy_cat <- suppressMessages(data_pick %>% 
-  dist_gran_cat(gran1 = "month_year", response = "general_supply_kwh"))%>%
-  mutate(gran = "moy")
-
-wkndwday_cat <- suppressMessages(data_pick %>% 
-  dist_gran_cat(gran1 = "wknd_wday", response = "general_supply_kwh"))%>% 
-  mutate(gran = "wnwd")
-
-data_validation_cat <- bind_rows(hod_cat, moy_cat, wkndwday_cat)
-
-gran_cat_dist <- data_validation_cat %>% 
-  left_join(cluster_result, by = c("customer_from" = "customer_id")) %>% 
-  rename("group_from" = "group") %>% 
-  left_join(cluster_result, by = c("customer_to" = "customer_id")) %>%  
-  rename("group_to" = "group") %>% 
-  group_by(gran,
-           category,
-           group_from, 
-           group_to) %>% 
-  summarise(sum = sum(distance),.groups = 'drop') %>% 
-  pivot_wider(names_from = group_to, values_from = sum) %>% 
-  mutate(distance = sum(`1`, `2`, `3`, `4`, `5`)) %>% 
-  select(-(4:8)) %>% 
-  arrange(-distance)
-
-```
-
-```{r hod-bubble}
-gran_cat_hod <-   gran_cat_dist %>% 
-  filter(gran=="hod")
-
-gran_cat_hod$category <- factor(gran_cat_hod$category, levels = 0:23)
-
-group_total <- gran_cat_hod %>% 
-  ungroup() %>% 
-    group_by(group_from) %>% 
-    summarise(group_total = sum(distance),.groups = 'drop')
-
-gran_cat_hod %>% 
-ungroup %>% 
-    left_join(group_total, by = "group_from") %>% 
-    mutate(cat_contibution = distance*100/group_total) %>% 
-    group_by(group_from) %>% 
-    arrange(-distance) %>% slice(c(1:4)) %>% 
-      ggplot(aes(group_from, category, color = as.factor(group_from))) +
-      geom_point(aes(size = distance)) +
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+ theme_light()
-```
-
-```{r hod-cont}
-data_pcp <- gran_cat_hod %>% 
-    ungroup %>% 
-    left_join(group_total, by = "group_from") %>% 
-    mutate(cat_contibution = distance*100/group_total) %>% 
-    group_by(group_from) %>% 
-    arrange(-distance) %>% 
-  select(group_from, category, cat_contibution) %>% 
-  pivot_wider(names_from = "category", values_from = "cat_contibution") %>% mutate(group_from = as.factor(group_from))
-
-parcoord <- GGally::ggparcoord(data_pcp ,
-                   columns = 2:ncol(data_pcp),
-                   groupColumn = "group_from",
-                   showPoints = TRUE, 
-                   alphaLines = 1,
-                   order = "anyClass",
-                   scale = "globalminmax"
-) + 
-  ggplot2::theme(
-    plot.title = ggplot2::element_text(size=10)
-  )+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
-  theme(legend.position = "bottom") +
-  xlab("") +
-  ylab("wpd")+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black")) + theme_bw()
-   parcoord 
-```
-
-```{r moy-bubble, eval = FALSE}
-gran_cat_moy <-   gran_cat_dist %>% 
-  filter(gran=="moy")
-
-gran_cat_moy$category <- factor(gran_cat_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
-
-group_total <- gran_cat_moy %>% 
-  ungroup() %>% 
-    group_by(group_from) %>% 
-    summarise(group_total = sum(distance),.groups = 'drop')
-
-gran_cat_moy %>% 
-ungroup %>% 
-left_join(group_total, by = "group_from") %>% 
-mutate(cat_contibution = distance*100/group_total) %>% 
-group_by(group_from) %>% 
-arrange(-distance) %>% slice(c(1:3)) %>% 
-    ggplot(aes(group_from, category, color = as.factor(group_from))) +
-      geom_point(aes(size = distance))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+ theme_light()
-
-```
-
-```{r moy-contr, eval = FALSE}
-
-data_pcp <- gran_cat_moy %>% 
-ungroup %>% 
-left_join(group_total, by = "group_from") %>% 
-mutate(cat_contibution = distance*100/group_total) %>% 
-group_by(group_from) %>% 
-arrange(-distance) %>% 
-  select(group_from, category, cat_contibution) %>% 
-  pivot_wider(names_from = "category", values_from = "cat_contibution") %>% mutate(group_from = as.factor(group_from))
-
-parcoord <- GGally::ggparcoord(data_pcp ,
-                   columns = 2:ncol(data_pcp),
-                   groupColumn = "group_from",
-                   showPoints = TRUE, 
-                   alphaLines = 1,
-                   order = "anyClass",
-                   scale = "globalminmax"
-) + 
-  ggplot2::theme(
-    plot.title = ggplot2::element_text(size=10)
-  )+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
-  theme(legend.position = "bottom") +
-  xlab("") +
-  ylab("wpd")+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black")) + theme_bw()
-    parcoord
-
-```
-
-```{r wkndwday-bubble, eval = FALSE}
-gran_cat_wkndwday <-   gran_cat_dist %>% 
-  filter(gran=="wnwd")
-
-gran_cat_wkndwday$category <- factor(gran_cat_wkndwday$category, levels = c("Weekday", "Weekend"))
-
-group_total <- gran_cat_wkndwday %>% 
-  ungroup() %>% 
-    group_by(group_from) %>% 
-    summarise(group_total = sum(distance),.groups = 'drop')
-
-gran_cat_wkndwday %>% 
-ungroup %>% 
-left_join(group_total, by = "group_from") %>% 
-mutate(cat_contibution = distance*100/group_total) %>% 
-group_by(group_from) %>% 
-arrange(-distance) %>% slice(c(1)) %>% 
- ggplot(aes(group_from, category, color = as.factor(group_from))) +
- geom_point(aes(size = distance))+
- scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+ theme_light()
-```
-
-```{r data-heatmap-hod-group}
-legend_title <- "group"
-
-data_group <- data_pick  %>% 
-  left_join(cluster_result, by = c("customer_id"))
-
-data_heatmap_hod_group <- quantile_gran(data_group,
-                                  gran1="hour_day",
-                                  quantile_prob_val = c(0.25, 0.5, 0.75),
-                                  group="group") %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values)
-
-data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
-
-data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")
-
-hod_group <- data_heatmap_hod_group %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`,
-                  group=group,
-                  fill = as.factor(group), alpha = 0.5),
-              alpha = 0.5) +
-  geom_line(aes(y = `50%`,
-                group=group, 
-                color = as.factor(group)), size = 1)+
-  facet_wrap(~group, 
-             scales = "free_y",  
-             ncol = 5) + 
-              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
-  ylab("demand (in Kwh)") + 
-  theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank())+
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
-  scale_x_discrete(breaks = seq(1, 24, 3))+ 
-  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
-  scale_fill_manual(legend_title, values = c("#E69F00",
-  "#009E73","#0072B2", "#D55E00", "grey")) + 
+  theme_bw() + theme_application() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +
   theme(legend.position = "bottom") 
-
-```
-
-```{r data-heatmap-moy-group}
-data_heatmap_moy_group <- quantile_gran(data_group,
-                                  gran1="month_year",
-                                  quantile_prob_val = c(0.25, 0.5, 0.75),
-                                  group="group") %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values)
-
-data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
-
-data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")
-
-moy_group <- data_heatmap_moy_group %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
-  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
-  facet_wrap(~group, 
-             scales = "free_y", 
-             labeller = "label_value",
-             ncol = 5) +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
-  ylab("demand (in Kwh)") +
-  theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + 
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
-  theme(legend.position = "bottom") +
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
-```
-
-```{r combined-groups, fig.cap = The distribution of electricity demand for the clusters across hour-of-day, month-of-year and wknd-wday.
-The median is represented by a line and the shaded region represents the area between $25^{th}$ and $75^{th}$ percentile.
-Group 1 and 3 have a stronger hour-of-day pattern, while group 2, 3, 4 have a month-of-year pattern.
-For wknd-wday differences across different groups are not distinct suggesting that it might not be that important a variable to distinguish different clusters."}
-combined <- (hod_group/moy_group) & theme(legend.position = "bottom")
-combined + plot_layout(guides = "collect")
 ```
@@
   #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
-  geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
+  geom_boxplot(aes(fill = group, color = group),alpha = 0.5, outlier.alpha = 0.05)+
   #geom_boxplot(outlier.size = 1) + 
@@
              labeller = "label_both") + 
-  theme_bw() +
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + 
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
-  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
-  scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
-  theme(legend.position = "bottom") +
-  theme(panel.spacing =unit(0, "lines")) + 
-  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
-  wkndwday_group
-```
-
-##  wpd-based clustering
-
-```{r ggpairs, fig.cap = "The ggpairs plot of electricity demand against the three granularities hod, moy and wkndwday are shown.
-The correlation between any two granularities is not so high so as to pose a threat to the clustering methodology."}
-
+  theme_bw() + theme_application() +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +
+  theme(legend.position = "none") 
+```
+
+```{r combined-groups-js, fig.cap = "The distribution of electricity demand for the clusters across hod (a), moy (b) and wkndwday (c).
+It seems like group 2 and 5 have a hod pattern across its members, while group 1, 3, 5 have a moy pattern.
+Wknd-wday variations across groups are not distinguishable, indicating that it is not a critical variable for clustering.
+It is helpful to compare the summarised distributions of groups to that of individuals to confirm that the most of individuals in the group have the same characterisation."}
+
+ (hod_group + moy_group + wkndwday_group) +
+  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
+  plot_layout(guides = "collect")& theme(legend.position = 'bottom')
+```
+
+### wpd-based distances
+
+We chose the optimal number of clusters using (@Hennig2014-ah) as $3$.
+A parallel coordinate plot with the three significant cyclic granularities used for wpd-based clustering.
+The variables are sorted according to their separation across classes (rather than their overall variation between classes).
+This means that $moy$ is the most important variable in distinguishing the designs followed by $hod$ and $wkndwday$.
+There is only one customer who has significant $wpd$ across $wkndwday$ and stands out from the rest of the customers.
+Group 3 has a higher wpd for hod than moy or wkndwday.
+Group 2 has the most distinct pattern across moy.
+Group 1 is a mixed group that has strong patterns on at least one of the three variables.
+The findings vary from js-based clustering, yet it is a helpful grouping.
+
+```{r data-pick-wpd}
 elec_600_wpd <- read_rds(here::here("data/algo2-cust600-wpd-rawdata.rds"))
@@
 
-```
-
-```{r, fig.cap=" A ggpairsplot and parallel coordinate plot are used to depict each of the 24 customers.
-The ggpairs plot four distinct clusters across the month-of-year, which are less prominent across the hour-of-day and wknd-wday.
-The parallel coordinate plot ranks the variables in order of importance, indicating that the month-of-year is the most important in identifying clusters, whereas wkdn-wday is the least significant and has the least variability among the three variables.", message=FALSE, warning=FALSE}
-
 scaled_var <- elec_pick_wide
@@
 
-group <- f%>% hclust (method = "ward.D") %>% cutree(k=5)
+k = array()
+for(i in 2:20)
+{
+group <- f %>% hclust (method = "ward.D") %>% cutree(k=i)
+p <- cluster.stats(f, clustering = group, silhouette = TRUE)
+k[i]=p$sindex
+}
+# ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 20, 1))
+
+group <- f%>% hclust (method = "ward.D") %>% cutree(k=3)
 
@@
 
-pairsplot <- ggpairs(data_pcp, 
-                     columns = 3:5,
-                     aes(fill=group, 
-                     color = group)) +
-  scale_fill_viridis_d(direction = -1) +
-  scale_color_viridis_d(direction = -1) +
-  theme_light()
+data_table <- data_pcp %>% group_by(group) %>% 
+  summarise(nobs = n(),
+            moy = round(median(month_year),2),
+            hod = round(median(hour_day),2),
+            wdwn = round(median(wknd_wday),2)) %>% 
+  select(-group)
+
+rownames(data_table) <- c("group-1", "group-2", "group-3")
+
+```
+
+```{r parcoord, fig.cap="Each of the 24 customers is represented by a parallel coordinate plot (a) with three wpd-based groupings.
+The plot shows that moy is the most important variable in identifying clusters, whereas wkdn-wday is the least significant and has the least fluctuation.
+One particular customer with high wpd across wknwday stands out in this display.
+Group 3 has a higher wpd for hod than moy or wkndwday.
+Group 2 has most discernible pattern across moy.
+Group 1 is a mixed group with strong patterns on atleast one of the three variables.
+All of these could be observed from the plot or the table (b) which shows median wpd values for each group.", message=FALSE, warning=FALSE}
 
@@
   xlab("") +
-  ylab("wpd") + scale_fill_viridis_d(direction = -1) +
-  scale_color_viridis_d(direction = -1) + theme_light()
-
-pairsplot
-
-```
-
-```{r parcoord, fig.cap = "A parallel coordinate plot with the three significant cyclic granularities used for wpd-based clustering.
-The variables are sorted according to their separation across classes (rather than their overall variation between classes).
-This means that moy is the most important variable in distinguishing the designs followed by hod and wkndwday.
-It can be observed that cluster 3 and 4 are distinguished by moy while 1 and 2 are distinguished by hod.
-Here, wkndwday is still acting as the nuisance variable."}
-parcoord
-```
-
-A parallel coordinate plot with the three significant cyclic granularities used for wpd-based clustering.
-The variables are sorted according to their separation across classes (rather than their overall variation between classes).
-This means that moy is the most important variable in distinguishing the designs followed by hod and wkndwday.
-It can be observed that clusters are well separated by month-of-year, while hour-of-day and weekend-weekday are not useful distinguishing the clusters produced with this clustering method.
-There is only one customer who has significant wpd across weekend-weekday and stands out from the rest of the customers.
-3 and 4 are distinguished by moy while 1 and 2 are distinguished by hod.
-Here, wkndwday is still acting as the nuisance variable.
-The ggpairs plot four distinct clusters across the month-of-year, which are less prominent across the hour-of-day and wknd-wday.
-The parallel coordinate plot ranks the variables in order of importance, indicating that the month-of-year is the most important in identifying clusters, whereas wkdn-wday is the least significant and has the least variability among the three variables.
-
-<!-- \noindent We discover 4 qualitative clusters of varying shapes in the distribution of all consumers in the first panel of Figure \ref{fig:combined}.
-Group-1 includes consumers who work 9-5, get up and conduct morning activities from 7-10am, and then depart.
-Then they return home in the evening to cook supper and perform other activities, giving the evening a greater peak than the morning.
-Group 2 is the group that rushes out of the house in the morning to get to work.
-They only return at night and do all activities at night, so there is no morning peak.
-The third one has a strong early morning and late night hours.
-These consumers may be flexible students or elderly retirees who are night owls.
-Presence of children or stay-at-home parents is indicated by Group-4's almost equivalent morning, afternoon and evening profile.
-All of this may be validated with further information about the customer.
--->
-
-<!-- \noindent The second panel of Figure \ref{fig:combined} shows that month-of-year qualitative clusters are not as distinguishable as hour-of-day.
-Group 2 is the most distinct and uses the most power during the summer, possibly owing to the use of air conditioners.
-Group 4 has a flat profile, indicating no significant month-to-month changes.
-Groups 1 and 3 have heaters on in the winter but consume less energy in the summer.
-Since gas is not available in all of NSW LGAs, it is possible that customers' heater usage is recorded in electricity rather than gas.
--->
-
-<!-- \noindent The third panel of Figure \ref{fig:combined} shows that the wknd-wday groups exhibit no significant changes across clusters, indicating that they may be a nuisance variable for these consumers.
--->
-
-<!-- \noindent The plotting scales are not displayed since we want to emphasise comparable shapes rather than scales.
-A customer in the cluster may have low daily or total energy usage, but their behaviour may be quite similar to a customer with high usage.
-That places them in the same group.
--->
+  ylab("wpd") + scale_fill_viridis_d(direction = 1) +
+  scale_color_viridis_d(direction = 1) + theme_light()
+
+(parcoord + gridExtra::tableGrob(data_table))+ plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')') & theme(legend.position = "bottom")
+
+```
+
+Things become far more complicated when we consider a larger data set with more uncertainty, as they do with any clustering problem.
+Summarizing distributions across clusters with varied or outlying customers can result in a shape that does not represent the group.
+Furthermore, combining hetergeneous customers may result in similar-looking final clusters that are not effective for visually differentiating them.
+It is also worth noting that the wknd-wday behavior in the given case does not characterize any cluster.
+This, however, will not be true for all of the customers in the data set.
+If more extensive prototype selection is used, resulting in more comprehensive prototypes in the data set, this method might be used to classify the entire data set into these prototype behaviors.
+However, the goal of this section was to have a few customers that have significant patterns over one or more cyclic granularities, apply our clustering methodology to cluster them, and demonstrate that the method produces useful clusters.
 
@@
 
-```{r}
-library(ochRe)
-  scale_colour_ochre(
-    palette="nolan_ned")
-```
+s

### 2021-10-29 main_application.Rmd
major rewrite application complete

@@
 library(viridis)
+library(pals)
 ```
@@
 However, the goal of this section was to have a few customers that have significant patterns over one or more cyclic granularities, apply our clustering methodology to cluster them, and demonstrate that the method produces useful clusters.
+
+```{r data-load-entire}
+# Read the nqt distances
+
+wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_nqt.rds")) %>% broom::tidy()
+
+moy <- read_rds(here("data/dist_gran_moy_356cust_nqt.rds")) %>% broom::tidy()
+
+hod <- read_rds(here("data/dist_gran_hod_356cust_nqt.rds")) %>% broom::tidy()
+
+# Make the distance metrics
+
+distance <- wkndwday %>% 
+  left_join(moy, by = c("item1", "item2")) %>% 
+  left_join(hod, by = c("item1", "item2")) %>% 
+  rename("wkndwday" ="distance.x",
+         "moy" = "distance.y",
+         "hod" = "distance") %>%
+  mutate(item1 = as.integer(as.character(item1)),
+         item2 = as.integer(as.character(item2)))
+
+filtered_distance <- distance %>%
+  filter(!(item1 %in% c(8196183, 8508008, 8680538))) %>% 
+  filter(!(item2 %in% c(8196183, 8508008, 8680538)))
+
+total_distance <- filtered_distance %>% 
+  mutate(total = wkndwday/2 + moy/12 + hod/24)
+
+total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
+                                                      names_from = item2,
+                                                      values_from = total)
+
+rownames(total_distance_wide) <- total_distance_wide$item1
+
+mds_data <- total_distance_wide %>% 
+  mutate_all(~replace(., is.na(.), 0)) %>%
+  tibble::rownames_to_column() %>%  
+  select(-item1) %>% 
+  pivot_longer(-rowname) %>% 
+  pivot_wider(names_from=rowname, values_from=value)
+
+rownames(mds_data) <- total_distance_wide$item1
+
+df <- mds_data[-1] %>% as.matrix()
+DM <- matrix(0, ncol(mds_data), ncol(mds_data))
+DM[lower.tri(DM)] = df[lower.tri(df, diag=TRUE)] # distance metric
+f = as.dist(DM)
+
+first_lot <- mds_data %>% names()
+
+id <- c(first_lot[-1], mds_data$name[nrow(mds_data)])
+
+# Find optimal number of clusters
+
+library(fpc)
+library(cluster)
+k = array()
+for(i in 5:50)
+{
+  group <- f %>% hclust (method = "ward.D") %>% cutree(k=i)
+  p <- cluster.stats(f, clustering = group, silhouette = TRUE)
+  k[i]=p$sindex
+}
+
+# ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 50, 2))
+# 
+# 
+# plot(k, type = "l")
+# 6 coming as the number of clusters with maximum silwidth
+
+group <- f %>% hclust (method = "ward.D") %>% cutree(k=16)
+cluster_result <- bind_cols(customer_id = id, group = group)
+
+# cluster_result %>% group_by(group) %>% count()
+
+```
+
+```{r data-heatmap-hod-group-entire}
+legend_title <- "group"
+
+data_group <- data_pick  %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+   gracsr::scale_gran( method = "robust",
+                       response = "general_supply_kwh") %>% 
+  left_join(cluster_result, by = c("customer_id"))
+
+data_heatmap_hod_group <- quantile_gran(data_group,
+                                  gran1="hour_day",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
+
+data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")
+
+hod_group_entire <- data_heatmap_hod_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=group,
+                  fill = as.factor(group), alpha = 0.5),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=group, 
+                color = as.factor(group)), size = 1)+
+  facet_wrap(~group, 
+             scales = "free_y",  
+             ncol = 16) + 
+              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
+  ylab("demand (in Kwh)") + 
+  theme_bw()  +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ 
+  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme_application() +
+  scale_fill_manual(values=as.vector(polychrome(17)))+
+  scale_color_manual(values=as.vector(polychrome(17)))+
+  theme(legend.position = "bottom")
+
+```
+
+```{r data-heatmap-moy-group-entire}
+data_heatmap_moy_group <- quantile_gran(data_group,
+                                  gran1="month_year",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")
+
+moy_group_entire <- data_heatmap_moy_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 16) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
+  ylab("demand (in Kwh)") +
+  theme_bw() + theme_application() +
+  scale_fill_manual(values=as.vector(polychrome(17)))+
+  scale_color_manual(values=as.vector(polychrome(17)))+
+  theme(legend.position = "bottom")
+```
+
+```{r data-wnwd-group-entire}
+wkndwday_data <- data_group %>%
+  create_gran("wknd_wday") 
+wkndwday_data$group <- as.factor(wkndwday_data$group)
+# %>% 
+#   create_gran("hour_day")
+
+ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]
+
+wkndwday_group_entire <- wkndwday_data%>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
+  geom_boxplot(aes(fill = group, color = group),alpha = 0.5, outlier.alpha = 0.05)+
+  #geom_boxplot(outlier.size = 1) + 
+  coord_cartesian(ylim = ylim1*1.05)+
+  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
+#geom_boxplot(aes(fill = as.factor(group))) +
+  #scale_fill_lv() +
+ xlab("wknd-wday") + 
+  ylab("demand (in Kwh)") +
+   facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_both",
+             ncol = 16) + 
+  theme_bw() + theme_application() +
+  scale_fill_manual(values=as.vector(polychrome(17)))+
+  scale_color_manual(values=as.vector(polychrome(17)))+
+  theme(legend.position = "none")
+```
+
+```{r combined-groups-js-entire, fig.cap = "The distribution of electricity demand for the clusters across hod (a), moy (b) and wkndwday (c).
+It seems like group 2 and 5 have a hod pattern across its members, while group 1, 3, 5 have a moy pattern.
+Wknd-wday variations across groups are not distinguishable, indicating that it is not a critical variable for clustering.
+It is helpful to compare the summarised distributions of groups to that of individuals to confirm that the most of individuals in the group have the same characterisation."}
+
+ (hod_group_entire/moy_group_entire/ wkndwday_group_entire) +
+  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
+  plot_layout(guides = "collect")& theme(legend.position = 'bottom')
+```
+
+```{r, eval=FALSE}
+library(ochRe)
+  scale_colour_ochre(
+    palette="nolan_ned")
+
+  scale_color_pals
+```

### 2021-10-29 clustering_all.R

@@ -66,22 +66,6 @@ id <- c(first_lot[-1], mds_data$name[nrow(mds_data)])
 # Find optimal number of clusters
 
 library(fpc)
-cluster.stats(f, clustering = group)
-
-koptimal = koptimal$kopt
-
-# Cluster them
-# 
-
-k = array()
-for(i in 5:50)
-{
-group <- f %>% hclust (method = "ward.D") %>% cutree(k=i)
-p <- cluster.stats(f, clustering = group)
-k[i]=p$avg.silwidth
-}
-
-library(fpc)
 library(cluster)
 k = array()
 for(i in 5:50)
@@ -91,11 +75,13 @@ for(i in 5:50)
   k[i]=p$sindex
 }
 
+ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 50, 2))
+
 
 plot(k, type = "l")
 # 6 coming as the number of clusters with maximum silwidth
 
-group <- f %>% hclust (method = "ward.D") %>% cutree(k=16)
+group <- f %>% hclust (method = "ward.D") %>% cutree(k=17)
 cluster_result <- bind_cols(customer_id = id, group = group) 
 
 cluster_result %>% group_by(group) %>% count()
@@ -136,14 +122,16 @@ data_heatmap_hod_group %>%
   facet_wrap(~group, 
              scales = "free_y", 
              labeller = "label_value",
-             nrow = 3) +
+             ncol = 17) +
   theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) +
   xlab("hour-of-day") + ylab("demand (in Kwh)") + 
   theme_bw() + theme(panel.spacing =unit(0, "lines")) +
   theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
   scale_x_discrete(breaks = seq(1, 24, 3))+ theme(
     strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm") ) + theme(legend.position = "bottom")
+  theme(plot.margin = margin(0, 0, 0, 0, "cm") ) + theme(legend.position = "bottom") +
+  scale_fill_manual(values=as.vector(polychrome(17)))+
+  theme_application() +theme(legend.position = "bottom")
 
 
 
@@ -164,10 +152,12 @@ data_heatmap_moy_group %>%
   facet_wrap(~group, 
[…]

### 2021-10-29 main.Rmd
minor changes in validation section

@@
 We spiked several attributes in the data to see where one method works better than the other and where they might give us the same outcome or the effect of missing data <!---and trends--> on the proposed methods.
-Three circular granularities $g1$, $g2$ and $g3$ are considered with categories denoted by ${g10,g11}$, ${g20, g21, g22}$ and ${g30, g31, g32, g33, g34}$ and levels $l_{g_1}=2$, $l_{g_2}=3$ and $l_{g_3}=5$.
+Three circular granularities $g1$, $g2$ and $g3$ are considered with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $l_{g_1}=2$, $l_{g_2}=3$ and $l_{g_3}=5$.
 These categories could be integers or some more meaningful labels.
@@
 So scenarios with all, few and just one significant granularities are considered.
-The code for creating these designs and the detailed results can be found in the Supplementary section (link to github repo).
+The code for creating these designs and the detailed results can be found in the Supplementary section (https://github.com/Sayani07/paper-gracsR).
 
@@
 As the methods ignore the linear progression of time, there is little value in adding time dependency in the data generating process.
-It is often reasonable to construct a time series using properties such as trend, seasonality, and auto-correlation.
-However, when examining distributions across categories of cyclic granularities, these time series features are lost or addressed independently by considering seasonal fluctuations through cyclic granularities.
-Because the time span during which an entity is observed in order to ascertain its behavior is not very long,
-the behavior of the entity will not change drastically and hence the time series can be assumed to remain stationary throughout the observation period.
-If the observation period is very long (for e.g more than 3 years), property, physical or geographical attributes might change leading to a non-stationary time series.
-But such a scenario is not considered here and the resulting clusters are assumed to be time invariant in the observation period.
 The data type is set to be "continuous," and the setup is assumed to be Gaussian.
@@
 To eliminate the effect of starting values, the first 500 observations in each series are discarded.
+-->
+
+<!-- It is often reasonable to construct a time series using properties such as trend, seasonality, and auto-correlation.
+However, when examining distributions across categories of cyclic granularities, these time series features are lost or addressed independently by considering seasonal fluctuations through cyclic granularities.
+Because the time span during which an entity is observed in order to ascertain its behavior is not very long, -->
+<!-- the behavior of the entity will not change drastically and hence the time series can be assumed to remain stationary throughout the observation period.
+If the observation period is very long (for e.g more than 3 years), property, physical or geographical attributes might change leading to a non-stationary time series.
+But such a scenario is not considered here and the resulting clusters are assumed to be time invariant in the observation period.
 -->
@@
 The difference in structure becomes quite clear when we see the distribution across cyclic granularities.
-Hence, for the consequent scenarios, only graphical displays across cyclic granularities are provided to emphasize the difference in structure.
+Hence, for the consequent scenarios, only graphical displays across cyclic granularities are provided to emphasize the distributional differences in categories.
 
@@
 
-All the methods were fitted to each data designs and results are reported through confusion matrices.
+All the methods were fitted to each data designs and results are reported through cluster statistics or plots.
 With increasing difference between categories, it gets easier for the methods to correctly distinguish the designs.
-For $mean_diff=1$, the performances are pretty bad for js-robust methods and wpd method for lower nT.
+For ${diff}=1$, the performances are pretty bad for js-robust methods and wpd method for lower nT.
 Although, with the kind of residential load datasets, a full year of load is the minimal requirement to capture expected variations in winter and summer profiles, for example.

### 2021-10-29 main.Rmd
characterisation of clusters subsection done

@@
 
-Characterization of clusters both statistically and qualitatively is an important stage of a cluster analysis.
-A potential way is to look at the findings from all the groups in graphs, and enhance our qualitative descriptions of the groupings.
-@Cook2007-qe provides several ways to characterize clusters.
-
-(a) Parallel coordinate plot: Parallel coordinate plot (@wegman1990hyperdimensional) are widely used to display high-dimensional and multivariate data, allowing visual grouping to detect patterns.
-In a Parallel Coordinates Plot, each variable has its own axis, which are all parallel.
-Each axis is connected by a series of lines.
-That is, each line is made up of connected points on each axis.
-The order of the axes can affect how the reader interprets the data.
-This is because adjacent variables are more easily perceived than non-adjacent variables and rearranging the axes can reveal patterns or correlations between variables.
-Scattered plots of the p variables are arranged in a scatterplot matrix.
-It's a neat way to show multiple relationships at once, and it allows us to compare all the plots at once.
-
-(b) Scatterplot matrix: The scatter plot matrix (draftsman's plot) is a matrix that comprises pairwise scatter plots of the p variables.Pairwise scatter plots are excellent for determining relationships between variables and determining which factors have contributed the most to clustering.
-
-(c) Plotting cluster statistics: For larger problems, parallel coordinate plots may become cluttered and difficult to read, therefore we may opt to display cluster statistics instead.
-(@Dasu_undated-gw)
-
-(c) MDS, PCA and t-SNE: While all of the techniques examine a matrix of distances or dissimilarities to give a representation of the data points in a reduced-dimension space, their goals are not the same.
-The principal component analysis (Johnson & Wichern 2002) attempts to retain data variance.
+\noindent Cluster characterization, both quantitatively and qualitatively, is a crucial aspect in cluster analysis.
+@Cook2007-qe lists numerous methods for characterising clusters.
+Listed below are a few techniques and R packages that are utilized in this study.
+
+(a) _Parallel coordinate plots_ (@wegman1990hyperdimensional) are often used to visualise high-dimensional and multivariate data, allowing visual grouping and pattern detection..
+A Parallel Coordinates Plot features parallel axes for each variable.
+Each axis is linked by lines.
+The axes' arrangement may affect the reader's interpretation of the data.
+Changing the axes may reveal patterns or relationships between variables for categorical variables.
+However, for categories with cyclic temporal granularities, preserving the underlying ordering is more desirable.
+
+(b) _Scatterplot matrix_ contains pairwise scatter plots of the p variables.
+Pairwise scatter plots are useful for figuring out how variables relate to each other and how factors determine the clustering.
+
+(c) _Displaying cluster statistics_ are useful when we have larger problems and it is difficult to read the Parallel coordinate plots due to congestion. (@dasu2005grouping)
+
+(d) _MDS, PCA and t-SNE_ While all of them use a distance or dissimilarity matrix to construct a reduced-dimension space representation, their goals are diverse.
+PCA seeks to retain data variance (Johnson and Wichern 2002).
 Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
 t-SNE, on the other hand, is concerned with preserving neighbourhood data points.
-Close data points in high-dimensional space will be condensed in the t-SNE embeddings.
-
-(d) Tour: A tour (@wickham2011tourr) is a collection of interpolated linear projections of multivariate data into a lower-dimensional space.
-The sequence is seen as a dynamic visualization, enabling the viewer to observe the shadows cast by the high-dimensional data in a lower-dimensional view.
-<!-- and also provides examples on how one might spin and brush using a tour, to build up a cluster analysis manually. -->
-
-Depending on the distance measure utilized for the study, the cluster characterization technique will differ.
-Clusters that utilize wpd-based distances are characterised using multi-dimensional scaling and parallel coordinate displays.
-For JS-based distances, the distribution across major granularities may be presented to ensure that the goal of similar shapes within clusters and distinct shapes across clusters is met.
-This technique may potentially make advantage of multi-dimensional scaling.
+The t-SNE embeddings will compress data points which are close in high-dimensional space.
+
+(e) _Tour_ is a collection of interpolated linear projections of multivariate data into lower-dimensional space.
+As a result, the viewer may observe the high-dimensional data's shadows from a low-dimensional perspective.
+
+The cluster characterization approach varies depending on the distance metric used.
+Parallel coordinate plots, scatter plot matrices, MDS or PCA are potentially useful ways to characterize clusters using wpd-based distances.
+For JS-based distances, plotting cluster statistics is beneficial for characterization and variable importance could be displayed through parallel coordinate plots.
+This part of the work uses R packages `GGally` (@R-GGally), `Rtsne` (@R-tsne), `ggplot2` (Wickham2009pk), `tour` (@wickham2011tourr), `stats` (@R-language).
 

### 2021-10-30 main.Rmd
cluster characterisation typo
added validation section graphs

@@
 (d) _MDS, PCA and t-SNE_ While all of them use a distance or dissimilarity matrix to construct a reduced-dimension space representation, their goals are diverse.
-PCA seeks to retain data variance (Johnson and Wichern 2002).
+PCA seeks to retain data variance.
 Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
@@
 Figure
-\ref{#fig:plot-linear}(left) shows the considered design.
+\ref{#fig:gran2and1-clubbed}(left) shows the considered design.
 
@@
 
-\noindent Here only one granularity is responsible for distinguishing the designs. Designs change significantly only for the granularity $g3$. Figure \ref{fig:plot-1gran}(right) shows this.
+\noindent Here only one granularity is responsible for distinguishing the designs. Designs change significantly only for the granularity $g3$. Figure \ref{fig:gran2and1-clubbed}(right) shows this.
 
@@
 
-All the methods were fitted to each data designs and results are reported through cluster statistics or plots.
-With increasing difference between categories, it gets easier for the methods to correctly distinguish the designs.
-For ${diff}=1$, the performances are pretty bad for js-robust methods and wpd method for lower nT.
+All the methods were fitted to each data designs to check if the formed clusters matches the design, if they are separated well and have low intra-cluster variation.
+Many of the potentially desired features of clusterings such as separation between clusters, high density within clusters, and distributional shapes can be explored graphically in a more holistic (if subjective) way than by looking at index values.
+So we use mds and parallel coordinate plots to demonstrate the results.
+and also for understanding variable importance.
+Results are summarized below:
+
+- In Figure \ref{fig:mds-plot-validation}, we tried to see how separated our clusters are.
+We observe that in all scenarios and for different mean differences, cluster is separated.
+However, the separation increases with increase in mean differences across scenarios.
+This is intuitive because with increasing difference between categories, it gets easier for the methods to correctly distinguish the designs.
+
+- Since the designs are constructed in a way that gives importance to three, two and one granularities respectively, we plot the inter cluster differences for each group under the iterations and plot them in a parallel coordinate plot.
+Figure \ref{fig:parcoord-sim} shows that in the first plot, no variable stands out in the clustering, whereas for the next two designs it is clear that g1, and g1 and g2 have little role to play in the cluster and thus the inter cluster distances are small compared to those variables.
+For more detailed plots and tables, please refer to the supplementary paper
+Each line represents each scenario, hence the point is not to look at cluster separation but more at the fact that which variable led to the clustering.
+The lines are not colored by group because the goal is not to highlight class division but rather the contribution of the factors to categorization.
+
+ - The performances get worse for js-robust and wpd method for $nT=300$, after which it picks up for larger $nT$ considered in the study.
 Although, with the kind of residential load datasets, a full year of load is the minimal requirement to capture expected variations in winter and summer profiles, for example.
@@
 The performance is promising except when the number of observations for a customer is really small.
-For smaller difference between categories, it is expected that method js-nqt would preform better than the other two.
+
+- For smaller difference between categories, it is expected that method js-nqt would preform better than the other methods.
+
+```{r mds-plot-validation, fig.cap = "Observations are projected through MDS and the first two dimensions are shown to display the relative position of clusters for different simulation parameters.
+The rows of the grid represent the different mean differences in designs and the columns represent different designs.
+It can be observed that clusters become more compact and separated for higher distributional differences between categories."}
+
+```
+
+```{r  parcoord-sim, fig.cap="Parallel coordinate plots are used to identify key variables for classification.
+Each line in this figure represents a formed group for a certain set of simulated parameters.
+(a), (b) and (c) corresponds to Scenarios (a), (b) and (c).
+The cyclic granularities $g1$, $g2$, and $g3$ are plotted on the x-axis, while total inter-cluster distances are plotted on the y-axis.
+Each line in this figure represents a group for a certain configuration of the various parameters ($nT$, $diff$, $R$).
+(c) demonstrates that intercluster distances are only significant for the variable $g3$, meaning that the clusters formed are primarily due to $g3$.
+This corresponds to our design in Scenario (c), in which distributional differences were implemented solely for g3.
+(b) demonstrates the role of g2 and g3 in clustering.
+(a) depicts a heterogeneous pattern of inter-cluster distances across variables, indicating that no single variable is to account for the clusters.
+"}
+
+```
+
 <!-- A confusion table can come alive with linked brushing, so that mismatches and agreements between methods can be explored. -->
@@
 f = as.dist(distance)
+
 ```

### 2021-10-30 main_application.Rmd
added validation section graphs

@@
   tibble::rownames_to_column() %>%  
-  select(-item1) %>% 
+  dplyr::select(-item1) %>% 
   pivot_longer(-rowname) %>% 

### 2021-10-30 CVI.R

@@ -1,0 +1,95 @@
+library(readr)
+library(tidyverse)
+library(fpc)
+library(patchwork)
+
+append_files_plot <- function(folder_name, path){
+  all_files = list.files(path = paste0(folder_name,path), 
+                         pattern = "data_validation_")
+  
+  names_levels <- map_dfr(all_files, 
+                          function(x){
+                            z = str_split(str_remove(x, ".rds"), "_") %>% 
+                              unlist()
+                            bind_cols(index = z[3])
+                          })
+  
+  all_files_path <- paste0(folder_name, path,
+                           all_files)  
+  
+  
+  all_data <- lapply(1:length(all_files_path), function(x){
+    
+    data = all_files_path %>% magrittr::extract2(x) %>% 
+      readRDS()
+    
+    names = names_levels %>% magrittr::extract(x,)
+    names_rep =   names %>% slice(rep(1:n(), each = nrow(data)))
+    bind_cols(names_rep, data)
+  }) %>% bind_rows() 
+# 
+# all_data <- append_files("js-nqt")
+
+# compute all inter cluster distance "from" group
+all_mat_mat_from <- all_data%>%
+  group_by(index, gran, group_item1, group_item2) %>% 
+  filter(group_item1!= group_item2) %>% 
+  summarise(sum = sum(distance)) %>% 
+  pivot_wider(names_from = group_item2, values_from = sum) %>% ungroup %>% 
+  replace(is.na(.), 0) %>% 
+  mutate(distance = rowSums(across(where(is.numeric)))) %>% 
+  dplyr::select(c(index, gran, group_item1, distance)) %>% 
+  rename("group"="group_item1")
+
+
+# compute all inter cluster distance "to" group
+
+all_mat_mat_to <- all_data%>%
+  group_by(index, gran, group_item1, group_item2) %>% 
+  filter(group_item1!= group_item2) %>% 
+  summarise(sum = sum(distance)) %>% 
+  pivot_wider(names_from = group_item1, values_from = sum) %>% ungroup %>% 
+  replace(is.na(.), 0) %>% 
+  mutate(distance = rowSums(across(where(is.numeric)))) %>% 
+  dplyr::select(c(index, gran, group_item2, distance))%>% 
+  rename("group"="group_item2")
+
+# compute all inter cluster distance to and from
+
+data_pcp <- bind_rows(all_mat_mat_from, all_mat_mat_to) %>% 
+  group_by(index, gran, group) %>% 
[…]

### 2021-10-30 old/CVI_old.R

@@ -1,0 +1,153 @@
+# compute dunn and avg silhoutte
+
+library(tidyverse)
+library(readr)
+g = read_rds("js-nqt/3gran_change_5D/data_validation_25.rds")
+
+total_distance <- g %>%
+  group_by(item1, item2) %>% 
+  summarise(total = sum(distance)) %>% ungroup()
+  
+total_distance_wide <- total_distance%>% 
+  pivot_wider(names_from = item2, values_from = total)
+
+
+rownames(total_distance_wide) <- total_distance_wide$item1
+
+mds_data <- total_distance_wide %>% 
+  mutate_all(~replace(., is.na(.), 0)) %>%
+  tibble::rownames_to_column() %>%  
+  dplyr::select(-item1) %>% 
+  pivot_longer(-rowname) %>% 
+  pivot_wider(names_from=rowname, values_from=value) 
+
+rownames(mds_data) <- total_distance_wide$item1
+
+df <- mds_data[-1] %>% as.matrix()
+DM <- matrix(0, ncol(mds_data), ncol(mds_data))
+DM[lower.tri(DM)] = df[lower.tri(df, diag=TRUE)] # distance metric
+f = as.dist(DM)
+
+mds <- f %>%
+  cmdscale() %>%
+  as_tibble()
+
+colnames(mds) <- c("Dim.1", "Dim.2")
+
+groups<-bind_rows(g %>%
+  distinct(item1, group_item1) %>% 
+  rename("item" = "item1",
+         "group" = "group_item1"),
+g %>% 
+  distinct(item2, group_item2) %>%
+  rename("item" = "item2",
+         "group" = "group_item2")) %>% distinct(item, group) 
+  
+
+
+all_data_cluster <- cbind(groups, mds) %>%
+  mutate(group = as.factor(group)) %>% as_tibble()
+
+
+mds_plot_10 <- ggplot(all_data_cluster,
+                      aes(x = Dim.1,
+                          y = Dim.2,
+                          color = group,
+                          label = rownames(mds))) +
+  geom_point(size = 2, alpha = 0.8) +
+  geom_text(check_overlap = TRUE)  +
+  theme_classic()+
+  theme(legend.position = "bottom") +
+  scale_color_brewer(palette = "Dark2")
+
+mds_plot_10
+
+# not symmetric because distance from one group to another might not be symmetric
+m <- g %>%
[…]

### 2021-10-30 validation/append_3gran_change_5D.R

@@ -5,11 +5,7 @@
 library(tidyverse)
 library(readr)
 
-
-
 folder_name = "js_robust"
-
-
 
 append_files <- function(folder_name){
 all_files = list.files(path = paste0(folder_name, "/3gran_change_5D/"), 

### 2021-10-30 main.Rmd
Visual exploration in validation section

@@
 
-## Results
-
-All the methods were fitted to each data designs to check if the formed clusters matches the design, if they are separated well and have low intra-cluster variation.
-Many of the potentially desired features of clusterings such as separation between clusters, high density within clusters, and distributional shapes can be explored graphically in a more holistic (if subjective) way than by looking at index values.
-So we use mds and parallel coordinate plots to demonstrate the results.
-and also for understanding variable importance.
-Results are summarized below:
+## Visual exploration
+
+All of the approaches were fitted to each data design and for each combination of the considered parameters.
+The formed clusters have to match the design, be well separated, and have minimal intra-cluster variation.
+It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values.
+So we use MDS and parallel coordinate graphs to demonstrate the findings:
 
@@
 This is intuitive because with increasing difference between categories, it gets easier for the methods to correctly distinguish the designs.
-
-- Since the designs are constructed in a way that gives importance to three, two and one granularities respectively, we plot the inter cluster differences for each group under the iterations and plot them in a parallel coordinate plot.
-Figure \ref{fig:parcoord-sim} shows that in the first plot, no variable stands out in the clustering, whereas for the next two designs it is clear that g1, and g1 and g2 have little role to play in the cluster and thus the inter cluster distances are small compared to those variables.
-For more detailed plots and tables, please refer to the supplementary paper
-Each line represents each scenario, hence the point is not to look at cluster separation but more at the fact that which variable led to the clustering.
-The lines are not colored by group because the goal is not to highlight class division but rather the contribution of the factors to categorization.
-
- - The performances get worse for js-robust and wpd method for $nT=300$, after which it picks up for larger $nT$ considered in the study.
-Although, with the kind of residential load datasets, a full year of load is the minimal requirement to capture expected variations in winter and summer profiles, for example.
-It is likely that $nT$ would be at least $1000$ with half-hourly data, even if data is only available just for a month.
-The performance is promising except when the number of observations for a customer is really small.
-
-- For smaller difference between categories, it is expected that method js-nqt would preform better than the other methods.
+Results corresponding to Scenario (a) is shown here.
+
+- Figure \ref{fig:parcoord-sim} depicts a parallel coordinate plot with the vertical bar showing total inter-cluster distances with regard to granularities $g1$, $g2$, and $g3$.
+For all simulation settings and scenarios, values are represented as a sequence of lines connected across each axis.
+One line in the figure, for example, shows the inter-cluster distances for one simulation scenario.
+The lines are not coloured by group since the purpose is to highlight the contribution of the factors to categorization rather than class separation.
+The first plot shows that no variable stands out in the clustering, but the following two designs show that {g1} and {g1, g2} have very low inter cluster distances, meaning that they did not contribute to the clustering.
+It is worth noting that these facts correspond to our original assumptions when developing the scenarios, which incorporate distributional differences over three (a), two (b), and one (c) significant granularities.
+Hence, Figure \ref{fig:parcoord-sim} (a), (b), and (c) validate the construction of scenarios (a), (b), and (c) respectively.
+
+ - The js-robust and wpd methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
+Although, with the type of residential load data sets, a complete year of load is the minimum requirement to capture predicted differences in winter and summer profiles, for example.
+Even if the data is only available for a month, $nT$ is expected to be at least $1000$ with half-hourly data.
+Hence, practically this is not a challenge as long as the performance is promising for higher $nT=300$.
+
+- For smaller difference between categories, it is seen that method js-nqt would preform better js-robust in our study sample.
+In our study sample, method js-nqt outperforms method js-robust for smaller differences between categories.
+More testing, however, is required to corroborate this.
+
+For more detailed plots and tables, please refer to the supplementary paper.
 

### 2021-10-30 main.Rmd
data design section done

@@
 
-\noindent Consider the scenario when all three granularities $g1$, $g2$, and $g3$ are responsible for distinguishing the designs.
-This implies that the patterns across each granularity will change significantly for at least one among the to-be-grouped designs.
-We consider different distributions across categories (as in Table \ref{tab:tab-dist-design} top) that will lead to different designs (as in Table \ref{tab:tab-dist-design} below).
-Figure \ref{fig:plot-3gran-new} shows the linear and cyclic representation of the simulated variable under these five designs.
-As could be seen from the plot, it is impossible to decipher the structural difference in the time series variable just by looking at the linear view.
-The difference in structure becomes quite clear when we see the distribution across cyclic granularities.
-Hence, for the consequent scenarios, only graphical displays across cyclic granularities are provided to emphasize the distributional differences in categories.
+\noindent Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction.
+Meaning that at least one of the designs to be categorized will have significantly different patterns for each granularity.
+In Table \ref{tab:tab-dist-design} (top), we explore various distributions across categories (as shown in Table Table \ref{tab:tab-dist-design} (bottom).
+Figure \ref{fig:plot-3gran-new} shows the simulated variable's linear and cyclic representations for each of these five designs.
+The structural difference in the time series variable is impossible to discern from the linear view.
+The shift in structure may be seen clearly in the distribution of cyclic granularities (Figure \ref{fig:plot-3gran-new} (right)).
+The following scenarios use solely graphical displays across cyclic granularities to highlight distributional differences in categories.
 
@@
 \noindent This is the case where one granularity will remain the same across all designs.
-We consider the case where the distribution of $v$ would vary across levels of $g2$ for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs.
-So $g3$ is not responsible for distinguishing across designs.
-Figure
-\ref{#fig:gran2and1-clubbed}(left) shows the considered design.
+We consider the case where the distribution of $v$ would vary across levels of $g2$ for all designs, across levels of $g3$ for few designs and $g1$ does not change across designs.
+The proposed design is shown in Figure \ref{fig:gran2and1-clubbed}(left).
 
@@
 
-\noindent Here only one granularity is responsible for distinguishing the designs. Designs change significantly only for the granularity $g3$. Figure \ref{fig:gran2and1-clubbed}(right) shows this.
+\noindent Only one granularity is responsible for identifying the designs in this case.
+This is depicted in Figure \ref{fig:gran2and1-clubbed} (right) where only $g3$ affects the designs significantly.
 
@@
 
-The proposed methods could be extended when two granularities of interest interact and we are interested to group subjects based on the interaction of the two granularities.
-For example, consider a group having a different weekday, weekend behavior in summer months, but not across winter.
-This type of joint behavior across granularities wknd-wday and month-of-year can be discovered by examining the distribution across combination of categories for different interacting granularities.
-Hence, in this scenario, we consider combination of categories to be generated from different distributions.
-For simplicity, consider a case with just two interacting granularities $g1$ and $g2$ of interest.
-As opposed to the last case, where we could examine distributions across $l_{g_1} + l_{g_2} = 5$ individual categories, with interaction, we need to examine the distribution of $l_{g_1}*l_{g_2}=6$ combination of categories.
-Consider $4$ designs in Figure \ref{fig:interaction-gran} where different distributions are assumed for different combination of categories resulting in different designs.
-Design-1 has no change in distributions across $g1$ or $g2$, while Design-2 and Design-3 change across only $g1$ and $g2$ respectively.
-Design-4 changes across categories of both $g1$ and $g2$.
-Design-3 and Design-4 looks similar according to their relative difference between consecutive categories, but Design-4 also changes across facets, unlike Design-3 where all facets look the same.
+The proposed methods could be extended when two granularities of interest interact and we want to group subjects based on the interaction of the two granularities.
+Consider a group that has a different weekday and weekend behavior in the summer but not in the winter.
+This type of combined behaviour across granularities  can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (wknd-wday and month-of-year in this example).
+As a result, in this scenario, we analyse a combination of categories generated from different distributions.
+Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$.
+In contrast to the previous situation, when we could study distributions across $l_{g_1} + l_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $l_{g_1}*l_{g_2}=6$ combination of categories.
+Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs.
+Design-1 exhibits no change in distributions across $g1$ or $g2$, whereas Design-2 and Design-3 alter across only $g1$ and $g2$, respectively.
+Design-4 varies across both $g1$ and $g2$ categories.
+Design-3 and Design-4 appear similar based on their relative difference across consecutive categories, however Design-4 also changes across facets, unlike Design-3, which has all facets look the same.
 
@@
 
-## Visual exploration
+## Visual exploration of findings
 

### 2021-10-30 main.Rmd
validation section done

@@
 A continuous measured variable $v$ of length $T$ indexed by $\{0, 1, \dots T-1\}$ is simulated such that it follows the structure across $g1$, $g2$ and $g3$.
-We created independent replications $R = \{25, 250, 500\}$ of all data designs to see if our proposed clustering approaches can detect distinct designs in various groups for small, medium and large number of series.
-A sample size of $T=\{300, 1000, 5000\}$ is used in all designs to test small, medium and large sized series.
-The methods could preform differently with different jumps between consecutive categories.
-So a mean difference of $diff = \{1, 2, 5\}$ for corresponding categories are also considered.
-The performance of the methods can vary with different number of significant granularities.
-So scenarios with all, few and just one significant granularities are considered.
-The code for creating these designs and the detailed results can be found in the Supplementary section (https://github.com/Sayani07/paper-gracsR).
+We constructed independent replications of all data designs $R = 25, 250, 500$ to investigate if our proposed clustering method can discover distinct designs in small, medium, and big number of series.
+All designs employ $T=300, 1000, 5000$ sample sizes to evaluate small, medium, and large sized series.
+Variations in method performance may be due to different jumps between categories.
+So a mean difference of $diff = 1, 2, 5$ is examined.
+The approaches' performance varies with the number of significant granularities.
+So all, few, and one major granularity scenarios are considered.
+Please see the Supplementary section (https://github.com/Sayani07/paper-gracsR) for the code and findings.
 
@@
 
-```{r mds-plot-validation, fig.cap = "Observations are projected through MDS and the first two dimensions are shown to display the relative position of clusters for different simulation parameters.
+```{r  parcoord-sim, fig.cap="The parallel coordinate plot illustrates the total inter-cluster distances for granularities $g1$, $g2$, and $g3$.
+One line in the figure depicts the inter-cluster distances for a single simulation scenario.
+While the first plot indicates that no variable stands out during clustering, the next two designs demonstrate that {g1} and {g1, g2} have extremely low inter-cluster distances, indicating that they did not contribute to clustering.
+It is worth emphasising that these facts are consistent with our initial assumptions when designing the scenarios.
+"}
+
+```
+
+```{r mds-plot-validation, fig.cap = "Cases are projected through MDS and the first two dimensions are shown to display the relative position of clusters for different simulation parameters.
 The rows of the grid represent the different mean differences in designs and the columns represent different designs.
@@
 
-```{r  parcoord-sim, fig.cap="Parallel coordinate plots are used to identify key variables for classification.
+<!-- "Parallel coordinate plots are used to identify key variables for classification.
 Each line in this figure represents a formed group for a certain set of simulated parameters.
@@
 (b) demonstrates the role of g2 and g3 in clustering.
-(a) depicts a heterogeneous pattern of inter-cluster distances across variables, indicating that no single variable is to account for the clusters.
-"}
-
-```
+(a) depicts a heterogeneous pattern of inter-cluster distances across variables, indicating that no single variable is to account for the clusters." -->
 
@@
 -->
-
-s

### 2021-10-31 main.Rmd
intro of methodology fixed

@@
 The proposed methodology aim to leverage the intrinsic temporal data structure hidden in time series data.
-The foundation of our method is unsupervised clustering algorithms based exclusively on the time-series data.
-The similarity measure is the most essential ingredient of time series clustering.
-The (dis) similarity measure in this paper focuses on looking at the (dis) similarity between underlying distributions that may have resulted in different patterns across different cyclic temporal granularities.
+The foundation of our method is unsupervised clustering algorithms based exclusively on the time-series data.The similarity measure is the most essential ingredient of time series clustering.
+First step is to decide what we mean by similar.
+The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.
+In this work, instead of considering the probability distributions of the linear time series, we assume that the measured variables across different categories of any cyclic granularity are from different data generating processes.
+Hence, we want to be able to cluster distributions of the form $f_{i,A,B \dots, {N_C}}(v)$, where $A, B$ represent the cyclic granularities under consideration such that $A = \{a_j: j=1, 2, \dots J\}$,  $B = \{b_k: k  = 1, 2, \dots K\}$ and so on.
+We consider individual category of a cyclic granularity ($A$) or combination of categories for interaction of cyclic granularities (for e.g.
+$A*B$) to have a distribution.
+For example, let us consider we have two cyclic  granularities of interest, $A = {0, 1, 2, \dots, 23}$ representing hour-of-day and  $B = \{Mon, Tue, Wed, \dots, Sun\}$ representing day-of-week.
+Each customer $i$ consist of a collection of probability distributions.
+In case individual granularities ($A$ or $B$) are considered, there are  $J = 24$ distributions of the form $f_{i,j}(v)$ or $K = 7$ distributions of the form $f_{i,k}(v)$ for each customer $i$.
+In case of interaction, $J*K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
+As a result, we need to decide how to measure similarities between these collections of univariate probability distributions.
+There are multiple ways to measure similarities between time series based on probability distributions, depending on the objective of the problem.
+This paper focuses on looking at the (dis) similarity between underlying distributions that may have resulted in different patterns across different cyclic temporal granularities, that eventually have resulted in the (dis) similarity between time series.
+It considers two methodologies for finding distances between time series, which are explained in the next segment.
+
+<!-- Once we have chosen relevant features to characterize the probability distributions, we have to decide how to combine them into a single number, which serves as a distance between two observations.
+-->
+
+<!-- which are described below. -->
+
+<!-- Consequently, the distance measure might also change depending on the similarity measure chosen. -->
+
+<!-- The (dis) similarity measure in this paper focuses on looking at the (dis) similarity between underlying distributions that may have resulted in different patterns across different cyclic temporal granularities and considers two methodology for finding distances between observations, which are explained in the next segment.
+-->
+
 It is worth noting that when studying these similarities, a variety of objectives may be pursued.

### 2021-10-31 main.Rmd
JS-distance better fluency

@@
 It considers two methodologies for finding distances between time series, which are explained in the next segment.
+
+**JS-based distances**
+
+This distance matrix considers two time series to be similar if every category of an individual cyclic granularity or combination of categories for interacting cyclic granularities have similar distributions.
+In this study, the distribution for each category is characterized using deciles (can potentially consider any list of quantiles) and the distances between distributions are computed by using the Jensen-Shannon distances (@Menendez1997-in), which is symmetric and hence could be used as a distance measure.
+
+\noindent The total distance between two series $x$ and $y$ with respect to a cyclic granularity $A$ is then defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity A) or  $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
+Once we have obtained the distance between two series with respect to on granularity, we have to combine them to obtain a distance based on all significant granularities.When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels, $$S_{x, y} = \sum_lS^{C_l}_{x,y}/n_l$$ is used, which is also a distance metric being the sum of JS distances.
+
+This distance matrix considers two time series to be similar if the distributions of each category of an individual cyclic granularity or combination of categories for interacting cyclic granularities are similar.
+In this study, the distribution for each category is characterised using deciles (can potentially consider any list of quantiles), and the distances between distributions are calculated using the Jensen-Shannon distances (@Menendez1997-in), which are symmetric and thus could be used as a distance measure.
+
+\noindent The sum of the distances between two series $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
+After determining the distance between two series in terms of one granularity, we must combine them to produce a distance based on all major granularities.
+When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels,  $$S_{x, y} = \sum_lS^{C_l}_{x,y}/n_l$$ is employed, which is also a distance metric since it is the sum of JS distances.
+
+**wpd-based distances**
+
+Compute weighted pairwise distances $wpd$ (@Gupta2021-hakear) for all considered granularities for all observations.
+$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalized by appropriate parameters.
+A higher value of $wpd$ indicates that some interesting pattern is expected, whereas a lower value would indicate otherwise.
+
+\noindent Once we have chosen $wpd$ as a relevant feature for characterizing the time series across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
+We choose the euclidean distance between them with the granularities being the variables and $wpd$ being the value under each variable.
+With this approach, we should expect the observations with similar $wpd$ values to be clustered together.
+Thus this approach is useful to group customers that have similar significance of patterns across different granularities.
+Similar significance do not imply similar pattern, and this how this approach differs from the JS-based distances which picks up differences in patters across categories.
 

### 2021-10-31 main.Rmd
better wording for wpd based distance

--- 2021-10-31 main.Rmd
@@
 It considers two methodologies for finding distances between time series, which are explained in the next segment.
+
+**JS-based distances**
+
+This distance metric considers two time series to be similar if the distributions of each category of an individual cyclic granularity or combination of categories for interacting cyclic granularities are similar.
+In this study, the distribution for each category is characterised using deciles (can potentially consider any list of quantiles), and the distances between distributions are calculated using the Jensen-Shannon distances (@Menendez1997-in), which are symmetric and thus could be used as a distance measure.
+
+\noindent The sum of the distances between two series $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
+After determining the distance between two series in terms of one granularity, we must combine them to produce a distance based on all significant granularities.
+When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels,  $$S_{x, y} = \sum_lS^{C_l}_{x,y}/n_l$$ is employed, which is also a distance metric since it is the sum of JS distances.
+
+**wpd-based distances**
+
+Compute weighted pairwise distances $wpd$ (@Gupta2021-hakear) for all considered granularities for all observations.
+$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalised by appropriate parameters.
+A higher value of $wpd$ indicates that some interesting pattern is expected, whereas a lower value would indicate otherwise.
+
+\noindent Once we have chosen $wpd$ as a relevant feature for characterizing the time series across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
+The euclidean distance between them is chosen, with the granularities acting as variables and $wpd$ representing the value under each variable.
+With this approach, we should expect the observations with similar $wpd$ values to be clustered together.
+Thus, this approach is useful for grouping observations that have similar significance of patterns across different granularities.
+Similar significance does not imply similar pattern, which is where this technique varies from JS-based distances, which detect differences in patterns across categories.
 

### 2021-10-31 main.Rmd
methodology flowchart introduced

@@
 This paper focuses on looking at the (dis) similarity between underlying distributions that may have resulted in different patterns across different cyclic temporal granularities, that eventually have resulted in the (dis) similarity between time series.
-It considers two methodologies for finding distances between time series, which are explained in the next segment.
-
-**JS-based distances**
-
-This distance metric considers two time series to be similar if the distributions of each category of an individual cyclic granularity or combination of categories for interacting cyclic granularities are similar.
-In this study, the distribution for each category is characterised using deciles (can potentially consider any list of quantiles), and the distances between distributions are calculated using the Jensen-Shannon distances (@Menendez1997-in), which are symmetric and thus could be used as a distance measure.
-
-\noindent The sum of the distances between two series $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
-After determining the distance between two series in terms of one granularity, we must combine them to produce a distance based on all significant granularities.
-When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels,  $$S_{x, y} = \sum_lS^{C_l}_{x,y}/n_l$$ is employed, which is also a distance metric since it is the sum of JS distances.
-
-**wpd-based distances**
-
-Compute weighted pairwise distances $wpd$ (@Gupta2021-hakear) for all considered granularities for all observations.
-$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalised by appropriate parameters.
-A higher value of $wpd$ indicates that some interesting pattern is expected, whereas a lower value would indicate otherwise.
-
-\noindent Once we have chosen $wpd$ as a relevant feature for characterizing the time series across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
-The euclidean distance between them is chosen, with the granularities acting as variables and $wpd$ representing the value under each variable.
-With this approach, we should expect the observations with similar $wpd$ values to be clustered together.
-Thus, this approach is useful for grouping observations that have similar significance of patterns across different granularities.
-Similar significance does not imply similar pattern, which is where this technique varies from JS-based distances, which detect differences in patterns across categories.
-
-<!-- Once we have chosen relevant features to characterize the probability distributions, we have to decide how to combine them into a single number, which serves as a distance between two observations.
--->
-
-<!-- which are described below. -->
-
-<!-- Consequently, the distance measure might also change depending on the similarity measure chosen. -->
-
-<!-- The (dis) similarity measure in this paper focuses on looking at the (dis) similarity between underlying distributions that may have resulted in different patterns across different cyclic temporal granularities and considers two methodology for finding distances between observations, which are explained in the next segment.
--->
-
-It is worth noting that when studying these similarities, a variety of objectives may be pursued.
+It considers a methodology with two appraoches for  finding distances between time series.
+Both of these appraoches may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
+The obtained distances could be fed into a clustering algorithm to break large data sets into subgroups that can then be analyzed separately.
+These clusters may be commonly associated with real-world data segmentation.
+However, since the data is unlabeled a priori, more information is required to corroborate this.
+The methodology is explained in the Figure \ref{fig:flowchart} and each element of the pipeline is discussed.
+
+```{r flowchart, fig.cap = "Flow chart illustrating the pipeline for methodology", out.width="100%"}
+knitr::include_graphics("figs/flowchart.png")
+
+```
+
+<!-- It is worth noting that when studying these similarities, a variety of objectives may be pursued.
 One objective could be to group time series with similar shapes over all relevant cyclic granularities.
 In this scenario, the variation in customers within each group is in magnitude rather than shape, while the variation between groups is only in shape.
-There
-are distance measures are used for shape-based clustering [Ding et al. 2008; Wang
-et al.
+There -->
+<!-- are distance measures are used for shape-based clustering [Ding et al. 2008; Wang -->
+<!-- et al.
 2013] and many more but none of them look at the probability distributions while computing similarity.
@@
 However, since the data is unlabeled a priori, more information is required to corroborate this.
-This section presents the work flow of the methodology:
+This section presents the work flow of the methodology:  -->
 
@@
 
-This distance matrix considers two objects to be similar if every category of an individual cyclic granularity or combination of categories for interacting cyclic granularities have similar distributions.
-In this study, the distribution for each category is characterized using deciles and the distances between distributions are computed by using the Jensen-Shannon distance, which is symmetric and hence could be used as a distance measure.
-
-\noindent The total distance between two elements $x$ and $y$ is then defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity A) or  $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
-When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels, $$S_{x, y} = \sum_lS^{C_l}_{x,y}/n_l$$ is used, which is also a distance metric being the sum of JS distances.
+This distance metric considers two time series to be similar if the distributions of each category of an individual cyclic granularity or combination of categories for interacting cyclic granularities are similar.
+In this study, the distribution for each category is characterised using deciles (can potentially consider any list of quantiles), and the distances between distributions are calculated using the Jensen-Shannon distances (@Menendez1997-in), which are symmetric and thus could be used as a distance measure.
+
+\noindent The sum of the distances between two series $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
+After determining the distance between two series in terms of one granularity, we must combine them to produce a distance based on all significant granularities.
+When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels,  $$S_{x, y} = \sum_lS^{C_l}_{x,y}/n_l$$ is employed, which is also a distance metric since it is the sum of JS distances.
+In this approach, the variation in time series within each group is in magnitude rather than distributional pattern, while the variation between groups is only in distributional pattern across categories.
 
@@
 
-Compute weighted pairwise distances ($wpd$) for all considered granularities for all objects.
-$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalized by appropriate parameters.
+Compute weighted pairwise distances $wpd$ (@Gupta2021-hakear) for all considered granularities for all observations.
+$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalised by appropriate parameters.
 A higher value of $wpd$ indicates that some interesting pattern is expected, whereas a lower value would indicate otherwise.
 
-\noindent Distance between elements is then taken as the euclidean distances between them with the granularities being the variables and $wpd$ being the value under each variable.
-Since Euclidean distance is chosen, the observations with high values of features ($wpd$ values) will be clustered together.
-The same holds true for observations with low values of features.
-Thus this distance matrix would be useful to group customers that have similar significance of patterns across different granularities.
+\noindent Once we have chosen $wpd$ as a relevant feature for characterizing the time series across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
+The euclidean distance between them is chosen, with the granularities acting as variables and $wpd$ representing the value under each variable.
+With this approach, we should expect the observations with similar $wpd$ values to be clustered together.
+Thus, this approach is useful for grouping observations that have similar significance of patterns across different granularities.
+Similar significance does not imply similar pattern, which is where this technique varies from JS-based distances, which detect differences in patterns across categories.
 

### 2021-10-31 main.Rmd
data pre-processing wording improved

@@
 
+- _Data transformation_
+
+Time series often have a somewhat skewed distribution.
+Furthermore, their ranges might vary greatly.
+It is helpful to do a statistical transformation on the data to bring all of them to the same range or at least normalise each series such that the CDF of the resultant variable is Gaussian.
+For the JS-based approaches, two data transformation techniques are utilised viz, Normal-Quantile Transform (NQT) and Robust scaling.
+NQT is a built-in transformation for computing $wpd$, which is the foundation of wpd-based distances.
+
+_Robust scaling_ To get a Gaussian probability distribution for an input variable, most scaling methods remove the mean from values and divide by the standard deviation (zero mean and unit variance).
+Standardization may be distorted or biased if the input variable contains outliers.
+To address this, robust scaling methods could be utilized (value – median) / (p75 – p25)) which results in a variable with a zero mean and median, as well as a standard deviation of one, while the outliers are still there with the same relative connections to other values.
+
+_Normal-Quantile transform_ First, the original data is ranked in ascending order and the probabilities $P(Y<=y(i)) = i/(n+1)$ are attached to $y(i)$, in terms of their ranking order.
+A NQT based transformation is applied by computing from a standard normal distribution a variable $\eta(i)$, which corresponds to the same probability $P(\eta< \eta(i)) = i/n+1$.
+By doing this, the new variables $\eta(i)$ will be marginally distributed according to standard Normal, N(0,1).
+NQT will make the skewed distributions bell-shaped.
+So it's hard to tell from the modified distribution which raw distribution was which.
+Also, multimodality is disguised or inverted.
+This, however, is not a problem for this methodology as distributions are characterised by quantiles and the order of the quantiles are reserved under NQT.
+
 <!-- It is worth noting that when studying these similarities, a variety of objectives may be pursued.
@@
 
-- _Individual or combined categories of cyclic granularities as DGP_
-
-The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
-In our work, we are using $i$ as denoting a customer and the underlying variable as the electricity demand.
-So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
-In this work, instead of considering the probability distributions of the linear time series, we assume that the measured variables across different categories of any cyclic granularity are from different data generating processes.
-Hence, we want to be able to cluster distributions of the form $f_{i,A,B \dots, {N_C}}(v)$, where $A, B$ represent the cyclic granularities under consideration such that $A = \{a_j: j=1, 2, \dots J\}$,  $B = \{b_k: k  = 1, 2, \dots K\}$ and so on.
-We consider individual category of a cyclic granularity ($A$) or combination of categories for interaction of cyclic granularities (for e.g.
-$A*B$) to have a distribution.
-For example, let us consider we have two cyclic  granularities of interest, $A = {0, 1, 2, \dots, 23}$ representing hour-of-day and  $B = \{Mon, Tue, Wed, \dots, Sun\}$ representing day-of-week.
-Each customer $i$ consist of a collection of probability distributions.
-In case individual granularities ($A$ or $B$) are considered, there are  $J = 24$ distributions of the form $f_{i,j}(v)$ or $K = 7$ distributions of the form $f_{i,k}(v)$ for each customer $i$.
-In case of interaction, $J*K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
-
-\noindent As a result, a distance between collections of these univariate probability distributions is required.
-Depending on the objective of the problem, there could be many approaches to considering such distances.
-This paper considers two approaches, which are explained in the next segment.
-
 - _Distance metrics_
@@
 -->
-
-- _Pre-processing steps_
-
-<!--Handling trend, seasonality, non-stationarity and auto-correlation:** Trend and seasonality are fundamental characteristics of time series data, and it is reasonable to define a time series according to its degree of trend and seasonality.
-These characteristics of the time series are lost or handled independently by considering probability distributions (trend is lost) across categories of cyclic granularities (by independently modeling all seasonal fluctuations), and so there is no need to de-trend or de-seasonalize the data before conducting the clustering method.
-There is no need to omit holiday or weekend patterns for similar reasons-->
-
-Practically most problems will have a very skewed distribution, it is often helpful to bring them to a normal-like shape before clustering.
-Two data transformation techniques are employed for the JS-based methods and NQT is built-in transformation used for computation of $wpd$, which forms the basis of wpd-based distances.
-
-_Robust scaling_ Standardizing is a common scaling method that subtracts the mean from values and divides by the standard deviation, resulting in a conventional Gaussian probability distribution for an input variable (zero mean and unit variance).
-If the input variable includes outlier values, standardisation may become skewed or prejudiced.
-To address this, robust scaling methods could be utilized  (value – median) / (p75 – p25)) which results in a variable with a zero mean and median, as well as a standard deviation of one, while the outliers are still there with the same relative connections to other values.
-
-_Normal-Quantile transform_  First as a data pre-processing step to make all assymetrical real world variables more symmetric, we perform a quantile-normal transform on the data.
-This makes sure that the CDF of the resulting variable in Gaussian.
-The original data is ranked in ascending order and the probabilities $P(Y<=y(i)) = i/(n+1)$ are attached to $y(i)$, in terms of their ranking order.
-A NQT based transformation is applied by computing from a standard normal distribution a variable $\eta(i)$, which corresponds to the same probability $P(\eta< \eta(i)) = i/n+1$.
-By doing this, the new variables $\eta(i)$ will be marginally distributed according to standard Normal, N(0,1).
-NQT will transform the positively and negatively skewed distribution to a similar bell-shaped.
-From the transformed distribution, it is difficult to understand that raw distribution was of which shape.
-Also, multimodality gets hidden or magnitude get reversed with NQT.
-But deciles from the distribution will move in a similar manner as the raw distribution and hence the final distance matrix seem to be unaffected.
-Hence, this could be used.
 

### 2021-10-31 main.Rmd
transformation section done

@@
 
+- _Find significant granularities or harmonies_
+
+[@Gupta2021-hakear] proposes a method for choosing significant cyclic granularities and harmonies, which is used in this work.
+We define "significant" granularities as those with significant distributional differences across categories.
+It is better to select only those granularities because it is expected that there would be some fascinating repetitive behaviour that we are interested in studying.
+It is worth noting that not all of the observations in the study may have the same set of important granularities.
+A method for selecting a list ($S_c$) of significant granularities for all observations may be as follows:
+
+ (a) eliminate from the comprehensive list the granularities that are inconsequential for all observations.
+
+ (b) consider only those granularities which are significant for most observations.
+
+In both circumstances, there will be observations for which one or a few selected granularities are uninteresting.
+Even in that situation, having this group of observations that show no intriguing patterns over a granularity that frequently detects patterns may be useful.
+In contrast, if the granularities under consideration are indeed significant for a set of observations, distinct patterns could be detected while clustering them.
+
 - _Data transformation_
 
-Time series often have a somewhat skewed distribution.
-Furthermore, their ranges might vary greatly.
-It is helpful to do a statistical transformation on the data to bring all of them to the same range or at least normalise each series such that the CDF of the resultant variable is Gaussian.
+Time series often have a somewhat skewed distribution and their ranges might vary greatly. It
+is helpful to do a statistical transformation on the data to bring all of them to the same range or normalize each series.
 For the JS-based approaches, two data transformation techniques are utilised viz, Normal-Quantile Transform (NQT) and Robust scaling.
@@
 
-_Robust scaling_ To get a Gaussian probability distribution for an input variable, most scaling methods remove the mean from values and divide by the standard deviation (zero mean and unit variance).
-Standardization may be distorted or biased if the input variable contains outliers.
-To address this, robust scaling methods could be utilized (value – median) / (p75 – p25)) which results in a variable with a zero mean and median, as well as a standard deviation of one, while the outliers are still there with the same relative connections to other values.
-
-_Normal-Quantile transform_ First, the original data is ranked in ascending order and the probabilities $P(Y<=y(i)) = i/(n+1)$ are attached to $y(i)$, in terms of their ranking order.
-A NQT based transformation is applied by computing from a standard normal distribution a variable $\eta(i)$, which corresponds to the same probability $P(\eta< \eta(i)) = i/n+1$.
-By doing this, the new variables $\eta(i)$ will be marginally distributed according to standard Normal, N(0,1).
+_Robust scaling_ The normalised $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - p_{0.50}}{p_{0.75}-p_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $p_{0.25}$, $p_{0.50}$ and $p_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
+$v_{norm}$ has zero mean and median, as well as a standard deviation of one, while the outliers are still there with the same relative connections to other values.
+
+_Normal-Quantile transform_ The raw data for all observations is individually normal-quantile transformed (NQT) [@Krzysztofowicz1997-bv], so that the transformed data follows a standard normal distribution.
 NQT will make the skewed distributions bell-shaped.
-So it's hard to tell from the modified distribution which raw distribution was which.
+As a result, determining which raw distribution was used is difficult using the modified distribution.
 Also, multimodality is disguised or inverted.
-This, however, is not a problem for this methodology as distributions are characterised by quantiles and the order of the quantiles are reserved under NQT.
+This, however, is not a problem for implementation of this methodology as distributions are characterised by quantiles and the order of the quantiles is reserved under NQT.
+
+<!-- First, the original data is ranked in ascending order and the probabilities $P(Y<=y(i)) = i/(n+1)$ are attached to $y(i)$, in terms of their ranking order.
+A NQT based transformation is applied by computing from a standard normal distribution a variable $\eta(i)$, which corresponds to the same probability $P(\eta< \eta(i)) = i/n+1$.By doing this, the new variables $\eta(i)$ will be marginally distributed according to standard Normal, N(0,1).
+-->
 
@@
 So, each category of a cyclic granularity corresponds to a list of numbers which is essentially few chosen quantiles of the multiple observations.
-
-- _Finding significant cyclic granularities or harmonies_
-
-\noindent These cyclic granularities are useful for exploring repetitive patterns in time series data that get lost in the linear representation of time.
-It is advantageous to consider only those cyclic granularities across which there is a significant repetitive pattern for the majority of customers or noteworthy in an electricity-behavior context.
-In that case, when the customers are grouped, we can expect to observe some interesting patterns across the categories of the cyclic granularities considered.
-[@Gupta2021-hakear] proposes a way to select significant cyclic granularities and harmonies which is used for this paper.
 

### 2021-10-31 main.Rmd
methodology done

@@
 
-- _Data preparation_
-
-@wang2020tsibble introduced the tidy "tsibble" data structure to support exploration and modeling of temporal data comprising of an index, optional key(s), and measured variables.
+- _Data pre-preprocessing_
+
+@wang2020tsibble introduced the tidy "tsibble" data structure to assist temporal data exploration and modeling.
+To start with, the measured variable for each key variable (observation in this context) is a time-indexed sequence of values for various measurement variables at each time point.
+This sequence, however, could be shown in several ways.
+A shuffle of the raw sequence may represent hourly consumption throughout a day, a week, or a year.
+Cyclic granularities like hour-of-day, hour-of-week can be expressed in terms of the index set in the "tsibble" data structure.
+But the data structure changes while transporting from linear to cyclic scale of time as multiple observations of the measured variable would correspond to each category of the cyclic granularities.
+In this paper, quantiles are chosen to characterize the distributions induced by the multiple observations for each category of the cyclic granularity.
+So, each category of a cyclic granularity corresponds to a list of numbers which is essentially few chosen quantiles of the multiple observations.
+
 <!-- An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
 A linear granularity is a mapping of the index set to subsets of the time domain.
-For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.--> For each key variable, the raw smart meter data is a sequence that is indexed by time and comprises values of several measurement variables at each time point.
-This sequence, though, could be depicted in a variety of ways.
-A shuffling of the raw sequence could reflect the distribution of hourly consumption over a single day, while another could indicate consumption over a week or a year.
-These temporal deconstructions of a time period into units such as hour-of-day, work-day/weekend are called cyclic temporal granularities.
-All cyclic granularities can be expressed in terms of the index set and could be augmented with the initial tsibble structure (index, key, measurements).
-It is worthwhile to note that the data structure changes while transporting from linear to cyclic scale of time as multiple observations of the measured variable would correspond to each category of the cyclic granularities.
-In this paper, quantiles are chosen to characterize the distributions for each category of the cyclic granularity.
-So, each category of a cyclic granularity corresponds to a list of numbers which is essentially few chosen quantiles of the multiple observations.
+For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.-->
 
@@
 
-In the analysis of energy smart meter data, K-Means or hierarchical clustering are often employed.
-These are simple and effective techniques that work well in a range of scenarios.
-For clustering, both employ a distance measure, and the distance measure chosen has a major influence on the structure of the clusters.
+With a way to obtain pairwise distances, any clustering algorithm can be employed that supports the given distance metric as input.
+A good comprehensive list of algorithms can be found in @Xu2015-ja based on traditional ways like partition, hierarchy or more recent approaches like distribution, density and others.
 We employ agglomerative hierarchical clustering in conjunction with Ward's criteria (XXX reference).
-The pair of clusters with minimum between-cluster distance are are sequentially merged in this using this agglomerative algorithms.
-A good comprehensive list of algorithms can be found in @.Xu2015-ja.
-We can possibly employ any clustering method that supports the given distance metric as input.
+Hierarchical cluster techniques fuse neighbouring points sequentially to form bigger clusters, beginning with a full pairwise distance matrix.
+The distance between clusters is described using a "linkage technique".
+This agglomerative approach successively merges the pair of clusters with the shortest between-cluster distance using Ward's linkage method.
+Hierarchical algorithms is one of the most widely used, can operate with data of any form, has reasonable scalability, and the number of clusters is not needed as a parameter.
 

### 2021-10-31 main.Rmd
gitignore update

@@
 This agglomerative approach successively merges the pair of clusters with the shortest between-cluster distance using Ward's linkage method.
-Hierarchical algorithms is one of the most widely used, can operate with data of any form, has reasonable scalability, and the number of clusters is not needed as a parameter.
+Hierarchical algorithms are one of the most widely used, can operate with data of any shape, has reasonable scalability, and the number of clusters is not needed as a parameter.
 

### 2021-10-31 main.Rmd
cleaning folder

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
+```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
+# knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
 # Make sure you have the latest version of rmarkdown and bookdown
@@
 #devtools::install_github("rstudio/bookdown")
-library(ggplot2)
-library(gravitas)
-library(gracsr)
-library(ggdendro)
-library(dplyr)
-library(readr)
-library(visdat)
-library(ggplot2)
-library(tidyverse)
-library(naniar)
-library(here)
-library(tsibble)
+options("knitr.graphics.auto_pdf" = TRUE,
+    tinytex.verbose = FALSE)
 library(knitr)
-library(patchwork)
-library(GGally)
-library(distributional)
-library(viridis)
+opts_chunk$set(
+ echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
+ fig.path = "figure/", fig.align = "center", fig.show = "hold",
+ cache = TRUE, cache.path = "cache/",
+ out.width = ifelse(is_html_output(), "100%", "\\textwidth")
+)
+knitr::opts_knit$set(root.dir = here::here())
 ```
@@
 ```{r external, include = FALSE, cache = FALSE}
-knitr::read_chunk(here('script/smart_meter.R'))
+knitr::read_chunk('script/smart_meter.R')
+```
+
+```{r load-lib}
+
 ```

### 2021-10-31 supplementary/prototype_selection/check_with3anchors.Rmd
cleaning folder

--- 2021-10-21 check_work_Di.Rmd
@@
 theme_characterisation <- function() {
-
-  theme_bw() + # seeting theme
-    theme(strip.text = element_text(size = 9,
-                                    margin = margin(b = 0, t = 0))) + # narrow facet space
+theme_bw() + 
+      theme(strip.text = element_text(size = 7, margin = margin(b = 0, t = 0))) +  + theme_bw() +
    theme(axis.title.y=element_blank(),
         axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + # no axis ticks 
-  theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
-    theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
-    theme(legend.position = "bottom")+
+        axis.ticks.y=element_blank()) +
+  theme(panel.spacing =unit(0, "lines")) + 
+    theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) + theme(legend.position = "bottom")+ theme(
+    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
   theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-  theme(axis.text.x = element_text(size=5))
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))+
+  theme(legend.position = "bottom")
 }
@@
 anchor3 <- 8629763
-anchor4 <- 11013154
-kopt <- 4 # number of clusters
+#anchor4 <- 11013154
+kopt <- 3 # number of clusters
 ```
@@
 
+data_pick_one <- nn_pick_function(anchor1)
+data_pick_two <- nn_pick_function(anchor2)
+data_pick_three <- nn_pick_function(anchor3)
+#data_pick_four <- nn_pick_function(anchor4)
+
 # c(11013154, 9345642, 8328008, 8636035,8627007,8454235)
 
-data_pick_one <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble
-data_pick_two <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble
-data_pick_three <- c(8221762, 8273636, 10359424, 8232822)%>% as_tibble
-data_pick_four <- c(10590714,8495194,8589936, 8454235) %>% as_tibble
-
 data_pick_cust <- bind_rows(
-data_pick_one, data_pick_two, data_pick_three,data_pick_four,
+data_pick_one, data_pick_two, data_pick_three,
+#data_pick_four,
 .id = "design")
@@
 
+```{r mds-plots}
+
+data_pick_char <- data_pick_cust %>% 
+  rename("customer_id" = "value") %>% 
+  mutate(customer_id = as.character(customer_id))
+
+mds_all <- read_rds(here("data/mds_all.rds"))  %>% 
+  left_join(data_pick_char)
+
+highlight_df <- mds_all %>% filter(customer_id %in% data_pick_char$customer_id)
+
+total_plot <- ggplot(mds_all,
+aes(x = mds1.total, y = mds2.total)) +
+  geom_point(alpha = 0.3) +
+  geom_point(data = highlight_df, 
+             aes(x = mds1.total, y = mds2.total, color = as.factor(design)))
+
+wpd_plot <- ggplot(mds_all,
+aes(x = mds1.wpd, y = mds2.wpd)) +
+  geom_point(alpha = 0.3) +
+  geom_point(data = highlight_df, 
+             aes(x = mds1.wpd, y = mds2.wpd, color = as.factor(design)))
+
+hod_plot <- ggplot(mds_all,
+aes(x = mds1.hod, y = mds2.hod)) +
+  geom_point(alpha = 0.3) +
+  geom_point(data = highlight_df, 
+             aes(x = mds1.hod, y = mds2.hod, color = as.factor(design)))
+
+moy_plot <- ggplot(mds_all,
+aes(x = mds1.moy, y = mds2.moy)) +
+  geom_point(alpha = 0.3) +
+  geom_point(data = highlight_df, 
+             aes(x = mds1.moy, y = mds2.moy, color = as.factor(design)))
+
+wkndwday_plot <- ggplot(mds_all,
+aes(x = mds1.wkndwday, y = mds2.wkndwday)) +
+  geom_point(alpha = 0.3) +
+  geom_point(data = highlight_df, 
+             aes(x =  mds1.wkndwday, 
+                 y =  mds2.wkndwday, 
+                 color = as.factor(design)))
+
+p <- (total_plot + wpd_plot)/(hod_plot + moy_plot + wkndwday_plot)
+
+p +
+  plot_layout(guides = "collect")
+
+```
+
 ```{r clustering}
@@
 total_distance <- filtered_distance %>% 
-mutate(total = wkndwday + moy + hod) 
+mutate(total = wkndwday/2 + moy/12 + hod/24) 
 #mutate(total = hod)
@@
                   ymax = `75%`,
-                  group=customer_id, fill = as.factor(group)),
+                  group=customer_id,
+                  fill = as.factor(group)),
               alpha = 0.5) +
   geom_line(aes(y = `50%`,
-                group=customer_id, color = as.factor(group)), size = 1) +
+                group=customer_id, 
+                color = as.factor(group)), size = 1) +
   facet_wrap(design~customer_id, 
@@
                   ymax = `75%`, 
-                  group=customer_id, fill = as.factor(group)), alpha = 0.5) +
+                  group=customer_id, 
+              fill = as.factor(group)), alpha = 0.5) +
   geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
@@
                   ymax = `75%`, 
-                  group=customer_id, fill = as.factor(group)), alpha = 0.5) +
-  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group))) +
+                  group=customer_id, 
+              fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
   facet_wrap(design~customer_id, 
@@
              labeller = "label_value",
-             ncol = 6)  +
+             ncol = 6)+
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
@@
 combined <- hod_group + moy_group + wkndwday_group & theme(legend.position = "bottom")
-combined + plot_layout(guides = "collect")
+combined + plot_layout(guides = "collect") + theme(legend.position = "bottom")
 
@@
                    scale = "globalminmax"
-) + theme_characterisation()+
+) + 
   ylab("wpd") + scale_fill_viridis_d(direction = -1) +
@@
   ylab("demand (in Kwh)") + 
-  theme_light()+ 
-  theme_characterisation() +
+  theme_light() +
  scale_fill_viridis_d(direction = -1) +
@@
              nrow = 4)   + 
-  theme_light()+ 
-  theme_characterisation() +
+  theme_light() +
  scale_fill_viridis_d(direction = -1) +
@@
 combined <- hod_group_wpd + moy_group_wpd + wkndwday_group_wpd & theme(legend.position = "bottom")
-combined + plot_layout(guides = "collect")
+combined + plot_layout(guides = "collect") + theme(legend.position = "bottom")
 

### 2021-10-31 supplementary/prototype_selection/check_work_nqt.Rmd
cleaning folder

--- 2021-10-18 check_work.Rmd
@@
 ```{r anchor}
-anchor1 <- 11156027
-anchor2 <- 10420683
-anchor3 <- 8629763
-anchor4 <- 11013154
+anchor1 <- 8302280 # 1Q
+anchor2 <- 8640717  # 2Q
+anchor3 <- 8402996 # 3Q
+anchor4 <- 8308172 # 4Q
 kopt <- 4 # number of clusters
@@
 
-wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_robust.rds"))
+wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_nqt.rds")) %>% broom::tidy()
 
-moy <- read_rds(here("data/dist_gran_moy_356cust.rds"))
+moy <- read_rds(here("data/dist_gran_moy_356cust_nqt.rds")) %>% broom::tidy()
 
-hod <- read_rds(here("data/dist_gran_hod_356cust_robust.rds"))
+hod <- read_rds(here("data/dist_gran_hod_356cust_nqt.rds")) %>% broom::tidy()
 
@@
 
-mds_all <- read_rds(here("data/mds_all.rds"))  %>% 
+mds_all <- read_rds(here("data/mds_all_nqt.rds"))  %>% 
   left_join(data_pick_char)
@@
   geom_point(data = highlight_df, 
-             aes(x = mds1.total, y = mds2.total, color = as.factor(design)), size = 2)
+             aes(x = mds1.total, y = mds2.total, color = as.factor(design)))
 
@@
   geom_point(data = highlight_df, 
-             aes(x = mds1.wpd, y = mds2.wpd, color = as.factor(design)), size = 2)
+             aes(x = mds1.wpd, y = mds2.wpd, color = as.factor(design)))
 
@@
   geom_point(data = highlight_df, 
-             aes(x = mds1.hod, y = mds2.hod, color = as.factor(design)), size = 2)
+             aes(x = mds1.hod, y = mds2.hod, color = as.factor(design)))
 
@@
   geom_point(data = highlight_df, 
-             aes(x = mds1.moy, y = mds2.moy, color = as.factor(design)), size = 2)
+             aes(x = mds1.moy, y = mds2.moy, color = as.factor(design)))
 
@@
                  y =  mds2.wkndwday, 
-                 color = as.factor(design)),
-             size = 2)
+                 color = as.factor(design)))
 
@@
 
-p* +
+p*theme_characterisation() +
   plot_layout(guides = "collect")

### 2021-10-31 old/MDS.R

@@ -1,51 +1,0 @@
-library(ggpubr)
-library(gravitas)
-library(tibble)
-clust_var <- "customer_id"
-ncust = 10
-niter = 2
-response = "general_supply_kwh"
-gran1 = "wknd_wday"
-gran2 = "hour_day"
-
-dis_cus <- smart_meter10 %>%
-  dplyr::distinct(.data[[clust_var]]) %>%
-  dplyr::sample_n(ncust)
-
-cust_data <- (1:nrow(dis_cus)) %>%
-  purrr::map_df(function(i){
-    make_cust_iter(smart_meter10, gran1, gran2, response, custid = dis_cus[[clust_var]][i], niter)
-  })
-
-data_whole <-  cust_data %>%
-  dplyr::group_by(cust_iter_id) %>%
-  dplyr::mutate(index = dplyr::row_number()) %>%
-  dplyr::ungroup() %>%
-  tsibble::as_tsibble(index = index, key = cust_iter_id)
-
-.data = data_whole
-distance_mat <- sim_dist(.data, gran1, gran2, response = "sim_data", key = "cust_iter_id",dist_ordered = TRUE,method = "max")
-mds <- distance_mat %>%
-dist() %>%
-cmdscale() %>%
-as_tibble()
-colnames(mds) <- c("Dim.1", "Dim.2")
-
-# Plot MDS
-
-ggscatter(mds, x = "Dim.1", y = "Dim.2",
-label = colnames(distance_mat),
-size = 1,
-repel = TRUE)
-
-# clustering
-
-.data <- smart_meter10
-gran1 = "wknd_wday"
-gran2 = "hour_day"
- key = "customer_id"
- response = "general_supply_kwh"
- distance_mat_all <- sim_dist(.data, gran1, gran2, response, key,dist_ordered = TRUE,method = "max")
- d <- stats::as.dist(distance_mat_all)
- hc <- stats::hclust(d)
- ggdendro::ggdendrogram(hc)

### 2021-10-31 old/algo2.R

@@ -1,117 +1,0 @@
-library(knitr)
-library(tidyverse)
-library(lubridate)
-library(lvplot)
-#library(ggridges)
-library(tsibble)
-library(gravitas)
-library(ggpubr)
-library(readr)
-library(kableExtra)
-library(distributional)
-library(ggplot2)
-library(sugrrants)
-library(here)
-library(ggplot2)
-library(patchwork)
-library(scales)
-library(GGally)
-library(viridis)
-
-# take customers who do not have gaps in their data
-elec_nogap <- read_rds("data/elec_nogap.rds")
-
-set.seed(12345)
-
-# just take 100 customers from them
-sm_100 <- elec_nogap %>%
-  as_tibble() %>% 
-  distinct(customer_id) %>%
-  slice_sample(n = 100)
-
-# elec <- elec_ts %>%
-#   as_tibble() %>% 
-#   distinct(customer_id) %>%
-#   slice_sample(n = 2)
-
-# take data for those customers only for 2013
-elec <- elec_nogap %>% 
-  filter(customer_id %in% sm_100$customer_id) %>% 
-  ungroup() %>% 
-  filter(year(reading_datetime)==2013)
-
-# consider harmonies which are generally significant for 
-
-harmonies <- read_rds("../paper-hakear/paper/data/harmonies.rds")
-
-harmonies_acro <- harmonies %>% 
-  mutate(facet_variable = case_when(
-    facet_variable == "hour_day" ~ "hod" ,
-    facet_variable == "day_month" ~ "dom" ,
-    facet_variable == "day_week" ~ "dow" ,
-    facet_variable == "week_month" ~ "wom" ,
-    facet_variable == "wknd_wday" ~ "wdwnd"
-  )) %>% 
-  mutate(x_variable = case_when(
-    x_variable == "hour_day" ~ "hod" ,
-    x_variable == "day_month" ~ "dom" ,
-    x_variable == "day_week" ~ "dow" ,
-    x_variable == "week_month" ~ "wom" ,
-    x_variable == "wknd_wday" ~ "wdwnd"
-  )) %>% mutate(comb = paste(facet_variable, x_variable, sep = "-")) %>% 
-  filter(comb %in% c("hod-wdwnd", "dom-hod", "wdwnd-hod", "hod-dow", "dow-hod")) %>% 
-  select(-comb)
-
-# take customers who do not have gaps in their data
-
-library(tictoc)
-#tic()
[…]

### 2021-10-31 old/approaches.R

@@ -1,249 +1,0 @@
-# harmony table considered for all
-
-#This script calculates raw mmpd for each simulation scenario
-##Read Simulation Table
-
-.libPaths(c("~/R/libs", .libPaths()))
-library(distributional)
-library(readr)
-library(drake)
-library(tidyverse)
-library(hakear)
-library(gravitas)
-library(parallel)
-library(here)
-library(lubridate)
-library(tsibble)
-library(patchwork)
-
-set.seed(9999)
-
-# parameter set
-nsamp = 100 # number of samples based on which thresholds are decided
-ncust = 8 # number of customers for trial
-nrep = 3 # number of error added usage for each customer
-kstep1 = 3 # number of clusters in step1/ only step
-kstep2 = 2 # number of clusters in step2 (this will actually depend on the number of observations in the cluster after step1)
-
-# change path while running it on HPC
-simtable<-read_csv(here('../gracsR/sim_table/sim_table.csv'))
-
-
-# <needed>
-harmonies<-read_csv(here('../gracsR/sim_table/harmonies.csv'))
-
-elec <- read_rds(here("data/elec_all-8.rds")) %>% 
-  dplyr::filter(date(reading_datetime) >= ymd("20190701"), date(reading_datetime) < ymd("20191231"), meter_id==1) %>% 
-  select(-meter_id) %>% 
-  rename("customer_id" = "household_id",
-         "general_supply_kwh" = "kwh") %>% 
-  as_tsibble(key = customer_id, index = reading_datetime)
-
-
-sm_cust_data <- elec
-
-sm_cust_data_list <- sm_cust_data %>% group_split(customer_id)
-
-
-### Extract flags from simulation scenario
-# Run wpd_threshold on all of them, each household can form a separate seed on cluster
-
-all_house_nrep_orig_list_5 <-  sm_cust_data_list
-
-all_house_harmony_wpd <- mclapply(all_house_nrep_orig_list_5,
-                                  function(x)
-                                    wpd = wpd_threshold(x,
-                                                  harmony_tbl = harmonies,
-                                                  response = general_supply_kwh, nsamp = nsamp)) %>%
-  bind_rows(.id = "customer_id")
-
-# Find harmonies that are not significant for all, whichever columns are integer in the following table drop them
-
-
-significant_tag <- function(x){
-  !all(is.na(str_extract(x, "\\*")))
-}
-
-
-m <- all_house_harmony_wpd %>%
-  mutate(harmony_name = paste(facet_variable, x_variable, sep = "-")) %>%
[…]

### 2021-10-31 old/change_index.R

@@ -1,174 +1,0 @@
-change_index_data <- function(ntimes_val = NULL,
-                         nx_val = NULL,
-                         nfacet_val = NULL,
-                         sim_function = sim_varx_normal){
-
-  data <- sim_panel(
-    nx = nx_val, nfacet =  nfacet_val,
-    ntimes = ntimes_val,
-    # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
-    sim_dist = sim_function(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
-  ) %>% unnest(data)
-
-
-  index_new <- map(seq_len(ntimes_val), function(i){
-    map((seq_len(nx_val*nfacet_val)), function(j)
-    {
-      value = i + (j-1)*ntimes_val
-    })
-  }) %>% unlist()
-
-  data_new = data %>%
-    ungroup() %>%
-    mutate(index_old = row_number(),
-           index_new = index_new)
-
-  y = data_new[match(index_new, data_new$index_old),]
-
-  y <- y %>%
-    mutate(time = row_number()) %>%
-    select(-c(index_old, index_new))
-
-  return(y)
-}
-##----making 10 iterations of 4 designs
-
-sample_seed <-  seq(10, 100, 10)
-
-data_null <- map(1:10, function(seed){
-  set.seed(sample_seed[seed])
-  change_index_data(5, 2, 3, sim_null_normal) %>%
-    compute_quantiles(quantile_prob_val)
-}) %>%
-  bind_rows(.id = "seed_id") %>%
-  mutate(data_type = "data_null")
-
-
-data_varf <- map(1:10, function(seed){
-  set.seed(sample_seed[seed])
-  change_index_data(5, 2, 3, sim_varf_normal)%>%
-    compute_quantiles(quantile_prob_val)
-}) %>%
-  bind_rows(.id = "seed_id") %>%
-  mutate(data_type = "data_varf")
-
-data_varx <- map(1:10, function(seed){
-  set.seed(sample_seed[seed])
-  change_index_data(5, 2, 3, sim_varx_normal)%>%
-    compute_quantiles(quantile_prob_val)
-}) %>%
-  bind_rows(.id = "seed_id") %>%
-  mutate(data_type = "data_varx")
-
-data_varall <- map(1:10, function(seed){
-  set.seed(sample_seed[seed])
-  change_index_data(5, 2, 3, sim_varall_normal)%>%
-    compute_quantiles(quantile_prob_val)
-}) %>%
[…]

### 2021-10-31 old/check.R

@@ -1,171 +1,0 @@
-library(conflicted)
-library(tidyverse)
-library(gracsr)
-library(tsibble)
-library(gravitas)
-library(dplyr)
-
-generate_design <- function(n = 300, #length of time series
-         mu11 = 0,
-         mu12 = 0,
-         mu21 = 0,
-         mu22 = 0 ,
-         mu23 = 0,
-         mu31 = 0,
-         mu32 = 0,
-         mu33 = 0,
-         mu34 = 0,
-         mu35 = 0)
-{
-  
-t <- seq(0, n-1, 1)
-
-g1 <- t %%2
-g2 <- t %%3
-g3 <- t %%5
-
-str_gran <- bind_cols(index = t, g1 = g1, g2 = g2, g3 = g3)
-
-# calculation of g1
-g1_table <- bind_cols(g1 = unique(g1), dist_mean = c(mu11, mu12))
-
-g1_tally <- str_gran %>% group_by(g1) %>%
-  count() %>%
-  left_join(g1_table, by = "g1")
-     
-g1_dist <- g1_tally %>%
-  mutate(g1_d = list(rep(g1, each = n)),
-                    g1_dist = list(rnorm(n, dist_mean, 1))) %>% 
-  ungroup() %>% 
-  select(g1_d, g1_dist) %>% 
-  unnest(cols = c(g1_d, g1_dist))
-
-g1_data <- str_gran %>% arrange(g1) %>% bind_cols(g1_dist = g1_dist$g1_dist) %>% arrange(index)
-
-# calculation of g2
-
-g2_table<- bind_cols(g2 = unique(g2),
-                     dist_mean = c(mu21, mu22, mu23))
-
-g2_tally <- str_gran %>%
-  group_by(g2) %>%
-  count() %>% 
-  left_join(g2_table, by = "g2")
-
-g2_dist <- g2_tally %>%
-  mutate(g2_d = list(rep(g2, each = n)),
-         g2_dist = list(rnorm(n, dist_mean, 1))) %>% 
-  ungroup() %>% 
-  select(g2_d, g2_dist) %>% 
-  unnest(cols = c(g2_d, g2_dist))
-
-g2_data <- str_gran %>% arrange(g2) %>% bind_cols(g2_dist = g2_dist$g2_dist) %>% arrange(index)
-
-# calculation of g3
-
-g3_table<- bind_cols(g3 = unique(g3), dist_mean = c(mu31, mu32, mu33, mu34, mu35))
-
[…]

### 2021-10-31 old/check_DGP.R

@@ -1,276 +1,0 @@
-
-set.seed(123)
-# DGP2
-# 
-generate_design_DGP2 <- function(n = 300, #length of time series
-                            mu11 = 0,
-                            mu12 = 0,
-                            mu21 = 0,
-                            mu22 = 0 ,
-                            mu23 = 0,
-                            mu31 = 0,
-                            mu32 = 0,
-                            mu33 = 0,
-                            mu34 = 0,
-                            mu35 = 0)
-{
-  
-  n <- n+500 # 500 burning observations
-  t <- seq(0, n-1, 1)
-  
-  g1 <- t %%2
-  g2 <- t %%3
-  g3 <- t %%5
-  
-  str_gran <- bind_cols(index = t, g1 = g1, g2 = g2, g3 = g3)
-  
-  # calculation of g1
-  g1_table <- bind_cols(g1 = unique(g1), dist_mean = c(mu11, mu12))
-  
-  g1_tally <- str_gran %>% group_by(g1) %>%
-    count() %>%
-    left_join(g1_table, by = "g1")
-  
-  g1_dist <- g1_tally %>%
-    mutate(g1_d = list(rep(g1, each = n)),
-           g1_dist = list(rnorm(n, dist_mean, 1))) %>% 
-    ungroup() %>% 
-    select(g1_d, g1_dist) %>% 
-    unnest(cols = c(g1_d, g1_dist))
-  
-  g1_data <- str_gran %>% arrange(g1) %>% bind_cols(g1_dist = g1_dist$g1_dist) %>% arrange(index)
-  
-  # calculation of g2
-  
-  g2_table<- bind_cols(g2 = unique(g2),
-                       dist_mean = c(mu21, mu22, mu23))
-  
-  g2_tally <- str_gran %>%
-    group_by(g2) %>%
-    count() %>% 
-    left_join(g2_table, by = "g2")
-  
-  g2_dist <- g2_tally %>%
-    mutate(g2_d = list(rep(g2, each = n)),
-           g2_dist = list(rnorm(n, dist_mean, 1))) %>% 
-    ungroup() %>% 
-    select(g2_d, g2_dist) %>% 
-    unnest(cols = c(g2_d, g2_dist))
-  
[…]

### 2021-10-31 old/job-algo2-500cust.sh

@@ -1,13 +1,0 @@
-#! /bin/bash
-#SBATCH --job-name=algo2cust500
-#SBATCH --output=trace_%a.out
-#SBATCH --error=error_%a.err
-#SBATCH --mem-per-cpu=4096
-#SBATCH --time=72:00:00
-#SBATCH --array=1-500
-#SBATCH --ntasks=1
-#SBATCH --cpus-per-task=1
-#SBATCH --export=NONE
-module load R/4.0.5
-Rscript job-algo2-500cust.R $SLURM_ARRAY_TASK_ID 
-

### 2021-10-31 old/job-algo2.sh

@@ -1,13 +1,0 @@
-#! /bin/bash
-#SBATCH --job-name=algo2
-#SBATCH --output=trace_%a.out
-#SBATCH --error=error_%a.err
-#SBATCH --mem-per-cpu=4096
-#SBATCH --time=72:00:00
-#SBATCH --array=1-100
-#SBATCH --ntasks=1
-#SBATCH --cpus-per-task=1
-#SBATCH --export=NONE
-module load R/4.0.5
-Rscript job-algo2.R $SLURM_ARRAY_TASK_ID 
-

### 2021-10-31 old/makedata01.R

@@ -1,263 +1,0 @@
-# harmony table considered for all
-
-#This script calculates raw mmpd for each simulation scenario
-##Read Simulation Table
-
-.libPaths(c("~/R/libs", .libPaths()))
-library(distributional)
-library(readr)
-library(drake)
-library(tidyverse)
-library(hakear)
-library(gravitas)
-library(parallel)
-library(here)
-
-set.seed(9999)
-
-# parameter set
-nsamp = 100 # number of samples based on which thresholds are decided
-ncust = 6 # number of customers for trial
-nrep = 3 # number of error added usage for each customer
-kstep1 = 3 # number of clusters in step1/ only step
-kstep2 = 2 # number of clusters in step2 (this will actually depend on the number of observations in the cluster after step1)
-
-
-# change path while running it on HPC
-simtable<-read_csv(here('sim_table/sim_table.csv'))
-
-# <needed>
-harmonies<-read_csv(here('sim_table/harmonies.csv'))
-
-sm_cust_data <- read_rds(here("data/sm_cust_data.rds"))
-
-sm_cust_data_list <- sm_cust_data %>% group_split(customer_id)
-
-# Add some error distribution to each of them before trying out the clustering process
-
-# function to generate some error to the usage for each customer
-# number of times to repeat
-# names of customer id when repeated
-
-add_error_unif <- function(x){
- y = runif(nrow(x), 0.01, 0.05)
- x$general_supply_kwh + y
-}
-
-# generate nrep times distribution for one household
-
-one_house_nrep <- function(x, nrep = 3)
-{
-  onerep = lapply(1:nrep, function(y){
-    bind_cols(customer_id  =  paste(x$customer_id, y, sep = "-"),
-              reading_datetime =  x$reading_datetime,
-              general_supply_kwh = rep(add_error_unif(x), 1))
-  })
-    onerep
-}
-
-
-# generate nrep times distribution for many households
-
-many_house_nrep <- lapply(sm_cust_data_list, function(x, nrep = 3){
-  one_house_nrep(x, nrep = nrep)
-}
-  ) %>% bind_rows
-
-
-all_house_nrep_orig <- bind_rows(sm_cust_data, many_house_nrep)
-
-
-saveRDS(all_house_nrep_orig, "data/all_house_nrep_orig.rds") # seed and dist not included yet
-
-
-
-all_house_nrep_orig_list <- all_house_nrep_orig %>%
[…]

### 2021-10-31 old/parallel_param.R

@@ -1,5 +1,0 @@
-sample_seed.all <- list(seq(10, 100, 10))
-
-for (sample_seed in sample_seed.all) {
-  rmarkdown::render('index.Rmd', params = list(sample_seed = sample_seed))
-}

### 2021-10-31 old/sim_dist.R

@@ -1,221 +1,0 @@
-#' Similarity distances based on probability distributions
-#'
-#' @param .data a tsibble.
-#' @param gran1,gran2 the granularity against which similarity is measured. For the granularity hour of the week, value is "hour_week".
-#' @param response response variable.
-#' @param key column in which subject lies
-#' @param method "sum" or "max" of distances
-#' @param ... Other arguments passed on to individual methods.
-#' @return matrix of similarity distances based on probability distributions
-#'
-#' @examples
-#' library(tsibble)
-#' library(dplyr)
-#' library(gravitas)
-#' library(purrr)
-#' .data <- smart_meter10
-#' gran1 <- "hour_day"
-#' gran2 <- "day_week"
-#' key = "customer_id"
-#' response = "general_supply_kwh"
-#' distance_mat <- sim_dist(.data, gran1, gran2, response, key,dist_ordered = TRUE,method = "max")
-#' @export sim_dist
-
-# distance matrix of all subjects for a harmony pair based on probability distribution
-
-sim_dist <- function(.data = NULL,
-                         gran1 = NULL,
-                         gran2 = NULL,
-                         response = NULL,
-                         key = NULL,
-                         method = "max",...
-)
-{
-  distinct_key <- .data %>%
-    tibble::as_tibble() %>%
-    dplyr::select(-tsibble::index(.data)) %>%
-    dplyr::select(key) %>%
-    dplyr::distinct()
-
-  distinct_key2 <- distinct_key %>%
-    dplyr::mutate(key2 = distinct_key[[key]]) %>%
-    expand.grid() %>%
-    dplyr::rename(key1 = !!key) %>%
-    dplyr::mutate(value = 1) %>%
-    tidyr::pivot_wider(names_from = key2,
-                       values_from = value) %>%
-    dplyr::mutate(key1 = as.character(key1))
-
-   #  filter(key1!=key2) %>%
-   #  mutate(conc_key = paste(key1,"-",key2)) %>%
-   #  mutate(conc_reverse = paste(key2,"-",key1)) %>%
-   # filter(conc_key = conc_reverse)
-
-  colNames <- colnames(distinct_key2)
-  colNames <-colNames[2:length(colNames)]
-
[…]

### 2021-10-31 old/simdata01.R

@@ -1,256 +1,0 @@
-# scripts to see how the time series for the four designs look like
-
-##----load-lib
-library(hakear)
-library(tidyr)
-library(distributional)
-library(ggplot2)
-library(patchwork)
-library(purrr)
-library(dplyr)
-library(kableExtra)
-library(tidyverse)
-library(gravitas)
-library(magrittr)
-
-## ----designs
-
-sim_null_normal <- function(nxj, nfacetj, mean, sd, w1 = 0, w2=0) {
-  rep(distributional::dist_normal(mu = mean, sigma = sd),
-      times = nxj * nfacetj
-  )
-}
-
-sim_varf_normal <- function(nx, nfacet, mean, sd, w1, w2) {
-  rep(dist_normal((mean + seq(0, nfacet - 1, by = 1) * w1),
-                  (sd + seq(0, nfacet - 1, by = 1) * w2)), each = nx)
-}
-
-sim_varx_normal <- function(nx, nfacet, mean, sd, w1, w2) {
-  rep(dist_normal((mean + seq(0, nx - 1, by = 1) * w1),
-                  (sd + seq(0, nx - 1, by = 1) * w2)), nfacet)
-}
-
-sim_varall_normal <- function(nx, nfacet, mean, sd, w1, w2) {
-  dist_normal((mean + seq(0, (nx *nfacet - 1), by = 1) * w1),
-              (sd + seq(0,  (nx *  nfacet - 1), by = 1) * w2))
-}
-
-
-## ----data
-
-
-sim_panel_varall <- sim_panel(
-  nx = nx_val, nfacet =  nfacet_val,
-  ntimes = ntimes_val,
-  # sim_dist = sim_varall_normal(2, 3, 5, 10, 5, -1.5)
-  sim_dist = sim_varall_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
-) %>% unnest(data)
-
-sim_panel_varx <- sim_panel(
-  nx = nx_val, nfacet =  nfacet_val,
-  ntimes = ntimes_val,
-  # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
-  sim_dist = sim_varx_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
-) %>% unnest(data)
-
-sim_panel_varf <- sim_panel(
-  nx = nx_val, nfacet =  nfacet_val,
-  ntimes = ntimes_val,
[…]

### 2021-10-31 old/simdata_cluster02.R

@@ -1,7 +1,0 @@
-# generate 4 series, two from design varx and another two from varf and see if clustering happens properly through the approaches
-
-
-# 2 time series from varx
-
-
-# 2 time series from varf

### 2021-10-31 script/smart_meter.R

@@ -1,13 +1,28 @@
 ## ---- load-lib
-library(tidyverse)
+
+#theme_set(theme_bw())
+library(ggplot2)
 library(lubridate)
-library(tsibble)
-library(fable)
-library(patchwork)
-library(here)
-theme_set(theme_bw())
-library(forcats)
+library(gravitas)
+library(gracsr)
+library(ggdendro)
+library(dplyr)
 library(readr)
+library(visdat)
+library(ggplot2)
+library(tidyverse)
+library(naniar)
+library(here)
+library(tsibble)
+library(knitr)
+library(patchwork)
+library(GGally)
+library(distributional)
+library(viridis)
+library(forcats)
+library(tidyr)
+library(purrr)
+library(stringr)
 
 ##----load-data
 load("data/ALL_DATA.Rdata")
@@ -882,9 +897,9 @@ append_files_plot <- function(folder_name, path){
   #   theme_light()
   
 }
-design1 <- append_files_plot(folder_name = "js-nqt", path = "/3gran_change_5D/")
-design2 <- append_files_plot(folder_name = "js-nqt", path = "/2gran_change_4D/")
-design3 <- append_files_plot(folder_name = "js-nqt", path = "/1gran_change_5D/")
+design1 <- append_files_plot(folder_name = "validation/js-nqt", path = "/3gran_change_5D/")
+design2 <- append_files_plot(folder_name = "validation/js-nqt", path = "/2gran_change_4D/")
+design3 <- append_files_plot(folder_name = "validation/js-nqt", path = "/1gran_change_5D/")
 
 # (design1 + design2 + design3)&theme(legend.position = "bottom") 
 # 
@@ -961,18 +976,18 @@ simtable <- expand.grid(mean_diff = mean_diff,
                         #time_series = time_series, 
                         nT = nT)
 
-data1 <- mds_data(read_rds("js-nqt/3gran_change_5D/data_validation_25.rds")) %>% mutate(diff = 1 , design= "a")
-data2 <-mds_data(read_rds("js-nqt/3gran_change_5D/data_validation_26.rds")) %>% mutate(diff = 2 , design= "a")
-data3 <-mds_data(read_rds("js-nqt/3gran_change_5D/data_validation_27.rds")) %>% mutate(diff = 5 , design= "a")
[…]

### 2021-10-31 smart_meter.R

@@ -1,25 +1,0 @@
-smart_meter_data_raw <- read_csv("data-raw/sm_raw.csv", n_max = 3e6)
-
-sm_cust <- smart_meter_data_raw %>% 
-  distinct(CUSTOMER_ID) %>%
-  .$CUSTOMER_ID %>% 
-  sample(size= 50)
-
-smart_meter_data <- smart_meter_data_raw %>%
-  dplyr::filter(CUSTOMER_ID %in% sm_cust) %>%
-  dplyr::rename_all(tolower) %>%
-  dplyr::arrange(customer_id, reading_datetime) %>%
-  dplyr::group_by(customer_id) %>%
-  dplyr::mutate(reading_datetime = case_when(
-    duplicated(reading_datetime) ~ reading_datetime + lubridate::hours(1),
-    TRUE ~ reading_datetime
-  ))
-
-smart_meter_data$customer_id <- as.character(smart_meter_data$customer_id)
-
-sm_cust50 <- smart_meter_data %>%
-  as_tsibble(index = reading_datetime, key = customer_id) %>%
-  ungroup() %>%
-  dplyr::select(-calendar_key)
-
-save(sm_cust50, file = "data/sm_cust50.rds", compress = "xz")

### 2021-11-01 main.Rmd
add readme, changed name from main.R from smart_meter.R
embellish abstract

@@
  With more and more time series data being collected at much finer temporal resolution, for a longer length of time, and for a larger number of individuals/entities, time series clustering research is getting a lot of traction.
-The sort of noisy, patchy, uneven, and asynchronous time series that is typical in many disciplines limits similarity searches among these lengthy time series.
-In this work, we suggest a method for overcoming these constraints by grouping time series based on probability distributions over cyclic temporal granularities.
-Cyclic granularities are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, and so on, and can be helpful for detecting repeating patterns.
-Looking at probability distributions across cyclic granularities results in an approach that is robust to missing or noisy data, aids in dimension reduction, and ensures small pockets of similar repeated behaviours.
-The proposed method was tested using a collection of residential electricity customers.
-The simulated and empirical evidence demonstrates that our method is capable of producing meaningful clusters.
+Long, noisy, patchy, uneven, and asynchronous time series are common in many fields, limiting similarity searches or lowering method efficiency when clustering is based on a distance metric.
+In this work, we suggest two approaches for obtaining similarity between time series based on probability distributions over cyclic temporal granularities for distance-based clustering approaches.
+Cyclic granularities like hour-of-the-day, work-day/weekend, month-of-the-year and so on, are useful for finding repeated patterns in the data.
+Looking at probability distributions across cyclic granularities serves two purposes: (a) "Probability distributions" characterise the inherent temporal data structure of these large unequal-length time series and are robust to missing or noisy data.
+(b) Using probability distributions over "cyclic granularities" ensures small pockets of similar "repeated" behaviours.
+Our method is capable of producing useful clusters, as demonstrated by testing on validation data designs and a sample of residential smart metee consumers.
 
@@
 ```{r external, include = FALSE, cache = FALSE}
-knitr::read_chunk('script/smart_meter.R')
+knitr::read_chunk('script/main.R')
 ```

### 2021-11-01 main.Rmd
disucssion what our work does

@@
 
-We propose different clustering methodology for grouping noisy, patchy time series data available at a fine temporal scale.
-Depending on the aim of clustering, they produce different clustering.
-The clustering is done based on probability distributions of the time series variable measured across several cyclic granularirties.
-There is issue with scaling it up to many customers as anomalies need to be removed before such classification would be useful.
+We offer two methods for calculating pairwise distances between time series based on probability distributions at various cyclic granularities.
+Depending on the goal of the clustering, these distance metrics, when fed into a hierarchical clustering algorithm using Ward's linkage, yield meaningful clusters.
+Probability distributions provide an intuitive method to characterise noisy, patchy, long, and unequal-length time series data.
+Distributions over cyclic granularities help to characterise the formed clusters in terms of their repeating behaviour over these cyclic granularities.
+Furthermore, unlike earlier efforts that group customers based on behavaior across only one cyclic granularity (such as hour-of-day), our method is more comprehensive in detecting clusters with repeated patterns at all relevant granularities.
+The JS-based approaches are faster than wpd-based approaches.
+
+There are few areas to extend this research.
+Firstly, things become far more complicated when we consider a larger data set with more uncertainty, as they do with any clustering problem.
+Characterising clusters with varied or outlying customers can result in a shape that does not represent the group.
+Furthermore, combining hetergeneous customers may result in similar-looking final clusters that are not effective for visually differentiating them.
+Hence, a method for appropriately scaling it up to many customers such that anomalies can be removed before implementing the clustering method, would be useful for bringing out meaningful, compact and separated groups.
+
+Secondly, we have assumed the time series to be stationary and hence the distributions are assumed to remaining constant for the observation period.
+In reality however, it might change.
+For the smart meter example, the distribution for a customer moving to a different house or changing electrical equipments can change drastically.
+These dynamic changes can not be picked up by our method.
+
+Thirdly, it is possible that for few customers, data for some categories from the list of significant granularities are missing.
+In our application, we have removed those customers and done the analysis but the metrics used should be able to incorporate those customers in the clustering by handling their missing categories.
+
+$wpd$ is computationally heavy even under parallel computation. Future work can make the computations more efficient so that it is easily scalable to a large number of customers.
 

### 2021-11-01 main.Rmd
Discussion: future work

@@
 
-There are few areas to extend this research.
-Firstly, things become far more complicated when we consider a larger data set with more uncertainty, as they do with any clustering problem.
+\noindent There are few areas to extend this research.
+First, larger data sets with more uncertainty complicate matters, as is true for any clustering task.
 Characterising clusters with varied or outlying customers can result in a shape that does not represent the group.
-Furthermore, combining hetergeneous customers may result in similar-looking final clusters that are not effective for visually differentiating them.
-Hence, a method for appropriately scaling it up to many customers such that anomalies can be removed before implementing the clustering method, would be useful for bringing out meaningful, compact and separated groups.
-
-Secondly, we have assumed the time series to be stationary and hence the distributions are assumed to remaining constant for the observation period.
-In reality however, it might change.
+Moreover, integrating heterogeneous consumers may result in visually identical end clusters, which are potentially not useful.
+Hence, a way of appropriately scaling it up to many customers such that anomalies are removed before clustering would be useful for bringing forth meaningful, compact and separated clusters.
+
+\noindent Secondly, we have assumed the time series to be stationary, and hence the distributions are assumed to remain constant for the observation period.
+In reality, however, it might change.
 For the smart meter example, the distribution for a customer moving to a different house or changing electrical equipments can change drastically.
-These dynamic changes can not be picked up by our method.
-
-Thirdly, it is possible that for few customers, data for some categories from the list of significant granularities are missing.
+Our current approach can not detect these dynamic changes.
+
+\noindent Thirdly, it is possible that for few customers, data for some categories from the list of considered significant granularities are missing.
 In our application, we have removed those customers and done the analysis but the metrics used should be able to incorporate those customers in the clustering by handling their missing categories.
 
-$wpd$ is computationally heavy even under parallel computation. Future work can make the computations more efficient so that it is easily scalable to a large number of customers.
+\noindent Finally, $wpd$ is computationally heavy even under parallel computation. Future work can make the computations more efficient so that they are easily scalable to a large number of customers.
 

### 2021-11-01 main.Rmd
Acknowledgment and supplementary information

@@
 
-\noindent There are few areas to extend this research.
+There are few areas to extend this research.
 First, larger data sets with more uncertainty complicate matters, as is true for any clustering task.
@@
 Hence, a way of appropriately scaling it up to many customers such that anomalies are removed before clustering would be useful for bringing forth meaningful, compact and separated clusters.
-
-\noindent Secondly, we have assumed the time series to be stationary, and hence the distributions are assumed to remain constant for the observation period.
+Secondly, we have assumed the time series to be stationary, and hence the distributions are assumed to remain constant for the observation period.
 In reality, however, it might change.
@@
 Our current approach can not detect these dynamic changes.
-
-\noindent Thirdly, it is possible that for few customers, data for some categories from the list of considered significant granularities are missing.
+Thirdly, it is possible that for few customers, data for some categories from the list of considered significant granularities are missing.
 In our application, we have removed those customers and done the analysis but the metrics used should be able to incorporate those customers in the clustering by handling their missing categories.
-
-\noindent Finally, $wpd$ is computationally heavy even under parallel computation. Future work can make the computations more efficient so that they are easily scalable to a large number of customers.
+Finally, $wpd$ is computationally heavy even under parallel computation.
+Future work can make the computations more efficient so that they are easily scalable to a large number of customers.
+
+# Acknowledgments {-}
+
+The authors thank the ARC Centre of Excellence for Mathematical and Statistical Frontiers [(ACEMS)](https://acems.org.au/home) for supporting this research.
+Sayani Gupta was partially funded by [Data61 CSIRO](https://data61.csiro.au/) during her PhD.
+The Monash eResearch Centre and eSolutions-Study Support Services supported this research in part through the resource usage of the MonARCH HPC Cluster.
+The Github repository, [github.com/Sayani07/paper-gracsr](https://github.com/Sayani07/paper-gracsr), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
+This article was created with R [@R-language], `knitr` [@knitr2015, @R-knitr] and `rmarkdown` [@rmarkdown2018, @R-rmarkdown].
+Graphics are produced with `ggplot2` [@Wickham2009pk] and GGally [@R-GGally].
+
+# Supplementary Materials
+
+**Data and scripts:** Data sets and R code to reproduce all figures in this article (main.R).
+
+\noindent
+**Supplementary paper:** Additional table, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd)
+
+\noindent
+**R-package:** To implement the ideas provided in this research, the open-source R package 'gracsr' is available on Github (https://github.com/Sayani07/gracsr).
+
+# Bibliography
+
+```{r write-bib, eval = FALSE}
+write_bib(c("rmarkdown", "bookdown", "rticles", "knitr", "ggplot2", "dplyr", "tsibble", "gravitas","Sugrrants", "rlang", "tidyr", "GGally","R-tsne", "ggpubr", "kableExtra", "magrittr", "readr", "lubridate", "renv", "archive", "hakear", "patchwork", "ggpubr", "purrr"), file = "rpkgs.bib")
+```
 

### 2021-11-01 main.Rmd
keywords update

@@
 keywords:
-- data visualization
-- statistical distributions
+- clustering
 - time granularities
-- calendar algebra
+- probability distributions
+- Jensen-Shannon distances
 - periodic data
-- grammar of graphics
+- smart meter
+- electricity consumption behaviour
 - R

### 2021-11-01 main.Rmd
shuffling intro structure

@@
 The time-series features, length of time-series, representation technique, and, of course, the purpose of clustering time-series all influence the suitable similarity measure or distance strategy to a meaningful level.
-The three primary methods to time series clustering (@liao2005clustering) are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based) (model-based).
-The efficacy of distance-based techniques is highly dependent on the distance measure utilised.
+The three primary methods to time series clustering (@liao2005clustering) are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based).
+The efficacy of distance-based techniques is highly dependent on the distance measure utilized.
 Defining an appropriate distance measure for the raw time series may be a difficult task since it must take into account noise, variable lengths of time series, asynchronous time series, different scales, and missing data.
-Commonly used Distance-based similarity measures as suggested by a review of time series clustering approaches (@Aghabozorgi2015-ct) are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping, Autocorrelation, Short time series distance, Piecewise regularisation, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances (@liao2007clustering).
-Euclidean distance and DTW are often used in time series clustering.
-When it comes to time-series clustering accuracy, the Euclidean distance beats DTW, but DTW has its own advantages (@corradini2001dynamic).Euclidean distance requires time series of equal length.
-while DTW can assist cluster time series of varying lengths (@ratanamahatana2005multimedia), only if there are no missing observations.
-
-We are motivated by the residential smart meter data.
-These long time series are asynchronous, with varying time lengths for different houses and missing observations and characterised by noisy and patchy behavior that can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
-Choosing probability distributions instead of raw data seems to be a natural way to analyze these types of data sets.
-Hence this paper proposes a distance metric based on Jensen-Shannon distances between probability distributions across significant cyclic granularities.
-Cyclic temporal granularities, which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
-Since cyclic granularities are considered instead of linear granularities, the resulting clusters are expected to group customers that have similar repetitive behaviors.
+Commonly used Distance-based similarity measures as suggested by a decade review of time series clustering approaches (@Aghabozorgi2015-ct) are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping, Autocorrelation, Short time series distance, Piecewise regularisation, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances (@liao2007clustering) but on a vector time series data.
+Among these alternatives, Euclidean distances have high performance but need the same length of data over the same period, resulting in information loss regardless of whether it is on raw data or a smaller collection of features.
+DTW works well with time series of different lengths (@corradini2001dynamic), but it is incapable of handling missing observations.
+Surprisingly, probability distributions, which may reflect the inherent temporal structure of a time series have not been considered in determining time series similarity.
+
+We consider the problem of clustering a large number of univariate time series of continuous values which are available at fine temporal scales, being motivated by the residential smart meter data.
+These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and missing observations and characterised by noisy and patchy behavior that can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
+Choosing probability distributions seem to be a natural way to analyze these types of data sets since they are robust to uneven length, missing data, or noise.
+Hence this paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across significant cyclic granularities.
+Cyclic temporal granularities (@Gupta2021-hd), which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
+The resulting clusters are expected to group customers that have similar repetitive behaviors across each of the interesting cyclic granularities.
 Below are some of the benefits of our method, which will be detailed in further depth in subsequent sections.
 
-- Some clustering algorithms become problematic with the very high dimensionality of the time series resulting from the frequency at which they are recorded and the length of time for which they are observed.
-We can efficiently cluster long length time series by reducing dimensionality by characterising through probability distributions;
-
-- By utilising Jensen-Shannon distances, we are evaluating the distance between two distributions rather than raw data, which is less susceptible to missing observations and outliers compared to other traditional distance measures;
-
-- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar forms across all key cyclic granularities.
-Because cyclic granularities are used rather than linear granularities, clustering would group consumers who exhibit similar repeating behaviour over many cyclic granularities where patterns are predicted to be important.
-
-- It is reasonable to define a time series based on its degree of trend and seasonality and to take these characteristics into account while clustering it.
-The change in data structure by considering probability distributions across cyclic granularities ensures there is no trend and seasonal fluctuations are handled separately.
-Thus there is no need to de-trend or de-seasonalize the data prior to performing the clustering method.
+<!-- since similarity between time series is based on the distances between probability distributions for each of the time granularites across which patterns are interesting. -->
+
+<!-- - Some clustering algorithms become problematic with the very high dimensionality of the time series resulting from the frequency at which they are recorded and the length of time for which they are observed.
+We can efficiently cluster long length time series by reducing dimensionality by characterising through probability distributions; -->
+
+- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern).
+
+<!-- (for example, some customers observed only for holidays, others observed over non holidays or year of observation is different which can lead to technological advancement.) -->
+
+- Jensen-Shannon distances evaluate the distance between two distributions rather than raw data, which is less sensitive to missing observations and outliers than other conventional distance methods;
+
+- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar distributions across all interesting cyclic granularities;
+
+<!-- Because cyclic granularities are used rather than linear granularities, clustering would group consumers who exhibit similar repeating behavior over many cyclic granularities where patterns are predicted to be important.
+-->
+
+- It is reasonable to define a time series based on its degree of trend and seasonality, and to take these characteristics into account while clustering it.
+The modification of the data structure by taking into account probability distributions across cyclic granularities assures that there is no trend and that seasonal variations are handled independently.
+As a result, there is no need to de-trend or de-seasonalize the data before applying the clustering method.
 For similar reasons, there is no need to exclude holiday or weekend routines.
@@
 
-<!-- significance of load clustering -->
 Large spatio-temporal data sets, both from open and administrative sources, offer up a world of possibilities for research.
@@
 
-_Related work_
+<!-- _Related work_ -->
 
@@
 @Melnykov2013-sp disucusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
-
-\noindent Given the limitations of the similarity measures in dealing with large volumes of this complicated time series data, we present a similarity measure based on probability distributions that seems to be a more organic option for coping with time series data with aforementioned characteristics.
-The remainder of the paper is organized as follows: Section&nbsp;\ref{sec:methodology} provides the clustering methodology introducing the features and distance metrics.
-Section&nbsp;\ref{sec:validation} shows data designs to validate our methods and draw comparisons against several methods.Section&nbsp;\ref{sec:application} discusses the application of the method to a subset of the real data.
+Given none of the methods use probability distributions across cyclic granularity to explore heterogeneity in different repetitive behaviors, we present a way to do that by presenting similarity through probability distributions across cyclic granularities.
+
+The remainder of the paper is organized as follows: Section&nbsp;\ref{sec:methodology} provides the clustering methodology.
+Section&nbsp;\ref{sec:validation} shows data designs to validate our methods.
+Section&nbsp;\ref{sec:application} discusses the application of the method to a subset of the real data.
 Finally, we summarize our results and discuss possible future directions in Section&nbsp;\ref{sec:discussion}.
@@
 But the data structure changes while transporting from linear to cyclic scale of time as multiple observations of the measured variable would correspond to each category of the cyclic granularities.
-In this paper, quantiles are chosen to characterize the distributions induced by the multiple observations for each category of the cyclic granularity.
+Directly computing Jensen-Shannon distances between probability distributions can be very costly.
+Hence, in this paper, quantiles are chosen to characterize the probability distributions induced by the multiple observations for each category of the cyclic granularity.
 So, each category of a cyclic granularity corresponds to a list of numbers which is essentially few chosen quantiles of the multiple observations.

### 2021-11-01 main.Rmd
stop working on introduction

@@
 
-We consider the problem of clustering a large number of univariate time series of continuous values which are available at fine temporal scales, being motivated by the residential smart meter data.
+\noindent We consider the problem of clustering a large number of univariate time series of continuous values which are available at fine temporal scales, being motivated by the residential smart meter data.
 These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and missing observations and characterised by noisy and patchy behavior that can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
-Choosing probability distributions seem to be a natural way to analyze these types of data sets since they are robust to uneven length, missing data, or noise.
-Hence this paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across significant cyclic granularities.
-Cyclic temporal granularities (@Gupta2021-hd), which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
-The resulting clusters are expected to group customers that have similar repetitive behaviors across each of the interesting cyclic granularities.
-Below are some of the benefits of our method, which will be detailed in further depth in subsequent sections.
-
-<!-- since similarity between time series is based on the distances between probability distributions for each of the time granularites across which patterns are interesting. -->
-
-<!-- - Some clustering algorithms become problematic with the very high dimensionality of the time series resulting from the frequency at which they are recorded and the length of time for which they are observed.
-We can efficiently cluster long length time series by reducing dimensionality by characterising through probability distributions; -->
-
-- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern).
-
-<!-- (for example, some customers observed only for holidays, others observed over non holidays or year of observation is different which can lead to technological advancement.) -->
-
-- Jensen-Shannon distances evaluate the distance between two distributions rather than raw data, which is less sensitive to missing observations and outliers than other conventional distance methods;
-
-- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar distributions across all interesting cyclic granularities;
-
-<!-- Because cyclic granularities are used rather than linear granularities, clustering would group consumers who exhibit similar repeating behavior over many cyclic granularities where patterns are predicted to be important.
--->
-
-- It is reasonable to define a time series based on its degree of trend and seasonality, and to take these characteristics into account while clustering it.
-The modification of the data structure by taking into account probability distributions across cyclic granularities assures that there is no trend and that seasonal variations are handled independently.
-As a result, there is no need to de-trend or de-seasonalize the data before applying the clustering method.
-For similar reasons, there is no need to exclude holiday or weekend routines.
-
-<!-- . Often the objective is finding similarities in time (Fourier transforms, Piecewise aggregate approximation), in  shape (Dynamic time Warping (DTW)), in structure, based on global features. -->
-
-<!-- However, the process of determining a similarity measure is challenging since time-series data are naturally noisy and include outliers and missing observations, and the length of time-series changes and can be asynchronous.
--->
-
-<!-- Time-series clustering is the most-used approach as an exploratory technique  for discovery of interesting patterns in time-series datasets.
-However, the process of finding similarity measure is complicated, because time-series data are naturally noisy and include outliers and shifts [18], at the other hand the length of time-series varies and the distance among them needs to be calculated.
-These common issues have made the -->
-<!-- similarity measure a major challenge for data miners. -->
-
-_Background and motivation_
-
-<!-- Description of data available -->
-
-Large spatio-temporal data sets, both from open and administrative sources, offer up a world of possibilities for research.
-<!--Beyea (2010) has pointed out, there has been little exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data.--> One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee)<!--and Data61 CSIRO-->.
-The project provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014.
-<!---The massive amount of data generated in such projects could be overwhelming for analysis.Raw data of these asynchronous time series can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
-Electricity utilities can utilize the smart meter usage patterns to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.
---><!-- Sources of variation in large data sets --><!-- The enormous quantity of data provides for greater individual level clarity and analysis.
-However, -->
-<!-- due to the growing variety of consumers,  -->.
 Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.
@@
 
-<!-- _Related work_ -->
-
-A multitude of papers have emerged around smart meter time series clustering for deepening our knowledge of consumption patterns.
+\noindent A multitude of papers have emerged around smart meter time series clustering for deepening our knowledge of consumption patterns.
+@Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
+None of the $34$ articles chosen for their emphasis use Australian smart meter data.
+The most often used algorithm is K-Means.
+Using K-Means without considering time series structure or correlation results in inefficient clusters.
+Principal Component Analysis (PCA) or Self-Organizing Maps (SOM) eliminate correlation patterns and decrease feature space, but lose interpretability.
+To reduce dimensionality, several studies use principal component analysis or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).
+Other algorithms utilised in the literature include k-means variants, hierarchical approaches, and greedy k-medoids.
+Time series data, such as smart metre data, are not well-suited to any of the techniques mentioned in @Tureczek2017-pb.
+Only one study [@ozawa2016determining] identified time series characteristics using Fourier transformation, which converts data from time to frequency and then uses K-Means to cluster by greatest frequency .
+@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential datasets and addresses model-based clustering  by transforming the series into other other objects such as structure or set of parameters which can be more easily characterised and clustered.
+[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
+@Melnykov2013-sp disucusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
+None of the methods focus on exploring heterogeneity in repetitive behaviors across different temporal dependencies, we present a way to do that in this paper.
+
+Choosing probability distributions seem to be a natural way to analyze these types of data sets since they are robust to uneven length, missing data, or noise.
+Hence this paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across significant cyclic granularities.
+Cyclic temporal granularities, which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
+The resulting clusters are expected to group customers that have similar repetitive behaviors across each of the interesting cyclic granularities.
+Below are some of the benefits of our method, which will be detailed in further depth in subsequent sections.
+
+- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern).
+
+<!-- (for example, some customers observed only for holidays, others observed over non holidays or year of observation is different which can lead to technological advancement.) -->
+
+- Jensen-Shannon distances evaluate the distance between two distributions rather than raw data, which is less sensitive to missing observations and outliers than other conventional distance methods;
+
+- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar distributions across all interesting cyclic granularities;
+
+- It is reasonable to define a time series based on its degree of trend and seasonality, and to take these characteristics into account while clustering it.
+The modification of the data structure by taking into account probability distributions across cyclic granularities assures that there is no trend and that seasonal variations are handled independently.
+As a result, there is no need to de-trend or de-seasonalize the data before applying the clustering method.
+For similar reasons, there is no need to exclude holiday or weekend routines.
+
+<!-- _Background and motivation_ -->
+
+<!-- Large spatio-temporal data sets, both from open and administrative sources, offer up a world of possibilities for research.
+One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) The project provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014.
+Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.
+Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
+The behavioural patterns differ amongst customers due to many temporal dependencies.
+Some households use a dryer, while others dry their clothes on a line.
+Their weekly profile may reflect this.
+They may vary monthly, with some customers using more air conditioners or heaters than others, while having equivalent electrical equipment and weather circumstances.
+Some customers are night owls, while others are morning larks.
+Day-off energy use varies depending on whether customers stay home or go outside.
+Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behaviour a fascinating one.
+This challenge is worsened when all we know about our consumers is their energy use history (@Ushakova2020-rl).
+To safeguard the customers' privacy, it is probable that such information is not accessible.
+Also, energy suppliers may not always update client information, such as property features, in a timely manner.
+Thus, there is a growing need to have research that examines how much energy usage heterogeneity can be found in smart meter data and what are some of the most common power consumption patterns, rather than explaining why consumption differs.
+-->
+
+<!-- <!-- _Related work_ -->
+
+<!-- A multitude of papers have emerged around smart meter time series clustering for deepening our knowledge of consumption patterns.
 @Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
@@
 Given none of the methods use probability distributions across cyclic granularity to explore heterogeneity in different repetitive behaviors, we present a way to do that by presenting similarity through probability distributions across cyclic granularities.
+-->
 
@@
 The Github repository, [github.com/Sayani07/paper-gracsr](https://github.com/Sayani07/paper-gracsr), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-This article was created with R [@R-language], `knitr` [@knitr2015, @R-knitr] and `rmarkdown` [@rmarkdown2018, @R-rmarkdown].
+This article was created with R [@R-language], `knitr` [@knitr, @R-knitr] and `rmarkdown` [@rmarkdown, @R-rmarkdown].
 Graphics are produced with `ggplot2` [@Wickham2009pk] and GGally [@R-GGally].

### 2021-11-01 main.Rmd
OK now don't touch it

@@
 These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and missing observations and characterised by noisy and patchy behavior that can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
+Choosing probability distributions seem to be a natural way to analyze these types of data sets since they are robust to uneven length, missing data, or noise.
+Hence this paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across significant cyclic granularities.
+Cyclic temporal granularities, which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
+The resulting clusters are expected to group customers that have similar repetitive behaviors across each of the interesting cyclic granularities.
+Below are some of the benefits of our method, which will be detailed in further depth in subsequent sections.
+
+- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern).
+
+<!-- (for example, some customers observed only for holidays, others observed over non holidays or year of observation is different which can lead to technological advancement.) -->
+
+- Jensen-Shannon distances evaluate the distance between two distributions rather than raw data, which is less sensitive to missing observations and outliers than other conventional distance methods;
+
+- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar distributions across all interesting cyclic granularities;
+
+- It is reasonable to define a time series based on its degree of trend and seasonality, and to take these characteristics into account while clustering it.
+The modification of the data structure by taking into account probability distributions across cyclic granularities assures that there is no trend and that seasonal variations are handled independently.
+As a result, there is no need to de-trend or de-seasonalize the data before applying the clustering method.
+For similar reasons, there is no need to exclude holiday or weekend routines.
+
+_Background and motivation_
+
+Large spatio-temporal data sets, both from open and administrative sources, offer up a world of possibilities for research.
+One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) The project provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014.
 Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.
 Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
-The behavioural patterns differ amongst customers due to many temporal dependencies.
+The behavioral patterns differ amongst customers due to many temporal dependencies.
 Some households use a dryer, while others dry their clothes on a line.
@@
 
-\noindent A multitude of papers have emerged around smart meter time series clustering for deepening our knowledge of consumption patterns.
-@Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
-None of the $34$ articles chosen for their emphasis use Australian smart meter data.
-The most often used algorithm is K-Means.
-Using K-Means without considering time series structure or correlation results in inefficient clusters.
-Principal Component Analysis (PCA) or Self-Organizing Maps (SOM) eliminate correlation patterns and decrease feature space, but lose interpretability.
-To reduce dimensionality, several studies use principal component analysis or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).
-Other algorithms utilised in the literature include k-means variants, hierarchical approaches, and greedy k-medoids.
-Time series data, such as smart metre data, are not well-suited to any of the techniques mentioned in @Tureczek2017-pb.
-Only one study [@ozawa2016determining] identified time series characteristics using Fourier transformation, which converts data from time to frequency and then uses K-Means to cluster by greatest frequency .
-@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential datasets and addresses model-based clustering  by transforming the series into other other objects such as structure or set of parameters which can be more easily characterised and clustered.
-[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
-@Melnykov2013-sp disucusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
-None of the methods focus on exploring heterogeneity in repetitive behaviors across different temporal dependencies, we present a way to do that in this paper.
-
-Choosing probability distributions seem to be a natural way to analyze these types of data sets since they are robust to uneven length, missing data, or noise.
-Hence this paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across significant cyclic granularities.
-Cyclic temporal granularities, which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
-The resulting clusters are expected to group customers that have similar repetitive behaviors across each of the interesting cyclic granularities.
-Below are some of the benefits of our method, which will be detailed in further depth in subsequent sections.
-
-- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern).
-
-<!-- (for example, some customers observed only for holidays, others observed over non holidays or year of observation is different which can lead to technological advancement.) -->
-
-- Jensen-Shannon distances evaluate the distance between two distributions rather than raw data, which is less sensitive to missing observations and outliers than other conventional distance methods;
-
-- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar distributions across all interesting cyclic granularities;
-
-- It is reasonable to define a time series based on its degree of trend and seasonality, and to take these characteristics into account while clustering it.
-The modification of the data structure by taking into account probability distributions across cyclic granularities assures that there is no trend and that seasonal variations are handled independently.
-As a result, there is no need to de-trend or de-seasonalize the data before applying the clustering method.
-For similar reasons, there is no need to exclude holiday or weekend routines.
-
-<!-- _Background and motivation_ -->
-
-<!-- Large spatio-temporal data sets, both from open and administrative sources, offer up a world of possibilities for research.
-One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) The project provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014.
-Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.
-Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
-The behavioural patterns differ amongst customers due to many temporal dependencies.
-Some households use a dryer, while others dry their clothes on a line.
-Their weekly profile may reflect this.
-They may vary monthly, with some customers using more air conditioners or heaters than others, while having equivalent electrical equipment and weather circumstances.
-Some customers are night owls, while others are morning larks.
-Day-off energy use varies depending on whether customers stay home or go outside.
-Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behaviour a fascinating one.
-This challenge is worsened when all we know about our consumers is their energy use history (@Ushakova2020-rl).
-To safeguard the customers' privacy, it is probable that such information is not accessible.
-Also, energy suppliers may not always update client information, such as property features, in a timely manner.
-Thus, there is a growing need to have research that examines how much energy usage heterogeneity can be found in smart meter data and what are some of the most common power consumption patterns, rather than explaining why consumption differs.
--->
-
-<!-- <!-- _Related work_ -->
-
-<!-- A multitude of papers have emerged around smart meter time series clustering for deepening our knowledge of consumption patterns.
+_Related work_
+
+A multitude of papers have emerged around smart meter time series clustering for deepening our knowledge of consumption patterns.
 @Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
@@
 @Melnykov2013-sp disucusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
-Given none of the methods use probability distributions across cyclic granularity to explore heterogeneity in different repetitive behaviors, we present a way to do that by presenting similarity through probability distributions across cyclic granularities.
+None of the methods focus on exploring heterogeneity in repetitive behaviors across different temporal dependencies, we present a way to do that in this paper.
+
+<!-- Given none of the methods use probability distributions across cyclic granularity to explore heterogeneity in different repetitive behaviors, we present a way to do that by presenting similarity through probability distributions across cyclic granularities.
 -->

### 2021-11-01 main.Rmd
notation of level change

@@
 We spiked several attributes in the data to see where one method works better than the other and where they might give us the same outcome or the effect of missing data <!---and trends--> on the proposed methods.
-Three circular granularities $g1$, $g2$ and $g3$ are considered with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $l_{g_1}=2$, $l_{g_2}=3$ and $l_{g_3}=5$.
+Three circular granularities $g1$, $g2$ and $g3$ are considered with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $n_{g_1}=2$, $n_{g_2}=3$ and $l_{g_3}=5$.
 These categories could be integers or some more meaningful labels.
@@
 Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$.
-In contrast to the previous situation, when we could study distributions across $l_{g_1} + l_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $l_{g_1}*l_{g_2}=6$ combination of categories.
+In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories.
 Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs.

### 2021-11-01 main.Rmd
validation section done

@@
 
-To validate the clustering approaches, we create data designs that replicate prototype behaviors that might be seen in electricity data contexts.
-We spiked several attributes in the data to see where one method works better than the other and where they might give us the same outcome or the effect of missing data <!---and trends--> on the proposed methods.
-Three circular granularities $g1$, $g2$ and $g3$ are considered with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $n_{g_1}=2$, $n_{g_2}=3$ and $l_{g_3}=5$.
+To validate the clustering approaches, we spiked several attributes in the data to see where one method works better than the other and where they might give us the same outcome <!--or the effect of missing data <!---and trends--> on the proposed methods.
+Three circular granularities $g1$, $g2$ and $g3$ are considered with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $n_{g_1}=2$, $n_{g_2}=3$ and $n_{g_3}=5$.
 These categories could be integers or some more meaningful labels.
-For example, the granularity "day-of-week" could be either represented by ${0, 1, 2, \dots, 6}$ or ${Mon, Tue, \dots, Sun}$.
+For example, the granularity "day-of-week" could be either represented by $\{0, 1, 2, \dots, 6\}$ or $\{Mon, Tue, \dots, Sun\}$.
 Here categories of $g1$, $g2$ and $g3$ are represented by  $\{0, 1\}$, $\{0, 1, 2\}$ and $\{0, 1, 2, 3, 4\}$ respectively.
 A continuous measured variable $v$ of length $T$ indexed by $\{0, 1, \dots T-1\}$ is simulated such that it follows the structure across $g1$, $g2$ and $g3$.
-We constructed independent replications of all data designs $R = 25, 250, 500$ to investigate if our proposed clustering method can discover distinct designs in small, medium, and big number of series.
-All designs employ $T=300, 1000, 5000$ sample sizes to evaluate small, medium, and large sized series.
+We constructed independent replications of all data designs $R = \{25, 250, 500\}$ to investigate if our proposed clustering method can discover distinct designs in small, medium, and big number of series.
+All designs employ $T=\{300, 1000, 5000\}$ sample sizes to evaluate small, medium, and large sized series.
 Variations in method performance may be due to different jumps between categories.
-So a mean difference of $diff = 1, 2, 5$ is examined.
-The approaches' performance varies with the number of significant granularities.
-So all, few, and one major granularity scenarios are considered.
-Please see the Supplementary section (https://github.com/Sayani07/paper-gracsR) for the code and findings.
+So a mean difference of $\mu = \{1, 2, 5\}$ is examined.
+The performance of the approaches varies with the number of granularities which has interesting patterns across its categories.
+So scenarios with all, few, and one significant granularities are considered.
 
@@
 Each category or combination of categories from $g1$, $g2$ and $g3$ are assumed to come from the same distribution, a subset of them from the same distribution, a subset of them from separate distributions, or all from different distributions, resulting in various data designs.
-As the methods ignore the linear progression of time, there is little value in adding time dependency in the data generating process.
+As the methods ignore the linear progression of time, there is little value in adding time dependency to the data generating process.
 The data type is set to be "continuous," and the setup is assumed to be Gaussian.
 When the distribution of a granularity is "fixed", it means distributions across categories do not vary and are considered to be from N (0,1).
-The mean of different categories are altered in the "varying" designs, leading to varying distributions across categories.
+$\mu$ alters in the "varying" designs, leading to varying distributions across categories.
 
@@
 \noindent Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction.
-Meaning that at least one of the designs to be categorized will have significantly different patterns for each granularity.
-In Table \ref{tab:tab-dist-design} (top), we explore various distributions across categories (as shown in Table Table \ref{tab:tab-dist-design} (bottom).
-Figure \ref{fig:plot-3gran-new} shows the simulated variable's linear and cyclic representations for each of these five designs.
-The structural difference in the time series variable is impossible to discern from the linear view.
-The shift in structure may be seen clearly in the distribution of cyclic granularities (Figure \ref{fig:plot-3gran-new} (right)).
+This means that at least one of the designs to be categorized will have significantly different patterns for each granularity.
+In Table \ref{tab:tab-dist-design} (top), we explore various distributions across categories (as shown in Table \ref{tab:tab-dist-design} (bottom).
+Figure \ref{fig:plot-3gran-new} shows the simulated variable's linear (left) and cyclic (right) representations for each of these five designs.
+The structural difference in the time series variable is impossible to discern from the linear view, with all of them looking very similar.
+The shift in structure may be seen clearly in the distribution of cyclic granularities.
 The following scenarios use solely graphical displays across cyclic granularities to highlight distributional differences in categories.
@@
 \noindent This is the case where one granularity will remain the same across all designs.
-We consider the case where the distribution of $v$ would vary across levels of $g2$ for all designs, across levels of $g3$ for few designs and $g1$ does not change across designs.
-The proposed design is shown in Figure \ref{fig:gran2and1-clubbed}(left).
-
-_(c) One signifiant granularity_
+We consider the case where the distribution of $v$ varies across $g2$ levels for all designs, across $g3$ levels for a few designs, and $g1$ does not vary across designs.
+The proposed design is shown in Figure \ref{fig:gran2and1-clubbed}(b).
+
+_Scenario (c): One signifiant granularity_
 
@@
 
-```{r generate-design-3change}
-
-```
-
-```{r plot-3gran-new, out.width="100%", fig.cap="The linear (left) and cyclic (right) representation of the measured variable is shown.
+```{r generate-D3change}
+
+```
+
+```{r plot-3gran-new, out.width="100%", fig.cap="The linear (left) and cyclic (right) representation of the simulated variable is shown.
+Each row represents a design in Scenario (a).
 In this scenario, all of $g1$, $g2$ and $g3$ changes across at least one design.
-Also, it is not possible to comprehend these patterns across cyclic granularities or group similar series just by looking at the linear plots."}
+Also, it is not possible to comprehend these differences in patterns just by looking at or considering the linear representation."}
 
@@
 
-```{r gran2and1-clubbed, fig.cap=" For the left scenario $g1$, $g2$ would change across atleast one design but $g3$ change remains same across all design.
-For the right one, only $g3$ changes across different designs."}
+```{r gran2and1-clubbed, fig.cap=" Plots (b) and (c) correspond to Design (b) and (c) respectively.
+In (b) $g2$, $g3$ changes across atleast one design but $g1$ remains constant.
+Only $g3$ changes across different designs in (c)."}
 #gran2_change  + gran1_change + plot_layout(widths = c(1,1))
@@
 Consider a group that has a different weekday and weekend behavior in the summer but not in the winter.
-This type of combined behaviour across granularities  can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (wknd-wday and month-of-year in this example).
+This type of combined behaviour across granularities  can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (Weekend/Weekday and month-of-year in this example).
 As a result, in this scenario, we analyse a combination of categories generated from different distributions.
@@
 Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs.
-Design-1 exhibits no change in distributions across $g1$ or $g2$, whereas Design-2 and Design-3 alter across only $g1$ and $g2$, respectively.
-Design-4 varies across both $g1$ and $g2$ categories.
-Design-3 and Design-4 appear similar based on their relative difference across consecutive categories, however Design-4 also changes across facets, unlike Design-3, which has all facets look the same.
-
-```{r interaction-gran, fig.cap = "Design-1 (a) has no change in distributions across different categories of $g1$ or $g2$, while Design-2 (b) and Design-3 (c) change across only $g1$ and $g2$ respectively.
-Design-4 (d) changes across categories of both $g1$ and $g2$."}
+Design $D1$ exhibits no change in distributions across $g1$ or $g2$, whereas  Designs $D2$ and $D3$ alter across only $g1$ and $g2$, respectively.
+D4 varies across both $g1$ and $g2$ categories.
+D3 and D4 appear similar based on their relative differences across consecutive categories, but $D4$ also changes across facets, unlike $D3$, which has all facets look the same.
+
+```{r interaction-gran, fig.cap = "Distribution of the simulated variable across $g1$ conditional on $g2$ is shown through boxplots for 4 designs.
+D1 has no change in distributions across different categories of $g1$ or $g2$, while D2 and D3 change across only $g1$ and $g2$ respectively.
+D4 changes across categories of both $g1$ and $g2$."}
 
@@
 We observe that in all scenarios and for different mean differences, cluster is separated.
-However, the separation increases with increase in mean differences across scenarios.
-This is intuitive because with increasing difference between categories, it gets easier for the methods to correctly distinguish the designs.
-Results corresponding to Scenario (a) is shown here.
+However, the separation increases with an increase in mean differences across scenarios.
+This is intuitive because, as the difference between categories increases, it gets easier for the methods to correctly distinguish the designs.
 
@@
 For all simulation settings and scenarios, values are represented as a sequence of lines connected across each axis.
-One line in the figure, for example, shows the inter-cluster distances for one simulation scenario.
+One line in the figure, for example, shows the inter-cluster distances for one simulation setting.
 The lines are not coloured by group since the purpose is to highlight the contribution of the factors to categorization rather than class separation.
@@
  - The js-robust and wpd methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
-Although, with the type of residential load data sets, a complete year of load is the minimum requirement to capture predicted differences in winter and summer profiles, for example.
-Even if the data is only available for a month, $nT$ is expected to be at least $1000$ with half-hourly data.
-Hence, practically this is not a challenge as long as the performance is promising for higher $nT=300$.
-
-- For smaller difference between categories, it is seen that method js-nqt would preform better js-robust in our study sample.
-In our study sample, method js-nqt outperforms method js-robust for smaller differences between categories.
-More testing, however, is required to corroborate this.
-
-For more detailed plots and tables, please refer to the supplementary paper.
-
-```{r  parcoord-sim, fig.cap="The parallel coordinate plot illustrates the total inter-cluster distances for granularities $g1$, $g2$, and $g3$.
+Although, a complete year of data is the minimum requirement to capture distributional differences in winter and summer profiles, for example.
+Even if the data is only available for a month, $nT$ with half-hourly data is expected to be at least $1000$.
+As a result, as long as the performance is promising for higher $nT=300$, this is not a challenge.
+
+- In our study sample, the method js-nqt outperforms the method js-robust for smaller differences between categories. More testing, however, is required to corroborate this.
+
+For more detailed results, please refer to the supplementary paper.
+The code for creating these data designs and running the methodologies is available at (https://github.com/Sayani07/paper-gracsR/Validation).
+
+```{r  parcoord-sim, fig.cap = "The parallel coordinate plot illustrates the total inter-cluster distances for granularities $g1$, $g2$, and $g3$.
 One line in the figure depicts the inter-cluster distances for a single simulation scenario.
 While the first plot indicates that no variable stands out during clustering, the next two designs demonstrate that {g1} and {g1, g2} have extremely low inter-cluster distances, indicating that they did not contribute to clustering.
-It is worth emphasising that these facts are consistent with our initial assumptions when designing the scenarios.
-"}
-
-```
-
-```{r mds-plot-validation, fig.cap = "Cases are projected through MDS and the first two dimensions are shown to display the relative position of clusters for different simulation parameters.
-The rows of the grid represent the different mean differences in designs and the columns represent different designs.
-It can be observed that clusters become more compact and separated for higher distributional differences between categories."}
+It is worth emphasising that these facts are consistent with our initial assumptions when designing the scenarios and (a), (b), and (c) correspond to Scenario (a), (b) and (c) respectively."}
+
+```
+
+```{r mds-plot-validation, fig.cap = "Relative positions of clusters corresponding to different scenarios (columns) for different values of mean differences between categories (rows) are shown using the first two dimensions of MDS.
+It can be observed that clusters become more compact and separated for higher mean differences between categories across all designs.
+Between designs, separation is least prominent corresponding to scenario (c) where only granularity is responsible for the clusters."}
 
@@
 Figure \ref{fig:combined-groups-js} shows Groups $2$ and $5$ show a stronger hod pattern with a typical morning and evening peak, whereas groups 1, 3, and 5 show a moy pattern  with higher usage in winter months.
-Differences in wknd-wday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters.
+Differences in Weekend/Weekday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters.
 
@@
 It seems like group 2 and 5 have a hod pattern across its members, while group 1, 3, 5 have a moy pattern.
-Wknd-wday variations across groups are not distinguishable, indicating that it is not a critical variable for clustering.
+Weekend/Weekday variations across groups are not distinguishable, indicating that it is not a critical variable for clustering.
 It is helpful to compare the summarised distributions of groups to that of individuals to confirm that the most of individuals in the group have the same characterisation."}
@@
 Furthermore, combining hetergeneous customers may result in similar-looking final clusters that are not effective for visually differentiating them.
-It is also worth noting that the wknd-wday behavior in the given case does not characterize any cluster.
+It is also worth noting that the Weekend/Weekday behavior in the given case does not characterize any cluster.
 This, however, will not be true for all of the customers in the data set.

### 2021-11-01 main.Rmd
methodology 1st para

@@
 <!-- time series clustering and its challenges -->
-Time-series clustering is the process of unsupervised partitioning of $n$ time-series data into $k$ ($k<n$) groups such that homogeneous time-series are grouped together based on a certain similarity measure.
+Time-series clustering is the process of unsupervised partitioning of $n$ time-series data into $k$ ($k<n$) meaningful groups such that homogeneous time-series are grouped together based on a certain similarity measure.
 The time-series features, length of time-series, representation technique, and, of course, the purpose of clustering time-series all influence the suitable similarity measure or distance strategy to a meaningful level.
@@
 
-The proposed methodology aim to leverage the intrinsic temporal data structure hidden in time series data.
-The foundation of our method is unsupervised clustering algorithms based exclusively on the time-series data.The similarity measure is the most essential ingredient of time series clustering.
-First step is to decide what we mean by similar.
+The foundation of our method is unsupervised clustering algorithms based exclusively on the time-series data.
+The proposed methodology aims to leverage the intrinsic data structure hidden within cyclic temporal granularities.
 The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
-So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.
-In this work, instead of considering the probability distributions of the linear time series, we assume that the measured variables across different categories of any cyclic granularity are from different data generating processes.
-Hence, we want to be able to cluster distributions of the form $f_{i,A,B \dots, {N_C}}(v)$, where $A, B$ represent the cyclic granularities under consideration such that $A = \{a_j: j=1, 2, \dots J\}$,  $B = \{b_k: k  = 1, 2, \dots K\}$ and so on.
-We consider individual category of a cyclic granularity ($A$) or combination of categories for interaction of cyclic granularities (for e.g.
-$A*B$) to have a distribution.
-For example, let us consider we have two cyclic  granularities of interest, $A = {0, 1, 2, \dots, 23}$ representing hour-of-day and  $B = \{Mon, Tue, Wed, \dots, Sun\}$ representing day-of-week.
-Each customer $i$ consist of a collection of probability distributions.
+<!--So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.--> In this work, instead of considering the probability distributions of the linear time series, we assume it across different categories of any cyclic granularity.
+We can consider categories of an individual cyclic granularity ($A$) or combination of categories for two interacting granularities ($A*B$) to have a distribution, where  $A, B$ represent two cyclic granularities defined by $A = \{a_j: j=1, 2, \dots J\}$ and  $B = \{b_k: k  = 1, 2, \dots K\}$.
+For example, let us consider two cyclic  granularities $hod$ and $dow$ representing hour-of-day and day-of-week and defined by, $hod = \{0, 1, 2, \dots, 23\}$ and $dow = \{Mon, Tue, Wed, \dots, Sun\}$.
 In case individual granularities ($A$ or $B$) are considered, there are  $J = 24$ distributions of the form $f_{i,j}(v)$ or $K = 7$ distributions of the form $f_{i,k}(v)$ for each customer $i$.
 In case of interaction, $J*K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
-As a result, we need to decide how to measure similarities between these collections of univariate probability distributions.
-There are multiple ways to measure similarities between time series based on probability distributions, depending on the objective of the problem.
-This paper focuses on looking at the (dis) similarity between underlying distributions that may have resulted in different patterns across different cyclic temporal granularities, that eventually have resulted in the (dis) similarity between time series.
-It considers a methodology with two appraoches for  finding distances between time series.
-Both of these appraoches may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
+Hence clustering these customers is equivalent to clustering these collections of conditional distributions.
+Towards this goal, we need to decide how to measure similarities between collections of univariate probability distributions.
+There are multiple ways to measure similarities depending on the aim of the analysis.
+<!--This paper focuses on looking at the (dis) similarity between underlying distributions that may have resulted in different patterns across different cyclic temporal granularities, that eventually have resulted in the (dis) similarity between time series.
+--> This paper considers a methodology with two approaches for finding distances between time series.
+Both of these approaches may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
 The obtained distances could be fed into a clustering algorithm to break large data sets into subgroups that can then be analyzed separately.
-These clusters may be commonly associated with real-world data segmentation.
-However, since the data is unlabeled a priori, more information is required to corroborate this.
-The methodology is explained in the Figure \ref{fig:flowchart} and each element of the pipeline is discussed.
+<!--These clusters may be commonly associated with real-world data segmentation.
+However, since the data is unlabeled a priori, more information is required to corroborate this.--> The methodology is explained in the Figure \ref{fig:flowchart} and each element of the pipeline is discussed.
 

### 2021-11-01 main.Rmd
freeze finding significant granularities

@@
 
-[@Gupta2021-hakear] proposes a method for choosing significant cyclic granularities and harmonies, which is used in this work.
+[@Gupta2021-hakear] proposes a method for choosing significant cyclic granularities and harmonies (interacting granularities that can be studied together) [@Gupta2021-gravitas], which is used in this work.
 We define "significant" granularities as those with significant distributional differences across categories.
-It is better to select only those granularities because it is expected that there would be some fascinating repetitive behaviour that we are interested in studying.
-It is worth noting that not all of the observations in the study may have the same set of important granularities.
-A method for selecting a list ($S_c$) of significant granularities for all observations may be as follows:
-
- (a) eliminate from the comprehensive list the granularities that are inconsequential for all observations.
-
- (b) consider only those granularities which are significant for most observations.
-
-In both circumstances, there will be observations for which one or a few selected granularities are uninteresting.
-Even in that situation, having this group of observations that show no intriguing patterns over a granularity that frequently detects patterns may be useful.
-In contrast, if the granularities under consideration are indeed significant for a set of observations, distinct patterns could be detected while clustering them.
+It is preferable to choose only those granularities since it is more probable that there will be some intriguing repeated behaviour worth investigating.
+It should be noted that all of the observations in the study may not have the same set of important granularities.
+The following is a method for generating a list ($S_c$) of significant granularities for all observations:
+
+ (a) Remove granularities from the comprehensive list that are insignificant for all observations.
+
+ (b) select only the granularities that are significant for the majority of observations.
+
+There will be observations in both cases where one or a few selected granularities are boring.
+Even in that case, having this group of observations with no interesting patterns at a granularity that regularly discovers patterns may be valuable.
+If, on the other hand, the granularities under examination are truly important for a group of data, unique patterns may be identified while clustering them.
 

### 2021-11-01 main.Rmd
Freeze data pre-preprocessing

@@
 
-Time series often have a somewhat skewed distribution and their ranges might vary greatly. It
-is helpful to do a statistical transformation on the data to bring all of them to the same range or normalize each series.
+Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
+Statistical transformations are employed to bring all of them to the same range or normalize each observation.
+This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.
 For the JS-based approaches, two data transformation techniques are utilised viz, Normal-Quantile Transform (NQT) and Robust scaling.
@@
 
-_Robust scaling_ The normalised $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - p_{0.50}}{p_{0.75}-p_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $p_{0.25}$, $p_{0.50}$ and $p_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
-$v_{norm}$ has zero mean and median, as well as a standard deviation of one, while the outliers are still there with the same relative connections to other values.
+_Robust scaling_ The normalised $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
+$v_{norm}$ has zero mean and median, as well as a standard deviation of one, with the outliers having same relative connections to other values.
 
@@
 As a result, determining which raw distribution was used is difficult using the modified distribution.
-Also, multimodality is disguised or inverted.
-This, however, is not a problem for implementation of this methodology as distributions are characterised by quantiles and the order of the quantiles is reserved under NQT.
+Multimodality is also concealed or reversed.
+As a result, while displaying transformed data using NQT, one must exercise caution.
+This, on the other hand, is a very useful strategy for making assymetric distributions symmetric, which often improves the clustering performance.
+
+<!-- is not a problem for implementation of this methodology as distributions are characterised by quantiles and the order of the quantiles is reserved under NQT. -->
 
@@
 
-@wang2020tsibble introduced the tidy "tsibble" data structure to assist temporal data exploration and modeling.
-To start with, the measured variable for each key variable (observation in this context) is a time-indexed sequence of values for various measurement variables at each time point.
+We start with a "tsibble" (@wang2020tsibble) data structure with an index variable representing inherent ordering from past to present and a key variable that define observational units over time.
+The measured variable for observation is a time-indexed sequence of values.
 This sequence, however, could be shown in several ways.
@@
 Cyclic granularities like hour-of-day, hour-of-week can be expressed in terms of the index set in the "tsibble" data structure.
-But the data structure changes while transporting from linear to cyclic scale of time as multiple observations of the measured variable would correspond to each category of the cyclic granularities.
+But the data structure changes while transporting from linear to cyclic scale of time as multiple observations are generated for each category and induces a probability distribution.
 Directly computing Jensen-Shannon distances between probability distributions can be very costly.
-Hence, in this paper, quantiles are chosen to characterize the probability distributions induced by the multiple observations for each category of the cyclic granularity.
-So, each category of a cyclic granularity corresponds to a list of numbers which is essentially few chosen quantiles of the multiple observations.
-
-<!-- An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
-A linear granularity is a mapping of the index set to subsets of the time domain.
-For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.-->
+Hence, in this paper, deciles are chosen to characterize the probability distributions.
+So, in the final data structure, each category of a cyclic granularity corresponds to a list of numbers which is essentially few chosen quantiles of the multiple observations.
 

### 2021-11-01 main.Rmd
methodology done + spell check

@@
 Defining an appropriate distance measure for the raw time series may be a difficult task since it must take into account noise, variable lengths of time series, asynchronous time series, different scales, and missing data.
-Commonly used Distance-based similarity measures as suggested by a decade review of time series clustering approaches (@Aghabozorgi2015-ct) are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping, Autocorrelation, Short time series distance, Piecewise regularisation, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances (@liao2007clustering) but on a vector time series data.
+Commonly used Distance-based similarity measures as suggested by a decade review of time series clustering approaches (@Aghabozorgi2015-ct) are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping, Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances (@liao2007clustering) but on a vector time series data.
 Among these alternatives, Euclidean distances have high performance but need the same length of data over the same period, resulting in information loss regardless of whether it is on raw data or a smaller collection of features.
@@
 \noindent We consider the problem of clustering a large number of univariate time series of continuous values which are available at fine temporal scales, being motivated by the residential smart meter data.
-These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and missing observations and characterised by noisy and patchy behavior that can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
+These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and missing observations and characterized by noisy and patchy behavior that can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
 Choosing probability distributions seem to be a natural way to analyze these types of data sets since they are robust to uneven length, missing data, or noise.
@@
 Day-off energy use varies depending on whether customers stay home or go outside.
-Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behaviour a fascinating one.
+Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior a fascinating one.
 This challenge is worsened when all we know about our consumers is their energy use history (@Ushakova2020-rl).
@@
 To reduce dimensionality, several studies use principal component analysis or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).
-Other algorithms utilised in the literature include k-means variants, hierarchical approaches, and greedy k-medoids.
-Time series data, such as smart metre data, are not well-suited to any of the techniques mentioned in @Tureczek2017-pb.
+Other algorithms utilized in the literature include k-means variants, hierarchical approaches, and greedy k-medoids.
+Time series data, such as smart meter data, are not well-suited to any of the techniques mentioned in @Tureczek2017-pb.
 Only one study [@ozawa2016determining] identified time series characteristics using Fourier transformation, which converts data from time to frequency and then uses K-Means to cluster by greatest frequency .
-@Motlagh2019-yj suggests that the time feature extraction islimited by the type of noisy, patchy, and unequal time-series common in residential datasets and addresses model-based clustering  by transforming the series into other other objects such as structure or set of parameters which can be more easily characterised and clustered.
+@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential  and addresses model-based clustering  by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered.
 [@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
-@Melnykov2013-sp disucusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
+@Melnykov2013-sp discusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
 None of the methods focus on exploring heterogeneity in repetitive behaviors across different temporal dependencies, we present a way to do that in this paper.
@@
 
-<!-- A model-based clustering works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterised and clustered (@Motlagh2019-yj).
+<!-- A model-based clustering works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered (@Motlagh2019-yj).
 [@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
@@
 
-<!-- The data set solely contains readings from smart meters and no information about the consumers' specific physical, geographical, or behavioural attributes.
+<!-- The data set solely contains readings from smart meters and no information about the consumers' specific physical, geographical, or behavioral attributes.
 As a result, no attempt is made to explain why consumption varies.
@@
 We define "significant" granularities as those with significant distributional differences across categories.
-It is preferable to choose only those granularities since it is more probable that there will be some intriguing repeated behaviour worth investigating.
+It is preferable to choose only those granularities since it is more probable that there will be some intriguing repeated behavior worth investigating.
 It should be noted that all of the observations in the study may not have the same set of important granularities.
@@
 This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.
-For the JS-based approaches, two data transformation techniques are utilised viz, Normal-Quantile Transform (NQT) and Robust scaling.
+For the JS-based approaches, two data transformation techniques are utilized viz, Normal-Quantile Transform (NQT) and Robust scaling.
 NQT is a built-in transformation for computing $wpd$, which is the foundation of wpd-based distances.
 
-_Robust scaling_ The normalised $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
+_Robust scaling_ The normalized $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
 $v_{norm}$ has zero mean and median, as well as a standard deviation of one, with the outliers having same relative connections to other values.
@@
 
-<!-- is not a problem for implementation of this methodology as distributions are characterised by quantiles and the order of the quantiles is reserved under NQT. -->
+<!-- is not a problem for implementation of this methodology as distributions are characterized by quantiles and the order of the quantiles is reserved under NQT. -->
 
@@
 Another purpose of clustering could be to group customers that have similar differences in patterns across all major cyclic granularities, capturing similar jumps across categories regardless of the overall shape.
-For example, in the first goal, similar shapes across hours of the day will be grouped together, resulting in customers with similar behaviour across all hours of the day, whereas in the second goal, any similar big-enough jumps across hours of the day will be clubbed together, regardless of which hour of the day it is.
+For example, in the first goal, similar shapes across hours of the day will be grouped together, resulting in customers with similar behavior across all hours of the day, whereas in the second goal, any similar big-enough jumps across hours of the day will be clubbed together, regardless of which hour of the day it is.
 Both of these objectives may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
@@
 Directly computing Jensen-Shannon distances between probability distributions can be very costly.
-Hence, in this paper, deciles are chosen to characterize the probability distributions.
+Hence, in this paper, quantiles are chosen to characterize the probability distributions.
 So, in the final data structure, each category of a cyclic granularity corresponds to a list of numbers which is essentially few chosen quantiles of the multiple observations.
@@
 
-Considering each individual or combined categories of cyclic granularities as a data generating process lead to a collection of conditional distributions for each customer $i$.
-The (dis) similarity between each pair of customers should be obtained by combining the distances between these collections of conditional distributions such that the resulting metric is a distance metric, which could be fed into the clustering algorithm.
+<!-- Considering each individual or combined categories of cyclic granularities to have an underlying distribution lead to a collection of conditional distributions for each customer $i$. -->
+
+The total (dis) similarity between each pair of customers is obtained by combining the distances between the collections of conditional distributions.
+This needs to be done in a way such that the resulting metric is a distance metric, and could be fed into the clustering algorithm.
 Two types of distance metric is considered:
@@
 This distance metric considers two time series to be similar if the distributions of each category of an individual cyclic granularity or combination of categories for interacting cyclic granularities are similar.
-In this study, the distribution for each category is characterised using deciles (can potentially consider any list of quantiles), and the distances between distributions are calculated using the Jensen-Shannon distances (@Menendez1997-in), which are symmetric and thus could be used as a distance measure.
-
-\noindent The sum of the distances between two series $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
+In this study, the distribution for each category is characterized using deciles (can potentially consider any list of quantiles), and the distances between distributions are calculated using the Jensen-Shannon distances (@Menendez1997-in), which are symmetric and thus could be used as a distance measure.
+
+\noindent The sum of the distances between two observations $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
 After determining the distance between two series in terms of one granularity, we must combine them to produce a distance based on all significant granularities.
 When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels,  $$S_{x, y} = \sum_lS^{C_l}_{x,y}/n_l$$ is employed, which is also a distance metric since it is the sum of JS distances.
-In this approach, the variation in time series within each group is in magnitude rather than distributional pattern, while the variation between groups is only in distributional pattern across categories.
+This approach is expected to yield groups, such that the variation in observations within each group is in magnitude rather than distributional pattern, while the variation between groups is only in distributional pattern across categories.
 
@@
 Compute weighted pairwise distances $wpd$ (@Gupta2021-hakear) for all considered granularities for all observations.
-$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalised by appropriate parameters.
+$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalized by appropriate parameters.
 A higher value of $wpd$ indicates that some interesting pattern is expected, whereas a lower value would indicate otherwise.
 
-\noindent Once we have chosen $wpd$ as a relevant feature for characterizing the time series across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
-The euclidean distance between them is chosen, with the granularities acting as variables and $wpd$ representing the value under each variable.
+\noindent Once we have chosen $wpd$ as a relevant feature for characterizing the distributions across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
+The Euclidean distance between them is chosen, with the granularities acting as variables and $wpd$ representing the value under each variable.
 With this approach, we should expect the observations with similar $wpd$ values to be clustered together.
 Thus, this approach is useful for grouping observations that have similar significance of patterns across different granularities.
-Similar significance does not imply similar pattern, which is where this technique varies from JS-based distances, which detect differences in patterns across categories.
+Similar significance does not imply a similar pattern, which is where this technique varies from JS-based distances, which detect differences in patterns across categories.
 
@@
 If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
-We aim to have pockets of households showing similar periodic behavior by conidering $wpd$ vlaues for different harmonies and some time series features.
+We aim to have pockets of households showing similar periodic behavior by considering $wpd$ values for different harmonies and some time series features.
 The features should also characterize probability distributions of different household.
@@
 A good comprehensive list of algorithms can be found in @Xu2015-ja based on traditional ways like partition, hierarchy or more recent approaches like distribution, density and others.
-We employ agglomerative hierarchical clustering in conjunction with Ward's criteria (XXX reference).
-Hierarchical cluster techniques fuse neighbouring points sequentially to form bigger clusters, beginning with a full pairwise distance matrix.
+We employ agglomerative hierarchical clustering in conjunction with Ward's linkage.
+Hierarchical cluster techniques fuse neighboring points sequentially to form bigger clusters, beginning with a full pairwise distance matrix.
 The distance between clusters is described using a "linkage technique".
@@
 
-\noindent Cluster characterization, both quantitatively and qualitatively, is a crucial aspect in cluster analysis.
-@Cook2007-qe lists numerous methods for characterising clusters.
-Listed below are a few techniques and R packages that are utilized in this study.
-
-(a) _Parallel coordinate plots_ (@wegman1990hyperdimensional) are often used to visualise high-dimensional and multivariate data, allowing visual grouping and pattern detection..
-A Parallel Coordinates Plot features parallel axes for each variable.
-Each axis is linked by lines.
-The axes' arrangement may affect the reader's interpretation of the data.
+\noindent Cluster characterization is an important element of cluster analysis.
+@Cook2007-qe provides several methods for characterizing clusters.
+_Parallel coordinate plots_ (@wegman1990hyperdimensional), _Scatterplot matrix_, _Displaying cluster statistics_ (@dasu2005grouping), _MDS_ (@borg2005modern), PCA, _t-SNE_(@R-tsne), _Tour_ (@wickham2011tourr) are some of the graphical approaches used in this study.
+A Parallel Coordinates Plot features parallel axes for each variable and each axis is linked by lines.
 Changing the axes may reveal patterns or relationships between variables for categorical variables.
-However, for categories with cyclic temporal granularities, preserving the underlying ordering is more desirable.
-
-(b) _Scatterplot matrix_ contains pairwise scatter plots of the p variables.
-Pairwise scatter plots are useful for figuring out how variables relate to each other and how factors determine the clustering.
-
-(c) _Displaying cluster statistics_ are useful when we have larger problems and it is difficult to read the Parallel coordinate plots due to congestion. (@dasu2005grouping)
-
-(d) _MDS, PCA and t-SNE_ While all of them use a distance or dissimilarity matrix to construct a reduced-dimension space representation, their goals are diverse.
-PCA seeks to retain data variance.
+However, for categories with cyclic temporal granularities, preserving the underlying ordering of time is more desirable.
+Displaying cluster statistics is useful when we have larger problems and it is difficult to read the parallel coordinate plots due to congestion.
+All of MDS, PCA and t-SNE use a distance or dissimilarity matrix to construct a reduced-dimension space representation, their goals are diverse.
 Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
-t-SNE, on the other hand, is concerned with preserving neighbourhood data points.
-The t-SNE embeddings will compress data points which are close in high-dimensional space.
-
-(e) _Tour_ is a collection of interpolated linear projections of multivariate data into lower-dimensional space.
-As a result, the viewer may observe the high-dimensional data's shadows from a low-dimensional perspective.
-
+The t-SNE embedding will compress data points which are close in high-dimensional space.
+Tour is a collection of interpolated linear projections of multivariate data into lower-dimensional space.
 The cluster characterization approach varies depending on the distance metric used.
@@
 For JS-based distances, plotting cluster statistics is beneficial for characterization and variable importance could be displayed through parallel coordinate plots.
-This part of the work uses R packages `GGally` (@R-GGally), `Rtsne` (@R-tsne), `ggplot2` (Wickham2009pk), `tour` (@wickham2011tourr), `stats` (@R-language).
+
+<!-- CUT IT SHORT -->
+
+<!-- (a) _Parallel coordinate plots_ (@wegman1990hyperdimensional) are often used to visualize high-dimensional and multivariate data, allowing visual grouping and pattern detection.
+A Parallel Coordinates Plot features parallel axes for each variable.
+Each axis is linked by lines.
+Changing the axes may reveal patterns or relationships between variables for categorical variables.
+However, for categories with cyclic temporal granularities, preserving the underlying ordering of time is more desirable.
+-->
+
+<!-- (b) _Scatterplot matrix_ contains pairwise scatter plots of the p variables.
+Pairwise scatter plots are useful for figuring out how variables relate to each other and how factors determine the clustering.
+-->
+
+<!-- (c) _Displaying cluster statistics_ are useful when we have larger problems and it is difficult to read the Parallel coordinate plots due to congestion. (@dasu2005grouping) -->
+
+<!-- (d) _MDS, PCA and t-SNE_ While all of them use a distance or dissimilarity matrix to construct a reduced-dimension space representation, their goals are diverse.
+PCA seeks to retain data variance.
+Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
+t-SNE, on the other hand, is concerned with preserving neighborhood data points.
+The t-SNE embedding will compress data points which are close in high-dimensional space.
+-->
+
+<!-- (e) _Tour_ is a collection of interpolated linear projections of multivariate data into lower-dimensional space.
+As a result, the viewer may observe the high-dimensional data's shadows from a low-dimensional perspective.
+-->
+
+<!-- The cluster characterization approach varies depending on the distance metric used.
+Parallel coordinate plots, scatter plot matrices, MDS or PCA are potentially useful ways to characterize clusters using wpd-based distances.
+For JS-based distances, plotting cluster statistics is beneficial for characterization and variable importance could be displayed through parallel coordinate plots.
+-->
+
+<!-- This part of the work uses R packages `GGally` (@R-GGally), `Rtsne` (@R-tsne), `ggplot2` (Wickham2009pk), `tour` (@wickham2011tourr), `stats` (@R-language). -->
 
@@
 
-_Scenario (a): All signifiant granularities_
+_Scenario (a): All significant granularities_
 
@@
 
-_Scenario (c): One signifiant granularity_
+_Scenario (c): One significant granularity_
 
@@
 Consider a group that has a different weekday and weekend behavior in the summer but not in the winter.
-This type of combined behaviour across granularities  can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (Weekend/Weekday and month-of-year in this example).
-As a result, in this scenario, we analyse a combination of categories generated from different distributions.
+This type of combined behavior across granularities  can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (Weekend/Weekday and month-of-year in this example).
+As a result, in this scenario, we analyze a combination of categories generated from different distributions.
 Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$.
@@
 One line in the figure, for example, shows the inter-cluster distances for one simulation setting.
-The lines are not coloured by group since the purpose is to highlight the contribution of the factors to categorization rather than class separation.
+The lines are not colored by group since the purpose is to highlight the contribution of the factors to categorization rather than class separation.
 The first plot shows that no variable stands out in the clustering, but the following two designs show that {g1} and {g1, g2} have very low inter cluster distances, meaning that they did not contribute to the clustering.
@@
 - Obtain $wpd$ for all cyclic granularities considered for these customers.
-It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weeknd/weekday) are coming out to be significant for most customers.
+It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weekend/weekday) are coming out to be significant for most customers.
 We use these three granularities while clustering.
@@
 - Remove customers whose data for an entire category of a significant granularity is empty.
-For example, a customer who does not have data for an entire month is excluded because their monthly behaviour cannot be analyzed.
+For example, a customer who does not have data for an entire month is excluded because their monthly behavior cannot be analyzed.
 
@@
 There are several ways to approach the prototype selection.
-Following @Fan2021-bq's idea of picking related examples (neighbours) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2D space.
+Following @Fan2021-bq's idea of picking related examples (neighbors) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2D space.
 Then pick a few "anchor" customers who are far apart in 2D space and pick a few neighbors for each.
@@
 Summarizing distributions across clusters with varied or outlying customers can result in a shape that does not represent the group.
-Furthermore, combining hetergeneous customers may result in similar-looking final clusters that are not effective for visually differentiating them.
+Furthermore, combining heterogeneous customers may result in similar-looking final clusters that are not effective for visually differentiating them.
 It is also worth noting that the Weekend/Weekday behavior in the given case does not characterize any cluster.
@@
 Probability distributions provide an intuitive method to characterise noisy, patchy, long, and unequal-length time series data.
-Distributions over cyclic granularities help to characterise the formed clusters in terms of their repeating behaviour over these cyclic granularities.
-Furthermore, unlike earlier efforts that group customers based on behavaior across only one cyclic granularity (such as hour-of-day), our method is more comprehensive in detecting clusters with repeated patterns at all relevant granularities.
+Distributions over cyclic granularities help to characterise the formed clusters in terms of their repeating behavior over these cyclic granularities.
+Furthermore, unlike earlier efforts that group customers based on behavior across only one cyclic granularity (such as hour-of-day), our method is more comprehensive in detecting clusters with repeated patterns at all relevant granularities.
 The JS-based approaches are faster than wpd-based approaches.
@@
 First, larger data sets with more uncertainty complicate matters, as is true for any clustering task.
-Characterising clusters with varied or outlying customers can result in a shape that does not represent the group.
+Characterizing clusters with varied or outlying customers can result in a shape that does not represent the group.
 Moreover, integrating heterogeneous consumers may result in visually identical end clusters, which are potentially not useful.
@@
 In reality, however, it might change.
-For the smart meter example, the distribution for a customer moving to a different house or changing electrical equipments can change drastically.
+For the smart meter example, the distribution for a customer moving to a different house or changing electrical equipment can change drastically.
 Our current approach can not detect these dynamic changes.
@@
 In specific applications the user may be interested in some of these criteria rather than others.
-A focus of the paper is on methodology to standardise the different characteristics so that users can aggregate them in a suitable way specifying weights for the various criteria that are relevant in the clustering application at hand.
--->
+A focus of the paper is on methodology to standardize the different characteristics so that users can aggregate them in a suitable way specifying weights for the various criteria that are relevant in the clustering application at hand.
+-->

### 2021-11-02 main.Rmd
proof reading changes

@@
 (b) Using probability distributions over "cyclic granularities" ensures small pockets of similar "repeated" behaviours.
-Our method is capable of producing useful clusters, as demonstrated by testing on validation data designs and a sample of residential smart metee consumers.
+Our method is capable of producing useful clusters, as demonstrated by testing on validation data designs and a sample of residential smart meter consumers.
 
@@
 
-```{r mytheme}
-theme_validation <- function() {
-  theme_bw() +
-      theme(
-    strip.text = element_text(size = 8, margin = margin(b = 0, t = 1)),
-    plot.margin = margin(0, 0, 0, 0, "cm"),
-    axis.title.y = element_blank(),
-    #axis.text.x = element_blank(),
-  axis.text.y = element_blank(),
-  axis.ticks = element_blank(),
-      legend.position = "bottom"
-    )
-}
-
-theme_characterisation <- function() {
-
-  theme_bw() + # seeting theme
-    theme(strip.text = element_text(size = 10,
-                                    margin = margin(b = 0, t = 0))) + # narrow facet space
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + # no axis ticks 
-  theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
-    theme(axis.text.x = element_text(angle=90, hjust=1, size = 10)) + # rotate the x-axis text
-    theme(legend.position = "bottom")+
-  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-  theme(axis.text.x = element_text(size=5))
-}
-```
-
 ```{r mytheme-application}
-theme_application <- function() {
-
-  theme_light() + # setting theme
-    #theme(strip.text = element_text(margin = margin(b = 0, t = 0))) + # narrow facet space
-   theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + # no axis ticks 
-  theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
-    theme(axis.ticks.x = element_blank(),
-        axis.text.x = element_blank()) + # no x-axis labels to further reduce the gap between facets
-    #theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
-  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-  #theme(axis.text.x = element_text(size=5)) +
-    theme(strip.background = element_blank(),
-  strip.text.x = element_blank())
-}
+```
+
+```{r theme_characterisation}
+```
+
+```{r theme-validation}
+
 ```
@@
 Time-series clustering is the process of unsupervised partitioning of $n$ time-series data into $k$ ($k<n$) meaningful groups such that homogeneous time-series are grouped together based on a certain similarity measure.
-The time-series features, length of time-series, representation technique, and, of course, the purpose of clustering time-series all influence the suitable similarity measure or distance strategy to a meaningful level.
+The time-series features, length of time-series, representation technique, and, of course, the purpose of clustering time-series all influence the suitable similarity measure or distance metric to a meaningful level.
 The three primary methods to time series clustering (@liao2005clustering) are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based).
@@
 Choosing probability distributions seem to be a natural way to analyze these types of data sets since they are robust to uneven length, missing data, or noise.
-Hence this paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across significant cyclic granularities.
-Cyclic temporal granularities, which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
+This paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across significant cyclic granularities.
+Cyclic temporal granularities [@Gupta2021-hakear], which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
 The resulting clusters are expected to group customers that have similar repetitive behaviors across each of the interesting cyclic granularities.
-Below are some of the benefits of our method, which will be detailed in further depth in subsequent sections.
-
-- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern).
+Below are some of the benefits of our method:
+
+- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern);
 
@@
 Time series data, such as smart meter data, are not well-suited to any of the techniques mentioned in @Tureczek2017-pb.
-Only one study [@ozawa2016determining] identified time series characteristics using Fourier transformation, which converts data from time to frequency and then uses K-Means to cluster by greatest frequency .
+Only one study [@ozawa2016determining] identified time series characteristics using Fourier transformation, which converts data from time to frequency and then uses K-Means to cluster by greatest frequency.
 @Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential  and addresses model-based clustering  by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered.
@@
 @Melnykov2013-sp discusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
-None of the methods focus on exploring heterogeneity in repetitive behaviors across different temporal dependencies, we present a way to do that in this paper.
+In our knowledge, none of the methods focus on exploring heterogeneity in repetitive behaviors based on the dynamics of multiple temporal dependencies using probability distributions.
 
@@
 
-- Remove customers whose data for an entire category of a significant granularity is empty.
+- Remove customers whose data for an entire category of `hod`, `moy` or `wnwd` is empty.
 For example, a customer who does not have data for an entire month is excluded because their monthly behavior cannot be analyzed.
@@
 Hence we perform a linked tour with t-SNE layout using the R package `liminal` (@R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
-Please see the Supplementary article for further details on how the prototypes are chosen.
-Figure \ref{fig:hod-ind-group} (a, b, c) shows the raw time plot, distribution across `hod`, `moy` and `wkndwday` for the set of chosen $24$ customers.
+(Refer to Supplementary article for further details).
+Figure \ref{fig:hod-ind-group} (a, b, c) shows the raw time plot, distribution across `hod`, `moy` and `wnwd` for the set of chosen $24$ customers.
 Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution.
@@
     aes(group = 1), size = 1, color = "black")+
-  xlab("wdwn")+
+  xlab("wnwd")+
   theme_application()
@@
 
-Cluster characterization is a crucial aspect of cluster analysis.
-The $24$ prototypes <!--and the data subset consisting of $353$ customers--> are clustered using the methodology described in \ref{sec:methodology} and results are reported below.
+The $24$ prototypes <!--and the data subset consisting of $353$ customers--> are clustered using the methodology described in Section \ref{sec:methodology} and results are reported below.
 In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.
@@
 The plotting scales are not displayed since we want to emphasize comparable shapes rather than scales.
-The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to shape or significance across cyclic granularities.
+The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.
 
@@
 For clustering based on JS-based distances, we chose the optimal number of clusters using (@Hennig2014-ah) as $5$.
-The distribution of electricity demand for the selected $24$ customers across hod, moy and wdwn are shown in Figure \ref{fig:hod-ind-group} (d, e, f).
+The distribution of electricity demand for the selected $24$ customers across `hod`, `moy` and `wnwd` are shown in Figure \ref{fig:hod-ind-group} (d, e, f).
 Our methodology is useful for grouping similar distributions over hod and moy and they are placed closely for easy comparison.
@@
 Figure \ref{fig:combined-groups-js} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
-Figure \ref{fig:combined-groups-js} shows Groups $2$ and $5$ show a stronger hod pattern with a typical morning and evening peak, whereas groups 1, 3, and 5 show a moy pattern  with higher usage in winter months.
-Differences in Weekend/Weekday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters.
+It shows Groups $2$ and $5$ show a stronger hod pattern with a typical morning and evening peak, whereas groups 1, 3, and 5 show a moy pattern  with higher usage in winter months.
+Differences in Weekend/Weekday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters unless maybe conditioned by `moy` or `hod`.
 
@@
   scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7"))  + theme(legend.position = "none")+
-  xlab("wdwn")
+  xlab("wnwd")
 
@@
 We chose the optimal number of clusters using (@Hennig2014-ah) as $3$.
-A parallel coordinate plot with the three significant cyclic granularities used for wpd-based clustering.
+A parallel coordinate plot with the three significant cyclic granularities is used to characterise the groups here.
 The variables are sorted according to their separation across classes (rather than their overall variation between classes).
-This means that $moy$ is the most important variable in distinguishing the designs followed by $hod$ and $wkndwday$.
-There is only one customer who has significant $wpd$ across $wkndwday$ and stands out from the rest of the customers.
+This means that $moy$ is the most important variable in distinguishing the designs followed by $hod$ and $wnwd$.
+There is only one customer who has significant $wpd$ across $wnwd$ and stands out from the rest of the customers.
 Group 3 has a higher wpd for hod than moy or wkndwday.
@@
             hod = round(median(hour_day),2),
-            wdwn = round(median(wknd_wday),2)) %>% 
+            wnwd = round(median(wknd_wday),2)) %>% 
   select(-group)
@@
 If more extensive prototype selection is used, resulting in more comprehensive prototypes in the data set, this method might be used to classify the entire data set into these prototype behaviors.
-However, the goal of this section was to have a few customers that have significant patterns over one or more cyclic granularities, apply our clustering methodology to cluster them, and demonstrate that the method produces useful clusters.
+However, the goal of this section was to have a few customers that have significant patterns over one or more cyclic granularities, apply our methodology to cluster them, and demonstrate that the method produces useful clusters.
 
@@
 
-We offer two methods for calculating pairwise distances between time series based on probability distributions at various cyclic granularities.
+We offer two approaches for calculating pairwise distances between time series based on probability distributions over multiple cyclic granularities at once.
 Depending on the goal of the clustering, these distance metrics, when fed into a hierarchical clustering algorithm using Ward's linkage, yield meaningful clusters.
@@
 Furthermore, unlike earlier efforts that group customers based on behavior across only one cyclic granularity (such as hour-of-day), our method is more comprehensive in detecting clusters with repeated patterns at all relevant granularities.
-The JS-based approaches are faster than wpd-based approaches.
+<!--The JS-based approaches are faster than wpd-based approaches.-->
 
@@
 Future work can make the computations more efficient so that they are easily scalable to a large number of customers.
+Moreoever, experiments can also be run with non-hierarchy based clustering algorithms to verify if these distances work better with other algorithms.
 

### 2021-11-02 paper-supplementary.Rmd
proof reading changes

@@
 ---
+
+```{r external, include = FALSE, cache = FALSE}
+knitr::read_chunk('script/main.R')
+```
+
+```{r load-lib}
+
+```
+
+```{r mytheme-application}
+```
+
+```{r theme_characterisation}
+```
+
+```{r theme-validation}
+
+```
 
@@
 A dual-level progressive similar instance selection (DPSIS) approach could also be used -->
+
+## Clustering on 350 customers
+
+```{r data-load-entire}
+# Read the nqt distances
+
+wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_nqt.rds")) %>% broom::tidy()
+
+moy <- read_rds(here("data/dist_gran_moy_356cust_nqt.rds")) %>% broom::tidy()
+
+hod <- read_rds(here("data/dist_gran_hod_356cust_nqt.rds")) %>% broom::tidy()
+
+# Make the distance metrics
+
+distance <- wkndwday %>% 
+  left_join(moy, by = c("item1", "item2")) %>% 
+  left_join(hod, by = c("item1", "item2")) %>% 
+  rename("wkndwday" ="distance.x",
+         "moy" = "distance.y",
+         "hod" = "distance") %>%
+  mutate(item1 = as.integer(as.character(item1)),
+         item2 = as.integer(as.character(item2)))
+
+filtered_distance <- distance %>%
+  filter(!(item1 %in% c(8196183, 8508008, 8680538))) %>% 
+  filter(!(item2 %in% c(8196183, 8508008, 8680538)))
+
+total_distance <- filtered_distance %>% 
+  mutate(total = wkndwday/2 + moy/12 + hod/24)
+
+total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
+                                                      names_from = item2,
+                                                      values_from = total)
+
+rownames(total_distance_wide) <- total_distance_wide$item1
+
+mds_data <- total_distance_wide %>% 
+  mutate_all(~replace(., is.na(.), 0)) %>%
+  tibble::rownames_to_column() %>%  
+  dplyr::select(-item1) %>% 
+  pivot_longer(-rowname) %>% 
+  pivot_wider(names_from=rowname, values_from=value)
+
+rownames(mds_data) <- total_distance_wide$item1
+
+df <- mds_data[-1] %>% as.matrix()
+DM <- matrix(0, ncol(mds_data), ncol(mds_data))
+DM[lower.tri(DM)] = df[lower.tri(df, diag=TRUE)] # distance metric
+f = as.dist(DM)
+
+first_lot <- mds_data %>% names()
+
+id <- c(first_lot[-1], mds_data$name[nrow(mds_data)])
+
+# Find optimal number of clusters
+
+library(fpc)
+library(cluster)
+k = array()
+for(i in 5:50)
+{
+  group <- f %>% hclust (method = "ward.D") %>% cutree(k=i)
+  p <- cluster.stats(f, clustering = group, silhouette = TRUE)
+  k[i]=p$sindex
+}
+
+ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 50, 2))
+
+# plot(k, type = "l")
+# 6 coming as the number of clusters with maximum silwidth
+
+group <- f %>% hclust (method = "ward.D") %>% cutree(k=17)
+cluster_result <- bind_cols(customer_id = id, group = group)
+
+# cluster_result %>% group_by(group) %>% count()
+
+```
+
+```{r data-heatmap-hod-group-entire}
+legend_title <- "group"
+
+data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
+  mutate(customer_id = as.character(customer_id)) %>%
+  gracsr::scale_gran( method = "robust",
+                      response = "general_supply_kwh")
+
+data_group <- data_pick  %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+   gracsr::scale_gran( method = "robust",
+                       response = "general_supply_kwh") %>% 
+  left_join(cluster_result, by = c("customer_id"))
+
+data_heatmap_hod_group <- quantile_gran(data_group,
+                                  gran1="hour_day",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
+
+data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")
+
+hod_group_entire <- data_heatmap_hod_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=group,
+                  fill = as.factor(group), alpha = 0.5),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=group, 
+                color = as.factor(group)), size = 1)+
+  facet_wrap(~group, 
+             scales = "free_y",  
+             ncol = 17) + 
+              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
+  ylab("demand (in Kwh)") + 
+  theme_bw()  +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ 
+  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme_application() +
+  scale_fill_manual(values=as.vector(polychrome(20)))+
+  scale_color_manual(values=as.vector(polychrome(20)))+
+  theme(legend.position = "bottom")
+
+```
+
+```{r data-heatmap-moy-group-entire}
+data_heatmap_moy_group <- quantile_gran(data_group,
+                                  gran1="month_year",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")
+
+moy_group_entire <- data_heatmap_moy_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             ncol = 17) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
+  ylab("demand (in Kwh)") +
+  theme_bw() + theme_application() +
+  scale_fill_manual(values=as.vector(polychrome(20)))+
+  scale_color_manual(values=as.vector(polychrome(20)))+
+  theme(legend.position = "bottom")
+```
+
+```{r data-wnwd-group-entire}
+wkndwday_data <- data_group %>%
+  create_gran("wknd_wday") 
+wkndwday_data$group <- as.factor(wkndwday_data$group)
+# %>% 
+#   create_gran("hour_day")
+
+ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]
+
+wkndwday_group_entire <- wkndwday_data%>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
+  geom_boxplot(aes(fill = group, color = group),alpha = 0.5, outlier.alpha = 0.05)+
+  #geom_boxplot(outlier.size = 1) + 
+  coord_cartesian(ylim = ylim1*1.05)+
+  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
+#geom_boxplot(aes(fill = as.factor(group))) +
+  #scale_fill_lv() +
+ xlab("wknd-wday") + 
+  ylab("demand (in Kwh)") +
+   facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_both",
+             ncol = 17) + 
+  theme_bw() + theme_application() +
+  scale_fill_manual(values=as.vector(polychrome(20)))+
+  scale_color_manual(values=as.vector(polychrome(20)))+
+  theme(legend.position = "none")
+```
+
+```{r combined-groups-js-entire, fig.cap = "The distribution of electricity demand for the clusters across hod (a), moy (b) and wkndwday (c).
+It seems like group 2 and 5 have a hod pattern across its members, while group 1, 3, 5 have a moy pattern.
+Wknd-wday variations across groups are not distinguishable, indicating that it is not a critical variable for clustering.
+It is helpful to compare the summarised distributions of groups to that of individuals to confirm that the most of individuals in the group have the same characterisation."}
+
+ (hod_group_entire/moy_group_entire/ wkndwday_group_entire) +
+  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
+  plot_layout(guides = "collect")& theme(legend.position = 'none')
+```
+
+# contribution
+
+```{r contribution}
+# data_validation <- hod %>%  
+#   rename("hod" = "distance") %>% 
+#   left_join(moy, 
+#             by = c("item1", "item2"))%>% 
+#   rename("moy" = "distance") %>% 
+#   left_join((wkndwday), 
+#             by = c("item1", "item2"))%>% 
+#   rename("wkndwday" = "distance") %>% 
+#   left_join(cluster_result, by = c("item1" = "customer_id")) %>% 
+#   rename("group_item1" = "group") %>% 
+#   left_join(cluster_result, by = c("item2" = "customer_id")) %>%  
+#   rename("group_item2" = "group") %>% 
+#   mutate(hod = hod/24, moy = moy/12, wkndwday = wkndwday/2) %>% 
+#   filter(!is.na(group_item1)| !is.na(group_item2)) %>% 
+#   pivot_longer(3:5,names_to="gran",
+#                values_to = "distance")  
+#
+
+data_validation_cat <- read_rds(here::here("validation/data_validation_cat_all_cluster.rds"))
+
+# hod_cat <- data_pick %>%
+#   dist_gran_cat(gran1 = "hour_day", response = "general_supply_kwh") %>%
+#   mutate(gran = "hod")
+# 
+# moy_cat <- data_pick %>%
+#   dist_gran_cat(gran1 = "month_year", response = "general_supply_kwh")%>%
+#   mutate(gran = "moy")
+# 
+# wkndwday_cat <- data_pick %>%
+#   dist_gran_cat(gran1 = "wknd_wday", response = "general_supply_kwh")%>%
+#   mutate(gran = "wnwd")
+# 
+# 
+# data_validation_cat <- bind_rows(hod_cat, moy_cat, wkndwday_cat) %>%
+#   mutate(customer_from = as.character(customer_from),
+#          customer_to = as.character(customer_to))
+
+gran_cat_dist <- data_validation_cat %>% 
+  mutate(customer_from = as.character(customer_from),
+         customer_to = as.character(customer_to)) %>% 
+  left_join(cluster_result, by = c("customer_from" = "customer_id")) %>% 
+  rename("group_from" = "group") %>% 
+  left_join(cluster_result, by = c("customer_to" = "customer_id")) %>%  
+  rename("group_to" = "group") %>% 
+  group_by(gran,
+           category,
+           group_from, 
+           group_to) %>% 
+  summarise(sum = sum(distance)) %>% 
+  pivot_wider(names_from = group_to, values_from = sum) %>% 
+  mutate(distance = sum(`1`, `2`, `3`, `4`, `5`,`6`, `7`, `8`, `9`, `10`, `11`, `12`, `13`, `14`, `15`, `16`, `17`)) %>% 
+  select(-(4:8)) %>% 
+  arrange(-distance)
+
+# Contribution of individual category for hod
+
+gran_cat_hod <-   gran_cat_dist %>% 
+  filter(gran=="hod") %>% 
+  mutate(category = as.integer(category)) %>% 
+  arrange(group_from, category)
+
+gran_cat_hod$category <- factor(gran_cat_hod$category, levels = 0:23)
+
+# for each group, percent contribution of hours
+
+group_total <- gran_cat_hod %>% 
+  ungroup() %>% 
+  group_by(group_from) %>% 
+  summarise(group_total = sum(distance))
+
+data_pcp <- gran_cat_hod %>% 
+  ungroup %>% 
+  left_join(group_total, by = "group_from") %>% 
+  mutate(cat_contibution = distance*100/group_total) %>% 
+  group_by(group_from) %>% 
+  #arrange(-distance) %>% 
+  select(group_from, category, cat_contibution) %>% 
+  pivot_wider(names_from = "category", 
+              values_from = "cat_contibution") %>% mutate(group_from = as.factor(group_from))
+
+parcoord <- GGally::ggparcoord(data_pcp ,
+                               columns = 2:ncol(data_pcp),
+                               groupColumn = "group_from",
+                               showPoints = TRUE, 
+                               alphaLines = 1,
+                               scale = "globalminmax"
+) + 
+  ggplot2::theme(
+    plot.title = ggplot2::element_text(size=10)
+  )+
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
+  theme(legend.position = "bottom") +
+  xlab("") +
+  ylab("wpd")+
+  scale_fill_manual(values=as.vector(polychrome(20)))+
+  theme_bw()
+parcoord 
+```

### 2021-11-02 supplementary/application/main_application.Rmd
proof reading changes

--- 2021-10-30 main_application.Rmd
@@
 
-# ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 50, 2))
-# 
-# 
+ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 50, 2))
+
 # plot(k, type = "l")
@@
 legend_title <- "group"
+
+data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
+  mutate(customer_id = as.character(customer_id)) %>%
+  gracsr::scale_gran( method = "robust",
+                      response = "general_supply_kwh")
 
@@
              scales = "free_y",  
-             ncol = 16) + 
+             ncol = 17) + 
               #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
@@
              labeller = "label_value",
-             ncol = 16) +
+             ncol = 17) +
     theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
@@
              labeller = "label_both",
-             ncol = 16) + 
+             ncol = 17) + 
   theme_bw() + theme_application() +
@@
   plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
-  plot_layout(guides = "collect")& theme(legend.position = 'bottom')
+  plot_layout(guides = "collect")& theme(legend.position = 'none')
 ```

### 2021-11-02 supplementary/prototype_selection/selecting_customers.Rmd
proof reading changes

--- 2021-10-28 selecting_customers.Rmd
@@
 ```{r tsne, eval = FALSE}
-data_356cust_wide <- read_rds("data/data_356cust_wide.rds")
+data_356cust_wide <- read_rds(here::here("data/data_356cust_wide.rds"))
 
@@
                    alphaLines = 1,
-                   order = "anyClass",
                    scale = "globalminmax"

### 2021-11-02 script/main.R

@@ -23,9 +23,63 @@ library(forcats)
 library(tidyr)
 library(purrr)
 library(stringr)
+library(pals)
 
 ##----load-data
 load("data/ALL_DATA.Rdata")
+
+
+
+##----theme-validation
+theme_validation <- function() {
+  theme_bw() +
+    theme(
+      strip.text = element_text(size = 8, margin = margin(b = 0, t = 1)),
+      plot.margin = margin(0, 0, 0, 0, "cm"),
+      axis.title.y = element_blank(),
+      #axis.text.x = element_blank(),
+      axis.text.y = element_blank(),
+      axis.ticks = element_blank(),
+      legend.position = "bottom"
+    )
+}
+
+
+##----theme_characterisation
+
+theme_characterisation <- function() {
+  
+  theme_bw() + # seeting theme
+    theme(strip.text = element_text(size = 10,
+                                    margin = margin(b = 0, t = 0))) + # narrow facet space
+    theme(axis.title.y=element_blank(),
+          axis.text.y=element_blank(),
+          axis.ticks.y=element_blank()) + # no axis ticks 
+    theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
+    theme(axis.text.x = element_text(angle=90, hjust=1, size = 10)) + # rotate the x-axis text
+    theme(legend.position = "bottom")+
+    theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+    theme(axis.text.x = element_text(size=5))
+}
+
+
+##----mytheme-application
+theme_application <- function() {
+  
+  theme_light() + # setting theme
+    #theme(strip.text = element_text(margin = margin(b = 0, t = 0))) + # narrow facet space
+    theme(axis.title.y=element_blank(),
+          axis.text.y=element_blank(),
+          axis.ticks.y=element_blank()) + # no axis ticks 
+    theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
+    theme(axis.ticks.x = element_blank(),
+          axis.text.x = element_blank()) + # no x-axis labels to further reduce the gap between facets
+    #theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
+    theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+    #theme(axis.text.x = element_text(size=5)) +
+    theme(strip.background = element_blank(),
+          strip.text.x = element_blank())
+}
 
 
 ##----format-data

### 2021-11-02 main.Rmd
proof read done

@@
 <!--So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.--> In this work, instead of considering the probability distributions of the linear time series, we assume it across different categories of any cyclic granularity.
-We can consider categories of an individual cyclic granularity ($A$) or combination of categories for two interacting granularities ($A*B$) to have a distribution, where  $A, B$ represent two cyclic granularities defined by $A = \{a_j: j=1, 2, \dots J\}$ and  $B = \{b_k: k  = 1, 2, \dots K\}$.
-For example, let us consider two cyclic  granularities $hod$ and $dow$ representing hour-of-day and day-of-week and defined by, $hod = \{0, 1, 2, \dots, 23\}$ and $dow = \{Mon, Tue, Wed, \dots, Sun\}$.
-In case individual granularities ($A$ or $B$) are considered, there are  $J = 24$ distributions of the form $f_{i,j}(v)$ or $K = 7$ distributions of the form $f_{i,k}(v)$ for each customer $i$.
+We can consider categories of an individual cyclic granularity ($A$) or combination of categories for two interacting granularities ($A*B$) to have a distribution, where  $A, B$ are defined as $A = \{a_j: j=1, 2, \dots J\}$ and  $B = \{b_k: k  = 1, 2, \dots K\}$.
+For example, let us consider two cyclic granularities $A$ and $B$ representing hour-of-day and day-of-week.
+Then $A = \{0, 1, 2, \dots, 23\}$ and $B = \{Mon, Tue, Wed, \dots, Sun\}$.
+In case individual granularities are considered, there are  $J = 24$ distributions of the form $f_{i,j}(v)$ or $K = 7$ distributions of the form $f_{i,k}(v)$ for each customer $i$.
 In case of interaction, $J*K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
-Hence clustering these customers is equivalent to clustering these collections of conditional distributions.
+Hence clustering these customers is equivalent to clustering these collections of conditional univariate probability distributions.
 Towards this goal, we need to decide how to measure similarities between collections of univariate probability distributions.
@@
 <!--This paper focuses on looking at the (dis) similarity between underlying distributions that may have resulted in different patterns across different cyclic temporal granularities, that eventually have resulted in the (dis) similarity between time series.
---> This paper considers a methodology with two approaches for finding distances between time series.
+--> This paper considers a two approaches for finding distances between time series.
 Both of these approaches may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
@@
 We define "significant" granularities as those with significant distributional differences across categories.
-It is preferable to choose only those granularities since it is more probable that there will be some intriguing repeated behavior worth investigating.
+It is preferable to consider those since it is more probable that there will be some intriguing repeated behavior worth investigating.
 It should be noted that all of the observations in the study may not have the same set of important granularities.
@@
 As a result, while displaying transformed data using NQT, one must exercise caution.
-This, on the other hand, is a very useful strategy for making assymetric distributions symmetric, which often improves the clustering performance.
+But NQT is an useful transformation that often improves clustering performance.
+
+<!-- NQT, however, is an useful transformation which often improves the clustering performance.
+Since our variables are cyclic granularities which are derived from linear granularities, NQT ensures that the conditional distribution across each variable is also normalised.
+-->
 
@@
 We start with a "tsibble" (@wang2020tsibble) data structure with an index variable representing inherent ordering from past to present and a key variable that define observational units over time.
-The measured variable for observation is a time-indexed sequence of values.
+The measured variable for an observation is a time-indexed sequence of values.
 This sequence, however, could be shown in several ways.
 A shuffle of the raw sequence may represent hourly consumption throughout a day, a week, or a year.
-Cyclic granularities like hour-of-day, hour-of-week can be expressed in terms of the index set in the "tsibble" data structure.
-But the data structure changes while transporting from linear to cyclic scale of time as multiple observations are generated for each category and induces a probability distribution.
-Directly computing Jensen-Shannon distances between probability distributions can be very costly.
+Cyclic granularities can be expressed in terms of the index set in the "tsibble" data structure.
+But the data structure changes while transporting from linear to cyclic scale of time as multiple observations now correspond to each category and induces a probability distribution.
+Directly computing Jensen-Shannon distances between the entire probability distributions can be very costly.
 Hence, in this paper, quantiles are chosen to characterize the probability distributions.
-So, in the final data structure, each category of a cyclic granularity corresponds to a list of numbers which is essentially few chosen quantiles of the multiple observations.
+So, in the final data structure, each category of a cyclic granularity corresponds to a list of numbers which is essentially few chosen quantiles.
 
@@
 This agglomerative approach successively merges the pair of clusters with the shortest between-cluster distance using Ward's linkage method.
-Hierarchical algorithms are one of the most widely used, can operate with data of any shape, has reasonable scalability, and the number of clusters is not needed as a parameter.
+
+<!--Hierarchical algorithms are one of the most widely used, can operate with data of any shape, has reasonable scalability, and the number of clusters is not needed as a parameter.-->
 
@@
 
-To validate the clustering approaches, we spiked several attributes in the data to see where one method works better than the other and where they might give us the same outcome <!--or the effect of missing data <!---and trends--> on the proposed methods.
-Three circular granularities $g1$, $g2$ and $g3$ are considered with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $n_{g_1}=2$, $n_{g_2}=3$ and $n_{g_3}=5$.
+To validate our clustering methods, we spiked many attributes in the data to create different data designs.
+<!--to see where one method works better than the other and where they might give us the same outcome --or the effect of missing data <!---and trends on the proposed methods.--> Three circular granularities $g1$, $g2$ and $g3$ are considered with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $n_{g_1}=2$, $n_{g_2}=3$ and $n_{g_3}=5$.
 These categories could be integers or some more meaningful labels.
@@
 Variations in method performance may be due to different jumps between categories.
-So a mean difference of $\mu = \{1, 2, 5\}$ is examined.
+So a mean difference of $\mu = \{1, 2, 5\}$ between categories is considered.
 The performance of the approaches varies with the number of granularities which has interesting patterns across its categories.
-So scenarios with all, few, and one significant granularities are considered.
+So three scenarios are considered to accomodate that.
 
@@
 \noindent Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction.
-This means that at least one of the designs to be categorized will have significantly different patterns for each granularity.
-In Table \ref{tab:tab-dist-design} (top), we explore various distributions across categories (as shown in Table \ref{tab:tab-dist-design} (bottom).
+This means that each granularity will have significantly different patterns at least across one of the designs to be clustered.
+In Table \ref{tab:tab-dist-design} various distributions across categories are considered (top) which lead to different designs (bottom).
 Figure \ref{fig:plot-3gran-new} shows the simulated variable's linear (left) and cyclic (right) representations for each of these five designs.
@@
 - In Figure \ref{fig:mds-plot-validation}, we tried to see how separated our clusters are.
-We observe that in all scenarios and for different mean differences, cluster is separated.
+We observe that in all scenarios and for different mean differences, clusters are separated.
 However, the separation increases with an increase in mean differences across scenarios.
@@
 
-- Figure \ref{fig:parcoord-sim} depicts a parallel coordinate plot with the vertical bar showing total inter-cluster distances with regard to granularities $g1$, $g2$, and $g3$.
-For all simulation settings and scenarios, values are represented as a sequence of lines connected across each axis.
-One line in the figure, for example, shows the inter-cluster distances for one simulation setting.
+- Figure \ref{fig:parcoord-sim} depicts a parallel coordinate plot with the vertical bar showing total inter-cluster distances with regard to granularities $g1$, $g2$, and $g3$ for all simulation settings and scenarios.
+So one line in the figure shows the inter-cluster distances for one simulation setting and scenarios vary across facets.
 The lines are not colored by group since the purpose is to highlight the contribution of the factors to categorization rather than class separation.
@@
 Of course, certain customers in each group have distributions that differ from other members in the same group.
-However, it appears that the aim of grouping comparable distributions over considered variables has been accomplished to some extent.
+However, it appears that the aim of grouping comparable distributions across categories of multiple temporal dependencies has been accomplished to some extent.
 Figure \ref{fig:combined-groups-js} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
-It shows Groups $2$ and $5$ show a stronger hod pattern with a typical morning and evening peak, whereas groups 1, 3, and 5 show a moy pattern  with higher usage in winter months.
+It shows Groups $2$ and $4$ show a stronger hod pattern with a typical morning and evening peak, whereas groups 1, 3, and 5 show a moy pattern  with higher usage in winter months.
 Differences in Weekend/Weekday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters unless maybe conditioned by `moy` or `hod`.
@@
 The variables are sorted according to their separation across classes (rather than their overall variation between classes).
-This means that $moy$ is the most important variable in distinguishing the designs followed by $hod$ and $wnwd$.
+This means that $moy$ is the most important variable in distinguishing the groups followed by $hod$ and $wnwd$.
 There is only one customer who has significant $wpd$ across $wnwd$ and stands out from the rest of the customers.
@@
 \noindent
-**Supplementary paper:** Additional table, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd)
+**Supplementary paper:** Additional tables, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd)
 

### 2021-11-02 paper-supplementary.Rmd
proof read done

@@
 ```
+
+```{r data-heatmap-hod-group-entire-1col}
+
+hod_group_entire <- data_heatmap_hod_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=group,
+                  fill = as.factor(group), alpha = 0.5),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=group, 
+                color = as.factor(group)), size = 1)+
+  facet_wrap(~group, 
+             scales = "free_y",  
+             nrow = 17) + 
+              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
+  ylab("demand (in Kwh)") + 
+  theme_bw()  +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ 
+  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme_application() +
+  scale_fill_manual(values=as.vector(polychrome(20)))+
+  scale_color_manual(values=as.vector(polychrome(20)))+
+  theme(legend.position = "bottom")
+
+```
+
+```{r data-heatmap-moy-group-entire-1col}
+
+moy_group_entire <- data_heatmap_moy_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow = 17) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
+  ylab("demand (in Kwh)") +
+  theme_bw() + theme_application() +
+  scale_fill_manual(values=as.vector(polychrome(20)))+
+  scale_color_manual(values=as.vector(polychrome(20)))+
+  theme(legend.position = "bottom")
+```
+
+```{r data-wnwd-group-entire-1col}
+
+wkndwday_group_entire <- wkndwday_data%>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
+  geom_boxplot(aes(fill = group, color = group),alpha = 0.5, outlier.alpha = 0.05)+
+  #geom_boxplot(outlier.size = 1) + 
+  coord_cartesian(ylim = ylim1*1.05)+
+  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
+#geom_boxplot(aes(fill = as.factor(group))) +
+  #scale_fill_lv() +
+ xlab("wknd-wday") + 
+  ylab("demand (in Kwh)") +
+   facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_both",
+             nrow = 17) + 
+  theme_bw() + theme_application() +
+  scale_fill_manual(values=as.vector(polychrome(20)))+
+  scale_color_manual(values=as.vector(polychrome(20)))+
+  theme(legend.position = "none")
+```
+
+```{r combined-groups-js-entire-1col}
+
+ (hod_group_entire + moy_group_entire+ wkndwday_group_entire) +
+  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
+  plot_layout(guides = "collect")& theme(legend.position = 'none')
+```

### 2021-11-02 main.Rmd
some grammar and spelling fixes

@@
 - smart meter
-- electricity consumption behaviour
+- electricity consumption behavior
 - R
@@
 Looking at probability distributions across cyclic granularities serves two purposes: (a) "Probability distributions" characterise the inherent temporal data structure of these large unequal-length time series and are robust to missing or noisy data.
-(b) Using probability distributions over "cyclic granularities" ensures small pockets of similar "repeated" behaviours.
+(b) Using probability distributions over "cyclic granularities" ensures small pockets of similar "repeated" behaviors.
 Our method is capable of producing useful clusters, as demonstrated by testing on validation data designs and a sample of residential smart meter consumers.
@@
 @Melnykov2013-sp discusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
-In our knowledge, none of the methods focus on exploring heterogeneity in repetitive behaviors based on the dynamics of multiple temporal dependencies using probability distributions.
+To our knowledge, none of the methods focus on exploring heterogeneity in repetitive behaviors based on the dynamics of multiple temporal dependencies using probability distributions.
 
@@
 As a result, while displaying transformed data using NQT, one must exercise caution.
-But NQT is an useful transformation that often improves clustering performance.
+But NQT is a useful transformation that often improves clustering performance.
 
@@
 
-We start with a "tsibble" (@wang2020tsibble) data structure with an index variable representing inherent ordering from past to present and a key variable that define observational units over time.
+We start with a "tsibble" (@wang2020tsibble) data structure with an index variable representing inherent ordering from past to present and a key variable that defines observational units over time.
 The measured variable for an observation is a time-indexed sequence of values.
@@
 Cyclic granularities can be expressed in terms of the index set in the "tsibble" data structure.
-But the data structure changes while transporting from linear to cyclic scale of time as multiple observations now correspond to each category and induces a probability distribution.
+But the data structure changes while transporting from linear to cyclic scale of time as multiple observations now correspond to each category and induce a probability distribution.
 Directly computing Jensen-Shannon distances between the entire probability distributions can be very costly.
 Hence, in this paper, quantiles are chosen to characterize the probability distributions.
-So, in the final data structure, each category of a cyclic granularity corresponds to a list of numbers which is essentially few chosen quantiles.
+So, in the final data structure, each category of a cyclic granularity corresponds to a list of numbers which is essentially a few chosen quantiles.
 
@@
 With this approach, we should expect the observations with similar $wpd$ values to be clustered together.
-Thus, this approach is useful for grouping observations that have similar significance of patterns across different granularities.
+Thus, this approach is useful for grouping observations that have a similar significance of patterns across different granularities.
 Similar significance does not imply a similar pattern, which is where this technique varies from JS-based distances, which detect differences in patterns across categories.
@@
 With a way to obtain pairwise distances, any clustering algorithm can be employed that supports the given distance metric as input.
-A good comprehensive list of algorithms can be found in @Xu2015-ja based on traditional ways like partition, hierarchy or more recent approaches like distribution, density and others.
+A good comprehensive list of algorithms can be found in @Xu2015-ja based on traditional ways like partition, hierarchy, or more recent approaches like distribution, density, and others.
 We employ agglomerative hierarchical clustering in conjunction with Ward's linkage.
@@
 Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
-The t-SNE embedding will compress data points which are close in high-dimensional space.
+The t-SNE embedding will compress data points that are close in high-dimensional space.
 Tour is a collection of interpolated linear projections of multivariate data into lower-dimensional space.
@@
 We constructed independent replications of all data designs $R = \{25, 250, 500\}$ to investigate if our proposed clustering method can discover distinct designs in small, medium, and big number of series.
-All designs employ $T=\{300, 1000, 5000\}$ sample sizes to evaluate small, medium, and large sized series.
+All designs employ $T=\{300, 1000, 5000\}$ sample sizes to evaluate small, medium, and large-sized series.
 Variations in method performance may be due to different jumps between categories.
@@
 The performance of the approaches varies with the number of granularities which has interesting patterns across its categories.
-So three scenarios are considered to accomodate that.
+So three scenarios are considered to accommodate that.
 
@@
 Instance selection in unsupervised learning has received little attention in the literature, yet it could be a useful tool for evaluating model or method performance.
-There are several ways to approach the prototype selection.
+There are several ways to approach prototype selection.
 Following @Fan2021-bq's idea of picking related examples (neighbors) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2D space.
@@
 Tours can reveal variable separation that was hidden in a single variable display better than static projections.
-Hence we perform a linked tour with t-SNE layout using the R package `liminal` (@R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
+Hence we perform a linked tour with a t-SNE layout using the R package `liminal` (@R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
 (Refer to Supplementary article for further details).
@@
 In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.
-All customers with the same color represent same clustered groups.
+All customers with the same color represent the same clustered groups.
 Groups by JS-based distances and wpd-based distances are colored differently as they represent different groupings.
@@
 Our current approach can not detect these dynamic changes.
-Thirdly, it is possible that for few customers, data for some categories from the list of considered significant granularities are missing.
+Thirdly, it is possible that for a few customers, data for some categories from the list of considered significant granularities are missing.
 In our application, we have removed those customers and done the analysis but the metrics used should be able to incorporate those customers in the clustering by handling their missing categories.
@@
 Future work can make the computations more efficient so that they are easily scalable to a large number of customers.
-Moreoever, experiments can also be run with non-hierarchy based clustering algorithms to verify if these distances work better with other algorithms.
+Moreover, experiments can also be run with non-hierarchy based clustering algorithms to verify if these distances work better with other algorithms.
 

### 2021-11-02 main.Rmd
Do not embellish the plots mormore before Thursday

@@
 
+```{r generate-design-3change}
+
+```
+
 ```{r plot-3gran-new, out.width="100%", fig.cap="The linear (left) and cyclic (right) representation of the simulated variable is shown.
@@
 (Refer to Supplementary article for further details).
-Figure \ref{fig:hod-ind-group} (a, b, c) shows the raw time plot, distribution across `hod`, `moy` and `wnwd` for the set of chosen $24$ customers.
+Figure \ref{fig:hod-ind-group} shows the distribution across `hod` (a), `moy`(b) and `wnwd` (c) for the set of chosen $24$ customers that were chosen.
 Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution.
@@
 ```{r prototype-data-pick}
-quantile_prob_graph <- c(0.25, 0.5, 0.75)
-
-# data_pick_one <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
-# data_pick_two <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
-# data_pick_three <- c(8221762, 8273636, 10359424, 8232822)%>% as_tibble %>% set_names("customer_id")
-
-data_pick_one <- c(8541744, 9355808, 8603880, 8619309, 10542667) %>% as_tibble %>% set_names("customer_id")
-#data_pick_two <- c(8688242, 8643837, 8184707, 10534355, 8684420) %>% as_tibble%>% set_names("customer_id")
-data_pick_three <- c(9792072, 8589936, 8454235, 10692366, 8603828)%>% as_tibble%>% set_names("customer_id")
-data_pick_four <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
-data_pick_five <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
-data_pick_six <- c(8221762, 8273636, 10359424, 8232822, 11450499)%>% as_tibble %>% set_names("customer_id")
 
@@
 ```{r assemble}
-data_pick_cust <- bind_rows(
-data_pick_one, 
-#data_pick_two, 
-data_pick_three,
-data_pick_four,
-data_pick_five, 
-data_pick_six,
-.id = "design") %>% 
- mutate(customer_id = as.character(customer_id))
 
@@
 ```{r data-pick}
-data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
-  mutate(customer_id = as.character(customer_id)) %>% 
-  dplyr::filter(customer_id %in% data_pick_cust$customer_id) %>% 
-  gracsr::scale_gran( method = "robust",
-                      response = "general_supply_kwh")
 ```
@@
 Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most households, there is one household for whom weekend behavior is standing out from the rest."}
-
-data_hod <- quantile_gran(data_pick,
-                                  "hour_day",
-                                  quantile_prob_val = quantile_prob_graph) %>% 
-  pivot_wider(names_from = quantiles,
-              values_from = quantiles_values) %>% 
-  left_join(data_pick_cust, by = c("customer_id"))
-
-#data_heatmap_hod$customer_id = factor(data_heatmap_hod$customer_id, levels = data_pick_cust$value)
-data_hod$category <- factor(data_hod$category, levels = 0:23)
-
-hod_ind_design <- data_hod %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`,
-                  group=customer_id),
-              alpha = 0.5) +
-  geom_line(aes(y = `50%`,
-                group=customer_id), 
-            size = 1) +
-  facet_wrap(~customer_id, 
-             scales = "free_y",
-             nrow = 24) + 
- theme_application() +
-  xlab("hod") +
-  scale_fill_viridis_d()+
-  scale_color_viridis_d()+
-  scale_x_discrete(breaks = seq(0, 23, 3))
-
-data_moy <- quantile_gran(data_pick,
-                                  "month_year", 
-                                  quantile_prob_val = quantile_prob_graph) %>% 
-  pivot_wider(names_from = quantiles, 
-              values_from = quantiles_values) %>% 
-  left_join(data_pick_cust, by = c("customer_id"))
-
-data_moy$category <- factor(data_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
-
-moy_ind_design <- data_moy %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, 
-                  group=customer_id), 
-              alpha = 0.5) +
-  geom_line(aes(y = `50%`, group=customer_id), size = 1) +
-  facet_wrap(~customer_id, 
-             scales = "free_y", 
-             nrow = 24) +
-    ylab("demand (in Kwh)") +
-    xlab("moy")  +
-  theme_application()
-
-data_wkndwday <- data_pick  %>%
-   create_gran("wknd_wday")  %>% 
-  left_join(data_pick_cust, by = c("customer_id"))
-
-ylim1 = boxplot.stats(data_wkndwday$general_supply_kwh)$stats[c(1, 5)]
-
-wkndwday_ind_design <- data_wkndwday%>% 
-  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
-  #lvplot::geom_lv(aes(fill = as.factor(design), 
-   #                   color = as.factor(design)), k=5, alpha = 0.5) +
-  geom_boxplot(alpha = 0.5, fill = "black")+
-  #geom_boxplot(outlier.shape = NA) + 
-  coord_cartesian(ylim = ylim1*1.05)+
-  facet_wrap(~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-              nrow = 24)  +
-  stat_summary(
-    fun = median,
-    geom = 'line',
-    aes(group = 1), size = 1, color = "black")+
-  xlab("wnwd")+
-  theme_application()
-
 ```
@@
 For clustering based on JS-based distances, we chose the optimal number of clusters using (@Hennig2014-ah) as $5$.
-The distribution of electricity demand for the selected $24$ customers across `hod`, `moy` and `wnwd` are shown in Figure \ref{fig:hod-ind-group} (d, e, f).
-Our methodology is useful for grouping similar distributions over hod and moy and they are placed closely for easy comparison.
-Of course, certain customers in each group have distributions that differ from other members in the same group.
-However, it appears that the aim of grouping comparable distributions across categories of multiple temporal dependencies has been accomplished to some extent.
+The groupings are shown in Figure \ref{fig:hod-ind-group}.
+Customers with the same color represent the same clustered groups.
+Our methodology is useful for grouping similar distributions over `hod` and `moy` and they are placed closely for easy comparison.
+Few groups have mixed patterns across `hod` and `moy`, but few have all customers in the group having a similar profile.
 Figure \ref{fig:combined-groups-js} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
-It shows Groups $2$ and $4$ show a stronger hod pattern with a typical morning and evening peak, whereas groups 1, 3, and 5 show a moy pattern with higher usage in winter months.
+It shows Groups $2$ and $4$ have `hod` pattern with a typical morning and evening peak, whereas groups $1$, $3$, and $5$ show a `moy` pattern with higher usage in winter months.
 Differences in Weekend/Weekday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters unless maybe conditioned by `moy` or `hod`.
+It may be interesting to compare these two plots to verify if the summarized distributions across groups correctly characterised the groupings.
+If it has, then the majority of the group's members should share a similar profile.
+
+<!--  It appears that the aim of grouping comparable distributions across categories of multiple temporal dependencies has been accomplished to some extent. -->
 
 ```{r all-data}
-data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
-  mutate(customer_id = as.character(customer_id)) %>% 
-  dplyr::filter(customer_id %in% data_pick_cust$customer_id) %>% 
-  gracsr::scale_gran( method = "nqt",
-                      response = "general_supply_kwh")
 ```
@@
 ```{r clustering}
-hod <- suppressMessages(data_pick %>% 
-  dist_gran(gran1 = "hour_day", response = "general_supply_kwh"))
-
-moy <- suppressMessages(data_pick %>% 
-  dist_gran(gran1 = "month_year", response = "general_supply_kwh"))
-
-wkndwday <- suppressMessages(data_pick %>% 
-  dist_gran(gran1 = "wknd_wday", response = "general_supply_kwh"))
-
-distance <- wkndwday/2 + moy/12 + hod/24
-
-f = as.dist(distance)
 
@@
 ```{r opt-clusters}
-library(fpc)
-library(cluster)
-k = array()
-for(i in 2:20)
-{
-group <- f %>% hclust (method = "ward.D") %>% cutree(k=i)
-p <- cluster.stats(f, clustering = group, silhouette = TRUE)
-k[i]=p$sindex
-}
-# 
-# ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 20, 1))
 ```
@@
 ```{r groups-24}
-cluster_result <- suppressMessages(f %>% 
-  clust_gran(kopt = 5)) %>% 
-  rename("customer_id" = "id") %>% 
-  mutate(group = as.factor(group))
-```
-
-```{r hod-ind-group, fig.cap="The distribution of selected consumers over hod (a, d), moy (b, e), and wkndwday (d, f).
-In each case, the same colour denotes the same group in plots (d), (e), (f) and are placed together to facilitate comparison.
-That means the customer orderings are different for (a, b, c) and (d, e, f).
-Our clustering methodology is useful for grouping similar distributions over hod and moy.
-Of course, certain customers in each group have distributions that differ from those of other members in the same group.
-However, it appears that the aim of grouping comparable distributions over considered variables has been accomplished to some extent." }
-
-hod_ind_group <- data_hod %>% 
-  left_join(cluster_result, by = c("customer_id")) %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`,
-                  group=customer_id, fill = group),
-              alpha = 0.5) +
-  geom_line(aes(y = `50%`,
-                group=customer_id, 
-                color = group), size = 1) +
-  facet_wrap(group~customer_id, 
-             scales = "free_y",
-              nrow=24) + 
-  theme_application()+
-  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +    xlab("hod")  +
-  scale_x_discrete(breaks = seq(0, 23, 3))+ theme(legend.position = "bottom")
-# + scale_y_continuous(breaks = NULL)
-
-moy_ind_group <- data_moy %>% 
-  left_join(cluster_result, by = c("customer_id")) %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, 
-                  group=customer_id, fill = group), alpha = 0.5) +
-  geom_line(aes(y = `50%`, group=customer_id, color = group), size = 1) +
-  facet_wrap(group~customer_id, 
-             scales = "free_y", 
-             nrow=24) +
-    ylab("demand (in Kwh)") +
-    xlab("moy")  +
-  theme_application() +
-  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) + theme(legend.position = "bottom")
-
-wkndwday_ind_group <- data_wkndwday%>% 
-  left_join(cluster_result, by = c("customer_id")) %>% 
-  mutate(group  = as.factor(group)) %>% 
-  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
-  #lvplot::geom_lv(aes(fill = as.factor(group), 
-          #            color = as.factor(group)), k=5, alpha = 0.5) +
-  geom_boxplot(aes(color = group, fill = group),alpha = 0.5)+
-  #geom_boxplot(outlier.shape = NA) + 
-  coord_cartesian(ylim = ylim1*1.05)+
-  facet_wrap(group~customer_id, 
-             scales = "free_y", 
-             labeller = "label_value",
-             nrow=24)  +
-  theme_application()+
-  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7"))  + theme(legend.position = "none")+
-  xlab("wnwd")
-
-hod_ind_design + moy_ind_design + wkndwday_ind_design + hod_ind_group + moy_ind_group + wkndwday_ind_group +  plot_layout(guides = "collect", ncol = 6) + plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')') & theme(legend.position = 'none')
-```
+```
+
+```{r hod-ind-group, fig.cap = "The distribution of selected customers over hod (a), moy (b), and wnwd (c) for the $24$ selected customers is shown.
+In each case, the same colour denotes the same group using js-based clustering and are placed together to facilitate comparison.
+Individuals with similar distributional differences across categories of `hod` or `moy` are grouped together, with some exceptions.
+" }
+```
+
+<!-- It appears that the aim of grouping comparable distributions over considered variables has been accomplished to some extent. -->
 
 ```{r data-heatmap-hod-group}
-legend_title <- "group"
-
-data_group <- data_pick  %>% 
-  left_join(cluster_result, by = c("customer_id"))
-
-data_heatmap_hod_group <- quantile_gran(data_group,
-                                  gran1="hour_day",
-                                  quantile_prob_val = c(0.25, 0.5, 0.75),
-                                  group="group") %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values)
-
-data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
-
-data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")
-
-hod_group <- data_heatmap_hod_group %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`,
-                  group=group,
-                  fill = as.factor(group), alpha = 0.5),
-              alpha = 0.5) +
-  geom_line(aes(y = `50%`,
-                group=group, 
-                color = as.factor(group)), size = 1)+
-  facet_wrap(~group, 
-             scales = "free_y",  
-             nrow = 5) + 
-              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
-  ylab("demand (in Kwh)") + 
-  theme_bw()  +
-  scale_x_discrete(breaks = seq(1, 24, 3))+ 
-  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme_application() +
-  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +
-  theme(legend.position = "bottom")
-
 ```
@@
 ```{r data-heatmap-moy-group}
-data_heatmap_moy_group <- quantile_gran(data_group,
-                                  gran1="month_year",
-                                  quantile_prob_val = c(0.25, 0.5, 0.75),
-                                  group="group") %>% 
-  pivot_wider(names_from = quantiles, values_from = quantiles_values)
-
-data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
-
-data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")
-
-moy_group <- data_heatmap_moy_group %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
-  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
-  facet_wrap(~group, 
-             scales = "free_y", 
-             labeller = "label_value",
-             nrow = 5) +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
-  ylab("demand (in Kwh)") +
-  theme_bw() + theme_application() +
-  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +
-  theme(legend.position = "bottom") 
-```
-
-```{r}
-wkndwday_data <- data_group %>% create_gran("wknd_wday") %>% 
-  create_gran("hour_day")
-
-ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]
-
-wkndwday_group <- wkndwday_data%>% 
-  ggplot(aes(x=hour_day, y = general_supply_kwh)) +
-  #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
-  geom_boxplot(aes(fill = group, color = group),alpha = 0.5, outlier.alpha = 0.05)+
-  #geom_boxplot(outlier.size = 1) + 
-  coord_cartesian(ylim = ylim1*1.05)+
-  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
-#geom_boxplot(aes(fill = as.factor(group))) +
-  #scale_fill_lv() +
- xlab("wknd-wday") + 
-  ylab("demand (in Kwh)") +
-   facet_grid(group~wknd_wday, 
-             scales = "free_y", 
-             labeller = "label_both") + 
-  theme_bw() + theme_application() +
-  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
-  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +
-  theme(legend.position = "none") 
-```
-
-```{r combined-groups-js, fig.cap = "The distribution of electricity demand for the clusters across hod (a), moy (b) and wkndwday (c).
-It seems like group 2 and 5 have a hod pattern across its members, while group 1, 3, 5 have a moy pattern.
-Weekend/Weekday variations across groups are not distinguishable, indicating that it is not a critical variable for clustering.
-It is helpful to compare the summarised distributions of groups to that of individuals to confirm that the most of individuals in the group have the same characterisation."}
-
- (hod_group + moy_group + wkndwday_group) +
-  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
-  plot_layout(guides = "collect")& theme(legend.position = 'bottom')
+
+```
+
+```{r data-heatmap-wkndwday-group}
+
+```
+
+```{r combined-groups-js, fig.cap = "The distribution of electricity demand over hod (a), moy (b), and wnwd*hod (c).
+Groups 2 and 5 appear to have a hod pattern among its members, whereas groups 1, 3, and 5 appear to have a moy pattern.
+Weekend/weekday differences across groups are not discernible, showing that it is not a critical variable for clustering on its own.
+It is probable that the distribution over Weekend/Weekday will differ conditional on the categories of hod or moy.
+(c) demonstrates that the distribution is not different, except for the tails.
+It is useful to compare the summarised distributions of groups to those of individuals in order to establish that the majority of individuals in the group share the same characteristics."}
 ```
@@
 ```{r data-pick-wpd}
-elec_600_wpd <- read_rds(here::here("data/algo2-cust600-wpd-rawdata.rds"))
-
-elec_pick <- elec_600_wpd %>% 
-  filter(customer_id %in% data_pick_cust$customer_id)
-
-elec_pick_wide <- elec_pick %>% pivot_wider(-c(1, 2), names_from = "x_variable", values_from = wpd)
-
-scaled_var <- elec_pick_wide
-
-f <- elec_pick_wide[-1] %>% dist()
-
-k = array()
-for(i in 2:20)
-{
-group <- f %>% hclust (method = "ward.D") %>% cutree(k=i)
-p <- cluster.stats(f, clustering = group, silhouette = TRUE)
-k[i]=p$sindex
-}
-# ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 20, 1))
-
-group <- f%>% hclust (method = "ward.D") %>% cutree(k=3)
-
-cluster_result_wpd <- bind_cols(id = elec_pick_wide$customer_id, group = group)
-
-data_pcp <- scaled_var %>% 
-  #bind_cols(customer_id =  elec_pick_wide$customer_id) %>%
-  left_join(cluster_result_wpd , by = c("customer_id" = "id")) %>% 
-  select(customer_id, group, everything()) %>% 
-  mutate(group = as.factor(group))
-
-data_table <- data_pcp %>% group_by(group) %>% 
-  summarise(nobs = n(),
-            moy = round(median(month_year),2),
-            hod = round(median(hour_day),2),
-            wnwd = round(median(wknd_wday),2)) %>% 
-  select(-group)
-
-rownames(data_table) <- c("group-1", "group-2", "group-3")
-
 ```
@@
 All of these could be observed from the plot or the table (b) which shows median wpd values for each group.", message=FALSE, warning=FALSE}
-
-parcoord <- GGally::ggparcoord(data_pcp ,
-                   columns = 3:ncol(data_pcp),
-                   groupColumn = "group",
-                   showPoints = FALSE, 
-                   alphaLines = 0.8,
-                   order = "anyClass",
-                   scale = "globalminmax"
-) + 
-  ggplot2::theme(
-    plot.title = ggplot2::element_text(size=10)
-  )+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
-  theme(legend.position = "bottom") +
-  xlab("") +
-  ylab("wpd") + scale_fill_viridis_d(direction = 1) +
-  scale_color_viridis_d(direction = 1) + theme_light()
-
-(parcoord + gridExtra::tableGrob(data_table))+ plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')') & theme(legend.position = "bottom")
 

### 2021-11-02 script/main.R

@@ -78,7 +78,11 @@ theme_application <- function() {
     theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
     #theme(axis.text.x = element_text(size=5)) +
     theme(strip.background = element_blank(),
-          strip.text.x = element_blank())
+          strip.text.x = element_blank()) +
+    theme(
+            panel.grid.major.y = element_blank(),
+            panel.grid.minor.y = element_blank()
+          )
 }
 
 
@@ -1067,4 +1071,378 @@ mds_plot_10 <- ggplot(all_data,
   ylab("mds1") + xlab("mds2")
 
 mds_plot_10
+
+##### application
+##----prototype-data-pick
+quantile_prob_graph <- c(0.25, 0.5, 0.75)
+
+# data_pick_one <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
+# data_pick_two <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
+# data_pick_three <- c(8221762, 8273636, 10359424, 8232822)%>% as_tibble %>% set_names("customer_id")
+
+
+data_pick_one <- c(8541744, 9355808, 8603880, 8619309, 10542667) %>% as_tibble %>% set_names("customer_id")
+#data_pick_two <- c(8688242, 8643837, 8184707, 10534355, 8684420) %>% as_tibble%>% set_names("customer_id")
+data_pick_three <- c(9792072, 8589936, 8454235, 10692366, 8603828)%>% as_tibble%>% set_names("customer_id")
+data_pick_four <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
+data_pick_five <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
+data_pick_six <- c(8221762, 8273636, 10359424, 8232822, 11450499)%>% as_tibble %>% set_names("customer_id")
+
+
+##----assemble
+data_pick_cust <- bind_rows(
+data_pick_one, 
+#data_pick_two, 
+data_pick_three,
+data_pick_four,
+data_pick_five, 
+data_pick_six,
+.id = "design") %>% 
+  mutate(customer_id = as.character(customer_id))
+
+##----data-pick
+data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
+  mutate(customer_id = as.character(customer_id)) %>% 
[…]

### 2021-11-03 main.Rmd
add changes minor

@@
 In this scenario, all of $g1$, $g2$ and $g3$ changes across at least one design.
-Also, it is not possible to comprehend these differences in patterns just by looking at or considering the linear representation."}
+Also, it is not possible to comprehend these distributional differences in patterns just by looking at or considering the linear representation."}
 
@@
 
-```{r gran2and1-clubbed, fig.cap=" Plots (b) and (c) correspond to Design (b) and (c) respectively.
+```{r gran2and1-clubbed, fig.cap=" Plots (b) and (c) correspond to Scenarios (b) and (c) respectively.
 In (b) $g2$, $g3$ changes across atleast one design but $g1$ remains constant.
@@
 One line in the figure depicts the inter-cluster distances for a single simulation scenario.
-While the first plot indicates that no variable stands out during clustering, the next two designs demonstrate that {g1} and {g1, g2} have extremely low inter-cluster distances, indicating that they did not contribute to clustering.
+While the first plot indicates that no variable stands out during clustering, the next two designs demonstrate that g1 and (g1, g2) have relatively lower inter-cluster distances, indicating that they did not contribute to clustering.
 It is worth emphasising that these facts are consistent with our initial assumptions when designing the scenarios and (a), (b), and (c) correspond to Scenario (a), (b) and (c) respectively."}
@@
 ```{r mds-plot-validation, fig.cap = "Relative positions of clusters corresponding to different scenarios (columns) for different values of mean differences between categories (rows) are shown using the first two dimensions of MDS.
-It can be observed that clusters become more compact and separated for higher mean differences between categories across all designs.
-Between designs, separation is least prominent corresponding to scenario (c) where only granularity is responsible for the clusters."}
+It can be observed that clusters become more compact and separated for higher mean differences between categories across all scenarios.
+Between scenarios, separation is least prominent corresponding to scenario (c) where only one granularity is responsible for distinguishing the clusters."}
 
@@
 In each case, the same colour denotes the same group using js-based clustering and are placed together to facilitate comparison.
-Individuals with similar distributional differences across categories of `hod` or `moy` are grouped together, with some exceptions.
-" }
+Individuals with similar distributional differences across categories of hod or moy are grouped together, with some exceptions.
+Distributional differences across categories of wnwday are not discernable except for a couple of customers." }
 ```
@@
 
-```{r combined-groups-js, fig.cap = "The distribution of electricity demand over hod (a), moy (b), and wnwd*hod (c).
-Groups 2 and 5 appear to have a hod pattern among its members, whereas groups 1, 3, and 5 appear to have a moy pattern.
-Weekend/weekday differences across groups are not discernible, showing that it is not a critical variable for clustering on its own.
-It is probable that the distribution over Weekend/Weekday will differ conditional on the categories of hod or moy.
-(c) demonstrates that the distribution is not different, except for the tails.
+```{r combined-groups-js, fig.cap = "The distribution of electricity demand over hod (a), moy (b), and hod (conditional on wnwd) (c).
+Groups 2 and 5 appear to have a hod pattern among its members (morning and evening peak), whereas groups 1, 3, and 5 appear to have a moy pattern (higher usage in mid-year).
+With their conditional dependencies on hod or moy categories, wnwd differences may not be discernible on their own.
+(c) demonstrates that the distribution of hod conditional on wnwd is also not different, except for the tails.
+This might be because most of these chosen customers do not have a wnwd pattern.
 It is useful to compare the summarised distributions of groups to those of individuals in order to establish that the majority of individuals in the group share the same characteristics."}

### 2021-11-03 paper-supplementary.Rmd
add changes minor

--- 2021-11-02 paper-supplementary.Rmd
@@
 ---
+
+```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
+# knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
+# Make sure you have the latest version of rmarkdown and bookdown
+#devtools::install_github("rstudio/rmarkdown")
+#devtools::install_github("rstudio/bookdown")
+options("knitr.graphics.auto_pdf" = TRUE,
+    tinytex.verbose = FALSE)
+library(knitr)
+opts_chunk$set(
+ echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
+ fig.path = "figure/", fig.align = "center", fig.show = "hold",
+ cache = TRUE, cache.path = "cache/",
+ out.width = ifelse(is_html_output(), "100%", "\\textwidth")
+)
+knitr::opts_knit$set(root.dir = here::here())
+```
 
@@
 
-```{r combined-groups-js-entire, fig.cap = "The distribution of electricity demand for the clusters across hod (a), moy (b) and wkndwday (c).
+```{r combined-groups-js-entire, eval = FALSE, fig.cap = "The distribution of electricity demand for the clusters across hod (a), moy (b) and wkndwday (c).
 It seems like group 2 and 5 have a hod pattern across its members, while group 1, 3, 5 have a moy pattern.
@@
 
-# contribution
+# contribution of hod
 
@@
   ylab("wpd")+
-  scale_fill_manual(values=as.vector(polychrome(20)))+
+  scale_color_manual(values=as.vector(polychrome(20)))+
   theme_bw()
@@
 ```
+
+# contribution of moy
+
+```{r contribution-moy}
+
+gran_cat_hod <-   gran_cat_dist %>% 
+  filter(gran == "moy") %>% 
+  #mutate(category = as.integer(category)) %>% 
+  arrange(group_from, category)
+
+# gran_cat_hod$category <- factor(gran_cat_hod$category, levels = 0:23)
+
+gran_cat_hod$category <- factor(gran_cat_hod$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+# for each group, percent contribution of hours
+
+group_total <- gran_cat_hod %>% 
+  arrange(category) %>% 
+  ungroup() %>% 
+  group_by(group_from) %>% 
+  summarise(group_total = sum(distance))
+
+data_pcp <- gran_cat_hod %>% 
+  arrange(category) %>% 
+  ungroup %>% 
+  left_join(group_total, by = "group_from") %>% 
+  mutate(cat_contibution = distance*100/group_total) %>% 
+  group_by(group_from) %>% 
+  #arrange(-distance) %>% 
+  select(group_from, category, cat_contibution) %>% 
+  pivot_wider(names_from = "category", 
+              values_from = "cat_contibution") %>% mutate(group_from = as.factor(group_from))
+
+parcoord <- GGally::ggparcoord(data_pcp ,
+                               columns = 2:ncol(data_pcp),
+                               groupColumn = "group_from",
+                               showPoints = TRUE, 
+                               alphaLines = 1,
+                               scale = "globalminmax"
+) + 
+  ggplot2::theme(
+    plot.title = ggplot2::element_text(size=10)
+  )+
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
+  theme(legend.position = "bottom") +
+  xlab("") +
+  ylab("wpd")+
+  scale_color_manual(values=as.vector(polychrome(20)))+
+  theme_bw()
+parcoord 
+```
+
+```{r data-heatmap-hod-group-entire-1col}
+
+hod_group_entire <- data_heatmap_hod_group %>% 
+  filter(group!="group-NA") %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=group,
+                  fill = as.factor(group), alpha = 0.5),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=group, 
+                color = as.factor(group)), size = 1)+
+  facet_wrap(~group, 
+             scales = "free_y",  
+             nrow = 8) + 
+              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
+  ylab("demand (in Kwh)") + 
+  theme_bw()  +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ 
+  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme_application() +
+  scale_fill_manual(values=as.vector(polychrome(20)))+
+  scale_color_manual(values=as.vector(polychrome(20)))+
+  theme(legend.position = "bottom")
+
+```
+
+```{r data-heatmap-moy-group-entire-1col}
+
+moy_group_entire <- data_heatmap_moy_group %>% 
+  filter(group!="group-NA") %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_value",
+             nrow = 8) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
+  ylab("demand (in Kwh)") +
+  theme_bw() + theme_application() +
+  scale_fill_manual(values=as.vector(polychrome(20)))+
+  scale_color_manual(values=as.vector(polychrome(20)))+
+  theme(legend.position = "bottom")
+```
+
+```{r data-wnwd-group-entire-1col}
+
+wkndwday_group_entire <- wkndwday_data%>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
+  geom_boxplot(aes(fill = group, color = group),alpha = 0.5, outlier.alpha = 0.05)+
+  #geom_boxplot(outlier.size = 1) + 
+  coord_cartesian(ylim = ylim1*1.05)+
+  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
+#geom_boxplot(aes(fill = as.factor(group))) +
+  #scale_fill_lv() +
+ xlab("wknd-wday") + 
+  ylab("demand (in Kwh)") +
+   facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_both",
+             nrow = 8) + 
+  theme_bw() + theme_application() +
+  scale_fill_manual(values=as.vector(polychrome(20)))+
+  scale_color_manual(values=as.vector(polychrome(20)))+
+  theme(legend.position = "none")
+```
+
+```{r combined-groups-js-entire-1col}
+
+ (hod_group_entire + moy_group_entire+ wkndwday_group_entire) +
+  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
+  plot_layout(guides = "collect")& theme(legend.position = 'none')
+```

### 2021-11-07 main.Rmd
files upload

@@
   thanks: "Email: Sayani.Gupta@monash.edu"
-
+- name: Dianne Cook
+  affiliation: Department of Econometrics and Business Statistics, Monash University
 - name: Rob J Hyndman
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-
-- name: Dianne Cook
   affiliation: Department of Econometrics and Business Statistics, Monash University

### 2021-11-07 paper-supplementary.Rmd
files upload

@@
 
-gran_cat_dist <- data_validation_cat %>% 
+gran_cat_dist_from <- data_validation_cat %>% 
   mutate(customer_from = as.character(customer_from),
@@
   select(-(4:8)) %>% 
-  arrange(-distance)
+  arrange(-distance) %>% 
+    rename("group"="group_from")
+
+gran_cat_dist_to <- data_validation_cat %>% 
+  mutate(customer_from = as.character(customer_from),
+         customer_to = as.character(customer_to)) %>% 
+  left_join(cluster_result, by = c("customer_from" = "customer_id")) %>% 
+  rename("group_from" = "group") %>% 
+  left_join(cluster_result, by = c("customer_to" = "customer_id")) %>%  
+  rename("group_to" = "group") %>% 
+  group_by(gran,
+           category,
+           group_from, 
+           group_to) %>% 
+  summarise(sum = sum(distance)) %>% 
+  pivot_wider(names_from = group_from, values_from = sum) %>% 
+  mutate(distance = sum(`1`, `2`, `3`, `4`, `5`,`6`, `7`, `8`, `9`, `10`, `11`, `12`, `13`, `14`, `15`, `16`, `17`)) %>% 
+  select(-(4:8)) %>% 
+  arrange(-distance) %>% 
+    rename("group"="group_to")
 

### 2021-11-07 supplementary/algorithm1/real-data/cluster4-iter100.Rmd
files upload

@@
 library(knitr)
-read_chunk("R/simdata01.R")
+library(here)
+read_chunk(here::here("validation/old/simdata01.R"))
 ```

### 2021-11-07 supplementary/old/analysis.Rmd
files upload

---
title: "analysis"
author: "Sayani Gupta"
date: "05/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-lib}
library(readr)
library(visdat)
library(ggplot2)
library(tidyverse)
library(naniar)
```

```{r load-data}
load("data/ALL_DATA.Rdata")
```

```{r miss-data}
empty_as_na <- function(x){
    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
    ifelse(as.character(x)!="", x, NA)
}

## transform all columns
CustomerData <- CustomerData %>% mutate_each(funs(empty_as_na))

vis_miss(CustomerData, warn_large_data = FALSE) + 
  theme(axis.text.x = element_text(angle = 90, size = 5))
```

```{r miss-var}
gg_miss_var(CustomerData) + theme_bw() +
  theme(axis.text.y = element_text(size = 4))
```

### 2021-11-07 supplementary/old/cluster4-iter500.Rmd
files upload

--- 2021-07-08 supplementary/cluster4designs.Rmd
@@
 ---
-title: "Clustering four designs"
+title: "Clustering four designs 500 iter"
 output:
@@
 ```{r, echo=TRUE}
-sample_seed <-  seq(10, 100, 10)
+sample_seed <-  seq(1, 500, 1)
 ```

### 2021-11-07 supplementary/old/introduction.Rmd
files upload

---
title: "introduction_section"
author: "Sayani Gupta"
date: "08/09/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Paper: Clustering of residential electricity customers using load time series

- Time feature extraction is a potential remedy, however, it is limited by the type of noisy, patchy, and unequal time-series common in residential datasets.

 - In this paper we propose a strategy to alleviate these limitations by
converting any types of load time series into map models that can be readily clustered.

- higher cluster distinction and robustness

- The experiment with 12 clusters results in around 61% distinction, improved coincidence
factor by around 6.75% relative to a random grouping, and robustness of around 59% against the applied noise.

- stochastic and not deterministic segmentation

- The processes of stacking daily
curves and applying PCA to extract core features from the pile of
curves, do not suit parallel computing. This makes the algorithm unsuitable for a large number of customers.

-  Alternatively, Shannonian entropy [62,63], can be used to optimise the embedding
parameters.

- feature based sensitive to outlier days

- The method feature based  is static in that any
addition of new data (of additional customers and or additional days)
require repeating both feature extraction and clustering

- model based : Regardless of its length, dates and times, each
continuous series translates to a map function or yields a geometry
 Alternatively,
Shannonian entropy [62,63], can be used to optimise the embedding
parameters.

 A distinction metric is the silhouette width [65,50], which measures how
similar homes are to their own cluster members (cohesion) compared
with homes in other clusters (separation), and is bounded within (−1,
1).

This is a desirable outcome; and clustering methods which are
able to isolate outliers are found to exhibit high performance as indicated in

Though, with a finite set of clusters the outliers stand
out, forcing the rest of the samples to lump together in big clusters with
high variability and vague profiles. To remedy this situation, we can
exclude outliers and re-cluster samples.

To remedy this situation, we can
exclude outliers and re-cluster samples. Though exclusion of outliers
often gives way to new outliers to emerge. Alternatively, we reduce
clusters to between 8 and 12 to retain all samples. The optima are
obtained at 9 and 12 clusters with ∼62.4% and ∼61.1% distinctions,
respectively.

There are at least two key areas to extend this research. These include load prediction based on clustering repeatability, and customer
privacy protection by noise encryption. The current repeatability metric
determines if the same clusters of homes are formed by clustering over
successive periods of time, such as successive seasons, or years. To reinforce repeatability, future work can also focus on cross validation by
bootstrapping, i.e., the same clusters of homes should be formed based
on randomly selected days within a certain period.

### 2021-11-07 supplementary/old/approaches.R

@@ -1,0 +1,249 @@
+# harmony table considered for all
+
+#This script calculates raw mmpd for each simulation scenario
+##Read Simulation Table
+
+.libPaths(c("~/R/libs", .libPaths()))
+library(distributional)
+library(readr)
+library(drake)
+library(tidyverse)
+library(hakear)
+library(gravitas)
+library(parallel)
+library(here)
+library(lubridate)
+library(tsibble)
+library(patchwork)
+
+set.seed(9999)
+
+# parameter set
+nsamp = 100 # number of samples based on which thresholds are decided
+ncust = 8 # number of customers for trial
+nrep = 3 # number of error added usage for each customer
+kstep1 = 3 # number of clusters in step1/ only step
+kstep2 = 2 # number of clusters in step2 (this will actually depend on the number of observations in the cluster after step1)
+
+# change path while running it on HPC
+simtable<-read_csv(here('../gracsR/sim_table/sim_table.csv'))
+
+
+# <needed>
+harmonies<-read_csv(here('../gracsR/sim_table/harmonies.csv'))
+
+elec <- read_rds(here("data/elec_all-8.rds")) %>% 
+  dplyr::filter(date(reading_datetime) >= ymd("20190701"), date(reading_datetime) < ymd("20191231"), meter_id==1) %>% 
+  select(-meter_id) %>% 
+  rename("customer_id" = "household_id",
+         "general_supply_kwh" = "kwh") %>% 
+  as_tsibble(key = customer_id, index = reading_datetime)
+
+
+sm_cust_data <- elec
+
+sm_cust_data_list <- sm_cust_data %>% group_split(customer_id)
+
+
+### Extract flags from simulation scenario
+# Run wpd_threshold on all of them, each household can form a separate seed on cluster
+
+all_house_nrep_orig_list_5 <-  sm_cust_data_list
+
+all_house_harmony_wpd <- mclapply(all_house_nrep_orig_list_5,
+                                  function(x)
+                                    wpd = wpd_threshold(x,
+                                                  harmony_tbl = harmonies,
+                                                  response = general_supply_kwh, nsamp = nsamp)) %>%
+  bind_rows(.id = "customer_id")
+
+# Find harmonies that are not significant for all, whichever columns are integer in the following table drop them
+
+
+significant_tag <- function(x){
+  !all(is.na(str_extract(x, "\\*")))
+}
+
+
+m <- all_house_harmony_wpd %>%
+  mutate(harmony_name = paste(facet_variable, x_variable, sep = "-")) %>%
[…]

### 2021-11-07 supplementary/old/check.R

@@ -1,0 +1,171 @@
+library(conflicted)
+library(tidyverse)
+library(gracsr)
+library(tsibble)
+library(gravitas)
+library(dplyr)
+
+generate_design <- function(n = 300, #length of time series
+         mu11 = 0,
+         mu12 = 0,
+         mu21 = 0,
+         mu22 = 0 ,
+         mu23 = 0,
+         mu31 = 0,
+         mu32 = 0,
+         mu33 = 0,
+         mu34 = 0,
+         mu35 = 0)
+{
+  
+t <- seq(0, n-1, 1)
+
+g1 <- t %%2
+g2 <- t %%3
+g3 <- t %%5
+
+str_gran <- bind_cols(index = t, g1 = g1, g2 = g2, g3 = g3)
+
+# calculation of g1
+g1_table <- bind_cols(g1 = unique(g1), dist_mean = c(mu11, mu12))
+
+g1_tally <- str_gran %>% group_by(g1) %>%
+  count() %>%
+  left_join(g1_table, by = "g1")
+     
+g1_dist <- g1_tally %>%
+  mutate(g1_d = list(rep(g1, each = n)),
+                    g1_dist = list(rnorm(n, dist_mean, 1))) %>% 
+  ungroup() %>% 
+  select(g1_d, g1_dist) %>% 
+  unnest(cols = c(g1_d, g1_dist))
+
+g1_data <- str_gran %>% arrange(g1) %>% bind_cols(g1_dist = g1_dist$g1_dist) %>% arrange(index)
+
+# calculation of g2
+
+g2_table<- bind_cols(g2 = unique(g2),
+                     dist_mean = c(mu21, mu22, mu23))
+
+g2_tally <- str_gran %>%
+  group_by(g2) %>%
+  count() %>% 
+  left_join(g2_table, by = "g2")
+
+g2_dist <- g2_tally %>%
+  mutate(g2_d = list(rep(g2, each = n)),
+         g2_dist = list(rnorm(n, dist_mean, 1))) %>% 
+  ungroup() %>% 
+  select(g2_d, g2_dist) %>% 
+  unnest(cols = c(g2_d, g2_dist))
+
+g2_data <- str_gran %>% arrange(g2) %>% bind_cols(g2_dist = g2_dist$g2_dist) %>% arrange(index)
+
+# calculation of g3
+
+g3_table<- bind_cols(g3 = unique(g3), dist_mean = c(mu31, mu32, mu33, mu34, mu35))
+
[…]

### 2021-11-07 supplementary/old/conditional_dendro.R

@@ -1,0 +1,83 @@
+#' Dendogram of simulated objects
+#'
+#' @param .data a tsibble.
+#' @param gran1,gran2 the granularity against which similarity is measured. For the granularity hour of the week, value is "hour_week".
+#' @param response response variable.
+#' @param clust_var column in which subject lies
+#' @param ncust number of subjects taken
+#' @param niter numbre of simulations per customer
+#' @return dendogram describing the tree produced by the clustering process
+#'
+#' @examples
+#'library(ggplot2)
+#'library(dplyr)
+#'library(ggdendro)
+#'library(gravitas)
+#'.data = smart_meter10
+#'gran1 = "wknd_wday"
+#'gran2 = "hour_day"
+#'response = "general_supply_kwh"
+#'clust_var = "customer_id"
+#' @examples
+#' conditional_dendro(.data, gran1,
+#'                       gran2, response,
+#'                       clust_var, niter = 5,
+#'                       ncust = 2)
+#' @export
+conditional_dendro <- function(.data, gran1,
+                                   gran2, response,
+                                   clust_var, niter = 5,
+                                   ncust = 2)
+{
+dis_cus <- .data %>%
+  dplyr::distinct(.data[[clust_var]]) %>%
+  dplyr::sample_n(ncust)
+
+cust_data <- (1:nrow(dis_cus)) %>%
+  purrr::map_df(function(i){
+    make_cust_iter(.data, gran1, gran2, response, custid = dis_cus[[clust_var]][i], niter)
+    })
+
+data_whole <-  cust_data %>%
+  dplyr::group_by(cust_iter_id) %>%
+  dplyr::mutate(index = dplyr::row_number()) %>%
+  dplyr::ungroup() %>%
+  tsibble::as_tsibble(index = index, key = cust_iter_id)
+
+distance_sum10cust <- sim_dist(data_whole,
+                              gran1,
+                              gran2,
+                              response = "sim_data",
+                              key = "cust_iter_id",
+                              dist_ordered = TRUE,
+                              method = "sum")
+
+d <- stats::as.dist(distance_sum10cust)
+hc <- stats::hclust(d)
+simulate_5each_sum <- ggdendro::ggdendrogram(hc)
+simulate_5each_sum + ggplot2::ggtitle(paste(nrow(dis_cus),"customer simulated", niter, "times"))
+}
+
+library(dplyr)
[…]

### 2021-11-07 supplementary/old/interact_gran_4D.R

@@ -1,0 +1,301 @@
+
+library(knitr)
+library(here)
+read_chunk(here::here("old/simdata01.R"))
+
+# How the time series plot of the designs look like for different sample sizes?
+
+set.seed(9999)
+nx_val = 2 # number of x-axis levels
+nfacet_val = 3 # number of facet levels
+w1_val = 3 # increment in mean
+w2_val = 0 # increment in sd
+mean_val = 0 # mean of normal distribution of starting combination
+sd_val = 1 # sd of normal distribution of starting combination
+quantile_prob_val = seq(0.1, 0.9, 0.1)
+```
+
+```{r load-lib}
+```
+
+
+```{r designs}
+```
+
+
+## nobs per combination is 50
+
+```{r nobs50, echo = TRUE}
+ntimes_val = 50 # nobs per combination 
+```
+
+
+```{r data}
+```
+
+```{r category-plots}
+```
+
+```{r line-graph}
+```
+
+```{r plot50}
+(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)
+
+```
+
+# Clustering on simulated datasets
+
+```{r, echo=TRUE}
+sample_seed <-  seq(1, 100, 1)
+```
+
+_DGP: Generate 10 time series from each designs_ 
+Time series are simulated from each of these designs with $50$ observations in each group. So we have $50$ observations each for the six combination of categories $(1,1), (1, 2), (2, 1), (2, 2), (3, 1), (3, 2)$. Time series are simulated for ten different seeds (from `sample_seed`) for each design. `data_varall-1` represents a data set from a design $D_{var_{all}}$ (distributions change across both facet and x) with a seed $10$. `data_varall-2` represents a data set from a design $D_{var_{all}}$ with a seed $20$ and so on.
+`data_null`, `data_varx` and `data_varf `corresponds to designs $D_{null}$, $D_{var_{x}}$ and $D_{var_{f}}$ designs respectively.   
+
+_Compute quantiles of conditional distributions_
+Conditional quantiles are obtained for each combination of categories.  
+
+_JS Pairwise distances between datasets_
+Distance between the data sets is computed as the sum of JS distances across different categories.  
+
+_Hierarchical clustering with 4 clusters_
[…]

### 2021-11-07 supplementary/old/job-algo2-500cust.R

@@ -1,0 +1,83 @@
+library(knitr)
+library(dplyr)
+library(lubridate)
+#library(lvplot)
+#library(ggridges)
+library(tsibble)
+library(gravitas)
+#library(ggpubr)
+library(readr)
+#library(kableExtra)
+#library(sugrrants)
+#library(here)
+#library(patchwork)
+#library(scales)
+#library(GGally)
+#library(viridis)
+
+
+ncust = 100 # number of sampled customers
+nperm = 200 # number of permutations for normalization
+nsamp = 200 # number of samples for threshold
+
+#take customers who do not have gaps in their data
+# elec_nogap <- read_rds("data/elec_nogap.rds")
+# 
+# elec_nogap_2013 <- elec_nogap %>%
+#   filter(year(reading_datetime)==2013)
+#   
+# write_rds(elec_nogap_2013, "data/elec_nogap_2013.rds")
+
+# elec_nogap_2013 <- read_rds("data/elec_nogap_2013.rds")
+
+#just take 100 customers from them
+# set.seed(12345)
+# 
+# sm_100 <- elec_nogap_2013 %>%
+#   as_tibble() %>%
+#   distinct(customer_id) %>%
+#   slice_sample(n = ncust)
+# 
+# # take data for those customers only for 2013
+# elec <- elec_nogap_2013 %>%
+#   filter(customer_id %in% sm_100$customer_id) %>%
+#   ungroup()
+# 
+# write_rds(elec, "data/elec_nogap_2013_100.rds")
+
+elec <- read_rds("data/elec_nogap_2013_500.rds")
+cust_id <- unique(elec$customer_id)
+scen <- as.numeric(commandArgs()[[6]])
+#scen <- 2
+custj <- cust_id[scen]
+# consider harmonies which are generally significant for electricity data
+
+#harmonies <- read_rds("../paper-hakear/paper/data/harmonies.rds")
+
+# harmonies <- harmonies %>% 
+#   mutate(comb = paste(facet_variable, 
+#                              x_variable, 
+#                              sep = "-")) %>% 
+#   filter(comb %in% c("hour_day-wknd_wday", "day_month-hour_day", "wknd_wday-hour_day", "hour_day-day_week", "day_week-hour_day")) %>% 
+#   select(-comb)
+# 
+# write_rds(harmonies, "data/harmonies.rds")
+
+harmonies <- read_rds("data/harmonies.rds")
+
+# take customers who do not have gaps in their data
+  
+  data_id <-  elec %>% 
[…]

### 2021-11-07 supplementary/old/job-algo2.R

@@ -1,0 +1,90 @@
+library(knitr)
+library(tidyverse)
+library(lubridate)
+#library(lvplot)
+#library(ggridges)
+library(tsibble)
+library(gravitas)
+#library(ggpubr)
+library(readr)
+#library(kableExtra)
+library(distributional)
+#library(ggplot2)
+#library(sugrrants)
+#library(here)
+#library(patchwork)
+#library(scales)
+#library(GGally)
+#library(viridis)
+
+
+ncust = 100 # number of sampled customers
+nperm = 2 # number of permutations for normalization
+nsamp = 2 # number of samples for threshold
+
+#take customers who do not have gaps in their data
+# elec_nogap <- read_rds("data/elec_nogap.rds")
+# 
+# elec_nogap_2013 <- elec_nogap %>%
+#   filter(year(reading_datetime)==2013)
+#   
+# write_rds(elec_nogap_2013, "data/elec_nogap_2013.rds")
+
+# elec_nogap_2013 <- read_rds("data/elec_nogap_2013.rds")
+# 
+# #just take 100 customers from them
+# set.seed(12345)
+# 
+# sm_500 <- elec_nogap_2013 %>%
+#   filter(!customer_id %in% elec$customer_id) %>% 
+#   as_tibble() %>%
+#   distinct(customer_id) %>%
+#   slice_sample(n = 500)
+
+#elec <- read_rds("data/elec_nogap_2013_100.rds")
+
+# take data for those customers only for 2013
+# new_elec <- elec_nogap_2013 %>%
+#   filter(customer_id %in% sm_500$customer_id) %>%
+#   ungroup()
+
+#write_rds(elec, "data/elec_nogap_2013_100.rds")
+#write_rds(new_elec, "data/elec_nogap_2013_500.rds")
+
+#elec <- read_rds("data/elec_nogap_2013_100.rds")
+elec <- read_rds("data/elec_nogap_2013_500.rds")
+cust_id <- unique(elec$customer_id)
+#scen <- as.numeric(commandArgs()[[6]])
+scen <- 2
+custj <- cust_id[scen]
+# consider harmonies which are generally significant for electricity data
+
+#harmonies <- read_rds("../paper-hakear/paper/data/harmonies.rds")
+
+# harmonies <- harmonies %>% 
+#   mutate(comb = paste(facet_variable, 
+#                              x_variable, 
+#                              sep = "-")) %>% 
[…]

### 2021-11-07 supplementary/old/parallel_param.R

@@ -1,0 +1,5 @@
+sample_seed.all <- list(seq(10, 100, 10))
+
+for (sample_seed in sample_seed.all) {
+  rmarkdown::render('index.Rmd', params = list(sample_seed = sample_seed))
+}

### 2021-11-07 main.Rmd
mds plot square aspect ratio and caption change

@@
 
-```{r mds-plot-validation, fig.cap = "Relative positions of clusters corresponding to different scenarios (columns) for different values of mean differences between categories (rows) are shown using the first two dimensions of MDS.
+```{r mds-plot-validation, fig.cap = "MDS summary plots to illustrate the cluster sparation for the range of mean differences (rows) under the different scenarios (columns).
 It can be observed that clusters become more compact and separated for higher mean differences between categories across all scenarios.

### 2021-11-07 script/main.R

@@ -1068,7 +1068,8 @@ mds_plot_10 <- ggplot(all_data,
   theme_bw()+
   theme(legend.position = "bottom") +
   scale_color_brewer(palette = "Dark2")+
-  ylab("mds1") + xlab("mds2")
+  ylab("mds1") + xlab("mds2")+
+  coord_fixed()
 
 mds_plot_10
 

### 2021-11-07 main.Rmd
parcoord-sim parcoord in validation section, scenario (a), (b) and (c) changed to S1, S2 and S3

@@
 As a result, in this scenario, we analyze a combination of categories generated from different distributions.
-Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$.
-In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories.
-Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs.
-Design $D1$ exhibits no change in distributions across $g1$ or $g2$, whereas Designs $D2$ and $D3$ alter across only $g1$ and $g2$, respectively.
-D4 varies across both $g1$ and $g2$ categories.
-D3 and D4 appear similar based on their relative differences across consecutive categories, but $D4$ also changes across facets, unlike $D3$, which has all facets look the same.
-
-```{r interaction-gran, fig.cap = "Distribution of the simulated variable across $g1$ conditional on $g2$ is shown through boxplots for 4 designs.
-D1 has no change in distributions across different categories of $g1$ or $g2$, while D2 and D3 change across only $g1$ and $g2$ respectively.
-D4 changes across categories of both $g1$ and $g2$."}
-
-```
+Display of design and related results can be found in supplementary.
 
@@
 While the first plot indicates that no variable stands out during clustering, the next two designs demonstrate that g1 and (g1, g2) have relatively lower inter-cluster distances, indicating that they did not contribute to clustering.
-It is worth emphasising that these facts are consistent with our initial assumptions when designing the scenarios and (a), (b), and (c) correspond to Scenario (a), (b) and (c) respectively."}
+It is worth emphasising that these facts are consistent with our initial assumptions when designing the scenarios and S1, S2, S3 correspond to scenarios with three, two and one significant granularity respectively."}
 

### 2021-11-07 paper-supplementary.Rmd
parcoord-sim parcoord in validation section, scenario (a), (b) and (c) changed to S1, S2 and S3

@@
 A dual-level progressive similar instance selection (DPSIS) approach could also be used -->
+
+### Interaction of granularities
+
+Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$.
+In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories.
+Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs.
+Design $D1$ exhibits no change in distributions across $g1$ or $g2$, whereas Designs $D2$ and $D3$ alter across only $g1$ and $g2$, respectively.
+D4 varies across both $g1$ and $g2$ categories.
+D3 and D4 appear similar based on their relative differences across consecutive categories, but $D4$ also changes across facets, unlike $D3$, which has all facets look the same.
+
+```{r interaction-gran, fig.cap = "Distribution of the simulated variable across $g1$ conditional on $g2$ is shown through boxplots for 4 designs.
+D1 has no change in distributions across different categories of $g1$ or $g2$, while D2 and D3 change across only $g1$ and $g2$ respectively.
+D4 changes across categories of both $g1$ and $g2$."}
+
+```
 

### 2021-11-07 script/main.R

@@ -981,7 +981,7 @@ design3 <- append_files_plot(folder_name = "validation/js-nqt", path = "/1gran_c
 # design2 <- append_files_plot(folder_name = "js-nqt", path = "/2gran_change_4D/")
 # design3 <- append_files_plot(folder_name = "js-nqt", path = "/1gran_change_5D/")
 
-(design1 + design2 + design3 + plot_layout(guides = "collect")+ plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')'))*theme_characterisation()&theme(legend.position = "none")&xlab("")
+(design1 + design2 + design3 + plot_layout(guides = "collect")+ plot_annotation(tag_levels = '1', tag_prefix = 'S', tag_suffix = ''))*theme_characterisation()*theme(plot.title = element_text(hjust = 0.5))&theme(legend.position = "none")&xlab("")
 
 
 ##----mds-plot-validation

### 2021-11-07 main.Rmd
S2 and S3 plots caption and title revised

@@
 
-```{r gran2and1-clubbed, fig.cap=" Plots (b) and (c) correspond to Scenarios (b) and (c) respectively.
-In (b) $g2$, $g3$ changes across atleast one design but $g1$ remains constant.
-Only $g3$ changes across different designs in (c)."}
+```{r gran2and1-clubbed, fig.cap=" Boxplots showing distributions of categories across different designs (rows) and granularities (columns) for scenarios S2 and S3.
+In S2, $g2$, $g3$ change across atleast one design but $g1$ remains constant.
+Only $g3$ changes across different designs in S3."}
 #gran2_change  + gran1_change + plot_layout(widths = c(1,1))

### 2021-11-07 main.Rmd
plots S1, S2 and S3 fixed

@@
 
-_Scenario (a): All significant granularities_
+_Scenario (S1): All significant granularities_
 
@@
 
-_Scenario (b): Few significant granularities_
+_Scenario (S2): Few significant granularities_
 
@@
 
-_Scenario (c): One significant granularity_
+_Scenario (S3): One significant granularity_
 
@@
 
-```{r plot-3gran-new, out.width="100%", fig.cap="The linear (left) and cyclic (right) representation of the simulated variable is shown.
-Each row represents a design in Scenario (a).
-In this scenario, all of $g1$, $g2$ and $g3$ changes across at least one design.
-Also, it is not possible to comprehend these distributional differences in patterns just by looking at or considering the linear representation."}
+```{r plot-3gran-new, out.width="100%", fig.cap = "The linear (left) and cyclic (right) representation is shown under scenario S1 using line plots and boxplots respectively.
+Each row represents a design.
+Distributions of categories across $g1$, $g2$ and $g3$ change across at least one design as can be observed in the cyclic representation.
+It is not possible to comprehend these structural differences in patterns just by looking at or considering the linear representation."}
 
@@
 It can be observed that clusters become more compact and separated for higher mean differences between categories across all scenarios.
-Between scenarios, separation is least prominent corresponding to scenario (c) where only one granularity is responsible for distinguishing the clusters."}
+Between scenarios, separation is least prominent corresponding to Scenario (S3) where only one granularity is responsible for distinguishing the clusters."}
 
@@
 (c) demonstrates that intercluster distances are only significant for the variable $g3$, meaning that the clusters formed are primarily due to $g3$.
-This corresponds to our design in Scenario (c), in which distributional differences were implemented solely for g3.
+This corresponds to our design in Scenario (S3), in which distributional differences were implemented solely for g3.
 (b) demonstrates the role of g2 and g3 in clustering.

### 2021-11-07 main.Rmd
All scenario (a), (b) and (c) changed to S1, S2, S3

@@
 The first plot shows that no variable stands out in the clustering, but the following two designs show that {g1} and {g1, g2} have very low inter cluster distances, meaning that they did not contribute to the clustering.
-It is worth noting that these facts correspond to our original assumptions when developing the scenarios, which incorporate distributional differences over three (a), two (b), and one (c) significant granularities.
-Hence, Figure \ref{fig:parcoord-sim} (a), (b), and (c) validate the construction of scenarios (a), (b), and (c) respectively.
+It is worth noting that these facts correspond to our original assumptions when developing the scenarios, which incorporate distributional differences over three (S1), two (S2), and one (S3) significant granularities.
+Hence, Figure \ref{fig:parcoord-sim} (S1), (S2), and (S3) validate the construction of scenarios (S1), (S2), and (S3) respectively.
 

### 2021-11-07 script/main.R

@@ -332,7 +332,7 @@ table <- tibble(design, g1, g2, g3)
 
 knitr::kable(
   list(tab_dist,  tibble(design, g1, g2, g3)),
-  caption = 'For Scenario (a), distributions of different categories when they vary (top). If distributions are fixed, they are set to N(0, 1). 5 designs resulting from different distributions across categories (below)',
+  caption = 'For S1, distributions of different categories when they vary (top). If distributions are fixed, they are set to N(0, 1). 5 designs resulting from different distributions across categories (below)',
   booktabs = TRUE, valign = 't'
 )
 
@@ -1016,7 +1016,7 @@ mds_data <- function(data){
   
   colnames(mds) <- c("Dim.1", "Dim.2")
   
-  groups<-bind_rows(data %>%
+  groups <- bind_rows(data %>%
                       distinct(item1, group_item1) %>% 
                       rename("item" = "item1",
                              "group" = "group_item1"),
@@ -1042,18 +1042,18 @@ simtable <- expand.grid(mean_diff = mean_diff,
                         #time_series = time_series, 
                         nT = nT)
 
-data1 <- mds_data(read_rds("validation/js-nqt/3gran_change_5D/data_validation_25.rds")) %>% mutate(diff = 1 , design= "a")
-data2 <-mds_data(read_rds("validation/js-nqt/3gran_change_5D/data_validation_26.rds")) %>% mutate(diff = 2 , design= "a")
-data3 <-mds_data(read_rds("validation/js-nqt/3gran_change_5D/data_validation_27.rds")) %>% mutate(diff = 5 , design= "a")
+data1 <- mds_data(read_rds("validation/js-nqt/3gran_change_5D/data_validation_25.rds")) %>% mutate(diff = 1 , design= "S1")
+data2 <-mds_data(read_rds("validation/js-nqt/3gran_change_5D/data_validation_26.rds")) %>% mutate(diff = 2 , design= "S1")
+data3 <-mds_data(read_rds("validation/js-nqt/3gran_change_5D/data_validation_27.rds")) %>% mutate(diff = 5 , design= "S1")
 
-data4 <- mds_data(read_rds("validation/js-nqt/2gran_change_4D/data_validation_7.rds")) %>% mutate(diff = 1 , design= "b")
[…]

### 2021-11-07 script/main.R
change the application graph

@@ -67,7 +67,7 @@ theme_characterisation <- function() {
 theme_application <- function() {
   
   theme_light() + # setting theme
-    #theme(strip.text = element_text(margin = margin(b = 0, t = 0))) + # narrow facet space
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)), strip.text.x = element_text(size = 8)) + # narrow facet space
     theme(axis.title.y=element_blank(),
           axis.text.y=element_blank(),
           axis.ticks.y=element_blank()) + # no axis ticks 
@@ -1224,36 +1224,78 @@ cluster_result <- suppressMessages(f %>%
   rename("customer_id" = "id") %>% 
   mutate(group = as.factor(group))
 
+cluster_result_id <- cluster_result %>% arrange(group) %>% mutate(divide_cust = rep(c(1,2), each = 12))
+
 ##----hod-ind-group
-hod_ind_group <- data_hod %>% 
-  left_join(cluster_result, by = c("customer_id")) %>% 
+hod_ind_group1 <- data_hod %>% 
+  left_join(cluster_result_id, by = c("customer_id")) %>% 
+  filter(divide_cust == 1) %>% 
   ggplot(aes(x = category)) + 
   geom_ribbon(aes(ymin = `25%`, 
                   ymax = `75%`,
-                  group=customer_id, fill = group),
+                  group=customer_id),
               alpha = 0.5) +
   geom_line(aes(y = `50%`,
-                group=customer_id, 
-                color = group), size = 1) +
-  facet_wrap(group~customer_id, 
+                group=customer_id), size = 1) +
+  facet_wrap(~customer_id, 
              scales = "free_y",
-             nrow=8) + 
+             ncol=1) + 
   theme_application()+
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
   scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +    xlab("hod")  +
   scale_x_discrete(breaks = seq(0, 23, 3))+ theme(legend.position = "bottom")+theme(plot.margin = unit(c(0,0,0,-1), "cm"))
[…]

### 2021-11-08 main.Rmd
change the application graph
adding tsne plot

@@
 
+```{r tsne-plot}
+```
+
 <!-- - can be done in several ways -->
@@
 
-```{r opt-clusters}
+```{r opt-clusters-jsd}
+
+```
+
+```{r opt-cluster-tsne-jsd, fig.cap = "(a) shows t-SNE summary of the 24 customers chosen through prototype selection, (b) shows a plot of sindex for different cluster size $k$"}
+
 ```
@@
 Individuals with similar distributional differences across categories of hod or moy are grouped together, with some exceptions.
-Distributional differences across categories of wnwday are not discernable except for a couple of customers." }
+Distributional differences across categories of wnwday are not discernable except for a couple of customers.", fig.height =  3.5}
 ```

### 2021-11-08 main.Rmd
add tsne and sindex plot

@@
 
-```{r opt-cluster-tsne-jsd, fig.cap = "(a) shows t-SNE summary of the 24 customers chosen through prototype selection, (b) shows a plot of sindex for different cluster size $k$"}
+```{r opt-cluster-tsne-jsd, fig.cap = "(a) shows t-SNE summary of the $24$ selected customers chosen through prototype selection, (b) shows a plot of sindex for different cluster size."}
 

### 2021-11-08 supplementary/prototype_selection/quantiles_raw.R

@@ -89,6 +89,9 @@ load("data/data_356cust_wide")
 
 # without scaling
 
+
+
+
 data_356cust_pc <- prcomp(data_356cust_wide[,-1],
                           center = FALSE, scale = FALSE, retx = TRUE)
 
@@ -112,10 +115,16 @@ sort(abs(data_356cust_pc$rotation[,1]))
 library(Rtsne)
 set.seed(2099)
 
+
+
 tSNE_fit <- data_356cust_wide%>% 
   select(-customer_id) %>% 
   Rtsne(pca = FALSE,
          perplexity = 30)
+
+
+
+
 
 # data_pick_one <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble 
 # data_pick_two <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble
@@ -152,6 +161,8 @@ tsne_df <- data.frame(tsneX = tSNE_fit$Y[, 1],
   data_pick_cust, by = c("customer_id"
 )) %>% mutate(design = if_else(is.na(design), "0", design))
 
+
+
 #rownames(tsne_df) <- data_356cust_wide$customer_id
 
 rownames(data_356cust_pc10) <- data_356cust_wide$customer_id
@@ -164,7 +175,9 @@ tsne_xy <- ggplot(tsne_df, aes(x = tsneX, y = tsneY, color = design)) +
   guides(color = FALSE) +
   labs(caption = "tSNE") +
   theme(aspect.ratio = 1) +
-  theme_light()
+  theme_light()+coord_fixed(ratio = 1)
+
+
 
 ## ----highlight-customer--------------------------------------------------------
 

### 2021-11-08 paper-supplementary.Rmd
deleted contribution of hod and moy

@@
 ## Plot displaying missing observations
+
+\noindent The data set contains 13,735 customers, of which 8,685 have no missing values and the remaining $5,050 customers have at least one missing value.
+We wanted to see if there was any structure to these missing observations, so we plotted them over time in Figure \ref{fig:missing-data}.
+Each row represents a customer, with the blank indicating non-missing values and the shapes indicating missing values.
+Figure 9 of @wang2019tsibble inspired this plot.
+For the most part, it appears that data for September and October 2012 are missing.
+Otherwise, missingness can occur at any time and there is no pattern.
+
+```{r missing-data}
+knitr::include_graphics("figs/missing-data-5050.png")
+```
 
@@
 group <- f %>% hclust (method = "ward.D") %>% cutree(k=17)
-cluster_result <- bind_cols(customer_id = id, group = group)
+
+cluster_result <- bind_cols(customer_id = id, group = group) %>%
+  arrange(group) %>% 
+  mutate(divide_cust = c(rep(1, 177), rep(2, 176)))
 
@@
 
-# contribution of hod
-
-```{r contribution}
-# data_validation <- hod %>%  
-#   rename("hod" = "distance") %>% 
-#   left_join(moy, 
-#             by = c("item1", "item2"))%>% 
-#   rename("moy" = "distance") %>% 
-#   left_join((wkndwday), 
-#             by = c("item1", "item2"))%>% 
-#   rename("wkndwday" = "distance") %>% 
-#   left_join(cluster_result, by = c("item1" = "customer_id")) %>% 
-#   rename("group_item1" = "group") %>% 
-#   left_join(cluster_result, by = c("item2" = "customer_id")) %>%  
-#   rename("group_item2" = "group") %>% 
-#   mutate(hod = hod/24, moy = moy/12, wkndwday = wkndwday/2) %>% 
-#   filter(!is.na(group_item1)| !is.na(group_item2)) %>% 
-#   pivot_longer(3:5,names_to="gran",
-#                values_to = "distance")  
-#
-
-data_validation_cat <- read_rds(here::here("validation/data_validation_cat_all_cluster.rds"))
-
-# hod_cat <- data_pick %>%
-#   dist_gran_cat(gran1 = "hour_day", response = "general_supply_kwh") %>%
-#   mutate(gran = "hod")
-# 
-# moy_cat <- data_pick %>%
-#   dist_gran_cat(gran1 = "month_year", response = "general_supply_kwh")%>%
-#   mutate(gran = "moy")
-# 
-# wkndwday_cat <- data_pick %>%
-#   dist_gran_cat(gran1 = "wknd_wday", response = "general_supply_kwh")%>%
-#   mutate(gran = "wnwd")
-# 
-# 
-# data_validation_cat <- bind_rows(hod_cat, moy_cat, wkndwday_cat) %>%
-#   mutate(customer_from = as.character(customer_from),
-#          customer_to = as.character(customer_to))
-
-gran_cat_dist_from <- data_validation_cat %>% 
-  mutate(customer_from = as.character(customer_from),
-         customer_to = as.character(customer_to)) %>% 
-  left_join(cluster_result, by = c("customer_from" = "customer_id")) %>% 
-  rename("group_from" = "group") %>% 
-  left_join(cluster_result, by = c("customer_to" = "customer_id")) %>%  
-  rename("group_to" = "group") %>% 
-  group_by(gran,
-           category,
-           group_from, 
-           group_to) %>% 
-  summarise(sum = sum(distance)) %>% 
-  pivot_wider(names_from = group_to, values_from = sum) %>% 
-  mutate(distance = sum(`1`, `2`, `3`, `4`, `5`,`6`, `7`, `8`, `9`, `10`, `11`, `12`, `13`, `14`, `15`, `16`, `17`)) %>% 
-  select(-(4:8)) %>% 
-  arrange(-distance) %>% 
-    rename("group"="group_from")
-
-gran_cat_dist_to <- data_validation_cat %>% 
-  mutate(customer_from = as.character(customer_from),
-         customer_to = as.character(customer_to)) %>% 
-  left_join(cluster_result, by = c("customer_from" = "customer_id")) %>% 
-  rename("group_from" = "group") %>% 
-  left_join(cluster_result, by = c("customer_to" = "customer_id")) %>%  
-  rename("group_to" = "group") %>% 
-  group_by(gran,
-           category,
-           group_from, 
-           group_to) %>% 
-  summarise(sum = sum(distance)) %>% 
-  pivot_wider(names_from = group_from, values_from = sum) %>% 
-  mutate(distance = sum(`1`, `2`, `3`, `4`, `5`,`6`, `7`, `8`, `9`, `10`, `11`, `12`, `13`, `14`, `15`, `16`, `17`)) %>% 
-  select(-(4:8)) %>% 
-  arrange(-distance) %>% 
-    rename("group"="group_to")
-
-# Contribution of individual category for hod
-
-gran_cat_hod <-   gran_cat_dist %>% 
-  filter(gran=="hod") %>% 
-  mutate(category = as.integer(category)) %>% 
-  arrange(group_from, category)
-
-gran_cat_hod$category <- factor(gran_cat_hod$category, levels = 0:23)
-
-# for each group, percent contribution of hours
-
-group_total <- gran_cat_hod %>% 
-  ungroup() %>% 
-  group_by(group_from) %>% 
-  summarise(group_total = sum(distance))
-
-data_pcp <- gran_cat_hod %>% 
-  ungroup %>% 
-  left_join(group_total, by = "group_from") %>% 
-  mutate(cat_contibution = distance*100/group_total) %>% 
-  group_by(group_from) %>% 
-  #arrange(-distance) %>% 
-  select(group_from, category, cat_contibution) %>% 
-  pivot_wider(names_from = "category", 
-              values_from = "cat_contibution") %>% mutate(group_from = as.factor(group_from))
-
-parcoord <- GGally::ggparcoord(data_pcp ,
-                               columns = 2:ncol(data_pcp),
-                               groupColumn = "group_from",
-                               showPoints = TRUE, 
-                               alphaLines = 1,
-                               scale = "globalminmax"
-) + 
-  ggplot2::theme(
-    plot.title = ggplot2::element_text(size=10)
-  )+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
-  theme(legend.position = "bottom") +
-  xlab("") +
-  ylab("wpd")+
-  scale_color_manual(values=as.vector(polychrome(20)))+
-  theme_bw()
-parcoord 
-```
-
-# contribution of moy
-
-```{r contribution-moy}
-
-gran_cat_hod <-   gran_cat_dist %>% 
-  filter(gran == "moy") %>% 
-  #mutate(category = as.integer(category)) %>% 
-  arrange(group_from, category)
-
-# gran_cat_hod$category <- factor(gran_cat_hod$category, levels = 0:23)
-
-gran_cat_hod$category <- factor(gran_cat_hod$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
-
-# for each group, percent contribution of hours
-
-group_total <- gran_cat_hod %>% 
-  arrange(category) %>% 
-  ungroup() %>% 
-  group_by(group_from) %>% 
-  summarise(group_total = sum(distance))
-
-data_pcp <- gran_cat_hod %>% 
-  arrange(category) %>% 
-  ungroup %>% 
-  left_join(group_total, by = "group_from") %>% 
-  mutate(cat_contibution = distance*100/group_total) %>% 
-  group_by(group_from) %>% 
-  #arrange(-distance) %>% 
-  select(group_from, category, cat_contibution) %>% 
-  pivot_wider(names_from = "category", 
-              values_from = "cat_contibution") %>% mutate(group_from = as.factor(group_from))
-
-parcoord <- GGally::ggparcoord(data_pcp ,
-                               columns = 2:ncol(data_pcp),
-                               groupColumn = "group_from",
-                               showPoints = TRUE, 
-                               alphaLines = 1,
-                               scale = "globalminmax"
-) + 
-  ggplot2::theme(
-    plot.title = ggplot2::element_text(size=10)
-  )+
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
-  theme(legend.position = "bottom") +
-  xlab("") +
-  ylab("wpd")+
-  scale_color_manual(values=as.vector(polychrome(20)))+
-  theme_bw()
-parcoord 
-```
-
 ```{r data-heatmap-hod-group-entire-1col}

### 2021-11-09 paper-supplementary.Rmd
raw plot update

@@
 
-## Raw time series plot
-
-\noindent To get a sense of how the raw time series data looks, we plot the energy usage for $50$ sampled households is plotted along the y-axis versus time from past to future in Figure ref\{fig:raw-data-50}.
+```{r raw-data-50, out.width = "100%", fig.cap=" The raw half-hourly energy usage for $50$ sampled households is plotted along the y-axis versus time in a linear scale.
 Each of these series is associated with a single customer.
-Energy consumption for each customer is given at fine temporal resolution (every 30 minutes) for a period of 2-3 years.
-
-```{r raw-data-50, out.width = "100%", fig.cap="The raw data for 50 households are shown.
 It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed.
@@
 cluster_result <- bind_cols(customer_id = id, group = group) %>%
-  arrange(group) %>% 
-  mutate(divide_cust = c(rep(1, 177), rep(2, 176)))
+  arrange(group)
+# %>% 
+#   mutate(divide_cust = c(rep(1, 177), rep(2, 176)))
 
@@
 
-data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")
+# data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")
 
-hod_group_entire <- data_heatmap_hod_group %>% 
+hod_group_entire1 <- data_heatmap_hod_group %>% 
+  filter(group<9) %>% 
   ggplot(aes(x = category)) + 
@@
              scales = "free_y",  
-             ncol = 17) + 
+             nrow = 8,labeller = "label_both") + 
               #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
@@
   scale_color_manual(values=as.vector(polychrome(20)))+
+  theme(legend.position = "bottom")
+
+hod_group_entire2 <- data_heatmap_hod_group %>% 
+  filter(group>=9) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=group,
+                  fill = as.factor(group), alpha = 0.5),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=group, 
+                color = as.factor(group)), size = 1)+
+  facet_wrap(~group, 
+             scales = "free_y",  
+             nrow = 9,labeller = "label_both") + 
+              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
+  ylab("demand (in Kwh)") + 
+  theme_bw()  +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ 
+  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme_application() +
+  scale_fill_manual(values=as.vector(tableau20(20)))+
+  scale_color_manual(values=as.vector(tableau20(20)))+
   theme(legend.position = "bottom")
@@
 
-data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")
+# data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")
 
-moy_group_entire <- data_heatmap_moy_group %>% 
+moy_group_entire1 <- data_heatmap_moy_group %>% 
+  filter(group<9) %>% 
   ggplot(aes(x = category)) + 
@@
              scales = "free_y", 
-             labeller = "label_value",
-             ncol = 17) +
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
-  ylab("demand (in Kwh)") +
-  theme_bw() + theme_application() +
-  scale_fill_manual(values=as.vector(polychrome(20)))+
-  scale_color_manual(values=as.vector(polychrome(20)))+
-  theme(legend.position = "bottom")
-```
-
-```{r data-wnwd-group-entire}
-wkndwday_data <- data_group %>%
-  create_gran("wknd_wday") 
-wkndwday_data$group <- as.factor(wkndwday_data$group)
-# %>% 
-#   create_gran("hour_day")
-
-ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]
-
-wkndwday_group_entire <- wkndwday_data%>% 
-  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
-  #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
-  geom_boxplot(aes(fill = group, color = group),alpha = 0.5, outlier.alpha = 0.05)+
-  #geom_boxplot(outlier.size = 1) + 
-  coord_cartesian(ylim = ylim1*1.05)+
-  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
-#geom_boxplot(aes(fill = as.factor(group))) +
-  #scale_fill_lv() +
- xlab("wknd-wday") + 
-  ylab("demand (in Kwh)") +
-   facet_wrap(~group, 
-             scales = "free_y", 
              labeller = "label_both",
-             ncol = 17) + 
-  theme_bw() + theme_application() +
-  scale_fill_manual(values=as.vector(polychrome(20)))+
-  scale_color_manual(values=as.vector(polychrome(20)))+
-  theme(legend.position = "none")
-```
-
-```{r combined-groups-js-entire, eval = FALSE, fig.cap = "The distribution of electricity demand for the clusters across hod (a), moy (b) and wkndwday (c).
-It seems like group 2 and 5 have a hod pattern across its members, while group 1, 3, 5 have a moy pattern.
-Wknd-wday variations across groups are not distinguishable, indicating that it is not a critical variable for clustering.
-It is helpful to compare the summarised distributions of groups to that of individuals to confirm that the most of individuals in the group have the same characterisation."}
-
- (hod_group_entire/moy_group_entire/ wkndwday_group_entire) +
-  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
-  plot_layout(guides = "collect")& theme(legend.position = 'none')
-```
-
-```{r data-heatmap-hod-group-entire-1col}
-
-hod_group_entire <- data_heatmap_hod_group %>% 
-  filter(group!="group-NA") %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`,
-                  group=group,
-                  fill = as.factor(group), alpha = 0.5),
-              alpha = 0.5) +
-  geom_line(aes(y = `50%`,
-                group=group, 
-                color = as.factor(group)), size = 1)+
-  facet_wrap(~group, 
-             scales = "free_y",  
-             nrow = 8) + 
-              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
-    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
-  ylab("demand (in Kwh)") + 
-  theme_bw()  +
-  scale_x_discrete(breaks = seq(1, 24, 3))+ 
-  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
-  theme_application() +
-  scale_fill_manual(values=as.vector(polychrome(20)))+
-  scale_color_manual(values=as.vector(polychrome(20)))+
-  theme(legend.position = "bottom")
-
-```
-
-```{r data-heatmap-moy-group-entire-1col}
-
-moy_group_entire <- data_heatmap_moy_group %>% 
-  filter(group!="group-NA") %>% 
-  ggplot(aes(x = category)) + 
-  geom_ribbon(aes(ymin = `25%`, 
-                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
-  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
-  facet_wrap(~group, 
-             scales = "free_y", 
-             labeller = "label_value",
              nrow = 8) +
@@
   theme(legend.position = "bottom")
+
+moy_group_entire2 <- data_heatmap_moy_group %>% 
+  filter(group>=9) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_both",
+             nrow =  9) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
+  ylab("demand (in Kwh)") +
+  theme_bw() + theme_application() +
+  scale_fill_manual(values=as.vector(tableau20(20)))+
+  scale_color_manual(values=as.vector(tableau20(20)))+
+  theme(legend.position = "bottom")
 ```
 
-```{r data-wnwd-group-entire-1col}
+```{r data-wnwd-group-entire}
+wkndwday_data <- data_group %>%
+  create_gran("wknd_wday") 
+# wkndwday_data$group <- as.factor(wkndwday_data$group)
+# %>% 
+#   create_gran("hour_day")
 
-wkndwday_group_entire <- wkndwday_data%>% 
+ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]
+
+wkndwday_group_entire1 <- wkndwday_data%>% 
+  filter(group<9) %>% 
+  mutate(group = as.factor(group)) %>% 
   ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
@@
   theme(legend.position = "none")
+
+wkndwday_group_entire2 <- wkndwday_data%>% 
+  filter(group>=9) %>% 
+  mutate(group = as.factor(group)) %>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
+  geom_boxplot(aes(fill = group, color = group),alpha = 0.5, outlier.alpha = 0.05)+
+  #geom_boxplot(outlier.size = 1) + 
+  coord_cartesian(ylim = ylim1*1.05)+
+  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
+#geom_boxplot(aes(fill = as.factor(group))) +
+  #scale_fill_lv() +
+ xlab("wknd-wday") + 
+  ylab("demand (in Kwh)") +
+   facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_both",
+             nrow = 9) + 
+  theme_bw() + theme_application() +
+  scale_fill_manual(values=as.vector(tableau20(20)))+
+  scale_color_manual(values=as.vector(tableau20(20)))+
+  theme(legend.position = "none")
 ```
 
-```{r combined-groups-js-entire-1col}
+```{r combined-groups-js-entire, fig.cap = "The distribution of electricity demand for the clusters across hod (a), moy (b) and wkndwday (c).
+It seems like group 2 and 5 have a hod pattern across its members, while group 1, 3, 5 have a moy pattern.
+Wknd-wday variations across groups are not distinguishable, indicating that it is not a critical variable for clustering.
+It is helpful to compare the summarised distributions of groups to that of individuals to confirm that the most of individuals in the group have the same characterisation."}
 
- (hod_group_entire + moy_group_entire+ wkndwday_group_entire) +
-  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
-  plot_layout(guides = "collect")& theme(legend.position = 'none')
+group_entire1 <- (hod_group_entire1 + moy_group_entire1 + wkndwday_group_entire1)&theme_characterisation()& theme(legend.position = 'none')
+
+group_entire2 <- (hod_group_entire2 + moy_group_entire2 + wkndwday_group_entire2)&theme_characterisation()& theme(legend.position = 'none')
+
+ggpubr::ggarrange(group_entire1, group_entire2)
+ #  
+ #  
+ #  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
+ #  plot_layout(guides = "collect")& theme(legend.position = 'none')
+ # 
+ # 
+ # (hod_group_entire/moy_group_entire/ wkndwday_group_entire) +
+ #  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
+ #  plot_layout(guides = "collect")& theme(legend.position = 'none')
 ```

### 2021-11-09 paper-supplementary.Rmd
supplementary material till interaction updated

@@
 
-## Plot displaying missing observations
+(ref:missing-data) Investigating the temporal location of missing values for customers who have implicit missing values.
+There are $13,735$ customers in the data set, with 8,685 having no missing values and the remaining $5,050$ having at least one missing value.
+Each cross represents a missed observation in time, while the line connecting two dots represents continuous missingness over time.
+Missing values occur at random times and do not appear to follow a pattern, although there is a higher concentration of missing values in September and October 2012 for the majority of customers.
+This plot is inspired by @wang2020tsibble.
 
-\noindent The data set contains 13,735 customers, of which 8,685 have no missing values and the remaining $5,050 customers have at least one missing value.
-We wanted to see if there was any structure to these missing observations, so we plotted them over time in Figure \ref{fig:missing-data}.
-Each row represents a customer, with the blank indicating non-missing values and the shapes indicating missing values.
-Figure 9 of @wang2019tsibble inspired this plot.
-For the most part, it appears that data for September and October 2012 are missing.
-Otherwise, missingness can occur at any time and there is no pattern.
-
-```{r missing-data}
+```{r missing-data, fig.cap="(ref:missing-data)"}
 knitr::include_graphics("figs/missing-data-5050.png")
@@
 
-### Interaction of granularities
+<!-- ### Interaction of granularities -->
 
-Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$.
+<!-- Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$.
 In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories.
@@
 D3 and D4 appear similar based on their relative differences across consecutive categories, but $D4$ also changes across facets, unlike $D3$, which has all facets look the same.
+-->
 
-```{r interaction-gran, fig.cap = "Distribution of the simulated variable across $g1$ conditional on $g2$ is shown through boxplots for 4 designs.
+```{r interaction-gran, fig.cap = "The distribution of simulated variable across $g1$ conditional on $g2$ is shown through boxplots for 4 designs to extend the proposed validation designs when two granularities of interest interact.
 D1 has no change in distributions across different categories of $g1$ or $g2$, while D2 and D3 change across only $g1$ and $g2$ respectively.

### 2021-11-09 paper-supplementary.Rmd
sindex for 353 customers update

@@
 
-## Clustering on 350 customers
-
 ```{r data-load-entire}
@@
 id <- c(first_lot[-1], mds_data$name[nrow(mds_data)])
+```
+
+```{r opt-clust-350, fig.cap = "Cluster separation for the 353 customers across the number of clusters is shown.
+When the cluster size changes from 17 to 18, the separation index drops sharply and then flattens out, resulting in the appearance of the elbow.
+Hence, when grouping the 353 customers, the number of clusters is taken to be 17."}
 
@@
 
-library(fpc)
-library(cluster)
 k = array()
@@
 
-ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 50, 2))
+ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 50, 2), minor_breaks = 1) + xlab("number of clusters") + ylab ("separation index")
+```
+
+```{r}
 

### 2021-11-09 paper-supplementary.Rmd
353 customers grouping updated in supplementary

@@
 
-```{r}
+```{r group-353}
 
@@
   theme_application() +
-  scale_fill_manual(values=as.vector(polychrome(20)))+
-  scale_color_manual(values=as.vector(polychrome(20)))+
+  scale_fill_manual(values=as.vector(okabe(n = 8)))+
+  scale_color_manual(values=as.vector(okabe(n = 8)))+
   theme(legend.position = "bottom")
@@
   theme_bw() + theme_application() +
-  scale_fill_manual(values=as.vector(polychrome(20)))+
-  scale_color_manual(values=as.vector(polychrome(20)))+
+  scale_fill_manual(values=as.vector(okabe(n = 8)))+
+  scale_color_manual(values=as.vector(okabe(n = 8)))+
   theme(legend.position = "bottom")
@@
   theme_bw() + theme_application() +
-  scale_fill_manual(values=as.vector(polychrome(20)))+
-  scale_color_manual(values=as.vector(polychrome(20)))+
+  scale_fill_manual(values=as.vector(okabe(n = 8)))+
+  scale_color_manual(values=as.vector(okabe(n = 8)))+
   theme(legend.position = "none")
@@
 
-```{r combined-groups-js-entire, fig.cap = "The distribution of electricity demand for the clusters across hod (a), moy (b) and wkndwday (c).
-It seems like group 2 and 5 have a hod pattern across its members, while group 1, 3, 5 have a moy pattern.
-Wknd-wday variations across groups are not distinguishable, indicating that it is not a critical variable for clustering.
-It is helpful to compare the summarised distributions of groups to that of individuals to confirm that the most of individuals in the group have the same characterisation."}
+```{r combined-groups-js-entire, fig.cap = "The distribution of electricity demand for the clusters across hod, moy and wkndwday for the $17$ groups from $353$ customers.
+Wknd-wday variations across groups are not distinguishable, but ideally each group should have an unique combination of hod and moy.", fig.height=6}
 

### 2021-11-09 paper-supplementary.Rmd
tsne with ellipse update

@@
 knitr::include_graphics("figs/liminal_view.png")
+```
+
+```{r prototype-data-pick}
+```
+
+```{r assemble}
+```
+
+```{r data-pick}
+```
+
+```{r all-data}
+
+```
+
+```{r clustering}
+```
+
+```{r groups-24}
+
+```
+
+```{r tsne-fit}
+```
+
+```{r tsne-plot-supplementary, fig.cap="t-SNE summary of selected $24$ prototype customers.
+Each ellipse corresponds to a group after clustering using our methodology.
+Few customers are not enclosed inside an ellipse as they are from mixed groups and there are too few points to calculate an ellipse around them.
+It is important to note that the selection of prototypes is based on only $50$ percentiles, whereas, the clustering is based on all the deciles."}
+
 ```

### 2021-11-09 script/main.R
tsne with ellipse updated

@@ -1096,7 +1096,7 @@ data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
   gracsr::scale_gran( method = "nqt",
                       response = "general_supply_kwh")
 
-##---tsne-plot
+##---tsne-fit
 
 data_356cust_hod <- read_rds("data/quantile_data_356cust_hod_robust.rds") %>% 
   filter(quantiles %in% "50%")
@@ -1135,6 +1135,9 @@ tSNE_fit <- data_24cust_wide%>%
   Rtsne(pca = FALSE, perplexity = 2)
 
 
+##---tsne-plot
+
+
 tsne_df <- data.frame(tsneX = tSNE_fit$Y[, 1], 
                       tsneY = tSNE_fit$Y[, 2], 
                       customer_id = as.character(data_24cust_wide$customer_id))
@@ -1147,6 +1150,7 @@ tsne_xy <- ggplot(tsne_df, aes(x = tsneX, y = tsneY)) +
   #labs(caption = "tSNE") +
   theme(aspect.ratio = 1) +
   theme_light()+ coord_fixed(ratio=1)
+
 
 
 ##----hod-moy-wkndwday plots
@@ -1679,4 +1683,19 @@ opt_clusters <- ggplot(k %>% as_tibble %>% mutate(k = row_number()),
 tsne_xy + opt_clusters +
   ggpubr::plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')
 
+##----tsne-plot-supplementary
 
+
+tsne_df <- data.frame(tsneX = tSNE_fit$Y[, 1], 
+                      tsneY = tSNE_fit$Y[, 2], 
+                      customer_id = as.character(data_24cust_wide$customer_id)) %>% left_join(cluster_result, by = c("customer_id")) %>% 
+  mutate(group = as.factor(group))
+
+tsne_xy_supplementary <- ggplot(tsne_df) +
+  aes(tsneX, tsneY) +
+  geom_point(size = 1) + 
+  theme_light()+ 
+  coord_fixed(ratio=1)+
+  stat_ellipse( aes(tsneX, tsneY, group = group ), level = 0.85)
+
+tsne_xy_supplementary

### 2021-11-09 paper-supplementary.Rmd
code ready but document not running

@@
 knitr::read_chunk('script/main.R')
+read_chunk(here::here("validation/simdata01.R"))
 ```
@@
 
+# Raw data plot
+
 ```{r raw-data-50, out.width = "100%", fig.cap=" The raw half-hourly energy usage for $50$ sampled households is plotted along the y-axis versus time in a linear scale.
@@
 
+# Missing data plot
 (ref:missing-data) Investigating the temporal location of missing values for customers who have implicit missing values.
@@
 
-_Prototype selection method_
+# Prototype selection
 
@@
 
-```{r mds, out.width="100%", fig.cap="Instance selection using tours and projecting the points in a lower dimensional tsne cloud."}
+```{r mds, out.width="100%", fig.cap="Instance selection using tours and projecting the points in a lower dimensional tsne cloud., eval = FALSE"}
 knitr::include_graphics("figs/liminal_view.png")
@@
 
-<!-- ### Interaction of granularities -->
+# Interaction of granularities
 
-<!-- Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$.
+Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$.
 In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories.
@@
 D3 and D4 appear similar based on their relative differences across consecutive categories, but $D4$ also changes across facets, unlike $D3$, which has all facets look the same.
--->
 
@@
 ```
+
+```{r sim-val}
+set.seed(9999)
+nx_val = 2 # number of x-axis levels
+nfacet_val = 3 # number of facet levels
+w1_val = 3 # increment in mean
+w2_val = 0 # increment in sd
+mean_val = 0 # mean of normal distribution of starting combination
+sd_val = 1 # sd of normal distribution of starting combination
+quantile_prob_val = seq(0.1, 0.9, 0.1)
+```
+
+```{r nobs50}
+ntimes_val = 300 # nobs per combination 
+```
+
+<!-- # Clustering on simulated datasets -->
+
+```{r sample-seed-few}
+sample_seed <-  seq(10, 100, 10)
+```
+
+<!-- _DGP: Generate 10 time series from each designs_  -->
+<!-- Time series are simulated from each of these designs with $50$ observations in each group.
+So we have $50$ observations each for the six combination of categories $(1,1), (1, 2), (2, 1), (2, 2), (3, 1), (3, 2)$.
+Time series are simulated for ten different seeds (from `sample_seed`) for each design.
+`data_varall-1` represents a data set from a design $D_{var_{all}}$ (distributions change across both facet and x) with a seed $10$.
+`data_varall-2` represents a data set from a design $D_{var_{all}}$ with a seed $20$ and so on.
+-->
+<!-- `data_null`, `data_varx` and `data_varf `corresponds to designs $D_{null}$, $D_{var_{x}}$ and $D_{var_{f}}$ designs respectively.    -->
+
+```{r change-index-data-few}
+change_index_data <- function(ntimes_val = NULL,
+                         nx_val = NULL,
+                         nfacet_val = NULL,
+                         sim_function = sim_varx_normal){
+
+  data <- sim_panel(
+    nx = nx_val, nfacet =  nfacet_val,
+    ntimes = ntimes_val,
+    # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
+    sim_dist = sim_function(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+  ) %>% unnest(data)
+
+  index_new <- map(seq_len(ntimes_val), function(i){
+    map((seq_len(nx_val*nfacet_val)), function(j)
+    {
+      value = i + (j-1)*ntimes_val
+    })
+  }) %>% unlist()
+
+  data_new = data %>%
+    ungroup() %>%
+    mutate(index_old = row_number(),
+           index_new = index_new)
+
+  y = data_new[match(index_new, data_new$index_old),]
+
+  y <- y %>%
+    mutate(time = row_number()) %>%
+    select(-c(index_old, index_new))
+
+  return(y)
+}
+```
+
+```{r data-make-few}
+data_null <- map(sample_seed, function(seed){
+  set.seed(seed)
+  change_index_data(5, 2, 3, sim_null_normal) %>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_null")
+
+data_varf <- map(sample_seed, function(seed){
+   set.seed(seed)
+  change_index_data(5, 2, 3, sim_varf_normal)%>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_varf")
+
+data_varx <- map(sample_seed, function(seed){
+   set.seed(seed)
+  change_index_data(5, 2, 3, sim_varx_normal)%>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_varx")
+
+data_varall <- map(sample_seed, function(seed){
+   set.seed(seed)
+  change_index_data(5, 2, 3, sim_varall_normal)%>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_varall")
+
+data_q <- bind_rows(data_null,
+                      data_varf,
+                      data_varx,
+                      data_varall) %>%
+  mutate(seed_id = sprintf("%02d", 
+                              as.numeric(seed_id))) %>% 
+  mutate(unique_data = paste(data_type, seed_id, sep = "-"))
+
+data_q$data_type <- factor(data_q$data_type,
+                                    levels = c("data_null" , "data_varf",
+                                               "data_varx","data_varall"))
+
+```
+
+```{r JS-few}
+
+```
+
+```{r JS-mydata-few}
+data_q_wide <- data_q %>%
+  select(id_facet, id_x, unique_data, sim_data_quantile) %>%
+  pivot_wider(names_from = unique_data,
+              values_from = sim_data_quantile) %>%
+  select(-c(1, 2))
+
+ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
+ldata <- nrow(ndata)
+lcomb <-  nx_val*nfacet_val
+
+dist_data <- map(1:ldata, function(x){ # first data
+  map(1:ldata, function(y){ # 2nd data
+    map(1:lcomb, function(z){ # number of combinations nx*nfacet
+      JS(
+        prob = quantile_prob_val,
+        unlist(data_q_wide[z,x]),
+        unlist(data_q_wide[z,y])
+      ) %>% as_tibble()
+    })%>% bind_rows(.id = "combinations")
+  })%>% bind_rows(.id = "data_type1")
+}) %>% bind_rows(.id = "data_type2")
+
+dist_mat <- dist_data %>%
+  group_by(data_type1, data_type2) %>%
+  summarise(dist = sum(value)) %>%
+  pivot_wider(names_from = data_type2,
+              values_from = dist) %>%
+  mutate(data_type1 = as.numeric(data_type1)) %>%
+  left_join(ndata, by = c("data_type1" = "index"))
+
+dist_mat_format <- dist_mat %>%
+  ungroup() %>%
+  select(-data_type1, -unique_data)
+
+rownames(dist_mat_format) <- dist_mat$unique_data
+```
+
+```{r hier-clust-few}
+d = stats::as.dist(dist_mat_format)
+par(cex=0.5)
+hc = stats::hclust(d,method="complete")
+plot(hc)
+```
+
+```{r mds-plot-interact-few}
+mds <- d %>%
+  cmdscale() %>%
+  as_tibble()
+
+colnames(mds) <- c("Dim.1", "Dim.2")
+rownames(mds) <- dist_mat$unique_data
+
+groups<-cutree(hc, k=4)
+
+all_data_cluster <- cbind(dist_mat_format, groups) %>%
+  cbind(mds)%>%
+  mutate(groups = as.factor(groups)) %>% as_tibble()
+
+mds_plot_10 <- ggplot(all_data_cluster,
+       aes(x = Dim.1,
+           y = Dim.2,
+       color = groups, label = rownames(mds))) +
+  geom_point(size = 1) +
+  theme_classic()+
+  theme(legend.position = "bottom") +
+  scale_color_brewer(palette = "Dark2") +
+  coord_fixed(ratio = 1)+
+  theme(text = element_text(size = 6))
+
+mds_plot_10
+```
+
+<!-- # Clustering on simulated datasets -->
+
+```{r sample-seed}
+sample_seed <-  seq(1, 100, 1)
+```
+
+<!-- _DGP: Generate 10 time series from each designs_  -->
+<!-- Time series are simulated from each of these designs with $50$ observations in each group.
+So we have $50$ observations each for the six combination of categories $(1,1), (1, 2), (2, 1), (2, 2), (3, 1), (3, 2)$.
+Time series are simulated for ten different seeds (from `sample_seed`) for each design.
+`data_varall-1` represents a data set from a design $D_{var_{all}}$ (distributions change across both facet and x) with a seed $10$.
+`data_varall-2` represents a data set from a design $D_{var_{all}}$ with a seed $20$ and so on.
+-->
+<!-- `data_null`, `data_varx` and `data_varf `corresponds to designs $D_{null}$, $D_{var_{x}}$ and $D_{var_{f}}$ designs respectively.    -->
+
+```{r change-index-data}
+change_index_data <- function(ntimes_val = NULL,
+                         nx_val = NULL,
+                         nfacet_val = NULL,
+                         sim_function = sim_varx_normal){
+
+  data <- sim_panel(
+    nx = nx_val, nfacet =  nfacet_val,
+    ntimes = ntimes_val,
+    # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
+    sim_dist = sim_function(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+  ) %>% unnest(data)
+
+  index_new <- map(seq_len(ntimes_val), function(i){
+    map((seq_len(nx_val*nfacet_val)), function(j)
+    {
+      value = i + (j-1)*ntimes_val
+    })
+  }) %>% unlist()
+
+  data_new = data %>%
+    ungroup() %>%
+    mutate(index_old = row_number(),
+           index_new = index_new)
+
+  y = data_new[match(index_new, data_new$index_old),]
+
+  y <- y %>%
+    mutate(time = row_number()) %>%
+    select(-c(index_old, index_new))
+
+  return(y)
+}
+```
+
+```{r data-make}
+data_null <- map(sample_seed, function(seed){
+  set.seed(sample_seed[seed])
+  change_index_data(5, 2, 3, sim_null_normal) %>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_null")
+
+data_varf <- map(sample_seed, function(seed){
+  set.seed(sample_seed[seed])
+  change_index_data(5, 2, 3, sim_varf_normal)%>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_varf")
+
+data_varx <- map(sample_seed, function(seed){
+  set.seed(sample_seed[seed])
+  change_index_data(5, 2, 3, sim_varx_normal)%>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_varx")
+
+data_varall <- map(sample_seed, function(seed){
+  set.seed(sample_seed[seed])
+  change_index_data(5, 2, 3, sim_varall_normal)%>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_varall")
+
+data_q <- bind_rows(data_null,
+                      data_varf,
+                      data_varx,
+                      data_varall) %>%
+  mutate(seed_id = sprintf("%02d", 
+                              as.numeric(seed_id))) %>% 
+  mutate(unique_data = paste(data_type, seed_id, sep = "-"))
+
+data_q$data_type <- factor(data_q$data_type,
+                                    levels = c("data_null" , "data_varf",
+                                               "data_varx","data_varall"))
+
+```
+
+```{r JS}
+
+```
+
+```{r JS-mydata}
+data_q_wide <- data_q %>%
+  select(id_facet, id_x, unique_data, sim_data_quantile) %>%
+  pivot_wider(names_from = unique_data,
+              values_from = sim_data_quantile) %>%
+  select(-c(1, 2))
+
+ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
+ldata <- nrow(ndata)
+lcomb <-  nx_val*nfacet_val
+
+dist_data <- map(1:ldata, function(x){ # first data
+  map(1:ldata, function(y){ # 2nd data
+    map(1:lcomb, function(z){ # number of combinations nx*nfacet
+      JS(
+        prob = quantile_prob_val,
+        unlist(data_q_wide[z,x]),
+        unlist(data_q_wide[z,y])
+      ) %>% as_tibble()
+    })%>% bind_rows(.id = "combinations")
+  })%>% bind_rows(.id = "data_type1")
+}) %>% bind_rows(.id = "data_type2")
+
+dist_mat <- dist_data %>%
+  group_by(data_type1, data_type2) %>%
+  summarise(dist = sum(value)) %>%
+  pivot_wider(names_from = data_type2,
+              values_from = dist) %>%
+  mutate(data_type1 = as.numeric(data_type1)) %>%
+  left_join(ndata, by = c("data_type1" = "index"))
+
+dist_mat_format <- dist_mat %>%
+  ungroup() %>%
+  select(-data_type1, -unique_data)
+
+rownames(dist_mat_format) <- dist_mat$unique_data
+```
+
+```{r hier-clust}
+d = stats::as.dist(dist_mat_format)
+par(cex=0.3)
+hc = stats::hclust(d,method="complete")
+plot(hc)
+```
+
+```{r mds-plot-interact}
+mds <- d %>%
+  cmdscale() %>%
+  as_tibble()
+
+colnames(mds) <- c("Dim.1", "Dim.2")
+rownames(mds) <- dist_mat$unique_data
+
+groups<-cutree(hc, k=4)
+
+all_data_cluster <- cbind(dist_mat_format, groups) %>%
+  cbind(mds)%>%
+  mutate(groups = as.factor(groups)) %>% as_tibble()
+
+mds_plot_10 <- ggplot(all_data_cluster,
+       aes(x = Dim.1,
+           y = Dim.2,
+       color = groups,
+       label = rownames(mds))) +
+  geom_point(size = 0.5) +
+  theme_classic()+
+  theme(legend.position = "bottom") +
+  scale_color_brewer(palette = "Dark2") +
+  coord_fixed(ratio = 1)
+
+mds_plot_10
+```
+
+```{r heatmap-interact, fig.height=8}
+heatmap_raw <- data_q %>% 
+  #select(unique_data, sim_data_quantile, id_facet, id_x) %>% 
+  mutate(category = paste0("facet= ", id_facet, ",x= ", id_x)) %>% 
+  unnest(sim_data_quantile) %>% 
+  group_by(unique_data, category) %>% 
+  mutate(D = row_number()) %>% 
+  arrange(D, category) %>% 
+  left_join(dist_mat, by = c("unique_data")) %>% 
+  select(c(2:9))
+
+data2 <- cbind(dist_mat_format, groups)
+
+heatmap_final <- bind_cols(serial_data = rownames(data2) %>%
+                             as_tibble(),
+          group = data2$groups) %>%
+  left_join(heatmap_raw, by = c("value" = "unique_data")) %>% select(-id_facet, - id_x, - data_type1) %>% 
+  arrange(D, category, group)
+
+heatmap_final$data_type <- factor(heatmap_final$data_type,
+                                   levels = c("data_null" , "data_varf",
+                                              "data_varx","data_varall"))
+
+heatmap_final$category <- reorder(heatmap_final$category, heatmap_final$data_type)
+
+ggplot(heatmap_final,
+       aes(x= reorder(value, data_type), y = as.factor(D))) +
+  geom_tile(aes(fill = sim_data_quantile)) +
+  scale_fill_distiller(palette = "YlGnBu", direction = 1)+ 
+  facet_wrap(~category, ncol = 1) +
+  theme(axis.text.x = element_text(angle = 90)) +
+  ylab("Deciles") +
+  xlab("data sets") +
+  theme(legend.position = "bottom")+
+  scale_x_discrete(breaks = c("data_null" , "data_varf",
+                                              "data_varx","data_varall"))
+
+```
+
+# Clustering all 350 customers
 

### 2021-11-09 paper-supplementary.Rmd
supplementary done

@@
 
-# Raw data plot
-
 ```{r raw-data-50, out.width = "100%", fig.cap=" The raw half-hourly energy usage for $50$ sampled households is plotted along the y-axis versus time in a linear scale.
@@
 
-# Missing data plot
 (ref:missing-data) Investigating the temporal location of missing values for customers who have implicit missing values.
@@
 
-# Prototype selection
+_Prototype selection_
 
@@
 Run t-SNE using the default arguments on the complete data (sets the perplexity to equal 30 and performs random initialisation).
-We then create a linked tour with t-SNE layout with liminal as shown in Figure 4.
+We then create a linked tour with t-SNE layout with R package liminal.
 
@@
 When we brush the corners of the simplex, we find they fall on the edge of the t-SNE point cloud.
-Hall, Marron, and Neeman (2005) have shown that in the extreme case of high-dimension, low-sample size data, observations are on the vertices of a simplex.
 
-This is because in high-dimensional data analysis the curse of dimensionality reasons that points tend to be far away from the center of the distribution and on the edge of high-dimensional space.
+<!-- Hall, Marron, and Neeman (2005) have shown that in the extreme case of high-dimension, low-sample size data, observations are on the vertices of a simplex. -->
+
+<!-- This is because in high-dimensional data analysis the curse of dimensionality reasons that points tend to be far away from the center of the distribution and on the edge of high-dimensional space.
 Contrary to this, is that projected data tends to clump at the center.
+-->
 
@@
 
-```{r mds, out.width="100%", fig.cap="Instance selection using tours and projecting the points in a lower dimensional tsne cloud., eval = FALSE"}
-knitr::include_graphics("figs/liminal_view.png")
+```{r mds, out.width="100%", fig.cap="One instance of brushing in tours (right) and projecting the points in a lower dimensional tsne cloud (left).", out.height="200px"}
+knitr::include_graphics("figs/tour-tsne.png")
 ```
@@
 
-```{r JS-few}
+```{r JS}
 
@@
 rownames(dist_mat_format) <- dist_mat$unique_data
+# write_rds(dist_mat_format, "validation/interact-4D/distmat_40series.rds")
+# write_rds(dist_mat, "validation/interact-4D/distmat_40series_withnames.rds")
 ```
 
-```{r hier-clust-few}
+```{r hier-clust-few, fig.cap= "A dendogram showing the branching of $40$ series with each of the $4$ designs repeated $10$ times."}
 d = stats::as.dist(dist_mat_format)
@@
 plot(hc)
+rect.hclust(hc, k = 4, border = "red")
+
 ```
 
-```{r mds-plot-interact-few}
+```{r mds-plot-interact-few, fig.cap= "MDS summary plot of $40$ series with each of the $4$ designs repeated $10$ times.
+It can be observed that all the series belonging to same design lie closer in this plot and all the ones in different designs are placed further away."}
 mds <- d %>%
@@
            y = Dim.2,
-       color = groups, label = rownames(mds))) +
+       color = groups)) +
+  geom_text(label = rownames(mds),check_overlap = TRUE)+
   geom_point(size = 1) +
@@
 
-```{r JS}
-
-```
-
 ```{r JS-mydata}
@@
 rownames(dist_mat_format) <- dist_mat$unique_data
+# write_rds(dist_mat_format, "validation/interact-4D/distmat_400series.rds")
+# write_rds(dist_mat, "validation/interact-4D/distmat_400series_withnames.rds")
 ```
 
-```{r hier-clust}
+```{r hier-clust, fig.cap = "A dendogram showing the branching of $400$ series with each of the $4$ designs repeated $100$ times."}
 d = stats::as.dist(dist_mat_format)
-par(cex=0.3)
+par(cex=0.5)
 hc = stats::hclust(d,method="complete")
 plot(hc)
+rect.hclust(hc, k = 4, border = "red")
 ```
 
-```{r mds-plot-interact}
+```{r mds-plot-interact, fig.cap= "MDS summary plot of $400$ series with each of the $4$ designs repeated $100$ times.
+It can be observed that all the series belonging to same design lie closer in this plot and all the ones in different designs are placed further away."}
 mds <- d %>%
@@
            y = Dim.2,
-       color = groups,
-       label = rownames(mds))) +
+       color = groups)) +
+  #geom_text(label = rownames(mds),check_overlap = TRUE)+
   geom_point(size = 0.5) +
@@
 
-```{r heatmap-interact, fig.height=8}
+```{r heatmap-interact, fig.height=8, fig.cap="A heatmap showing the cluster summary of $400$ series.
+Deciles of the clusters are plotted along the y-axis with each group on the x-axis faceted by each combination of the interacting granularity.
+The four distinct behavior across clusters are evident in all the combinations, with the first block corresponding to $data_null$.
+The deciles of the second block changes across both facet and x-axis.
+For similar reasons, the second, third and forth blocks correspond $data_varall$, $data_varf$ and $data_varx$ respectively." }
 heatmap_raw <- data_q %>% 
@@
 
-# Clustering all 350 customers
+_Clustering all 350 customers_
 

### 2021-11-09 supplementary/algorithm1/real-data/cluster4-iter100.Rmd
supplementary done

--- 2021-08-18 supplementary/algorithm1/real-data/cluster4-iter100.Rmd
@@
 library(knitr)
-read_chunk("R/simdata01.R")
+library(here)
+read_chunk(here::here("validation/simdata01.R"))
 ```
@@
 ```{r nobs50, echo = TRUE}
-ntimes_val = 50 # nobs per combination 
+ntimes_val = 300 # nobs per combination 
 ```

### 2021-11-09 validation/simdata01.R

@@ -1,0 +1,254 @@
+# scripts to see how the time series for the four designs look like
+
+##----load-lib
+library(hakear)
+library(tidyr)
+library(distributional)
+library(ggplot2)
+library(patchwork)
+library(purrr)
+library(dplyr)
+library(tidyverse)
+library(gravitas)
+library(magrittr)
+## ----designs
+
+sim_null_normal <- function(nxj, nfacetj, mean, sd, w1 = 0, w2=0) {
+  rep(distributional::dist_normal(mu = mean, sigma = sd),
+      times = nxj * nfacetj
+  )
+}
+
+sim_varf_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  rep(dist_normal((mean + seq(0, nfacet - 1, by = 1) * w1),
+                  (sd + seq(0, nfacet - 1, by = 1) * w2)), each = nx)
+}
+
+sim_varx_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  rep(dist_normal((mean + seq(0, nx - 1, by = 1) * w1),
+                  (sd + seq(0, nx - 1, by = 1) * w2)), nfacet)
+}
+
+sim_varall_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  dist_normal((mean + seq(0, (nx *nfacet - 1), by = 1) * w1),
+              (sd + seq(0,  (nx *  nfacet - 1), by = 1) * w2))
+}
+
+
+## ----data
+
+
+sim_panel_varall <- sim_panel(
+  nx = nx_val, nfacet =  nfacet_val,
+  ntimes = ntimes_val,
+  # sim_dist = sim_varall_normal(2, 3, 5, 10, 5, -1.5)
+  sim_dist = sim_varall_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+) %>% unnest(data)
+
+sim_panel_varx <- sim_panel(
+  nx = nx_val, nfacet =  nfacet_val,
+  ntimes = ntimes_val,
+  # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
+  sim_dist = sim_varx_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+) %>% unnest(data)
+
+sim_panel_varf <- sim_panel(
+  nx = nx_val, nfacet =  nfacet_val,
+  ntimes = ntimes_val,
+  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
[…]

### 2021-11-10 paper-supplementary.Rmd
supplementary plots done

@@
 It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed.
-No insightful behavioral pattern could be discerned from this view other than when the customer is not at home."}
+No insightful behavioral pattern could be discerned from this view other than when the customer is not at home.", fig.pos="h"}
 
@@
 
-```{r missing-data, fig.cap="(ref:missing-data)"}
+```{r missing-data, fig.cap="(ref:missing-data)", fig.pos="h"}
 knitr::include_graphics("figs/missing-data-5050.png")
@@
 
-# Interaction of granularities
+<!-- # Interaction of granularities -->
 

### 2021-11-10 script/gracsr-supplementary.R

@@ -661,7 +661,7 @@ wkndwday_group_entire2 <- wkndwday_data%>%
   scale_color_manual(values=as.vector(tableau20(20)))+
   theme(legend.position = "none")
 
-##----combined-groups-js-entire
+## ---- combined-groups-js-entire
 group_entire1 <- (hod_group_entire1 + moy_group_entire1 + wkndwday_group_entire1)&theme_characterisation()& theme(legend.position = 'none')
 
 group_entire2 <- (hod_group_entire2 + moy_group_entire2 + wkndwday_group_entire2)&theme_characterisation()& theme(legend.position = 'none')

### 2021-11-10 main.Rmd
tsne with ellipse updated
Adding what is sindex

@@
 
-- _Clustering algorithm_
+- _Clustering_
 
-With a way to obtain pairwise distances, any clustering algorithm can be employed that supports the given distance metric as input.
+- Number of clusters: Different approaches to determining cluster size are required for varied applications and research objectives.
+Some typical ways of determining the number of clusters in the literature (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy]) are inspired primarily by defining and balancing "within-cluster homogeneity" and "between-cluster separation." Cluster separation index ($sindex$) proposed by @Hennig2014-ah is employed in this paper to have less dependency on a single or few ambiguous points.
+$sindex$ [@R-fpc] is determined by calculating the distance between each observation and the cluster to which it does not belong.
+Then, the least 10% of these distances are averaged to only evaluate places near the cluster edges.
+$sindex$, like other statistics in the literature, cannot be optimized over the number of clusters, since increasing the number of clusters reduces it.
+The number of clusters required is determined by the slope going from steep to shallow (an elbow).
+
+- ALgorithm: With a way to obtain pairwise distances, any clustering algorithm can be employed that supports the given distance metric as input.
 A good comprehensive list of algorithms can be found in @Xu2015-ja based on traditional ways like partition, hierarchy, or more recent approaches like distribution, density, and others.
@@
 
+```{r tsne-fit}
+
+```
+
 ```{r tsne-plot}

### 2021-11-10 main.Rmd
added labels for raw distribution plot

@@
 (Refer to Supplementary article for further details).
-Figure \ref{fig:hod-ind-group} shows the distribution across `hod` (a), `moy`(b) and `wnwd` (c) for the set of chosen $24$ customers that were chosen.
+Figure \ref{fig:hod-ind-group-png} shows the distribution across `hod` (a), `moy`(b) and `wnwd` (c) for the set of chosen $24$ customers that were chosen.
 Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution.
@@
 For clustering based on JS-based distances, we chose the optimal number of clusters using (@Hennig2014-ah) as $5$.
-The groupings are shown in Figure \ref{fig:hod-ind-group}.
+The groupings are shown in Figure \ref{fig:hod-ind-group-png}.
 Customers with the same color represent the same clustered groups.
@@
 
-```{r hod-ind-group, fig.cap = "The distribution of selected customers over hod (a), moy (b), and wnwd (c) for the $24$ selected customers is shown.
+```{r hod-ind-group-png, fig.cap = "The distribution of selected customers over hod (a), moy (b), and wnwd (c) for the $24$ selected customers is shown.
 In each case, the same colour denotes the same group using js-based clustering and are placed together to facilitate comparison.
 Individuals with similar distributional differences across categories of hod or moy are grouped together, with some exceptions.
-Distributional differences across categories of wnwday are not discernable except for a couple of customers.", fig.height =  3.5}
+Distributional differences across categories of wnwday are not discernable except for a couple of customers."}
 ```

### 2021-11-10 script/main.R
remove sindex for wpd

@@ -55,16 +55,17 @@ theme_validation <- function() {
 theme_characterisation <- function() {
   
   theme_bw() + # seeting theme
-    theme(strip.text = element_text(size = 10,
+    theme(strip.text = element_text(size = 14,
                                     margin = margin(b = 0, t = 0))) + # narrow facet space
     theme(axis.title.y=element_blank(),
           axis.text.y=element_blank(),
           axis.ticks.y=element_blank()) + # no axis ticks 
     theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
-    theme(axis.text.x = element_text(angle=90, hjust=1, size = 10)) + # rotate the x-axis text
+    theme(axis.text.x = element_text(angle=90, hjust=1, size = 14)) + # rotate the x-axis text
     theme(legend.position = "bottom")+
     theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-    theme(axis.text.x = element_text(size=5))
+    theme(axis.text.x = element_text(size=5)) +
+    theme(axis.text = element_text(size = 12))  
 }
 
 
@@ -93,9 +94,24 @@ theme_application <- function() {
 
 ##----mytheme-application2
 theme_application2 <- function() {
-  theme_characterisation() +     theme(strip.background = element_blank(),
+  theme_characterisation() +   
+    theme(strip.background = element_blank(),
                                        strip.text.x = element_blank())
 
+}
+
+##----mytheme-application3
+theme_application3 <- function() {
+  theme_characterisation() +
+    theme(
+      panel.grid.major.y = element_blank(),
+      panel.grid.minor.y = element_blank()
+    )+ 
+    theme(axis.ticks.x = element_blank(),
+          axis.text.x = element_blank()) +
+    theme(
+          strip.text.y = element_blank())
+  
 }
 
 ##----format-data
@@ -1385,14 +1401,14 @@ hod_group <- data_heatmap_hod_group %>%
                 color = as.factor(group)), size = 1)+
   facet_wrap(~group, 
              scales = "free_y",  
-             nrow = 5) + 
+             nrow = 5, labeller = "label_value") + 
[…]

### 2021-11-10 main.Rmd
remove sindex for wpd
PCP of application update

@@
 ```{r mytheme-application}
+```
+
+```{r mytheme-application3}
+
 ```
@@
 
-```{r combined-groups-js, fig.cap = "The distribution of electricity demand over hod (a), moy (b), and hod (conditional on wnwd) (c).
+```{r combined-groups-js-big2, fig.cap = "The distribution of electricity demand over hod (a), moy (b), and hod (conditional on wnwd) (c).
 Groups 2 and 5 appear to have a hod pattern among its members (morning and evening peak), whereas groups 1, 3, and 5 appear to have a moy pattern (higher usage in mid-year).
@@
 All of these could be observed from the plot or the table (b) which shows median wpd values for each group.", message=FALSE, warning=FALSE}
-
+include_graphics("figs/parcoord-application.png")
 ```

### 2021-11-10 script/main.R

@@ -1457,7 +1457,7 @@ wkndwday_group <- wkndwday_data%>%
   #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
   #geom_boxplot(aes(fill = as.factor(group))) +
   #scale_fill_lv() +
-  xlab("wnwd") + 
+  xlab("hod conditional by wdwnd") + 
   ylab("demand (in Kwh)") +
   facet_grid(group~wknd_wday, 
              scales = "free_y", 
@@ -1465,7 +1465,7 @@ wkndwday_group <- wkndwday_data%>%
   theme_bw() + theme_application3() +
   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
   scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +
-  theme(legend.position = "none") 
+  theme(legend.position = "none") + theme(strip.text.x = element_text(size=0))
 
 
 ##----combined-groups-js-big2
@@ -1516,10 +1516,10 @@ data_table <- data_pcp %>% group_by(group) %>%
 
 rownames(data_table) <- c("group-1", "group-2", "group-3")
 
-##----parcoord
+##----parcoord-application
 parcoord <- GGally::ggparcoord(data_pcp %>% left_join(cluster_result_id, by = "customer_id") ,
                                columns = 3:5,
-                               groupColumn = "group",
+                               groupColumn = "group.x",
                                showPoints = FALSE, 
                                alphaLines = 0.8,
                                order = "anyClass",
@@ -1534,9 +1534,9 @@ parcoord <- GGally::ggparcoord(data_pcp %>% left_join(cluster_result_id, by = "c
   ylab("wpd") + scale_fill_viridis_d(direction = 1) +
   scale_color_viridis_d(direction = 1) + theme_light()
 
-parcoord + geom_text(aes(label = id))
 
-(parcoord + gridExtra::tableGrob(data_table))+ plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')') & theme(legend.position = "bottom")
+
+(parcoord + geom_text(aes(label = id)) + gridExtra::tableGrob(data_table))+ plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')') & theme(legend.position = "bottom")
 
 ##----opt-clusters-jsd
 

### 2021-11-10 main.Rmd
hod-ind-group-png  update

@@
 
-```{r hod-ind-group-png, fig.cap = "The distribution of selected customers over hod (a), moy (b), and wnwd (c) for the $24$ selected customers is shown.
+```{r hod-ind-group-png, fig.cap = "The distribution of selected customers over multiple granularities `hod`, `moy`, and `wnwd` are shown for the $24$ selected customers using quantile and box plots.
+The $24$ customers are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer.
+Each customer is identified by their profile across all the granularities.
+A few of these customers have similar distributions across `moy` and some are similar in their `hod distributions.
+Distributional differences across categories of wnwd are not discernable except for a couple of customers (id: 16 and 18)."}
+```
+
  In each case, the same colour denotes the same group using js-based clustering and are placed together to facilitate comparison.
-Individuals with similar distributional differences across categories of hod or moy are grouped together, with some exceptions.
-Distributional differences across categories of wnwday are not discernable except for a couple of customers."}
-```
 

### 2021-11-10 main.Rmd
summarized distribution caption update

@@
 Few groups have mixed patterns across `hod` and `moy`, but few have all customers in the group having a similar profile.
-Figure \ref{fig:combined-groups-js} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
+Figure \ref{fig:combined-groups-js-big2} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
 It shows Groups $2$ and $4$ have `hod` pattern with a typical morning and evening peak, whereas groups $1$, $3$, and $5$ show a `moy` pattern with higher usage in winter months.
@@
 
-```{r hod-ind-group-png, fig.cap = "The distribution of selected customers over multiple granularities `hod`, `moy`, and `wnwd` are shown for the $24$ selected customers using quantile and box plots.
-The $24$ customers are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer.
+```{r hod-ind-group-png, fig.cap = "The distribution of electricity demand across individual customers over multiple granularities `hod`, `moy`, and `wnwd` are shown for the $24$ selected customers using quantile and box plots.
+They are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer.
 Each customer is identified by their profile across all the granularities.
@@
 ```
-
- In each case, the same colour denotes the same group using js-based clustering and are placed together to facilitate comparison.
 
@@
 
-```{r combined-groups-js-big2, fig.cap = "The distribution of electricity demand over hod (a), moy (b), and hod (conditional on wnwd) (c).
+```{r combined-groups-js-big2, fig.cap = "The distribution of electricity demand across groups over `hod` (a), `moy` (b), and `hod` (conditional on `wnwd`) (c) are shown through quantile plots.
+Group 1 (1-3), Group 2 (4-9), Group 3 (10 - 15), Group 4 (16- 20), Group 5 (21 - 24).
 Groups 2 and 5 appear to have a hod pattern among its members (morning and evening peak), whereas groups 1, 3, and 5 appear to have a moy pattern (higher usage in mid-year).
@@
 This might be because most of these chosen customers do not have a wnwd pattern.
-It is useful to compare the summarised distributions of groups to those of individuals in order to establish that the majority of individuals in the group share the same characteristics."}
+It is useful to compare the raw distributions with the summarized distributions to establish that the majority of individuals in the group share the same characteristics.
+To facilitate this comparison, the individual distributions are sorted by their JS-based groupings."}
 ```

### 2021-11-10 main.Rmd
parcoord-application update

@@
 
-```{r opt-cluster-tsne-jsd, fig.cap = "(a) shows t-SNE summary of the $24$ selected customers chosen through prototype selection, (b) shows a plot of sindex for different cluster size."}
+```{r opt-cluster-tsne-jsd, fig.cap = "(a) shows t-SNE summary of the $24$ selected customers chosen through prototype selection, (b) shows a plot of sindex for different cluster size for clustering based on JS-based distances.
+The separation index decreases sharply from $5$ to $6$ clusters, so $5$ clusters are used here."}
 
@@
 
-```{r parcoord, fig.cap="Each of the 24 customers is represented by a parallel coordinate plot (a) with three wpd-based groupings.
-The plot shows that moy is the most important variable in identifying clusters, whereas wkdn-wday is the least significant and has the least fluctuation.
-One particular customer with high wpd across wknwday stands out in this display.
+```{r parcoord, fig.cap="Each of the $24$ customers is represented by a parallel coordinate plot (a) with three wpd-based groupings.
+Plot (b) which shows median wpd values for each group.
+The plot shows that `moy` is the most important variable in identifying clusters, whereas `wnwd` is the least significant and has the least fluctuation.
+Two customers (id: 16 and 18) with high wpd across `wnwd` stand out in this display.
 Group 3 has a higher wpd for hod than moy or wkndwday.
@@
 Group 1 is a mixed group with strong patterns on atleast one of the three variables.
-All of these could be observed from the plot or the table (b) which shows median wpd values for each group.", message=FALSE, warning=FALSE}
+All of these could be observed from the plot or the table (b).", message=FALSE, warning=FALSE}
 include_graphics("figs/parcoord-application.png")

### 2021-11-10 main.Rmd
caption in application updated

@@
 Few groups have mixed patterns across `hod` and `moy`, but few have all customers in the group having a similar profile.
-Figure \ref{fig:combined-groups-js-big2} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
+Figure \ref{fig:combined-groups-js} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
 It shows Groups $2$ and $4$ have `hod` pattern with a typical morning and evening peak, whereas groups $1$, $3$, and $5$ show a `moy` pattern with higher usage in winter months.
@@
 
-```{r hod-ind-group-png, fig.cap = "The distribution of electricity demand across individual customers over multiple granularities `hod`, `moy`, and `wnwd` are shown for the $24$ selected customers using quantile and box plots.
+```{r hod-ind-group-png, fig.cap = "The distribution of electricity demand across individual customers over multiple granularities hod, moy, and wnwd are shown for the $24$ selected customers using quantile and box plots.
 They are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer.
 Each customer is identified by their profile across all the granularities.
-A few of these customers have similar distributions across `moy` and some are similar in their `hod distributions.
+A few of these customers have similar distributions across moy and some are similar in their hod distributions.
 Distributional differences across categories of wnwd are not discernable except for a couple of customers (id: 16 and 18)."}
@@
 
-```{r combined-groups-js-big2, fig.cap = "The distribution of electricity demand across groups over `hod` (a), `moy` (b), and `hod` (conditional on `wnwd`) (c) are shown through quantile plots.
+```{r combined-groups-js, fig.cap = "The distribution of electricity demand across groups over hod (a), moy (b), and hod (conditional on wnwd) (c) are shown through quantile plots.
 Group 1 (1-3), Group 2 (4-9), Group 3 (10 - 15), Group 4 (16- 20), Group 5 (21 - 24).
@@
 Plot (b) which shows median wpd values for each group.
-The plot shows that `moy` is the most important variable in identifying clusters, whereas `wnwd` is the least significant and has the least fluctuation.
-Two customers (id: 16 and 18) with high wpd across `wnwd` stand out in this display.
+The plot shows that moy is the most important variable in identifying clusters, whereas wnwd is the least significant and has the least fluctuation.
+Two customers (id: 16 and 18) with high wpd across wnwd stand out in this display.
 Group 3 has a higher wpd for hod than moy or wkndwday.

### 2021-11-10 check.R

@@ -1,0 +1,14 @@
+collapse_x <- function(x)
+{
+  y = unlist(x)
+  z = paste(y, collapse = ",")
+}
+
+x = list(1, 5, 2, c(1,5))
+cluster_summary <- cluster_result_id %>% select(-customer_id, -divide_cust) %>%  group_by(group) %>% nest(data = c(id))
+
+apply(cluster_summary, 1, collapse_x)
+z =  rbind(group = cluster_summary$group, rbind(members = map(cluster_summary$data, collapse_x))) %>% t()
+
+z =  rbind(group = cluster_summary$group, rbind(members = map(cluster_summary$data, collapse_x))) %>% t()
+

### 2021-11-10 main.Rmd
abstract revised

@@
 abstract: |
- With more and more time series data being collected at much finer temporal resolution, for a longer length of time, and for a larger number of individuals/entities, time series clustering research is getting a lot of traction.
-Long, noisy, patchy, uneven, and asynchronous time series are common in many fields, limiting similarity searches or lowering method efficiency when clustering is based on a distance metric.
-In this work, we suggest two approaches for obtaining similarity between time series based on probability distributions over cyclic temporal granularities for distance-based clustering approaches.
-Cyclic granularities like hour-of-the-day, work-day/weekend, month-of-the-year and so on, are useful for finding repeated patterns in the data.
-Looking at probability distributions across cyclic granularities serves two purposes: (a) "Probability distributions" characterise the inherent temporal data structure of these large unequal-length time series and are robust to missing or noisy data.
-(b) Using probability distributions over "cyclic granularities" ensures small pockets of similar "repeated" behaviors.
-Our method is capable of producing useful clusters, as demonstrated by testing on validation data designs and a sample of residential smart meter consumers.
+ Clustering is a potential approach for organising large collections of times series into small homogeneous groups, but a difficult step is determining an appropriate metric to measure similarity between time series.
+The similarity metric needs to be capable of accommodating long, noisy, and asynchronous time series and also capture cyclical patterns.
+In this paper, two approaches for measuring similarity between time series are presented, based on probability distributions over cyclic temporal granularities.
+Both are compatible with a variety of clustering algorithms.
+Cyclic granularities like hour-of-the-day, work-day/weekend, month-of-the-year, are useful for finding repeated patterns in the data.
+Measuring similarity based on probability distributions across cyclic granularities serves two purposes: (a) characterising the inherent temporal data structure of long, unequal-length time series in a manner robust to missing or noisy data; (b) small pockets of similar repeated behaviors can be captured.
+This approach is capable of producing useful clusters, as demonstrated on validation data designs and a sample of residential smart meter records.
 

### 2021-11-10 supplementary/distance-computation/real-data/01select_clean_cust_from600cust.R

@@ -13,7 +13,9 @@ elec <- read_rds(here::here("data/elec_nogap_2013_600.rds"))%>%
 elec_iqr0 <- elec %>% as_tibble() %>% group_by(customer_id) %>%
   summarise(q2 = stats::median(general_supply_kwh, na.rm = TRUE),
             iqr = stats::IQR(general_supply_kwh, na.rm = TRUE),
-            quantile = stats::quantile(general_supply_kwh, na.rm = TRUE, prob = seq(0.1, 0.9, 0.1))) %>% filter(iqr==0)
+            quantile = stats::quantile(general_supply_kwh, 
+                  na.rm = TRUE, prob = seq(0.1, 0.9, 0.1))) %>% 
+  filter(iqr==0)
 
 elec_sub <- elec %>%
   filter(!(customer_id %in%

### 2021-11-11 main.Rmd
done with intro

@@
 Defining an appropriate distance measure for the raw time series may be a difficult task since it must take into account noise, variable lengths of time series, asynchronous time series, different scales, and missing data.
-Commonly used Distance-based similarity measures as suggested by a decade review of time series clustering approaches (@Aghabozorgi2015-ct) are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping, Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances (@liao2007clustering) but on a vector time series data.
+Commonly used distance-based similarity measures as suggested by a decade review of time series clustering approaches (@Aghabozorgi2015-ct) are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping (DTW), Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances (@liao2007clustering) but on a vector time series data.
 Among these alternatives, Euclidean distances have high performance but need the same length of data over the same period, resulting in information loss regardless of whether it is on raw data or a smaller collection of features.
@@
 
-\noindent We consider the problem of clustering a large number of univariate time series of continuous values which are available at fine temporal scales, being motivated by the residential smart meter data.
-These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and missing observations and characterized by noisy and patchy behavior that can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
-Choosing probability distributions seem to be a natural way to analyze these types of data sets since they are robust to uneven length, missing data, or noise.
-This paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across significant cyclic granularities.
-Cyclic temporal granularities [@Gupta2021-hakear], which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
-The resulting clusters are expected to group customers that have similar repetitive behaviors across each of the interesting cyclic granularities.
-Below are some of the benefits of our method:
+\noindent This work is motivated by a need to cluster a large collection of residential smart meter data, so that households can be grouped into similar energy usage patterns.
+These can be considered to be univariate time series of continuous values which are available at fine temporal scales.
+These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and sporadic missing values.
+Using probability distributions is a natural way to analyze this types of data because they are robust to uneven length, missing data, or noise.
+This paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across a selection of cyclic granularities.
+Cyclic temporal granularities [@Gupta2021-hakear], which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can measure repetitive patterns in large univariate time series data.
+The resulting clusters are expected to group customers that have similar repetitive behaviors across cyclic granularities.
+The benefits of this approach are as follows.
 
-- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern);
+- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern).
 
@@
 
-- Jensen-Shannon distances evaluate the distance between two distributions rather than raw data, which is less sensitive to missing observations and outliers than other conventional distance methods;
+- Jensen-Shannon distances evaluate the distance between two distributions rather than raw data, which is less sensitive to missing observations and outliers than other conventional distance methods.
 
-- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar distributions across all interesting cyclic granularities;
+- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar distributions across all interesting cyclic granularities.
 

### 2021-11-11 script/main.R

@@ -33,7 +33,6 @@ library(grid)
 load("data/ALL_DATA.Rdata")
 
 
-
 ##----theme-validation
 theme_validation <- function() {
   theme_bw() +

### 2021-11-11 main.Rmd
background

@@
 
-_Background and motivation_
+<!--_Background and motivation_-->
 
-Large spatio-temporal data sets, both from open and administrative sources, offer up a world of possibilities for research.
-One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) The project provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014.
-Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.
-Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
+The primary application of this work is data from the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
+Half-hourly measurements of usage for more than 13,000 household electricity smart meters is provided from from October 2011 to March 2014.
+<!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
 The behavioral patterns differ amongst customers due to many temporal dependencies.
 Some households use a dryer, while others dry their clothes on a line.
-Their weekly profile may reflect this.
+Their weekly usage profile may reflect this.
 They may vary monthly, with some customers using more air conditioners or heaters than others, while having equivalent electrical equipment and weather circumstances.
 Some customers are night owls, while others are morning larks.
-Day-off energy use varies depending on whether customers stay home or go outside.
-Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior a fascinating one.
-This challenge is worsened when all we know about our consumers is their energy use history (@Ushakova2020-rl).
-To safeguard the customers' privacy, it is probable that such information is not accessible.
-Also, energy suppliers may not always update client information, such as property features, in a timely manner.
-Thus, there is a growing need to have research that examines how much energy usage heterogeneity can be found in smart meter data and what are some of the most common power consumption patterns, rather than explaining why consumption differs.
+Daily energy usage varies depending on whether customers stay home or work away from home.
+Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior complex.
+The challenge is to be able to cluster consumers into these type of expected patterns, and other unexpected patterns, using only their energy usage history (@Ushakova2020-rl).
+<!--To safeguard the customers' privacy, it is probable that such information is not accessible.
+Also, energy suppliers may not always update client information, such as property features, in a timely manner.--> There is a growing need to have methods that can examine the energy usage heterogeneity observed in smart meter data and what are some of the most common power consumption patterns.
 

### 2021-11-11 main.Rmd
done with intro - full intro

@@
 
-_Related work_
+<!-- _Related work_-->
 
-A multitude of papers have emerged around smart meter time series clustering for deepening our knowledge of consumption patterns.
+There is a growing body of literature focused on time series clustering related to smart meter data.
 @Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
-None of the $34$ articles chosen for their emphasis use Australian smart meter data.
-The most often used algorithm is K-Means.
-Using K-Means without considering time series structure or correlation results in inefficient clusters.
-Principal Component Analysis (PCA) or Self-Organizing Maps (SOM) eliminate correlation patterns and decrease feature space, but lose interpretability.
-To reduce dimensionality, several studies use principal component analysis or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).
-Other algorithms utilized in the literature include k-means variants, hierarchical approaches, and greedy k-medoids.
+<!--None of the $34$ articles chosen for their emphasis use Australian smart meter data.--> The most often used algorithm is $k$-means (XXX reference something here so the reader can learn about k-means if they don't already know).
+Using $k$-means without considering time series structure or correlation fails to extract cyclical patterns (XXX???
+or do you mean to say that it can suffer from being a large p, small n problem??).
+To reduce dimensionality, several studies use principal component analysis (PCA) or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).
+PCA eliminates correlation patterns and decreases feature space, but loses interpretability.
+Other algorithms utilized in the literature include $k$-means variants, hierarchical clustering, and greedy $k$-medoids.
 Time series data, such as smart meter data, are not well-suited to any of the techniques mentioned in @Tureczek2017-pb.
-Only one study [@ozawa2016determining] identified time series characteristics using Fourier transformation, which converts data from time to frequency and then uses K-Means to cluster by greatest frequency.
-@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential and addresses model-based clustering by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered.
+Only one study [@ozawa2016determining] identified time series characteristics by first conducting a Fourier transformation, to convert data from time to frequency domain, followed by $k$-means to cluster by greatest frequency.
+@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential customers and addresses model-based clustering by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered.
 [@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
 @Melnykov2013-sp discusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
-To our knowledge, none of the methods focus on exploring heterogeneity in repetitive behaviors based on the dynamics of multiple temporal dependencies using probability distributions.
+None of these methods focuses on exploring heterogeneity in repetitive patterns based on the dynamics of multiple temporal dependencies using probability distributions, which forms the basis of the methodology reported here.
 
@@
 
-The remainder of the paper is organized as follows: Section&nbsp;\ref{sec:methodology} provides the clustering methodology.
-Section&nbsp;\ref{sec:validation} shows data designs to validate our methods.
-Section&nbsp;\ref{sec:application} discusses the application of the method to a subset of the real data.
-Finally, we summarize our results and discuss possible future directions in Section&nbsp;\ref{sec:discussion}.
+This paper is organized as follows.
+Section~\ref{sec:methodology} provides the clustering methodology.
+Section~\ref{sec:validation} shows data designs to validate our methods.
+Section~\ref{sec:application} discusses the application of the method to a subset of the real data.
+Finally, we summarize our results and discuss possible future directions in Section~\ref{sec:discussion}.
 

### 2021-11-11 main.Rmd
finished selecting granularities

@@
 
-The foundation of our method is unsupervised clustering algorithms based exclusively on the time-series data.
-The proposed methodology aims to leverage the intrinsic data structure hidden within cyclic temporal granularities.
-The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
-<!--So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.--> In this work, instead of considering the probability distributions of the linear time series, we assume it across different categories of any cyclic granularity.
+<!--The foundation of our method is unsupervised clustering algorithms based exclusively on the time-series data.
+The proposed methods leverage the intrinsic data structure hidden within cyclic temporal granularities.--> The existing work on clustering probability distributions assumes we have an independent and identically distributed samples $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+<!--So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.--> In our approach, instead of considering the probability distributions of the linear time series, we compare them across different categories of a cyclic granularity.
 We can consider categories of an individual cyclic granularity ($A$) or combination of categories for two interacting granularities ($A*B$) to have a distribution, where $A, B$ are defined as $A = \{a_j: j=1, 2, \dots J\}$ and $B = \{b_k: k = 1, 2, \dots K\}$.
-For example, let us consider two cyclic granularities $A$ and $B$ representing hour-of-day and day-of-week.
+For example, let us consider two cyclic granularities, $A$ and $B$, representing hour-of-day and day-of-week, repsectively.
 Then $A = \{0, 1, 2, \dots, 23\}$ and $B = \{Mon, Tue, Wed, \dots, Sun\}$.
@@
 Hence clustering these customers is equivalent to clustering these collections of conditional univariate probability distributions.
-Towards this goal, we need to decide how to measure similarities between collections of univariate probability distributions.
-There are multiple ways to measure similarities depending on the aim of the analysis.
+Towards this goal, the next step is to decide how to measure similarities between collections of univariate probability distributions.
+<!--There are multiple ways to measure similarities depending on the aim of the analysis.
 <!--This paper focuses on looking at the (dis) similarity between underlying distributions that may have resulted in different patterns across different cyclic temporal granularities, that eventually have resulted in the (dis) similarity between time series.
---> This paper considers a two approaches for finding distances between time series.
-Both of these approaches may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
-The obtained distances could be fed into a clustering algorithm to break large data sets into subgroups that can then be analyzed separately.
+--> Here, we describe two approaches for finding distances between time series.
+Both of these approaches may be useful in a practical context, and produce very different but equally useful customer groupings.
+The distances can be supplied to any usual clustering algorithm, including $k$-means or hierarchical clustering, to group observations into a smaller more homogeneous collection.
 <!--These clusters may be commonly associated with real-world data segmentation.
-However, since the data is unlabeled a priori, more information is required to corroborate this.--> The methodology is explained in the Figure \ref{fig:flowchart} and each element of the pipeline is discussed.
+However, since the data is unlabeled a priori, more information is required to corroborate this.--> The flow of the procedures is illustrated in Figure \ref{fig:flowchart}.
 
-```{r flowchart, fig.cap = "Flow chart illustrating the pipeline for methodology", out.width="100%"}
+```{r flowchart, fig.cap = "Flow chart illustrating the pipeline for out method for clustering time series.", out.width="100%"}
 knitr::include_graphics("figs/flowchart.png")
@@
 
-- _Find significant granularities or harmonies_
+## Selecting granularities
 
-[@Gupta2021-hakear] proposes a method for choosing significant cyclic granularities and harmonies (interacting granularities that can be studied together) [@Gupta2021-gravitas], which is used in this work.
-We define "significant" granularities as those with significant distributional differences across categories.
-It is preferable to consider those since it is more probable that there will be some intriguing repeated behavior worth investigating.
-It should be noted that all of the observations in the study may not have the same set of important granularities.
-The following is a method for generating a list ($S_c$) of significant granularities for all observations:
+[@Gupta2021-hakear] provides a method for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
+(This extends to harmonies, pairs of granularities that might interact with each other.) We define "significant" granularities as those with significant distributional differences across at least one category.
+The reason for subsetting granularities in this way, is that, clustering algorithms perform badly in the presence of nuisance variables.
+Granularities that do not have some difference between categories are likely to be nuisance variables.
+It should be noted that all of the time series in a collection may not have the same set of significant granularities.
+This is the approach for managing this generate a subset ($S_c$) of significant granularities across a collection of time series:
 
- (a) Remove granularities from the comprehensive list that are insignificant for all observations.
+ (a) Remove granularities from the comprehensive list that are not significant for any time series.
 
- (b) select only the granularities that are significant for the majority of observations.
+ (b) Select only the granularities that are significant for the majority of time series.
 
-There will be observations in both cases where one or a few selected granularities are boring.
+<!--There will be time series in both cases where one or a few selected granularities are boring.
 Even in that case, having this group of observations with no interesting patterns at a granularity that regularly discovers patterns may be valuable.
-If, on the other hand, the granularities under examination are truly important for a group of data, unique patterns may be identified while clustering them.
+If, on the other hand, the granularities under examination are truly important for a group of data, unique patterns may be identified while clustering them.-->
 

### 2021-11-11 main.Rmd
data transformation

@@
 
-- _Data transformation_
+## Data transformation
 
-Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
+The shape and scale of the distribution of the measured variable (e.g.
+energy usage) affects distance calculations.
+Skewed distributions need to be symmetrised.
+Scales of individuals needs to be standardised, because clustering is to select similar patterns, not magnitude of usage.
+(Organising individuals based on magnitude can achieved simply by sorting on a statistic like the average value across time.) <!--Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
 Statistical transformations are employed to bring all of them to the same range or normalize each observation.
-This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.
-For the JS-based approaches, two data transformation techniques are utilized viz, Normal-Quantile Transform (NQT) and Robust scaling.
-NQT is a built-in transformation for computing $wpd$, which is the foundation of wpd-based distances.
+This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS).
+NQT is used in @@Gupta2021-hakear prior sorting granularities.
+<!-- $wpd$, which is the foundation of wpd-based distances.
+XX wpd was not yet introduced-->
 
-_Robust scaling_ The normalized $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
+- RS: The normalized $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
 $v_{norm}$ has zero mean and median, as well as a standard deviation of one, with the outliers having same relative connections to other values.
 
-_Normal-Quantile transform_ The raw data for all observations is individually normal-quantile transformed (NQT) [@Krzysztofowicz1997-bv], so that the transformed data follows a standard normal distribution.
-NQT will make the skewed distributions bell-shaped.
-As a result, determining which raw distribution was used is difficult using the modified distribution.
-Multimodality is also concealed or reversed.
-As a result, while displaying transformed data using NQT, one must exercise caution.
-But NQT is a useful transformation that often improves clustering performance.
+- NQT: The raw data for all observations is individually transformed [@Krzysztofowicz1997-bv], so that the transformed data follows a standard normal distribution.
+NQT will symmetrize skewed distributions.
+<!--As a result, determining which raw distribution was used is difficult using the modified distribution.--> A drawback is that any multimodality will be concealed.
+This, this should be checked prior to applying NQT.
 

### 2021-11-11 main.Rmd
data pre-processing

@@
 - RS: The normalized $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
-$v_{norm}$ has zero mean and median, as well as a standard deviation of one, with the outliers having same relative connections to other values.
+Note that, $v_{norm}$ has zero mean and median, but otherwise the shape does not change.
 
@@
 
-- _Data pre-preprocessing_
+## Data pre-preprocessing
 
-We start with a "tsibble" (@wang2020tsibble) data structure with an index variable representing inherent ordering from past to present and a key variable that defines observational units over time.
+Computationally in R, the data is assumed to be a "tsibble object" (@wang2020tsibble) equipped with an index variable representing inherent ordering from past to present and a key variable that defines observational units over time.
 The measured variable for an observation is a time-indexed sequence of values.
@@
 A shuffle of the raw sequence may represent hourly consumption throughout a day, a week, or a year.
-Cyclic granularities can be expressed in terms of the index set in the "tsibble" data structure.
-But the data structure changes while transporting from linear to cyclic scale of time as multiple observations now correspond to each category and induce a probability distribution.
-Directly computing Jensen-Shannon distances between the entire probability distributions can be very costly.
-Hence, in this paper, quantiles are chosen to characterize the probability distributions.
-So, in the final data structure, each category of a cyclic granularity corresponds to a list of numbers which is essentially a few chosen quantiles.
+Cyclic granularities can be expressed in terms of the index set in the tsibble data structure.
 
-- _Distance metrics_
+The data object will change when cyclic granularities are computed, as multiple observations will be caetgorized into levels of the granularity, thus inducing multiple probability distributions.
+Directly computing Jensen-Shannon distances between the entire probability distributions can be computationally intensive.
+Thus it is recommended that quantiles are used to characterize the probability distributions.
+In the final data object, each category of a cyclic granularity corresponds to a list of numbers which is composed of a few quantiles.
+
+## Distance metrics
 

### 2021-11-11 main.Rmd
sindex plot added

@@
 
+- sindex is plotted against the number of clusters ($k$) in Figure \ref{fig:cluster-number-selection-validation}.
+This can be used to determine the number of clusters for each scenario.
+Each line in the plots represents a different simulation setting for the same scenario.
+When sindex for each scenario are examined, it appears that $k = 5, 4, 4$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in sindex from that value of $k$.
+The number of clusters corresponds to the number of designs that were originally considered in each scenario.
+
 - In Figure \ref{fig:mds-plot-validation}, we tried to see how separated our clusters are.
@@
 The code for creating these data designs and running the methodologies is available at (https://github.com/Sayani07/paper-gracsR/Validation).
+
+```{r cluster-number-selection-validation, fig.cap="sindex plotted as a function of the number of clusters for S1, S2, and S3.
+S1 has a sharp decrease in sindex from 5 to 6, whereas S2 and S3 have a decrease from 4 to 5.
+As a result, the number of clusters considered for S1, S2, and S3 is 5, 4, 4, respectively.
+This corresponds to the number of designs taken into account in each scenario."}
+```
 

### 2021-11-11 script/main.R

@@ -713,7 +713,7 @@ append_files_plot <- function(folder_name, path){
   # all_data <- append_files("js-nqt")
   
   # compute all inter cluster distance "from" group
-  all_mat_mat_from <- all_datas%>%
+  all_mat_mat_from <- all_data%>%
     group_by(index, gran, group_item1, group_item2) %>% 
     filter(group_item1!= group_item2) %>% 
     summarise(sum = sum(distance),.groups = "drop") %>% 
@@ -1557,19 +1557,28 @@ sindex_all <- map(index_set, function(i){
 }) %>% bind_rows(.id = "index")
 
   
+
+}
+
+S1_nclust <- append_files_nclust(folder_name = "validation/js-nqt", path = "/3gran_change_5D/") %>% mutate(Scenario = "S1")
+
+S2_nclust <- append_files_nclust(folder_name = "validation/js-nqt", path = "/2gran_change_4D/")%>% mutate(Scenario = "S2")
+
+S3_nclust <- append_files_nclust(folder_name = "validation/js-nqt", path = "/1gran_change_5D/")%>% mutate(Scenario = "S3")
+
+sindex_all <- bind_rows(S1_nclust, S2_nclust, S3_nclust)
+
+
 opt_clusters_validation <- ggplot(sindex_all, 
                                   aes(x=k, y = sindex)) +
   geom_line(alpha = 0.3, aes(group = index), color = "blue") + 
   scale_x_continuous(breaks = seq(2, 20, 1)) + 
-  theme_bw() + ylab("sindex") + xlab("number of clusters")
+  facet_wrap(~Scenario, ncol = 1)+
+  ylab("sindex") + xlab("number of clusters")+
+  theme_characterisation()
 
-}
+opt_clusters_validation
 
-S1_nclust <- append_files_nclust(folder_name = "validation/js-nqt", path = "/3gran_change_5D/")
-
-S2_nclust <- append_files_nclust(folder_name = "validation/js-nqt", path = "/2gran_change_4D/")
-
-S3_nclust <- append_files_nclust(folder_name = "validation/js-nqt", path = "/1gran_change_5D/")
-
-S1_nclust/S2_nclust/S3_nclust + plot_annotation(tag_levels = '1', tag_prefix = 'S', tag_suffix = '')
+# 
+# S1_nclust/S2_nclust/S3_nclust + plot_annotation(tag_levels = '1', tag_prefix = 'S', tag_suffix = '')
 

### 2021-11-11 main.Rmd
added XXX info

@@
 @Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
-<!--None of the $34$ articles chosen for their emphasis use Australian smart meter data.--> The most often used algorithm is $k$-means (XXX reference something here so the reader can learn about k-means if they don't already know).
-Using $k$-means without considering time series structure or correlation fails to extract cyclical patterns (XXX???
-or do you mean to say that it can suffer from being a large p, small n problem??).
+<!--None of the $34$ articles chosen for their emphasis use Australian smart meter data.--> The most often used algorithm is $k$-means [@Rhodes2014-it].
+$k$-means can be made to perform better by explicitly incorporating time series features such as correlation or cyclic patterns rather than performing it on raw data.
 To reduce dimensionality, several studies use principal component analysis (PCA) or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).

### 2021-11-11 script/main.R

@@ -713,7 +713,7 @@ append_files_plot <- function(folder_name, path){
   # all_data <- append_files("js-nqt")
   
   # compute all inter cluster distance "from" group
-  all_mat_mat_from <- all_datas%>%
+  all_mat_mat_from <- all_data%>%
     group_by(index, gran, group_item1, group_item2) %>% 
     filter(group_item1!= group_item2) %>% 
     summarise(sum = sum(distance),.groups = "drop") %>% 
@@ -1557,19 +1557,29 @@ sindex_all <- map(index_set, function(i){
 }) %>% bind_rows(.id = "index")
 
   
+
+}
+
+S1_nclust <- append_files_nclust(folder_name = "validation/js-nqt", path = "/3gran_change_5D/") %>% mutate(Scenario = "S1")
+
+S2_nclust <- append_files_nclust(folder_name = "validation/js-nqt", path = "/2gran_change_4D/")%>% mutate(Scenario = "S2")
+
+S3_nclust <- append_files_nclust(folder_name = "validation/js-nqt", path = "/1gran_change_5D/")%>% mutate(Scenario = "S3")
+
+sindex_all <- bind_rows(S1_nclust, S2_nclust, S3_nclust)
+
+
 opt_clusters_validation <- ggplot(sindex_all, 
                                   aes(x=k, y = sindex)) +
   geom_line(alpha = 0.3, aes(group = index), color = "blue") + 
   scale_x_continuous(breaks = seq(2, 20, 1)) + 
-  theme_bw() + ylab("sindex") + xlab("number of clusters")
+  facet_wrap(~Scenario, ncol = 1)+
+  ylab("sindex") + xlab("number of clusters")+
+  theme_light() +
+  scale_x_continuous(breaks = seq(2, 15, 1), minor_breaks = 1)
 
-}
+opt_clusters_validation
 
-S1_nclust <- append_files_nclust(folder_name = "validation/js-nqt", path = "/3gran_change_5D/")
-
-S2_nclust <- append_files_nclust(folder_name = "validation/js-nqt", path = "/2gran_change_4D/")
-
-S3_nclust <- append_files_nclust(folder_name = "validation/js-nqt", path = "/1gran_change_5D/")
-
-S1_nclust/S2_nclust/S3_nclust + plot_annotation(tag_levels = '1', tag_prefix = 'S', tag_suffix = '')
+# 
+# S1_nclust/S2_nclust/S3_nclust + plot_annotation(tag_levels = '1', tag_prefix = 'S', tag_suffix = '')
 

### 2021-11-11 main.Rmd
minor re-write of sindex plot
noindent delete

@@
 
-\noindent The sum of the distances between two observations $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
+The sum of the distances between two observations $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
 After determining the distance between two series in terms of one granularity, we must combine them to produce a distance based on all significant granularities.
@@
 
-\noindent Once we have chosen $wpd$ as a relevant feature for characterizing the distributions across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
+Once we have chosen $wpd$ as a relevant feature for characterizing the distributions across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
 The Euclidean distance between them is chosen, with the granularities acting as variables and $wpd$ representing the value under each variable.
@@
 
-- _Clustering_
+## Clustering
 
-- Number of clusters: Different approaches to determining cluster size are required for varied applications and research objectives.
+_Number of clusters_: Different approaches to determining cluster size are required for varied applications and research objectives.
 Some typical ways of determining the number of clusters in the literature (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy]) are inspired primarily by defining and balancing "within-cluster homogeneity" and "between-cluster separation." Cluster separation index ($sindex$) proposed by @Hennig2014-ah is employed in this paper to have less dependency on a single or few ambiguous points.
@@
 
-- ALgorithm: With a way to obtain pairwise distances, any clustering algorithm can be employed that supports the given distance metric as input.
+_ALgorithm_: With a way to obtain pairwise distances, any clustering algorithm can be employed that supports the given distance metric as input.
 A good comprehensive list of algorithms can be found in @Xu2015-ja based on traditional ways like partition, hierarchy, or more recent approaches like distribution, density, and others.
@@
 
-- _Characterization of clusters_
+## Characterization of clusters
 
-\noindent Cluster characterization is an important element of cluster analysis.
+Cluster characterization is an important element of cluster analysis.
 @Cook2007-qe provides several methods for characterizing clusters.
@@
 
-\noindent Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction.
+Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction.
 This means that each granularity will have significantly different patterns at least across one of the designs to be clustered.
@@
 
-\noindent This is the case where one granularity will remain the same across all designs.
+This is the case where one granularity will remain the same across all designs.
 We consider the case where the distribution of $v$ varies across $g2$ levels for all designs, across $g3$ levels for a few designs, and $g1$ does not vary across designs.
@@
 
-\noindent Only one granularity is responsible for identifying the designs in this case.
-This is depicted in Figure \ref{fig:gran2and1-clubbed} (right) where only $g3$ affects the designs significantly.
+Only one granularity is responsible for identifying the designs in this case. This is depicted in Figure \ref{fig:gran2and1-clubbed} (right) where only $g3$ affects the designs significantly.
 
@@
 
-- sindex is plotted against the number of clusters ($k$) in Figure \ref{fig:cluster-number-selection-validation}.
+- Figure \ref{fig:cluster-number-selection-validation} shows sindex plotted against the number of clusters ($k$).
 This can be used to determine the number of clusters for each scenario.
@@
 
-\noindent
 **Supplementary paper:** Additional tables, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd)
 
-\noindent
 **R-package:** To implement the ideas provided in this research, the open-source R package 'gracsr' is available on Github (https://github.com/Sayani07/gracsr).

### 2021-11-11 main.Rmd
change similarity measure and distance terminology

@@
 The similarity metric needs to be capable of accommodating long, noisy, and asynchronous time series and also capture cyclical patterns.
-In this paper, two approaches for measuring similarity between time series are presented, based on probability distributions over cyclic temporal granularities.
+In this paper, two approaches for measuring distances between time series are presented, based on probability distributions over cyclic temporal granularities.
 Both are compatible with a variety of clustering algorithms.
@@
 Hence clustering these customers is equivalent to clustering these collections of conditional univariate probability distributions.
-Towards this goal, the next step is to decide how to measure similarities between collections of univariate probability distributions.
+Towards this goal, the next step is to decide how to measure distances between collections of univariate probability distributions.
 <!--There are multiple ways to measure similarities depending on the aim of the analysis.

### 2021-11-11 main.Rmd
group summary updated only with wnwd
added sindex plot for wpd approach

@@
 
+```{r groups-24}
+```
+
 ```{r opt-clusters-jsd}
@@
 
-```{r opt-cluster-tsne-jsd, fig.cap = "(a) shows t-SNE summary of the $24$ selected customers chosen through prototype selection, (b) shows a plot of sindex for different cluster size for clustering based on JS-based distances.
-The separation index decreases sharply from $5$ to $6$ clusters, so $5$ clusters are used here."}
-
+```{r data-pick-wpd}
 ```
 
-```{r groups-24}
+```{r opt-cluster-tsne-jsd-png, fig.cap = "(a) shows t-SNE summary of the $24$ selected customers chosen through prototype selection, (b) and (c) shows a plot of sindex for different cluster size for clustering based on JS-based distances and wpd-distances respectively.
+The plot suggests optimal number of clusters as $4$ and $7$ respectively."}
+include_graphics("figs/opt_clust_parenthesis.png")
 ```
@@
 
-```{r data-heatmap-hod-group}
+```{r data-heatmap-hod-group-new}
 ```
 
-```{r data-heatmap-moy-group}
+```{r data-heatmap-moy-group-new}
 
@@
 The findings vary from js-based clustering, yet it is a helpful grouping.
-
-```{r data-pick-wpd}
-```
 

### 2021-11-12 main.Rmd

@@
 
-**JS-based distances**
+### JS-based distances
 
@@
 
-**wpd-based distances**
+### wpd-based distances
 
@@
 
-_Number of clusters_: Different approaches to determining cluster size are required for varied applications and research objectives.
+### Number of clusters
+
+Different approaches to determining cluster size are required for varied applications and research objectives.
 Some typical ways of determining the number of clusters in the literature (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy]) are inspired primarily by defining and balancing "within-cluster homogeneity" and "between-cluster separation." Cluster separation index ($sindex$) proposed by @Hennig2014-ah is employed in this paper to have less dependency on a single or few ambiguous points.
@@
 
-_ALgorithm_: With a way to obtain pairwise distances, any clustering algorithm can be employed that supports the given distance metric as input.
+### Algorithm
+
+With a way to obtain pairwise distances, any clustering algorithm can be employed that supports the given distance metric as input.
 A good comprehensive list of algorithms can be found in @Xu2015-ja based on traditional ways like partition, hierarchy, or more recent approaches like distribution, density, and others.
@@
 
-## Characterization of clusters
+### Characterization of clusters
 
@@
 
-## Data generating processes
+## Data generation
 
@@
 
-_Scenario (S1): All significant granularities_
+_Scenario (S1): All granularities significant_
 
@@
 
-_Scenario (S3): One significant granularity_
+_Scenario (S3): Only one significant granularity_
 
@@
 
-## Visual exploration of findings
+## Visual exploration of results
 
@@
 
-_Data filtering and variable selection_
+## Data filtering and variable selection
 
@@
 
-_Prototype selection_
+## Prototype selection
 

### 2021-11-13 main.Rmd
adding index plots and summary plot for 2 groupings

@@
 
-The use of our methodology is illustrated on smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
-It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
+The use of our methodology is illustrated on smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.<!-- It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
 In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
 We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation.
-Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.
-The raw data for these consumers is of unequal length, with varying start and finish dates.
+Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.--> The raw data for these consumers is of unequal length, with varying start and finish dates.
 Because our proposed methods evaluate probability distributions rather than raw data, neither of these data features would pose any threat to our methodology unless they contained any structure or systematic patterns.
@@
 ## Data filtering and variable selection
+
+Things become far more complicated when we consider a larger data set with more uncertainty, as they do with any clustering problem.
 
@@
 It is useful to compare the raw distributions with the summarized distributions to establish that the majority of individuals in the group share the same characteristics.
-To facilitate this comparison, the individual distributions are sorted by their JS-based groupings."}
+To facilitate this comparison, the individual distributions are sorted by their JS-based groupings.", eval = FALSE}
+
+```
+
+```{r opt-clusters}
+knitr::include_graphics("figs/opt_clusters.png")
+```
+
+```{r combined4and5}
+knitr::include_graphics("figs/combined4and5.png")
 ```

### 2021-11-13 group5.R

@@ -1,0 +1,103 @@
+
+##----data-heatmap-hod-group-new-4and5
+
+data_group <- data_pick_robust  %>% 
+  left_join(cluster_result_id4and5, by = c("customer_id"))
+
+data_heatmap_hod_group <- quantile_gran(data_group,
+                                        gran1="hour_day",
+                                        quantile_prob_val = c(0.25, 0.5, 0.75),
+                                        group="group5") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values) 
+
+
+data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
+
+data_heatmap_hod_group$group5 <- paste("groupB", data_heatmap_hod_group$group5, sep = "-")
+
+hod_group <- data_heatmap_hod_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=group5,
+                  fill = as.factor(group5), alpha = 0.5),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=group5, 
+                color = as.factor(group5)), size = 1)+
+  facet_wrap(~group5, 
+             scales = "free_y",  
+             nrow = 5, labeller = "label_value") + 
+  #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
+  theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hod") + 
+  ylab("demand (in Kwh)") + 
+  theme_bw()  +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ 
+  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme_application3()  +
+  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00","#CC79A7"))+
+  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "#CC79A7")) +
+  theme(legend.position = "bottom") 
+
+##----data-heatmap-moy-group-new-4and5
+data_heatmap_moy_group <- quantile_gran(data_group,
+                                        gran1="month_year",
[…]

### 2021-11-14 main.Rmd
Say why you choose a clean sample

@@
 
-The use of our methodology is illustrated on smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.<!-- It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
-In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
+The use of our methodology is illustrated on smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
+It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
 We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation.
@@
 
-## Data filtering and variable selection
+Clustering huge data sets presents a slew of difficulties.
+As a result, we dissect the larger problem and test our solutions on a small sample of prototype customers.
+To do this, data is first filtered to generate a clean sample, and then significant cyclic granularities (variables) for them are chosen.
+This is described in Section \ref{sec:datafilter}.
+The clean set is subsequently subsampled along all dimensions of interest, to ensure that the sampled data reveals some patterns across at least one specified variable, which is described in Section \ref{sec:prototype}.
+Section \ref{sec:clustering} details the clustering results for the subsample.
+
+## Data filtering and variable selection {#sec:datafilter}
 
@@
 
-## Prototype selection
+## Prototype selection {#sec:prototype}
 
@@
 Unfortunately, this does not ensure that consumers with significant patterns across all variables are chosen.
-Tours can reveal variable separation that was hidden in a single variable display better than static projections.
+Tours can reveal variable separation that gets hidden in a single variable display better than static projections.
 Hence we perform a linked tour with a t-SNE layout using the R package `liminal` (@R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
@@
 
-## Clustering
+## Clustering {#sec:clustering}
 

### 2021-11-14 main.Rmd
number of clusters: several indices to be examined

@@
 
-Different approaches to determining cluster size are required for varied applications and research objectives.
-Some typical ways of determining the number of clusters in the literature (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy]) are inspired primarily by defining and balancing "within-cluster homogeneity" and "between-cluster separation." Cluster separation index ($sindex$) proposed by @Hennig2014-ah is employed in this paper to have less dependency on a single or few ambiguous points.
-$sindex$ [@R-fpc] is determined by calculating the distance between each observation and the cluster to which it does not belong.
-Then, the least 10% of these distances are averaged to only evaluate places near the cluster edges.
-$sindex$, like other statistics in the literature, cannot be optimized over the number of clusters, since increasing the number of clusters reduces it.
+Different approaches to determining number of clusters are required for varied applications and research objectives.
+It is often beneficial to investigate numerous cluster statistics in order to arrive at a number of clusters, since clustering has a variety of objectives (e.g., between-cluster separation, within-cluster homogeneity, representation by centroids, etc.), which could be in conflict with one another.
+Some typical ways of determining the number of clusters in the literature (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy]) are inspired by defining and balancing "within-cluster homogeneity" and "between-cluster separation".
+The Cluster separation index ($sindex$) proposed by @Hennig2014-ah is used as a guide to select the number of clusters in Section \ref{sec:validation}.
+$sindex$ [@R-fpc] is a compromise between minimum cluster separation and average cluster separation and is less sensitive to a single or few ambiguous points.
+$sindex$ cannot be optimized over the number of clusters, since increasing the number of clusters reduces it.
 The number of clusters required is determined by the slope going from steep to shallow (an elbow).
+
+.<!--determined by calculating the distance between each observation and the cluster to which it does not belong.
+Then, the least 10% of these distances are averaged to only evaluate places near the cluster edges.-->
 
@@
 The $24$ prototypes <!--and the data subset consisting of $353$ customers--> are clustered using the methodology described in Section \ref{sec:methodology} and results are reported below.
-In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.
-All customers with the same color represent the same clustered groups.
-Groups by JS-based distances and wpd-based distances are colored differently as they represent different groupings.
-The plotting scales are not displayed since we want to emphasize comparable shapes rather than scales.
+In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.<!--All customers with the same color represent the same clustered groups.
+--> Groups by JS-based distances and wpd-based distances are colored differently as they represent different groupings.
+The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales.
 The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.

### 2021-11-14 main.Rmd
delete repeating data filtering

@@
 Additionally, there were missing values in the database but further investigation revealed that there is no structure in the missingness (see Supplementary paper for raw data features and missingness).
-The study begins by subsetting a data set along all dimensions of interest using data filtering and prototyping.
-By grouping the prototypes using our methods and assessing their meaning, the study hopes to unravel some of the heterogeneities observed in energy usage data.
-Because our application does not employ additional customer data, we cannot explain why consumption varies, but rather try to identify how it varies.
 
@@
 The clean set is subsequently subsampled along all dimensions of interest, to ensure that the sampled data reveals some patterns across at least one specified variable, which is described in Section \ref{sec:prototype}.
-Section \ref{sec:clustering} details the clustering results for the subsample.
+By grouping the prototypes using our methods in Section \ref{sec:clustering} and assessing their meaning, the study hopes to unravel some of the heterogeneities observed in energy usage data.
+Because our application does not employ additional customer data, we cannot explain why consumption varies, but rather try to identify how it varies.
 

### 2021-11-14 main.Rmd
Figure 9 update

@@
 The $24$ prototypes <!--and the data subset consisting of $353$ customers--> are clustered using the methodology described in Section \ref{sec:methodology} and results are reported below.
-In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.<!--All customers with the same color represent the same clustered groups.
+In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups.
 --> Groups by JS-based distances and wpd-based distances are colored differently as they represent different groupings.
@@
 The groupings are shown in Figure \ref{fig:hod-ind-group-png}.
-Customers with the same color represent the same clustered groups.
-Our methodology is useful for grouping similar distributions over `hod` and `moy` and they are placed closely for easy comparison.
+<!--Customers with the same color represent the same clustered groups.--> Our methodology is useful for grouping similar distributions over `hod` and `moy` and they are placed closely for easy comparison.
 Few groups have mixed patterns across `hod` and `moy`, but few have all customers in the group having a similar profile.
@@
 
-```{r combined4and5}
+<!-- GroupA and Group B may represent different set of customers as they are groupings based on different number of clusters. -->
+
+```{r combined4and5, fig.cap = "For $k = 4$ (a) and $k=5$ (b), the distribution of electricity demand across groups over hod, moy, and wnwd is shown.
+groupA-2, groupA-3, and groupA-4 profiles correspond to groupB-2, groupB-3, and groupB-5, respectively.
+groupA-1 is split into groupB-1 and groupB-4, each of which has a distinct shape across moy and wnwd.
+groupB-4 (id: 16-20) is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage.
+groupB-1 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.
+When $k = 4$ is used, these two groups merge to form groupA-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group."}
 knitr::include_graphics("figs/combined4and5.png")

### 2021-11-14 main.Rmd
no additional gain in increasing nclust  from 3 for wpd-based

@@
 
-```{r opt-clusters}
-knitr::include_graphics("figs/opt_clusters.png")
+<!-- The widest within-cluster gap (widest_gap) is high till $k=5$ and then have a gradual decrease. Number of -->
+
+```{r opt-clusters, fig.cap=" Different cluster statistics plotted as a function of number of clusters for JS-based (a) and wpd-based (b) approaches.
+avg within is not too useful for revealing any sharp change in average distance between clusters with increasing number of clusters.
+For (a), sindex sees a sharp fall at $k = 4$ and then do not decrease until $k=16$ and the widest gap is high till $k=5$.
+For (b), sindex sees a fall at $k=3$ and then at $k=7$ and widest gap has a sharp fall at $k=6$."}
+
+#knitr::include_graphics("figs/opt_clusters.png")
 ```
@@
 
-```{r parcoord, fig.cap="Each of the $24$ customers is represented by a parallel coordinate plot (a) with three wpd-based groupings.
+```{r parcoord-application, fig.cap="Each of the $24$ customers is represented by a parallel coordinate plot (a) with three wpd-based groupings.
 Plot (b) which shows median wpd values for each group.
@@
 Group 1 is a mixed group with strong patterns on atleast one of the three variables.
-All of these could be observed from the plot or the table (b).", message=FALSE, warning=FALSE}
-include_graphics("figs/parcoord-application.png")
+All of these could be observed from the plot or the table (b).", message=FALSE, warning=FALSE, fig.height=6}
+#include_graphics("figs/parcoord-application.png")
 ```

### 2021-11-14 script/main.R

@@ -913,18 +913,21 @@ f = as.dist(distance)
 
 ##----opt-clusters-jsd
 
-all_index <- map_dfr(2:20, function(x){
+all_index <- map_dfr(2:23, function(x){
   group <- f %>% hclust (method = "ward.D") %>% cutree(k=x)
-  p <- cluster.stats(f, clustering = group, silhouette = TRUE)
-  index <- c(k = x, sindex = p$sindex, avg_silwidth = p$avg.silwidth, gamma =  p$pearsongamma, widegap = p$widestgap)
+  p <- cluster.stats(f, clustering = group, silhouette = TRUE, wgap = TRUE)
+  index <- c(k = x,
+             sindex = p$sindex,  
+             avg_within =  p$average.within,
+             widest_gap = p$widestgap)
 }) 
 
-  # , gamma = p$pearsongamma, widegap = p$widestgap, dunn = p$dunn), gamma =  p$pearsongamma,
-data_index <- all_index %>% pivot_longer(2:5, names_to = "index_type", values_to = "index_value")
+  # , gamma = p$pearsongamma, widegap = p$widestgap, dunn = p$dunn), gamma =  p$pearsongamma,avg_silwidth = p$avg.silwidth,
+data_index <- all_index %>% pivot_longer(2:4, names_to = "index_type", values_to = "index_value")
   
   
 opt_clusters <- data_index %>% ggplot(aes(x = k, y = index_value, group = index_type))+
-  geom_line() + scale_x_continuous(breaks = seq(2, 20, 1), minor_breaks = 1) + theme_bw() + ylab("index") + xlab("number of clusters") + facet_wrap(~index_type, ncol = 1, scales = "free_y")
+  geom_line() + scale_x_continuous(breaks = seq(2, 23, 1), minor_breaks = 1) + theme_bw() + ylab("index") + xlab("number of clusters") + facet_wrap(~index_type, ncol = 1, scales = "free_y") +  theme(axis.text.x = element_text(angle=90, hjust=1, size = 8))
 
 ##----groups-24
 cluster_result <- suppressMessages(f %>% 
@@ -1282,18 +1285,18 @@ f <- elec_pick_wide[-1] %>% dist()
 # }
 
 
-all_index <- map_dfr(2:20, function(x){
+all_index <- map_dfr(2:23, function(x){
   group <- f %>% hclust (method = "ward.D") %>% cutree(k=x)
[…]

### 2021-11-14 main.Rmd
fig 4 caption

@@
 
-```{r cluster-number-selection-validation, fig.cap="sindex plotted as a function of the number of clusters for S1, S2, and S3.
-S1 has a sharp decrease in sindex from 5 to 6, whereas S2 and S3 have a decrease from 4 to 5.
+```{r cluster-number-selection-validation, fig.cap="Separation index plotted as a function of the number of clusters for S1, S2, and S3.
+S1 has a sharp decrease from 5 to 6, whereas S2 and S3 have a decrease from 4 to 5.
 As a result, the number of clusters considered for S1, S2, and S3 is 5, 4, 4, respectively.

### 2021-11-14 main.Rmd
adding range of parameters table

--- 2021-11-14 main.Rmd
@@
 So three scenarios are considered to accommodate that.
+Figure \ref{fig:range-parameter} shows the range of parameters considered for each scenario.
+
+```{r range-parameter, fig.cap="Range of parameters for different scenarios"}
+knitr::include_graphics("figs/simulation_table.png")
+```
 
@@
 For more detailed results, please refer to the supplementary paper.
-The code for creating these data designs and running the methodologies is available at (https://github.com/Sayani07/paper-gracsR/Validation).
 
@@
 
-**Supplementary paper:** Additional tables, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd)
+**Supplementary paper:** Additional tables, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd).
+The code for creating validation designs and running the methodologies is available at (https://github.com/Sayani07/paper-gracsR/Validation).
 

### 2021-11-14 main.Rmd
remove all index plot

@@
 For (a), sindex sees a sharp fall at $k = 4$ and then do not decrease until $k=16$ and the widest gap is high till $k=5$.
-For (b), sindex sees a fall at $k=3$ and then at $k=7$ and widest gap has a sharp fall at $k=6$."}
-
+For (b), sindex sees a fall at $k=3$ and then at $k=7$ and widest gap has a sharp fall at $k=6$.", eval = FALSE}
 #knitr::include_graphics("figs/opt_clusters.png")

### 2021-11-14 main.Rmd
fixing sindex plot

@@
 
-```{r tsne-fit}
-
-```
-
-```{r tsne-plot}
-```
-
 <!-- - can be done in several ways -->
@@
 
-```{r opt-cluster-tsne-jsd-png, fig.cap = "(a) shows t-SNE summary of the $24$ selected customers chosen through prototype selection, (b) and (c) shows a plot of sindex for different cluster size for clustering based on JS-based distances and wpd-distances respectively.
+```{r tsne-fit}
+```
+
+```{r tsne-df}
+```
+
+```{r tsne-xy}
+
+```
+
+```{r opt-cluster-tsne-jsd, fig.cap = "(a) shows t-SNE summary of the $24$ selected customers chosen through prototype selection, (b) and (c) shows a plot of sindex for different cluster size for clustering based on JS-based distances and wpd-distances respectively.
 The plot suggests optimal number of clusters as $4$ and $7$ respectively."}
-include_graphics("figs/opt_clust_parenthesis.png")
+#include_graphics("figs/opt_clust_parenthesis.png")
 ```

### 2021-11-14 script/main.R
parcoord application format

@@ -1306,6 +1306,9 @@ opt_clusters_wpd <- all_index %>% ggplot(aes(x = k, y = sindex))+
 
 # opt_clusters_wpd <- ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 20, 1), minor_breaks = 1) + theme_bw() + ylab("sindex") + xlab("number of clusters")
 
+
+
+##----wpd-clustering
 group <- f%>% hclust (method = "ward.D") %>% cutree(k=4)
 
 
@@ -1316,16 +1319,19 @@ data_pcp <- scaled_var %>%
   left_join(cluster_result_wpd , by = c("customer_id" = "id")) %>% 
   select(customer_id, group, everything()) %>% 
   mutate(group = as.factor(group)) %>% 
-  mutate(customer_id = as.character(customer_id))
+  mutate(customer_id = as.character(customer_id)) %>% 
+  rename("moy" = "month_year",
+         "hod" = "hour_day",
+         "wnwd" = "wknd_wday")
 
 data_table <- data_pcp %>% group_by(group) %>% 
   summarise(nobs = n(),
-            moy = round(median(month_year),2),
-            hod = round(median(hour_day),2),
-            wnwd = round(median(wknd_wday),2)) %>% 
-  select(-group)
+            moy = round(median(moy),2),
+            hod = round(median(hod),2),
+            wnwd = round(median(wnwd),2)) %>% 
+  select(-group) 
 
-rownames(data_table) <- c("group-1", "group-2", "group-3")
+rownames(data_table) <- c("group-1", "group-2", "group-3", "group-4")
 
 ##----parcoord-application
 parcoord <- GGally::ggparcoord(data_pcp %>% left_join(cluster_result_id, by = "customer_id") ,
@@ -1343,11 +1349,13 @@ parcoord <- GGally::ggparcoord(data_pcp %>% left_join(cluster_result_id, by = "c
   theme(legend.position = "bottom") +
   xlab("") +
   ylab("wpd") + scale_fill_viridis_d(direction = 1) +
-  scale_color_viridis_d(direction = 1) + theme_light()
+  scale_color_viridis_d(direction = 1) + theme_light()  +labs(colour="group") +
+  theme(panel.border = element_blank())
 
+# 
[…]

### 2021-11-14 main.Rmd
parcoord application format

@@
 
+```{r wpd-clustering}
+
+```
+
 ```{r parcoord-application, fig.cap="Each of the $24$ customers is represented by a parallel coordinate plot (a) with three wpd-based groupings.

### 2021-11-14 main.Rmd
plots update along with writeup in validation section

--- 2021-11-14 main.Rmd
@@
 
-All of the approaches were fitted to each data design and for each combination of the considered parameters.
+All of the approaches were fitted to each data design and to each combination of the considered parameters.
 The formed clusters have to match the design, be well separated, and have minimal intra-cluster variation.
 It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values.
-So we use MDS and parallel coordinate graphs to demonstrate the findings:
-
-- Figure \ref{fig:cluster-number-selection-validation} shows sindex plotted against the number of clusters ($k$).
+MDS and parallel coordinate graphs are used to demonstrate the findings, along with an index value plot for guidance on the number of clusters.
+JS-based approaches corresponding to NQT and RS are referred to as JS-NQT and JS-RS respectively.
+In the following plots, results for JS-NQT are reported.
+For similar results with JS-RS or wpd-based distances, please refer to the supplementary paper.
+
+- Figure \ref{fig:sindex-plot-validation} shows sindex plotted against the number of clusters ($k$) for the range of mean differences (rows) under the different scenarios (columns).
 This can be used to determine the number of clusters for each scenario.
-Each line in the plots represents a different simulation setting for the same scenario.
 When sindex for each scenario are examined, it appears that $k = 5, 4, 4$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in sindex from that value of $k$.
@@
 
- - The js-robust and wpd methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
+ - The JS-RS and wpd-based methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
 Although, a complete year of data is the minimum requirement to capture distributional differences in winter and summer profiles, for example.
 Even if the data is only available for a month, $nT$ with half-hourly data is expected to be at least $1000$.
-As a result, as long as the performance is promising for higher $nT=300$, this is not a challenge.
-
-- In our study sample, the method js-nqt outperforms the method js-robust for smaller differences between categories. More testing, however, is required to corroborate this.
-
-For more detailed results, please refer to the supplementary paper.
+As a result, as long as the performance is promising for higher $nT$, this is not a challenge.
+
+- In our study sample, the method JS-NQT outperforms the method JS-RS for smaller differences between categories. More testing, however, is required to corroborate this.
+
+```{r sindex-data-validation}
+
+```
+
+```{r sindex-plot-validation, fig.cap = "The cluster separation index (sindex) is plotted as a function of the number of clusters for the range of mean differences (rows) under the different scenarios (columns).
+S1 has a sharp decrease in sindex from 5 to 6, whereas S2 and S3 have a decrease from 4 to 5.
+As a result, the number of clusters considered for S1, S2, and S3 is 5, 4, 4, respectively.
+This corresponds to the number of designs taken into account in each scenario."}
+
+```
 
@@
 As a result, the number of clusters considered for S1, S2, and S3 is 5, 4, 4, respectively.
-This corresponds to the number of designs taken into account in each scenario."}
+This corresponds to the number of designs taken into account in each scenario.", eval=FALSE}
 ```
@@
 It is worth emphasising that these facts are consistent with our initial assumptions when designing the scenarios and S1, S2, S3 correspond to scenarios with three, two and one significant granularity respectively."}
+
+```
+
+```{r mds-data-validation}
 
@@
 
-For clustering based on JS-based distances, we chose the optimal number of clusters using (@Hennig2014-ah) as $5$.
+For clustering based on JS-based (JS-NQT) distances, we chose the optimal number of clusters using (@Hennig2014-ah) as $5$.
 The groupings are shown in Figure \ref{fig:hod-ind-group-png}.
@@
 Few groups have mixed patterns across `hod` and `moy`, but few have all customers in the group having a similar profile.
-Figure \ref{fig:combined-groups-js} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
+Figure \ref{fig:hod-ind-group-png} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
 It shows Groups $2$ and $4$ have `hod` pattern with a typical morning and evening peak, whereas groups $1$, $3$, and $5$ show a `moy` pattern with higher usage in winter months.
@@
 Group 1 is a mixed group that has strong patterns on at least one of the three variables.
-The findings vary from js-based clustering, yet it is a helpful grouping.
+The findings vary from JS-based clustering, yet it is a helpful grouping.
+
+```{r wpd-clustering}
+
+```
 

### 2021-11-15 main.Rmd
Figure 9 update

@@
 --> Groups by JS-based distances and wpd-based distances are colored differently as they represent different groupings.
+While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominate the shape of the summarized clusters because of magnitude.
 The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales.
@@
 
-For clustering based on JS-based (JS-NQT) distances, we chose the optimal number of clusters using (@Hennig2014-ah) as $5$.
-The groupings are shown in Figure \ref{fig:hod-ind-group-png}.
-<!--Customers with the same color represent the same clustered groups.--> Our methodology is useful for grouping similar distributions over `hod` and `moy` and they are placed closely for easy comparison.
+The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in \ref{fig:opt-cluster-tsne-jsd}.
+The $sindex$ falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases.
+In our study, we use $k=4, 5$.
+Figure \ref{fig:combined4and5} depicts the summarised distributions across $4$ and  $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster.
+A-2, A-3, and A-4 have profiles that correspond to groups B-2, B-3, and B-5, respectively.
+B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-5 (id: 21-24) have a 'hod' pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day.
+A-1 is subdivided into B-1 and B-4, each of which has a distinct shape across 'moy' and 'wnwd'.
+B-4 (id: 16-20) is distinguished by higher energy consumption in the first few months of the year and greater variability in weekend usage.
+B-1 (id: 1-3) has higher consumption in the middle of the year (winter months) and a similar weekday-weekend inter-quartile range.
+When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals.
+It may be worthwhile to compare Figures \ref{fig:opt-cluster-tsne-jsd} and \ref{fig:hod-ind-group-png} to see if the summarized distributions across groups accurately characterized the groupings.
+If it has, then the majority of the members of the group should have a similar profile.
+
+<!-- Our methodology is useful for grouping similar distributions over `hod` and `moy` and they are placed closely for easy comparison.
 Few groups have mixed patterns across `hod` and `moy`, but few have all customers in the group having a similar profile.
-Figure \ref{fig:hod-ind-group-png} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
-It shows Groups $2$ and $4$ have `hod` pattern with a typical morning and evening peak, whereas groups $1$, $3$, and $5$ show a `moy` pattern with higher usage in winter months.
-Differences in Weekend/Weekday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters unless maybe conditioned by `moy` or `hod`.
-It may be interesting to compare these two plots to verify if the summarized distributions across groups correctly characterised the groupings.
-If it has, then the majority of the group's members should share a similar profile.
+-->
 
@@
 ```{r combined4and5, fig.cap = "For $k = 4$ (a) and $k=5$ (b), the distribution of electricity demand across groups over hod, moy, and wnwd is shown.
-groupA-2, groupA-3, and groupA-4 profiles correspond to groupB-2, groupB-3, and groupB-5, respectively.
-groupA-1 is split into groupB-1 and groupB-4, each of which has a distinct shape across moy and wnwd.
-groupB-4 (id: 16-20) is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage.
-groupB-1 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.
-When $k = 4$ is used, these two groups merge to form groupA-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group."}
+Groups A-2, A-3, and A-4 profiles correspond to Groups B-2, B-3, and B-5, respectively.
+A-1 is split into B-1 and B-4, each of which has a distinct shape across moy and wnwd.
+B-4 (id: 16-20) is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage.
+B-1 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.
+When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group."}
 knitr::include_graphics("figs/combined4and5.png")

### 2021-11-15 group4.R

@@ -5,7 +5,8 @@ cluster_result_kopt4 <- suppressMessages(f %>%
   mutate(group = as.factor(group))
 
 cluster_result_id4and5 <- cluster_result_id %>% left_join(cluster_result_kopt4, by="customer_id") %>% rename("group4" = "group.y") %>% rename("group5" = "group.x")
-##----data-heatmap-hod-group-new-4and5
+
+#data-heatmap-hod-group-new-4
 
 data_group <- data_pick_robust  %>% 
   left_join(cluster_result_id4and5, by = c("customer_id"))
@@ -19,7 +20,7 @@ data_heatmap_hod_group <- quantile_gran(data_group,
 
 data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
 
-data_heatmap_hod_group$group4 <- paste("groupA", data_heatmap_hod_group$group4, sep = "-")
+data_heatmap_hod_group$group4 <- paste("A", data_heatmap_hod_group$group4, sep = "-")
 
 hod_group <- data_heatmap_hod_group %>% 
   ggplot(aes(x = category)) + 
@@ -55,7 +56,7 @@ data_heatmap_moy_group <- quantile_gran(data_group,
 data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
 
 
-data_heatmap_moy_group$group4 <- paste("groupA", data_heatmap_moy_group$group4, sep = "-")
+data_heatmap_moy_group$group4 <- paste("A", data_heatmap_moy_group$group4, sep = "-")
 
 
 moy_group <- data_heatmap_moy_group %>% 
@@ -74,13 +75,13 @@ moy_group <- data_heatmap_moy_group %>%
   scale_color_manual(values=as.vector(tableau20(20)))+
   theme(legend.position = "bottom") 
 
-##----data-heatmap-wkndwday-group-4and5
+#data-heatmap-wkndwday-group-4and5
 wkndwday_data <- data_group %>% create_gran("wknd_wday") %>% 
   create_gran("hour_day")
 
 ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]
 
-wkndwday_data$group4 <- paste("groupA", wkndwday_data$group4, sep = "-")
[…]

### 2021-11-15 main.Rmd
wpd-based clustering update

@@
 
-We chose the optimal number of clusters using (@Hennig2014-ah) as $3$.
-A parallel coordinate plot with the three significant cyclic granularities is used to characterise the groups here.
+The $sindex$ is plotted against the number of clusters for clustering based on wpd-based distances in \ref{fig:opt-cluster-tsne-jsd}.
+The scales are very different from JS-based clustering and also there is more gradual decrease in the separation index.
+The $sindex$ falls from around $20$ when $k=3$ to $19$ when $k=4$ and further decreases to $17.5$ with $k>7$.
+Thus, there are few choices available for number of clusters.
+The groupings with $k=4$ is shown through a parallel coordinate plot with the three significant cyclic granularities.
 The variables are sorted according to their separation across classes (rather than their overall variation between classes).
 This means that $moy$ is the most important variable in distinguishing the groups followed by $hod$ and $wnwd$.
-There is only one customer who has significant $wpd$ across $wnwd$ and stands out from the rest of the customers.
-Group 3 has a higher wpd for hod than moy or wkndwday.
-Group 2 has the most distinct pattern across moy.
-Group 1 is a mixed group that has strong patterns on at least one of the three variables.
+The two customers (id: 16 and 18) which has significant differences between $wnwd$ catgeories (in Figure \ref{fig:hod-ind-group-png}) stand out from the rest of the customers.
+Group-4 has a higher $wpd$ for `hod` than `moy` or `wkndwday`.
+Group-2 and Group-3 have the most distinct pattern across `moy`.
+Group-1 is a mixed group that has strong patterns on at least one of the three variables.
+When $k=3$ is used, Group-2 and Group-3 are merged together because their patterns relative to `moy`, '`hod` and `wnwd` are same.
 The findings vary from JS-based clustering, yet it is a helpful grouping.
@@
 
-```{r parcoord-application, fig.cap="Each of the $24$ customers is represented by a parallel coordinate plot (a) with three wpd-based groupings.
-Plot (b) which shows median wpd values for each group.
-The plot shows that moy is the most important variable in identifying clusters, whereas wnwd is the least significant and has the least fluctuation.
-Two customers (id: 16 and 18) with high wpd across wnwd stand out in this display.
-Group 3 has a higher wpd for hod than moy or wkndwday.
-Group 2 has most discernible pattern across moy.
-Group 1 is a mixed group with strong patterns on atleast one of the three variables.
+```{r parcoord-application, fig.cap="Each of the $24$ customers is represented by a parallel coordinate plot (a) with four wpd-based groupings.
+Plot (b) which shows median $wpd$ values for each group.
+The plot shows that `moy` is the most important variable in identifying clusters, whereas `wnwd` is the least significant and has the least fluctuation.
+Two customers (id: 16 and 18) with high $wpd$ across wnwd stand out in this display.
+Group-4 has a higher $wpd$ for `hod` than `moy` or `wnwd`.
+Group-2 and Group-3 has most discernible pattern across `moy`.
+Group-1 is a mixed group with strong patterns across atleast one of the three variables.
 All of these could be observed from the plot or the table (b).", message=FALSE, warning=FALSE, fig.height=6}

### 2021-11-15 main.Rmd
summarized dist -jsd plot

@@
 In our study, we use $k=4, 5$.
-Figure \ref{fig:combined4and5} depicts the summarised distributions across $4$ and  $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster.
+Figure \ref{fig:groups-4and5} depicts the summarised distributions across $4$ and  $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster.
 A-2, A-3, and A-4 have profiles that correspond to groups B-2, B-3, and B-5, respectively.
@@
 When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group."}
-knitr::include_graphics("figs/combined4and5.png")
+#knitr::include_graphics("figs/combined4and5.png")
+```
+
+```{r groups-4and5, fig.cap = "For $k = 4$ (a) and $k=5$ (b), the distribution of electricity demand across groups over hod, moy, and wnwd is shown.
+Groups A-2, A-3, and A-4 profiles correspond to Groups B-2, B-3, and B-5, respectively.
+A-1 is split into B-1 and B-4, each of which has a distinct shape across moy and wnwd.
+B-4 (id: 16-20) is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage.
+B-1 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.
+When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group."}
+
 ```

### 2021-11-15 script/main.R

@@ -998,6 +998,8 @@ distance <- wkndwday/2 + moy/12 + hod/24
 
 f = as.dist(distance)
 
+distance_jsd = f
+
 
 ##----opt-clusters-jsd
 
@@ -1679,3 +1681,217 @@ sindex_plot <- ggplot(all_data,
   scale_x_continuous(breaks = seq(2, 10, 1), minor_breaks = 1)
 
 sindex_plot
+
+
+
+##----groups-4and5
+cluster_result_kopt4 <- suppressMessages(distance_jsd %>% 
+                                           clust_gran(kopt = 4)) %>% 
+  rename("customer_id" = "id") %>% 
+  mutate(group = as.factor(group))
+
+cluster_result_id4and5 <- cluster_result_id %>% left_join(cluster_result_kopt4, by="customer_id") %>% rename("group4" = "group.y") %>% rename("group5" = "group.x")
+
+#data-heatmap-hod-group-new-4
+
+data_group <- data_pick_robust  %>% 
+  left_join(cluster_result_id4and5, by = c("customer_id"))
+
+data_heatmap_hod_group <- quantile_gran(data_group,
+                                        gran1="hour_day",
+                                        quantile_prob_val = c(0.25, 0.5, 0.75),
+                                        group="group4") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values) 
+
+
+data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
+
+data_heatmap_hod_group$group4 <- paste("A", data_heatmap_hod_group$group4, sep = "-")
+
+hod_group <- data_heatmap_hod_group %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=group4,
+                  fill = as.factor(group4), alpha = 0.5),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=group4, 
+                color = as.factor(group4)), size = 1)+
+  facet_wrap(~group4, 
+             scales = "free_y",  
+             nrow = 5, labeller = "label_value") + 
+  #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
[…]

### 2021-11-15 main.Rmd

@@
 
-```
-
-```{r cluster-number-selection-validation, fig.cap="sindex plotted as a function of the number of clusters for S1, S2, and S3.
-S1 has a sharp decrease in sindex from 5 to 6, whereas S2 and S3 have a decrease from 4 to 5.
+```{r cluster-number-selection-validation, fig.cap="Separation index plotted as a function of the number of clusters for S1, S2, and S3.
+S1 has a sharp decrease from 5 to 6, whereas S2 and S3 have a decrease from 4 to 5.
 As a result, the number of clusters considered for S1, S2, and S3 is 5, 4, 4, respectively.
-This corresponds to the number of designs taken into account in each scenario.", eval=FALSE}
+This corresponds to the number of designs taken into account in each scenario."}
 ```

### 2021-11-15 main.Rmd
typo in application fixed

@@
 This corresponds to the number of designs taken into account in each scenario."}
-
-```{r cluster-number-selection-validation, fig.cap="Separation index plotted as a function of the number of clusters for S1, S2, and S3.
-S1 has a sharp decrease from 5 to 6, whereas S2 and S3 have a decrease from 4 to 5.
-As a result, the number of clusters considered for S1, S2, and S3 is 5, 4, 4, respectively.
-This corresponds to the number of designs taken into account in each scenario."}
 ```
@@
 
-Things become far more complicated when we consider a larger data set with more uncertainty, as they do with any clustering problem.
-
 - Choose a smaller subset of randomly selected $600$ customers with no implicit missing values for 2013.
@@
 - Remove customers whose energy consumption is 0 in all deciles.
-These are the clients whose consumption is likely to remain essentially flat and with no intriguing repeated patterns that we are interested in studying.
+These are the clients whose consumption is likely to remain essenatially flat and with no intriguing repeated patterns that we are interested in studying.
 
@@
 
-The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in \ref{fig:opt-cluster-tsne-jsd}.
+The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}.
 The $sindex$ falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases.
@@
 A-2, A-3, and A-4 have profiles that correspond to groups B-2, B-3, and B-5, respectively.
-B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-5 (id: 21-24) have a 'hod' pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day.
+B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-5 (id: 21-24) have a `hod` pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day.
 A-1 is subdivided into B-1 and B-4, each of which has a distinct shape across 'moy' and 'wnwd'.
@@
 
-The $sindex$ is plotted against the number of clusters for clustering based on wpd-based distances in \ref{fig:opt-cluster-tsne-jsd}.
-The scales are very different from JS-based clustering and also there is more gradual decrease in the separation index.
+In Figure \ref{fig:opt-cluster-tsne-jsd}, the $sindex$ is plotted against the number of clusters for clustering based on wpd-based distances.
+The scales are very different from those used in JS-based clustering.
 The $sindex$ falls from around $20$ when $k=3$ to $19$ when $k=4$ and further decreases to $17.5$ with $k>7$.
-Thus, there are few choices available for number of clusters.
+Thus, there are few choices available for the number of clusters.
 The groupings with $k=4$ is shown through a parallel coordinate plot with the three significant cyclic granularities.
 The variables are sorted according to their separation across classes (rather than their overall variation between classes).
-This means that $moy$ is the most important variable in distinguishing the groups followed by $hod$ and $wnwd$.
-The two customers (id: 16 and 18) which has significant differences between $wnwd$ catgeories (in Figure \ref{fig:hod-ind-group-png}) stand out from the rest of the customers.
+This means that $moy$ is the most important variable in distinguishing the groups, followed by $hod$ and $wnwd$.
+The two customers (id: 16 and 18) that have significant differences between $wnwd$ categories (in Figure \ref{fig:hod-ind-group-png}) stand out from the rest of the customers.
 Group-4 has a higher $wpd$ for `hod` than `moy` or `wkndwday`.
-Group-2 and Group-3 have the most distinct pattern across `moy`.
+Group-2 and Group-3 have the most distinct patterns across `moy`.
 Group-1 is a mixed group that has strong patterns on at least one of the three variables.
-When $k=3$ is used, Group-2 and Group-3 are merged together because their patterns relative to `moy`, '`hod` and `wnwd` are same.
+When $k=3$ is used, Group-2 and Group-3 are merged because their patterns relative to `moy`, `hod` and `wnwd` are the same.
 The findings vary from JS-based clustering, yet it is a helpful grouping.
@@
 Plot (b) which shows median $wpd$ values for each group.
-The plot shows that `moy` is the most important variable in identifying clusters, whereas `wnwd` is the least significant and has the least fluctuation.
+The plot shows that moy is the most important variable in identifying clusters, whereas wnwd is the least significant and has the least fluctuation.
 Two customers (id: 16 and 18) with high $wpd$ across wnwd stand out in this display.
-Group-4 has a higher $wpd$ for `hod` than `moy` or `wnwd`.
-Group-2 and Group-3 has most discernible pattern across `moy`.
+Group-4 has a higher $wpd$ for hod than moy or wnwd.
+Group-2 and Group-3 has most discernible pattern across moy.
 Group-1 is a mixed group with strong patterns across atleast one of the three variables.

### 2021-11-15 main.Rmd
remove overlap of point and text in figure 9

@@
 This is analogous to subsampling the population along all dimensions of interest such that the sampled data represents the primary features of the underlying distribution.
-Instance selection in unsupervised learning has received little attention in the literature, yet it could be a useful tool for evaluating model or method performance.
+Instance selection in unsupervised learning has received little attention in the literature.
+However, it could be a useful tool for evaluating model or method performance.
 There are several ways to approach prototype selection.
 Following @Fan2021-bq's idea of picking related examples (neighbors) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2D space.
-Then pick a few "anchor" customers who are far apart in 2D space and pick a few neighbors for each.
+Then pick a few "anchor" customers far apart in 2D space and pick a few neighbors for each.
 Unfortunately, this does not ensure that consumers with significant patterns across all variables are chosen.
 Tours can reveal variable separation that gets hidden in a single variable display better than static projections.
-Hence we perform a linked tour with a t-SNE layout using the R package `liminal` (@R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
-(Refer to Supplementary article for further details).
-Figure \ref{fig:hod-ind-group-png} shows the distribution across `hod` (a), `moy`(b) and `wnwd` (c) for the set of chosen $24$ customers that were chosen.
-Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution.
+Hence, we perform a linked tour with a t-SNE layout using the R package `liminal` (@R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
+(Refer to the supplementary article for further details).
+Figure \ref{fig:hod-ind-group-png} shows the distribution across `hod`, `moy` and `wnwd` for the set of chosen $24$ customers that were chosen.
+The customers are split into batches of $12$ in (a) and (b).
+Each row in (a) and that in (b) represents the profile of one customer across different variables.
+<!--Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution.
+-->
 
@@
 In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups.
---> Groups by JS-based distances and wpd-based distances are colored differently as they represent different groupings.
-While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominate the shape of the summarized clusters because of magnitude.
+--> Groups are colored differently whenever they represent different groupings.
+While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude.
 The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales.
@@
 Figure \ref{fig:groups-4and5} depicts the summarised distributions across $4$ and  $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster.
-A-2, A-3, and A-4 have profiles that correspond to groups B-2, B-3, and B-5, respectively.
+A-2, Groups A-3, and A-4 have profiles that correspond to B-2, B-3, and B-5, respectively.
 B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-5 (id: 21-24) have a `hod` pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day.
-A-1 is subdivided into B-1 and B-4, each of which has a distinct shape across 'moy' and 'wnwd'.
+A-1 is subdivided into B-1 and B-4, each of which has a distinct shape across `moy` and `wnwd`.
 B-4 (id: 16-20) is distinguished by higher energy consumption in the first few months of the year and greater variability in weekend usage.
@@
 Furthermore, combining heterogeneous customers may result in similar-looking final clusters that are not effective for visually differentiating them.
-It is also worth noting that the Weekend/Weekday behavior in the given case does not characterize any cluster.
+It is also worth noting that the Weekend/Weekday behavior in the given case does not characterize most selected customers.
 This, however, will not be true for all of the customers in the data set.
@@
 Thirdly, it is possible that for a few customers, data for some categories from the list of considered significant granularities are missing.
-In our application, we have removed those customers and done the analysis but the metrics used should be able to incorporate those customers in the clustering by handling their missing categories.
+In our application, we have removed those customers and done the analysis but the metrics used should be able to incorporate those customers with such structured missingness.
 Finally, $wpd$ is computationally heavy even under parallel computation.

### 2021-11-15 main.Rmd
scanning typos throughout the paper

@@
 
-The primary application of this work is data from the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
+The primary application of this work is data from the Smart Grid, Smart City (SGSC) project (2010–2014) available through the [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
 Half-hourly measurements of usage for more than 13,000 household electricity smart meters is provided from from October 2011 to March 2014.
@@
 Only one study [@ozawa2016determining] identified time series characteristics by first conducting a Fourier transformation, to convert data from time to frequency domain, followed by $k$-means to cluster by greatest frequency.
-@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential customers and addresses model-based clustering by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered.
+@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered.
 [@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
@@
 This needs to be done in a way such that the resulting metric is a distance metric, and could be fed into the clustering algorithm.
-Two types of distance metric is considered:
+Two types of distance metrics are considered:
 
@@
 
-The sum of the distances between two observations $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
+The sum of the distances between two observations $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$).
 After determining the distance between two series in terms of one granularity, we must combine them to produce a distance based on all significant granularities.
@@
 $wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalized by appropriate parameters.
-A higher value of $wpd$ indicates that some interesting pattern is expected, whereas a lower value would indicate otherwise.
+A higher value of $wpd$ indicates that some interesting patterns are expected, whereas a lower value would indicate otherwise.
 
@@
 Displaying cluster statistics is useful when we have larger problems and it is difficult to read the parallel coordinate plots due to congestion.
-All of MDS, PCA and t-SNE use a distance or dissimilarity matrix to construct a reduced-dimension space representation, their goals are diverse.
+All of MDS, PCA and t-SNE use a distance or dissimilarity matrix to construct a reduced-dimension space representation, but their goals are diverse.
 Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
@@
 A continuous measured variable $v$ of length $T$ indexed by $\{0, 1, \dots T-1\}$ is simulated such that it follows the structure across $g1$, $g2$ and $g3$.
-We constructed independent replications of all data designs $R = \{25, 250, 500\}$ to investigate if our proposed clustering method can discover distinct designs in small, medium, and big number of series.
+We constructed independent replications of all data designs $R = \{25, 250, 500\}$ to investigate if our proposed clustering method can discover distinct designs in small, medium, and big numbers of series.
 All designs employ $T=\{300, 1000, 5000\}$ sample sizes to evaluate small, medium, and large-sized series.
@@
 ```{r gran2and1-clubbed, fig.cap=" Boxplots showing distributions of categories across different designs (rows) and granularities (columns) for scenarios S2 and S3.
-In S2, $g2$, $g3$ change across atleast one design but $g1$ remains constant.
+In S2, $g2$, $g3$ change across at least one design but $g1$ remains constant.
 Only $g3$ changes across different designs in S3."}
@@
 Consider a group that has a different weekday and weekend behavior in the summer but not in the winter.
-This type of combined behavior across granularities can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (Weekend/Weekday and month-of-year in this example).
+This type of combined behavior across granularities can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (weekend/weekday and month-of-year in this example).
 As a result, in this scenario, we analyze a combination of categories generated from different distributions.
@@
 The lines are not colored by group since the purpose is to highlight the contribution of the factors to categorization rather than class separation.
-The first plot shows that no variable stands out in the clustering, but the following two designs show that {g1} and {g1, g2} have very low inter cluster distances, meaning that they did not contribute to the clustering.
+The first plot shows that no variable stands out in the clustering, but the following two designs show that {g1} and {g1, g2} have very low inter-cluster distances, meaning that they did not contribute to the clustering.
 It is worth noting that these facts correspond to our original assumptions when developing the scenarios, which incorporate distributional differences over three (S1), two (S2), and one (S3) significant granularities.
@@
  - The JS-RS and wpd-based methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
-Although, a complete year of data is the minimum requirement to capture distributional differences in winter and summer profiles, for example.
+However, a complete year of data is the minimum requirement to capture distributional differences in winter and summer profiles, for example.
 Even if the data is only available for a month, $nT$ with half-hourly data is expected to be at least $1000$.
@@
 - Obtain $wpd$ for all cyclic granularities considered for these customers.
-It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weekend/weekday) are coming out to be significant for most customers.
+It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weekend/weekday) are turning out to be significant for most customers.
 We use these three granularities while clustering.
@@
 There are several ways to approach prototype selection.
-Following @Fan2021-bq's idea of picking related examples (neighbors) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2D space.
+Following the idea of @Fan2021-bq of picking related examples (neighbors) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2D space.
 Then pick a few "anchor" customers far apart in 2D space and pick a few neighbors for each.
@@
 Group-4 has a higher $wpd$ for hod than moy or wnwd.
-Group-2 and Group-3 has most discernible pattern across moy.
-Group-1 is a mixed group with strong patterns across atleast one of the three variables.
+Group-2 and Group-3 have the most discernible pattern across moy.
+Group-1 is a mixed group with strong patterns across at least one of the three variables.
 All of these could be observed from the plot or the table (b).", message=FALSE, warning=FALSE, fig.height=6}
@@
 Furthermore, combining heterogeneous customers may result in similar-looking final clusters that are not effective for visually differentiating them.
-It is also worth noting that the Weekend/Weekday behavior in the given case does not characterize most selected customers.
+It is also worth noting that the weekend/weekday behavior in the given case does not characterize most selected customers.
 This, however, will not be true for all of the customers in the data set.
@@
 
-There are few areas to extend this research.
+There are a few areas to extend this research.
 First, larger data sets with more uncertainty complicate matters, as is true for any clustering task.
@@
 This article was created with R [@R-language], `knitr` [@knitr, @R-knitr] and `rmarkdown` [@rmarkdown, @R-rmarkdown].
-Graphics are produced with `ggplot2` [@Wickham2009pk] and GGally [@R-GGally].
+Graphics are produced with `ggplot2` [@Wickham2009pk] and `GGally` [@R-GGally].
 
@@
 
-**Supplementary paper:** Additional tables, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd).
+**Supplementary paper:** Additional tables, graphics and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd).
 The code for creating validation designs and running the methodologies is available at (https://github.com/Sayani07/paper-gracsR/Validation).

### 2021-11-15 main.Rmd
some more typo fixed

@@
 It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values.
-MDS and parallel coordinate graphs are used to demonstrate the findings, along with an index value plot for guidance on the number of clusters.
+MDS and parallel coordinate graphs are used to demonstrate the findings, as well as an index value plot to provide direction on the number of clusters.
 JS-based approaches corresponding to NQT and RS are referred to as JS-NQT and JS-RS respectively.
@@
 
-- Figure \ref{fig:sindex-plot-validation} shows sindex plotted against the number of clusters ($k$) for the range of mean differences (rows) under the different scenarios (columns).
+- Figure \ref{fig:sindex-plot-validation} shows $sindex$ plotted against the number of clusters ($k$) for the range of mean differences (rows) under the different scenarios (columns).
 This can be used to determine the number of clusters for each scenario.
-When sindex for each scenario are examined, it appears that $k = 5, 4, 4$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in sindex from that value of $k$.
-The number of clusters corresponds to the number of designs that were originally considered in each scenario.
-
-- In Figure \ref{fig:mds-plot-validation}, we tried to see how separated our clusters are.
-We observe that in all scenarios and for different mean differences, clusters are separated.
+When $sindex$ for each scenario are examined, it appears that $k = 5, 4, 4$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in $sindex$ from that value of $k$.
+Thus, tshe number of clusters corresponds to the number of designs that were originally considered in each scenario.
+
+- Figure \ref{fig:mds-plot-validation} shows separation of our clusters.
+It can be observed that in all scenarios and for different mean differences, clusters are separated.
 However, the separation increases with an increase in mean differences across scenarios.
@@
 The lines are not colored by group since the purpose is to highlight the contribution of the factors to categorization rather than class separation.
-The first plot shows that no variable stands out in the clustering, but the following two designs show that {g1} and {g1, g2} have very low inter-cluster distances, meaning that they did not contribute to the clustering.
+Panel S1 shows that no variable stands out in the clustering, but the following two panels show that {g1} and {g1, g2} have very low inter-cluster distances, meaning that they did not contribute to the clustering.
 It is worth noting that these facts correspond to our original assumptions when developing the scenarios, which incorporate distributional differences over three (S1), two (S2), and one (S3) significant granularities.
@@
 
-```{r parcoord-sim, fig.cap = "The parallel coordinate plot illustrates the total inter-cluster distances for granularities $g1$, $g2$, and $g3$.
-One line in the figure depicts the inter-cluster distances for a single simulation scenario.
-While the first plot indicates that no variable stands out during clustering, the next two designs demonstrate that g1 and (g1, g2) have relatively lower inter-cluster distances, indicating that they did not contribute to clustering.
-It is worth emphasising that these facts are consistent with our initial assumptions when designing the scenarios and S1, S2, S3 correspond to scenarios with three, two and one significant granularity respectively."}
-
-```
-
 ```{r mds-data-validation}
@@
 Between scenarios, separation is least prominent corresponding to Scenario (S3) where only one granularity is responsible for distinguishing the clusters."}
+
+```
+
+```{r parcoord-sim, fig.cap = "The parallel coordinate figure depicts total inter-cluster distances for the granularities $g1$, $g2$, and $g3$.
+The inter-cluster distances for a single simulation scenario are represented by one line in the figure.
+While panel S1 shows that no variable stands out during clustering, panels S2 and S3 show that g1 and (g1, g2) had smaller inter-cluster distances, indicating that they did not contribute to clustering.
+These facts are consistent with our initial assumptions when building the scenarios, and S1, S2, S3 correspond to scenarios with three, two, and one significant granularity, respectively."}
 

### 2021-11-15 main.Rmd
split code + clean code

@@
 ```{r external, include = FALSE, cache = FALSE}
-knitr::read_chunk('script/main.R')
+#knitr::read_chunk('script/main.R')
+knitr::read_chunk('script/mylibrary.R')
+knitr::read_chunk('script/mytheme.R')
+knitr::read_chunk('script/application.R')
+knitr::read_chunk('script/application.R')
+knitr::read_chunk('script/validation.R')
 ```
@@
 
+```{r theme_characterisation}
+```
+
+```{r mytheme-application2}
+
+```
+
 ```{r mytheme-application3}
 
-```
-
-```{r theme_characterisation}
 ```
@@
 
-```{r hod-moy-wkndwday plots, fig.cap="The distribution across moy, hod and wkndwday for the selected designs.
+<!-- # ```{r hod-moy-wkndwday plots, fig.cap="The distribution across moy, hod and wkndwday for the selected designs.
 Few are similar in their hod pattern, while others are similar in moy behavior.
-Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most households, there is one household for whom weekend behavior is standing out from the rest."}
-```
+Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most households, there is one household for whom weekend behavior is standing out from the rest."} -->
+<!-- # ``` -->
 
@@
 
+```{r data-pick-robust}
+
+```
+
+```{r hod-moy-wkndwday-data}
+
+```
+
 ```{r opt-clusters-jsd}
@@
 #include_graphics("figs/opt_clust_parenthesis.png")
+```
+
+```{r hod-ind-group}
 ```
@@
 
-```{r combined4and5, fig.cap = "For $k = 4$ (a) and $k=5$ (b), the distribution of electricity demand across groups over hod, moy, and wnwd is shown.
-Groups A-2, A-3, and A-4 profiles correspond to Groups B-2, B-3, and B-5, respectively.
-A-1 is split into B-1 and B-4, each of which has a distinct shape across moy and wnwd.
-B-4 (id: 16-20) is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage.
-B-1 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.
-When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group."}
-#knitr::include_graphics("figs/combined4and5.png")
-```
-
 ```{r groups-4and5, fig.cap = "For $k = 4$ (a) and $k=5$ (b), the distribution of electricity demand across groups over hod, moy, and wnwd is shown.
@@
 B-1 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.
-When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group."}
+When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group.", fig.height=6}
 

### 2021-11-15 paper-supplementary.Rmd
split code + clean code

@@
 read_chunk(here::here("validation/simdata01.R"))
+knitr::read_chunk('script/gracsr-supplementary.R')
 ```
@@
 ```
+
+```{r sindex-data-validation}
+
+```
+
+```{r sindex-plot-validation-wpd}
+
+```

### 2021-11-15 application_Check.R

@@ -1,63 +1,0 @@
-library(gracsr)
-library(tidyverse)
-library(readr)
-library(tsibble)
-library(gravitas)
-
-
-data_356cust_hod <- read_rds("data/quantile_data_356cust_hod_robust.rds") %>% 
-  filter(quantiles %in% "50%")
-
-data_356cust_moy <- read_rds("data/quantile_data_356cust_moy_robust.rds") %>% 
-  filter(quantiles %in% "50%")
-
-data_356cust_wkndwday <- read_rds("data/quantile_data_356cust_wkndwday_robust.rds") %>% 
-  filter(quantiles %in% "50%")
-
-
-data_356cust_hod_wide <- data_356cust_hod %>%
-  pivot_wider(names_from = c("gran", "category", "quantiles"),
-              values_from = "quantiles_values")
-
-data_356cust_moy_wide <- data_356cust_moy %>%
-  pivot_wider(names_from = c("gran", "category", "quantiles"),
-              values_from = "quantiles_values")
-
-data_356cust_wkndwday_wide <- data_356cust_wkndwday %>%
-  pivot_wider(names_from = c("gran", "category", "quantiles"),
-              values_from = "quantiles_values")
-
-data_356cust_wide <- left_join(data_356cust_hod_wide,
-                               data_356cust_moy_wide, by="customer_id") %>% 
-  left_join(data_356cust_wkndwday_wide,  by="customer_id"
-  )
-
-
-data_356cust_pc <- prcomp(data_356cust_wide[,-1],
-                          center = FALSE, scale = FALSE, retx = TRUE)
-
-plot(data_356cust_pc, type="l", npcs=50)
-
-data_24cust_wide <- data_356cust_wide %>% filter(customer_id %in% data_pick$customer_id)
-
-set.seed(2935)
-
-tSNE_fit <- data_24cust_wide%>% 
-  select(-customer_id) %>% 
-  Rtsne(pca = FALSE, perplexity = 2)
-
-
-tsne_df <- data.frame(tsneX = tSNE_fit$Y[, 1], 
-                      tsneY = tSNE_fit$Y[, 2], 
-                      customer_id = as.character(data_24cust_wide$customer_id))
-
-tsne_xy <- ggplot(tsne_df, aes(x = tsneX, y = tsneY)) +
-  geom_point(aes(text = customer_id), size =2) +
-  #scale_color_manual(values = limn_pal_tableau10()) +
[…]

### 2021-11-15 check.R

@@ -1,14 +1,0 @@
-collapse_x <- function(x)
-{
-  y = unlist(x)
-  z = paste(y, collapse = ",")
-}
-
-x = list(1, 5, 2, c(1,5))
-cluster_summary <- cluster_result_id %>% select(-customer_id, -divide_cust) %>%  group_by(group) %>% nest(data = c(id))
-
-apply(cluster_summary, 1, collapse_x)
-z =  rbind(group = cluster_summary$group, rbind(members = map(cluster_summary$data, collapse_x))) %>% t()
-
-z =  rbind(group = cluster_summary$group, rbind(members = map(cluster_summary$data, collapse_x))) %>% t()
-

### 2021-11-15 script/application.R

@@ -1,0 +1,841 @@
+##### application
+
+## ----prototype-data-pick-------------------------------------------------------
+quantile_prob_graph <- c(0.25, 0.5, 0.75)
+
+data_pick_one <- c(8541744, 9355808, 8603880, 8619309, 10542667) %>%
+  as_tibble() %>%
+  set_names("customer_id")
+# data_pick_two <- c(8688242, 8643837, 8184707, 10534355, 8684420) %>% as_tibble%>% set_names("customer_id")
+data_pick_three <- c(9792072, 8589936, 8454235, 10692366, 8603828) %>%
+  as_tibble() %>%
+  set_names("customer_id")
+data_pick_four <- c(8618759, 8291696, 10357256, 8290374) %>%
+  as_tibble() %>%
+  set_names("customer_id")
+data_pick_five <- c(9044864, 8642053, 10534367, 9021526, 11162275) %>%
+  as_tibble() %>%
+  set_names("customer_id")
+data_pick_six <- c(8221762, 8273636, 10359424, 8232822, 11450499) %>%
+  as_tibble() %>%
+  set_names("customer_id")
+
+
+
+## ----assemble------------------------------------------------------------------
+data_pick_cust <- bind_rows(
+  data_pick_one,
+  # data_pick_two,
+  data_pick_three,
+  data_pick_four,
+  data_pick_five,
+  data_pick_six,
+  .id = "design"
+) %>%
+  mutate(customer_id = as.character(customer_id))
+
+## ----all-data-----------------------------------------------------------------
+data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
+  mutate(customer_id = as.character(customer_id)) %>%
+  dplyr::filter(customer_id %in% data_pick_cust$customer_id) %>%
+  gracsr::scale_gran(
+    method = "nqt",
+    response = "general_supply_kwh"
+  )
+
+
+
+## ----clustering---------------------------------------------------------------
+hod <- suppressMessages(data_pick %>%
+  dist_gran(
+    gran1 = "hour_day",
+    response = "general_supply_kwh"
+  ))
+
+moy <- suppressMessages(data_pick %>%
+  dist_gran(
+    gran1 = "month_year",
[…]

### 2021-11-15 script/gran2_orig.R

@@ -1,323 +1,0 @@
-# Rscripts for simulation chapter 3, patterns across 3 grans changing atleast across one design amongst 5 designs
-
-library(conflicted)
-library(tidyverse)
-library(gracsr)
-library(tsibble)
-library(gravitas)
-library(dplyr)
-library(here)
-
-# fix parameters
-
-# niter
-#seed_len = seq(12345, 12350, 1)
-# number of iterations corresponding to
-niter <- c(5, 50, 100) #number of series you are clustering
-nT <-  c(300, 1000, 5000) # length of the time series
-mean_diff <- c(1, 2, 5) # difference between consecutive categories
-
-simtable <- expand.grid(mean_diff = mean_diff,
-                        niter = niter,
-                        #time_series = time_series, 
-                        nT = nT)
-
-# parallel::mclapply(seq_len(nrow(simtable)), function(scen){
-
-scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
-#scen <- 10
-
-
-simj<-simtable[scen,] #Extract row of table
-mean_diffj <- simj$mean_diff
-niterj <- simj$niter
-nTj <- simj$nT
-# time_seriesj <- simj$time_series
-#seed_lenj <- simj$seed_len
-
-generate_design <- function(n = 300, #length of time series
-                            mu11 = 0,
-                            mu12 = 0,
-                            mu21 = 0,
-                            mu22 = 0 ,
-                            mu23 = 0,
-                            mu31 = 0,
-                            mu32 = 0,
-                            mu33 = 0,
-                            mu34 = 0,
-                            mu35 = 0)
-{
-  
-  n <- n+500 # 500 burning observations
-  t <- seq(0, n-1, 1)
-  
-  g1 <- t %%2
-  g2 <- t %%3
-  g3 <- t %%5
-  
-  str_gran <- bind_cols(index = t, g1 = g1, g2 = g2, g3 = g3)
-  
-  # calculation of g1
-  g1_table <- bind_cols(g1 = unique(g1), dist_mean = c(mu11, mu12))
-  
-  g1_tally <- str_gran %>% group_by(g1) %>%
-    count() %>%
-    left_join(g1_table, by = "g1")
-  
[…]

### 2021-11-15 script/gran3_orig.R

@@ -1,281 +1,0 @@
-# Rscripts for simulation chapter 3, patterns across 3 grans changing atleast across one design amongst 5 designs
-
-library(conflicted)
-library(tidyverse)
-library(gracsr)
-library(tsibble)
-library(gravitas)
-library(dplyr)
-library(here)
-
-
-# fix parameters
-
-# niter
-#seed_len = seq(12345, 12350, 1)
-# number of iterations corresponding to
-niter <- c(5, 50, 100) #number of series you are clustering
-nT <-  c(300, 1000, 5000) # length of the time series
-mean_diff <- c(1, 2, 5) # difference between consecutive categories
-
-# number of iterations corresponding to
-# niter <- c(5) #number of series you are clustering
-# nT <-  c(300, 1000, 5000) # length of the time series
-# mean_diff <- c(1) # difference between consecutive categories
-
-#time_series <- c("ar1", "arma22")
-
-simtable <- expand.grid(mean_diff = mean_diff,
-                        niter = niter,
-                        #time_series = time_series, 
-                        nT = nT)
-
-# parallel::mclapply(seq_len(nrow(simtable)), function(scen){
-
-scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
-#scen <- 27
-
-
-simj<-simtable[scen,] #Extract row of table
-mean_diffj <- simj$mean_diff
-niterj <- simj$niter
-nTj <- simj$nT
-# time_seriesj <- simj$time_series
-#seed_lenj <- simj$seed_len
-
-generate_design <- function(n = 300, #length of time series
-                            mu11 = 0,
-                            mu12 = 0,
-                            mu21 = 0,
-                            mu22 = 0 ,
-                            mu23 = 0,
-                            mu31 = 0,
-                            mu32 = 0,
-                            mu33 = 0,
-                            mu34 = 0,
-                            mu35 = 0)
-{
-  
-  n <- n+500 # 500 burning observations
-  t <- seq(0, n-1, 1)
-  
[…]

### 2021-11-15 script/main.R

@@ -1,1897 +1,0 @@
-## ---- load-lib
-
-#theme_set(theme_bw())
-library(ggplot2)
-library(lubridate)
-library(gravitas)
-library(gracsr)
-library(ggdendro)
-library(dplyr)
-library(readr)
-library(visdat)
-library(ggplot2)
-library(tidyverse)
-library(naniar)
-library(here)
-library(tsibble)
-library(knitr)
-library(patchwork)
-library(GGally)
-library(distributional)
-library(viridis)
-library(forcats)
-library(tidyr)
-library(purrr)
-library(stringr)
-library(pals)
-library(Rtsne)
-library(fpc)
-library(cluster)
-library(ggpubr)
-library(grid)
-library(ggrepel)
-
-
-##----theme-validation
-theme_validation <- function() {
-  theme_bw() +
-    theme(
-      strip.text = element_text(size = 8, 
-                                margin = margin(b = 0, t = 1)),
-      plot.margin = margin(0, 0, 0, 0, "cm"),
-      axis.title.y = element_blank(),
-      #axis.text.x = element_blank(),
-      axis.text.y = element_blank(),
-      axis.ticks = element_blank(),
-      legend.position = "bottom"
-    )
-}
-
-
-##----theme_characterisation
-
-theme_characterisation <- function() {
-  
-  theme_bw() + # seeting theme
-    theme(strip.text = element_text(size = 10,
-                                    margin = margin(b = 0, t = 0))) + # narrow facet space
-    theme(axis.title.y=element_blank(),
-          axis.text.y=element_blank(),
-          axis.ticks.y=element_blank()) + # no axis ticks 
-    theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
-    theme(axis.text.x = element_text(angle=90, hjust=1, size = 10)) + # rotate the x-axis text
-    theme(legend.position = "bottom")+
-    theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-    theme(axis.text.x = element_text(size=5)) +
-    theme(axis.text = element_text(size = 12))  
-}
-
-
-##----mytheme-application
-theme_application <- function() {
-  
-  theme_light() + # setting theme
-    theme(strip.text = element_text(size = 10,
[…]

### 2021-11-15 split_sindex.R

@@ -1,81 +1,0 @@
-sindex_data <- function(data){
-  
-  total_distance <- data %>%
-  group_by(item1, item2) %>% 
-  summarise(total = sum(distance), .groups = "drop") %>% ungroup()
-
-total_distance_wide <- total_distance%>% 
-  pivot_wider(names_from = item2, values_from = total)
-
-rownames(total_distance_wide) <- total_distance_wide$item1
-
-distance_lowertriangle <- total_distance_wide %>% 
-  mutate_all(~replace(., is.na(.), 0)) %>%
-  tibble::rownames_to_column() %>%  
-  dplyr::select(-item1) %>% 
-  pivot_longer(-rowname) %>% 
-  pivot_wider(names_from=rowname, values_from=value) 
-
-rownames(distance_lowertriangle) <- total_distance_wide$item1
-
-df <- distance_lowertriangle[-1] %>% as.matrix()
-DM <- matrix(0, ncol(distance_lowertriangle), ncol(distance_lowertriangle)) 
-DM[lower.tri(DM)] = df[lower.tri(df, diag=TRUE)] # distance metric
-#print(i)
-distance_matrix = as.dist(DM)
-
-groups <- bind_rows(data %>%
-                      distinct(item1, group_item1) %>% 
-                      rename("item" = "item1",
-                             "group" = "group_item1"),
-                    data %>% 
-                      distinct(item2, group_item2) %>%
-                      rename("item" = "item2",
-                             "group" = "group_item2")) %>% distinct(item, group) 
-
-
-k = array()
-for(j in 2:10)
-{
-  group <- distance_matrix %>% 
-    hclust (method = "ward.D") %>% 
-    cutree(k=j)
-  #print(j)
-  p <- cluster.stats(distance_matrix,
-                     clustering = group, 
-                     silhouette = TRUE)
-  k[j]=p$sindex
-}
-k %>% as_tibble() %>% set_names(c("sindex"))%>%
-  mutate(k = row_number())
-}
-
-
-
-
-data1 <- sindex_data(read_rds("validation/js-robust/3gran_change_5D/data_validation_25.rds")) %>% mutate(diff = 1 , design= "S1")
-data2 <-sindex_data(read_rds("validation/js-robust/3gran_change_5D/data_validation_26.rds")) %>% mutate(diff = 2 , design= "S1")
[…]

### 2021-11-15 paper-supplementary.Rmd
supplementary update

@@
 ```{r external, include = FALSE, cache = FALSE}
-knitr::read_chunk('script/main.R')
+knitr::read_chunk('script/mylibrary.R')
+knitr::read_chunk('script/mytheme.R')
+knitr::read_chunk('script/validation.R')
 read_chunk(here::here("validation/simdata01.R"))
-knitr::read_chunk('script/gracsr-supplementary.R')
+knitr::read_chunk('script/supplementary.R')
 ```

### 2021-11-15 paper-supplementary.Rmd
supplementary document update

@@
 ```{r combined-groups-js-entire, fig.cap = "The distribution of electricity demand for the clusters across hod, moy and wkndwday for the $17$ groups from $353$ customers.
-Wknd-wday variations across groups are not distinguishable, but ideally each group should have an unique combination of hod and moy.", fig.height=6}
+Wknd-wday variations across groups are not distinguishable, but ideally each group should have an unique combination of hod and moy.", fig.height=8}
 
@@
 
-```{r sindex-data-validation}
+```{r sindex-data-valid-supplement}
 
@@
 
-```{r sindex-plot-validation-wpd}
+```{r sindex-plot-valid-supplement, fig.cap="The cluster separation index (sindex) is plotted as a function of the number of clusters for the range of mean differences (rows) under the different scenarios (columns) for JS-RS based distance clustering.
+S1 has a sharp decrease in sindex from 5 to 6, whereas S2 and S3 have a decrease from 4 to 5.
+As a result, the number of clusters considered for S1, S2, and S3 is 5, 4, 4, respectively.
+This corresponds to the number of designs taken into account in each scenario"}
 

### 2021-11-16 main.Rmd
small tweak

@@
 knitr::read_chunk('script/mytheme.R')
-knitr::read_chunk('script/application.R')
 knitr::read_chunk('script/application.R')

### 2021-11-16 script/application.R

@@ -35,7 +35,7 @@ data_pick_cust <- bind_rows(
   mutate(customer_id = as.character(customer_id))
 
 ## ----all-data-----------------------------------------------------------------
-data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
+data_pick <- readr::read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
   mutate(customer_id = as.character(customer_id)) %>%
   dplyr::filter(customer_id %in% data_pick_cust$customer_id) %>%
   gracsr::scale_gran(

### 2021-11-16 main.Rmd
added files on gitignore

--- 2021-11-15 main.Rmd
@@
 abstract: |
- Clustering is a potential approach for organising large collections of times series into small homogeneous groups, but a difficult step is determining an appropriate metric to measure similarity between time series.
+ Clustering is a potential approach for Organizing large collections of times series into small homogeneous groups, but a difficult step is determining an appropriate metric to measure similarity between time series.
 The similarity metric needs to be capable of accommodating long, noisy, and asynchronous time series and also capture cyclical patterns.
@@
 Cyclic granularities like hour-of-the-day, work-day/weekend, month-of-the-year, are useful for finding repeated patterns in the data.
-Measuring similarity based on probability distributions across cyclic granularities serves two purposes: (a) characterising the inherent temporal data structure of long, unequal-length time series in a manner robust to missing or noisy data; (b) small pockets of similar repeated behaviors can be captured.
+Measuring similarity based on probability distributions across cyclic granularities serves two purposes: (a) characterizing the inherent temporal data structure of long, unequal-length time series in a manner robust to missing or noisy data; (b) small pockets of similar repeated behaviors can be captured.
 This approach is capable of producing useful clusters, as demonstrated on validation data designs and a sample of residential smart meter records.
@@
 energy usage) affects distance calculations.
-Skewed distributions need to be symmetrised.
-Scales of individuals needs to be standardised, because clustering is to select similar patterns, not magnitude of usage.
-(Organising individuals based on magnitude can achieved simply by sorting on a statistic like the average value across time.) <!--Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
+Skewed distributions need to be symmetrized.
+Scales of individuals needs to be standardized, because clustering is to select similar patterns, not magnitude of usage.
+(Organizing individuals based on magnitude can achieved simply by sorting on a statistic like the average value across time.) <!--Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
 Statistical transformations are employed to bring all of them to the same range or normalize each observation.
@@
 
-The data object will change when cyclic granularities are computed, as multiple observations will be caetgorized into levels of the granularity, thus inducing multiple probability distributions.
+The data object will change when cyclic granularities are computed, as multiple observations will be categorized into levels of the granularity, thus inducing multiple probability distributions.
 Directly computing Jensen-Shannon distances between the entire probability distributions can be computationally intensive.
@@
 When $sindex$ for each scenario are examined, it appears that $k = 5, 4, 4$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in $sindex$ from that value of $k$.
-Thus, tshe number of clusters corresponds to the number of designs that were originally considered in each scenario.
+Thus, the number of clusters corresponds to the number of designs that were originally considered in each scenario.
 
@@
 - Remove customers whose energy consumption is 0 in all deciles.
-These are the clients whose consumption is likely to remain essenatially flat and with no intriguing repeated patterns that we are interested in studying.
+These are the clients whose consumption is likely to remain essentially flat and with no intriguing repeated patterns that we are interested in studying.
 
@@
 In our study, we use $k=4, 5$.
-Figure \ref{fig:groups-4and5} depicts the summarised distributions across $4$ and  $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster.
+Figure \ref{fig:groups-4and5} depicts the summarized distributions across $4$ and  $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster.
 A-2, Groups A-3, and A-4 have profiles that correspond to B-2, B-3, and B-5, respectively.
@@
 Depending on the goal of the clustering, these distance metrics, when fed into a hierarchical clustering algorithm using Ward's linkage, yield meaningful clusters.
-Probability distributions provide an intuitive method to characterise noisy, patchy, long, and unequal-length time series data.
-Distributions over cyclic granularities help to characterise the formed clusters in terms of their repeating behavior over these cyclic granularities.
+Probability distributions provide an intuitive method to characterize noisy, patchy, long, and unequal-length time series data.
+Distributions over cyclic granularities help to characterize the formed clusters in terms of their repeating behavior over these cyclic granularities.
 Furthermore, unlike earlier efforts that group customers based on behavior across only one cyclic granularity (such as hour-of-day), our method is more comprehensive in detecting clusters with repeated patterns at all relevant granularities.
@@
 There are various different characteristics of a clustering that can be relevant in practice, depending on the aim of clustering, such as low within-cluster distances and high between-cluster separation.
-In this paper, a number of validation criteria will be introduced that refer to different desirable characteristics of a clustering, and that characterise a clustering in a multidimensional way.
+In this paper, a number of validation criteria will be introduced that refer to different desirable characteristics of a clustering, and that characterize a clustering in a multidimensional way.
 In specific applications the user may be interested in some of these criteria rather than others.

### 2021-11-16 main.Rmd
re-worked clustering, choosing metrics

@@
 
-Different approaches to determining number of clusters are required for varied applications and research objectives.
-It is often beneficial to investigate numerous cluster statistics in order to arrive at a number of clusters, since clustering has a variety of objectives (e.g., between-cluster separation, within-cluster homogeneity, representation by centroids, etc.), which could be in conflict with one another.
-Some typical ways of determining the number of clusters in the literature (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy]) are inspired by defining and balancing "within-cluster homogeneity" and "between-cluster separation".
-The Cluster separation index ($sindex$) proposed by @Hennig2014-ah is used as a guide to select the number of clusters in Section \ref{sec:validation}.
-$sindex$ [@R-fpc] is a compromise between minimum cluster separation and average cluster separation and is less sensitive to a single or few ambiguous points.
-$sindex$ cannot be optimized over the number of clusters, since increasing the number of clusters reduces it.
-The number of clusters required is determined by the slope going from steep to shallow (an elbow).
+Determining the number of clusters is typically a difficult task.
+Many metrics have been defined for choosing clusters.
+Most metrics for choosing the optimal number of clusters, are based on comparing distances between observations within a class to those distances between observations between classes, which makes the assumption that there are some separated clusters.
+Some common procedures include the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy] and the separation index ($sindex$) [@Hennig2014-ah].
+These are constructed by balancing within-cluster homogeneity and between-cluster separation.
+
+These can all give contradictory suggestions when the data does not naturally break into groups, or in the presence of nuisance variables (no contribution to clustering) or nuisance observations (inlying outlying observations falling between clusters).
+There is no one best metric, perhaps a reason why so many exist.
+
+In this work, we have chosen to use $sindex$.
+It is a very simple, but effective metric.
+This is computed by averaging the minimum 10% of distances between observations between clusters, making it relatively robust to nuisance observations.
+The value of $sindex$ always decreases, and where values drop substantially when an extra cluster are candidates for optimal clustering.
+The number of clusters corresponding to the value before the drop is the recommendation.
 

### 2021-11-16 main.Rmd

@@
 
-```{r parcoord-sim, fig.cap = "The parallel coordinate figure depicts total inter-cluster distances for the granularities $g1$, $g2$, and $g3$.
-The inter-cluster distances for a single simulation scenario are represented by one line in the figure.
-While panel S1 shows that no variable stands out during clustering, panels S2 and S3 show that g1 and (g1, g2) had smaller inter-cluster distances, indicating that they did not contribute to clustering.
-These facts are consistent with our initial assumptions when building the scenarios, and S1, S2, S3 correspond to scenarios with three, two, and one significant granularity, respectively."}
+```{r parcoord-sim, fig.cap = "Exploring the contribution of granularities in the clustering for scenarios S1, S2, S3, using parallel coordinate plots.
+Inter-cluster distances are displayed vertically.
+All three granularities $g1$, $g2$, and $g3$ have high inter-cluster distances for S1, suggesting all are important.
+In S2 g1 and in S3 both g1 and g2 have smaller inter-cluster distances, indicating that they did not contribute to clustering."}
 

### 2021-11-16 main.Rmd

@@
 
-Cluster characterization is an important element of cluster analysis.
-@Cook2007-qe provides several methods for characterizing clusters.
-_Parallel coordinate plots_ (@wegman1990hyperdimensional), _Scatterplot matrix_, _Displaying cluster statistics_ (@dasu2005grouping), _MDS_ (@borg2005modern), PCA, _t-SNE_(@R-tsne), _Tour_ (@wickham2011tourr) are some of the graphical approaches used in this study.
-A Parallel Coordinates Plot features parallel axes for each variable and each axis is linked by lines.
+Cluster characterization is an important final stage of a cluster analysis.
+The primary purpose is to compare the homogeneity within a cluster to the heterogeneity of clusters.
+This can be done numerically, by tabulating cluster means and standard deviations [@dasu2005grouping], and visually using methods for graphics multivariate data.
+@Cook2007-qe provides visual examples using both tours [@asimov] and parallel coordinate plots [@wegman1990hyperdimensional].
+Dimension reduction techniques like principal component analysis [@joliffe], multidimensional scaling (MDS) [@borg2005modern], t-distributed stochastic neighbor embedding (t-SNE) [@maarten-tsne] and linear discrimant analysis (LDA) [@lda] are also useful.
+<!-- A Parallel Coordinates Plot features parallel axes for each variable and each axis is linked by lines.
 Changing the axes may reveal patterns or relationships between variables for categorical variables.
@@
 For JS-based distances, plotting cluster statistics is beneficial for characterization and variable importance could be displayed through parallel coordinate plots.
+-->
 

### 2021-11-16 main.Rmd
caption of table 2

@@
 
-```{r range-parameter, fig.cap="Range of parameters for different scenarios"}
+```{r range-parameter, fig.cap="The range of parameters used for the validation study, for the three different scenarios, number of simulations for each design, differences between means across granularities and series lengths.
+"}
 knitr::include_graphics("figs/simulation_table.png")

### 2021-11-16 main.Rmd
typos in intro

@@
 This paper is organized as follows.
-Section~\ref{sec:methodology} provides the clustering methodology.
-Section~\ref{sec:validation} shows data designs to validate our methods.
-Section~\ref{sec:application} discusses the application of the method to a subset of the real data.
-Finally, we summarize our results and discuss possible future directions in Section~\ref{sec:discussion}.
+Section \ref{sec:methodology} provides the clustering methodology.
+Section \ref{sec:validation} shows data designs to validate our methods.
+Section \ref{sec:application} discusses the application of the method to a subset of the real data.
+Finally, we summarize our results and discuss possible future directions in Section \ref{sec:discussion}.
 
@@
 This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS).
-NQT is used in @@Gupta2021-hakear prior sorting granularities.
+NQT is used in @Gupta2021-hakear prior sorting granularities.
 <!-- $wpd$, which is the foundation of wpd-based distances.

### 2021-11-16 main.Rmd
references fixed

@@
 The time-series features, length of time-series, representation technique, and, of course, the purpose of clustering time-series all influence the suitable similarity measure or distance metric to a meaningful level.
-The three primary methods to time series clustering (@liao2005clustering) are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based).
+The three primary methods to time series clustering [@liao2005clustering] are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based).
 The efficacy of distance-based techniques is highly dependent on the distance measure utilized.
 Defining an appropriate distance measure for the raw time series may be a difficult task since it must take into account noise, variable lengths of time series, asynchronous time series, different scales, and missing data.
-Commonly used distance-based similarity measures as suggested by a decade review of time series clustering approaches (@Aghabozorgi2015-ct) are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping (DTW), Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances (@liao2007clustering) but on a vector time series data.
+Commonly used distance-based similarity measures as suggested by a decade review of time series clustering approaches [@Aghabozorgi2015-ct] are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping (DTW), Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances [@liao2007clustering] but on a vector time series data.
 Among these alternatives, Euclidean distances have high performance but need the same length of data over the same period, resulting in information loss regardless of whether it is on raw data or a smaller collection of features.
-DTW works well with time series of different lengths (@corradini2001dynamic), but it is incapable of handling missing observations.
+DTW works well with time series of different lengths [@corradini2001dynamic], but it is incapable of handling missing observations.
 Surprisingly, probability distributions, which may reflect the inherent temporal structure of a time series have not been considered in determining time series similarity.
@@
 Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior complex.
-The challenge is to be able to cluster consumers into these type of expected patterns, and other unexpected patterns, using only their energy usage history (@Ushakova2020-rl).
+The challenge is to be able to cluster consumers into these type of expected patterns, and other unexpected patterns, using only their energy usage history [@Ushakova2020-rl].
 <!--To safeguard the customers' privacy, it is probable that such information is not accessible.
@@
 $k$-means can be made to perform better by explicitly incorporating time series features such as correlation or cyclic patterns rather than performing it on raw data.
-To reduce dimensionality, several studies use principal component analysis (PCA) or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).
+To reduce dimensionality, several studies use principal component analysis (PCA) or factor analysis to pre-process smart-meter data before clustering [@Ndiaye2011-pf].
 PCA eliminates correlation patterns and decreases feature space, but loses interpretability.
@@
 
-<!-- A model-based clustering works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered (@Motlagh2019-yj).
+<!-- A model-based clustering works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered [@Motlagh2019-yj].
 [@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
@@
 <!-- This work -->
-<!-- This is similar to a stochastic approach (@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time-series characteristics, or a model of the series, to enable unsupervised clustering.
+<!-- This is similar to a stochastic approach [@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time-series characteristics, or a model of the series, to enable unsupervised clustering.
 Unsupervised clustering is only as good as the features that are extracted/selected or the distance metrics that were utilized.
@@
 
-Computationally in R, the data is assumed to be a "tsibble object" (@wang2020tsibble) equipped with an index variable representing inherent ordering from past to present and a key variable that defines observational units over time.
+Computationally in R, the data is assumed to be a "tsibble object" [@wang2020tsibble] equipped with an index variable representing inherent ordering from past to present and a key variable that defines observational units over time.
 The measured variable for an observation is a time-indexed sequence of values.
@@
 This distance metric considers two time series to be similar if the distributions of each category of an individual cyclic granularity or combination of categories for interacting cyclic granularities are similar.
-In this study, the distribution for each category is characterized using deciles (can potentially consider any list of quantiles), and the distances between distributions are calculated using the Jensen-Shannon distances (@Menendez1997-in), which are symmetric and thus could be used as a distance measure.
+In this study, the distribution for each category is characterized using deciles (can potentially consider any list of quantiles), and the distances between distributions are calculated using the Jensen-Shannon distances [@Menendez1997-in], which are symmetric and thus could be used as a distance measure.
 
@@
 
-Compute weighted pairwise distances $wpd$ (@Gupta2021-hakear) for all considered granularities for all observations.
+Compute weighted pairwise distances $wpd$ [@Gupta2021-hakear] for all considered granularities for all observations.
 $wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalized by appropriate parameters.
@@
 All of MDS, PCA and t-SNE use a distance or dissimilarity matrix to construct a reduced-dimension space representation, but their goals are diverse.
-Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
+Multidimensional scaling [@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
 The t-SNE embedding will compress data points that are close in high-dimensional space.
@@
 
-<!-- (a) _Parallel coordinate plots_ (@wegman1990hyperdimensional) are often used to visualize high-dimensional and multivariate data, allowing visual grouping and pattern detection.
+<!-- (a) _Parallel coordinate plots_ [@wegman1990hyperdimensional) are often used to visualize high-dimensional and multivariate data, allowing visual grouping and pattern detection.
 A Parallel Coordinates Plot features parallel axes for each variable.
@@
 
-<!-- (c) _Displaying cluster statistics_ are useful when we have larger problems and it is difficult to read the Parallel coordinate plots due to congestion. (@dasu2005grouping) -->
+<!-- (c) _Displaying cluster statistics_ are useful when we have larger problems and it is difficult to read the Parallel coordinate plots due to congestion. [@dasu2005grouping) -->
 
@@
 PCA seeks to retain data variance.
-Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
+Multidimensional scaling [@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
 t-SNE, on the other hand, is concerned with preserving neighborhood data points.
@@
 
-<!-- This part of the work uses R packages `GGally` (@R-GGally), `Rtsne` (@R-tsne), `ggplot2` (Wickham2009pk), `tour` (@wickham2011tourr), `stats` (@R-language). -->
+<!-- This part of the work uses R packages `GGally` [@R-GGally), `Rtsne` [@R-tsne), `ggplot2` (Wickham2009pk), `tour` [@wickham2011tourr), `stats` [@R-language). -->
 
@@
 
-```{r range-parameter, fig.cap="The range of parameters used for the validation study, for the three different scenarios, number of simulations for each design, differences between means across granularities and series lengths.
-"}
-knitr::include_graphics("figs/simulation_table.png")
+<!-- # ```{r range-parameter, fig.cap="The range of parameters used for the validation study, for the three different scenarios, number of simulations for each design, differences between means across granularities and series lengths.
+"} -->
+<!-- # knitr::include_graphics("figs/simulation_table.png") -->
+<!-- # ``` -->
+
+```{r range-parameter}
+Scenario = c("S1", "S2","S3")
+designs = c(5, 4, 4)
+niter = list (c(25, 20, 500), c(25, 20, 500), c(25, 20, 500))
+mu = list (c(1, 2, 5), c(1, 2, 5), c(1, 2, 5))
+nT = list (c(300, 1000, 5000), c(300, 1000, 5000), c(300, 1000, 5000))
+
+df = tibble::tibble(Scenario, designs, niter, mu, nT) %>%  
+  set_names("Scenario", 
+            " \\# designs", 
+            "niter", 
+            "$\\mu$", 
+            "nT") %>%
+  knitr::kable(format = "latex",
+             escape = FALSE,  booktabs = TRUE,
+             caption = "The range of parameters used for the validation study, for the three different scenarios, number of simulations ($niter$) for each design, differences between means ($\\mu$) across granularities and series lengths ($nT$).")%>%
+  kableExtra::collapse_rows(columns = 3:5)
+
+df
 ```
@@
 Supervised learning uses a training set of known information to categorize new events through instance selection.
-Instance selection (@olvera2010review) is a method of rejecting instances that are not helpful for classification.
+Instance selection [@olvera2010review] is a method of rejecting instances that are not helpful for classification.
 This is analogous to subsampling the population along all dimensions of interest such that the sampled data represents the primary features of the underlying distribution.
@@
 Tours can reveal variable separation that gets hidden in a single variable display better than static projections.
-Hence, we perform a linked tour with a t-SNE layout using the R package `liminal` (@R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
+Hence, we perform a linked tour with a t-SNE layout using the R package `liminal` [@R-liminal] to identify customers who are more likely to have distinct patterns across the variables studied.
 (Refer to the supplementary article for further details).

### 2021-11-16 main.Rmd
start on Application intro

--- 2021-11-16 main.Rmd
@@
 
-_Scenario (S2): Few significant granularities_
+\noindent _Scenario (S2): Few significant granularities_
 
@@
 
-_Scenario (S3): Only one significant granularity_
+\noindent _Scenario (S3): Only one significant granularity_
 
@@
 
-The use of our methodology is illustrated on smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
-It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
+Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
+Teh full data contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
 We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation.
-Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.--> The raw data for these consumers is of unequal length, with varying start and finish dates.
-Because our proposed methods evaluate probability distributions rather than raw data, neither of these data features would pose any threat to our methodology unless they contained any structure or systematic patterns.
-Additionally, there were missing values in the database but further investigation revealed that there is no structure in the missingness (see Supplementary paper for raw data features and missingness).
-
-Clustering huge data sets presents a slew of difficulties.
+Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.--> The raw data for these consumers is of unequal length, with varying starting and end dates.
+Additionally, there were missing values in many series.
+(The supplementary material contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless any systematic structure unrelated to granularities.
+
+Huge data sets presents more complications for clustering.
+Clustering algorithms work well when there well-separated clusters, with no nuisance variables or nuisance observations.
+When converting a series to granularities, many variables (each level of a granularity) are generated, possibly creating a slew of nuisance variables.
+Some households may have a mix of energy use patterns, which could be considered nuisance observations, between major clusters.
+For this reason, we have chosen to select a small group of households with relatively distinct and different patterns in order to illustrate the clustering, simply.
+This is often a good approach to tackling a big analysis task, so start with a simpler task.
+The approach, however, is applicable using on all households, but it would be more complicated to make final choice of the number of clusters and explain the results.
+
 As a result, we dissect the larger problem and test our solutions on a small sample of prototype customers.

### 2021-11-16 main.Rmd
updated range of parameterss table

--- 2021-11-16 main.Rmd
@@
 ```{r range-parameter}
-Scenario = c("S1", "S2","S3")
+scenario = c("S1", "S2","S3")
 designs = c(5, 4, 4)
@@
 
-df = tibble::tibble(Scenario, designs, niter, mu, nT) %>%  
-  set_names("Scenario", 
-            " \\# designs", 
-            "niter", 
-            "$\\mu$", 
-            "nT") %>%
+df = tibble::tibble(scenario, " \\# designs" = designs, niter,  "$\\mu$" = mu, nT) %>%
   knitr::kable(format = "latex",
-             escape = FALSE,  booktabs = TRUE,
+             escape = FALSE,
              caption = "The range of parameters used for the validation study, for the three different scenarios, number of simulations ($niter$) for each design, differences between means ($\\mu$) across granularities and series lengths ($nT$).")%>%

### 2021-11-16 script/mylibrary.R

@@ -30,5 +30,6 @@ library(cluster)
 library(ggpubr)
 library(grid)
 library(ggrepel)
+library(kableExtra)
 
 

### 2021-11-16 main.Rmd
table range parameter update
mylibrary and mytheme changed to library and theme - main

@@
 #knitr::read_chunk('script/main.R')
-knitr::read_chunk('script/mylibrary.R')
-knitr::read_chunk('script/mytheme.R')
+knitr::read_chunk('script/library.R')
+knitr::read_chunk('script/theme.R')
 knitr::read_chunk('script/application.R')
@@
 So three scenarios are considered to accommodate that.
-Figure \ref{fig:range-parameter} shows the range of parameters considered for each scenario.
+Table \ref{tab:range-parameter} shows the range of parameters considered for each scenario.
 

### 2021-11-16 paper-supplementary.Rmd
mylibrary and mytheme changed to library and theme - main

@@
 ```{r external, include = FALSE, cache = FALSE}
-knitr::read_chunk('script/mylibrary.R')
-knitr::read_chunk('script/mytheme.R')
+knitr::read_chunk('script/library.R')
+knitr::read_chunk('script/theme.R')
 knitr::read_chunk('script/validation.R')

### 2021-11-16 main.Rmd
Fig 9 update

@@
 
-df = tibble::tibble(scenario, " \\# designs" = designs, niter,  "$\\mu$" = mu, nT) %>%
+df = tibble::tibble(scenario, " designs" = designs, niter,  "$\\mu$" = mu, nT) %>%
   knitr::kable(format = "latex",
@@
 Figure \ref{fig:groups-4and5} depicts the summarized distributions across $4$ and $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster.
-A-2, Groups A-3, and A-4 have profiles that correspond to B-2, B-3, and B-5, respectively.
-B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-5 (id: 21-24) have a `hod` pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day.
-A-1 is subdivided into B-1 and B-4, each of which has a distinct shape across `moy` and `wnwd`.
+A-1, Groups A-2, and A-3 have profiles that correspond to B-1, B-2, and B-3, respectively.
+B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-1 (id: 21-24) have a `hod` pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day.
+A-4 is subdivided into B-4 and B-5, each of which has a distinct shape across `moy` and `wnwd`.
 B-4 (id: 16-20) is distinguished by higher energy consumption in the first few months of the year and greater variability in weekend usage.
-B-1 (id: 1-3) has higher consumption in the middle of the year (winter months) and a similar weekday-weekend inter-quartile range.
-When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals.
+B-5 (id: 1-3) has higher consumption in the middle of the year (winter months) and a similar weekday-weekend inter-quartile range.
+When $k = 4$ is used, these two groups merge to form A-4, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals.
 It may be worthwhile to compare Figures \ref{fig:opt-cluster-tsne-jsd} and \ref{fig:hod-ind-group-png} to see if the summarized distributions across groups accurately characterized the groupings.
@@
 ```{r groups-4and5, fig.cap = "For $k = 4$ (a) and $k=5$ (b), the distribution of electricity demand across groups over hod, moy, and wnwd is shown.
-Groups A-2, A-3, and A-4 profiles correspond to Groups B-2, B-3, and B-5, respectively.
-A-1 is split into B-1 and B-4, each of which has a distinct shape across moy and wnwd.
+Groups A-1, A-2, and A-3 profiles correspond to Groups B-1, B-2, and B-3, respectively.
+A-4 is split into B-4 and B-5, each of which has a distinct shape across moy and wnwd.
 B-4 (id: 16-20) is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage.
-B-1 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.
-When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group.", fig.height=6}
+B-5 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.
+When $k = 4$ is used, these two groups merge to form A-4, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group.", fig.height=5}
 

### 2021-11-16 main.Rmd
typos till Section 2 fixed

@@
 abstract: |
- Clustering is a potential approach for Organizing large collections of times series into small homogeneous groups, but a difficult step is determining an appropriate metric to measure similarity between time series.
+ Clustering is a potential approach for organizing large collections of times series into small homogeneous groups, but a difficult step is determining an appropriate metric to measure similarity between time series.
 The similarity metric needs to be capable of accommodating long, noisy, and asynchronous time series and also capture cyclical patterns.
@@
 <!-- time series clustering and its challenges -->
-Time-series clustering is the process of unsupervised partitioning of $n$ time-series data into $k$ ($k<n$) meaningful groups such that homogeneous time-series are grouped together based on a certain similarity measure.
-The time-series features, length of time-series, representation technique, and, of course, the purpose of clustering time-series all influence the suitable similarity measure or distance metric to a meaningful level.
+Time series clustering is the process of unsupervised partitioning of $n$ time series data into $k$ ($k<n$) meaningful groups such that homogeneous time series are grouped together based on a certain similarity measure.
+The time series features, length of time series, representation technique, and, of course, the purpose of clustering time series all influence the suitable similarity measure or distance metric to a meaningful level.
 The three primary methods to time series clustering [@liao2005clustering] are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based).
@@
 
-There is a growing body of literature focused on time series clustering related to smart meter data.
+There is an extensive body of literature focused on time series clustering related to smart meter data.
 @Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
@@
 Only one study [@ozawa2016determining] identified time series characteristics by first conducting a Fourier transformation, to convert data from time to frequency domain, followed by $k$-means to cluster by greatest frequency.
-@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered.
+@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered.
 [@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
@@
 <!-- This work -->
-<!-- This is similar to a stochastic approach [@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time-series characteristics, or a model of the series, to enable unsupervised clustering.
+<!-- This is similar to a stochastic approach [@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time series characteristics, or a model of the series, to enable unsupervised clustering.
 Unsupervised clustering is only as good as the features that are extracted/selected or the distance metrics that were utilized.
@@
 
-<!--The foundation of our method is unsupervised clustering algorithms based exclusively on the time-series data.
+<!--The foundation of our method is unsupervised clustering algorithms based exclusively on the time series data.
 The proposed methods leverage the intrinsic data structure hidden within cyclic temporal granularities.--> The existing work on clustering probability distributions assumes we have an independent and identically distributed samples $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
@@
 Granularities that do not have some difference between categories are likely to be nuisance variables.
-It should be noted that all of the time series in a collection may not have the same set of significant granularities.
-This is the approach for managing this generate a subset ($S_c$) of significant granularities across a collection of time series:
+It should be noted that all of the time series in a collection may not have the same set of significant granularities.<!--This is the approach for managing this generate a subset ($S_c$) of significant granularities across a collection of time series:--> This is the approach for generating a subset ($S_c$) of significant granularities across a collection of time series:
 
@@
 Scales of individuals needs to be standardized, because clustering is to select similar patterns, not magnitude of usage.
-(Organizing individuals based on magnitude can achieved simply by sorting on a statistic like the average value across time.) <!--Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
+(Organizing individuals based on magnitude can be achieved simply by sorting on a statistic like the average value across time.) <!--Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
 Statistical transformations are employed to bring all of them to the same range or normalize each observation.
@@
 <!--As a result, determining which raw distribution was used is difficult using the modified distribution.--> A drawback is that any multimodality will be concealed.
-This, this should be checked prior to applying NQT.
+This should be checked prior to applying NQT.
 
@@
 Compute weighted pairwise distances $wpd$ [@Gupta2021-hakear] for all considered granularities for all observations.
-$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalized by appropriate parameters.
+$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction.
+It is estimated by the maximum pairwise distances between distributions across consecutive categories normalized by appropriate parameters.
 A higher value of $wpd$ indicates that some interesting patterns are expected, whereas a lower value would indicate otherwise.

### 2021-11-16 main.Rmd
making moy, hod and wnwd font consistent

@@
 
-df = tibble::tibble(scenario, " designs" = designs, niter,  "$\\mu$" = mu, nT) %>%
+df = tibble::tibble(scenario, designs, niter, mu, nT) %>% rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
+  #tibble::tibble(scenario, " designs" = designs, "R" =niter,  "$\\mu$" = mu, "T"=nT) %>% 
   knitr::kable(format = "latex",
              escape = FALSE,
-             caption = "The range of parameters used for the validation study, for the three different scenarios, number of simulations ($niter$) for each design, differences between means ($\\mu$) across granularities and series lengths ($nT$).")%>%
+             caption = "The range of parameters used for the validation study, for the three different scenarios, number of simulations ($R$) for each design, differences between means ($\\mu$) across granularities and series lengths ($T$).")%>%
   kableExtra::collapse_rows(columns = 3:5)
@@
 Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
-Teh full data contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
+The full data contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
 We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation.
@@
 - Obtain $wpd$ for all cyclic granularities considered for these customers.
-It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weekend/weekday) are turning out to be significant for most customers.
+It was found that $hod$ (hour-of-day), $moy$ (month-of-year) and $wnwd$ (weekend/weekday) are turning out to be significant for most customers.
 We use these three granularities while clustering.
 
-- Remove customers whose data for an entire category of `hod`, `moy` or `wnwd` is empty.
+- Remove customers whose data for an entire category of $hod$, $moy$ or $wnwd$ is empty.
 For example, a customer who does not have data for an entire month is excluded because their monthly behavior cannot be analyzed.
@@
 (Refer to the supplementary article for further details).
-Figure \ref{fig:hod-ind-group-png} shows the distribution across `hod`, `moy` and `wnwd` for the set of chosen $24$ customers that were chosen.
+Figure \ref{fig:hod-ind-group-png} shows the distribution across $hod$, $moy$ and $wnwd$ for the set of chosen $24$ customers that were chosen.
 The customers are split into batches of $12$ in (a) and (b).
 Each row in (a) and that in (b) represents the profile of one customer across different variables.
-<!--Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution.
+<!--Few of these customers have similar distribution across $moy$ and some are similar in their $hod$ distribution.
 -->
@@
 Figure \ref{fig:groups-4and5} depicts the summarized distributions across $4$ and $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster.
-A-1, Groups A-2, and A-3 have profiles that correspond to B-1, B-2, and B-3, respectively.
-B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-1 (id: 21-24) have a `hod` pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day.
-A-4 is subdivided into B-4 and B-5, each of which has a distinct shape across `moy` and `wnwd`.
+Groups sA-1, A-2, and A-3 have profiles that correspond to B-1, B-2, and B-3 respectively.
+B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-1 (id: 21-24) have a $hod$ pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day.
+A-4 is subdivided into B-4 and B-5, each of which has a distinct shape across $moy$ and $wnwd$.
 B-4 (id: 16-20) is distinguished by higher energy consumption in the first few months of the year and greater variability in weekend usage.
@@
 
-<!-- Our methodology is useful for grouping similar distributions over `hod` and `moy` and they are placed closely for easy comparison.
-Few groups have mixed patterns across `hod` and `moy`, but few have all customers in the group having a similar profile.
+<!-- Our methodology is useful for grouping similar distributions over $hod$ and $moy$ and they are placed closely for easy comparison.
+Few groups have mixed patterns across $hod$ and $moy$, but few have all customers in the group having a similar profile.
 -->
@@
 The two customers (id: 16 and 18) that have significant differences between $wnwd$ categories (in Figure \ref{fig:hod-ind-group-png}) stand out from the rest of the customers.
-Group-4 has a higher $wpd$ for `hod` than `moy` or `wkndwday`.
-Group-2 and Group-3 have the most distinct patterns across `moy`.
+Group-4 has a higher $wpd$ for $hod$ than $moy$ or $wnwd$.
+Group-2 and Group-3 have the most distinct patterns across $moy$.
 Group-1 is a mixed group that has strong patterns on at least one of the three variables.
-When $k=3$ is used, Group-2 and Group-3 are merged because their patterns relative to `moy`, `hod` and `wnwd` are the same.
+When $k=3$ is used, Group-2 and Group-3 are merged because their patterns relative to $moy$, $hod$ and $wnwd$ are the same.
 The findings vary from JS-based clustering, yet it is a helpful grouping.

### 2021-11-17 main.Rmd
2.1 typos

--- 2021-11-16 main.Rmd
@@
 
-[@Gupta2021-hakear] provides a method for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
+@Gupta2021-hakear provides a method for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
 (This extends to harmonies, pairs of granularities that might interact with each other.) We define "significant" granularities as those with significant distributional differences across at least one category.
@@
 It should be noted that all of the time series in a collection may not have the same set of significant granularities.
-This is the approach for managing this generate a subset ($S_c$) of significant granularities across a collection of time series:
+This is the approach for managing this to generate a subset ($S_c$) of significant granularities across a collection of time series:
 

### 2021-11-17 main.Rmd
number of clusters language updated

--- 2021-11-16 main.Rmd
@@
 
-These can all give contradictory suggestions when the data does not naturally break into groups, or in the presence of nuisance variables (no contribution to clustering) or nuisance observations (inlying outlying observations falling between clusters).
-There is no one best metric, perhaps a reason why so many exist.
+All of the common approaches can give contradictory suggestions for the optimal number of clusters, particularly when the data does not naturally break into groups, or in the presence of nuisance variables (no contribution to clustering) or nuisance observations (inlying outlying observations falling between clusters).
+There is no one best metric, which is perhaps a reason why metrics so many exist.
 
@@
 It is a very simple, but effective metric.
-This is computed by averaging the minimum 10% of distances between observations between clusters, making it relatively robust to nuisance observations.
-The value of $sindex$ always decreases, and where values drop substantially when an extra cluster are candidates for optimal clustering.
-The number of clusters corresponding to the value before the drop is the recommendation.
+This is computed by averaging the smallest 10% of inter-cluster distances.
+It is relatively robust to nuisance observations.
+The value of $sindex$ always decreases, and sharp drops in value indicate candidates for the optimal number of clusters.
+The number of clusters corresponding to the value **before the drop** is the recommendation.
 

### 2021-11-17 main.Rmd
validation language

@@
 
-To validate our clustering methods, we spiked many attributes in the data to create different data designs.
-<!--to see where one method works better than the other and where they might give us the same outcome --or the effect of missing data <!---and trends on the proposed methods.--> Three circular granularities $g1$, $g2$ and $g3$ are considered with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $n_{g_1}=2$, $n_{g_2}=3$ and $n_{g_3}=5$.
+To validate our clustering methods, have created several different data designs containing different granularity features.
+<!--to see where one method works better than the other and where they might give us the same outcome --or the effect of missing data <!---and trends on the proposed methods.--> There are three circular granularities $g1$, $g2$ and $g3$ with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $n_{g_1}=2$, $n_{g_2}=3$ and $n_{g_3}=5$.
 These categories could be integers or some more meaningful labels.

### 2021-11-17 main.Rmd
section 3.3 visual exploration

@@
 The formed clusters have to match the design, be well separated, and have minimal intra-cluster variation.
-It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values.
-MDS and parallel coordinate graphs are used to demonstrate the findings, as well as an index value plot to provide direction on the number of clusters.
+<!--It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values.--> MDS and parallel coordinate graphs are used to demonstrate the findings, as well as an index value plot to provide direction on the number of clusters.
 JS-based approaches corresponding to NQT and RS are referred to as JS-NQT and JS-RS respectively.
-In the following plots, results for JS-NQT are reported.
-For similar results with JS-RS or wpd-based distances, please refer to the supplementary paper.
-
-- Figure \ref{fig:sindex-plot-validation} shows $sindex$ plotted against the number of clusters ($k$) for the range of mean differences (rows) under the different scenarios (columns).
+In the following plots, results for JS-NQT are reported, and results with JS-RS or wpd-based distances, which are similar are in the  supplementary material.
+
+Figure \ref{fig:sindex-plot-validation} shows $sindex$ plotted against the number of clusters ($k$) for the range of mean differences (rows) under the different scenarios (columns).
 This can be used to determine the number of clusters for each scenario.
@@
 
-- Figure \ref{fig:mds-plot-validation} shows separation of our clusters.
+Figure \ref{fig:mds-plot-validation} shows separation of our clusters.
 It can be observed that in all scenarios and for different mean differences, clusters are separated.
@@
 
-- Figure \ref{fig:parcoord-sim} depicts a parallel coordinate plot with the vertical bar showing total inter-cluster distances with regard to granularities $g1$, $g2$, and $g3$ for all simulation settings and scenarios.
+Figure \ref{fig:parcoord-sim} depicts a parallel coordinate plot with the vertical bar showing total inter-cluster distances with regard to granularities $g1$, $g2$, and $g3$ for all simulation settings and scenarios.
 So one line in the figure shows the inter-cluster distances for one simulation setting and scenarios vary across facets.
@@
 
- - The JS-RS and wpd-based methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
+The JS-RS and wpd-based methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
 However, a complete year of data is the minimum requirement to capture distributional differences in winter and summer profiles, for example.
@@
 
-- In our study sample, the method JS-NQT outperforms the method JS-RS for smaller differences between categories. More testing, however, is required to corroborate this.
+In our study sample, the method JS-NQT outperforms the method JS-RS for smaller differences between categories. More testing, however, would be needed to to be confident in this conclusion.
 

### 2021-11-17 main.Rmd
updated pdf, table 1 headers need to fix

@@
 
-df = tibble::tibble(scenario, designs, niter, mu, nT) %>% rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
+df = tibble::tibble(scenario, designs, niter, mu, nT) %>% #rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
   #tibble::tibble(scenario, " designs" = designs, "R" =niter,  "$\\mu$" = mu, "T"=nT) %>% 

### 2021-11-17 main.Rmd
fig 4 caption

--- 2021-11-17 main.Rmd
@@
 
-```{r sindex-plot-validation, fig.cap = "The cluster separation index (sindex) is plotted as a function of the number of clusters for the range of mean differences (rows) under the different scenarios (columns).
-S1 has a sharp decrease in sindex from 5 to 6, whereas S2 and S3 have a decrease from 4 to 5.
-As a result, the number of clusters considered for S1, S2, and S3 is 5, 4, 4, respectively.
-This corresponds to the number of designs taken into account in each scenario."}
+```{r sindex-plot-validation, fig.cap = "Choosing optimal cluster number across the range of scenarios and mean differences used in the validation study, using the cluster separation index (sindex) for the JS-NQT.
+S1 has a sharp decrease in sindex from 5 to 6, whereas S2 and S3 have a decrease from 4 to 5, especially when mean difference is large, providing the recommended number of clusters to be 5, 4, 4, respectively.
+This precisely reflects the structure in designs that we would hope the clustering could recover."}
 ```
@@
 
-Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
+Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO [@make-these-two-URLs-into-proper-citations].
 The full data contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
@@
 Additionally, there were missing values in many series.
-(The supplementary material contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless any systematic structure unrelated to granularities.
-
-Huge data sets presents more complications for clustering.
-Clustering algorithms work well when there well-separated clusters, with no nuisance variables or nuisance observations.
+(The supplementary material contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless there is any systematic structure is related to granularities.
+
+Huge data sets present more complications for clustering.
+Clustering algorithms work well when there are well-separated clusters, with no nuisance variables or nuisance observations.
 When converting a series to granularities, many variables (each level of a granularity) are generated, possibly creating a slew of nuisance variables.
-Some households may have a mix of energy use patterns, which could be considered nuisance observations, between major clusters.
-For this reason, we have chosen to select a small group of households with relatively distinct and different patterns in order to illustrate the clustering, simply.
-This is often a good approach to tackling a big analysis task, so start with a simpler task.
-The approach, however, is applicable using on all households, but it would be more complicated to make final choice of the number of clusters and explain the results.
+Some households may have a mix of energy use patterns, which could be considered nuisance observations located between major clusters.
+For this reason, we have chosen to select a small group of households with relatively distinct and different patterns in order to illustrate the clustering more simply.
+This is often a good approach to tackling a big analysis task, to start with a simpler task.
+The approach, however, is applicable to all households.
 
 As a result, we dissect the larger problem and test our solutions on a small sample of prototype customers.
-To do this, data is first filtered to generate a clean sample, and then significant cyclic granularities (variables) for them are chosen.
-This is described in Section \ref{sec:datafilter}.
-The clean set is subsequently subsampled along all dimensions of interest, to ensure that the sampled data reveals some patterns across at least one specified variable, which is described in Section \ref{sec:prototype}.
-By grouping the prototypes using our methods in Section \ref{sec:clustering} and assessing their meaning, the study hopes to unravel some of the heterogeneities observed in energy usage data.
-Because our application does not employ additional customer data, we cannot explain why consumption varies, but rather try to identify how it varies.
+To do this, data is first filtered to generate a small sample, and then significant cyclic granularities (variables) for them are chosen (as described in Section \ref{sec:datafilter}).
+The sample set is subsequently examined along all dimensions of interest, to ensure that they reveal some patterns across at least one specified variable (as described in Section \ref{sec:prototype}).
+<!--By grouping the prototypes using our methods in Section \ref{sec:clustering} and assessing their meaning, the study hopes to unravel some of the heterogeneities observed in energy usage data.--> Because the data does not contain additional customer characteristics, we cannot explain why consumption varies, but can only identify how it varies.
 
@@
 
-- Choose a smaller subset of randomly selected $600$ customers with no implicit missing values for 2013.
-
-- Obtain $wpd$ for all cyclic granularities considered for these customers.
+The steps for household filtering and variable selection were:
+
+1. Choose a smaller subset of randomly selected $600$ customers with no implicit missing values for 2013.
+
+2.
+Obtain $wpd$ for all cyclic granularities considered for these customers.
 It was found that $hod$ (hour-of-day), $moy$ (month-of-year) and $wnwd$ (weekend/weekday) are turning out to be significant for most customers.
@@
 
-- Remove customers whose data for an entire category of $hod$, $moy$ or $wnwd$ is empty.
+3.
+Remove customers whose data for an entire category of $hod$, $moy$ or $wnwd$ is empty.
 For example, a customer who does not have data for an entire month is excluded because their monthly behavior cannot be analyzed.
 
-- Remove customers whose energy consumption is 0 in all deciles.
+4.
+Remove customers whose energy consumption is 0 in all deciles.
 These are the clients whose consumption is likely to remain essentially flat and with no intriguing repeated patterns that we are interested in studying.
 
-## Prototype selection {#sec:prototype}
+## Choosing households {#sec:prototype}
 
 <!-- Why instance selection -->
+Figure \ref{fig:hod-ind-group-png} shows the distribution across $hod$, $moy$ and $wnwd$ for the set of $24$ customers is the sample to illustrate clustering.
+The customers are displayed in two columns of $12$ for space reasons.
+Each row of each column represents the profile of a single customer across different variables.
+
 Supervised learning uses a training set of known information to categorize new events through instance selection.
@@
 (Refer to the supplementary article for further details).
-Figure \ref{fig:hod-ind-group-png} shows the distribution across $hod$, $moy$ and $wnwd$ for the set of chosen $24$ customers that were chosen.
-The customers are split into batches of $12$ in (a) and (b).
-Each row in (a) and that in (b) represents the profile of one customer across different variables.
 <!--Few of these customers have similar distribution across $moy$ and some are similar in their $hod$ distribution.

### 2021-11-17 main.Rmd
all references fixed

--- 2021-11-17 main.Rmd
@@
 
-\noindent This work is motivated by a need to cluster a large collection of residential smart meter data, so that households can be grouped into similar energy usage patterns.
+\noindent This work is motivated by a need to cluster a large collection of residential smart meter data, so that customers can be grouped into similar energy usage patterns.
 These can be considered to be univariate time series of continuous values which are available at fine temporal scales.
@@
 Half-hourly measurements of usage for more than 13,000 household electricity smart meters is provided from from October 2011 to March 2014.
-<!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
+<!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Customers vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
 The behavioral patterns differ amongst customers due to many temporal dependencies.
-Some households use a dryer, while others dry their clothes on a line.
+Some customers use a dryer, while others dry their clothes on a line.
 Their weekly usage profile may reflect this.
@@
 If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
-We aim to have pockets of households showing similar periodic behavior by considering $wpd$ values for different harmonies and some time series features.
+We aim to have pockets of customers showing similar periodic behavior by considering $wpd$ values for different harmonies and some time series features.
 The features should also characterize probability distributions of different household.
@@
 This can be done numerically, by tabulating cluster means and standard deviations [@dasu2005grouping], and visually using methods for graphics multivariate data.
-@Cook2007-qe provides visual examples using both tours [@asimov] and parallel coordinate plots [@wegman1990hyperdimensional].
-Dimension reduction techniques like principal component analysis [@joliffe], multidimensional scaling (MDS) [@borg2005modern], t-distributed stochastic neighbor embedding (t-SNE) [@maarten-tsne] and linear discrimant analysis (LDA) [@lda] are also useful.
+@Cook2007-qe provides visual examples using both tours [@asimov1985grand] and parallel coordinate plots [@wegman1990hyperdimensional].
+Dimension reduction techniques like principal component analysis [@jolliffe2016principal], multidimensional scaling (MDS) [@borg2005modern], t-distributed stochastic neighbor embedding (t-SNE) [@van2008visualizing] and linear discrimant analysis (LDA) [@fisher1936use] are also useful.
 <!-- A Parallel Coordinates Plot features parallel axes for each variable and each axis is linked by lines.
@@
 <!-- 1. Compute quantiles of distributions across each category of the cyclic granularity -->
-<!-- 2. Compute JS distance between households for each each category of the cyclic granularity -->
-<!-- 3. Total distance between households computed as sum of JS distances for all hours -->
+<!-- 2. Compute JS distance between customers for each each category of the cyclic granularity -->
+<!-- 3. Total distance between customers computed as sum of JS distances for all hours -->
 <!-- 4. Cluster using this distance with hierarchical clustering algorithm (method "Ward.D") -->
@@
 
-<!-- # ```{r range-parameter, fig.cap="The range of parameters used for the validation study, for the three different scenarios, number of simulations for each design, differences between means across granularities and series lengths.
+<!-- ```{r range-parameter, fig.cap="The range of parameters used for the validation study, for the three different scenarios, number of simulations for each design, differences between means across granularities and series lengths.
 "} -->
-<!-- # knitr::include_graphics("figs/simulation_table.png") -->
-<!-- # ``` -->
+<!-- knitr::include_graphics("figs/simulation_table.png") -->
+<!-- ``` -->
 
@@
 designs = c(5, 4, 4)
-niter = list (c(25, 20, 500), c(25, 20, 500), c(25, 20, 500))
-mu = list (c(1, 2, 5), c(1, 2, 5), c(1, 2, 5))
-nT = list (c(300, 1000, 5000), c(300, 1000, 5000), c(300, 1000, 5000))
+niter = list (c(25, 20, 500),
+              c(25, 20, 500),
+              c(25, 20, 500))
+mu = list (c(1, 2, 5),
+           c(1, 2, 5),
+           c(1, 2, 5))
+nT = list (c(300, 1000, 5000),
+           c(300, 1000, 5000),
+           c(300, 1000, 5000))
 
@@
 When converting a series to granularities, many variables (each level of a granularity) are generated, possibly creating a slew of nuisance variables.
-Some households may have a mix of energy use patterns, which could be considered nuisance observations, between major clusters.
-For this reason, we have chosen to select a small group of households with relatively distinct and different patterns in order to illustrate the clustering, simply.
+Some customers may have a mix of energy use patterns, which could be considered nuisance observations, between major clusters.
+For this reason, we have chosen to select a small group of customers with relatively distinct and different patterns in order to illustrate the clustering, simply.
 This is often a good approach to tackling a big analysis task, so start with a simpler task.
-The approach, however, is applicable using on all households, but it would be more complicated to make final choice of the number of clusters and explain the results.
+The approach, however, is applicable using on all customers, but it would be more complicated to make final choice of the number of clusters and explain the results.
 
@@
 Few are similar in their hod pattern, while others are similar in moy behavior.
-Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most households, there is one household for whom weekend behavior is standing out from the rest."} -->
+Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most customers, there is one household for whom weekend behavior is standing out from the rest."} -->
 <!-- # ``` -->

### 2021-11-17 main.Rmd

--- 2021-11-17 main.Rmd
@@
 The primary application of this work is data from the Smart Grid, Smart City (SGSC) project (2010–2014) available through the [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
-Half-hourly measurements of usage for more than 13,000 household electricity smart meters is provided from from October 2011 to March 2014.
-<!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
+Half-hourly measurements of usage for more than 13,000 electricity smart meter customers is provided from from October 2011 to March 2014.
+<!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Customers vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
 The behavioral patterns differ amongst customers due to many temporal dependencies.
-Some households use a dryer, while others dry their clothes on a line.
+Some customers use a dryer, while others dry their clothes on a line.
 Their weekly usage profile may reflect this.
@@
 designs = c(5, 4, 4)
-niter = list (c(25, 20, 500), c(25, 20, 500), c(25, 20, 500))
-mu = list (c(1, 2, 5), c(1, 2, 5), c(1, 2, 5))
-nT = list (c(300, 1000, 5000), c(300, 1000, 5000), c(300, 1000, 5000))
-
-df = tibble::tibble(scenario, designs, niter, mu, nT) %>% rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
+niter = list (c(25, 20, 500),
+              c(25, 20, 500),
+              c(25, 20, 500))
+mu = list (c(1, 2, 5),
+           c(1, 2, 5),
+           c(1, 2, 5))
+nT = list (c(300, 1000, 5000),
+           c(300, 1000, 5000),
+           c(300, 1000, 5000))
+
+tibble::tibble(scenario, designs, niter, mu, nT) %>% #rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
   #tibble::tibble(scenario, " designs" = designs, "R" =niter,  "$\\mu$" = mu, "T"=nT) %>% 
@@
   kableExtra::collapse_rows(columns = 3:5)
-
-df
 ```
@@
 When converting a series to granularities, many variables (each level of a granularity) are generated, possibly creating a slew of nuisance variables.
-Some households may have a mix of energy use patterns, which could be considered nuisance observations located between major clusters.
-For this reason, we have chosen to select a small group of households with relatively distinct and different patterns in order to illustrate the clustering more simply.
+Some customers may have a mix of energy use patterns, which could be considered nuisance observations located between major clusters.
+For this reason, we have chosen to select a small group of customers with relatively distinct and different patterns in order to illustrate the clustering more simply.
 This is often a good approach to tackling a big analysis task, to start with a simpler task.
-The approach, however, is applicable to all households.
+The approach, however, is applicable to all customers.
 
@@
 
-The steps for household filtering and variable selection were:
+The steps for customer filtering and variable selection were:
 
@@
 
-## Choosing households {#sec:prototype}
+## Choosing customers {#sec:prototype}
 

### 2021-11-17 main.Rmd
added standardization of wpd

@@
 
-@Gupta2021-hakear provides a method for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
+@Gupta2021-hakear provides a distance measure ($wpd$) for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
 (This extends to harmonies, pairs of granularities that might interact with each other.) We define "significant" granularities as those with significant distributional differences across at least one category.
@@
 Statistical transformations are employed to bring all of them to the same range or normalize each observation.
-This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS).
-NQT is used in @Gupta2021-hakear prior sorting granularities.
-<!-- $wpd$, which is the foundation of wpd-based distances.
-XX wpd was not yet introduced-->
+This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS).<!-- NQT is used in @Gupta2021-hakear prior sorting granularities.-- $wpd$, which is the foundation of wpd-based distances.
+XX wpd was not yet introduced--> While @Gupta2021-hakear already uses NQT when computing $wpd$, it could be useful to standardize it for the selected set of significant granularities prior to computing the distances.
 
@@
 
-tibble::tibble(scenario, designs, niter, mu, nT) %>% #rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
+tibble::tibble(scenario, designs, niter, mu, nT) %>% rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
   #tibble::tibble(scenario, " designs" = designs, "R" =niter,  "$\\mu$" = mu, "T"=nT) %>% 
@@
              escape = FALSE,
-             caption = "The range of parameters used for the validation study, for the three different scenarios, number of simulations ($R$) for each design, differences between means ($\\mu$) across granularities and series lengths ($T$).")%>%
-  kableExtra::collapse_rows(columns = 3:5)
+             caption = "The range of parameters used for the validation study, for the three different scenarios, number of simulations ($R$) for each design, differences between means ($\\mu$) across granularities and series lengths ($T$).")
+#%>%  kableExtra::collapse_rows(columns = 3:5)
 ```

### 2021-11-17 main.Rmd
working on section 4.2

--- 2021-11-17 main.Rmd
@@
 
-\noindent This work is motivated by a need to cluster a large collection of residential smart meter data, so that customers can be grouped into similar energy usage patterns.
+\noindent This work is motivated by a need to cluster a large collection of residential smart meter data, so that households can be grouped into similar energy usage patterns.
 These can be considered to be univariate time series of continuous values which are available at fine temporal scales.
@@
 The primary application of this work is data from the Smart Grid, Smart City (SGSC) project (2010–2014) available through the [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
-Half-hourly measurements of usage for more than 13,000 electricity smart meter customers is provided from from October 2011 to March 2014.
-<!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Customers vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
+Half-hourly measurements of usage for more than 13,000 household electricity smart meters is provided from from October 2011 to March 2014.
+<!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
 The behavioral patterns differ amongst customers due to many temporal dependencies.
-Some customers use a dryer, while others dry their clothes on a line.
+Some households use a dryer, while others dry their clothes on a line.
 Their weekly usage profile may reflect this.
@@
 If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
-We aim to have pockets of customers showing similar periodic behavior by considering $wpd$ values for different harmonies and some time series features.
+We aim to have pockets of households showing similar periodic behavior by considering $wpd$ values for different harmonies and some time series features.
 The features should also characterize probability distributions of different household.
@@
 This can be done numerically, by tabulating cluster means and standard deviations [@dasu2005grouping], and visually using methods for graphics multivariate data.
-@Cook2007-qe provides visual examples using both tours [@asimov1985grand] and parallel coordinate plots [@wegman1990hyperdimensional].
-Dimension reduction techniques like principal component analysis [@jolliffe2016principal], multidimensional scaling (MDS) [@borg2005modern], t-distributed stochastic neighbor embedding (t-SNE) [@van2008visualizing] and linear discrimant analysis (LDA) [@fisher1936use] are also useful.
+@Cook2007-qe provides visual examples using both tours [@asimov] and parallel coordinate plots [@wegman1990hyperdimensional].
+Dimension reduction techniques like principal component analysis [@joliffe], multidimensional scaling (MDS) [@borg2005modern], t-distributed stochastic neighbor embedding (t-SNE) [@maarten-tsne] and linear discrimant analysis (LDA) [@lda] are also useful.
 <!-- A Parallel Coordinates Plot features parallel axes for each variable and each axis is linked by lines.
@@
 <!-- 1. Compute quantiles of distributions across each category of the cyclic granularity -->
-<!-- 2. Compute JS distance between customers for each each category of the cyclic granularity -->
-<!-- 3. Total distance between customers computed as sum of JS distances for all hours -->
+<!-- 2. Compute JS distance between households for each each category of the cyclic granularity -->
+<!-- 3. Total distance between households computed as sum of JS distances for all hours -->
 <!-- 4. Cluster using this distance with hierarchical clustering algorithm (method "Ward.D") -->
@@
 
-<!-- ```{r range-parameter, fig.cap="The range of parameters used for the validation study, for the three different scenarios, number of simulations for each design, differences between means across granularities and series lengths.
+<!-- # ```{r range-parameter, fig.cap="The range of parameters used for the validation study, for the three different scenarios, number of simulations for each design, differences between means across granularities and series lengths.
 "} -->
-<!-- knitr::include_graphics("figs/simulation_table.png") -->
-<!-- ``` -->
+<!-- # knitr::include_graphics("figs/simulation_table.png") -->
+<!-- # ``` -->
 
@@
 designs = c(5, 4, 4)
-niter = list (c(25, 20, 500),
-              c(25, 20, 500),
-              c(25, 20, 500))
-mu = list (c(1, 2, 5),
-           c(1, 2, 5),
-           c(1, 2, 5))
-nT = list (c(300, 1000, 5000),
-           c(300, 1000, 5000),
-           c(300, 1000, 5000))
-
-tibble::tibble(scenario, designs, niter, mu, nT) %>% #rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
+niter = list (c(25, 20, 500), c(25, 20, 500), c(25, 20, 500))
+mu = list (c(1, 2, 5), c(1, 2, 5), c(1, 2, 5))
+nT = list (c(300, 1000, 5000), c(300, 1000, 5000), c(300, 1000, 5000))
+
+df = tibble::tibble(scenario, designs, niter, mu, nT) %>% #rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
   #tibble::tibble(scenario, " designs" = designs, "R" =niter,  "$\\mu$" = mu, "T"=nT) %>% 
@@
   kableExtra::collapse_rows(columns = 3:5)
+
+df
 ```
@@
 When converting a series to granularities, many variables (each level of a granularity) are generated, possibly creating a slew of nuisance variables.
-Some customers may have a mix of energy use patterns, which could be considered nuisance observations located between major clusters.
-For this reason, we have chosen to select a small group of customers with relatively distinct and different patterns in order to illustrate the clustering more simply.
+Some households may have a mix of energy use patterns, which could be considered nuisance observations located between major clusters.
+For this reason, we have chosen to select a small group of households with relatively distinct and different patterns in order to illustrate the clustering more simply.
 This is often a good approach to tackling a big analysis task, to start with a simpler task.
-The approach, however, is applicable to all customers.
+The approach, however, is applicable to all households.
 
@@
 
-The steps for customer filtering and variable selection were:
+The steps for household filtering and variable selection were:
 
 1. Choose a smaller subset of randomly selected $600$ customers with no implicit missing values for 2013.
-
 2.
@@
 We use these three granularities while clustering.
-
 3.
@@
 For example, a customer who does not have data for an entire month is excluded because their monthly behavior cannot be analyzed.
-
 4.
@@
 
-## Choosing customers {#sec:prototype}
+## Choosing households {#sec:prototype}
 
@@
 
-Supervised learning uses a training set of known information to categorize new events through instance selection.
-Instance selection [@olvera2010review] is a method of rejecting instances that are not helpful for classification.
-This is analogous to subsampling the population along all dimensions of interest such that the sampled data represents the primary features of the underlying distribution.
+It is common to pre-process data prior to fitting a supervised learning model using instance selection [@olvera2010review] which removes observations that might impede the model building.
+For clustering, there has been little attention given, but it would be analogous to identifying and removing nuisance observations.
 Instance selection in unsupervised learning has received little attention in the literature.
@@
 Few are similar in their hod pattern, while others are similar in moy behavior.
-Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most customers, there is one household for whom weekend behavior is standing out from the rest."} -->
+Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most households, there is one household for whom weekend behavior is standing out from the rest."} -->
 <!-- # ``` -->

### 2021-11-17 main.Rmd

--- 2021-11-17 main.Rmd
@@
 1. Choose a smaller subset of randomly selected $600$ customers with no implicit missing values for 2013.
-
 2.

### 2021-11-17 main.Rmd
prototype selection

@@
 
-## Choosing customers {#sec:prototype}
+## Selecting prototypes {#sec:prototype}
 
@@
 It is common to pre-process data prior to fitting a supervised learning model using instance selection [@olvera2010review] which removes observations that might impede the model building.
-For clustering, there has been little attention given, but it would be analogous to identifying and removing nuisance observations.
-Instance selection in unsupervised learning has received little attention in the literature.
-However, it could be a useful tool for evaluating model or method performance.
-There are several ways to approach prototype selection.
-Following the idea of @Fan2021-bq of picking related examples (neighbors) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2D space.
-Then pick a few "anchor" customers far apart in 2D space and pick a few neighbors for each.
-Unfortunately, this does not ensure that consumers with significant patterns across all variables are chosen.
-Tours can reveal variable separation that gets hidden in a single variable display better than static projections.
-Hence, we perform a linked tour with a t-SNE layout using the R package `liminal` [@R-liminal] to identify customers who are more likely to have distinct patterns across the variables studied.
-(Refer to the supplementary article for further details).
-<!--Few of these customers have similar distribution across $moy$ and some are similar in their $hod$ distribution.
+For clustering, this is analogous to identifying and removing nuisance observations.
+Prototype selection is more severe than instance selection, because a handful of cases is selected.
+@cutler proposed a method called archetypal analysis which has inspired this approach but the procedure we have used follows @Fan2021-bq.
+First  dimension reduction such as t-SNE, MDS or PCA is used to project the data into a 2D space.
+Second a few  "anchor" customers far apart in 2D space are selected.
+Additional close neighbors to the anchors are selected.
+To check the selections relative to the full set of variables, we used a tour linked to a t-SNE layout using the R package `liminal` [@R-liminal].
+This ensured that the final sample of clustered customers were also far apart in the high-dimensional space.
+(See the supplementary materials for further details.) <!--Few of these customers have similar distribution across $moy$ and some are similar in their $hod$ distribution.
 -->
@@
 
-## Clustering {#sec:clustering}
+## Clustering results {#sec:clustering}
 

### 2021-11-17 main.Rmd
added more reference for instance selection

--- 2021-11-17 main.Rmd
@@
   \setlength {\marginparwidth }{2cm}
-  \usepackage{mathtools,amssymb,booktabs,longtable,todonotes,amsthm}
+    \usepackage{mathtools,amssymb,booktabs,longtable,todonotes,amsthm,multirow,array}
   \def\mod{~\text{mod}~}
@@
              escape = FALSE,
-             caption = "The range of parameters used for the validation study, for the three different scenarios, number of simulations ($R$) for each design, differences between means ($\\mu$) across granularities and series lengths ($T$).")
-#%>%  kableExtra::collapse_rows(columns = 3:5)
+             caption = "The range of parameters used for the validation study, for the three different scenarios, number of simulations ($R$) for each design, differences between means ($\\mu$) across granularities and series lengths ($T$).")%>%  kableExtra::collapse_rows(columns = 3:5)
 ```
@@
 
-It is common to pre-process data prior to fitting a supervised learning model using instance selection [@olvera2010review] which removes observations that might impede the model building.
+It is common to pre-process data prior to fitting a supervised learning model using instance selection [@olvera2010review; @Cutler1994-hm] which removes observations that might impede the model building.
 For clustering, there has been little attention given, but it would be analogous to identifying and removing nuisance observations.

### 2021-11-17 main.Rmd

--- 2021-11-17 main.Rmd
@@
 Prototype selection is more severe than instance selection, because a handful of cases is selected.
-@cutler proposed a method called archetypal analysis which has inspired this approach but the procedure we have used follows @Fan2021-bq.
+@Cutler1994-hm proposed a method called archetypal analysis which has inspired this approach but the procedure we have used follows @Fan2021-bq.
 First  dimension reduction such as t-SNE, MDS or PCA is used to project the data into a 2D space.
@@
 This ensured that the final sample of clustered customers were also far apart in the high-dimensional space.
-(See the supplementary materials for further details.) <!--Few of these customers have similar distribution across $moy$ and some are similar in their $hod$ distribution.
+(See the supplementary materials for further details.)<!--Few of these customers have similar distribution across $moy$ and some are similar in their $hod$ distribution.
 -->

### 2021-11-17 main.Rmd
adding smart meter data and Hennig paper reference

@@
 Most metrics for choosing the optimal number of clusters, are based on comparing distances between observations within a class to those distances between observations between classes, which makes the assumption that there are some separated clusters.
-Some common procedures include the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy] and the separation index ($sindex$) [@Hennig2014-ah].
+Some common procedures include the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy] and the separation index ($sindex$) [@Hennig2019-eq; @Hennig2014-ah].
 These are constructed by balancing within-cluster homogeneity and between-cluster separation.
@@
 
-Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO [@make-these-two-URLs-into-proper-citations].
+Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from @smart-meter.
 The full data contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.

### 2021-11-17 main.Rmd
scenario section format

@@
 
-_Scenario (S1): All granularities significant_
-
+**Scenario 1 (S1) - All granularities significant:**
 <!-- Consider a case where all the three granularities $g1$, $g2$ and $g3$ would be responsible for making the designs distinct.
@@
 -->
-
 Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction.
@@
 
-\noindent _Scenario (S2): Few significant granularities_
-
+\noindent **Scenario 2 (S2) - Few significant granularities:**
 This is the case where one granularity will remain the same across all designs.
@@
 
-\noindent _Scenario (S3): Only one significant granularity_
-
+\noindent **Scenario 3 (S3) - Only one significant granularity:**
 Only one granularity is responsible for identifying the designs in this case. This is depicted in Figure \ref{fig:gran2and1-clubbed} (right) where only $g3$ affects the designs significantly.

### 2021-11-17 main.Rmd
ordering of figures for application

@@
 <!-- Why instance selection -->
-Figure \ref{fig:hod-ind-group-png} shows the distribution across $hod$, $moy$ and $wnwd$ for the set of $24$ customers is the sample to illustrate clustering.
+Figure \ref{fig:hod-ind-group-png} shows the distribution across $hod$, $moy$ and $wnwd$ for the set of $24$ customers used to illustrate clustering.
 The customers are displayed in two columns of $12$ for space reasons.
-Each row of each column represents the profile of a single customer across different variables.
+Each row, of each column, represents the profile of a single customer across different variables.
 
@@
 
-The $24$ prototypes <!--and the data subset consisting of $353$ customers--> are clustered using the methodology described in Section \ref{sec:methodology} and results are reported below.
+Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:hod-ind-group-png} and \ref{fig:parcoord-application}.
 In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups.
@@
 
-```{r opt-cluster-tsne-jsd, fig.cap = "(a) shows t-SNE summary of the $24$ selected customers chosen through prototype selection, (b) and (c) shows a plot of sindex for different cluster size for clustering based on JS-based distances and wpd-distances respectively.
-The plot suggests optimal number of clusters as $4$ and $7$ respectively."}
-#include_graphics("figs/opt_clust_parenthesis.png")
-```
-
 ```{r hod-ind-group}
@@
 
-```{r hod-ind-group-png, fig.cap = "The distribution of electricity demand across individual customers over multiple granularities hod, moy, and wnwd are shown for the $24$ selected customers using quantile and box plots.
+```{r hod-ind-group-png, fig.cap = "The distribution of electricity demand across individual customers over three granularities $hod$, $moy$, and $wnwd$ are shown for the $24$ selected customers using quantile and box plots.
 They are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer.
@@
 Distributional differences across categories of wnwd are not discernable except for a couple of customers (id: 16 and 18)."}
+```
+
+```{r opt-cluster-tsne-jsd, fig.cap = "(a) shows t-SNE summary of the $24$ selected customers chosen through prototype selection, (b) and (c) shows a plot of sindex for different cluster size for clustering based on JS-based distances and wpd-distances respectively.
+The plot suggests optimal number of clusters as $4$ and $7$ respectively."}
+#include_graphics("figs/opt_clust_parenthesis.png")
 ```

### 2021-11-17 main.Rmd
language in application section

@@
 
-Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:hod-ind-group-png} and \ref{fig:parcoord-application}.
+Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, and \ref{fig:parcoord-application}.
 In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups.

### 2021-11-17 main.Rmd
fig 7 caption

@@
 
-```{r opt-cluster-tsne-jsd, fig.cap = "(a) shows t-SNE summary of the $24$ selected customers chosen through prototype selection, (b) and (c) shows a plot of sindex for different cluster size for clustering based on JS-based distances and wpd-distances respectively.
-The plot suggests optimal number of clusters as $4$ and $7$ respectively."}
+```{r opt-cluster-tsne-jsd, fig.cap = "Clustering summaries: t-SNE computed on the $24$ selected customers (a), and separation index ($sindex$) for 2-10 clusters using JS (b) and WPD (c).
+Various choices in number of clusters would be recommended.
+Four clusters are visible in t-SNE, although it might hide a fifth cluster because dimension reduction to 2D may be insufficient to see the difference.
+JS suggests 3 and wpd suggests 3 or 6."}
 #include_graphics("figs/opt_clust_parenthesis.png")

### 2021-11-17 main.Rmd
cluster summaries

@@
 
-Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, and \ref{fig:parcoord-application}.
-In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups.
+Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:combined-groups-js} and \ref{fig:parcoord-application}.
+(XXX move the next sentences down to where you talk about the clustering results for JS) In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups.
 --> Groups are colored differently whenever they represent different groupings.
@@
 
-```{r combined-groups-js, fig.cap = "The distribution of electricity demand across groups over hod (a), moy (b), and hod (conditional on wnwd) (c) are shown through quantile plots.
-Group 1 (1-3), Group 2 (4-9), Group 3 (10 - 15), Group 4 (16- 20), Group 5 (21 - 24).
+```{r combined-groups-js, fig.cap = "The distribution of electricity demand across clusters over $hod$ (a), $moy$ (b), and $wnwd$ (c) are shown through quantile plots.
+Group 1 (customers 1-3), Group 2 (customers 4-9), Group 3 (customers 10-15), Group 4 (customers 16-20), Group 5 (customers 21-24).
 Groups 2 and 5 appear to have a hod pattern among its members (morning and evening peak), whereas groups 1, 3, and 5 appear to have a moy pattern (higher usage in mid-year).
@@
 
-```{r groups-4and5, fig.cap = "For $k = 4$ (a) and $k=5$ (b), the distribution of electricity demand across groups over hod, moy, and wnwd is shown.
-Groups A-1, A-2, and A-3 profiles correspond to Groups B-1, B-2, and B-3, respectively.
-A-4 is split into B-4 and B-5, each of which has a distinct shape across moy and wnwd.
-B-4 (id: 16-20) is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage.
+```{r groups-4and5, fig.cap = "Summary plots for four and five clusters.
+The distribution of electricity demand combined for all members over $hod$, $moy$, and $wnwd$ is shown.
+Groups A-1 (customers XXX), A-2 (customers XXX), and A-3 (customers XXX) profiles correspond to Groups B-1, B-2, and B-3, respectively.
+Cluster A-4 splits into B-4 (customers XXX) and B-5 (customers XXX)to produce the five clusters, whci h better resolves the $moy$ distribution.
+B-4 is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage.
 B-5 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.

### 2021-11-17 main.Rmd
Fig 8 update with prototype labels

--- 2021-11-17 main.Rmd
@@
 
-**Scenario 1 (S1) - All granularities significant:**
+_Scenario (S1): All granularities significant_
+
 <!-- Consider a case where all the three granularities $g1$, $g2$ and $g3$ would be responsible for making the designs distinct.
@@
 -->
+
 Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction.
@@
 
-\noindent **Scenario 2 (S2) - Few significant granularities:**
+\noindent _Scenario (S2): Few significant granularities_
+
 This is the case where one granularity will remain the same across all designs.
@@
 
-\noindent **Scenario 3 (S3) - Only one significant granularity:**
+\noindent _Scenario (S3): Only one significant granularity_
+
 Only one granularity is responsible for identifying the designs in this case. This is depicted in Figure \ref{fig:gran2and1-clubbed} (right) where only $g3$ affects the designs significantly.
@@
 <!-- Why instance selection -->
-Figure \ref{fig:hod-ind-group-png} shows the distribution across $hod$, $moy$ and $wnwd$ for the set of $24$ customers used to illustrate clustering.
+Figure \ref{fig:hod-ind-group-png} shows the distribution across $hod$, $moy$ and $wnwd$ for the set of $24$ customers is the sample to illustrate clustering.
 The customers are displayed in two columns of $12$ for space reasons.
-Each row, of each column, represents the profile of a single customer across different variables.
+Each row of each column represents the profile of a single customer across different variables.
 
@@
 
-Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, and \ref{fig:parcoord-application}.
+The $24$ prototypes <!--and the data subset consisting of $353$ customers--> are clustered using the methodology described in Section \ref{sec:methodology} and results are reported below.
 In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups.
@@
 
+```{r opt-cluster-tsne-jsd, fig.cap = "(a) shows t-SNE summary of the $24$ selected customers chosen through prototype selection, (b) and (c) shows a plot of sindex for different cluster size for clustering based on JS-based distances and wpd-distances respectively.
+The plot suggests optimal number of clusters as $4$ and $7$ respectively."}
+#include_graphics("figs/opt_clust_parenthesis.png")
+```
+
 ```{r hod-ind-group}
@@
 
-```{r hod-ind-group-png, fig.cap = "The distribution of electricity demand across individual customers over three granularities $hod$, $moy$, and $wnwd$ are shown for the $24$ selected customers using quantile and box plots.
+```{r hod-ind-group-png, fig.cap = "The distribution of electricity demand across individual customers over multiple granularities hod, moy, and wnwd are shown for the $24$ selected customers using quantile and box plots.
 They are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer.
@@
 A few of these customers have similar distributions across moy and some are similar in their hod distributions.
-Distributional differences across categories of wnwd are not discernable except for a couple of customers (id: 16 and 18)."}
-```
-
-```{r opt-cluster-tsne-jsd, fig.cap = "(a) shows t-SNE summary of the $24$ selected customers chosen through prototype selection, (b) and (c) shows a plot of sindex for different cluster size for clustering based on JS-based distances and wpd-distances respectively.
-The plot suggests optimal number of clusters as $4$ and $7$ respectively."}
-#include_graphics("figs/opt_clust_parenthesis.png")
+Distributional differences across categories of wnwd are not discernable except for a couple of customers (id: 16 and 18).", fig.height=6}
 ```

### 2021-11-17 main.Rmd

--- 2021-11-17 main.Rmd
@@
 Each customer is identified by their profile across all the granularities.
-A few of these customers have similar distributions across moy and some are similar in their hod distributions.
-Distributional differences across categories of wnwd are not discernable except for a couple of customers (id: 16 and 18)."}
+A few of these customers have similar distributions across $moy$ and some are similar in their $hod$ distributions.
+Distributional differences across categories of $wnwd$ are not discernable except for a couple of customers (id: 16 and 18)."}
 ```

### 2021-11-17 main.Rmd
Fig 7 and 8 updated

@@
 Most metrics for choosing the optimal number of clusters, are based on comparing distances between observations within a class to those distances between observations between classes, which makes the assumption that there are some separated clusters.
-Some common procedures include the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy] and the separation index ($sindex$) [@Hennig2019-eq; @Hennig2014-ah].
+Some common procedures include the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@Dunn1973-gk] and the separation index ($sindex$) [@Hennig2019-eq; @Hennig2014-ah].
 These are constructed by balancing within-cluster homogeneity and between-cluster separation.
@@
 
+```{r all-data-rs}
+
+```
+
+```{r clustering-rs}
+
+```
+
+```{r opt-clusters-jsd-rs}
+
+```
+
+```{r opt-clusters-all}
+
+```
+
 ```{r hod-ind-group}
@@
 
-```{r opt-cluster-tsne-jsd, fig.cap = "Clustering summaries: t-SNE computed on the $24$ selected customers (a), and separation index ($sindex$) for 2-10 clusters using JS (b) and WPD (c).
+```{r opt-cluster-tsne-jsd, fig.cap = "Clustering summaries: t-SNE computed on the $24$ selected customers (a), and separation index ($sindex$) for $2$-$10$ clusters using JS-NQT, JS-RS and WPD in (b).
 Various choices in number of clusters would be recommended.
 Four clusters are visible in t-SNE, although it might hide a fifth cluster because dimension reduction to 2D may be insufficient to see the difference.
-JS suggests 3 and wpd suggests 3 or 6."}
+JS-NQT suggests $3$, JS-RS suggests $3$ or $5$ and WPD suggests $3$ or $6$."}
 #include_graphics("figs/opt_clust_parenthesis.png")

### 2021-11-17 main.Rmd
XXX for Fig 9 addressed

@@
 Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:combined-groups-js} and \ref{fig:parcoord-application}.
-(XXX move the next sentences down to where you talk about the clustering results for JS) In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups.
---> Groups are colored differently whenever they represent different groupings.
+Groups are colored differently whenever they represent different groupings.
 While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude.
@@
 Figure \ref{fig:groups-4and5} depicts the summarized distributions across $4$ and $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster.
-Groups sA-1, A-2, and A-3 have profiles that correspond to B-1, B-2, and B-3 respectively.
+In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.
+Groups A-1, A-2, and A-3 have profiles that correspond to B-1, B-2, and B-3 respectively.
 B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-1 (id: 21-24) have a $hod$ pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day.
@@
 The distribution of electricity demand combined for all members over $hod$, $moy$, and $wnwd$ is shown.
-Groups A-1 (customers XXX), A-2 (customers XXX), and A-3 (customers XXX) profiles correspond to Groups B-1, B-2, and B-3, respectively.
-Cluster A-4 splits into B-4 (customers XXX) and B-5 (customers XXX)to produce the five clusters, whci h better resolves the $moy$ distribution.
+Groups A-1 (customers 21-24), A-2 (customers 4-9), and A-3 (customers 10-15) profiles correspond to Groups B-1, B-2, and B-3, respectively.
+Cluster A-4 splits into B-4 (customers 16-20) and B-5 (customers 1-3) to produce the five clusters, which better resolves the $moy$ distribution.
 B-4 is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage.
-B-5 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.
+B-5 is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.
 When $k = 4$ is used, these two groups merge to form A-4, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group.", fig.height=5}

### 2021-11-17 main.Rmd
consistent italisation of hod, ...

@@
 <!-- # ```{r hod-moy-wkndwday plots, fig.cap="The distribution across moy, hod and wkndwday for the selected designs.
-Few are similar in their hod pattern, while others are similar in moy behavior.
+Few are similar in their hod pattern, while others are similar in $moy$ behavior.
 Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most customers, there is one household for whom weekend behavior is standing out from the rest."} -->
@@
 B-5 (id: 1-3) has higher consumption in the middle of the year (winter months) and a similar weekday-weekend inter-quartile range.
-When $k = 4$ is used, these two groups merge to form A-4, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals.
+When $k = 4$ is used, these two groups merge to form A-4, which has a $moy$ profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals.
 It may be worthwhile to compare Figures \ref{fig:opt-cluster-tsne-jsd} and \ref{fig:hod-ind-group-png} to see if the summarized distributions across groups accurately characterized the groupings.
@@
 Group 1 (customers 1-3), Group 2 (customers 4-9), Group 3 (customers 10-15), Group 4 (customers 16-20), Group 5 (customers 21-24).
-Groups 2 and 5 appear to have a hod pattern among its members (morning and evening peak), whereas groups 1, 3, and 5 appear to have a moy pattern (higher usage in mid-year).
-With their conditional dependencies on hod or moy categories, wnwd differences may not be discernible on their own.
-(c) demonstrates that the distribution of hod conditional on wnwd is also not different, except for the tails.
-This might be because most of these chosen customers do not have a wnwd pattern.
+Groups 2 and 5 appear to have a $hod$ pattern among its members (morning and evening peak), whereas groups 1, 3, and 5 appear to have a $moy$ pattern (higher usage in mid-year).
+With their conditional dependencies on $hod$ or $moy$ categories,  $wnwd$ differences may not be discernible on their own.
+(c) demonstrates that the distribution of $hod$ conditional on  $wnwd$ is also not different, except for the tails.
+This might be because most of these chosen customers do not have a  $wnwd$ pattern.
 It is useful to compare the raw distributions with the summarized distributions to establish that the majority of individuals in the group share the same characteristics.
@@
 B-5 is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.
-When $k = 4$ is used, these two groups merge to form A-4, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group.", fig.height=5}
+When $k = 4$ is used, these two groups merge to form A-4, which has a $moy$ profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group.", fig.height=5}
 
@@
 Plot (b) which shows median $wpd$ values for each group.
-The plot shows that moy is the most important variable in identifying clusters, whereas wnwd is the least significant and has the least fluctuation.
-Two customers (id: 16 and 18) with high $wpd$ across wnwd stand out in this display.
-Group-4 has a higher $wpd$ for hod than moy or wnwd.
-Group-2 and Group-3 have the most discernible pattern across moy.
+The plot shows that $moy$ is the most important variable in identifying clusters, whereas  $wnwd$ is the least significant and has the least fluctuation.
+Two customers (id: 16 and 18) with high $wpd$ across  $wnwd$ stand out in this display.
+Group-4 has a higher $wpd$ for $hod$ than $moy$ or  $wnwd$.
+Group-2 and Group-3 have the most discernible pattern across $moy$.
 Group-1 is a mixed group with strong patterns across at least one of the three variables.

### 2021-11-17 main.Rmd
caption for fig 9

@@
 
-```{r groups-4and5, fig.cap = "Summary plots for four and five clusters.
-The distribution of electricity demand combined for all members over $hod$, $moy$, and $wnwd$ is shown.
+```{r groups-4and5, fig.cap = "Summary plots for four and five clusters showing the distribution of electricity demand combined for all members over $hod$, $moy$, and $wnwd$.
 Groups A-1 (customers 21-24), A-2 (customers 4-9), and A-3 (customers 10-15) profiles correspond to Groups B-1, B-2, and B-3, respectively.

### 2021-11-17 main.Rmd
fig 10 solution

@@
 
-```{r groups-4and5, fig.cap = "Summary plots for four and five clusters showing the distribution of electricity demand combined for all members over $hod$, $moy$, and $wnwd$.
+```{r groups-4and5, fig.cap = "Summary plots for four and five clusters from JS-NQT showing the distribution of electricity demand combined for all members over $hod$, $moy$, and $wnwd$.
 Groups A-1 (customers 21-24), A-2 (customers 4-9), and A-3 (customers 10-15) profiles correspond to Groups B-1, B-2, and B-3, respectively.
-Cluster A-4 splits into B-4 (customers 16-20) and B-5 (customers 1-3) to produce the five clusters, which better resolves the $moy$ distribution.
-B-4 is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage.
-B-5 is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.
-When $k = 4$ is used, these two groups merge to form A-4, which has a $moy$ profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group.", fig.height=5}
+Cluster A-4 splits into B-4 (customers 16-20) and B-5 (customers 1-3) to produce the five clusters, which better resolves the $moy$ distribution.", fig.height=5}
 
@@
 
-```{r parcoord-application, fig.cap="Each of the $24$ customers is represented by a parallel coordinate plot (a) with four wpd-based groupings.
-Plot (b) which shows median $wpd$ values for each group.
-The plot shows that $moy$ is the most important variable in identifying clusters, whereas  $wnwd$ is the least significant and has the least fluctuation.
-Two customers (id: 16 and 18) with high $wpd$ across  $wnwd$ stand out in this display.
-Group-4 has a higher $wpd$ for $hod$ than $moy$ or  $wnwd$.
-Group-2 and Group-3 have the most discernible pattern across $moy$.
-Group-1 is a mixed group with strong patterns across at least one of the three variables.
-All of these could be observed from the plot or the table (b).", message=FALSE, warning=FALSE, fig.height=6}
+```{r parcoord-application, fig.cap="Summary of the four cluster results from WPD, shown as a parallel coordinate plot and table.
+XXX TO BE REVISED WHEN DIFFERENT CLUSTER SOLUTION IS SHOWN", message=FALSE, warning=FALSE, fig.height=6}
 #include_graphics("figs/parcoord-application.png")

### 2021-11-17 main.Rmd
figure ref fixes

@@
 
-Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:combined-groups-js} and \ref{fig:parcoord-application}.
+Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:groups-4and5} and \ref{fig:parcoord-application}.
 Groups are colored differently whenever they represent different groupings.

### 2021-11-17 main.Rmd
start of 4.3

@@
 Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:groups-4and5} and \ref{fig:parcoord-application}.
-Groups are colored differently whenever they represent different groupings.
-While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude.
+The t-SNE visualization suggests that there are four well-separated clusters.
+It is possible that because the representation is only 2D, that the fifth group from the original prototype selection is distinctly different in high dimensions.
+The $sindex$ plots for the three methods indicate some disagreement: JS-NQT suggests 3, JS-RS suggests 3 or 5 and WPD suggests 3 or 6.
+JS-RS would appear to match the original prototypes with the five cluster solution, but it actually differs.
+Even though the $sindex$ for JS-NQT suggests three clusters, the five cluster solution more closely matches the original prototypes.
+The WPD clustering provides a different grouping of the customers, and even though it disagrees with the original prototypes it is a useful grouping.
+<!--While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude.
 The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales.
-The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.
+The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.-->
 
@@
 
-The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}.
+Figure \ref{fig:opt-cluster-tsne-jsd} The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}.
 The $sindex$ falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases.
@@
 
-```{r opt-cluster-tsne-jsd, fig.cap = "Clustering summaries: t-SNE computed on the $24$ selected customers (a), and separation index ($sindex$) for $2$-$10$ clusters using JS-NQT, JS-RS and WPD in (b).
+```{r opt-cluster-tsne-jsd, fig.cap = "Clustering summaries:  (a) t-SNE computed on the $24$ selected customers, and  (b) separation index ($sindex$) for $2$-$10$ clusters using JS-NQT, JS-RS and WPD.
 Various choices in number of clusters would be recommended.

### 2021-11-17 main.Rmd
JS distances in application

@@
 
-### JS-based distances
-
-Figure \ref{fig:opt-cluster-tsne-jsd} The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}.
+<!--### JS-NQT and JS-RS-->
+
+<!--Figure \ref{fig:opt-cluster-tsne-jsd} The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}.
 The $sindex$ falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases.
-In our study, we use $k=4, 5$.
-Figure \ref{fig:groups-4and5} depicts the summarized distributions across $4$ and $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster.
-In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.
-Groups A-1, A-2, and A-3 have profiles that correspond to B-1, B-2, and B-3 respectively.
-B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-1 (id: 21-24) have a $hod$ pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day.
-A-4 is subdivided into B-4 and B-5, each of which has a distinct shape across $moy$ and $wnwd$.
+In our study, we use $k=4, 5$.--> Figure \ref{fig:groups-4and5} displays the summarized distributions across $4$ and $5$ clusters in (a) and (b) respectively, and helps to characterize each cluster.
+In the quantile plots the line represents the median, and the region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.
+The only difference between the four and five cluster solution is that A-4 divides further into B-4 and B-5.
+This additional division makes a clearer clustering, because it resolves the heterogeneity in $moy$ creating a group (B-5, customers 1-3) which has a winter peak in usage, and a group (B-4, customers 16-20) which has a start of the year peak in usage.
+B-2 (customers 4-9) and B-1 (customers 21-24) have distinctive $hod$ patterns but are both heterogeneous in $moy$ and $wnwd$.
+B-3 (customers 10-15) has peak usage at the end of the year, but is heterogeneous on $hod$ and $wdwn$.
+This clustering almost agrees with the clusters visible in the t-SNE plot.
+
+<!-- A-4 is subdivided into B-4 and B-5, each of which has a distinct shape across $moy$ and $wnwd$.
 B-4 (id: 16-20) is distinguished by higher energy consumption in the first few months of the year and greater variability in weekend usage.
@@
 It may be worthwhile to compare Figures \ref{fig:opt-cluster-tsne-jsd} and \ref{fig:hod-ind-group-png} to see if the summarized distributions across groups accurately characterized the groupings.
-If it has, then the majority of the members of the group should have a similar profile.
+If it has, then the majority of the members of the group should have a similar profile.-->
 
@@
 
-### wpd-based distances
+<!--### WPD-->
 

### 2021-11-17 main.Rmd
fig 7 caption

@@
 They are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer.
-Each customer is identified by their profile across all the granularities.
+The number indicates a unique customer id and a prototype id.
+<!--Each customer is identified by their profile across all the granularities.
 A few of these customers have similar distributions across $moy$ and some are similar in their $hod$ distributions.
-Distributional differences across categories of $wnwd$ are not discernable except for a couple of customers (id: 16 and 18)."}
+Distributional differences across categories of $wnwd$ are not discernable except for a couple of customers (id: 16 and 18).-->"}
 ```

### 2021-11-17 main.Rmd
updated pdf

@@
 
-In Figure \ref{fig:opt-cluster-tsne-jsd}, the $sindex$ is plotted against the number of clusters for clustering based on wpd-based distances.
+<!-- In Figure \ref{fig:opt-cluster-tsne-jsd}, the $sindex$ is plotted against the number of clusters for clustering based on wpd-based distances.
 The scales are very different from those used in JS-based clustering.
 The $sindex$ falls from around $20$ when $k=3$ to $19$ when $k=4$ and further decreases to $17.5$ with $k>7$.
-Thus, there are few choices available for the number of clusters.
-The groupings with $k=4$ is shown through a parallel coordinate plot with the three significant cyclic granularities.
+Thus, there are few choices available for the number of clusters.--> Figure \ref{fig:parcoord-application} whos the wpd values for the three different granularities of the 24 customers, coloured by the 6 cluster solution.
+XXX REWORK THIS WHEN NEW CLUSTER RESULTS ARE SHOWN<!--The groupings with $k=4$ is shown through a parallel coordinate plot with the three significant cyclic granularities.
 The variables are sorted according to their separation across classes (rather than their overall variation between classes).
@@
 When $k=3$ is used, Group-2 and Group-3 are merged because their patterns relative to $moy$, $hod$ and $wnwd$ are the same.
-The findings vary from JS-based clustering, yet it is a helpful grouping.
+The findings vary from JS-based clustering, yet it is a helpful grouping.-->
 
@@
 ```
+
+In summary, none of the methods captured the five original prototypes exactly.
 

### 2021-11-18 main.Rmd
last paragraph of application

@@
 In summary, none of the methods captured the five original prototypes exactly.
-
-Things become far more complicated when we consider a larger data set with more uncertainty, as they do with any clustering problem.
+JS-NQT was almost identical, though, but WPD produced quite a different grouping.
+This is quite a reasonable result and illustrates both the difficulties of clustering to obtain a particularly expected grouping and the ability to learn unexpected patterns in the data.
+It is possible that the JS-based distances were distracted by the presence of nuisance variables, levels of the granularities, that do not contribute to clustering.
+This would also be supported by the results of the validation study, where clustering was less effective in S2 and S3, where only some granularities had differences between levels.
+Clustering using WPD is expected to produce quite different results because it will group only by overall value of a granularity, not a particular pattern.
+A cluster summarization like Figure \ref{fig:groups-4and5} is not possible because there may be different but equally interesting patterns (e.g.
+high evening $hod$ and high day time $hod$) in the same cluster.
+Simply it provides information that across a collection of customers this specific cluster has interesting patterns in a granularity (eg $hod$).
+One would need to post-process the cluster to separate specific patterns.
+XXX Might still need some work to summarise the application section
+
+<!--Things become far more complicated when we consider a larger data set with more uncertainty, as they do with any clustering problem.
 Summarizing distributions across clusters with varied or outlying customers can result in a shape that does not represent the group.
@@
 If more extensive prototype selection is used, resulting in more comprehensive prototypes in the data set, this method might be used to classify the entire data set into these prototype behaviors.
-However, the goal of this section was to have a few customers that have significant patterns over one or more cyclic granularities, apply our methodology to cluster them, and demonstrate that the method produces useful clusters.
+However, the goal of this section was to have a few customers that have significant patterns over one or more cyclic granularities, apply our methodology to cluster them, and demonstrate that the method produces useful clusters.-->
 

### 2021-11-18 main.Rmd
wpd solution summary plot update - still need to reduce font

@@
 
+```{r opt_clusters-wpd}
+```
+
+```{r summary-group-wpd}
+```
+
+```{r summary-table-wpd}
+
+```
+
+```{r summary-plot-wpd, fig.cap = "fill this up", fig.height=6}
+```
+
 ```{r tsne-fit}
@@
 Clustering using WPD is expected to produce quite different results because it will group only by overall value of a granularity, not a particular pattern.
-A cluster summarization like Figure \ref{fig:groups-4and5} is not possible because there may be different but equally interesting patterns (e.g.
+A cluster summary like Figure \ref{fig:groups-4and5} is not possible because there may be different but equally interesting patterns (e.g.
 high evening $hod$ and high day time $hod$) in the same cluster.

### 2021-11-18 main.Rmd
something is a bit strange with the scenario 2 and 3 bold text, might still need more fixing

@@
 
-\noindent **Scenario 2 (S2) - Few significant granularities:**
+**Scenario 2 (S2) - Few significant granularities:**
 This is the case where one granularity will remain the same across all designs.
@@
 
-\noindent **Scenario 3 (S3) - Only one significant granularity:**
+**Scenario 3 (S3) - Only one significant granularity:**
 Only one granularity is responsible for identifying the designs in this case. This is depicted in Figure \ref{fig:gran2and1-clubbed} (right) where only $g3$ affects the designs significantly.

### 2021-11-18 main.Rmd
removed comment from fig 8 caption

@@
 They are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer.
-The number indicates a unique customer id and a prototype id.
-<!--Each customer is identified by their profile across all the granularities.
-A few of these customers have similar distributions across $moy$ and some are similar in their $hod$ distributions.
-Distributional differences across categories of $wnwd$ are not discernable except for a couple of customers (id: 16 and 18).-->"}
+The number indicates a unique customer id and a prototype id."}
 ```

### 2021-11-18 main.Rmd
fix position of WPD summary plot

--- 2021-11-18 main.Rmd
@@
 
-```{r summary-plot-wpd, fig.cap = "fill this up", fig.height=6}
-```
-
 ```{r tsne-fit}
@@
 
+```
+
+```{r summary-plot-wpd, fig.cap = "Summary plots for three (a) and five (b) clusters from WPD showing the $wpd$ values of each customers across $hod$, $moy$, and $wnwd$ through a parallel coordinate plot.
+Group P-1 and P-2 correspond to Q-1 and Q-2 respectively.
+Cluster P-3 splits into Q-3, Q-4 and Q-5.
+", fig.height=6}
 ```

### 2021-11-18 main.Rmd
added id in table
minor corrections

--- 2021-11-18 main.Rmd
@@
 @Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered.
-[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
+@chicco2010renyi addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
 @Melnykov2013-sp discusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
@@
 
-Compute weighted pairwise distances $wpd$ [@Gupta2021-hakear] for all considered granularities for all observations.
+We compute weighted pairwise distances $wpd$ [@Gupta2021-hakear] for all considered granularities for all observations.
 $wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction.
@@
 All of the common approaches can give contradictory suggestions for the optimal number of clusters, particularly when the data does not naturally break into groups, or in the presence of nuisance variables (no contribution to clustering) or nuisance observations (inlying outlying observations falling between clusters).
-There is no one best metric, which is perhaps a reason why metrics so many exist.
+There is no one best metric, which is perhaps a reason why so many metrics exist.
 
@@
 The value of $sindex$ always decreases, and sharp drops in value indicate candidates for the optimal number of clusters.
-The number of clusters corresponding to the value **before the drop** is the recommendation.
+The recommended number of clusters corresponding to the value *before the drop*.
 
@@
 
-To validate our clustering methods, have created several different data designs containing different granularity features.
+To validate our clustering methods, we have created several different data designs containing different granularity features.
 <!--to see where one method works better than the other and where they might give us the same outcome --or the effect of missing data <!---and trends on the proposed methods.--> There are three circular granularities $g1$, $g2$ and $g3$ with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $n_{g_1}=2$, $n_{g_2}=3$ and $n_{g_3}=5$.
@@
 All three granularities $g1$, $g2$, and $g3$ have high inter-cluster distances for S1, suggesting all are important.
-In S2 g1 and in S3 both g1 and g2 have smaller inter-cluster distances, indicating that they did not contribute to clustering."}
+In S2 $g1$ and in S3 both $g1$ and $g2$ have smaller inter-cluster distances, indicating that they did not contribute to clustering."}
 
@@
 Obtain $wpd$ for all cyclic granularities considered for these customers.
-It was found that $hod$ (hour-of-day), $moy$ (month-of-year) and $wnwd$ (weekend/weekday) are turning out to be significant for most customers.
+It was found that $hod$ (hour-of-day), $moy$ (month-of-year) and $wnwd$ (weekend/weekday) are significant for most customers.
 We use these three granularities while clustering.
@@
 
-```{r opt_clusters-wpd}
-```
-
-```{r summary-group-wpd}
-```
-
-```{r summary-table-wpd}
-
-```
-
 ```{r tsne-fit}
@@
 
-```
-
-```{r summary-plot-wpd, fig.cap = "Summary plots for three (a) and five (b) clusters from WPD showing the $wpd$ values of each customers across $hod$, $moy$, and $wnwd$ through a parallel coordinate plot.
-Group P-1 and P-2 correspond to Q-1 and Q-2 respectively.
-Cluster P-3 splits into Q-3, Q-4 and Q-5.
-", fig.height=6}
 ```
@@
 The Github repository, [github.com/Sayani07/paper-gracsr](https://github.com/Sayani07/paper-gracsr), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-This article was created with R [@R-language], `knitr` [@knitr, @R-knitr] and `rmarkdown` [@rmarkdown, @R-rmarkdown].
+This article was created with R [@R-language], `knitr` [@knitr; @R-knitr] and `rmarkdown` [@rmarkdown; @R-rmarkdown].
 Graphics are produced with `ggplot2` [@Wickham2009pk] and `GGally` [@R-GGally].

### 2021-11-18 main.Rmd
changed column name and WPD suggestion

--- 2021-11-18 main.Rmd
@@
 It is possible that because the representation is only 2D, that the fifth group from the original prototype selection is distinctly different in high dimensions.
-The $sindex$ plots for the three methods indicate some disagreement: JS-NQT suggests 3, JS-RS suggests 3 or 5 and WPD suggests 3 or 6.
+The $sindex$ plots for the three methods indicate some disagreement: JS-NQT suggests $3$, JS-RS suggests $2$ or $5$ and WPD suggests $3$ or $5$.
 JS-RS would appear to match the original prototypes with the five cluster solution, but it actually differs.
@@
 B-2 (customers 4-9) and B-1 (customers 21-24) have distinctive $hod$ patterns but are both heterogeneous in $moy$ and $wnwd$.
-B-3 (customers 10-15) has peak usage at the end of the year, but is heterogeneous on $hod$ and $wdwn$.
+B-3 (customers 10-15) has peak usage at the end of the year, but is heterogeneous on $hod$ and $wnwd$.
 This clustering almost agrees with the clusters visible in the t-SNE plot.
@@
 
-```{r summary-table-wpd}
-
-```
-
 ```{r tsne-fit}
@@
 Four clusters are visible in t-SNE, although it might hide a fifth cluster because dimension reduction to 2D may be insufficient to see the difference.
-JS-NQT suggests $3$, JS-RS suggests $3$ or $5$ and WPD suggests $3$ or $6$."}
+JS-NQT suggests $3$, JS-RS suggests $2$ or $5$ and WPD suggests $3$ or $5$."}
 #include_graphics("figs/opt_clust_parenthesis.png")
@@
 ", fig.height=6}
+```
+
+```{r summary-table-wpd}
+
 ```

### 2021-11-18 script/application.R

@@ -823,7 +823,7 @@ q5 <- data_pcp_id %>% filter(group5 =="Q-5") %>% pull(sort_group_id) %>% paste(c
 
 data_table <- bind_rows(data_table3,
                         data_table5) %>%
-  bind_cols(id = c(p1, p2, p3, q1, q2, q3, q4, q5)) %>% 
+  bind_cols(`customer-prototype id` = c(p1, p2, p3, q1, q2, q3, q4, q5)) %>% 
   select(k, group, everything()) %>% 
   bind_cols()
 
@@ -950,8 +950,9 @@ tsne_df <- data.frame(
 
 ## ----tsne-xy------------------------------------------------------------------
 tsne_xy <- ggplot(tsne_df, aes(x = tsneX, y = tsneY)) +
-  #geom_point(size = 2, shape = ".") +
-  ggrepel::geom_text_repel(aes(label = sort_group_id), size = 2, seed = 2935, max.overlaps = 10) +
+  geom_point(size = 2, shape = ".") +
+geom_text(aes(label = sort_group_id), size = 2)+
+#, seed = 2935, max.overlaps = 10) +
   # scale_color_manual(values = limn_pal_tableau10()) +
   scale_colour_viridis_d(direction = -1) +
   # guides(color = FALSE) +

### 2021-11-18 main.Rmd
caption wpd parallel coordinate plot

@@
 Cluster P-3 splits into Q-3, Q-4 and Q-5.
-", fig.height=6}
+P-1 (customers 16, 18) is characterized by high values of $wpd$ on $wnwd$.
+P-2 has lower $wpd$ for $hod$ than $moy$ and $wnwd$.
+P-3 behaves exactly opposite to P-2 with higher $wpd$ for $hod$ compared to $moy$ and $wnwd$.
+For $5$ cluster solution, this groups gets split into Q3-Q5 characterizing different relative significance over $moy$ and $wnwd$.", fig.height=6}
 ```

### 2021-11-18 main.Rmd
fixing plot reference

@@
 
-Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:groups-4and5} and \ref{fig:parcoord-application}.
+Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:groups-4and5} and \ref{fig:summary-plot-wpd}.
 The t-SNE visualization suggests that there are four well-separated clusters.
@@
 The $sindex$ falls from around $20$ when $k=3$ to $19$ when $k=4$ and further decreases to $17.5$ with $k>7$.
-Thus, there are few choices available for the number of clusters.--> Figure \ref{fig:parcoord-application} whos the wpd values for the three different granularities of the 24 customers, coloured by the 6 cluster solution.
+Thus, there are few choices available for the number of clusters.--> Figure \ref{fig:summary-plot-wpd} whos the wpd values for the three different granularities of the 24 customers, coloured by the 6 cluster solution.
 XXX REWORK THIS WHEN NEW CLUSTER RESULTS ARE SHOWN<!--The groupings with $k=4$ is shown through a parallel coordinate plot with the three significant cyclic granularities.
@@
 
-```{r parcoord-application, fig.cap="Summary of the four cluster results from WPD, shown as a parallel coordinate plot and table.
-XXX TO BE REVISED WHEN DIFFERENT CLUSTER SOLUTION IS SHOWN", message=FALSE, warning=FALSE, fig.height=6}
-#include_graphics("figs/parcoord-application.png")
-```
+<!-- # ```{r summary-plot-wpd, fig.cap="Summary of the four cluster results from WPD, shown as a parallel coordinate plot and table.
+XXX TO BE REVISED WHEN DIFFERENT CLUSTER SOLUTION IS SHOWN", message=FALSE, warning=FALSE, fig.height=6} -->
+<!-- # #include_graphics("figs/summary-plot-wpd.png") -->
+<!-- # ``` -->
 

### 2021-11-18 main.Rmd
filled all XXX except for Discusssion

@@
 The $sindex$ falls from around $20$ when $k=3$ to $19$ when $k=4$ and further decreases to $17.5$ with $k>7$.
-Thus, there are few choices available for the number of clusters.--> Figure \ref{fig:summary-plot-wpd} whos the wpd values for the three different granularities of the 24 customers, coloured by the 6 cluster solution.
+Thus, there are few choices available for the number of clusters.
+Figure \ref{fig:summary-plot-wpd} whos the wpd values for the three different granularities of the 24 customers, coloured by the 6 cluster solution.
 XXX REWORK THIS WHEN NEW CLUSTER RESULTS ARE SHOWN<!--The groupings with $k=4$ is shown through a parallel coordinate plot with the three significant cyclic granularities.
@@
 The findings vary from JS-based clustering, yet it is a helpful grouping.-->
+
+Figure \ref{fig:summary-plot-wpd} shows the $wpd$ values $hod$, $moy$, and $wnwd$ of the $24$ customers, colored by $3$ (a) and $5$ (b) cluster solution as suggested by Figure \ref{fig:opt-cluster-tsne-jsd} through a parallel coordinate plot.
+The variables are
+standardized prior to clustering using WPD.
+In the display, the variables are sorted according to their separation across groups.
+This means that $wnwd$ is the most important variable in distinguishing the groups, followed by $hod$ and $moy$ for both (a) and (b).
+Group P-1 and P-2 correspond to Q-1 and Q-2 respectively.
+Cluster P-3 splits into Q-3, Q-4 and Q-5.
+Customers 16 and 18 are characterized by unusual high values of $wpd$ on $wnwd$ compared to the rest of the customers and hence form a group.
+This could again be verified from figure \ref{fig:hod-ind-group-png}, where these were the only two customers with a difference in their $wnwd$ behavior.
+They are represented by P-1.
+P-2 has lower $wpd$ for $hod$ than $moy$ and $wnwd$.
+This can also be verified from Table \ref{tab:summary-table-wpd} where $hod$ has very low median values of $wpd$ over the members in the cluster, implying, at least half of these customers do not have an interesting $hod$ pattern.
+P-3 behaves exactly opposite to P-2 with higher $wpd$ for $hod$ compared to $moy$ and $wnwd$.
+These are the customers who have some significant pattern across $hod$ and this can again be validated by looking at \ref{fig:hod-ind-group-png}.
+For $5$ cluster solution, this groups gets split into Q3-Q5 characterizing different relative significance over $moy$ and $wnwd$.
+For example, Q-4 and Q-5 have almost no pattern across $moy$, but Q-3 has a $moy$ pattern and thus it is reasonable to split them.
+The patterns could be different, but they are significant.
+Q-4 and Q-5 are separated because of their significance across $hod$.
 

### 2021-11-18 main.Rmd
done tweaking the application section

@@
 For this reason, we have chosen to select a small group of customers with relatively distinct and different patterns in order to illustrate the clustering more simply.
+Figure \ref{fig:hod-ind-group-png} shows the distribution across $hod$, $moy$ and $wnwd$ for the set of $24$ customers used to illustrate clustering.
+The customers are displayed in two columns of $12$ for space reasons.
+Each row, of each column, represents the profile of a single customer across different variables.
 This is often a good approach to tackling a big analysis task, to start with a simpler task.
@@
 <!-- Why instance selection -->
-Figure \ref{fig:hod-ind-group-png} shows the distribution across $hod$, $moy$ and $wnwd$ for the set of $24$ customers used to illustrate clustering.
-The customers are displayed in two columns of $12$ for space reasons.
-Each row, of each column, represents the profile of a single customer across different variables.
-
-It is common to pre-process data prior to fitting a supervised learning model using instance selection [@olvera2010review] which removes observations that might impede the model building.
+
+It is common to filter data prior to fitting a supervised classification model using instance selection [@olvera2010review] which removes observations that might impede the model building.
 For clustering, this is analogous to identifying and removing nuisance observations.
-Prototype selection is more severe than instance selection, because a handful of cases is selected.
+Prototype selection is more severe than instance selection, because only a handful of cases is selected.
 @Cutler1994-hm proposed a method called archetypal analysis which has inspired this approach but the procedure we have used follows @Fan2021-bq.
@@
 Figure \ref{fig:summary-plot-wpd} shows the $wpd$ values $hod$, $moy$, and $wnwd$ of the $24$ customers, colored by $3$ (a) and $5$ (b) cluster solution as suggested by Figure \ref{fig:opt-cluster-tsne-jsd} through a parallel coordinate plot.
-The variables are
+The variables (wpd's for different granularities) are
 standardized prior to clustering using WPD.
@@
 One would need to post-process the cluster to separate specific patterns.
-XXX Might still need some work to summarise the application section
 

### 2021-11-18 main.Rmd
Proofread Section 1 and 2

@@
 abstract: |
- Clustering is a potential approach for organizing large collections of times series into small homogeneous groups, but a difficult step is determining an appropriate metric to measure similarity between time series.
+ Clustering is a potential approach for organizing large collections of time series into small homogeneous groups, but a difficult step is determining an appropriate metric to measure similarity between time series.
 The similarity metric needs to be capable of accommodating long, noisy, and asynchronous time series and also capture cyclical patterns.
@@
 Both are compatible with a variety of clustering algorithms.
-Cyclic granularities like hour-of-the-day, work-day/weekend, month-of-the-year, are useful for finding repeated patterns in the data.
+Cyclic granularities like hour-of-the-day, work-day/weekend, and month-of-the-year, are useful for finding repeated patterns in the data.
 Measuring similarity based on probability distributions across cyclic granularities serves two purposes: (a) characterizing the inherent temporal data structure of long, unequal-length time series in a manner robust to missing or noisy data; (b) small pockets of similar repeated behaviors can be captured.
@@
 Defining an appropriate distance measure for the raw time series may be a difficult task since it must take into account noise, variable lengths of time series, asynchronous time series, different scales, and missing data.
-Commonly used distance-based similarity measures as suggested by a decade review of time series clustering approaches [@Aghabozorgi2015-ct] are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping (DTW), Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances [@liao2007clustering] but on a vector time series data.
+Commonly used distance-based similarity measures as suggested by a decade review of time series clustering approaches [@Aghabozorgi2015-ct] are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping (DTW), Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances [@liao2007clustering] but on vector time series data.
 Among these alternatives, Euclidean distances have high performance but need the same length of data over the same period, resulting in information loss regardless of whether it is on raw data or a smaller collection of features.
 DTW works well with time series of different lengths [@corradini2001dynamic], but it is incapable of handling missing observations.
-Surprisingly, probability distributions, which may reflect the inherent temporal structure of a time series have not been considered in determining time series similarity.
+Surprisingly, probability distributions, which may reflect the inherent temporal structure of a time series, have not been considered in determining time series similarity.
 
@@
 These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and sporadic missing values.
-Using probability distributions is a natural way to analyze this types of data because they are robust to uneven length, missing data, or noise.
+Using probability distributions is a natural way to analyze these types of data because they are robust to uneven length, missing data, or noise.
 This paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across a selection of cyclic granularities.
-Cyclic temporal granularities [@Gupta2021-hakear], which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can measure repetitive patterns in large univariate time series data.
+Cyclic temporal granularities [@Gupta2021-hakear], which are temporal deconstructions of a time period into units such as hour-of-the-day or work-day/weekend, can measure repetitive patterns in large univariate time series data.
 The resulting clusters are expected to group customers that have similar repetitive behaviors across cyclic granularities.
@@
 The primary application of this work is data from the Smart Grid, Smart City (SGSC) project (2010–2014) available through the [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
-Half-hourly measurements of usage for more than 13,000 electricity smart meter customers is provided from from October 2011 to March 2014.
+Half-hourly measurements of usage for more than 13,000 electricity smart meter customers is provided from October 2011 to March 2014.
 <!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Customers vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
@@
 Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior complex.
-The challenge is to be able to cluster consumers into these type of expected patterns, and other unexpected patterns, using only their energy usage history [@Ushakova2020-rl].
+The challenge is to be able to cluster consumers into these types of expected patterns, and other unexpected patterns, using only their energy usage history [@Ushakova2020-rl].
 <!--To safeguard the customers' privacy, it is probable that such information is not accessible.
@@
 Only one study [@ozawa2016determining] identified time series characteristics by first conducting a Fourier transformation, to convert data from time to frequency domain, followed by $k$-means to cluster by greatest frequency.
-@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered.
-[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
+@Motlagh2019-yj suggest that the time feature extraction is limited by the type of noisy, patchy, and unequal time series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structures or sets of parameters which can be more easily characterized and clustered.
+@chicco2010renyi address information theory-based clustering such as Shannon or Renyi entropy and its variations.
 @Melnykov2013-sp discusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
@@
 <!--The foundation of our method is unsupervised clustering algorithms based exclusively on the time series data.
-The proposed methods leverage the intrinsic data structure hidden within cyclic temporal granularities.--> The existing work on clustering probability distributions assumes we have an independent and identically distributed samples $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+The proposed methods leverage the intrinsic data structure hidden within cyclic temporal granularities.--> The existing work on clustering probability distributions assumes we have independent and identically distributed samples $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
 <!--So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.--> In our approach, instead of considering the probability distributions of the linear time series, we compare them across different categories of a cyclic granularity.
 We can consider categories of an individual cyclic granularity ($A$) or combination of categories for two interacting granularities ($A*B$) to have a distribution, where $A, B$ are defined as $A = \{a_j: j=1, 2, \dots J\}$ and $B = \{b_k: k = 1, 2, \dots K\}$.
-For example, let us consider two cyclic granularities, $A$ and $B$, representing hour-of-day and day-of-week, repsectively.
+For example, let us consider two cyclic granularities, $A$ and $B$, representing hour-of-day and day-of-week, respectively.
 Then $A = \{0, 1, 2, \dots, 23\}$ and $B = \{Mon, Tue, Wed, \dots, Sun\}$.
@@
 
-@Gupta2021-hakear provides a distance measure ($wpd$) for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
+@Gupta2021-hakear provide a distance measure ($wpd$) for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
 (This extends to harmonies, pairs of granularities that might interact with each other.) We define "significant" granularities as those with significant distributional differences across at least one category.
@@
 Skewed distributions need to be symmetrized.
-Scales of individuals needs to be standardized, because clustering is to select similar patterns, not magnitude of usage.
+Scales of individuals need to be standardized, because clustering is to select similar patterns, not magnitude of usage.
 (Organizing individuals based on magnitude can be achieved simply by sorting on a statistic like the average value across time.) <!--Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
@@
 This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS).<!-- NQT is used in @Gupta2021-hakear prior sorting granularities.-- $wpd$, which is the foundation of wpd-based distances.
-XX wpd was not yet introduced--> While @Gupta2021-hakear already uses NQT when computing $wpd$, it could be useful to standardize it for the selected set of significant granularities prior to computing the distances.
-
-- RS: The normalized $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
-Note that, $v_{norm}$ has zero mean and median, but otherwise the shape does not change.
+XX wpd was not yet introduced--> While @Gupta2021-hakear already use NQT when computing $wpd$, it could be useful to standardize it for the selected set of significant granularities prior to computing the distances.
+
+- RS: The normalized $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentiles of the time series for the $i^{th}$ observation.
+Note that $v_{norm}$ has zero mean and median, but otherwise the shape does not change.
 
@@
 Many metrics have been defined for choosing clusters.
-Most metrics for choosing the optimal number of clusters, are based on comparing distances between observations within a class to those distances between observations between classes, which makes the assumption that there are some separated clusters.
+Most metrics for choosing the optimal number of clusters are based on comparing distances between observations within a class to those distances between observations between classes, which makes the assumption that there are some separated clusters.
 Some common procedures include the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@Dunn1973-gk] and the separation index ($sindex$) [@Hennig2019-eq; @Hennig2014-ah].
@@
 
-All of the common approaches can give contradictory suggestions for the optimal number of clusters, particularly when the data does not naturally break into groups, or in the presence of nuisance variables (no contribution to clustering) or nuisance observations (inlying outlying observations falling between clusters).
-There is no one best metric, which is perhaps a reason why metrics so many exist.
+All of the common approaches can give contradictory suggestions for the optimal number of clusters, particularly when the data does not naturally break into groups, or in the presence of nuisance variables (no contribution to clustering) or nuisance observations (inlying and outlying observations falling between clusters).
+There is no one best metric, which is perhaps a reason why so many metrics exist.
 
 In this work, we have chosen to use $sindex$.
-It is a very simple, but effective metric.
+It is a very simple but effective metric.
 This is computed by averaging the smallest 10% of inter-cluster distances.
@@
 This can be done numerically, by tabulating cluster means and standard deviations [@dasu2005grouping], and visually using methods for graphics multivariate data.
-@Cook2007-qe provides visual examples using both tours [@asimov1985grand] and parallel coordinate plots [@wegman1990hyperdimensional].
+@Cook2007-qe provide visual examples using both tours [@asimov1985grand] and parallel coordinate plots [@wegman1990hyperdimensional].
 Dimension reduction techniques like principal component analysis [@jolliffe2016principal], multidimensional scaling (MDS) [@borg2005modern], t-distributed stochastic neighbor embedding (t-SNE) [@van2008visualizing] and linear discrimant analysis (LDA) [@fisher1936use] are also useful.

### 2021-11-18 main.Rmd

--- 2021-11-18 main.Rmd
@@
 Only one study [@ozawa2016determining] identified time series characteristics by first conducting a Fourier transformation, to convert data from time to frequency domain, followed by $k$-means to cluster by greatest frequency.
-@Motlagh2019-yj suggest that the time feature extraction is limited by the type of noisy, patchy, and unequal time series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structures or sets of parameters which can be more easily characterized and clustered.
-@chicco2010renyi address information theory-based clustering such as Shannon or Renyi entropy and its variations.
+@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered.
+@chicco2010renyi addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
 @Melnykov2013-sp discusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.

### 2021-11-18 main.Rmd
Notational change following Emi's recommendation

@@
 <!--So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.--> In our approach, instead of considering the probability distributions of the linear time series, we compare them across different categories of a cyclic granularity.
-We can consider categories of an individual cyclic granularity ($A$) or combination of categories for two interacting granularities ($A*B$) to have a distribution, where $A, B$ are defined as $A = \{a_j: j=1, 2, \dots J\}$ and $B = \{b_k: k = 1, 2, \dots K\}$.
+We can consider categories of an individual cyclic granularity ($A$) or combination of categories for two interacting granularities ($A, B$) to have a distribution, where $A$ and $B$ are defined as $A = \{a_j: j=1, 2, \dots J\}$ and $B = \{b_k: k = 1, 2, \dots K\}$.
 For example, let us consider two cyclic granularities, $A$ and $B$, representing hour-of-day and day-of-week, respectively.
@@
 
-The sum of the distances between two observations $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$).
+The sum of the distances between two observations $x$ and $y$ in terms of a cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j\in A} D(x_j,y_j)$$ where $D$ is the Jensen-Shannon distances, $x_j$ is the set of quantiles over the values filtered by $j^{th}$ level of granularity $A$ for observation $x$ (similar for $y$).
+
+The sum of the distances between two observations $x$ and $y$ in terms of a pair of cyclic granularities $(A, B)$ is defined as $$S^{A,B}_{x,y} = \sum_{(j, k) \in (A, B)} D(x_{jk},y_{jk})$$ $x_{jk}$ is the set of quantiles over the values filtered by the combination of $j^{th}$ level of granularity $A$ and $k^{th}$ level of granularity $B$ for observation $x$ (similar for $y$).
+
 After determining the distance between two series in terms of one granularity, we must combine them to produce a distance based on all significant granularities.
-When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels, $$S_{x, y} = \sum_lS^{C_l}_{x,y}/n_l$$ is employed, which is also a distance metric since it is the sum of JS distances.
+When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels, $$S_{x, y} = \sum_{l \in L}S^{C_l}_{x,y}/n_l$$ is employed, which is also a distance metric since it is the sum of JS distances.
 This approach is expected to yield groups, such that the variation in observations within each group is in magnitude rather than distributional pattern, while the variation between groups is only in distributional pattern across categories.

### 2021-11-18 main.Rmd
validation proofread and positioning table for improved reading

@@
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
- fig.path = "figure/", fig.align = "center", fig.show = "hold",
- cache = TRUE, cache.path = "cache/",
+ fig.path = "figure/",
+ fig.align = "center",
+ fig.show = "hold",
+ cache = TRUE,
+ cache.path = "cache/",
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
@@
 As a result, in this scenario, we analyze a combination of categories generated from different distributions.
-Display of design and related results can be found in supplementary.
+Display of design and related results can be found in supplementary material.
 
@@
 This can be used to determine the number of clusters for each scenario.
-When $sindex$ for each scenario are examined, it appears that $k = 5, 4, 4$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in $sindex$ from that value of $k$.
+When $sindex$ for each scenario are examined, it appears that $k = \{5, 4, 4\}$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in $sindex$ from that value of $k$.
 Thus, the number of clusters corresponds to the number of designs that were originally considered in each scenario.
@@
 Additionally, there were missing values in many series.
-(The supplementary material contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless there is any systematic structure is related to granularities.
+(The supplementary material contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless there is any systematic structure related to granularities.
 

### 2021-11-18 main.Rmd
double spaced removed

@@
 Even though the $sindex$ for JS-NQT suggests three clusters, the five cluster solution more closely matches the original prototypes.
-The WPD clustering provides a different grouping of the customers, and even though it disagrees with the original prototypes it is a useful grouping.
-<!--While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude.
+The WPD clustering provides a different grouping of the customers, and even though it disagrees with the original prototypes it is a useful grouping.<!--While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude.
 The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales.
@@
 <!-- Our methodology is useful for grouping similar distributions over $hod$ and $moy$ and they are placed closely for easy comparison.
-Few groups have mixed patterns across $hod$ and $moy$, but few have all customers in the group having a similar profile.
--->
+Few groups have mixed patterns across $hod$ and $moy$, but few have all customers in the group having a similar profile.-->
 

### 2021-11-18 main.Rmd
some more proofread

@@
 In case individual granularities are considered, there are $J = 24$ distributions of the form $f_{i,j}(v)$ or $K = 7$ distributions of the form $f_{i,k}(v)$ for each customer $i$.
-In case of interaction, $J*K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
+In case of interaction, $J \times K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
 Hence clustering these customers is equivalent to clustering these collections of conditional univariate probability distributions.
@@
 JS-based approaches corresponding to NQT and RS are referred to as JS-NQT and JS-RS respectively.
-In the following plots, results for JS-NQT are reported, and results with JS-RS or wpd-based distances, which are similar are in the  supplementary material.
+In the following plots, results for JS-NQT are reported, and results with JS-RS or wpd-based distances are in the supplementary material.
 

### 2021-11-18 main.Rmd
proofread punctuations

@@
 @Cutler1994-hm proposed a method called archetypal analysis which has inspired this approach but the procedure we have used follows @Fan2021-bq.
-First dimension reduction such as t-SNE, MDS or PCA is used to project the data into a 2D space.
-Second a few "anchor" customers far apart in 2D space are selected.
+First, dimension reduction such as t-SNE, MDS or PCA is used to project the data into a 2D space.
+Second, a few "anchor" customers far apart in 2D space are selected.
 Additional close neighbors to the anchors are selected.
@@
 The t-SNE visualization suggests that there are four well-separated clusters.
-It is possible that because the representation is only 2D, that the fifth group from the original prototype selection is distinctly different in high dimensions.
+It is possible that because the representation is only 2D, the fifth group from the original prototype selection is distinctly different in high dimensions.
 The $sindex$ plots for the three methods indicate some disagreement: JS-NQT suggests $3$, JS-RS suggests $2$ or $5$ and WPD suggests $3$ or $5$.
@@
 P-3 behaves exactly opposite to P-2 with higher $wpd$ for $hod$ compared to $moy$ and $wnwd$.
-For $5$ cluster solution, this groups gets split into Q3-Q5 characterizing different relative significance over $moy$ and $wnwd$.", fig.height=6}
+For $5$ cluster solution, this groups gets split into Q3-Q5 characterizing different relative significance of $moy$ and $wnwd$.", fig.height=6}
 ```
@@
 Customers 16 and 18 are characterized by unusual high values of $wpd$ on $wnwd$ compared to the rest of the customers and hence form a group.
-This could again be verified from figure \ref{fig:hod-ind-group-png}, where these were the only two customers with a difference in their $wnwd$ behavior.
+This could again be verified from Figure \ref{fig:hod-ind-group-png}, where these were the only two customers with a difference in their $wnwd$ behavior.
 They are represented by P-1.
 P-2 has lower $wpd$ for $hod$ than $moy$ and $wnwd$.
-This can also be verified from Table \ref{tab:summary-table-wpd} where $hod$ has very low median values of $wpd$ over the members in the cluster, implying, at least half of these customers do not have an interesting $hod$ pattern.
+This can also be verified from Table \ref{tab:summary-table-wpd} where $hod$ has very low median values of $wpd$ over the members in the cluster, implying at least half of these customers do not have an interesting $hod$ pattern.
 P-3 behaves exactly opposite to P-2 with higher $wpd$ for $hod$ compared to $moy$ and $wnwd$.
 These are the customers who have some significant pattern across $hod$ and this can again be validated by looking at \ref{fig:hod-ind-group-png}.
-For $5$ cluster solution, this groups gets split into Q3-Q5 characterizing different relative significance over $moy$ and $wnwd$.
+For $5$ cluster solution, this groups gets split into Q3-Q5 characterizing different relative significance of $moy$ and $wnwd$.
 For example, Q-4 and Q-5 have almost no pattern across $moy$, but Q-3 has a $moy$ pattern and thus it is reasonable to split them.
@@
 This is quite a reasonable result and illustrates both the difficulties of clustering to obtain a particularly expected grouping and the ability to learn unexpected patterns in the data.
-It is possible that the JS-based distances were distracted by the presence of nuisance variables, levels of the granularities, that do not contribute to clustering.
+It is possible that the JS-based distances were distracted by the presence of nuisance variables, levels of the granularities that do not contribute to clustering.
 This would also be supported by the results of the validation study, where clustering was less effective in S2 and S3, where only some granularities had differences between levels.
@@
 A cluster summary like Figure \ref{fig:groups-4and5} is not possible because there may be different but equally interesting patterns (e.g.
-high evening $hod$ and high day time $hod$) in the same cluster.
+high evening $hod$ and high daytime $hod$) in the same cluster.
 Simply it provides information that across a collection of customers this specific cluster has interesting patterns in a granularity (eg $hod$).

### 2021-11-18 main.Rmd
say what the number in Figure 7 represents

@@
 Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from @smart-meter.
-The full data contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
+The full data contains half-hourly general supply in kWh for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
 We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation.
@@
 Each row, of each column, represents the profile of a single customer across different variables.
+Each customer is associated with an identifier of the form [$a$-$b$], where $a \in \{1, 2, \dots, 24\}$ represents the customer-prototype id and $b \in \{1, 2, \dots, 5\}$ indicates the label of the prototype in which a customer was placed.
 This is often a good approach to tackling a big analysis task, to start with a simpler task.
@@
 
-```{r summary-plot-wpd, fig.cap = "Summary plots for three (a) and five (b) clusters from WPD showing the $wpd$ values of each customers across $hod$, $moy$, and $wnwd$ through a parallel coordinate plot.
-Group P-1 and P-2 correspond to Q-1 and Q-2 respectively.
-Cluster P-3 splits into Q-3, Q-4 and Q-5.
-P-1 (customers 16, 18) is characterized by high values of $wpd$ on $wnwd$.
-P-2 has lower $wpd$ for $hod$ than $moy$ and $wnwd$.
-P-3 behaves exactly opposite to P-2 with higher $wpd$ for $hod$ compared to $moy$ and $wnwd$.
-For $5$ cluster solution, this groups gets split into Q3-Q5 characterizing different relative significance of $moy$ and $wnwd$.", fig.height=6}
+```{r summary-plot-wpd, fig.cap = "Summary plots for three (a) and five (b) clusters from WPD approach showing the $wpd$ values of each customers across $hod$, $moy$, and $wnwd$ through a parallel coordinate plot.
+P-1 and P-2 groups correspond to Q-1 and Q-2, respectively.
+Cluster P-3 is subdivided into Q-3, Q-4, and Q-5.
+P-1 (customers 16, 18) is distinguished by high $wpd$ on $wnwd$ values.
+P-2 has lower $wpd$ than $moy$ and $wnwd$ for $hod$.
+P-3 operates in the opposite way as P-2, with larger $wpd$ for $hod$ in comparison to $moy$ and $wnwd$.
+For the $5$ cluster solution, this group is divided into Q3-Q5, which are distinct due to their different relative significance of $moy$ and $wnwd$.", fig.height=6}
 ```

### 2021-11-18 main.Rmd
application rewording again

@@
 The $sindex$ falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases.
-In our study, we use $k=4, 5$.--> Figure \ref{fig:groups-4and5} displays the summarized distributions across $4$ and $5$ clusters in (a) and (b) respectively, and helps to characterize each cluster.
+In our study, we use $k=4, 5$.--> Figure \ref{fig:groups-4and5} displays the summarized distributions across $4$ and $5$ clusters from JS-NQT clustering in (a) and (b) respectively, and helps to characterize each cluster.
 In the quantile plots the line represents the median, and the region shows the area between the $25^{th}$ and $75^{th}$ percentiles.
@@
 
-Figure \ref{fig:summary-plot-wpd} shows the $wpd$ values $hod$, $moy$, and $wnwd$ of the $24$ customers, colored by $3$ (a) and $5$ (b) cluster solution as suggested by Figure \ref{fig:opt-cluster-tsne-jsd} through a parallel coordinate plot.
+Figure \ref{fig:summary-plot-wpd} shows the $wpd$ values of the $24$ customers over $hod$, $moy$, and $wnwd$ , colored by $3$ (a) and $5$ (b) clusters from WPD clustering using a parallel coordinate plot.
 The variables (wpd's for different granularities) are
@@
 This means that $wnwd$ is the most important variable in distinguishing the groups, followed by $hod$ and $moy$ for both (a) and (b).
-Group P-1 and P-2 correspond to Q-1 and Q-2 respectively.
+Groups P-1 and P-2 correspond to Q-1 and Q-2 respectively.
 Cluster P-3 splits into Q-3, Q-4 and Q-5.
@@
 P-2 has lower $wpd$ for $hod$ than $moy$ and $wnwd$.
-This can also be verified from Table \ref{tab:summary-table-wpd} where $hod$ has very low median values of $wpd$ over the members in the cluster, implying at least half of these customers do not have an interesting $hod$ pattern.
-P-3 behaves exactly opposite to P-2 with higher $wpd$ for $hod$ compared to $moy$ and $wnwd$.
-These are the customers who have some significant pattern across $hod$ and this can again be validated by looking at \ref{fig:hod-ind-group-png}.
-For $5$ cluster solution, this groups gets split into Q3-Q5 characterizing different relative significance of $moy$ and $wnwd$.
+P-3 behaves opposite to P-2 with higher $wpd$ for $hod$ compared to $moy$ and $wnwd$.
+These are the customers who have some significant pattern across $hod$ and this can again be validated by looking at Figure \ref{fig:hod-ind-group-png}.
+For a $5$ cluster solution, this group gets split into Q-3, Q-4 and Q-5 characterizing different relative significance of $moy$ and $wnwd$.
 For example, Q-4 and Q-5 have almost no pattern across $moy$, but Q-3 has a $moy$ pattern and thus it is reasonable to split them.
 The patterns could be different, but they are significant.
-Q-4 and Q-5 are separated because of their significance across $hod$.
+Q-4 and Q-5 are separated because of their different significance of $hod$.
+All of these can also be verified from Table \ref{tab:summary-table-wpd} which shows the cluster summaries with members and median values of $wpd$ for the three variables.
+
+<!-- $hod$ has very low median values of $wpd$ for the members in the cluster, implying at least half of these customers do not have an interesting $hod$ pattern. -->
 

### 2021-11-18 main.Rmd
supplementary materials section update

@@
 The Monash eResearch Centre and eSolutions-Study Support Services supported this research in part through the resource usage of the MonARCH HPC Cluster.
-The Github repository, [github.com/Sayani07/paper-gracsr](https://github.com/Sayani07/paper-gracsr), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
+The Github repository, [github.com/Sayani07/paper-gracsr](https://github.com/Sayani07/paper-gracsr), contains all materials required to reproduce this article and the code is also available online in the supplementary materials.
 This article was created with R [@R-language], `knitr` [@knitr; @R-knitr] and `rmarkdown` [@rmarkdown; @R-rmarkdown].
@@
 
-# Supplementary Materials
-
-**Data and scripts:** Data sets and R code to reproduce all figures in this article (main.R).
-
-**Supplementary paper:** Additional tables, graphics and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd).
-The code for creating validation designs and running the methodologies is available at (https://github.com/Sayani07/paper-gracsR/Validation).
-
-**R-package:** To implement the ideas provided in this research, the open-source R package 'gracsr' is available on Github (https://github.com/Sayani07/gracsr).
+# Supplementary materials
+
+**Data and scripts:** Data sets and R codes to reproduce all figures in this article  
+(validation.R, application.R)
+
+\noindent
+**Supplementary paper:** Additional graphics and R codes to reproduce it  
+(paper-supplementary.pdf, paper-supplementary.Rmd, supplementary.R)
+
+\noindent
+**R-package:** To implement the ideas provided in this research, the open-source R package `gracsr` is available on Github (https://github.com/Sayani07/gracsr).
 
